### 2022-09-16 background.tex
mainly the Colin's stuff

\section{Background}
\label{sec:background}
\subsection{Virtual Memory Managers}
\label{sec:backgroundonvmm}
\subsection{Modal Logic}
\label{sec:backgroundonmodallogic}
Work on modal logic extends back for decades with many widely-varied variations; we discuss here only the pieces most relevant to our work.
Broadly speaking modal logics incorporate \emph{modal operators}, which take as arguments a proposition expected to be true in another time~\cite{pnueli1977temporal}, place~\cite{gordon2019modal,goranko1996hierarchies,areces2001hybrid,gargov1993modal}, or circumstance~\cite{hintikka1962knowledge,halpern1985guide}, and result in a proposition true in the \emph{current} time, place, or circumstance at which the truth of the use of the modal operator is being evaluated. Classic examples include modal necessity $\square P$ describing that $P$ is \emph{necessarily} true, $\mathsf{G}~P$ meaning $P$ is true \emph{globally} (i.e., forever from this time onwards), or $K_i(P)$ describing that a particular participant $i$ \emph{knows} that $P$ is true. The latter is an example of \emph{multimodal} logic, where there is an indexed family of modalities (modal operators) parameterized by some dimension of interest (there, participants).
A closely related variant of multimodal logic is \emph{dynamic logic}~\cite{pratt1976semantical}, a logic of weakest preconditions~\cite{dijkstra-75} which works with modalities of the form $[p](P)$, which states that \emph{in the current program state}, \emph{if} program $p$ is run then afterwards $P$ will hold (modulo non-termination).
This same idea is used to encode Hoare triples in the Iris program logic~\cite{krebbers2017essence}, using the same encoding as in Pratt's original presentation, where a Hoare triple $\{P\}C\{Q\}$ is encoded as $P\rightarrow[C](Q)$. Unlike classic work on dynamic logic~\cite{harel2000dynamic}, Iris applies these ideas in a \emph{substructural} setting (separation logic) where distributivity laws over substructural connectives must be considered. For both the dynamic modality and the later modality $\blacktriangleright$ used for guarded recursive predicates, these modalities satisfy axioms of the form $M(P)\ast Q\vdash M(P\ast Q)$, but not the reverse.
Iris is not the first combination of modal and substructural logic~\cite{dovsen1992modal,restall1993modalities,d1997grafting,kamide2002kripke,licata2017fibrational}, but is certainly the most heavily tested.
A hallmark of a unary logical operator $M$ being a modality is if $M$ satisfies a property akin to $(\upvarphi\rightarrow\psi)\rightarrow M(\upvarphi)\rightarrow M(\psi)$, which roughly states that modus ponens holds under the modality.\footnote{Afficionados of modal logic will note that this property is not quite Axiom K (which requires the initial implication to also hold under $M$), but follows from K and a necessitation rule $\upvarphi\rightarrow M(\upvarphi)$. Non-necessitive modalities typically satisfy the weaker property we call out above.}
% Most pertinent to our work is that Iris itself has a modalities built into the logic to solve problems with step-indexing / guardedness (the later modality) and general infrastructure for duplicable assumptions / resources (the persistent modality, reminiscent of linear logic's exponential). And of course the central concept in the use of Iris as a program logic is the use of the weakest precondition modality, which like our modalities is defined \emph{within} the logic itself, rather than being part of the base logic.

Another pillar of our work is hybrid logic, a branch of modal logic with \emph{names} (called \emph{nominals}) for states in Kripke models~\cite{blackburn1995hybrid,goranko1996hierarchies,areces2001hybrid,gargov1993modal}, in contrast to the typical flavor of modality which refers to an \emph{unspecified} (in the assertion) set of alternative circumstances.

Our work draws on ideas from hybrid modal logics, applied in the context of Iris's higher-order separation logic~\cite{jung2018iris}. We define our modalities and prove their rules directly within Iris, so there is no model-theoretic novelty in our work. Instead, we show that targeted use of modalities combining the ideas of \emph{named resources} with the power of substructural reasoning enable clear specification of general reasoning principles for data structures. Along the way we take advantage of the fact that modalities need not be injections between the \emph{same} logics, but in fact the propositions that modal operators contextualize can come from \emph{other} logics; in our case we exploit the fact that Iris itself is parameterized by a choice of step-indexed~\cite{ahmed-appel-virga-02} \textsf{BI}-algebra~\cite{ohearn1999bunched} to allow convenient specifications in embedded data-structure-relative logics.

### 2022-09-16 cpp_paper.tex
mainly the Colin's stuff

%%
%% This is file `sample-sigplan.tex',
%% generated with the docstrip utility.
%%
%% The original source files were:
%%
%% samples.dtx  (with options: `sigplan')
%% 
%% IMPORTANT NOTICE:
%% 
%% For the copyright see the source file.
%% 
%% Any modified versions of this file must be renamed
%% with new filenames distinct from sample-sigplan.tex.
%% 
%% For distribution of the original source see the terms
%% for copying and modification in the file samples.dtx.
%% 
%% This generated file may be distributed as long as the
%% original source files, as listed above, are part of the
%% same distribution. (The sources need not necessarily be
%% in the same archive or directory.)
%%
%%
%% Commands for TeXCount
%TC:macro \cite [option:text,text]
%TC:macro \citep [option:text,text]
%TC:macro \citet [option:text,text]
%TC:envir table 0 1
%TC:envir table* 0 1
%TC:envir tabular [ignore] word
%TC:envir displaymath 0 word
%TC:envir math 0 word
%TC:envir comment 0 0
%%
%%
%% The first command in your LaTeX source must be the \documentclass command.

\documentclass[sigplan,10pt,anonymous,review]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
%%
%% \BibTeX command to typeset BibTeX logo in the docs
\AtBeginDocument{%
  \providecommand\BibTeX{{%
    Bib\TeX}}}

%% Rights management information.  This information is sent to you
%% when you complete the rights form.  These commands have SAMPLE
%% values in them; it is your responsibility as an author to replace
%% the commands and values with those provided to you when you
%% complete the rights form.
\setcopyright{acmcopyright}
\copyrightyear{2018}
\acmYear{2018}
\acmDOI{XXXXXXX.XXXXXXX}

%% These commands are for a PROCEEDINGS abstract or paper.
\acmConference[Conference acronym 'XX]{Make sure to enter the correct
  conference title from your rights confirmation emai}{June 03--05,
  2018}{Woodstock, NY}
\acmPrice{15.00}
\acmISBN{978-1-4503-XXXX-X/18/06}

%%
%% Submission ID.
%% Use this when submitting an article to a sponsored event. You'll
%% receive a unique submission ID from the organizers
%% of the event, and this ID should be used as the parameter to this command.
%%\acmSubmissionID{123-A56-BU3}

%%
%% For managing citations, it is recommended to use bibliography
%% files in BibTeX format.
%%
%% You can then either use BibTeX with the ACM-Reference-Format style,
%% or BibLaTeX with the acmnumeric or acmauthoryear sytles, that include
%% support for advanced citation of software artefact from the
%% biblatex-software package, also separately available on CTAN.
%%
%% Look at the sample-*-biblatex.tex files for templates showcasing
%% the biblatex styles.
%%

%%
%% The majority of ACM publications use numbered citations and
%% references.  The command \citestyle{authoryear} switches to the
%% "author year" style.
%%
%% If you are preparing content for an event
%% sponsored by ACM SIGGRAPH, you must use the "author year" style of
%% citations and references.
%% Uncommenting
%% the next command will enable that style.
%%\citestyle{acmauthoryear}

%%
%% end of the preamble, start of the body of the document source.
\begin{document}

%%
%% The "title" command has an optional parameter,
%% allowing the author to define a "short title" to be used in page headers.
\title{Modalities for Location Virtualization}
\subtitle{Virtualizing Memory Addresses}
%%
%% The "author" command and its associated commands are used to define
%% the authors and their affiliations.
%% Of note is the shared affiliation of the first two authors, and the
%% "authornote" and "authornotemark" commands
%% used to denote shared contribution to the research.
\author{X X}
\authornote{Both authors contributed equally to this research.}
\email{trovato@corporation.com}
\orcid{1234-5678-9012}
\author{G.K.M. Tobin}
\authornotemark[1]
\email{webmaster@marysville-ohio.com}
\affiliation{%
  \institution{Institute for Clarity in Documentation}
  \streetaddress{P.O. Box 1212}
  \city{Dublin}
  \state{Ohio}
  \country{USA}
  \postcode{43017-6221}
}

%%
%% By default, the full list of authors will be used in the page
%% headers. Often, this list is too long, and will overlap
%% other information printed in the page headers. This command allows
%% the author to define a more concise list
%% of authors' names for this purpose.
\renewcommand{\shortauthors}{Trovato et al.}

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
Operating system kernels employs virtual memory subsystem which employs memory management units on each CPU to virtualize the addresses of memory regions:a logical (virtual) address which is translated to a physical address in memory by the MMU based on
kernel-controlled page tables -- a hardware-defined sparse tree-map structure -- stored in memory, which itself is accessed by the kernel through virtual addresses.
With this core mechanism, operating system kernels generalizes its memory-resource service to different contexts, mainly through the manipulating and aliasing 
same physical memory to multiple virtual addresses through the tree structure: subtrees or data pages may be shared across processes (e.g., the kernel mapping),
accessing an unmapped address produces a fault which the OS can use to kill the process or take corrective action (like add a mapping), and resume the interrupted code which constitutes the kernel's capability to restrict which memory is accessible to different processes, hide memory limits from user programs, ensure process isolation, implement paging and copy-on-write behaviors for performance
and resource controls.

So, verifying a full-featured version of the essential virtual memory manager systems is a crucial step towards achieving the progress towards handling fully verified real kernels written using these designs. 
Regarding this effort, in this paper, we introduce a modal abstraction to describe
the truth of assertions in the virtual resources corresponding to virtual address spaces: [r]P indicating that P holds in the
virtual address space rooted at r, which implicitly treats the memory state as a modal frame and the choice of
page table root as a world in Kripke-style semantics. Moreover, we define virtual points-to assertions defined in terms of physical points-to assertions
and hardware-specific address translation assertions, and paged points-to assertions. Finally, we experiment our abstractions in specifying and proving demand paging of a fresh page of zeros and context
switches between address spaces. All definitions and theorems mentioned in this paper including the operational model of a RISC-like fragment of x86-64, a simple language run on this operational model, and a logic as an instantiation of the Iris framework are mechanized inside Coq.
\end{abstract}

%%
%% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
%% Please copy and paste the code instead of the example below.
%%
\begin{CCSXML}
<ccs2012>
 <concept>
  <concept_id>10010520.10010553.10010562</concept_id>
  <concept_desc>Computer systems organization~Embedded systems</concept_desc>
  <concept_significance>500</concept_significance>
 </concept>
 <concept>
  <concept_id>10010520.10010575.10010755</concept_id>
  <concept_desc>Computer systems organization~Redundancy</concept_desc>
  <concept_significance>300</concept_significance>
 </concept>
 <concept>
  <concept_id>10010520.10010553.10010554</concept_id>
  <concept_desc>Computer systems organization~Robotics</concept_desc>
  <concept_significance>100</concept_significance>
 </concept>
 <concept>
  <concept_id>10003033.10003083.10003095</concept_id>
  <concept_desc>Networks~Network reliability</concept_desc>
  <concept_significance>100</concept_significance>
 </concept>
</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Computer systems organization~Embedded systems}
\ccsdesc[300]{Computer systems organization~Redundancy}
\ccsdesc{Computer systems organization~Robotics}
\ccsdesc[100]{Networks~Network reliability}

%%
%% Keywords. The author(s) should pick words that accurately describe
%% the work being presented. Separate the keywords with commas.
\keywords{datasets, neural networks, gaze detection, text tagging}

\maketitle
\input{intro.tex}
\input{background.tex}
\input{syntax.tex}
\input{semantics.tex}
\input{programlogic.tex}
\input{examples.tex}
\input{discussion.tex}
%% The next two lines define the bibliography style to be used, and
%% the bibliography file.
\bibliographystyle{ACM-Reference-Format}
\bibliography{sample-base}

\end{document}
%%
%% End of file `sample-sigplan.tex'.

### 2022-09-16 intro.tex
mainly the Colin's stuff

\section{Introduction}
\label{sec:intro}
In the UNIX-inspired designs typical of modern general-purpose kernels [93, 82, 89, 91], virtual memory management, the filesystem, and device subsystem are tightly
coupled, rely upon other core OS services (scheduling, interrupt dispatch, etc.), and are central
to higher-level functionality. For example, in Solaris Internals [91], McDougall and Mauro write:
“The virtual memory sub-system can be considered the core of a Solaris instance, and the implementation of Solaris virtual memory affects just about every other subsystem in the operating
system.” So by verifying a kernel with full-featured versions of these essential systems, following the
designs kernel developers have found valuable enough to preserve for decades, we can make essential
progress towards handling real kernels written using these designs. This goal is feasible because
the verification challenges faced by these systems are all instances of just three core concepts from
systems research and development, which apply more broadly than just in OS kernels: location
virtualization, caching, and using asymmetric protocols to coordinate shared resource usage. We
believe all three concepts are amenable to direct treatment using tools from modal logic.

Virtual memory managers (VMMs) use hardware memory management units (MMUs) on each
CPU to virtualize the addresses of memory regions. Software actually accesses memory in terms of
a logical (virtual) address which is translated to a physical address in memory by the MMU based
on kernel-controlled page tables stored in memory: a hardware-defined sparse tree-map structure,
which itself is accessed by the kernel through virtual addresses. This tree structure may alias the
same physical memory to multiple virtual addresses. Whole trees are shared between threads of a
process, and subtrees or data pages may be shared across processes (e.g., the kernel mapping ~\cite{91,
93}). Accessing an unmapped address produces a fault which the OS can use to kill the process
or take corrective action (like add a mapping) and resume the interrupted code. This allows the
kernel to restrict which memory is accessible to different processes, hide memory limits from user
programs, ensure process isolation, implement paging and copy-on-write behaviors for performance
and resource controls, and even implement live migration of virtual machines in hypervisors ~\cite{110,
31}. Modern hardware extends this idea with an IOMMU ~\cite{71, 11} the OS can use to similarly
mediate hardware device access to memory.

We plan to use families of modalities to describe
the truth of assertions in the virtual resources corresponding to virtual address spaces or logical
locations in a file system: a modality for each concept, with corresponding assertions about the
underlying physical resource.
For virtual memory, this might correspond to an assertion [r]P indicating that P holds in the
virtual address space rooted at r. This is essentially what hybrid logic [19, 49, 10, 48] calls a
satisfaction operator, which evaluates the truth of a predicate in a named alternative state (here,
address space). An example usage of this would be as part of the invariant for a sleeping thread
— ∃R : registerset. Context(ctxt, R) ∗ ∃P : Prop. [R.cr3]P ∗ {P ∗ CurrentRegs(R)}(R.rip) — that the
context structure at virtual address ctxt holds a certain register set, and that the precondition for
the stored instruction pointer rip is that register set and the virtual invariant P, which is true in
the saved context’s address space rooted at cr3. This would be required to ensure safety of context
switches between address spaces.
Virtual points-to assertions would need to be defined in terms of physical points-to assertions
and hardware-specific address translation assertions (as in Kolanski and Klein’s work), and paged
points-to assertions indicating the VMM had committed swap space to hold the appropriate value
on disk. The semantics of address translation assertions would be given as an extension to Kolanski
and Klein’s approach, which implicitly treats the memory state as a modal frame and the choice of
page table root as a world in Kripke-style semantics. They stopped short, however, of incorporating
the choice of world into the logic’s assertions, or treating overlapping address spaces, or transitions

between address spaces. Overlapping address spaces — both shared data, and shared page tables
which may see updates — can be coordinated with protocols to ensure those updates preserve
address space invariants. For filesystem addressing modalities, the interpretation is dictated by the
entirely by filesystem design rather than the hardware, but the concept is similar.
Both systems have an implicit notion of a “current” world. CPUs have a current address space
based on their page table register (CR3 for x86-64), and the logic must support reasoning about
changing the current address space. Copy-on-write filesystems have a corresponding fixed location
on disk pointing to the physical block of the latest filesystem snapshot. For example, for CPUs, we
would ideally like proof rules close to the following:
{P ∗ CR3 = r1 ∗ [r2](Q) ∗ rax = r2}mov cr3 rax{[r1](P) ∗ CR3 = r2 ∗ Q ∗ rax = r2}
This rule captures the notion that switching address spaces pulls information out of one modality
into the “current view” of memory, and leaves other assertions true relative to the previous address
space. But real code requires stronger reasoning. Code that performs task switches typically
establishes an invariant in one address space — such as that the save-context of the thread being
taken off the CPU is valid — that must still be known true in the new address space. The right
way to expose this knowledge transfer is not obvious

### 2022-09-16 examples.tex
All the files TODOs separated

\section{Experiment}
\label{sec:experiment}
\mytodo{Ismail's Part}

### 2022-09-16 programlogic.tex
All the files TODOs separated

\section{Program Logic}
\label{sec:programlogic}
\mytodo{Ismail's Part}

### 2022-09-16 semantics.tex
All the files TODOs separated

\section{Semantics}
\label{sec:semantics}
\mytodo{Ismail's Part}

### 2022-09-16 syntax.tex
All the files TODOs separated

\section{Syntax}
\label{sec:syntax}
\mytodo{Ismail's Part}

### 2022-09-17 background.tex
separted stuff

@@
 \label{sec:background}
+\subsection{Machine Model}
+\label{sec:backgroundonmachinemodel}
+\mytodo{Colin's part}
 \subsection{Virtual Memory Managers}
+\mytodo{Colin's part}
 \label{sec:backgroundonvmm}
@@
 \label{sec:backgroundonmodallogic}
-Work on modal logic extends back for decades with many widely-varied variations; we discuss here only the pieces most relevant to our work.
-Broadly speaking modal logics incorporate \emph{modal operators}, which take as arguments a proposition expected to be true in another time~\cite{pnueli1977temporal}, place~\cite{gordon2019modal,goranko1996hierarchies,areces2001hybrid,gargov1993modal}, or circumstance~\cite{hintikka1962knowledge,halpern1985guide}, and result in a proposition true in the \emph{current} time, place, or circumstance at which the truth of the use of the modal operator is being evaluated.
-Classic examples include modal necessity $\square P$ describing that $P$ is \emph{necessarily} true, $\mathsf{G}~P$ meaning $P$ is true \emph{globally} (i.e., forever from this time onwards), or $K_i(P)$ describing that a particular participant $i$ \emph{knows} that $P$ is true.
-The latter is an example of \emph{multimodal} logic, where there is an indexed family of modalities (modal operators) parameterized by some dimension of interest (there, participants).
-A closely related variant of multimodal logic is \emph{dynamic logic}~\cite{pratt1976semantical}, a logic of weakest preconditions~\cite{dijkstra-75} which works with modalities of the form $[p](P)$, which states that \emph{in the current program state}, \emph{if} program $p$ is run then afterwards $P$ will hold (modulo non-termination).
-This same idea is used to encode Hoare triples in the Iris program logic~\cite{krebbers2017essence}, using the same encoding as in Pratt's original presentation, where a Hoare triple $\{P\}C\{Q\}$ is encoded as $P\rightarrow[C](Q)$.
-Unlike classic work on dynamic logic~\cite{harel2000dynamic}, Iris applies these ideas in a \emph{substructural} setting (separation logic) where distributivity laws over substructural connectives must be considered.
-For both the dynamic modality and the later modality $\blacktriangleright$ used for guarded recursive predicates, these modalities satisfy axioms of the form $M(P)\ast Q\vdash M(P\ast Q)$, but not the reverse.
-Iris is not the first combination of modal and substructural logic~\cite{dovsen1992modal,restall1993modalities,d1997grafting,kamide2002kripke,licata2017fibrational}, but is certainly the most heavily tested.
-A hallmark of a unary logical operator $M$ being a modality is if $M$ satisfies a property akin to $(\upvarphi\rightarrow\psi)\rightarrow M(\upvarphi)\rightarrow M(\psi)$, which roughly states that modus ponens holds under the modality.\footnote{Afficionados of modal logic will note that this property is not quite Axiom K (which requires the initial implication to also hold under $M$), but follows from K and a necessitation rule $\upvarphi\rightarrow M(\upvarphi)$.
-Non-necessitive modalities typically satisfy the weaker property we call out above.}
-% Most pertinent to our work is that Iris itself has a modalities built into the logic to solve problems with step-indexing / guardedness (the later modality) and general infrastructure for duplicable assumptions / resources (the persistent modality, reminiscent of linear logic's exponential).
-And of course the central concept in the use of Iris as a program logic is the use of the weakest precondition modality, which like our modalities is defined \emph{within} the logic itself, rather than being part of the base logic.
-
-Another pillar of our work is hybrid logic, a branch of modal logic with \emph{names} (called \emph{nominals}) for states in Kripke models~\cite{blackburn1995hybrid,goranko1996hierarchies,areces2001hybrid,gargov1993modal}, in contrast to the typical flavor of modality which refers to an \emph{unspecified} (in the assertion) set of alternative circumstances.
-
-Our work draws on ideas from hybrid modal logics, applied in the context of Iris's higher-order separation logic~\cite{jung2018iris}.
-We define our modalities and prove their rules directly within Iris, so there is no model-theoretic novelty in our work.
-Instead, we show that targeted use of modalities combining the ideas of \emph{named resources} with the power of substructural reasoning enable clear specification of general reasoning principles for data structures.
-Along the way we take advantage of the fact that modalities need not be injections between the \emph{same} logics, but in fact the propositions that modal operators contextualize can come from \emph{other} logics; in our case we exploit the fact that Iris itself is parameterized by a choice of step-indexed~\cite{ahmed-appel-virga-02} \textsf{BI}-algebra~\cite{ohearn1999bunched} to allow convenient specifications in embedded data-structure-relative logics.
+\mytodo{Colin's part}

### 2022-09-17 intro.tex
separted stuff

@@
 \label{sec:intro}
-In the UNIX-inspired designs typical of modern general-purpose kernels [93, 82, 89, 91], virtual memory management, the filesystem, and device subsystem are tightly
-coupled, rely upon other core OS services (scheduling, interrupt dispatch, etc.), and are central
-to higher-level functionality. For example, in Solaris Internals [91], McDougall and Mauro write:
-“The virtual memory sub-system can be considered the core of a Solaris instance, and the implementation of Solaris virtual memory affects just about every other subsystem in the operating
-system.” So by verifying a kernel with full-featured versions of these essential systems, following the
-designs kernel developers have found valuable enough to preserve for decades, we can make essential
-progress towards handling real kernels written using these designs. This goal is feasible because
-the verification challenges faced by these systems are all instances of just three core concepts from
-systems research and development, which apply more broadly than just in OS kernels: location
-virtualization, caching, and using asymmetric protocols to coordinate shared resource usage. We
-believe all three concepts are amenable to direct treatment using tools from modal logic.
-
-Virtual memory managers (VMMs) use hardware memory management units (MMUs) on each
-CPU to virtualize the addresses of memory regions. Software actually accesses memory in terms of
-a logical (virtual) address which is translated to a physical address in memory by the MMU based
-on kernel-controlled page tables stored in memory: a hardware-defined sparse tree-map structure,
-which itself is accessed by the kernel through virtual addresses. This tree structure may alias the
-same physical memory to multiple virtual addresses. Whole trees are shared between threads of a
-process, and subtrees or data pages may be shared across processes (e.g., the kernel mapping ~\cite{91,
-93}). Accessing an unmapped address produces a fault which the OS can use to kill the process
-or take corrective action (like add a mapping) and resume the interrupted code. This allows the
-kernel to restrict which memory is accessible to different processes, hide memory limits from user
-programs, ensure process isolation, implement paging and copy-on-write behaviors for performance
-and resource controls, and even implement live migration of virtual machines in hypervisors ~\cite{110,
-31}. Modern hardware extends this idea with an IOMMU ~\cite{71, 11} the OS can use to similarly
-mediate hardware device access to memory.
-
-We plan to use families of modalities to describe
-the truth of assertions in the virtual resources corresponding to virtual address spaces or logical
-locations in a file system: a modality for each concept, with corresponding assertions about the
-underlying physical resource.
-For virtual memory, this might correspond to an assertion [r]P indicating that P holds in the
-virtual address space rooted at r. This is essentially what hybrid logic [19, 49, 10, 48] calls a
-satisfaction operator, which evaluates the truth of a predicate in a named alternative state (here,
-address space). An example usage of this would be as part of the invariant for a sleeping thread
-— ∃R : registerset. Context(ctxt, R) ∗ ∃P : Prop. [R.cr3]P ∗ {P ∗ CurrentRegs(R)}(R.rip) — that the
-context structure at virtual address ctxt holds a certain register set, and that the precondition for
-the stored instruction pointer rip is that register set and the virtual invariant P, which is true in
-the saved context’s address space rooted at cr3. This would be required to ensure safety of context
-switches between address spaces.
-Virtual points-to assertions would need to be defined in terms of physical points-to assertions
-and hardware-specific address translation assertions (as in Kolanski and Klein’s work), and paged
-points-to assertions indicating the VMM had committed swap space to hold the appropriate value
-on disk. The semantics of address translation assertions would be given as an extension to Kolanski
-and Klein’s approach, which implicitly treats the memory state as a modal frame and the choice of
-page table root as a world in Kripke-style semantics. They stopped short, however, of incorporating
-the choice of world into the logic’s assertions, or treating overlapping address spaces, or transitions
-
-between address spaces. Overlapping address spaces — both shared data, and shared page tables
-which may see updates — can be coordinated with protocols to ensure those updates preserve
-address space invariants. For filesystem addressing modalities, the interpretation is dictated by the
-entirely by filesystem design rather than the hardware, but the concept is similar.
-Both systems have an implicit notion of a “current” world. CPUs have a current address space
-based on their page table register (CR3 for x86-64), and the logic must support reasoning about
-changing the current address space. Copy-on-write filesystems have a corresponding fixed location
-on disk pointing to the physical block of the latest filesystem snapshot. For example, for CPUs, we
-would ideally like proof rules close to the following:
-{P ∗ CR3 = r1 ∗ [r2](Q) ∗ rax = r2}mov cr3 rax{[r1](P) ∗ CR3 = r2 ∗ Q ∗ rax = r2}
-This rule captures the notion that switching address spaces pulls information out of one modality
-into the “current view” of memory, and leaves other assertions true relative to the previous address
-space. But real code requires stronger reasoning. Code that performs task switches typically
-establishes an invariant in one address space — such as that the save-context of the thread being
-taken off the CPU is valid — that must still be known true in the new address space. The right
-way to expose this knowledge transfer is not obvious
+\mytodo{Colin's Part}

### 2022-09-17 syntax.tex
separted stuff
+

@@
 \label{sec:syntax}
-\mytodo{Ismail's Part}
+\subsection{Registers, Memory Indices}
+There exist two pieces of resources we access in our model: registers and memory.
+A register identifier, $\reg$, is chosen from a fixed finite set of register identifiers,$\regset$.
+We use these identifiers to access the register values, $\regval \in \regvaltype$.
+Unlike registers, we do not abstract the memory indices as a special type but instead, for the sake of clarity and ease of representation, we show differently masked machine words, $\loc \in \Loc$, with the subscripts showing the length of a 64-bit machine word after masking, e.g.
+$\kw{w}_{12}$ is a 12-bit resource which can be obtained after masking 52-bit of a 64-bit word.
+This simple syntax in \ref{fig:syntax} includes a simple syntax for values because our language is purely \textit{imperative}, i.e.
+an instruction does not return a value.
+Intentionally, we restrict any stream of instruction, $\instrs$, to eventually reduce to $\iskip$ instruction which is injected into unit.
+\input{fig-syntax}
+\subsection{State}
+\label{state}

### 2022-09-18 soundness.tex
shaping the structure more and more

\section{Soundness}
\label{sec:soundness}

Our ... is an instance of Iris~\cite{iris}. This means that our logic ... is set up
via the following steps:

\begin{enumerate}
\item define a certain amount of \emph{ghost state}
      that is needed in the next two steps;
\item define a central \emph{invariant}
      that ties the machine's physical state
      to the ghost state;
\item
  define the \emph{assertions} that an end user of our ... exploits, namely
  points-to assertions, deallocation witnesses, space credits, and pointed-by
  assertions; all of these assertions are defined as the ownership
  of a fragment of the ghost state.
\end{enumerate}

Once these steps have been carried out, Iris provides a semantic definition of
the Hoare triple $\triple\pre\instr\post$. We can then use Hoare triples to
state the reasoning rules of ...., to prove that
these rules are valid, and to state and prove the fact that ... is sound.
This is the topic of the next subsection (\sref{def:soundness:statement}).
%
After presenting these results, in \sref{sec:invariant}, we come back to the
three steps listed above and describe them at an informal level, without
diving into the concepts of Iris, which would be difficult to explain in a
thorough and concise way.
% For more information on Iris, the reader is referred to Jung \etal.'s
% paper~(\citeyear{iris}).

\subsection{The Soundness Statement}
\label{def:soundness:statement}

The operational semantics of ...lang guarantees that ... (\sref{sec:valid}). A~program that attempts to ... stuck~(\sref{sec:spacelang:instructions}).
% Thus, the space consumption of every \spacelang program is bounded.
Thus, in order to prove that ... is sound, all we have to do is prove that
it rules out the possibility of a runtime failure. This is guaranteed by the
following theorem:

\begin{theorem}[Soundness of the Logic ...]
\label{th:adequacy}
 The execution of the instruction~$\instrs$, beginning in an initial state, cannot result in a configuration where a thread is stuck.
\end{theorem}

This theorem states that if the program~$\instr$ satisfies a semantic Hoare
triple, then this program cannot crash. In
particular, it cannot reach a situation where DOING X ....
%
The proof of this theorem is short: it is a direct consequence of Iris's
adequacy theorem~\cite[\S6.4]{iris}.

Separately, we prove that the reasoning rules of~our logic are valid, that is,
each rule is a valid lemma:

\begin{theorem}[Validity of the Reasoning Rules]
\label{th:validity}
  Each of the rules in Figures~\ref{fig:laws}
  and~\ref{fig:reasoning} is valid.
\end{theorem}

The proof of this theorem represents more work: one lemma per reasoning rule
is required, not to mention a large number of auxiliary lemmas. The proofs of
these lemmas are fairly uninteresting; the key insights lie in the definitions
presented in the next subsection (\sref{sec:invariant}).

Together, Theorems~\ref{th:adequacy} and~\ref{th:validity} guarantee that, if
the Hoare triple $\iTrue\instrs\iTrue$ can be obtained by applying
the reasoning rules of our logic, then the program~$\instrs$ is safe, that is, it cannot crash ....

\subsection{Key Definitions and Invariants}
\label{sec:invariant}

For the definitions that follow, we need some more terminology about stores.

\begin{definition}[More Terminology]
\label{def:closed}

.............
Next, we must define the relation that ties the physical store to the
\logical store, two concepts that we have informally described earlier
(\sref{sec:logical}). From here on, we write~$\store$ for the physical
store and~$\logicalstore$ for the \logical store.

\begin{definition}[Relation between Physical and \Logical Stores]
\label{def:related}
  Two stores $\store$ and $\logicalstore$ are \emph{related},
  which we write $\related\store\logicalstore$, when
  they have the same domain,
  the same reachable locations,
  and the same content at every reachable location.
\end{definition}

Equipped with the above definitions, we are now prepared to enter the realm of
Iris and perform the three steps described at the beginning of
\sref{sec:soundness}, that is, define the ghost state, the central invariant,
and the assertions of our .. . We give the definitions first and explain them
afterwards.

\begin{comment}
% fp: this is the beginning of an explanation of ghost state in Iris
%     but it quickly grows too long, even though many aspects are missing.
A ghost memory cell in Iris is a memory cell that does not exist at runtime.
The ownership and the content of a physical memory cell are described via a
points-to predicate such as $\loc\pointsto\val$. Similarly, the ownership and
content of a ghost cell are described by an Iris assertion, usually written
$\ownGhost\gamma{a}$, where $\gamma$ is the name of the ghost cell and $a$ is
its content. Whereas the content of a physical memory cell is a
programming-language value~$\val$, the content of a ghost cell inhabits a
mathematical structure, a \emph{camera}~\cite[\S4.4]{iris}, which the user
chooses when the ghost cell is created. A camera is equipped with a
composition operation~$\cdot$ and the ghost state assertion satisfies the law
$\ownGhost\gamma{a\cdot b} \equiv \ownGhost\gamma{a} \star
\ownGhost\gamma{b}$. Thus, the choice of a suitable camera determines in what
ways the ghost state can be split.

A key Iris idiom involves the use of the \emph{authoritative
  camera}~\cite[\S6.3.3]{iris} together with an Iris invariant.
The elements of the authoritative camera include
\emph{authoritative elements} $\authfull{c}$ and
\emph{fragmentary elements} $\authfrag{b}$.
The composition law is defined in such a way that
there always exists at most one authoritative element
and the composition of all fragmentary elements in existence is
contained in the authoritative element, etc.
\end{comment}

\begin{assumption}
\label{assumption}
  Iris defines a certain piece of ghost state,
  defines a predicate $\genheapinterp\store$
  that ties a store~$\store$ to this ghost state,
  and defines the points-to assertion $\loc\fpointsto\qv\blk$
  in terms of this ghost state.
  This is visible in the paper~\cite[\S6.3.2]{iris}
  and in Iris's \texttt{gen\_heap} library~\cite{genheap}.
  %
  We re-use this machinery without change,
  so we do not repeat these definitions.
  We mention the predicate $\genheapinterp\!$
  in our own invariant (Definition~\ref{def:invariant}),
  where it is applied to the \logical store~$\logicalstore$.
\end{assumption}

\begin{definition}[Ghost State]
...
stores an element of the monoid
\newcommand\fpfn{\rightarrow_{\textrm{fin}}}
\(
  \authm(\;
    \Loc \;\fpfn\;
    (\textsc{Frac}, \mathord{+})
    \times
    (\textsc{Multiset}(\Loc), \mathord\uplus)
  \;)
\)
% \emph{authoritative camera}
\cite[\S6.3.3]{iris}.
\end{definition}

\begin{definition}[Central Invariant]
\label{def:invariant}
The central invariant of our logic,
or \emph{state interpretation invariant} \cite[\S7.3]{iris},
is the following Iris assertion:
\[
\centralinvariant\;\store \triangleq
\exists\logicalstore. \left\{
\def\arraystretch{1.2}
\begin{array}{l@{\quad\star\quad}l@{\quad}l}
  \genheapinterp\logicalstore & \star \\
   \genheapinterp\logicalstore & \star
\end{array}
\right.
\]
\end{definition}

\begin{definition}[Assertions]
\label{def:assertions}
  The assertions of our ... are defined as follows:
  \begin{enumerate}
  \item
  % The definition of the points-to predicate is inherited from gen_heap.
  % It is implicitly parameterized by a ghost cell that holds the heap.
  The \emph{points-to assertion} ..
  is inherited from Iris (Assumption~\ref{assumption}).
  \item
  The \emph{map-of-map} for a single physical memory location,

  \item
  The \emph{location-virtualization-modal}
  \item
  The \emph{pointed-by assertion} .
  \end{enumerate}
\end{definition}

% What do these definitions mean?

Let us try and explain the most important aspects of these definitions.

In Iris, the state interpretation invariant~$\centralinvariant\;\store$ is an assertion that
holds of the \emph{physical store}~$\store$ in between every two steps of
computation. It is used by Iris in the definition of Hoare triples; we inherit
this definition from Iris.

Our definition of $\centralinvariant\;\store$ begins with an existential quantification over
a \emph{\logical store}~$\logicalstore$.
..........
%.......

### 2022-09-18 semantics.tex
separted stuff
+

@@
 \label{sec:semantics}
-\mytodo{Ismail's Part}
+\subsection{Instructions}
+\label{sec:instrsemantics}
+The syntax of instructions is given in \fref{fig:syntax}. Their
+small-step operational semantics, a~reduction relation of the form
+$\hsfork{\instr}{\store}{\instr'}{\store'}{\instrs}$,
+is defined in \fref{fig:semantics}.
+%
+The last argument~$\instrs$ is a list of newly-spawned threads;
+when it is empty, we omit it
+and write just $\hs{\instr}{\store}{\instr'}{\store'}$.
+%
+% Also, when it is a singleton, we omit the square brackets (StepFork).
+
+% We briefly review the syntax and reduction rules.
+%
+The no-op ``$\iskip$'' and the sequencing construct
+``$\iseq{\instr_1}{\instr_2}$'' are standard.
+%
+
+\input{fig-semantics}

### 2022-09-18 soundness.tex
+

@@
   define the \emph{assertions} that an end user of our ... exploits, namely
-  points-to assertions, deallocation witnesses, space credits, and pointed-by
-  assertions; all of these assertions are defined as the ownership
+  points-to assertions that are defined as the ownership
   of a fragment of the ghost state.
@@
 
-The operational semantics of ...lang guarantees that ... (\sref{sec:valid}). A~program that attempts to ... stuck~(\sref{sec:spacelang:instructions}).
+The operational semantics of ...lang guarantees that ... (\sref{sec:valid}). A~program that attempts to ... stuck~(\sref{sec:instrsemantics}).
 % Thus, the space consumption of every \spacelang program is bounded.
@@
 
-\begin{definition}[More Terminology]
-\label{def:closed}
-
-.............
 Next, we must define the relation that ties the physical store to the
@@
 store and~$\logicalstore$ for the \logical store.
-
-\begin{definition}[Relation between Physical and \Logical Stores]
-\label{def:related}
-  Two stores $\store$ and $\logicalstore$ are \emph{related},
-  which we write $\related\store\logicalstore$, when
-  they have the same domain,
-  the same reachable locations,
-  and the same content at every reachable location.
-\end{definition}
 
@@
 \end{comment}
+% The predicate gen_heap_interp.
+\newcommand{\genheapinterp}[1]{\mathit{Heap}\;#1}
+% Our predicate pred (defined in ph.v), expanded.
+\newcommand{\pred}[1]{\ownGhost\gammaPred{\authfull{(\mapone\predstore)}}}
+% A notation for assigning fraction 1 to every element of \predstore.
+\newcommand{\mapone}[1]{1.#1}
+% The predicate mapsfrom_exact, expanded.
+\newcommand{\mapsfromexact}[3]{
+  \ownGhost\gammaPred{\authfrag{\singletonMap{#1}{(#2, #3)}}}
+}
+% A metavariable for a share.
+\newcommand{\sh}{L'}
+% The predicate mapsfrom, expanded.
+\newcommand{\mapsfromdef}[3]{
+  \exists\sh.\;
+  \mapsfromexact{#1}{#2}{\sh} \star \pure{\sh \subseteq #3}
+}
 
@@
 \centralinvariant\;\store \triangleq
-\exists\logicalstore. \left\{
+\left\{
 \def\arraystretch{1.2}
 \begin{array}{l@{\quad\star\quad}l@{\quad}l}
-  \genheapinterp\logicalstore & \star \\
-   \genheapinterp\logicalstore & \star
+  \genheapinterp\store & \star \\
+   \genheapinterp\store & \star
 \end{array}

### 2022-09-18 soundness.tex
moreless the structure is fine: now fill in the gaps

@@
   in our own invariant (Definition~\ref{def:invariant}),
-  where it is applied to the \logical store~$\logicalstore$.
+  where it is applied to the \logical store~$\store$.
 \end{assumption}

### 2022-09-18 semantics.tex
+

@@
 \label{sec:semantics}
+\subsection{State}
+\label{sec:state}
+We represent the machine state mainly as a finite map of registers to register values and a map of masked physical memory addresses to 64-bit physical memory values.
+Together with default def\textsf{CPU} instantiation, we end up having our state with following pieces:
+\begin{itemize}
+\item A constant cpu initialized with default value: $\sigma.\mathcal{C}$
+\item A register map: $\sigma.\mathcal{R}: \kw{greg} \rightarrow_{\textrm{fin}} \kw{regval} $
+\item A memory map: $\sigma.\mathcal{M}: \Loc \rightharpoonup_{\textrm{fin}} (\rightharpoonup_{\textrm{fin}} )$
+\end{itemize}
+
+\subsection{Indexed Memory \& Register References}
+\label{sec:}
+As one might have already anticipated from the syntax we introduce, we do not bind any value of an evaluated expression.
+All the indices and accessed values are treated as globally referenced.
+In align with this design choice, our expression is a stream of instructions, which is not evaluated to a value to be bound, but changes the machine state through the indices (e.g.
+$\kw{r}\in\kw{greg}$) -- to the global maps.
 \subsection{Instructions}
@@
 %
-The no-op ``$\iskip$'' and the sequencing construct
-``$\iseq{\instr_1}{\instr_2}$'' are standard.
-%
 

### 2022-09-18 syntax.tex
+

@@
 \subsection{Registers, Memory Indices}
-There exist two pieces of resources we access in our model: registers and memory.
+Programs we demonstrate in this paper requires accessing two types of computer resource: registers and memory.
 A register identifier, $\reg$, is chosen from a fixed finite set of register identifiers,$\regset$.
@@
 Intentionally, we restrict any stream of instruction, $\instrs$, to eventually reduce to $\iskip$ instruction which is injected into unit.
+Our syntax capture the no-op ``$\iskip$'' and the sequencing construct ``$\iseq{\instr}{\instrs}$'' in their standard forms.
 \input{fig-syntax}

### 2022-09-18 semantics.tex
selected ins for ops
+

@@
 \item A register map: $\sigma.\mathcal{R}: \kw{greg} \rightarrow_{\textrm{fin}} \kw{regval} $
-\item A memory map: $\sigma.\mathcal{M}: \Loc \rightharpoonup_{\textrm{fin}} (\rightharpoonup_{\textrm{fin}} )$
+\item A memory map: $\sigma.\mathcal{M}: \Locft \rightharpoonup_{\textrm{fin}} (\Loctw \rightharpoonup_{\textrm{fin}} \Locsf )$
 \end{itemize}
-
-\subsection{Indexed Memory \& Register References}
+\subsection{Indexed Memory \& Register by References}
 \label{sec:}
@@
 and write just $\hs{\instr}{\store}{\instr'}{\store'}$.
-%
-% Also, when it is a singleton, we omit the square brackets (StepFork).
-
-% We briefly review the syntax and reduction rules.
-%
+\subsection{A Set of Selected Instructions Changing the Machine State}
+\label{sec:selectedinstrsemantics}
+In addition to the reduction rules in Figure \fref{fig:reductionsemantics} for the light-weight integration of our machine model to our simple language, and, finally, to our reasoning principles, we think that it is worth building the intuition on the way some of the instructions -- specifically the ones used in the examples contained in this paper -- change the machine state when reduced with \RULE{StepContext}.
+\RULE{StepMovRR} in \fref{fig:instrsemantics} shows how machine state change when we copy the source register's value ($\rvsrc$) referenced by the register name ($\rgsrc$) to the destination register ($\rgdst$).
 

### 2022-09-18 semantics.tex
+

@@
 \RULE{StepMovRR} in \fref{fig:instrsemantics} shows how machine state change when we copy the source register's value ($\rvsrc$) referenced by the register name ($\rgsrc$) to the destination register ($\rgdst$).
+We follow the convention of showing only the pieces of the machine state changed after the instruction execution -- only the register value of the destiantion register ($ \readlval\rgdst\storeregprime\rvsrc$) has changed in the state (\sigmaprime) after the execution of the instruction $\textsf{mvrr}$.
 

### 2022-09-18 semantics.tex
+

--- 2022-09-18 semantics.tex
@@
 \RULE{StepMovRR} in \fref{fig:instrsemantics} shows how machine state change when we copy the source register's value ($\rvsrc$) referenced by the register name ($\rgsrc$) to the destination register ($\rgdst$).
-
+We follow the convention of showing only the pieces of the machine state changed after the instruction execution -- only the register value of the destiantion register ($ \readlval\rgdst\storeregprime\rvsrc$) has changed in the state ($\storeprime$) after the execution of the instruction $\textsf{mvrr}$.
+All the other rules in Figure \fref{fig:instrsemantics}, presents the change in the state when a value is loaded/stored to/from register/memory, and are differentiated with respect to the addressing mode ($amode$) which determines whether the accessed memory address is computed based on some offset (\RULE{StepMovRMOff}, \RULE{StepMovMROff}) or not (\RULE{StepMovRMBase}) and \RULE{StepMovMRBase}).
+The most common essential aspect of loading the source register's value ($\rvsrc$) to the memory, \RULE{StepMovRMBase} and \RULE{StepMovRMOff}, is the hierarchical translation of memory location, $\readlval{\plusaddr\maddr\offs}{\storememstar\crval}\locsf$, which is hidden in the rules, expanded as the following:
+\begin{itemize}
+\item
+\item
+\item
+\item
+\end{itemize}
+where the first three levels of translation allows address-translation .
 \input{fig-semantics}

### 2022-09-18 semantics.tex
checkpoint

@@
 All the other rules in Figure \fref{fig:instrsemantics}, presents the change in the state when a value is loaded/stored to/from register/memory, and are differentiated with respect to the addressing mode ($amode$) which determines whether the accessed memory address is computed based on some offset (\RULE{StepMovRMOff}, \RULE{StepMovMROff}) or not (\RULE{StepMovRMBase}) and \RULE{StepMovMRBase}).
-The most common essential aspect of loading the source register's value ($\rvsrc$) to the memory, \RULE{StepMovRMBase} and \RULE{StepMovRMOff}, is the hierarchical translation of memory location, $\readlval{\plusaddr\maddr\offs}{\storememstar\crval}\locsf$, which is hidden in the rules, expanded as the following:
+The most common essential aspect of loading the source register's value ($\rvsrc$) to the memory, \RULE{StepMovRMBase} and \RULE{StepMovRMOff}, is the hierarchical translation of memory location, $\readlval\maddr{\storememstar\crval}\locsf$, which is hidden in the rules, expanded as the following:
 \begin{itemize}
-\item
-\item
-\item
-\item
+\item top-level-address translation: with the given root address (64-bit $\crval$) of the address space, performs the address translation, handling the first level (to get the PML4 entry) itself.
+The next level table address is computed with the fetched PML4 offset value which exhibits itself as 9-bit offset in $\kw{maddr}$.
+\item translating from PML4 entry: performs the second level of address translation, to retrieve starting at the PML4 table entry, and interprets the PML4 entry that references a Page-Directory-Pointer Table (PDPT).
+We obtain the PDPT offset which exhibits itself as 9 bit offset in $\kw{maddr}$ to obtain the address of the next level page directory table (PDT)
+\item translating from PD entry: performs the third level of address translation, to retrieve starting at the PDP table entry, and interprets a PDPTE that references a PD table.
+Likewise, we obtain the PD offset which exhibits itself as 9 bit offset in $\kw{maddr}$ to obtain the address of the next level page directory table (PT)
+\item translate from PT entry: performs final level of address translation, starting from the PT entry, with a given 12 bit page offset, we can compute the physical address referencing $\locsf$
 \end{itemize}
-where the first three levels of translation allows address-translation .
+In fact $\readlval\maddr{\storememstar\crval}\locsf$ is unfoled into multiple physical memory lookups for the final page address retrieval -- i.e. address translation traversal. 
 \input{fig-semantics}

### 2022-09-19 programlogic.tex
+
start explaining logic

@@
-\section{Program Logic}
-\label{sec:programlogic}
-\mytodo{Ismail's Part}
+\section{Program Logic for Location Virtualization}
+\label{sec:logic}
+In this section, we describe our high-level reasoning principles for location virtualization whose detailed justification is shown in \sref{sec:soundness}. 
+\subsection{Points-To Assertions}
+\label{sec:pointsto}
+
+We present three main points-to relations as part of utilization of separation logic \ref{} as an ambient logic:
+\begin{enumerate}
+\item Physical address points-to, $\pfpointsto\locsf\locsf\ppts\qfrac$
+\item Register points-to, $\pfpointsto\rg\rv\rpts\qfrac$
+\item Virtual memory address points-to, $\pfpointsto\vaddr\locsf\vpts\qfrac$
+\end{enumerate}
+\paragraph{Register points-to} The assertion $\pfpointsto\rg\rv\rpts\qfrac$ ensures the ownership of the register $\rg$ naming the register value $\rv$.
+The fraction $\qfrac$ with value 1 asserts the unique ownership of the register mapping, and grants update permission on it, otherwise, any value $0 < \qfrac <1$ represents partial ownership granting readonly permission on the mapping.
+\paragraph{Physical-Memory points-to} The our physical memory points-to relation ($\pfpointsto\locsf\locsf\ppts\qfrac$) exhibits nested-mappings due to masking applied on the indexing address ($\locsf$).
+The unfolded definition of the physical mapsto
+\[
+\begin{array}{cl}
+\pfpointsto\locsf\locsf\ppts\qfrac \stackrel{def}{=} & \nfpointsto{\mask\locsf\tw}{\mask\locsf\ft}\locsf\qfrac\naddr
+\end{array}
+\]
+where we simply abstract the generation of hierarchical mapping of a physical address ($\locsf$) through the different masks of the address ($\locsf|^{12}$ and $\locsf|^{52}$) as nested map resources.
+\paragraph{Virtual-Memory points-to}
+\subsection{Modal Abstraction for Virtual Memory}
+\label{sec:credits}
+\input{fig-laws}
+
+\subsection{Memory Deallocation as a Ghost Operation}
+\label{sec:free}
+
+\subsection{Reasoning Rules for Selected Instructions}
+\label{sec:reasoning}
+\input{fig-reasoning}

### 2022-09-19 programlogic.tex
+

@@
 where we simply abstract the generation of hierarchical mapping of a physical address ($\locsf$) through the different masks of the address ($\locsf|^{12}$ and $\locsf|^{52}$) as nested map resources.
-\paragraph{Virtual-Memory points-to}
-\subsection{Modal Abstraction for Virtual Memory}
+\paragraph{Virtual-Memory points-to} As we see in Figure \fref{fig:}, virtual address mapping is obtained via traversal of page table, i.e.
+indirection provided by physical memory lookups.
+As expected, we define the virtual-points to relation, ($\pfpointsto\vaddr\locsf\vpts\qfrac$), in terms of multiple physical memory mappings representing the indirection shown in Figure \fref{fig:}
+\[
+\begin{array}{cl}
+\pfpointsto\locsf\locsf\ppts\qfrac \stackrel{def}{=} & \nfpointsto{\mask\locsf\tw}{\mask\locsf\ft}\locsf\qfrac\naddr
+\end{array}
+\]
+\subsection{Modal Abstraction for Virtualizing Memory Locations}
 \label{sec:credits}

### 2022-09-19 background.tex

@@
 \label{sec:backgroundonmachinemodel}
-\mytodo{Colin's part}
+
+In typical system configurations, all memory addresses seen by programs running on modern computers are \emph{virtualized}: the address observed by a running program generally will not correspond directly to the physical location in memory, and may not even correspond to a physical location that \emph{exists} in the machine.
+Instead, these \emph{virtual} addresses are translated to \emph{physical} addresses that correspond directly to locations in RAM.
+On most modern architectures, this translation is performed through cooperation of the hardware and OS kernel: while executing an instruction that dereferences a (virtual) address, the CPU's \emph{memory management unit} (\textsc{MMU}) implements an architecture-specific process of \emph{address translation}, resulting in a physical address used to access the cache\footnote{Technically, for performance reasons most caches are indexed with parts of the virtual address, but tagged with the physical data addresses, so cache lookups and address translations can proceed in parallel.} and/or memory-bus.
+
+\begin{figure}
+    \includegraphics[width=\columnwidth]{pagetables.png}
+    \caption{x86-64 page table lookups.}
+    \label{fig:pagetables}
+\end{figure}
+
+On the x86-64 architecture, the input to the \textsc{MMU}'s address translation is a sparse hierarchical set of tables calle d \emph{page tables} (referring to pages of memory).
+As Figure \ref{fig:pagetables}\mytodo{cite AMD ref} shows, address translation proceeds by repeatedly taking designated slices of the virtual address and indexing into a table.
+On x86-64, standard configurations use 4 levels of page tables, labelled levels 4 through 1, with lookups in the level 1 page table resulting in the actual page of physical memory holding the requested data, and the low-order 12 bits being used to index into this page.\footnote{Technically levels 1--3 have designated names originating when there were fewer levels.
+We omit those names to avoid confusion, and for greater consistency with the recent addition of a 5th level on the most recent processors, where levels 4 and 5 are simply known by their numbers.}  Because address translation traverses a series of tables, the translation process or algorithm is sometimes referred to as a \emph{page-table walk}.
+While Figure \ref{fig:pagetables} and most of our details are specific to the x86-64 architecture, ARMv8 (a.k.a.\ \texttt{aarch64}), RISC-V, and PowerPC use similar hierarchical page tables for address translation.
+
+The entries of each table are 64 bits wide, but each points to a physical address aligned to 4KB (4096 byte) boundaries, which leaves 12 bits to spare to control a validity bit (called the \emph{present} bit), a read-write bit (which permits write access through the entry if and only if it is set), and a range of additional bits which can be used to control caching, write-through, and more.
+This paper will only consider the present bit (0), the read-write bit (1), and the accessed bit (5), which is set to 1 by the hardware every time an entry is used, allowing the OS to determine which pieces of memory are used or unused.
+
+The page tables mapping virtual addresses to physical addresses are managed by the OS.
+Typically each process has its on virtul-to-physical mappings represented by its own page table, which the OS registers with the CPU by storing it in a specific register (\texttt{cr3}) as part of switching to a new process.
+Using different mappings, which map only disjoint portions of physical memory (with some exceptions in the next section) is how the OS ensures memory isolation between processes.
+
+If an instruction is executed that accesses a virtual address that either has no mapping, or does not have a mapping permitting the kind of access that was performed (e.g., the instruction was a memory write, but the relevant address range was marked read-only in the relevant page table entry), the hardware triggers a \emph{page fault}, transferring control to a \emph{page fault handler} registered with the hardware by the OS, allowing it to take corrective action (described in the next section).
+
 \subsection{Virtual Memory Managers}
-\mytodo{Colin's part}
 \label{sec:backgroundonvmm}
+Most OS kernels have a component called a \emph{Virtual Memory Manager} (VMM)\footnote{Not to be confused with Virtual Machine Monitor.
+In this paper we focus on non-virtualized scenarios, but hardware virtualization extensions for both x86-64 and ARM make use of an additional set of page tables translating what a \emph{guest} considers to be its (virutalized) physical memory to actual physical memory.
+Our contributions should offer value in this scenario as well.}
+which is responsible for setting up page table mappings and for taking action when a page fault occurs.
+Often, a page fault indicates a bug in a program --- for example, a null pointer dereference crashes a program not because the hardware designates it an error, but because most OSes refuse to map the first 4K worth of virtual addresses, so dereferencing \texttt{NULL} results in a page fault.
+In such cases, the OS will terminate the program.
+
+In other cases the OS uses page faults to implement specialized functionality and optimizations.
+A key example of this is (confusingly) called \emph{paging}: saving room in physical RAM by waiting until memory addresses for code are accessed before copying them to RAM (saving process startup time); or copying memory that has not recently been accessed to disk (i.e., in a \emph{swap} file or partition) and marking those page table entries invalid so a future page fault will give the OS the chance to copy the relevant data back into memory when the program tries to access it again.
+
+The key pieces of functionality a VMM must implement itself, or have available in another OS component (depending on design), are
+adding a new page mapping (whether the mapped page contains zeros, file data, or swap data), and removing an existing page mapping.
+While this initially sounds like relatively modest functionality whose implementation may be complicated by hardware subtleties, a number of more serious complexities must be handled.
+
+For one, virtual addresses may \emph{alias} when more than one is mapped to the same physical memory.
+This can occur when a process requests it (e.g., via \texttt{mmap} or \texttt{shm\_open} on UNIX-style kernels).
+It also occurs automatically in most general-purpose kernels: while for normal programs the OS uses the MMU to isolate and abstract physical memory, the kernel's own functionality is often easier to implement if the kernel can directly access memory given its physical address, for example when interacting with hardware devices.
+For this reason, most kernels contain an \emph{identity map} of all physical memory in a machine.
+In some cases this is a literal identity map, where for every virtual address $p$, interpreting $p$ as a virtual address instead will map back to $p$ as a physical address.
+In other cases there is an offset involved, where to access physical address $p$ the kernel can access virtual address $p+\mathit{offset}$.
+This identity map saves the kernel the performance cost of having to continually add and remove transient mappings simply to briefly access a specific physical memory region.
+But this coexists with additional virtual addresses (those used by most kernel code and data structures), in a form of virtual address aliasing.
+Note that virtual address aliasing is both prevalent in general-purpose kernels, and violates the core assumption of most memory models assumed by verified compilers like CompCert, which assume no virtual address aliasing.
+
+Another complicating factor is that portions of page tables can be shared, i.e., trees of page tables can overlap.
+One common way this used to happen was for one subtree of page tables to map the kernel if linked in appropriately to a higher-level page table, and then to reuse that kernel mapping portion in every process's page tables.
+This was mostly phased out in the wake of Spectre\mytodo{Colin finish}
+
+These dove-tail with functionality in the OS scheduler, which since it is responsible for dealing with multiple address spaces, must keep track of which virtual addresses are valid (and in what way) in which address spaces, where some virtual addresses are valid in only a single address space (e.g., a code address for a particular usermode process), while others are valid in all address spaces (e.g., kernel data structure pointers).
+The VMM must maintain some of these assumptions on behalf of the rest of the kernel, for example by guaranteeing that a certain range of virtual addresses (corresponding to the kernel's code and data) are valid in every address space.
+
 \subsection{Modal Logic}
 \label{sec:backgroundonmodallogic}
-\mytodo{Colin's part}
+The problem of needing to keep track of things being true in some contexts and not in others is hardly unique to virtual memory management, and is the general insight behind most flavors of modal logic.
+\mytodo{Colin finish}

### 2022-09-19 intro.tex
some intro stuff, bibliography

@@
 \mytodo{Colin's Part}
+Virtual memory management lies at the core of modern OS kernel implementation.
+It is deeply intertwined with most other parts of a typical general-purpose OS kernel design, including scheduling, hardware drivers, and even the filesystem buffer cache.
+In writing the authoritative reference on the internals of the Solaris kernel, McDougal and Mauro went so far as to claim that ``\emph{the virtual memory sub-system can be considered the core of a Solaris instance, and the implementation of Solaris virtual memory affects just about every other subsystem in the operating system}''~\cite{mcdougal2006solaris}.
+This makes rich support for verification the virtual memory management subsystem of an OS kernel critical to the correctness of every other piece of an OS or any software running atop it.
+
+At its core, the virtual memory functionality of modern CPUs is about \emph{location virtualization}: the memory locations (addresses) seen by most code are not, in fact, the exact location in physical memory where data resides.
+Instead these are \emph{virtual} addresses, which are mapped to actual physical resources by the cooperation of the hardware and OS.
+This is what enables separation of process memory resources:
+the OS manipulates hardware functionality to ensure that any attempt by a process to access memory not explicitly granted to it by the kernel will fail.
+But this is complicated by the fact that the OS and hardware can also enable, shared (overlapping) access to physical memory regions; the fact that the kernel data structures themselves are accessed via virtual memory addresses; and the fact that control over these mappings of virtual to physical addresses is itself mediated by \emph{in-memory data structures}, which the kernel still accesses via virtual address, leading to indirect cycles.
+
+Further complicating matters is that addresses themselves bear no information about which address space they originate from.
+For user processes this is of little concern, as these have access to only their own address space.
+But the kernel has access to all address spaces.
+Mixing up addresses from different address spaces leads to severe bugs.
+More concerning, keeping track of which \emph{assertions} hold in different address spaces during kernel verification is difficult: some assertions should hold across all address spaces, while others hold in only one, and others may hold in multiple but still not all.
+
+This kind of context-dependent assertion, where a fact may be true in one address space but not others, has a modal flavor.
+We propose tackling the verification of virtual memory subsystems (and kernels more broadly) by adapting ideas from hybrid modal logic, which can label assertions true under \emph{other} circumstances (i.e., in another address space) with a modality indexed by a name for that space (in our case, the root of the page tables for an address space).
+This offers a convenient and powerful way to modularly (1) isolate assertions specific to a particular address space, and (2) explicitly state when an assertion is true across address spaces.
+
+We develop these ideas in an Iris-based\mytodo{cite JFP} separation logic, and verify simplified versions of several critical virtual-memory-related pieces of OS functionality, including mapping and unmapping pages, and describing the self-referential identity maps used in most kernels to allow convenient access to physical memory.\mytodo{colin finish}

### 2022-09-19 programlogic.tex

@@
 The unfolded definition of the physical mapsto
+\begin{figure}[!ht]
 \[
@@
 \]
+\caption{Physical Points-to with Nested Masking}
+  \label{fig:physicalpointsto}
+\end{figure}
 where we simply abstract the generation of hierarchical mapping of a physical address ($\locsf$) through the different masks of the address ($\locsf|^{12}$ and $\locsf|^{52}$) as nested map resources.
@@
 As expected, we define the virtual-points to relation, ($\pfpointsto\vaddr\locsf\vpts\qfrac$), in terms of multiple physical memory mappings representing the indirection shown in Figure \fref{fig:}
+\begin{figure*}
 \[
 \begin{array}{cl}
-\pfpointsto\locsf\locsf\ppts\qfrac \stackrel{def}{=} & \nfpointsto{\mask\locsf\tw}{\mask\locsf\ft}\locsf\qfrac\naddr
+  \pfpointsto\locsf\locsf\ppts\qfrac \stackrel{def}{=} \lambda \tlf\Loc.
+  & \exists_{ \tlfoff\Locn \;, \tltoff\Locn \;, \tltwoff\Locn \;, \tlooff\Locn \;, \tpgoff\Loctw \;, \tlt\Loc \;, \tltw\Loc \;, \tlo\Loc \tpg\Loc   }
 \end{array}
 \]
+\caption{Virtual Points-to Relation}
+  \label{fig:virtualpointsto}
+\end{figure*}
+
+\begin{comment}
+   Definition virt_mapsto (a:word 64) (v:word 64) (q:Qp) : word 64 -> iProp Σ :=
+λ (cr:word 64) ,
+     ( ⌜ aligned a ⌝ ∗
+         ∃ (l3 l2 l1 page w :word 64) (pageoff:word 12) (l4off l3off l2off l1off:word 9) ,
+           ⌜ (WordImpl.concat (wzero 16) (WordImpl.concat l4off (WordImpl.concat l3off (WordImpl.concat l2off (WordImpl.concat l1off pageoff))))) = a ⌝ ∗
+                          (natToWord 64 ((wordToNat l4)+(8*(wordToNat l4off)))) ↦p{q} (l3  ^|  (natToWord 64 3)) ∗
+                          (natToWord 64 ((wordToNat l3)+(8*(wordToNat l3off)))) ↦p{q} (l2  ^|  (natToWord 64 3)) ∗
+                          (natToWord 64 ((wordToNat l2)+(8*(wordToNat l2off)))) ↦p{q} ( l1  ^|  (natToWord 64 3)) ∗
+                          (natToWord 64 ((wordToNat l2)+(8*(wordToNat l1off)))) ↦p{q} ( page  ^|  (natToWord 64 3)) ∗
+                          (natToWord 64 ((wordToNat page) +(wordToNat pageoff))) ↦p v )%I.
+\end{comment}
+
 \subsection{Modal Abstraction for Virtualizing Memory Locations}
-\label{sec:credits}
+\label{sec:modallocationvirtualization}
 \input{fig-laws}
-
-\subsection{Memory Deallocation as a Ghost Operation}
-\label{sec:free}
 

### 2022-09-19 intro.tex
citations work better when spelled correctly

@@
 It is deeply intertwined with most other parts of a typical general-purpose OS kernel design, including scheduling, hardware drivers, and even the filesystem buffer cache.
-In writing the authoritative reference on the internals of the Solaris kernel, McDougal and Mauro went so far as to claim that ``\emph{the virtual memory sub-system can be considered the core of a Solaris instance, and the implementation of Solaris virtual memory affects just about every other subsystem in the operating system}''~\cite{mcdougal2006solaris}.
+In writing the authoritative reference on the internals of the Solaris kernel, McDougall and Mauro went so far as to claim that ``\emph{the virtual memory sub-system can be considered the core of a Solaris instance, and the implementation of Solaris virtual memory affects just about every other subsystem in the operating system}''~\cite{mcdougall2006solaris}.
 This makes rich support for verification the virtual memory management subsystem of an OS kernel critical to the correctness of every other piece of an OS or any software running atop it.

### 2022-09-19 programlogic.tex
to fix vptsto

--- 2022-09-19 programlogic.tex
@@
 As expected, we define the virtual-points to relation, ($\pfpointsto\vaddr\locsf\vpts\qfrac$), in terms of multiple physical memory mappings representing the indirection shown in Figure \fref{fig:}
+\begin{figure*}
 \[
 \begin{array}{cl}
-\pfpointsto\locsf\locsf\ppts\qfrac \stackrel{def}{=} & \nfpointsto{\mask\locsf\tw}{\mask\locsf\ft}\locsf\qfrac\naddr
+  \ppointsto\locsf\locsf'\vpts \stackrel{def}{=} \lambda \tlf\Loc.
+  & \exists_{\tlfoff\Locn \;, \tltoff\Locn \;, \tltwoff\Locn \;, \tlooff\Locn \;, \tpgoff\Loctw \;, \tlt\Loc \;, \tltw\Loc \;, \tlo\Loc \tpg\Loc}\\
+  & \textsf{aligned } \locsf \\
+  & \locsf = \locsx ++ \lfoff ++ \ltoff ++ \ltwoff ++ \looff ++ \pgoff \\
+  &  \pfpointsto{\lvlsum\crt\lfoff}{\lvlbor\lt}\ppts\qfrac \\
+  &  \pfpointsto{\lvlsum\lt\ltoff}{\lvlbor\ltw}\ppts\qfrac \\
+  & \pfpointsto{\lvlsum\ltw\ltwoff}{\lvlbor\lo}\ppts\qfrac \\
+  & \pfpointsto{\lvlsum\lo\looff}{\lvlbor\pg}\ppts\qfrac \\
+  & \ppointsto{\pageptstosum\pg\pgoff}\locsf'\ppts \\
+
 \end{array}
 \]
+\caption{Virtual Points-to Relation}
+  \label{fig:virtualpointsto}
+\end{figure*}
+
 \subsection{Modal Abstraction for Virtualizing Memory Locations}
-\label{sec:credits}
+\label{sec:modallocationvirtualization}
 \input{fig-laws}
-
-\subsection{Memory Deallocation as a Ghost Operation}
-\label{sec:free}
 
@@
 \input{fig-reasoning}
+As is usual in \SL, the reasoning rules of the program logic allow deriving
+Hoare triples of the form $\triple\pre\instr\post$, where the
+precondition~$\pre$ and the postcondition~$\post$ are assertions,
+and where $i$ is an instruction.
+Such a triple receives a standard interpretation as a statement of partial
+correctness: the assertion $\triple\pre\instr\post$ means that, in a state
+where $\pre$ holds, it is safe to execute the instruction~$\instr$, and if
+this execution terminates, then it leads to a final state where $\post$ holds.
+A more precise statement is given in the next section (\sref{sec:soundness}).
+
+\mytodo{iso todo: talk about each rule}

### 2022-09-19 programlogic.tex
+

@@
 \begin{enumerate}
-\item Physical address points-to, $\pfpointsto\locsf\locsf\ppts\qfrac$
-\item Register points-to, $\pfpointsto\rg\rv\rpts\qfrac$
-\item Virtual memory address points-to, $\pfpointsto\vaddr\locsf\vpts\qfrac$
+\item Physical address points-to, $\pfpointsto\locsf\locsf\qfrac\ppts$
+\item Register points-to, $\pfpointsto\rg\rv\qfrac\rpts$
+\item Virtual memory address points-to, $\pfpointsto\vaddr\locsf\qfrac\vpts$
 \end{enumerate}
@@
 The fraction $\qfrac$ with value 1 asserts the unique ownership of the register mapping, and grants update permission on it, otherwise, any value $0 < \qfrac <1$ represents partial ownership granting readonly permission on the mapping.
-\paragraph{Physical-Memory points-to} The our physical memory points-to relation ($\pfpointsto\locsf\locsf\ppts\qfrac$) exhibits nested-mappings due to masking applied on the indexing address ($\locsf$).
+\paragraph{Physical-Memory points-to} The our physical memory points-to relation ($\pfpointsto\locsf\locsf\qfrac\ppts$) exhibits nested-mappings due to masking applied on the indexing address ($\locsf$).
 The unfolded definition of the physical mapsto
@@
 \begin{array}{cl}
-\pfpointsto\locsf\locsf\ppts\qfrac \stackrel{def}{=} & \nfpointsto{\mask\locsf\tw}{\mask\locsf\ft}\locsf\qfrac\naddr
+\pfpointsto\locsf\locsf\qfrac\ppts \stackrel{def}{=} & \nfpointsto{\mask\locsf\tw}{\mask\locsf\ft}\locsf\qfrac\naddr
 \end{array}
@@
 indirection provided by physical memory lookups.
-As expected, we define the virtual-points to relation, ($\pfpointsto\vaddr\locsf\vpts\qfrac$), in terms of multiple physical memory mappings representing the indirection shown in Figure \fref{fig:}
+As expected, we define the virtual-points to relation, ($\pfpointsto\vaddr\locsf\qfrac\vpts$), in terms of multiple physical memory mappings representing the indirection shown in Figure \fref{fig:}
 \begin{figure*}
@@
   & \locsf = \locsx ++ \lfoff ++ \ltoff ++ \ltwoff ++ \looff ++ \pgoff \\
-  &  \pfpointsto{\lvlsum\crt\lfoff}{\lvlbor\lt}\ppts\qfrac \\
-  &  \pfpointsto{\lvlsum\lt\ltoff}{\lvlbor\ltw}\ppts\qfrac \\
-  & \pfpointsto{\lvlsum\ltw\ltwoff}{\lvlbor\lo}\ppts\qfrac \\
-  & \pfpointsto{\lvlsum\lo\looff}{\lvlbor\pg}\ppts\qfrac \\
+  &  \pfpointsto{\lvlsum\crt\lfoff}{\lvlbor\lt}\qfrac\ppts \\
+  &  \pfpointsto{\lvlsum\lt\ltoff}{\lvlbor\ltw}\qfrac\ppts \\
+  & \pfpointsto{\lvlsum\ltw\ltwoff}{\lvlbor\lo}\qfrac\ppts \\
+  & \pfpointsto{\lvlsum\lo\looff}{\lvlbor\pg}\qfrac\ppts \\
   & \ppointsto{\pageptstosum\pg\pgoff}\locsf'\ppts \\

### 2022-09-19 programlogic.tex
checkpoint before makingg wps precise
+

@@
 \begin{array}{cl}
-  \ppointsto\locsf\locsf'\vpts \stackrel{def}{=} \lambda \tlf\Loc.
+  \ppointsto\vaddr\locsf\vpts \stackrel{def}{=} \lambda \tlf\Loc.
   & \exists_{\tlfoff\Locn \;, \tltoff\Locn \;, \tltwoff\Locn \;, \tlooff\Locn \;, \tpgoff\Loctw \;, \tlt\Loc \;, \tltw\Loc \;, \tlo\Loc \tpg\Loc}\\
-  & \textsf{aligned } \locsf \\
-  & \locsf = \locsx ++ \lfoff ++ \ltoff ++ \ltwoff ++ \looff ++ \pgoff \\
+  & \textsf{aligned } \vaddr \\
+  & \vaddr = \locsx ++ \lfoff ++ \ltoff ++ \ltwoff ++ \looff ++ \pgoff \\
   &  \pfpointsto{\lvlsum\crt\lfoff}{\lvlbor\lt}\qfrac\ppts \\
-  &  \pfpointsto{\lvlsum\lt\ltoff}{\lvlbor\ltw}\qfrac\ppts \\
-  & \pfpointsto{\lvlsum\ltw\ltwoff}{\lvlbor\lo}\qfrac\ppts \\
-  & \pfpointsto{\lvlsum\lo\looff}{\lvlbor\pg}\qfrac\ppts \\
-  & \ppointsto{\pageptstosum\pg\pgoff}\locsf'\ppts \\
+  &  \pfpointsto{\lvlsum\lt\ltoff}{\lvlbor\ltw}\qfractw\ppts \\
+  & \pfpointsto{\lvlsum\ltw\ltwoff}{\lvlbor\lo}\qfract\ppts \\
+  & \pfpointsto{\lvlsum\lo\looff}{\lvlbor\pg}\qfracf\ppts \\
+  & \ppointsto{\pageptstosum\pg\pgoff}\locsf\ppts \\
 
@@
 \end{figure*}
-
+\todo{iso todo: stress fractions and aligment}
 \subsection{Modal Abstraction for Virtualizing Memory Locations}
@@
 \input{fig-laws}
+The most essential part of our reasoning principles is shaped around abstractions for location virtualization, more concretely virtual memory points-to relation in Figure \fref{fig:virtualpointsto}.
+One might have already noticed that the resource represented by the virtual-points to is not only a simple immediately mapped singly resource but also
+\begin{itemize}
+\item a tree-like structured resrouce rooted at the certain physical memory address ($\crt$) -- not a single resource mapping abstraction 
+\item the shared indirection levels abstracted as mappings with fractional permissions
+\item through partially owned set of resource mappings grants the eventual uniquely ownable single resource, i.e. page address 
+\end{itemize}
 
@@
 A more precise statement is given in the next section (\sref{sec:soundness}).
-
-\mytodo{iso todo: talk about each rule}
+\mytodo{iso todo: talk about each rule and change the intro to this subsection}

### 2022-09-19 background.tex
../..

@@
 The problem of needing to keep track of things being true in some contexts and not in others is hardly unique to virtual memory management, and is the general insight behind most flavors of modal logic.
-\mytodo{Colin finish}
+
+Most modal logics use unary operators on propositions to express that a logical claim $P$ is \emph{contingently} true in certain other circumstances, such as in other times~\cite{pnueli1977} or places~\cite{gordon2019modal}.
+That is, for different modal operators $M$, $M(P)$ may mean that $P$ is true in another time or place (which may be a physical place like another network node~\cite{murphy,gordon2019modal} or a conceptual place like a person's mind~\cite{hintikka1962knowledge}).
+A unifying concept across any modality is that they behave as applicative functors, typically satisfying (directly, or as a derived law):
+\[ (P\rightarrow Q) \rightarrow M(P) \rightarrow M(Q)\]
+\mytodo{Ismail, I think this means we have pure intro, $\square P \mathrel{-\ast} [r](\square P)$ if I'm using the right symbol for pure assertions }
+Many modalities, so-called \emph{normal} modalities also possess introduction rules of the form $P\rightarrow M(P)$, the classic example being that if $P$ is true, then $P$ is \emph{necessarily} true ($\square P$)
+
+Of particular interest for reasoning about virtual memory are modalities that permit \emph{naming} the alternate circumstances, prominently \emph{hybrid} modal logics~\cite{blackburn1995hybrid,areces2001hybrid}, which come equipped with assertions of the form $@_\ell(P)$ indicating that $P$ is true in the specific alternate circumstance (Kripke world) named by the \emph{nominal} $\ell$.
+
+For our purposes, these are natural candidates to adapt for virtual memory management.
+We can reinterpret the notion of naming an alternate world slightly more loosely, and instead name \emph{address spaces} by the physical address of the page table root, since these structures are the physical representations of page tables.
+Thus in this paper we develop the notion that we can represent contingent truth of an assertion via $[r](P)$ indicating that $P$ holds in the address space rooted at physical address $r$.
+The typical hybrid modality introduction rule, that $P$ and $\ell$ (indicating the current world is $\ell$) imply $@_\ell(P)$, has a natural analogue: knowing $P$ and that the current address space is $r$ (i.e., that $r$ is the current \texttt{cr3} value) suggests a way to construct $[r](P)$.
+
+A relatively under-explored space of modal logics is the interaction of modal and substructural logics, in particular hybrid-style modalities in substructural logics, which has seen only minimal exploration and no prior application.
+We develop our ideas in the Iris program logic~\cite{jung2018}, as a result exploring additional subtleties that arise where the modality itself may entail ownership of resources, as well as interactions between our hybrid-style modality and substructural rules.
+For example, Iris contains a number of modalities that distribute over separating conjunction, or for which resources can freely move into the modality (e.g., $\blacktriangleright(P)\ast Q \mathrel{-\ast} \blacktriangleright(P\ast Q)$).
+In our setting some of these rules apply while others do not.
+For example, in our setting, an assertion that involves no modalities is interpreted as holding in the current (active-on-the-CPU) address space, so clearly cannot move into arbitrary other address spaces --- unless guarded by another address space modality.

### 2022-09-19 syntax.tex
+
../..

@@
 \label{sec:syntax}
-\subsection{Registers, Memory Indices}
+To develop our core logical ideas and explain their intepretation, we need only a handful of instructions, and a machine model corresponding to execution of x86-64 assembly instructions with virtual memory enabled on the CPU.
+
+\subsection{Registers and Memory}
 Programs we demonstrate in this paper requires accessing two types of computer resource: registers and memory.
@@
 \input{fig-syntax}
-\subsection{State}
-\label{state}

### 2022-09-19 programlogic.tex
+

@@
 In this section, we describe our high-level reasoning principles for location virtualization whose detailed justification is shown in \sref{sec:soundness}.
+Here we present our reasoning principle shaped around understanding the virtual address space, specifially the opeations using virtual points-to mappings.
 \subsection{Points-To Assertions}
@@
 \end{itemize}
+which we think can be well-summarized with the modal abstraction shown in \RULE{ModalLocVirt} which allows us to have
+\begin{itemize}
+  \item \textbf{contingency}: the fact $\modalP$ holds \textit{at} on the resrouces \textit{hidden-under} the physical location $\ell$ (e.g. $\crt$)
+  \item \textbf{hiding states}: hide the indirection abstracted as page-table mappings through the local-state modality
+  \item \textbf{composing local-states}: allows c
+\end{itemize}
 

### 2022-09-19 intro.tex

@@
 
-We develop these ideas in an Iris-based\mytodo{cite JFP} separation logic, and verify simplified versions of several critical virtual-memory-related pieces of OS functionality, including mapping and unmapping pages, and describing the self-referential identity maps used in most kernels to allow convenient access to physical memory.\mytodo{colin finish}
+We develop these ideas in an Iris-based\mytodo{cite JFP} separation logic, and verify simplified versions of several critical virtual-memory-related pieces of OS functionality, including mapping and unmapping pages, and describing the self-referential identity maps used in most kernels to allow convenient access to physical memory.

### 2022-09-19 programlogic.tex
+

--- 2022-09-19 programlogic.tex
@@
 As expected, we define the virtual-points to relation, ($\pfpointsto\vaddr\locsf\qfrac\vpts$), in terms of multiple physical memory mappings representing the indirection shown in Figure \fref{fig:}
-\begin{figure*}
+\begin{figure}
 \[
-\begin{array}{cl}
-  \ppointsto\vaddr\locsf\vpts \stackrel{def}{=} \lambda \tlf\Loc.
-  & \exists_{\tlfoff\Locn \;, \tltoff\Locn \;, \tltwoff\Locn \;, \tlooff\Locn \;, \tpgoff\Loctw \;, \tlt\Loc \;, \tltw\Loc \;, \tlo\Loc \tpg\Loc}\\
-  & \textsf{aligned } \vaddr \\
-  & \vaddr = \locsx ++ \lfoff ++ \ltoff ++ \ltwoff ++ \looff ++ \pgoff \\
-  &  \pfpointsto{\lvlsum\crt\lfoff}{\lvlbor\lt}\qfrac\ppts \\
-  &  \pfpointsto{\lvlsum\lt\ltoff}{\lvlbor\ltw}\qfractw\ppts \\
-  & \pfpointsto{\lvlsum\ltw\ltwoff}{\lvlbor\lo}\qfract\ppts \\
-  & \pfpointsto{\lvlsum\lo\looff}{\lvlbor\pg}\qfracf\ppts \\
-  & \ppointsto{\pageptstosum\pg\pgoff}\locsf\ppts \\
-
+\begin{array}{l}
+  \ppointsto\vaddr\locsf\vpts \stackrel{def}{=} \lambda \tlf\Loc.\\
+  \quad \exists_{\tlfoff\Locn \;, \tltoff\Locn \;, \tltwoff\Locn \;,} \\
+  \quad \exists_{\tlooff\Locn \;, \tpgoff\Loctw \;, \tlt\Loc \;, \tltw\Loc \;, \tlo\Loc \tpg\Loc} \ldotp\\
+  \quad \ulcorner \textsf{aligned } \vaddr \urcorner \star\\
+  \quad \ulcorner \vaddr = \locsx ++ \lfoff ++ \ltoff ++ \\
+  \qquad\qquad \qquad \ltwoff ++ \looff ++ \pgoff \urcorner \star\\
+  \quad  \pfpointsto{\lvlsum\crt\lfoff}{\lvlbor\lt}\qfrac\ppts \star \\
+  \quad  \pfpointsto{\lvlsum\lt\ltoff}{\lvlbor\ltw}\qfractw\ppts \star \\
+  \quad \pfpointsto{\lvlsum\ltw\ltwoff}{\lvlbor\lo}\qfract\ppts \star \\
+  \quad \pfpointsto{\lvlsum\lo\looff}{\lvlbor\pg}\qfracf\ppts \star \\
+  \quad \ppointsto{\pageptstosum\pg\pgoff}\locsf\ppts 
 \end{array}
@@
   \label{fig:virtualpointsto}
-\end{figure*}
+\end{figure}
 \todo{iso todo: stress fractions and aligment}
-\subsection{Modal Abstraction for Virtualizing Memory Locations}
-\label{sec:modallocationvirtualization}
-\input{fig-laws}
-The most essential part of our reasoning principles is shaped around abstractions for location virtualization, more concretely virtual memory points-to relation in Figure \fref{fig:virtualpointsto}.
-One might have already noticed that the resource represented by the virtual-points to is not only a simple immediately mapped singly resource but also
-\begin{itemize}
-\item a tree-like structured resrouce rooted at the certain physical memory address ($\crt$) -- not a single resource mapping abstraction 
-\item the shared indirection levels abstracted as mappings with fractional permissions
-\item through partially owned set of resource mappings grants the eventual uniquely ownable single resource, i.e. page address 
-\end{itemize}
-
 \subsection{Reasoning Rules for Selected Instructions}
@@
 \mytodo{iso todo: talk about each rule and change the intro to this subsection}
+\subsection{Modal Abstraction for Virtualizing Memory Locations}
+\label{sec:modallocationvirtualization}
+\input{fig-laws}
+The most essential part of our reasoning principles is shaped around abstractions for location virtualization, more concretely virtual memory points-to relation in Figure \fref{fig:virtualpointsto}.
+One might have already noticed that the resource represented by the virtual-points to is not only a simple immediately mapped singly resource but also
+\begin{itemize}
+\item a tree-like structured resrouce rooted at the certain physical memory address ($\crt$) -- not a single resource mapping abstraction 
+\item the shared indirection levels abstracted as mappings with fractional permissions
+\item through partially owned set of resource mappings grants the eventual uniquely ownable single resource, i.e. page address 
+\end{itemize}
+which we think can be well-summarized with the modal abstraction shown in \RULE{ModalLocVirt} which allows us to have
+\begin{itemize}
+  \item \textbf{contingency}: the fact $\modalP$ abstracts the memory address mappings \textit{under} the address space rooted at $\ell$ (e.g. $\crt$)
+  \item \textbf{hiding states}: \textit{controlled-independence} from the indirection abstracted as page-table mappings is achieved through hiding them under the virtual address space modality
+  \item \textbf{interaction with ambient logic for a local-state}: \mytodo{talk about this issue's vitality in this paper with Colin}
+\end{itemize}
+
+To concretize the understanding of these principles achieved with the modal abstraction for address-space, $\modaldef\ell\modalP$, let us try to specify \textsf{movctl} in addition to the specifications of general \textsf{mov} instructions in Figure \fref{fig:reasoning}.

### 2022-09-20 programlogic.tex
movctl

@@
 
-To concretize the understanding of these principles achieved with the modal abstraction for address-space, $\modaldef\ell\modalP$, let us try to specify \textsf{movctl} in addition to the specifications of general \textsf{mov} instructions in Figure \fref{fig:reasoning}.
+To concretize the understanding of these principles achieved with the modal abstraction for address-space, $\modaldef\ell\modalP$, let us first investigate the specification of the instruction ($\textsf{movctl\_instr}$) switching address spaces.
+Assume that our CPU keeps the current address space based on its page table register CR3, as we see in Figure \fref{fig:addressspaceswitching}
+\input{fig-ctlreasoning}

### 2022-09-20 programlogic.tex
+

@@
 To concretize the understanding of these principles achieved with the modal abstraction for address-space, $\modaldef\ell\modalP$, let us first investigate the specification of the instruction ($\textsf{movctl\_instr}$) switching address spaces.
-Assume that our CPU keeps the current address space based on its page table register CR3, as we see in Figure \fref{fig:addressspaceswitching}
+Assume that our CPU keeps the current address space based on its page table register CR3, as we see in Figure \fref{fig:addressspaceswitching}.
 \input{fig-ctlreasoning}
+Thinking of $\modalP$ and $\modalQ$ as resources representing the indirection mappings ($\mapsto_{\kw{p}}$) in the address spaces rooted at $\rvsrc$ and $\rvdst$ respectively, the rule interacts with the ambient logic, \SL, via pulls local state out of the address-space modality rooted at $\rvdst$, and make it the observable view of the memory.

### 2022-09-20 background.tex

@@
 \[ (P\rightarrow Q) \rightarrow M(P) \rightarrow M(Q)\]
-\mytodo{Ismail, I think this means we have pure intro, $\square P \mathrel{-\ast} [r](\square P)$ if I'm using the right symbol for pure assertions }
+\todo[inline]{Ismail, I think this means we have pure intro, $\square P \mathrel{-\ast} [r](\square P)$ if I'm using the right symbol for pure assertions }
 Many modalities, so-called \emph{normal} modalities also possess introduction rules of the form $P\rightarrow M(P)$, the classic example being that if $P$ is true, then $P$ is \emph{necessarily} true ($\square P$)

### 2022-09-20 intro.tex
a couple notes

@@
 We propose tackling the verification of virtual memory subsystems (and kernels more broadly) by adapting ideas from hybrid modal logic, which can label assertions true under \emph{other} circumstances (i.e., in another address space) with a modality indexed by a name for that space (in our case, the root of the page tables for an address space).
-This offers a convenient and powerful way to modularly (1) isolate assertions specific to a particular address space, and (2) explicitly state when an assertion is true across address spaces.
+This offers a convenient and powerful way to modularly (1) isolate assertions specific to a particular address space, (2) explicitly state when an assertion is true across address spaces, (3) manipulate address spaces from within other address spaces, and (4) reason about change in address spaces.
+These advantages make this approach to reasoning about virtual memory more flexible than prior program logic techniques~\cite{kolanski08vstte,kolanski09tphols}, which was only able to work with a single address space (the current address space on the CPU).
 

### 2022-09-20 semantics.tex
+
a couple notes

@@
 \label{sec:state}
+\todo[inline]{I don't think we need or want to describe the operational semantics in detail in the submission.
+It adds a lot of formalization that doen'st necessarily contribute to reader understanding.
+We should describe the runtime \emph{state} $\sigma$ of memory and registers, since that is directly used in interpreting assertions.
+But the reduction rules will mostly add mental overhead here.}
 We represent the machine state mainly as a finite map of registers to register values and a map of masked physical memory addresses to 64-bit physical memory values.
@@
 \end{itemize}
-\subsection{Indexed Memory \& Register by References}
+\subsection{Memory Addresses \& Naming Registers}
 \label{sec:}
@@
 \end{itemize}
-In fact $\readlval\maddr{\storememstar\crval}\locsf$ is unfoled into multiple physical memory lookups for the final page address retrieval -- i.e. address translation traversal. 
+In fact $\readlval\maddr{\storememstar\crval}\locsf$ is unfoled into multiple physical memory lookups for the final page address retrieval -- i.e.
+address translation traversal as shown in Figure \ref{fig:pagetables}  \mytodo{1.address more about the challenging aspects of this traversal and the last level update}
 \input{fig-semantics}

### 2022-09-20 relwork.tex
write most of related work

\section{Related Work}
\label{sec:relwork}

There has been relatively little prior work on formal verification of virtual memory.
Instead, much OS verification work has focused on minimizing reasoning about virtual memory management.
The original Verisoft project~\cite{alkassar2008verisoft,alkassar2010pervasive,alkassar2008formal,dalinger2005verification,hillebrand2005address,alkassar2008formal,starostin2010formal} relied on custom hardware which, among other things, always ran kernel code with virtual memory disabled, removing the circularity that is a key challenge of verifying actual virtual memory code: at that point page tables become a basic partial map data structure to represent user program address translations.
Other work on OS verification either never progressed far enough to address VMM verification (Verisoft XT~\cite{cohen2009vcc,cohen2010local,dahlweid2009vcc,cohen2013SOFSEM}), or uses memory-safe languages to enable safe co-habitation of a single address space by all processes (Singularity~\cite{Fahndrich2006language,Hunt2007singularity,Hunt2007sealing,Barnett2011specsharp}, Verve~\cite{Yang2010Verve}, and Tock~\cite{levy2017multiprogramming}).

The work that does address the core challenges of VMM verification is all associated with either \textsc{seL4} or \textsc{CertiKOS}.

\textsc{CertiKOS}~\cite{gu15,gu2016certikos,gu2018certikos,chen2016interrupts} is a microkernel intended for use as a hypervisor, and its papers do not explicitly detail verification of the VMM, so we do not know the full space of which VMM functionality is verified (e.g., we do not know if it includes lazy page loading or copy-on-write behavior, or is a minimalistic VMM that eagerly maps address ranges for guessts). The overall approach in that body of work is many layers of refinement proofs, using a proliferation of layers with small differences to keep most individual refinements tractable. In keeping with precursor work on the project from the same group~\cite{vaynberg2012compositional}, the purpose of some layers is to abstract away from virtual memory, so the proof is essentially a simulation proof covering for example a proof that execution with page-in on page faults is a valid refinement of an execution model where.
Another key aspect of their approach is that the OS is written in Clight and compiled with \textsc{CompCert}~\cite{blazy2006formal,leroy2009formally,leroy2008formal}.
CompCert's memory abstraction~\cite{leroy2008formal} assumes
memory is a set of disjoint chunks of bytes with no overlap, so the lowest levels of CertiKOS must provide a matching machine model as a layer. This prohibits virtual address aliasing, so CertiKOS cannot support simultaneous memory-mapped (\texttt{mmap}) and stream-oriented (\texttt{read}/\texttt{write}) IO to a single file\todo{should we go into this detail?}, and cannot use
the common kernel design choice of mapping all physical memory into the bottom of the kernel's address space for direct access while the kernel code is simultaneously mapped (and executed) at higher virtual addresses.
This is not necessary for \textsc{CertiKOS}'s intended primary use case (a hypervisor), but means that \textsc{CertiKOS}'s approach cannot be used to support this functionality in other systems, without major surgery to \textsc{CompCert}.

\textsc{seL4}~\cite{Klein2009seL4,seL4TOCS,Sewell2013translation} is a formally verified L4 microkernel~\cite{Liedtke1995,Liedtke1996} (and the first verified OS kernel to run on real-world hardware), verified with a mix of refinement proofs and program logic reasoning down to the assembly level.
Because \textsc{seL4} is a microkernel, most VMM functionality actually lives in usermode and is unverified, and moreover, their hardware model omits address translation entirely and the MMU entirely~\cite{Klein2009seL4,seL4TOCS}. As a result, the limited page table management present in the microkernel treats page tables as idiosyncratic tree-maps, ignoring the risks posed by even transient inconsistencies that would crash the kernel on real hardware (like ``temporarily'' unmapping the kernel). This is mitigated primarily by manually identifying some trusted invariants (e.g., that the address range designated for the kernel is appropriately mapped) and setting up the proof to ensure those invariants are maintained (i.e., as an extra proof obligation not required by their hardware model).

\todo[inline]{Need to come back and ack common weaknesses of above work and this work, like skipping the TLB (though there's newer stuff not integrated into seL4), and not treating the relationship between memory and instructions precisely.}

One important outgrowth of the \textsc{seL4} project, not integrated into the main project's proof, was work by Kolanski and Klein which studied verification of code against a hardware model that \emph{did} include address translation --- the only work aside from ours to do so --- initially in terms of basic memory~\cite{kolanski08vstte} and subsequently integrating source-level types into the interpretation~\cite{kolanski09tphols}. 
They were the first work to model physical and virtual points-to assertions separately, defining virtual points-to assertions in terms of physical points-to assertions mimicking page table walks, and defining all of their assertions as predicates on a pair of (physical) machine memory and a page table root, an approach we improve on.

Kolanski and Klein's points-to assertions do not model that page table entries for nearby addresses typically \emph{share} entries in higher layers of the page tables,\todo{actually, not sure, need to confirm} a key limitation that our definitions avoid.
They do not consider virtuallv address aliasing (another source of sharing physical resources), which we\todo{todo!} support.
They verify only one address space manipulation: mapping a single unmapped page into the current address space (in both papers).
The mapping write is verified in an ad hoc way by unfolding the machine semantics, because the logic lacks proper reusable rules for
updating page tables.  
Critically, they do not consider
manipulating \emph{other} address spaces besides the one in use (essential for managing
processes).
So their means for sharing access to intermediate page table levels does not extend across address spaces, and there is no way to share data pages between address spaces.  But every major OS includes widely-used facilities for sharing pages between processes (e.g., UNIX/POSIX \texttt{MAP\_SHARED}), and until the recent Spectre~\cite{kocher2018spectre} and Meltdown~\cite{lipp2018meltdown} attacks, FreeBSD, Solaris, and others shared the kernel mapping across processes~\cite{mckusick2014design,mcdougall2006solaris,wininternals6a,wininternals6b} (they still do so for performance on unaffected hardware; e.g., Meltdown affects Intel but not AMD~\cite{amdmeltdown}).

Our approach in this paper uses modalities to distinguish virtual address based assertions that hold only in specific address spaces, making it possible to manipulate other address spaces, and equally critically, to \emph{change} address spaces while reasoning about correctness. In addition, while Kolanski and Klein work in a traditional Hoare logic, we work in a separation logic, inheriting its modularity gains at the cost of working out some interactions between modalities and substructural logics.

Unlike our work, Kolanski and Klein prove very useful embedding theorems stating that code that does not modify page table entries can be verified in a VM-ignorant program logic, and that proofs in that logic can be embedded into the VM-aware logic (essentially by interpreting ``normal'' points-to relations as virtual points-to facts). While we have not proven such a result, an analagous result \emph{should} hold of our work: consider that the triples for the \texttt{mov} instruction that access memory behave just as one would expect for a VM-ignorant logic~\todo{cite some example logic}.
With our general approach to virtual points-to assertions being inspired by Kolanski and Klein, \emph{both} of our approaches could in principle be extended to account for pageable points-to assertions by adding additional disjunctions to an extended points-to definition.

\todo[inline]{discuss hybrid+substructural logics}

### 2022-09-20 programlogic.tex
a couple notes
+
quick comment

@@
 \caption{Virtual Points-to Relation}
+\todo[inline]{These fractions aren't quite right (though I see you added the variance between levels), I can walk you through in our meeting tomorrow.}
   \label{fig:virtualpointsto}
@@
 Thinking of $\modalP$ and $\modalQ$ as resources representing the indirection mappings ($\mapsto_{\kw{p}}$) in the address spaces rooted at $\rvsrc$ and $\rvdst$ respectively, the rule interacts with the ambient logic, \SL, via pulls local state out of the address-space modality rooted at $\rvdst$, and make it the observable view of the memory.
+
+\todo[inline]{We don't need to talk about sexec at all.
+Let's also talk tomorrow about adjusting the syntax in the paper to mimic actual textual assembly more closely (e.g., all of the instructions are mov, with different argument shapes)}
+
+\mytodo{STOP HERE}

### 2022-09-20 examples.tex
update example's work division
another asm snippet

@@
 \label{sec:experiment}
-\mytodo{Ismail's Part}
+\mytodo{Mechanization is Ismail's Part -- Discussing and putting in here is Colin's}
+
+To both validate and demonstrate the value of the modal approach to reasoning about virtual memory management, we study several distillations of real concerns of virtual memory managers.
+Recall from Section \ref{sec:logic} that virtual points-to assertions work just like regular points-to assertions, by design.
+
+\subsection{Identity Mappings}
+\todo[inline]{need to describe this}
+
+\subsection{Mapping a New Page}
+One of the key tasks of a page fault handler in a general-purpose OS kernel is to map new pages into an address space by writing into an existing page table.
+\todo[inline]{Ismail, see notes about intermediate assertions on the Notion page}
+\begin{figure}
+\begin{lstlisting}
+;; Assume the faulting address f is in r12,
+;; a callee-save register, and f with the 
+;; lower 12 bits zeroed out (so, the virtual
+;; page containing f) is in r13 (also callee-save)
+
+;; Ensure there's a place to put the L1 mapping
+mov %rdi, r13
+call ensure_L1_page
+;; Returns the virtual address of the L1 entry in rax
+mov %r14, %rax ;; Save that before another call
+call alloc_phys_page_or_panic
+;; Calculate new PTE by setting the low-order bits
+;; of the physical page
+add %rax, 3
+;; update the page table entry, mapping the page
+mov (%rcx), %rax
+;; memset the newly-mapped page
+mov %rdi, r13
+mov %rsi, 0
+mov %rdx, 4096
+call memset
+;; The newly mapped page is full of 0s
+\end{lstlisting}
+\caption{Distilled code for mapping a new page}
+\label{fig:mapping_code}
+\todo[inline]{this is terribly ugly, should switch to minted}
+\end{figure}
+
+\todo[inline]{emphasize the write to the L1 page table is validated by the fact that the identity mapping exists}
+
+\subsection{Unmapping a Page}
+The reverse operation, unmapping a designated page that is currently mapped, is pleasingly the reverse: by flipping the entry back to an invalid entry, full ownership of the updated entry reverts back to the now-invalid cell of the L1 table.
+
+\subsection{Change of Address Space}
+A critical piece of \emph{trusted} code in verified OS kernels is the assembly code to change the current address space; current verified OS kernels currently lack effective ways to specify and reason about this low-level operation, for resaons outlined in Section \ref{sec:relwork}.
+
+\todo[inline]{probably not much more than just the cr3 update, but could be extended to a full context switch since the instruction pointer isn't updated in a swtch}
+
+\begin{lstlisting}
+;; Assume the save-space is in rdi,
+;; load-space in rsi
+;; No need to save/restore caller-save regs
+;; Save yielding context
+mov 0[%rdi], %rbx
+mov 8[%rdi], %rsp
+mov 16[%rdi], %rbp
+mov 24[%rdi], %r12
+mov 32[%rdi], %r13
+mov 40[%rdi], %r14
+mov 48[%rdi], %r15
+mov 56[%rdi], %cr3
+;; Restore target context
+mov %rbx, 0[%rsi] 
+;; This switches to a new stack
+;; This stack may not be mapped in the
+;; current address space!
+mov %rsp, 8[%rsi] 
+mov %rbp, 16[%rsi]
+mov %r12, 24[%rsi]
+mov %r13, 32[%rsi]
+mov %r14, 40[%rsi]
+mov %r15, 48[%rsi]
+;; Switch to the new address space
+mov %cr3, 56[%rsi]
+\end{lstlisting}
+\todo[inline]{technically we weren't planning to prove reads of cr3...}
+\todo[inline]{need to figure out the right pre/post condition to highlight the change in address space, beyond the vpoints-to of the two structures}

### 2022-09-20 cpp_paper.tex
separted stuff
+
stuff from colin
some intro stuff, bibliography
../..
write most of related work
try and fail to make the listings non-hideous
marginally less ugly assembly listings

@@
 \documentclass[sigplan,10pt,anonymous,review]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
+\usepackage[utf8]{inputenc}
+\usepackage{todonotes}
+\usepackage{blindtext}
+\usepackage{listings}
+\lstset{
+  columns=fullflexible,
+  numbers=left,
+  basicstyle=\ttfamily,
+  keywordstyle=\color{blue}\bfseries,
+  morekeywords={mov,add,call},
+  emph={rdx,rax,rbx,rbp,rsi,rdi,rcx,r8,r9,r10,r11,r12,r13,r14,r15},
+  emphstyle=\color{green},
+  emph={[2]cr3},
+  emphstyle={[2]\color{violet}},
+  morecomment=[l]{;;}
+}
+% Hack for Colin to edit in low light: https://tex.stackexchange.com/questions/40495/invert-background-and-text-colours-across-whole-document-with-pdflatex
+%Commands definitions
+\newcommand{\setbackgroundcolour}{\pagecolor[rgb]{0.19,0.19,0.19}}  
+\newcommand{\settextcolour}{\color[rgb]{0.77,0.77,0.77}}    
+\newcommand{\invertbackgroundtext}{\setbackgroundcolour\settextcolour}
+%Command execution. 
+%If this line is commented, then the appearance remains as usual.
+\invertbackgroundtext
+
+% Packages.
+\usepackage{amsmath}
+% If we need multiple references to a footnote, via \cref:
+\usepackage{cleveref}
+\crefformat{footnote}{#2\footnotemark[#1]#3}
+\usepackage[T1]{fontenc}
+\usepackage{fontawesome}
+\usepackage[utf8]{inputenc}
+\usepackage{iris}
+\usepackage{mathpartir}
+\renewcommand{\TirNameStyle}[1]{\hypertarget{#1}{\textsc{#1}}}
+\newcommand{\RULE}[1]{\hyperlink{#1}{\textsc{#1}}\xspace}
+\usepackage{xspace}
+\input{macros}
+
+\newcommand{\mytodo}[1]{% <==========================================
+  \todo[linecolor=white, bordercolor=white, textcolor=white]{#1}%
+}
+
 %%
@@
 %% the work being presented. Separate the keywords with commas.
-\keywords{datasets, neural networks, gaze detection, text tagging}
+\keywords{program verification}
 
 \maketitle
+
+{
+\theoremstyle{acmdefinition}
+\newtheorem{assumption}[theorem]{Assumption}
+}
+
 \input{intro.tex}
+%\input{motivation.tex}
 \input{background.tex}
@@
 \input{programlogic.tex}
+\input{soundness.tex}
 \input{examples.tex}
 \input{discussion.tex}
+\input{relwork.tex}
 %% The next two lines define the bibliography style to be used, and
@@
 \bibliographystyle{ACM-Reference-Format}
-\bibliography{sample-base}
+\bibliography{vmm}
 

### 2022-09-20 soundness.tex
will continue on soundness tommorrwo

@@
-\section{Soundness}
+\section{Implementing Logical Machinery \& Soundness}
+We instantiate our program logic, $[@]$, as an instantiation of Iris~\cite{iris}, and build our modal abstractions on top of it. 
+\subsection{BI-Lifting for Address Space Modality}
+\label{sec:bilifting}
+\mytodo{we should put some stuff here}
+\subsection{Soundness}
 \label{sec:soundness}
+Our logic, $[@]$, operates on the machine state, which means we do not need to augment the machine state, i.e.
+our logical state is exact logical representation of our machine state.
+Consequently, we do not require an invariant, $\mathcal{I}$, except from the fact that logical/physical associated maps representing the register/memory mappings have the same domain.
+This ensures the correct lifting of mappings in the machine state to the \emph{assertions} that the client of our logic uses, i.e.
+points-to assertions that are defined as the ownership of a fragment of the logical state.
 
-Our ... is an instance of Iris~\cite{iris}. This means that our logic ... is set up
-via the following steps:
-
-\begin{enumerate}
-\item define a certain amount of \emph{ghost state}
-      that is needed in the next two steps;
-\item define a central \emph{invariant}
-      that ties the machine's physical state
-      to the ghost state;
-\item
-  define the \emph{assertions} that an end user of our ... exploits, namely
-  points-to assertions that are defined as the ownership
-  of a fragment of the ghost state.
-\end{enumerate}
-
+The crucial pragmatic aspect of instantiating Iris for your reasoning principles \todo{Tired stop here.} 
 Once these steps have been carried out, Iris provides a semantic definition of

### 2022-09-20 soundness.tex
+add more

@@
 
-The crucial pragmatic aspect of instantiating Iris for your reasoning principles \todo{Tired stop here.} 
-Once these steps have been carried out, Iris provides a semantic definition of
-the Hoare triple $\triple\pre\instr\post$. We can then use Hoare triples to
-state the reasoning rules of ...., to prove that
-these rules are valid, and to state and prove the fact that ... is sound.
-This is the topic of the next subsection (\sref{def:soundness:statement}).
-%
-After presenting these results, in \sref{sec:invariant}, we come back to the
-three steps listed above and describe them at an informal level, without
-diving into the concepts of Iris, which would be difficult to explain in a
-thorough and concise way.
-% For more information on Iris, the reader is referred to Jung \etal.'s
-% paper~(\citeyear{iris}).
+Iris instantiation comes with the semantic definition for the weakest-precondition which we can refactor into Hoare triples, $\triple\pre\instr\post$, to specify actions using our reasoning principles, show that these triples are sound.
 
@@
 \label{def:soundness:statement}
-
-The operational semantics of ...lang guarantees that ... (\sref{sec:valid}). A~program that attempts to ... stuck~(\sref{sec:instrsemantics}).
-% Thus, the space consumption of every \spacelang program is bounded.
-Thus, in order to prove that ... is sound, all we have to do is prove that
-it rules out the possibility of a runtime failure. This is guaranteed by the
-following theorem:
-
-\begin{theorem}[Soundness of the Logic ...]
+The operational semantics of our simple lang just executes the instruction of our x86-64 model.
+Therefore, our soundness argument is to show that any execution composed of actions in our machine model (some of which are shown in Figure \sref{sec:semantics}) does not end-up in a invalid state.guarantees that  (\sref{sec:valid}).
+A~program that attempts to ...
+stuck~(\sref{sec:instrsemantics}).
+\begin{theorem}[Soundness of the Logic]
   \label{th:adequacy}
+  Assuming that  Suppose $\triple{\textsf{valid\_init}}\instr\iTrue$ holds.
  The execution of the instruction~$\instrs$, beginning in an initial state, cannot result in a configuration where a thread is stuck.
 \end{theorem}
+which states that if the program~$\instr$, with the given \textsf{valid\_init} asserting a valid state initialization, satisfies a semantic Hoare
+triple, then this program cannot crash: by a direct consequence of Iris's adequacy theorem~\cite[\S6.4]{iris}.
 
-This theorem states that if the program~$\instr$ satisfies a semantic Hoare
-triple, then this program cannot crash. In
-particular, it cannot reach a situation where DOING X ....
-%
-The proof of this theorem is short: it is a direct consequence of Iris's
-adequacy theorem~\cite[\S6.4]{iris}.
-
-Separately, we prove that the reasoning rules of~our logic are valid, that is,
-each rule is a valid lemma:
-
+Moreover we need to show the validity of each rules in Figures \fref{fig:reasoning} and \fref{fig:laws}.
 \begin{theorem}[Validity of the Reasoning Rules]
@@
 \end{theorem}
-
-The proof of this theorem represents more work: one lemma per reasoning rule
-is required, not to mention a large number of auxiliary lemmas. The proofs of
-these lemmas are fairly uninteresting; the key insights lie in the definitions
-presented in the next subsection (\sref{sec:invariant}).
+Due to the space limits in this paper, we do not mention each proof for these rules explicitly (which are already mechanized in Coq),
+but the definitions and constructions used requires explanation.
 
 Together, Theorems~\ref{th:adequacy} and~\ref{th:validity} guarantee that, if
-the Hoare triple $\iTrue\instrs\iTrue$ can be obtained by applying
-the reasoning rules of our logic, then the program~$\instrs$ is safe, that is, it cannot crash ....
+the Hoare triple $\textsf{valid\_init}\instrs\iTrue$ can be obtained by applying
+the reasoning rules of our logic, then the program~$\instrs$ is safe.
 
-\subsection{Key Definitions and Invariants}
+\subsection{Logical Constructions}
 \label{sec:invariant}

### 2022-09-20 soundness.tex
+

@@
 
-For the definitions that follow, we need some more terminology about stores.
+Since our semantics operates right on top of the x86-64 machine state (with Iris terminology, physical store),
+we do not need to augment the physical store for more abstract reasoning, consequently, we do not need to come up
+with a complicated logical relations that ties the physical store to the
+\logical store. 
+\begin{definition}[Relation between Physical and \Logical Stores]
+\label{def:related}
+  Two stores $\store$ and $\logicalstore$ are \emph{related},
+  which we write $\related\store\logicalstore$, when
+  they have the same domain.
+\end{definition}
 
-Next, we must define the relation that ties the physical store to the
-\logical store, two concepts that we have informally described earlier
-(\sref{sec:logical}). From here on, we write~$\store$ for the physical
-store and~$\logicalstore$ for the \logical store.
-
-Equipped with the above definitions, we are now prepared to enter the realm of
-Iris and perform the three steps described at the beginning of
-\sref{sec:soundness}, that is, define the ghost state, the central invariant,
-and the assertions of our .. . We give the definitions first and explain them
-afterwards.
-
-\begin{comment}
-% fp: this is the beginning of an explanation of ghost state in Iris
-%     but it quickly grows too long, even though many aspects are missing.
-A ghost memory cell in Iris is a memory cell that does not exist at runtime.
-The ownership and the content of a physical memory cell are described via a
-points-to predicate such as $\loc\pointsto\val$. Similarly, the ownership and
-content of a ghost cell are described by an Iris assertion, usually written
-$\ownGhost\gamma{a}$, where $\gamma$ is the name of the ghost cell and $a$ is
-its content. Whereas the content of a physical memory cell is a
-programming-language value~$\val$, the content of a ghost cell inhabits a
-mathematical structure, a \emph{camera}~\cite[\S4.4]{iris}, which the user
-chooses when the ghost cell is created. A camera is equipped with a
-composition operation~$\cdot$ and the ghost state assertion satisfies the law
-$\ownGhost\gamma{a\cdot b} \equiv \ownGhost\gamma{a} \star
-\ownGhost\gamma{b}$. Thus, the choice of a suitable camera determines in what
-ways the ghost state can be split.
-
-A key Iris idiom involves the use of the \emph{authoritative
-  camera}~\cite[\S6.3.3]{iris} together with an Iris invariant.
-The elements of the authoritative camera include
-\emph{authoritative elements} $\authfull{c}$ and
-\emph{fragmentary elements} $\authfrag{b}$.
-The composition law is defined in such a way that
-there always exists at most one authoritative element
-and the composition of all fragmentary elements in existence is
-contained in the authoritative element, etc.
-\end{comment}
+We already have the physical and logical stores and a simple invariant between them. Now, we can rely on the following Assumption \ref{assumption} from Iris to utilize its logical constructions.
 % The predicate gen_heap_interp.

### 2022-09-20 soundness.tex
+

@@
 Together, Theorems~\ref{th:adequacy} and~\ref{th:validity} guarantee that, if
-the Hoare triple $\textsf{valid\_init}\instrs\iTrue$ can be obtained by applying
+the Hoare triple $\textsf{valid\_init }\instrs\;\iTrue$ can be obtained by applying
 the reasoning rules of our logic, then the program~$\instrs$ is safe.
@@
   \label{def:related}
-  Two stores $\store$ and $\logicalstore$ are \emph{related},
-  which we write $\related\store\logicalstore$, when
-  they have the same domain.
+  Since we pick the logical store exactly same with physical store,  
+  two stores are \emph{related}, identical.
 \end{definition}

### 2022-09-20 soundness.tex
soundness done

@@
 \newcommand{\genheapinterp}[1]{\mathit{Heap}\;#1}
+\newcommand{\genmemheapinterp}[1]{\mathit{MemHeap}\;#1}
 % Our predicate pred (defined in ph.v), expanded.
@@
 \label{assumption}
-  Iris defines a certain piece of ghost state,
-  defines a predicate $\genheapinterp\store$
-  that ties a store~$\store$ to this ghost state,
-  and defines the points-to assertion $\loc\fpointsto\qv\blk$
+Iris defines two pieces of ghost state
+\begin{enumerate}
+\item  defines a predicate $\genheapinterp\store.\mathcal{R}$
+  that ties a store~$\store.\mathcal{R}$ to this ghost state,
+  and defines the points-to assertion $\ppointsto\rg\rv\qfrac\rpts$
   in terms of this ghost state.
@@
   where it is applied to the \logical store~$\store$.
+\item unlike the points-to relation obtained by direct interpretion of $\genheapinterp\store.\mathcal{R}$ using Iris existing \textsf{gen\_heap\_interp},
+  we define a new algebra (\textsf{gen\_mem\_UR}) for abstracting the nested maps due to different levels of masking in memory mappings. 
+\end{enumerate}
 \end{assumption}
 
-\begin{definition}[Ghost State]
-...
-stores an element of the monoid
+\begin{definition}[Ghost State - Memory with Nested Mappings]
+We define our custom-tailored algebra 
+\newcommand\fpfn{\rightarrow_{\textrm{fin}}}
+\( \textsf{gen\_memUR} \stackrel{def}{=}
+  \authm(\;
+  \Locft \;\fpfn\;
+  (\Loctw \;\fpfn\;  (\textsc{Frac }, \mathord{+}) \times (\textsc{Agree } \Loc,\mathord{=}) )
+  \)
+  for our nested memory mapping abstracting two different machine word masking. Then, we register our \emph{authoritative camera} \cite[\S6.3.3]{iris} from this algebra
+\[\textsf{inG } \Sigma \textsf{ gen\_memUR} \]
+As a final step, we allocate an empty ghost cell $\gamma$ at the beggining of a program execution to store the element of the monoid $\textsf{inG } \Sigma \textsf{ gen\_memUR}$.
+\end{definition}
+
+\begin{definition}[Ghost State - Register Mappings]
+We allocate $\theta$ ghost cell which stores an
+element of the monoid 
 \newcommand\fpfn{\rightarrow_{\textrm{fin}}}
@@
   \authm(\;
-    \Loc \;\fpfn\;
+    \regset \;\fpfn\;
     (\textsc{Frac}, \mathord{+})
     \times
-    (\textsc{Multiset}(\Loc), \mathord\uplus)
+    (\regvaltype, \mathord{=})
   \;)
@@
 \label{def:invariant}
-The central invariant of our logic,
-or \emph{state interpretation invariant} \cite[\S7.3]{iris},
-is the following Iris assertion:
+The central invariant of our logic is, due to lack of need for augmenting the machine state, simply the state interpretation \cite[\S7.3]{iris} : 
 \[
@@
   \genheapinterp\store & \star \\
-   \genheapinterp\store & \star
+   \genmemheapinterp\store
 \end{array}
@@
 
-\begin{definition}[Assertions]
-\label{def:assertions}
-  The assertions of our ... are defined as follows:
-  \begin{enumerate}
-  \item
-  % The definition of the points-to predicate is inherited from gen_heap.
-  % It is implicitly parameterized by a ghost cell that holds the heap.
-  The \emph{points-to assertion} ..
-  is inherited from Iris (Assumption~\ref{assumption}).
-  \item
-  The \emph{map-of-map} for a single physical memory location,
+As noted earlier (Assumption~\ref{assumption}), the predicate
+$\genheapinterp\!$ and the points-to assertion for register mappings is inherited from Iris. What
+this means, as far as we are concerned, is that the points-to assertion is
+defined in a standard manner: $\genheapinterp\!$ is applied to the physical store~$\store$, as is usually the case~\cite[\S6.3.2]{iris}.
+Likewise, we define nested interpretation $\genmemheapinterp\!$ in terms of $\genheapinterp\!$ to obtain physical memory points-to relation.
 
-  \item
-  The \emph{location-virtualization-modal}
-  \item
-  The \emph{pointed-by assertion} .
-  \end{enumerate}
-\end{definition}
-
-% What do these definitions mean?
-
-Let us try and explain the most important aspects of these definitions.
-
-In Iris, the state interpretation invariant~$\centralinvariant\;\store$ is an assertion that
-holds of the \emph{physical store}~$\store$ in between every two steps of
-computation. It is used by Iris in the definition of Hoare triples; we inherit
-this definition from Iris.
-
-Our definition of $\centralinvariant\;\store$ begins with an existential quantification over
-a \emph{\logical store}~$\logicalstore$.
-..........
-%.......
+In fact, long story short, all assertions in our logic, i.e.
+points-to assertions for registers and memory, really about the logical store, which is in our case identical to the physical store, so that we can reason about the physical store.

### 2022-09-20 cpp_paper.tex
more swtch stuff

@@
   morekeywords={mov,add,call},
-  emph={rdx,rax,rbx,rbp,rsi,rdi,rcx,r8,r9,r10,r11,r12,r13,r14,r15},
+  emph={rsp,rdx,rax,rbx,rbp,rsi,rdi,rcx,r8,r9,r10,r11,r12,r13,r14,r15},
   emphstyle=\color{green},

### 2022-09-20 examples.tex
more swtch stuff

@@
 ;; Ensure there's a place to put the L1 mapping
-mov %rdi, r13
+mov %rdi, %r13
 call ensure_L1_page
@@
 ;; memset the newly-mapped page
-mov %rdi, r13
+mov %rdi, %r13
 mov %rsi, 0
@@
 
+\begin{figure}
 \begin{lstlisting}
@@
 \end{lstlisting}
+\caption{Basic task switch code that switches address spaces.}
+\label{fig:swtch}
+\end{figure}
+
+Figure \ref{fig:swtch} gives simplified code for a basic task switch, the heart of an OS scheduler implementation.
+This is code that saves the context (registers and stack) of the running thread (here in a structure pointed to by \lstinline|rdi|'s value) and restores the context of an existing thread (from \lstinline|rsi|), including the corresponding change of address space for a target thread in another process.
+This code assumes the System V AMD64 ABI calling convention, where the normal registers not mentioned are caller-save, and therefore saved on the stack of the thread that calls this code, as well as on the new stack of the thread that is restored, thus only the callee-save registers and \texttt{cr3} must be restored.\footnote{We are simplifying in a couple basic ways.
+First, we are ignoring non-integer registers (e.g., floating point, vector registers) entirely.
+Second, we are ignoring that the caller-save registers should still be initialized to 0 to avoid leaking information across processes.
+We focus on the core logical requirements.}
+With the addition of a return instruction, this code would satisfy the C function signature\footnote{The name comes from the UNIX 6th Edition \lstinline|swtch| function, the source of the infamous ``You are not expected to understand this'' comment~\cite{lions1996lions}.}
+\begin{lstlisting}[language=C]
+void swtch(context_t* save, context_t* restore);
+\end{lstlisting}
+A call to this code begins executing one thread in one address space (whose information will be saved in \lstinline[language=C]|save|) and finishes execution executing a different thread in a different address space (whose information is initially in \lstinline[language=C]|restore|).
+
+Because this code does not directly update the instruction pointer, it is worth explaining \emph{how} this switches threads: by switching stacks.
+This is meant to be called with a return address for the current thread stored on the current stack when called --- which must be reflected in the calling convention.
+In particular, the precondition of the return address on the initial stack requires the callee-save register values at the time of the call: those stored in the first half of the code.
+Likewise, part of the invariant of the stack of the second thread, the one being restored, is that the return address on \emph{that} stack requires the saved callee-save registers stored in that context to be in registers as its precondition.
+
+The wrinkle, and the importance of the modal treatment of assertions, is that the target thread's precondition is \emph{relative to its address space}, not the address space of the calling thread.
+Thus the precondition of this code is that the then-current stack pointer has a return address expecting the then-current callee-save register values and suitablly updated (i.e., post-return) stack in the \emph{current} (initial) address space, while the stack pointer saved in the context to restore expects the same of the saved registers and stack \emph{in the other address space}.
+And the postcondition is analagous, but interpreted \emph{in a different address space}: the then-current (updated) stack has a return address expecting the new (restored) register values, and the saved context's invariant captures the precondition for restoring its execution \emph{in the previous address space}.
+At that point, it is safe to execute a return (or in a variant on this for user-mode and/or preemptive scheduling, a return-from-interrupt instruction).
+
+In addition, immediately after the page table switch, the points-to information about the saved and restored contexts is \emph{also} guarded by a modality for the old address space, since there is no automatic guarantee that that memory is mapped in the new address space.
+The ability to transfer that points-to information out of that modality is specific to a given kernel's design.
+Kernels that map kernel memory into all address spaces would need to ensure and specify enough specific details about memory mappings to allow a proof of an elimination rule for specific modally-constrained points-to assertions.
+Following Spectre and Meltdown, this kernel design became less prevalent because speculative execution of accesses to kernel addresses could leak information even if the access did eventually cause a fault (the user/kernel mode permission check was done after fetching data from memory).
+Thus many modern kernels have reverted to the older kernel design where the kernel inhabits its own unique address space, and user processes have only enough extra material mapped in their address spaces to switch into the kernel (CPUs do not speculate past updates to \texttt{cr3}).
+
 \todo[inline]{technically we weren't planning to prove reads of cr3...}

### 2022-09-20 examples.tex
recording discussions from Tuesday meeting

@@
 \subsection{Identity Mappings}
-\todo[inline]{need to describe this}
+\todo[inline]{Identity mappings are difficult, and our current approach won't quite work.
+Consider trying to have a virtual pointsto for an actual page table entry (i.e., that one could use to update a page table mapping), while also having a virtual pointsto for an address that entry mapped.
+With the current (let's call it v1) solution, we can't actually have both of those simultaneously!
+That's because the PTE pointsto will assert full ownership of the physical memory cell holding the PTE as its data value, while the virtual pointsto for the data mapped by that entry will \emph{also} assert (fractional) ownership of all entries a page table walk would traverse.
+}
+\todo[inline,color=violet]{This doesn't seem to cause issues with the mapping/unmapping examples, only with changing intermediate page table pointers.
+The mapping example requires a virtual pointsto for the blank PTE, and once filled in that ownership can be immediately split to create the 512 new virtual pointsto assertions for the newly mapped page.
+Conversely, for unmapping we'd assume ownership of all the relevant virtual pointsto assertions for the page we're unmapping, at which point we can (with a bit of work) show that they all correspond to the same L1 PTE, and extract the 512 fractional shares of that entry from the pointsto assertions.
+But changing intermediate page tables, as one would do for coallescing or splitting a superpage while preserving the virtual-to-physical mappings, couldn't be done without some really complicated separating implication tricks.}
+\todo[inline,color=green]{One possible approach to resolving this, which we came up with in our Tuesday meeting, is to recognize that the current (v1) virtual points-to is too strong, because it really doesn't care about \emph{owning} those fractional resources, it only cares that \emph{something} ensures the correct page table walk exists.
+Iris has a ghost map resource where authoritative ownership of an individual key-value pair can be handled as a resource.
+(Colin was using this in the filesystem cache.)
+We can use that mechanism to separate the virtual-to-physical translation from the physical memory involved (Kolanski and Klein may have done something similar for different reasons): (fractional) virtual points-to assertions can be defined in terms of (fractional) ownership of these authoritative ghost map entry assertions, plus sharing an invariant that the current installed page table respects all entries of the mapping.
+Unmapping collects the authoritative map kvpairs from collecting the assertions, and then can remove them from the ghost map and update the page tables.
+Critically, physical ownership of the page tables then lives in the invariant on the current page table, so some virtual pointsto assertions can refer to memory in those page tables.
+This still works with the modality, since that invariant is also semantically a predicate on a page table root.
+Let's call this v2.
+}
+\todo[inline]{the current course of action is to finish the WPs for v1 and try an example that should work with either approach (like the task switch).
+This is a good basic sanity check on our approaches.
+Once that's done, then we work out the details of v2, which should require \emph{refactoring} the same sorts of reasoning we had for v1, rather than totally different proofs from v1.}
 

### 2022-09-23 relwork.tex
some todos and some notes about Kolanski and Klein

@@
 updating page tables.
-Critically, they do not consider
+
+They also define their virtual points-to assertions such that a virtual points-to $p\mapsto_\mathsf{v} a$ owns the full lookup path to virtual address $p$.
+This means that given two virtual points-to assertions at the same time, such as $p\mapsto_\mathsf{v}a \ast p'\mapsto_\mathsf{v}b$, the memory locations traversed to translate $p$ and $p'$ must be disjoint.
+This means the logic has a peculiar limit on how many virtual points-to assertions can coexist in a proof.
+Since page tables fan out, the bottleneck is the number of entries in the root table.
+For their 32-bit ARMv6 example, the top-level address is still 4Kb (4096 bytes), and each entry (consumed entirely by a virtual points-to in their scheme) is 4 bytes, so they have a maximum of 1024 virtual points-tos in their ARMv6 configuration.
+(They do formulate their logic over an abstract model, but every architecture would incur a similar limitation.) Na\"ively transferring their model to x86-64 4-level tables would yield a limit of 512 assertions (also a 4Kb root page, but 8-byte entries).
+Our initial prototype worked around this limitation using fractional permissions on the intermediate entries, so entries in the L1 table were split 512 ways, entries in the L2 were split $512^2$, and so on --- but this results in problems when considering identity mappings of page tables (which Kolanski and Klein's approach also does not support), where the data portion of a virtual points-to to a an intermediate table entry must also coexist with the shares for the ``traversal'' portion of the virtual points-tos those entries support.
+In principle knowing how many such identity-like mappings might exist could allow shifting the fractions, but this quickly becomes very complex.
+Our approach\todo{v2} of separating the physical support for a virtual address mapping from the \emph{guarantee of} physical support for the virtual address avoids these issues.
+
+Of critical relevance to the distinction between our work and theirs, they do not consider
 manipulating \emph{other} address spaces besides the one in use (essential for managing

### 2022-09-25 cpp_paper.tex
some todos and some notes about Kolanski and Klein
new format
+

@@
 %%
-%% This is file `sample-sigplan.tex',
+%% This is file `sample-acmsmall-conf.tex',
 %% generated with the docstrip utility.
@@
 %%
-%% samples.dtx  (with options: `sigplan')
+%% samples.dtx  (with options: `acmsmall-conf')
 %% 
@@
 %% Any modified versions of this file must be renamed
-%% with new filenames distinct from sample-sigplan.tex.
+%% with new filenames distinct from sample-acmsmall-conf.tex.
 %% 
@@
 %%
-%% The first command in your LaTeX source must be the \documentclass command.
+%% The first command in your LaTeX source must be the \documentclass
+%% command.
+%%
+%% For submission and review of your manuscript please change the
+%% command to \documentclass[manuscript, screen, review]{acmart}.
+%%
+%% When submitting camera ready or to TAPS, please change the command
+%% to \documentclass[sigconf]{acmart} or whichever template is required
+%% for your publication.
+%%
+%%
+\documentclass[acmsmall]{acmart}
 
-\documentclass[sigplan,10pt,anonymous,review]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
+%%
+%% \BibTeX command to typeset BibTeX logo in the docs
+\AtBeginDocument{%
+  \providecommand\BibTeX{{%
+    Bib\TeX}}}
+
+%% Rights management information.  This information is sent to you
+%% when you complete the rights form.  These commands have SAMPLE
+%% values in them; it is your responsibility as an author to replace
+%% the commands and values with those provided to you when you
+%% complete the rights form.
+\setcopyright{acmcopyright}
+\copyrightyear{2018}
+\acmYear{2018}
+\acmDOI{XXXXXXX.XXXXXXX}
+
+%% These commands are for a PROCEEDINGS abstract or paper.
+\acmConference[Conference acronym 'XX]{Make sure to enter the correct
+  conference title from your rights confirmation emai}{June 03--05,
+  2018}{Woodstock, NY}
+%%
+%%  Uncomment \acmBooktitle if the title of the proceedings is different
+%%  from ``Proceedings of ...''!
+%%
+%%\acmBooktitle{Woodstock '18: ACM Symposium on Neural Gaze Detection,
+%%  June 03--05, 2018, Woodstock, NY}
+\acmPrice{15.00}
+\acmISBN{978-1-4503-XXXX-X/18/06}
+
+%%
+%% Submission ID.
+%% Use this when submitting an article to a sponsored event. You'll
+%% receive a unique submission ID from the organizers
+%% of the event, and this ID should be used as the parameter to this command.
+%%\acmSubmissionID{123-A56-BU3}
+
+%%
+%% For managing citations, it is recommended to use bibliography
+%% files in BibTeX format.
+%%
+%% You can then either use BibTeX with the ACM-Reference-Format style,
+%% or BibLaTeX with the acmnumeric or acmauthoryear sytles, that include
+%% support for advanced citation of software artefact from the
+%% biblatex-software package, also separately available on CTAN.
+%%
+%% Look at the sample-*-biblatex.tex files for templates showcasing
+%% the biblatex styles.
+%%
+
+%%
+%% The majority of ACM publications use numbered citations and
+%% references.  The command \citestyle{authoryear} switches to the
+%% "author year" style.
+%%
+%% If you are preparing content for an event
+%% sponsored by ACM SIGGRAPH, you must use the "author year" style of
+%% citations and references.
+%% Uncommenting
+%% the next command will enable that style.
+%%\citestyle{acmauthoryear}
+
 \usepackage[utf8]{inputenc}
@@
 %If this line is commented, then the appearance remains as usual.
-\invertbackgroundtext
+%\invertbackgroundtext
 
@@
 }
-
-%%
-%% \BibTeX command to typeset BibTeX logo in the docs
-\AtBeginDocument{%
-  \providecommand\BibTeX{{%
-    Bib\TeX}}}
-
-%% Rights management information.  This information is sent to you
-%% when you complete the rights form.  These commands have SAMPLE
-%% values in them; it is your responsibility as an author to replace
-%% the commands and values with those provided to you when you
-%% complete the rights form.
-\setcopyright{acmcopyright}
-\copyrightyear{2018}
-\acmYear{2018}
-\acmDOI{XXXXXXX.XXXXXXX}
-
-%% These commands are for a PROCEEDINGS abstract or paper.
-\acmConference[Conference acronym 'XX]{Make sure to enter the correct
-  conference title from your rights confirmation emai}{June 03--05,
-  2018}{Woodstock, NY}
-\acmPrice{15.00}
-\acmISBN{978-1-4503-XXXX-X/18/06}
 
@@
 %% allowing the author to define a "short title" to be used in page headers.
-\title{Modalities for Location Virtualization}
-\subtitle{Virtualizing Memory Addresses}
+\title{Modalities for Virtualizing Memory Addresses}
+\subtitle{Understanding the Locality with Modals}
 %%
@@
 
+\todo[inline]{Colin think through instruction pointer updates vs logic,
+think about framing plus cr3 update and modality shifts
+}
+
 \input{intro.tex}
@@
 \bibliography{vmm}
+\begin{appendix}
+\section{Complete Syntax}
+\label{appendix:syntax}
+\begin{figure}[!ht]
+\newcommand{\commentary}[1]{ & \text{\small\it #1} \\}
+\[
+  \begin{array}{r@{\;}c@{\;}l}
+    \loc & \in & \Loc \\
+    \reg & \in & \regset \\
+    \regval & \in & \regvaltype \\
+    \val & ::= & \vunit
+
+\\
+    \instrs & ::= &
+    \begin{array}[t]{@{}l@{\hspace{10mm}}l@{}}
+    \begin{array}[t]{@{}ll@{}}
+      \iskip
+                   \commentary{no-op}
+      \iseq\instr\instrs
+                   \commentary{sequencing}
+      \ising\instr
+                   \commentary{executing}             
+    \end{array}
+    %&
+    %\begin{array}[t]{@{}ll@{}}
+     % \ialloc\lval\allocsize
+      %             \commentary{heap allocation}
+    %\end{array}
+    \end{array}
+    \\
+
+    \ectx & ::= &
+      \hole \mid
+      \iseq\ectx\instrs 
+    \\
+  \end{array}
+\]
+\caption{Syntax}
+\Description{Syntax}
+\label{fig:syntax}
+\end{figure}
+\section{Operational Semantics of Selected Mov Instructions}
+\label{appendix:movops}
+\input{fig-semantics}
+
+\end{appendix}
 

### 2022-09-25 programlogic.tex
+

@@
 In this section, we describe our high-level reasoning principles for location virtualization whose detailed justification is shown in \sref{sec:soundness}.
-Here we present our reasoning principle shaped around understanding the virtual address space, specifially the opeations using virtual points-to mappings.
+The principles shown here are mainly shaped around understanding and expressing the fact related the virtual address space, most specifically virtualization and translation of memory locations .
 \subsection{Points-To Assertions}

### 2022-09-25 semantics.tex
+

--- 2022-09-18 semantics.tex
@@
-\section{Semantics}
-\label{sec:semantics}
 \subsection{State}
 \label{sec:state}
+\todo[inline]{I don't think we need or want to describe the operational semantics in detail in the submission.
+It adds a lot of formalization that doen'st necessarily contribute to reader understanding.
+We should describe the runtime \emph{state} $\sigma$ of memory and registers, since that is directly used in interpreting assertions.
+But the reduction rules will mostly add mental overhead here.}
 We represent the machine state mainly as a finite map of registers to register values and a map of masked physical memory addresses to 64-bit physical memory values.
@@
 and write just $\hs{\instr}{\store}{\instr'}{\store'}$.
-\subsection{A Set of Selected Instructions Changing the Machine State}
+\subsection{Machine State under Address Translation}
 \label{sec:selectedinstrsemantics}
-In addition to the reduction rules in Figure \fref{fig:reductionsemantics} for the light-weight integration of our machine model to our simple language, and, finally, to our reasoning principles, we think that it is worth building the intuition on the way some of the instructions -- specifically the ones used in the examples contained in this paper -- change the machine state when reduced with \RULE{StepContext}.
-\RULE{StepMovRR} in \fref{fig:instrsemantics} shows how machine state change when we copy the source register's value ($\rvsrc$) referenced by the register name ($\rgsrc$) to the destination register ($\rgdst$).
-We follow the convention of showing only the pieces of the machine state changed after the instruction execution -- only the register value of the destiantion register ($ \readlval\rgdst\storeregprime\rvsrc$) has changed in the state ($\storeprime$) after the execution of the instruction $\textsf{mvrr}$.
-All the other rules in Figure \fref{fig:instrsemantics}, presents the change in the state when a value is loaded/stored to/from register/memory, and are differentiated with respect to the addressing mode ($amode$) which determines whether the accessed memory address is computed based on some offset (\RULE{StepMovRMOff}, \RULE{StepMovMROff}) or not (\RULE{StepMovRMBase}) and \RULE{StepMovMRBase}).
-The most common essential aspect of loading the source register's value ($\rvsrc$) to the memory, \RULE{StepMovRMBase} and \RULE{StepMovRMOff}, is the hierarchical translation of memory location, $\readlval{\plusaddr\maddr\offs}{\storememstar\crval}\locsf$, which is hidden in the rules, expanded as the following:
+Althought we give the complete set of operational semantics rule in \sref{appendix:movops}, It is worth building the intuition on the the way some of these rules bahave in the context of address translation.
+In fact $\readlval\maddr{\storememstar\crval}\locsf$ is unfoled into multiple physical memory lookups for the final page address retrieval -- i.e.
+address translation traversal as shown in Figure \ref{fig:pagetables}:
 \begin{itemize}
-\item
-\item
-\item
-\item
+\item top-level-address translation: with the given root address (64-bit $\crval$) of the address space, performs the address translation, handling the first level (to get the PML4 entry) itself.
+The next level table address is computed with the fetched PML4 offset value which exhibits itself as 9-bit offset in $\kw{maddr}$.
+\mytodo{iso: put move}
+\item translating from PML4 entry: performs the second level of address translation, to retrieve starting at the PML4 table entry, and interprets the PML4 entry that references a Page-Directory-Pointer Table (PDPT).
+We obtain the PDPT offset which exhibits itself as 9 bit offset in $\kw{maddr}$ to obtain the address of the next level page directory table (PDT) \mytodo{iso: put move}
+\item translating from PD entry: performs the third level of address translation, to retrieve starting at the PDP table entry, and interprets a PDPTE that references a PD table.
+Likewise, we obtain the PD offset which exhibits itself as 9 bit offset in $\kw{maddr}$ to obtain the address of the next level page directory table (PT) \mytodo{iso:put move }
+\item translate from PT entry: performs final level of address translation, starting from the PT entry, with a given 12 bit page offset, we can compute the physical address referencing $\locsf$ \mytodo{ismo: put mov}
 \end{itemize}
-where the first three levels of translation allows address-translation .
-\input{fig-semantics}

### 2022-09-26 programlogic.tex
+
STOP: until discussing with Colin

@@
 As expected, we define the virtual-points to relation, ($\pfpointsto\vaddr\locsf\qfrac\vpts$), in terms of multiple physical memory mappings representing the indirection shown in Figure \fref{fig:}
-\begin{figure}
+\begin{figure*}
 \[
@@
   \ppointsto\vaddr\locsf\vpts \stackrel{def}{=} \lambda \tlf\Loc.\\
-  \quad \exists_{\tlfoff\Locn \;, \tltoff\Locn \;, \tltwoff\Locn \;,} \\
-  \quad \exists_{\tlooff\Locn \;, \tpgoff\Loctw \;, \tlt\Loc \;, \tltw\Loc \;, \tlo\Loc \tpg\Loc} \ldotp\\
-  \quad \ulcorner \textsf{aligned } \vaddr \urcorner \star\\
-  \quad \ulcorner \vaddr = \locsx ++ \lfoff ++ \ltoff ++ \\
-  \qquad\qquad \qquad \ltwoff ++ \looff ++ \pgoff \urcorner \star\\
-  \quad  \pfpointsto{\lvlsum\crt\lfoff}{\lvlbor\lt}\qfrac\ppts \star \\
-  \quad  \pfpointsto{\lvlsum\lt\ltoff}{\lvlbor\ltw}\qfractw\ppts \star \\
-  \quad \pfpointsto{\lvlsum\ltw\ltwoff}{\lvlbor\lo}\qfract\ppts \star \\
-  \quad \pfpointsto{\lvlsum\lo\looff}{\lvlbor\pg}\qfracf\ppts \star \\
-  \quad \ppointsto{\pageptstosum\pg\pgoff}\locsf\ppts 
+  \exists_{\tlfoff\Locn \;, \tltoff\Locn \;, \tltwoff\Locn \;,\tlooff\Locn \;, \tpgoff\Loctw \;, \tlt\Loc \;, \tltw\Loc \;, \tlo\Loc \tpg\Loc} \ldotp \\
+  \ulcorner \textsf{aligned } \vaddr \urcorner \star 
+   \ulcorner \vaddr = \locsx :: \lfoff :: \ltoff :: 
+   \ltwoff :: \looff :: \pgoff \urcorner \star\\
+  \pfpointsto{\lvlsum\crt\lfoff}{\lvlbor\lt}\qfracfotsss\ppts \star 
+  \pfpointsto{\lvlsum\lt\ltoff}{\lvlbor\ltw}\qfracfotss\ppts \star \\
+  \pfpointsto{\lvlsum\ltw\ltwoff}{\lvlbor\lo}\qfracfots\ppts \star 
+  \pfpointsto{\lvlsum\lo\looff}{\lvlbor\pg}\qfracfot\ppts \star \\
+  \ppointsto{\pageptstosum\pg\pgoff}\locsf\ppts 
 \end{array}
@@
   \label{fig:virtualpointsto}
-\end{figure}
+\end{figure*}
+\paragraph{Coalescing and Coexistence} \mytodo{Super Huge pages}
 \todo{iso todo: stress fractions and aligment}
-\subsection{Reasoning Rules for Selected Instructions}
-\label{sec:reasoning}
-\input{fig-reasoning}
-As is usual in \SL, the reasoning rules of the program logic allow deriving
-Hoare triples of the form $\triple\pre\instr\post$, where the
-precondition~$\pre$ and the postcondition~$\post$ are assertions,
-and where $i$ is an instruction.
-Such a triple receives a standard interpretation as a statement of partial
-correctness: the assertion $\triple\pre\instr\post$ means that, in a state
-where $\pre$ holds, it is safe to execute the instruction~$\instr$, and if
-this execution terminates, then it leads to a final state where $\post$ holds.
-A more precise statement is given in the next section (\sref{sec:soundness}).
-\mytodo{iso todo: talk about each rule and change the intro to this subsection}
-\subsection{Modal Abstraction for Virtualizing Memory Locations}
+\subsection{Reasoning with Modal Abstraction for Virtual Memory Locations}
 \label{sec:modallocationvirtualization}
@@
 \begin{itemize}
-  \item \textbf{contingency}: the fact $\modalP$ abstracts the memory address mappings \textit{under} the address space rooted at $\ell$ (e.g. $\crt$)
+  \item \textbf{locality}: the fact $\modalP$ abstracts the memory address mappings \textit{under} the address space rooted \textit{at} $\ell$ (e.g. $\crt$)
   \item \textbf{hiding states}: \textit{controlled-independence} from the indirection abstracted as page-table mappings is achieved through hiding them under the virtual address space modality
-  \item \textbf{interaction with ambient logic for a local-state}: \mytodo{talk about this issue's vitality in this paper with Colin}
 \end{itemize}
 
-To concretize the understanding of these principles achieved with the modal abstraction for address-space, $\modaldef\ell\modalP$, let us first investigate the specification of the instruction ($\textsf{movctl\_instr}$) switching address spaces.
+\paragraph{Interaction with Ambient Logic for Local State} To concretize the understanding of these principles achieved with the modal abstraction for address-space, $\modaldef\ell\modalP$, let us first investigate the specification of the instruction ($\textsf{movctl\_instr}$) switching address spaces.
 Assume that our CPU keeps the current address space based on its page table register CR3, as we see in Figure \fref{fig:addressspaceswitching}.
@@
 
+\paragraph{Copying, Expanding and Shrinking}
 \todo[inline]{We don't need to talk about sexec at all.
@@
 
-\mytodo{STOP HERE}
+\paragraph{Self-Mapping Location Virtualization}

### 2022-09-28 cpp_paper.tex
write up brief discussion of framing and cr3 updates

@@
 
+\section{Framing and Change-of-Address Space}
+An important subtlety arises with supporting \lstinline|mov|s into \lstinline|%cr3|. Consider the hypothetical rule:
+\begin{mathpar}
+\inferrule[Broken]{ }{
+  \{P \ast cr3=r_1 \ast r=r_2 \ast [r_2](Q)\}
+  \texttt{mov}~\texttt{\%cr3},~r%\lstinline|mov %cr3, r| 
+  \{[r_1](P) \ast cr3=r_2 \ast r = r_2 \ast Q\}
+}
+\end{mathpar}
+This rule captures the intuitive change of address space in a Hoare triple, rather than double, form.
+The problem with this is that it interacts quite poorly with the traditional frame rule and the modal flavor of virtual points-to assertions:
+\begin{mathpar}
+  \inferrule*[right=Frame]{
+    \inferrule*[right=Broken]{ }{
+    \{\mathsf{emp} \ast cr3=r_1 \ast r=r_2 \ast [r_2](Q)\}
+    \texttt{mov}~\texttt{\%cr3},~r%\lstinline|mov %cr3, r| 
+    \{[r_1](\mathsf{emp}) \ast cr3=r_1 \ast r=r_2 \ast Q\}
+    }
+  }{
+    \{a\mapsto_\mathsf{v} x \ast \mathsf{emp} \ast cr3=r_1 \ast r=r_2 \ast [r_2](Q)\}
+    \texttt{mov}~\texttt{\%cr3},~r%\lstinline|mov %cr3, r| 
+    \{a\mapsto_\mathsf{v} x \ast [r_1](\mathsf{emp}) \ast cr3=r_1 \ast r=r_2 \ast Q\}
+  }
+\end{mathpar}
+Notice that both the precondition and postcondition assert that $a\mapsto_\mathsf{v} x$ in the current address space, but we have no basis for concluding that address translation is preserved by the change of address space.
+So this derivation clearly leads to an unsound conclusion.
+This suggestss that the traditional frame rule and the Hoare triple presentation of the change-of-address-space rule cannot soundly coexist in the same system.
+The heart of the problem is that while updating \lstinline|cr3| is \emph{physically} local, it globally changes the interpretation of virtual addresses.
+So it is simply unsound to frame around \lstinline|cr3| updates.
+
+Switching to Hoare doubles resolves this problem because an under-appreciated subtlety of Hoare doubles is that typically \emph{there is no frame rule}.
+Instead each verification essentially includes a local frame, giving each overall rule a \emph{global} (rather than local) precondition.
+For most rules this is not that important, but it does permit rules that have global effects on their preconditions.
+
+This is then how we justify our actual rule for \lstinline|cr3| updates:
+\begin{mathpar}
+\inferrule[ChangeAddressSpace]{
+  \{[r_1](P) \ast cr3=r_2 \ast r = r_2 \ast Q\}\overline{is}
+}{
+  \{P \ast cr3=r_1 \ast r=r_2 \ast [r_2](Q)\}
+  \texttt{mov}~\texttt{\%cr3},~r;\;\overline{is}
+  %\lstinline|mov %cr3, r| 
+}
+\end{mathpar}
+
 \end{appendix}

### 2022-09-28 examples.tex
fix register typo

@@
 ;; update the page table entry, mapping the page
-mov (%rcx), %rax
+mov (%r14), %rax
 ;; memset the newly-mapped page

### 2022-09-28 cpp_paper.tex
extend discussion

@@
 Switching to Hoare doubles resolves this problem because an under-appreciated subtlety of Hoare doubles is that typically \emph{there is no frame rule}.
-Instead each verification essentially includes a local frame, giving each overall rule a \emph{global} (rather than local) precondition.
+Instead each verification essentially includes a local frame that it passes to the next instructions (think continuation-passing style), giving each overall rule a \emph{global} (rather than local) precondition.
 For most rules this is not that important, but it does permit rules that have global effects on their preconditions.
@@
 \end{mathpar}
+Because the precondition on this rule is global, we avoid issues with framing.
+
+If we wanted to consider a frame rule that would work for this logic, we could consider:
+\begin{mathpar}
+  \inferrule[Cr3Frame]{
+    \{P\ast cr3=v\}\;C\;\{ Q \ast cr3=v\}
+  }{
+    \{R\ast P\ast cr3=v\}\;C\;\{ R \ast Q \ast cr3=v\}
+  }
+\end{mathpar}
+By demanding that \lstinline|cr3| be held constant (or rather, at least restored to its original value) we could frame almost traditionally.
+In particular, this rule would work with framing around calls that might lead to address space switches, such as calling blocking operations in the kernel.
+
+Readers familiar with dynamic frames~\cite{parkinson2011relationship} might find it useful to notice that a different perspective on this matter is that virtual points-to assertions are self-stable \emph{except} for changes in \lstinline|cr3|, so framing would then naturally require other means of holding \lstinline|cr3| constant (or saving and restoring it).
+Virtual points-to assertions could be made self-stable by also giving them partial ownership over \lstinline|cr3| assertions, but this would require explicitly plumbing that ownership from \emph{all} assertions back to any place an address space change might occur; this would seem to be a far graver loss of modularity than this extra quirk in framing discussions.
 

### 2022-12-02 intro.tex
some points on intro

@@
 We develop these ideas in an Iris-based\mytodo{cite JFP} separation logic, and verify simplified versions of several critical virtual-memory-related pieces of OS functionality, including mapping and unmapping pages, and describing the self-referential identity maps used in most kernels to allow convenient access to physical memory.
+Our contributions are:
+\begin{itemize}
+\item We present the first Separatoin Logic for reasoning about address-space management including mapping and unmapping pages, and describing the self-referential identity maps used in most kernels to allow convenient access to physical memory
+\item We develop a novel high-level modal abstraction for address-space management which allows us to assert certain facts whose validity are indexed by the address of a certain address-space.
+\item We define logical relations, such as introduction and elimination, to capture the logical facts about address-space operations, e.g. context-switch
+\item We discuss how our methodology is a high-level, light-weight mechanism that differ from defining contingent truth as predicate qualifier, e.g.
+\textit{Stackable} ~\cite{}, and build a mechanism, a custom-tailored weakest-precondition, to operated on these certain type predicates.
+ \end{itemize}

### 2022-12-04 intro.tex
+

@@
 This kind of context-dependent assertion, where a fact may be true in one address space but not others, has a modal flavor.
+%Despite its modal flavor, up until now, program logic have taken the contingency to be asserted as a qualified-predicate, e.g.
+\textit{Stackable} ~\cite{}, and built a mechanism, a decorated weakest-precondition, to operated on these certain type predicates.
+On the other, we, in this paper, propose
+%\begin{itemize}
+%\item {more light-weight} such that
+%  \begin{itemize}
+%  \item modal abstractions remove the burden of carrying the resource that ---
+%  \item >>>contingency as a modal abstractions
+%  \item 
+%  \item we do no
+%  \end{itemize}
+%\item {more intuitive} modal abstraction definition, together with relations defined
+%\begin{itemized}
+%  \end{itemize}
+%\end{itemize}
 We propose tackling the verification of virtual memory subsystems (and kernels more broadly) by adapting ideas from hybrid modal logic, which can label assertions true under \emph{other} circumstances (i.e., in another address space) with a modality indexed by a name for that space (in our case, the root of the page tables for an address space).
@@
 We develop these ideas in an Iris-based\mytodo{cite JFP} separation logic, and verify simplified versions of several critical virtual-memory-related pieces of OS functionality, including mapping and unmapping pages, and describing the self-referential identity maps used in most kernels to allow convenient access to physical memory.
-Our contributions are:
+Our contributions are as the following:
 \begin{itemize}
@@
 \item We define logical relations, such as introduction and elimination, to capture the logical facts about address-space operations, e.g. context-switch
-\item We discuss how our methodology is a high-level, light-weight mechanism that differ from defining contingent truth as predicate qualifier, e.g.
-\textit{Stackable} ~\cite{}, and build a mechanism, a custom-tailored weakest-precondition, to operated on these certain type predicates.
+\item We verified X,Y,Z
  \end{itemize}

### 2022-12-04 intro.tex
+

@@
 
-This kind of context-dependent assertion, where a fact may be true in one address space but not others, has a modal flavor.
+This kind of context-dependent assertion, where a fact may be true in one address space but not others, has a modal flavor%., and manifets itself at many other verification challenges besides the one, virtual-memory-management, we take as a challenge in this paper.
+For example, Monie et.al.
+~\cite{}, introduces a theory for heap management under the existence of garbage collection where  which relies on two pillars: 1.
+a predicate
 %Despite its modal flavor, up until now, program logic have taken the contingency to be asserted as a qualified-predicate, e.g.

### 2022-12-04 intro.tex
+

--- 2022-12-04 intro.tex
@@
 This kind of context-dependent assertion, where a fact may be true in one address space but not others, has a modal flavor.
+Up until now, program logics ~\cite{} has already approached some verification challenges manifesting itself in this (or similar) forms ~\cite{}.
+One of the most recent ones introduces reasoning principles for heap management under garbage collection ~\cite{}.
+In this work, a set of \textit{already iterated} memory locations (\texit{roots}) identified based on their \textit{stackability}, i.e.
+a metric determines a memory location is safe to allocate/reclaim or not, more specifically implemented as an abstraction holding mappings from memory locations to factions where a complete fraction (1) asserts the fact that the relevant memory location is safe for allocation.
+This work, in fact, distinguishes itself as \textit{"High Level"} from its predecessor fundamentally by its new logical constructions that remove the burden of keeping explicit roots of each memory location, and it is achieved through ~\cite{}
+\begin{itemize}
+\item a custom-tailored weakest-precondition modality defined tp operate with \textit{Stackable} assertion.
+\item a set of logical rules defined over this modality allows us
+  \begin{itemize}
+  \item  a light-weight reasoning in the non-existence of explicit root tracking
+  \item  an intuitive reasoning over the data structure (a set of roots), e.g. \textit{allocation} as introduction (expansion of Stackable) into its modal weakest-precondition
+  \end{itemize}
+\end{itemize}
 %Despite its modal flavor, up until now, program logic have taken the contingency to be asserted as a qualified-predicate, e.g.

### 2022-12-04 intro.tex
+

@@
 \begin{itemize}
-\item a custom-tailored weakest-precondition modality defined tp operate with \textit{Stackable} assertion.
+\item a custom-tailored weakest-precondition modality defined to operate with \textit{Stackable} assertion.
 \item a set of logical rules defined over this modality allows us
   \begin{itemize}
-  \item  a light-weight reasoning in the non-existence of explicit root tracking
-  \item  an intuitive reasoning over the data structure (a set of roots), e.g. \textit{allocation} as introduction (expansion of Stackable) into its modal weakest-precondition
+  \item  a light-weight reasoning by the replacement of explicit root tracking ~\cite{} with a set of \textit{stackable} bookkeeps already \textit{iterated} (logically \textit{framed-out} in the pre/post condition)
+  \item  an intuitive reasoning over the data structure (a set of roots), e.g.
+\textit{allocation} as introduction (expansion of \textit{Stackable} set), and binding of a variable to a value as \textit{elimination} of individual \textit{stackable} assertion for a single memory location into its modal weakest-precondition
   \end{itemize}
 \end{itemize}
+
 %Despite its modal flavor, up until now, program logic have taken the contingency to be asserted as a qualified-predicate, e.g.

### 2022-12-04 intro.tex
+

@@
   \begin{itemize}
-  \item  a light-weight reasoning by the replacement of explicit root tracking ~\cite{} with a set of \textit{stackable} bookkeeps already \textit{iterated} (logically \textit{framed-out} in the pre/post condition)
+  \item  a light-weight reasoning by the replacement of explicit root tracking ~\cite{} with a set of \textit{stackable} bookkeeps already \textit{iterated} (logically \textit{framed-out} in the pre/post condition).
   \item  an intuitive reasoning over the data structure (a set of roots), e.g.
@@
 \end{itemize}
+Despite its modal flavor, and the abstraction of \texit{already iterated stackable memory locations} set enabling a good summarization of resources to be framed out, we think that it is worth raising attention on the methodology followed for bringing the logic Madiot et.al ~\cite{} the the \textit{higher-level} version from Moine et.al.
+~\cite{} around
+\begin{itemize}
+\item {more light-weight} such that
+  \begin{itemize}
+  \item modal abstractions remove the burden of carrying the resource that ---
+  \item >>>contingency as a modal abstractions
+  \item 
+  \item we do no
+  \end{itemize}
+\item {more intuitive} modal abstraction definition, together with relations defined
+\begin{itemized}
+  \end{itemize}
+\end{itemize}
 
-%Despite its modal flavor, up until now, program logic have taken the contingency to be asserted as a qualified-predicate, e.g.
-\textit{Stackable} ~\cite{}, and built a mechanism, a decorated weakest-precondition, to operated on these certain type predicates.
-On the other, we, in this paper, propose
-%\begin{itemize}
-%\item {more light-weight} such that
-%  \begin{itemize}
-%  \item modal abstractions remove the burden of carrying the resource that ---
-%  \item >>>contingency as a modal abstractions
-%  \item 
-%  \item we do no
-%  \end{itemize}
-%\item {more intuitive} modal abstraction definition, together with relations defined
-%\begin{itemized}
-%  \end{itemize}
-%\end{itemize}
 We propose tackling the verification of virtual memory subsystems (and kernels more broadly) by adapting ideas from hybrid modal logic, which can label assertions true under \emph{other} circumstances (i.e., in another address space) with a modality indexed by a name for that space (in our case, the root of the page tables for an address space).

### 2022-12-04 intro.tex
+

@@
 One of the most recent ones introduces reasoning principles for heap management under garbage collection ~\cite{}.
-In this work, a set of \textit{already iterated} memory locations (\texit{roots}) identified based on their \textit{stackability}, i.e.
+In this work, a set of \textit{already iterated} memory locations (\textit{roots}) identified based on their \textit{stackability}, i.e.
 a metric determines a memory location is safe to allocate/reclaim or not, more specifically implemented as an abstraction holding mappings from memory locations to factions where a complete fraction (1) asserts the fact that the relevant memory location is safe for allocation.
@@
 \end{itemize}
-Despite its modal flavor, and the abstraction of \texit{already iterated stackable memory locations} set enabling a good summarization of resources to be framed out, we think that it is worth raising attention on the methodology followed for bringing the logic Madiot et.al ~\cite{} the the \textit{higher-level} version from Moine et.al.
-~\cite{} around
-\begin{itemize}
-\item {more light-weight} such that
-  \begin{itemize}
-  \item modal abstractions remove the burden of carrying the resource that ---
-  \item >>>contingency as a modal abstractions
-  \item 
-  \item we do no
-  \end{itemize}
-\item {more intuitive} modal abstraction definition, together with relations defined
-\begin{itemized}
-  \end{itemize}
-\end{itemize}
+Despite its modal flavor, and the abstraction of \textit{already iterated stackable memory locations} set enabling a good summarization of resources to be framed out, we think that it is worth raising attention on the methodology followed for bringing the logic Madiot et.al ~\cite{} the the \textit{higher-level} version from Moine et.al.
+~\cite{} around the following issues.
+\begin{remark}[Hiding What is Stackable in a Modal Context?] 
+\end{remark}
+\begin{remark}[Custom-Tailored Modal Abstractions for a Each Stackable Data-Structure]
+\end{remark}
+\begin{remark}[Not only Being Stackable But How?]
+\end{remark}
 

### 2022-12-05 intro.tex
forgot to remove the intro change

@@
 This kind of context-dependent assertion, where a fact may be true in one address space but not others, has a modal flavor. 
-Up until now, program logics ~\cite{} has already approached some verification challenges manifesting itself in this (or similar) forms ~\cite{}.
-One of the most recent ones introduces reasoning principles for heap management under garbage collection ~\cite{}.
-In this work, a set of \textit{already iterated} memory locations (\textit{roots}) identified based on their \textit{stackability}, i.e.
-a metric determines a memory location is safe to allocate/reclaim or not, more specifically implemented as an abstraction holding mappings from memory locations to factions where a complete fraction (1) asserts the fact that the relevant memory location is safe for allocation.
-This work, in fact, distinguishes itself as \textit{"High Level"} from its predecessor fundamentally by its new logical constructions that remove the burden of keeping explicit roots of each memory location, and it is achieved through ~\cite{}
-\begin{itemize}
-\item a custom-tailored weakest-precondition modality defined to operate with \textit{Stackable} assertion.
-\item a set of logical rules defined over this modality allows us
-  \begin{itemize}
-  \item  a light-weight reasoning by the replacement of explicit root tracking ~\cite{} with a set of \textit{stackable} bookkeeps already \textit{iterated} (logically \textit{framed-out} in the pre/post condition).
-  \item  an intuitive reasoning over the data structure (a set of roots), e.g.
-\textit{allocation} as introduction (expansion of \textit{Stackable} set), and binding of a variable to a value as \textit{elimination} of individual \textit{stackable} assertion for a single memory location into its modal weakest-precondition
-  \end{itemize}
-\end{itemize}
-Despite its modal flavor, and the abstraction of \textit{already iterated stackable memory locations} set enabling a good summarization of resources to be framed out, we think that it is worth raising attention on the methodology followed for bringing the logic Madiot et.al ~\cite{} the the \textit{higher-level} version from Moine et.al.
-~\cite{} around the following issues.
-\begin{remark}[Hiding What is Stackable in a Modal Context?] 
-\end{remark}
-\begin{remark}[Custom-Tailored Modal Abstractions for a Each Stackable Data-Structure]
-\end{remark}
-\begin{remark}[Not only Being Stackable But How?]
-\end{remark}
-
 We propose tackling the verification of virtual memory subsystems (and kernels more broadly) by adapting ideas from hybrid modal logic, which can label assertions true under \emph{other} circumstances (i.e., in another address space) with a modality indexed by a name for that space (in our case, the root of the page tables for an address space).
@@
 We develop these ideas in an Iris-based\mytodo{cite JFP} separation logic, and verify simplified versions of several critical virtual-memory-related pieces of OS functionality, including mapping and unmapping pages, and describing the self-referential identity maps used in most kernels to allow convenient access to physical memory.
-Our contributions are as the following:
-\begin{itemize}
-\item We present the first Separatoin Logic for reasoning about address-space management including mapping and unmapping pages, and describing the self-referential identity maps used in most kernels to allow convenient access to physical memory
-\item We develop a novel high-level modal abstraction for address-space management which allows us to assert certain facts whose validity are indexed by the address of a certain address-space.
-\item We define logical relations, such as introduction and elimination, to capture the logical facts about address-space operations, e.g. context-switch
-\item We verified X,Y,Z
- \end{itemize}

### 2023-02-06 programlogic.tex
+

@@
 \end{figure*}
-\paragraph{Coalescing and Coexistence} \mytodo{Super Huge pages}
-\todo{iso todo: stress fractions and aligment}
-\subsection{Reasoning with Modal Abstraction for Virtual Memory Locations}
-\label{sec:modallocationvirtualization}
-\input{fig-laws}
-The most essential part of our reasoning principles is shaped around abstractions for location virtualization, more concretely virtual memory points-to relation in Figure \fref{fig:virtualpointsto}.
-One might have already noticed that the resource represented by the virtual-points to is not only a simple immediately mapped singly resource but also
-\begin{itemize}
-\item a tree-like structured resrouce rooted at the certain physical memory address ($\crt$) -- not a single resource mapping abstraction 
-\item the shared indirection levels abstracted as mappings with fractional permissions
-\item through partially owned set of resource mappings grants the eventual uniquely ownable single resource, i.e. page address 
-\end{itemize}
-which we think can be well-summarized with the modal abstraction shown in \RULE{ModalLocVirt} which allows us to have
-\begin{itemize}
-  \item \textbf{locality}: the fact $\modalP$ abstracts the memory address mappings \textit{under} the address space rooted \textit{at} $\ell$ (e.g. $\crt$)
-  \item \textbf{hiding states}: \textit{controlled-independence} from the indirection abstracted as page-table mappings is achieved through hiding them under the virtual address space modality
-\end{itemize}
+\section{Considering Hoare Doubles in the Context of Switching Address-Space}
+An important subtlety arises with supporting \lstinline|mov|s into \lstinline|%cr3|. Consider the hypothetical rule:
+\begin{mathpar}
+\inferrule[Broken]{ }{
+  \{P \ast cr3=r_1 \ast r=r_2 \ast [r_2](Q)\}
+  \texttt{mov}~\texttt{\%cr3},~r%\lstinline|mov %cr3, r| 
+  \{[r_1](P) \ast cr3=r_2 \ast r = r_2 \ast Q\}
+}
+\end{mathpar}
+This rule captures the intuitive change of address space in a Hoare triple, rather than double, form.
+The problem with this is that it interacts quite poorly with the traditional frame rule and the modal flavor of virtual points-to assertions:
+\begin{mathpar}
+  \inferrule*[right=Frame]{
+    \inferrule*[right=Broken]{ }{
+    \{\mathsf{emp} \ast cr3=r_1 \ast r=r_2 \ast [r_2](Q)\}
+    \texttt{mov}~\texttt{\%cr3},~r%\lstinline|mov %cr3, r| 
+    \{[r_1](\mathsf{emp}) \ast cr3=r_1 \ast r=r_2 \ast Q\}
+    }
+  }{
+    \{a\mapsto_\mathsf{v} x \ast \mathsf{emp} \ast cr3=r_1 \ast r=r_2 \ast [r_2](Q)\}
+    \texttt{mov}~\texttt{\%cr3},~r%\lstinline|mov %cr3, r| 
+    \{a\mapsto_\mathsf{v} x \ast [r_1](\mathsf{emp}) \ast cr3=r_1 \ast r=r_2 \ast Q\}
+  }
+\end{mathpar}
+Notice that both the precondition and postcondition assert that $a\mapsto_\mathsf{v} x$ in the current address space, but we have no basis for concluding that address translation is preserved by the change of address space.
+So this derivation clearly leads to an unsound conclusion.
+This suggestss that the traditional frame rule and the Hoare triple presentation of the change-of-address-space rule cannot soundly coexist in the same system.
+The heart of the problem is that while updating \lstinline|cr3| is \emph{physically} local, it globally changes the interpretation of virtual addresses.
+So it is simply unsound to frame around \lstinline|cr3| updates.
 
-\paragraph{Interaction with Ambient Logic for Local State} To concretize the understanding of these principles achieved with the modal abstraction for address-space, $\modaldef\ell\modalP$, let us first investigate the specification of the instruction ($\textsf{movctl\_instr}$) switching address spaces.
-Assume that our CPU keeps the current address space based on its page table register CR3, as we see in Figure \fref{fig:addressspaceswitching}.
-\input{fig-ctlreasoning}
-Thinking of $\modalP$ and $\modalQ$ as resources representing the indirection mappings ($\mapsto_{\kw{p}}$) in the address spaces rooted at $\rvsrc$ and $\rvdst$ respectively, the rule interacts with the ambient logic, \SL, via pulls local state out of the address-space modality rooted at $\rvdst$, and make it the observable view of the memory.
+Switching to Hoare doubles resolves this problem because an under-appreciated subtlety of Hoare doubles is that typically \emph{there is no frame rule}.
+Instead each verification essentially includes a local frame that it passes to the next instructions (think continuation-passing style), giving each overall rule a \emph{global} (rather than local) precondition.
+For most rules this is not that important, but it does permit rules that have global effects on their preconditions.
 
-\paragraph{Copying, Expanding and Shrinking}
-\todo[inline]{We don't need to talk about sexec at all.
-Let's also talk tomorrow about adjusting the syntax in the paper to mimic actual textual assembly more closely (e.g., all of the instructions are mov, with different argument shapes)}
+This is then how we justify our actual rule for \lstinline|cr3| updates:
+\begin{mathpar}
+\inferrule[ChangeAddressSpace]{
+  \{[r_1](P) \ast cr3=r_2 \ast r = r_2 \ast Q\}\overline{is}
+}{
+  \{P \ast cr3=r_1 \ast r=r_2 \ast [r_2](Q)\}
+  \texttt{mov}~\texttt{\%cr3},~r;\;\overline{is}
+  %\lstinline|mov %cr3, r| 
+}
+\end{mathpar}
+Because the precondition on this rule is global, we avoid issues with framing.
 
-\paragraph{Self-Mapping Location Virtualization}
+If we wanted to consider a frame rule that would work for this logic, we could consider:
+\begin{mathpar}
+  \inferrule[Cr3Frame]{
+    \{P\ast cr3=v\}\;C\;\{ Q \ast cr3=v\}
+  }{
+    \{R\ast P\ast cr3=v\}\;C\;\{ R \ast Q \ast cr3=v\}
+  }
+\end{mathpar}
+By demanding that \lstinline|cr3| be held constant (or rather, at least restored to its original value) we could frame almost traditionally.
+In particular, this rule would work with framing around calls that might lead to address space switches, such as calling blocking operations in the kernel.
+
+Readers familiar with dynamic frames~\cite{parkinson2011relationship} might find it useful to notice that a different perspective on this matter is that virtual points-to assertions are self-stable \emph{except} for changes in \lstinline|cr3|, so framing would then naturally require other means of holding \lstinline|cr3| constant (or saving and restoring it).
+Virtual points-to assertions could be made self-stable by also giving them partial ownership over \lstinline|cr3| assertions, but this would require explicitly plumbing that ownership from \emph{all} assertions back to any place an address space change might occur; this would seem to be a far graver loss of modularity than this extra quirk in framing discussions.

### 2023-02-06 programlogic.tex
removed the old stuff

@@
 The principles shown here are mainly shaped around understanding and expressing the fact related the virtual address space, most specifically virtualization and translation of memory locations .
+\subsection{State}
+\label{sec:state}
+\todo[inline]{I don't think we need or want to describe the operational semantics in detail in the submission.
+It adds a lot of formalization that doen'st necessarily contribute to reader understanding.
+We should describe the runtime \emph{state} $\sigma$ of memory and registers, since that is directly used in interpreting assertions.
+But the reduction rules will mostly add mental overhead here.}
+We represent the machine state mainly as a finite map of registers to register values and a map of masked physical memory addresses to 64-bit physical memory values.
+Together with default def\textsf{CPU} instantiation, we end up having our state with following pieces:
+\begin{itemize}
+\item A constant cpu initialized with default value: $\sigma.\mathcal{C}$
+\item A register map: $\sigma.\mathcal{R}: \kw{greg} \rightarrow_{\textrm{fin}} \kw{regval} $
+\item A memory map: $\sigma.\mathcal{M}: \Locft \rightharpoonup_{\textrm{fin}} (\Loctw \rightharpoonup_{\textrm{fin}} \Locsf )$
+\end{itemize}
+\subsection{Memory Addresses \& Naming Registers}
+\label{sec:}
+As one might have already anticipated from the syntax we introduce, we do not bind any value of an evaluated expression.
+All the indices and accessed values are treated as globally referenced.
+In align with this design choice, our expression is a stream of instructions, which is not evaluated to a value to be bound, but changes the machine state through the indices (e.g.
+$\kw{r}\in\kw{greg}$) -- to the global maps.
+
+\subsection{Machine State under Address Translation}
+\label{sec:selectedinstrsemantics}
+Althought we give the complete set of operational semantics rule in \sref{appendix:movops}, It is worth building the intuition on the the way some of these rules bahave in the context of address translation.
+In fact $\readlval\maddr{\storememstar\crval}\locsf$ is unfoled into multiple physical memory lookups for the final page address retrieval -- i.e.
+address translation traversal as shown in Figure \ref{fig:pagetables}:
+\begin{itemize}
+\item top-level-address translation: with the given root address (64-bit $\crval$) of the address space, performs the address translation, handling the first level (to get the PML4 entry) itself.
+The next level table address is computed with the fetched PML4 offset value which exhibits itself as 9-bit offset in $\kw{maddr}$.
+\mytodo{iso: put move}
+\item translating from PML4 entry: performs the second level of address translation, to retrieve starting at the PML4 table entry, and interprets the PML4 entry that references a Page-Directory-Pointer Table (PDPT).
+We obtain the PDPT offset which exhibits itself as 9 bit offset in $\kw{maddr}$ to obtain the address of the next level page directory table (PDT) \mytodo{iso: put move}
+\item translating from PD entry: performs the third level of address translation, to retrieve starting at the PDP table entry, and interprets a PDPTE that references a PD table.
+Likewise, we obtain the PD offset which exhibits itself as 9 bit offset in $\kw{maddr}$ to obtain the address of the next level page directory table (PT) \mytodo{iso:put move }
+\item translate from PT entry: performs final level of address translation, starting from the PT entry, with a given 12 bit page offset, we can compute the physical address referencing $\locsf$ \mytodo{ismo: put mov}
+\end{itemize}
+
 \subsection{Points-To Assertions}
@@
 \end{figure*}
-\section{Considering Hoare Doubles in the Context of Switching Address-Space}
+\subsection{Considering Hoare Doubles in the Context of Switching Address-Space}
 An important subtlety arises with supporting \lstinline|mov|s into \lstinline|%cr3|. Consider the hypothetical rule:

### 2023-02-10 programlogic.tex
+

--- 2023-02-06 programlogic.tex
@@
 \label{sec:logic}
-In this section, we describe our high-level reasoning principles for location virtualization whose detailed justification is shown in \sref{sec:soundness}.
-The principles shown here are mainly shaped around understanding and expressing the fact related the virtual address space, most specifically virtualization and translation of memory locations .
+%\subsection{Machine State under Address Translation}
+%\label{sec:selectedinstrsemantics}
+%Althought we give the complete set of operational semantics rule in \sref{appendix:movops}, It is worth building the intuition on the the way some of these rules bahave in the context of address translation.
+In fact $\readlval\maddr{\storememstar\crval}\locsf$ is unfoled into multiple physical memory lookups for the final page address retrieval -- i.e.
+address translation traversal as shown in Figure \ref{fig:pagetables}:
+%\begin{itemize}
+%\item top-level-address translation: with the given root address (64-bit $\crval$) of the address space, performs the address translation, handling the first level (to get the PML4 entry) itself.
+The next level table address is computed with the fetched PML4 offset value which exhibits itself as 9-bit offset in $\kw{maddr}$.
+\mytodo{iso: put move}
+%\item translating from PML4 entry: performs the second level of address translation, to retrieve starting at the PML4 table entry, and interprets the PML4 entry that references a Page-Directory-Pointer Table (PDPT).
+We obtain the PDPT offset which exhibits itself as 9 bit offset in $\kw{maddr}$ to obtain the address of the next level page directory table (PDT) \mytodo{iso: put move}
+%\item translating from PD entry: performs the third level of address translation, to retrieve starting at the PDP table entry, and interprets a PDPTE that references a PD table.
+Likewise, we obtain the PD offset which exhibits itself as 9 bit offset in $\kw{maddr}$ to obtain the address of the next level page directory table (PT) \mytodo{iso:put move }
+%\item translate from PT entry: performs final level of address translation, starting from the PT entry, with a given 12 bit page offset, we can compute the physical address referencing $\locsf$ \mytodo{ismo: put mov}
+%\end{itemize}
 \subsection{Points-To Assertions}

### 2023-02-10 syntaxsemantics.tex
+

\section{Machine State, Instructions, Naming Registers and Memory Translation}
\label{sec:syntax}
To develop our core logical ideas and explain their intepretation, we need only a handful of instructions, and a machine model corresponding to execution of x86-64 assembly instructions with virtual memory enabled on the CPU.

\subsection{Registers and Memory}
Programs we demonstrate in this paper requires accessing two types of computer resource: registers and memory. A register identifier, $\reg$, is chosen from a fixed finite set of register identifiers,$\regset$. We use these identifiers to access the register values, $\regval \in \regvaltype$.

\input{fig-syntax}

Unlike registers, we do not abstract the memory indices as a special type but instead, for the sake of clarity and ease of representation, we show differently masked machine words, $\loc \in \Loc$, with the subscripts showing the length of a 64-bit machine word after masking, e.g. $\kw{w}_{12}$ is a 12-bit resource which can be obtained after masking 52-bit of a 64-bit word.
This simple syntax in \ref{fig:syntax} includes a simple syntax for values because our language is purely \textit{imperative}, i.e. an instruction does not return a value. Intentionally, we restrict any stream of instruction, $\instrs$, to eventually reduce to $\iskip$ instruction. Our syntax capture the no-op ``$\iskip$'' and the sequencing construct ``$\iseq{\instr}{\instrs}$'' in their standard forms.
\todo[inline]{Ismail: this paper is not about the binding of the machine model to Iris. Low-level details like ``injected into unit'' are not really important unless someone asks.}
\subsection{State}
\label{sec:state}
\todo[inline]{I don't think we need or want to describe the operational semantics in detail in the submission. It adds a lot of formalization that doen'st necessarily contribute to reader understanding. We should describe the runtime \emph{state} $\sigma$ of memory and registers, since that is directly used in interpreting assertions. But the reduction rules will mostly add mental overhead here.}
We represent the machine state mainly as a finite map of registers to register values and a map of masked physical memory addresses to 64-bit physical memory values. Together with default def\textsf{CPU} instantiation, we end up having our state with following pieces:
\begin{itemize}
\item A constant cpu initialized with default value: $\sigma.\mathcal{C}$
\item A register map: $\sigma.\mathcal{R}: \kw{greg} \rightarrow_{\textrm{fin}} \kw{regval} $
\item A memory map: $\sigma.\mathcal{M}: \Locft \rightharpoonup_{\textrm{fin}} (\Loctw \rightharpoonup_{\textrm{fin}} \Locsf )$
\end{itemize}
As one might have already anticipated from the syntax we introduce, we do not bind any value of an evaluated expression. All the indices and accessed values are treated as globally referenced. In align with this design choice, our expression is a stream of instructions, which is not evaluated to a value to be bound, but changes the machine state through the indices (e.g. $\kw{r}\in\kw{greg}$) -- to the global maps. 
%
%\subsection{Machine State under Address Translation}
%\label{sec:selectedinstrsemantics}
%Althought we give the complete set of operational semantics rule in \sref{appendix:movops}, It is worth building the intuition on the the way some of these rules bahave in the context of address translation. In fact $\readlval\maddr{\storememstar\crval}\locsf$ is unfoled into multiple physical memory lookups for the final page address retrieval -- i.e. address translation traversal as shown in Figure \ref{fig:pagetables}:
%\begin{itemize}
%\item top-level-address translation: with the given root address (64-bit $\crval$) of the address space, performs the address translation, handling the first level (to get the PML4 entry) itself. The next level table address is computed with the fetched PML4 offset value which exhibits itself as 9-bit offset in $\kw{maddr}$. \mytodo{iso: put move}
%\item translating from PML4 entry: performs the second level of address translation, to retrieve starting at the PML4 table entry, and interprets the PML4 entry that references a Page-Directory-Pointer Table (PDPT). We obtain the PDPT offset which exhibits itself as 9 bit offset in $\kw{maddr}$ to obtain the address of the next level page directory table (PDT) \mytodo{iso: put move}
%\item translating from PD entry: performs the third level of address translation, to retrieve starting at the PDP table entry, and interprets a PDPTE that references a PD table. Likewise, we obtain the PD offset which exhibits itself as 9 bit offset in $\kw{maddr}$ to obtain the address of the next level page directory table (PT) \mytodo{iso:put move }
%\item translate from PT entry: performs final level of address translation, starting from the PT entry, with a given 12 bit page offset, we can compute the physical address referencing $\locsf$ \mytodo{ismo: put mov}
%\end{itemize}

### 2023-02-20 programlogic.tex
+

@@
 \label{sec:logic}
+Our reasoning principles are shaped around understanding the following issues exhibiting themselves withing the context of vritual-memory-management:
+\begin{enumerate}
+\item abstracting the relation between memory (virtual) addresses and the physical page addresses backing these virtual addresses
+\item regarding the address-space management, we propose reasoning principles for
+  \begin{itemize}
+  \item aliasing/sharing Physical Pages:
+  \item abstracting address-space as a modality
+    \begin{itemize}
+    \item modal resource context:
+    \item interaction:
+    \end{itemize}
+    \item Hoare-Doubles: 
+  \end{itemize}
+\end{enumerate}
 %\subsection{Machine State under Address Translation}
@@
 \label{sec:pointsto}
-
-We present three main points-to relations as part of utilization of separation logic \ref{} as an ambient logic:
+As part of assertions on our machine state ($\sigma$), we abstract memory addresses and registers naming their relevant values through well-known separation logic assertion \textit{points-to}.
 \begin{enumerate}
@@
 \item Register points-to, $\pfpointsto\rg\rv\qfrac\rpts$
-\item Virtual memory address points-to, $\pfpointsto\vaddr\locsf\qfrac\vpts$
 \end{enumerate}
@@
 \begin{array}{cl}
-\pfpointsto\locsf\locsf\qfrac\ppts \stackrel{def}{=} & \nfpointsto{\mask\locsf\tw}{\mask\locsf\ft}\locsf\qfrac\naddr
+\pfpointsto\locsf\locsf\qfrac\ppts \stackrel{def}{=} & \nfpointsto{\mask\locsf\ft}{\mask\locsf\tw}\locsf\qfrac\naddr
 \end{array}
@@
 \end{figure}
-where we simply abstract the generation of hierarchical mapping of a physical address ($\locsf$) through the different masks of the address ($\locsf|^{12}$ and $\locsf|^{52}$) as nested map resources.
+where we simply abstract the generation of hierarchical mapping of a physical address ($\locsf$) through two different levels of masking:
+
+\begin{itemize}
+\item  After a series of masking imposed on the virtual memory address (e.g. \textsf{vaddr} 64bits address [0-64]) :
+  \begin{itemize}
+  \item Page Map Level-4 Translation (L4): Calculating the offset to the 4 \textsf{Kbyte Physical Page} table to obtain the  physical address by  which resides in the last 12 bits of the virtual memory address, $\locsf|^{12}$,  shown in Figure \ref{}.
+  \item Page Directory Pointer Level Translation (L3):
+  \item Page Directory Table Level Translation (L2):  we obtain the address of the page table via \textsf{PDE}$|^{52}$ and picking the offset from the virtual memory address via imposing the mask to obtain page table offset bits (9 bits [20-12] shown in Figure \ref{}) to obtain the page table entry (\textsf{PTE}).
+  \item Page Table Level Translation (L1): 
+  \end{itemize}
+ \item then after obtainini the physical page address, we obtain the value 
+\end{itemize}
+
 \paragraph{Virtual-Memory points-to} As we see in Figure \fref{fig:}, virtual address mapping is obtained via traversal of page table, i.e.

### 2023-02-20 programlogic.tex
+

@@
   \item Page Map Level-4 Translation (L4): Calculating the offset to the 4 \textsf{Kbyte Physical Page} table to obtain the  physical address by  which resides in the last 12 bits of the virtual memory address, $\locsf|^{12}$,  shown in Figure \ref{}.
+    \[  \pfpointsto{\lvlsum\crt\lfoff}{\lvlbor\lt}\qfracfotsss\ppts \star \]
   \item Page Directory Pointer Level Translation (L3):
+    \[ \pfpointsto{\lvlsum\lt\ltoff}{\lvlbor\ltw}\qfracfotss\ppts \star \]
   \item Page Directory Table Level Translation (L2):  we obtain the address of the page table via \textsf{PDE}$|^{52}$ and picking the offset from the virtual memory address via imposing the mask to obtain page table offset bits (9 bits [20-12] shown in Figure \ref{}) to obtain the page table entry (\textsf{PTE}).
+    \[  \pfpointsto{\lvlsum\ltw\ltwoff}{\lvlbor\lo}\qfracfots\ppts \star \]
   \item Page Table Level Translation (L1):
+    \[  \pfpointsto{\lvlsum\lo\looff}{\lvlbor\pg}\qfracfot\ppts \star \]
   \end{itemize}
 \item then after obtainini the physical page address, we obtain the value
+  \[  \ppointsto{\pageptstosum\pg\pgoff}\locsf\ppts \]
 \end{itemize}
 
-\paragraph{Virtual-Memory points-to} As we see in Figure \fref{fig:}, virtual address mapping is obtained via traversal of page table, i.e.
+\subsection{Aliasing/Sharing Physical Pages} As we see in Figure \fref{fig:}, virtual address mapping is obtained via traversal of page table, i.e.
 indirection provided by physical memory lookups.
 As expected, we define the virtual-points to relation, ($\pfpointsto\vaddr\locsf\qfrac\vpts$), in terms of multiple physical memory mappings representing the indirection shown in Figure \fref{fig:}
-\begin{figure*}
-\[
-\begin{array}{l}
-  \ppointsto\vaddr\locsf\vpts \stackrel{def}{=} \lambda \tlf\Loc.\\
-  \exists_{\tlfoff\Locn \;, \tltoff\Locn \;, \tltwoff\Locn \;,\tlooff\Locn \;, \tpgoff\Loctw \;, \tlt\Loc \;, \tltw\Loc \;, \tlo\Loc \tpg\Loc} \ldotp \\
-  \ulcorner \textsf{aligned } \vaddr \urcorner \star 
-   \ulcorner \vaddr = \locsx :: \lfoff :: \ltoff :: 
-   \ltwoff :: \looff :: \pgoff \urcorner \star\\
-  \pfpointsto{\lvlsum\crt\lfoff}{\lvlbor\lt}\qfracfotsss\ppts \star 
-  \pfpointsto{\lvlsum\lt\ltoff}{\lvlbor\ltw}\qfracfotss\ppts \star \\
-  \pfpointsto{\lvlsum\ltw\ltwoff}{\lvlbor\lo}\qfracfots\ppts \star 
-  \pfpointsto{\lvlsum\lo\looff}{\lvlbor\pg}\qfracfot\ppts \star \\
-  \ppointsto{\pageptstosum\pg\pgoff}\locsf\ppts 
-\end{array}
-\]
-\caption{Virtual Points-to Relation}
-\todo[inline]{These fractions aren't quite right (though I see you added the variance between levels), I can walk you through in our meeting tomorrow.}
-  \label{fig:virtualpointsto}
-\end{figure*}
+
+%\begin{figure*}
+%\[
+%\begin{array}{l}
+%  \ppointsto\vaddr\locsf\vpts \stackrel{def}{=} \lambda \tlf\Loc.\\
+%  \exists_{\tlfoff\Locn \;, \tltoff\Locn \;, \tltwoff\Locn \;,\tlooff\Locn \;, \tpgoff\Loctw \;, \tlt\Loc \;, \tltw\Loc \;, \tlo\Loc \tpg\Loc} \ldotp \\
+%  \ulcorner \textsf{aligned } \vaddr \urcorner \star 
+%   \ulcorner \vaddr = \locsx :: \lfoff :: \ltoff :: 
+%   \ltwoff :: \looff :: \pgoff \urcorner \star\\
+%  \pfpointsto{\lvlsum\crt\lfoff}{\lvlbor\lt}\qfracfotsss\ppts \star 
+%  \pfpointsto{\lvlsum\lt\ltoff}{\lvlbor\ltw}\qfracfotss\ppts \star \\
+%  \pfpointsto{\lvlsum\ltw\ltwoff}{\lvlbor\lo}\qfracfots\ppts \star 
+%  \pfpointsto{\lvlsum\lo\looff}{\lvlbor\pg}\qfracfot\ppts \star \\
+%  \ppointsto{\pageptstosum\pg\pgoff}\locsf\ppts 
+%\end{array}
+%\]
+%\caption{Virtual Points-to Relation}
+%\todo[inline]{These fractions aren't quite right (though I see you added the variance between levels), I can walk you through in our meeting tomorrow.}
+%  \label{fig:virtualpointsto}
+%\end{figure*}
+\subsection{Address-Spaces as a Modality}
+\subsubsection{Modal Resource Context}
+\subsubsection{Interaction: address-space switch as a modal-entailment}
+
 \subsection{Considering Hoare Doubles in the Context of Switching Address-Space}

### 2023-02-20 soundness.tex
+

@@
 \section{Implementing Logical Machinery \& Soundness}
-We instantiate our program logic, $[@]$, as an instantiation of Iris~\cite{iris}, and build our modal abstractions on top of it. 
-\subsection{BI-Lifting for Address Space Modality}
-\label{sec:bilifting}
+We build our program logic, as an instantiation of Iris~\cite{iris}, and build our modal abstractions on top of it. 
 \mytodo{we should put some stuff here}
@@
 \label{sec:soundness}
-Our logic, $[@]$, operates on the machine state, which means we do not need to augment the machine state, i.e.
-our logical state is exact logical representation of our machine state.
-Consequently, we do not require an invariant, $\mathcal{I}$, except from the fact that logical/physical associated maps representing the register/memory mappings have the same domain.
-This ensures the correct lifting of mappings in the machine state to the \emph{assertions} that the client of our logic uses, i.e.
+Our logic, operates on the machine state, which means we do not need to augment the machine state.
+The invarian that we pick, \textit{central invariant} ($\mathcal{I}$), is just semantic interpretation of stores in the machine state, $\sigma$.R and $\sigma$.M.
+This semantic interpratation ensures the correct lifting of mappings in the machine state to the assertions that the client of our logic uses, i.e.
 points-to assertions that are defined as the ownership of a fragment of the logical state.
 
-Iris instantiation comes with the semantic definition for the weakest-precondition which we can refactor into Hoare triples, $\triple\pre\instr\post$, to specify actions using our reasoning principles, show that these triples are sound.
+We prefer to skip explaining the steps used in instantiation of Iris because it is an almost standard procedure, has already been explained for many other logic \ref{}, and we are concerned with the page-count limitation.
+However, it is worh noting that once you instantiate Iris for your language, it comes with the semantic definition for the weakest-precondition which we can refactor into Hoare triples, $\triple\pre\instr\post$, to specify our  selected \textsf{AMD64} instructions, and show that these triples are sound.
 
@@
   \label{th:adequacy}
-  Assuming that  Suppose $\triple{\textsf{valid\_init}}\instr\iTrue$ holds.
+  Assuming that $\triple{\textsf{valid\_init}}\instr\iTrue$ holds,
  The execution of the instruction~$\instrs$, beginning in an initial state, cannot result in a configuration where a thread is stuck.
@@
 the reasoning rules of our logic, then the program~$\instrs$ is safe.
+\subsection{BI-Lifting for Address Space Modality}
+\label{sec:bilifting}
 
@@
 \label{sec:invariant}
-
-Since our semantics operates right on top of the x86-64 machine state (with Iris terminology, physical store),
-we do not need to augment the physical store for more abstract reasoning, consequently, we do not need to come up
-with a complicated logical relations that ties the physical store to the
-\logical store. 
-\begin{definition}[Relation between Physical and \Logical Stores]
-  \label{def:related}
-  Since we pick the logical store exactly same with physical store,  
-  two stores are \emph{related}, identical.
-\end{definition}
 
@@
   where it is applied to the \logical store~$\store$.
-\item unlike the points-to relation obtained by direct interpretion of $\genheapinterp\store.\mathcal{R}$ using Iris existing \textsf{gen\_heap\_interp},
-  we define a new algebra (\textsf{gen\_mem\_UR}) for abstracting the nested maps due to different levels of masking in memory mappings. 
+\item unlike the points-to relation obtained by direct interpretion of $\genheapinterp\store.\mathcal{R}$ using Iris, the existing \textsf{gen\_heap\_interp},
+  does not directly helps introducing the ghost we define a new algebra (\textsf{gen\_mem\_UR}) for abstracting the nested maps due to different levels of masking in memory mappings. 
 \end{enumerate}
@@
 \begin{array}{l@{\quad\star\quad}l@{\quad}l}
-  \genheapinterp\store & \star \\
+  \genheapinterp\store &  \\
    \genmemheapinterp\store

### 2023-02-20 soundness.tex
+

@@
 \label{def:soundness:statement}
-The operational semantics of our simple lang just executes the instruction of our x86-64 model.
-Therefore, our soundness argument is to show that any execution composed of actions in our machine model (some of which are shown in Figure \sref{sec:semantics}) does not end-up in a invalid state.guarantees that  (\sref{sec:valid}).
-A~program that attempts to ...
+The operational semantics of our simple lang just executes sequences of instructions in our x86-64 model.
+Therefore, our soundness argument is to show that any execution composed of instructions in our machine model (some of which are shown in Figure \sref{sec:semantics}) does not end-up in a invalid state.
+In other words, a~program that attempts to ...
 stuck~(\sref{sec:instrsemantics}).
@@
   \label{th:adequacy}
-  Assuming that $\triple{\textsf{valid\_init}}\instr\iTrue$ holds,
- The execution of the instruction~$\instrs$, beginning in an initial state, cannot result in a configuration where a thread is stuck.
+ Together with the assumptions on the initial state hold,
+ the execution of the instruction~$\instrs$, beginning with this initial state, cannot result in a configuration where the execution is stuck.
 \end{theorem}
@@
 triple, then this program cannot crash: by a direct consequence of Iris's adequacy theorem~\cite[\S6.4]{iris}.
-
+\todo{ismail, give the real adeq. theorem here}
 Moreover we need to show the validity of each rules in Figures \fref{fig:reasoning} and \fref{fig:laws}.

### 2023-02-21 programlogic.tex
+

--- 2023-02-20 programlogic.tex
@@
   \item Page Map Level-4 Translation (L4): Calculating the offset to the 4 \textsf{Kbyte Physical Page} table to obtain the  physical address by  which resides in the last 12 bits of the virtual memory address, $\locsf|^{12}$,  shown in Figure \ref{}.
+    \[  \pfpointsto{\lvlsum\crt\lfoff}{\lvlbor\lt}\qfracfotsss\ppts \star \]
   \item Page Directory Pointer Level Translation (L3):
+    \[ \pfpointsto{\lvlsum\lt\ltoff}{\lvlbor\ltw}\qfracfotss\ppts \star \]
   \item Page Directory Table Level Translation (L2):  we obtain the address of the page table via \textsf{PDE}$|^{52}$ and picking the offset from the virtual memory address via imposing the mask to obtain page table offset bits (9 bits [20-12] shown in Figure \ref{}) to obtain the page table entry (\textsf{PTE}).
+    \[  \pfpointsto{\lvlsum\ltw\ltwoff}{\lvlbor\lo}\qfracfots\ppts \star \]
   \item Page Table Level Translation (L1):
+    \[  \pfpointsto{\lvlsum\lo\looff}{\lvlbor\pg}\qfracfot\ppts \star \]
   \end{itemize}
 \item then after obtainini the physical page address, we obtain the value
+  \[  \ppointsto{\pageptstosum\pg\pgoff}\locsf\ppts \]
 \end{itemize}
 
-\paragraph{Virtual-Memory points-to} As we see in Figure \fref{fig:}, virtual address mapping is obtained via traversal of page table, i.e.
-indirection provided by physical memory lookups.
-As expected, we define the virtual-points to relation, ($\pfpointsto\vaddr\locsf\qfrac\vpts$), in terms of multiple physical memory mappings representing the indirection shown in Figure \fref{fig:}
+\subsection{Aliasing/Sharing Physical Pages}
+As we see in Figure \fref{fig:}, virtual address mapping is obtained simply in two fold: 1.
+traversing the page tables to reach the page table entry 2.
+fetch the physical address and the value it is adressing (Section \ref{fig:physicalpointsto}).
+This two fold access pattern enabling sharing of pages by virtual memory addresses exhibits logically sound representation against virtual memory operations some of which ((e.g.
+mapping a new page) we explain in detail in Section \ref{}.
+
+The very first assurance our logical constructions need to be able assert is the \textit{existence} of the page table walk.
+A naive attempt to assert the existence of the page-walk for a certain virtual memory translation would look like the one in Figure \ref{figure:strongvirtualpointsto}, in which we see that a virtual address with a certain fractional permissions at each level of the page-table walk gives us sound assurance of the path existence reaching to a page table entry, consequently, with a full-ownership of page address, we can obtain and update the value mapped value at this physical page address through our virtual memory address.
 \begin{figure*}
@@
 \]
-\caption{Virtual Points-to Relation}
+\caption{A Strong Virtual Points-to Relation}
 \todo[inline]{These fractions aren't quite right (though I see you added the variance between levels), I can walk you through in our meeting tomorrow.}
-  \label{fig:virtualpointsto}
+  \label{fig:strongvirtualpointsto}
 \end{figure*}
+One might have already observed that this virtual points-to definition is too strong to specify any virtual memory operations, some of which we explain within as part of experiement in Section \ref{}, require reasoning of sharing of physicall pages accross virtual memory addresses because this strong virtual points-to relation doesn’t care about owning those fractional resources, it only cares that something ensures the correct page table walk exists.
+Therefore we need to weaken the virtual-pointsto to capture the acutal sharing accross the virtual addresses.
+
+\subsection{Address-Spaces as a Modality}
+\subsubsection{Modal Resource Context}
+\subsubsection{Interaction: address-space switch as a modal-entailment}
+
 \subsection{Considering Hoare Doubles in the Context of Switching Address-Space}

### 2023-02-21 programlogic.tex
+

@@
 
+One possible approach to resolving this is to identify and separate the type of ownership within and outside the virtual points-to relation.
+Iris has a ghost map resource where authoritative ownership of an individual key-value pair can be handled as a resource.
+We use this mechanism to separate the virtual-to-physical translation from the physical memory involved: (fractional) virtual points-to assertions can be defined in terms of (fractional) ownership of these authoritative ghost map entry assertions, plus sharing an in-variant that the current installed page table respects all entries of the mapping as shown in Figure.
+Unmapping collects the authoritative map kvpairs from collecting the assertions, and then can remove them from the ghost map and update the page tables.
+Critically, physical ownership of the page tables then lives in the invariant on the current page table, so some virtual pointsto as- sertions can refer to memory in those page tables.
+This still works with the modality, since that invariant is also semantically a predicate on a page table root.
+
 \subsection{Address-Spaces as a Modality}

### 2023-02-21 programlogic.tex
+

--- 2023-02-21 programlogic.tex
@@
 
+One possible approach to resolving this is to identify and separate the type of ownership within and outside the virtual points-to relation.
+Iris has a ghost map resource where fragmental-ownership of an individual key-value pair (per-element) can be handled as a resource.
+We use this mechanism to separate the virtual-to-physical translation from the physical memory involved: virtual points-to assertions can be defined in terms of fragmental ownership (ownership of an element) of these partial ghost map entry assertions, plus sharing an invariant that the current installed page table respects all entries of the mapping as shown in Figure \ref{}.
+Unmapping collects the authoritative map kvpairs from collecting the assertions, and then can remove them from the ghost map and update the page tables.
+Critically, physical full ownership of the page tables then lives in the invariant on the current page table, so some virtual pointsto as- sertions can refer to memory in those page tables.
+This still works with the modality, since that invariant is also semantically a predicate on a page table root.
+
 \subsection{Address-Spaces as a Modality}

### 2023-02-22 programlogic.tex
+

@@
 \subsection{Aliasing/Sharing Physical Pages}
+\label{sec:sharingpages}
 As we see in Figure \fref{fig:}, virtual address mapping is obtained simply in two fold: 1.
@@
 
+\todo[inline,color=green]{ismail, put the fancy Iris figure for this ghost thing and expand the text above}
 \subsection{Address-Spaces as a Modality}
-\subsubsection{Modal Resource Context}
-\subsubsection{Interaction: address-space switch as a modal-entailment}
+\label{sec:aspacemodal}
+Abstractions we introduce for virtual address spaces are based on two pillars: 1.
+a global address space invariant asserting the authorative ownership of the ghost map, , summarizing the existence of the page-table walks, and for each physical page address reached by one of these abstract page-table walks we know that there exists concrete physical page-table accesses, i.e.
+L4-L1 pointsto  2.
+a modal definition describing the truth of assertions for the resources inside the address space, $[r]P$ indicating that $P$ holds in the virtual address space rooted at $r$.
+\subsubsection{Global Addressspace Invariant}
+
+\subsubsection{Address-Space Modality}
+\paragraph{Modal Resource Context}
+\paragraph{Interaction: address-space switch as a modal-entailment}
 

### 2023-02-22 programlogic.tex
+

@@
 \todo[inline,color=green]{ismail, put the fancy Iris figure for this ghost thing and expand the text above}
-\subsection{Address-Spaces as a Modality}
+\subsection{Address-Space Abstraction}
 \label{sec:aspacemodal}
@@
 a modal definition describing the truth of assertions for the resources inside the address space, $[r]P$ indicating that $P$ holds in the virtual address space rooted at $r$.
-\subsubsection{Global Addressspace Invariant}
+This is essentially what hybrid logic \ref{} calls a satisfaction operator, which evaluates the truth of a predicate in a named alternative state (here, address space).
+\subsubsection{Global Per-Addressspace Invariant}
+\label{sec:peraspaceinvariant}
 
 \subsubsection{Address-Space Modality}
+\label{sec:aspacemodalist}
 \paragraph{Modal Resource Context}
-\paragraph{Interaction: address-space switch as a modal-entailment}
-
+\label{sec:resourcecontext}
+\todo[inline,color=green]{ismail, speak about virtual points-to established under a certain root of an address space}
+\paragraph{Interaction}
+\label{sec:interaction}
+\todo[inline,color=green]{ismail: speak about the interection with ambient logic}
 \subsection{Considering Hoare Doubles in the Context of Switching Address-Space}

### 2023-02-22 programlogic.tex
+

@@
 This still works with the modality, since that invariant is also semantically a predicate on a page table root.
-
+\begin{figure*}
+\[
+\begin{array}{l}
+  \textsf{VirtPtsTo}\stackrel{def}{=} \lambda \tlf\Loc.\\
+  \ownGhost\gamma{a}
+\end{array}
+\]
+\caption{Virtual-Pointsto for Sharing Pages}
+\todo[inline]{These fractions aren't quite right (though I see you added the variance between levels), I can walk you through in our meeting tomorrow.}
+  \label{fig:virtualpointstosharing}
+\end{figure*}
 \todo[inline,color=green]{ismail, put the fancy Iris figure for this ghost thing and expand the text above}
@@
 \label{sec:peraspaceinvariant}
-
+\begin{figure*}
+\[
+\begin{array}{l}
+  \textsf{ASpaceInv}\stackrel{def}{=} \lambda \tlf\Loc.\\
+  \ownGhost\gamma{a}
+\end{array}
+\]
+\caption{Global Address-Space Invarian}
+\todo[inline]{These fractions aren't quite right (though I see you added the variance between levels), I can walk you through in our meeting tomorrow.}
+  \label{fig:peraspaceinvariant}
+\end{figure*}
 \subsubsection{Address-Space Modality}

### 2023-02-26 programlogic.tex
change logic shape aronud

@@
 \label{sec:logic}
-Our reasoning principles are shaped around understanding the following issues exhibiting themselves withing the context of vritual-memory-management:
+Our reasoning principles are shaped around understanding the following two main pillars of vritual-memory-management:
 \begin{enumerate}
-\item abstracting the relation between memory (virtual) addresses and the physical page addresses backing these virtual addresses
-\item regarding the address-space management, we propose reasoning principles for
-  \begin{itemize}
-  \item aliasing/sharing Physical Pages:
-  \item abstracting address-space as a modality
-    \begin{itemize}
-    \item modal resource context:
-    \item interaction:
-    \end{itemize}
-    \item Hoare-Doubles: 
-  \end{itemize}
+\item page-table walk: L4-L1 walk on shared page tables to reach a physical page address
+\item allocating/updating a pages backing a virtual address 
+\item switching from one address-space to another 
 \end{enumerate}

### 2023-02-26 programlogic.tex
+

@@
 \label{sec:logic}
-Our reasoning principles are shaped around understanding the following two main pillars of vritual-memory-management:
+Our reasoning principles are shaped around understanding the following main pillars of vritual-memory-management:
 \begin{enumerate}
@@
 The fraction $\qfrac$ with value 1 asserts the unique ownership of the register mapping, and grants update permission on it, otherwise, any value $0 < \qfrac <1$ represents partial ownership granting readonly permission on the mapping.
-\paragraph{Physical-Memory points-to} The our physical memory points-to relation ($\pfpointsto\locsf\locsf\qfrac\ppts$) exhibits nested-mappings due to masking applied on the indexing address ($\locsf$).
+\paragraph{Physical-Memory points-to} Our physical memory points-to relation ($\pfpointsto\locsf\locsf\qfrac\ppts$) exhibits nested-mappings abstracting the  masks applied on the indexing virtual address ($\textsf{vaddr}\in\mathcal{W}_{64}$).
 The unfolded definition of the physical mapsto
@@
 \begin{array}{cl}
-\pfpointsto\locsf\locsf\qfrac\ppts \stackrel{def}{=} & \nfpointsto{\mask\locsf\ft}{\mask\locsf\tw}\locsf\qfrac\naddr
+\pfpointsto\locsf\locsf\qfrac\ppts \stackrel{def}{=} & \nfpointsto{\mask\locsf\ft\generalentry}{\mask\locsf\tw\generalentry}\locsf\qfrac\naddr
 \end{array}
@@
 \end{figure}
-where we simply abstract the generation of hierarchical mapping of a physical address ($\locsf$) through two different levels of masking:
-
-\begin{itemize}
-\item  After a series of masking imposed on the virtual memory address (e.g. \textsf{vaddr} 64bits address [0-64]) :
+where, for an example virtual address \textsf{va} and given \textsf{CR3} in Figure \ref{}, we can abstract the L4-L1 table traversal as the following:
 \begin{itemize}
   \item Page Map Level-4 Translation (L4): Calculating the offset to the 4 \textsf{Kbyte Physical Page} table to obtain the  physical address by  which resides in the last 12 bits of the virtual memory address, $\locsf|^{12}$,  shown in Figure \ref{}.
-    \[  \pfpointsto{\lvlsum\crt\lfoff}{\lvlbor\lt}\qfracfotsss\ppts \star \]
+    % \pfpointsto{\lvlsum\crt\lfoff}{\lvlbor\lt}\qfracfotsss\ppts \star
+    \[ \nfpointsto{\mask\vaddr\ft\crthree}{\mask\vaddr\tw\crthree}\entryf\qfracfotsss\naddr \]
   \item Page Directory Pointer Level Translation (L3):
-    \[ \pfpointsto{\lvlsum\lt\ltoff}{\lvlbor\ltw}\qfracfotss\ppts \star \]
+    \[ \nfpointsto{\mask\vaddr\ft\entryf}{\mask\vaddr\tw\entryf}\entrytr\qfracfotss\naddr \]
   \item Page Directory Table Level Translation (L2):  we obtain the address of the page table via \textsf{PDE}$|^{52}$ and picking the offset from the virtual memory address via imposing the mask to obtain page table offset bits (9 bits [20-12] shown in Figure \ref{}) to obtain the page table entry (\textsf{PTE}).
-    \[  \pfpointsto{\lvlsum\ltw\ltwoff}{\lvlbor\lo}\qfracfots\ppts \star \]
+    \[ \nfpointsto{\mask\vaddr\ft\entrytr}{\mask\vaddr\tw\entrytr}\entrytw\qfracfots\naddr \]
   \item Page Table Level Translation (L1):
-    \[  \pfpointsto{\lvlsum\lo\looff}{\lvlbor\pg}\qfracfot\ppts \star \]
+    \[ \nfpointsto{\mask\vaddr\ft\entrytw}{\mask\vaddr\tw\entrytw}\entryo\qfracfot\naddr \]
+  \item Page Address Level Translation
+    \[ \nfpointsto{\mask\vaddr\ft\entryo}{\mask\vaddr\tw\entryo}\vpage\qfracone\naddr \]
 \end{itemize}
-\item then after obtainini the physical page address, we obtain the value
-  \[  \ppointsto{\pageptstosum\pg\pgoff}\locsf\ppts \]
-\end{itemize}
+\begin{remark}[A Strong Definition for Virtual Memory Addressing]
+  \label{rem:strongvmem}
 
-\subsection{Aliasing/Sharing Physical Pages}
-\label{sec:sharingpages}
-As we see in Figure \fref{fig:}, virtual address mapping is obtained simply in two fold: 1.
-traversing the page tables to reach the page table entry 2.
-fetch the physical address and the value it is adressing (Section \ref{fig:physicalpointsto}).
-This two fold access pattern enabling sharing of pages by virtual memory addresses exhibits logically sound representation against virtual memory operations some of which ((e.g.
-mapping a new page) we explain in detail in Section \ref{}.
-
-The very first assurance our logical constructions need to be able assert is the \textit{existence} of the page table walk.
-A naive attempt to assert the existence of the page-walk for a certain virtual memory translation would look like the one in Figure \ref{figure:strongvirtualpointsto}, in which we see that a virtual address with a certain fractional permissions at each level of the page-table walk gives us sound assurance of the path existence reaching to a page table entry, consequently, with a full-ownership of page address, we can obtain and update the value mapped value at this physical page address through our virtual memory address.
 \begin{figure*}
@@
 \end{figure*}
+\end{remark}
+%then we reach to a physical page addr (e.g. \textsf{pa}) 
+%\[  \paddr\mapsto_{\textsf{a}}\{1\}\vsome \ast \nfpointsto{\mask\vaddr\ft\entryo}{\mask\vaddr\tw\entryo}\vpage\qfrac\naddr \]
+  \subsection{Aliasing/Sharing Physical Pages}
+\label{sec:sharingpages}
+As we see in Figure \fref{fig:}, virtual address mapping is obtained simply in two fold: 1.
+traversing the page tables to reach the page table entry 2.
+fetch the physical address and the value it is adressing (Section \ref{fig:physicalpointsto}).
+This two fold access pattern enabling sharing of pages by virtual memory addresses exhibits logically sound representation against virtual memory operations some of which ((e.g.
+mapping a new page) we explain in detail in Section \ref{}.
+
+The very first assurance our logical constructions need to be able assert is the \textit{existence} of the page table walk.
+A naive attempt to assert the existence of the page-walk for a certain virtual memory translation would look like the one in Figure \ref{figure:strongvirtualpointsto}, in which we see that a virtual address with a certain fractional permissions at each level of the page-table walk gives us sound assurance of the path existence reaching to a page table entry, consequently, with a full-ownership of page address, we can obtain and update the value mapped value at this physical page address through our virtual memory address.
+
 One might have already observed that this virtual points-to definition is too strong to specify any virtual memory operations, some of which we explain within as part of experiement in Section \ref{}, require reasoning of sharing of physicall pages accross virtual memory addresses because this strong virtual points-to relation doesn’t care about owning those fractional resources, it only cares that something ensures the correct page table walk exists.

### 2023-02-27 programlogic.tex
synch
+

@@
     % \pfpointsto{\lvlsum\crt\lfoff}{\lvlbor\lt}\qfracfotsss\ppts \star
-    \[ \nfpointsto{\mask\vaddr\ft\crthree}{\mask\vaddr\tw\crthree}\entryf\qfracfotsss\naddr \]
+    \[ \textsf{L4translate} \stackrel{\triangle}{=} \exists \entryf \ldotp \nfpointsto{\mask\vaddr\ft\crthree}{\mask\vaddr\tw\crthree}\entryf\qfracfotsss\naddr \]
   \item Page Directory Pointer Level Translation (L3):
-    \[ \nfpointsto{\mask\vaddr\ft\entryf}{\mask\vaddr\tw\entryf}\entrytr\qfracfotss\naddr \]
+    \[ \textsf{L3translate} \stackrel{\triangle}{=} \exists \entrytr \ldotp \nfpointsto{\mask\vaddr\ft\entryf}{\mask\vaddr\tw\entryf}\entrytr\qfracfotss\naddr \]
   \item Page Directory Table Level Translation (L2):  we obtain the address of the page table via \textsf{PDE}$|^{52}$ and picking the offset from the virtual memory address via imposing the mask to obtain page table offset bits (9 bits [20-12] shown in Figure \ref{}) to obtain the page table entry (\textsf{PTE}).
-    \[ \nfpointsto{\mask\vaddr\ft\entrytr}{\mask\vaddr\tw\entrytr}\entrytw\qfracfots\naddr \]
+    \[ \textsf{L2translate} \stackrel{\triangle}{=} \exists \entrytw \ldotp \nfpointsto{\mask\vaddr\ft\entrytr}{\mask\vaddr\tw\entrytr}\entrytw\qfracfots\naddr \]
   \item Page Table Level Translation (L1):
-    \[ \nfpointsto{\mask\vaddr\ft\entrytw}{\mask\vaddr\tw\entrytw}\entryo\qfracfot\naddr \]
+    \[ \textsf{L1translate}  \stackrel{\triangle}{=} \exists \entryo \ldotp \nfpointsto{\mask\vaddr\ft\entrytw}{\mask\vaddr\tw\entrytw}\entryo\qfracfot\naddr \]
   \item Page Address Level Translation
-    \[ \nfpointsto{\mask\vaddr\ft\entryo}{\mask\vaddr\tw\entryo}\vpage\qfracone\naddr \]
+    \[ \textsf{PageAccess} \stackrel{\triangle}{=} \exists \vpage \ldotp \nfpointsto{\mask\vaddr\ft\entryo}{\mask\vaddr\tw\entryo}\vpage\qfracone\naddr \]
+
 \end{itemize}
@@
   \label{rem:strongvmem}
+  One could give a naive logical definition to the virtual mememory addressing as conjuctions of each level translation (Figure \fref{fig:strongvirtualpointsto}).
+The very first assurance our logical constructions need to be able assert is the \textit{existence} of the page table walk.
+Our naive attempt shown in Figure \ref{figure:strongvirtualpointsto}, we see that a virtual address with a certain fractional permissions at each level of the page-table walk gives us sound assurance of the path existence reaching to a page table entry, consequently, with a full-ownership of page address, we can obtain and update the value mapped value at this physical page address through our virtual memory address.
 
@@
 \begin{array}{l}
-  \ppointsto\vaddr\locsf\vpts \stackrel{def}{=} \lambda \tlf\Loc.\\
-  \exists_{\tlfoff\Locn \;, \tltoff\Locn \;, \tltwoff\Locn \;,\tlooff\Locn \;, \tpgoff\Loctw \;, \tlt\Loc \;, \tltw\Loc \;, \tlo\Loc \tpg\Loc} \ldotp \\
-  \ulcorner \textsf{aligned } \vaddr \urcorner \star 
-   \ulcorner \vaddr = \locsx :: \lfoff :: \ltoff :: 
-   \ltwoff :: \looff :: \pgoff \urcorner \star\\
-  \pfpointsto{\lvlsum\crt\lfoff}{\lvlbor\lt}\qfracfotsss\ppts \star 
-  \pfpointsto{\lvlsum\lt\ltoff}{\lvlbor\ltw}\qfracfotss\ppts \star \\
-  \pfpointsto{\lvlsum\ltw\ltwoff}{\lvlbor\lo}\qfracfots\ppts \star 
-  \pfpointsto{\lvlsum\lo\looff}{\lvlbor\pg}\qfracfot\ppts \star \\
-  \ppointsto{\pageptstosum\pg\pgoff}\locsf\ppts 
+  \vaddr\mapsto_{\textsf{v}}\{\textsf{q}\}\vpage \stackrel{def}{=} \lambda \crthree.\\
+  \exists_{\entryf \;, \entrytr \;, \entrytw \;,\entryo} \ldotp
+
+  \ulcorner \textsf{aligned } \vaddr \urcorner \star  \\
+  \nfpointsto{\mask\vaddr\ft\crthree}{\mask\vaddr\tw\crthree}\entryf\qfracfotsss\naddr \star \\ 
+   \nfpointsto{\mask\vaddr\ft\entryf}{\mask\vaddr\tw\entryf}\entrytr\qfracfotss\naddr  \star  \\
+   \nfpointsto{\mask\vaddr\ft\entrytr}{\mask\vaddr\tw\entrytr}\entrytw\qfracfots\naddr \star \\
+  \nfpointsto{\mask\vaddr\ft\entrytw}{\mask\vaddr\tw\entrytw}\entryo\qfracfot\naddr \star \\
+   \nfpointsto{\mask\vaddr\ft\entryo}{\mask\vaddr\tw\entryo}\vpage\qfracone\naddr 
 \end{array}
@@
   \label{sec:sharingpages}
-As we see in Figure \fref{fig:}, virtual address mapping is obtained simply in two fold: 1.
+  Here we see that virtual address mapping is obtained simply in two fold: 1.
 traversing the page tables to reach the page table entry 2.
 fetch the physical address and the value it is adressing (Section \ref{fig:physicalpointsto}).
-This two fold access pattern enabling sharing of pages by virtual memory addresses exhibits logically sound representation against virtual memory operations some of which ((e.g.
+This two fold access-pattern enabling sharing of pages by virtual memory addresses exhibits logically sound representation against virtual memory operations some of which ((e.g.
 mapping a new page) we explain in detail in Section \ref{}.
-
-The very first assurance our logical constructions need to be able assert is the \textit{existence} of the page table walk.
-A naive attempt to assert the existence of the page-walk for a certain virtual memory translation would look like the one in Figure \ref{figure:strongvirtualpointsto}, in which we see that a virtual address with a certain fractional permissions at each level of the page-table walk gives us sound assurance of the path existence reaching to a page table entry, consequently, with a full-ownership of page address, we can obtain and update the value mapped value at this physical page address through our virtual memory address.
 

### 2023-02-27 programlogic.tex
+

@@
   \label{sec:sharingpages}
-  Here we see that virtual address mapping is obtained simply in two fold: 1.
+  We already build the intuition on what a virtual address mapping is composed of: 1.
 traversing the page tables to reach the page table entry 2.
-fetch the physical address and the value it is adressing (Section \ref{fig:physicalpointsto}).
+fetching the physical address, and accessing the value mapped.
 This two fold access-pattern enabling sharing of pages by virtual memory addresses exhibits logically sound representation against virtual memory operations some of which ((e.g.

### 2023-02-27 programlogic.tex
+

@@
 
-One might have already observed that this virtual points-to definition is too strong to specify any virtual memory operations, some of which we explain within as part of experiement in Section \ref{}, require reasoning of sharing of physicall pages accross virtual memory addresses because this strong virtual points-to relation doesn’t care about owning those fractional resources, it only cares that something ensures the correct page table walk exists.
+One might have already observed that this virtual points-to definition is too strong to specify virtual memory operations, some of which we explain within as part of experiement in Section \ref{}, which  requires reasoning of sharing physical pages across different virtual memory addresses.
+Although there seems to fractional permissions for fragmental ownership of page-table entries, strong virtual points-to (in Figure \ref{strongvirtualpointsto}) relation doesn’t care about owning those fractional resources, it only cares that something ensures the correct page table walk exists.
 Therefore we need to weaken the virtual-pointsto to capture the acutal sharing accross the virtual addresses.
 
-One possible approach to resolving this is to identify and separate the type of ownership within and outside the virtual points-to relation.
+One possible approach to is to just abstracting the \textit{valid} page-table walk existence from a each virtual address (e.g.
+\textsf{va}) to a physical page address, as \textit{per-element} ownership of a logical \textit{page-table-walk-summarization-map}
+\[\mathcal{M}\textsf{PTableWalk} \stackrel{\triangle}{=} \textsc{Auth} (\; \mathcal{W}_{64} \;\rightarrow_{\textrm{fin}} \;  (\mathcal{W}_{64} \; \rightarrow_{\textrm{fin}}\;  (\textsc{Frac }, \mathord{+}) \times (\textsc{Agree } \Loc,\mathord{=}) ))\]
+provided by a fragmentery ghost resrource ownership of $\mathcal{M}\textsf{PTableWalk}$.
+\todo[inline,color=red]{stop here}tranresolving this is to identify and separate the type of ownership within and outside the virtual points-to relation.
 Iris has a ghost map resource where fragmental-ownership of an individual key-value pair (per-element) can be handled as a resource.

### 2023-02-27 programlogic.tex
+

@@
 \label{sec:logic}
+% The predicate gen_heap_interp.
+\newcommand{\gammaPred}{\delta}
+
+\newcommand{\sumwalkabs}[3]{
+  \ownGhost\gammaPred{\authfrag{\singletonMap{#1}{(#2, #3)}}}
+}
+\newcommand{\ptableabswalk}[1]{\mathcal{A}\textsf{bsPTableWalk}(#1)}
+\newcommand{\ptablestore}{\theta}
 Our reasoning principles are shaped around understanding the following main pillars of vritual-memory-management:
@@
 %\[  \paddr\mapsto_{\textsf{a}}\{1\}\vsome \ast \nfpointsto{\mask\vaddr\ft\entryo}{\mask\vaddr\tw\entryo}\vpage\qfrac\naddr \]
+
   \subsection{Aliasing/Sharing Physical Pages}
@@
 \[\mathcal{M}\textsf{PTableWalk} \stackrel{\triangle}{=} \textsc{Auth} (\; \mathcal{W}_{64} \;\rightarrow_{\textrm{fin}} \;  (\mathcal{W}_{64} \; \rightarrow_{\textrm{fin}}\;  (\textsc{Frac }, \mathord{+}) \times (\textsc{Agree } \Loc,\mathord{=}) ))\]
-provided by a fragmentery ghost resrource ownership of $\mathcal{M}\textsf{PTableWalk}$.
-\todo[inline,color=red]{stop here}tranresolving this is to identify and separate the type of ownership within and outside the virtual points-to relation.
-Iris has a ghost map resource where fragmental-ownership of an individual key-value pair (per-element) can be handled as a resource.
-We use this mechanism to separate the virtual-to-physical translation from the physical memory involved: virtual points-to assertions can be defined in terms of fragmental ownership (ownership of an element) of these partial ghost map entry assertions, plus sharing an invariant that the current installed page table respects all entries of the mapping as shown in Figure \ref{}.
-Unmapping collects the authoritative map kvpairs from collecting the assertions, and then can remove them from the ghost map and update the page tables.
-Critically, physical full ownership of the page tables then lives in the invariant on the current page table, so some virtual pointsto as- sertions can refer to memory in those page tables.
-This still works with the modality, since that invariant is also semantically a predicate on a page table root.
+provided by a fragmentery ghost resrource ownership of $\mathcal{A}\textsf{bsPTableWalk}$.
+\begin{remark}[Exchanging Page-Table-Walk-Summarization Mappings (Tokens) with Physical Page-Table-Walk Pointstos ]
+$\mathcal{A}\textsf{PTableWalk}$ is a ghost map resource where fragmental-ownership of an individual key-value pair (per-element) can be handled as a resource.
+We use this mechanism to separate the virtual-to-physical translation from the physical memory involved: virtual points-to assertions can be defined in terms of fragmental ownership (ownership of an element) of these partial ghost map entry assertions , $\sumwalkabs\vaddr\qfrac\paddr$, plus sharing an invariant ($\mathcal{I}$\textsf{ASpace} in Figure \ref{fig:peraspaceinvariant}) that the current installed page table respects all entries of the mapping, $\ownGhost\gammaPred{\authfull{\ptableabswalk\ptablestore}}$.
+From one another perspective, unmapping pages collects the fragmental map kvpairs, e.g.
+$\sumwalkabs\vaddr\qfrac\paddr$, from collecting the assertions, and then can remove them from the ghost map and update the page tables.
+Critically, physical full ownership of the page tables lives, $\ownGhost\gammaPred{\authfull{\ptableabswalk\ptablestore}}$, in the central invariant on the current page table, so some virtual pointsto assertions can refer to memory in those page tables.% This still works with the modality, since that invariant is also semantically a predicate on a page table root.
+\end{remark}
 \begin{figure*}
@@
 \end{figure*}
-\todo[inline,color=green]{ismail, put the fancy Iris figure for this ghost thing and expand the text above}
+
 \subsection{Address-Space Abstraction}

### 2023-02-27 programlogic.tex
+

@@
 traversing the page tables to reach the page table entry 2.
-fetching the physical address, and accessing the value mapped.
-This two fold access-pattern enabling sharing of pages by virtual memory addresses exhibits logically sound representation against virtual memory operations some of which ((e.g.
-mapping a new page) we explain in detail in Section \ref{}.
+obtaining the physical address, and accessing the value mapped.
 
-One might have already observed that this virtual points-to definition is too strong to specify virtual memory operations, some of which we explain within as part of experiement in Section \ref{}, which  requires reasoning of sharing physical pages across different virtual memory addresses.
+One might have already observed that the virtual points-to definition shown in Figure \ref{fig:strongvirtualpointsto} is too strong to specify virtual memory operations updating a page table entry.
 Although there seems to fractional permissions for fragmental ownership of page-table entries, strong virtual points-to (in Figure \ref{strongvirtualpointsto}) relation doesn’t care about owning those fractional resources, it only cares that something ensures the correct page table walk exists.
-Therefore we need to weaken the virtual-pointsto to capture the acutal sharing accross the virtual addresses.
+Any virtual memory operations such as mapping a new page which updates an page table entry would violate the soundness of other virtual-pointsto relations in the existence of page-sharing across the virtual addresses in the same address space.
 
@@
 \textsf{va}) to a physical page address, as \textit{per-element} ownership of a logical \textit{page-table-walk-summarization-map}
-\[\mathcal{M}\textsf{PTableWalk} \stackrel{\triangle}{=} \textsc{Auth} (\; \mathcal{W}_{64} \;\rightarrow_{\textrm{fin}} \;  (\mathcal{W}_{64} \; \rightarrow_{\textrm{fin}}\;  (\textsc{Frac }, \mathord{+}) \times (\textsc{Agree } \Loc,\mathord{=}) ))\]
+\[\mathcal{A}\textsf{bsPTableWalk} \stackrel{\triangle}{=} \textsc{Auth} (\; \mathcal{W}_{64} \;\rightarrow_{\textrm{fin}} \;  (\mathcal{W}_{64} \; \rightarrow_{\textrm{fin}}\;  (\textsc{Frac }, \mathord{+}) \times (\textsc{Agree } \Loc,\mathord{=}) ))\]
 provided by a fragmentery ghost resrource ownership of $\mathcal{A}\textsf{bsPTableWalk}$.
@@
 \end{remark}
+Then, our virtual memory addressing would be the fragmental ownership of the abstract page-table-walk mapping with the physical page address mapping as shown in Figure \ref{fig:virtualpointstosharing}.
 \begin{figure*}
@@
 \begin{array}{l}
-  \textsf{VirtPtsTo}\stackrel{def}{=} \lambda \tlf\Loc.\\
-  \ownGhost\gamma{a} 
+  \textsf{VPointsTo}(\vaddr \;, \vpage)\stackrel{\triangle}{=} 
+  \exists \entryo \ldotp 
+  \sumwalkabs\vaddr\qfrac\paddr \star 
+   \nfpointsto{\mask\vaddr\ft\entryo}{\mask\vaddr\tw\entryo}\vpage\qfracone\naddr
 \end{array}
@@
 \caption{Virtual-Pointsto for Sharing Pages}
-\todo[inline]{These fractions aren't quite right (though I see you added the variance between levels), I can walk you through in our meeting tomorrow.}
   \label{fig:virtualpointstosharing}
@@
 \label{sec:aspacemodal}
-Abstractions we introduce for virtual address spaces are based on two pillars: 1.
-a global address space invariant asserting the authorative ownership of the ghost map, , summarizing the existence of the page-table walks, and for each physical page address reached by one of these abstract page-table walks we know that there exists concrete physical page-table accesses, i.e.
-L4-L1 pointsto  2.
-a modal definition describing the truth of assertions for the resources inside the address space, $[r]P$ indicating that $P$ holds in the virtual address space rooted at $r$.
+Abstractions we introduce for virtual address spaces are based on two pillars:
+\begin{enumerate}
+\item a global address space invariant asserting the authorative ownership of the ghost map, , summarizing the existence of the page-table walks, and for each physical page address reached by one of these abstract page-table walks we know that there exists concrete physical page-table accesses, i.e.
+L4-L1 pointsto
+  \item  a modal definition describing the truth of assertions for the resources inside the address space, $[r]P$ indicating that $P$ holds in the virtual address space rooted at $r$.
 This is essentially what hybrid logic \ref{} calls a satisfaction operator, which evaluates the truth of a predicate in a named alternative state (here, address space).
+\end{enumerate}
     \subsubsection{Global Per-Addressspace Invariant}
+
 \label{sec:peraspaceinvariant}
@@
 \begin{array}{l}
-  \textsf{ASpaceInv}\stackrel{def}{=} \lambda \tlf\Loc.\\
+  \mathcal{I}\textsf{ASpace} \stackrel{def}{=} \lambda \tlf\Loc.\\
   \ownGhost\gamma{a}

### 2023-02-27 programlogic.tex
+

@@
   \begin{array}{l}
+\begin{array}{l}
   \vaddr\mapsto_{\textsf{v}}\{\textsf{q}\}\vpage \stackrel{def}{=} \lambda \crthree.\\
@@
   \ulcorner \textsf{aligned } \vaddr \urcorner \star  \\
-  \nfpointsto{\mask\vaddr\ft\crthree}{\mask\vaddr\tw\crthree}\entryf\qfracfotsss\naddr \star \\ 
-   \nfpointsto{\mask\vaddr\ft\entryf}{\mask\vaddr\tw\entryf}\entrytr\qfracfotss\naddr  \star  \\
-   \nfpointsto{\mask\vaddr\ft\entrytr}{\mask\vaddr\tw\entrytr}\entrytw\qfracfots\naddr \star \\
-  \nfpointsto{\mask\vaddr\ft\entrytw}{\mask\vaddr\tw\entrytw}\entryo\qfracfot\naddr \star \\
+  \textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\crthree,\entryf,\entrytr,\entrytw,\entryo)\\
    \nfpointsto{\mask\vaddr\ft\entryo}{\mask\vaddr\tw\entryo}\vpage\qfracone\naddr 
+\end{array} \\
+
+\textsf{where   }\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\crthree,\entryf,\entrytr,\entrytw,\entryo) \stackrel{\triangle}{=} \\
+ \qquad\qquad \nfpointsto{\mask\vaddr\ft\crthree}{\mask\vaddr\tw\crthree}\entryf\qfracfotsss\naddr \star \\ 
+ \qquad\qquad  \nfpointsto{\mask\vaddr\ft\entryf}{\mask\vaddr\tw\entryf}\entrytr\qfracfotss\naddr  \star  \\
+  \qquad\qquad \nfpointsto{\mask\vaddr\ft\entrytr}{\mask\vaddr\tw\entrytr}\entrytw\qfracfots\naddr \star \\
+  \qquad\qquad \nfpointsto{\mask\vaddr\ft\entrytw}{\mask\vaddr\tw\entrytw}\entryo\qfracfot\naddr 
    \end{array}
@@
 \begin{enumerate}
-\item a global address space invariant asserting the authorative ownership of the ghost map, , summarizing the existence of the page-table walks, and for each physical page address reached by one of these abstract page-table walks we know that there exists concrete physical page-table accesses, i.e.
-L4-L1 pointsto
+\item a global address space invariant asserting the authorative ownership of the ghost map, $\ownGhost\gammaPred{\authfull{\ptableabswalk\ptablestore}}$ 
+\item for each ghost page-table-walk mapping,  $\sumwalkabs\vaddr\qfrac\paddr$, there exists concrete physical page-table accesses path
+  \[\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\crthree,\entryf,\entrytr,\entrytw,\entryo)\]
   \item  a modal definition describing the truth of assertions for the resources inside the address space, $[r]P$ indicating that $P$ holds in the virtual address space rooted at $r$.

### 2023-03-01 programlogic.tex
+

@@
 \label{sec:aspacemodal}
-Abstractions we introduce for virtual address spaces are based on two pillars:
+Abstractions we introduce for virtual address spaces are based on the following aspects:
 \begin{enumerate}
@@
   \[\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\crthree,\entryf,\entrytr,\entrytw,\entryo)\]
-  \item  a modal definition describing the truth of assertions for the resources inside the address space, $[r]P$ indicating that $P$ holds in the virtual address space rooted at $r$.
-This is essentially what hybrid logic \ref{} calls a satisfaction operator, which evaluates the truth of a predicate in a named alternative state (here, address space).
+  \item all virtual pointsto assertions happen to be true in a certain address space, i.e.
+the truth on address virtualization is indexed by CPUs page table register ($\crthree$ for x86-64), and we present modal reasoning principles for updating $\crthree$ (switching from the current address space to another)
 \end{enumerate}
-    \subsubsection{Global Per-Addressspace Invariant}
-
+    \subsubsection{Global Per-Address-Space Invariant}
     \label{sec:peraspaceinvariant}
+    Virtual points-to assertions would need to be defined in terms of physical points-to assertions and hardware-specific address translation assertions for mapped virtual addresses ($\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\crthree,\entryf,\entrytr,\entrytw,\entryo)$ in Figure \ref{fig:peraspaceinvariant}).
+For each virtual address \textit{mapped} with a physical page table entry, abstracted with ghost mappings ($\sumwalkabs\vaddr\qfrac\paddr$) in virtul-pointsto relations as shown in Figure \ref{fig:virtualpointstosharing}), there exsits a physical page table walk.
+    \[ \bigast{(\vaddr, \entryo)\in \ptablestore}{\exists\;(\entryf\;,\entrytr\;,\entrytw)\ldotp \textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\crthree,\vaddr,\entryf,\entrytr,\entrytw,\entryo)} \]
+    By separating the virtual-points Overlapping address spaces — both shared data, and shared page tables
+which may see updates — can be coordinated with protocols to ensure those updates preserve
+address space invariants.
 \begin{figure*}
@@
 \begin{array}{l}
-  \mathcal{I}\textsf{ASpace} \stackrel{def}{=} \lambda \tlf\Loc.\\
-  \ownGhost\gamma{a}
+  \mathcal{I}\textsf{ASpace} \stackrel{\triangle}{=} \lambda \; \crthree \ldotp
+  \exists\;\gammaPred \;\ldotp\; \ownGhost\gammaPred{\authfull{\ptableabswalk\ptablestore}} \star \\
+  \bigast{(\vaddr, \entryo)\in \ptablestore}{\exists\;(\entryf\;,\entrytr\;,\entrytw)\ldotp \textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\crthree,\vaddr,\entryf,\entrytr,\entrytw,\entryo)}
 \end{array}
 \]
-\caption{Global Address-Space Invarian}
-\todo[inline]{These fractions aren't quite right (though I see you added the variance between levels), I can walk you through in our meeting tomorrow.}
+\caption{Global Address-Space Invariant}
   \label{fig:peraspaceinvariant}
@@
 \label{sec:aspacemodalist}
+
+The truth on an address space, exhibits itself as a contingent truth: a location virtualization assertion (a virtual points-to in \ref{fig:virtualpointstosharing}),  happens to valid in the \textit{current world} (in the current address space), and switching address spaces pulls information out of one \textit{world} into the “current view” of memory, and leaves other assertions true relative to the previous address space.
+
+Therefore, an address-space, as an abstraction, can be treated as naming the memory state as a modal frame and the choice of page table root as a world in Kripke-style semantics.
+Not suprisingly, transition between two address-spaces, then, can be just an entailment relation \textit{alternating} the \textit{named-state} against the address-space invariant ($\mathcal{I}\textsf{ASpace}$ in Figure \ref{fig:peraspaceinvariant}).
+
+Being inspired by what hybrid logic \ref{} calls a satisfaction operator, which evaluates the truth of a predicate in a named alternative state (here, address space), we give a modal definition describing the truth of assertions for the resources inside the address space, $[r]P$ indicating that $P$ holds in the virtual address space rooted at $r$.
+While doing so, we explain what a modal resource context is, and how it interacts with the ambient logic, i.e.
+separation logic.
 \paragraph{Modal Resource Context}
 \label{sec:resourcecontext}
-\todo[inline,color=green]{ismail, speak about virtual points-to established under a certain root of an address space}
+
 \paragraph{Interaction}
 \label{sec:interaction}
-\todo[inline,color=green]{ismail: speak about the interection with ambient logic}
+
 \subsection{Considering Hoare Doubles in the Context of Switching Address-Space}

### 2023-03-01 programlogic.tex
+

@@
 \label{sec:resourcecontext}
-
+When we pick up a certain contingency, which is \textit{"a certain virtual-pointsto facts whose validity is indexed by the root address of an address space"}for address space modality, we, in fact, make our choice around some logical resources for which we care about reasoning, which is virtual pointsto relation shown in Figure \ref{fig:virtualpointstosharing}.
 \paragraph{Interaction}

### 2023-03-01 programlogic.tex
+

@@
 When we pick up a certain contingency, which is \textit{"a certain virtual-pointsto facts whose validity is indexed by the root address of an address space"}for address space modality, we, in fact, make our choice around some logical resources for which we care about reasoning, which is virtual pointsto relation shown in Figure \ref{fig:virtualpointstosharing}.
+In other words, our modality definition is the logical virtual pointsto resources: introduction/elimination of logical virtual pointsto resources are the ones essential to the reasoning capabilities the address space modality provides, i.e.
+validity of virtual addresses under/out of modality -- resources under/out of \textit{modal resource context}.
 \paragraph{Interaction}

### 2023-03-01 programlogic.tex
+

@@
 validity of virtual addresses under/out of modality -- resources under/out of \textit{modal resource context}.
+
+The choice modal context itself already exhibits itself as a reasoning principle (Figure \ref{fig:modalcontextinvariance}) in which we can intro/eliminate any resource that is not virtual pointsto into/from modal context without dischargin any obligation.
+We can use this entailment for  \todo[inline,color=red]{Ask Colin}.
+\begin{figure}
+  \[
+  \begin{array}{l}
+\hbox{(\TirNameStyle{ModalContextInvariance})} \qquad
+ [r](Q \star \ell \mapsto_{\textsf{p}} \vpage)  \;\vdash [r](Q) \star \ell \mapsto_{\textsf{p}} \vpage 
+  \end{array}
+  \]
+\caption{Modal Context Invariance}
+\Description{Modal Context Invariance}
+\label{fig:modalcontextinvariance}
+\end{figure}
+\begin{remark}[The Choice of Modal Context as a Pattern of Verification]
+  By choosing the contingency and the modal contex, we simply introduce 
+  \begin{enumerate}
+  \item the pattern of what the essential abstraction (e.g. address spaces) we care about to do verification is
+  \item how a modal context (e.g.
+a address-space modality), logically, draws the attention to the certain facts in the interest of reasoning (e.g.
+virtual pointsto relations in the current address-space)
+  \item and, offloads the burden of individual bookkeeping of these facts being interesting under different context (e.g.
+virtual pointsto relations in other address-spaces) via utilizing the \textit{summarization} aspect of the contingency it represents (e.g.
+other virtual pointsto relations are valid under other certain address-space page-table root addresseses).
+  \end{enumerate}
+\end{remark}
 \paragraph{Interaction}
 \label{sec:interaction}
+After defining the modal context, we can now use introduction/elimination of a resource into/from a modal context for the sake of real low-level systems verification.
+For example, we can extend the modal context with some new virtual address mapping after backing them with new pages...
+
+After defining the mod 
+This rule captures the notion that switching address spaces pulls information out of one modality into the \textit{“current view”} of memory, and leaves other assertions true relative to the previous address
+space.
 

### 2023-03-02 programlogic.tex
+

--- 2023-03-01 programlogic.tex
@@
 \begin{array}{l}
-  \textsf{VPointsTo}(\vaddr \;, \vpage)\stackrel{\triangle}{=} 
+    \vaddr\mapsto_{\textsf{v}}\{\textsf{q}\}\vpage \stackrel{\triangle}{=} 
   \exists \entryo \ldotp 
@@
 \label{sec:aspacemodalist}
-
 The truth on an address space, exhibits itself as a contingent truth: a location virtualization assertion (a virtual points-to in \ref{fig:virtualpointstosharing}),  happens to valid in the \textit{current world} (in the current address space), and switching address spaces pulls information out of one \textit{world} into the “current view” of memory, and leaves other assertions true relative to the previous address space.
@@
 Not suprisingly, transition between two address-spaces, then, can be just an entailment relation \textit{alternating} the \textit{named-state} against the address-space invariant ($\mathcal{I}\textsf{ASpace}$ in Figure \ref{fig:peraspaceinvariant}).
-
-Being inspired by what hybrid logic \ref{} calls a satisfaction operator, which evaluates the truth of a predicate in a named alternative state (here, address space), we give a modal definition describing the truth of assertions for the resources inside the address space, $[r]P$ indicating that $P$ holds in the virtual address space rooted at $r$.
-While doing so, we explain what a modal resource context is, and how it interacts with the ambient logic, i.e.
-separation logic.
+\[
+  \begin{array}{l}
+\hbox{(\TirNameStyle{ModalAddressSpace})} \qquad
+ [r](P)  \;\stackrel{\triangle}{=} (\mathcal{W}_{64} -\textsf{b}> \textsf{uPredI} (\textsf{iResUR} \;\Sigma)) -\textsf{b}> \textsf{uPredI} \_ := \lambda \textsf{P},\; \textsf{P r}. 
+  \end{array}
+\]
+Being inspired by what hybrid logic \ref{} calls a satisfaction operator, which evaluates the truth of a predicate in a named alternative state (here, address space), we give a modal definition describing the truth of assertions for the resources (i.e.
+virtual pointsto relations) inside the address space.
+Ignoring the predicate types for a while, $[r]P$ in Definition $\hbox{(\TirNameStyle{ModalAddressSpace})}$, indicates that $P$ holds in the virtual address space rooted at $r$, the truth ($P$) on an address-space is indexed by the root page-table address $r$ of the address-space.
+In the rest of this section, we explain the structural aspects -- the modal resource context of address space modality -- , and how we lift the interaction of address-space modality with the ambient logic, i.e.
+separation logic as an entailment for specifying \textit{address-space-switch}.
 \paragraph{Modal Resource Context}
@@
 validity of virtual addresses under/out of modality -- resources under/out of \textit{modal resource context}.
+
+The choice of resource defining the modal context already exhibits itself as a reasoning principle (Figure \ref{fig:modalcontextinvariance}) in which we can intro/eliminate any resource that is not virtual pointsto into/from modal context without dischargin any obligation.
+We can use this entailment for  \todo[inline,color=red]{Ask Colin}.
+  \[
+  \begin{array}{l}
+\hbox{(\TirNameStyle{ModalContextInvariance})} \qquad
+ [r](Q \star \ell \mapsto_{\textsf{p}} \vpage)  \;\vdash [r](Q) \star \ell \mapsto_{\textsf{p}} \vpage 
+  \end{array}
+  \]
+\begin{remark}[The Choice of Modal Context Definition as a Pattern of Verification Context]
+  \label{remark:pattern}
+  By choosing the contingency and what defines the modal contex, we simply introduce 
+  \begin{enumerate}
+  \item the pattern of what the essential abstraction (e.g. address spaces) we care about to do verification for is
+  \item how a modal context (e.g.
+a address-space modality $[r]P$), logically, draws the attention to the certain facts in the interest of reasoning (e.g.
+virtual pointsto relations in the current address-space in Figure \ref{fig:virtualpointstosharing})
+  \item and, offloads the burden of individual bookkeeping of these facts (e.g.
+virtual pointsto relations per address-space) under different context (e.g.
+virtual pointsto relations in other address-spaces) via utilizing the \textit{summarization} aspect of the contingency it represents (e.g.
+other virtual pointsto relations are valid under other certain address-space page-table root addresseses).
+  \end{enumerate}
+\end{remark}
 \paragraph{Interaction}
 \label{sec:interaction}
+So far, we have just mentioned the structural properties of our modal definition (Remark \ref{remark:pattern}), and what inherent entailments (Rule \hbox{(\TirNameStyle{ModalContextInvariance})}) these properties provide.
+However, all these structural aspects of using the address-space modality ($[r]P$) considers the fixed index ($r$) of on the truth $P$.
+From the client's perspective, structural properties of the modal abstraction mainly constitutes the modal frame with summarizing the facts for one another address-space under the address-space modality roughly as it is in the following spec
+\[
+\{ \crthree \mapsto_{r} r0 \star \textsf{va1} \mapsto_{v} \textsf{v}_1 \star \textsf{va2} \mapsto_{v} \textsf{v}_2 \star \textsf{va3} \mapsto_{v} \textsf{v}_3 \star [r1]\;P
+\]
+where we see the \textit{current address-space} rooted at $r0$ is loaded into the  \textit{current-view} of the memory with its resources ($ \textsf{va1} \mapsto_{v} \textsf{v}_1 \star \textsf{va2} \mapsto_{v} \textsf{v}_2 \star \textsf{va3} \mapsto_{v} \textsf{v}_3$), and framed by the address-space rooted at $r1$ with resources $P$.
+Now, we need to a rule capturing the notion that switching the address spaces by pulling the resources ($P$) out of the one under modality into the \textit{“current view”} of memory, and leaves other assertions true relative to the previous address space (the one rooted at $r0$).
+To do so, we need to have a proof rule for the instruction (\lstinline|mov\_ctl|) updating the $\crthree$ register.
 
@@
 Virtual points-to assertions could be made self-stable by also giving them partial ownership over \lstinline|cr3| assertions, but this would require explicitly plumbing that ownership from \emph{all} assertions back to any place an address space change might occur; this would seem to be a far graver loss of modularity than this extra quirk in framing discussions.
+
+\begin{figure}
+\begin{mathpar}
+\inferrule[ChangeAddressSpace]{
+  \{[r_1](P) \ast cr3=r_2 \ast r = r_2 \ast Q\}\overline{is}
+}{
+  \{P \ast cr3=r_1 \ast r=r_2 \ast [r_2](Q)\}
+  \texttt{mov}~\texttt{\%cr3},~r;\;\overline{is}
+  %\lstinline|mov %cr3, r| 
+}
+\end{mathpar}
+\caption{Selected Reasoning Rules for ..}
+\label{fig:wpdamd}
+\end{figure}

### 2023-03-02 programlogic.tex
+
sync with Desk

@@
   \begin{array}{l}
-\hbox{(\TirNameStyle{ModalAddressSpace})} \qquad
+    \hbox{(\TirNameStyle{ModalAddressSpace})} \\ \qquad
          [r](P)  \;\stackrel{\triangle}{=}  (\mathcal{W}_{64} -\textsf{b}> \textsf{uPredI} (\textsf{iResUR} \;\Sigma)) -\textsf{b}> \textsf{uPredI} \_ := \lambda \textsf{P},\; \textsf{P r}. 
@@
 \[
-\{ \crthree \mapsto_{r} r0 \star \textsf{va1} \mapsto_{v} \textsf{v}_1 \star \textsf{va2} \mapsto_{v} \textsf{v}_2 \star \textsf{va3} \mapsto_{v} \textsf{v}_3 \star [r1]\;P
+\{ \crthree \mapsto_{r} r0 \star \textsf{va1} \mapsto_{v} \textsf{v}_1 \star \textsf{va2} \mapsto_{v} \textsf{v}_2 \star \textsf{va3} \mapsto_{v} \textsf{v}_3 \star [r1]\;P \}
 \]
@@
 Now, we need to a rule capturing the notion that switching the address spaces by pulling the resources ($P$) out of the one under modality into the \textit{“current view”} of memory, and leaves other assertions true relative to the previous address space (the one rooted at $r0$).
-To do so, we need to have a proof rule for the instruction (\lstinline|mov\_ctl|) updating the $\crthree$ register.
+To do so, we need to have a proof rule for the instruction (\lstinline|mov_ctl|) updating the $\crthree$.
 
@@
 \inferrule[Broken]{ }{
-  \{P \ast cr3=r_1 \ast r=r_2 \ast [r_2](Q)\}
+  \{P \ast cr3\mapsto_{r} \;r_1 \ast r \mapsto_{r} \;r_2 \ast [r_2](Q)\}
   \texttt{mov}~\texttt{\%cr3},~r%\lstinline|mov %cr3, r| 
-  \{[r_1](P) \ast cr3=r_2 \ast r = r_2 \ast Q\}
+  \{[r_1](P) \ast cr3 \mapsto_{r} r_2 \ast r \mapsto_{r} r_2 \ast Q\}
 }
@@
     \inferrule*[right=Broken]{ }{
-    \{\mathsf{emp} \ast cr3=r_1 \ast r=r_2 \ast [r_2](Q)\}
+    \{\mathsf{emp} \ast cr3 \mapsto_{r} r_1 \ast r\mapsto_{r} r_2 \ast [r_2](Q)\}
     \texttt{mov}~\texttt{\%cr3},~r%\lstinline|mov %cr3, r| 
-    \{[r_1](\mathsf{emp}) \ast cr3=r_1 \ast r=r_2 \ast Q\}
+    \{[r_1](\mathsf{emp}) \ast cr3 \mapsto_{r} r_1 \ast r \mapsto_{r} r_2 \ast Q\}
     }
   }{
-    \{a\mapsto_\mathsf{v} x \ast \mathsf{emp} \ast cr3=r_1 \ast r=r_2 \ast [r_2](Q)\}
+    \{a\mapsto_\mathsf{v} x \ast \mathsf{emp} \ast cr3\mapsto_{r}r_1 \ast r \mapsto_{r}r_2 \ast [r_2](Q)\}
     \texttt{mov}~\texttt{\%cr3},~r%\lstinline|mov %cr3, r| 
-    \{a\mapsto_\mathsf{v} x \ast [r_1](\mathsf{emp}) \ast cr3=r_1 \ast r=r_2 \ast Q\}
+    \{a\mapsto_\mathsf{v} x \ast [r_1](\mathsf{emp}) \ast cr3 \mapsto_{r} r_1 \ast r \mapsto_{r} r_2 \ast Q\}
   }
@@
 \inferrule[ChangeAddressSpace]{
-  \{[r_1](P) \ast cr3=r_2 \ast r = r_2 \ast Q\}\overline{is}
+  \{[r_1](P) \ast cr3 \mapsto_{r} r_2 \ast r \mapsto_{r} r_2 \ast Q\}\overline{is}
 }{
-  \{P \ast cr3=r_1 \ast r=r_2 \ast [r_2](Q)\}
+  \{P \ast cr3 \mapsto_{r} r_1 \ast r \mapsto_{r} r_2 \ast [r_2](Q)\}
   \texttt{mov}~\texttt{\%cr3},~r;\;\overline{is}
@@
 Virtual points-to assertions could be made self-stable by also giving them partial ownership over \lstinline|cr3| assertions, but this would require explicitly plumbing that ownership from \emph{all} assertions back to any place an address space change might occur; this would seem to be a far graver loss of modularity than this extra quirk in framing discussions.
-
+\todo{inlined, color= red}[STOP HERE, go back fix the stuff discussed with COLIN]
 \begin{figure}
 \begin{mathpar}
-\inferrule[ChangeAddressSpace]{
+\inferrule[WriteToMemoryFromReg]{
+  \{[r_1](P) \ast cr3=r_2 \ast r = r_2 \ast Q\}\overline{is}
+}{
+  \{P \ast cr3=r_1 \ast r=r_2 \ast [r_2](Q)\}
+  \texttt{mov}~\texttt{\%cr3},~r;\;\overline{is}
+  %\lstinline|mov %cr3, r| 
+} 
+\inferrule[WriteToRegFromReg]{
+  \{[r_1](P) \ast cr3=r_2 \ast r = r_2 \ast Q\}\overline{is}
+}{
+  \{P \ast cr3=r_1 \ast r=r_2 \ast [r_2](Q)\}
+  \texttt{mov}~\texttt{\%cr3},~r;\;\overline{is}
+  %\lstinline|mov %cr3, r| 
+} \\ 
+\inferrule[WriteToRegFromMem]{
+  \{[r_1](P) \ast cr3=r_2 \ast r = r_2 \ast Q\}\overline{is}
+}{
+  \{P \ast cr3=r_1 \ast r=r_2 \ast [r_2](Q)\}
+  \texttt{mov}~\texttt{\%cr3},~r;\;\overline{is}
+  %\lstinline|mov %cr3, r| 
+}
+\inferrule[WriteToCtlFromReg]{
   \{[r_1](P) \ast cr3=r_2 \ast r = r_2 \ast Q\}\overline{is}

### 2023-03-03 programlogic.tex
+

@@
   \item Page Map Level-4 Translation (L4): Calculating the offset to the 4 \textsf{Kbyte Physical Page} table to obtain the  physical address by  which resides in the last 12 bits of the virtual memory address, $\locsf|^{12}$,  shown in Figure \ref{}.
-    % \pfpointsto{\lvlsum\crt\lfoff}{\lvlbor\lt}\qfracfotsss\ppts \star
+    % \pfpointsto{\lvlsum\crt\lfoff}{\lvlbor\lt}\qfracfotsss\ppts \ast
     \[ \textsf{L4translate} \stackrel{\triangle}{=} \exists \entryf \ldotp \nfpointsto{\mask\vaddr\ft\crthree}{\mask\vaddr\tw\crthree}\entryf\qfracfotsss\naddr \]
@@
 
-  \ulcorner \textsf{aligned } \vaddr \urcorner \star  \\
+  \ulcorner \textsf{aligned } \vaddr \urcorner \ast  \\
   \textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\crthree,\entryf,\entrytr,\entrytw,\entryo)\\
@@
 \textsf{where   }\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\crthree,\entryf,\entrytr,\entrytw,\entryo) \stackrel{\triangle}{=} \\
- \qquad\qquad \nfpointsto{\mask\vaddr\ft\crthree}{\mask\vaddr\tw\crthree}\entryf\qfracfotsss\naddr \star \\ 
- \qquad\qquad  \nfpointsto{\mask\vaddr\ft\entryf}{\mask\vaddr\tw\entryf}\entrytr\qfracfotss\naddr  \star  \\
-  \qquad\qquad \nfpointsto{\mask\vaddr\ft\entrytr}{\mask\vaddr\tw\entrytr}\entrytw\qfracfots\naddr \star \\
+ \qquad\qquad \nfpointsto{\mask\vaddr\ft\crthree}{\mask\vaddr\tw\crthree}\entryf\qfracfotsss\naddr \ast \\ 
+ \qquad\qquad  \nfpointsto{\mask\vaddr\ft\entryf}{\mask\vaddr\tw\entryf}\entrytr\qfracfotss\naddr  \ast  \\
+  \qquad\qquad \nfpointsto{\mask\vaddr\ft\entrytr}{\mask\vaddr\tw\entrytr}\entrytw\qfracfots\naddr \ast \\
   \qquad\qquad \nfpointsto{\mask\vaddr\ft\entrytw}{\mask\vaddr\tw\entrytw}\entryo\qfracfot\naddr 
@@
   \exists \entryo \ldotp 
-  \sumwalkabs\vaddr\qfrac\paddr \star 
+  \sumwalkabs\vaddr\qfrac\paddr \ast 
    \nfpointsto{\mask\vaddr\ft\entryo}{\mask\vaddr\tw\entryo}\vpage\qfracone\naddr
@@
   \mathcal{I}\textsf{ASpace} \stackrel{\triangle}{=} \lambda \; \crthree \ldotp
-  \exists\;\gammaPred \;\ldotp\; \ownGhost\gammaPred{\authfull{\ptableabswalk\ptablestore}} \star \\
+  \exists\;\gammaPred \;\ldotp\; \ownGhost\gammaPred{\authfull{\ptableabswalk\ptablestore}} \ast \\
   \bigast{(\vaddr, \entryo)\in \ptablestore}{\exists\;(\entryf\;,\entrytr\;,\entrytw)\ldotp \textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\crthree,\vaddr,\entryf,\entrytr,\entrytw,\entryo)}
@@
 \hbox{(\TirNameStyle{ModalContextInvariance})} \qquad
- [r](Q \star \ell \mapsto_{\textsf{p}} \vpage)  \;\vdash [r](Q) \star \ell \mapsto_{\textsf{p}} \vpage 
+ [r](Q \ast \ell \mapsto_{\textsf{p}} \vpage)  \;\vdash [r](Q) \ast \ell \mapsto_{\textsf{p}} \vpage 
   \end{array}
@@
 \[
-\{ \crthree \mapsto_{r} r0 \star \textsf{va1} \mapsto_{v} \textsf{v}_1 \star \textsf{va2} \mapsto_{v} \textsf{v}_2 \star \textsf{va3} \mapsto_{v} \textsf{v}_3 \star [r1]\;P \}
+\{ \crthree \mapsto_{r} r0 \ast \textsf{va1} \mapsto_{v} \textsf{v}_1 \ast \textsf{va2} \mapsto_{v} \textsf{v}_2 \ast \textsf{va3} \mapsto_{v} \textsf{v}_3 \ast [r1]\;P \}
 \]
-where we see the \textit{current address-space} rooted at $r0$ is loaded into the  \textit{current-view} of the memory with its resources ($ \textsf{va1} \mapsto_{v} \textsf{v}_1 \star \textsf{va2} \mapsto_{v} \textsf{v}_2 \star \textsf{va3} \mapsto_{v} \textsf{v}_3$), and framed by the address-space rooted at $r1$ with resources $P$.
+where we see the \textit{current address-space} rooted at $r0$ is loaded into the  \textit{current-view} of the memory with its resources ($ \textsf{va1} \mapsto_{v} \textsf{v}_1 \ast \textsf{va2} \mapsto_{v} \textsf{v}_2 \ast \textsf{va3} \mapsto_{v} \textsf{v}_3$), and framed by the address-space rooted at $r1$ with resources $P$.
 Now, we need to a rule capturing the notion that switching the address spaces by pulling the resources ($P$) out of the one under modality into the \textit{“current view”} of memory, and leaves other assertions true relative to the previous address space (the one rooted at $r0$).
@@
 Virtual points-to assertions could be made self-stable by also giving them partial ownership over \lstinline|cr3| assertions, but this would require explicitly plumbing that ownership from \emph{all} assertions back to any place an address space change might occur; this would seem to be a far graver loss of modularity than this extra quirk in framing discussions.
-\todo{inlined, color= red}[STOP HERE, go back fix the stuff discussed with COLIN]
 \begin{figure}
 \begin{mathpar}
-\inferrule[WriteToMemoryFromReg]{
-  \{[r_1](P) \ast cr3=r_2 \ast r = r_2 \ast Q\}\overline{is}
+\inferrule[WriteToRegFromReg]{
+  \{P \ast r_d \mapsto_{r} \texttt{rvs} \ast r_s \mapsto_{r}\{q\} \texttt{ rvs} \}\;\overline{ is}
 }{
-  \{P \ast cr3=r_1 \ast r=r_2 \ast [r_2](Q)\}
-  \texttt{mov}~\texttt{\%cr3},~r;\;\overline{is}
+  \{P \ast r_d \mapsto_{r} \texttt{rvd} \ast r_s \mapsto_{r}\{q\} \texttt{ rvs} \}
+  \texttt{ mov}~\texttt{r}_d,~\texttt{r}_s;\;\overline{is}
   %\lstinline|mov %cr3, r| 
 }
-\inferrule[WriteToRegFromReg]{
-  \{[r_1](P) \ast cr3=r_2 \ast r = r_2 \ast Q\}\overline{is}
+\\
+\inferrule[WriteToMemFromReg]{
+  \{P \ast r_s \mapsto_{r}\{q\}  \texttt{rvs} \ast \lstinline|cr3| \mapsto_{r} \texttt{cr3val}  \ast r_a \mapsto_{r} \{q\} \texttt{ vaddr} \ast \texttt{vaddr} \mapsto_{t,\texttt{cr3val}} \texttt{v} \}\;\overline{ is}
 }{
-  \{P \ast cr3=r_1 \ast r=r_2 \ast [r_2](Q)\}
-  \texttt{mov}~\texttt{\%cr3},~r;\;\overline{is}
-  %\lstinline|mov %cr3, r| 
-} \\ 
+  \{P \ast r_s \mapsto_{r}\{q\}  \texttt{rvs} \ast \lstinline|cr3| \mapsto_{r} \texttt{cr3val}  \ast r_a \mapsto_{r}\{q\} \texttt{ vaddr} \ast \texttt{vaddr} \mapsto_{t,\texttt{cr3val}} \texttt{rvs} \}
+  \texttt{ mov}~\texttt{r}_a,~\texttt{r}_s;\;\overline{is}
+}
+\\
 \inferrule[WriteToRegFromMem]{
-  \{[r_1](P) \ast cr3=r_2 \ast r = r_2 \ast Q\}\overline{is}
+  \{P \ast r_d \mapsto_{r}  \texttt{v} \ast \lstinline|cr3| \mapsto_{r} \texttt{cr3val}  \ast r_a \mapsto_{r} \{q\} \texttt{ vaddr} \ast \texttt{vaddr} \mapsto_{t,\texttt{cr3val}} \texttt{v} \}\;\overline{is}
 }{
-  \{P \ast cr3=r_1 \ast r=r_2 \ast [r_2](Q)\}
-  \texttt{mov}~\texttt{\%cr3},~r;\;\overline{is}
-  %\lstinline|mov %cr3, r| 
+  \{P \ast r_d \mapsto_{r}  \texttt{rvd} \ast \lstinline|cr3| \mapsto_{r} \texttt{cr3val}  \ast r_a \mapsto_{r} \{q\} \texttt{ vaddr} \ast \texttt{vaddr} \mapsto_{t,\texttt{cr3val}} \texttt{v} \}
+  \texttt{mov}~\texttt{r}_d,~\texttt{r}_a;\;\overline{is}
 }
-\inferrule[WriteToCtlFromReg]{
-  \{[r_1](P) \ast cr3=r_2 \ast r = r_2 \ast Q\}\overline{is}
+\\
+\inferrule[WriteToRegCtlFromReg]{
+  \{P \ast r_s \mapsto_{r}\{q\}  \texttt{ rvs} \ast \lstinline|cr3| \mapsto_{r} \texttt{rvs}  \}\overline{is}
 }{
-  \{P \ast cr3=r_1 \ast r=r_2 \ast [r_2](Q)\}
-  \texttt{mov}~\texttt{\%cr3},~r;\;\overline{is}
+  \{P \ast r_s \mapsto_{r}\{q\}  \texttt{ rvs} \ast \lstinline|cr3| \mapsto_{r} \texttt{cr3val}  \}
+  \texttt{mov}~\texttt{cr3},~r_s;\;\overline{is}
   %\lstinline|mov %cr3, r| 

### 2023-03-03 programlogic.tex
fixed compilation

@@
 \begin{array}{l}
-  \vaddr\mapsto_{\textsf{v}}\{\textsf{q}\}\vpage \stackrel{def}{=} \lambda \crthree.\\
+  \vaddr\mapsto_{\textsf{t}}\{\textsf{q}\}\vpage \stackrel{def}{=} \lambda \crthree.\\
   \exists_{\entryf \;, \entrytr \;, \entrytw \;,\entryo} \ldotp
@@
 \inferrule[WriteToRegFromReg]{
-  \{P \ast r_d \mapsto_{r} \texttt{rvs} \ast r_s \mapsto_{r}\{q\} \texttt{ rvs} \}\;\overline{ is}
+  \{P \ast r_d \mapsto_{r} \textsf{rvs} \ast r_s \mapsto_{r}\{q\} \textsf{ rvs} \}\;\overline{ is}
 }{
-  \{P \ast r_d \mapsto_{r} \texttt{rvd} \ast r_s \mapsto_{r}\{q\} \texttt{ rvs} \}
-  \texttt{ mov}~\texttt{r}_d,~\texttt{r}_s;\;\overline{is}
+  \{P \ast r_d \mapsto_{r} \textsf{rvd} \ast r_s \mapsto_{r}\{q\} \textsf{ rvs} \}
+  \textsf{ mov}~\textsf{r}_d,~\textsf{r}_s;\;\overline{is}
   %\lstinline|mov %cr3, r| 
@@
 \inferrule[WriteToMemFromReg]{
-  \{P \ast r_s \mapsto_{r}\{q\}  \texttt{rvs} \ast \lstinline|cr3| \mapsto_{r} \texttt{cr3val}  \ast r_a \mapsto_{r} \{q\} \texttt{ vaddr} \ast \texttt{vaddr} \mapsto_{t,\texttt{cr3val}} \texttt{v} \}\;\overline{ is}
+  \{P \ast r_s \mapsto_{r}\{q\}  \textsf{rvs} \ast \textsf{cr3} \mapsto_{r} \textsf{cr3val}  \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{t},\textsf{cr3val}} \textsf{v} \}\;\overline{ is}
 }{
-  \{P \ast r_s \mapsto_{r}\{q\}  \texttt{rvs} \ast \lstinline|cr3| \mapsto_{r} \texttt{cr3val}  \ast r_a \mapsto_{r}\{q\} \texttt{ vaddr} \ast \texttt{vaddr} \mapsto_{t,\texttt{cr3val}} \texttt{rvs} \}
-  \texttt{ mov}~\texttt{r}_a,~\texttt{r}_s;\;\overline{is}
+  \{P \ast r_s \mapsto_{r}\{q\}  \textsf{rvs} \ast \textsf{cr3} \mapsto_{r} \textsf{cr3val}  \ast r_a \mapsto_{r}\{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{t},\textsf{cr3val}} \textsf{rvs} \}
+  \textsf{ mov}~\textsf{r}_a,~\textsf{r}_s;\;\overline{is}
 }
@@
 \inferrule[WriteToRegFromMem]{
-  \{P \ast r_d \mapsto_{r}  \texttt{v} \ast \lstinline|cr3| \mapsto_{r} \texttt{cr3val}  \ast r_a \mapsto_{r} \{q\} \texttt{ vaddr} \ast \texttt{vaddr} \mapsto_{t,\texttt{cr3val}} \texttt{v} \}\;\overline{is}
+  \{P \ast r_d \mapsto_{r}  \textsf{v} \ast \textsf{cr3} \mapsto_{r} \textsf{cr3val}  \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{t},\textsf{cr3val}} \textsf{v} \}\;\overline{is}
 }{
-  \{P \ast r_d \mapsto_{r}  \texttt{rvd} \ast \lstinline|cr3| \mapsto_{r} \texttt{cr3val}  \ast r_a \mapsto_{r} \{q\} \texttt{ vaddr} \ast \texttt{vaddr} \mapsto_{t,\texttt{cr3val}} \texttt{v} \}
-  \texttt{mov}~\texttt{r}_d,~\texttt{r}_a;\;\overline{is}
+  \{P \ast r_d \mapsto_{r}  \textsf{rvd} \ast \textsf{cr3} \mapsto_{r} \textsf{cr3val}  \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{t},\textsf{cr3val}} \textsf{v} \}
+  \textsf{mov}~\textsf{r}_d,~\textsf{r}_a;\;\overline{is}
 }
@@
 \inferrule[WriteToRegCtlFromReg]{
-  \{P \ast r_s \mapsto_{r}\{q\}  \texttt{ rvs} \ast \lstinline|cr3| \mapsto_{r} \texttt{rvs}  \}\overline{is}
+  \{P \ast r_s \mapsto_{r}\{q\}  \textsf{ rvs} \ast \textsf{cr3} \mapsto_{r} \textsf{rvs}  \}\overline{is}
 }{
-  \{P \ast r_s \mapsto_{r}\{q\}  \texttt{ rvs} \ast \lstinline|cr3| \mapsto_{r} \texttt{cr3val}  \}
-  \texttt{mov}~\texttt{cr3},~r_s;\;\overline{is}
+  \{P \ast r_s \mapsto_{r}\{q\}  \textsf{ rvs} \ast \textsf{cr3} \mapsto_{r} \textsf{cr3val}  \}
+  \textsf{mov}~\textsf{cr3},~r_s;\;\overline{is}
   %\lstinline|mov %cr3, r| 
@@
 \end{figure}
+\paragraph{Register To Memory}
+\paragraph{Memory To Register}
+\paragraph{Register To Register}
+\todo[inline,color=red]{STOP HERE}

### 2023-03-03 programlogic.tex
sync with desk

@@
 \inferrule[WriteToMemFromReg]{
-  \{P \ast r_s \mapsto_{r}\{q\}  \textsf{rvs} \ast \textsf{cr3} \mapsto_{r} \textsf{cr3val}  \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{t},\textsf{cr3val}} \textsf{v} \}\;\overline{ is}
+  \{P \ast r_s \mapsto_{r}\{q\}  \textsf{rvs} \ast \textsf{cr3} \mapsto_{r} \textsf{cr3val}  \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast (\textsf{vaddr} \mapsto_{\textsf{t}} \textsf{v})\;\textsf{cr3val} \}\;\overline{ is}
 }{
-  \{P \ast r_s \mapsto_{r}\{q\}  \textsf{rvs} \ast \textsf{cr3} \mapsto_{r} \textsf{cr3val}  \ast r_a \mapsto_{r}\{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{t},\textsf{cr3val}} \textsf{rvs} \}
+  \{P \ast r_s \mapsto_{r}\{q\}  \textsf{rvs} \ast \textsf{cr3} \mapsto_{r} \textsf{cr3val}  \ast r_a \mapsto_{r}\{q\} \textsf{ vaddr} \ast (\textsf{vaddr} \mapsto_{\textsf{t}} \textsf{rvs})\;\textsf{cr3val} \}
   \textsf{ mov}~\textsf{r}_a,~\textsf{r}_s;\;\overline{is}
@@
 \inferrule[WriteToRegFromMem]{
-  \{P \ast r_d \mapsto_{r}  \textsf{v} \ast \textsf{cr3} \mapsto_{r} \textsf{cr3val}  \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{t},\textsf{cr3val}} \textsf{v} \}\;\overline{is}
+  \{P \ast r_d \mapsto_{r}  \textsf{v} \ast \textsf{cr3} \mapsto_{r} \textsf{cr3val}  \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast (\textsf{vaddr} \mapsto_{\textsf{t}} \textsf{v})\;\textsf{cr3val} \}\;\overline{is}
 }{
-  \{P \ast r_d \mapsto_{r}  \textsf{rvd} \ast \textsf{cr3} \mapsto_{r} \textsf{cr3val}  \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{t},\textsf{cr3val}} \textsf{v} \}
+  \{P \ast r_d \mapsto_{r}  \textsf{rvd} \ast \textsf{cr3} \mapsto_{r} \textsf{cr3val}  \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast (\textsf{vaddr} \mapsto_{\textsf{t}} \textsf{v})\;\textsf{cr3val} \}
   \textsf{mov}~\textsf{r}_d,~\textsf{r}_a;\;\overline{is}

### 2023-03-05 programlogic.tex
+

@@
 The truth on an address space, exhibits itself as a contingent truth: a location virtualization assertion (a virtual points-to in \ref{fig:virtualpointstosharing}),  happens to valid in the \textit{current world} (in the current address space), and switching address spaces pulls information out of one \textit{world} into the “current view” of memory, and leaves other assertions true relative to the previous address space.
-
+\begin{figure}
+   \includegraphics[width=\columnwidth]{addr_space_switch.png}
+  \caption{Loading an Address-Space into the Current Memory View}
+  \label{fig:addrswitch}
+  \end{figure}
 Therefore, an address-space, as an abstraction, can be treated as naming the memory state as a modal frame and the choice of page table root as a world in Kripke-style semantics.

### 2023-03-23 programlogic.tex
+

@@
     % \pfpointsto{\lvlsum\crt\lfoff}{\lvlbor\lt}\qfracfotsss\ppts \ast
-    \[ \textsf{L4translate} \stackrel{\triangle}{=} \exists \entryf \ldotp \nfpointsto{\mask\vaddr\ft\crthree}{\mask\vaddr\tw\crthree}\entryf\qfracfotsss\naddr \]
+    \[ \begin{array}{l}
+      \hbox{(\TirNameStyle{{L4translate}})} \quad
+      \exists \entryf \ldotp \nfpointsto{\mask\vaddr\ft\crthree}{\mask\vaddr\tw\crthree}\entryf\qfracfotsss\naddr
+      \end{array}
+      \]
   \item Page Directory Pointer Level Translation (L3):
-    \[ \textsf{L3translate} \stackrel{\triangle}{=} \exists \entrytr \ldotp \nfpointsto{\mask\vaddr\ft\entryf}{\mask\vaddr\tw\entryf}\entrytr\qfracfotss\naddr \]
+    \[\begin{array}{l}
+    \hbox{(\TirNameStyle{{L3translate}})} \quad
+    \exists \entrytr \ldotp \nfpointsto{\mask\vaddr\ft\entryf}{\mask\vaddr\tw\entryf}\entrytr\qfracfotss\naddr
+    \end{array}
+    \]
   \item Page Directory Table Level Translation (L2):  we obtain the address of the page table via \textsf{PDE}$|^{52}$ and picking the offset from the virtual memory address via imposing the mask to obtain page table offset bits (9 bits [20-12] shown in Figure \ref{}) to obtain the page table entry (\textsf{PTE}).
-    \[ \textsf{L2translate} \stackrel{\triangle}{=} \exists \entrytw \ldotp \nfpointsto{\mask\vaddr\ft\entrytr}{\mask\vaddr\tw\entrytr}\entrytw\qfracfots\naddr \]
+    \[ \begin{array}{l}
+            \hbox{(\TirNameStyle{{L2translate}})} \quad 
+       \exists \entrytw \ldotp \nfpointsto{\mask\vaddr\ft\entrytr}{\mask\vaddr\tw\entrytr}\entrytw\qfracfots\naddr
+      \end{array}
+      \]
   \item Page Table Level Translation (L1):
-    \[ \textsf{L1translate}  \stackrel{\triangle}{=} \exists \entryo \ldotp \nfpointsto{\mask\vaddr\ft\entrytw}{\mask\vaddr\tw\entrytw}\entryo\qfracfot\naddr \]
+    \[ \begin{array}{l}
+      \hbox{(\TirNameStyle{{L1translate}})} \quad
+      \exists \entryo \ldotp \nfpointsto{\mask\vaddr\ft\entrytw}{\mask\vaddr\tw\entrytw}\entryo\qfracfot\naddr
+      \end{array}
+      \]
   \item Page Address Level Translation
-    \[ \textsf{PageAccess} \stackrel{\triangle}{=} \exists \vpage \ldotp \nfpointsto{\mask\vaddr\ft\entryo}{\mask\vaddr\tw\entryo}\vpage\qfracone\naddr \]
+    \[ \begin{array}{l}
+      \hbox{(\TirNameStyle{{PageLebelAccess}})} \quad
+      \exists \vpage \ldotp \nfpointsto{\mask\vaddr\ft\entryo}{\mask\vaddr\tw\entryo}\vpage\qfracone\naddr
+      \end{array}
+      \]
 
@@
 
-\textsf{where   }\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\crthree,\entryf,\entrytr,\entrytw,\entryo) \stackrel{\triangle}{=} \\
+\textsf{where   }\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\vaddr,\crthree,\entryf,\entrytr,\entrytw,\entryo) \stackrel{\triangle}{=} \\
  \qquad\qquad \nfpointsto{\mask\vaddr\ft\crthree}{\mask\vaddr\tw\crthree}\entryf\qfracfotsss\naddr \ast \\ 
@@
   \label{sec:sharingpages}  
-  We already build the intuition on what a virtual address mapping is composed of: 1.
-traversing the page tables to reach the page table entry 2.
-obtaining the physical address, and accessing the value mapped.
+  One might have already observed that the virtual points-to definition shown in Figure \ref{fig:strongvirtualpointsto} is too strong to specify virtual memory operations updating a page table entry.
+To speak more concretely, the fractional permissions on the page-table-walk assertion (\textsf{L}$_4$\_\textsf{L}$_{1}$ in Figure \ref{fig:strongvirtualpointsto}) \textit{only} assure the existence of the certain number of mappings -- 512 at each level -- abstracting virtual address translations, and updating any of these mappings would make the other mappings, address translations sharing the page table mappings.
+Therefore, we need to have logical abstractions ensuring sound updates to the page-table entries in the existence of sharing.
 
-One might have already observed that the virtual points-to definition shown in Figure \ref{fig:strongvirtualpointsto} is too strong to specify virtual memory operations updating a page table entry.
-Although there seems to fractional permissions for fragmental ownership of page-table entries, strong virtual points-to (in Figure \ref{strongvirtualpointsto}) relation doesn’t care about owning those fractional resources, it only cares that something ensures the correct page table walk exists.
-Any virtual memory operations such as mapping a new page which updates an page table entry would violate the soundness of other virtual-pointsto relations in the existence of page-sharing across the virtual addresses in the same address space.
-
-One possible approach to is to just abstracting the \textit{valid} page-table walk existence from a each virtual address (e.g.
-\textsf{va}) to a physical page address, as \textit{per-element} ownership of a logical \textit{page-table-walk-summarization-map}
+ One possible approach is to abstract away the physical page-table-walk pointsto (\textsf{L}$_4$\_\textsf{L}$_{1}$ in Figure \ref{fig:strongvirtualpointsto}) from the virtual pointsto relation via replacing it with read-only abstract map entries -- a ghost (abstract) map entry per \textsf{L}$_4$\_\textsf{L}$_{1}$ mappings -- which can be used as a resource ($ \sumwalkabs\vaddr\qfrac\paddr$) whose ownership ensures that the virtual address ($\vaddr$) is mapped to a page (referenced by $\paddr$) ,and as a token to access to the physical page-table-walk pointsto (\textsf{L}$_4$\_\textsf{L}$_{1}$) stored in an address-space invariant.
+We use \textsf{Iris}'s authorative ghost-map logical machinery
 \[\mathcal{A}\textsf{bsPTableWalk} \stackrel{\triangle}{=} \textsc{Auth} (\; \mathcal{W}_{64} \;\rightarrow_{\textrm{fin}} \;  (\mathcal{W}_{64} \; \rightarrow_{\textrm{fin}}\;  (\textsc{Frac }, \mathord{+}) \times (\textsc{Agree } \Loc,\mathord{=}) ))\]
-provided by a fragmentery ghost resrource ownership of $\mathcal{A}\textsf{bsPTableWalk}$.
+to realize these reasoning principles ($\mathcal{A}\textsf{bsPTableWalk}$).
 \begin{remark}[Exchanging Page-Table-Walk-Summarization Mappings (Tokens) with Physical Page-Table-Walk Pointstos ]
-$\mathcal{A}\textsf{PTableWalk}$ is a ghost map resource where fragmental-ownership of an individual key-value pair (per-element) can be handled as a resource.
-We use this mechanism to separate the virtual-to-physical translation from the physical memory involved: virtual points-to assertions can be defined in terms of fragmental ownership (ownership of an element) of these partial ghost map entry assertions , $\sumwalkabs\vaddr\qfrac\paddr$, plus sharing an invariant ($\mathcal{I}$\textsf{ASpace} in Figure \ref{fig:peraspaceinvariant}) that the current installed page table respects all entries of the mapping, $\ownGhost\gammaPred{\authfull{\ptableabswalk\ptablestore}}$.
-From one another perspective, unmapping pages collects the fragmental map kvpairs, e.g.
-$\sumwalkabs\vaddr\qfrac\paddr$, from collecting the assertions, and then can remove them from the ghost map and update the page tables.
-Critically, physical full ownership of the page tables lives, $\ownGhost\gammaPred{\authfull{\ptableabswalk\ptablestore}}$, in the central invariant on the current page table, so some virtual pointsto assertions can refer to memory in those page tables.% This still works with the modality, since that invariant is also semantically a predicate on a page table root.
+$\mathcal{A}\textsf{PTableWalk}$'s fragmental-ownership represents an individual key-value pair (per-element) which can be handled as a resource.
+Critically, physical full ownership of the page tables lives, $\ownGhost\gammaPred{\authfull{\ptableabswalk\ptablestore}}$, in the central invariant on the current page table, so some virtual pointsto assertions can refer to memory in those page tables.
+We use this mechanism to separate the virtual-to-physical translation from the physical memory involved: virtual points-to assertions can be defined in terms of fragmental ownership (ownership of an element) of these partial ghost map entry assertions , $\sumwalkabs\vaddr\qfrac\paddr$, and physical page mapping ($nfpointsto{\mask\vaddr\ft\entryo}{\mask\vaddr\tw\entryo}\vpage\qfracone\naddr$).
+Physical page-table-walk pointsto, unlike the case in the strong virtual-pointsto relation in Figure \ref{fig:strongvirtualpointsto}, is moved into an invariant ($\mathcal{I}$\textsf{ASpace} in Figure \ref{fig:peraspaceinvariant}), and resides with authorative ownership of $\mathcal{A}\textsf{PTableWalk}$ ($\ownGhost\gammaPred{\authfull{\ptableabswalk\ptablestore}}$) which respects all entries of the abstract virtual-address translation mappings ($\sumwalkabs\vaddr\qfrac\paddr$).
+To build more intuition, we can see the use-case \textit{unmapping-pages} as collecting the fragmental map pairs, e.g.
+$\sumwalkabs\vaddr\qfrac\paddr$, from collecting the assertions, removing them from the ghost map, and then updating the page tables.
+% This still works with the modality, since that invariant is also semantically a predicate on a page table root.
 \end{remark}
@@
 \end{enumerate}
+\begin{figure}
+   \includegraphics[width=0.75\columnwidth]{logical_addr_space.png}
+  \caption{Logical Resources of Address-Space Abstraction}
+  \label{fig:logicaladdrspace}
+  \end{figure}
     \subsubsection{Global Per-Address-Space Invariant}
@@
 For each virtual address \textit{mapped} with a physical page table entry, abstracted with ghost mappings ($\sumwalkabs\vaddr\qfrac\paddr$) in virtul-pointsto relations as shown in Figure \ref{fig:virtualpointstosharing}), there exsits a physical page table walk.
-    \[ \bigast{(\vaddr, \entryo)\in \ptablestore}{\exists\;(\entryf\;,\entrytr\;,\entrytw)\ldotp \textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\crthree,\vaddr,\entryf,\entrytr,\entrytw,\entryo)} \]
+    \[ \bigast{(\vaddr, \entryo)\in \ptablestore}{\exists\;(\entryf\;,\entrytr\;,\entrytw)\ldotp \textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\vaddr,\crthree,\entryf,\entrytr,\entrytw,\entryo)} \]
     By separating the virtual-points Overlapping address spaces — both shared data, and shared page tables
@@
   \exists\;\gammaPred \;\ldotp\; \ownGhost\gammaPred{\authfull{\ptableabswalk\ptablestore}} \ast \\
-  \bigast{(\vaddr, \entryo)\in \ptablestore}{\exists\;(\entryf\;,\entrytr\;,\entrytw)\ldotp \textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\crthree,\vaddr,\entryf,\entrytr,\entrytw,\entryo)}
+  \bigast{(\vaddr, \entryo)\in \ptablestore}{\exists\;(\entryf\;,\entrytr\;,\entrytw)\ldotp \textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\vaddr,\crthree,\entryf,\entrytr,\entrytw,\entryo)}
 \end{array}
@@
     \hbox{(\TirNameStyle{ModalAddressSpace})} \\ \qquad
-         [r](P)  \;\stackrel{\triangle}{=}  (\mathcal{W}_{64} -\textsf{b}> \textsf{uPredI} (\textsf{iResUR} \;\Sigma)) -\textsf{b}> \textsf{uPredI} \_ := \lambda \textsf{P},\; \textsf{P r}. 
+         [r](P)  \;\stackrel{\triangle}{=}  (\mathcal{W}_{64} \rightarrow_{\textsf{b}} \textsf{uPredI} (\textsf{iResUR} \;\Sigma)) \rightarrow_{\textsf{b}} \textsf{uPredI } \_ := \lambda \textsf{P},\; \textsf{P r}.
   \end{array}

### 2023-03-23 programlogic.tex
+

@@
 \label{sec:aspacemodal}
-Abstractions we introduce for virtual address spaces are based on the following aspects:
+Abstractions we introduce for virtual address spaces are based on the following aspects pictorially shown in Figure \ref{fig:logicaladdrspace}:
 \begin{enumerate}
-\item a global address space invariant asserting the authorative ownership of the ghost map, $\ownGhost\gammaPred{\authfull{\ptableabswalk\ptablestore}}$ 
-\item for each ghost page-table-walk mapping,  $\sumwalkabs\vaddr\qfrac\paddr$, there exists concrete physical page-table accesses path
+
+\item a global address space invariant (solid round square labelled as $\mathcal{I}$\textsf{ASpace} in Figure \ref{fig:peraspaceinvariant}) asserting the authorative ownership of the ghost map, $\ownGhost\gammaPred{\authfull{\ptableabswalk\ptablestore}}$ \textit{represents} the full-ownership (solid-round-head arrow) of $\mathcal{A}\textsf{PTableWalk}$
+\item per-virtual-pointsto relation ((cutted round square labelled as \textsf{Virtual PointsTo} in Figure \ref{fig:peraspaceinvariant})) asserting the fragmental ownership of the ghost map, $\sumwalkabs\vaddr\qfrac\paddr$ \textit{represents} the fragmental-ownership (cutted-diamond-head arrow) of $\mathcal{A}\textsf{PTableWalk}$
+\item for each ghost page-table-walk mapping,  $\sumwalkabs\vaddr\qfrac\paddr$, there exists one shared persistent invariant (shown as double-ended solid arrow 1 to N)
   \[\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\crthree,\entryf,\entrytr,\entrytw,\entryo)\]

### 2023-03-24 programlogic.tex
+

@@
   One might have already observed that the virtual points-to definition shown in Figure \ref{fig:strongvirtualpointsto} is too strong to specify virtual memory operations updating a page table entry.
-To speak more concretely, the fractional permissions on the page-table-walk assertion (\textsf{L}$_4$\_\textsf{L}$_{1}$ in Figure \ref{fig:strongvirtualpointsto}) \textit{only} assure the existence of the certain number of mappings -- 512 at each level -- abstracting virtual address translations, and updating any of these mappings would make the other mappings, address translations sharing the page table mappings.
+To speak more concretely, the fractional permissions on the page-table-walk assertion ($textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$ in Figure \ref{fig:strongvirtualpointsto}) \textit{only} assure the existence of the certain number of mappings -- 512 at each level -- abstracting virtual address translations, and updating any of these mappings would make the other mappings, address translations sharing the page table mappings.
 Therefore, we need to have logical abstractions ensuring sound updates to the page-table entries in the existence of sharing.
 
- One possible approach is to abstract away the physical page-table-walk pointsto (\textsf{L}$_4$\_\textsf{L}$_{1}$ in Figure \ref{fig:strongvirtualpointsto}) from the virtual pointsto relation via replacing it with read-only abstract map entries -- a ghost (abstract) map entry per \textsf{L}$_4$\_\textsf{L}$_{1}$ mappings -- which can be used as a resource ($ \sumwalkabs\vaddr\qfrac\paddr$) whose ownership ensures that the virtual address ($\vaddr$) is mapped to a page (referenced by $\paddr$) ,and as a token to access to the physical page-table-walk pointsto (\textsf{L}$_4$\_\textsf{L}$_{1}$) stored in an address-space invariant.
+ One possible approach is to abstract away the physical page-table-walk pointsto ($\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$ in Figure \ref{fig:strongvirtualpointsto}) from the virtual pointsto relation via replacing it with read-only abstract map entries -- a ghost (abstract) map entry per \textsf{L}$_4$\_\textsf{L}$_{1}$ mappings -- which can be used as a resource ($ \sumwalkabs\vaddr\qfrac\paddr$) whose ownership ensures that the virtual address ($\vaddr$) is mapped to a page (referenced by $\paddr$) ,and as a token to access to the physical page-table-walk pointsto (\textsf{L}$_4$\_\textsf{L}$_{1}$) stored in an address-space invariant ($\mathcal{I}$\textsf{ASpace} in Figure \ref{fig:peraspaceinvariant}).
 We use \textsf{Iris}'s authorative ghost-map logical machinery
@@
 \begin{remark}[Exchanging Page-Table-Walk-Summarization Mappings (Tokens) with Physical Page-Table-Walk Pointstos ]
-$\mathcal{A}\textsf{PTableWalk}$'s fragmental-ownership represents an individual key-value pair (per-element) which can be handled as a resource.
-Critically, physical full ownership of the page tables lives, $\ownGhost\gammaPred{\authfull{\ptableabswalk\ptablestore}}$, in the central invariant on the current page table, so some virtual pointsto assertions can refer to memory in those page tables.
-We use this mechanism to separate the virtual-to-physical translation from the physical memory involved: virtual points-to assertions can be defined in terms of fragmental ownership (ownership of an element) of these partial ghost map entry assertions , $\sumwalkabs\vaddr\qfrac\paddr$, and physical page mapping ($nfpointsto{\mask\vaddr\ft\entryo}{\mask\vaddr\tw\entryo}\vpage\qfracone\naddr$).
-Physical page-table-walk pointsto, unlike the case in the strong virtual-pointsto relation in Figure \ref{fig:strongvirtualpointsto}, is moved into an invariant ($\mathcal{I}$\textsf{ASpace} in Figure \ref{fig:peraspaceinvariant}), and resides with authorative ownership of $\mathcal{A}\textsf{PTableWalk}$ ($\ownGhost\gammaPred{\authfull{\ptableabswalk\ptablestore}}$) which respects all entries of the abstract virtual-address translation mappings ($\sumwalkabs\vaddr\qfrac\paddr$).
-To build more intuition, we can see the use-case \textit{unmapping-pages} as collecting the fragmental map pairs, e.g.
+  $\mathcal{A}\textsf{PTableWalk}$'s fragmental-ownership represents an individual key-value pair (per-virtual-address) which can be handled as a \textit{look-up} resource for an individual virtual address.
+  To build more intuition, we can think of the use-case \textit{unmapping-pages} as collecting the fragmental map pairs, e.g.
 $\sumwalkabs\vaddr\qfrac\paddr$, from collecting the assertions, removing them from the ghost map, and then updating the page tables.
-% This still works with the modality, since that invariant is also semantically a predicate on a page table root.
+  Critically, physical full ownership ($\ownGhost\gammaPred{\authfull{\ptableabswalk\ptablestore}}$) of the page tables, as a \textit{singleton resource}, lives in the central invariant keeping physical-page-table pointstos per-virtual address translation.
+  With this mechanism we can separate the virtual-to-physical translation from the physical memory involved, and consequently
+  \begin{itemize}
+  \item virtual pointsto assertions, whose definition is shown in Figure \ref{fig:virtualpointstosharing}, can refer to their physical memory assertions representing their page-address translation in the physical-page-tables ($\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$) which reside in the address-space invariant ($\mathcal{I}$\textsf{ASpace} in Figure \ref{fig:peraspaceinvariant}) using individual key-value pair, i.e.
+instances of ($\sumwalkabs\vaddr\qfrac\paddr$)
+  \item unlike the case in the strong virtual-pointsto relation in Figure \ref{fig:strongvirtualpointsto}, physical-page-table-walk pointsto assertions ($\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$) for all virtual addresses are moved into the address-space invariant ($\mathcal{I}$\textsf{ASpace} in Figure \ref{fig:peraspaceinvariant}), and resides with authorative ownership of $\mathcal{A}\textsf{PTableWalk}$ ($\ownGhost\gammaPred{\authfull{\ptableabswalk\ptablestore}}$) which respects all entries of the abstract virtual-address translation mappings ($\sumwalkabs\vaddr\qfrac\paddr$).
+    \end{itemize}
 \end{remark}
-Then, our virtual memory addressing would be the fragmental ownership of the abstract page-table-walk mapping with the physical page address mapping as shown in Figure \ref{fig:virtualpointstosharing}.
 \begin{figure*}
@@
 \end{figure*}
-
 \subsection{Address-Space Abstraction}

### 2023-03-24 programlogic.tex
+

@@
 \end{figure*}
-\subsection{Address-Space Abstraction}
-\label{sec:aspacemodal}
-Abstractions we introduce for virtual address spaces are based on the following aspects pictorially shown in Figure \ref{fig:logicaladdrspace}:
+%\subsection{Address-Space Abstraction}
+%\label{sec:aspacemodal}
+
+Then, once we want to outline the distribution of the facts per-address, as we see it pictorially  in Figure \ref{fig:logicaladdrspace}, the outline is composed of the following aspects:
 \begin{enumerate}
-
-\item a global address space invariant (solid round square labelled as $\mathcal{I}$\textsf{ASpace} in Figure \ref{fig:peraspaceinvariant}) asserting the authorative ownership of the ghost map, $\ownGhost\gammaPred{\authfull{\ptableabswalk\ptablestore}}$ \textit{represents} the full-ownership (solid-round-head arrow) of $\mathcal{A}\textsf{PTableWalk}$
+\item a global address space invariant (solid round square labelled as $\mathcal{I}$\textsf{ASpace} in Figure \ref{fig:peraspaceinvariant}) asserting the authorative ownership (i.e.
+update capability), $\ownGhost\gammaPred{\authfull{\ptableabswalk\ptablestore}}$, of the ghost map -- shown as solid-round-head arrow to $\mathcal{A}\textsf{PTableWalk}$ Figure \ref{fig:logicaladdrspace}
 \item per-virtual-pointsto relation ((cutted round square labelled as \textsf{Virtual PointsTo} in Figure \ref{fig:peraspaceinvariant})) asserting the fragmental ownership of the ghost map, $\sumwalkabs\vaddr\qfrac\paddr$ \textit{represents} the fragmental-ownership (cutted-diamond-head arrow) of $\mathcal{A}\textsf{PTableWalk}$
@@
   \[\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\crthree,\entryf,\entrytr,\entrytw,\entryo)\]
-  \item all virtual pointsto assertions happen to be true in a certain address space, i.e.
-the truth on address virtualization is indexed by CPUs page table register ($\crthree$ for x86-64), and we present modal reasoning principles for updating $\crthree$ (switching from the current address space to another)
 \end{enumerate}
+
 \begin{figure}
@@
 \end{figure*}
+\subsection{Address-Space Management}
+\label{sec:aspacemanagement}
+% \item abstractions shaped-around address-space switch where all virtual pointsto assertions happen to be true in a certain address space, i.e.
+the truth on address virtualization is indexed by CPUs page table register ($\crthree$ for x86-64), and we present modal reasoning principles for updating $\crthree$ (switching from the current address space to another)
 \subsubsection{Address-Space Modality}

### 2023-03-29 programlogic.tex
+

@@
 \newcommand{\ptablestore}{\theta}
-Our reasoning principles are shaped around understanding the following main pillars of vritual-memory-management:
+Our reasoning principles are shaped around understanding the following perspectives on the vritual-memory abstraction:
 \begin{enumerate}
-\item page-table walk: L4-L1 walk on shared page tables to reach a physical page address
-\item allocating/updating a pages backing a virtual address 
+\item a bag of physical pages reached through a set of physical page-table (L4-L1 page-tables) acceses which are shared amongst different virtual addresses indexed by the root address of the address-space
+\item allocating/updating pages backing a virtual address 
 \item switching from one address-space to another 
@@
 \label{sec:aspacemanagement}
+With the reasoning principles introduced for the essential resource of address spaces, i.e. virtual memory addresses, now we can ...
 % \item abstractions shaped-around address-space switch where all virtual pointsto assertions happen to be true in a certain address space, i.e.

### 2023-03-29 programlogic.tex
+

@@
 \end{enumerate}
+Based on these perspectives, we coneptualize truth on an address-space as indexed by its roots address, therefore.
+we lift standard bi proposition to be indexed with $\mathcal{W}_{64}$ as shown in Figure \ref{fig:vprop}.
+Contesequently.
+all abstractions we explain in this section as part of our address-space logic are type $\textsf{vProp }\Sigma$.
+\begin{figure}[!ht]
+\[
+\begin{array}{cl}
+\textsf{vProp } \Sigma \; : \; \textsf{bi} \; \stackrel{def}{=} \mathcal{W}_{64} -\textsf{b}> \textsf{iPropI } \Sigma
+\end{array}
+\]
+\caption{$\textsf{vProp }\Sigma$: Root-Address Indexed Address-Space Proposition}
+  \label{fig:vprop}
+\end{figure}
 %\subsection{Machine State under Address Translation}
@@
 \label{sec:pointsto}
-As part of assertions on our machine state ($\sigma$), we abstract memory addresses and registers naming their relevant values through well-known separation logic assertion \textit{points-to}.
+We abstract memory addresses and registers naming their relevant values through well-known separation logic assertion \textit{points-to}.
 \begin{enumerate}

### 2023-03-29 intro.tex
arch file

@@
 
+\begin{figure}
+   \includegraphics[width=0.5\columnwidth]{architecture.png}
+  \caption{Layers of Implementation}
+  \label{fig:architecture}
+  \end{figure}
 We develop these ideas in an Iris-based\mytodo{cite JFP} separation logic, and verify simplified versions of several critical virtual-memory-related pieces of OS functionality, including mapping and unmapping pages, and describing the self-referential identity maps used in most kernels to allow convenient access to physical memory.

### 2023-03-29 programlogic.tex
+

@@
 \end{enumerate}
-Based on these perspectives, we coneptualize truth on an address-space as indexed by its roots address, therefore.
+Based on these perspectives, we coneptualize truth on an address-space through the contingency it exhibits: \textit{assertion P holds on this address-space indexed by its roots address}, Therefore.
 we lift standard bi proposition to be indexed with $\mathcal{W}_{64}$ as shown in Figure \ref{fig:vprop}.
-Contesequently.
-all abstractions we explain in this section as part of our address-space logic are type $\textsf{vProp }\Sigma$.
+In other words, all abstractions we explain in this section as part of our address-space logic are type $\textsf{vProp }\Sigma$.
 \begin{figure}[!ht]

### 2023-04-03 intro.tex
+

@@
 
-At its core, the virtual memory functionality of modern CPUs is about \emph{location virtualization}: the memory locations (addresses) seen by most code are not, in fact, the exact location in physical memory where data resides.
+At its core, the virtual memory functionality of modern CPUs is about \emph{location virtualization}: the memory locations (addresses) seen by most code are not, in fact, the exact location in physical memory where data reside.
 Instead these are \emph{virtual} addresses, which are mapped to actual physical resources by the cooperation of the hardware and OS.
@@
 We propose tackling the verification of virtual memory subsystems (and kernels more broadly) by adapting ideas from hybrid modal logic, which can label assertions true under \emph{other} circumstances (i.e., in another address space) with a modality indexed by a name for that space (in our case, the root of the page tables for an address space).
-This offers a convenient and powerful way to modularly (1) isolate assertions specific to a particular address space, (2) explicitly state when an assertion is true across address spaces, (3) manipulate address spaces from within other address spaces, and (4) reason about change in address spaces.
+This offers a \textit{convenient} and \textit{powerful way to modularly}
+\begin{itemize}
+\item isolate assertions specific to a particular address space,
+\item explicitly state when an assertion is true across address spaces,
+\item manipulate address spaces from within other address spaces, and
+\item reason about change in address spaces.
+\end{itemize}
   These advantages make this approach to reasoning about virtual memory more flexible than prior program logic techniques~\cite{kolanski08vstte,kolanski09tphols}, which was only able to work with a single address space (the current address space on the CPU).
@@
 \end{figure}
-We develop these ideas in an Iris-based\mytodo{cite JFP} separation logic, and verify simplified versions of several critical virtual-memory-related pieces of OS functionality, including mapping and unmapping pages, and describing the self-referential identity maps used in most kernels to allow convenient access to physical memory.
+We realize and show our reasoning principles, showin in Figure \ref{fig:architecture}
+\begin{itemize}
+\item we develop these ideas in an Iris-based\mytodo{cite JFP} separation logic, a logic for address-virtualization (\textsf{vProp} in Figure \ref{fig:architecture})
+\item we model \textsf{AMD64} in Coq, and specify and prove its instructions in \textsf{vProp}
+\item and verify simplified versions of several critical virtual-memory-related pieces of OS functionality, including mapping and unmapping pages, and describing the self-referential identity maps used in most kernels to allow convenient access to physical memory.
+\end{itemize}

### 2023-04-03 syntaxsemantics.tex
+
replace the figure

@@
-\section{Machine State, Instructions, Naming Registers and Memory Translation}
+\section{Machine State \& Syntax}
 \label{sec:syntax}
-To develop our core logical ideas and explain their intepretation, we need only a handful of instructions, and a machine model corresponding to execution of x86-64 assembly instructions with virtual memory enabled on the CPU.
+To develop our core logical ideas, we instantiate \textsf{Iris} with a simple language for streams of instructions, and a logical machine model corresponding to execution of x86-64 assembly instructions with virtual memory enabled on the CPU.
 
@@
 We use these identifiers to access the register values, $\regval \in \regvaltype$.
+\begin{figure}[!ht]
+\newcommand{\commentary}[1]{ & \text{\small\it #1} \\}
+\[
+  \begin{array}{r@{\;}c@{\;}l}
+    \loc & \in & \Loc \\
+    \reg & \in & \regset \\
+    \regval & \in & \regvaltype \\
+    \val & ::= & \vunit
 
-\input{fig-syntax}
+\\
+    \instrs & ::= &
+    \begin{array}[t]{@{}l@{\hspace{10mm}}l@{}}
+    \begin{array}[t]{@{}ll@{}}
+      \iskip
+                   \commentary{no-op}
+      \iseq\instr\instrs
+                   \commentary{sequencing}
+      \ising\instr
+                   \commentary{executing}             
+    \end{array}
+    %&
+    %\begin{array}[t]{@{}ll@{}}
+     % \ialloc\lval\allocsize
+      %             \commentary{heap allocation}
+    %\end{array}
+    \end{array}
+    \\
 
+    \ectx & ::= &
+      \hole \mid
+      \iseq\ectx\instrs 
+    \\
+  \end{array}
+\]
+\caption{Syntax}
+\Description{Syntax}
+\label{fig:syntax}
+\end{figure}
 Unlike registers, we do not abstract the memory indices as a special type but instead, for the sake of clarity and ease of representation, we show differently masked machine words, $\loc \in \Loc$, with the subscripts showing the length of a 64-bit machine word after masking, e.g.

### 2023-04-04 programlogic.tex
+

--- 2023-03-29 programlogic.tex
@@
 \begin{enumerate}
-\item a bag of physical pages reached through a set of physical page-table (L4-L1 page-tables) acceses which are shared amongst different virtual addresses indexed by the root address of the address-space
-\item allocating/updating pages backing a virtual address 
-\item switching from one address-space to another 
+\item \textit{logical representation of addressing}: we present physical memory addressing as separation-logic points-to relations.
+\item \textit{sharing}: a bag of physical pages are reached through a set of physical page-table (L4-L1 page-tables) acceses that are shared amongst different virtual addresses.
+This sharing imposes constraints on giving a definition to the virtual-address in terms of physical (L4-L1) page-table memory accesses
+\item \textit{context-agnostic-resources}: each virtual address is valid under a certain address-space, but it does not represent this \textit{knowledge} on its address-space. 
+\item \textit{address-spaces as modal-contexts}: we call these context-agnostic resources (virtual-address pointstos) as \textit{context-resource}, and the context identfied by a unique \textit{namespace}, and  defining the truth for these context resources as \text{modal-context}
+\item \textit{updating inter/intra address-spaces}: we present logical abstractions to enable allocating/updating pages in the existence of \textit{sharing}  
+\item \textit{address-space switch as changing the "World" of truth}: switching from one address-space to another logically becomes introduction of \textit{context-resources} to the \textit{modal-context} under which they are valid, and loading \textit{context-resources} that are valid under the \textit{currently} loaded address-space
 \end{enumerate}
-Based on these perspectives, we coneptualize truth on an address-space as indexed by its roots address, therefore.
-we lift standard bi proposition to be indexed with $\mathcal{W}_{64}$ as shown in Figure \ref{fig:vprop}.
-Contesequently.
-all abstractions we explain in this section as part of our address-space logic are type $\textsf{vProp }\Sigma$.
+Based on these perspectives, we coneptualize truth on an address-space through the contingency it exhibits: \textit{assertion P holds on this address-space indexed by its roots address}.
+Therefore, as a first step, we lift standard bi proposition to be indexed with $\mathcal{W}_{64}$ as shown in Figure \ref{fig:vprop}, and, most notably, all abstractions we explain in this section as part of our custom-tailored address-space logic are type $\textsf{vProp }\Sigma$.
 \begin{figure}[!ht]

### 2023-04-04 programlogic.tex
+

@@
 \begin{array}{cl}
-\textsf{vProp } \Sigma \; : \; \textsf{bi} \; \stackrel{def}{=} \mathcal{W}_{64} -\textsf{b}> \textsf{iPropI } \Sigma
+\textsf{vProp } \Sigma \; : \; \textsf{bi} \; \stackrel{def}{=} \mathcal{W}_{64} \rightarrow_{\textsf{b}} \textsf{iPropI } \Sigma
 \end{array}
@@
 \label{sec:pointsto}
-We abstract memory addresses and registers naming their relevant values through well-known separation logic assertion \textit{points-to}.
+We abstract physical memory addressing and registers naming with well-known separation logic assertion \textit{points-to} assertions.
 \begin{enumerate}
@@
 The fraction $\qfrac$ with value 1 asserts the unique ownership of the register mapping, and grants update permission on it, otherwise, any value $0 < \qfrac <1$ represents partial ownership granting readonly permission on the mapping.
-\paragraph{Physical-Memory points-to} Our physical memory points-to relation ($\pfpointsto\locsf\locsf\qfrac\ppts$) exhibits nested-mappings abstracting the  masks applied on the indexing virtual address ($\textsf{vaddr}\in\mathcal{W}_{64}$).
-The unfolded definition of the physical mapsto
+\paragraph{Physical-Memory points-to} Our physical memory points-to relation for a page table entry (e.g.
+\textsf{PDE} in Figure \ref{fig:physicalpointsto}) abstracts masking -- $|^{52}$ for frame for and $|^{12}$ for offset -- through logical nested maps.
 \begin{figure}[!ht]
@@
 \end{figure}
-where, for an example virtual address \textsf{va} and given \textsf{CR3} in Figure \ref{}, we can abstract the L4-L1 table traversal as the following:
+Given the definition of physical page-pointsto assertion and the root address of virtual-address space kept in the control register \textsf{CR3} as shown in Figure \ref{fig:pagetable},  one can build the physical address-translation for a virtual address (e.g.
+\textsf{va}) via abstracting the L4-L1 table traversal as the following:
 \begin{itemize}

### 2023-04-04 programlogic.tex
+

@@
 \begin{enumerate}
-\item Physical address points-to, $\pfpointsto\locsf\locsf\qfrac\ppts$
+\item Physical address points-to, $\pfpointsto\locsf\vpts\qfrac\ppts$
 \item Register points-to, $\pfpointsto\rg\rv\qfrac\rpts$
@@
 The fraction $\qfrac$ with value 1 asserts the unique ownership of the register mapping, and grants update permission on it, otherwise, any value $0 < \qfrac <1$ represents partial ownership granting readonly permission on the mapping.
-\paragraph{Physical-Memory points-to} Our physical memory points-to relation for a page table entry (e.g.
-\textsf{PDE} in Figure \ref{fig:physicalpointsto}) abstracts masking -- $|^{52}$ for frame for and $|^{12}$ for offset -- through logical nested maps.
+\paragraph{Physical-Memory points-to} Our physical memory points-to relation in Figure \ref{fig:physicalpointsto} for a page table entry (e.g.
+\textsf{PDE} in Figure \ref{fig:pagetables}) abstracts masking -- $|^{52}$ for frame for and $|^{12}$ for offset -- through logical nested maps.
 \begin{figure}[!ht]
@@
 \begin{array}{cl}
-\pfpointsto\locsf\locsf\qfrac\ppts \stackrel{def}{=} & \nfpointsto{\mask\locsf\ft\generalentry}{\mask\locsf\tw\generalentry}\locsf\qfrac\naddr
+\pfpointsto\locsf\vpts\qfrac\ppts \stackrel{def}{=} & \nfpointsto{\mask\locsf\ft\generalentry}{\mask\locsf\tw\generalentry}\vpts\qfrac\naddr
 \end{array}

### 2023-04-04 programlogic.tex
+

@@
 \begin{itemize}
-  \item Page Map Level-4 Translation (L4): Calculating the offset to the 4 \textsf{Kbyte Physical Page} table to obtain the  physical address by  which resides in the last 12 bits of the virtual memory address, $\locsf|^{12}$,  shown in Figure \ref{}.
-    % \pfpointsto{\lvlsum\crt\lfoff}{\lvlbor\lt}\qfracfotsss\ppts \ast
+  \item Page Map Level-4 Translation (L4): Address translation, handling the first level (to get the PML4E entry) itself by calculating the offset (\textsf{CR3}$|^{12}$) and frame (PML4E$|^{52}$) via masking with the root address space address currently stored in \textsf{CR3}.
     \[ \begin{array}{l}
@@
       \]
-  \item Page Directory Pointer Level Translation (L3):
+  \item Page Directory Pointer Level Translation (L3): Performs the second level of address translation via interpreting, in other words masking, the PML4 entry (PML4E$|^{52}$) to retrieve the starting address of a page-directory-pointer-table and the offset ( (PML4E$|^{9}$)) to the entry
     \[\begin{array}{l}
@@
     \]
-  \item Page Directory Table Level Translation (L2):  we obtain the address of the page table via \textsf{PDE}$|^{52}$ and picking the offset from the virtual memory address via imposing the mask to obtain page table offset bits (9 bits [20-12] shown in Figure \ref{}) to obtain the page table entry (\textsf{PTE}).
+  \item Page Directory Table Level Translation (L2):  Performs the third level of address translation by retrieving the starting address at the PD table (PDPE$|^{52}$), and interprets a PDPE ($|^{9}$) that references the PD table entry.
     \[ \begin{array}{l}

### 2023-04-04 programlogic.tex
+

@@
 \begin{itemize}
-  \item Page Map Level-4 Translation (L4): Address translation, handling the first level (to get the PML4E entry) itself by calculating the offset (\textsf{CR3}$|^{12}$) and frame (PML4E$|^{52}$) via masking with the root address space address currently stored in \textsf{CR3}.
+  \item Page Map Level-4 Translation (L4): Address the first level address translation to get the PML4E entry in Figure \ref{fig:pagetables} by using the masks(\textsf{va}$|^{52}$(\textsf{CR3}) to get the starting address of the PML4 table and using the offset mask \textsf{va}$|^{12}$(\textsf{CR3}).
+It is worth noting that our offset indices are 12 bits as all physical memory pointsto assertions has to satisfy being \textsf{\textit{aligned}}, e.g.
+[47-39] plus 3 alignment bits in Figure \ref{fig:pagetables}.
     \[ \begin{array}{l}

### 2023-04-04 programlogic.tex
+

--- 2023-04-04 programlogic.tex
@@
 \begin{itemize}
-  \item Page Map Level-4 Translation (L4): Calculating the offset to the 4 \textsf{Kbyte Physical Page} table to obtain the  physical address by  which resides in the last 12 bits of the virtual memory address, $\locsf|^{12}$,  shown in Figure \ref{}.
-    % \pfpointsto{\lvlsum\crt\lfoff}{\lvlbor\lt}\qfracfotsss\ppts \ast
+  \item Page Map Level-4 Translation (L4): Performs the first level address translation to get the PML4 entry (PML4E) in Figure \ref{fig:pagetables} by using the masks(\textsf{va}$|^{52}$(\textsf{CR3}) to get the starting address of the PML4 table (PLM4T) and using the offset mask \textsf{va}$|^{12}$(\textsf{CR3}).
+It is worth noting that our offset indices are 12 bits as all physical memory pointsto assertions has to satisfy being \textsf{\textit{aligned}}, e.g.
+[47-39] plus 3 alignment bits in Figure \ref{fig:pagetables}
     \[ \begin{array}{l}
@@
       \]
-  \item Page Directory Pointer Level Translation (L3):
+  \item Page Directory Pointer Level Translation (L3): Performs the second level address translation to get the PDP entry (PDPE) in Figure \ref{fig:pagetables} by using the masks(\textsf{va}$|^{52}$(\textsf{PLM4E}) to get the starting address of the PDP table (PDPT) and using the offset mask \textsf{va}$|^{12}$(\textsf{PLM4E})
     \[\begin{array}{l}
@@
     \]
-  \item Page Directory Table Level Translation (L2):  we obtain the address of the page table via \textsf{PDE}$|^{52}$ and picking the offset from the virtual memory address via imposing the mask to obtain page table offset bits (9 bits [20-12] shown in Figure \ref{}) to obtain the page table entry (\textsf{PTE}).
+  \item Page Directory Table Level Translation (L2): Performs the second level address translation to get the PD entry (PDE) in Figure \ref{fig:pagetables} by using the masks(\textsf{va}$|^{52}$(\textsf{PDPE}) to get the starting address of the PD table and using the offset mask \textsf{va}$|^{12}$(\textsf{PDPE})
     \[ \begin{array}{l}
@@
       \]
-  \item Page Table Level Translation (L1):
+  \item Page Table Level Translation (L1): Performs the second level address translation to get the PT entry (PTE) in Figure \ref{fig:pagetables} by using the masks(\textsf{va}$|^{52}$(\textsf{PDE}) to get the starting address of the Page table (PT) and using the offset mask \textsf{va}$|^{12}$(\textsf{PDE})
     \[ \begin{array}{l}
@@
       \]
-  \item Page Address Level Translation
+  \item Page Address Level Translation: Performs the second level address translation to get the physical page address in Figure \ref{fig:pagetables} by using the masks(\textsf{va}$|^{52}$(\textsf{PTE}) to get the starting address of the physical page address table and using the offset mask \textsf{va}$|^{12}$(\textsf{PTE})
     \[ \begin{array}{l}

### 2023-04-04 programlogic.tex
+

@@
 \newcommand{\gammaPred}{\delta}
+\newcommand{\gammaPreds}{\delta\textsf{s}}
 
@@
   \ownGhost\gammaPred{\authfrag{\singletonMap{#1}{(#2, #3)}}}
+}
+
+\newcommand{\sumapaces}[2]{
+  \ownGhost\gammaPreds{\authfrag{\singletonMap{#1}{#2}}}
 }
@@
 \begin{array}{l}
-  \vaddr\mapsto_{\textsf{t}}\{\textsf{q}\}\vpage \stackrel{def}{=} \lambda \crthree.\\
+  \vaddr\mapsto_{\textsf{t}}\{\textsf{q}\}\vpage\; : \mathsf{vProp} \; \Sigma \;  \stackrel{def}{:=}  \lambda \crthree. \;\; \\
   \exists_{\entryf \;, \entrytr \;, \entrytw \;,\entryo} \ldotp
@@
   \ulcorner \textsf{aligned } \vaddr \urcorner \ast  \\
-  \textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\crthree,\entryf,\entrytr,\entrytw,\entryo)\\
-   \nfpointsto{\mask\vaddr\ft\entryo}{\mask\vaddr\tw\entryo}\vpage\qfracone\naddr 
+ \quad \textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\crthree,\entryf,\entrytr,\entrytw,\entryo) \ast \nfpointsto{\mask\vaddr\ft\entryo}{\mask\vaddr\tw\entryo}\vpage\qfracone\naddr 
 \end{array} \\
-
+\\
 \textsf{where   }\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\vaddr,\crthree,\entryf,\entrytr,\entrytw,\entryo) \stackrel{\triangle}{=} \\
@@
   One might have already observed that the virtual points-to definition shown in Figure \ref{fig:strongvirtualpointsto} is too strong to specify virtual memory operations updating a page table entry.
-To speak more concretely, the fractional permissions on the page-table-walk assertion ($textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$ in Figure \ref{fig:strongvirtualpointsto}) \textit{only} assure the existence of the certain number of mappings -- 512 at each level -- abstracting virtual address translations, and updating any of these mappings would make the other mappings, address translations sharing the page table mappings.
-Therefore, we need to have logical abstractions ensuring sound updates to the page-table entries in the existence of sharing.
+To speak more concretely, the fractional permissions on the page-table-walk assertion ($\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$ in Figure \ref{fig:strongvirtualpointsto}) \textit{only} assure the existence of the certain number of mappings -- 512 at each level -- abstracting virtual address translations, and updating any of these mappings would make the other mappings, address translations sharing the page table mappings.
+Therefore, we need to consturct logical abstractions ensuring sound updates to the page-table entries in the existence of sharing.
 
@@
     \exists \entryo \ldotp
+    \sumapaces\crthree\delta \ast 
   \sumwalkabs\vaddr\qfrac\paddr \ast 

### 2023-04-04 programlogic.tex
+

@@
 
- One possible approach is to abstract away the physical page-table-walk pointsto ($\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$ in Figure \ref{fig:strongvirtualpointsto}) from the virtual pointsto relation via replacing it with read-only abstract map entries -- a ghost (abstract) map entry per \textsf{L}$_4$\_\textsf{L}$_{1}$ mappings -- which can be used as a resource ($ \sumwalkabs\vaddr\qfrac\paddr$) whose ownership ensures that the virtual address ($\vaddr$) is mapped to a page (referenced by $\paddr$) ,and as a token to access to the physical page-table-walk pointsto (\textsf{L}$_4$\_\textsf{L}$_{1}$) stored in an address-space invariant ($\mathcal{I}$\textsf{ASpace} in Figure \ref{fig:peraspaceinvariant}).
-We use \textsf{Iris}'s authorative ghost-map logical machinery
+  One possible approach is to abstract away the physical page-table-walk pointsto ($\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$ in Figure \ref{fig:strongvirtualpointsto}) from the virtual pointsto relation.
+To do so, we introduce the protocol separating read-only (fragmental) and authorative (full ownership) accesses to the page tables, specifically $\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$.
+We realize this protocol by embedding it into an already existing \textsf{Iris}'s ghost-map (or view\_maps) construction
   \[\mathcal{A}\textsf{bsPTableWalk} \stackrel{\triangle}{=} \textsc{Auth} (\; \mathcal{W}_{64} \;\rightarrow_{\textrm{fin}} \;  (\mathcal{W}_{64} \; \rightarrow_{\textrm{fin}}\;  (\textsc{Frac }, \mathord{+}) \times (\textsc{Agree } \Loc,\mathord{=}) ))\]
-to realize these reasoning principles ($\mathcal{A}\textsf{bsPTableWalk}$).
-\begin{remark}[Exchanging Page-Table-Walk-Summarization Mappings (Tokens) with Physical Page-Table-Walk Pointstos ]
+ where
+  \begin{itemize}
+  \item we decorate the fragmental ownership of the ghost-map ($ \sumwalkabs\vaddr\qfrac\paddr$) with the physical $\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$ whose ownership just ensures that the virtual address ($\vaddr$) is mapped to a page (referenced by $\paddr$)
+  \item we place the authorative ownership $\mathcal{A}\textsf{PTableWalk}$ of the ghost-map in an address-space invariant ($\mathcal{I}$\textsf{ASpace} in Figure \ref{fig:peraspaceinvariant}) whose ownership grants sound update on the page table entries
+  \end{itemize}
+
+\begin{remark}[Exchanging Page-Table-Walk-Summarization Mappings (Tokens) ]
   $\mathcal{A}\textsf{PTableWalk}$'s fragmental-ownership represents an individual key-value pair (per-virtual-address) which can be handled as a \textit{look-up} resource for an individual virtual address.
-  To build more intuition, we can think of the use-case \textit{unmapping-pages} as collecting the fragmental map pairs, e.g.
+  To build more intuition on the convenience of this protocol, we can think of the use-case \textit{unmapping-pages} as collecting the fragmental map pairs, e.g.
 $\sumwalkabs\vaddr\qfrac\paddr$, from collecting the assertions, removing them from the ghost map, and then updating the page tables.
-  Critically, physical full ownership ($\ownGhost\gammaPred{\authfull{\ptableabswalk\ptablestore}}$) of the page tables, as a \textit{singleton resource}, lives in the central invariant keeping physical-page-table pointstos per-virtual address translation.
-  With this mechanism we can separate the virtual-to-physical translation from the physical memory involved, and consequently
-  \begin{itemize}
-  \item virtual pointsto assertions, whose definition is shown in Figure \ref{fig:virtualpointstosharing}, can refer to their physical memory assertions representing their page-address translation in the physical-page-tables ($\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$) which reside in the address-space invariant ($\mathcal{I}$\textsf{ASpace} in Figure \ref{fig:peraspaceinvariant}) using individual key-value pair, i.e.
-instances of ($\sumwalkabs\vaddr\qfrac\paddr$)
-  \item unlike the case in the strong virtual-pointsto relation in Figure \ref{fig:strongvirtualpointsto}, physical-page-table-walk pointsto assertions ($\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$) for all virtual addresses are moved into the address-space invariant ($\mathcal{I}$\textsf{ASpace} in Figure \ref{fig:peraspaceinvariant}), and resides with authorative ownership of $\mathcal{A}\textsf{PTableWalk}$ ($\ownGhost\gammaPred{\authfull{\ptableabswalk\ptablestore}}$) which respects all entries of the abstract virtual-address translation mappings ($\sumwalkabs\vaddr\qfrac\paddr$).
-    \end{itemize}
 \end{remark}

### 2023-04-04 programlogic.tex
+

@@
 $\sumwalkabs\vaddr\qfrac\paddr$, from collecting the assertions, removing them from the ghost map, and then updating the page tables.
+while Besides, y separating the virtual-points Overlapping address spaces — both shared data, and shared page tables
+which may see updates — can be coordinated with protocols to ensure those updates preserve
+address space invariants.
 \end{remark}
@@
   \end{figure}
-    \subsubsection{Global Per-Address-Space Invariant}
-    \label{sec:peraspaceinvariant}
-    Virtual points-to assertions would need to be defined in terms of physical points-to assertions and hardware-specific address translation assertions for mapped virtual addresses ($\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\crthree,\entryf,\entrytr,\entrytw,\entryo)$ in Figure \ref{fig:peraspaceinvariant}).
-For each virtual address \textit{mapped} with a physical page table entry, abstracted with ghost mappings ($\sumwalkabs\vaddr\qfrac\paddr$) in virtul-pointsto relations as shown in Figure \ref{fig:virtualpointstosharing}), there exsits a physical page table walk.
-    \[ \bigast{(\vaddr, \entryo)\in \ptablestore}{\exists\;(\entryf\;,\entrytr\;,\entrytw)\ldotp \textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\vaddr,\crthree,\entryf,\entrytr,\entrytw,\entryo)} \]
-    By separating the virtual-points Overlapping address spaces — both shared data, and shared page tables
-which may see updates — can be coordinated with protocols to ensure those updates preserve
-address space invariants.
+
 \begin{figure*}
@@
 the truth on address virtualization is indexed by CPUs page table register ($\crthree$ for x86-64), and we present modal reasoning principles for updating $\crthree$ (switching from the current address space to another)
+ \[\mathcal{A}\textsf{spaces} \stackrel{\triangle}{=} \textsc{Auth} (\; \mathcal{W}_{64} \;\rightarrow_{\textrm{fin}} \; (\textsc{Agree } \mathsf{GName},\mathord{=}) ) \]
 \subsubsection{Address-Space Modality}

### 2023-04-04 programlogic.tex
+

@@
 
-Then, once we want to outline the distribution of the facts per-address, as we see it pictorially  in Figure \ref{fig:logicaladdrspace}, the outline is composed of the following aspects:
-\begin{enumerate}
-\item a global address space invariant (solid round square labelled as $\mathcal{I}$\textsf{ASpace} in Figure \ref{fig:peraspaceinvariant}) asserting the authorative ownership (i.e.
+Relations amongst these logical constructions, for the sake making it clearly understood, outlined pictorially  in Figure \ref{fig:logicaladdrspace}:
+\begin{itemize}
+\item a global address space invariant (solid round square labelled as $\mathcal{I}$\textsf{ASpace} in Figure \ref{fig:logicaladdrspace}) asserting the authorative ownership (i.e.
 update capability), $\ownGhost\gammaPred{\authfull{\ptableabswalk\ptablestore}}$, of the ghost map -- shown as solid-round-head arrow to $\mathcal{A}\textsf{PTableWalk}$ Figure \ref{fig:logicaladdrspace}
 \item per-virtual-pointsto relation ((cutted round square labelled as \textsf{Virtual PointsTo} in Figure \ref{fig:peraspaceinvariant})) asserting the fragmental ownership of the ghost map, $\sumwalkabs\vaddr\qfrac\paddr$ \textit{represents} the fragmental-ownership (cutted-diamond-head arrow) of $\mathcal{A}\textsf{PTableWalk}$
-\item for each ghost page-table-walk mapping,  $\sumwalkabs\vaddr\qfrac\paddr$, there exists one shared persistent invariant (shown as double-ended solid arrow 1 to N)
-  \[\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\crthree,\entryf,\entrytr,\entrytw,\entryo)\]
-\end{enumerate}
+\item there exists only one persistent authorative ownership ($\ownGhost\gammaPred{\authfull{\ptableabswalk\ptablestore}}$) for all fragmental ghost page-table-walk mapping ($\sumwalkabs\vaddr\qfrac\paddr$) shown as double-ended solid arrow 1 to N
+\end{itemize}
 
@@
 
-\begin{figure*}
-\[
-\begin{array}{l}
-  \mathcal{I}\textsf{ASpace} \stackrel{\triangle}{=} \lambda \; \crthree \ldotp
-  \exists\;\gammaPred \;\ldotp\; \ownGhost\gammaPred{\authfull{\ptableabswalk\ptablestore}} \ast \\
-  \bigast{(\vaddr, \entryo)\in \ptablestore}{\exists\;(\entryf\;,\entrytr\;,\entrytw)\ldotp \textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\vaddr,\crthree,\entryf,\entrytr,\entrytw,\entryo)}
-\end{array}
-\]
-\caption{Global Address-Space Invariant}
-  \label{fig:peraspaceinvariant}
-\end{figure*}
 \subsection{Address-Space Management}
 \label{sec:aspacemanagement}
-With the reasoning principles introduced for the essential resource of address spaces, i.e. virtual memory addresses, now we can ...
-% \item abstractions shaped-around address-space switch where all virtual pointsto assertions happen to be true in a certain address space, i.e.
-the truth on address virtualization is indexed by CPUs page table register ($\crthree$ for x86-64), and we present modal reasoning principles for updating $\crthree$ (switching from the current address space to another)
+So far, we have introduced logical abstractions for a single address space.
+However, \textsf{VMM} implementations handle more than one address-space which we abstract as a ghost map from root address of an address-space to a namespace labelling the address space
 \[\mathcal{A}\textsf{spaces} \stackrel{\triangle}{=} \textsc{Auth} (\; \mathcal{W}_{64} \;\rightarrow_{\textrm{fin}} \; (\textsc{Agree } \mathsf{GName},\mathord{=}) ) \]
-\subsubsection{Address-Space Modality}
-\label{sec:aspacemodalist}
-The truth on an address space, exhibits itself as a contingent truth: a location virtualization assertion (a virtual points-to in \ref{fig:virtualpointstosharing}),  happens to valid in the \textit{current world} (in the current address space), and switching address spaces pulls information out of one \textit{world} into the “current view” of memory, and leaves other assertions true relative to the previous address space.
+
+\paragraph{Valid Context-Resources} Fragmental ownership of $\mathcal{A}\textsf{spaces}$ ($\sumapaces\crthree\gammaPred$) conveys the fact that a current address-space ($\gammaPred$) is the one selected, i.e.
+loaded into the $\crthree$.
+This simple fact becomes essential when we observe it inside virtual-pointsto definition in Figure \ref{fig:virtualpointstosharing} and pictorially Figure \ref{logicaladdrspace}.
+Existence of this fragmental ownership ensures that the contextual resource, i.e.
+a virtual pointsto assertion, resides under the correct address-space, i.e.
+modal-context, that makes it valid.
+
+\paragraph{Modal Resource Context\&Interaction with Ambient Logic \textsf{vProp}}
+\label{sec:resourcecontext}
+The relation between the address-space and and the validity of virtual-pointsto assertions exhibits itself as a contingency, i.e.
+\textit{"a certain virtual-pointsto facts whose validity is indexed by the root address of an address space"}.
+But, what sort of convenience this pattern provides when we do reasoning about more than one address-space.
+
 \begin{figure}
@@
   \end{figure}
+
+In Figure \ref{fig:addrswitch}, we see two different address-spaces rooted at R0 and R1 in the pre/post condition.
+To be put into words, while switching an address-space by executing \lstinline|mov_ctl|) updating the $\crthree$, it is expected to mention resources of two different address-spaces in the specification.
+In the precondition, the dashed white background denotes the resources of the current address space: $\crthree$ is loaded with R0 and all the virtual-pointsto assertions ($a_0 \mapsto_{\textsf{v}}p_0$ and $a_1 \mapsto_{\textsf{v}}p_1$).
+We need one more address-space -- the solid square with the root address R1 -- which has its own resources ($b_0 \mapsto_{\textsf{v}}p_3$ and $b_1 \mapsto_{\textsf{v}}p_4$) which are invalid under the current address space rooted at R0.
+This pattern of specification requires resources being mentioned under the \textit{world} they are valid, i.e.
+a modal-framing within the specification.
+Once the current address-space becomes the one with root address R1, i.e.
+$\crthree$ is loaded with R1, the context-resources $b_0 \mapsto_{\textsf{v}}p_3$ and $b_1 \mapsto_{\textsf{v}}p_4$ are loaded into the current-view of the memory, and all the other resources that are valid under R0 are introduced into the modal context rooted at R0
+\[
+\begin{array}{l}
+  \{ \crthree \mapsto_{r} R0 \ast a_0 \mapsto_{\textsf{v}}p_0  \ast a_1 \mapsto_{\textsf{v}}p_1  \ast [r1]\;( b_0 \mapsto_{\textsf{v}}p_3  \ast b_1 \mapsto_{\textsf{v}}p_4 )\}\\
+  \qquad \qquad \qquad \qquad\qquad \qquad \mathsf{mov\_ctl} \;\; \crthree \; \; R1\\
+
+  \{ \crthree \mapsto_{r} R1 \ast [r0](a_0 \mapsto_{\textsf{v}}p_0  \ast a_1 \mapsto_{\textsf{v}}p_1)  \ast  b_0 \mapsto_{\textsf{v}}p_3  \ast b_1 \mapsto_{\textsf{v}}p_4 \}
+  \end{array}
+\]
+
+The choice of resource defining the modal context already exhibits itself as a reasoning principle (Figure \ref{fig:modalcontextinvariance}) in which we can intro/eliminate any resource that is not virtual pointsto into/from modal context without dischargin any obligation.
+We can use this entailment for  \todo[inline,color=red]{Ask Colin}.
+
+\begin{remark}[The Choice of Modal Context Definition as a Pattern of Verification Context]
+  \label{remark:pattern}
+  By choosing the contingency and what defines the modal contex, we simply introduce 
+  \begin{enumerate}
+  \item how a modal context (e.g.
+a address-space modality $[r]P$), logically, draws the attention to the certain facts in the interest of reasoning (e.g.
+virtual pointsto relations in the current address-space in Figure \ref{fig:virtualpointstosharing})
+  \item and, offloads the burden of individual bookkeeping of these facts (e.g.
+virtual pointsto relations per address-space) under different context (e.g.
+virtual pointsto relations in other address-spaces) via utilizing the \textit{summarization} aspect of the contingency it represents (e.g.
+other virtual pointsto relations are valid under other certain address-space page-table root addresseses).
+  \end{enumerate}
+\end{remark}
+\paragraph{Interaction}
+\label{sec:interaction}
+
+\subsubsection{Address-Space Modality}
+\label{sec:aspacemodalist}
+The truth on an address space, exhibits itself as a contingent truth: a location virtualization assertion (a virtual points-to in \ref{fig:virtualpointstosharing}),  happens to valid in the \textit{current world} (in the current address space), and switching address spaces pulls information out of one \textit{world} into the “current view” of memory, and leaves other assertions true relative to the previous address space.
+
 Therefore, an address-space, as an abstraction, can be treated as naming the memory state as a modal frame and the choice of page table root as a world in Kripke-style semantics.
@@
 separation logic as an entailment for specifying \textit{address-space-switch}.
-\paragraph{Modal Resource Context}
-\label{sec:resourcecontext}
-When we pick up a certain contingency, which is \textit{"a certain virtual-pointsto facts whose validity is indexed by the root address of an address space"}for address space modality, we, in fact, make our choice around some logical resources for which we care about reasoning, which is virtual pointsto relation shown in Figure \ref{fig:virtualpointstosharing}.
-In other words, our modality definition is the logical virtual pointsto resources: introduction/elimination of logical virtual pointsto resources are the ones essential to the reasoning capabilities the address space modality provides, i.e.
-validity of virtual addresses under/out of modality -- resources under/out of \textit{modal resource context}.
-
-The choice of resource defining the modal context already exhibits itself as a reasoning principle (Figure \ref{fig:modalcontextinvariance}) in which we can intro/eliminate any resource that is not virtual pointsto into/from modal context without dischargin any obligation.
-We can use this entailment for  \todo[inline,color=red]{Ask Colin}.
-  \[
-  \begin{array}{l}
-\hbox{(\TirNameStyle{ModalContextInvariance})} \qquad
- [r](Q \ast \ell \mapsto_{\textsf{p}} \vpage)  \;\vdash [r](Q) \ast \ell \mapsto_{\textsf{p}} \vpage 
-  \end{array}
-  \]
-\begin{remark}[The Choice of Modal Context Definition as a Pattern of Verification Context]
-  \label{remark:pattern}
-  By choosing the contingency and what defines the modal contex, we simply introduce 
-  \begin{enumerate}
-  \item the pattern of what the essential abstraction (e.g. address spaces) we care about to do verification for is
-  \item how a modal context (e.g.
-a address-space modality $[r]P$), logically, draws the attention to the certain facts in the interest of reasoning (e.g.
-virtual pointsto relations in the current address-space in Figure \ref{fig:virtualpointstosharing})
-  \item and, offloads the burden of individual bookkeeping of these facts (e.g.
-virtual pointsto relations per address-space) under different context (e.g.
-virtual pointsto relations in other address-spaces) via utilizing the \textit{summarization} aspect of the contingency it represents (e.g.
-other virtual pointsto relations are valid under other certain address-space page-table root addresseses).
-  \end{enumerate}
-\end{remark}
-\paragraph{Interaction}
-\label{sec:interaction}
-So far, we have just mentioned the structural properties of our modal definition (Remark \ref{remark:pattern}), and what inherent entailments (Rule \hbox{(\TirNameStyle{ModalContextInvariance})}) these properties provide.
-However, all these structural aspects of using the address-space modality ($[r]P$) considers the fixed index ($r$) of on the truth $P$.
-From the client's perspective, structural properties of the modal abstraction mainly constitutes the modal frame with summarizing the facts for one another address-space under the address-space modality roughly as it is in the following spec
-\[
-\{ \crthree \mapsto_{r} r0 \ast \textsf{va1} \mapsto_{v} \textsf{v}_1 \ast \textsf{va2} \mapsto_{v} \textsf{v}_2 \ast \textsf{va3} \mapsto_{v} \textsf{v}_3 \ast [r1]\;P \}
-\]
-where we see the \textit{current address-space} rooted at $r0$ is loaded into the  \textit{current-view} of the memory with its resources ($ \textsf{va1} \mapsto_{v} \textsf{v}_1 \ast \textsf{va2} \mapsto_{v} \textsf{v}_2 \ast \textsf{va3} \mapsto_{v} \textsf{v}_3$), and framed by the address-space rooted at $r1$ with resources $P$.
-Now, we need to a rule capturing the notion that switching the address spaces by pulling the resources ($P$) out of the one under modality into the \textit{“current view”} of memory, and leaves other assertions true relative to the previous address space (the one rooted at $r0$).
-To do so, we need to have a proof rule for the instruction (\lstinline|mov_ctl|) updating the $\crthree$.
 

### 2023-04-04 programlogic.tex
+

@@
 \]
-
-The choice of resource defining the modal context already exhibits itself as a reasoning principle (Figure \ref{fig:modalcontextinvariance}) in which we can intro/eliminate any resource that is not virtual pointsto into/from modal context without dischargin any obligation.
-We can use this entailment for  \todo[inline,color=red]{Ask Colin}.
-
 \begin{remark}[The Choice of Modal Context Definition as a Pattern of Verification Context]
@@
 \label{sec:interaction}
-
 \subsubsection{Address-Space Modality}

### 2023-04-04 programlogic.tex
+

@@
 
-\begin{figure}
-   \includegraphics[width=\columnwidth]{addr_space_switch.png}
-  \caption{Loading an Address-Space into the Current Memory View}
-  \label{fig:addrswitch}
-  \end{figure}
-
 In Figure \ref{fig:addrswitch}, we see two different address-spaces rooted at R0 and R1 in the pre/post condition.
 To be put into words, while switching an address-space by executing \lstinline|mov_ctl|) updating the $\crthree$, it is expected to mention resources of two different address-spaces in the specification.
-In the precondition, the dashed white background denotes the resources of the current address space: $\crthree$ is loaded with R0 and all the virtual-pointsto assertions ($a_0 \mapsto_{\textsf{v}}p_0$ and $a_1 \mapsto_{\textsf{v}}p_1$).
-We need one more address-space -- the solid square with the root address R1 -- which has its own resources ($b_0 \mapsto_{\textsf{v}}p_3$ and $b_1 \mapsto_{\textsf{v}}p_4$) which are invalid under the current address space rooted at R0.
+In the precondition, $\crthree$ is loaded with R0 and all the virtual-pointsto assertions $a_0 \mapsto_{\textsf{v}}p_0$ and $a_1 \mapsto_{\textsf{v}}p_1$ are in the current-view of the memory and valid in the ambient logic \textsf{vProp}.
+The other address-space with the root address R1 has its own resources ($b_0 \mapsto_{\textsf{v}}p_3$ and $b_1 \mapsto_{\textsf{v}}p_4$) which are invalid under the current address space rooted at R0.
 This pattern of specification requires resources being mentioned under the \textit{world} they are valid, i.e.
@@
 $\crthree$ is loaded with R1, the context-resources $b_0 \mapsto_{\textsf{v}}p_3$ and $b_1 \mapsto_{\textsf{v}}p_4$ are loaded into the current-view of the memory, and all the other resources that are valid under R0 are introduced into the modal context rooted at R0
+
+\begin{figure}
 \[
@@
 \]
-\begin{remark}[The Choice of Modal Context Definition as a Pattern of Verification Context]
+  \label{fig:addrswitch}
+  \end{figure}
+
+\begin{remark}[The Choice of Modal Context, Contingency and  Pattern of Verification Context]
   \label{remark:pattern}
-  By choosing the contingency and what defines the modal contex, we simply introduce 
-  \begin{enumerate}
-  \item how a modal context (e.g.
-a address-space modality $[r]P$), logically, draws the attention to the certain facts in the interest of reasoning (e.g.
-virtual pointsto relations in the current address-space in Figure \ref{fig:virtualpointstosharing})
-  \item and, offloads the burden of individual bookkeeping of these facts (e.g.
-virtual pointsto relations per address-space) under different context (e.g.
-virtual pointsto relations in other address-spaces) via utilizing the \textit{summarization} aspect of the contingency it represents (e.g.
-other virtual pointsto relations are valid under other certain address-space page-table root addresseses).
-  \end{enumerate}
-\end{remark}
-\paragraph{Interaction}
-\label{sec:interaction}
-\subsubsection{Address-Space Modality}
-\label{sec:aspacemodalist}
 The truth on an address space, exhibits itself as a contingent truth: a location virtualization assertion (a virtual points-to in \ref{fig:virtualpointstosharing}),  happens to valid in the \textit{current world} (in the current address space), and switching address spaces pulls information out of one \textit{world} into the “current view” of memory, and leaves other assertions true relative to the previous address space.
@@
 separation logic as an entailment for specifying \textit{address-space-switch}.
+
+Pragmatically, 
+  \begin{enumerate}
+  \item a modal context (e.g.
+a address-space modality $[r]P$), logically, draws the attention to the certain facts in the interest of reasoning (e.g.
+virtual pointsto relations in the current address-space in Figure \ref{fig:virtualpointstosharing})
+  \item and, offloads the burden of individual bookkeeping of these facts (e.g.
+virtual pointsto relations per address-space) under different context (e.g.
+virtual pointsto relations in other address-spaces) via utilizing the \textit{summarization} aspect of the contingency it represents (e.g.
+other virtual pointsto relations are valid under other certain address-space page-table root addresseses).
+  \end{enumerate}
+\end{remark}
 

### 2023-04-05 programlogic.tex
+

@@
 $\crthree$ is loaded with R1, the context-resources $b_0 \mapsto_{\textsf{v}}p_3$ and $b_1 \mapsto_{\textsf{v}}p_4$ are loaded into the current-view of the memory, and all the other resources that are valid under R0 are introduced into the modal context rooted at R0
+\begin{figure}
+
+\[
+  \begin{array}{l}
+    \hbox{(\TirNameStyle{ModalAddressSpace)}} \qquad
+         [r](P) \; \mathsf{vProp} \; \Sigma \; := \; \lambda \mathsf{P},\; \mathsf{P r}. 
+  \end{array}
+  \]
+  \label{fig:modaldef}
+  \end{figure}
 
@@
 Therefore, an address-space, as an abstraction, can be treated as naming the memory state as a modal frame and the choice of page table root as a world in Kripke-style semantics.
-Not suprisingly, transition between two address-spaces, then, can be just an entailment relation \textit{alternating} the \textit{named-state} against the address-space invariant ($\mathcal{I}\textsf{ASpace}$ in Figure \ref{fig:peraspaceinvariant}).
-\[
-  \begin{array}{l}
-    \hbox{(\TirNameStyle{ModalAddressSpace})} \\ \qquad
-         [r](P)  \;\stackrel{\triangle}{=}  (\mathcal{W}_{64} \rightarrow_{\textsf{b}} \textsf{uPredI} (\textsf{iResUR} \;\Sigma)) \rightarrow_{\textsf{b}} \textsf{uPredI } \_ := \lambda \textsf{P},\; \textsf{P r}.
-  \end{array}
-\]
+Not suprisingly, transition between two address-spaces, then, can be just an entailment relation \textit{alternating} the \textit{named-state}
+
 Being inspired by what hybrid logic \ref{} calls a satisfaction operator, which evaluates the truth of a predicate in a named alternative state (here, address space), we give a modal definition describing the truth of assertions for the resources (i.e.
 virtual pointsto relations) inside the address space.
-Ignoring the predicate types for a while, $[r]P$ in Definition $\hbox{(\TirNameStyle{ModalAddressSpace})}$, indicates that $P$ holds in the virtual address space rooted at $r$, the truth ($P$) on an address-space is indexed by the root page-table address $r$ of the address-space.
+Ignoring the predicate types for a while, $[r]P$ in Figure \ref{fig:modaldef}, indicates that $P$ holds in the virtual address space rooted at $r$, the truth ($P$) on an address-space is indexed by the root page-table address $r$ of the address-space.
 In the rest of this section, we explain the structural aspects -- the modal resource context of address space modality -- , and how we lift the interaction of address-space modality with the ambient logic, i.e.
@@
   \begin{enumerate}
-  \item a modal context (e.g.
-a address-space modality $[r]P$), logically, draws the attention to the certain facts in the interest of reasoning (e.g.
-virtual pointsto relations in the current address-space in Figure \ref{fig:virtualpointstosharing})
+  \item a modal context with its context-resources and picked contingency defines a modality (e.g.
+a address-space modality $[r]P$).
+For the convenience of verification, it enables focusing on the certain facts in the interest of reasoning (e.g.
+virtual pointsto relations in the current address-space)
   \item and, offloads the burden of individual bookkeeping of these facts (e.g.
@@
 \subsection{Considering Hoare Doubles in the Context of Switching Address-Space}
+\subsubsection{Issues with Using Hoare Triples in Address-Space Specification}
 An important subtlety arises with supporting \lstinline|mov|s into \lstinline|%cr3|. Consider the hypothetical rule:
@@
 Virtual points-to assertions could be made self-stable by also giving them partial ownership over \lstinline|cr3| assertions, but this would require explicitly plumbing that ownership from \emph{all} assertions back to any place an address space change might occur; this would seem to be a far graver loss of modularity than this extra quirk in framing discussions.
+\subsubsection{Defining Hoare Doubles for Address-Space Specification}
+
+\begin{figure}
+  \[
+  \begin{array}{l}
+    \textsf{wpd\_def e s E1 } \Phi \;\mathsf{ rootaddr } : \textsf{iProp }\Sigma := \\
+   \qquad ((\textsf{cr cr3}) \mapsto_{\textsf{r}} (\textsf{num rootaddr)}) \textsf{rootaddr}) \ast (\Phi \textsf{ rootaddr}) \wand \textsf{WP e E } @ \textsf{s ; E1} \{\{\_, \textsf{True} \}\}
+    \end{array}
+  \]
+\caption{Unfolded Hoare-Double Definition for \textsf{vProp} Logic }
+\label{fig:wpddefinition}
+\end{figure}
+
 \begin{figure}

### 2023-04-05 programlogic.tex
+

@@
 \subsubsection{Issues with Using Hoare Triples in Address-Space Specification}
+\label{sec:issues}
 An important subtlety arises with supporting \lstinline|mov|s into \lstinline|%cr3|. Consider the hypothetical rule:
@@
 \subsubsection{Defining Hoare Doubles for Address-Space Specification}
-
+Based on the issues we discuss in Section \ref{sec:issues}, and the design of the abstractions for \textsf{vProp} logic, e.g.
+virtual-pointsto assertion and the modal address-space abstraction, we give the unfolded Hoare-double definition in Figure \ref{fig:wpddefinition}.
 \begin{figure}
@@
 \end{figure}
-
+This definition carries only the facts ($\Phi$) that are valid under the current addresss-space -- $\Phi \; \textsf{rootaddr}$.
+It also helps to the agnosticism of the client on what the current address-space is by carrying the \textsf{cr3} value on the definition.
 \begin{figure}

### 2023-04-05 programlogic.tex
+

@@
 \inferrule[WriteToMemFromReg]{
-  \{P \ast r_s \mapsto_{r}\{q\}  \textsf{rvs} \ast \textsf{cr3} \mapsto_{r} \textsf{cr3val}  \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast (\textsf{vaddr} \mapsto_{\textsf{t}} \textsf{v})\;\textsf{cr3val} \}\;\overline{ is}
+  \{P \ast r_s \mapsto_{r}\{q\}  \textsf{rvs}  \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast (\textsf{vaddr} \mapsto_{\textsf{t}} \textsf{v})\;\textsf{cr3val} \}\;\overline{ is}  
 }{
-  \{P \ast r_s \mapsto_{r}\{q\}  \textsf{rvs} \ast \textsf{cr3} \mapsto_{r} \textsf{cr3val}  \ast r_a \mapsto_{r}\{q\} \textsf{ vaddr} \ast (\textsf{vaddr} \mapsto_{\textsf{t}} \textsf{rvs})\;\textsf{cr3val} \}
+  \{P \ast r_s \mapsto_{r}\{q\}  \textsf{rvs}   \ast r_a \mapsto_{r}\{q\} \textsf{ vaddr} \ast (\textsf{vaddr} \mapsto_{\textsf{t}} \textsf{rvs})\;\textsf{cr3val} \}
   \textsf{ mov}~\textsf{r}_a,~\textsf{r}_s;\;\overline{is}
@@
 \inferrule[WriteToRegFromMem]{
-  \{P \ast r_d \mapsto_{r}  \textsf{v} \ast \textsf{cr3} \mapsto_{r} \textsf{cr3val}  \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast (\textsf{vaddr} \mapsto_{\textsf{t}} \textsf{v})\;\textsf{cr3val} \}\;\overline{is}
+  \{P \ast r_d \mapsto_{r}  \textsf{v} \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast (\textsf{vaddr} \mapsto_{\textsf{t}} \textsf{v})\;\textsf{cr3val} \}\;\overline{is}
 }{
-  \{P \ast r_d \mapsto_{r}  \textsf{rvd} \ast \textsf{cr3} \mapsto_{r} \textsf{cr3val}  \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast (\textsf{vaddr} \mapsto_{\textsf{t}} \textsf{v})\;\textsf{cr3val} \}
+  \{P \ast r_d \mapsto_{r}  \textsf{rvd} \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast (\textsf{vaddr} \mapsto_{\textsf{t}} \textsf{v})\;\textsf{cr3val} \}
   \textsf{mov}~\textsf{r}_d,~\textsf{r}_a;\;\overline{is}
@@
 \inferrule[WriteToRegCtlFromReg]{
-  \{P \ast r_s \mapsto_{r}\{q\}  \textsf{ rvs} \ast \textsf{cr3} \mapsto_{r} \textsf{rvs}  \}\overline{is}
+  \{P \ast r_s \mapsto_{r}\{q\}  \textsf{ rvs}  \}\overline{is}
 }{
-  \{P \ast r_s \mapsto_{r}\{q\}  \textsf{ rvs} \ast \textsf{cr3} \mapsto_{r} \textsf{cr3val}  \}
+  \{P \ast r_s \mapsto_{r}\{q\}  \textsf{ rvs}   \}
   \textsf{mov}~\textsf{cr3},~r_s;\;\overline{is}
@@
 \end{figure}
-\paragraph{Register To Memory}
-\paragraph{Memory To Register}
-\paragraph{Register To Register}
 \todo[inline,color=red]{STOP HERE}
+\subsubsection{Register To Memory}
+\subsubsection{Memory To Register}
+\subsubsection{Register To Register}

### 2023-04-05 programlogic.tex
+

@@
 \newcommand{\gammaPreds}{\delta\textsf{s}}
+\newcommand{\rtv}{\textsf{rtv}}
 
@@
 \begin{array}{l}
-    \vaddr\mapsto_{\textsf{v}}\{\textsf{q}\}\vpage \stackrel{\triangle}{=} 
+    \vaddr\mapsto_{\textsf{v,rtv}}\{\textsf{q}\}\vpage \stackrel{\triangle}{=} 
     \exists \entryo \ldotp
-    \sumapaces\crthree\delta \ast 
+    \sumapaces\rtv\delta \ast 
   \sumwalkabs\vaddr\qfrac\paddr \ast 
@@
 
-\paragraph{Valid Context-Resources} Fragmental ownership of $\mathcal{A}\textsf{spaces}$ ($\sumapaces\crthree\gammaPred$) conveys the fact that a current address-space ($\gammaPred$) is the one selected, i.e.
-loaded into the $\crthree$.
+\paragraph{Valid Context-Resources} Fragmental ownership of $\mathcal{A}\textsf{spaces}$ ($\sumapaces\rtv\gammaPred$) conveys the fact that a current address-space ($\gammaPred$) is the one selected, i.e.
+loaded into the $\rtv$.
 This simple fact becomes essential when we observe it inside virtual-pointsto definition in Figure \ref{fig:virtualpointstosharing} and pictorially Figure \ref{logicaladdrspace}.
@@
 In Figure \ref{fig:addrswitch}, we see two different address-spaces rooted at R0 and R1 in the pre/post condition.
-To be put into words, while switching an address-space by executing \lstinline|mov_ctl|) updating the $\crthree$, it is expected to mention resources of two different address-spaces in the specification.
-In the precondition, $\crthree$ is loaded with R0 and all the virtual-pointsto assertions $a_0 \mapsto_{\textsf{v}}p_0$ and $a_1 \mapsto_{\textsf{v}}p_1$ are in the current-view of the memory and valid in the ambient logic \textsf{vProp}.
+To be put into words, while switching an address-space by executing \lstinline|mov_ctl|) updating the  \lstinline|cr3|, it is expected to mention resources of two different address-spaces in the specification.
+In the precondition,  \lstinline|cr3| is loaded with R0 and all the virtual-pointsto assertions $a_0 \mapsto_{\textsf{v}}p_0$ and $a_1 \mapsto_{\textsf{v}}p_1$ are in the current-view of the memory and valid in the ambient logic \textsf{vProp}.
 The other address-space with the root address R1 has its own resources ($b_0 \mapsto_{\textsf{v}}p_3$ and $b_1 \mapsto_{\textsf{v}}p_4$) which are invalid under the current address space rooted at R0.
@@
 Once the current address-space becomes the one with root address R1, i.e.
-$\crthree$ is loaded with R1, the context-resources $b_0 \mapsto_{\textsf{v}}p_3$ and $b_1 \mapsto_{\textsf{v}}p_4$ are loaded into the current-view of the memory, and all the other resources that are valid under R0 are introduced into the modal context rooted at R0
+\lstinline|cr3| is loaded with R1, the context-resources $b_0 \mapsto_{\textsf{v}}p_3$ and $b_1 \mapsto_{\textsf{v}}p_4$ are loaded into the current-view of the memory, and all the other resources that are valid under R0 are introduced into the modal context rooted at R0
 \begin{figure}
@@
 \begin{array}{l}
-  \{ \crthree \mapsto_{r} R0 \ast a_0 \mapsto_{\textsf{v}}p_0  \ast a_1 \mapsto_{\textsf{v}}p_1  \ast [r1]\;( b_0 \mapsto_{\textsf{v}}p_3  \ast b_1 \mapsto_{\textsf{v}}p_4 )\}\\
-  \qquad \qquad \qquad \qquad\qquad \qquad \mathsf{mov\_ctl} \;\; \crthree \; \; R1\\
+  \{ \textsf{cr3 } \mapsto_{r} R0 \ast a_0 \mapsto_{\textsf{v}}p_0  \ast a_1 \mapsto_{\textsf{v}}p_1  \ast [r1]\;( b_0 \mapsto_{\textsf{v}}p_3  \ast b_1 \mapsto_{\textsf{v}}p_4 )\}\\
+  \qquad \qquad \qquad \qquad \qquad \mathsf{mov\_ctl} \;\; \texttt{cr3} \; \; R1\\
 
-  \{ \crthree \mapsto_{r} R1 \ast [r0](a_0 \mapsto_{\textsf{v}}p_0  \ast a_1 \mapsto_{\textsf{v}}p_1)  \ast  b_0 \mapsto_{\textsf{v}}p_3  \ast b_1 \mapsto_{\textsf{v}}p_4 \}
+  \{ \textsf{cr3 } \mapsto_{r} R1 \ast [r0](a_0 \mapsto_{\textsf{v}}p_0  \ast a_1 \mapsto_{\textsf{v}}p_1)  \ast  b_0 \mapsto_{\textsf{v}}p_3  \ast b_1 \mapsto_{\textsf{v}}p_4 \}
   \end{array}
@@
 \inferrule[Broken]{ }{
-  \{P \ast cr3\mapsto_{r} \;r_1 \ast r \mapsto_{r} \;r_2 \ast [r_2](Q)\}
+  \{P \ast \textsf{cr3}\mapsto_{r} \;r_1 \ast r \mapsto_{r} \;r_2 \ast [r_2](Q)\}
   \texttt{mov}~\texttt{\%cr3},~r%\lstinline|mov %cr3, r| 
-  \{[r_1](P) \ast cr3 \mapsto_{r} r_2 \ast r \mapsto_{r} r_2 \ast Q\}
+  \{[r_1](P) \ast \textsf{cr3} \mapsto_{r} r_2 \ast r \mapsto_{r} r_2 \ast Q\}
 }
@@
     \inferrule*[right=Broken]{ }{
-    \{\mathsf{emp} \ast cr3 \mapsto_{r} r_1 \ast r\mapsto_{r} r_2 \ast [r_2](Q)\}
+    \{\mathsf{emp} \ast \textsf{cr3} \mapsto_{r} r_1 \ast r\mapsto_{r} r_2 \ast [r_2](Q)\}
     \texttt{mov}~\texttt{\%cr3},~r%\lstinline|mov %cr3, r| 
-    \{[r_1](\mathsf{emp}) \ast cr3 \mapsto_{r} r_1 \ast r \mapsto_{r} r_2 \ast Q\}
+    \{[r_1](\mathsf{emp}) \ast \textsf{cr3} \mapsto_{r} r_1 \ast r \mapsto_{r} r_2 \ast Q\}
     }
   }{
-    \{a\mapsto_\mathsf{v} x \ast \mathsf{emp} \ast cr3\mapsto_{r}r_1 \ast r \mapsto_{r}r_2 \ast [r_2](Q)\}
+    \{a\mapsto_\mathsf{v} x \ast \mathsf{emp} \ast \textsf{cr3}\mapsto_{r}r_1 \ast r \mapsto_{r}r_2 \ast [r_2](Q)\}
     \texttt{mov}~\texttt{\%cr3},~r%\lstinline|mov %cr3, r| 
-    \{a\mapsto_\mathsf{v} x \ast [r_1](\mathsf{emp}) \ast cr3 \mapsto_{r} r_1 \ast r \mapsto_{r} r_2 \ast Q\}
+    \{a\mapsto_\mathsf{v} x \ast [r_1](\mathsf{emp}) \ast \textsf{cr3} \mapsto_{r} r_1 \ast r \mapsto_{r} r_2 \ast Q\}
   }
@@
 \subsubsection{Defining Hoare Doubles for Address-Space Specification}
-Based on the issues we discuss in Section \ref{sec:issues}, and the design of the abstractions for \textsf{vProp} logic, e.g.
-virtual-pointsto assertion and the modal address-space abstraction, we give the unfolded Hoare-double definition in Figure \ref{fig:wpddefinition}.
+Based on the issues we discuss in Section \ref{sec:issues}, we make sure our Hoare-Double definition in Figure \ref{fig:wpddefinition} enables
+\begin{itemize}
+\item virtual-pointsto relations to be independent of \lstinline{cr3} ownership
+\item and full-ownership of \lstinline to be propagated inside the \textsf{wpd\_def} in order to
+\end{itemize}
+This act of separation allows client-specification to be agnostic of \lstinline|cr3| physical resource, (i.e.
+\lstinline|cr3| $\mapsto_{\textsf{r}} \rtv$).
+The context-resource (virtual-pointsto) has only the \textit{knowledge} of being in a namespace in which it is valid ($\sumapaces\rtv\gammaPred$), nothing more.
+
+In the rest of this, we explain how specifications of some selected \textsf{AMD64} instructions look like under the decisions made on the reasoning principles.
 \begin{figure} 
@@
   \begin{array}{l}
-    \textsf{wpd\_def e s E1 } \Phi \;\mathsf{ rootaddr } : \textsf{iProp }\Sigma := \\
-   \qquad ((\textsf{cr cr3}) \mapsto_{\textsf{r}} (\textsf{num rootaddr)}) \textsf{rootaddr}) \ast (\Phi \textsf{ rootaddr}) \wand \textsf{WP e E } @ \textsf{s ; E1} \{\{\_, \textsf{True} \}\}
+    \textsf{wpd\_def e s E1 } \Phi \;\mathsf{ cr3val } : \textsf{iProp }\Sigma := \\
+   \qquad ((\textsf{cr cr3}) \mapsto_{\textsf{r}} (\textsf{num cr3val)}) \textsf{ cr3val}) \ast (\Phi \textsf{ cr3val}) \wand \textsf{WP e E } @ \textsf{s ; E1 } \{\{\_, \textsf{True} \}\}
     \end{array}
@@
 \end{figure}
-This definition carries only the facts ($\Phi$) that are valid under the current addresss-space -- $\Phi \; \textsf{rootaddr}$.
-It also helps to the agnosticism of the client on what the current address-space is by carrying the \textsf{cr3} value on the definition.
+
 \begin{figure}
@@
 \\
-\inferrule[WriteToMemFromReg]{
-  \{P \ast r_s \mapsto_{r}\{q\}  \textsf{rvs}  \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast (\textsf{vaddr} \mapsto_{\textsf{t}} \textsf{v})\;\textsf{cr3val} \}\;\overline{ is}  
+\inferrule[WriteToPhysMemFromReg]{
+  \{P \ast r_s \mapsto_{r}\{q\}  \textsf{rvs}  \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{t}} \textsf{v} \}\;\overline{ is}  
 }{
-  \{P \ast r_s \mapsto_{r}\{q\}  \textsf{rvs}   \ast r_a \mapsto_{r}\{q\} \textsf{ vaddr} \ast (\textsf{vaddr} \mapsto_{\textsf{t}} \textsf{rvs})\;\textsf{cr3val} \}
+  \{P \ast r_s \mapsto_{r}\{q\}  \textsf{rvs}   \ast r_a \mapsto_{r}\{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{t}} \textsf{rvs} \}
   \textsf{ mov}~\textsf{r}_a,~\textsf{r}_s;\;\overline{is}
@@
 \\
-\inferrule[WriteToRegFromMem]{
-  \{P \ast r_d \mapsto_{r}  \textsf{v} \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast (\textsf{vaddr} \mapsto_{\textsf{t}} \textsf{v})\;\textsf{cr3val} \}\;\overline{is}
+\inferrule[WriteToRegFromPhysMem]{
+  \{P \ast r_d \mapsto_{r}  \textsf{v} \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{t}} \textsf{v} \}\;\overline{is}
 }{
-  \{P \ast r_d \mapsto_{r}  \textsf{rvd} \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast (\textsf{vaddr} \mapsto_{\textsf{t}} \textsf{v})\;\textsf{cr3val} \}
+  \{P \ast r_d \mapsto_{r}  \textsf{rvd} \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{t}} \textsf{v} \}
   \textsf{mov}~\textsf{r}_d,~\textsf{r}_a;\;\overline{is}
+}
+\\
+\inferrule[WriteToRegFromVirtMem]{
+  \{P \ast r_d \mapsto_{r}  \textsf{v} \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{v,rtv}} \textsf{v} \}\;\overline{is}
+}{
+  \{P \ast r_d \mapsto_{r}  \textsf{rvd} \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{v,rtv}} \textsf{v} \}
+  \textsf{mov}~\textsf{r}_d,~\textsf{r}_a;\;\overline{is}
+}
+\\
+\inferrule[WriteToVirtMemFromReg]{
+  \{P \ast r_s \mapsto_{r}\{q\}  \textsf{rvs}  \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{v,rtv}} \textsf{v} \}\;\overline{ is}  
+}{
+  \{P \ast r_s \mapsto_{r}\{q\}  \textsf{rvs}   \ast r_a \mapsto_{r}\{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{v,rtv}} \textsf{rvs} \}
+  \textsf{ mov}~\textsf{r}_a,~\textsf{r}_s;\;\overline{is}
 }
@@
   %\lstinline|mov %cr3, r| 
+}\\
+\inferrule[WriteToRegFromRegCtl]{
+  \{P \ast r_d \mapsto_{r} \textsf{rvs} \ast r_s \mapsto_{r}\{q\} \textsf{ rvs} \}\;\overline{ is}
+}{
+  \{P \ast r_d \mapsto_{r} \textsf{rvd} \ast r_s \mapsto_{r}\{q\} \textsf{ rvs} \}
+  \textsf{ mov}~\textsf{r}_d,~\textsf{r}_s;\;\overline{is}
+  %\lstinline|mov %cr3, r| 
 }
 \end{mathpar}
-\caption{Selected Reasoning Rules for ..}
+\caption{Reasoning Rules for Selected \textsf{AMD64} Instructions}
 \label{fig:wpdamd}
 \end{figure}
-\todo[inline,color=red]{STOP HERE}
 \subsubsection{Register To Memory}
+\paragraph{Register To Physical Memory}
+\paragraph{Register To Virtual Memory}
 \subsubsection{Memory To Register}
+\paragraph{Physical Memory To Register}
+\paragraph{Virtual Memory To Register}
 \subsubsection{Register To Register}
+\paragraph{Updating \lstinline|cr3|}

### 2023-04-05 programlogic.tex
+

@@
 \end{figure}
-Given the definition of physical page-pointsto assertion and the root address of virtual-address space kept in the control register \textsf{CR3} as shown in Figure \ref{fig:pagetable},  one can build the physical address-translation for a virtual address (e.g.
+Given the definition of physical page-pointsto assertion and the root address of virtual-address space kept in the control register \lstinline{cr3} as shown in Figure \ref{fig:pagetable},  one can build the physical address-translation for a virtual address (e.g.
 \textsf{va}) via abstracting the L4-L1 table traversal as the following:
 \begin{itemize}
-  \item Page Map Level-4 Translation (L4): Performs the first level address translation to get the PML4 entry (PML4E) in Figure \ref{fig:pagetables} by using the masks(\textsf{va}$|^{52}$(\textsf{CR3}) to get the starting address of the PML4 table (PLM4T) and using the offset mask \textsf{va}$|^{12}$(\textsf{CR3}).
+  \item Page Map Level-4 Translation (L4): Performs the first level address translation to get the PML4 entry (PML4E) in Figure \ref{fig:pagetables} by using the masks(\textsf{va}$|^{52}$(\lstinline{cr3}) to get the starting address of the PML4 table (PLM4T) and using the offset mask \textsf{va}$|^{12}$(\lstinline{cr3}).
 It is worth noting that our offset indices are 12 bits as all physical memory pointsto assertions has to satisfy being \textsf{\textit{aligned}}, e.g.
@@
   \end{itemize}
+
+  \begin{figure*}
+\[
+\begin{array}{l}
+  \mathcal{I}\textsf{ASpace} \stackrel{\triangle}{=} \lambda \; \crthree \ldotp
+  \exists\;\gammaPred \;\ldotp\; \ownGhost\gammaPred{\authfull{\ptableabswalk\ptablestore}} \ast \\
+  \bigast{(\vaddr, \entryo)\in \ptablestore}{\exists\;(\entryf\;,\entrytr\;,\entrytw)\ldotp \textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\vaddr,\crthree,\entryf,\entrytr,\entrytw,\entryo)}
+\end{array}
+\]
+\caption{Global Address-Space Invariant}
+  \label{fig:peraspaceinvariant}
+  \end{figure*}
 
@@
 \end{figure}
+\subsubsection{Memory To Register}
+In Figure \ref{fig:wpdamd}, we see the rules for updating the memory address ($r_a$).
+To do so, we essentially need page-table-walk pointsto ($\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$ in Figure \ref{fig:strongvirtualpointsto}) to locate the value to be written.
+We mainly have two options to achieve this.
+\paragraph{Physical Memory To Register} First one is to use  page-table-walk pointsto ($\textsf{vaddr} \mapsto_{\textsf{t}} \textsf{v}$), which is consists of all-physical page-table memory assertions, to prove the page-table traversal \textsf{AMD64} implements, update the register $r_d$ the value obtained ($\textsf{v}$) as shown in Figure (\TirNameStyle{WriteToRegFromPhysMem}).
+By doing so, we expose the resources ($\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$) needed only for the proof of \textsf{AMD64} page-table traversal to the client of the rule.
+This breaks the modularity of reasoning we want to maintain in location virtualization.
+  \paragraph{Virtual Memory To Register}  Therefore, we would like to use a virtual-pointsto assertion ($\textsf{vaddr} \mapsto_{\textsf{v},\textsf{rtv}}$), which hides physical resources needed for the proof of page-table-traversal from the client, to obtain the the value ($\textsf{v}$) to update the register $r_d$ as shown in Rule (\TirNameStyle{WriteToRegFromVirtMem}).
+While doing the proof of page-table-traversal with a given virtual-pointsto definition in Figure \ref{fig:virtualpointstosharing}, we unfold and obtain our token ($\sumwalkabs\vaddr\qfrac\paddr$) to exchange for a physical-table mapsto assertion inside the $\mathcal{I}\textsf{ASpace}$ ($\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$) in Figure \ref{fig:peraspaceinvariant}, and apply Rule (\TirNameStyle{WriteToRegFromPhysMem}).
 \subsubsection{Register To Memory}
-\paragraph{Register To Physical Memory}
+ \paragraph{Register To Physical Memory} First one is to
+
 \paragraph{Register To Virtual Memory}
-\subsubsection{Memory To Register}
-\paragraph{Physical Memory To Register}
-\paragraph{Virtual Memory To Register}
 \subsubsection{Register To Register}

### 2023-04-05 programlogic.tex
+

@@
 Therefore, we need to consturct logical abstractions ensuring sound updates to the page-table entries in the existence of sharing.
-
+\begin{figure*}
+\[
+\begin{array}{l}
+    \vaddr\mapsto_{\textsf{v,rtv}}\{\textsf{q}\}\vpage \stackrel{\triangle}{=} 
+    \exists \entryo \ldotp
+    \sumapaces\rtv\delta \ast 
+  \sumwalkabs\vaddr\qfrac\paddr \ast 
+   \nfpointsto{\mask\vaddr\ft\entryo}{\mask\vaddr\tw\entryo}\vpage\qfracone\naddr
+\end{array}
+\]
+\caption{Virtual-Pointsto for Sharing Pages}
+  \label{fig:virtualpointstosharing}
+\end{figure*}  
   One possible approach is to abstract away the physical page-table-walk pointsto ($\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$ in Figure \ref{fig:strongvirtualpointsto}) from the virtual pointsto relation.
@@
 
-\begin{remark}[Exchanging Page-Table-Walk-Summarization Mappings (Tokens) ]
+\begin{remark}[Exchanging Page-Table-Walk-Summarization Mappings (Tokens) under Mutation of Page Tables ]
   $\mathcal{A}\textsf{PTableWalk}$'s fragmental-ownership represents an individual key-value pair (per-virtual-address) which can be handled as a \textit{look-up} resource for an individual virtual address.
-  To build more intuition on the convenience of this protocol, we can think of the use-case \textit{unmapping-pages} as collecting the fragmental map pairs, e.g.
-$\sumwalkabs\vaddr\qfrac\paddr$, from collecting the assertions, removing them from the ghost map, and then updating the page tables.
-while Besides, y separating the virtual-points Overlapping address spaces — both shared data, and shared page tables
-which may see updates — can be coordinated with protocols to ensure those updates preserve
-address space invariants.
+  To build more intuition on the convenience of this protocol, we can think of the use-case \textit{unmapping-pages}.
+Normally, in the existence of shared page-tables, we need to first collect all the fragmental map pairs, e.g.
+$\sumwalkabs\vaddr\qfrac\paddr$, then, after making sure of non-existence of \textit{sharing}, we can \textit{unmap} the page.
+In this case, it would be ok to put $\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$ and $ \sumwalkabs\vaddr\qfrac\paddr$ side-by-side inside the virtual-pointsto definition in Figure \ref{fig:virtualpointstosharing}.
+However, this pattern of verifying is not only inconvenient due to requirement of extra bookkeeping, but also violates our soundness argument based on considering  mutation on the shared page-tables -- due to coalescing of page tables etc.
+--.
+In other words, physical $\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$ sitting aside to  $ \sumwalkabs\vaddr\qfrac\paddr$ would be made invalid by any concurrent update on the physical page-table.
+So there should only be one physical memory resource inside the virtual pointsto definition: from physical page address to the value stored ($   \nfpointsto{\mask\vaddr\ft\entryo}{\mask\vaddr\tw\entryo}\vpage\qfracone\naddr$).
 \end{remark}
-\begin{figure*}
-\[
-\begin{array}{l}
-    \vaddr\mapsto_{\textsf{v,rtv}}\{\textsf{q}\}\vpage \stackrel{\triangle}{=} 
-    \exists \entryo \ldotp
-    \sumapaces\rtv\delta \ast 
-  \sumwalkabs\vaddr\qfrac\paddr \ast 
-   \nfpointsto{\mask\vaddr\ft\entryo}{\mask\vaddr\tw\entryo}\vpage\qfracone\naddr
-\end{array}
-\]
-\caption{Virtual-Pointsto for Sharing Pages}
-  \label{fig:virtualpointstosharing}
-\end{figure*}
+
 %\subsection{Address-Space Abstraction}
@@
 \item per-virtual-pointsto relation ((cutted round square labelled as \textsf{Virtual PointsTo} in Figure \ref{fig:peraspaceinvariant})) asserting the fragmental ownership of the ghost map, $\sumwalkabs\vaddr\qfrac\paddr$ \textit{represents} the fragmental-ownership (cutted-diamond-head arrow) of $\mathcal{A}\textsf{PTableWalk}$
-\item there exists only one persistent authorative ownership ($\ownGhost\gammaPred{\authfull{\ptableabswalk\ptablestore}}$) for all fragmental ghost page-table-walk mapping ($\sumwalkabs\vaddr\qfrac\paddr$) shown as double-ended solid arrow 1 to N
+\item there exists only one persistent authorative ownership ($\ownGhost\gammaPred{\authfull{\ptableabswalk\ptablestore}}$) for all fragmental ghost page-table-walk mapping ($\sumwalkabs\vaddr\qfrac\paddr$) shown as double-ended solid arrow 1 to N imposing the update protocol
 \end{itemize}

### 2023-04-05 programlogic.tex
+

@@
 While doing the proof of page-table-traversal with a given virtual-pointsto definition in Figure \ref{fig:virtualpointstosharing}, we unfold and obtain our token ($\sumwalkabs\vaddr\qfrac\paddr$) to exchange for a physical-table mapsto assertion inside the $\mathcal{I}\textsf{ASpace}$ ($\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$) in Figure \ref{fig:peraspaceinvariant}, and apply Rule (\TirNameStyle{WriteToRegFromPhysMem}).
-\subsubsection{Register To Memory}
- \paragraph{Register To Physical Memory} First one is to
-
-\paragraph{Register To Virtual Memory}
-\subsubsection{Register To Register}
-\paragraph{Updating \lstinline|cr3|}
+\subsubsection{Register To Memory}\todo[inline,color=red]{Iso:explain}
+ \paragraph{Register To Physical Memory}\todo[inline,color=red]{Iso:explain}
+\paragraph{Register To Virtual Memory} \todo[inline,color=red]{Iso:explain}
+\subsubsection{Register To Register}\todo[inline,color=red]{Iso:explain}
+\paragraph{Updating \lstinline|cr3|}\todo[inline,color=red]{Iso:explain}

### 2023-04-05 soundness.tex
+

@@
 We prefer to skip explaining the steps used in instantiation of Iris because it is an almost standard procedure, has already been explained for many other logic \ref{}, and we are concerned with the page-count limitation.
-However, it is worh noting that once you instantiate Iris for your language, it comes with the semantic definition for the weakest-precondition which we can refactor into Hoare triples, $\triple\pre\instr\post$, to specify our  selected \textsf{AMD64} instructions, and show that these triples are sound.
-
+However, it is worh noting that once you instantiate Iris for your language, it comes with the semantic definition for the weakest-precondition which we can refactor into Hoare triples, $\triple\pre\instr\post$, to specify our  selected \textsf{AMD64} instructions shown in Figure \ref{fig:wpdamd}, and show that these triples are sound.
 \subsection{The Soundness Statement}
@@
 Therefore, our soundness argument is to show that any execution composed of instructions in our machine model (some of which are shown in Figure \sref{sec:semantics}) does not end-up in a invalid state.
-In other words, a~program that attempts to ...
-stuck~(\sref{sec:instrsemantics}).
+\begin{figure*}
+\small
+\begin{mathpar}
+\inferrule[Skip]{}{
+  \triple\post\iskip\post
+}
+
+\inferrule[Seq]{
+  \triple\pre{\instr_1}\midpoint \\
+  \triple\midpoint{\instr_2}\post
+}{
+  \triple\pre{(\iseq{\instr_1}{\instr_2})}\post
+}
+\end{mathpar}
+\caption{Structural Rules for Executing Instructions}
+\Description{Reasoning Rules}
+\label{fig:structural}
+\end{figure*}
+
 \begin{theorem}[Soundness of the Logic]
@@
 triple, then this program cannot crash: by a direct consequence of Iris's adequacy theorem~\cite[\S6.4]{iris}.
-\todo{ismail, give the real adeq. theorem here}
+
 Moreover we need to show the validity of each rules in Figures \fref{fig:reasoning} and \fref{fig:laws}.
@@
 \label{th:validity}
-  Each of the rules in Figures~\ref{fig:laws}
-  and~\ref{fig:reasoning} is valid.
+  Each of the rules in Figures~\ref{fig:wpdamd}
+  and~\ref{fig:structural} is valid.
 \end{theorem}
@@
 the reasoning rules of our logic, then the program~$\instrs$ is safe.
-\subsection{BI-Lifting for Address Space Modality}
-\label{sec:bilifting}
 

### 2023-04-05 soundness.tex
+

@@
 \end{theorem}
-Due to the space limits in this paper, we do not mention each proof for these rules explicitly (which are already mechanized in Coq),
-but the definitions and constructions used requires explanation.
+Due to the space limits in this paper, we do not mention each proof for the rules in Figures \ref{fig:wpdamd} and \ref{fig:structural} within this section, but we provide mechanized proofs for all these in Coq as a part of our artifact submission.
+However, we would like to give the definitions and constructions used in our proofs, and would like to give an outline of paper proof for \TirNameStyle{WriteToRegFromVirtMem} in Figure \ref{fig:wpdamd} within this section.
 

### 2023-04-05 soundness.tex
+

@@
 \label{sec:invariant}
-
 We already have the physical and logical stores and a simple invariant between them. Now, we can rely on the following Assumption \ref{assumption} from Iris to utilize its logical constructions.
@@
 \item unlike the points-to relation obtained by direct interpretion of $\genheapinterp\store.\mathcal{R}$ using Iris, the existing \textsf{gen\_heap\_interp},
-  does not directly helps introducing the ghost we define a new algebra (\textsf{gen\_mem\_UR}) for abstracting the nested maps due to different levels of masking in memory mappings. 
+  does not directly helps introducing the ghost we define a new algebra (\textsf{gen\_mem\_UR}) for abstracting the nested maps due to different levels of masking in memory mappings and an interpretation for this algebra.
 \end{enumerate}
@@
   \[\textsf{inG } \Sigma \textsf{ gen\_memUR} \]
+  To interpret this nested ghost map we define \textsf{to\_gen\_mem}
+  \[
+  \begin{array}{l}
+    \mathsf{to\_gen\_mem} : \textsf{ gmap L1 }(\textsf{gmap L2 V})\rightarrow \textsf{gen\_memUR L1 L2 V } := \textsf{ fmap} (\lambda \textsf{m . to\_gen\_heap m}). \\
+     \mathsf{to\_gen\_heap } : \textsf{gmap L V} \rightarrow \textsf{gen\_heapUR L V} :=  (\lambda \textsf{v }\ldotp (1, \textsf{ to\_agree } (\textsf{v} : \textsf{leibnizO V}))).
+  \end{array}
+  \]
+ throuh using \textsf{to\_gen\_heap} from previous Iris version. \todo[inline,color=red]{Ismail give exact version commit etc.}
 As a final step, we allocate an empty ghost cell $\gamma$ at the beggining of a program execution to store the element of the monoid $\textsf{inG } \Sigma \textsf{ gen\_memUR}$.

### 2023-04-05 soundness.tex
+

@@
 \begin{enumerate}
-\item  defines a predicate $\genheapinterp\store.\mathcal{R}$
+\item  defines a predicate $\textsf{to\_gen\_heap }\store.\mathcal{R}$
   that ties a store~$\store.\mathcal{R}$ to this ghost state,
@@
   where it is applied to the \logical store~$\store$.
-\item unlike the points-to relation obtained by direct interpretion of $\genheapinterp\store.\mathcal{R}$ using Iris, the existing \textsf{gen\_heap\_interp},
+\item unlike the register points-to relation obtained by direct interpretion of $\textsf{to\_gen\_heap }\store.\mathcal{R}$ using Iris, the existing \textsf{gen\_heap\_heap},
   does not directly helps introducing the ghost we define a new algebra (\textsf{gen\_mem\_UR}) for abstracting the nested maps due to different levels of masking in memory mappings and an interpretation for this algebra.
@@
 \begin{array}{l@{\quad\star\quad}l@{\quad}l}
-  \genheapinterp\store &  \\
-   \genmemheapinterp\store
+  \textsf{to\_gen\_heap} \;\store.\mathcal{R} &  \\
+   \textsf{to\_gen\_mem} \; \store.\mathcal{M}
 \end{array}
@@
 
-As noted earlier (Assumption~\ref{assumption}), the predicate
-$\genheapinterp\!$ and the points-to assertion for register mappings is inherited from Iris. What
-this means, as far as we are concerned, is that the points-to assertion is
-defined in a standard manner: $\genheapinterp\!$ is applied to the physical store~$\store$, as is usually the case~\cite[\S6.3.2]{iris}.
-Likewise, we define nested interpretation $\genmemheapinterp\!$ in terms of $\genheapinterp\!$ to obtain physical memory points-to relation.
-
-In fact, long story short, all assertions in our logic, i.e.
-points-to assertions for registers and memory, really about the logical store, which is in our case identical to the physical store, so that we can reason about the physical store.
+Now, we have well-enough definition for giving an outline for the proof of \TirNameStyle{WriteToRegFromVirtMem}.
+\todo[inline,color=red]{ISMAIL give the outlin}

### 2023-04-05 soundness.tex
stop at the end of the soundness

@@
 \label{def:invariant}
-The central invariant of our logic is, due to lack of need for augmenting the machine state, simply the state interpretation \cite[\S7.3]{iris} : 
+The central invariant of our logic is, due to lack of need for augmenting the machine state, simply the state interpretation: 
 \[
@@
 
+As a last definition, we give the head step relation required for the Iris instantiation for our simple language in Figure \ref{} to sequence instructions.
+This relation allows lifting the program expression to enables the application of changes imposed by the operational semantics on the program state $\sigma$ when applied for an insturction (\textsf{i}).
+\[
+\begin{array}{l}
+\textsf{exec\_step} (\textsf{i}: \textsf{ instr}) (\sigma:\textsf{ state}):\textsf{ option state} \stackrel{def}{=} \\
+\qquad \textsf{exec\_state} (\textsf{exec\_instr i } \sigma.(\textsf{cpu})  \\ \qquad (\textsf{regToregset } \sigma.(\mathcal{R}))  \\ \qquad ((\textsf{memToPhysMem }\sigma.(\mathcal{M})) (\textsf{exist \_}  \\ \qquad (\textsf{ regset\_get\_num }(\textsf{regToregset }\sigma.(\mathcal{R}))  (\textsf{ cr cr3}))))).
+\end{array}
+\]
+\todo[inline,color=yellow]{Colin, incase needed,not proven, could you simply say that we assume these map equalities}.
+Since the focus of this paper is not explaining our \textsf{AMD64} model, we prefer not to get into the details due to space limits.
+However, Coq formalization of our operational semantics is a part of our submission artifact.
+
 Now, we have well-enough definition for giving an outline for the proof of \TirNameStyle{WriteToRegFromVirtMem}.
-\todo[inline,color=red]{ISMAIL give the outlin}
+ \begin{lemma}[\textsc{\TirNameStyle{WriteToRegFromVirtMem}}]
+   \label{lemma:unlink}
+\begin{align*}
+\inferrule{
+  \{P \ast r_d \mapsto_{r}  \textsf{v} \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{v,rtv}} \textsf{v} \}\;\overline{is}
+}{
+  \{P \ast r_d \mapsto_{r}  \textsf{rvd} \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{v,rtv}} \textsf{v} \}
+  \textsf{mov}~\textsf{r}_d,~\textsf{r}_a;\;\overline{is}
+}
+\end{align*}
+ \end{lemma}
+ \begin{proof}
+   Assuming the inference rule realized with \textsf{wpd\_def}, we expand the precondition with $\textsf{cr3} \mapsto_{\textsf{r}} \rtv$.
+   Then we do two proofs, one for head reducibility of atomic step \textsf{mov\_reg64\_mem64}, the second one for the executing the expression and obtaining the new state.
+Steps taken in the first one are subset of the second one, so we outline the second portion of the proof, but in case of an interest in details of the proof, Coq artifact can be consulted.
+
+   \begin{itemize}
+   \item Step 1: we apply the head step relation and obtain the current valid state $\sigma1$
+   \item Step 2: we unfold the \textsf{stop here}
+   \end{itemize}
+
+   \end{proof}
+
+% Lemma wexploit_table_traversal (r:greg) maddr entry entry0 entry1 paddr cr3val  (v:word 64) σ :
+ % x64h_interp  σ.(mem)  σ.(reg) -∗
+  %     ((cr cr3) ↦ᵣ (num cr3val)) cr3val -∗
+   %    (r ↦ᵣ (num maddr)) cr3val -∗
+    %   (wtraversal_mapsto maddr entry entry0 entry1 paddr v) cr3val

### 2023-04-05 programlogic.tex
+

@@
 \begin{itemize}
-  \item Page Map Level-4 Translation (L4): Performs the first level address translation to get the PML4 entry (PML4E) in Figure \ref{fig:pagetables} by using the masks(\textsf{va}$|^{52}$(\lstinline{cr3}) to get the starting address of the PML4 table (PLM4T) and using the offset mask \textsf{va}$|^{12}$(\lstinline{cr3}).
+  \item Page Map Level-4 Translation (L4): Performs the first level address translation to get the PML4 entry (PML4E) in Figure \ref{fig:pagetables} by using the masks (\textsf{va}$|^{52}$(\lstinline{cr3}) to get the starting address of the PML4 table (PLM4T) and using the offset mask \textsf{va}$|^{12}$(\lstinline{cr3}).
 It is worth noting that our offset indices are 12 bits as all physical memory pointsto assertions has to satisfy being \textsf{\textit{aligned}}, e.g.

### 2023-04-05 soundness.tex
+

@@
 \[
-\centralinvariant\;\store \triangleq
+\textsf{x64\_h}\;\store \triangleq
 \left\{
@@
 \]
-\todo[inline,color=yellow]{Colin, incase needed,not proven, could you simply say that we assume these map equalities}.
+\todo[inline,color=yellow]{Colin, in case needed,not proven, could you simply say that we assume these map equalities or a paper proof etc.}.
 Since the focus of this paper is not explaining our \textsf{AMD64} model, we prefer not to get into the details due to space limits.
@@
  \end{lemma}
+
  \begin{proof}
@@
    \begin{itemize}
-   \item Step 1: we apply the head step relation and obtain the current valid state $\sigma1$
-   \item Step 2: we unfold the \textsf{stop here}
+   \item Step 1: we apply the head step relation and obtain the current valid state $\sigma1$ interpretation : $\textsf{x64\_h}\;\store$
+   \item Step 2: we unfold the virtual-pointsto ($\vaddr \mapsto_{\textsf{v,rtv}} \textsf{v}$) definition, and for an existential physical page address $\paddr$, we exchange our fragmental toke ($\sumwalkabs\vaddr\qfrac\paddr$) to obtain physical table-pointsto ($\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$ in Figure \ref{fig:strongvirtualpointsto}) relation
+   \item Step 3: for each physical pointsto inside $\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$, we obtain
+\[
+       \ulcorner \sigma.\mathcal{M } !! (\textsf{ va}|^{52}(\textsf{ent}_i) = \textsf{Some } \sigma\textsf{d}_i \land
+       \sigma\textsf{d }_i !! (\textsf{ va}|^{12}(\textsf{ent}_i)  = \textsf{Some entry}_{i+1} \urcorner
+       \]
+       where $\sigma.\textsf{d}_i$ in type \textsf{gmap (word 12) (word 64)} for our offset mappings. 
+     \item Step 4: use these concrete lookups to traverse the page tables to obtain the value \textsf{v}
+     \item Step 5: do the map update the $\sigma.\mathcal{R}$ for relevant register mapping ($r_d$) with the value \textsf{v} 
    \end{itemize}
@@
    \end{proof}
-
-% Lemma wexploit_table_traversal (r:greg) maddr entry entry0 entry1 paddr cr3val  (v:word 64) σ :
- % x64h_interp  σ.(mem)  σ.(reg) -∗
-  %     ((cr cr3) ↦ᵣ (num cr3val)) cr3val -∗
-   %    (r ↦ᵣ (num maddr)) cr3val -∗
-    %   (wtraversal_mapsto maddr entry entry0 entry1 paddr v) cr3val

### 2023-04-05 examples.tex
+
+start working on examples

@@
+\newcommand{\assert}[1]{\mathit{assert}\ #1}
+\newcommand{\specline}[1]{{\color{blue}\left\{#1\right\}}}
+
 \section{Experiment}
 \label{sec:experiment}
-\mytodo{Mechanization is Ismail's Part -- Discussing and putting in here is Colin's}
-
 To both validate and demonstrate the value of the modal approach to reasoning about virtual memory management, we study several distillations of real concerns of virtual memory managers.
@@
 \subsection{Identity Mappings}
+\begin{comment}
 \todo[inline]{Identity mappings are difficult, and our current approach won't quite work.
@@
 Once that's done, then we work out the details of v2, which should require \emph{refactoring} the same sorts of reasoning we had for v1, rather than totally different proofs from v1.}
-
+\end{comment}
 \subsection{Mapping a New Page}
@@
   \begin{lstlisting}
-;; Assume the faulting address f is in r12,
-;; a callee-save register, and f with the 
-;; lower 12 bits zeroed out (so, the virtual
-;; page containing f) is in r13 (also callee-save)
-
 ;; Ensure there's a place to put the L1 mapping
@@
 mov (%r14), %rax
-;; memset the newly-mapped page
-mov %rdi, %r13
-mov %rsi, 0
-mov %rdx, 4096
-call memset
-;; The newly mapped page is full of 0s
 \end{lstlisting}

### 2023-04-05 examples.tex
+

--- 2022-09-28 examples.tex
@@
+%$\assert{\ulcorner \mathsf{aligned maddr} \urcorner \ast \mathsf{r14} \mapsto_{\textsf{r}} \textsf{r14v} \ast \mathsf{r13} \mapsto_{\textsf{r}} \textsf{maddr} \ast \mathsf{rdi} \mapsto_{\textsf{r}} \textsf{rdiv} \ast \mathsf{rax} \mapsto_{\textsf{r}} \textsf{raxv}}$
+%$\assert{\ulcorner \ptablestore !! \textsf{maddr} = \textsf{None} \urcorner \ast \ownGhost\gammaPred{\authfull{\ptableabswalk\ptablestore}} \ast \textsf{Pf}} $
 \section{Experiment}
@@
 
-\subsection{Identity Mappings}
+\begin{comment}
 \todo[inline]{Identity mappings are difficult, and our current approach won't quite work.

### 2023-04-05 examples.tex
First have a short discussion on the examples section

--- 2023-04-05 examples.tex
@@
-\newcommand{\assert}[1]{\mathit{assert}\ #1}
-\newcommand{\specline}[1]{{\color{blue}\left\{#1\right\}}}
-
+%$\assert{\ulcorner \mathsf{aligned maddr} \urcorner \ast \mathsf{r14} \mapsto_{\textsf{r}} \textsf{r14v} \ast \mathsf{r13} \mapsto_{\textsf{r}} \textsf{maddr} \ast \mathsf{rdi} \mapsto_{\textsf{r}} \textsf{rdiv} \ast \mathsf{rax} \mapsto_{\textsf{r}} \textsf{raxv}}$
+%$\assert{\ulcorner \ptablestore !! \textsf{maddr} = \textsf{None} \urcorner \ast \ownGhost\gammaPred{\authfull{\ptableabswalk\ptablestore}} \ast \textsf{Pf}} $
 \section{Experiment}
 \label{sec:experiment}
+\mytodo{Mechanization is Ismail's Part -- Discussing and putting in here is Colin's}
+
 To both validate and demonstrate the value of the modal approach to reasoning about virtual memory management, we study several distillations of real concerns of virtual memory managers.
@@
 
-\subsection{Identity Mappings}
 \begin{comment}

### 2023-04-06 programlogic.tex
+

@@
 In the rest of this, we explain how specifications of some selected \textsf{AMD64} instructions look like under the decisions made on the reasoning principles.
+All rules, e.g.
+the ones with different addressing mode, can be found in our artifact.
+
+In the general outline of the proof in Figure \ref{wpdamd}, we see that each rule annotated with an address value ($\{\rtv\}$), i.e.
+a root address of an address-space, under which the resources mentioned in the specification are valid.
+Moreover, there is a frame resource $P$ in each rule.
 \begin{figure} 
@@
 \begin{mathpar}
-\inferrule[WriteToRegFromReg]{
+  \inferrule[WriteToRegFromReg | \{\textsf{rtv}\}]{
   \{P \ast r_d \mapsto_{r} \textsf{rvs} \ast r_s \mapsto_{r}\{q\} \textsf{ rvs} \}\;\overline{ is}
@@
 \\
-\inferrule[WriteToPhysMemFromReg]{
+\inferrule[WriteToPhysMemFromReg | \{\textsf{rtv}\}]{
   \{P \ast r_s \mapsto_{r}\{q\}  \textsf{rvs}  \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{t}} \textsf{v} \}\;\overline{ is}  
@@
 \\
-\inferrule[WriteToRegFromPhysMem]{
+\inferrule[WriteToRegFromPhysMem | \{\textsf{rtv}\}]{
   \{P \ast r_d \mapsto_{r}  \textsf{v} \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{t}} \textsf{v} \}\;\overline{is}
@@
 \\
-\inferrule[WriteToRegFromVirtMem]{
-  \{P \ast r_d \mapsto_{r}  \textsf{v} \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{v,rtv}} \textsf{v} \}\;\overline{is}
+\inferrule[WriteToRegFromVirtMem | \{\textsf{rtv}\}]{
+  \{P \ast \mathcal{I}\textsf{ASpace}\ast \mathcal{I}\textsf{ASpace} \ast r_d \mapsto_{r}  \textsf{v} \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{v,rtv}} \textsf{v} \}\;\overline{is}
 }{
-  \{P \ast r_d \mapsto_{r}  \textsf{rvd} \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{v,rtv}} \textsf{v} \}
+  \{P \ast \mathcal{I}\textsf{ASpace}\ast r_d \mapsto_{r}  \textsf{rvd} \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{v,rtv}} \textsf{v} \}
   \textsf{mov}~\textsf{r}_d,~\textsf{r}_a;\;\overline{is}
@@
 \\
-\inferrule[WriteToVirtMemFromReg]{
-  \{P \ast r_s \mapsto_{r}\{q\}  \textsf{rvs}  \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{v,rtv}} \textsf{v} \}\;\overline{ is}  
+\inferrule[WriteToVirtMemFromReg | \{\textsf{rtv}\}]{
+  \{P \ast \mathcal{I}\textsf{ASpace}\ast r_s \mapsto_{r}\{q\}  \textsf{rvs}  \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{v,rtv}} \textsf{v} \}\;\overline{ is}  
 }{
-  \{P \ast r_s \mapsto_{r}\{q\}  \textsf{rvs}   \ast r_a \mapsto_{r}\{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{v,rtv}} \textsf{rvs} \}
+  \{P \ast \mathcal{I}\textsf{ASpace}\ast r_s \mapsto_{r}\{q\}  \textsf{rvs}   \ast r_a \mapsto_{r}\{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{v,rtv}} \textsf{rvs} \}
   \textsf{ mov}~\textsf{r}_a,~\textsf{r}_s;\;\overline{is}
@@
 \\
-\inferrule[WriteToRegCtlFromReg]{
+\inferrule[WriteToRegCtlFromReg | \{\textsf{rtv}\} / \{\textsf{rvs}\}]{
   \{P \ast r_s \mapsto_{r}\{q\}  \textsf{ rvs}  \}\overline{is}
@@
 }\\
-\inferrule[WriteToRegFromRegCtl]{
+\inferrule[WriteToRegFromRegCtl | \{\textsf{rtv}\}]{
   \{P \ast r_d \mapsto_{r} \textsf{rvs} \ast r_s \mapsto_{r}\{q\} \textsf{ rvs} \}\;\overline{ is}

### 2023-04-06 programlogic.tex
For Colin's review

@@
 While doing the proof of page-table-traversal with a given virtual-pointsto definition in Figure \ref{fig:virtualpointstosharing}, we unfold and obtain our token ($\sumwalkabs\vaddr\qfrac\paddr$) to exchange for a physical-table mapsto assertion inside the $\mathcal{I}\textsf{ASpace}$ ($\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$) in Figure \ref{fig:peraspaceinvariant}, and apply Rule (\TirNameStyle{WriteToRegFromPhysMem}).
-\subsubsection{Register To Memory}\todo[inline,color=red]{Iso:explain}
- \paragraph{Register To Physical Memory}\todo[inline,color=red]{Iso:explain}
-\paragraph{Register To Virtual Memory} \todo[inline,color=red]{Iso:explain}
-\subsubsection{Register To Register}\todo[inline,color=red]{Iso:explain}
-\paragraph{Updating \lstinline|cr3|}\todo[inline,color=red]{Iso:explain}
+\paragraph{Updating \lstinline|cr3|} Unlike other rules, \TirNameStyle{WriteToRegCtlFromReg} updates the root address of the address-space determining the validity of resources, $\{\rtv\} / \{\textsf{rvs}\}$.

### 2023-04-08 soundness.tex
some cr3 removal from defs

@@
 Our logic, operates on the machine state, which means we do not need to augment the machine state.
-The invarian that we pick, \textit{central invariant} ($\mathcal{I}$), is just semantic interpretation of stores in the machine state, $\sigma$.R and $\sigma$.M.
+The invarian that we pick, \textit{central invariant} ($\mathcal{I}$\textsf{ASpace}), is just semantic interpretation of stores in the machine state, $\sigma$.R and $\sigma$.M.
 This semantic interpratation ensures the correct lifting of mappings in the machine state to the assertions that the client of our logic uses, i.e.

### 2023-04-09 examples.tex
+

@@
 \todo[inline]{Ismail, see notes about intermediate assertions on the Notion page}
+%\begin{figure*}\small
+%\[
+%\proofoutstep{x:\isolated\;\texttt{Node},y:\INT,z:\bool}
+%{
+ %   \proofoutstep{x:\isolated\;\texttt{Node},y:\INT,z:\bool}
+  %  {
+   %     \proofoutstep{x:\writable\;\texttt{Node},y:\INT,z:\bool}
+    %    {
+     %       \proofoutstep{x:\writable\;\texttt{Node},y:\INT,z:\bool}
+      %      {
+       %       mov\texttt{ }rdi \; r13
+        %    }{\textsc{T-Alloc}}
+         %   {x:\writable\;\texttt{Node},y:\writable\;\texttt{Node},z:\bool}\\
+          %  \;\;\;z\texttt{=}x\texttt{.next;}\\
+           % \specline{x:\writable\;\texttt{Node},y:\writable\;\texttt{Node},z:\writable\;\texttt{Node}}\\
+            %\;\;\;y\texttt{.next=}z\texttt{;}\\
+            %\specline{x:\writable\;\texttt{Node},y:\writable\;\texttt{Node},z:\writable\;\texttt{Node}}\\
+            %\;\;\;z\texttt{.prev=}y\texttt{;}\\
+            %\specline{x:\writable\;\texttt{Node},y:\writable\;\texttt{Node},z:\writable\;\texttt{Node}}\\
+            %\;\;\;y\texttt{.prev=}x\texttt{;}\\
+         %\specline{x:\writable\;\texttt{Node},y:\writable\;\texttt{Node},z:\writable\;\texttt{Node}}\\
+          %  \;\;\;x\texttt{.next=}y\texttt{;}\\
+         %\specline{x:\writable\;\texttt{Node},y:\writable\;\texttt{Node},z:\writable\;\texttt{Node}}
+        %}
+        %{\textsc{T-Seq}*~\textrm{across}~\textsc{T-FieldRead}\\\textrm{and}~\textsc{T-FieldWrite}}
+        %{x:\writable\;\texttt{Node},y:\writable\;\texttt{Node},z:\writable\;\texttt{Node}}
+    %}
+    %{\textsc{T-SubEnv}}
+    %{x:\writable\;\texttt{Node}}
+%}
+%{\textsc{T-RecovIso}}
+%{x:\isolated\;\texttt{Node}}
+%\]
+%\caption{Typing derivation for adding a node to an isolated doubly-linked list.}
+%\label{fig:recoviso_example}
+%\end{figure*}
+\lstset{
+  columns=fullflexible,
+  numbers=left,
+  basicstyle=\ttfamily,
+  keywordstyle=\color{blue}\bfseries,
+  morekeywords={mov,add,call},
+  emph={rsp,rdx,rax,rbx,rbp,rsi,rdi,rcx,r8,r9,r10,r11,r12,r13,r14,r15},
+  emphstyle=\color{green},
+  emph={[2]cr3},
+  emphstyle={[2]\color{violet}},
+  morecomment=[l]{;;}
+}
+\newcommand{\fpaddr}{\texttt{fpaddr}}
 \begin{figure}
   \begin{lstlisting}
-;; Ensure there's a place to put the L1 mapping
-mov %rdi, %r13
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,\gammaPred,\rtv)  \ast  \texttt{r14}\mapsto_{r} \_ \ast \texttt{rdi}\mapsto_{r} \vaddr \ast \texttt{rax}\mapsto_{r} \_ \ast \ulcorner \texttt{aligned va} \land  \theta \; !!\; \texttt{maddr} = \texttt{None}\urcorner}$
 call ensure_L1_page
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,\gammaPred,\rtv) \ast  \texttt{r14}\mapsto_{r} \_ \ast \texttt{rdi}\mapsto_{r} \vaddr \ast \texttt{rax}\mapsto_{r} \_ \ast }$
+$\specline{\exists (\entryf ,\;\entrytr,\; \entrytw,\; \entryo: \texttt{word 64}) \ldotp \ulcorner (\texttt{addr\_L1 }\vaddr) = \entryo\urcorner \ast }$
+$\specline{\nfpointsto{\mask\vaddr\ft\rtv}{\mask\vaddr\tw\rtv}\entryf\qone\naddr \ast}$ 
+$\specline{  \nfpointsto{\mask\vaddr\ft\entryf}{\mask\vaddr\tw\entryf}\entrytr\qtwo\naddr \ast \nfpointsto{\mask\vaddr\ft\entrytr}{\mask\vaddr\tw\entrytr}\entrytw\qthree\naddr \ast}$
+$\specline{\texttt{pte\_addr} \mapsto_{\texttt{vpte},\rtv} \entryo \;(\texttt{wzero 64}) \ast \texttt{rax}\mapsto_{r} \texttt{pte\_addr}  }$
 ;; Returns the virtual address of the L1 entry in rax
 mov %r14, %rax ;; Save that before another call
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,\gammaPred,\rtv) \ast  \texttt{r14}\mapsto_{r} \texttt{pte\_addr} \ast \texttt{rdi}\mapsto_{r} \vaddr \ast \texttt{rax}\mapsto_{r} \_ \ast }$
+$\specline{ \nfpointsto{\mask\vaddr\ft\rtv}{\mask\vaddr\tw\rtv}\entryf\qone\naddr \ast}$ 
+$\specline{  \nfpointsto{\mask\vaddr\ft\entryf}{\mask\vaddr\tw\entryf}\entrytr\qtwo\naddr \ast \nfpointsto{\mask\vaddr\ft\entrytr}{\mask\vaddr\tw\entrytr}\entrytw\qthree\naddr \ast}$
+$\specline{\texttt{pte\_addr} \mapsto_{\texttt{vpte},\rtv} \entryo \;(\texttt{wzero 64}) \ast \texttt{rax}\mapsto_{r} \texttt{pte\_addr}  }$
 call alloc_phys_page_or_panic
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,\gammaPred,\rtv) \ast  \texttt{r14}\mapsto_{r} \texttt{pte\_addr} \ast \texttt{rdi}\mapsto_{r} \vaddr \ast \texttt{rax}\mapsto_{r} \_ \ast}$
+$\specline{ \nfpointsto{\mask\vaddr\ft\rtv}{\mask\vaddr\tw\rtv}\entryf\qone\naddr \ast}$ 
+$\specline{  \nfpointsto{\mask\vaddr\ft\entryf}{\mask\vaddr\tw\entryf}\entrytr\qtwo\naddr \ast \nfpointsto{\mask\vaddr\ft\entrytr}{\mask\vaddr\tw\entrytr}\entrytw\qthree\naddr \ast}$
+$\specline{\texttt{pte\_addr} \mapsto_{\texttt{vpte},\rtv} \entryo \;(\texttt{wzero 64}) \ast \texttt{rax}\mapsto_{r} \texttt{pte\_addr}  }$
+$\specline{\exists \texttt{ fpaddr} \ldotp \ulcorner \texttt{aligned fpaddr} \urcorner \ast \texttt{rax}\mapsto_{r} \texttt{fpaddr} \ast \texttt{fpaddr} \mapsto_{a} (\texttt{wzero 64})}$
 ;; Calculate new PTE by setting the low-order bits
@@
 mov (%r14), %rax
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,\gammaPred,\rtv) \ast  \texttt{r14}\mapsto_{r} \texttt{pte\_addr} \ast \texttt{rdi}\mapsto_{r} \vaddr \ast \texttt{rax}\mapsto_{r} \_ \ast}$
+$\specline{\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\vaddr,\rtv,\entryf,\entrytr,\entrytw,\entryo) \ast \ulcorner \theta \;!!\; \texttt{maddr} = \texttt{None}\urcorner}$
+$\specline{\ulcorner \texttt{aligned fpaddr} \urcorner \ast \texttt{rax}\mapsto_{r} \texttt{fpaddr} \ast \texttt{fpaddr} \mapsto_{a} (\texttt{wzero 64})}$
+$\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\; \sqsubseteq $
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta<[\vaddr:=\texttt{fpaddr}]>),\gammaPred,\rtv) \ast  \texttt{r14}\mapsto_{r} \texttt{pte\_addr} \ast \texttt{rdi}\mapsto_{r} \vaddr \ast \texttt{rax}\mapsto_{r} \_ }$
+$\specline{\ulcorner \texttt{aligned fpaddr} \urcorner \ast \texttt{rax}\mapsto_{r} \texttt{fpaddr} \ast \texttt{aligned va} \ast\sumwalkabs\vaddr\qfrac\fpaddr}$
 \end{lstlisting}
-\caption{Distilled code for mapping a new page}
+\caption{Specification and proof of distilled code for mapping a new page with $\textsf{wpd\_def }_{\rtv}$}
 \label{fig:mapping_code}
@@
 ;; Save yielding context
+$\specline{}$  
 mov 0[%rdi], %rbx

### 2023-04-10 examples.tex
+

@@
   \caption{Specification and proof of distilled code for mapping a new page with $\textsf{wpd\_def }_{\rtv}$}
+  \todo[inline,color=yellow]{Colin, could you speak about the fact that we have wpd in iProp so that's the reason to have rtval whereever table traversal is needed}
 \label{fig:mapping_code}
@@
 ;; Save yielding context
-$\specline{}$  
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,\gammaPred,\rtv)}$
+$\specline{ \texttt{rsi}\mapsto_{r} \rtv' \ast \texttt{rdi}\mapsto_{r} \_ \ast \texttt{rbx}\mapsto_{r} \texttt{rbxv} \ast  \texttt{rsp}\mapsto_{r} \texttt{rspv} \ast \texttt{rbp}\mapsto_{r} \texttt{rbpv} }$
+$\specline{\texttt{r12}\mapsto_{r} \texttt{r12v} \ast \texttt{r13}\mapsto_{r} \texttt{r13v} \ast \texttt{r14}\mapsto_{r} \texttt{r14v} \ast \texttt{r15}\mapsto_{r} \texttt{r15v}}$
+$\specline{\texttt{rdi+0} \mapsto_{v,\rtv} \_ \ast \texttt{rdi+8} \mapsto_{v,\rtv} \_ \ast \texttt{rdi+16} \mapsto_{v,\rtv} \_ \ast \texttt{rdi+24} \mapsto_{v,\rtv} \_ \ast \texttt{rdi+32} \mapsto_{v,\rtv} \_}$
+$\specline{\texttt{rdi+40} \mapsto_{v,\rtv} \_ \ast \texttt{rdi+48} \mapsto_{v,\rtv} \_\ast \texttt{rdi+56} \mapsto_{v,\rtv} \_}$
 mov 0[%rdi], %rbx
@@
 mov 48[%rdi], %r15
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,\gammaPred,\rtv)}$
+$\specline{ \texttt{rdi}\mapsto_{r} \_ \ast \texttt{rbx}\mapsto_{r} \texttt{rbxv} \ast  \texttt{rsp}\mapsto_{r} \texttt{rspv} \ast \texttt{rbp}\mapsto_{r} \texttt{rbpv} }$
+$\specline{\texttt{r12}\mapsto_{r} \texttt{r12v} \ast \texttt{r13}\mapsto_{r} \texttt{r13v} \ast \texttt{r14}\mapsto_{r} \texttt{r14v} \ast \texttt{r15}\mapsto_{r} \texttt{r15v}}$
+$\specline{\texttt{rdi+0} \mapsto_{v,\rtv} \texttt{rbxv} \ast \texttt{rdi+8} \mapsto_{v,\rtv} \texttt{rspv} \ast \texttt{rdi+16} \mapsto_{v,\rtv} \texttt{rbpv} \ast \texttt{rdi+24} \mapsto_{v,\rtv} \texttt{r12v} \ast \texttt{rdi+32} \mapsto_{v,\rtv} \texttt{r13v}}$
+$\specline{\texttt{rdi+40} \mapsto_{v,\rtv} \texttt{r14v} \ast \texttt{rdi+48} \mapsto_{v,\rtv} \texttt{r15v}\ast \texttt{rdi+56} \mapsto_{v,\rtv} \_}$
 mov 56[%rdi], %cr3
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,\gammaPred,\rtv)}$
+$\specline{ \texttt{rdi}\mapsto_{r} \_ \ast \texttt{rbx}\mapsto_{r} \texttt{rbxv} \ast  \texttt{rsp}\mapsto_{r} \texttt{rspv} \ast \texttt{rbp}\mapsto_{r} \texttt{rbpv} }$
+$\specline{\texttt{r12}\mapsto_{r} \texttt{r12v} \ast \texttt{r13}\mapsto_{r} \texttt{r13v} \ast \texttt{r14}\mapsto_{r} \texttt{r14v} \ast \texttt{r15}\mapsto_{r} \texttt{r15v}}$
+$\specline{\texttt{rdi+0} \mapsto_{v,\rtv} \texttt{rbxv} \ast \texttt{rdi+8} \mapsto_{v,\rtv} \texttt{rspv} \ast \texttt{rdi+16} \mapsto_{v,\rtv} \texttt{rbpv} \ast \texttt{rdi+24} \mapsto_{v,\rtv} \texttt{r12v} \ast \texttt{rdi+32} \mapsto_{v,\rtv} \texttt{r13v}}$
+$\specline{\texttt{rdi+40} \mapsto_{v,\rtv} \texttt{r14v} \ast \texttt{rdi+48} \mapsto_{v,\rtv} \texttt{r15v}\ast \texttt{rdi+56} \mapsto_{v,\rtv} \rtv}$    
 ;; Restore target context
@@
 mov %r15, 48[%rsi]
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,\gammaPred,\rtv)}$
+$\specline{ \texttt{rdi}\mapsto_{r} \_ \ast \texttt{rbx}\mapsto_{r} \texttt{rbxv} \ast  \texttt{rsp}\mapsto_{r} \texttt{rspv} \ast \texttt{rbp}\mapsto_{r} \texttt{rbpv} }$
+$\specline{\texttt{r12}\mapsto_{r} \texttt{r12v} \ast \texttt{r13}\mapsto_{r} \texttt{r13v} \ast \texttt{r14}\mapsto_{r} \texttt{r14v} \ast \texttt{r15}\mapsto_{r} \texttt{r15v}}$
+$\specline{\texttt{rdi+0} \mapsto_{v,\rtv} \texttt{rbxv} \ast \texttt{rdi+8} \mapsto_{v,\rtv} \texttt{rspv} \ast \texttt{rdi+16} \mapsto_{v,\rtv} \texttt{rbpv} \ast \texttt{rdi+24} \mapsto_{v,\rtv} \texttt{r12v} \ast \texttt{rdi+32} \mapsto_{v,\rtv} \texttt{r13v}}$
+$\specline{\texttt{rdi+40} \mapsto_{v,\rtv} \texttt{r14v} \ast \texttt{rdi+48} \mapsto_{v,\rtv} \texttt{r15v}\ast \texttt{rdi+56} \mapsto_{v,\rtv} \rtv}$    
 ;; Switch to the new address space

### 2023-04-10 examples.tex
+

--- 2023-04-05 examples.tex
@@
 \todo[inline]{Ismail, see notes about intermediate assertions on the Notion page}
+%\begin{figure*}\small
+%\[
+%\proofoutstep{x:\isolated\;\texttt{Node},y:\INT,z:\bool}
+%{
+ %   \proofoutstep{x:\isolated\;\texttt{Node},y:\INT,z:\bool}
+  %  {
+   %     \proofoutstep{x:\writable\;\texttt{Node},y:\INT,z:\bool}
+    %    {
+     %       \proofoutstep{x:\writable\;\texttt{Node},y:\INT,z:\bool}
+      %      {
+       %       mov\texttt{ }rdi \; r13
+        %    }{\textsc{T-Alloc}}
+         %   {x:\writable\;\texttt{Node},y:\writable\;\texttt{Node},z:\bool}\\
+          %  \;\;\;z\texttt{=}x\texttt{.next;}\\
+           % \specline{x:\writable\;\texttt{Node},y:\writable\;\texttt{Node},z:\writable\;\texttt{Node}}\\
+            %\;\;\;y\texttt{.next=}z\texttt{;}\\
+            %\specline{x:\writable\;\texttt{Node},y:\writable\;\texttt{Node},z:\writable\;\texttt{Node}}\\
+            %\;\;\;z\texttt{.prev=}y\texttt{;}\\
+            %\specline{x:\writable\;\texttt{Node},y:\writable\;\texttt{Node},z:\writable\;\texttt{Node}}\\
+            %\;\;\;y\texttt{.prev=}x\texttt{;}\\
+         %\specline{x:\writable\;\texttt{Node},y:\writable\;\texttt{Node},z:\writable\;\texttt{Node}}\\
+          %  \;\;\;x\texttt{.next=}y\texttt{;}\\
+         %\specline{x:\writable\;\texttt{Node},y:\writable\;\texttt{Node},z:\writable\;\texttt{Node}}
+        %}
+        %{\textsc{T-Seq}*~\textrm{across}~\textsc{T-FieldRead}\\\textrm{and}~\textsc{T-FieldWrite}}
+        %{x:\writable\;\texttt{Node},y:\writable\;\texttt{Node},z:\writable\;\texttt{Node}}
+    %}
+    %{\textsc{T-SubEnv}}
+    %{x:\writable\;\texttt{Node}}
+%}
+%{\textsc{T-RecovIso}}
+%{x:\isolated\;\texttt{Node}}
+%\]
+%\caption{Typing derivation for adding a node to an isolated doubly-linked list.}
+%\label{fig:recoviso_example}
+%\end{figure*}
+\lstset{
+  columns=fullflexible,
+  numbers=left,
+  basicstyle=\ttfamily,
+  keywordstyle=\color{blue}\bfseries,
+  morekeywords={mov,add,call},
+  emph={rsp,rdx,rax,rbx,rbp,rsi,rdi,rcx,r8,r9,r10,r11,r12,r13,r14,r15},
+  emphstyle=\color{green},
+  emph={[2]cr3},
+  emphstyle={[2]\color{violet}},
+  morecomment=[l]{;;}
+}
+\newcommand{\fpaddr}{\texttt{fpaddr}}
 \begin{figure}
   \begin{lstlisting}
-;; Ensure there's a place to put the L1 mapping
-mov %rdi, %r13
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,\gammaPred)  \ast  \texttt{r14}\mapsto_{r} \_ \ast \texttt{rdi}\mapsto_{r} \vaddr \ast \texttt{rax}\mapsto_{r} \_ \ast \ulcorner \texttt{aligned va} \land  \theta \; !!\; \texttt{maddr} = \texttt{None}\urcorner}_{\rtv}$
 call ensure_L1_page
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,\gammaPred) \ast  \texttt{r14}\mapsto_{r} \_ \ast \texttt{rdi}\mapsto_{r} \vaddr \ast \texttt{rax}\mapsto_{r} \_ \ast }_{\rtv}$
+$\specline{\exists (\entryf ,\;\entrytr,\; \entrytw,\; \entryo: \texttt{word 64}) \ldotp \ulcorner (\texttt{addr\_L1 }\vaddr) = \entryo\urcorner \ast }_{\rtv}$
+$\specline{\nfpointsto{\mask\vaddr\ft\rtv}{\mask\vaddr\tw\rtv}\entryf\qone\naddr \ast}_{\rtv}$ 
+$\specline{  \nfpointsto{\mask\vaddr\ft\entryf}{\mask\vaddr\tw\entryf}\entrytr\qtwo\naddr \ast \nfpointsto{\mask\vaddr\ft\entrytr}{\mask\vaddr\tw\entrytr}\entrytw\qthree\naddr \ast}_{\rtv}$
+$\specline{\texttt{pte\_addr} \mapsto_{\texttt{vpte},\rtv} \entryo \;(\texttt{wzero 64}) \ast \texttt{rax}\mapsto_{r} \texttt{pte\_addr}  }_{\rtv}$
 ;; Returns the virtual address of the L1 entry in rax
 mov %r14, %rax ;; Save that before another call
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,\gammaPred) \ast  \texttt{r14}\mapsto_{r} \texttt{pte\_addr} \ast \texttt{rdi}\mapsto_{r} \vaddr \ast \texttt{rax}\mapsto_{r} \_ \ast }_{\rtv}$
+$\specline{ \nfpointsto{\mask\vaddr\ft\rtv}{\mask\vaddr\tw\rtv}\entryf\qone\naddr \ast}_{\rtv}$ 
+$\specline{  \nfpointsto{\mask\vaddr\ft\entryf}{\mask\vaddr\tw\entryf}\entrytr\qtwo\naddr \ast \nfpointsto{\mask\vaddr\ft\entrytr}{\mask\vaddr\tw\entrytr}\entrytw\qthree\naddr \ast}_{\rtv}$
+$\specline{\texttt{pte\_addr} \mapsto_{\texttt{vpte},\rtv} \entryo \;(\texttt{wzero 64}) \ast \texttt{rax}\mapsto_{r} \texttt{pte\_addr}  }_{\rtv}$
 call alloc_phys_page_or_panic
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,\gammaPred) \ast  \texttt{r14}\mapsto_{r} \texttt{pte\_addr} \ast \texttt{rdi}\mapsto_{r} \vaddr \ast \texttt{rax}\mapsto_{r} \_ \ast}_{\rtv}$
+$\specline{ \nfpointsto{\mask\vaddr\ft\rtv}{\mask\vaddr\tw\rtv}\entryf\qone\naddr \ast}_{\rtv}$ 
+$\specline{  \nfpointsto{\mask\vaddr\ft\entryf}{\mask\vaddr\tw\entryf}\entrytr\qtwo\naddr \ast \nfpointsto{\mask\vaddr\ft\entrytr}{\mask\vaddr\tw\entrytr}\entrytw\qthree\naddr \ast}_{\rtv}$
+$\specline{\texttt{pte\_addr} \mapsto_{\texttt{vpte},\rtv} \entryo \;(\texttt{wzero 64}) \ast \texttt{rax}\mapsto_{r} \texttt{pte\_addr}  }_{\rtv}$
+$\specline{\exists \texttt{ fpaddr} \ldotp \ulcorner \texttt{aligned fpaddr} \urcorner \ast \texttt{rax}\mapsto_{r} \texttt{fpaddr} \ast \texttt{fpaddr} \mapsto_{a} (\texttt{wzero 64})}_{\rtv}$
 ;; Calculate new PTE by setting the low-order bits
@@
 mov (%r14), %rax
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,\gammaPred) \ast  \texttt{r14}\mapsto_{r} \texttt{pte\_addr} \ast \texttt{rdi}\mapsto_{r} \vaddr \ast \texttt{rax}\mapsto_{r} \_ \ast}_{\rtv}$
+$\specline{\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\vaddr,\rtv,\entryf,\entrytr,\entrytw,\entryo) \ast \ulcorner \theta \;!!\; \texttt{maddr} = \texttt{None}\urcorner}_{\rtv}$
+$\specline{\ulcorner \texttt{aligned fpaddr} \urcorner \ast \texttt{rax}\mapsto_{r} \texttt{fpaddr} \ast \texttt{fpaddr} \mapsto_{a} (\texttt{wzero 64})}_{\rtv}$
+$\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\; \sqsubseteq $
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta<[\vaddr:=\texttt{fpaddr}]>),\gammaPred,\rtv) \ast  \texttt{r14}\mapsto_{r} \texttt{pte\_addr} \ast \texttt{rdi}\mapsto_{r} \vaddr \ast \texttt{rax}\mapsto_{r} \_ }_{\rtv}$
+$\specline{\ulcorner \texttt{aligned fpaddr} \urcorner \ast \texttt{rax}\mapsto_{r} \texttt{fpaddr} \ast \texttt{aligned va} \ast\sumwalkabs\vaddr\qfrac\fpaddr}_{\rtv}$
 \end{lstlisting}
-\caption{Distilled code for mapping a new page}
+  \caption{Specification and proof of distilled code for mapping a new page with $\textsf{wpd\_def }_{\rtv}$}
+  \todo[inline,color=yellow]{Colin, could you speak about the fact that we have wpd in iProp so that's the reason to have rtval whereever table traversal is needed}
 \label{fig:mapping_code}
@@
 ;; Save yielding context
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,\gammaPred,\rtv) \ast [\rtv'+56](\mathcal{I}\texttt{ASpace}(\theta,\gammaPred') \ast \texttt{Pother})}_{\rtv}$
+$\specline{ \texttt{rsi}\mapsto_{r} \rtv' \ast \texttt{rdi}\mapsto_{r} \_ \ast \texttt{rbx}\mapsto_{r} \texttt{rbxv} \ast  \texttt{rsp}\mapsto_{r} \texttt{rspv} \ast \texttt{rbp}\mapsto_{r} \texttt{rbpv} }_{\rtv}$
+$\specline{\texttt{r12}\mapsto_{r} \texttt{r12v} \ast \texttt{r13}\mapsto_{r} \texttt{r13v} \ast \texttt{r14}\mapsto_{r} \texttt{r14v} \ast \texttt{r15}\mapsto_{r} \texttt{r15v}}_{\rtv}$
+$\specline{\texttt{rdi+0} \mapsto_{v,\rtv} \_ \ast \texttt{rdi+8} \mapsto_{v,\rtv} \_ \ast \texttt{rdi+16} \mapsto_{v,\rtv} \_ \ast \texttt{rdi+24} \mapsto_{v,\rtv} \_ \ast \texttt{rdi+32} \mapsto_{v,\rtv} \_}_{\rtv}$
+$\specline{\texttt{rdi+40} \mapsto_{v,\rtv} \_ \ast \texttt{rdi+48} \mapsto_{v,\rtv} \_\ast \texttt{rdi+56} \mapsto_{v,\rtv} \_}_{\rtv}$
 mov 0[%rdi], %rbx
@@
 mov 48[%rdi], %r15
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,\gammaPred,\rtv) \ast [\rtv'+56](\mathcal{I}\texttt{ASpace}(\theta,\gammaPred') \ast \texttt{Pother})}_{\rtv}$
+$\specline{ \texttt{rdi}\mapsto_{r} \_ \ast \texttt{rbx}\mapsto_{r} \texttt{rbxv} \ast  \texttt{rsp}\mapsto_{r} \texttt{rspv} \ast \texttt{rbp}\mapsto_{r} \texttt{rbpv} }_{\rtv}$
+$\specline{\texttt{r12}\mapsto_{r} \texttt{r12v} \ast \texttt{r13}\mapsto_{r} \texttt{r13v} \ast \texttt{r14}\mapsto_{r} \texttt{r14v} \ast \texttt{r15}\mapsto_{r} \texttt{r15v}}_{\rtv}$
+$\specline{\texttt{rdi+0} \mapsto_{v,\rtv} \texttt{rbxv} \ast \texttt{rdi+8} \mapsto_{v,\rtv} \texttt{rspv} \ast \texttt{rdi+16} \mapsto_{v,\rtv} \texttt{rbpv} \ast \texttt{rdi+24} \mapsto_{v,\rtv} \texttt{r12v} \ast}_{\rtv}$
+$\specline{ \texttt{rdi+32} \mapsto_{v,\rtv} \texttt{r13v} \ast \texttt{rdi+40} \mapsto_{v,\rtv} \texttt{r14v} \ast \texttt{rdi+48} \mapsto_{v,\rtv} \texttt{r15v}\ast \texttt{rdi+56} \mapsto_{v,\rtv} \_}_{\rtv}$
 mov 56[%rdi], %cr3
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,\gammaPred,\rtv) \ast [\rtv'+56](\mathcal{I}\texttt{ASpace}(\theta,\gammaPred') \ast \texttt{Pother})}_{\rtv}$
+$\specline{ \texttt{rdi}\mapsto_{r} \_ \ast \texttt{rbx}\mapsto_{r} \texttt{rbxv} \ast  \texttt{rsp}\mapsto_{r} \texttt{rspv} \ast \texttt{rbp}\mapsto_{r} \texttt{rbpv} }_{\rtv}$
+$\specline{\texttt{r12}\mapsto_{r} \texttt{r12v} \ast \texttt{r13}\mapsto_{r} \texttt{r13v} \ast \texttt{r14}\mapsto_{r} \texttt{r14v} \ast \texttt{r15}\mapsto_{r} \texttt{r15v}}_{\rtv}$
+$\specline{\texttt{rdi+0} \mapsto_{v,\rtv} \texttt{rbxv} \ast \texttt{rdi+8} \mapsto_{v,\rtv} \texttt{rspv} \ast \texttt{rdi+16} \mapsto_{v,\rtv} \texttt{rbpv} \ast \texttt{rdi+24} \mapsto_{v,\rtv} \texttt{r12v} \ast }_{\rtv}$
+$\specline{\texttt{rdi+32} \mapsto_{v,\rtv} \texttt{r13v} \ast \texttt{rdi+40} \mapsto_{v,\rtv} \texttt{r14v} \ast \texttt{rdi+48} \mapsto_{v,\rtv} \texttt{r15v}\ast \texttt{rdi+56} \mapsto_{v,\rtv} \rtv}_{\rtv}$
 ;; Restore target context
@@
 mov %r15, 48[%rsi]
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,\gammaPred,\rtv) \ast [\rtv'+56](\mathcal{I}\texttt{ASpace}(\theta,\gammaPred') \ast \texttt{Pother})}_{\rtv}$
+$\specline{ \texttt{rdi}\mapsto_{r} \_ \ast \texttt{rbx}\mapsto_{r} \rtv'+0 \ast  \texttt{rsp}\mapsto_{r} \rtv'+8 \ast \texttt{rbp}\mapsto_{r} \rtv'+16 }_{\rtv}$
+$\specline{\texttt{r12}\mapsto_{r} \rtv'+24 \ast \texttt{r13}\mapsto_{r} \rtv'+32 \ast \texttt{r14}\mapsto_{r} \rtv'+40 \ast \texttt{r15}\mapsto_{r} \rtv'+48}_{\rtv}$
+$\specline{\texttt{rdi+0} \mapsto_{v,\rtv} \texttt{rbxv} \ast \texttt{rdi+8} \mapsto_{v,\rtv} \texttt{rspv} \ast \texttt{rdi+16} \mapsto_{v,\rtv} \texttt{rbpv} \ast \texttt{rdi+24} \mapsto_{v,\rtv} \texttt{r12v} \ast }_{\rtv}$
+$\specline{\texttt{rdi+32} \mapsto_{v,\rtv} \texttt{r13v} \ast \texttt{rdi+40} \mapsto_{v,\rtv} \texttt{r14v} \ast \texttt{rdi+48} \mapsto_{v,\rtv} \texttt{r15v}\ast \texttt{rdi+56} \mapsto_{v,\rtv} \rtv}_{\rtv}$
 ;; Switch to the new address space
 mov %cr3, 56[%rsi]
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,\gammaPred) \ast [\rtv'+56](\mathcal{I}\texttt{ASpace}(\theta,\gammaPred') \ast \texttt{Pother})}_{\rtv}$
+$\specline{ \texttt{rdi}\mapsto_{r} \_ \ast \texttt{rbx}\mapsto_{r} \rtv'+0 \ast  \texttt{rsp}\mapsto_{r} \rtv'+8 \ast \texttt{rbp}\mapsto_{r} \rtv'+16 }_{\rtv}$
+$\specline{\texttt{r12}\mapsto_{r} \rtv'+24 \ast \texttt{r13}\mapsto_{r} \rtv'+32 \ast \texttt{r14}\mapsto_{r} \rtv'+40 \ast \texttt{r15}\mapsto_{r} \rtv'+48}_{\rtv}$
+$\specline{\texttt{rdi+0} \mapsto_{v,\rtv} \texttt{rbxv} \ast \texttt{rdi+8} \mapsto_{v,\rtv} \texttt{rspv} \ast \texttt{rdi+16} \mapsto_{v,\rtv} \texttt{rbpv} \ast  \texttt{rdi+24} \mapsto_{v,\rtv} \texttt{r12v} \ast }_{\rtv}$
+$\specline{\texttt{rdi+32} \mapsto_{v,\rtv} \texttt{r13v} \ast \texttt{rdi+40} \mapsto_{v,\rtv} \texttt{r14v} \ast \texttt{rdi+48} \mapsto_{v,\rtv} \texttt{r15v}\ast \texttt{rdi+56} \mapsto_{v,\rtv} \rtv}_{\rtv}$
+$\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\; \sqsubseteq $
+$\specline{ [\rtv](\texttt{Pcurrent}) \ast \mathcal{I}\texttt{ASpace}(\theta,\gammaPred') \ast \texttt{Pother}}_{\rtv'+56}$
 \end{lstlisting}

### 2023-04-10 background.tex
lots of comments

--- 2022-09-19 background.tex
@@
     \caption{x86-64 page table lookups.}
+        \todo[inline]{Colin says: I recommend the cr3 caption say "Physical Base Address" rather than just
+                base address.
+            If we're saying in the text that we just to L4, L3, L2, L1, then the figure should do the same.
+        }
     \label{fig:pagetables}
@@
 
-On the x86-64 architecture, the input to the \textsc{MMU}'s address translation is a sparse hierarchical set of tables calle d \emph{page tables} (referring to pages of memory).
-As Figure \ref{fig:pagetables}\mytodo{cite AMD ref} shows, address translation proceeds by repeatedly taking designated slices of the virtual address and indexing into a table.
-On x86-64, standard configurations use 4 levels of page tables, labelled levels 4 through 1, with lookups in the level 1 page table resulting in the actual page of physical memory holding the requested data, and the low-order 12 bits being used to index into this page.\footnote{Technically levels 1--3 have designated names originating when there were fewer levels.
-We omit those names to avoid confusion, and for greater consistency with the recent addition of a 5th level on the most recent processors, where levels 4 and 5 are simply known by their numbers.}  Because address translation traverses a series of tables, the translation process or algorithm is sometimes referred to as a \emph{page-table walk}.
-While Figure \ref{fig:pagetables} and most of our details are specific to the x86-64 architecture, ARMv8 (a.k.a.\ \texttt{aarch64}), RISC-V, and PowerPC use similar hierarchical page tables for address translation.
+On the x86-64 architecture, the input to the \textsc{MMU}'s address translation is a sparse hierarchical set of tables 
+called \emph{page tables} (referring to pages of memory). As Figure \ref{fig:pagetables} (based on Figure 5-17 of the 
+AMD64 architecture manual~\cite{amd64_manual_vol2}\footnote{While x86 up through its 32-bit incarnation were due to Intel,
+the x86-64 architecture as a 64-bit extension to x86 was originally due to AMD. As a result, it is sometimes also referred to as the \texttt{amd64} architecture.}) shows, 
+address translation proceeds by repeatedly taking designated slices of the virtual address and indexing into a table.
+The final lookup in the page tables gives the base physical address of a page of physical memory, to which the low-order
+bits of the accessed virtual address are added to determine the actual physical address retrieved. 
+On x86-64, standard configurations use 4 levels of page tables, labelled levels 4 through 1, with lookups in the level 
+1 page table resulting in the actual page of physical memory holding the requested data, and the low-order 12 bits 
+being used to index into this page.\footnote{Technically levels 1--3 have designated names originating when there were 
+fewer levels. We omit those names to avoid confusion for those not already steeped in the arcana of x86 history,
+ and for greater consistency with the recent addition of a 5th level 
+on the most recent processors, where levels 4 and 5 are simply known by their numbers. Our formalization
+only deals with 4-level page tables, but is straightforwardly extensible to 5.}  
+Because address translation 
+traverses a series of tables, the translation process or algorithm is sometimes referred to as a \emph{page-table walk}. 
+While Figure \ref{fig:pagetables} and most of our details are specific to the x86-64 architecture, ARMv8 (a.k.a.\ 
+\texttt{aarch64}), RISC-V, and PowerPC use similar hierarchical page tables for address translation.
 
@@
 
-The page tables mapping virtual addresses to physical addresses are managed by the OS.
-Typically each process has its on virtul-to-physical mappings represented by its own page table, which the OS registers with the CPU by storing it in a specific register (\texttt{cr3}) as part of switching to a new process.
-Using different mappings, which map only disjoint portions of physical memory (with some exceptions in the next section) is how the OS ensures memory isolation between processes.
+The page tables mapping virtual addresses to physical addresses are managed by the OS. Typically each process has its own 
+virtual-to-physical mappings represented by its own page table, which the OS registers with the CPU by storing the (page-aligned)
+physical address of the root of the page table tree (in our context, the start of the L4 table) in a 
+specific register (\texttt{cr3}) as part of switching to a new process. 
+Using different mappings, which map only disjoint portions of physical memory (with some exceptions in the next section) 
+is how the OS ensures memory isolation between processes.
 
@@
 \label{sec:backgroundonvmm}
-Most OS kernels have a component called a \emph{Virtual Memory Manager} (VMM)\footnote{Not to be confused with Virtual Machine Monitor.
-In this paper we focus on non-virtualized scenarios, but hardware virtualization extensions for both x86-64 and ARM make use of an additional set of page tables translating what a \emph{guest} considers to be its (virutalized) physical memory to actual physical memory.
-Our contributions should offer value in this scenario as well.}
-which is responsible for setting up page table mappings and for taking action when a page fault occurs.
-Often, a page fault indicates a bug in a program --- for example, a null pointer dereference crashes a program not because the hardware designates it an error, but because most OSes refuse to map the first 4K worth of virtual addresses, so dereferencing \texttt{NULL} results in a page fault.
-In such cases, the OS will terminate the program.
+Most OS kernels have a component called a \emph{Virtual Memory Manager} (VMM)\footnote{Not to be confused with Virtual 
+Machine Monitor. In this paper we focus on non-hypervisor scenarios, but hardware virtualization extensions for both 
+x86-64 and ARM make use of an additional set of page tables translating what a \emph{guest} considers to be its 
+(virtualized) physical memory to actual physical memory. Our contributions should offer value in this scenario as well.}
+which is responsible for setting up page table mappings and for taking action when a page fault occurs. Often, a page 
+fault indicates a bug in a program --- for example, a null pointer dereference crashes a program not because the hardware
+ designates it an error, but because most OSes refuse to map the first 4K worth of virtual addresses, so dereferencing 
+\texttt{NULL} results in a page fault. In such cases, the OS will terminate the program.
 
-In other cases the OS uses page faults to implement specialized functionality and optimizations.
-A key example of this is (confusingly) called \emph{paging}: saving room in physical RAM by waiting until memory addresses for code are accessed before copying them to RAM (saving process startup time); or copying memory that has not recently been accessed to disk (i.e., in a \emph{swap} file or partition) and marking those page table entries invalid so a future page fault will give the OS the chance to copy the relevant data back into memory when the program tries to access it again.
+In other cases the OS uses page faults to implement specialized functionality and optimizations. A key example of this is 
+(confusingly) called \emph{paging}: saving room in physical RAM and avoiding unnecessary IO 
+by waiting until memory addresses for code are accessed 
+before copying them to RAM (in particular, saving process startup time); 
+or copying memory that has not recently been accessed to disk 
+(i.e., in a \emph{swap} file or partition) and marking those page table entries invalid so a future page fault will give 
+the OS the chance to copy the relevant data back into memory when the program tries to access it again.
 
-The key pieces of functionality a VMM must implement itself, or have available in another OS component (depending on design), are
-adding a new page mapping (whether the mapped page contains zeros, file data, or swap data), and removing an existing page mapping.
-While this initially sounds like relatively modest functionality whose implementation may be complicated by hardware subtleties, a number of more serious complexities must be handled.
+The key pieces of functionality a VMM must implement itself, or have available in another OS component (depending on 
+design), are
+adding a new page mapping (whether the mapped page contains zeros, file data, or swap data), and removing an existing 
+page mapping.
+While this initially sounds like relatively modest functionality whose implementation may be complicated by hardware 
+subtleties, a number of more serious complexities must be handled.
 
+\todo[inline]{Colin says: we should chat about the paragraph below. We don't want to advertise too many
+challenges of VMM verification that we don't actually solve (or rather, we'd push them to the and of the paper
+as future work). But some of these I think we could split the difference with, for example by defining
+a variant of virtual-pointsto that allowed virtual address aliasing, but not actually verifying any code examples
+with it (just showing it's compatible with our approach).
+}
 For one, virtual addresses may \emph{alias} when more than one is mapped to the same physical memory. 
 This can occur when a process requests it (e.g., via \texttt{mmap} or \texttt{shm\_open} on UNIX-style kernels). 
-It also occurs automatically in most general-purpose kernels: while for normal programs the OS uses the MMU to isolate and abstract physical memory, the kernel's own functionality is often easier to implement if the kernel can directly access memory given its physical address, for example when interacting with hardware devices.
-For this reason, most kernels contain an \emph{identity map} of all physical memory in a machine.
-In some cases this is a literal identity map, where for every virtual address $p$, interpreting $p$ as a virtual address instead will map back to $p$ as a physical address.
-In other cases there is an offset involved, where to access physical address $p$ the kernel can access virtual address $p+\mathit{offset}$.
-This identity map saves the kernel the performance cost of having to continually add and remove transient mappings simply to briefly access a specific physical memory region.
-But this coexists with additional virtual addresses (those used by most kernel code and data structures), in a form of virtual address aliasing.
-Note that virtual address aliasing is both prevalent in general-purpose kernels, and violates the core assumption of most memory models assumed by verified compilers like CompCert, which assume no virtual address aliasing.
+It also occurs automatically in most general-purpose kernels: while for normal programs the OS uses the MMU to isolate 
+and abstract physical memory, the kernel's own functionality is often easier to implement if the kernel can directly 
+access memory given its physical address, for example when interacting with hardware devices. For this reason, most 
+kernels contain an \emph{identity map} of all physical memory in a machine. 
+\todo[inline]{We should try defining an identity map on a whiteboard and see what breaks. If it's feasible we should
+tweak and get a definition in the paper. If not, we should cut the identity map discussion.}
+In some cases this is a literal identity map, 
+where for every virtual address $p$, interpreting $p$ as a virtual address instead will map back to $p$ as a physical 
+address. In other cases there is an offset involved, where to access physical address $p$ the kernel can access virtual
+ address $p+\mathit{offset}$. This identity map saves the kernel the performance cost of having to continually add and
+ remove transient mappings simply to briefly access a specific physical memory region. But this coexists with additional
+ virtual addresses (those used by most kernel code and data structures), in a form of virtual address aliasing. Note 
+that virtual address aliasing is both prevalent in general-purpose kernels, and violates the core assumption of most 
+memory models assumed by verified compilers like CompCert, which assume no virtual address aliasing.
 
-Another complicating factor is that portions of page tables can be shared, i.e., trees of page tables can overlap.
-One common way this used to happen was for one subtree of page tables to map the kernel if linked in appropriately to a higher-level page table, and then to reuse that kernel mapping portion in every process's page tables.
-This was mostly phased out in the wake of Spectre\mytodo{Colin finish}
+\todo[inline]{Colin says: Again, let's chat about the next paragraph. This might be another where we can
+split the difference on extension by defining something and proving a $\ast-\ast$ kind of result for sharing
+page table fragments, rather than actually proving any new code. I'm less sure about this one, though.
+}
+Another complicating factor is that portions of page tables can be shared, i.e., trees of page tables can overlap and
+become forests. 
+One common way this used to happen was for one subtree of page tables to map the kernel if linked in appropriately to 
+a higher-level page table, and then to reuse that kernel mapping portion in every process's page tables. This was mostly 
+phased out in the wake of Spectre, but remains relevant on correctly-implemented hardware.
 
-These dove-tail with functionality in the OS scheduler, which since it is responsible for dealing with multiple address spaces, must keep track of which virtual addresses are valid (and in what way) in which address spaces, where some virtual addresses are valid in only a single address space (e.g., a code address for a particular usermode process), while others are valid in all address spaces (e.g., kernel data structure pointers).
-The VMM must maintain some of these assumptions on behalf of the rest of the kernel, for example by guaranteeing that a certain range of virtual addresses (corresponding to the kernel's code and data) are valid in every address space.
+These dove-tail with functionality in the OS scheduler, which since it is responsible for dealing with multiple 
+address spaces, must keep track of which virtual addresses are valid (and in what way) in which address spaces, 
+where some virtual addresses are valid in only a single address space (e.g., a code address for a particular usermode 
+process), while others are valid in all address spaces (e.g., kernel data structure pointers). 
+The VMM must maintain some of these assumptions on behalf of the rest of the kernel, for example by guaranteeing that 
+a certain range of virtual addresses (corresponding to the kernel's code and data) are valid in every address space.
 
@@
 \label{sec:backgroundonmodallogic}
-The problem of needing to keep track of things being true in some contexts and not in others is hardly unique to virtual memory management, and is the general insight behind most flavors of modal logic.
+The problem of needing to keep track of things being true in some contexts and not in others is hardly unique to virtual 
+memory management, and is the general insight behind most flavors of modal logic.
 
-Most modal logics use unary operators on propositions to express that a logical claim $P$ is \emph{contingently} true in certain other circumstances, such as in other times~\cite{pnueli1977} or places~\cite{gordon2019modal}.
-That is, for different modal operators $M$, $M(P)$ may mean that $P$ is true in another time or place (which may be a physical place like another network node~\cite{murphy,gordon2019modal} or a conceptual place like a person's mind~\cite{hintikka1962knowledge}).
-A unifying concept across any modality is that they behave as applicative functors, typically satisfying (directly, or as a derived law):
+Most modal logics use unary operators on propositions to express that a logical claim $P$ is \emph{contingently} true 
+in certain other circumstances, such as in other times~\cite{pnueli1977temporal} or places~\cite{gordon2019modal}. That is, for 
+different modal operators $M$, $M(P)$ may mean that $P$ is true in another time or place (which may be a physical place 
+like another network node~\cite{murphy2008type,gordon2019modal} or a conceptual place like a person's 
+mind~\cite{hintikka1962knowledge}). A unifying concept across any modality is that they behave as applicative functors, 
+typically satisfying (directly, or as a derived law):
 \[ (P\rightarrow Q) \rightarrow M(P) \rightarrow M(Q)\]
-\mytodo{Ismail, I think this means we have pure intro, $\square P \mathrel{-\ast} [r](\square P)$ if I'm using the right symbol for pure assertions }
-Many modalities, so-called \emph{normal} modalities also possess introduction rules of the form $P\rightarrow M(P)$, the classic example being that if $P$ is true, then $P$ is \emph{necessarily} true ($\square P$)
+\todo[inline]{Ismail, I think this means we have pure intro, $\square P \mathrel{-\ast} [r](\square P)$ if I'm using the right symbol for pure assertions }
+\todo[inline,color=red]{The above todo comment isn't quite right.
+Purity is about being able to duplicate. You had defined another typeclass/property-of-assertions that meant it didn't
+care what address space it was in (like physical pointstos). That's orthogonal to purity.
+}
+Many modalities, so-called \emph{normal} modalities also possess introduction rules of the form $P\rightarrow M(P)$, 
+the classic example being that if $P$ is true, then $P$ is \emph{necessarily} true ($\square P$).
 
-Of particular interest for reasoning about virtual memory are modalities that permit \emph{naming} the alternate circumstances, prominently \emph{hybrid} modal logics~\cite{blackburn1995hybrid,areces2001hybrid}, which come equipped with assertions of the form $@_\ell(P)$ indicating that $P$ is true in the specific alternate circumstance (Kripke world) named by the \emph{nominal} $\ell$.
+Of particular interest for reasoning about virtual memory are modalities that permit \emph{naming} the alternate 
+circumstances, prominently \emph{hybrid} modal logics~\cite{blackburn1995hybrid,areces2001hybrid}, which come equipped 
+with assertions of the form $@_\ell(P)$ indicating that $P$ is true in the specific alternate circumstance (Kripke world)
+ named by the \emph{nominal} $\ell$.
 
-For our purposes, these are natural candidates to adapt for virtual memory management.
-We can reinterpret the notion of naming an alternate world slightly more loosely, and instead name \emph{address spaces} by the physical address of the page table root, since these structures are the physical representations of page tables.
-Thus in this paper we develop the notion that we can represent contingent truth of an assertion via $[r](P)$ indicating that $P$ holds in the address space rooted at physical address $r$.
-The typical hybrid modality introduction rule, that $P$ and $\ell$ (indicating the current world is $\ell$) imply $@_\ell(P)$, has a natural analogue: knowing $P$ and that the current address space is $r$ (i.e., that $r$ is the current \texttt{cr3} value) suggests a way to construct $[r](P)$.
+For our purposes, these are natural candidates to adapt for virtual memory management. We can reinterpret the notion of 
+naming an alternate world slightly more loosely, and instead name \emph{address spaces} by the physical address of the 
+page table root, since these structures are the physical representations of page tables. Thus in this paper we develop 
+the notion that we can represent contingent truth of an assertion via $[r](P)$ indicating that $P$ holds in the address 
+space rooted at physical address $r$.
+The typical hybrid modality introduction rule, that $P$ and $\ell$ (indicating the current world is $\ell$) imply 
+$@_\ell(P)$, has a natural analogue: knowing $P$ and that the current address space is $r$ (i.e., that $r$ is the 
+current \texttt{cr3} value) suggests a way to construct $[r](P)$.
 
-A relatively under-explored space of modal logics is the interaction of modal and substructural logics, in particular hybrid-style modalities in substructural logics, which has seen only minimal exploration and no prior application.
-We develop our ideas in the Iris program logic~\cite{jung2018}, as a result exploring additional subtleties that arise where the modality itself may entail ownership of resources, as well as interactions between our hybrid-style modality and substructural rules.
-For example, Iris contains a number of modalities that distribute over separating conjunction, or for which resources can freely move into the modality (e.g., $\blacktriangleright(P)\ast Q \mathrel{-\ast} \blacktriangleright(P\ast Q)$).
-In our setting some of these rules apply while others do not.
-For example, in our setting, an assertion that involves no modalities is interpreted as holding in the current (active-on-the-CPU) address space, so clearly cannot move into arbitrary other address spaces --- unless guarded by another address space modality.
+A relatively under-explored space of modal logics is the interaction of modal and substructural logics, 
+in particular hybrid-style modalities in substructural logics, which has seen only minimal exploration~\todo{copy refs from fs paper} and no prior 
+application. We develop our ideas in the Iris program logic~\cite{jung2018iris}, as a result exploring additional subtleties 
+that arise where the modality itself may entail ownership of resources, as well as interactions between our hybrid-style 
+modality and substructural rules.  For example, Iris contains a number of modalities that distribute over separating 
+conjunction, or for which resources can freely move into the modality 
+(e.g., $\blacktriangleright(P)\ast Q \wand \blacktriangleright(P\ast Q)$). In our setting some of these rules 
+apply while others do not. For example, in our setting, an assertion that involves no modalities is interpreted as 
+holding in the current (active-on-the-CPU) address space, so clearly cannot move into arbitrary other address spaces 
+--- unless guarded by another address space modality.

### 2023-04-10 cpp_paper.tex
+
removed the old stuff
replace the figure
lots of comments

--- 2022-09-25 cpp_paper.tex
@@
 %%
-\documentclass[acmsmall]{acmart}
+\documentclass[acmsmall,screen,review,anonymous]{acmart}
 
@@
 %% complete the rights form.
-\setcopyright{acmcopyright}
+\setcopyright{none}
+%\setcopyright{acmcopyright}
 \copyrightyear{2018}
@@
 %%\citestyle{acmauthoryear}
-
+\usepackage{lipsum}
 \usepackage[utf8]{inputenc}
@@
 %\invertbackgroundtext
+\usepackage{wrapfig}
+\newtheorem*{remark}{Remark}
 
@@
 \usepackage{amsmath}
+% swiped from iris.sty in the iris repo
+\newcommand{\wand}{\mathrel{-\!\!\ast}}
+\newcommand{\wandIff}{\mathrel{\ast\!\!{-}\!\!\ast}}
 % If we need multiple references to a footnote, via \cref:
@@
 %% allowing the author to define a "short title" to be used in page headers.
-\title{Modalities for Virtualizing Memory Addresses}
-\subtitle{Understanding the Locality with Modals}
+\title{Modal Abstractions for Virtualizing Memory Addresses}
 %%
@@
 \begin{abstract}
-Operating system kernels employs virtual memory subsystem which employs memory management units on each CPU to virtualize the addresses of memory regions:a logical (virtual) address which is translated to a physical address in memory by the MMU based on
-kernel-controlled page tables -- a hardware-defined sparse tree-map structure -- stored in memory, which itself is accessed by the kernel through virtual addresses.
-With this core mechanism, operating system kernels generalizes its memory-resource service to different contexts, mainly through the manipulating and aliasing 
-same physical memory to multiple virtual addresses through the tree structure: subtrees or data pages may be shared across processes (e.g., the kernel mapping),
-accessing an unmapped address produces a fault which the OS can use to kill the process or take corrective action (like add a mapping), and resume the interrupted code which constitutes the kernel's capability to restrict which memory is accessible to different processes, hide memory limits from user programs, ensure process isolation, implement paging and copy-on-write behaviors for performance
+Operating system kernels employ virtual memory subsystems, which use a CPU's memory management units (MMUs) to virtualize the addresses of memory regions:
+a logical (virtual) address is translated to a physical address in memory by the MMU based on
+kernel-controlled page tables -- a hardware-defined sparse tree-map structure -- stored in memory, which itself is 
+accessed by the kernel through virtual addresses.
+Operating systems manipulate these virtualized memory mappings to isolate untrusted processes,
+ restrict which memory is accessible to different processes, 
+hide memory limits from user programs, 
+ensure process isolation, implement demand-paging and copy-on-write behaviors for performance
 and resource controls.
+At the same time, misuse of MMU hardware can lead to kernel crashes.
 
-So, verifying a full-featured version of the essential virtual memory manager systems is a crucial step towards achieving the progress towards handling fully verified real kernels written using these designs.
-Regarding this effort, in this paper, we introduce a modal abstraction to describe
-the truth of assertions in the virtual resources corresponding to virtual address spaces: [r]P indicating that P holds in the
-virtual address space rooted at r, which implicitly treats the memory state as a modal frame and the choice of
-page table root as a world in Kripke-style semantics. Moreover, we define virtual points-to assertions defined in terms of physical points-to assertions
-and hardware-specific address translation assertions, and paged points-to assertions.
-Finally, we experiment our abstractions in specifying and proving demand paging of a fresh page of zeros and context
-switches between address spaces.
-All definitions and theorems mentioned in this paper including the operational model of a RISC-like fragment of x86-64, a simple language run on this operational model, and a logic as an instantiation of the Iris framework are mechanized inside Coq.
+Virtual memory management (VMM) code is a critical piece of general-purpose OS kernels, but verification of this functionality
+is challenging due to the complexity of the hardware interface (the page tables are updated via writes to those
+memory locations, using addresses which are themselves virtualized).
+Prior work on verification of VMM code has either only handled a single address space, trusted significant
+pieces of assembly code, or resorted to direct reasoning over machine semantics rather than
+exposing a clean logical interface.
+
+In this paper, we introduce a modal abstraction to describe
+the truth of assertions relative to a specific virtual address space: [r]P indicating that P holds in the
+virtual address space rooted at r. Such modal assertions 
+allow different address spaces to refer to each other, enabling complete verification of instruction sequences
+manipulating multiple address spaces. Using them effectively requires working with other assertions,
+such as points-to assertions in our separation logic, as relative to a given address space.
+We therefore define virtual points-to relations, which mimic hardware address translation,
+relative to a page table root.
+We demonstrate our approach with challenging fragments of VMM code showing that our approach
+handles examples beyond what prior work can address, including reasoning about
+a sequence of instructions as it changes address spaces.
+All definitions and theorems mentioned in this paper including the operational model of a RISC-like fragment of x86-64, 
+a simple language run on this operational model, and a logic as an instantiation of the Iris framework are mechanized 
+inside Coq.
 \end{abstract}
@@
 %%
-\begin{CCSXML}
-<ccs2012>
- <concept>
-  <concept_id>10010520.10010553.10010562</concept_id>
-  <concept_desc>Computer systems organization~Embedded systems</concept_desc>
-  <concept_significance>500</concept_significance>
- </concept>
- <concept>
-  <concept_id>10010520.10010575.10010755</concept_id>
-  <concept_desc>Computer systems organization~Redundancy</concept_desc>
-  <concept_significance>300</concept_significance>
- </concept>
- <concept>
-  <concept_id>10010520.10010553.10010554</concept_id>
-  <concept_desc>Computer systems organization~Robotics</concept_desc>
-  <concept_significance>100</concept_significance>
- </concept>
- <concept>
-  <concept_id>10003033.10003083.10003095</concept_id>
-  <concept_desc>Networks~Network reliability</concept_desc>
-  <concept_significance>100</concept_significance>
- </concept>
-</ccs2012>
-\end{CCSXML}
+% \begin{CCSXML}
+% <ccs2012>
+%  <concept>
+%   <concept_id>10010520.10010553.10010562</concept_id>
+%   <concept_desc>Computer systems organization~Embedded systems</concept_desc>
+%   <concept_significance>500</concept_significance>
+%  </concept>
+%  <concept>
+%   <concept_id>10010520.10010575.10010755</concept_id>
+%   <concept_desc>Computer systems organization~Redundancy</concept_desc>
+%   <concept_significance>300</concept_significance>
+%  </concept>
+%  <concept>
+%   <concept_id>10010520.10010553.10010554</concept_id>
+%   <concept_desc>Computer systems organization~Robotics</concept_desc>
+%   <concept_significance>100</concept_significance>
+%  </concept>
+%  <concept>
+%   <concept_id>10003033.10003083.10003095</concept_id>
+%   <concept_desc>Networks~Network reliability</concept_desc>
+%   <concept_significance>100</concept_significance>
+%  </concept>
+% </ccs2012>
+% \end{CCSXML}
 
-\ccsdesc[500]{Computer systems organization~Embedded systems}
-\ccsdesc[300]{Computer systems organization~Redundancy}
-\ccsdesc{Computer systems organization~Robotics}
-\ccsdesc[100]{Networks~Network reliability}
+% \ccsdesc[500]{Computer systems organization~Embedded systems}
+% \ccsdesc[300]{Computer systems organization~Redundancy}
+% \ccsdesc{Computer systems organization~Robotics}
+% \ccsdesc[100]{Networks~Network reliability}
 
@@
 %% the work being presented. Separate the keywords with commas.
-\keywords{program verification}
+% \keywords{program verification}
 
@@
 \input{background.tex}
-\input{syntax.tex}
-\input{semantics.tex}
+\input{syntaxsemantics.tex}
 \input{programlogic.tex}
@@
 \bibliography{vmm}
-\begin{appendix}
-\section{Complete Syntax}
-\label{appendix:syntax}
-\begin{figure}[!ht]
-\newcommand{\commentary}[1]{ & \text{\small\it #1} \\}
-\[
-  \begin{array}{r@{\;}c@{\;}l}
-    \loc & \in & \Loc \\
-    \reg & \in & \regset \\
-    \regval & \in & \regvaltype \\
-    \val & ::= & \vunit
-
-\\
-    \instrs & ::= &
-    \begin{array}[t]{@{}l@{\hspace{10mm}}l@{}}
-    \begin{array}[t]{@{}ll@{}}
-      \iskip
-                   \commentary{no-op}
-      \iseq\instr\instrs
-                   \commentary{sequencing}
-      \ising\instr
-                   \commentary{executing}             
-    \end{array}
-    %&
-    %\begin{array}[t]{@{}ll@{}}
-     % \ialloc\lval\allocsize
-      %             \commentary{heap allocation}
-    %\end{array}
-    \end{array}
-    \\
-
-    \ectx & ::= &
-      \hole \mid
-      \iseq\ectx\instrs 
-    \\
-  \end{array}
-\]
-\caption{Syntax}
-\Description{Syntax}
-\label{fig:syntax}
-\end{figure}
-\section{Operational Semantics of Selected Mov Instructions}
-\label{appendix:movops}
-\input{fig-semantics}
-
-\end{appendix}
 

### 2023-04-10 intro.tex
+
lots of comments

@@
 
-At its core, the virtual memory functionality of modern CPUs is about \emph{location virtualization}: the memory locations (addresses) seen by most code are not, in fact, the exact location in physical memory where data reside.
-Instead these are \emph{virtual} addresses, which are mapped to actual physical resources by the cooperation of the hardware and OS.
+At its core, the virtual memory functionality of modern CPUs is about \emph{location virtualization}: the memory locations
+(addresses) seen by most code are not, in fact, the exact location in physical memory where data reside. Instead these 
+are \emph{virtual} addresses, which are mapped to actual physical resources by the cooperation of the hardware and OS. 
 This is what enables separation of process memory resources:
-the OS manipulates hardware functionality to ensure that any attempt by a process to access memory not explicitly granted to it by the kernel will fail.
-But this is complicated by the fact that the OS and hardware can also enable, shared (overlapping) access to physical memory regions; the fact that the kernel data structures themselves are accessed via virtual memory addresses; and the fact that control over these mappings of virtual to physical addresses is itself mediated by \emph{in-memory data structures}, which the kernel still accesses via virtual address, leading to indirect cycles.
+the OS manipulates hardware functionality to ensure that any attempt by a process to access memory not explicitly granted 
+to it by the kernel will fail. But this is complicated by the fact that 
+%the OS and hardware can also enable, shared (overlapping) access to physical memory regions; the fact that the kernel data structures themselves are accessed via virtual memory addresses; 
+%and the fact that 
+control over these mappings of virtual to physical addresses is itself mediated by \emph{in-memory data structures}, 
+which the kernel still accesses via virtual address, leading to indirect cycles.
 
-Further complicating matters is that addresses themselves bear no information about which address space they originate from.
-For user processes this is of little concern, as these have access to only their own address space.
-But the kernel has access to all address spaces.
-Mixing up addresses from different address spaces leads to severe bugs.
-More concerning, keeping track of which \emph{assertions} hold in different address spaces during kernel verification is difficult: some assertions should hold across all address spaces, while others hold in only one, and others may hold in multiple but still not all.
+Further complicating matters is that addresses themselves bear no information about which address space they originate 
+from. For user processes this is of little concern, as these have access to only their own address space. But the kernel has
+(or can grant itself) access to all address spaces. Mixing up addresses from different address spaces leads to severe bugs.
+More concerning, keeping track of which \emph{assertions} hold in different address spaces during kernel verification is 
+difficult: some assertions should hold across all address spaces, while others hold in only one, and others may hold in 
+multiple but still not all.
 
 This kind of context-dependent assertion, where a fact may be true in one address space but not others, has a modal flavor. 
-We propose tackling the verification of virtual memory subsystems (and kernels more broadly) by adapting ideas from hybrid modal logic, which can label assertions true under \emph{other} circumstances (i.e., in another address space) with a modality indexed by a name for that space (in our case, the root of the page tables for an address space).
-This offers a \textit{convenient} and \textit{powerful way to modularly}
+We propose tackling the verification of virtual memory subsystems (and kernels more broadly) by adapting ideas from hybrid
+modal logic, which can label assertions true under \emph{other, named} circumstances (i.e., in another address space) with a 
+modality indexed by a name for that space (in our case, the root of the page tables for an address space). This offers a 
+\textit{convenient} and \textit{powerful} way to \emph{modularly}
 \begin{itemize}
@@
 \end{itemize}
-  These advantages make this approach to reasoning about virtual memory more flexible than prior program logic techniques~\cite{kolanski08vstte,kolanski09tphols}, which was only able to work with a single address space (the current address space on the CPU).
+These advantages make this approach to reasoning about virtual memory more flexible than prior program logic techniques~\cite{kolanski08vstte,kolanski09tphols}, 
+which were only able to work with a single address space (the current address space on the CPU) because they were unable
+to speak directly \emph{within the logic} about other address spaces.
 
@@
   \caption{Layers of Implementation}
+  \todo[inline]{Colin says: I don't think this figure helps readers; it is too abstract}
   \label{fig:architecture}
 \end{figure}
-We realize and show our reasoning principles, showin in Figure \ref{fig:architecture}
+
+We realize and show our reasoning principles, shown in Figure \ref{fig:architecture}
 \begin{itemize}
-\item we develop these ideas in an Iris-based\mytodo{cite JFP} separation logic, a logic for address-virtualization (\textsf{vProp} in Figure \ref{fig:architecture})
-\item we model \textsf{AMD64} in Coq, and specify and prove its instructions in \textsf{vProp}
-\item and verify simplified versions of several critical virtual-memory-related pieces of OS functionality, including mapping and unmapping pages, and describing the self-referential identity maps used in most kernels to allow convenient access to physical memory.
+\item We develop these ideas in the form of a logic for working with virtual-address-space-relative assertions,
+      implemented as an embedded separation logic within Iris~\cite{jung2018iris} separation logic.
+\item We prove the soundness of our \textsf{vProp} logic with respect to a RISC-like fragment of \textsf{AMD64} instructions.
+\item We verify simplified versions of several critical virtual-memory-related pieces of OS functionality, 
+      including mapping  pages, the assembly code for switching address spaces, and code which manipulates page
+      tables for \emph{currently mapped memory} without disturbing mappings.\todo[inline]{Colin says: I have optimistically
+      added this last example to the text here, because I think we can probably do it. Leave this TODO
+      here until we either do or cut it.}
+      The latter in particular goes beyond the capabilities of prior work.
 \end{itemize}

### 2023-04-10 programlogic.tex
small stuff on figures
some cr3 removal from defs
+
lots of comments

@@
 \newcommand{\ptablestore}{\theta}
-Our reasoning principles are shaped around understanding the following perspectives on the vritual-memory abstraction:
+Our reasoning principles derive from the following perspectives on the virtual-memory abstraction:
 \begin{enumerate}
 \item \textit{logical representation of addressing}: we present physical memory addressing as separation-logic points-to relations.
-\item \textit{sharing}: a bag of physical pages are reached through a set of physical page-table (L4-L1 page-tables) acceses that are shared amongst different virtual addresses.
-This sharing imposes constraints on giving a definition to the virtual-address in terms of physical (L4-L1) page-table memory accesses
-\item \textit{context-agnostic-resources}: each virtual address is valid under a certain address-space, but it does not represent this \textit{knowledge} on its address-space. 
-\item \textit{address-spaces as modal-contexts}: we call these context-agnostic resources (virtual-address pointstos) as \textit{context-resource}, and the context identfied by a unique \textit{namespace}, and  defining the truth for these context resources as \text{modal-context}
+\item \textit{sharing}: the physical location backing a virtual address is located (during a page table walk) through a 
+      set of physical page-table (L4-L1 page-tables) acceses that are shared amongst different virtual addresses (specifically,
+      those on the same page of memory). This sharing imposes constraints on giving a definition to the virtual-address 
+      in terms of physical (L4-L1) page-table memory accesses
+\item \textit{context-agnostic-resources}: each virtual address is valid under a certain address-space, but it does not represent this \textit{knowledge} of its address-space. 
+\item \textit{address-spaces as modal contexts}: we call these context-dependent\todo{Colin switched away from `agnostic', which is the opposite!} resources (virtual-address pointstos) as \textit{context-resource}, and the context identfied by a unique \textit{namespace}, and  defining the truth for these context resources as \text{modal-context}
 \item \textit{updating inter/intra address-spaces}: we present logical abstractions to enable allocating/updating pages in the existence of \textit{sharing}  
@@
 \end{enumerate}
-Based on these perspectives, we coneptualize truth on an address-space through the contingency it exhibits: \textit{assertion P holds on this address-space indexed by its roots address}.
-Therefore, as a first step, we lift standard bi proposition to be indexed with $\mathcal{W}_{64}$ as shown in Figure \ref{fig:vprop}, and, most notably, all abstractions we explain in this section as part of our custom-tailored address-space logic are type $\textsf{vProp }\Sigma$.
-\begin{figure}[!ht]
+Based on these perspectives, we coneptualize truth in an address-space through the contingency it exhibits: 
+\textit{assertion P holds only relative to a certain address space}. 
+Therefore, as a first step, we work in a pointwise lifting of Iris's standard BI proposition,
+essentially working with separation logic assertions indexed by a page table root as a $\mathcal{W}_{64}$
+(Figure \ref{fig:vprop}), which we call $\textsf{vProp }\Sigma$; this is the (\textsc{Coq}) type of assertions in our logic.
+\begin{figure}[t]
 \[
 \begin{array}{cl}
-\textsf{vProp } \Sigma \; : \; \textsf{bi} \; \stackrel{def}{=} \mathcal{W}_{64} \rightarrow_{\textsf{b}} \textsf{iPropI } \Sigma
+\textsf{vProp } \Sigma \; : \; \textsf{bi} \; \stackrel{{\triangle}}{:=} \mathcal{W}_{64} \rightarrow_{\textsf{b}} \textsf{iPropI } \Sigma
 \end{array}
@@
 %\end{itemize}
+
 \subsection{Points-To Assertions}
 \label{sec:pointsto}
+\todo[inline]{Colin says: Current limit of \emph{thorough} revisions}
+\todo[inline,color=red]{Colin says: This section at a glance looks on the right track, but none of the assertions are actually
+written as vProps, they're all still iProps. I can fix that as I go}
+\todo[inline]{Colin says: There's really no need to discuss the 52-12 split in the physical pointsto definition.
+That level of detail of the operational semantics \emph{is} too detailed for the paper (but should be mentioned in the
+artifact docs). I also don't know what the `ent' and `rtv' and similar are in those split physical pointstos}
 We abstract physical memory addressing and registers naming with well-known separation logic assertion \textit{points-to} assertions.
@@
 The fraction $\qfrac$ with value 1 asserts the unique ownership of the register mapping, and grants update permission on it, otherwise, any value $0 < \qfrac <1$ represents partial ownership granting readonly permission on the mapping.
+
+\todo[inline]{We should drop all talk about PDE, PDPT, etc., and just use L1-L4}
 \paragraph{Physical-Memory points-to} Our physical memory points-to relation in Figure \ref{fig:physicalpointsto} for a page table entry (e.g.
 \textsf{PDE} in Figure \ref{fig:pagetables}) abstracts masking -- $|^{52}$ for frame for and $|^{12}$ for offset -- through logical nested maps.
-\begin{figure}[!ht]
+\begin{figure}[t]
 \[
 \begin{array}{cl}
-\pfpointsto\locsf\vpts\qfrac\ppts \stackrel{def}{=} & \nfpointsto{\mask\locsf\ft\generalentry}{\mask\locsf\tw\generalentry}\vpts\qfrac\naddr
+\pfpointsto\locsf\vpts\qfrac\ppts \stackrel{{\triangle}}{:=} & \nfpointsto{\mask\locsf\ft\generalentry}{\mask\locsf\tw\generalentry}\vpts\qfrac\naddr
 \end{array}
@@
       \hbox{(\TirNameStyle{{L4translate}})} \quad
-      \exists \entryf \ldotp \nfpointsto{\mask\vaddr\ft\crthree}{\mask\vaddr\tw\crthree}\entryf\qfracfotsss\naddr
+      \exists \entryf \ldotp \nfpointsto{\mask\vaddr\ft\rtv}{\mask\vaddr\tw\rtv}\entryf\qfracfotsss\naddr
       \end{array}
@@
 \end{itemize}
-\begin{remark}[A Strong Definition for Virtual Memory Addressing]
-  \label{rem:strongvmem}
-  One could give a naive logical definition to the virtual mememory addressing as conjuctions of each level translation (Figure \fref{fig:strongvirtualpointsto}).
-The very first assurance our logical constructions need to be able assert is the \textit{existence} of the page table walk.
-Our naive attempt shown in Figure \ref{figure:strongvirtualpointsto}, we see that a virtual address with a certain fractional permissions at each level of the page-table walk gives us sound assurance of the path existence reaching to a page table entry, consequently, with a full-ownership of page address, we can obtain and update the value mapped value at this physical page address through our virtual memory address.
+\subsection{An Overly-Restrictive Definition for Virtual Memory Addressing}
+One could give a naive logical definition to the virtual mememory addressing as conjuctions of each level translation 
+(Figure \fref{fig:strongvirtualpointsto}). The very first assurance our logical constructions need to be able assert is 
+the \textit{existence} of the page table walk. Our naive attempt shown in Figure \ref{figure:strongvirtualpointsto}, 
+we see that a virtual address with a certain fractional permissions at each level of the page-table walk gives us sound 
+assurance of the path existence reaching to a page table entry, consequently, with a full-ownership of page address, 
+we can obtain and update the value mapped value at this physical page address through our virtual memory address.
+
+\todo[inline]{Cite Kolanski here; not all of them, but some actually defined it exactly this way, for the current
+address space.}
 
@@
 \begin{array}{l}
-  \vaddr\mapsto_{\textsf{t}}\{\textsf{q}\}\vpage\; : \mathsf{vProp} \; \Sigma \;  \stackrel{def}{:=}  \lambda \crthree. \;\; \\
+  \vaddr\mapsto_{\textsf{t}}\{\textsf{q}\}\vpage\; : \mathsf{vProp} \; \Sigma \;  \stackrel{{\triangle}}{:=} 
   \exists_{\entryf \;, \entrytr \;, \entrytw \;,\entryo} \ldotp
@@
   \ulcorner \textsf{aligned } \vaddr \urcorner \ast  \\
- \quad \textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\crthree,\entryf,\entrytr,\entrytw,\entryo) \ast \nfpointsto{\mask\vaddr\ft\entryo}{\mask\vaddr\tw\entryo}\vpage\qfracone\naddr 
+ \quad \textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\rtv,\entryf,\entrytr,\entrytw,\entryo) \ast \nfpointsto{\mask\vaddr\ft\entryo}{\mask\vaddr\tw\entryo}\vpage\qfracone\naddr 
 \end{array} \\
 \\
-\textsf{where   }\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\vaddr,\crthree,\entryf,\entrytr,\entrytw,\entryo) \stackrel{\triangle}{=} \\
- \qquad\qquad \nfpointsto{\mask\vaddr\ft\crthree}{\mask\vaddr\tw\crthree}\entryf\qfracfotsss\naddr \ast \\ 
+\textsf{where   }\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\vaddr,\rtv,\entryf,\entrytr,\entrytw,\entryo) \stackrel{\triangle}{=} \\
+ \qquad\qquad \nfpointsto{\mask\vaddr\ft\rtv}{\mask\vaddr\tw\rtv}\entryf\qfracfotsss\naddr \ast \\ 
  \qquad\qquad  \nfpointsto{\mask\vaddr\ft\entryf}{\mask\vaddr\tw\entryf}\entrytr\qfracfotss\naddr  \ast  \\
@@
 \caption{A Strong Virtual Points-to Relation}
-\todo[inline]{These fractions aren't quite right (though I see you added the variance between levels), I can walk you through in our meeting tomorrow.}
+\todo[inline,color=red]{Colin says: shouldn't that fraction be a q, not a 1, in the second line?}
   \label{fig:strongvirtualpointsto}
@@
   \label{sec:sharingpages}  
-  One might have already observed that the virtual points-to definition shown in Figure \ref{fig:strongvirtualpointsto} is too strong to specify virtual memory operations updating a page table entry.
-To speak more concretely, the fractional permissions on the page-table-walk assertion ($\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$ in Figure \ref{fig:strongvirtualpointsto}) \textit{only} assure the existence of the certain number of mappings -- 512 at each level -- abstracting virtual address translations, and updating any of these mappings would make the other mappings, address translations sharing the page table mappings.
-Therefore, we need to consturct logical abstractions ensuring sound updates to the page-table entries in the existence of sharing.
+  One might have already observed that the virtual points-to definition shown in Figure \ref{fig:strongvirtualpointsto} 
+  is too strong to specify some operations updating a page table entry.
+   To speak more concretely, the fractional permissions on the page-table-walk assertion ($\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$ in Figure \ref{fig:strongvirtualpointsto}) 
+  ensure fractional ownership of the four physical page table entries accessed by the hardware during address translation,
+  as well as the physical memory holding the data of the virtual points-to.
+  However, this formulation stores knowledge of the page table walk's details with ownership of the backing physical
+  memory.
+  Updating any of these mappings (e.g., move the page tables in physical memory, as in coallescing for
+  superpages or hugepages)
+  would require explicitly collecting all virtual points-to facts that traverse affected entries.
+  It is preferable to permit the page tables themselves to be updated independently of the virtual points-to assertions,
+  so long as those updates preserve the same virtual-to-physical translations.
+
 \begin{figure*}
@@
 \begin{array}{l}
-    \vaddr\mapsto_{\textsf{v,rtv}}\{\textsf{q}\}\vpage \stackrel{\triangle}{=} 
+    \vaddr\mapsto_{\textsf{v,rtv}}\{\textsf{q}\}\vpage \stackrel{\triangle}{:=} 
     \exists \entryo \ldotp
@@
 \begin{array}{l}
-  \mathcal{I}\textsf{ASpace} \stackrel{\triangle}{=} \lambda \; \crthree \ldotp
+  \mathcal{I}\textsf{ASpace} \stackrel{\triangle}{:=} 
   \exists\;\gammaPred \;\ldotp\; \ownGhost\gammaPred{\authfull{\ptableabswalk\ptablestore}} \ast \\
-  \bigast{(\vaddr, \entryo)\in \ptablestore}{\exists\;(\entryf\;,\entrytr\;,\entrytw)\ldotp \textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\vaddr,\crthree,\entryf,\entrytr,\entrytw,\entryo)}
+  \bigast{(\vaddr, \entryo)\in \ptablestore}{\exists\;(\entryf\;,\entrytr\;,\entrytw)\ldotp \textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\vaddr,\rtv,\entryf,\entrytr,\entrytw,\entryo)}
 \end{array}
@@
     \hbox{(\TirNameStyle{ModalAddressSpace)}} \qquad
-         [r](P) \; \mathsf{vProp} \; \Sigma \; := \; \lambda \mathsf{P},\; \mathsf{P r}. 
+         [r](P) \;: \mathsf{vProp} \; \Sigma \; \stackrel{\triangle}{:=} \; \lambda \mathsf{P},\; \mathsf{P r}. 
   \end{array}

### 2023-04-10 syntaxsemantics.tex
lots of comments

@@
 \label{sec:syntax}
-To develop our core logical ideas, we instantiate \textsf{Iris} with a simple language for streams of instructions, and a logical machine model corresponding to execution of x86-64 assembly instructions with virtual memory enabled on the CPU.
+To develop our core logical ideas, we instantiate \textsf{Iris} with a simple language for streams of instructions, 
+and a logical machine model corresponding to execution of x86-64 assembly instructions with virtual memory enabled on the 
+CPU.
 
@@
 Programs we demonstrate in this paper requires accessing two types of computer resource: registers and memory.
-A register identifier, $\reg$, is chosen from a fixed finite set of register identifiers,$\regset$.
+ A register identifier, $\reg$, is chosen from a fixed finite set of register identifiers, $\regset$. 
 We use these identifiers to access the register values, $\regval \in \regvaltype$.
-\begin{figure}[!ht]
+\begin{figure}[t]
 \newcommand{\commentary}[1]{ & \text{\small\it #1} \\}
@@
                    \commentary{sequencing}
-      \ising\instr
-                   \commentary{executing}             
+      % \ising\instr
+                   % \commentary{executing}             
     \end{array}
@@
 \caption{Syntax}
+\todo[inline]{Colin says: I removed the 'sexec' thing because I'm pretty sure it's an old remnant of an old Iris instantiation}
 \Description{Syntax}
@@
 $\kw{w}_{12}$ is a 12-bit resource which can be obtained after masking 52-bit of a 64-bit word.
-This simple syntax in \ref{fig:syntax} includes a simple syntax for values because our language is purely \textit{imperative}, i.e.
+This simple syntax in Figure \ref{fig:syntax} includes a simple syntax for values because our language is purely \textit{imperative}, i.e.
 an instruction does not return a value.
@@
 Our syntax capture the no-op ``$\iskip$'' and the sequencing construct ``$\iseq{\instr}{\instrs}$'' in their standard forms.
-\todo[inline]{Ismail: this paper is not about the binding of the machine model to Iris. Low-level details like ``injected into unit'' are not really important unless someone asks.}
+
 \subsection{State}
@@
 But the reduction rules will mostly add mental overhead here.}
+\todo[inline,color=red]{Colin says: On the contrary, we must describe assembly instructions in enough detail for readers
+who aren't already assembly programmers to decode our examples! We don't have to talk about details of the formalization
+of the semantics, but we do need to talk about basic instructions we cover, basic patterns (like destination on the left),
+and just a bit about how all memory accesses implement Figure \ref{fig:pagetables}'s semantics.}
 We represent the machine state mainly as a finite map of registers to register values and a map of masked physical memory addresses to 64-bit physical memory values.

### 2023-04-10 background.tex
+

@@
     \caption{x86-64 page table lookups.}
-        \todo[inline]{Colin says: I recommend the cr3 caption say "Physical Base Address" rather than just
-                base address.
-            If we're saying in the text that we just to L4, L3, L2, L1, then the figure should do the same.
-        }
     \label{fig:pagetables}
@@
 \[ (P\rightarrow Q) \rightarrow M(P) \rightarrow M(Q)\]
-\todo[inline]{Ismail, I think this means we have pure intro, $\square P \mathrel{-\ast} [r](\square P)$ if I'm using the right symbol for pure assertions }
-\todo[inline,color=red]{The above todo comment isn't quite right.
-Purity is about being able to duplicate. You had defined another typeclass/property-of-assertions that meant it didn't
-care what address space it was in (like physical pointstos). That's orthogonal to purity.
-}
+%\todo[inline]{Ismail, I think this means we have pure intro, $\square P \mathrel{-\ast} [r](\square P)$ if I'm using the right symbol for pure assertions }
+%\todo[inline,color=red]{The above todo comment isn't quite right.
+%Purity is about being able to duplicate. You had defined another typeclass/property-of-assertions that meant it didn't
+%care what address space it was in (like physical pointstos). That's orthogonal to purity.
+%}
 Many modalities, so-called \emph{normal} modalities also possess introduction rules of the form $P\rightarrow M(P)$, 
-the classic example being that if $P$ is true, then $P$ is \emph{necessarily} true ($\square P$).
+the classic example being that if $P$ is true, then $P$ is \emph{necessarily} true with the contingency picked up.%($\square P$).
 
@@
 circumstances, prominently \emph{hybrid} modal logics~\cite{blackburn1995hybrid,areces2001hybrid}, which come equipped 
-with assertions of the form $@_\ell(P)$ indicating that $P$ is true in the specific alternate circumstance (Kripke world)
+with assertions of the form $[\ell](P)$ indicating that $P$ is true in the specific alternate circumstance (Kripke world)
  named by the \emph{nominal} $\ell$.
@@
 The typical hybrid modality introduction rule, that $P$ and $\ell$ (indicating the current world is $\ell$) imply 
-$@_\ell(P)$, has a natural analogue: knowing $P$ and that the current address space is $r$ (i.e., that $r$ is the 
-current \texttt{cr3} value) suggests a way to construct $[r](P)$.
+$[\ell](P)$, has a natural analogue: knowing $P$ and that the current address space is $r$ (i.e., that $r$ is the 
+current \texttt{cr3} value) suggests a way to construct $[r](P)$. By the choice of \texttt{cr3} identifying the contingency,
+we can also identify a fact as a \textit{pure} fact as long as it does not \textit{necessarily} depend on \texttt{cr3}: a truth representing a raw physical memory addressing, unlike virtual-address translation, does not need the \textit{knowledge} of \texttt{cr3}, therefore it can be introduce into the modal context as a pure fact.
 

### 2023-04-10 background.tex
+

@@
 $[\ell](P)$, has a natural analogue: knowing $P$ and that the current address space is $r$ (i.e., that $r$ is the 
-current \texttt{cr3} value) suggests a way to construct $[r](P)$. By the choice of \texttt{cr3} identifying the contingency,
-we can also identify a fact as a \textit{pure} fact as long as it does not \textit{necessarily} depend on \texttt{cr3}: a truth representing a raw physical memory addressing, unlike virtual-address translation, does not need the \textit{knowledge} of \texttt{cr3}, therefore it can be introduce into the modal context as a pure fact.
+current \texttt{cr3} value) suggests a way to construct $[r](P)$. By the choice of \texttt{cr3} identifying the contingency, we can identify a fact
+\begin{itemize}
+  \item as a \textit{context} fact if its validity depends on \texttt{cr3} when it is outside the modality (i.e.
+modal context), and made \texttt{cr3} independent as a part of custom-tailored modal logic for virtual-memory: a truth representing a virtual-memory addressing depends \texttt{cr3} due to address-translation operation, is dependent on \texttt{cr3} in the ambient-logic (e.g.
+separation-logic), and is made \textit{independent} of the facts related to \texttt{cr3} by being introduced into the modal context identified with \texttt{cr3} under the assumption that it conveys the knowledge related to the \texttt{cr3} in the ambient logic.
+  \item as a \textit{pure} fact as long as it does not \textit{necessarily} depend on any fact related to \texttt{cr3}: a truth representing a raw physical memory addressing, unlike virtual-address translation, does not need the \textit{knowledge} of \texttt{cr3}, therefore it can be introduce into the modal context as a pure fact
+\end{itemize}
 

### 2023-04-10 background.tex
+

@@
 modal context), and made \texttt{cr3} independent as a part of custom-tailored modal logic for virtual-memory: a truth representing a virtual-memory addressing depends \texttt{cr3} due to address-translation operation, is dependent on \texttt{cr3} in the ambient-logic (e.g.
-separation-logic), and is made \textit{independent} of the facts related to \texttt{cr3} by being introduced into the modal context identified with \texttt{cr3} under the assumption that it conveys the knowledge related to the \texttt{cr3} in the ambient logic.
+separation-logic), and is made \textit{independent} of the facts related to \texttt{cr3} by being introduced into the modal context under the assumption that it exhibits the knowledge on its validity with respect to \texttt{cr3} in the ambient logic.
   \item as a \textit{pure} fact as long as it does not \textit{necessarily} depend on any fact related to \texttt{cr3}: a truth representing a raw physical memory addressing, unlike virtual-address translation, does not need the \textit{knowledge} of \texttt{cr3}, therefore it can be introduce into the modal context as a pure fact

### 2023-04-10 background.tex
+

@@
 A relatively under-explored space of modal logics is the interaction of modal and substructural logics, 
-in particular hybrid-style modalities in substructural logics, which has seen only minimal exploration~\todo{copy refs from fs paper} and no prior 
+in particular hybrid-style modalities in substructural logics, which has seen only minimal exploration~\cite{dovsen1992modal,restall1993modalities,d1997grafting,kamide2002kripke,licata2017fibrational} and no prior
 application. We develop our ideas in the Iris program logic~\cite{jung2018iris}, as a result exploring additional subtleties 
@@
 --- unless guarded by another address space modality.
+%
+%
+%Work on modal logic extends back for decades with many widely-varied variations; we discuss here only the pieces most relevant to our work.
+%Broadly speaking modal logics incorporate \emph{modal operators}, which take as arguments a proposition expected to be true in another time~\cite{pnueli1977temporal}, place~\cite{gordon2019modal}\todo{other cites}, or circumstance~\cite{hintikka1962knowledge,halpern1985guide}, and result in a proposition true in the \emph{current} time, place, or circumstance at which the truth of the use of the modal operator is being evaluated.
+Classic examples include modal necessity $\square P$ describing that $P$ is \emph{necessarily} true, $\mathsf{G}~P$ meaning $P$ is true \emph{globally} (i.e., forever from this time onwards), or $K_i(P)$ describing that a particular participant $i$ \emph{knows} that $P$ is true.
+The latter is an example of of \emph{multimodal} logic, where there is an indexed family of modalities (modal operators) parameterized by some dimension of interest (there, participants).
+%A closely related variant of multimodal logic is \emph{dynamic logic}~\cite{pratt1976semantical}, a logic of weakest preconditions~\cite{dijkstra-75} which works with modalities of the form $[p](P)$, which states that \emph{in the current program state}, \emph{if} program $p$ is run then afterwards $P$ will hold (modulo non-termination).
+%This same idea is used to encode Hoare triples in the Iris program logic~\cite{krebbers2017essence}, using the same encoding as in Pratt's original presentation, where a Hoare triple $\{P\}C\{Q\}$ is encoded as $P\rightarrow[C](Q)$.
+Unlike classic work on dynamic logic~\cite{harel2000dynamic}, Iris applies these ideas in a \emph{substructural} setting (separation logic) where distributivity laws over substructural connectives must be considered.
+For both the dynamic modality and the later modality $\blacktriangleright$ used for guarded recursive predicates, these modalities satisfy axioms of the form $M(P)\ast Q\vdash M(P\ast Q)$, but not the reverse.
+%Iris is not the first combination of modal and substructural logic~\cite{dovsen1992modal,restall1993modalities,d1997grafting,kamide2002kripke,licata2017fibrational}, but is certainly the most heavily tested.
+%A hallmark of a unary logical operator $M$ being a modality is if $M$ satisfies an property akin to $(\upvarphi\rightarrow\psi)\rightarrow M(\upvarphi)\rightarrow\psi$, which roughly states that modus ponens holds under the modality.\footnote{Afficionados of modal logic will note that this property is not quite Axiom K (which requires the initial implication to also hold under $M$), but follows from K and a necessitation rule $\upvarphi\rightarrow M(\upvarphi)$.
+Non-necessitive modalities typically satisfy the weaker property we call out above.}
+%Another pillar of our work is hybrid logic, a branch of modal logic with \emph{names} (called \emph{nominals} for states in Kripke models~\cite{blackburn1995hybrid,goranko1996hierarchies,areces2001hybrid,gargov1993modal}, in contrast to the typical flavor of modality which refers to an \emph{unspecified} (in the assertion) set of alternative circumstances.
+We are aware of only two prior works combining hybrid logics with program logics specifically.
+Brotherston and Villard~\cite{brotherston2014parametric} demonstrated that may properties true of various separation logics are not definable in boolean \BI (\BBI), and showed that a hybrid extension \HyBBI allows most such properties to be defined (e.g., the fact that separating conjunction is cancellative is unprovable in boolean \BI, but provable in \HyBBI).
+There, nominals named resources (roughly, but not exactly, heap fragments).
+Gordon~\cite{gordon2019modal} described a use of hybrid logic in the verification of actor programs, where nominals named the local state of individual actors (with such assertions stabilized with a rely/guarantee approach).
+Beyond these, there is limited work on the interaction of specifically \emph{hybrid logic} with substructural logics.
+Primarily there is a line of work on hybrid linear logic (\HyLL)~\cite{despeyroux2014hybrid}, originally used as a way to more conveniently express aspects of transition systems in linear logic.
+However, \HyLL's proof rules offer no non-trivial interactions with multiplicative connectives (every \HyLL proof can in fact be embedded into regular linear logic~\cite{chaudhuri2019hybrid}, unlike Brotherston and Villard's \HyBBI, which demonstrably increases expressive power over its base \BBI.\colinsays{some of this is too detailed for this point in the paper, needs to go to related work.
+But I'll split it later.}
+%
+%In both \HyLL and \HyBBI, nominals denote worlds with monadic structure (as worlds in Kripke semantics for either LL or \BBI necessarily have monoidal structure).
+Our nominals, by contrast, do not name worlds in the same sense with respect to Iris's CMRAs, but in fact \emph{classes} of worlds, because the names are locations (a means of \emph{selecting} resources) rather than resources.
+A key difference is that the use of nominals in those logics corresponds specifically to hypothetical reasoning about resources (until a nominal is connected to a current resource, in which case conclusions can be drawn about the current resource), which means the modalities themselves do not ``own'' resources.
+Instead, our hybrid modalities directly own portions of the data structure...\colinsays{also need to move to related work}
+%
+%Our work draws on ideas from hybrid modal logics, applied in the context of Iris's higher-order separation logic~\cite{jung2018iris}.
+We define our modalities and prove their rules directly within Iris, so there is no model-theoretic novelty in our work.
+Instead, we show that targeted use of modalities combining the ideas of \emph{named resources} with the power of substructural reasoning enable clear specification of general reasoning principles for data structures.
+Along the way we take advantage of the fact that modalities need not be injections between the \emph{same} logics, but in fact the propositions that modal operators contextualize can come from \emph{other} logics; in our case we exploit the fact that Iris itself is parameterized by a choice of step-indexed~\cite{ahmed-appel-virga-02} \textsf{BI}-algebra~\cite{ohearn1999bunched} to allow convenient specifications in embedded data-structure-relative logics.

### 2023-04-11 syntaxsemantics.tex
+

--- 2023-04-03 syntaxsemantics.tex
@@
+\definecolor{dkgreen}{rgb}{0,0.6,0}
+\definecolor{ltblue}{rgb}{0,0.4,0.4}
+\definecolor{dkviolet}{rgb}{0.3,0,0.5}
+
+% lstlisting coq style (inspired from a file of Assia Mahboubi)
+\lstdefinelanguage{Coq}{ 
+    % Anything betweeen $ becomes LaTeX math mode
+    mathescape=true,
+    % Comments may or not include Latex commands
+    texcl=false, 
+    % Vernacular commands
+    morekeywords=[1]{Section, Module, End, Require, Import, Export,
+        Variable, Variables, Parameter, Parameters, Axiom, Hypothesis,
+        Hypotheses, Notation, Local, Tactic, Reserved, Scope, Open, Close,
+        Bind, Delimit, Definition, Let, Ltac, Fixpoint, CoFixpoint, Add,
+        Morphism, Relation, Implicit, Arguments, Unset, Contextual,
+        Strict, Prenex, Implicits, Inductive, CoInductive, Record,
+        Structure, Canonical, Coercion, Context, Class, Global, Instance,
+        Program, Infix, Theorem, Lemma, Corollary, Proposition, Fact,
+        Remark, Example, Proof, Goal, Save, Qed, Defined, Hint, Resolve,
+        Rewrite, View, Search, Show, Print, Printing, All, Eval, Check,
+        Projections, inside, outside, Def},
+    % Gallina
+    morekeywords=[2]{forall, exists, exists2, fun, fix, cofix, struct,
+        match, with, end, as, in, return, let, if, is, then, else, for, of,
+        nosimpl, when},
+    % Sorts
+    morekeywords=[3]{Type, Prop, Set, true, false, option},
+    % Various tactics, some are std Coq subsumed by ssr, for the manual purpose
+    morekeywords=[4]{pose, set, move, case, elim, apply, clear, hnf,
+        intro, intros, generalize, rename, pattern, after, destruct,
+        induction, using, refine, inversion, injection, rewrite, congr,
+        unlock, compute, ring, field, fourier, replace, fold, unfold,
+        change, cutrewrite, simpl, have, suff, wlog, suffices, without,
+        loss, nat_norm, assert, cut, trivial, revert, bool_congr, nat_congr,
+        symmetry, transitivity, auto, split, left, right, autorewrite},
+    % Terminators
+    morekeywords=[5]{by, done, exact, reflexivity, tauto, romega, omega,
+        assumption, solve, contradiction, discriminate},
+    % Control
+    morekeywords=[6]{do, last, first, try, idtac, repeat},
+    % Comments delimiters, we do turn this off for the manual
+    morecomment=[s]{(*}{*)},
+    % Spaces are not displayed as a special character
+    showstringspaces=false,
+    % String delimiters
+    morestring=[b]",
+    morestring=[d],
+    % Size of tabulations
+    tabsize=3,
+    % Enables ASCII chars 128 to 255
+    extendedchars=false,
+    % Case sensitivity
+    sensitive=true,
+    % Automatic breaking of long lines
+    breaklines=false,
+    % Default style fors listings
+    basicstyle=\small,
+    % Position of captions is bottom
+    captionpos=b,
+    % flexible columns
+    columns=[l]flexible,
+    % Style for (listings') identifiers
+    identifierstyle={\ttfamily\color{black}},
+    % Style for declaration keywords
+    keywordstyle=[1]{\ttfamily\color{dkviolet}},
+    % Style for gallina keywords
+    keywordstyle=[2]{\ttfamily\color{dkgreen}},
+    % Style for sorts keywords
+    keywordstyle=[3]{\ttfamily\color{ltblue}},
+    % Style for tactics keywords
+    keywordstyle=[4]{\ttfamily\color{dkblue}},
+    % Style for terminators keywords
+    keywordstyle=[5]{\ttfamily\color{dkred}},
+    %Style for iterators
+    %keywordstyle=[6]{\ttfamily\color{dkpink}},
+    % Style for strings
+    stringstyle=\ttfamily,
+    % Style for comments
+    commentstyle={\ttfamily\color{dkgreen}},
+    %moredelim=**[is][\ttfamily\color{red}]{/&}{&/},
+    literate=
+    {\\forall}{{\color{dkgreen}{$\forall\;$}}}1
+    {\\exists}{{$\exists\;$}}1
+    {<-}{{$\leftarrow\;$}}1
+    {=>}{{$\Rightarrow\;$}}1
+    {==}{{\code{==}\;}}1
+    {==>}{{\code{==>}\;}}1
+    %    {:>}{{\code{:>}\;}}1
+    {->}{{$\rightarrow\;$}}1
+    {<->}{{$\leftrightarrow\;$}}1
+    {<==}{{$\leq\;$}}1
+    {\#}{{$^\star$}}1 
+    {\\o}{{$\circ\;$}}1 
+    {\@}{{$\cdot$}}1 
+    {\/\\}{{$\wedge\;$}}1
+    {\\\/}{{$\vee\;$}}1
+    {++}{{\code{++}}}1
+    {~}{{\ }}1
+    {\@\@}{{$@$}}1
+    {\\mapsto}{{$\mapsto\;$}}1
+    {\\hline}{{\rule{\linewidth}{0.5pt}}}1
+    %
+}[keywords,comments,strings]
+
 \section{Machine State \& Syntax}
 \label{sec:syntax}
-To develop our core logical ideas, we instantiate \textsf{Iris} with a simple language for streams of instructions, and a logical machine model corresponding to execution of x86-64 assembly instructions with virtual memory enabled on the CPU.
+To develop our core logical ideas, we instantiate \textsf{Iris} with a simple language for streams of instructions, 
+and a logical machine model corresponding to execution of x86-64 assembly instructions with virtual memory enabled on the 
+CPU.
 
@@
 Our syntax capture the no-op ``$\iskip$'' and the sequencing construct ``$\iseq{\instr}{\instrs}$'' in their standard forms.
-\todo[inline]{Ismail: this paper is not about the binding of the machine model to Iris. Low-level details like ``injected into unit'' are not really important unless someone asks.}
 \subsection{State}
 \label{sec:state}
-\todo[inline]{I don't think we need or want to describe the operational semantics in detail in the submission.
+%\todo[inline]{I don't think we need or want to describe the operational semantics in detail in the submission.
 It adds a lot of formalization that doen'st necessarily contribute to reader understanding.
@@
 But the reduction rules will mostly add mental overhead here.}
+%\todo[inline,color=red]{Colin says: On the contrary, we must describe assembly instructions in enough detail for readers
+%who aren't already assembly programmers to decode our examples! We don't have to talk about details of the formalization
+%of the semantics, but we do need to talk about basic instructions we cover, basic patterns (like destination on the left),
+%and just a bit about how all memory accesses implement Figure \ref{fig:pagetables}'s semantics.}
 We represent the machine state mainly as a finite map of registers to register values and a map of masked physical memory addresses to 64-bit physical memory values.
-Together with default def\textsf{CPU} instantiation, we end up having our state with following pieces:
+Together with default \textsf{CPU} instantiation, we end up having our state with following pieces:
 \begin{itemize}
@@
 $\kw{r}\in\kw{greg}$) -- to the global maps.
+\subsection{Instructions}
+\label{sec:instructions}
+Page-table walk shown in Figure \ref{fig:pagetables} is realized with physical memory \texttt{load}s and \textsf{store}s for the entries in the page-tables which are simply modelled as lookup and updates to an instance of a memory map with 64 bit entries.
+In the address translation,
+\begin{lstlisting}[language=Coq]
+ Definition translate (m: $\sigma.\mathcal{M}$) (root: word 64) (w: word 64): address :=
+      let plm4e := translate_top_level m root  w in
+      let pdpte := translate_from_l4e m pml4e w in 
+      let pde := translate_from_l3e m pdpte w in
+      let pte: = translate_from_l2e m pde w in
+      let addr := translate_from_l1e m pte w in
+      concat (shift_dropn addr 3 three_lt64) low3.
+    end.
+\end{lstlisting}
+we see translation physical memory $\mathsf{load}$ for each level of page-table-walk, which is basically a table lookup with
+\begin{lstlisting}[language=Coq]
+   Definition load (mem: $\sigma.\mathcal{M}$) (w: { w: word 64 | aligned w }):word 64 + MemFail :=
+    match  $\sigma$.mem !! (pfn (proj1_sig w)) with
+      | None => inl (zero 64)
+      | Some (inl page) => inl (page (page_off (proj1_sig w)))
+      | Some (inr fault) => inr fault
+      end.
+
+  Definition store (mem: $\sigma.\mathcal{M}$) (w: { w: word 64 | aligned w }) (v: word 64):mem64 * (unit + MemFail) :=
+    match $\sigma$.mem !! (pfn (proj1_sig w)) mem with
+      | None =>
+        (($\sigma.\mathcal{M}$<[(pfn (proj1_sig w)) := 
+             (inl (fun off =>
+                  if (word_dec off (page_off (proj1_sig w))) then v else zero 64))]> mem
+         ), inl tt)
+      | Some (inl page) => (* update page contents *)
+        (($\sigma.\mathcal{M}$<[(pfn (proj1_sig w)) := 
+             (inl (fun off =>
+                  if (word_dec off (page_off (proj1_sig w))) then v else zero 64))]> mem
+         ), inl tt)
+      | Some (inr fault) => (mem, inr fault)
+    end.
+
+  Program Definition pfn (w : word 64) := shift_dropn w 12 _.
+  Program Definition page_off (w : word 64) := bottomn w 12 _.
+\end{lstlisting}
+with frame and offset masking applied on the physical address being loaded. As expected \textsf{store} would be the update on the memory map entry
+\todo[inline,color=yellow]{Colin, could you mention that we simplified the maps here to get rid of MemFail?}
+After defining the translation in terms of physical load and store action, now we can define our \lstinline|mov| instructions.
+STOP HERE
+
 %

### 2023-04-11 syntaxsemantics.tex
+

@@
       end.
-
   Definition store (mem: $\sigma.\mathcal{M}$) (w: { w: word 64 | aligned w }) (v: word 64):mem64 * (unit + MemFail) :=
@@
       | None =>
-        (($\sigma.\mathcal{M}$<[(pfn (proj1_sig w)) := 
-             (inl (fun off =>
-                  if (word_dec off (page_off (proj1_sig w))) then v else zero 64))]> mem
-         ), inl tt)
+        (($\sigma.\mathcal{M}$<[(pfn (proj1_sig w)) := (inl (fun off =>
+           if (word_dec off (page_off (proj1_sig w))) then v else zero 64))]> mem), inl tt)
       | Some (inl page) => (* update page contents *)
-        (($\sigma.\mathcal{M}$<[(pfn (proj1_sig w)) := 
-             (inl (fun off =>
-                  if (word_dec off (page_off (proj1_sig w))) then v else zero 64))]> mem
-         ), inl tt)
+        (($\sigma.\mathcal{M}$<[(pfn (proj1_sig w)) := (inl (fun off =>
+           if (word_dec off (page_off (proj1_sig w))) then v else zero 64))]> mem), inl tt)
       | Some (inr fault) => (mem, inr fault)
     end.
-
   Program Definition pfn (w : word 64) := shift_dropn w 12 _.
@@
 After defining the translation in terms of physical load and store action, now we can define our \lstinline|mov| instructions.
-STOP HERE
 

### 2023-04-11 syntaxsemantics.tex
+

@@
 \todo[inline,color=yellow]{Colin, could you mention that we simplified the maps here to get rid of MemFail?}
-After defining the translation in terms of physical load and store action, now we can define our \lstinline|mov| instructions.
+After defining the translation in terms of physical load and store action, now we can define our memory \lstinline|mov| instructions
+\[
+\begin{array}{l}
+  \hbox{(\TirNameStyle{WriteToMemFromReg})} \qquad \qquad\qquad\qquad 
+  \textsf{mov mem64 srcreg64} \\
+  \hbox{(\TirNameStyle{WriteToRegFromMem})} \qquad \qquad\qquad\qquad 
+  \textsf{mov reg64 mem64}
+  \end{array}
+  \]
+  which are realized by first translating the virtual-memory address to store a value from a register (\TirNameStyle{WriteToMemFromReg}) or load a value at the translated address to be stored in a register (\TirNameStyle{WriteToRegFromMem}).
+More \lstinline|mov| instructions operating with memory addresses can be found in the submission artifact for different addressing schemes and data sizes.
 
+  Register instructions, on the other hand, are implemented as $\sigma.\mathcal{R}$ lookup and store
+\[
+\begin{array}{l}
+  \hbox{(\TirNameStyle{WriteToRegFromReg})} \qquad \qquad\qquad\qquad 
+  \textsf{mov ireg64 ireg64} \\
+  \hbox{(\TirNameStyle{WriteToCtlRegFromReg})} \qquad \qquad\qquad\qquad 
+  \textsf{mov creg64 ireg64}
+    \hbox{(\TirNameStyle{WriteToRegFromCtlReg})} \qquad \qquad\qquad\qquad 
+  \textsf{mov ireg64 creg64}
+  \end{array}
+\]
+for different types of registers (e.g. control and general).
 %

### 2023-04-11 soundness.tex
+

--- 2023-04-05 soundness.tex
@@
+\definecolor{dkgreen}{rgb}{0,0.6,0}
+\definecolor{ltblue}{rgb}{0,0.4,0.4}
+\definecolor{dkviolet}{rgb}{0.3,0,0.5}
+
+% lstlisting coq style (inspired from a file of Assia Mahboubi)
+\lstdefinelanguage{Coq}{ 
+    % Anything betweeen $ becomes LaTeX math mode
+    mathescape=true,
+    % Comments may or not include Latex commands
+    texcl=false, 
+    % Vernacular commands
+    morekeywords=[1]{Section, Module, End, Require, Import, Export,
+        Variable, Variables, Parameter, Parameters, Axiom, Hypothesis,
+        Hypotheses, Notation, Local, Tactic, Reserved, Scope, Open, Close,
+        Bind, Delimit, Definition, Let, Ltac, Fixpoint, CoFixpoint, Add,
+        Morphism, Relation, Implicit, Arguments, Unset, Contextual,
+        Strict, Prenex, Implicits, Inductive, CoInductive, Record,
+        Structure, Canonical, Coercion, Context, Class, Global, Instance,
+        Program, Infix, Theorem, Lemma, Corollary, Proposition, Fact,
+        Remark, Example, Proof, Goal, Save, Qed, Defined, Hint, Resolve,
+        Rewrite, View, Search, Show, Print, Printing, All, Eval, Check,
+        Projections, inside, outside, Def},
+    % Gallina
+    morekeywords=[2]{forall, exists, exists2, fun, fix, cofix, struct,
+        match, with, end, as, in, return, let, if, is, then, else, for, of,
+        nosimpl, when},
+    % Sorts
+    morekeywords=[3]{Type, Prop, Set, true, false, option},
+    % Various tactics, some are std Coq subsumed by ssr, for the manual purpose
+    morekeywords=[4]{pose, set, move, case, elim, apply, clear, hnf,
+        intro, intros, generalize, rename, pattern, after, destruct,
+        induction, using, refine, inversion, injection, rewrite, congr,
+        unlock, compute, ring, field, fourier, replace, fold, unfold,
+        change, cutrewrite, simpl, have, suff, wlog, suffices, without,
+        loss, nat_norm, assert, cut, trivial, revert, bool_congr, nat_congr,
+        symmetry, transitivity, auto, split, left, right, autorewrite},
+    % Terminators
+    morekeywords=[5]{by, done, exact, reflexivity, tauto, romega, omega,
+        assumption, solve, contradiction, discriminate},
+    % Control
+    morekeywords=[6]{do, last, first, try, idtac, repeat},
+    % Comments delimiters, we do turn this off for the manual
+    morecomment=[s]{(*}{*)},
+    % Spaces are not displayed as a special character
+    showstringspaces=false,
+    % String delimiters
+    morestring=[b]",
+    morestring=[d],
+    % Size of tabulations
+    tabsize=3,
+    % Enables ASCII chars 128 to 255
+    extendedchars=false,
+    % Case sensitivity
+    sensitive=true,
+    % Automatic breaking of long lines
+    breaklines=false,
+    % Default style fors listings
+    basicstyle=\small,
+    % Position of captions is bottom
+    captionpos=b,
+    % flexible columns
+    columns=[l]flexible,
+    % Style for (listings') identifiers
+    identifierstyle={\ttfamily\color{black}},
+    % Style for declaration keywords
+    keywordstyle=[1]{\ttfamily\color{dkviolet}},
+    % Style for gallina keywords
+    keywordstyle=[2]{\ttfamily\color{dkgreen}},
+    % Style for sorts keywords
+    keywordstyle=[3]{\ttfamily\color{ltblue}},
+    % Style for tactics keywords
+    keywordstyle=[4]{\ttfamily\color{dkblue}},
+    % Style for terminators keywords
+    keywordstyle=[5]{\ttfamily\color{dkred}},
+    %Style for iterators
+    %keywordstyle=[6]{\ttfamily\color{dkpink}},
+    % Style for strings
+    stringstyle=\ttfamily,
+    % Style for comments
+    commentstyle={\ttfamily\color{dkgreen}},
+    %moredelim=**[is][\ttfamily\color{red}]{/&}{&/},
+    literate=
+    {\\forall}{{\color{dkgreen}{$\forall\;$}}}1
+    {\\exists}{{$\exists\;$}}1
+    {<-}{{$\leftarrow\;$}}1
+    {=>}{{$\Rightarrow\;$}}1
+    {==}{{\code{==}\;}}1
+    {==>}{{\code{==>}\;}}1
+    %    {:>}{{\code{:>}\;}}1
+    {->}{{$\rightarrow\;$}}1
+    {<->}{{$\leftrightarrow\;$}}1
+    {<==}{{$\leq\;$}}1
+    {\#}{{$^\star$}}1 
+    {\\o}{{$\circ\;$}}1 
+    {\@}{{$\cdot$}}1 
+    {\/\\}{{$\wedge\;$}}1
+    {\\\/}{{$\vee\;$}}1
+    {++}{{\code{++}}}1
+    {~}{{\ }}1
+    {\@\@}{{$@$}}1
+    {\\mapsto}{{$\mapsto\;$}}1
+    {\\hline}{{\rule{\linewidth}{0.5pt}}}1
+    %
+}[keywords,comments,strings]
+
 \section{Implementing Logical Machinery \& Soundness}
-We build our program logic, as an instantiation of Iris~\cite{iris}, and build our modal abstractions on top of it. 
+We build our program logic, as an instantiation of Iris~\cite{iris}, and construct our reasoning principles including the modal one on top of it. 
 \mytodo{we should put some stuff here}
@@
 Our logic, operates on the machine state, which means we do not need to augment the machine state.
-The invarian that we pick, \textit{central invariant} ($\mathcal{I}$), is just semantic interpretation of stores in the machine state, $\sigma$.R and $\sigma$.M.
+The invarian that we pick, \textit{central invariant} ($\textsf{x64\_h}$), is just semantic interpretation of stores in the machine state, $\sigma.\mathcal{R}$ and $\sigma.\mathcal{M}$.
 This semantic interpratation ensures the correct lifting of mappings in the machine state to the assertions that the client of our logic uses, i.e.
@@
 We prefer to skip explaining the steps used in instantiation of Iris because it is an almost standard procedure, has already been explained for many other logic \ref{}, and we are concerned with the page-count limitation.
-However, it is worh noting that once you instantiate Iris for your language, it comes with the semantic definition for the weakest-precondition which we can refactor into Hoare triples, $\triple\pre\instr\post$, to specify our  selected \textsf{AMD64} instructions shown in Figure \ref{fig:wpdamd}, and show that these triples are sound.
+However, it is worh noting that once you instantiate Iris for your language, it comes with the semantic definition for the weakest-precondition which we can refactor into Hoare-Doubles to specify our  selected \textsf{AMD64} instructions shown in Figure \ref{fig:wpdamd}, and show that these triples are sound.
 \subsection{The Soundness Statement}
@@
 \newcommand\fpfn{\rightarrow_{\textrm{fin}}}
-\( \textsf{gen\_memUR} \stackrel{def}{=}
+\( \textsf{gen\_memUR} \stackrel{\triangle}{=}
   \authm(\;
@@
   \)
-  for our nested memory mapping abstracting two different machine word masking. Then, we register our \emph{authoritative camera} \cite[\S6.3.3]{iris} from this algebra
-  \[\textsf{inG } \Sigma \textsf{ gen\_memUR} \]
-  To interpret this nested ghost map we define \textsf{to\_gen\_mem}
-  \[
-  \begin{array}{l}
-    \mathsf{to\_gen\_mem} : \textsf{ gmap L1 }(\textsf{gmap L2 V})\rightarrow \textsf{gen\_memUR L1 L2 V } := \textsf{ fmap} (\lambda \textsf{m . to\_gen\_heap m}). \\
-     \mathsf{to\_gen\_heap } : \textsf{gmap L V} \rightarrow \textsf{gen\_heapUR L V} :=  (\lambda \textsf{v }\ldotp (1, \textsf{ to\_agree } (\textsf{v} : \textsf{leibnizO V}))).
-  \end{array}
-  \]
+  for our nested memory mapping abstracting two different machine word masking.
+To interpret this nested ghost map, i.e.
+obtain points-to assertions out of mappings in an ordinary \textsf{gmap}, we define \textsf{to\_gen\_mem}
+  \begin{lstlisting}[language=Coq]
+    Definition to_gen_mem : gmap L1 ( gmap L2 V) $\rightarrow$ gen_memUR L1 L2 V := fmap ($\lambda$ m . to_gen_heap m).
+    Definition to_gen_heap : gmap L V $\rightarrow$ gen_heapUR L V :=  $\lambda$  v $\ldotp$ (1, to_agree (v :leibnizO V)).
+  \end{lstlisting}
  throuh using \textsf{to\_gen\_heap} from previous Iris version. \todo[inline,color=red]{Ismail give exact version commit etc.}
@@
 \begin{array}{l@{\quad\star\quad}l@{\quad}l}
-  \textsf{to\_gen\_heap} \;\store.\mathcal{R} &  \\
-   \textsf{to\_gen\_mem} \; \store.\mathcal{M}
+  \textsf{to\_gen\_heap} \;\store.\mathcal{R} & \textsf{to\_gen\_mem} \; \store.\mathcal{M}
 \end{array}
@@
 This relation allows lifting the program expression to enables the application of changes imposed by the operational semantics on the program state $\sigma$ when applied for an insturction (\textsf{i}).
-\[
-\begin{array}{l}
-\textsf{exec\_step} (\textsf{i}: \textsf{ instr}) (\sigma:\textsf{ state}):\textsf{ option state} \stackrel{def}{=} \\
-\qquad \textsf{exec\_state} (\textsf{exec\_instr i } \sigma.(\textsf{cpu})  \\ \qquad (\textsf{regToregset } \sigma.(\mathcal{R}))  \\ \qquad ((\textsf{memToPhysMem }\sigma.(\mathcal{M})) (\textsf{exist \_}  \\ \qquad (\textsf{ regset\_get\_num }(\textsf{regToregset }\sigma.(\mathcal{R}))  (\textsf{ cr cr3}))))).
-\end{array}
-\]
+\begin{lstlisting}[language=Coq]
+Definition exec_step (i: instr) ($\sigma$:state) : option state :=
+    exec_state (exec_instr i $\sigma\mathcal{C}$ $\sigma.\mathcal{R}$ ((memToPhysMem $\sigma.\mathcal{M}$) (exist _ ($\sigma.\mathcal{R}$ !!  cr3)))).
+
+\end{lstlisting}
 \todo[inline,color=yellow]{Colin, in case needed,not proven, could you simply say that we assume these map equalities or a paper proof etc.}.
@@
    \item Step 3: for each physical pointsto inside $\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$, we obtain
-\[
-       \ulcorner \sigma.\mathcal{M } !! (\textsf{ va}|^{52}(\textsf{ent}_i) = \textsf{Some } \sigma\textsf{d}_i \land
-       \sigma\textsf{d }_i !! (\textsf{ va}|^{12}(\textsf{ent}_i)  = \textsf{Some entry}_{i+1} \urcorner
-       \]
-       where $\sigma.\textsf{d}_i$ in type \textsf{gmap (word 12) (word 64)} for our offset mappings. 
    \item Step 4: use these concrete lookups to traverse the page tables to obtain the value \textsf{v}

### 2023-04-11 soundness.tex
+

@@
  throuh using \textsf{to\_gen\_heap} from previous Iris version. \todo[inline,color=red]{Ismail give exact version commit etc.}
-As a final step, we allocate an empty ghost cell $\gamma$ at the beggining of a program execution to store the element of the monoid $\textsf{inG } \Sigma \textsf{ gen\_memUR}$.
 \end{definition}
@@
 \textsf{x64\_h}\;\store \triangleq
-\left\{
 \def\arraystretch{1.2}
@@
 \end{array}
-\right.
 \]
@@
 \todo[inline,color=yellow]{Colin, in case needed,not proven, could you simply say that we assume these map equalities or a paper proof etc.}.
-Since the focus of this paper is not explaining our \textsf{AMD64} model, we prefer not to get into the details due to space limits.
-However, Coq formalization of our operational semantics is a part of our submission artifact.
+Due to the page limitation, we can only give the outline of a selected instruction (\TirNameStyle{WriteToRegFromVirtMem}) from our \textsf{AMD64} model.
+However, mechanized soundness proofs of other instructions can be found as a part of our submission artifact.
 
-Now, we have well-enough definition for giving an outline for the proof of \TirNameStyle{WriteToRegFromVirtMem}.
+After obtaining ownership (points-to) predicates from application of state interpretation, now, we have well-enough definition for giving an outline for the proof of \TirNameStyle{WriteToRegFromVirtMem}.
  \begin{lemma}[\textsc{\TirNameStyle{WriteToRegFromVirtMem}}]
@@
    \item Step 2: we unfold the virtual-pointsto ($\vaddr \mapsto_{\textsf{v,rtv}} \textsf{v}$) definition, and for an existential physical page address $\paddr$, we exchange our fragmental toke ($\sumwalkabs\vaddr\qfrac\paddr$) to obtain physical table-pointsto ($\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$ in Figure \ref{fig:strongvirtualpointsto}) relation
-   \item Step 3: for each physical pointsto inside $\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$, we obtain
+   \item Step 3: for each physical pointsto inside $\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$, there exists a \textsf{load}, i.e. a concrete memory lookup
    \item Step 4: use these concrete lookups to traverse the page tables to obtain the value \textsf{v}

### 2023-04-11 examples.tex
+

@@
+\newcommand{\specline}[1]{{\color{blue}\left\{#1\right\}}}
 %$\assert{\ulcorner \mathsf{aligned maddr} \urcorner \ast \mathsf{r14} \mapsto_{\textsf{r}} \textsf{r14v} \ast \mathsf{r13} \mapsto_{\textsf{r}} \textsf{maddr} \ast \mathsf{rdi} \mapsto_{\textsf{r}} \textsf{rdiv} \ast \mathsf{rax} \mapsto_{\textsf{r}} \textsf{raxv}}$
@@
 \label{sec:experiment}
-\mytodo{Mechanization is Ismail's Part -- Discussing and putting in here is Colin's}
-
 To both validate and demonstrate the value of the modal approach to reasoning about virtual memory management, we study several distillations of real concerns of virtual memory managers.
@@
 \newcommand{\fpaddr}{\texttt{fpaddr}}
-\begin{figure}
+\begin{figure}\footnotesize
   \begin{lstlisting}
@@
 
-\begin{figure}
+\begin{figure}\footnotesize
 \begin{lstlisting}
@@
 mov %cr3, 56[%rsi]
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,\gammaPred) \ast [\rtv'+56](\mathcal{I}\texttt{ASpace}(\theta,\gammaPred') \ast \texttt{Pother})}_{\rtv}$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,\gammaPred) \ast [\rtv'+56](\mathcal{I}\texttt{ASpace}(\theta,\gammaPred') \ast \texttt{Pother}) }_{\rtv}$
 $\specline{ \texttt{rdi}\mapsto_{r} \_ \ast \texttt{rbx}\mapsto_{r} \rtv'+0 \ast  \texttt{rsp}\mapsto_{r} \rtv'+8 \ast \texttt{rbp}\mapsto_{r} \rtv'+16 }_{\rtv}$

### 2023-04-11 examples.tex
+

@@
-\newcommand{\specline}[1]{{\color{blue}\left\{#1\right\}}}
 %$\assert{\ulcorner \mathsf{aligned maddr} \urcorner \ast \mathsf{r14} \mapsto_{\textsf{r}} \textsf{r14v} \ast \mathsf{r13} \mapsto_{\textsf{r}} \textsf{maddr} \ast \mathsf{rdi} \mapsto_{\textsf{r}} \textsf{rdiv} \ast \mathsf{rax} \mapsto_{\textsf{r}} \textsf{raxv}}$
@@
 
-\begin{comment}
-\todo[inline]{Identity mappings are difficult, and our current approach won't quite work.
+%\begin{comment}
+%\todo[inline]{Identity mappings are difficult, and our current approach won't quite work.
 Consider trying to have a virtual pointsto for an actual page table entry (i.e., that one could use to update a page table mapping), while also having a virtual pointsto for an address that entry mapped.
@@
 That's because the PTE pointsto will assert full ownership of the physical memory cell holding the PTE as its data value, while the virtual pointsto for the data mapped by that entry will \emph{also} assert (fractional) ownership of all entries a page table walk would traverse.
-}
-\todo[inline,color=violet]{This doesn't seem to cause issues with the mapping/unmapping examples, only with changing intermediate page table pointers.
+%}
+%\todo[inline,color=violet]{This doesn't seem to cause issues with the mapping/unmapping examples, only with changing intermediate page table pointers.
 The mapping example requires a virtual pointsto for the blank PTE, and once filled in that ownership can be immediately split to create the 512 new virtual pointsto assertions for the newly mapped page.
@@
 But changing intermediate page tables, as one would do for coallescing or splitting a superpage while preserving the virtual-to-physical mappings, couldn't be done without some really complicated separating implication tricks.}
-\todo[inline,color=green]{One possible approach to resolving this, which we came up with in our Tuesday meeting, is to recognize that the current (v1) virtual points-to is too strong, because it really doesn't care about \emph{owning} those fractional resources, it only cares that \emph{something} ensures the correct page table walk exists.
+%\todo[inline,color=green]{One possible approach to resolving this, which we came up with in our Tuesday meeting, is to recognize that the current (v1) virtual points-to is too strong, because it really doesn't care about \emph{owning} those fractional resources, it only cares that \emph{something} ensures the correct page table walk exists.
 Iris has a ghost map resource where authoritative ownership of an individual key-value pair can be handled as a resource.
 (Colin was using this in the filesystem cache.)
-We can use that mechanism to separate the virtual-to-physical translation from the physical memory involved (Kolanski and Klein may have done something similar for different reasons): (fractional) virtual points-to assertions can be defined in terms of (fractional) ownership of these authoritative ghost map entry assertions, plus sharing an invariant that the current installed page table respects all entries of the mapping.
+%We can use that mechanism to separate the virtual-to-physical translation from the physical memory involved (Kolanski and Klein may have done something similar for different reasons): (fractional) virtual points-to assertions can be defined in terms of (fractional) ownership of these authoritative ghost map entry assertions, plus sharing an invariant that the current installed page table respects all entries of the mapping.
 Unmapping collects the authoritative map kvpairs from collecting the assertions, and then can remove them from the ghost map and update the page tables.
 Critically, physical ownership of the page tables then lives in the invariant on the current page table, so some virtual pointsto assertions can refer to memory in those page tables.
-This still works with the modality, since that invariant is also semantically a predicate on a page table root.
-Let's call this v2.
-}
-\todo[inline]{the current course of action is to finish the WPs for v1 and try an example that should work with either approach (like the task switch).
-This is a good basic sanity check on our approaches.
-Once that's done, then we work out the details of v2, which should require \emph{refactoring} the same sorts of reasoning we had for v1, rather than totally different proofs from v1.}
-\end{comment}
+%This still works with the modality, since that invariant is also semantically a predicate on a page table root.
+%Let's call this v2.
+%}
+%\end{comment}
 \subsection{Mapping a New Page}
@@
 \newcommand{\fpaddr}{\texttt{fpaddr}}
+\newcommand{\specline}[1]{{\color{blue}\left\{#1\right\}}}
+
 \begin{figure}\footnotesize
   \begin{lstlisting}
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,\gammaPred)  \ast  \texttt{r14}\mapsto_{r} \_ \ast \texttt{rdi}\mapsto_{r} \vaddr \ast \texttt{rax}\mapsto_{r} \_ \ast \ulcorner \texttt{aligned va} \land  \theta \; !!\; \texttt{maddr} = \texttt{None}\urcorner}_{\rtv}$
+\{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,\gammaPred)  \ast  \texttt{r14}\mapsto_{r} \_ \ast \texttt{rdi}\mapsto_{r} \vaddr \ast \texttt{rax}\mapsto_{r} \_ \ast \ulcorner \texttt{aligned va} \land  \theta \; !!\; \texttt{maddr} = \texttt{None}\urcorner\}_{\rtv}
 call ensure_L1_page
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,\gammaPred) \ast  \texttt{r14}\mapsto_{r} \_ \ast \texttt{rdi}\mapsto_{r} \vaddr \ast \texttt{rax}\mapsto_{r} \_ \ast }_{\rtv}$
+$\{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,\gammaPred) \ast  \texttt{r14}\mapsto_{r} \_ \ast \texttt{rdi}\mapsto_{r} \vaddr \ast \texttt{rax}\mapsto_{r} \_ \ast \}_{\rtv}$
 $\specline{\exists (\entryf ,\;\entrytr,\; \entrytw,\; \entryo: \texttt{word 64}) \ldotp \ulcorner (\texttt{addr\_L1 }\vaddr) = \entryo\urcorner \ast }_{\rtv}$

### 2023-04-11 syntaxsemantics.tex
section 3
+

@@
 \begin{lstlisting}[language=Coq]
- Definition translate (m: $\sigma.\mathcal{M}$) (root: word 64) (w: word 64): address :=
-      let plm4e := translate_top_level m root  w in
-      let pdpte := translate_from_l4e m pml4e w in 
-      let pde := translate_from_l3e m pdpte w in
-      let pte: = translate_from_l2e m pde w in
-      let addr := translate_from_l1e m pte w in
+ Definition translate (root: word 64) (w: word 64): address :=
+      let plm4e := translate_top_level $\sigma.\mathcal{M}$ root  w in
+      let pdpte := translate_from_l4e $\sigma.\mathcal{M}$ pml4e w in 
+      let pde := translate_from_l3e $\sigma.\mathcal{M}$ pdpte w in
+      let pte: = translate_from_l2e $\sigma.\mathcal{M}$ pde w in
+      let addr := translate_from_l1e $\sigma.\mathcal{M}$ pte w in
       concat (shift_dropn addr 3 three_lt64) low3.
@@
 \end{lstlisting}
-we see translation physical memory $\mathsf{load}$ for each level of page-table-walk, which is basically a table lookup with
+we see physical memory $\mathsf{load}$ for each level of page-table-walk concluded with alignment of the returned address.
 \begin{lstlisting}[language=Coq]
-   Definition load (mem: $\sigma.\mathcal{M}$) (w: { w: word 64 | aligned w }):word 64 + MemFail :=
-    match  $\sigma$.mem !! (pfn (proj1_sig w)) with
+   Definition load (w: { w: word 64 | aligned w }):word 64 + MemFail :=
+    match  $\sigma.\mathcal{M}$ !! (pfn (proj1_sig w)) with
       | None => inl (zero 64)
@@
       end.
-  Definition store (mem: $\sigma.\mathcal{M}$) (w: { w: word 64 | aligned w }) (v: word 64):mem64 * (unit + MemFail) :=
-    match $\sigma$.mem !! (pfn (proj1_sig w)) mem with
+  Definition store (w: { w: word 64 | aligned w }) (v: word 64):mem64 * (unit + MemFail) :=
+    match $\sigma.\mathcal{M}$ !! (pfn (proj1_sig w)) with
       | None =>
-        (($\sigma.\mathcal{M}$<[(pfn (proj1_sig w)) := (inl (fun off =>
-           if (word_dec off (page_off (proj1_sig w))) then v else zero 64))]> mem), inl tt)
+        ((<[(pfn (proj1_sig w)) := (inl (fun off => if (word_dec off (page_off (proj1_sig w)))
+                                        then v else zero 64))]> $\sigma.\mathcal{M}$), inl tt)
       | Some (inl page) => (* update page contents *)
-        (($\sigma.\mathcal{M}$<[(pfn (proj1_sig w)) := (inl (fun off =>
-           if (word_dec off (page_off (proj1_sig w))) then v else zero 64))]> mem), inl tt)
-      | Some (inr fault) => (mem, inr fault)
+        ((<[(pfn (proj1_sig w)) := (inl (fun off => if (word_dec off (page_off (proj1_sig w)))
+                                        then v else zero 64))]> $\sigma.\mathcal{M}$), inl tt)
+      | Some (inr fault) => ($\sigma.\mathcal{M}$, inr fault)
     end.
@@
 \begin{array}{l}
-  \hbox{(\TirNameStyle{WriteToMemFromReg})} \qquad \qquad\qquad\qquad 
+  \hbox{(\TirNameStyle{WriteToMemFromReg})} \qquad
   \textsf{mov mem64 srcreg64} \\
-  \hbox{(\TirNameStyle{WriteToRegFromMem})} \qquad \qquad\qquad\qquad 
+  \hbox{(\TirNameStyle{WriteToRegFromMem})} \qquad 
   \textsf{mov reg64 mem64}
@@
 \begin{array}{l}
-  \hbox{(\TirNameStyle{WriteToRegFromReg})} \qquad \qquad\qquad\qquad 
+  \hbox{(\TirNameStyle{WriteToRegFromReg})} \qquad 
   \textsf{mov ireg64 ireg64} \\
-  \hbox{(\TirNameStyle{WriteToCtlRegFromReg})} \qquad \qquad\qquad\qquad 
-  \textsf{mov creg64 ireg64}
-    \hbox{(\TirNameStyle{WriteToRegFromCtlReg})} \qquad \qquad\qquad\qquad 
+  \hbox{(\TirNameStyle{WriteToCtlRegFromReg})} \qquad 
+  \textsf{mov creg64 ireg64}\\
+    \hbox{(\TirNameStyle{WriteToRegFromCtlReg})} \qquad 
   \textsf{mov ireg64 creg64}

### 2023-04-11 syntaxsemantics.tex
+

@@
 \begin{lstlisting}[language=Coq]
-   Definition load (w: { w: word 64 | aligned w }):word 64 + MemFail :=
+   Definition load (w:{w: word 64 | aligned w}):word 64 + MemFail :=
     match  $\sigma.\mathcal{M}$ !! (pfn (proj1_sig w)) with
@@
       end.
-  Definition store (w: { w: word 64 | aligned w }) (v: word 64):mem64 * (unit + MemFail) :=
+  Definition store (w:{w: word 64 | aligned w}) (v:word 64):mem64 * (unit + MemFail) :=
     match $\sigma.\mathcal{M}$ !! (pfn (proj1_sig w)) with

### 2023-04-11 programlogic.tex
compiles
+

@@
 \newcommand{\rtv}{\textsf{rtv}}
+\newcommand{\qone}{\texttt{q1}}
+\newcommand{\qtwo}{\texttt{q2}}
+\newcommand{\qthree}{\texttt{q3}}
+\newcommand{\qfour}{\texttt{q4}}
 
@@
 \begin{figure}[t]
-\[
-\begin{array}{cl}
-\textsf{vProp } \Sigma \; : \; \textsf{bi} \; \stackrel{{\triangle}}{:=} \mathcal{W}_{64} \rightarrow_{\textsf{b}} \textsf{iPropI } \Sigma
-\end{array}
-\]
+\begin{lstlisting}[language=Coq]
+  Definition vProp  $\Sigma$ : bi := word 64 $\rightarrow$_{b} iPropI  $\Sigma$.
+  \end{lstlisting}
 \caption{$\textsf{vProp }\Sigma$: Root-Address Indexed Address-Space Proposition}
@@
       \hbox{(\TirNameStyle{{L4translate}})} \quad
-      \exists \entryf \ldotp \nfpointsto{\mask\vaddr\ft\rtv}{\mask\vaddr\tw\rtv}\entryf\qfracfotsss\naddr
+      \exists \entryf \ldotp \nfpointsto{\mask\vaddr\ft\rtv}{\mask\vaddr\tw\rtv}\entryf\qone\naddr
       \end{array}
@@
     \hbox{(\TirNameStyle{{L3translate}})} \quad
-    \exists \entrytr \ldotp \nfpointsto{\mask\vaddr\ft\entryf}{\mask\vaddr\tw\entryf}\entrytr\qfracfotss\naddr
+    \exists \entrytr \ldotp \nfpointsto{\mask\vaddr\ft\entryf}{\mask\vaddr\tw\entryf}\entrytr\qtwo\naddr
     \end{array}
@@
             \hbox{(\TirNameStyle{{L2translate}})} \quad 
-       \exists \entrytw \ldotp \nfpointsto{\mask\vaddr\ft\entrytr}{\mask\vaddr\tw\entrytr}\entrytw\qfracfots\naddr
+       \exists \entrytw \ldotp \nfpointsto{\mask\vaddr\ft\entrytr}{\mask\vaddr\tw\entrytr}\entrytw\qthree\naddr
       \end{array}
@@
       \hbox{(\TirNameStyle{{L1translate}})} \quad
-      \exists \entryo \ldotp \nfpointsto{\mask\vaddr\ft\entrytw}{\mask\vaddr\tw\entrytw}\entryo\qfracfot\naddr
+      \exists \entryo \ldotp \nfpointsto{\mask\vaddr\ft\entrytw}{\mask\vaddr\tw\entrytw}\entryo\qfour\naddr
       \end{array}
@@
 \begin{array}{l}
-\begin{array}{l}
   \vaddr\mapsto_{\textsf{t}}\{\textsf{q}\}\vpage\; : \mathsf{vProp} \; \Sigma \;  \stackrel{{\triangle}}{:=} 
@@
  \quad \textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\rtv,\entryf,\entrytr,\entrytw,\entryo) \ast \nfpointsto{\mask\vaddr\ft\entryo}{\mask\vaddr\tw\entryo}\vpage\qfracone\naddr 
-\end{array} \\
-\\
-\textsf{where   }\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\vaddr,\rtv,\entryf,\entrytr,\entrytw,\entryo) \stackrel{\triangle}{=} \\
- \qquad\qquad \nfpointsto{\mask\vaddr\ft\rtv}{\mask\vaddr\tw\rtv}\entryf\qfracfotsss\naddr \ast \\ 
- \qquad\qquad  \nfpointsto{\mask\vaddr\ft\entryf}{\mask\vaddr\tw\entryf}\entrytr\qfracfotss\naddr  \ast  \\
-  \qquad\qquad \nfpointsto{\mask\vaddr\ft\entrytr}{\mask\vaddr\tw\entrytr}\entrytw\qfracfots\naddr \ast \\
-  \qquad\qquad \nfpointsto{\mask\vaddr\ft\entrytw}{\mask\vaddr\tw\entrytw}\entryo\qfracfot\naddr 
 \end{array}
 \]
+\begin{lstlisting}[language=Coq]
+ Definition L4_L1_PointsTo (cr3val maddr entry entry0 entry1 pte :word 64) : vProp $\Sigma$ :=
+  ((l4M52 maddr cr3val) \$\sim$\ (l4off maddr cr3val) $\mapsto_{a}$ {q1} entry  $\ast$
+  (l3M52 maddr entry) \$\sim$\ (l3off maddr entry)  $\mapsto_{a}$ {q2} entry0 $\ast$ 
+  (l2M52 maddr entry0) \$\sim$\ (l2off maddr entry0) $\mapsto_{a}$ {q3} entry1 $\ast$
+  (l1M52 maddr entry1) \$\sim$\ (l1off maddr entry1) $\mapsto_{a}$ {q4} pte ).
+\end{lstlisting}
 \caption{A Strong Virtual Points-to Relation}
@@
 \end{figure*}
-\end{remark}
+
 %then we reach to a physical page addr (e.g. \textsf{pa}) 

### 2023-04-11 programlogic.tex
+

@@
 \begin{figure*}
-  \[
-\begin{array}{l}
-  \vaddr\mapsto_{\textsf{t}}\{\textsf{q}\}\vpage\; : \mathsf{vProp} \; \Sigma \;  \stackrel{{\triangle}}{:=} 
-  \exists_{\entryf \;, \entrytr \;, \entrytw \;,\entryo} \ldotp
-
-  \ulcorner \textsf{aligned } \vaddr \urcorner \ast  \\
- \quad \textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\rtv,\entryf,\entrytr,\entrytw,\entryo) \ast \nfpointsto{\mask\vaddr\ft\entryo}{\mask\vaddr\tw\entryo}\vpage\qfracone\naddr 
-\end{array}
-\]
 \begin{lstlisting}[language=Coq]
- Definition L4_L1_PointsTo (cr3val maddr entry entry0 entry1 pte :word 64) : vProp $\Sigma$ :=
-  ((l4M52 maddr cr3val) \$\sim$\ (l4off maddr cr3val) $\mapsto_{a}$ {q1} entry  $\ast$
+ Definition $\vaddr\mapsto_{\textsf{t}}\{\textsf{q}\}\vpage$ :vProp $\Sigma$ := 
+  $\exists_{\entryf \;, \entrytr \;, \entrytw \;,\entryo} \ldotp$  $\ulcorner$ aligned $\vaddr \urcorner \ast$ L4_L1_PointsTo($\rtv,\entryf,\entrytr,\entrytw,\entryo$) $\ast \entryo \mapsto_{p} \vpage$.
+ Definition L4_L1_PointsTo (rtv maddr entry entry0 entry1 pte :word 64) : vProp $\Sigma$ :=
+  ((l4M52 maddr rtv) \$\sim$\ (l4off maddr rtv) $\mapsto_{a}$ {q1} entry  $\ast$
   (l3M52 maddr entry) \$\sim$\ (l3off maddr entry)  $\mapsto_{a}$ {q2} entry0 $\ast$ 

### 2023-04-11 programlogic.tex
in the middle of refactoring

@@
 \textsf{PDE} in Figure \ref{fig:pagetables}) abstracts masking -- $|^{52}$ for frame for and $|^{12}$ for offset -- through logical nested maps.
-\begin{figure}[t]
-\[
-\begin{array}{cl}
-\pfpointsto\locsf\vpts\qfrac\ppts \stackrel{{\triangle}}{:=} & \nfpointsto{\mask\locsf\ft\generalentry}{\mask\locsf\tw\generalentry}\vpts\qfrac\naddr
-\end{array}
-\]
-\caption{Physical Points-to with Nested Masking}
-  \label{fig:physicalpointsto}
-\end{figure}
+
 Given the definition of physical page-pointsto assertion and the root address of virtual-address space kept in the control register \lstinline{cr3} as shown in Figure \ref{fig:pagetable},  one can build the physical address-translation for a virtual address (e.g.
@@
   \begin{lstlisting}[language=Coq]
+ Definition $\pfpointsto\locsf\vpts\qfrac\ppts$ :vProp $\Sigma$ := pmapsto_def_v (pfn w) (page_off w) q v.
  Definition $\vaddr\mapsto_{\textsf{t}}\{\textsf{q}\}\vpage$ :vProp $\Sigma$ := 
-  $\exists_{\entryf \;, \entrytr \;, \entrytw \;,\entryo} \ldotp$  $\ulcorner$ aligned $\vaddr \urcorner \ast$ L4_L1_PointsTo($\rtv,\entryf,\entrytr,\entrytw,\entryo$) $\ast \entryo \mapsto_{p} \vpage$.
+  $\exists_{\entryf \;, \entrytr \;, \entrytw \;,\entryo} \ldotp$  $\ulcorner$ aligned $\vaddr \urcorner \ast$ L4_L1_PointsTo($\rtv,\entryf,\entrytr,\entrytw,\entryo$) $\ast$ addr_L1$(\vaddr,\entryo) \mapsto_{p} \vpage$.
  Definition L4_L1_PointsTo (rtv maddr entry entry0 entry1 pte :word 64) : vProp $\Sigma$ :=
@@
 \end{figure*}
-
-%then we reach to a physical page addr (e.g. \textsf{pa}) 
-%\[  \paddr\mapsto_{\textsf{a}}\{1\}\vsome \ast \nfpointsto{\mask\vaddr\ft\entryo}{\mask\vaddr\tw\entryo}\vpage\qfrac\naddr \]
 

### 2023-04-11 syntaxsemantics.tex
+

@@
  Definition translate (root: word 64) (w: word 64): address :=
-      let plm4e := translate_top_level $\sigma.\mathcal{M}$ root  w in
-      let pdpte := translate_from_l4e $\sigma.\mathcal{M}$ pml4e w in 
-      let pde := translate_from_l3e $\sigma.\mathcal{M}$ pdpte w in
-      let pte: = translate_from_l2e $\sigma.\mathcal{M}$ pde w in
-      let addr := translate_from_l1e $\sigma.\mathcal{M}$ pte w in
+      l4e $\leftarrow$ translate_top_level $\sigma.\mathcal{M}$ root w; 
+      l3e $\leftarrow$ translate_from_l4e $\sigma.\mathcal{M}$ pml4e w; 
+      l2e $\leftarrow$ translate_from_l3e $\sigma.\mathcal{M}$ pdpte w;
+      l1e $\leftarrow$ translate_from_l2e $\sigma.\mathcal{M}$ pde w;
+      addr $\leftarrow$ translate_from_l1e $\sigma.\mathcal{M}$ pte w;
       concat (shift_dropn addr 3 three_lt64) low3.
-    end.
 \end{lstlisting}
 we see physical memory $\mathsf{load}$ for each level of page-table-walk concluded with alignment of the returned address.
-\begin{lstlisting}[language=Coq]
-   Definition load (w:{w: word 64 | aligned w}):word 64 + MemFail :=
-    match  $\sigma.\mathcal{M}$ !! (pfn (proj1_sig w)) with
-      | None => inl (zero 64)
-      | Some (inl page) => inl (page (page_off (proj1_sig w)))
-      | Some (inr fault) => inr fault
-      end.
-  Definition store (w:{w: word 64 | aligned w}) (v:word 64):mem64 * (unit + MemFail) :=
-    match $\sigma.\mathcal{M}$ !! (pfn (proj1_sig w)) with
-      | None =>
-        ((<[(pfn (proj1_sig w)) := (inl (fun off => if (word_dec off (page_off (proj1_sig w)))
-                                        then v else zero 64))]> $\sigma.\mathcal{M}$), inl tt)
-      | Some (inl page) => (* update page contents *)
-        ((<[(pfn (proj1_sig w)) := (inl (fun off => if (word_dec off (page_off (proj1_sig w)))
-                                        then v else zero 64))]> $\sigma.\mathcal{M}$), inl tt)
-      | Some (inr fault) => ($\sigma.\mathcal{M}$, inr fault)
-    end.
-  Program Definition pfn (w : word 64) := shift_dropn w 12 _.
-  Program Definition page_off (w : word 64) := bottomn w 12 _.
-\end{lstlisting}
-with frame and offset masking applied on the physical address being loaded. As expected \textsf{store} would be the update on the memory map entry
+
+As expected \textsf{store} would be the update on the memory map entry
 \todo[inline,color=yellow]{Colin, could you mention that we simplified the maps here to get rid of MemFail?}

### 2023-04-11 programlogic.tex
+

@@
 \begin{lstlisting}[language=Coq]
-  Definition vProp  $\Sigma$ : bi := word 64 $\rightarrow$_{b} iPropI  $\Sigma$.
+  Definition vProp  $\Sigma$ : bi := word 64 $\rightarrow_{b}$ iPropI  $\Sigma$.
   \end{lstlisting}
@@
 %\end{itemize}
-
 \subsection{Points-To Assertions}
@@
 
+\begin{figure*}
+  \begin{lstlisting}[language=Coq]
+ Definition $\pfpointsto\locsf\vpts\qfrac\ppts$ :vProp $\Sigma$ := pmapsto_def_v (pfn w) (page_off w) q v.
+ Definition $\vaddr\mapsto_{\textsf{t}}\{\textsf{q}\}\vpage$ :vProp $\Sigma$ := 
+  $\exists_{\entryf \;, \entrytr \;, \entrytw \;,\entryo} \ldotp$  $\ulcorner$ aligned $\vaddr \urcorner \ast$ L4_L1_PointsTo($\rtv,\entryf,\entrytr,\entrytw,\entryo$) $\ast$ addr_L1$(\vaddr,\entryo) \mapsto_{p} \vpage$.
+ Definition L4_L1_PointsTo (rtv maddr l4e l3e l2e l1e :word 64) : vProp $\Sigma$ :=
+  (l4M52 maddr rtv) \$\sim$\ (l4off maddr rtv) $\mapsto_{a}$ {q1} l4e  $\ast$
+  (l3M52 maddr entry) \$\sim$\ (l3off maddr entry)  $\mapsto_{a}$ {q2} l3e $\ast$ 
+  (l2M52 maddr entry0) \$\sim$\ (l2off maddr entry0) $\mapsto_{a}$ {q3} l2e $\ast$
+  (l1M52 maddr entry1) \$\sim$\ (l1off maddr entry1) $\mapsto_{a}$ {q4} l1e.
+\end{lstlisting}
+\caption{A Strong Virtual Points-to Relation}
+\todo[inline,color=red]{Colin says: shouldn't that fraction be a q, not a 1, in the second line?}
+  \label{fig:strongvirtualpointsto}
+\end{figure*}
 Given the definition of physical page-pointsto assertion and the root address of virtual-address space kept in the control register \lstinline{cr3} as shown in Figure \ref{fig:pagetable},  one can build the physical address-translation for a virtual address (e.g.
@@
 [47-39] plus 3 alignment bits in Figure \ref{fig:pagetables}
-    \[ \begin{array}{l}
-      \hbox{(\TirNameStyle{{L4translate}})} \quad
-      \exists \entryf \ldotp \nfpointsto{\mask\vaddr\ft\rtv}{\mask\vaddr\tw\rtv}\entryf\qone\naddr
-      \end{array}
-      \]
+    \begin{lstlisting}[language=Coq]
+      $\hbox{(\TirNameStyle{{L4translate}})} \quad$ (l4M52 maddr rtv) \$\sim$\ (l4off maddr rtv) $\mapsto_{a}$ {q1} l4e 
+    \end{lstlisting}
  \item Page Directory Pointer Level Translation (L3): Performs the second level address translation to get the PDP entry (PDPE) in Figure \ref{fig:pagetables} by using the masks(\textsf{va}$|^{52}$(\textsf{PLM4E}) to get the starting address of the PDP table (PDPT) and using the offset mask \textsf{va}$|^{12}$(\textsf{PLM4E})
-    \[\begin{array}{l}
-    \hbox{(\TirNameStyle{{L3translate}})} \quad
-    \exists \entrytr \ldotp \nfpointsto{\mask\vaddr\ft\entryf}{\mask\vaddr\tw\entryf}\entrytr\qtwo\naddr
-    \end{array}
-    \]
+    \begin{lstlisting}[language=Coq]
+    $\hbox{(\TirNameStyle{{L3translate}})}$ (l3M52 maddr entry) \$\sim$\ (l3off maddr entry)  $\mapsto_{a}$ {q2} l3e
+    \end{lstlisting}
   \item Page Directory Table Level Translation (L2): Performs the second level address translation to get the PD entry (PDE) in Figure \ref{fig:pagetables} by using the masks(\textsf{va}$|^{52}$(\textsf{PDPE}) to get the starting address of the PD table and using the offset mask \textsf{va}$|^{12}$(\textsf{PDPE})
-    \[ \begin{array}{l}
-            \hbox{(\TirNameStyle{{L2translate}})} \quad 
-       \exists \entrytw \ldotp \nfpointsto{\mask\vaddr\ft\entrytr}{\mask\vaddr\tw\entrytr}\entrytw\qthree\naddr
-      \end{array}
-      \]
+\begin{lstlisting}[language=Coq]
+    $\hbox{(\TirNameStyle{{L2translate}})}$ (l2M52 maddr entry) \$\sim$\ (l2off maddr entry)  $\mapsto_{a}$ {q2} l2e
+    \end{lstlisting}
   \item Page Table Level Translation (L1): Performs the second level address translation to get the PT entry (PTE) in Figure \ref{fig:pagetables} by using the masks(\textsf{va}$|^{52}$(\textsf{PDE}) to get the starting address of the Page table (PT) and using the offset mask \textsf{va}$|^{12}$(\textsf{PDE})
-    \[ \begin{array}{l}
-      \hbox{(\TirNameStyle{{L1translate}})} \quad
-      \exists \entryo \ldotp \nfpointsto{\mask\vaddr\ft\entrytw}{\mask\vaddr\tw\entrytw}\entryo\qfour\naddr
-      \end{array}
-      \]
+   \begin{lstlisting}[language=Coq]
+    $\hbox{(\TirNameStyle{{L1translate}})}$ (l1M52 maddr entry) \$\sim$\ (l1off maddr entry)  $\mapsto_{a}$ {q2} l1e
+    \end{lstlisting}
   \item Page Address Level Translation: Performs the second level address translation to get the physical page address in Figure \ref{fig:pagetables} by using the masks(\textsf{va}$|^{52}$(\textsf{PTE}) to get the starting address of the physical page address table and using the offset mask \textsf{va}$|^{12}$(\textsf{PTE})
-    \[ \begin{array}{l}
-      \hbox{(\TirNameStyle{{PageLebelAccess}})} \quad
-      \exists \vpage \ldotp \nfpointsto{\mask\vaddr\ft\entryo}{\mask\vaddr\tw\entryo}\vpage\qfracone\naddr
-      \end{array}
-      \]
-
+   \begin{lstlisting}[language=Coq]
+    $\hbox{(\TirNameStyle{{PageLabelAccess}})}$ addr_L1$(\vaddr$,\textsf{l1e}$) \mapsto_{p} \vpage$.
+    \end{lstlisting} 
 \end{itemize}
@@
 address space.}
-
-\begin{figure*}
-  \begin{lstlisting}[language=Coq]
- Definition $\pfpointsto\locsf\vpts\qfrac\ppts$ :vProp $\Sigma$ := pmapsto_def_v (pfn w) (page_off w) q v.
- Definition $\vaddr\mapsto_{\textsf{t}}\{\textsf{q}\}\vpage$ :vProp $\Sigma$ := 
-  $\exists_{\entryf \;, \entrytr \;, \entrytw \;,\entryo} \ldotp$  $\ulcorner$ aligned $\vaddr \urcorner \ast$ L4_L1_PointsTo($\rtv,\entryf,\entrytr,\entrytw,\entryo$) $\ast$ addr_L1$(\vaddr,\entryo) \mapsto_{p} \vpage$.
- Definition L4_L1_PointsTo (rtv maddr entry entry0 entry1 pte :word 64) : vProp $\Sigma$ :=
-  ((l4M52 maddr rtv) \$\sim$\ (l4off maddr rtv) $\mapsto_{a}$ {q1} entry  $\ast$
-  (l3M52 maddr entry) \$\sim$\ (l3off maddr entry)  $\mapsto_{a}$ {q2} entry0 $\ast$ 
-  (l2M52 maddr entry0) \$\sim$\ (l2off maddr entry0) $\mapsto_{a}$ {q3} entry1 $\ast$
-  (l1M52 maddr entry1) \$\sim$\ (l1off maddr entry1) $\mapsto_{a}$ {q4} pte ).
-\end{lstlisting}
-\caption{A Strong Virtual Points-to Relation}
-\todo[inline,color=red]{Colin says: shouldn't that fraction be a q, not a 1, in the second line?}
-  \label{fig:strongvirtualpointsto}
-\end{figure*}
-
 \subsection{Aliasing/Sharing Physical Pages}

### 2023-04-11 programlogic.tex
+

@@
 \begin{itemize}
-  \item Page Map Level-4 Translation (L4): Performs the first level address translation to get the PML4 entry (PML4E) in Figure \ref{fig:pagetables} by using the masks (\textsf{va}$|^{52}$(\lstinline{cr3}) to get the starting address of the PML4 table (PLM4T) and using the offset mask \textsf{va}$|^{12}$(\lstinline{cr3}).
-It is worth noting that our offset indices are 12 bits as all physical memory pointsto assertions has to satisfy being \textsf{\textit{aligned}}, e.g.
-[47-39] plus 3 alignment bits in Figure \ref{fig:pagetables}
+  \item Level-4 Translation (L4): Performs the first level address translation to get the L4 entry (L4E) in Figure \ref{fig:pagetables} by using the masks l4M52 and l4off with \textsf{rtv} virtual base address to get the starting address of the L4 table (L4T), and locate the entry respectively.
     \begin{lstlisting}[language=Coq]
@@
     \end{lstlisting}
- \item Page Directory Pointer Level Translation (L3): Performs the second level address translation to get the PDP entry (PDPE) in Figure \ref{fig:pagetables} by using the masks(\textsf{va}$|^{52}$(\textsf{PLM4E}) to get the starting address of the PDP table (PDPT) and using the offset mask \textsf{va}$|^{12}$(\textsf{PLM4E})
+ \item Level-3 Translation (L3): Performs the second level address translation to get the L3 entry (L3E) in Figure \ref{fig:pagetables} by using the masks l3M52 and l3off with L4 entry (l4e) to get the starting address of the L3 table (L3T), and locate the entry respectively.
     \begin{lstlisting}[language=Coq]
-    $\hbox{(\TirNameStyle{{L3translate}})}$ (l3M52 maddr entry) \$\sim$\ (l3off maddr entry)  $\mapsto_{a}$ {q2} l3e
+    $\hbox{(\TirNameStyle{{L3translate}})}$ (l3M52 maddr l4e) \$\sim$\ (l3off maddr l4e)  $\mapsto_{a}$ {q2} l3e
     \end{lstlisting}
-  \item Page Directory Table Level Translation (L2): Performs the second level address translation to get the PD entry (PDE) in Figure \ref{fig:pagetables} by using the masks(\textsf{va}$|^{52}$(\textsf{PDPE}) to get the starting address of the PD table and using the offset mask \textsf{va}$|^{12}$(\textsf{PDPE})
+  \item Level-2 Translation (L2): Performs the second level address translation to get the L2 entry (L2E) in Figure \ref{fig:pagetables} by using the masks l2M52 and l2off with L3 entry (l3e) to get the starting address of the L2 table (L2T), and locate the entry respectively.
 \begin{lstlisting}[language=Coq]
-    $\hbox{(\TirNameStyle{{L2translate}})}$ (l2M52 maddr entry) \$\sim$\ (l2off maddr entry)  $\mapsto_{a}$ {q2} l2e
+    $\hbox{(\TirNameStyle{{L2translate}})}$ (l2M52 maddr l3e) \$\sim$\ (l2off maddr l3e)  $\mapsto_{a}$ {q2} l2e
     \end{lstlisting}
-  \item Page Table Level Translation (L1): Performs the second level address translation to get the PT entry (PTE) in Figure \ref{fig:pagetables} by using the masks(\textsf{va}$|^{52}$(\textsf{PDE}) to get the starting address of the Page table (PT) and using the offset mask \textsf{va}$|^{12}$(\textsf{PDE})
+  \item Level-1 Translation (L1): Performs the second level address translation to get the L1 entry (L1E) in Figure \ref{fig:pagetables} by using the masks l1M52 and l1off with L2 entry (l2e) to get the starting address of the L1 table (L1T), and locate the entry respectively.
    \begin{lstlisting}[language=Coq]
-    $\hbox{(\TirNameStyle{{L1translate}})}$ (l1M52 maddr entry) \$\sim$\ (l1off maddr entry)  $\mapsto_{a}$ {q2} l1e
+    $\hbox{(\TirNameStyle{{L1translate}})}$ (l1M52 maddr l2e) \$\sim$\ (l1off maddr l2e)  $\mapsto_{a}$ {q2} l1e
     \end{lstlisting}
-  \item Page Address Level Translation: Performs the second level address translation to get the physical page address in Figure \ref{fig:pagetables} by using the masks(\textsf{va}$|^{52}$(\textsf{PTE}) to get the starting address of the physical page address table and using the offset mask \textsf{va}$|^{12}$(\textsf{PTE})
+  \item Page Address Level Translation: Final computed physical page-address ($\textsf{addr\_L1}(\vaddr$,$\textsf{l1e}$)) points-to the value stored in the address ($\vpage$).
    \begin{lstlisting}[language=Coq]

### 2023-04-11 programlogic.tex
+

@@
 The fraction $\qfrac$ with value 1 asserts the unique ownership of the register mapping, and grants update permission on it, otherwise, any value $0 < \qfrac <1$ represents partial ownership granting readonly permission on the mapping.
-
-\todo[inline]{We should drop all talk about PDE, PDPT, etc., and just use L1-L4}
-\paragraph{Physical-Memory points-to} Our physical memory points-to relation in Figure \ref{fig:physicalpointsto} for a page table entry (e.g.
-\textsf{PDE} in Figure \ref{fig:pagetables}) abstracts masking -- $|^{52}$ for frame for and $|^{12}$ for offset -- through logical nested maps.
-
 \begin{figure*}

### 2023-04-11 syntaxsemantics.tex
+

@@
 \begin{lstlisting}[language=Coq]
- Definition translate (root: word 64) (w: word 64): address :=
-      l4e $\leftarrow$ translate_top_level $\sigma.\mathcal{M}$ root w; 
-      l3e $\leftarrow$ translate_from_l4e $\sigma.\mathcal{M}$ pml4e w; 
-      l2e $\leftarrow$ translate_from_l3e $\sigma.\mathcal{M}$ pdpte w;
-      l1e $\leftarrow$ translate_from_l2e $\sigma.\mathcal{M}$ pde w;
-      addr $\leftarrow$ translate_from_l1e $\sigma.\mathcal{M}$ pte w;
+ Definition translate (rtv: word 64) (w: word 64): address :=
+      l4e $\leftarrow$ translate_top_level $\sigma.\mathcal{M}$ rtv w; 
+      l3e $\leftarrow$ translate_from_l4e $\sigma.\mathcal{M}$ l4e w; 
+      l2e $\leftarrow$ translate_from_l3e $\sigma.\mathcal{M}$ l3e w;
+      l1e $\leftarrow$ translate_from_l2e $\sigma.\mathcal{M}$ l2e w;
+      addr $\leftarrow$ translate_from_l1e $\sigma.\mathcal{M}$ l1e w;
       concat (shift_dropn addr 3 three_lt64) low3.

### 2023-04-11 programlogic.tex
+

@@
  Definition $\vaddr\mapsto_{\textsf{t}}\{\textsf{q}\}\vpage$ :vProp $\Sigma$ := 
-  $\exists_{\entryf \;, \entrytr \;, \entrytw \;,\entryo} \ldotp$  $\ulcorner$ aligned $\vaddr \urcorner \ast$ L4_L1_PointsTo($\rtv,\entryf,\entrytr,\entrytw,\entryo$) $\ast$ addr_L1$(\vaddr,\entryo) \mapsto_{p} \vpage$.
+  $\exists_{\textsf{l4e l3e l2e l1}} \ldotp$  $\ulcorner$ aligned $\vaddr \urcorner \ast$ L4_L1_PointsTo(rtv maddr l4e l3e l2e l1e) $\ast$ addr_L1$(\vaddr,\textsf{l1e}) \mapsto_{p} \vpage$.
  Definition L4_L1_PointsTo (rtv maddr l4e l3e l2e l1e :word 64) : vProp $\Sigma$ :=
   (l4M52 maddr rtv) \$\sim$\ (l4off maddr rtv) $\mapsto_{a}$ {q1} l4e  $\ast$
-  (l3M52 maddr entry) \$\sim$\ (l3off maddr entry)  $\mapsto_{a}$ {q2} l3e $\ast$ 
-  (l2M52 maddr entry0) \$\sim$\ (l2off maddr entry0) $\mapsto_{a}$ {q3} l2e $\ast$
-  (l1M52 maddr entry1) \$\sim$\ (l1off maddr entry1) $\mapsto_{a}$ {q4} l1e.
+  (l3M52 maddr l4e) \$\sim$\ (l3off maddr l4e)  $\mapsto_{a}$ {q2} l3e $\ast$ 
+  (l2M52 maddr l3e) \$\sim$\ (l2off maddr l3e) $\mapsto_{a}$ {q3} l2e $\ast$
+  (l1M52 maddr l2e) \$\sim$\ (l1off maddr l2e) $\mapsto_{a}$ {q4} l1e.
 \end{lstlisting}
@@
    \begin{lstlisting}[language=Coq]
-    $\hbox{(\TirNameStyle{{PageLabelAccess}})}$ addr_L1$(\vaddr$,\textsf{l1e}$) \mapsto_{p} \vpage$.
+    $\hbox{(\TirNameStyle{{PageLabelAccess}})} \qquad$ addr_L1($\vaddr,\textsf{l1e}$) $\mapsto_{p} \vpage$.
     \end{lstlisting} 
@@
     \vaddr\mapsto_{\textsf{v,rtv}}\{\textsf{q}\}\vpage \stackrel{\triangle}{:=} 
-    \exists \entryo \ldotp
+    \exists \paddr\ldotp
     \sumapaces\rtv\delta \ast 
-  \sumwalkabs\vaddr\qfrac\paddr \ast 
-   \nfpointsto{\mask\vaddr\ft\entryo}{\mask\vaddr\tw\entryo}\vpage\qfracone\naddr
+  \sumwalkabs\vaddr\qfrac\paddr \ast \paddr \mapsto_{p} \vpage
 \end{array}
@@
   \exists\;\gammaPred \;\ldotp\; \ownGhost\gammaPred{\authfull{\ptableabswalk\ptablestore}} \ast \\
-  \bigast{(\vaddr, \entryo)\in \ptablestore}{\exists\;(\entryf\;,\entrytr\;,\entrytw)\ldotp \textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\vaddr,\rtv,\entryf,\entrytr,\entrytw,\entryo)}
+  \bigast{(\vaddr, \textsf{l1e})\in \ptablestore}{\exists\;(\textsf{l4e l3e l2e})\ldotp \textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\rtv,\textsf{ maddr, l4e, l3e, l2e, l1e})}
 \end{array}
@@
 In other words, physical $\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$ sitting aside to  $ \sumwalkabs\vaddr\qfrac\paddr$ would be made invalid by any concurrent update on the physical page-table.
-So there should only be one physical memory resource inside the virtual pointsto definition: from physical page address to the value stored ($   \nfpointsto{\mask\vaddr\ft\entryo}{\mask\vaddr\tw\entryo}\vpage\qfracone\naddr$).
+So there should only be one physical memory resource inside the virtual pointsto definition: from physical page address to the value stored ($ \paddr \mapsto_{p} \vpage$).
 \end{remark}
@@
 \begin{figure}
-
 \[

### 2023-04-11 programlogic.tex
+

@@
   \begin{array}{l}
-    \textsf{wpd\_def e s E1 } \Phi \;\mathsf{ cr3val } : \textsf{iProp }\Sigma := \\
-   \qquad ((\textsf{cr cr3}) \mapsto_{\textsf{r}} (\textsf{num cr3val)}) \textsf{ cr3val}) \ast (\Phi \textsf{ cr3val}) \wand \textsf{WP e E } @ \textsf{s ; E1 } \{\{\_, \textsf{True} \}\}
+    \textsf{wpd\_def e s E1 } \Phi \;\mathsf{ rtv } : \textsf{iProp }\Sigma := \\
+   \qquad ((\textsf{cr cr3}) \mapsto_{\textsf{r}} (\textsf{num rtv)}) \textsf{ cr3val}) \ast (\Phi \textsf{ rtv}) \wand \textsf{WP e E } @ \textsf{s ; E1 } \{\{\_, \textsf{True} \}\}
     \end{array}
@@
 \begin{mathpar}
-  \inferrule[WriteToRegFromReg | \{\textsf{rtv}\}]{
-  \{P \ast r_d \mapsto_{r} \textsf{rvs} \ast r_s \mapsto_{r}\{q\} \textsf{ rvs} \}\;\overline{ is}
+  \inferrule[WriteToRegFromReg]{
+  \{P \ast r_d \mapsto_{r} \textsf{rvs} \ast r_s \mapsto_{r}\{q\} \textsf{ rvs} \}_{\textsf{rtv}}\;\overline{ is}
 }{
-  \{P \ast r_d \mapsto_{r} \textsf{rvd} \ast r_s \mapsto_{r}\{q\} \textsf{ rvs} \}
+  \{P \ast r_d \mapsto_{r} \textsf{rvd} \ast r_s \mapsto_{r}\{q\} \textsf{ rvs} \}_{\textsf{rtv}}
   \textsf{ mov}~\textsf{r}_d,~\textsf{r}_s;\;\overline{is}
@@
 \\
-\inferrule[WriteToPhysMemFromReg | \{\textsf{rtv}\}]{
-  \{P \ast r_s \mapsto_{r}\{q\}  \textsf{rvs}  \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{t}} \textsf{v} \}\;\overline{ is}  
+\inferrule[WriteToPhysMemFromReg]{
+  \{P \ast r_s \mapsto_{r}\{q\}  \textsf{rvs}  \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{t}} \textsf{v} \}_{\textsf{rtv}}\;\overline{ is}  
 }{
-  \{P \ast r_s \mapsto_{r}\{q\}  \textsf{rvs}   \ast r_a \mapsto_{r}\{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{t}} \textsf{rvs} \}
+  \{P \ast r_s \mapsto_{r}\{q\}  \textsf{rvs}   \ast r_a \mapsto_{r}\{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{t}} \textsf{rvs} \}_{\textsf{rtv}}
   \textsf{ mov}~\textsf{r}_a,~\textsf{r}_s;\;\overline{is}
@@
 \\
-\inferrule[WriteToRegFromPhysMem | \{\textsf{rtv}\}]{
-  \{P \ast r_d \mapsto_{r}  \textsf{v} \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{t}} \textsf{v} \}\;\overline{is}
+\inferrule[WriteToRegFromPhysMem]{
+  \{P \ast r_d \mapsto_{r}  \textsf{v} \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{t}} \textsf{v} \}_{\textsf{rtv}}\;\overline{is}
 }{
-  \{P \ast r_d \mapsto_{r}  \textsf{rvd} \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{t}} \textsf{v} \}
+  \{P \ast r_d \mapsto_{r}  \textsf{rvd} \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{t}} \textsf{v} \}_{\textsf{rtv}}
   \textsf{mov}~\textsf{r}_d,~\textsf{r}_a;\;\overline{is}
@@
 \\
-\inferrule[WriteToRegFromVirtMem | \{\textsf{rtv}\}]{
-  \{P \ast \mathcal{I}\textsf{ASpace}\ast \mathcal{I}\textsf{ASpace} \ast r_d \mapsto_{r}  \textsf{v} \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{v,rtv}} \textsf{v} \}\;\overline{is}
+\inferrule[WriteToRegFromVirtMem]{
+  \{P \ast \mathcal{I}\textsf{ASpace}\ast \mathcal{I}\textsf{ASpace} \ast r_d \mapsto_{r}  \textsf{v} \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{v,rtv}} \textsf{v} \}_{\textsf{rtv}}\;\overline{is}
 }{
-  \{P \ast \mathcal{I}\textsf{ASpace}\ast r_d \mapsto_{r}  \textsf{rvd} \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{v,rtv}} \textsf{v} \}
+  \{P \ast \mathcal{I}\textsf{ASpace}\ast r_d \mapsto_{r}  \textsf{rvd} \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{v,rtv}} \textsf{v} \}_{\textsf{rtv}}
   \textsf{mov}~\textsf{r}_d,~\textsf{r}_a;\;\overline{is}
@@
 \\
-\inferrule[WriteToVirtMemFromReg | \{\textsf{rtv}\}]{
-  \{P \ast \mathcal{I}\textsf{ASpace}\ast r_s \mapsto_{r}\{q\}  \textsf{rvs}  \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{v,rtv}} \textsf{v} \}\;\overline{ is}  
+\inferrule[WriteToVirtMemFromReg]{
+  \{P \ast \mathcal{I}\textsf{ASpace}\ast r_s \mapsto_{r}\{q\}  \textsf{rvs}  \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{v,rtv}} \textsf{v} \}_{\textsf{rtv}}\;\overline{ is}
 }{
-  \{P \ast \mathcal{I}\textsf{ASpace}\ast r_s \mapsto_{r}\{q\}  \textsf{rvs}   \ast r_a \mapsto_{r}\{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{v,rtv}} \textsf{rvs} \}
+  \{P \ast \mathcal{I}\textsf{ASpace}\ast r_s \mapsto_{r}\{q\}  \textsf{rvs}   \ast r_a \mapsto_{r}\{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{v,rtv}} \textsf{rvs} \}_{\textsf{rtv}}
   \textsf{ mov}~\textsf{r}_a,~\textsf{r}_s;\;\overline{is}
@@
 \\
-\inferrule[WriteToRegCtlFromReg | \{\textsf{rtv}\} / \{\textsf{rvs}\}]{
-  \{P \ast r_s \mapsto_{r}\{q\}  \textsf{ rvs}  \}\overline{is}
+\inferrule[WriteToRegCtlFromReg]{
+  \{P \ast r_s \mapsto_{r}\{q\}  \textsf{ rvs}  \}_{\textsf{rvs}} \overline{is}
 }{
-  \{P \ast r_s \mapsto_{r}\{q\}  \textsf{ rvs}   \}
+  \{P \ast r_s \mapsto_{r}\{q\}  \textsf{ rvs}   \}_{\textsf{rtv}}
   \textsf{mov}~\textsf{cr3},~r_s;\;\overline{is}
@@
 }\\
-\inferrule[WriteToRegFromRegCtl | \{\textsf{rtv}\}]{
-  \{P \ast r_d \mapsto_{r} \textsf{rvs} \ast r_s \mapsto_{r}\{q\} \textsf{ rvs} \}\;\overline{ is}
+\inferrule[WriteToRegFromRegCtl]{
+  \{P \ast r_d \mapsto_{r} \textsf{rvs} \ast r_s \mapsto_{r}\{q\} \textsf{ rvs} \}_{\textsf{rtv}}\;\overline{ is}
 }{
-  \{P \ast r_d \mapsto_{r} \textsf{rvd} \ast r_s \mapsto_{r}\{q\} \textsf{ rvs} \}
+  \{P \ast r_d \mapsto_{r} \textsf{rvd} \ast r_s \mapsto_{r}\{q\} \textsf{ rvs} \}_{\textsf{rtv}}
   \textsf{ mov}~\textsf{r}_d,~\textsf{r}_s;\;\overline{is}

### 2023-04-12 programlogic.tex
+

@@
 
-In the general outline of the proof in Figure \ref{wpdamd}, we see that each rule annotated with an address value ($\{\rtv\}$), i.e.
+In the general outline of the proof in Figure \ref{wpdamd}, we see that each rule annotated with an address value (\textsf{rtv}), i.e.
 a root address of an address-space, under which the resources mentioned in the specification are valid.
@@
     \textsf{wpd\_def e s E1 } \Phi \;\mathsf{ rtv } : \textsf{iProp }\Sigma := \\
-   \qquad ((\textsf{cr cr3}) \mapsto_{\textsf{r}} (\textsf{num rtv)}) \textsf{ cr3val}) \ast (\Phi \textsf{ rtv}) \wand \textsf{WP e E } @ \textsf{s ; E1 } \{\{\_, \textsf{True} \}\}
+   \qquad ((\textsf{cr cr3}) \mapsto_{\textsf{r}} (\textsf{num rtv)}) \textsf{ rtv}) \ast (\Phi \textsf{ rtv}) \wand \textsf{WP e E } @ \textsf{s ; E1 } \{\{\_, \textsf{True} \}\}
     \end{array}

### 2023-04-12 soundness.tex
+

@@
 \def\arraystretch{1.2}
-\begin{array}{l@{\quad\star\quad}l@{\quad}l}
+\begin{array}{l@{\quad\ast\quad}l@{\quad}l}
   \textsf{to\_gen\_heap} \;\store.\mathcal{R} & \textsf{to\_gen\_mem} \; \store.\mathcal{M}
@@
 As a last definition, we give the head step relation required for the Iris instantiation for our simple language in Figure \ref{} to sequence instructions.
-This relation allows lifting the program expression to enables the application of changes imposed by the operational semantics on the program state $\sigma$ when applied for an insturction (\textsf{i}).
+This relation allows lifting the program expression to enable the application of changes imposed by the operational semantics on the program state (for a certain cpu, a register map, and a selected memory for an address-space) when applied for an insturction (\textsf{i}).
 \begin{lstlisting}[language=Coq]
 Definition exec_step (i: instr) ($\sigma$:state) : option state :=
-    exec_state (exec_instr i $\sigma\mathcal{C}$ $\sigma.\mathcal{R}$ ((memToPhysMem $\sigma.\mathcal{M}$) (exist _ ($\sigma.\mathcal{R}$ !!  cr3)))).
-
+    exec_instr i $\sigma\mathcal{C}$ $\sigma.\mathcal{R}$ ((memToPhysMem $\sigma.\mathcal{M}$)  ($\sigma.\mathcal{R}$ !!  cr3)).
 \end{lstlisting}
 \todo[inline,color=yellow]{Colin, in case needed,not proven, could you simply say that we assume these map equalities or a paper proof etc.}.
-Due to the page limitation, we can only give the outline of a selected instruction (\TirNameStyle{WriteToRegFromVirtMem}) from our \textsf{AMD64} model.
+Again, due to the page limits, we can only give an outline of proof for a selected instruction (\TirNameStyle{WriteToRegFromVirtMem}) from our \textsf{AMD64} model.
 However, mechanized soundness proofs of other instructions can be found as a part of our submission artifact.
@@
 \inferrule{
-  \{P \ast r_d \mapsto_{r}  \textsf{v} \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{v,rtv}} \textsf{v} \}\;\overline{is}
+  \{P \ast r_d \mapsto_{r}  \textsf{v} \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{v,rtv}} \textsf{v} \}_{\textsf{rtv}}\;\overline{is}
 }{
-  \{P \ast r_d \mapsto_{r}  \textsf{rvd} \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{v,rtv}} \textsf{v} \}
+  \{P \ast r_d \mapsto_{r}  \textsf{rvd} \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{v,rtv}} \textsf{v} \}_{\textsf{rtv}}
   \textsf{mov}~\textsf{r}_d,~\textsf{r}_a;\;\overline{is}

### 2023-04-12 programlogic.tex
+

@@
 \begin{lstlisting}[language=Coq]
-  Definition vProp  $\Sigma$ : bi := word 64 $\rightarrow_{b}$ iPropI  $\Sigma$.
+  Definition vProp  $\Sigma$ : bi := word 64 -b> iPropI  $\Sigma$.
   \end{lstlisting}
@@
 \end{figure*}
-Given the definition of physical page-pointsto assertion and the root address of virtual-address space kept in the control register \lstinline{cr3} as shown in Figure \ref{fig:pagetable},  one can build the physical address-translation for a virtual address (e.g.
+Given the definition of physical page-pointsto assertion and the root address of virtual-address space as shown in Figure \ref{fig:pagetable}, one can build the physical address-translation for a virtual address (e.g.
 \textsf{va}) via abstracting the L4-L1 table traversal as the following:
 \begin{itemize}
-  \item Level-4 Translation (L4): Performs the first level address translation to get the L4 entry (L4E) in Figure \ref{fig:pagetables} by using the masks l4M52 and l4off with \textsf{rtv} virtual base address to get the starting address of the L4 table (L4T), and locate the entry respectively.
+  \item Level-4 Translation (L4): Performs the first level address translation to get the L4 entry (L4E) in Figure \ref{fig:pagetables} by using the masks l4M52 and l4off with \textsf{rtv} virtual base address to get the starting address of the L4 table (L4T), and locate the entry entry amongst 512 (q1) respectively.
     \begin{lstlisting}[language=Coq]
@@
     \end{lstlisting}
- \item Level-3 Translation (L3): Performs the second level address translation to get the L3 entry (L3E) in Figure \ref{fig:pagetables} by using the masks l3M52 and l3off with L4 entry (l4e) to get the starting address of the L3 table (L3T), and locate the entry respectively.
+ \item Level-3 Translation (L3): Performs the second level address translation to get the L3 entry (L3E) in Figure \ref{fig:pagetables} by using the masks l3M52 and l3off with L4 entry (l4e) to get the starting address of the L3 table (L3T), and locate the entry amongst 512 (q2) respectively.
     \begin{lstlisting}[language=Coq]
@@
     \end{lstlisting}
-  \item Level-2 Translation (L2): Performs the second level address translation to get the L2 entry (L2E) in Figure \ref{fig:pagetables} by using the masks l2M52 and l2off with L3 entry (l3e) to get the starting address of the L2 table (L2T), and locate the entry respectively.
+  \item Level-2 Translation (L2): Performs the second level address translation to get the L2 entry (L2E) in Figure \ref{fig:pagetables} by using the masks l2M52 and l2off with L3 entry (l3e) to get the starting address of the L2 table (L2T), and locate the entry entry amongst 512 (q3) respectively.
 \begin{lstlisting}[language=Coq]
-    $\hbox{(\TirNameStyle{{L2translate}})}$ (l2M52 maddr l3e) \$\sim$\ (l2off maddr l3e)  $\mapsto_{a}$ {q2} l2e
+    $\hbox{(\TirNameStyle{{L2translate}})}$ (l2M52 maddr l3e) \$\sim$\ (l2off maddr l3e)  $\mapsto_{a}$ {q3} l2e
     \end{lstlisting}
-  \item Level-1 Translation (L1): Performs the second level address translation to get the L1 entry (L1E) in Figure \ref{fig:pagetables} by using the masks l1M52 and l1off with L2 entry (l2e) to get the starting address of the L1 table (L1T), and locate the entry respectively.
+  \item Level-1 Translation (L1): Performs the second level address translation to get the L1 entry (L1E) in Figure \ref{fig:pagetables} by using the masks l1M52 and l1off with L2 entry (l2e) to get the starting address of the L1 table (L1T), and locate the entry entry amongst 512 (q4) respectively.
    \begin{lstlisting}[language=Coq]
-    $\hbox{(\TirNameStyle{{L1translate}})}$ (l1M52 maddr l2e) \$\sim$\ (l1off maddr l2e)  $\mapsto_{a}$ {q2} l1e
+    $\hbox{(\TirNameStyle{{L1translate}})}$ (l1M52 maddr l2e) \$\sim$\ (l1off maddr l2e)  $\mapsto_{a}$ {q4} l1e
     \end{lstlisting}

### 2023-04-12 soundness.tex
+

@@
 \begin{lstlisting}[language=Coq]
-Definition exec_step (i: instr) ($\sigma$:state) : option state :=
-    exec_instr i $\sigma\mathcal{C}$ $\sigma.\mathcal{R}$ ((memToPhysMem $\sigma.\mathcal{M}$)  ($\sigma.\mathcal{R}$ !!  cr3)).
+Definition exec_step (i: instr) ($\sigma$:state) : option state :=  exec_instr i $\sigma.\mathcal{C}$ $\sigma.\mathcal{R}$ ((memToPhysMem $\sigma.\mathcal{M}$)  ($\sigma.\mathcal{R}$ !! cr3)).
 \end{lstlisting}

### 2023-04-12 examples.tex
+
add mathescape

@@
 One of the key tasks of a page fault handler in a general-purpose OS kernel is to map new pages into an address space by writing into an existing page table.
-\todo[inline]{Ismail, see notes about intermediate assertions on the Notion page}
-%\begin{figure*}\small
-%\[
-%\proofoutstep{x:\isolated\;\texttt{Node},y:\INT,z:\bool}
-%{
- %   \proofoutstep{x:\isolated\;\texttt{Node},y:\INT,z:\bool}
-  %  {
-   %     \proofoutstep{x:\writable\;\texttt{Node},y:\INT,z:\bool}
-    %    {
-     %       \proofoutstep{x:\writable\;\texttt{Node},y:\INT,z:\bool}
-      %      {
-       %       mov\texttt{ }rdi \; r13
-        %    }{\textsc{T-Alloc}}
-         %   {x:\writable\;\texttt{Node},y:\writable\;\texttt{Node},z:\bool}\\
-          %  \;\;\;z\texttt{=}x\texttt{.next;}\\
-           % \specline{x:\writable\;\texttt{Node},y:\writable\;\texttt{Node},z:\writable\;\texttt{Node}}\\
-            %\;\;\;y\texttt{.next=}z\texttt{;}\\
-            %\specline{x:\writable\;\texttt{Node},y:\writable\;\texttt{Node},z:\writable\;\texttt{Node}}\\
-            %\;\;\;z\texttt{.prev=}y\texttt{;}\\
-            %\specline{x:\writable\;\texttt{Node},y:\writable\;\texttt{Node},z:\writable\;\texttt{Node}}\\
-            %\;\;\;y\texttt{.prev=}x\texttt{;}\\
-         %\specline{x:\writable\;\texttt{Node},y:\writable\;\texttt{Node},z:\writable\;\texttt{Node}}\\
-          %  \;\;\;x\texttt{.next=}y\texttt{;}\\
-         %\specline{x:\writable\;\texttt{Node},y:\writable\;\texttt{Node},z:\writable\;\texttt{Node}}
-        %}
-        %{\textsc{T-Seq}*~\textrm{across}~\textsc{T-FieldRead}\\\textrm{and}~\textsc{T-FieldWrite}}
-        %{x:\writable\;\texttt{Node},y:\writable\;\texttt{Node},z:\writable\;\texttt{Node}}
-    %}
-    %{\textsc{T-SubEnv}}
-    %{x:\writable\;\texttt{Node}}
-%}
-%{\textsc{T-RecovIso}}
-%{x:\isolated\;\texttt{Node}}
-%\]
-%\caption{Typing derivation for adding a node to an isolated doubly-linked list.}
-%\label{fig:recoviso_example}
-%\end{figure*}
 \lstset{
@@
   emphstyle={[2]\color{violet}},
-  morecomment=[l]{;;}
+  morecomment=[l]{;;},
+  mathescape
 }
@@
   \begin{lstlisting}
-\{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,\gammaPred)  \ast  \texttt{r14}\mapsto_{r} \_ \ast \texttt{rdi}\mapsto_{r} \vaddr \ast \texttt{rax}\mapsto_{r} \_ \ast \ulcorner \texttt{aligned va} \land  \theta \; !!\; \texttt{maddr} = \texttt{None}\urcorner\}_{\rtv}
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,\gammaPred,\rtv)  \ast  \texttt{r14}\mapsto_{r} \_ \ast \texttt{rdi}\mapsto_{r} \vaddr \ast \texttt{rax}\mapsto_{r} \_ \ast \ulcorner \texttt{aligned va} \land  \theta \; !!\; \texttt{maddr} = \texttt{None}\urcorner}_{\rtv}$
 call ensure_L1_page
-$\{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,\gammaPred) \ast  \texttt{r14}\mapsto_{r} \_ \ast \texttt{rdi}\mapsto_{r} \vaddr \ast \texttt{rax}\mapsto_{r} \_ \ast \}_{\rtv}$
-$\specline{\exists (\entryf ,\;\entrytr,\; \entrytw,\; \entryo: \texttt{word 64}) \ldotp \ulcorner (\texttt{addr\_L1 }\vaddr) = \entryo\urcorner \ast }_{\rtv}$
-$\specline{\nfpointsto{\mask\vaddr\ft\rtv}{\mask\vaddr\tw\rtv}\entryf\qone\naddr \ast}_{\rtv}$ 
-$\specline{  \nfpointsto{\mask\vaddr\ft\entryf}{\mask\vaddr\tw\entryf}\entrytr\qtwo\naddr \ast \nfpointsto{\mask\vaddr\ft\entrytr}{\mask\vaddr\tw\entrytr}\entrytw\qthree\naddr \ast}_{\rtv}$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,\gammaPred),\rtv \ast  \texttt{r14}\mapsto_{r} \_ \ast \texttt{rdi}\mapsto_{r} \vaddr \ast \texttt{rax}\mapsto_{r} \_ \; \ast }_{\rtv}$
+$\specline{\exists (\entryf ,\;\entrytr,\; \entrytw,\; \entryo: \texttt{word 64}) \ldotp \ulcorner (\texttt{addr\_L1 }\vaddr) = \entryo\urcorner \; \ast }_{\rtv}$
+$\specline{\nfpointsto{\mask\vaddr\maskfour\rtv}{\mask\vaddr\maskfouroff\rtv}\entryf\qone\naddr \; \ast}_{\rtv}$ 
+$\specline{  \nfpointsto{\mask\vaddr\maskthree\entryf}{\mask\vaddr\maskthreeoff\entryf}\entrytr\qtwo\naddr \ast \nfpointsto{\mask\vaddr\masktwo\entrytr}{\mask\vaddr\masktwooff\entrytr}\entrytw\qthree\naddr \;\ast}_{\rtv}$
 $\specline{\texttt{pte\_addr} \mapsto_{\texttt{vpte},\rtv} \entryo \;(\texttt{wzero 64}) \ast \texttt{rax}\mapsto_{r} \texttt{pte\_addr}  }_{\rtv}$
@@
 mov %r14, %rax ;; Save that before another call
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,\gammaPred) \ast  \texttt{r14}\mapsto_{r} \texttt{pte\_addr} \ast \texttt{rdi}\mapsto_{r} \vaddr \ast \texttt{rax}\mapsto_{r} \_ \ast }_{\rtv}$
-$\specline{ \nfpointsto{\mask\vaddr\ft\rtv}{\mask\vaddr\tw\rtv}\entryf\qone\naddr \ast}_{\rtv}$ 
-$\specline{  \nfpointsto{\mask\vaddr\ft\entryf}{\mask\vaddr\tw\entryf}\entrytr\qtwo\naddr \ast \nfpointsto{\mask\vaddr\ft\entrytr}{\mask\vaddr\tw\entrytr}\entrytw\qthree\naddr \ast}_{\rtv}$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,\gammaPred,\rtv) \ast  \texttt{r14}\mapsto_{r} \texttt{pte\_addr} \ast \texttt{rdi}\mapsto_{r} \vaddr \ast \texttt{rax}\mapsto_{r} \_ \; \ast }_{\rtv}$
+$\specline{ \nfpointsto{\mask\vaddr\maskfour\rtv}{\mask\vaddr\maskfouroff\rtv}\entryf\qone\naddr \; \ast}_{\rtv}$ 
+$\specline{  \nfpointsto{\mask\vaddr\maskthree\entryf}{\mask\vaddr\maskthreeoff\entryf}\entrytr\qtwo\naddr \ast \nfpointsto{\mask\vaddr\masktwo\entrytr}{\mask\vaddr\masktwooff\entrytr}\entrytw\qthree\naddr \;\ast}_{\rtv}$
 $\specline{\texttt{pte\_addr} \mapsto_{\texttt{vpte},\rtv} \entryo \;(\texttt{wzero 64}) \ast \texttt{rax}\mapsto_{r} \texttt{pte\_addr}  }_{\rtv}$
 call alloc_phys_page_or_panic
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,\gammaPred) \ast  \texttt{r14}\mapsto_{r} \texttt{pte\_addr} \ast \texttt{rdi}\mapsto_{r} \vaddr \ast \texttt{rax}\mapsto_{r} \_ \ast}_{\rtv}$
-$\specline{ \nfpointsto{\mask\vaddr\ft\rtv}{\mask\vaddr\tw\rtv}\entryf\qone\naddr \ast}_{\rtv}$ 
-$\specline{  \nfpointsto{\mask\vaddr\ft\entryf}{\mask\vaddr\tw\entryf}\entrytr\qtwo\naddr \ast \nfpointsto{\mask\vaddr\ft\entrytr}{\mask\vaddr\tw\entrytr}\entrytw\qthree\naddr \ast}_{\rtv}$
-$\specline{\texttt{pte\_addr} \mapsto_{\texttt{vpte},\rtv} \entryo \;(\texttt{wzero 64}) \ast \texttt{rax}\mapsto_{r} \texttt{pte\_addr}  }_{\rtv}$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,\gammaPred,\rtv) \ast  \texttt{r14}\mapsto_{r} \texttt{pte\_addr} \ast \texttt{rdi}\mapsto_{r} \vaddr \ast \texttt{rax}\mapsto_{r} \_ \;\ast}_{\rtv}$
+$\specline{ \nfpointsto{\mask\vaddr\maskfour\rtv}{\mask\vaddr\maskfouroff\rtv}\entryf\qone\naddr \ast}_{\rtv}$ 
+$\specline{  \nfpointsto{\mask\vaddr\maskthree\entryf}{\mask\vaddr\maskthreeoff\entryf}\entrytr\qtwo\naddr \ast \nfpointsto{\mask\vaddr\masktwo\entrytr}{\mask\vaddr\masktwooff\entrytr}\entrytw\qthree\naddr \;\ast}_{\rtv}$
+$\specline{\texttt{pte\_addr} \mapsto_{\texttt{vpte},\rtv} \entryo \;(\texttt{wzero 64}) \ast \texttt{rax}\mapsto_{r} \texttt{pte\_addr} \; \ast }_{\rtv}$
 $\specline{\exists \texttt{ fpaddr} \ldotp \ulcorner \texttt{aligned fpaddr} \urcorner \ast \texttt{rax}\mapsto_{r} \texttt{fpaddr} \ast \texttt{fpaddr} \mapsto_{a} (\texttt{wzero 64})}_{\rtv}$
@@
 mov (%r14), %rax
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,\gammaPred) \ast  \texttt{r14}\mapsto_{r} \texttt{pte\_addr} \ast \texttt{rdi}\mapsto_{r} \vaddr \ast \texttt{rax}\mapsto_{r} \_ \ast}_{\rtv}$
-$\specline{\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\vaddr,\rtv,\entryf,\entrytr,\entrytw,\entryo) \ast \ulcorner \theta \;!!\; \texttt{maddr} = \texttt{None}\urcorner}_{\rtv}$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,\gammaPred,\rtv) \ast  \texttt{r14}\mapsto_{r} \texttt{pte\_addr} \ast \texttt{rdi}\mapsto_{r} \vaddr \ast \texttt{rax}\mapsto_{r} \_ \;\ast}_{\rtv}$
+$\specline{\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\vaddr,\rtv,\entryf,\entrytr,\entrytw,\entryo) \ast \ulcorner \theta \;!!\; \texttt{maddr} = \texttt{None}\urcorner \; \ast}_{\rtv}$
 $\specline{\ulcorner \texttt{aligned fpaddr} \urcorner \ast \texttt{rax}\mapsto_{r} \texttt{fpaddr} \ast \texttt{fpaddr} \mapsto_{a} (\texttt{wzero 64})}_{\rtv}$
 $\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\; \sqsubseteq $
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta<[\vaddr:=\texttt{fpaddr}]>),\gammaPred,\rtv) \ast  \texttt{r14}\mapsto_{r} \texttt{pte\_addr} \ast \texttt{rdi}\mapsto_{r} \vaddr \ast \texttt{rax}\mapsto_{r} \_ }_{\rtv}$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta<[\vaddr:=\texttt{fpaddr}]>),\gammaPred,\rtv) \ast  \texttt{r14}\mapsto_{r} \texttt{pte\_addr} \ast \texttt{rdi}\mapsto_{r} \vaddr \ast \texttt{rax}\mapsto_{r} \_ \; \ast}_{\rtv}$
 $\specline{\ulcorner \texttt{aligned fpaddr} \urcorner \ast \texttt{rax}\mapsto_{r} \texttt{fpaddr} \ast \texttt{aligned va} \ast\sumwalkabs\vaddr\qfrac\fpaddr}_{\rtv}$

### 2023-04-12 syntaxsemantics.tex
+

@@
 \begin{lstlisting}[language=Coq]
- Definition translate (rtv: word 64) (w: word 64): address :=
+ Definition translate (rtv: word 64) (w: word 64): word 64 + MemFail :=
       l4e $\leftarrow$ translate_top_level $\sigma.\mathcal{M}$ rtv w; 
@@
 \end{lstlisting}
-we see physical memory $\mathsf{load}$ for each level of page-table-walk concluded with alignment of the returned address.
+we see a physical memory $\mathsf{load}$ (i.e.
+physical memory map ($\sigma.\mathcal{M}$) for each level) for each level of page-table-walk concluded with alignment of the returned address for each \textit{successful} page-table-walk traversal.
 

### 2023-04-12 examples.tex
+

@@
 \newcommand{\specline}[1]{{\color{blue}\left\{#1\right\}}}
-
 \begin{figure}\footnotesize

### 2023-04-12 syntaxsemantics.tex
+

--- 2023-04-11 syntaxsemantics.tex
@@
 \end{lstlisting}
-we see physical memory $\mathsf{load}$ for each level of page-table-walk concluded with alignment of the returned address.
+we see a physical memory $\mathsf{load}$ (i.e.
+physical memory map ($\sigma.\mathcal{M}$) for each level) for each level of page-table-walk concluded with alignment of the returned address for each \textit{successful} page-table-walk traversal what our reasoning principles only consider.
 

### 2023-04-12 syntaxsemantics.tex
+

@@
 we see a physical memory $\mathsf{load}$ (i.e.
-physical memory map ($\sigma.\mathcal{M}$) for each level) for each level of page-table-walk concluded with alignment of the returned address for each \textit{successful} page-table-walk traversal what our reasoning principles only consider.
+physical memory map ($\sigma.\mathcal{M}$) for each level) for each level of page-table-walk concluded with alignment of the returned address for each \textit{successful} page-table-walk traversal which is what our reasoning principles only consider.
 

### 2023-04-12 examples.tex
+

@@
 One of the key tasks of a page fault handler in a general-purpose OS kernel is to map new pages into an address space by writing into an existing page table.
+To do so, we first allocate a fresh page, then calculate the appropriate known-valid page table walks and update the appropriate L1 page table entry.
 \lstset{
@@
   \begin{lstlisting}
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,\gammaPred,\rtv)  \ast  \texttt{r14}\mapsto_{r} \_ \ast \texttt{rdi}\mapsto_{r} \vaddr \ast \texttt{rax}\mapsto_{r} \_ \ast \ulcorner \texttt{aligned va} \land  \theta \; !!\; \texttt{maddr} = \texttt{None}\urcorner}_{\rtv}$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,\gammaPred,\rtv)  \ast  \texttt{r14}\mapsto_{r} \_ \ast \texttt{rdi}\mapsto_{r} \vaddr \ast \texttt{rax}\mapsto_{r} \_ \ast \ulcorner \texttt{aligned va} \land  \theta \; !!\; \vaddr = \texttt{None}\urcorner}_{\rtv}$
 call ensure_L1_page
@@
 \end{figure}
-
-\todo[inline]{emphasize the write to the L1 page table is validated by the fact that the identity mapping exists}
+In Figure \ref{fig:mapping_code}, we see
+%\todo[inline]{emphasize the write to the L1 page table is validated by the fact that the identity mapping exists}
 

### 2023-04-12 programlogic.tex
+
some edits and comments and todos for myself on section 4

@@
 \item Physical address points-to, $\pfpointsto\locsf\vpts\qfrac\ppts$
+\todo[inline,color=red]{Colin says: This syntax is what most readers would expect for physical poinst-tos, but it's
+inconsistent with the split 52+12  stuff later... Whatever we do, we need to be consistent}
 \item Register points-to, $\pfpointsto\rg\rv\qfrac\rpts$
 \end{enumerate}
+
 \paragraph{Register points-to} The assertion $\pfpointsto\rg\rv\rpts\qfrac$ ensures the ownership of the register $\rg$ naming the register value $\rv$.
 The fraction $\qfrac$ with value 1 asserts the unique ownership of the register mapping, and grants update permission on it, otherwise, any value $0 < \qfrac <1$ represents partial ownership granting readonly permission on the mapping.
+
 \begin{figure*}
@@
 \caption{A Strong Virtual Points-to Relation}
-\todo[inline,color=red]{Colin says: shouldn't that fraction be a q, not a 1, in the second line?}
+\todo[inline,color=red]{Colin says: shouldn't ther be a fraction q in the body of the virtual pointsto def?
+I know this is just a sketch that we then move away from, but it's worth getting accurate.}
+\todo[inline,color=cyan]{Colin says: the virtual points-to here is not defined as a vprop! the body is an iProp,
+containing an unbound variable rtv, which should be the cr3val taken as an argument. (Also, rtv doesn't make sense to me
+as a shorthand, why not cr3val?). I'm also pretty sure the maddr in the virtual points-to definition should be va.}
   \label{fig:strongvirtualpointsto}
 \end{figure*}
+
+\todo[inline,color=red]{Colin says: Ismail, I think you accidentally deleted discussion of physical points-tos here?}
+
+\todo[inline,color=red]{Colin says: Why are we talking about assertions for address translations here? It seems like we should
+\todo[inelin]{Colin says: There's no explanation of the split 52+12 \emph{notation} here. If we added such an explanation
+(maybe to where physical points-tos are introduced?) and moved this discussion to \ref{sec:overly-restrictive}, this is otherwise
+a nice explanation}
+discuss physical memory pointsto assertions in general first, then talk about this in \ref{sec:overly-restrictive}}
 Given the definition of physical page-pointsto assertion and the root address of virtual-address space as shown in Figure \ref{fig:pagetable}, one can build the physical address-translation for a virtual address (e.g.
@@
 \end{itemize}
+
 \subsection{An Overly-Restrictive Definition for Virtual Memory Addressing}
+\label{sec:overly-restrictive}
 One could give a naive logical definition to the virtual mememory addressing as conjuctions of each level translation 
@@
   One might have already observed that the virtual points-to definition shown in Figure \ref{fig:strongvirtualpointsto} 
-  is too strong to specify some operations updating a page table entry.
+  is too strong to specify some operations a virtual memory manager may need to do, such as
+  move one level of the page table to a different physical location while preserving all virtual-to-physical mappings.\footnote{
+    x86-64 hardware, like other architectures, includes a feature (which we do not formalize assertions for) to
+    replace an L1 page table address in an L2 entry with a pointer to a \emph{larger} 2MB page (called super-pages), 
+    or replace an L2 page table address in an L3 entry with a pointer to a 1GB page (called huge-pages).
+  }
    To speak more concretely, the fractional permissions on the page-table-walk assertion ($\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$ in Figure \ref{fig:strongvirtualpointsto}) 
@@
 \begin{array}{l}
-    \vaddr\mapsto_{\textsf{v,rtv}}\{\textsf{q}\}\vpage \stackrel{\triangle}{:=} 
+    \vaddr\mapsto_{\textsf{v,rtv}}\{\textsf{q}\}\vpage \stackrel{\triangle}{=} 
     \exists \paddr\ldotp
@@
 \]
+\todo[inline]{Colin says: again, this is supposed to be a vprop, but isn't doing anything with a cr3val argument.
+I realize the GASI and per-address-space invariants, and therefore machinery to tie the virtual points-to
+to the GASI, hasn't been introduced yet. Is there a short-hand here we can use to be ``more complete'' until the later text
+that fills in these details?
+}
+\todo[inline,color=cyan]{Colin says: Oh!!! I see. You have that extremely subtle rtv subscript on the virtual points-to
+syntax! Don't do that. First, it's really subtle and wasn't explained anywhere, so even I missed it. Second,
+we're very explicitly describing vProps as functions from root to iProp. Sticking the rtv on the arrow there
+is partly weird because it obscures which argument is really the vProp argument (there are 2 other word64s in that syntax),
+partly weird because it's written next to something that \emph{isn't} a variable (the v subscript on the virtual points-to
+arrow) while using the same font, and partly weird because when you actually write these assertions, rtv doesn't show
+up in the syntax at all, which is the whole point --- virtual points-to defs shouldn't mention their cr3 or address space at
+all!
+Here's what I think it should look like (I also existentially quantified $\delta$ and fixed the p subscript on physical
+points-to):
+}
+\[
+\begin{array}{l}
+    \vaddr\mapsto_{\textsf{v}}\{\textsf{q}\}\vpage : \mathsf{vProp}~\Sigma \stackrel{\triangle}{=} 
+    \lambda \mathit{cr3val}\ldotp
+    \exists \paddr\ldotp
+    \exists \delta\ldotp
+    \sumapaces{\mathit{cr3val}}\delta \ast 
+  \sumwalkabs\vaddr\qfrac\paddr \ast \paddr \mapsto_{\mathsf{p}} \vpage
+\end{array}
+\]
+\todo[inline]{Colin says: The fact that we can't visually distinguish the variable v that is the value in memory
+from the subscript v on the virtual points-to arrow is not ideal: it makes it hard to tell what's concrete syntax
+and what's a variable. This is worth fixing, but near the bottom of the priority list}
 \caption{Virtual-Pointsto for Sharing Pages}
@@
 \end{figure*}
-  One possible approach is to abstract away the physical page-table-walk pointsto ($\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$ in Figure \ref{fig:strongvirtualpointsto}) from the virtual pointsto relation.
+
+  Intuitively, the definition in Figure \ref{fig:strongvirtualpointsto} is too strong because the virtual points-to
+  assertion there tracks too much information: when writing programs that access memory via virtual addresses,
+  most code does not care \emph{which physical memory locations are involved in address translation}: it only cares
+  that virtual address translation would succeed. The necessary information about the physical page table walk
+  must still be tracked, but can be tracked separately from the virtual points-to assertion itself.
+
+  We take the approach of separating the physical page-table-walk pointsto ($\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$ in Figure \ref{fig:strongvirtualpointsto}) 
+  from the virtual pointsto relation. 
+  To do so, we take advantage of an Iris construction which permits tracking a map in ghost state that maps
+  virtual addresses to the physical addresses they translate to, and has for each key in the map (i.e.,
+  each virtual address) a token. Updating the virtual-to-physical (ghost) mapping for a virtual address $va$
+  requires full ownership of the token. In addition, we keep an invariant that for each $va\mapsto pa$ mapping
+  in this map, there are \emph{physical} resources sufficient to ensure that address translation for $va$
+  will resolve in the hardware to $pa$. But these physical resources, which specify \emph{which} physical locations
+  are involved in the page table walk, are now separated from but consistent with
+  the knowledge that such resources exist (which is embodied by the token for $va$, which tracks that $va$ maps to $pa$
+  in the ghost map). Thus we can store the token which summarizes the translation and ensures it exists in the virtual
+  points-to, and keep the ghost map and the invariant that every mapping in the ghost map has corresponding physical resources
+  for translation in a separate global invariant.
+
+  \todo[inline]{Colin says: This requires explaining all the math here. Most reviewers of our paper will not be Iris
+  experts. Even the notion of an authoritative RA is something you'd need to explain (the general idea of).
+  So: write for non-Iris-expert readers. Above I explained the intuition behind the construction, so now we have to show
+  that we correctly implemented that intuition in Iris. So explain authoritative RAs, and how the line below
+  implements the intuition above.}
   To do so, we introduce the protocol separating read-only (fragmental) and authorative (full ownership) accesses to the page tables, specifically $\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$.
@@
 
+\todo[inline]{Colin says: I can rewrite this remark into flowing text after you fully explain the math above.}
 \begin{remark}[Exchanging Page-Table-Walk-Summarization Mappings (Tokens) under Mutation of Page Tables ]
@@
 \label{sec:aspacemanagement}
+\todo[inline]{Colin says: I will go back and add some foreshadowing in the explanation of virtual points-to,
+which already contains the Aspaces lookup a couple pages back.}
 So far, we have introduced logical abstractions for a single address space.

### 2023-04-12 examples.tex

@@
 \end{figure}
-In Figure \ref{fig:mapping_code}, we see
+In Figure \ref{fig:mapping_code}, we see an address ($\vaddr$) currently not mapped to a page ($\theta \; !!\; \vaddr = \texttt{None}$).
+The assurance of L1 table entry is provided by a page-worth \textit{virtual-pointsto} from an address (\texttt{pte\_addr}) stored in \textsf{rax} register (Specline 7) together with the assurance on the existence of L4, L3, and L2 table lookups (Speclines 4-6).
 %\todo[inline]{emphasize the write to the L1 page table is validated by the fact that the identity mapping exists}
-
 \subsection{Unmapping a Page}

### 2023-04-12 programlogic.tex
+

@@
  Definition $\pfpointsto\locsf\vpts\qfrac\ppts$ :vProp $\Sigma$ := pmapsto_def_v (pfn w) (page_off w) q v.
- Definition $\vaddr\mapsto_{\textsf{t}}\{\textsf{q}\}\vpage$ :vProp $\Sigma$ := 
+ Definition $\vaddr\mapsto_{\textsf{t}}\{\textsf{q}\}\; \vpage$ : vProp $\Sigma$ := 
   $\exists_{\textsf{l4e l3e l2e l1}} \ldotp$  $\ulcorner$ aligned $\vaddr \urcorner \ast$ L4_L1_PointsTo(rtv maddr l4e l3e l2e l1e) $\ast$ addr_L1$(\vaddr,\textsf{l1e}) \mapsto_{p} \vpage$.
@@
    \begin{lstlisting}[language=Coq]
-    $\hbox{(\TirNameStyle{{PageLabelAccess}})} \qquad$ addr_L1($\vaddr,\textsf{l1e}$) $\mapsto_{p} \vpage$.
+    $\hbox{(\TirNameStyle{{PageLevelAccess}})} \qquad$ addr_L1($\vaddr,\textsf{l1e}$) $\mapsto_{p} \vpage$.
     \end{lstlisting} 

### 2023-04-12 examples.tex
+

--- 2023-04-12 examples.tex
@@
   \begin{lstlisting}
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,\gammaPred,\rtv)  \ast  \texttt{r14}\mapsto_{r} \_ \ast \texttt{rdi}\mapsto_{r} \vaddr \ast \texttt{rax}\mapsto_{r} \_ \ast \ulcorner \texttt{aligned va} \land  \theta \; !!\; \vaddr = \texttt{None}\urcorner}_{\rtv}$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,\gammaPred)  \ast  \texttt{r14}\mapsto_{r} \_ \ast \texttt{rdi}\mapsto_{r} \vaddr \ast \texttt{rax}\mapsto_{r} \_ \ast \ulcorner \texttt{aligned va} \land  \theta \; !!\; \vaddr = \texttt{None}\urcorner}_{\rtv}$
 call ensure_L1_page
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,\gammaPred),\rtv \ast  \texttt{r14}\mapsto_{r} \_ \ast \texttt{rdi}\mapsto_{r} \vaddr \ast \texttt{rax}\mapsto_{r} \_ \; \ast }_{\rtv}$
-$\specline{\exists (\entryf ,\;\entrytr,\; \entrytw,\; \entryo: \texttt{word 64}) \ldotp \ulcorner (\texttt{addr\_L1 }\vaddr) = \entryo\urcorner \; \ast }_{\rtv}$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,\gammaPred) \ast  \texttt{r14}\mapsto_{r} \_ \ast \texttt{rdi}\mapsto_{r} \vaddr \ast \texttt{rax}\mapsto_{r} \textsf{ pte\_addr} \; \ast }_{\rtv}$
+$\specline{\exists (\entryf ,\;\entrytr,\; \entrytw,\; \entryo,\;\textsf{pte\_addr },\paddr \; : \texttt{word 64}) \ldotp \ulcorner \texttt{addr\_L1 }(\vaddr, \entryo) = \paddr \urcorner \; \ast }_{\rtv}$
 $\specline{\nfpointsto{\mask\vaddr\maskfour\rtv}{\mask\vaddr\maskfouroff\rtv}\entryf\qone\naddr \; \ast}_{\rtv}$ 
-$\specline{  \nfpointsto{\mask\vaddr\maskthree\entryf}{\mask\vaddr\maskthreeoff\entryf}\entrytr\qtwo\naddr \ast \nfpointsto{\mask\vaddr\masktwo\entrytr}{\mask\vaddr\masktwooff\entrytr}\entrytw\qthree\naddr \;\ast}_{\rtv}$
-$\specline{\texttt{pte\_addr} \mapsto_{\texttt{vpte},\rtv} \entryo \;(\texttt{wzero 64}) \ast \texttt{rax}\mapsto_{r} \texttt{pte\_addr}  }_{\rtv}$
+$\specline{  \nfpointsto{\mask\vaddr\maskthree\entryf}{\mask\vaddr\maskthreeoff\entryf}\entrytr\qtwo\naddr \ast \nfpointsto{\mask\vaddr\masktwo\entrytr}{\mask\vaddr\masktwooff\entrytr}\entryo\qthree\naddr \;\ast}_{\rtv}$
+$\specline{\texttt{pte\_addr} \mapsto_{\texttt{vpte}} \paddr \;(\texttt{wzero 64}) \ast \texttt{rax}\mapsto_{r} \texttt{pte\_addr}  }_{\rtv}$
 ;; Returns the virtual address of the L1 entry in rax
 mov %r14, %rax ;; Save that before another call
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,\gammaPred,\rtv) \ast  \texttt{r14}\mapsto_{r} \texttt{pte\_addr} \ast \texttt{rdi}\mapsto_{r} \vaddr \ast \texttt{rax}\mapsto_{r} \_ \; \ast }_{\rtv}$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,\gammaPred) \ast  \texttt{r14}\mapsto_{r} \texttt{pte\_addr} \ast \texttt{rdi}\mapsto_{r} \vaddr \ast \texttt{rax}\mapsto_{r} \textsf{ pte\_addr} \; \ast }_{\rtv}$
 $\specline{ \nfpointsto{\mask\vaddr\maskfour\rtv}{\mask\vaddr\maskfouroff\rtv}\entryf\qone\naddr \; \ast}_{\rtv}$ 
-$\specline{  \nfpointsto{\mask\vaddr\maskthree\entryf}{\mask\vaddr\maskthreeoff\entryf}\entrytr\qtwo\naddr \ast \nfpointsto{\mask\vaddr\masktwo\entrytr}{\mask\vaddr\masktwooff\entrytr}\entrytw\qthree\naddr \;\ast}_{\rtv}$
-$\specline{\texttt{pte\_addr} \mapsto_{\texttt{vpte},\rtv} \entryo \;(\texttt{wzero 64}) \ast \texttt{rax}\mapsto_{r} \texttt{pte\_addr}  }_{\rtv}$
+$\specline{  \nfpointsto{\mask\vaddr\maskthree\entryf}{\mask\vaddr\maskthreeoff\entryf}\entrytr\qtwo\naddr \ast \nfpointsto{\mask\vaddr\masktwo\entrytr}{\mask\vaddr\masktwooff\entrytr}\entryo\qthree\naddr \;\ast}_{\rtv}$
+$\specline{\texttt{pte\_addr} \mapsto_{\texttt{vpte}} \paddr \;(\texttt{wzero 64}) \ast \texttt{rax}\mapsto_{r} \texttt{pte\_addr}  }_{\rtv}$
 call alloc_phys_page_or_panic
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,\gammaPred,\rtv) \ast  \texttt{r14}\mapsto_{r} \texttt{pte\_addr} \ast \texttt{rdi}\mapsto_{r} \vaddr \ast \texttt{rax}\mapsto_{r} \_ \;\ast}_{\rtv}$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,\gammaPred) \ast  \texttt{r14}\mapsto_{r} \texttt{pte\_addr} \ast \texttt{rdi}\mapsto_{r} \vaddr \;\ast}_{\rtv}$
 $\specline{ \nfpointsto{\mask\vaddr\maskfour\rtv}{\mask\vaddr\maskfouroff\rtv}\entryf\qone\naddr \ast}_{\rtv}$ 
-$\specline{  \nfpointsto{\mask\vaddr\maskthree\entryf}{\mask\vaddr\maskthreeoff\entryf}\entrytr\qtwo\naddr \ast \nfpointsto{\mask\vaddr\masktwo\entrytr}{\mask\vaddr\masktwooff\entrytr}\entrytw\qthree\naddr \;\ast}_{\rtv}$
-$\specline{\texttt{pte\_addr} \mapsto_{\texttt{vpte},\rtv} \entryo \;(\texttt{wzero 64}) \ast \texttt{rax}\mapsto_{r} \texttt{pte\_addr} \; \ast }_{\rtv}$
+$\specline{  \nfpointsto{\mask\vaddr\maskthree\entryf}{\mask\vaddr\maskthreeoff\entryf}\entrytr\qtwo\naddr \ast \nfpointsto{\mask\vaddr\masktwo\entrytr}{\mask\vaddr\masktwooff\entrytr}\entryo\qthree\naddr \;\ast}_{\rtv}$
+$\specline{\texttt{pte\_addr} \mapsto_{\texttt{vpte}} \paddr \;(\texttt{wzero 64}) \; \ast }_{\rtv}$
 $\specline{\exists \texttt{ fpaddr} \ldotp \ulcorner \texttt{aligned fpaddr} \urcorner \ast \texttt{rax}\mapsto_{r} \texttt{fpaddr} \ast \texttt{fpaddr} \mapsto_{a} (\texttt{wzero 64})}_{\rtv}$
@@
 mov (%r14), %rax
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,\gammaPred,\rtv) \ast  \texttt{r14}\mapsto_{r} \texttt{pte\_addr} \ast \texttt{rdi}\mapsto_{r} \vaddr \ast \texttt{rax}\mapsto_{r} \_ \;\ast}_{\rtv}$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,\gammaPred) \ast  \texttt{r14}\mapsto_{r} \texttt{pte\_addr} \ast \texttt{rdi}\mapsto_{r} \vaddr \ast \texttt{rax}\mapsto_{r} \_ \;\ast}_{\rtv}$
 $\specline{\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\vaddr,\rtv,\entryf,\entrytr,\entrytw,\entryo) \ast \ulcorner \theta \;!!\; \texttt{maddr} = \texttt{None}\urcorner \; \ast}_{\rtv}$
@@
 $\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\; \sqsubseteq $
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta<[\vaddr:=\texttt{fpaddr}]>),\gammaPred,\rtv) \ast  \texttt{r14}\mapsto_{r} \texttt{pte\_addr} \ast \texttt{rdi}\mapsto_{r} \vaddr \ast \texttt{rax}\mapsto_{r} \_ \; \ast}_{\rtv}$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\<[\vaddr:=\texttt{fpaddr}]>\theta),\gammaPred) \ast  \texttt{r14}\mapsto_{r} \texttt{pte\_addr} \ast \texttt{rdi}\mapsto_{r} \vaddr \ast \texttt{rax}\mapsto_{r} \_ \; \ast}_{\rtv}$
 $\specline{\ulcorner \texttt{aligned fpaddr} \urcorner \ast \texttt{rax}\mapsto_{r} \texttt{fpaddr} \ast \texttt{aligned va} \ast\sumwalkabs\vaddr\qfrac\fpaddr}_{\rtv}$
@@
 \end{figure}
-In Figure \ref{fig:mapping_code}, we see
+In Figure \ref{fig:mapping_code}, we see an address ($\vaddr$) currently not mapped to a page ($\theta \; !!\; \vaddr = \texttt{None}$).
+Mapping a fresh page backing this virtual first requires an assurance on the existence L1 table entry which is realized in two folds:
+\begin{itemize}
+\item physical pointsto assertions on the tables L4, L3, and L2 reaching to the L1 level entry (l1e) ensures the successful table lookups (Specification Lines 4-6)
+\item ,and a virtual L1 pointsto ($\mapsto_{\textsf{vpte}}$) on the address computed from the reached L1 entry ($ \ulcorner \texttt{addr\_L1}(\vaddr,\entryo) = \paddr\urcorner$) which is almost exactly same with the virtual-pointsto definition in Figure \ref{fig:virtualpointstosharing} except the physical page address is already computed, not exsitentially quantified:
+ \[
+\begin{array}{l}
+    \vaddr\mapsto_{\textsf{vpte}} \;\{\textsf{q}\} \; \paddr \; \vpage : \mathsf{vProp}~\Sigma \stackrel{\triangle}{=} 
+    \lambda \mathit{cr3val}\ldotp
+    \exists \delta\ldotp
+    \sumapaces{\mathit{cr3val}}\delta \ast 
+  \sumwalkabs\vaddr\qfrac\paddr \ast \paddr \mapsto_{\mathsf{p}} \vpage
+\end{array}
+\]
+\end{itemize}
+After obtaining an virtual address \textsf{pte\_addr} in \textsf{rax} referring to a physical address, computed from already known L4-L2 physical memory pointstos ($\texttt{addr\_L1}(\vaddr,\entryo) = \paddr$),
+in Specification Line 7 ($\texttt{pte\_addr} \mapsto_{\texttt{vpte}}  \; \paddr\;(\texttt{wzero 64})$), we save it to \textsf{r14} to be updated later in Line 9.
+
+Then we allocate a fresh page initialized (i.e. zeroed) in Line 14 which return an \textsf{aligned} address of fresh page (\textsf{fpaddr}) in \textsf{rax} (Specification Line 19).
+
+Therefore, when we unfold the $\texttt{pte\_addr} \mapsto_{\texttt{vpte}}  \; \paddr\;(\texttt{wzero 64})$ we will be able obtain a full ownerhsip on the L1 table physical pointsto $\paddr \mapsto_{\mathsf{p}} \textsf{wzero 64}$ which can be split to obtain a fractional ownership (q4) on it ($\paddr \mapsto_{\mathsf{p}} \{q4\} \;\textsf{wzero 64}$) to connect it with the L4-L2 physical table pointstos (Specification Lines 4-6) to obtain the complete physical table-walk assertion (L4\_L1\_PointsTo(rtv maddr l4e l3e l2e pa)).
+
+After storing $\textsf{pte\_addr}$
 %\todo[inline]{emphasize the write to the L1 page table is validated by the fact that the identity mapping exists}
-
 \subsection{Unmapping a Page}
@@
 A critical piece of \emph{trusted} code in verified OS kernels is the assembly code to change the current address space; current verified OS kernels currently lack effective ways to specify and reason about this low-level operation, for resaons outlined in Section \ref{sec:relwork}.
-
 \todo[inline]{probably not much more than just the cr3 update, but could be extended to a full context switch since the instruction pointer isn't updated in a swtch}
-
 \begin{figure}\footnotesize
@@
 ;; Save yielding context
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,\gammaPred,\rtv) \ast [\rtv'+56](\mathcal{I}\texttt{ASpace}(\theta,\gammaPred') \ast \texttt{Pother})}_{\rtv}$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,\gammaPred) \ast [\rtv'+56](\mathcal{I}\texttt{ASpace}(\theta,\gammaPred') \ast \texttt{Pother})}_{\rtv}$
 $\specline{ \texttt{rsi}\mapsto_{r} \rtv' \ast \texttt{rdi}\mapsto_{r} \_ \ast \texttt{rbx}\mapsto_{r} \texttt{rbxv} \ast  \texttt{rsp}\mapsto_{r} \texttt{rspv} \ast \texttt{rbp}\mapsto_{r} \texttt{rbpv} }_{\rtv}$
 $\specline{\texttt{r12}\mapsto_{r} \texttt{r12v} \ast \texttt{r13}\mapsto_{r} \texttt{r13v} \ast \texttt{r14}\mapsto_{r} \texttt{r14v} \ast \texttt{r15}\mapsto_{r} \texttt{r15v}}_{\rtv}$
-$\specline{\texttt{rdi+0} \mapsto_{v,\rtv} \_ \ast \texttt{rdi+8} \mapsto_{v,\rtv} \_ \ast \texttt{rdi+16} \mapsto_{v,\rtv} \_ \ast \texttt{rdi+24} \mapsto_{v,\rtv} \_ \ast \texttt{rdi+32} \mapsto_{v,\rtv} \_}_{\rtv}$
-$\specline{\texttt{rdi+40} \mapsto_{v,\rtv} \_ \ast \texttt{rdi+48} \mapsto_{v,\rtv} \_\ast \texttt{rdi+56} \mapsto_{v,\rtv} \_}_{\rtv}$
+$\specline{\texttt{rdi+0} \mapsto_{v} \_ \ast \texttt{rdi+8} \mapsto_{v} \_ \ast \texttt{rdi+16} \mapsto_{v} \_ \ast \texttt{rdi+24} \mapsto_{v} \_ \ast \texttt{rdi+32} \mapsto_{v} \_}_{\rtv}$
+$\specline{\texttt{rdi+40} \mapsto_{v} \_ \ast \texttt{rdi+48} \mapsto_{v} \_\ast \texttt{rdi+56} \mapsto_{v} \_}_{\rtv}$
 mov 0[%rdi], %rbx
@@
 mov 48[%rdi], %r15
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,\gammaPred,\rtv) \ast [\rtv'+56](\mathcal{I}\texttt{ASpace}(\theta,\gammaPred') \ast \texttt{Pother})}_{\rtv}$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,\gammaPred) \ast [\rtv'+56](\mathcal{I}\texttt{ASpace}(\theta,\gammaPred') \ast \texttt{Pother})}_{\rtv}$
 $\specline{ \texttt{rdi}\mapsto_{r} \_ \ast \texttt{rbx}\mapsto_{r} \texttt{rbxv} \ast  \texttt{rsp}\mapsto_{r} \texttt{rspv} \ast \texttt{rbp}\mapsto_{r} \texttt{rbpv} }_{\rtv}$
 $\specline{\texttt{r12}\mapsto_{r} \texttt{r12v} \ast \texttt{r13}\mapsto_{r} \texttt{r13v} \ast \texttt{r14}\mapsto_{r} \texttt{r14v} \ast \texttt{r15}\mapsto_{r} \texttt{r15v}}_{\rtv}$
-$\specline{\texttt{rdi+0} \mapsto_{v,\rtv} \texttt{rbxv} \ast \texttt{rdi+8} \mapsto_{v,\rtv} \texttt{rspv} \ast \texttt{rdi+16} \mapsto_{v,\rtv} \texttt{rbpv} \ast \texttt{rdi+24} \mapsto_{v,\rtv} \texttt{r12v} \ast}_{\rtv}$
-$\specline{ \texttt{rdi+32} \mapsto_{v,\rtv} \texttt{r13v} \ast \texttt{rdi+40} \mapsto_{v,\rtv} \texttt{r14v} \ast \texttt{rdi+48} \mapsto_{v,\rtv} \texttt{r15v}\ast \texttt{rdi+56} \mapsto_{v,\rtv} \_}_{\rtv}$
+$\specline{\texttt{rdi+0} \mapsto_{v} \texttt{rbxv} \ast \texttt{rdi+8} \mapsto_{v} \texttt{rspv} \ast \texttt{rdi+16} \mapsto_{v} \texttt{rbpv} \ast \texttt{rdi+24} \mapsto_{v} \texttt{r12v} \ast}_{\rtv}$
+$\specline{ \texttt{rdi+32} \mapsto_{v} \texttt{r13v} \ast \texttt{rdi+40} \mapsto_{v} \texttt{r14v} \ast \texttt{rdi+48} \mapsto_{v} \texttt{r15v}\ast \texttt{rdi+56} \mapsto_{v} \_}_{\rtv}$
 mov 56[%rdi], %cr3
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,\gammaPred,\rtv) \ast [\rtv'+56](\mathcal{I}\texttt{ASpace}(\theta,\gammaPred') \ast \texttt{Pother})}_{\rtv}$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,\gammaPred) \ast [\rtv'+56](\mathcal{I}\texttt{ASpace}(\theta,\gammaPred') \ast \texttt{Pother})}_{\rtv}$
 $\specline{ \texttt{rdi}\mapsto_{r} \_ \ast \texttt{rbx}\mapsto_{r} \texttt{rbxv} \ast  \texttt{rsp}\mapsto_{r} \texttt{rspv} \ast \texttt{rbp}\mapsto_{r} \texttt{rbpv} }_{\rtv}$
 $\specline{\texttt{r12}\mapsto_{r} \texttt{r12v} \ast \texttt{r13}\mapsto_{r} \texttt{r13v} \ast \texttt{r14}\mapsto_{r} \texttt{r14v} \ast \texttt{r15}\mapsto_{r} \texttt{r15v}}_{\rtv}$
-$\specline{\texttt{rdi+0} \mapsto_{v,\rtv} \texttt{rbxv} \ast \texttt{rdi+8} \mapsto_{v,\rtv} \texttt{rspv} \ast \texttt{rdi+16} \mapsto_{v,\rtv} \texttt{rbpv} \ast \texttt{rdi+24} \mapsto_{v,\rtv} \texttt{r12v} \ast }_{\rtv}$
-$\specline{\texttt{rdi+32} \mapsto_{v,\rtv} \texttt{r13v} \ast \texttt{rdi+40} \mapsto_{v,\rtv} \texttt{r14v} \ast \texttt{rdi+48} \mapsto_{v,\rtv} \texttt{r15v}\ast \texttt{rdi+56} \mapsto_{v,\rtv} \rtv}_{\rtv}$
+$\specline{\texttt{rdi+0} \mapsto_{v} \texttt{rbxv} \ast \texttt{rdi+8} \mapsto_{v} \texttt{rspv} \ast \texttt{rdi+16} \mapsto_{v} \texttt{rbpv} \ast \texttt{rdi+24} \mapsto_{v} \texttt{r12v} \ast }_{\rtv}$
+$\specline{\texttt{rdi+32} \mapsto_{v} \texttt{r13v} \ast \texttt{rdi+40} \mapsto_{v} \texttt{r14v} \ast \texttt{rdi+48} \mapsto_{v} \texttt{r15v}\ast \texttt{rdi+56} \mapsto_{v} \rtv}_{\rtv}$    
 ;; Restore target context
@@
 $\specline{\texttt{r12}\mapsto_{r} \rtv'+24 \ast \texttt{r13}\mapsto_{r} \rtv'+32 \ast \texttt{r14}\mapsto_{r} \rtv'+40 \ast \texttt{r15}\mapsto_{r} \rtv'+48}_{\rtv}$
-$\specline{\texttt{rdi+0} \mapsto_{v,\rtv} \texttt{rbxv} \ast \texttt{rdi+8} \mapsto_{v,\rtv} \texttt{rspv} \ast \texttt{rdi+16} \mapsto_{v,\rtv} \texttt{rbpv} \ast \texttt{rdi+24} \mapsto_{v,\rtv} \texttt{r12v} \ast }_{\rtv}$
-$\specline{\texttt{rdi+32} \mapsto_{v,\rtv} \texttt{r13v} \ast \texttt{rdi+40} \mapsto_{v,\rtv} \texttt{r14v} \ast \texttt{rdi+48} \mapsto_{v,\rtv} \texttt{r15v}\ast \texttt{rdi+56} \mapsto_{v,\rtv} \rtv}_{\rtv}$
+$\specline{\texttt{rdi+0} \mapsto_{v} \texttt{rbxv} \ast \texttt{rdi+8} \mapsto_{v} \texttt{rspv} \ast \texttt{rdi+16} \mapsto_{v} \texttt{rbpv} \ast \texttt{rdi+24} \mapsto_{v} \texttt{r12v} \ast }_{\rtv}$
+$\specline{\texttt{rdi+32} \mapsto_{v} \texttt{r13v} \ast \texttt{rdi+40} \mapsto_{v} \texttt{r14v} \ast \texttt{rdi+48} \mapsto_{v} \texttt{r15v}\ast \texttt{rdi+56} \mapsto_{v} \rtv}_{\rtv}$
 ;; Switch to the new address space
@@
 $\specline{\texttt{r12}\mapsto_{r} \rtv'+24 \ast \texttt{r13}\mapsto_{r} \rtv'+32 \ast \texttt{r14}\mapsto_{r} \rtv'+40 \ast \texttt{r15}\mapsto_{r} \rtv'+48}_{\rtv}$
-$\specline{\texttt{rdi+0} \mapsto_{v,\rtv} \texttt{rbxv} \ast \texttt{rdi+8} \mapsto_{v,\rtv} \texttt{rspv} \ast \texttt{rdi+16} \mapsto_{v,\rtv} \texttt{rbpv} \ast  \texttt{rdi+24} \mapsto_{v,\rtv} \texttt{r12v} \ast }_{\rtv}$
-$\specline{\texttt{rdi+32} \mapsto_{v,\rtv} \texttt{r13v} \ast \texttt{rdi+40} \mapsto_{v,\rtv} \texttt{r14v} \ast \texttt{rdi+48} \mapsto_{v,\rtv} \texttt{r15v}\ast \texttt{rdi+56} \mapsto_{v,\rtv} \rtv}_{\rtv}$
+$\specline{\texttt{rdi+0} \mapsto_{v} \texttt{rbxv} \ast \texttt{rdi+8} \mapsto_{v} \texttt{rspv} \ast \texttt{rdi+16} \mapsto_{v} \texttt{rbpv} \ast  \texttt{rdi+24} \mapsto_{v} \texttt{r12v} \ast }_{\rtv}$
+$\specline{\texttt{rdi+32} \mapsto_{v} \texttt{r13v} \ast \texttt{rdi+40} \mapsto_{v} \texttt{r14v} \ast \texttt{rdi+48} \mapsto_{v} \texttt{r15v}\ast \texttt{rdi+56} \mapsto_{v} \rtv}_{\rtv}$
 $\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\; \sqsubseteq $

### 2023-04-12 examples.tex
+

@@
 $\specline{\nfpointsto{\mask\vaddr\maskfour\rtv}{\mask\vaddr\maskfouroff\rtv}\entryf\qone\naddr \; \ast}_{\rtv}$ 
-$\specline{  \nfpointsto{\mask\vaddr\maskthree\entryf}{\mask\vaddr\maskthreeoff\entryf}\entrytr\qtwo\naddr \ast \nfpointsto{\mask\vaddr\masktwo\entrytr}{\mask\vaddr\masktwooff\entrytr}\entryo\qthree\naddr \;\ast}_{\rtv}$
+$\specline{  \nfpointsto{\mask\vaddr\maskthree\entryf}{\mask\vaddr\maskthreeoff\entryf}\entrytr\qtwo\naddr \ast \nfpointsto{\mask\vaddr\masktwo\entrytr}{\mask\vaddr\masktwooff\entrytr}\paddr\qthree\naddr \;\ast}_{\rtv}$
 $\specline{\texttt{pte\_addr} \mapsto_{\texttt{vpte}} \paddr \;(\texttt{wzero 64}) \ast \texttt{rax}\mapsto_{r} \texttt{pte\_addr}  }_{\rtv}$
@@
 $\specline{ \nfpointsto{\mask\vaddr\maskfour\rtv}{\mask\vaddr\maskfouroff\rtv}\entryf\qone\naddr \; \ast}_{\rtv}$ 
-$\specline{  \nfpointsto{\mask\vaddr\maskthree\entryf}{\mask\vaddr\maskthreeoff\entryf}\entrytr\qtwo\naddr \ast \nfpointsto{\mask\vaddr\masktwo\entrytr}{\mask\vaddr\masktwooff\entrytr}\entryo\qthree\naddr \;\ast}_{\rtv}$
+$\specline{  \nfpointsto{\mask\vaddr\maskthree\entryf}{\mask\vaddr\maskthreeoff\entryf}\entrytr\qtwo\naddr \ast \nfpointsto{\mask\vaddr\masktwo\entrytr}{\mask\vaddr\masktwooff\entrytr}\paddr\qthree\naddr \;\ast}_{\rtv}$
 $\specline{\texttt{pte\_addr} \mapsto_{\texttt{vpte}} \paddr \;(\texttt{wzero 64}) \ast \texttt{rax}\mapsto_{r} \texttt{pte\_addr}  }_{\rtv}$
@@
 $\specline{ \nfpointsto{\mask\vaddr\maskfour\rtv}{\mask\vaddr\maskfouroff\rtv}\entryf\qone\naddr \ast}_{\rtv}$ 
-$\specline{  \nfpointsto{\mask\vaddr\maskthree\entryf}{\mask\vaddr\maskthreeoff\entryf}\entrytr\qtwo\naddr \ast \nfpointsto{\mask\vaddr\masktwo\entrytr}{\mask\vaddr\masktwooff\entrytr}\entryo\qthree\naddr \;\ast}_{\rtv}$
+$\specline{  \nfpointsto{\mask\vaddr\maskthree\entryf}{\mask\vaddr\maskthreeoff\entryf}\entrytr\qtwo\naddr \ast \nfpointsto{\mask\vaddr\masktwo\entrytr}{\mask\vaddr\masktwooff\entrytr}\paddr\qthree\naddr \;\ast}_{\rtv}$
 $\specline{\texttt{pte\_addr} \mapsto_{\texttt{vpte}} \paddr \;(\texttt{wzero 64}) \; \ast }_{\rtv}$
 $\specline{\exists \texttt{ fpaddr} \ldotp \ulcorner \texttt{aligned fpaddr} \urcorner \ast \texttt{rax}\mapsto_{r} \texttt{fpaddr} \ast \texttt{fpaddr} \mapsto_{a} (\texttt{wzero 64})}_{\rtv}$
-;; Calculate new PTE by setting the low-order bits
-;; of the physical page
-add %rax, 3
+;; Calculate new L1 entry
 ;; update the page table entry, mapping the page
@@
 
-Therefore, when we unfold the $\texttt{pte\_addr} \mapsto_{\texttt{vpte}}  \; \paddr\;(\texttt{wzero 64})$ we will be able obtain a full ownerhsip on the L1 table physical pointsto $\paddr \mapsto_{\mathsf{p}} \textsf{wzero 64}$ which can be split to obtain a fractional ownership (q4) on it ($\paddr \mapsto_{\mathsf{p}} \{q4\} \;\textsf{wzero 64}$) to connect it with the L4-L2 physical table pointstos (Specification Lines 4-6) to obtain the complete physical table-walk assertion (L4\_L1\_PointsTo(rtv maddr l4e l3e l2e pa)).
+\todo{ISMAIL STOP}
+The crux point in the proof is updating the L1 entry in Line 22 Therefore, when we unfold the $\texttt{pte\_addr} \mapsto_{\texttt{vpte}}  \; \paddr\;(\texttt{wzero 64})$ we will be able obtain a full ownerhsip on the L1 table physical pointsto $\paddr \mapsto_{\mathsf{p}} \textsf{wzero 64}$ which can be split to obtain a fractional ownership (q4) on it ($\paddr \mapsto_{\mathsf{p}} \{q4\} \;\textsf{wzero 64}$) to connect it with the L4-L2 physical table pointstos (Specification Lines 4-6) to obtain the complete physical table-walk assertion (L4\_L1\_PointsTo(rtv maddr l4e l3e l2e pa)).
 

### 2023-04-12 examples.tex
+

@@
 
-\todo{ISMAIL STOP}
-The crux point in the proof is updating the L1 entry in Line 22 Therefore, when we unfold the $\texttt{pte\_addr} \mapsto_{\texttt{vpte}}  \; \paddr\;(\texttt{wzero 64})$ we will be able obtain a full ownerhsip on the L1 table physical pointsto $\paddr \mapsto_{\mathsf{p}} \textsf{wzero 64}$ which can be split to obtain a fractional ownership (q4) on it ($\paddr \mapsto_{\mathsf{p}} \{q4\} \;\textsf{wzero 64}$) to connect it with the L4-L2 physical table pointstos (Specification Lines 4-6) to obtain the complete physical table-walk assertion (L4\_L1\_PointsTo(rtv maddr l4e l3e l2e pa)).
+The crux point in the proof is updating the virtual L1 address (\textsf{pte\_addr}) to show the fresly allocated page address (\textsf{fpaddr}) in Line 22 to obtain
+\[$\texttt{pte\_addr} \mapsto_{\texttt{vpte}}  \; \paddr\; \textsf{fpaddr}$\]
+so that when we unfold the $\texttt{pte\_addr} \mapsto_{\texttt{vpte}}  \; \paddr\; \textsf{fpaddr})$ we will be able obtain a full ownerhsip on the L1 table physical pointsto $\paddr \mapsto_{\mathsf{p}} \textsf{fpaddr}$ .
+Then, we can split this full-ownership into to obtain a fractional ownership (q4) on it ($\paddr \mapsto_{\mathsf{p}} \{q4\} \;\textsf{fpaddr}$).
+Finally, we can connect this fragmental ownership with the L4-L2 physical table pointstos (Specification Lines 4-6) to obtain the complete physical table-walk assertion (L4\_L1\_PointsTo(rtv maddr l4e l3e l2e pa)).
 

### 2023-04-12 examples.tex
+

@@
 mov (%r14), %rax
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,\gammaPred) \ast  \texttt{r14}\mapsto_{r} \texttt{pte\_addr} \ast \texttt{rdi}\mapsto_{r} \vaddr \ast \texttt{rax}\mapsto_{r} \_ \;\ast}_{\rtv}$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,\gammaPred) \ast  \texttt{r14}\mapsto_{r} \texttt{pte\_addr} \ast \texttt{rdi}\mapsto_{r} \vaddr \ast }_{\rtv}$
 $\specline{\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\vaddr,\rtv,\entryf,\entrytr,\entrytw,\entryo) \ast \ulcorner \theta \;!!\; \texttt{maddr} = \texttt{None}\urcorner \; \ast}_{\rtv}$
@@
 $\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\; \sqsubseteq $
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\<[\vaddr:=\texttt{fpaddr}]>\theta),\gammaPred) \ast  \texttt{r14}\mapsto_{r} \texttt{pte\_addr} \ast \texttt{rdi}\mapsto_{r} \vaddr \ast \texttt{rax}\mapsto_{r} \_ \; \ast}_{\rtv}$
-$\specline{\ulcorner \texttt{aligned fpaddr} \urcorner \ast \texttt{rax}\mapsto_{r} \texttt{fpaddr} \ast \texttt{aligned va} \ast\sumwalkabs\vaddr\qfrac\fpaddr}_{\rtv}$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace} (<[\vaddr:=\texttt{fpaddr}]> \theta,\gammaPred) \ast  \texttt{r14}\mapsto_{r} \texttt{pte\_addr} \ast \texttt{rdi}\mapsto_{r} \vaddr \; \ast}_{\rtv}$
+$\specline{\ulcorner \texttt{aligned fpaddr} \urcorner \ast \texttt{rax}\mapsto_{r} \texttt{fpaddr} \ast \texttt{aligned va} \ast\sumwalkabs\vaddr\qfrac\textsf{fpaddr}}_{\rtv}$
 \end{lstlisting}
@@
 After obtaining an virtual address \textsf{pte\_addr} in \textsf{rax} referring to a physical address, computed from already known L4-L2 physical memory pointstos ($\texttt{addr\_L1}(\vaddr,\entryo) = \paddr$),
-in Specification Line 7 ($\texttt{pte\_addr} \mapsto_{\texttt{vpte}}  \; \paddr\;(\texttt{wzero 64})$), we save it to \textsf{r14} to be updated later in Line 9.
+in Specification Line 7 ($\texttt{pte\_addr} \mapsto_{\textsf{vpte}}  \; \paddr\;(\texttt{wzero 64})$), we save it to \textsf{r14} to be updated later in Line 9.
 
@@
 The crux point in the proof is updating the virtual L1 address (\textsf{pte\_addr}) to show the fresly allocated page address (\textsf{fpaddr}) in Line 22 to obtain
-\[$\texttt{pte\_addr} \mapsto_{\texttt{vpte}}  \; \paddr\; \textsf{fpaddr}$\]
-so that when we unfold the $\texttt{pte\_addr} \mapsto_{\texttt{vpte}}  \; \paddr\; \textsf{fpaddr})$ we will be able obtain a full ownerhsip on the L1 table physical pointsto $\paddr \mapsto_{\mathsf{p}} \textsf{fpaddr}$ .
+\[\texttt{pte\_addr} \mapsto_{\texttt{vpte}}  \; \paddr\; \textsf{fpaddr}  \qquad\qquad (\textsf{by Rule} \TirNameStyle{WriteToVirtMemFromReg})\]
+by Rule \TirNameStyle{WriteToVirtMemFromReg} which allows proving virtual address updates without exposing its internal physical page-table memory accesses. 
+Consequently, right at this point in the proof, when we unfold the $\texttt{pte\_addr} \mapsto_{\texttt{vpte}}  \; \paddr\; \textsf{fpaddr})$ we will be able obtain a full ownerhsip on the L1 table physical pointsto $\paddr \mapsto_{\mathsf{p}} \textsf{fpaddr}$ .
 Then, we can split this full-ownership into to obtain a fractional ownership (q4) on it ($\paddr \mapsto_{\mathsf{p}} \{q4\} \;\textsf{fpaddr}$).

### 2023-04-12 examples.tex
+

@@
 mov (%r14), %rax
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,\gammaPred) \ast  \texttt{r14}\mapsto_{r} \texttt{pte\_addr} \ast \texttt{rdi}\mapsto_{r} \vaddr \;\ast}_{\rtv}$
+$\specline{ \nfpointsto{\mask\vaddr\maskfour\rtv}{\mask\vaddr\maskfouroff\rtv}\entryf\qone\naddr \ast}_{\rtv}$ 
+$\specline{  \nfpointsto{\mask\vaddr\maskthree\entryf}{\mask\vaddr\maskthreeoff\entryf}\entrytr\qtwo\naddr \ast \nfpointsto{\mask\vaddr\masktwo\entrytr}{\mask\vaddr\masktwooff\entrytr}\paddr\qthree\naddr \;\ast}_{\rtv}$
+$\specline{\texttt{pte\_addr} \mapsto_{\texttt{vpte}} \paddr \;(\texttt{fpaddr}) \; \ast }_{\rtv}$
+$\specline{\exists \texttt{ fpaddr} \ldotp \ulcorner \texttt{aligned fpaddr} \urcorner \ast \texttt{rax}\mapsto_{r} \texttt{fpaddr} \ast \texttt{fpaddr} \mapsto_{a} (\texttt{wzero 64})}_{\rtv}$
+$\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\; \sqsubseteq $
 $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,\gammaPred) \ast  \texttt{r14}\mapsto_{r} \texttt{pte\_addr} \ast \texttt{rdi}\mapsto_{r} \vaddr \ast }_{\rtv}$
-$\specline{\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\vaddr,\rtv,\entryf,\entrytr,\entrytw,\entryo) \ast \ulcorner \theta \;!!\; \texttt{maddr} = \texttt{None}\urcorner \; \ast}_{\rtv}$
-$\specline{\ulcorner \texttt{aligned fpaddr} \urcorner \ast \texttt{rax}\mapsto_{r} \texttt{fpaddr} \ast \texttt{fpaddr} \mapsto_{a} (\texttt{wzero 64})}_{\rtv}$
+$\specline{\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\vaddr,\entryf,\entrytr,\entrytw,\paddr) \ast \ulcorner \theta \;!!\;\vaddr = \texttt{None}\urcorner \; \ast}_{\rtv}$
+$\specline{\ulcorner \texttt{aligned fpaddr} \urcorner \ast \texttt{rax}\mapsto_{r} \texttt{fpaddr} \ast \texttt{fpaddr} \mapsto_{a} (\texttt{wzero 64}) }_{\rtv}$
 $\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\; \sqsubseteq $
 $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace} (<[\vaddr:=\texttt{fpaddr}]> \theta,\gammaPred) \ast  \texttt{r14}\mapsto_{r} \texttt{pte\_addr} \ast \texttt{rdi}\mapsto_{r} \vaddr \; \ast}_{\rtv}$
-$\specline{\ulcorner \texttt{aligned fpaddr} \urcorner \ast \texttt{rax}\mapsto_{r} \texttt{fpaddr} \ast \texttt{aligned va} \ast\sumwalkabs\vaddr\qfrac\textsf{fpaddr}}_{\rtv}$
+$\specline{\ulcorner \texttt{aligned fpaddr} \urcorner \ast \texttt{fpaddr} \mapsto_{\textsf{p}} \textsf{ wzero 64} \ast \sumapaces\rtv\delta  \ast\sumwalkabs\vaddr\qfrac\fpaddr}_{\rtv}$
+$\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\; \sqsubseteq $
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace} (<[\vaddr:=\texttt{fpaddr}]> \theta,\gammaPred) \ast \vaddr \mapsto_{\textsf{vpte}}\; \qfrac \;\fpaddr \textsf{ wzero 64}}_{\rtv}$
 \end{lstlisting}
@@
 \[\texttt{pte\_addr} \mapsto_{\texttt{vpte}}  \; \paddr\; \textsf{fpaddr}  \qquad\qquad (\textsf{by Rule} \TirNameStyle{WriteToVirtMemFromReg})\]
-by Rule \TirNameStyle{WriteToVirtMemFromReg} which allows proving virtual address updates without exposing its internal physical page-table memory accesses. 
+by Rule \TirNameStyle{WriteToVirtMemFromReg} which allows proving virtual address updates without exposing its internal physical page-table memory accesses (Specification Line 26). 
 Consequently, right at this point in the proof, when we unfold the $\texttt{pte\_addr} \mapsto_{\texttt{vpte}}  \; \paddr\; \textsf{fpaddr})$ we will be able obtain a full ownerhsip on the L1 table physical pointsto $\paddr \mapsto_{\mathsf{p}} \textsf{fpaddr}$ .
 Then, we can split this full-ownership into to obtain a fractional ownership (q4) on it ($\paddr \mapsto_{\mathsf{p}} \{q4\} \;\textsf{fpaddr}$).
-Finally, we can connect this fragmental ownership with the L4-L2 physical table pointstos (Specification Lines 4-6) to obtain the complete physical table-walk assertion (L4\_L1\_PointsTo(rtv maddr l4e l3e l2e pa)).
+Finally, we can connect this fragmental ownership with the L4-L2 physical table pointstos (Specification Lines 24 and 25) to obtain the complete physical table-walk assertion (L4\_L1\_PointsTo(maddr l4e l3e l2e pa)) in Line 30.
+
+Finally, we can insert $\vaddr$ into the ghost page-table-walk summarization map ($\theta$) to change its state from unmapped ($\ulcorner \theta \;!!\;\vaddr = \texttt{None}\urcorner$ in Specification Line 30) to mapped, and construct virtual-pointsto for $\sumwalkabs\vaddr\qfrac\fpaddr$ with fragmental ownership of
 

### 2023-04-12 examples.tex
+
done with Section 6.2

@@
 $\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\; \sqsubseteq $
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace} (<[\vaddr:=\texttt{fpaddr}]> \theta,\gammaPred) \ast  \texttt{r14}\mapsto_{r} \texttt{pte\_addr} \ast \texttt{rdi}\mapsto_{r} \vaddr \; \ast}_{\rtv}$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace} (<[\vaddr:=\texttt{fpaddr}]> \theta,\gammaPred) \ast}_{\rtv}$
 $\specline{\ulcorner \texttt{aligned fpaddr} \urcorner \ast \texttt{fpaddr} \mapsto_{\textsf{p}} \textsf{ wzero 64} \ast \sumapaces\rtv\delta  \ast\sumwalkabs\vaddr\qfrac\fpaddr}_{\rtv}$
@@
 
-Finally, we can insert $\vaddr$ into the ghost page-table-walk summarization map ($\theta$) to change its state from unmapped ($\ulcorner \theta \;!!\;\vaddr = \texttt{None}\urcorner$ in Specification Line 30) to mapped, and construct virtual-pointsto for $\sumwalkabs\vaddr\qfrac\fpaddr$ with fragmental ownership of
-
-After storing $\textsf{pte\_addr}$
-%\todo[inline]{emphasize the write to the L1 page table is validated by the fact that the identity mapping exists}
+Finally, we can insert $\vaddr$ into the ghost page-table-walk summarization map ($\theta$) to change its state from unmapped ($\ulcorner \theta \;!!\;\vaddr = \texttt{None}\urcorner$ in Specification Line 30) to mapped (Specification Line 33), and construct virtual-pointsto $\vaddr \mapsto_{\textsf{vpte}}\; \qfrac \;\fpaddr \textsf{ wzero 64}$ (Specification Line 36) with $\sumwalkabs\vaddr\qfrac\fpaddr$ obtained from ghost page-table-walk insertions, $\sumapaces\rtv\delta$ from unfolding the definition of $\mapsto_{\textsf{vpte}}$ in Specification Line 18, and $\fpaddr \mapsto_{\textsf{p}} \textsf{wzero 64}$ in Specification Line 34.
+fragmental ownership of
 \subsection{Unmapping a Page}

### 2023-04-13 programlogic.tex
+

@@
 }{
-  \{P \ast \mathcal{I}\textsf{ASpace}\ast r_d \mapsto_{r}  \textsf{rvd} \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{v,rtv}} \textsf{v} \}_{\textsf{rtv}}
+  \{P \ast \mathcal{I}\textsf{ASpace}\ast r_d \mapsto_{r}  \textsf{rvd} \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{v}} \textsf{v} \}_{\textsf{rtv}}
   \textsf{mov}~\textsf{r}_d,~\textsf{r}_a;\;\overline{is}
@@
 \inferrule[WriteToVirtMemFromReg]{
-  \{P \ast \mathcal{I}\textsf{ASpace}\ast r_s \mapsto_{r}\{q\}  \textsf{rvs}  \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{v,rtv}} \textsf{v} \}_{\textsf{rtv}}\;\overline{ is}
+  \{P \ast \mathcal{I}\textsf{ASpace}\ast r_s \mapsto_{r}\{q\}  \textsf{rvs}  \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{v}} \textsf{v} \}_{\textsf{rtv}}\;\overline{ is}
 }{
-  \{P \ast \mathcal{I}\textsf{ASpace}\ast r_s \mapsto_{r}\{q\}  \textsf{rvs}   \ast r_a \mapsto_{r}\{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{v,rtv}} \textsf{rvs} \}_{\textsf{rtv}}
+  \{P \ast \mathcal{I}\textsf{ASpace}\ast r_s \mapsto_{r}\{q\}  \textsf{rvs}   \ast r_a \mapsto_{r}\{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{v}} \textsf{rvs} \}_{\textsf{rtv}}
   \textsf{ mov}~\textsf{r}_a,~\textsf{r}_s;\;\overline{is}
@@
 \end{mathpar}
-\caption{Reasoning Rules for Selected \textsf{AMD64} Instructions}
+\caption{Reasoning Rules for Selected \textsf{AMD64} Instructions under}
 \label{fig:wpdamd}

### 2023-04-13 programlogic.tex
+

@@
   \begin{lstlisting}[language=Coq]
- Definition $\pfpointsto\locsf\vpts\qfrac\ppts$ :vProp $\Sigma$ := pmapsto_def_v (pfn w) (page_off w) q v.
  Definition $\vaddr\mapsto_{\textsf{t}}\{\textsf{q}\}\; \vpage$ : vProp $\Sigma$ := 
-  $\exists_{\textsf{l4e l3e l2e l1}} \ldotp$  $\ulcorner$ aligned $\vaddr \urcorner \ast$ L4_L1_PointsTo(rtv maddr l4e l3e l2e l1e) $\ast$ addr_L1$(\vaddr,\textsf{l1e}) \mapsto_{p} \vpage$.
+  $\exists_{\textsf{l4e l3e l2e l1}} \ldotp$  $\ulcorner$ aligned $\vaddr \urcorner \ast$ L4_L1_PointsTo(maddr l4e l3e l2e l1e) $\ast$ addr_L1$(\vaddr,\textsf{l1e}) \mapsto_{p} \vpage$.
  Definition L4_L1_PointsTo (rtv maddr l4e l3e l2e l1e :word 64) : vProp $\Sigma$ :=
@@
 \begin{figure*}
-\[
-\begin{array}{l}
-    \vaddr\mapsto_{\textsf{v,rtv}}\{\textsf{q}\}\vpage \stackrel{\triangle}{=} 
-    \exists \paddr\ldotp
-    \sumapaces\rtv\delta \ast 
-  \sumwalkabs\vaddr\qfrac\paddr \ast \paddr \mapsto_{p} \vpage
-\end{array}
-\]
 \todo[inline]{Colin says: again, this is supposed to be a vprop, but isn't doing anything with a cr3val argument.
@@
 \begin{array}{l}
-  \mathcal{I}\textsf{ASpace} \stackrel{\triangle}{:=} 
-  \exists\;\gammaPred \;\ldotp\; \ownGhost\gammaPred{\authfull{\ptableabswalk\ptablestore}} \ast \\
-  \bigast{(\vaddr, \textsf{l1e})\in \ptablestore}{\exists\;(\textsf{l4e l3e l2e})\ldotp \textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\rtv,\textsf{ maddr, l4e, l3e, l2e, l1e})}
+  \mathcal{I}\textsf{ASpace}(\ptablestore,m)\stackrel{\triangle}{:=} \exists\;\gammaPred \ldotp \; \lambda\textsf{ rtv} \ldotp \; \ulcorner m \; !!\; \textsf{rtv} = \textsf{Some } \gammPred \urcorner \ast
+    \ownGhost\gammaPred{\authfull{\ptableabswalk\ptablestore}} \ast \\
+  \bigast{(\vaddr, \textsf{l1e})\in \ptablestore}{\exists\;(\textsf{l4e l3e l2e})\ldotp \textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\textsf{ maddr, l4e, l3e, l2e, l1e})}
 \end{array}
 \]
-\caption{Global Address-Space Invariant}
+\caption{Global Address-Space Invariant with a fixed global map of address-space names $m$}
   \label{fig:peraspaceinvariant}

### 2023-04-13 programlogic.tex
+

@@
   $\exists_{\textsf{l4e l3e l2e l1}} \ldotp$  $\ulcorner$ aligned $\vaddr \urcorner \ast$ L4_L1_PointsTo(maddr l4e l3e l2e l1e) $\ast$ addr_L1$(\vaddr,\textsf{l1e}) \mapsto_{p} \vpage$.
- Definition L4_L1_PointsTo (rtv maddr l4e l3e l2e l1e :word 64) : vProp $\Sigma$ :=
-  (l4M52 maddr rtv) \$\sim$\ (l4off maddr rtv) $\mapsto_{a}$ {q1}  l4e  $\ast$
+ Definition L4_L1_PointsTo (maddr l4e l3e l2e l1e :word 64) : vProp $\Sigma$ :=
+ $\lambda$ cr3val.
+  (l4M52 maddr cr3val) \$\sim$\ (l4off maddr cr3val) $\mapsto_{a}$ {q1}  l4e  $\ast$
   (l3M52 maddr l4e) \$\sim$\ (l3off maddr l4e)  $\mapsto_{a}$ {q2}  l3e $\ast$ 
@@
 \begin{array}{l}
-  \mathcal{I}\textsf{ASpace}(\ptablestore,m)\stackrel{\triangle}{:=} \exists\;\gammaPred \ldotp \; \lambda\textsf{ rtv} \ldotp \; \ulcorner m \; !!\; \textsf{rtv} = \textsf{Some } \gammPred \urcorner \ast
-    \ownGhost\gammaPred{\authfull{\ptableabswalk\ptablestore}} \ast \\
-  \bigast{(\vaddr, \textsf{l1e})\in \ptablestore}{\exists\;(\textsf{l4e l3e l2e})\ldotp \textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\textsf{ maddr, l4e, l3e, l2e, l1e})}
+  \mathcal{I}\textsf{ASpace}(\ptablestore,m)\stackrel{\triangle}{=} \textsf{ASpace\_Lookup}(\ptablestore,m) \ast \\
+ \qquad\qquad\qquad \bigast{(\vaddr, \textsf{l1e})\in \ptablestore}{\exists\;(\textsf{l4e l3e l2e})\ldotp \textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\textsf{ maddr, l4e, l3e, l2e, l1e})} \\
+  \textsf{ where } \\
+   \textsf{ASpace\_Lookup} (\ptablestore,m) \stackrel{\triangle}{=} \lambda\textsf{ cr3val} \ldotp \; \ulcorner m \; !!\; \textsf{cr3val} = \textsf{Some } \gammPred \urcorner \ast
+    \ownGhost\gammaPred{\authfull{\ptableabswalk\ptablestore}}
+
 \end{array}

### 2023-04-13 programlogic.tex
+

@@
   so long as those updates preserve the same virtual-to-physical translations.
-
+\newcommand{\vale}{\textsf{val}}
 \begin{figure*}
@@
 \begin{array}{l}
-    \vaddr\mapsto_{\textsf{v}}\{\textsf{q}\}\vpage : \mathsf{vProp}~\Sigma \stackrel{\triangle}{=} 
-    \lambda \mathit{cr3val}\ldotp
+    \vaddr\mapsto_{\textsf{v}}\{\textsf{q}\}\vale : \mathsf{vProp}~\Sigma \stackrel{\triangle}{=} 
     \exists \paddr\ldotp
@@
     \sumapaces{\mathit{cr3val}}\delta \ast 
-  \sumwalkabs\vaddr\qfrac\paddr \ast \paddr \mapsto_{\mathsf{p}} \vpage
+  \sumwalkabs\vaddr\qfrac\paddr \ast \paddr \mapsto_{\mathsf{p}} \vale
 \end{array}
@@
   \textsf{ where } \\
-   \textsf{ASpace\_Lookup} (\ptablestore,m) \stackrel{\triangle}{=} \lambda\textsf{ cr3val} \ldotp \; \ulcorner m \; !!\; \textsf{cr3val} = \textsf{Some } \gammPred \urcorner \ast
+   \textsf{ASpace\_Lookup} (\ptablestore,m) \stackrel{\triangle}{=} \lambda\textsf{ cr3val} \ldotp \; \ulcorner m \; !!\; \textsf{cr3val} = \textsf{Some } \gammaPred \urcorner \ast
     \ownGhost\gammaPred{\authfull{\ptableabswalk\ptablestore}}

### 2023-04-13 programlogic.tex
+

@@
 \begin{itemize}
-  \item Level-4 Translation (L4): Performs the first level address translation to get the L4 entry (L4E) in Figure \ref{fig:pagetables} by using the masks l4M52 and l4off with \textsf{rtv} virtual base address to get the starting address of the L4 table (L4T), and locate the entry entry amongst 512 (q1) respectively.
+  \item Level-4 Translation (L4): Performs the first level address translation to get the L4 entry (L4E) in Figure \ref{fig:pagetables} by using the masks l4M52 and l4off with \textsf{rtv} virtual base address to get the starting address of the L4 table (L4T), and locate the entry amongst 512 (q1) ones respectively.
     \begin{lstlisting}[language=Coq]
@@
     \end{lstlisting}
- \item Level-3 Translation (L3): Performs the second level address translation to get the L3 entry (L3E) in Figure \ref{fig:pagetables} by using the masks l3M52 and l3off with L4 entry (l4e) to get the starting address of the L3 table (L3T), and locate the entry amongst 512 (q2) respectively.
+ \item Level-3 Translation (L3): Performs the second level address translation to get the L3 entry (L3E) in Figure \ref{fig:pagetables} by using the masks l3M52 and l3off with L4 entry (l4e) to get the starting address of the L3 table (L3T), and locate the entry amongst 512 ones (q2) respectively.
     \begin{lstlisting}[language=Coq]
@@
     \end{lstlisting}
-  \item Level-2 Translation (L2): Performs the second level address translation to get the L2 entry (L2E) in Figure \ref{fig:pagetables} by using the masks l2M52 and l2off with L3 entry (l3e) to get the starting address of the L2 table (L2T), and locate the entry entry amongst 512 (q3) respectively.
+  \item Level-2 Translation (L2): Performs the second level address translation to get the L2 entry (L2E) in Figure \ref{fig:pagetables} by using the masks l2M52 and l2off with L3 entry (l3e) to get the starting address of the L2 table (L2T), and locate the entry amongst 512 ones (q3) respectively.
 \begin{lstlisting}[language=Coq]
@@
     \end{lstlisting}
-  \item Level-1 Translation (L1): Performs the second level address translation to get the L1 entry (L1E) in Figure \ref{fig:pagetables} by using the masks l1M52 and l1off with L2 entry (l2e) to get the starting address of the L1 table (L1T), and locate the entry entry amongst 512 (q4) respectively.
+  \item Level-1 Translation (L1): Performs the second level address translation to get the L1 entry (L1E) in Figure \ref{fig:pagetables} by using the masks l1M52 and l1off with L2 entry (l2e) to get the starting address of the L1 table (L1T), and locate the entry amongst ones 512 (q4) respectively.
    \begin{lstlisting}[language=Coq]

### 2023-04-13 programlogic.tex
physical pointsto definition

@@
 
+\paragraph{Physical memory  points-to} Our memory design is shaped around page-table-walk accesses in Figure \ref{} where we need to mask the same address with different masks: 52 bit-masks for page-table-address and 12 bit masks offset masks for locating the entry.
+With being flexible on decorating these two masks with different the mask-computations, which is needed for different levels of page-tables (e.g.
+l4M52 and l3M52 for page table addresses of level 4 and 3 respectively), we give the physical-memory the following form
+\[ \textsf{w1} \ \sim \ \textsf{w2} \mapsto_{\textsf{a}} \; \{\textsf{q}\} \; \textsf{v} \]
+In this form, we see that \textsf{w1} could be an 52-bits masked address to level 4 table ($\textsf{w1 } =( \textsf{ l4M52 maddr cr3val}) $), and, expectedly \textsf{w2} is an address computed by page-offset computation (e.g.
+$\textsf{l4off maddr cr3val}$).
+Giving a raw 64-bits memory pointsto assertion becomes
+\[\textsf{w} \mapsto_{\textsf{p}} \{q\} \textsf{ v} \stackrel{\triangle}{=} (\textsf{drop 12} w \ \sim \ \textsf{bottom 12 } \mapsto_{\textsf{a}} \textsf{ v} \]
+trivially defining \textsf{w1} in terms of \textsf{w} by dropping its 12-bits, and taking the bottom 12-bits of \textsf{w} as \textsf{w2}.
 \begin{figure*}

### 2023-04-13 examples.tex
+
subseteq

--- 2023-04-12 examples.tex
@@
   \begin{lstlisting}
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,\gammaPred)  \ast  \texttt{r14}\mapsto_{r} \_ \ast \texttt{rdi}\mapsto_{r} \vaddr \ast \texttt{rax}\mapsto_{r} \_ \ast \ulcorner \texttt{aligned va} \land  \theta \; !!\; \vaddr = \texttt{None}\urcorner}_{\rtv}$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m)  \ast  \texttt{r14}\mapsto_{r} \_ \ast \texttt{rdi}\mapsto_{r} \vaddr \ast \texttt{rax}\mapsto_{r} \_ \ast \ulcorner \texttt{aligned va} \land  \theta \; !!\; \vaddr = \texttt{None}\urcorner}_{\rtv}$
 call ensure_L1_page
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,\gammaPred) \ast  \texttt{r14}\mapsto_{r} \_ \ast \texttt{rdi}\mapsto_{r} \vaddr \ast \texttt{rax}\mapsto_{r} \textsf{ pte\_addr} \; \ast }_{\rtv}$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast  \texttt{r14}\mapsto_{r} \_ \ast \texttt{rdi}\mapsto_{r} \vaddr \ast \texttt{rax}\mapsto_{r} \textsf{ pte\_addr} \; \ast }_{\rtv}$
 $\specline{\exists (\entryf ,\;\entrytr,\; \entrytw,\; \entryo,\;\textsf{pte\_addr },\paddr \; : \texttt{word 64}) \ldotp \ulcorner \texttt{addr\_L1 }(\vaddr, \entryo) = \paddr \urcorner \; \ast }_{\rtv}$
@@
 mov %r14, %rax ;; Save that before another call
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,\gammaPred) \ast  \texttt{r14}\mapsto_{r} \texttt{pte\_addr} \ast \texttt{rdi}\mapsto_{r} \vaddr \ast \texttt{rax}\mapsto_{r} \textsf{ pte\_addr} \; \ast }_{\rtv}$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast  \texttt{r14}\mapsto_{r} \texttt{pte\_addr} \ast \texttt{rdi}\mapsto_{r} \vaddr \ast \texttt{rax}\mapsto_{r} \textsf{ pte\_addr} \; \ast }_{\rtv}$
 $\specline{ \nfpointsto{\mask\vaddr\maskfour\rtv}{\mask\vaddr\maskfouroff\rtv}\entryf\qone\naddr \; \ast}_{\rtv}$ 
@@
 call alloc_phys_page_or_panic
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,\gammaPred) \ast  \texttt{r14}\mapsto_{r} \texttt{pte\_addr} \ast \texttt{rdi}\mapsto_{r} \vaddr \;\ast}_{\rtv}$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast  \texttt{r14}\mapsto_{r} \texttt{pte\_addr} \ast \texttt{rdi}\mapsto_{r} \vaddr \;\ast}_{\rtv}$
 $\specline{ \nfpointsto{\mask\vaddr\maskfour\rtv}{\mask\vaddr\maskfouroff\rtv}\entryf\qone\naddr \ast}_{\rtv}$ 
@@
 mov (%r14), %rax
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,\gammaPred) \ast  \texttt{r14}\mapsto_{r} \texttt{pte\_addr} \ast \texttt{rdi}\mapsto_{r} \vaddr \;\ast}_{\rtv}$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast  \texttt{r14}\mapsto_{r} \texttt{pte\_addr} \ast \texttt{rdi}\mapsto_{r} \vaddr \;\ast}_{\rtv}$
 $\specline{ \nfpointsto{\mask\vaddr\maskfour\rtv}{\mask\vaddr\maskfouroff\rtv}\entryf\qone\naddr \ast}_{\rtv}$ 
@@
 $\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\; \sqsubseteq $
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,\gammaPred) \ast  \texttt{r14}\mapsto_{r} \texttt{pte\_addr} \ast \texttt{rdi}\mapsto_{r} \vaddr \ast }_{\rtv}$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast  \texttt{r14}\mapsto_{r} \texttt{pte\_addr} \ast \texttt{rdi}\mapsto_{r} \vaddr \ast }_{\rtv}$
 $\specline{\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\vaddr,\entryf,\entrytr,\entrytw,\paddr) \ast \ulcorner \theta \;!!\;\vaddr = \texttt{None}\urcorner \; \ast}_{\rtv}$
@@
 $\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\; \sqsubseteq $
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace} (<[\vaddr:=\texttt{fpaddr}]> \theta,\gammaPred) \ast  \texttt{r14}\mapsto_{r} \texttt{pte\_addr} \ast \texttt{rdi}\mapsto_{r} \vaddr \; \ast}_{\rtv}$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace} (<[\vaddr:=\texttt{fpaddr}]> \theta,m) \ast}_{\rtv}$
 $\specline{\ulcorner \texttt{aligned fpaddr} \urcorner \ast \texttt{fpaddr} \mapsto_{\textsf{p}} \textsf{ wzero 64} \ast \sumapaces\rtv\delta  \ast\sumwalkabs\vaddr\qfrac\fpaddr}_{\rtv}$
 $\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\; \sqsubseteq $
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace} (<[\vaddr:=\texttt{fpaddr}]> \theta,\gammaPred) \ast \vaddr \mapsto_{\textsf{vpte}}\; \qfrac \;\fpaddr \textsf{ wzero 64}}_{\rtv}$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace} (<[\vaddr:=\texttt{fpaddr}]> \theta,m) \ast \vaddr \mapsto_{\textsf{vpte}}\; \qfrac \;\fpaddr \textsf{ wzero 64}}_{\rtv}$
 \end{lstlisting}
-  \caption{Specification and proof of distilled code for mapping a new page with $\textsf{wpd\_def }_{\rtv}$}
+  \caption{Specification and proof of distilled code for mapping a new page with $\textsf{wpd\_def }_{\rtv}$ with a fixed namespace $\gammaPreds$ and a global map of address-spaces $m$  }
   \todo[inline,color=yellow]{Colin, could you speak about the fact that we have wpd in iProp so that's the reason to have rtval whereever table traversal is needed}
@@
 
-Finally, we can insert $\vaddr$ into the ghost page-table-walk summarization map ($\theta$) to change its state from unmapped ($\ulcorner \theta \;!!\;\vaddr = \texttt{None}\urcorner$ in Specification Line 30) to mapped, and construct virtual-pointsto for $\sumwalkabs\vaddr\qfrac\fpaddr$ with fragmental ownership of
-
-After storing $\textsf{pte\_addr}$
-%\todo[inline]{emphasize the write to the L1 page table is validated by the fact that the identity mapping exists}
+Finally, we can insert $\vaddr$ into the ghost page-table-walk summarization map ($\theta$) to change its state from unmapped ($\ulcorner \theta \;!!\;\vaddr = \texttt{None}\urcorner$ in Specification Line 30) to mapped (Specification Line 33) by Iris' ghost-map update ($\sqsubseteq$), and construct virtual-pointsto $\vaddr \mapsto_{\textsf{vpte}}\; \qfrac \;\fpaddr \textsf{ wzero 64}$ (Specification Line 36) with $\sumwalkabs\vaddr\qfrac\fpaddr$ obtained from ghost page-table-walk insertions, $\sumapaces\rtv\delta$ from unfolding the definition of $\mapsto_{\textsf{vpte}}$ in Specification Line 18, and $\fpaddr \mapsto_{\textsf{p}} \textsf{wzero 64}$ in Specification Line 34.
+fragmental ownership of
 \subsection{Unmapping a Page}
@@
 ;; Save yielding context
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,\gammaPred) \ast [\rtv'+56](\mathcal{I}\texttt{ASpace}(\theta,\gammaPred') \ast \texttt{Pother})}_{\rtv}$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast [\rtv'+56](\mathcal{I}\texttt{ASpace}(\theta,m) \ast \texttt{Pother})}_{\rtv}$
 $\specline{ \texttt{rsi}\mapsto_{r} \rtv' \ast \texttt{rdi}\mapsto_{r} \_ \ast \texttt{rbx}\mapsto_{r} \texttt{rbxv} \ast  \texttt{rsp}\mapsto_{r} \texttt{rspv} \ast \texttt{rbp}\mapsto_{r} \texttt{rbpv} }_{\rtv}$
@@
 mov 48[%rdi], %r15
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,\gammaPred) \ast [\rtv'+56](\mathcal{I}\texttt{ASpace}(\theta,\gammaPred') \ast \texttt{Pother})}_{\rtv}$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast [\rtv'+56](\mathcal{I}\texttt{ASpace}(\theta,m) \ast \texttt{Pother})}_{\rtv}$
 $\specline{ \texttt{rdi}\mapsto_{r} \_ \ast \texttt{rbx}\mapsto_{r} \texttt{rbxv} \ast  \texttt{rsp}\mapsto_{r} \texttt{rspv} \ast \texttt{rbp}\mapsto_{r} \texttt{rbpv} }_{\rtv}$
@@
 mov 56[%rdi], %cr3
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,\gammaPred) \ast [\rtv'+56](\mathcal{I}\texttt{ASpace}(\theta,\gammaPred') \ast \texttt{Pother})}_{\rtv}$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast [\rtv'+56](\mathcal{I}\texttt{ASpace}(\theta,m) \ast \texttt{Pother})}_{\rtv}$
 $\specline{ \texttt{rdi}\mapsto_{r} \_ \ast \texttt{rbx}\mapsto_{r} \texttt{rbxv} \ast  \texttt{rsp}\mapsto_{r} \texttt{rspv} \ast \texttt{rbp}\mapsto_{r} \texttt{rbpv} }_{\rtv}$
@@
 mov %r15, 48[%rsi]
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,\gammaPred,\rtv) \ast [\rtv'+56](\mathcal{I}\texttt{ASpace}(\theta,\gammaPred') \ast \texttt{Pother})}_{\rtv}$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast [\rtv'+56](\mathcal{I}\texttt{ASpace}(\theta,m) \ast \texttt{Pother})}_{\rtv}$
 $\specline{ \texttt{rdi}\mapsto_{r} \_ \ast \texttt{rbx}\mapsto_{r} \rtv'+0 \ast  \texttt{rsp}\mapsto_{r} \rtv'+8 \ast \texttt{rbp}\mapsto_{r} \rtv'+16 }_{\rtv}$
@@
 mov %cr3, 56[%rsi]
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,\gammaPred) \ast [\rtv'+56](\mathcal{I}\texttt{ASpace}(\theta,\gammaPred') \ast \texttt{Pother}) }_{\rtv}$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast [\rtv'+56](\mathcal{I}\texttt{ASpace}(\theta,m) \ast \texttt{Pother}) }_{\rtv}$
 $\specline{ \texttt{rdi}\mapsto_{r} \_ \ast \texttt{rbx}\mapsto_{r} \rtv'+0 \ast  \texttt{rsp}\mapsto_{r} \rtv'+8 \ast \texttt{rbp}\mapsto_{r} \rtv'+16 }_{\rtv}$
@@
 $\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\; \sqsubseteq $
-$\specline{ [\rtv](\texttt{Pcurrent}) \ast \mathcal{I}\texttt{ASpace}(\theta,\gammaPred') \ast \texttt{Pother}}_{\rtv'+56}$
+$\specline{ [\rtv](\texttt{Pcurrent}) \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast \texttt{Pother}}_{\rtv'+56}$
 \end{lstlisting}

### 2023-04-13 examples.tex
compiles again
+

@@
   \begin{lstlisting}
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m)  \ast  \texttt{r14}\mapsto_{r} \_ \ast \texttt{rdi}\mapsto_{r} \vaddr \ast \texttt{rax}\mapsto_{r} \_ \ast \ulcorner \texttt{aligned va} \land  \theta \; !!\; \vaddr = \texttt{None}\urcorner}_{\rtv}$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m)  \ast  \texttt{r14}\mapsto_{\textsf{r}} \_ \ast \texttt{rdi}\mapsto_{r} \vaddr \ast \texttt{rax}\mapsto_{\textsf{r}} \_ \ast \ulcorner \texttt{aligned va} \land  \theta \; !!\; \vaddr = \texttt{None}\urcorner}_{\rtv}$
 call ensure_L1_page
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast  \texttt{r14}\mapsto_{r} \_ \ast \texttt{rdi}\mapsto_{r} \vaddr \ast \texttt{rax}\mapsto_{r} \textsf{ pte\_addr} \; \ast }_{\rtv}$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast  \texttt{r14}\mapsto_{\textsf{r}} \_ \ast \texttt{rdi}\mapsto_{r} \vaddr \ast \texttt{rax}\mapsto_{\textsf{r}} \textsf{ pte\_addr} \; \ast }_{\rtv}$
 $\specline{\exists (\entryf ,\;\entrytr,\; \entrytw,\; \entryo,\;\textsf{pte\_addr },\paddr \; : \texttt{word 64}) \ldotp \ulcorner \texttt{addr\_L1 }(\vaddr, \entryo) = \paddr \urcorner \; \ast }_{\rtv}$
@@
 $\specline{  \nfpointsto{\mask\vaddr\maskthree\entryf}{\mask\vaddr\maskthreeoff\entryf}\entrytr\qtwo\naddr \ast \nfpointsto{\mask\vaddr\masktwo\entrytr}{\mask\vaddr\masktwooff\entrytr}\paddr\qthree\naddr \;\ast}_{\rtv}$
-$\specline{\texttt{pte\_addr} \mapsto_{\texttt{vpte}} \paddr \;(\texttt{wzero 64}) \ast \texttt{rax}\mapsto_{r} \texttt{pte\_addr}  }_{\rtv}$
+$\specline{\texttt{pte\_addr} \mapsto_{\texttt{vpte}} \paddr \;(\texttt{wzero 64}) \ast \texttt{rax}\mapsto_{\textsf{r}} \texttt{pte\_addr}  }_{\rtv}$
 ;; Returns the virtual address of the L1 entry in rax
 mov %r14, %rax ;; Save that before another call
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast  \texttt{r14}\mapsto_{r} \texttt{pte\_addr} \ast \texttt{rdi}\mapsto_{r} \vaddr \ast \texttt{rax}\mapsto_{r} \textsf{ pte\_addr} \; \ast }_{\rtv}$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast  \texttt{r14}\mapsto_{\textsf{r}} \texttt{pte\_addr} \ast \texttt{rdi}\mapsto_{\textsf{r}} \vaddr \ast \texttt{rax}\mapsto_{\textsf{r}} \textsf{ pte\_addr} \; \ast }_{\rtv}$
 $\specline{ \nfpointsto{\mask\vaddr\maskfour\rtv}{\mask\vaddr\maskfouroff\rtv}\entryf\qone\naddr \; \ast}_{\rtv}$ 
 $\specline{  \nfpointsto{\mask\vaddr\maskthree\entryf}{\mask\vaddr\maskthreeoff\entryf}\entrytr\qtwo\naddr \ast \nfpointsto{\mask\vaddr\masktwo\entrytr}{\mask\vaddr\masktwooff\entrytr}\paddr\qthree\naddr \;\ast}_{\rtv}$
-$\specline{\texttt{pte\_addr} \mapsto_{\texttt{vpte}} \paddr \;(\texttt{wzero 64}) \ast \texttt{rax}\mapsto_{r} \texttt{pte\_addr}  }_{\rtv}$
+$\specline{\texttt{pte\_addr} \mapsto_{\texttt{vpte}} \paddr \;(\texttt{wzero 64}) \ast \texttt{rax}\mapsto_{\textsf{r}} \texttt{pte\_addr}  }_{\rtv}$
 call alloc_phys_page_or_panic
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast  \texttt{r14}\mapsto_{r} \texttt{pte\_addr} \ast \texttt{rdi}\mapsto_{r} \vaddr \;\ast}_{\rtv}$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast  \texttt{r14}\mapsto_{\textsf{r}} \texttt{pte\_addr} \ast \texttt{rdi}\mapsto_{\textsf{r}} \vaddr \;\ast}_{\rtv}$
 $\specline{ \nfpointsto{\mask\vaddr\maskfour\rtv}{\mask\vaddr\maskfouroff\rtv}\entryf\qone\naddr \ast}_{\rtv}$ 
@@
 $\specline{\texttt{pte\_addr} \mapsto_{\texttt{vpte}} \paddr \;(\texttt{wzero 64}) \; \ast }_{\rtv}$
-$\specline{\exists \texttt{ fpaddr} \ldotp \ulcorner \texttt{aligned fpaddr} \urcorner \ast \texttt{rax}\mapsto_{r} \texttt{fpaddr} \ast \texttt{fpaddr} \mapsto_{a} (\texttt{wzero 64})}_{\rtv}$
+$\specline{\exists \texttt{ fpaddr} \ldotp \ulcorner \texttt{aligned fpaddr} \urcorner \ast \texttt{rax}\mapsto_{\textsf{r}} \texttt{fpaddr} \ast \texttt{fpaddr} \mapsto_{\textsf{a}} (\texttt{wzero 64})}_{\rtv}$
 ;; Calculate new L1 entry
@@
 mov (%r14), %rax
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast  \texttt{r14}\mapsto_{r} \texttt{pte\_addr} \ast \texttt{rdi}\mapsto_{r} \vaddr \;\ast}_{\rtv}$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast  \texttt{r14}\mapsto_{\textsf{r}} \texttt{pte\_addr} \ast \texttt{rdi}\mapsto_{\textsf{r}} \vaddr \;\ast}_{\rtv}$
 $\specline{ \nfpointsto{\mask\vaddr\maskfour\rtv}{\mask\vaddr\maskfouroff\rtv}\entryf\qone\naddr \ast}_{\rtv}$ 
@@
 $\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\; \sqsubseteq $
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast  \texttt{r14}\mapsto_{r} \texttt{pte\_addr} \ast \texttt{rdi}\mapsto_{r} \vaddr \ast }_{\rtv}$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast  \texttt{r14}\mapsto_{\textsf{r}} \texttt{pte\_addr} \ast \texttt{rdi}\mapsto_{\textsf{r}} \vaddr \ast }_{\rtv}$
 $\specline{\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\vaddr,\entryf,\entrytr,\entrytw,\paddr) \ast \ulcorner \theta \;!!\;\vaddr = \texttt{None}\urcorner \; \ast}_{\rtv}$
-$\specline{\ulcorner \texttt{aligned fpaddr} \urcorner \ast \texttt{rax}\mapsto_{r} \texttt{fpaddr} \ast \texttt{fpaddr} \mapsto_{a} (\texttt{wzero 64}) }_{\rtv}$
+$\specline{\ulcorner \texttt{aligned fpaddr} \urcorner \ast \texttt{rax}\mapsto_{\textsf{r}} \texttt{fpaddr} \ast \texttt{fpaddr} \mapsto_{\textsf{a}} (\texttt{wzero 64}) }_{\rtv}$
 $\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\; \sqsubseteq $
@@
 $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast [\rtv'+56](\mathcal{I}\texttt{ASpace}(\theta,m) \ast \texttt{Pother})}_{\rtv}$
-$\specline{ \texttt{rsi}\mapsto_{r} \rtv' \ast \texttt{rdi}\mapsto_{r} \_ \ast \texttt{rbx}\mapsto_{r} \texttt{rbxv} \ast  \texttt{rsp}\mapsto_{r} \texttt{rspv} \ast \texttt{rbp}\mapsto_{r} \texttt{rbpv} }_{\rtv}$
-$\specline{\texttt{r12}\mapsto_{r} \texttt{r12v} \ast \texttt{r13}\mapsto_{r} \texttt{r13v} \ast \texttt{r14}\mapsto_{r} \texttt{r14v} \ast \texttt{r15}\mapsto_{r} \texttt{r15v}}_{\rtv}$
-$\specline{\texttt{rdi+0} \mapsto_{v} \_ \ast \texttt{rdi+8} \mapsto_{v} \_ \ast \texttt{rdi+16} \mapsto_{v} \_ \ast \texttt{rdi+24} \mapsto_{v} \_ \ast \texttt{rdi+32} \mapsto_{v} \_}_{\rtv}$
-$\specline{\texttt{rdi+40} \mapsto_{v} \_ \ast \texttt{rdi+48} \mapsto_{v} \_\ast \texttt{rdi+56} \mapsto_{v} \_}_{\rtv}$
+$\specline{ \texttt{rsi}\mapsto_{\textsf{r}} \rtv' \ast \texttt{rdi}\mapsto_{\textsf{r}} \_ \ast \texttt{rbx}\mapsto_{\textsf{r}} \texttt{rbxv} \ast  \texttt{rsp}\mapsto_{\textsf{r}} \texttt{rspv} \ast \texttt{rbp}\mapsto_{\textsf{r}} \texttt{rbpv} }_{\rtv}$
+$\specline{\texttt{r12}\mapsto_{\textsf{r}} \texttt{r12v} \ast \texttt{r13}\mapsto_{\textsf{r}} \texttt{r13v} \ast \texttt{r14}\mapsto_{\textsf{r}} \texttt{r14v} \ast \texttt{r15}\mapsto_{\textsf{r}} \texttt{r15v}}_{\rtv}$
+$\specline{\texttt{rdi+0} \mapsto_{\textsf{v}} \_ \ast \texttt{rdi+8} \mapsto_{\textsf{v}} \_ \ast \texttt{rdi+16} \mapsto_{\textsf{v}} \_ \ast \texttt{rdi+24} \mapsto_{\textsf{v}} \_ \ast \texttt{rdi+32} \mapsto_{\textsf{v}} \_}_{\rtv}$
+$\specline{\texttt{rdi+40} \mapsto_{\textsf{v}} \_ \ast \texttt{rdi+48} \mapsto_{\textsf{v}} \_\ast \texttt{rdi+56} \mapsto_{\textsf{v}} \_}_{\rtv}$
 mov 0[%rdi], %rbx
@@
 $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast [\rtv'+56](\mathcal{I}\texttt{ASpace}(\theta,m) \ast \texttt{Pother})}_{\rtv}$
-$\specline{ \texttt{rdi}\mapsto_{r} \_ \ast \texttt{rbx}\mapsto_{r} \texttt{rbxv} \ast  \texttt{rsp}\mapsto_{r} \texttt{rspv} \ast \texttt{rbp}\mapsto_{r} \texttt{rbpv} }_{\rtv}$
-$\specline{\texttt{r12}\mapsto_{r} \texttt{r12v} \ast \texttt{r13}\mapsto_{r} \texttt{r13v} \ast \texttt{r14}\mapsto_{r} \texttt{r14v} \ast \texttt{r15}\mapsto_{r} \texttt{r15v}}_{\rtv}$
-$\specline{\texttt{rdi+0} \mapsto_{v} \texttt{rbxv} \ast \texttt{rdi+8} \mapsto_{v} \texttt{rspv} \ast \texttt{rdi+16} \mapsto_{v} \texttt{rbpv} \ast \texttt{rdi+24} \mapsto_{v} \texttt{r12v} \ast}_{\rtv}$
-$\specline{ \texttt{rdi+32} \mapsto_{v} \texttt{r13v} \ast \texttt{rdi+40} \mapsto_{v} \texttt{r14v} \ast \texttt{rdi+48} \mapsto_{v} \texttt{r15v}\ast \texttt{rdi+56} \mapsto_{v} \_}_{\rtv}$
+$\specline{ \texttt{rdi}\mapsto_{\textsf{r}} \_ \ast \texttt{rbx}\mapsto_{\textsf{r}} \texttt{rbxv} \ast  \texttt{rsp}\mapsto_{\textsf{r}} \texttt{rspv} \ast \texttt{rbp}\mapsto_{\textsf{r}} \texttt{rbpv} }_{\rtv}$
+$\specline{\texttt{r12}\mapsto_{\textsf{r}} \texttt{r12v} \ast \texttt{r13}\mapsto_{\textsf{r}} \texttt{r13v} \ast \texttt{r14}\mapsto_{\textsf{r}} \texttt{r14v} \ast \texttt{r15}\mapsto_{\textsf{r}} \texttt{r15v}}_{\rtv}$
+$\specline{\texttt{rdi+0} \mapsto_{\textsf{v}} \texttt{rbxv} \ast \texttt{rdi+8} \mapsto_{\textsf{v}} \texttt{rspv} \ast \texttt{rdi+16} \mapsto_{\textsf{v}} \texttt{rbpv} \ast \texttt{rdi+24} \mapsto_{\textsf{v}} \texttt{r12v} \ast}_{\rtv}$
+$\specline{ \texttt{rdi+32} \mapsto_{\textsf{v}} \texttt{r13v} \ast \texttt{rdi+40} \mapsto_{v} \texttt{r14v} \ast \texttt{rdi+48} \mapsto_{\textsf{v}} \texttt{r15v}\ast \texttt{rdi+56} \mapsto_{\textsf{v}} \_}_{\rtv}$
 mov 56[%rdi], %cr3
 $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast [\rtv'+56](\mathcal{I}\texttt{ASpace}(\theta,m) \ast \texttt{Pother})}_{\rtv}$
-$\specline{ \texttt{rdi}\mapsto_{r} \_ \ast \texttt{rbx}\mapsto_{r} \texttt{rbxv} \ast  \texttt{rsp}\mapsto_{r} \texttt{rspv} \ast \texttt{rbp}\mapsto_{r} \texttt{rbpv} }_{\rtv}$
-$\specline{\texttt{r12}\mapsto_{r} \texttt{r12v} \ast \texttt{r13}\mapsto_{r} \texttt{r13v} \ast \texttt{r14}\mapsto_{r} \texttt{r14v} \ast \texttt{r15}\mapsto_{r} \texttt{r15v}}_{\rtv}$
-$\specline{\texttt{rdi+0} \mapsto_{v} \texttt{rbxv} \ast \texttt{rdi+8} \mapsto_{v} \texttt{rspv} \ast \texttt{rdi+16} \mapsto_{v} \texttt{rbpv} \ast \texttt{rdi+24} \mapsto_{v} \texttt{r12v} \ast }_{\rtv}$
-$\specline{\texttt{rdi+32} \mapsto_{v} \texttt{r13v} \ast \texttt{rdi+40} \mapsto_{v} \texttt{r14v} \ast \texttt{rdi+48} \mapsto_{v} \texttt{r15v}\ast \texttt{rdi+56} \mapsto_{v} \rtv}_{\rtv}$    
+$\specline{ \texttt{rdi}\mapsto_{\textsf{r}} \_ \ast \texttt{rbx}\mapsto_{\textsf{r}} \texttt{rbxv} \ast  \texttt{rsp}\mapsto_{\textsf{r}} \texttt{rspv} \ast \texttt{rbp}\mapsto_{\textsf{r}} \texttt{rbpv} }_{\rtv}$
+$\specline{\texttt{r12}\mapsto_{\textsf{r}} \texttt{r12v} \ast \texttt{r13}\mapsto_{\textsf{r}} \texttt{r13v} \ast \texttt{r14}\mapsto_{\textsf{r}} \texttt{r14v} \ast \texttt{r15}\mapsto_{\textsf{r}} \texttt{r15v}}_{\rtv}$
+$\specline{\texttt{rdi+0} \mapsto_{\textsf{v}} \texttt{rbxv} \ast \texttt{rdi+8} \mapsto_{\textsf{v}} \texttt{rspv} \ast \texttt{rdi+16} \mapsto_{\textsf{v}} \texttt{rbpv} \ast \texttt{rdi+24} \mapsto_{\textsf{v}} \texttt{r12v} \ast }_{\rtv}$
+$\specline{\texttt{rdi+32} \mapsto_{\textsf{v}} \texttt{r13v} \ast \texttt{rdi+40} \mapsto_{\textsf{v}} \texttt{r14v} \ast \texttt{rdi+48} \mapsto_{\textsf{v}} \texttt{r15v}\ast \texttt{rdi+56} \mapsto_{\textsf{v}} \rtv}_{\rtv}$
 ;; Restore target context
@@
 $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast [\rtv'+56](\mathcal{I}\texttt{ASpace}(\theta,m) \ast \texttt{Pother})}_{\rtv}$
-$\specline{ \texttt{rdi}\mapsto_{r} \_ \ast \texttt{rbx}\mapsto_{r} \rtv'+0 \ast  \texttt{rsp}\mapsto_{r} \rtv'+8 \ast \texttt{rbp}\mapsto_{r} \rtv'+16 }_{\rtv}$
-$\specline{\texttt{r12}\mapsto_{r} \rtv'+24 \ast \texttt{r13}\mapsto_{r} \rtv'+32 \ast \texttt{r14}\mapsto_{r} \rtv'+40 \ast \texttt{r15}\mapsto_{r} \rtv'+48}_{\rtv}$
-$\specline{\texttt{rdi+0} \mapsto_{v} \texttt{rbxv} \ast \texttt{rdi+8} \mapsto_{v} \texttt{rspv} \ast \texttt{rdi+16} \mapsto_{v} \texttt{rbpv} \ast \texttt{rdi+24} \mapsto_{v} \texttt{r12v} \ast }_{\rtv}$
-$\specline{\texttt{rdi+32} \mapsto_{v} \texttt{r13v} \ast \texttt{rdi+40} \mapsto_{v} \texttt{r14v} \ast \texttt{rdi+48} \mapsto_{v} \texttt{r15v}\ast \texttt{rdi+56} \mapsto_{v} \rtv}_{\rtv}$
+$\specline{ \texttt{rdi}\mapsto_{\textsf{r}} \_ \ast \texttt{rbx}\mapsto_{\textsf{r}} \rtv'+0 \ast  \texttt{rsp}\mapsto_{\textsf{r}} \rtv'+8 \ast \texttt{rbp}\mapsto_{\textsf{r}} \rtv'+16 }_{\rtv}$
+$\specline{\texttt{r12}\mapsto_{\textsf{r}} \rtv'+24 \ast \texttt{r13}\mapsto_{\textsf{r}} \rtv'+32 \ast \texttt{r14}\mapsto_{\textsf{r}} \rtv'+40 \ast \texttt{r15}\mapsto_{\textsf{r}} \rtv'+48}_{\rtv}$
+$\specline{\texttt{rdi+0} \mapsto_{\textsf{v}} \texttt{rbxv} \ast \texttt{rdi+8} \mapsto_{\textsf{v}} \texttt{rspv} \ast \texttt{rdi+16} \mapsto_{\textsf{v}} \texttt{rbpv} \ast \texttt{rdi+24} \mapsto_{\textsf{v}} \texttt{r12v} \ast }_{\rtv}$
+$\specline{\texttt{rdi+32} \mapsto_{\textsf{v}} \texttt{r13v} \ast \texttt{rdi+40} \mapsto_{\textsf{v}} \texttt{r14v} \ast \texttt{rdi+48} \mapsto_{\textsf{v}} \texttt{r15v}\ast \texttt{rdi+56} \mapsto_{\textsf{v}} \rtv}_{\rtv}$
 ;; Switch to the new address space
@@
 $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast [\rtv'+56](\mathcal{I}\texttt{ASpace}(\theta,m) \ast \texttt{Pother}) }_{\rtv}$
-$\specline{ \texttt{rdi}\mapsto_{r} \_ \ast \texttt{rbx}\mapsto_{r} \rtv'+0 \ast  \texttt{rsp}\mapsto_{r} \rtv'+8 \ast \texttt{rbp}\mapsto_{r} \rtv'+16 }_{\rtv}$
-$\specline{\texttt{r12}\mapsto_{r} \rtv'+24 \ast \texttt{r13}\mapsto_{r} \rtv'+32 \ast \texttt{r14}\mapsto_{r} \rtv'+40 \ast \texttt{r15}\mapsto_{r} \rtv'+48}_{\rtv}$
-$\specline{\texttt{rdi+0} \mapsto_{v} \texttt{rbxv} \ast \texttt{rdi+8} \mapsto_{v} \texttt{rspv} \ast \texttt{rdi+16} \mapsto_{v} \texttt{rbpv} \ast  \texttt{rdi+24} \mapsto_{v} \texttt{r12v} \ast }_{\rtv}$
-$\specline{\texttt{rdi+32} \mapsto_{v} \texttt{r13v} \ast \texttt{rdi+40} \mapsto_{v} \texttt{r14v} \ast \texttt{rdi+48} \mapsto_{v} \texttt{r15v}\ast \texttt{rdi+56} \mapsto_{v} \rtv}_{\rtv}$
+$\specline{ \texttt{rdi}\mapsto_{\textsf{r}} \_ \ast \texttt{rbx}\mapsto_{r} \rtv'+0 \ast  \texttt{rsp}\mapsto_{\textsf{r}} \rtv'+8 \ast \texttt{rbp}\mapsto_{\textsf{r}} \rtv'+16 }_{\rtv}$
+$\specline{\texttt{r12}\mapsto_{\textsf{r}} \rtv'+24 \ast \texttt{r13}\mapsto_{\textsf{r}} \rtv'+32 \ast \texttt{r14}\mapsto_{\textsf{r}} \rtv'+40 \ast \texttt{r15}\mapsto_{\textsf{r}} \rtv'+48}_{\rtv}$
+$\specline{\texttt{rdi+0} \mapsto_{\textsf{v}} \texttt{rbxv} \ast \texttt{rdi+8} \mapsto_{\textsf{v}} \texttt{rspv} \ast \texttt{rdi+16} \mapsto_{\textsf{v}} \texttt{rbpv} \ast  \texttt{rdi+24} \mapsto_{\textsf{v}} \texttt{r12v} \ast }_{\rtv}$
+$\specline{\texttt{rdi+32} \mapsto_{\textsf{v}} \texttt{r13v} \ast \texttt{rdi+40} \mapsto_{\textsf{v}} \texttt{r14v} \ast \texttt{rdi+48} \mapsto_{\textsf{v}} \texttt{r15v}\ast \texttt{rdi+56} \mapsto_{\textsf{v}} \rtv}_{\rtv}$
 $\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\; \sqsubseteq $
@@
 Figure \ref{fig:swtch} gives simplified code for a basic task switch, the heart of an OS scheduler implementation.
-This is code that saves the context (registers and stack) of the running thread (here in a structure pointed to by \lstinline|rdi|'s value) and restores the context of an existing thread (from \lstinline|rsi|), including the corresponding change of address space for a target thread in another process.
+This is code that saves the context (registers and stack) of the running thread (here in a structure pointed to by \lstinline|rdi|'s value shown in Lines 10-16 and Line 23 of Figure \ref{fig:swtch}) and restores the context of an existing thread (from \lstinline|rsi| shown in Lines 28-38 and Line 45), including the corresponding change of address space for a target thread in another process.
 This code assumes the System V AMD64 ABI calling convention, where the normal registers not mentioned are caller-save, and therefore saved on the stack of the thread that calls this code, as well as on the new stack of the thread that is restored, thus only the callee-save registers and \texttt{cr3} must be restored.\footnote{We are simplifying in a couple basic ways.
@@
 \end{lstlisting}
-A call to this code begins executing one thread in one address space (whose information will be saved in \lstinline[language=C]|save|) and finishes execution executing a different thread in a different address space (whose information is initially in \lstinline[language=C]|restore|).
+A call to this code begins executing one thread (Lines 10-16) in one address space (whose information will be saved in \lstinline[language=C]|save| shown in Line 22) and finishes execution executing a different thread in a different address space (whose information is initially in \lstinline[language=C]|restore| Line 29).
 
@@
 
-The wrinkle, and the importance of the modal treatment of assertions, is that the target thread's precondition is \emph{relative to its address space}, not the address space of the calling thread.
+The wrinkle, and the importance of the modal treatment of assertions, is that the target thread's precondition is \emph{relative to its address space}, not the address space of the calling thread shown as
+\[[\rtv'+56](\mathcal{I}\texttt{ASpace}(\theta,m) \ast \texttt{Pother})\]
+in the specfication.
 Thus the precondition of this code is that the then-current stack pointer has a return address expecting the then-current callee-save register values and suitablly updated (i.e., post-return) stack in the \emph{current} (initial) address space, while the stack pointer saved in the context to restore expects the same of the saved registers and stack \emph{in the other address space}.
@@
 
-In addition, immediately after the page table switch, the points-to information about the saved and restored contexts is \emph{also} guarded by a modality for the old address space, since there is no automatic guarantee that that memory is mapped in the new address space.
+In addition, immediately after the page table switch, the points-to information about the saved and restored contexts is \emph{also} guarded by a modality (Specification Line 52)
+\[[\rtv](\texttt{Pcurrent}) \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast \texttt{Pother}}_{\rtv'+56}\]
+for the old address space, since there is no automatic guarantee that that memory is mapped in the new address space.
 The ability to transfer that points-to information out of that modality is specific to a given kernel's design.

### 2023-04-13 examples.tex
+

@@
 \end{lstlisting}
-A call to this code begins executing one thread (Lines 10-16) in one address space (whose information will be saved in \lstinline[language=C]|save| shown in Line 22) and finishes execution executing a different thread in a different address space (whose information is initially in \lstinline[language=C]|restore| Line 29).
+A call to this code begins executing one thread (Lines 10-16) in one address space ($\rtv$ in Figure \ref{fig:swtch}) (whose information will be saved in \lstinline[language=C]|save| shown in Line 22) and finishes execution executing a different thread in a different address space (whose information is initially in \lstinline[language=C]|restore| Line 29).
 
@@
 in the specfication.
-Thus the precondition of this code is that the then-current stack pointer has a return address expecting the then-current callee-save register values and suitablly updated (i.e., post-return) stack in the \emph{current} (initial) address space, while the stack pointer saved in the context to restore expects the same of the saved registers and stack \emph{in the other address space}.
+\todo[Colin: could you associate the rest of this paragraph with spec?]Thus the precondition of this code is that the then-current stack pointer has a return address expecting the then-current callee-save register values and suitablly updated (i.e., post-return) stack in the \emph{current} (initial) address space, while the stack pointer saved in the context to restore expects the same of the saved registers and stack \emph{in the other address space}.
 And the postcondition is analagous, but interpreted \emph{in a different address space}: the then-current (updated) stack has a return address expecting the new (restored) register values, and the saved context's invariant captures the precondition for restoring its execution \emph{in the previous address space}.
@@
 In addition, immediately after the page table switch, the points-to information about the saved and restored contexts is \emph{also} guarded by a modality (Specification Line 52)
-\[[\rtv](\texttt{Pcurrent}) \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast \texttt{Pother}}_{\rtv'+56}\]
+\[ [\rtv](\texttt{Pcurrent}) \qquad \textsf{ where Pcurrent is all available logical assertions under addrress-space} \rtv\]
 for the old address space, since there is no automatic guarantee that that memory is mapped in the new address space.

### 2023-04-13 background.tex
revisions up through end of section 3

@@
     \includegraphics[width=\columnwidth]{pagetables.png}
+  \todo[inline]{this image got \emph{very} pixelated. is this possible to fix?}
     \caption{x86-64 page table lookups.}
@@
 While this initially sounds like relatively modest functionality whose implementation may be complicated by hardware 
-subtleties, a number of more serious complexities must be handled.
+subtleties, even these basic operations are actually quite intrictate.
+Most notably, updates to the page tables are performed as writes to memory --- which are themselves subject to address translation.
+In the case of changing the mappings for the currently-active set of page tables, the OS kernel is modifying the tables involved in its
+own access of the tables.
 
-\todo[inline]{Colin says: we should chat about the paragraph below. We don't want to advertise too many
-challenges of VMM verification that we don't actually solve (or rather, we'd push them to the and of the paper
-as future work). But some of these I think we could split the difference with, for example by defining
-a variant of virtual-pointsto that allowed virtual address aliasing, but not actually verifying any code examples
-with it (just showing it's compatible with our approach).
-}
-For one, virtual addresses may \emph{alias} when more than one is mapped to the same physical memory. 
-This can occur when a process requests it (e.g., via \texttt{mmap} or \texttt{shm\_open} on UNIX-style kernels). 
-It also occurs automatically in most general-purpose kernels: while for normal programs the OS uses the MMU to isolate 
-and abstract physical memory, the kernel's own functionality is often easier to implement if the kernel can directly 
-access memory given its physical address, for example when interacting with hardware devices. For this reason, most 
-kernels contain an \emph{identity map} of all physical memory in a machine. 
-\todo[inline]{We should try defining an identity map on a whiteboard and see what breaks. If it's feasible we should
-tweak and get a definition in the paper. If not, we should cut the identity map discussion.}
-In some cases this is a literal identity map, 
-where for every virtual address $p$, interpreting $p$ as a virtual address instead will map back to $p$ as a physical 
-address. In other cases there is an offset involved, where to access physical address $p$ the kernel can access virtual
- address $p+\mathit{offset}$. This identity map saves the kernel the performance cost of having to continually add and
- remove transient mappings simply to briefly access a specific physical memory region. But this coexists with additional
- virtual addresses (those used by most kernel code and data structures), in a form of virtual address aliasing. Note 
-that virtual address aliasing is both prevalent in general-purpose kernels, and violates the core assumption of most 
-memory models assumed by verified compilers like CompCert, which assume no virtual address aliasing.
+% \todo[inline]{Colin says: we should chat about the paragraph below. We don't want to advertise too many
+% challenges of VMM verification that we don't actually solve (or rather, we'd push them to the and of the paper
+% as future work). But some of these I think we could split the difference with, for example by defining
+% a variant of virtual-pointsto that allowed virtual address aliasing, but not actually verifying any code examples
+% with it (just showing it's compatible with our approach).
+% }
+% For one, virtual addresses may \emph{alias} when more than one is mapped to the same physical memory. 
+% This can occur when a process requests it (e.g., via \texttt{mmap} or \texttt{shm\_open} on UNIX-style kernels). 
+% It also occurs automatically in most general-purpose kernels: while for normal programs the OS uses the MMU to isolate 
+% and abstract physical memory, the kernel's own functionality is often easier to implement if the kernel can directly 
+% access memory given its physical address, for example when interacting with hardware devices. For this reason, most 
+% kernels contain an \emph{identity map} of all physical memory in a machine. 
+% \todo[inline]{We should try defining an identity map on a whiteboard and see what breaks. If it's feasible we should
+% tweak and get a definition in the paper. If not, we should cut the identity map discussion.}
+% In some cases this is a literal identity map, 
+% where for every virtual address $p$, interpreting $p$ as a virtual address instead will map back to $p$ as a physical 
+% address. In other cases there is an offset involved, where to access physical address $p$ the kernel can access virtual
+%  address $p+\mathit{offset}$. This identity map saves the kernel the performance cost of having to continually add and
+%  remove transient mappings simply to briefly access a specific physical memory region. But this coexists with additional
+%  virtual addresses (those used by most kernel code and data structures), in a form of virtual address aliasing. Note 
+% that virtual address aliasing is both prevalent in general-purpose kernels, and violates the core assumption of most 
+% memory models assumed by verified compilers like CompCert, which assume no virtual address aliasing.
 
-\todo[inline]{Colin says: Again, let's chat about the next paragraph. This might be another where we can
-split the difference on extension by defining something and proving a $\ast-\ast$ kind of result for sharing
-page table fragments, rather than actually proving any new code. I'm less sure about this one, though.
-}
-Another complicating factor is that portions of page tables can be shared, i.e., trees of page tables can overlap and
-become forests. 
-One common way this used to happen was for one subtree of page tables to map the kernel if linked in appropriately to 
-a higher-level page table, and then to reuse that kernel mapping portion in every process's page tables. This was mostly 
-phased out in the wake of Spectre, but remains relevant on correctly-implemented hardware.
+% \todo[inline]{Colin says: Again, let's chat about the next paragraph. This might be another where we can
+% split the difference on extension by defining something and proving a $\ast-\ast$ kind of result for sharing
+% page table fragments, rather than actually proving any new code. I'm less sure about this one, though.
+% }
+% Another complicating factor is that portions of page tables can be shared, i.e., trees of page tables can overlap and
+% become forests. 
+% One common way this used to happen was for one subtree of page tables to map the kernel if linked in appropriately to 
+% a higher-level page table, and then to reuse that kernel mapping portion in every process's page tables. This was mostly 
+% phased out in the wake of Spectre, but remains relevant on correctly-implemented hardware.
 
-These dove-tail with functionality in the OS scheduler, which since it is responsible for dealing with multiple 
-address spaces, must keep track of which virtual addresses are valid (and in what way) in which address spaces, 
-where some virtual addresses are valid in only a single address space (e.g., a code address for a particular usermode 
+Virtual memory concerns propagate to the OS scheduler, which is responsible for dealing with multiple 
+address spaces, so must keep track of which virtual addresses are valid (and in what way) in which address spaces. 
+Some virtual addresses are valid in only a single address space (e.g., a code address for a particular usermode 
 process), while others are valid in all address spaces (e.g., kernel data structure pointers). 
@@
 mind~\cite{hintikka1962knowledge}). A unifying concept across any modality is that they behave as applicative functors, 
-typically satisfying (directly, or as a derived law):
+typically satisfying (directly, or as a derived law, depending on the modality):
 \[ (P\rightarrow Q) \rightarrow M(P) \rightarrow M(Q)\]

### 2023-04-13 intro.tex
+
revisions up through end of section 3

@@
 \label{sec:intro}
-\mytodo{Colin's Part}
 Virtual memory management lies at the core of modern OS kernel implementation.
@@
 to speak directly \emph{within the logic} about other address spaces.
+%\begin{figure}
+%   \includegraphics[width=0.5\columnwidth]{architecture.png}
+%  \caption{Layers of Implementation}
+%  \todo[inline]{Colin says: I don't think this figure helps readers; it is too abstract}
+%  \label{fig:architecture}
+%\end{figure}
 
-\begin{figure}
-   \includegraphics[width=0.5\columnwidth]{architecture.png}
-  \caption{Layers of Implementation}
-  \todo[inline]{Colin says: I don't think this figure helps readers; it is too abstract}
-  \label{fig:architecture}
-\end{figure}
-
-We realize and show our reasoning principles, shown in Figure \ref{fig:architecture}
 \begin{itemize}

### 2023-04-13 syntaxsemantics.tex
revisions up through end of section 3

--- 2023-04-12 syntaxsemantics.tex
@@
 A register identifier, $\reg$, is chosen from a fixed finite set of register identifiers, $\regset$. 
-We use these identifiers to access the register values, $\regval \in \regvaltype$.
-\begin{figure}[t]
-\newcommand{\commentary}[1]{ & \text{\small\it #1} \\}
-\[
-  \begin{array}{r@{\;}c@{\;}l}
-    \loc & \in & \Loc \\
-    \reg & \in & \regset \\
-    \regval & \in & \regvaltype \\
-    \val & ::= & \vunit
-
-\\
-    \instrs & ::= &
-    \begin{array}[t]{@{}l@{\hspace{10mm}}l@{}}
-    \begin{array}[t]{@{}ll@{}}
-      \iskip
-                   \commentary{no-op}
-      \iseq\instr\instrs
-                   \commentary{sequencing}
-      % \ising\instr
-                   % \commentary{executing}             
-    \end{array}
-    %&
-    %\begin{array}[t]{@{}ll@{}}
-     % \ialloc\lval\allocsize
-      %             \commentary{heap allocation}
-    %\end{array}
-    \end{array}
-    \\
-
-    \ectx & ::= &
-      \hole \mid
-      \iseq\ectx\instrs 
-    \\
-  \end{array}
-\]
-\caption{Syntax}
-\Description{Syntax}
-\label{fig:syntax}
-\end{figure}
-Unlike registers, we do not abstract the memory indices as a special type but instead, for the sake of clarity and ease of representation, we show differently masked machine words, $\loc \in \Loc$, with the subscripts showing the length of a 64-bit machine word after masking, e.g.
-$\kw{w}_{12}$ is a 12-bit resource which can be obtained after masking 52-bit of a 64-bit word.
-This simple syntax in Figure \ref{fig:syntax} includes a simple syntax for values because our language is purely \textit{imperative}, i.e.
-an instruction does not return a value.
-Intentionally, we restrict any stream of instruction, $\instrs$, to eventually reduce to $\iskip$ instruction.
-Our syntax capture the no-op ``$\iskip$'' and the sequencing construct ``$\iseq{\instr}{\instrs}$'' in their standard forms.
+We use these identifiers $\reg$ for register names such as \texttt{rax}, \texttt{r8}, or \texttt{cr3}.
+% \begin{figure}[t]
+% \newcommand{\commentary}[1]{ & \text{\small\it #1} \\}
+% \[
+%   \begin{array}{r@{\;}c@{\;}l}
+%     \loc & \in & \Loc \\
+%   \end{array}
+%   \begin{array}{r@{\;}c@{\;}l}
+%     \reg & \in & \regset \\
+%   \end{array}
+%   \begin{array}{r@{\;}c@{\;}l}
+%     \regval & \in & \regvaltype \\
+%     % \val & ::= & \vunit
+%   \end{array}
+% \quad
+%   \begin{array}{r@{\;}c@{\;}l}
+%     \instrs & ::= &
+%     \begin{array}[t]{@{}l@{\hspace{10mm}}l@{}}
+%     \begin{array}[t]{@{}ll@{}}
+%       \iskip
+%                    \commentary{no-op}
+%       \iseq\instr\instrs
+%                    \commentary{sequencing}
+%       % \ising\instr
+%                    % \commentary{executing}             
+%     \end{array}
+%     %&
+%     %\begin{array}[t]{@{}ll@{}}
+%      % \ialloc\lval\allocsize
+%       %             \commentary{heap allocation}
+%     %\end{array}
+%     \end{array}
+%     % \\
+%     % \ectx & ::= &
+%     %   \hole \mid
+%     %   \iseq\ectx\instrs 
+%     % \\
+%   \end{array}
+% \]
+% \caption{Syntax}
+% \Description{Syntax}
+% \label{fig:syntax}
+% \end{figure}
+For clarity and ease of representation, we use machine words, $\loc \in \Loc$, with the subscripts showing the number of bits in a word,
+for memory addresses, values, and offsets, rather than distinct location types that wrap machine words.
+For example, $\kw{w}_{12}$ is a 12-bit word, which can be obtained for example truncating away 52 bits of a 64-bit word ($\kw{w}_{64}$).
 \subsection{State}
 \label{sec:state}
-%\todo[inline]{I don't think we need or want to describe the operational semantics in detail in the submission.
-It adds a lot of formalization that doen'st necessarily contribute to reader understanding.
-We should describe the runtime \emph{state} $\sigma$ of memory and registers, since that is directly used in interpreting assertions.
-But the reduction rules will mostly add mental overhead here.}
-%\todo[inline,color=red]{Colin says: On the contrary, we must describe assembly instructions in enough detail for readers
-%who aren't already assembly programmers to decode our examples! We don't have to talk about details of the formalization
-%of the semantics, but we do need to talk about basic instructions we cover, basic patterns (like destination on the left),
-%and just a bit about how all memory accesses implement Figure \ref{fig:pagetables}'s semantics.}
-We represent the machine state mainly as a finite map of registers to register values and a map of masked physical memory addresses to 64-bit physical memory values.
-Together with default \textsf{CPU} instantiation, we end up having our state with following pieces:
+We represent the machine state mainly as a finite map of registers to register values and a map of word-aligned physical memory addresses 
+to 64-bit physical memory values. 
+Thus our states $\sigma$ include:
+%Together with default \textsf{CPU} instantiation, we end up having our state with following pieces:
 \begin{itemize}
-\item A constant cpu initialized with default value: $\sigma.\mathcal{C}$
+%\item A constant cpu initialized with default value: $\sigma.\mathcal{C}$
 \item A register map: $\sigma.\mathcal{R}: \kw{greg} \rightarrow_{\textrm{fin}} \kw{regval} $
@@
 \end{itemize}
-As one might have already anticipated from the syntax we introduce, we do not bind any value of an evaluated expression.
+Of particular note, \texttt{cr3}, the page table register, is included in the machine state.
+% As one might have already anticipated from the syntax we introduce, we do not bind any value of an evaluated expression.
 All the indices and accessed values are treated as globally referenced.
@@
 \label{sec:instructions}
-Page-table walk shown in Figure \ref{fig:pagetables} is realized with physical memory \texttt{load}s and \textsf{store}s for the entries in the page-tables which are simply modelled as lookup and updates to an instance of a memory map with 64 bit entries.
-In the address translation,
-\begin{lstlisting}[language=Coq]
- Definition translate (rtv: word 64) (w: word 64): word 64 + MemFail :=
+
+\begin{figure}[t]
+\begin{lstlisting}[language=Coq,mathescape]
+ Definition translate (rtv: word 64) ($\sigma$: state) (w: word 64): word 64 + MemFail :=
       l4e $\leftarrow$ translate_top_level $\sigma.\mathcal{M}$ rtv w; 
@@
       addr $\leftarrow$ translate_from_l1e $\sigma.\mathcal{M}$ l1e w;
-      concat (shift_dropn addr 3 three_lt64) low3.
+      return concat (shift_dropn addr 3 three_lt64) low3.
 \end{lstlisting}
-we see a physical memory $\mathsf{load}$ (i.e.
-physical memory map ($\sigma.\mathcal{M}$) for each level) for each level of page-table-walk concluded with alignment of the returned address for each \textit{successful} page-table-walk traversal what our reasoning principles only consider.
+\vspace{-1em}
+\caption{\coq implementation of address translation (slightly simplified).}
+\label{fig:coq_addr_translation}
+\end{figure}
 
+Programs in our logic are instruction sequences \instrs, which are formed by either a basic instruction \iskip, or prefixing an existing instruction
+sequence with an additional instruction (\iseq\instr\instrs).
+We model (and later, give program logic rules for) instructions for basic register moves, and reading and writing memory.
+The latter require page table walks.
+Figure \ref{fig:coq_addr_translation} gives a slightly simplified version of our address translation code in monadic style:
+starting from the root page table address \textsf{rtv}, the virtual address \textsf{w} is either resolved to a physical address or an error is returned.
+The full set of possible errors is not relevant to this paper; our logic is tailored for kernel code which should not fault,
+so our proof rules guarantee no memory failures occur --- that the page table walk for any dereferenced address will succeed.
+
+% Page-table walk shown in Figure \ref{fig:pagetables} is realized with physical memory \texttt{load}s and \textsf{store}s for the entries in the 
+% page-tables which are simply modelled as lookup and updates to an instance of a memory map with 64 bit entries. In the address translation,
+% we see a physical memory $\mathsf{load}$ (i.e. physical memory map ($\sigma.\mathcal{M}$) for each level) for each level of page-table-walk 
+% concluded with alignment of the returned address for each \textit{successful} page-table-walk traversal which is what our reasoning principles 
+% only consider.
+
+The main instructions we implement are variants of the x86-64 \lstinline|mov| instruction, which we format in Intel syntax (destination on the left, source on the right).
 As expected \textsf{store} would be the update on the memory map entry
-\todo[inline,color=yellow]{Colin, could you mention that we simplified the maps here to get rid of MemFail?}
-After defining the translation in terms of physical load and store action, now we can define our memory \lstinline|mov| instructions
+% After defining the translation in terms of physical load and store action, now we can define our memory \lstinline|mov| instructions
+For memory operations, we treat
 \[
@@
   \hbox{(\TirNameStyle{WriteToMemFromReg})} \qquad
-  \textsf{mov mem64 srcreg64} \\
+  \textsf{mov}~[r_m]~r_r \\
   \hbox{(\TirNameStyle{WriteToRegFromMem})} \qquad 
-  \textsf{mov reg64 mem64}
+  \textsf{mov}~r_r~[r_m]
   \end{array}
   \]
-  which are realized by first translating the virtual-memory address to store a value from a register (\TirNameStyle{WriteToMemFromReg}) or load a value at the translated address to be stored in a register (\TirNameStyle{WriteToRegFromMem}).
-More \lstinline|mov| instructions operating with memory addresses can be found in the submission artifact for different addressing schemes and data sizes.
+which are realized by first translating the virtual-memory address of a memory location stored in $r_m$ to a physical memory location,
+then updating those memory contents with the contents of register $r_r$ (\TirNameStyle{WriteToMemFromReg}) 
+or loading the value at that physical memory location into the register $r_r$
+(\TirNameStyle{WriteToRegFromMem}). 
+% More \lstinline|mov| instructions operating with memory addresses can be found in the submission artifact for different addressing schemes and data sizes.
+Additional variants (such as instructions which access memory at a constant offset from the virtual address stored in a register) are also treated.
 
-  Register instructions, on the other hand, are implemented as $\sigma.\mathcal{R}$ lookup and store
-\[
-\begin{array}{l}
-  \hbox{(\TirNameStyle{WriteToRegFromReg})} \qquad 
-   \textsf{mov ireg64 ireg64} \\
-  \hbox{(\TirNameStyle{WriteToCtlRegFromReg})} \qquad 
-  \textsf{mov creg64 ireg64}\\
-    \hbox{(\TirNameStyle{WriteToRegFromCtlReg})} \qquad 
-  \textsf{mov ireg64 creg64}
-  \end{array}
-\]
-for different types of registers (e.g. control and general).
+For register operations, we support moves back and forth between any pair of registers, though the \coq formalization splits these according to whether
+the source or destination registers are standard data registers (e.g., \lstinline|rax|) or control registers (e.g., \lstinline|cr3|).
+% \[
+% \begin{array}{l}
+%   \hbox{(\TirNameStyle{WriteToRegFromReg})} \qquad 
+%    \textsf{mov ireg64 ireg64} \\
+%   \hbox{(\TirNameStyle{WriteToCtlRegFromReg})} \qquad 
+%   \textsf{mov creg64 ireg64}\\
+%     \hbox{(\TirNameStyle{WriteToRegFromCtlReg})} \qquad 
+%   \textsf{mov ireg64 creg64}
+%   \end{array}
+% \]
+% for different types of registers (e.g. control and general).
 %

### 2023-04-13 examples.tex

@@
 in the specfication.
-\todo[Colin: could you associate the rest of this paragraph with spec?]Thus the precondition of this code is that the then-current stack pointer has a return address expecting the then-current callee-save register values and suitablly updated (i.e., post-return) stack in the \emph{current} (initial) address space, while the stack pointer saved in the context to restore expects the same of the saved registers and stack \emph{in the other address space}.
+\todo{Colin: could you associate the rest of this paragraph with spec?}Thus the precondition of this code is that the then-current stack pointer has a return address expecting the then-current callee-save register values and suitablly updated (i.e., post-return) stack in the \emph{current} (initial) address space, while the stack pointer saved in the context to restore expects the same of the saved registers and stack \emph{in the other address space}.
 And the postcondition is analagous, but interpreted \emph{in a different address space}: the then-current (updated) stack has a return address expecting the new (restored) register values, and the saved context's invariant captures the precondition for restoring its execution \emph{in the previous address space}.

### 2023-04-13 programlogic.tex
compiles again
starting to clean up section 4

--- 2023-04-13 programlogic.tex
@@
 \newcommand{\ptablestore}{\theta}
-Our reasoning principles derive from the following perspectives on the virtual-memory abstraction:
+
+We derive a program logic (a separation logic) supporting the following stances and constraints:
 \begin{enumerate}
-\item \textit{logical representation of addressing}: we present physical memory addressing as separation-logic points-to relations.
-\item \textit{sharing}: the physical location backing a virtual address is located (during a page table walk) through a 
+\item \textit{address-spaces as modal contexts}: Assertions in our logic are context-dependent,
+  in the sense that their truth depends on which address space they are used in, due to the need to support virtual points-to assertions.
+\item \textit{sharing}: The physical location backing a virtual address's storage is located (during a page table walk) through a 
       set of physical page-table (L4-L1 page-tables) acceses that are shared amongst different virtual addresses (specifically,
-      those on the same page of memory). This sharing imposes constraints on giving a definition to the virtual-address 
+      those on the same page of memory\footnote{of in the case of L2 or higher levels of tables, within a given broader region.}).
+      This sharing imposes constraints on definiting points-to assertions
       in terms of physical (L4-L1) page-table memory accesses
-\item \textit{context-agnostic-resources}: each virtual address is valid under a certain address-space, but it does not represent this \textit{knowledge} of its address-space. 
-\item \textit{address-spaces as modal contexts}: we call these context-dependent\todo{Colin switched away from `agnostic', which is the opposite!} resources (virtual-address pointstos) as \textit{context-resource}, and the context identfied by a unique \textit{namespace}, and  defining the truth for these context resources as \text{modal-context}
-\item \textit{updating inter/intra address-spaces}: we present logical abstractions to enable allocating/updating pages in the existence of \textit{sharing}  
-\item \textit{address-space switch as changing the "World" of truth}: switching from one address-space to another logically becomes introduction of \textit{context-resources} to the \textit{modal-context} under which they are valid, and loading \textit{context-resources} that are valid under the \textit{currently} loaded address-space
+\item \textit{context-agnostic-resources}: each virtual address is valid under a certain address-space, 
+      but it does not represent this \textit{knowledge} of its address-space. 
+      That is, assertions are not explicitly tagged with their address space validity
+\item \textit{updating address-space mappings}: We present logical abstractions to enable 
+      updating not only pages of typical data in memory, but also page tables themselves.\footnote{Prior work relied on unfolding operational semantics
+      to verify page table updates.}
+\item \textit{explicitly-modal assertions}: Our logic includes a means to talk about facts being true
+      in another address space
+\item \textit{address-space switch as changing the "World" of truth}: Switching from one address-space to another logically
+      becomes a simultaneous introduction-and-elimination of a pair of modal assertions (for different address spaces)
 \end{enumerate}
-Based on these perspectives, we coneptualize truth in an address-space through the contingency it exhibits: 
-\textit{assertion P holds only relative to a certain address space}. 
-Therefore, as a first step, we work in a pointwise lifting of Iris's standard BI proposition,
+
+The first viewpoint, that the truth of an assertion is relative to an address space, has far-reaching consequences.
+Therefore, as a first step, we work entirely in a pointwise lifting of Iris's base BI logic,
 essentially working with separation logic assertions indexed by a page table root as a $\mathcal{W}_{64}$
 (Figure \ref{fig:vprop}), which we call $\textsf{vProp }\Sigma$; this is the (\textsc{Coq}) type of assertions in our logic.
+Most constructs in Iris's base logic are defined with respect to any BI-algebra (of \textsc{Coq} type \lstinline|bi|), so automatically
+carry over to our derived logic.
+However, be must still build up from existing Iris primitives to provide new primitives that depend on the address space --- primarily the notion
+of virtual points-to.
+
 \begin{figure}[t]
@@
 %\end{itemize}
-\subsection{Points-To Assertions}
+
+\subsection{Base Points-To Assertions}
 \label{sec:pointsto}
-\todo[inline]{Colin says: Current limit of \emph{thorough} revisions}
-\todo[inline,color=red]{Colin says: This section at a glance looks on the right track, but none of the assertions are actually
-written as vProps, they're all still iProps. I can fix that as I go}
-\todo[inline]{Colin says: There's really no need to discuss the 52-12 split in the physical pointsto definition.
-That level of detail of the operational semantics \emph{is} too detailed for the paper (but should be mentioned in the
-artifact docs). I also don't know what the `ent' and `rtv' and similar are in those split physical pointstos}
-We abstract physical memory addressing and registers naming with well-known separation logic assertion \textit{points-to} assertions.
-\begin{enumerate}
-\item Physical address points-to, $\pfpointsto\locsf\vpts\qfrac\ppts$
-\todo[inline,color=red]{Colin says: This syntax is what most readers would expect for physical poinst-tos, but it's
-inconsistent with the split 52+12  stuff later... Whatever we do, we need to be consistent}
-\item Register points-to, $\pfpointsto\rg\rv\qfrac\rpts$
-\end{enumerate}
+We build up our context-dependent assertions from two basic types of assertions that do not care about address spaces.
+% We abstract physical memory addressing and registers naming with well-known separation logic assertion \textit{points-to} assertions.
+% \begin{enumerate}
+% \item Physical address points-to, $\pfpointsto\locsf\vpts\qfrac\ppts$
+% \todo[inline,color=red]{Colin says: This syntax is what most readers would expect for physical poinst-tos, but it's
+% inconsistent with the split 52+12  stuff later... Whatever we do, we need to be consistent}
+% \item Register points-to, $\pfpointsto\rg\rv\qfrac\rpts$
+% \end{enumerate}
 
-\paragraph{Register points-to} The assertion $\pfpointsto\rg\rv\rpts\qfrac$ ensures the ownership of the register $\rg$ naming the register value $\rv$.
-The fraction $\qfrac$ with value 1 asserts the unique ownership of the register mapping, and grants update permission on it, otherwise, any value $0 < \qfrac <1$ represents partial ownership granting readonly permission on the mapping.
+\paragraph{Register points-to} 
+The assertion $\pfpointsto\rg\rv\rpts\qfrac$ ensures the ownership of the register $\rg$ naming the register value $\rv$. 
+The fraction $\qfrac$ with value 1 asserts the unique ownership of the register mapping, and grants update permission on it, 
+otherwise, any value $0 < \qfrac <1$ represents partial ownership granting readonly permission on the mapping.
+
+\paragraph{Physical memory  points-to} The soundness proofs for our logic's rules largely center around page-table-walk accesses as in Figure \ref{fig:coq_addr_translation},
+ where (inside the helper functions) we need to split a 64-bit word representing entries in these tables into 52-bit segments
+which specify the physical page number of another page in the walk, and 12-bit segments containing various control bits.
+For this reason we have two notions of physical points-to facts. The primitive notion closest to our machine model is:
+\[ \textsf{w1} \ \sim \ \textsf{w2} \mapsto_{\textsf{a}} \; \{\textsf{q}\} \; \textsf{v} \]
+In this form, we see that \textsf{w1} (a $\mathcal{W}_{52}$) essentially selects a 4KB page of physical memory,
+and \textsf{w2} (a $\mathcal{W}_{12}$) is an offset within that page.
+% could be an 52-bits masked address to level 4 table 
+% ($\textsf{w1 } =( \textsf{ l4M52 maddr cr3val}) $),
+%and, expectedly \textsf{w2} is an address computed by page-offset computation (e.g. $\textsf{l4off maddr cr3val}$). 
+From this we can derive a more concise physical points-to when the split is unimportant:
+% Giving a raw 64-bits memory pointsto assertion becomes
+\[\textsf{w} \mapsto_{\textsf{p}} \{q\} \textsf{ v} \stackrel{\triangle}{=} (\textsf{drop 12}~w) \ \sim \ (\textsf{bottom 12}~w)\mapsto_{\textsf{a}} \textsf{ v} \]
 
@@
 
-\todo[inline,color=red]{Colin says: Ismail, I think you accidentally deleted discussion of physical points-tos here?}
+\subsection{An Overly-Restrictive Definition for Virtual Memory Addressing}
+\label{sec:overly-restrictive}
+One could give a naive logical definition to the virtual mememory addressing as conjuctions of each level translation 
+(Figure \fref{fig:strongvirtualpointsto}). The very first assurance our logical constructions need to be able assert is 
+the \textit{existence} of the page table walk. Our naive attempt shown in Figure \ref{figure:strongvirtualpointsto}, 
+we see that a virtual address with a certain fractional permissions at each level of the page-table walk gives us sound 
+assurance of the path existence reaching to a page table entry, consequently, with a full-ownership of page address, 
+we can obtain and update the value mapped value at this physical page address through our virtual memory address.
 
-\todo[inline,color=red]{Colin says: Why are we talking about assertions for address translations here? It seems like we should
-\todo[inelin]{Colin says: There's no explanation of the split 52+12 \emph{notation} here. If we added such an explanation
-(maybe to where physical points-tos are introduced?) and moved this discussion to \ref{sec:overly-restrictive}, this is otherwise
-a nice explanation}
-discuss physical memory pointsto assertions in general first, then talk about this in \ref{sec:overly-restrictive}}
+The definition in Figure \ref{fig:strongvirtualpointsto} defines a virtual points-to assertion as
+existentially quantifying the physical memory addresses of the individual entries in each page table,
+asserting that a corresponding page table walk exists (\lstinline|L4_L1_PointsTo|), and then 
+\todo[inline]{snapshot... the last bit of that assertion is weird, but I'll fix. I'll also just remove the fraction from that def since it gets
+ignored anyways and we provide a better version shortly}
 Given the definition of physical page-pointsto assertion and the root address of virtual-address space as shown in Figure \ref{fig:pagetable}, one can build the physical address-translation for a virtual address (e.g.
@@
 \end{itemize}
-
-\subsection{An Overly-Restrictive Definition for Virtual Memory Addressing}
-\label{sec:overly-restrictive}
-One could give a naive logical definition to the virtual mememory addressing as conjuctions of each level translation 
-(Figure \fref{fig:strongvirtualpointsto}). The very first assurance our logical constructions need to be able assert is 
-the \textit{existence} of the page table walk. Our naive attempt shown in Figure \ref{figure:strongvirtualpointsto}, 
-we see that a virtual address with a certain fractional permissions at each level of the page-table walk gives us sound 
-assurance of the path existence reaching to a page table entry, consequently, with a full-ownership of page address, 
-we can obtain and update the value mapped value at this physical page address through our virtual memory address.
 

### 2023-04-14 programlogic.tex
getting there on section 4

@@
  Definition $\vaddr\mapsto_{\textsf{t}}\{\textsf{q}\}\; \vpage$ : vProp $\Sigma$ := 
-  $\exists_{\textsf{l4e l3e l2e l1}} \ldotp$  $\ulcorner$ aligned $\vaddr \urcorner \ast$ L4_L1_PointsTo(maddr l4e l3e l2e l1e) $\ast$ addr_L1$(\vaddr,\textsf{l1e}) \mapsto_{p} \vpage$.
+  $\exists_{\textsf{l4e l3e l2e l1}} \ldotp$  $\ulcorner$ aligned $\vaddr \urcorner \ast$ L4_L1_PointsTo(maddr l4e l3e l2e l1e) $\ast$ addr_L1$(\vaddr,\textsf{l1e}) \mapsto_{p}\{\mathsf{q}\} \vpage$.
  Definition L4_L1_PointsTo (maddr l4e l3e l2e l1e :word 64) : vProp $\Sigma$ :=
@@
 \end{lstlisting}
+\vspace{-1em}
 \caption{A Strong Virtual Points-to Relation}
-\todo[inline,color=red]{Colin says: shouldn't ther be a fraction q in the body of the virtual pointsto def?
-I know this is just a sketch that we then move away from, but it's worth getting accurate.}
-\todo[inline,color=cyan]{Colin says: the virtual points-to here is not defined as a vprop! the body is an iProp,
-containing an unbound variable rtv, which should be the cr3val taken as an argument. (Also, rtv doesn't make sense to me
-as a shorthand, why not cr3val?). I'm also pretty sure the maddr in the virtual points-to definition should be va.}
   \label{fig:strongvirtualpointsto}
@@
 \label{sec:overly-restrictive}
-One could give a naive logical definition to the virtual mememory addressing as conjuctions of each level translation 
-(Figure \fref{fig:strongvirtualpointsto}). The very first assurance our logical constructions need to be able assert is 
-the \textit{existence} of the page table walk. Our naive attempt shown in Figure \ref{figure:strongvirtualpointsto}, 
-we see that a virtual address with a certain fractional permissions at each level of the page-table walk gives us sound 
-assurance of the path existence reaching to a page table entry, consequently, with a full-ownership of page address, 
-we can obtain and update the value mapped value at this physical page address through our virtual memory address.
+A natural definition for a virtual points-to that depends on the current address space is to
+require that in order for a virtual address \textsf{va} to point to a value \textsf{v}, the assertion contains
+partial ownership of the physical memory involved in the page table walk that would translate \textsf{va} to
+its backing physical location --- with locations existentially quantified since a virtual points-to should not assert
+\emph{which} locations are accessed in a page table walk.
+Figure \ref{figure:strongvirtualpointsto} gives such a definition.
+It asserts the existence of for page table entries, one at each translation level, and via \lstinline|L4_L1_PointsTo|
+asserts from the current \lstinline|cr3| value, taking the offsets from \textsf{va} (as in Figure \ref{fig:pagetables})
+to index each level, finding the existentially quantified entries in each level, 
+ending with the L1 entry, whose physical page base is added to the page offset of \textsf{va} to obtain the physical address
+ --- and there (\lstinline|addr_L1(va,l1e)|) in physical memory, is the value.
+Most of the definition lives directly in \textsf{vProp}, using the separation logic structure lifted from Iris's \textsf{iProp}.
 
-The definition in Figure \ref{fig:strongvirtualpointsto} defines a virtual points-to assertion as
-existentially quantifying the physical memory addresses of the individual entries in each page table,
-asserting that a corresponding page table walk exists (\lstinline|L4_L1_PointsTo|), and then 
-\todo[inline]{snapshot... the last bit of that assertion is weird, but I'll fix. I'll also just remove the fraction from that def since it gets
-ignored anyways and we provide a better version shortly}
-Given the definition of physical page-pointsto assertion and the root address of virtual-address space as shown in Figure \ref{fig:pagetable}, one can build the physical address-translation for a virtual address (e.g.
+\lstinline|L4_L1_PointsTo| works by
+chaining together the entries for each level, using the sequence of table offsets from the address being translated to index
+each table level, and using the physical page address embedded in each entry.
+For example, 
+  the first level address translation to get the L4 entry (\lstinline|l4e|) 
+  uses the masks \textsf{l4M52} with the current \lstinline|cr3| to get the 
+  physical address of the start of the L4 table,
+  and \textsf{l4off} with the virtual address being translated to compute the correct offset within that table.
+  Then at that physical location is the appropriate entry in the L4 table \textsf{l4e}.
+  Subsequent levels of the page table walk work similarly.
+The statement of these assertions is simplified by the use of our split physical points-to assertions, since
+each level of tables is page-sized.\footnote{We do not address superpages and hugepages in this paper.}
+This helper definition is also a more explicit \textsf{vProp}, explicitly binding a value for \lstinline|cr3| and using it
+to start the translation process.
+
+% Given the definition of physical page-pointsto assertion and the root address of virtual-address space as shown in Figure \ref{fig:pagetable}, one can build the physical address-translation for a virtual address (e.g.
 \textsf{va}) via abstracting the L4-L1 table traversal as the following:
-\begin{itemize}
-  \item Level-4 Translation (L4): Performs the first level address translation to get the L4 entry (L4E) in Figure \ref{fig:pagetables} by using the masks l4M52 and l4off with \textsf{rtv} virtual base address to get the starting address of the L4 table (L4T), and locate the entry amongst 512 (q1) ones respectively.
-    \begin{lstlisting}[language=Coq]
-      $\hbox{(\TirNameStyle{{L4translate}})} \quad$ (l4M52 maddr rtv) \$\sim$\ (l4off maddr rtv) $\mapsto_{a}$ {q1} l4e 
-    \end{lstlisting}
- \item Level-3 Translation (L3): Performs the second level address translation to get the L3 entry (L3E) in Figure \ref{fig:pagetables} by using the masks l3M52 and l3off with L4 entry (l4e) to get the starting address of the L3 table (L3T), and locate the entry amongst 512 ones (q2) respectively.
-    \begin{lstlisting}[language=Coq]
-    $\hbox{(\TirNameStyle{{L3translate}})}$ (l3M52 maddr l4e) \$\sim$\ (l3off maddr l4e)  $\mapsto_{a}$ {q2} l3e
-    \end{lstlisting}
-  \item Level-2 Translation (L2): Performs the second level address translation to get the L2 entry (L2E) in Figure \ref{fig:pagetables} by using the masks l2M52 and l2off with L3 entry (l3e) to get the starting address of the L2 table (L2T), and locate the entry amongst 512 ones (q3) respectively.
-\begin{lstlisting}[language=Coq]
-    $\hbox{(\TirNameStyle{{L2translate}})}$ (l2M52 maddr l3e) \$\sim$\ (l2off maddr l3e)  $\mapsto_{a}$ {q3} l2e
-    \end{lstlisting}
-  \item Level-1 Translation (L1): Performs the second level address translation to get the L1 entry (L1E) in Figure \ref{fig:pagetables} by using the masks l1M52 and l1off with L2 entry (l2e) to get the starting address of the L1 table (L1T), and locate the entry amongst ones 512 (q4) respectively.
-   \begin{lstlisting}[language=Coq]
-    $\hbox{(\TirNameStyle{{L1translate}})}$ (l1M52 maddr l2e) \$\sim$\ (l1off maddr l2e)  $\mapsto_{a}$ {q4} l1e
-    \end{lstlisting}
-  \item Page Address Level Translation: Final computed physical page-address ($\textsf{addr\_L1}(\vaddr$,$\textsf{l1e}$)) points-to the value stored in the address ($\vpage$).
-   \begin{lstlisting}[language=Coq]
-    $\hbox{(\TirNameStyle{{PageLevelAccess}})} \qquad$ addr_L1($\vaddr,\textsf{l1e}$) $\mapsto_{p} \vpage$.
-    \end{lstlisting} 
-\end{itemize}
+% \begin{itemize}
+%   \item Level-4 Translation (L4): Performs 
+%   the first level address translation to get the L4 entry (L4E) in Figure \ref{fig:pagetables} by using the masks l4M52 and l4off with \textsf{rtv} virtual base address to get the starting address of the L4 table (L4T), and locate the entry amongst 512 (q1) ones respectively.
+%     \begin{lstlisting}[language=Coq]
+%       $\hbox{(\TirNameStyle{{L4translate}})} \quad$ (l4M52 maddr rtv) \$\sim$\ (l4off maddr rtv) $\mapsto_{a}$ {q1} l4e 
+%     \end{lstlisting}
+%  \item Level-3 Translation (L3): Performs the second level address translation to get the L3 entry (L3E) in Figure \ref{fig:pagetables} by using the masks l3M52 and l3off with L4 entry (l4e) to get the starting address of the L3 table (L3T), and locate the entry amongst 512 ones (q2) respectively.
+%     \begin{lstlisting}[language=Coq]
+%     $\hbox{(\TirNameStyle{{L3translate}})}$ (l3M52 maddr l4e) \$\sim$\ (l3off maddr l4e)  $\mapsto_{a}$ {q2} l3e
+%     \end{lstlisting}
+%   \item Level-2 Translation (L2): Performs the second level address translation to get the L2 entry (L2E) in Figure \ref{fig:pagetables} by using the masks l2M52 and l2off with L3 entry (l3e) to get the starting address of the L2 table (L2T), and locate the entry amongst 512 ones (q3) respectively.
+% \begin{lstlisting}[language=Coq]
+%     $\hbox{(\TirNameStyle{{L2translate}})}$ (l2M52 maddr l3e) \$\sim$\ (l2off maddr l3e)  $\mapsto_{a}$ {q3} l2e
+%     \end{lstlisting}
+%   \item Level-1 Translation (L1): Performs the second level address translation to get the L1 entry (L1E) in Figure \ref{fig:pagetables} by using the masks l1M52 and l1off with L2 entry (l2e) to get the starting address of the L1 table (L1T), and locate the entry amongst ones 512 (q4) respectively.
+%    \begin{lstlisting}[language=Coq]
+%     $\hbox{(\TirNameStyle{{L1translate}})}$ (l1M52 maddr l2e) \$\sim$\ (l1off maddr l2e)  $\mapsto_{a}$ {q4} l1e
+%     \end{lstlisting}
+%   \item Page Address Level Translation: Final computed physical page-address ($\textsf{addr\_L1}(\vaddr$,$\textsf{l1e}$)) points-to the value stored in the address ($\vpage$).
+%    \begin{lstlisting}[language=Coq]
+%     $\hbox{(\TirNameStyle{{PageLevelAccess}})} \qquad$ addr_L1($\vaddr,\textsf{l1e}$) $\mapsto_{p} \vpage$.
+%     \end{lstlisting} 
+% \end{itemize}
 
-\todo[inline]{Cite Kolanski here; not all of them, but some actually defined it exactly this way, for the current
-address space.}
+This solution is in fact very close to that of \citet{kolanski08vstte}, who define a separation logic from scratch in \textsc{Isabelle/HOL},
+where the semantics of all assertions are functions from pairs of heaps and page table root values to booleans.\footnote{
+  This was a typical explicit construction at the time; their work significantly predates Iris.
+}
+Our solution removes some restrictions in this definition by further abstracting the handling of address translation.
+
 \subsection{Aliasing/Sharing Physical Pages}
@@
   memory.
-  Updating any of these mappings (e.g., move the page tables in physical memory, as in coallescing for
+  Updating any of these mappings (e.g., moving the page tables in physical memory, as in coallescing for
   superpages or hugepages)
@@
 \begin{figure*}
-\todo[inline]{Colin says: again, this is supposed to be a vprop, but isn't doing anything with a cr3val argument.
-I realize the GASI and per-address-space invariants, and therefore machinery to tie the virtual points-to
-to the GASI, hasn't been introduced yet. Is there a short-hand here we can use to be ``more complete'' until the later text
-that fills in these details?
-}
-\todo[inline,color=cyan]{Colin says: Oh!!! I see. You have that extremely subtle rtv subscript on the virtual points-to
-syntax! Don't do that. First, it's really subtle and wasn't explained anywhere, so even I missed it. Second,
-we're very explicitly describing vProps as functions from root to iProp. Sticking the rtv on the arrow there
-is partly weird because it obscures which argument is really the vProp argument (there are 2 other word64s in that syntax),
-partly weird because it's written next to something that \emph{isn't} a variable (the v subscript on the virtual points-to
-arrow) while using the same font, and partly weird because when you actually write these assertions, rtv doesn't show
-up in the syntax at all, which is the whole point --- virtual points-to defs shouldn't mention their cr3 or address space at
-all!
-Here's what I think it should look like (I also existentially quantified $\delta$ and fixed the p subscript on physical
-points-to):
-}
 \[
 \begin{array}{l}
-    \vaddr\mapsto_{\textsf{v}}\{\textsf{q}\}\vale : \mathsf{vProp}~\Sigma \stackrel{\triangle}{=} 
+    \vaddr\mapsto_{\textsf{v}}\{\textsf{q}\}\vale : \mathsf{vProp}~\Sigma \stackrel{\triangle}{=} \\
+    \quad
     \exists \paddr\ldotp
     \exists \delta\ldotp
-    \sumapaces{\mathit{cr3val}}\delta \ast 
-  \sumwalkabs\vaddr\qfrac\paddr \ast \paddr \mapsto_{\mathsf{p}} \vale
+    \underbrace{\lambda\mathit{cr3val}\ldotp\sumapaces{\mathit{cr3val}}\delta}_\text{\begin{tabular}{c}Find ghost map for this\\address space\end{tabular}} \ast 
+  \underbrace{\sumwalkabs\vaddr\qfrac\paddr }_\text{\begin{tabular}{c}Fractional ownership of \\virt-to-phys translation\end{tabular}}\ast 
+  \underbrace{\paddr \mapsto_{\mathsf{p}} \vale}_\text{\begin{tabular}{c}Value at\\physical location\end{tabular}}
 \end{array}
 \]
-\todo[inline]{Colin says: The fact that we can't visually distinguish the variable v that is the value in memory
-from the subscript v on the virtual points-to arrow is not ideal: it makes it hard to tell what's concrete syntax
-and what's a variable. This is worth fixing, but near the bottom of the priority list}
 \caption{Virtual-Pointsto for Sharing Pages}
@@
 
-  \todo[inline]{Colin says: This requires explaining all the math here. Most reviewers of our paper will not be Iris
-  experts. Even the notion of an authoritative RA is something you'd need to explain (the general idea of).
-  So: write for non-Iris-expert readers. Above I explained the intuition behind the construction, so now we have to show
-  that we correctly implemented that intuition in Iris. So explain authoritative RAs, and how the line below
-  implements the intuition above.}
-  To do so, we introduce the protocol separating read-only (fragmental) and authorative (full ownership) accesses to the page tables, specifically $\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$.
-We realize this protocol by embedding it into an already existing \textsf{Iris}'s ghost-map (or view\_maps) construction
+  % To do so, we introduce the protocol separating read-only (fragmental) and authorative (full ownership) accesses to the page tables, specifically $\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$.
+  % We realize this protocol by embedding it into an already existing \textsf{Iris}'s ghost-map (or view\_maps) construction
+  In Iris this is realized by using an authoritative resource algebra: there is a single \emph{authoritative} global copy of the (ghost)
+  map caching virtual-to-physical address translations, and for each entry a read only \emph{partial} ownership of that key-value pair.
+  The resource algebra itself is instantiated as:
   \[\mathcal{A}\textsf{bsPTableWalk} \stackrel{\triangle}{=} \textsc{Auth} (\; \mathcal{W}_{64} \;\rightarrow_{\textrm{fin}} \;  (\mathcal{W}_{64} \; \rightarrow_{\textrm{fin}}\;  (\textsc{Frac }, \mathord{+}) \times (\textsc{Agree } \Loc,\mathord{=}) ))\]
- where
-  \begin{itemize}
-  \item we decorate the fragmental ownership of the ghost-map ($ \sumwalkabs\vaddr\qfrac\paddr$) with the physical $\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$ whose ownership just ensures that the virtual address ($\vaddr$) is mapped to a page (referenced by $\paddr$)
-  \item we place the authorative ownership $\mathcal{A}\textsf{PTableWalk}$ of the ghost-map in an address-space invariant ($\mathcal{I}$\textsf{ASpace} in Figure \ref{fig:peraspaceinvariant}) whose ownership grants sound update on the page table entries
-  \end{itemize}
+We replace the physical
+ $\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$
+with
+fragmental ownership of the ghost-map ($ \sumwalkabs\vaddr\qfrac\paddr$),
+ whose ownership just ensures that the virtual address ($\vaddr$) is mapped to a page (referenced by $\paddr$)
+
+We place the authorative ownership $\mathcal{A}\textsf{PTableWalk}$ of the ghost-map in a per-address-space invariant
+($\mathcal{I}$\textsf{ASpace} in Figure \ref{fig:peraspaceinvariant}) whose ownership, when combined with the
+fraction stored in the virtual points-to, permits changes in the virtual-to-physical translations.
+By itself, it allows changes to the page tables as long as they do not change the overall mappings from virtual
+to physical addresses.
+\todo[inline]{explain $\Theta$, $m$, and figure out where $\delta$ is bound in Figure \ref{fig:peraspaceinvariant}}
+
+Another major change is that, because our logic is the first to explicitly address the existence of multiple address spaces,
+the ghost map of virtual-to-physical translations (identified by the ghost name $\delta$ in Figure \ref{fig:virtualpointstosharing}
+ \emph{for each address space}. This means we also require a way to locate the map for the current address space
+from the current root page table pointer stored in \lstinline|cr3|.
+
+\
 
@@
 
-\todo[inline]{Colin says: I can rewrite this remark into flowing text after you fully explain the math above.}
-\begin{remark}[Exchanging Page-Table-Walk-Summarization Mappings (Tokens) under Mutation of Page Tables ]
-  $\mathcal{A}\textsf{PTableWalk}$'s fragmental-ownership represents an individual key-value pair (per-virtual-address) which can be handled as a \textit{look-up} resource for an individual virtual address.
-  To build more intuition on the convenience of this protocol, we can think of the use-case \textit{unmapping-pages}.
+  % Colin removed this remark because it's skwtching a *3rd* alternative design for virtual points-to,
+  % which could be useful for some readers, but will most likely just confuse
+% \begin{remark}[Exchanging Page-Table-Walk-Summarization Mappings (Tokens) under Mutation of Page Tables ]
+%   $\mathcal{A}\textsf{PTableWalk}$'s fragmental-ownership represents an individual key-value pair (per-virtual-address) which can be handled as a \textit{look-up} resource for an individual virtual address.
+%   To build more intuition on the convenience of this protocol, we can think of the use-case \textit{unmapping-pages}.
 Normally, in the existence of shared page-tables, we need to first collect all the fragmental map pairs, e.g.
@@
 So there should only be one physical memory resource inside the virtual pointsto definition: from physical page address to the value stored ($ \paddr \mapsto_{p} \vpage$).
-\end{remark}
+% \end{remark}
 

### 2023-04-14 relwork.tex
clean up related work, with a focus on comparing Kolanski & Klein to what we are submitting

--- 2022-09-20 relwork.tex
@@
 
-\textsc{CertiKOS}~\cite{gu15,gu2016certikos,gu2018certikos,chen2016interrupts} is a microkernel intended for use as a hypervisor, and its papers do not explicitly detail verification of the VMM, so we do not know the full space of which VMM functionality is verified (e.g., we do not know if it includes lazy page loading or copy-on-write behavior, or is a minimalistic VMM that eagerly maps address ranges for guessts).
-The overall approach in that body of work is many layers of refinement proofs, using a proliferation of layers with small differences to keep most individual refinements tractable.
-In keeping with precursor work on the project from the same group~\cite{vaynberg2012compositional}, the purpose of some layers is to abstract away from virtual memory, so the proof is essentially a simulation proof covering for example a proof that execution with page-in on page faults is a valid refinement of an execution model where.
-Another key aspect of their approach is that the OS is written in Clight and compiled with \textsc{CompCert}~\cite{blazy2006formal,leroy2009formally,leroy2008formal}.
-CompCert's memory abstraction~\cite{leroy2008formal} assumes
-memory is a set of disjoint chunks of bytes with no overlap, so the lowest levels of CertiKOS must provide a matching machine model as a layer.
-This prohibits virtual address aliasing, so CertiKOS cannot support simultaneous memory-mapped (\texttt{mmap}) and stream-oriented (\texttt{read}/\texttt{write}) IO to a single file\todo{should we go into this detail?}, and cannot use
-the common kernel design choice of mapping all physical memory into the bottom of the kernel's address space for direct access while the kernel code is simultaneously mapped (and executed) at higher virtual addresses.
-This is not necessary for \textsc{CertiKOS}'s intended primary use case (a hypervisor), but means that \textsc{CertiKOS}'s approach cannot be used to support this functionality in other systems, without major surgery to \textsc{CompCert}.
+\textsc{CertiKOS}~\cite{gu15,gu2016certikos,gu2018certikos,chen2016interrupts} is a microkernel intended for use as a hypervisor,
+and its papers do not explicitly detail verification of the VMM, so we do not know the full space of which VMM functionality 
+is verified, but we do know it includes the ability to map or unmap pages.
+The work is clear, however, that it trusts low-level assembly fragments such as the instruction sequence which actually
+switches address spaces, rather than verifying them.
+The overall approach in that body of work is many layers of refinement proofs, using a
+ proliferation of layers with small differences to keep most individual refinements tractable. In keeping with precursor work 
+on the project from the same group~\cite{vaynberg2012compositional}, the purpose of some layers is to abstract away from 
+virtual memory, so the proof is essentially a simulation proof covering for example a proof that execution with page-in on 
+page faults is a valid refinement of an execution model where no paging occurs.
+% Another key aspect of their approach is that the OS is written in Clight and compiled with \textsc{CompCert}~\cite{blazy2006formal,leroy2009formally,leroy2008formal}.
+% CompCert's memory abstraction~\cite{leroy2008formal} assumes
+% memory is a set of disjoint chunks of bytes with no overlap, so the lowest levels of CertiKOS must provide a matching 
+% machine model as a layer. This prohibits virtual address aliasing, so CertiKOS cannot support simultaneous memory-mapped 
+% (\texttt{mmap}) and stream-oriented (\texttt{read}/\texttt{write}) IO to a single file\todo{should we go into this detail?}, 
+% and cannot use
+% the common kernel design choice of mapping all physical memory into the bottom of the kernel's address space for direct access i
+% while the kernel code is simultaneously mapped (and executed) at higher virtual addresses.
+% This is not necessary for \textsc{CertiKOS}'s intended primary use case (a hypervisor), but means that \textsc{CertiKOS}'s
+% approach cannot be used to support this functionality in other systems, without major surgery to \textsc{CompCert}.
 
@@
 
-\todo[inline]{Need to come back and ack common weaknesses of above work and this work, like skipping the TLB (though there's newer stuff not integrated into seL4), and not treating the relationship between memory and instructions precisely.}
+One important outgrowth of the \textsc{seL4} project, not integrated into the main project's proof, was work by 
+Kolanski and Klein which studied verification of code against a hardware model that \emph{did} include address translation
+ --- the only work aside from ours to do so --- initially in terms of basic memory~\cite{kolanski08vstte} and subsequently 
+integrating source-level types into the interpretation~\cite{kolanski09tphols}. 
+They were the first work to model physical and virtual points-to assertions separately, defining virtual points-to assertions
+in terms of physical points-to assertions mimicking page table walks, and defining all of their assertions as predicates on a
+pair of (physical) machine memory and a page table root, an approach we improve on.
 
-One important outgrowth of the \textsc{seL4} project, not integrated into the main project's proof, was work by Kolanski and Klein which studied verification of code against a hardware model that \emph{did} include address translation --- the only work aside from ours to do so --- initially in terms of basic memory~\cite{kolanski08vstte} and subsequently integrating source-level types into the interpretation~\cite{kolanski09tphols}.
-They were the first work to model physical and virtual points-to assertions separately, defining virtual points-to assertions in terms of physical points-to assertions mimicking page table walks, and defining all of their assertions as predicates on a pair of (physical) machine memory and a page table root, an approach we improve on.
+Their work has a number of significant limitations which our work addresses.
+They also define their virtual points-to assertions such that a virtual points-to $p\mapsto_\mathsf{v} a$ owns the full 
+lookup path to virtual address $p$. This means that given two virtual points-to assertions at the same time, such as 
+$p\mapsto_\mathsf{v}a \ast p'\mapsto_\mathsf{v}b$, the memory locations traversed to translate $p$ and $p'$ must be disjoint. 
+This means the logic has a peculiar limit on how many virtual points-to assertions can coexist in a proof. Since page tables 
+fan out, the bottleneck is the number of entries in the root table. For their 32-bit ARMv6 example, the top-level address is 
+still 4Kb (4096 bytes), and each entry (consumed entirely by a virtual points-to in their scheme) is 4 bytes, so they have a 
+maximum of 1024 virtual points-tos in their ARMv6 configuration. Any assertion which implies more than that number
+of virtual addresses are mapped implies false in their logic.
+(They do formulate their logic over an abstract model, but every architecture would incur a similar limitation;
+Na\"ively transferring their model to x86-64 4-level tables would yield a limit of 512 assertions (also a 4Kb root page, 
+but 8-byte entries).
 
-Kolanski and Klein's points-to assertions do not model that page table entries for nearby addresses typically \emph{share} entries in higher layers of the page tables,\todo{actually, not sure, need to confirm} a key limitation that our definitions avoid.
-They do not consider virtuallv address aliasing (another source of sharing physical resources), which we\todo{todo!} support.
+% Kolanski and Klein's points-to assertions do not model that page table entries for nearby addresses typically 
+% \emph{share} entries in higher layers of the page tables --- a single L1 entry maps 4KB of memory on many architectures,
+% but their logics avoid fractional ownership, so they can in fact only use a single memory location per page of memory.
+% In fact, because their virtual points-to assertions contain \emph{full} ownership of all entries, even in the highest-level
+% page table (L4 in our case), each entry can contribute to only a single mapped address. Thus any assertion
+% in their logic that implies there are more virtual addresses mapped than entries in the top-level page table implies false.
+Our definitions make use of fractional permissions throughout; Figure \ref{fig:strongvirtualpointsto}'s definition
+of \lstinline|L4_L1_PointsTo| ellides the specific fractions used, but it in fact asserts 1/512 ownership of
+the L1 entry, 1/($512^2$) of the L2 entry, and so on, so each entry may map the appropriate number of machine words.
+
+As noted earlier, Kolanski and Klein's logics, by collocating both the physical ownership of the page table walk
+as part of the virtual points-to itself, preempt support for changes to page tables which do not actually affect 
+address translation.
+
+The other major distinction is that Kolanski and Klein have no accounting for other address spaces.
+Their logic does not deal with change of address space, and has no way to assert that certain facts hold
+in another address space.
 They verify only one address space manipulation: mapping a single unmapped page into the current address space (in both papers).
-The mapping write is verified in an ad hoc way by unfolding the machine semantics, because the logic lacks proper reusable rules for
-updating page tables.  
-Critically, they do not consider
-manipulating \emph{other} address spaces besides the one in use (essential for managing
-processes).
-So their means for sharing access to intermediate page table levels does not extend across address spaces, and there is no way to share data pages between address spaces.
-But every major OS includes widely-used facilities for sharing pages between processes (e.g., UNIX/POSIX \texttt{MAP\_SHARED}), and until the recent Spectre~\cite{kocher2018spectre} and Meltdown~\cite{lipp2018meltdown} attacks, FreeBSD, Solaris, and others shared the kernel mapping across processes~\cite{mckusick2014design,mcdougall2006solaris,wininternals6a,wininternals6b} (they still do so for performance on unaffected hardware; e.g., Meltdown affects Intel but not AMD~\cite{amdmeltdown}).
+We verify this, as well as a change-of-address-space, which requires us to introduce assertions for talking
+about other address spaces (we must know, for example, that the precondition of the code after the change must be true
+in the \emph{other} address space), and to deal with the fact that the standard frame rule
+for separation logic is unsound in the presence of address space changes and address-space-contingent assertions.
+% The mapping write is verified in an ad hoc way by unfolding the machine semantics, because the logic lacks proper reusable rules for
+% updating page tables.
 
-Our approach in this paper uses modalities to distinguish virtual address based assertions that hold only in specific address spaces, making it possible to manipulate other address spaces, and equally critically, to \emph{change} address spaces while reasoning about correctness.
-In addition, while Kolanski and Klein work in a traditional Hoare logic, we work in a separation logic, inheriting its modularity gains at the cost of working out some interactions between modalities and substructural logics.
+Our approach in this paper uses modalities to distinguish virtual-address-based assertions that hold only in specific 
+address spaces, making it possible to manipulate other address spaces, and equally critically, to \emph{change} address 
+spaces while reasoning about correctness.
 
-Unlike our work, Kolanski and Klein prove very useful embedding theorems stating that code that does not modify page table entries can be verified in a VM-ignorant program logic, and that proofs in that logic can be embedded into the VM-aware logic (essentially by interpreting ``normal'' points-to relations as virtual points-to facts).
-While we have not proven such a result, an analagous result \emph{should} hold of our work: consider that the triples for the \texttt{mov} instruction that access memory behave just as one would expect for a VM-ignorant logic~\todo{cite some example logic}.
-With our general approach to virtual points-to assertions being inspired by Kolanski and Klein, \emph{both} of our approaches could in principle be extended to account for pageable points-to assertions by adding additional disjunctions to an extended points-to definition.
-
-\todo[inline]{discuss hybrid+substructural logics}
+Unlike our work, Kolanski and Klein prove very useful embedding theorems stating that code that does not modify page table 
+entries can be verified in a VM-ignorant program logic, and that proofs in that logic can be embedded into the VM-aware logic 
+(essentially by interpreting ``normal'' points-to relations as virtual points-to facts). While we have not proven such a result,
+an analagous result {should} hold of our work: consider that the doubles for the \texttt{mov} instructions
+that access memory behave just as one would expect for a VM-ignorant logic~\cite{Chlipala2013Bedrock}.
+With our general approach to virtual points-to assertions being inspired by Kolanski and Klein, \emph{both}
+ our approach and theirs could in principle be extended to account for pageable points-to assertions by adding additional 
+disjunctions to an extended points-to definition; embedding ``regular'' separation logic into such a variant
+is the appropriate next step to extend reasoning to usermode programs running with a kernel that may demand-page the program's
+memory.

### 2023-04-14 programlogic.tex
tiny tweaks

@@
 \label{sec:aspacemanagement}
-\todo[inline]{Colin says: I will go back and add some foreshadowing in the explanation of virtual points-to,
-which already contains the Aspaces lookup a couple pages back.}
+\todo[inline]{Colin has not looked at this section in detail yet.
+It is clear the discussion of the cr3 update needs updating for doubles}
 So far, we have introduced logical abstractions for a single address space.
@@
     \hbox{(\TirNameStyle{ModalAddressSpace)}} \qquad
-         [r](P) \;: \mathsf{vProp} \; \Sigma \; \stackrel{\triangle}{:=} \; \lambda \mathsf{P},\; \mathsf{P r}. 
+         [r](P) \;: \mathsf{vProp} \; \Sigma \; \stackrel{\triangle}{:=} \; \lambda \_,\; \mathsf{P}~\mathsf{r}. 
   \end{array}

### 2023-04-14 soundness.tex
+
tiny tweaks

@@
 \section{Implementing Logical Machinery \& Soundness}
-We build our program logic, as an instantiation of Iris~\cite{iris}, and construct our reasoning principles including the modal one on top of it. 
-\mytodo{we should put some stuff here}
+We build our program logic, as an instantiation of Iris~\cite{jung2018iris}.
 \subsection{Soundness}
@@
 \inferrule{
-  \{P \ast r_d \mapsto_{r}  \textsf{v} \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{v,rtv}} \textsf{v} \}_{\textsf{rtv}}\;\overline{is}
+  \{P \ast r_d \mapsto_{r}  \textsf{v} \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{v}} \textsf{v} \}_{\textsf{rtv}}\;\overline{is}
 }{
-  \{P \ast r_d \mapsto_{r}  \textsf{rvd} \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{v,rtv}} \textsf{v} \}_{\textsf{rtv}}
-  \textsf{mov}~\textsf{r}_d,~\textsf{r}_a;\;\overline{is}
+  \{P \ast r_d \mapsto_{r}  \textsf{rvd} \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{v}} \textsf{v} \}_{\textsf{rtv}}
+  \textsf{ mov}~\textsf{r}_d,~\textsf{r}_a;\;\overline{is}
 }

### 2023-04-14 programlogic.tex
fix defs of ghost maps

@@
   The resource algebra itself is instantiated as:
-  \[\mathcal{A}\textsf{bsPTableWalk} \stackrel{\triangle}{=} \textsc{Auth} (\; \mathcal{W}_{64} \;\rightarrow_{\textrm{fin}} \;  (\mathcal{W}_{64} \; \rightarrow_{\textrm{fin}}\;  (\textsc{Frac }, \mathord{+}) \times (\textsc{Agree } \Loc,\mathord{=}) ))\]
+  \[\mathcal{A}\textsf{bsPTableWalk} \stackrel{\triangle}{=} \textsc{Auth} (\; \mathcal{W}_{64} \;\rightarrow_{\textrm{fin}} \;  ( (\textsc{Frac }, \mathord{+}) \times (\textsc{Agree } \Loc,\mathord{=}) ))\]
 We replace the physical
@@
 However, \textsf{VMM} implementations handle more than one address-space which we abstract as a ghost map from root address of an address-space to a namespace labelling the address space
-\[\mathcal{A}\textsf{spaces} \stackrel{\triangle}{=} \textsc{Auth} (\; \mathcal{W}_{64} \;\rightarrow_{\textrm{fin}} \; (\textsc{Agree } \mathsf{GName},\mathord{=}) ) \]
+\[\mathcal{A}\textsf{spaces} \stackrel{\triangle}{=} \textsc{Auth} (\; \mathcal{W}_{64} \;\rightarrow_{\textrm{fin}} \;  ( (\textsc{Frac }, \mathord{+}) \times (\textsc{Agree } \Loc,\mathord{=}) )) \]
 

### 2023-04-14 programlogic.tex
+

@@
  $\lambda$ cr3val.
+  $\ulcorner$ entry_present l4e $\land$ entry_present l3e $\land$ entry_present l2e $\land$ entry_present l1e$\urcorner$ $\ast$
   (l4M52 maddr cr3val) \$\sim$\ (l4off maddr cr3val) $\mapsto_{a}$ {q1}  l4e  $\ast$

### 2023-04-14 programlogic.tex
+

--- 2023-04-14 programlogic.tex
@@
  $\lambda$ cr3val.
-  (l4M52 maddr cr3val) \$\sim$\ (l4off maddr cr3val) $\mapsto_{a}$ {q1}  l4e  $\ast$
-  (l3M52 maddr l4e) \$\sim$\ (l3off maddr l4e)  $\mapsto_{a}$ {q2}  l3e $\ast$ 
-  (l2M52 maddr l3e) \$\sim$\ (l2off maddr l3e) $\mapsto_{a}$ {q3}  l2e $\ast$
-  (l1M52 maddr l2e) \$\sim$\ (l1off maddr l2e) $\mapsto_{a}$ {q4}  l1e.
+  $\ulcorner$ entry_present l4e $\land$ entry_present l3e $\land$ entry_present l2e $\land$ entry_present l1e$\urcorner$ $\ast$
+  (l4M52 maddr cr3val) $\sim$ (l4off maddr cr3val) $\mapsto_{a}$ {q1}  l4e  $\ast$
+  (l3M52 maddr l4e) $\sim$ (l3off maddr l4e)  $\mapsto_{a}$ {q2}  l3e $\ast$ 
+  (l2M52 maddr l3e) $\sim$ (l2off maddr l3e) $\mapsto_{a}$ {q3}  l2e $\ast$
+  (l1M52 maddr l2e) $\sim$ (l1off maddr l2e) $\mapsto_{a}$ {q4}  l1e.
 \end{lstlisting}

### 2023-04-14 programlogic.tex
+

@@
 \caption{A Strong Virtual Points-to Relation}
+\todo[inline,color=red]{Colin, I updated the definition with entry present}
   \label{fig:strongvirtualpointsto}
@@
   \begin{array}{l}
-    \hbox{(\TirNameStyle{ModalAddressSpace)}} \qquad
-         [r](P) \;: \mathsf{vProp} \; \Sigma \; \stackrel{\triangle}{:=} \; \lambda \_,\; \mathsf{P}~\mathsf{r}. 
+    \hbox{(\TirNameStyle{ModalAddressSpace})} \qquad
+         [r](\mathsf{P}) \;: \mathsf{vProp} \; \Sigma \; \stackrel{\triangle}{=} \; \lambda \_,\; \mathsf{P}~\mathsf{r}.\\
+    \hbox{(\TirNameStyle{ModalAddressSpaceMono})} \qquad (\mathsf{P} \vdash \mathsf{Q}) \vdash  ([r](\mathsf{P}) \vdash  [r](\mathsf{Q}))\\
+    \hbox{(\TirNameStyle{ModalAddressSpaceSep})} \qquad [r](\mathsf{P} \ast \mathsf{Q}) \dashv\vdash ([r](\mathsf{P}) \ast [r](\mathsf{Q})\\
+    \hbox{(\TirNameStyle{ModalAddressSpacePure})} \qquad  \mathsf{pure\_vprop P}(\mathsf{iPropI }\Sigma)\textsf{ : vProp }\Sigma \stackrel{\triangle}{=} \; \lambda \_ \ldotp \mathsf{P}\\
+    \hbox{(\TirNameStyle{ModalAddressSpaceFact})} \qquad  \mathsf{Fact P}(\mathsf{vProp }\Sigma) \stackrel{\triangle}{=} \;\forall \; ,r \; r' \ldotp  \mathsf{P r} \dashv\vdash \mathsf{P r'}\\
+    \hbox{(\TirNameStyle{ModalAddressSpaceFactIntroElim})} \\
+    \qquad\qquad \mathsf{Fact pure\_vprop P} \vdash  ([r](\mathsf{pure\_vprop P}) \dashv\vdash (\mathsf{pure\_vprop P}) (\mathsf{vProp }\Sigma))\\
+
+    \hbox{(\TirNameStyle{ModalAddressSpaceAnd})} \qquad   [r](\mathsf{P}\land \mathsf{Q}) \dashv\vdash [r](\mathsf{P})\land [r](\mathsf{Q})\\    
+    \hbox{(\TirNameStyle{ModalAddressSpaceOr})} \qquad    [r](\mathsf{P}\lor \mathsf{Q}) \dashv\vdash [r](\mathsf{P})\lor [r](\mathsf{Q})\\   
+    \hbox{(\TirNameStyle{ModalAddressSpaceFactIntroElim})} \\
+    \qquad\qquad \mathsf{Fact pure\_vprop P} \rightarrow  ([r](\mathsf{pure\_vprop P}) \dashv\vdash (\mathsf{pure\_vprop P}) (\mathsf{vProp }\Sigma))\\
+
   \end{array}

### 2023-04-14 examples.tex
tiny tweaks
+

@@
 \label{sec:experiment}
-To both validate and demonstrate the value of the modal approach to reasoning about virtual memory management, we study several distillations of real concerns of virtual memory managers.
+To both validate and demonstrate the value of the modal approach to reasoning about virtual memory management, we study several
+distillations of real concerns of virtual memory managers.
 Recall from Section \ref{sec:logic} that virtual points-to assertions work just like regular points-to assertions, by design.
@@
 call ensure_L1_page
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast  \texttt{r14}\mapsto_{\textsf{r}} \_ \ast \texttt{rdi}\mapsto_{r} \vaddr \ast \texttt{rax}\mapsto_{\textsf{r}} \textsf{ pte\_addr} \; \ast }_{\rtv}$
-$\specline{\exists (\entryf ,\;\entrytr,\; \entrytw,\; \entryo,\;\textsf{pte\_addr },\paddr \; : \texttt{word 64}) \ldotp \ulcorner \texttt{addr\_L1 }(\vaddr, \entryo) = \paddr \urcorner \; \ast }_{\rtv}$
+$\specline{\exists (\entryf ,\;\entrytr,\; \entrytw,\; \entryo,\;\textsf{pte\_addr },\paddr) \; \ldotp\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast  \texttt{r14}\mapsto_{\textsf{r}} \_ \ast \texttt{rdi}\mapsto_{r} \vaddr \ast \texttt{rax}\mapsto_{\textsf{r}} \textsf{ pte\_addr} \; \ast }_{\rtv}$
+$\specline{ \ulcorner  \texttt{addr\_L1 }(\vaddr, \entryo) = \paddr \urcorner \ast \ulcorner \texttt{entry\_present } \entryf \land \texttt{entry\_present } \entrytr \land  \texttt{entry\_present } \entrytw \urcorner \; \ast}_{\rtv}$
 $\specline{\nfpointsto{\mask\vaddr\maskfour\rtv}{\mask\vaddr\maskfouroff\rtv}\entryf\qone\naddr \; \ast}_{\rtv}$ 
@@
 $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast  \texttt{r14}\mapsto_{\textsf{r}} \texttt{pte\_addr} \ast \texttt{rdi}\mapsto_{\textsf{r}} \vaddr \ast \texttt{rax}\mapsto_{\textsf{r}} \textsf{ pte\_addr} \; \ast }_{\rtv}$
-$\specline{ \nfpointsto{\mask\vaddr\maskfour\rtv}{\mask\vaddr\maskfouroff\rtv}\entryf\qone\naddr \; \ast}_{\rtv}$ 
+$\specline{ \nfpointsto{\mask\vaddr\maskfour\rtv}{\mask\vaddr\maskfouroff\rtv}\entryf\qone\naddr \ast \ulcorner \texttt{entry\_present } \entryf \land \texttt{entry\_present } \entrytr \land  \texttt{entry\_present } \entrytw \urcorner \ast}_{\rtv}$
 $\specline{  \nfpointsto{\mask\vaddr\maskthree\entryf}{\mask\vaddr\maskthreeoff\entryf}\entrytr\qtwo\naddr \ast \nfpointsto{\mask\vaddr\masktwo\entrytr}{\mask\vaddr\masktwooff\entrytr}\paddr\qthree\naddr \;\ast}_{\rtv}$
@@
 $\specline{  \nfpointsto{\mask\vaddr\maskthree\entryf}{\mask\vaddr\maskthreeoff\entryf}\entrytr\qtwo\naddr \ast \nfpointsto{\mask\vaddr\masktwo\entrytr}{\mask\vaddr\masktwooff\entrytr}\paddr\qthree\naddr \;\ast}_{\rtv}$
-$\specline{\texttt{pte\_addr} \mapsto_{\texttt{vpte}} \paddr \;(\texttt{wzero 64}) \; \ast }_{\rtv}$
-$\specline{\exists \texttt{ fpaddr} \ldotp \ulcorner \texttt{aligned fpaddr} \urcorner \ast \texttt{rax}\mapsto_{\textsf{r}} \texttt{fpaddr} \ast \texttt{fpaddr} \mapsto_{\textsf{a}} (\texttt{wzero 64})}_{\rtv}$
+$\specline{\texttt{pte\_addr} \mapsto_{\texttt{vpte}} \paddr\; (\texttt{wzero 64})  \ast \ulcorner \texttt{entry\_present } \entryf \land \texttt{entry\_present } \entrytr \land  \texttt{entry\_present } \entrytw \urcorner}_{\rtv}$
+$\specline{\exists \texttt{ fpaddr} \ldotp \ulcorner \texttt{aligned fpaddr} \urcorner \ast \texttt{rax}\mapsto_{\textsf{r}} \texttt{fpaddr} \ast \texttt{fpaddr} \mapsto_{\textsf{a}} (\texttt{wzero 64})  \ulcorner \texttt{entry\_present fpaddr}\urcorner}_{\rtv}$
 ;; Calculate new L1 entry
@@
 $\specline{  \nfpointsto{\mask\vaddr\maskthree\entryf}{\mask\vaddr\maskthreeoff\entryf}\entrytr\qtwo\naddr \ast \nfpointsto{\mask\vaddr\masktwo\entrytr}{\mask\vaddr\masktwooff\entrytr}\paddr\qthree\naddr \;\ast}_{\rtv}$
-$\specline{\texttt{pte\_addr} \mapsto_{\texttt{vpte}} \paddr \;(\texttt{fpaddr}) \; \ast }_{\rtv}$
-$\specline{\exists \texttt{ fpaddr} \ldotp \ulcorner \texttt{aligned fpaddr} \urcorner \ast \texttt{rax}\mapsto_{r} \texttt{fpaddr} \ast \texttt{fpaddr} \mapsto_{a} (\texttt{wzero 64})}_{\rtv}$
+$\specline{\texttt{pte\_addr} \mapsto_{\texttt{vpte}} \paddr \;(\texttt{fpaddr}) \; \ast \ulcorner \texttt{entry\_present } \entryf \land \texttt{entry\_present } \entrytr \land  \texttt{entry\_present } \entrytw \urcorner }_{\rtv}$
+$\specline{\exists \texttt{ fpaddr} \ldotp \ulcorner \texttt{aligned fpaddr} \urcorner \ast \texttt{rax}\mapsto_{r} \texttt{fpaddr} \ast \texttt{fpaddr} \mapsto_{a} (\texttt{wzero 64})  \ulcorner \texttt{entry\_present fpaddr}\urcorner}_{\rtv}$
 $\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\; \sqsubseteq $
@@
 \begin{itemize}
-\item physical pointsto assertions on the tables L4, L3, and L2 reaching to the L1 level entry (l1e) ensures the successful table lookups (Specification Lines 4-6)
+\item physical pointsto assertions on the tables L4, L3, and L2 reaching to the L1 level entry (l1e) ensures the successful table lookups with the fact on \textit{presence} (\textsf{entry\_present}) of the relevant entries (\textsf{l4e l3e l2e}) (Specification Lines 4-6)
 \item ,and a virtual L1 pointsto ($\mapsto_{\textsf{vpte}}$) on the address computed from the reached L1 entry ($ \ulcorner \texttt{addr\_L1}(\vaddr,\entryo) = \paddr\urcorner$) which is almost exactly same with the virtual-pointsto definition in Figure \ref{fig:virtualpointstosharing} except the physical page address is already computed, not exsitentially quantified:
@@
 \[\texttt{pte\_addr} \mapsto_{\texttt{vpte}}  \; \paddr\; \textsf{fpaddr}  \qquad\qquad (\textsf{by Rule} \TirNameStyle{WriteToVirtMemFromReg})\]
-by Rule \TirNameStyle{WriteToVirtMemFromReg} which allows proving virtual address updates without exposing its internal physical page-table memory accesses (Specification Line 26). 
+
+by Rule \TirNameStyle{WriteToVirtMemFromReg} which allows proving the fact that virtual address updates without exposing its internal physical page-table memory accesses (Specification Line 26). 
 Consequently, right at this point in the proof, when we unfold the $\texttt{pte\_addr} \mapsto_{\texttt{vpte}}  \; \paddr\; \textsf{fpaddr})$ we will be able obtain a full ownerhsip on the L1 table physical pointsto $\paddr \mapsto_{\mathsf{p}} \textsf{fpaddr}$ .

### 2023-04-14 programlogic.tex
pure stuff on paper
+

--- 2023-04-14 programlogic.tex
@@
   \textsf{ where } \\
-   \textsf{ASpace\_Lookup} (\ptablestore,m) \stackrel{\triangle}{=} \lambda\textsf{ cr3val} \ldotp \; \ulcorner m \; !!\; \textsf{cr3val} = \textsf{Some } \gammaPred \urcorner \ast
+   \textsf{ASpace\_Lookup} (\ptablestore,m) \stackrel{\triangle}{=} \lambda\textsf{ cr3val} \ldotp \; \exists \gammaPred \; \ldotp \ulcorner m \; !!\; \textsf{cr3val} = \textsf{Some } \gammaPred \urcorner \ast
     \ownGhost\gammaPred{\authfull{\ptableabswalk\ptablestore}}
@@
   \begin{array}{l}
-    \hbox{(\TirNameStyle{ModalAddressSpace)}} \qquad
-         [r](P) \;: \mathsf{vProp} \; \Sigma \; \stackrel{\triangle}{:=} \; \lambda \_,\; \mathsf{P}~\mathsf{r}. 
+    \hbox{(\TirNameStyle{ModalAddressSpace})} \qquad
+         [r](\mathsf{P}) \;: \mathsf{vProp} \; \Sigma \; \stackrel{\triangle}{=} \; \lambda \_,\; \mathsf{P}~\mathsf{r}.\\
+    \hbox{(\TirNameStyle{ModalAddressSpaceMono})} \qquad (\textsf{P} \vdash \textsf{Q}) \vdash  ([r](\textsf{P}) \vdash  [r](\textsf{Q}))\\
+    \hbox{(\TirNameStyle{ModalAddressSpaceSep})} \qquad [r](\textsf{P} \ast \textsf{Q}) \dashv\vdash ([r](\textsf{P}) \ast [r](\textsf{Q})\\
+    \hbox{(\TirNameStyle{ModalAddressSpacePure})} \qquad  \textsf{pure\_vprop P}(\textsf{iPropI }\Sigma)\textsf{ : vProp }\Sigma \stackrel{\triangle}{=} \; \lambda \_ \ldotp \textsf{P}\\
+    \hbox{(\TirNameStyle{ModalAddressSpaceFact})} \qquad  \textsf{Fact P}(\textsf{vProp }\Sigma) \stackrel{\triangle}{=} \;\forall \; ,r \; r' \ldotp  \textsf{P r} \dashv\vdash \textsf{P r'}\\
+    \hbox{(\TirNameStyle{ModalAddressSpaceFactIntroElim})} \\
+    \qquad\qquad \textsf{Fact pure\_vprop P} \vdash  ([r](\textsf{pure\_vprop P}) \dashv\vdash (\textsf{pure\_vprop P}) (\textsf{vProp }\Sigma))\\
+
+    \hbox{(\TirNameStyle{ModalAddressSpaceAnd})} \qquad   [r](\textsf{P}\land \textsf{Q}) \dashv\vdash [r](\textsf{P})\land [r](\textsf{Q})\\    
+    \hbox{(\TirNameStyle{ModalAddressSpaceOr})} \qquad    [r](\textsf{P}\lor \textsf{Q}) \dashv\vdash [r](\textsf{P})\lor [r](\textsf{Q})\\   
+    \hbox{(\TirNameStyle{ModalAddressSpaceFactIntroElim})} \\
+    \qquad\qquad \textsf{Fact pure\_vprop P} \vdash  ([r](\textsf{pure\_vprop P}) \dashv\vdash (\textsf{pure\_vprop P}) (\textsf{vProp }\Sigma))\\
+
   \end{array}
   \]
+  \todo[inline,color=red]{Colin, story about any of the stuff above}
   \label{fig:modaldef}

### 2023-04-14 examples.tex
+

@@
 $\specline{\nfpointsto{\mask\vaddr\maskfour\rtv}{\mask\vaddr\maskfouroff\rtv}\entryf\qone\naddr \; \ast}_{\rtv}$ 
-$\specline{  \nfpointsto{\mask\vaddr\maskthree\entryf}{\mask\vaddr\maskthreeoff\entryf}\entrytr\qtwo\naddr \ast \nfpointsto{\mask\vaddr\masktwo\entrytr}{\mask\vaddr\masktwooff\entrytr}\paddr\qthree\naddr \;\ast}_{\rtv}$
+$\specline{  \nfpointsto{\mask\vaddr\maskthree\entryf}{\mask\vaddr\maskthreeoff\entryf}\entrytr\qtwo\naddr \ast \nfpointsto{\mask\vaddr\masktwo\entrytr}{\mask\vaddr\masktwooff\entrytr}\paddr\qthree\entryo \;\ast}_{\rtv}$
 $\specline{\texttt{pte\_addr} \mapsto_{\texttt{vpte}} \paddr \;(\texttt{wzero 64}) \ast \texttt{rax}\mapsto_{\textsf{r}} \texttt{pte\_addr}  }_{\rtv}$
@@
 $\specline{ \nfpointsto{\mask\vaddr\maskfour\rtv}{\mask\vaddr\maskfouroff\rtv}\entryf\qone\naddr \ast \ulcorner \texttt{entry\_present } \entryf \land \texttt{entry\_present } \entrytr \land  \texttt{entry\_present } \entrytw \urcorner \ast}_{\rtv}$
-$\specline{  \nfpointsto{\mask\vaddr\maskthree\entryf}{\mask\vaddr\maskthreeoff\entryf}\entrytr\qtwo\naddr \ast \nfpointsto{\mask\vaddr\masktwo\entrytr}{\mask\vaddr\masktwooff\entrytr}\paddr\qthree\naddr \;\ast}_{\rtv}$
+$\specline{  \nfpointsto{\mask\vaddr\maskthree\entryf}{\mask\vaddr\maskthreeoff\entryf}\entrytr\qtwo\naddr \ast \nfpointsto{\mask\vaddr\masktwo\entrytr}{\mask\vaddr\masktwooff\entrytr}\paddr\qthree\entryo \;\ast}_{\rtv}$
 $\specline{\texttt{pte\_addr} \mapsto_{\texttt{vpte}} \paddr \;(\texttt{wzero 64}) \ast \texttt{rax}\mapsto_{\textsf{r}} \texttt{pte\_addr}  }_{\rtv}$
@@
 $\specline{\texttt{pte\_addr} \mapsto_{\texttt{vpte}} \paddr\; (\texttt{wzero 64})  \ast \ulcorner \texttt{entry\_present } \entryf \land \texttt{entry\_present } \entrytr \land  \texttt{entry\_present } \entrytw \urcorner}_{\rtv}$
-$\specline{\exists \texttt{ fpaddr} \ldotp \ulcorner \texttt{aligned fpaddr} \urcorner \ast \texttt{rax}\mapsto_{\textsf{r}} \texttt{fpaddr} \ast \texttt{fpaddr} \mapsto_{\textsf{a}} (\texttt{wzero 64})  \ulcorner \texttt{entry\_present fpaddr}\urcorner}_{\rtv}$
+$\specline{\exists \texttt{ fpaddr} \ldotp \ulcorner \texttt{aligned fpaddr} \urcorner \ast \texttt{rax}\mapsto_{\textsf{r}} \texttt{fpaddr} \ast \texttt{fpaddr} \mapsto_{\textsf{p}} (\texttt{wzero 64})  \ulcorner \texttt{entry\_present fpaddr}\urcorner}_{\rtv}$
 ;; Calculate new L1 entry
@@
 $\specline{ \nfpointsto{\mask\vaddr\maskfour\rtv}{\mask\vaddr\maskfouroff\rtv}\entryf\qone\naddr \ast}_{\rtv}$ 
-$\specline{  \nfpointsto{\mask\vaddr\maskthree\entryf}{\mask\vaddr\maskthreeoff\entryf}\entrytr\qtwo\naddr \ast \nfpointsto{\mask\vaddr\masktwo\entrytr}{\mask\vaddr\masktwooff\entrytr}\paddr\qthree\naddr \;\ast}_{\rtv}$
+$\specline{  \nfpointsto{\mask\vaddr\maskthree\entryf}{\mask\vaddr\maskthreeoff\entryf}\entrytr\qtwo\naddr \ast \nfpointsto{\mask\vaddr\masktwo\entrytr}{\mask\vaddr\masktwooff\entrytr}\paddr\qthree\entryo \;\ast}_{\rtv}$
 $\specline{\texttt{pte\_addr} \mapsto_{\texttt{vpte}} \paddr \;(\texttt{fpaddr}) \; \ast \ulcorner \texttt{entry\_present } \entryf \land \texttt{entry\_present } \entrytr \land  \texttt{entry\_present } \entrytw \urcorner }_{\rtv}$
-$\specline{\exists \texttt{ fpaddr} \ldotp \ulcorner \texttt{aligned fpaddr} \urcorner \ast \texttt{rax}\mapsto_{r} \texttt{fpaddr} \ast \texttt{fpaddr} \mapsto_{a} (\texttt{wzero 64})  \ulcorner \texttt{entry\_present fpaddr}\urcorner}_{\rtv}$
+$\specline{\exists \texttt{ fpaddr} \ldotp \ulcorner \texttt{aligned fpaddr} \urcorner \ast \texttt{rax}\mapsto_{r} \texttt{fpaddr} \ast \texttt{fpaddr} \mapsto_{p} (\texttt{wzero 64})  \ulcorner \texttt{entry\_present fpaddr}\urcorner}_{\rtv}$
 $\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\; \sqsubseteq $
 $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast  \texttt{r14}\mapsto_{\textsf{r}} \texttt{pte\_addr} \ast \texttt{rdi}\mapsto_{\textsf{r}} \vaddr \ast }_{\rtv}$
-$\specline{\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\vaddr,\entryf,\entrytr,\entrytw,\paddr) \ast \ulcorner \theta \;!!\;\vaddr = \texttt{None}\urcorner \; \ast}_{\rtv}$
-$\specline{\ulcorner \texttt{aligned fpaddr} \urcorner \ast \texttt{rax}\mapsto_{\textsf{r}} \texttt{fpaddr} \ast \texttt{fpaddr} \mapsto_{\textsf{a}} (\texttt{wzero 64}) }_{\rtv}$
+$\specline{\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\vaddr,\entryf,\entrytr,\entrytw,\fpaddr) \ast \ulcorner \theta \;!!\;\vaddr = \texttt{None}\urcorner \; \ast}_{\rtv}$
+$\specline{\ulcorner \texttt{aligned fpaddr} \urcorner \ast \texttt{rax}\mapsto_{\textsf{r}} \texttt{fpaddr} \ast \texttt{fpaddr} \mapsto_{\textsf{p}} (\texttt{wzero 64}) }_{\rtv}$
 $\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\; \sqsubseteq $
@@
 
-The crux point in the proof is updating the virtual L1 address (\textsf{pte\_addr}) to show the fresly allocated page address (\textsf{fpaddr}) in Line 22 to obtain
+The crux point in the proof is updating L1 table through a virtual address (\textsf{pte\_addr}) through making a known physical address, in our example L1 table entry address ($\textsf{addr\_L1}(\textsf{va, l1e})$), to show the fresly allocated page address (\textsf{fpaddr}) in Line 22 to obtain
 \[\texttt{pte\_addr} \mapsto_{\texttt{vpte}}  \; \paddr\; \textsf{fpaddr}  \qquad\qquad (\textsf{by Rule} \TirNameStyle{WriteToVirtMemFromReg})\]
-
-by Rule \TirNameStyle{WriteToVirtMemFromReg} which allows proving the fact that virtual address updates without exposing its internal physical page-table memory accesses (Specification Line 26). 
+by Rule \TirNameStyle{WriteToVirtMemFromReg} (Specification Line 26). 
 Consequently, right at this point in the proof, when we unfold the $\texttt{pte\_addr} \mapsto_{\texttt{vpte}}  \; \paddr\; \textsf{fpaddr})$ we will be able obtain a full ownerhsip on the L1 table physical pointsto $\paddr \mapsto_{\mathsf{p}} \textsf{fpaddr}$ .
-Then, we can split this full-ownership into to obtain a fractional ownership (q4) on it ($\paddr \mapsto_{\mathsf{p}} \{q4\} \;\textsf{fpaddr}$).
-Finally, we can connect this fragmental ownership with the L4-L2 physical table pointstos (Specification Lines 24 and 25) to obtain the complete physical table-walk assertion (L4\_L1\_PointsTo(maddr l4e l3e l2e pa)) in Line 30.
+Then, we can first discard the allocated virtual-address pointsto assertion (i.e.
+dropping \textsf{pte\_addr}), and split the full-ownership of physical-page pointsto ($\textsf{fpaddr}\mapsto_{\textsf{p}} \;\textsf{wzero 64}$) to obtain a fractional ownership (q4) on it.
+Since we know $\texttt{addr\_L1 }(\vaddr, \entryo) \mapsto_{\mathsf{p}} \{q4\} \;\textsf{fpaddr}$) from $\mapsto_{\textsf{vpte}}$, after the drop of virtual-pointsto, we still can connect L4-L2 pointsto assertions (in Specification Lines 24 and 25) with newly allocated $\texttt{fpaddr} \mapsto_{a} (\texttt{wzero 64})$ to obtain the complete physical table-walk assertion (L4\_L1\_PointsTo(maddr l4e l3e l2e fpaddr)) in Line 30.
 

### 2023-04-14 programlogic.tex
4.3 revised

@@
 to physical addresses.
-\todo[inline]{explain $\Theta$, $m$, and figure out where $\delta$ is bound in Figure \ref{fig:peraspaceinvariant}}
-
-Another major change is that, because our logic is the first to explicitly address the existence of multiple address spaces,
-the ghost map of virtual-to-physical translations (identified by the ghost name $\delta$ in Figure \ref{fig:virtualpointstosharing}
- \emph{for each address space}. This means we also require a way to locate the map for the current address space
-from the current root page table pointer stored in \lstinline|cr3|.
-
-\
 
@@
 
-  % Colin removed this remark because it's skwtching a *3rd* alternative design for virtual points-to,
-  % which could be useful for some readers, but will most likely just confuse
-% \begin{remark}[Exchanging Page-Table-Walk-Summarization Mappings (Tokens) under Mutation of Page Tables ]
-%   $\mathcal{A}\textsf{PTableWalk}$'s fragmental-ownership represents an individual key-value pair (per-virtual-address) which can be handled as a \textit{look-up} resource for an individual virtual address.
-%   To build more intuition on the convenience of this protocol, we can think of the use-case \textit{unmapping-pages}.
-Normally, in the existence of shared page-tables, we need to first collect all the fragmental map pairs, e.g.
-$\sumwalkabs\vaddr\qfrac\paddr$, then, after making sure of non-existence of \textit{sharing}, we can \textit{unmap} the page.
-In this case, it would be ok to put $\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$ and $ \sumwalkabs\vaddr\qfrac\paddr$ side-by-side inside the virtual-pointsto definition in Figure \ref{fig:virtualpointstosharing}.
-However, this pattern of verifying is not only inconvenient due to requirement of extra bookkeeping, but also violates our soundness argument based on considering  mutation on the shared page-tables -- due to coalescing of page tables etc.
---.
-In other words, physical $\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$ sitting aside to  $ \sumwalkabs\vaddr\qfrac\paddr$ would be made invalid by any concurrent update on the physical page-table.
-So there should only be one physical memory resource inside the virtual pointsto definition: from physical page address to the value stored ($ \paddr \mapsto_{p} \vpage$).
-% \end{remark}
+Another major change is that, because our logic is the first to explicitly address the existence of multiple address spaces,
+the ghost map of virtual-to-physical translations (identified by the ghost name $\delta$ in Figure \ref{fig:virtualpointstosharing}
+ \emph{for each address space}. This means we also require a way to locate the map for the current address space
+from the current root page table pointer stored in \lstinline|cr3|.
 
-%\subsection{Address-Space Abstraction}
-%\label{sec:aspacemodal}
+Figure \ref{fig:peraspaceinvariant} describes the resources for each address space, which are collected
+for the full address space, but separated from virtual points-to assertions.
+$\Theta$ is the logical map from virtual addresses to the physical addresses they should translate to, 
+for the \emph{current} address space, corresponding to the currently installed page tables (the ones
+indicated by \lstinline|cr3|). In Figure \ref{fig:virtualpointstosharing}, the existentially-quantified $\delta$
+is the ghost name for the authoritative copy of this map (again, for the current address space).
+$m$ is the (full) logical map from the various page table bases to ghost names for the 
+per-address-space mappings like $\Theta$.
+It is the full version of the map named by the ghost name $\delta{}s$ in
+Figure \ref{fig:virtualpointstosharing}; there is only one such map in the system.
+This is why Figure \ref{fig:virtualpointstosharing} asserts there is some ghost name in that (global) map
+for the current address space.
 
-Relations amongst these logical constructions, for the sake making it clearly understood, outlined pictorially  in Figure \ref{fig:logicaladdrspace}:
-\begin{itemize}
-\item a global address space invariant (solid round square labelled as $\mathcal{I}$\textsf{ASpace} in Figure \ref{fig:logicaladdrspace}) asserting the authorative ownership (i.e.
-update capability), $\ownGhost\gammaPred{\authfull{\ptableabswalk\ptablestore}}$, of the ghost map -- shown as solid-round-head arrow to $\mathcal{A}\textsf{PTableWalk}$ Figure \ref{fig:logicaladdrspace}
-\item per-virtual-pointsto relation ((cutted round square labelled as \textsf{Virtual PointsTo} in Figure \ref{fig:peraspaceinvariant})) asserting the fragmental ownership of the ghost map, $\sumwalkabs\vaddr\qfrac\paddr$ \textit{represents} the fragmental-ownership (cutted-diamond-head arrow) of $\mathcal{A}\textsf{PTableWalk}$
-\item there exists only one persistent authorative ownership ($\ownGhost\gammaPred{\authfull{\ptableabswalk\ptablestore}}$) for all fragmental ghost page-table-walk mapping ($\sumwalkabs\vaddr\qfrac\paddr$) shown as double-ended solid arrow 1 to N imposing the update protocol
-\end{itemize}
+$\mathcal{I}\textsf{ASpace}(\Theta,m)$ is then the resources for the current address space:
+It ensures that the current address space is valid (it has an entry in $m$, whose corresponding
+authoritative map matches $\Theta$), and asserts ownership of the appropriate fractional ownership
+of the physical page table walk for each virtual-to-physical translation in $\Theta$ (for the current address space).
+
+Figure \ref{fig:logicaladdrspace} depicts the relationships between these pieces, for a single
+address space.
+The invariant for the currently active address space (solid round square labelled as 
+$\mathcal{I}$\textsf{ASpace} in Figure \ref{fig:logicaladdrspace}) asserts authorative ownership (i.e. 
+update capability), $\ownGhost\gammaPred{\authfull{\ptableabswalk\ptablestore}}$, of the ghost map $\Theta$
+describing virtual-to-physical address translations
+-- shown as a solid-round-head arrow to $\mathcal{A}\textsf{PTableWalk}$.
+Each virtual points-to assertion (dashed round square labelled as \textsf{Virtual PointsTo}
+ in Figure \ref{fig:peraspaceinvariant})
+asserts fragmental ownership of the ghost map $\Theta$, $\sumwalkabs\vaddr\qfrac\paddr$,
+shown via the dashed-diamond-head arrow to $\mathcal{A}\textsf{PTableWalk}$. Specifically, it asserts that
+the virtual address the virtual points-to assertion concerns corresponds to some physical address.
+There exists only one persistent authorative ownership 
+($\ownGhost\gammaPred{\authfull{\ptableabswalk\ptablestore}}$) which collects the fragmental 
+ghost page-table-walk mapping ($\sumwalkabs\vaddr\qfrac\paddr$),
+ shown as double-ended solid arrow 1 to N imposing the update protocol.
 
@@
    \includegraphics[width=0.75\columnwidth]{logical_addr_space.png}
-  \caption{Logical Resources of Address-Space Abstraction}
+  \caption{Logical Resources of Address-Space Abstraction (1 Address Space)}
   \label{fig:logicaladdrspace}
   \end{figure}
+
+\paragraph{From A Single Address Space to Many}
+In principle the resources from each individual address space should be collected into a single
+shared invariant, and for each memory access,
+a fragment of this corresponding to $\mathcal{I}\textsf{ASpace}(\Theta,m)$
+for the current address space should be extracted from this global resource, used to prove correctness of the
+memory access, and put back.
+In this paper we focus on the middle section, explicity identifying resources for each
+address space. The reason for this is that in practice, this global resource is also
+the place where kernel-specific assumptions, such as guaranteeing that certain virtual address range
+was mapped in all address spaces, would be enforced. This paper focuses on the
+reasoning principles behind the virtual memory access, and we leave the use of this within a larger
+kernel to future work.
 

### 2023-04-15 programlogic.tex
starting to restructure late section 4

@@
 for the current address space.
+The global map itself is tracked by a resource algebra 
+\[\mathcal{A}\textsf{spaces} \stackrel{\triangle}{=} \textsc{Auth} (\; \mathcal{W}_{64} \;\rightarrow_{\textrm{fin}} \;  ( (\textsc{Frac }, \mathord{+}) \times (\textsc{Agree } \Loc,\mathord{=}) )) \]
+\todo[inline]{shouldn't above be map to gname, not these pairs?}
 
@@
 \label{sec:aspacemanagement}
-\todo[inline]{Colin has not looked at this section in detail yet.
-It is clear the discussion of the cr3 update needs updating for doubles}
-So far, we have introduced logical abstractions for a single address space.
-However, \textsf{VMM} implementations handle more than one address-space which we abstract as a ghost map from root address of an address-space to a namespace labelling the address space
-\[\mathcal{A}\textsf{spaces} \stackrel{\triangle}{=} \textsc{Auth} (\; \mathcal{W}_{64} \;\rightarrow_{\textrm{fin}} \;  ( (\textsc{Frac }, \mathord{+}) \times (\textsc{Agree } \Loc,\mathord{=}) )) \]
+So far, we have introduced logical abstractions for a single address space. However, virtual memory management
+implementations typically handle more than one address-space,
+and doing this requires assertions which talk about other address spaces, and means to switch address spaces.
+% requiring a means to  which we abstract as a ghost map from root address of an address-space to a namespace labelling the address space
 
-\paragraph{Valid Context-Resources} Fragmental ownership of $\mathcal{A}\textsf{spaces}$ ($\sumapaces\rtv\gammaPred$) conveys the fact that a current address-space ($\gammaPred$) is the one selected, i.e.
+% \paragraph{Valid Context-Resources} Fragmental ownership of $\mathcal{A}\textsf{spaces}$ ($\sumapaces\rtv\gammaPred$) conveys the fact that a current address-space ($\gammaPred$) is the one selected, i.e.
 loaded into the $\rtv$.
@@
 
-\paragraph{Modal Resource Context\&Interaction with Ambient Logic \textsf{vProp}}
-\label{sec:resourcecontext}
-The relation between the address-space and and the validity of virtual-pointsto assertions exhibits itself as a contingency, i.e.
+% \paragraph{Modal Resource Context\&Interaction with Ambient Logic \textsf{vProp}}
+% \label{sec:resourcecontext}
+% The relation between the address-space and and the validity of virtual-pointsto assertions exhibits itself as a contingency, i.e.
 \textit{"a certain virtual-pointsto facts whose validity is indexed by the root address of an address space"}.
@@
 
-In Figure \ref{fig:addrswitch}, we see two different address-spaces rooted at R0 and R1 in the pre/post condition.
+% In Figure \ref{fig:addrswitch}, we see two different address-spaces rooted at R0 and R1 in the pre/post condition.
 To be put into words, while switching an address-space by executing \lstinline|mov_ctl|) updating the  \lstinline|cr3|, it is expected to mention resources of two different address-spaces in the specification.
@@
 \[
-  \begin{array}{l}
-    \hbox{(\TirNameStyle{ModalAddressSpace})} \qquad
+  \begin{array}{r@{\qquad}l}
+    \hbox{(\TirNameStyle{ModalAddressSpace})} &
          [r](\mathsf{P}) \;: \mathsf{vProp} \; \Sigma \; \stackrel{\triangle}{=} \; \lambda \_,\; \mathsf{P}~\mathsf{r}.\\
-    \hbox{(\TirNameStyle{ModalAddressSpaceMono})} \qquad (\textsf{P} \vdash \textsf{Q}) \vdash  ([r](\textsf{P}) \vdash  [r](\textsf{Q}))\\
-    \hbox{(\TirNameStyle{ModalAddressSpaceSep})} \qquad [r](\textsf{P} \ast \textsf{Q}) \dashv\vdash ([r](\textsf{P}) \ast [r](\textsf{Q})\\
-    \hbox{(\TirNameStyle{ModalAddressSpacePure})} \qquad  \textsf{pure\_vprop P}(\textsf{iPropI }\Sigma)\textsf{ : vProp }\Sigma \stackrel{\triangle}{=} \; \lambda \_ \ldotp \textsf{P}\\
-    \hbox{(\TirNameStyle{ModalAddressSpaceFact})} \qquad  \textsf{Fact P}(\textsf{vProp }\Sigma) \stackrel{\triangle}{=} \;\forall \; ,r \; r' \ldotp  \textsf{P r} \dashv\vdash \textsf{P r'}\\
-    \hbox{(\TirNameStyle{ModalAddressSpaceFactIntroElim})} \\
-    \qquad\qquad \textsf{Fact pure\_vprop P} \vdash  ([r](\textsf{pure\_vprop P}) \dashv\vdash (\textsf{pure\_vprop P}) (\textsf{vProp }\Sigma))\\
-
-    \hbox{(\TirNameStyle{ModalAddressSpaceAnd})} \qquad   [r](\textsf{P}\land \textsf{Q}) \dashv\vdash [r](\textsf{P})\land [r](\textsf{Q})\\    
-    \hbox{(\TirNameStyle{ModalAddressSpaceOr})} \qquad    [r](\textsf{P}\lor \textsf{Q}) \dashv\vdash [r](\textsf{P})\lor [r](\textsf{Q})\\   
-    \hbox{(\TirNameStyle{ModalAddressSpaceFactIntroElim})} \\
-    \qquad\qquad \textsf{Fact pure\_vprop P} \vdash  ([r](\textsf{pure\_vprop P}) \dashv\vdash (\textsf{pure\_vprop P}) (\textsf{vProp }\Sigma))\\
-
+    \hbox{(\TirNameStyle{ModalAddressSpaceMono})} & (\textsf{P} \vdash \textsf{Q}) \vdash  ([r](\textsf{P}) \vdash  [r](\textsf{Q}))\\
+    \hbox{(\TirNameStyle{ModalAddressSpaceSep})} & [r](\textsf{P} \ast \textsf{Q}) \dashv\vdash ([r](\textsf{P}) \ast [r](\textsf{Q})\\
+    \hbox{(\TirNameStyle{ModalAddressSpaceAnd})} &   [r](\textsf{P}\land \textsf{Q}) \dashv\vdash [r](\textsf{P})\land [r](\textsf{Q})\\    
+    \hbox{(\TirNameStyle{ModalAddressSpaceOr})} &    [r](\textsf{P}\lor \textsf{Q}) \dashv\vdash [r](\textsf{P})\lor [r](\textsf{Q})\\   
+    % \hbox{(\TirNameStyle{ModalAddressSpacePure})} &  \textsf{pure\_vprop P}(\textsf{iPropI }\Sigma)\textsf{ : vProp }\Sigma \stackrel{\triangle}{=} \; \lambda \_ \ldotp \textsf{P}\\
+    \hbox{(\TirNameStyle{ModalAddressSpaceFact})} &  \textsf{Fact P} \stackrel{\triangle}{=} \;\forall \; ,r \; r' \ldotp  \textsf{P r} \dashv\vdash \textsf{P r'}\\
+    % \hbox{(\TirNameStyle{ModalAddressSpaceFactIntroElim})} \\
+    % \qquad\qquad \textsf{Fact pure\_vprop P} \vdash  ([r](\textsf{pure\_vprop P}) \dashv\vdash (\textsf{pure\_vprop P}) (\textsf{vProp }\Sigma))\\
+    \hbox{(\TirNameStyle{ModalAddressSpaceFactIntroElim})} 
+    &\textsf{Fact P} \vdash  ([r](\textsf{P}) \dashv\vdash (\textsf{P}) %(\textsf{vProp }\Sigma))\\
   \end{array}
   \]
-  \todo[inline,color=red]{Colin, story about any of the stuff above}
+  \caption{Other-space Modality and Its Laws}
   \label{fig:modaldef}
   \end{figure}
+Figure \ref{fig:modaldef} gives the definition of our modal operator for asserting truth of a modal
+(address-space-contingent) assertion \emph{in another address space}, which we call
+the \emph{other-space} modality. The definition itself is not
+particularly surprising --- as our modal assertions are semantically predicates on a page table root (physical)
+address, the assertion $[r](P)$ is a modal assertion which ignores the (implicit) current page table root,
+and evaluates the truth of $P$ as if $r$ were the page table root. This simple definition is first of all
+enough for assertions in one address space to talk about assertions in another.
+
+More usefully, we can prove that this modality follows certain useful laws, and that it interacts with
+a class of assertions which are true regardless of which address space they are evaluated in.
+The other-space modality distributes over the various separation logic connectives, and follows the rule of
+consequence.
+We also define a class of \textsf{vProp}s we call \textsf{Fact}s, which are those assertions
+which ignore their address space entirely.
+Physical points-tos and register value assertions, when treated as \textsf{vProp}s, are \textsf{Fact}s.
+\textsf{Fact}s can move in and out of other-space modalities freely.
 
@@
 
-\subsection{Considering Hoare Doubles in the Context of Switching Address-Space}
-\subsubsection{Issues with Using Hoare Triples in Address-Space Specification}
+\subsection{The Impact of Changing Address Spaces on Modal Assertions}
+% \subsection{Considering Hoare Doubles in the Context of Switching Address-Space}
+% \subsubsection{Issues with Using Hoare Triples in Address-Space Specification}
 \label{sec:issues}
@@
 \end{mathpar}
-This rule captures the intuitive change of address space in a Hoare triple, rather than double, form.
-The problem with this is that it interacts quite poorly with the traditional frame rule and the modal flavor of virtual points-to assertions:
+This rule captures the intuitive change of address space in a Hoare triple.
+The problem with this is that it interacts quite poorly with the traditional frame
+rule and the modal flavor of virtual points-to assertions.
+If we abbreviate the pre- and post-conditions above as \textsf{Pre} and \textsf{Post}:
 \begin{mathpar}
@@
     \inferrule*[right=Broken]{ }{
-    \{\mathsf{emp} \ast \textsf{cr3} \mapsto_{r} r_1 \ast r\mapsto_{r} r_2 \ast [r_2](Q)\}
+    \{\textsf{Pre}\}
     \texttt{mov}~\texttt{\%cr3},~r%\lstinline|mov %cr3, r| 
-    \{[r_1](\mathsf{emp}) \ast \textsf{cr3} \mapsto_{r} r_1 \ast r \mapsto_{r} r_2 \ast Q\}
+    \{\textsf{Post}\}
     }
   }{
-    \{a\mapsto_\mathsf{v} x \ast \mathsf{emp} \ast \textsf{cr3}\mapsto_{r}r_1 \ast r \mapsto_{r}r_2 \ast [r_2](Q)\}
+    \{a\mapsto_\mathsf{v} x \ast \textsf{Pre}\}
     \texttt{mov}~\texttt{\%cr3},~r%\lstinline|mov %cr3, r| 
-    \{a\mapsto_\mathsf{v} x \ast [r_1](\mathsf{emp}) \ast \textsf{cr3} \mapsto_{r} r_1 \ast r \mapsto_{r} r_2 \ast Q\}
+    \{a\mapsto_\mathsf{v} x \ast \textsf{Post}\}
   }
@@
 So it is simply unsound to frame around \lstinline|cr3| updates.
+In hindsight this is not surprising, because we have essentially defined
+our assertions to depend on resources they do not explicitly mention --- the value of \lstinline|cr3|.
+However, it has important impacts on our formal development.
+
+\todo[inline]{Colin revision frontier}
 
@@
 Virtual points-to assertions could be made self-stable by also giving them partial ownership over \lstinline|cr3| assertions, but this would require explicitly plumbing that ownership from \emph{all} assertions back to any place an address space change might occur; this would seem to be a far graver loss of modularity than this extra quirk in framing discussions.
-\subsubsection{Defining Hoare Doubles for Address-Space Specification}
+\subsection{Selected Logical Rules}
+\label{sec:selected_rules}
 Based on the issues we discuss in Section \ref{sec:issues}, we make sure our Hoare-Double definition in Figure \ref{fig:wpddefinition} enables

### 2023-04-15 programlogic.tex
crank on section 4, nearly done, along the way adding some hybrid logic comparison stuff

@@
 address, the assertion $[r](P)$ is a modal assertion which ignores the (implicit) current page table root,
-and evaluates the truth of $P$ as if $r$ were the page table root. This simple definition is first of all
-enough for assertions in one address space to talk about assertions in another.
+and evaluates the truth of $P$ as if $r$ were the page table root. This simple definition is 
+what we need in order for assertions in one address space to talk about assertions in another.
 
-More usefully, we can prove that this modality follows certain useful laws, and that it interacts with
+The inspiration for this assertion comes from hybrid logic~\cite{blackburn1995hybrid,areces2001hybrid,goranko1996hierarchies,gargov1993modal},
+which as mentioned earlier uses as modalities the names of worlds in a Kripke model to state
+that an assertion is true in that specific world.
+Conceptually the address space rooted at some physical address (intended as the start of an L4 table)
+is a distinct world, and it makes sense to name it by referring to the root of the
+supporting page table structure. However, rather than explicitly building a Kripke model for our
+assertions, we inherit them by lifting the model of our ambient logic, Iris.
+
+This assertion is also useful in proofs: we can prove that this modality follows certain useful laws, and that it interacts with
 a class of assertions which are true regardless of which address space they are evaluated in.
@@
 
-\begin{figure}
-\[
-\begin{array}{l}
-  \{ \textsf{cr3 } \mapsto_{r} R0 \ast a_0 \mapsto_{\textsf{v}}p_0  \ast a_1 \mapsto_{\textsf{v}}p_1  \ast [r1]\;( b_0 \mapsto_{\textsf{v}}p_3  \ast b_1 \mapsto_{\textsf{v}}p_4 )\}\\
-  \qquad \qquad \qquad \qquad \qquad \mathsf{mov\_ctl} \;\; \texttt{cr3} \; \; R1\\
+% \begin{figure}
+% \[
+% \begin{array}{l}
+%   \{ \textsf{cr3 } \mapsto_{r} R0 \ast a_0 \mapsto_{\textsf{v}}p_0  \ast a_1 \mapsto_{\textsf{v}}p_1  \ast [r1]\;( b_0 \mapsto_{\textsf{v}}p_3  \ast b_1 \mapsto_{\textsf{v}}p_4 )\}\\
+%   \qquad \qquad \qquad \qquad \qquad \mathsf{mov\_ctl} \;\; \texttt{cr3} \; \; R1\\
 
-  \{ \textsf{cr3 } \mapsto_{r} R1 \ast [r0](a_0 \mapsto_{\textsf{v}}p_0  \ast a_1 \mapsto_{\textsf{v}}p_1)  \ast  b_0 \mapsto_{\textsf{v}}p_3  \ast b_1 \mapsto_{\textsf{v}}p_4 \}
-  \end{array}
-\]
-  \label{fig:addrswitch}
-  \end{figure}
+%   \{ \textsf{cr3 } \mapsto_{r} R1 \ast [r0](a_0 \mapsto_{\textsf{v}}p_0  \ast a_1 \mapsto_{\textsf{v}}p_1)  \ast  b_0 \mapsto_{\textsf{v}}p_3  \ast b_1 \mapsto_{\textsf{v}}p_4 \}
+%   \end{array}
+% \]
+%   \label{fig:addrswitch}
+%   \end{figure}
 
-\begin{remark}[The Choice of Modal Context, Contingency and  Pattern of Verification Context]
-  \label{remark:pattern}
-The truth on an address space, exhibits itself as a contingent truth: a location virtualization assertion (a virtual points-to in \ref{fig:virtualpointstosharing}),  happens to valid in the \textit{current world} (in the current address space), and switching address spaces pulls information out of one \textit{world} into the “current view” of memory, and leaves other assertions true relative to the previous address space.
+% \begin{remark}[The Choice of Modal Context, Contingency and  Pattern of Verification Context]
+%   \label{remark:pattern}
+% The truth on an address space, exhibits itself as a contingent truth: a location virtualization assertion (a virtual points-to in \ref{fig:virtualpointstosharing}),  happens to valid in the \textit{current world} (in the current address space), and switching address spaces pulls information out of one \textit{world} into the “current view” of memory, and leaves other assertions true relative to the previous address space.
 
-Therefore, an address-space, as an abstraction, can be treated as naming the memory state as a modal frame and the choice of page table root as a world in Kripke-style semantics.
+% Therefore, an address-space, as an abstraction, can be treated as naming the memory state as a modal frame and the choice of page table root as a world in Kripke-style semantics.
 Not suprisingly, transition between two address-spaces, then, can be just an entailment relation \textit{alternating} the \textit{named-state}
 
-Being inspired by what hybrid logic \ref{} calls a satisfaction operator, which evaluates the truth of a predicate in a named alternative state (here, address space), we give a modal definition describing the truth of assertions for the resources (i.e.
+% Being inspired by what hybrid logic \ref{} calls a satisfaction operator, which evaluates the truth of a predicate in a named alternative state (here, address space), we give a modal definition describing the truth of assertions for the resources (i.e.
 virtual pointsto relations) inside the address space.
@@
 
-Pragmatically, 
-  \begin{enumerate}
-  \item a modal context with its context-resources and picked contingency defines a modality (e.g.
+% Pragmatically, 
+%   \begin{enumerate}
+%   \item a modal context with its context-resources and picked contingency defines a modality (e.g.
 a address-space modality $[r]P$).
@@
 virtual pointsto relations in the current address-space)
-  \item and, offloads the burden of individual bookkeeping of these facts (e.g.
+%   \item and, offloads the burden of individual bookkeeping of these facts (e.g.
 virtual pointsto relations per address-space) under different context (e.g.
@@
 other virtual pointsto relations are valid under other certain address-space page-table root addresseses).
-  \end{enumerate}
-\end{remark}
+%   \end{enumerate}
+% \end{remark}
 
@@
 
-\todo[inline]{Colin revision frontier}
-
-Switching to Hoare doubles resolves this problem because an under-appreciated subtlety of Hoare doubles is that typically \emph{there is no frame rule}.
-Instead each verification essentially includes a local frame that it passes to the next instructions (think continuation-passing style), giving each overall rule a \emph{global} (rather than local) precondition.
+Switching to Hoare doubles resolves this problem because an under-appreciated subtlety of Hoare doubles 
+is that typically \emph{there is no primitive frame rule}. 
+Instead each verification essentially includes a local frame that it passes to the next instructions
+(think continuation-passing style), giving each overall rule a \emph{global} (rather than local) precondition.
 For most rules this is not that important, but it does permit rules that have global effects on their preconditions.
@@
 In particular, this rule would work with framing around calls that might lead to address space switches, such as calling blocking operations in the kernel.
+Attempting to derive a frame rule as some work using Hoare doubles does~\cite{Chlipala2013Bedrock}
+would enforce this restriction.
 
@@
 Virtual points-to assertions could be made self-stable by also giving them partial ownership over \lstinline|cr3| assertions, but this would require explicitly plumbing that ownership from \emph{all} assertions back to any place an address space change might occur; this would seem to be a far graver loss of modularity than this extra quirk in framing discussions.
+
 \subsection{Selected Logical Rules}
 \label{sec:selected_rules}
-Based on the issues we discuss in Section \ref{sec:issues}, we make sure our Hoare-Double definition in Figure \ref{fig:wpddefinition} enables
-\begin{itemize}
-\item virtual-pointsto relations to be independent of \lstinline{cr3} ownership
-\item and full-ownership of \lstinline to be propagated inside the \textsf{wpd\_def} in order to
-\end{itemize}
-This act of separation allows client-specification to be agnostic of \lstinline|cr3| physical resource, (i.e.
-\lstinline|cr3| $\mapsto_{\textsf{r}} \rtv$).
-The context-resource (virtual-pointsto) has only the \textit{knowledge} of being in a namespace in which it is valid ($\sumapaces\rtv\gammaPred$), nothing more.
-
-In the rest of this, we explain how specifications of some selected \textsf{AMD64} instructions look like under the decisions made on the reasoning principles.
-All rules, e.g.
-the ones with different addressing mode, can be found in our artifact.
-
-In the general outline of the proof in Figure \ref{wpdamd}, we see that each rule annotated with an address value (\textsf{rtv}), i.e.
-a root address of an address-space, under which the resources mentioned in the specification are valid.
-Moreover, there is a frame resource $P$ in each rule.
+To address the issues described in Section \ref{sec:issues}, we give our logic using
+the Hoare-Double definition in Figure \ref{fig:wpddefinition}.\footnote{The figure does
+omit some low-level Iris details related to stuckness and observations, which play no meaningful
+role in our development.}
 \begin{figure} 
@@
   \begin{array}{l}
-    \textsf{wpd\_def e s E1 } \Phi \;\mathsf{ rtv } : \textsf{iProp }\Sigma := \\
-   \qquad ((\textsf{cr cr3}) \mapsto_{\textsf{r}} (\textsf{num rtv)}) \textsf{ rtv}) \ast (\Phi \textsf{ rtv}) \wand \textsf{WP e E } @ \textsf{s ; E1 } \{\{\_, \textsf{True} \}\}
+    %\textsf{wpd\_def e s E1 } \Phi \;\mathsf{ rtv } : \textsf{iProp }\Sigma := \\
+    \{ \Phi \}_\mathsf{ rtv }\;\textsf{e} : \textsf{iProp }\Sigma := 
+   % \qquad
+   ((\textsf{cr3} \mapsto_{\textsf{r}} \textsf{rtv} \ast \Phi) \textsf{ rtv}) \wand \textsf{WP e } \{\_, \textsf{True} \}
     \end{array}
@@
 \end{figure}
+Our Hoare doubles $\{\Phi\}_\textsf{rtv}\;\textsf{e}$ state that the expression (i.e., sequence of instructions)
+\textsf{e} are safe to execute (will not fault)
+when executed with precondition $\Phi\ast\textsf{cr3}\mapsto_{\textsf{r}} \textsf{rtv}$.
+\textsf{WP} is Iris's own weakest precondition modality, unmodified~\cite{jung2018iris}.
+Making \textsf{rtv} an explicit parameter to the double, rather than simply using register assertions
+directly solves a technical problem with ensuring that the page table root used to evaluate
+the \textsf{vProp} (i.e., evaluating the assertion in the \emph{current}) address space
+is feasible.\footnote{Consider the difficulty of selecting the correct page table root value from an arbitrary
+opaque $\Phi$, which may even existentially quantify the page table root. An alternative is to
+require $\Phi$ to have a syntactic form where we can directly extract the value of \lstinline|cr3|,
+but this makes using Iris Proof Mode (IPM)~\cite{Krebbers:2017:IPH:3009837.3009855} with \textsf{vProps}
+  difficult; IPM works for any type matching the signature of an Iris \lstinline|bi|, which includes
+  \textsf{vProp}s, but manually guiding IPM to put an assertion in a specific position over and over adds
+  significant proof burden.
+}
+
+The rest of this section describes specifications of some selected \textsf{AMD64} instructions 
+in our logic. 
+These rules and others (e.g., including accessing memory at an instruction-specified offset from a register
+value, which is common in most ISAs)
+can be found in our artifact.
+Each rule in Figure \ref{fig:wpdamd}, 
+is annotated with an address value (\textsf{rtv}), 
+i.e. a root address of an address-space, under which the resources mentioned in the specification are valid,
+or equivalently, the assumed value of \lstinline|cr3| prior to executing the instructions.
+In general, we use $\textsf{r}_s$ and $\textsf{r}_d$ to specify source and destination registers
+for each instruction, and prefix various register value variables with \textsf{rv}.
+We sometimes use $\textsf{r}_a$ to emphasize when a register is expected to hold an address used
+for memory access, though the figure also uses typical assembler conventions of specifying
+memory access operands by bracketing the register holding the memory address.
+Standard for Hoare doubles, there is a frame resource $P$ in each rule for passing resources
+not used by the first instruction in sequence through and on to subsequent instructions.
 
@@
 \\
-\inferrule[WriteToPhysMemFromReg]{
-  \{P \ast r_s \mapsto_{r}\{q\}  \textsf{rvs}  \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{t}} \textsf{v} \}_{\textsf{rtv}}\;\overline{ is}  
-}{
-  \{P \ast r_s \mapsto_{r}\{q\}  \textsf{rvs}   \ast r_a \mapsto_{r}\{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{t}} \textsf{rvs} \}_{\textsf{rtv}}
-  \textsf{ mov}~\textsf{r}_a,~\textsf{r}_s;\;\overline{is}
-}
-\\
-\inferrule[WriteToRegFromPhysMem]{
-  \{P \ast r_d \mapsto_{r}  \textsf{v} \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{t}} \textsf{v} \}_{\textsf{rtv}}\;\overline{is}
-}{
-  \{P \ast r_d \mapsto_{r}  \textsf{rvd} \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{t}} \textsf{v} \}_{\textsf{rtv}}
-  \textsf{mov}~\textsf{r}_d,~\textsf{r}_a;\;\overline{is}
-}
-\\
+% \inferrule[WriteToPhysMemFromReg]{
+%   \{P \ast r_s \mapsto_{r}\{q\}  \textsf{rvs}  \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{t}} \textsf{v} \}_{\textsf{rtv}}\;\overline{ is}  
+% }{
+%   \{P \ast r_s \mapsto_{r}\{q\}  \textsf{rvs}   \ast r_a \mapsto_{r}\{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{t}} \textsf{rvs} \}_{\textsf{rtv}}
+%   \textsf{ mov}~[\textsf{r}_a],~\textsf{r}_s;\;\overline{is}
+% }
+% \\
+% \inferrule[WriteToRegFromPhysMem]{
+%   \{P \ast r_d \mapsto_{r}  \textsf{v} \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{t}} \textsf{v} \}_{\textsf{rtv}}\;\overline{is}
+% }{
+%   \{P \ast r_d \mapsto_{r}  \textsf{rvd} \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{t}} \textsf{v} \}_{\textsf{rtv}}
+%   \textsf{mov}~\textsf{r}_d,~[\textsf{r}_a];\;\overline{is}
+% }
+% \\
 \inferrule[WriteToRegFromVirtMem]{
@@
   \{P \ast \mathcal{I}\textsf{ASpace}\ast r_d \mapsto_{r}  \textsf{rvd} \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{v}} \textsf{v} \}_{\textsf{rtv}}
-  \textsf{mov}~\textsf{r}_d,~\textsf{r}_a;\;\overline{is}
+  \textsf{mov}~\textsf{r}_d,~[\textsf{r}_a];\;\overline{is}
 }
@@
   \{P \ast \mathcal{I}\textsf{ASpace}\ast r_s \mapsto_{r}\{q\}  \textsf{rvs}   \ast r_a \mapsto_{r}\{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{v}} \textsf{rvs} \}_{\textsf{rtv}}
-  \textsf{ mov}~\textsf{r}_a,~\textsf{r}_s;\;\overline{is}
+  \textsf{ mov}~[\textsf{r}_a],~\textsf{r}_s;\;\overline{is}
 }
@@
   \{P \ast r_s \mapsto_{r}\{q\}  \textsf{ rvs}   \}_{\textsf{rtv}}
+  \textsf{mov}~\textsf{cr3},~r_s;\;\overline{is}
+  %\lstinline|mov %cr3, r| 
+}\\
+\inferrule[WriteToRegCtlFromReg]{
+  \{[\textsf{rtv}](P \ast \mathcal{I}\textsf{ASpace})\ast \mathcal{I}\textsf{ASpace} \ast R\ast r_s \mapsto_{r}\{q\}  \textsf{ rvs}  \}_{\textsf{rvs}} \overline{is}
+}{
+  \{P \ast \mathcal{I}\textsf{ASpace} \ast [\textsf{rvs}](R \ast\mathcal{I}\textsf{ASpace})\ast r_s \mapsto_{r}\{q\}  \textsf{ rvs}   \}_{\textsf{rtv}}
   \textsf{mov}~\textsf{cr3},~r_s;\;\overline{is}
@@
 \end{figure}
-\subsubsection{Memory To Register}
-In Figure \ref{fig:wpdamd}, we see the rules for updating the memory address ($r_a$).
-To do so, we essentially need page-table-walk pointsto ($\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$ in Figure \ref{fig:strongvirtualpointsto}) to locate the value to be written.
-We mainly have two options to achieve this.
-\paragraph{Physical Memory To Register} First one is to use  page-table-walk pointsto ($\textsf{vaddr} \mapsto_{\textsf{t}} \textsf{v}$), which is consists of all-physical page-table memory assertions, to prove the page-table traversal \textsf{AMD64} implements, update the register $r_d$ the value obtained ($\textsf{v}$) as shown in Figure (\TirNameStyle{WriteToRegFromPhysMem}).
-By doing so, we expose the resources ($\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$) needed only for the proof of \textsf{AMD64} page-table traversal to the client of the rule.
-This breaks the modularity of reasoning we want to maintain in location virtualization.
-  \paragraph{Virtual Memory To Register}  Therefore, we would like to use a virtual-pointsto assertion ($\textsf{vaddr} \mapsto_{\textsf{v},\textsf{rtv}}$), which hides physical resources needed for the proof of page-table-traversal from the client, to obtain the the value ($\textsf{v}$) to update the register $r_d$ as shown in Rule (\TirNameStyle{WriteToRegFromVirtMem}).
-While doing the proof of page-table-traversal with a given virtual-pointsto definition in Figure \ref{fig:virtualpointstosharing}, we unfold and obtain our token ($\sumwalkabs\vaddr\qfrac\paddr$) to exchange for a physical-table mapsto assertion inside the $\mathcal{I}\textsf{ASpace}$ ($\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$) in Figure \ref{fig:peraspaceinvariant}, and apply Rule (\TirNameStyle{WriteToRegFromPhysMem}).
-\paragraph{Updating \lstinline|cr3|} Unlike other rules, \TirNameStyle{WriteToRegCtlFromReg} updates the root address of the address-space determining the validity of resources, $\{\rtv\} / \{\textsf{rvs}\}$.
+
+We briefly describe several key rules representative of the system.
+
+Figure \ref{fig:wpdamd}, includes two  rules for accessing memory at an address in a register $r_a$. 
+Ultimately, any memory access needs to ensure the relevant address translation would succeed,
+which can be ensured by what we informally call the page-table-walk points-to collection
+($\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$ in Figure \ref{fig:strongvirtualpointsto}).
+
+% We mainly have two options to achieve this. 
+% First one is to use  page-table-walk pointsto 
+% ($\textsf{vaddr} \mapsto_{\textsf{t}} \textsf{v}$), 
+% which is consists of all-physical page-table memory assertions, 
+% to prove the page-table traversal \textsf{AMD64} implements, update the register $r_d$ the value obtained 
+% ($\textsf{v}$) as shown in Figure (\TirNameStyle{WriteToRegFromPhysMem}). 
+% By doing so, we expose the resources ($\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$) needed only 
+% for the proof of \textsf{AMD64} page-table traversal to the client of the rule. 
+% This breaks the modularity of reasoning we want to maintain in location virtualization.
+
+\TirNameStyle{WriteToRegFromVirtMem} and \TirNameStyle{WriteToVirtMemFromReg}
+each use a virtual-pointsto assertion ($\textsf{vaddr} \mapsto_{\textsf{v},\textsf{rtv}}$),
+and are nearly-standard (assembly) separation logic rules for memory accesses.
+However, because we split the physical resources for the page table walk from the
+virtual points-to itself (in order to permit the physical page tables to be freely modified
+as long as they preserve virtual-to-physical translations), the rule requires $\mathcal{I}\textsf{ASpace}$
+for the current address space to be carried through.
+While doing the proof of soundness for this rule,
+the token ($\sumwalkabs\vaddr\qfrac\paddr$) from the virtual points-to
+is used to extract to extract page-table-traversal points-to collection
+for the relevant address from that resource,
+and then those  physical points-to assertions describing the page table walk
+(from $\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$) are used to prove
+the operational semantics succeed and have the desired effect.
+
+\paragraph{Updating \lstinline|cr3|} 
+Unlike other rules, \TirNameStyle{WriteToRegCtlFromReg} updates the root address of the 
+address-space determining the validity of resources, from $\rtv$ before the
+\lstinline|mov| to $\textsf{rvs}$ afterwards.
+\todo[inline]{Colin says: the cr3 update rule should have IASpace (x2) and modalities...?}

### 2023-04-15 relwork.tex
crank on section 4, nearly done, along the way adding some hybrid logic comparison stuff

@@
 memory.
+
+As noted throughout the paper, the inspiration for our other-space modality comes from hybrid logic~\cite{areces2001hybrid,blackburn1995hybrid,gargov1993modal,goranko1996hierarchies},
+where modalities are indexed by \emph{nominals} which are names for specific individual states in a Kripke model.
+We are aware of only two prior works combining hybrid logics with program logics specifically. 
+Brotherston and Villard~\cite{brotherston2014parametric} demonstrated that may properties true of various 
+separation logics are not definable in boolean \BI (\BBI), and showed that a hybrid extension \HyBBI allows
+most such properties to be defined (e.g., the fact that separating conjunction is cancellative is unprovable 
+in boolean \BI, but provable in \HyBBI). There, nominals named resources 
+(roughly, but not exactly, heap fragments). 
+Gordon~\cite{gordon2019modal} described a use of hybrid logic in the verification of actor programs, 
+where nominals named the local state of individual actors (with such assertions stabilized with a 
+rely/guarantee approach). Beyond these, there is limited work on the interaction of specifically 
+\emph{hybrid logic} with substructural logics. 
+Primarily there is a line of work on hybrid linear logic (\HyLL)~\cite{despeyroux2014hybrid}, 
+originally used as a way to more conveniently express aspects of transition systems in linear logic. 
+However, \HyLL's proof rules offer no non-trivial interactions with multiplicative connectives 
+(every \HyLL proof can in fact be embedded into regular linear logic~\cite{chaudhuri2019hybrid}, 
+unlike Brotherston and Villard's \HyBBI, which demonstrably increases expressive power over its base \BBI.
+
+In both \HyLL and \HyBBI, nominals denote worlds with monoidal structure (as worlds in Kripke semantics
+for either LL or \BBI necessarily have monoidal structure). Our nominals, by contrast, 
+do not name worlds in the same sense with respect to Iris's CMRAs, 
+but in fact \emph{classes} of worlds, because the names are locations 
+(a means of \emph{selecting} resources) rather than resources.  
+A key difference is that the use of nominals in those logics corresponds specifically to hypothetical 
+reasoning about resources (until a nominal is connected to a current resource, in which case conclusions 
+can be drawn about the current resource), which means the modalities themselves do not ``own'' resources. 
+Instead, assertions under our other-space modality can and do
+have resource footprints.
+Pleasantly, we sidestep most of the metatheoretical complexity of those other substructural hybrid
+systems by building our logic within a substructural metatheory (Iris).

### 2023-04-15 cpp_paper.tex
+
compiles
revisions up through end of section 3
remove discussion section
remove full soundness section, add bus-draft conclusion

@@
 %%\citestyle{acmauthoryear}
+%\usepackage{amsmath}
+%\usepackage{amssymb}
+%\usepackage{amsmath}
+%\usepackage{amssymb}
+
+\usepackage{hyperref}
+\usepackage{array}
+%\usepackage{amsthm}
+\usepackage{proof}
+\usepackage{stmaryrd}
+\usepackage{xspace}
+%\usepackage{listings}
+\usepackage{graphicx}
+
 \usepackage{lipsum}
@@
 \usepackage{amsmath}
-% swiped from iris.sty in the iris repo
-\newcommand{\wand}{\mathrel{-\!\!\ast}}
-\newcommand{\wandIff}{\mathrel{\ast\!\!{-}\!\!\ast}}
+
 % If we need multiple references to a footnote, via \cref:
@@
 \input{macros}
-
 \newcommand{\mytodo}[1]{% <==========================================
@@
 
-\todo[inline]{Colin think through instruction pointer updates vs logic,
-think about framing plus cr3 update and modality shifts
-}
-
 \input{intro.tex}
@@
 \input{programlogic.tex}
-\input{soundness.tex}
+%\input{soundness.tex}
 \input{examples.tex}
-\input{discussion.tex}
+%\input{discussion.tex} % not necessary
 \input{relwork.tex}
+\section{Conclusions}
+This paper advances the state of the art in formal verification of programs subject to address translation
+on hardware using virtual memory. We proposed to treat assertions about virtual
+memory locations explicitly as assertions in a modal logic, where the notion of context
+is the choice of current address space based on the page table root installed on the CPU.
+We improved the modularity of our virtual address translation to allow page table modifications that
+preserve mappings without collecting all affected virtual points-to assertions.
+To make specifications of code involving other address spaces cleaner, we borrowed the idea of
+modalities which explicitly name the conditions under which they are true from hybrid logics.
+We implemented these ideas in a derived separation logic within Iris, and proved soundness of
+the rules for essential memory- and address-space-change-related x86-64 instructions 
+sound against a hardware model of 64-bit 4-level address translation.
+Finally, we used our rules to verify the correctness of key VMM instruction sequences,
+including giving the first assembly-level proof of correctness in a program logic for a change
+of address space, against a specification which clearly identifies which assertions hold in which address space.
+
+% In the future we plan to 
+% extend our work in multiple directions, such as generalizing our assertions and proof rules
+% to support large-page functionality common to virtual memory subsystems (superpages, hugepages),
+% implementing
+
 %% The next two lines define the bibliography style to be used, and

### 2023-04-15 programlogic.tex
tiny soundness section

@@
 \todo[inline]{Colin says: the cr3 update rule should have IASpace (x2) and modalities...?}
+
+\subsection{Soundness}
+Our rules from Figure \ref{fig:wpdamd} are proven sound in Iris against an assembly-level hardware model
+implementing a fragment of x86-64 including 64-bit address translation with 4-level page tables.
+
+At the moment our proofs do rely on 11 small axioms of properties which should be provable, but
+are challenging to discharge due to some representation choices in our model.\footnote{See \texttt{srx/x64/machine/current_axioms.v}}

### 2023-04-15 examples.tex
stuff

@@
 \subsection{Mapping a New Page}
-One of the key tasks of a page fault handler in a general-purpose OS kernel is to map new pages into an address space by writing into an existing page table.
-To do so, we first allocate a fresh page, then calculate the appropriate known-valid page table walks and update the appropriate L1 page table entry.
+One of the key tasks of a page fault handler in a general-purpose OS kernel is
+to map new pages into an address space by writing into an existing page table.
+To do so, we first allocate a fresh page, then calculate the appropriate
+known-valid page table walks and update the appropriate L1 page table entry.
 \lstset{
@@
 $\specline{\texttt{pte\_addr} \mapsto_{\texttt{vpte}} \paddr\; (\texttt{wzero 64})  \ast \ulcorner \texttt{entry\_present } \entryf \land \texttt{entry\_present } \entrytr \land  \texttt{entry\_present } \entrytw \urcorner}_{\rtv}$
-$\specline{\exists \texttt{ fpaddr} \ldotp \ulcorner \texttt{aligned fpaddr} \urcorner \ast \texttt{rax}\mapsto_{\textsf{r}} \texttt{fpaddr} \ast \texttt{fpaddr} \mapsto_{\textsf{p}} (\texttt{wzero 64})  \ulcorner \texttt{entry\_present fpaddr}\urcorner}_{\rtv}$
+$\specline{\exists \texttt{ fpaddr} \ldotp \ulcorner \texttt{aligned fpaddr} \urcorner \ast \texttt{rax}\mapsto_{\textsf{r}} \texttt{fpaddr} \ast \texttt{fpaddr} \mapsto_{\textsf{p}} (\texttt{wzero 64}) \ast \ulcorner \texttt{entry\_present fpaddr}\urcorner}_{\rtv}$
 ;; Calculate new L1 entry
@@
 $\specline{\texttt{pte\_addr} \mapsto_{\texttt{vpte}} \paddr \;(\texttt{fpaddr}) \; \ast \ulcorner \texttt{entry\_present } \entryf \land \texttt{entry\_present } \entrytr \land  \texttt{entry\_present } \entrytw \urcorner }_{\rtv}$
-$\specline{\exists \texttt{ fpaddr} \ldotp \ulcorner \texttt{aligned fpaddr} \urcorner \ast \texttt{rax}\mapsto_{r} \texttt{fpaddr} \ast \texttt{fpaddr} \mapsto_{p} (\texttt{wzero 64})  \ulcorner \texttt{entry\_present fpaddr}\urcorner}_{\rtv}$
+$\specline{\exists \texttt{ fpaddr} \ldotp \ulcorner \texttt{aligned fpaddr} \urcorner \ast \texttt{rax}\mapsto_{r} \texttt{fpaddr} \ast \texttt{fpaddr} \mapsto_{p} (\texttt{wzero 64}) \ast \ulcorner \texttt{entry\_present fpaddr}\urcorner}_{\rtv}$
 $\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\; \sqsubseteq $
@@
 $\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\; \sqsubseteq $
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace} (<[\vaddr:=\texttt{fpaddr}]> \theta,m) \ast \vaddr \mapsto_{\textsf{vpte}}\; \qfrac \;\fpaddr \textsf{ wzero 64}}_{\rtv}$
+  $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace} (<[\vaddr:=\texttt{fpaddr}]> \theta,m) \ast \vaddr \mapsto_{\textsf{vpte}}\; \{\qfrac\} \;\fpaddr \textsf{ wzero 64}}_{\rtv}$
+$\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\; \sqsubseteq $
+  $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace} (<[\vaddr:=\texttt{fpaddr}]> \theta,m) \ast \vaddr \mapsto_{\textsf{v}}\; \{\qfrac\} \textsf{wzero 64}}_{\rtv}$
 \end{lstlisting}
   \caption{Specification and proof of distilled code for mapping a new page with $\textsf{wpd\_def }_{\rtv}$ with a fixed namespace $\gammaPreds$ and a global map of address-spaces $m$  }
-  \todo[inline,color=yellow]{Colin, could you speak about the fact that we have wpd in iProp so that's the reason to have rtval whereever table traversal is needed}
 \label{fig:mapping_code}
-\todo[inline]{this is terribly ugly, should switch to minted}
 \end{figure}
-In Figure \ref{fig:mapping_code}, we see an address ($\vaddr$) currently not mapped to a page ($\theta \; !!\; \vaddr = \texttt{None}$).
-Mapping a fresh page backing this virtual first requires an assurance on the existence L1 table entry which is realized in two folds:
+In Figure \ref{fig:mapping_code}, we see an address ($\vaddr$) currently not
+mapped to a page ($\theta \; !!\; \vaddr = \texttt{None}$). Mapping a fresh
+phyiscal page to back the desired virtual page first requires ensuring
+the existence of a memory location of an appropriate L1 table entry,
+which is realized in two pieces by the post-condition of \lstinline|ensure_L1|:
 \begin{itemize}
-\item physical pointsto assertions on the tables L4, L3, and L2 reaching to the L1 level entry (l1e) ensures the successful table lookups with the fact on \textit{presence} (\textsf{entry\_present}) of the relevant entries (\textsf{l4e l3e l2e}) (Specification Lines 4-6)
-\item ,and a virtual L1 pointsto ($\mapsto_{\textsf{vpte}}$) on the address computed from the reached L1 entry ($ \ulcorner \texttt{addr\_L1}(\vaddr,\entryo) = \paddr\urcorner$) which is almost exactly same with the virtual-pointsto definition in Figure \ref{fig:virtualpointstosharing} except the physical page address is already computed, not exsitentially quantified:
+\item physical pointsto assertions on the tables L4, L3, and L2 reaching to the
+	L1 level entry (l1e) ensure higher levels of the page table exist, 
+	are marked present (\textsf{entry\_present} for the relevant
+		entries \textsf{l4e l3e l2e}) (Specification Lines 4-6) 
+	\item a virtual \emph{PTE}
+		pointsto ($\mapsto_{\textsf{vpte}}$) allows access to the memory of the L1 entry
+		via virtual address \lstinline|pte_pointsto|.
+		A PTE points-to is defined just like the normal virtual points-to of Figure \ref{fig:virtualpointstosharing}, except the physical address is explicit in the assertion (here, \textsf{pa})
+		rather than existentially quantified:
  \[
@@
     \vaddr\mapsto_{\textsf{vpte}} \;\{\textsf{q}\} \; \paddr \; \vpage : \mathsf{vProp}~\Sigma \stackrel{\triangle}{=} 
-    \lambda \mathit{cr3val}\ldotp
     \exists \delta\ldotp
-    \sumapaces{\mathit{cr3val}}\delta \ast 
+	(\lambda \mathit{cr3val}\ldotp
+	\sumapaces{\mathit{cr3val}}\delta) \ast 
   \sumwalkabs\vaddr\qfrac\paddr \ast \paddr \mapsto_{\mathsf{p}} \vpage
@@
 \]
+		This supports rules for accessing memory
+		at that virtual address, but exposing the physical location being modified
+		makes this convenient to use for page table modifications, since we must ensure
+		the modified address is the correct physical address that will be used as the L1 entry.
+		This is also guaranteed ($ \ulcorner
+		\texttt{addr\_L1}(\vaddr,\entryo) = \paddr\urcorner$).
 \end{itemize}
-After obtaining an virtual address \textsf{pte\_addr} in \textsf{rax} referring to a physical address, computed from already known L4-L2 physical memory pointstos ($\texttt{addr\_L1}(\vaddr,\entryo) = \paddr$),
-in Specification Line 7 ($\texttt{pte\_addr} \mapsto_{\textsf{vpte}}  \; \paddr\;(\texttt{wzero 64})$), we save it to \textsf{r14} to be updated later in Line 9.
+After obtaining a virtual address \textsf{pte\_addr} in \textsf{rax} backed 
+by the physical memory for the L1 entry that will be used to translate the virtual addresses
+we are mapping, we save it to \textsf{r14} to be updated later in Line 9.
 
-Then we allocate a fresh page initialized (i.e. zeroed) in Line 14 which return an \textsf{aligned} address of fresh page (\textsf{fpaddr}) in \textsf{rax} (Specification Line 19).
+Then we allocate a fresh page initialized (i.e. zeroed) in Line 14 which return
+an \textsf{aligned} address of fresh page (\textsf{fpaddr}) in \textsf{rax}
+(Specification Line 19), plus 1 (so the return value is already a valid page table entry).
 
-The crux point in the proof is updating L1 table through a virtual address (\textsf{pte\_addr}) through making a known physical address, in our example L1 table entry address ($\textsf{addr\_L1}(\textsf{va, l1e})$), to show the fresly allocated page address (\textsf{fpaddr}) in Line 22 to obtain
-\[\texttt{pte\_addr} \mapsto_{\texttt{vpte}}  \; \paddr\; \textsf{fpaddr}  \qquad\qquad (\textsf{by Rule} \TirNameStyle{WriteToVirtMemFromReg})\]
-by Rule \TirNameStyle{WriteToVirtMemFromReg} (Specification Line 26). 
-Consequently, right at this point in the proof, when we unfold the $\texttt{pte\_addr} \mapsto_{\texttt{vpte}}  \; \paddr\; \textsf{fpaddr})$ we will be able obtain a full ownerhsip on the L1 table physical pointsto $\paddr \mapsto_{\mathsf{p}} \textsf{fpaddr}$ .
-Then, we can first discard the allocated virtual-address pointsto assertion (i.e.
-dropping \textsf{pte\_addr}), and split the full-ownership of physical-page pointsto ($\textsf{fpaddr}\mapsto_{\textsf{p}} \;\textsf{wzero 64}$) to obtain a fractional ownership (q4) on it.
-Since we know $\texttt{addr\_L1 }(\vaddr, \entryo) \mapsto_{\mathsf{p}} \{q4\} \;\textsf{fpaddr}$) from $\mapsto_{\textsf{vpte}}$, after the drop of virtual-pointsto, we still can connect L4-L2 pointsto assertions (in Specification Lines 24 and 25) with newly allocated $\texttt{fpaddr} \mapsto_{a} (\texttt{wzero 64})$ to obtain the complete physical table-walk assertion (L4\_L1\_PointsTo(maddr l4e l3e l2e fpaddr)) in Line 30.
+The crucial point in the proof is then actually updating the L1 entry,
+via the virtual address
+(\textsf{pte\_addr}) known to translate to the appropriate physical address, in our example the L1
+table entry address ($\textsf{addr\_L1}(\textsf{va, l1e})$), to hold the freshly
+allocated physical page address (\textsf{fpaddr}) in Line 22.
+After this write, we have (Specification Line 26)
+\[\texttt{pte\_addr} \mapsto_{\texttt{vpte}}  \; \paddr\; \textsf{fpaddr}  \qquad\qquad (\textsf{by the PTE variant of Rule}~\TirNameStyle{WriteToVirtMemFromReg})\]
 
-Finally, we can insert $\vaddr$ into the ghost page-table-walk summarization map ($\theta$) to change its state from unmapped ($\ulcorner \theta \;!!\;\vaddr = \texttt{None}\urcorner$ in Specification Line 30) to mapped (Specification Line 33) by Iris' ghost-map update ($\sqsubseteq$), and construct virtual-pointsto $\vaddr \mapsto_{\textsf{vpte}}\; \qfrac \;\fpaddr \textsf{ wzero 64}$ (Specification Line 36) with $\sumwalkabs\vaddr\qfrac\fpaddr$ obtained from ghost page-table-walk insertions, $\sumapaces\rtv\delta$ from unfolding the definition of $\mapsto_{\textsf{vpte}}$ in Specification Line 18, and $\fpaddr \mapsto_{\textsf{p}} \textsf{wzero 64}$ in Specification Line 34.
-fragmental ownership of
+At this point in the proof,  $\texttt{pte\_addr} \mapsto_{\texttt{vpte}}  \; \paddr\; \textsf{fpaddr})$ 
+contains  full ownerhsip of the L1 table entry physical pointsto $\paddr
+\mapsto_{\mathsf{p}} \textsf{fpaddr}$ . Then, we can split the full-ownership of that
+physical pointsto
+($\textsf{fpaddr}\mapsto_{\textsf{p}} \;\textsf{wzero 64}$) to obtain
+fractional ownership on it appropriate for a single address's share of an L1 entry (q4).
+Since we know $\texttt{addr\_L1 }(\vaddr,
+\entryo) \mapsto_{\mathsf{p}} \{q4\} \;\textsf{fpaddr}$) from
+$\mapsto_{\textsf{vpte}}$, after the split of the PTE points-to, we can
+connect L4-L2 points-to assertions (in Specification Lines 24 and 25) with 
+ $\texttt{fpaddr} \mapsto_{a} (\texttt{wzero 64})$ to obtain the
+complete physical table-walk assertion (L4\_L1\_PointsTo(maddr l4e l3e l2e
+fpaddr)) in Line 30.
+
+Finally, we can insert $\vaddr$ into the ghost page-table-walk summarization
+map ($\theta$) to change its state from unmapped ($\ulcorner \theta
+\;!!\;\vaddr = \texttt{None}\urcorner$ in Specification Line 30 and the
+precondition) to mapped (Specification Line 33) using Iris' ghost-map update
+($\sqsubseteq$), and construct a PTE -points-to $\vaddr
+\mapsto_{\textsf{vpte}}\; \qfrac \;\fpaddr (\textsf{ wzero 64})$ (Specification
+Line 36) using $\sumwalkabs\vaddr\qfrac\fpaddr$ obtained from ghost
+page-table-walk insertions, $\sumapaces\rtv\delta$ from unfolding the
+definition of $\mapsto_{\textsf{vpte}}$ in Specification Line 18, and $\fpaddr
+\mapsto_{\textsf{p}} \textsf{wzero 64}$ in Specification Line 34.
+Finally, by existentially quantifying away \textsf{fpaddr},
+we can shift the PTE points-to into a normal virtual points-to.
+
+This proof only maps the first word of the allocated page; the generalization to the full page
+is straightforward use of iterated separating conjunction to manage collections of clusters of 512 things
+(slices of the L1 entry physical ownership, virtual addresses at 8 byte increments, etc.).
+
 \subsection{Unmapping a Page}
-The reverse operation, unmapping a designated page that is currently mapped, is pleasingly the reverse: by flipping the entry back to an invalid entry, full ownership of the updated entry reverts back to the now-invalid cell of the L1 table.
+The reverse operation, unmapping a designated page that is currently mapped,
+would essentially be the reverse of
+the reasoning around line 22 above: given the virtual points-to assertions for all 512
+machine words of memory that the L1 entry would map,
+and information about the physical location, 
+full permission on the L1 entry could be obtained, allowing the construction of a
+full virtual PTE pointer for it, setting to 0, and reclaiming the now-unmapped physical memory.
 

### 2023-04-15 programlogic.tex
stuff

--- 2023-04-15 programlogic.tex
@@
 \todo[inline]{Colin says: the cr3 update rule should have IASpace (x2) and modalities...?}
+
+\subsection{Soundness}
+Our rules from Figure \ref{fig:wpdamd} are proven sound in Iris against an assembly-level hardware model
+implementing a fragment of x86-64 including 64-bit address translation with 4-level page tables.
+
+At the moment our proofs do rely on 11 small axioms of properties which should be provable, but
+are challenging to discharge due to some representation choices in our model.\footnote{See \lstinline|srx/x64/machine/current_axioms.v|}

### 2023-04-15 programlogic.tex
fix gname correction
settle the modal cr3 update

@@
 \end{figure}
-%\subsection{Machine State under Address Translation}
-%\label{sec:selectedinstrsemantics}
-%Althought we give the complete set of operational semantics rule in \sref{appendix:movops}, It is worth building the intuition on the the way some of these rules bahave in the context of address translation.
-In fact $\readlval\maddr{\storememstar\crval}\locsf$ is unfoled into multiple physical memory lookups for the final page address retrieval -- i.e.
-address translation traversal as shown in Figure \ref{fig:pagetables}:
-%\begin{itemize}
-%\item top-level-address translation: with the given root address (64-bit $\crval$) of the address space, performs the address translation, handling the first level (to get the PML4 entry) itself.
-The next level table address is computed with the fetched PML4 offset value which exhibits itself as 9-bit offset in $\kw{maddr}$.
-\mytodo{iso: put move}
-%\item translating from PML4 entry: performs the second level of address translation, to retrieve starting at the PML4 table entry, and interprets the PML4 entry that references a Page-Directory-Pointer Table (PDPT).
-We obtain the PDPT offset which exhibits itself as 9 bit offset in $\kw{maddr}$ to obtain the address of the next level page directory table (PDT) \mytodo{iso: put move}
-%\item translating from PD entry: performs the third level of address translation, to retrieve starting at the PDP table entry, and interprets a PDPTE that references a PD table.
-Likewise, we obtain the PD offset which exhibits itself as 9 bit offset in $\kw{maddr}$ to obtain the address of the next level page directory table (PT) \mytodo{iso:put move }
-%\item translate from PT entry: performs final level of address translation, starting from the PT entry, with a given 12 bit page offset, we can compute the physical address referencing $\locsf$ \mytodo{ismo: put mov}
-%\end{itemize}
 
@@
 We build up our context-dependent assertions from two basic types of assertions that do not care about address spaces.
-% We abstract physical memory addressing and registers naming with well-known separation logic assertion \textit{points-to} assertions.
-% \begin{enumerate}
-% \item Physical address points-to, $\pfpointsto\locsf\vpts\qfrac\ppts$
-% \todo[inline,color=red]{Colin says: This syntax is what most readers would expect for physical poinst-tos, but it's
-% inconsistent with the split 52+12  stuff later... Whatever we do, we need to be consistent}
-% \item Register points-to, $\pfpointsto\rg\rv\qfrac\rpts$
-% \end{enumerate}
 
@@
 \caption{A Strong Virtual Points-to Relation}
-\todo[inline,color=red]{Colin, I updated the definition with entry present}
   \label{fig:strongvirtualpointsto}
@@
 The global map itself is tracked by a resource algebra 
-\[\mathcal{A}\textsf{spaces} \stackrel{\triangle}{=} \textsc{Auth} (\; \mathcal{W}_{64} \;\rightarrow_{\textrm{fin}} \;  ( (\textsc{Frac }, \mathord{+}) \times (\textsc{Agree } \Loc,\mathord{=}) )) \]
-\todo[inline]{shouldn't above be map to gname, not these pairs?}
+\[\mathcal{A}\textsf{spaces} \stackrel{\triangle}{=} \textsc{Auth} (\; \mathcal{W}_{64} \;\rightarrow_{\textrm{fin}} \;  ( (\textsc{Frac }, \mathord{+}) \times (\textsc{Agree } \textsc{GhostName},\mathord{=}) )) \]
 
@@
 \label{sec:issues}
-An important subtlety arises with supporting \lstinline|mov|s into \lstinline|%cr3|. Consider the hypothetical rule:
+An important subtlety arises with supporting \lstinline|mov|s into \lstinline|%cr3|. Consider the hypothetical rule
 \begin{mathpar}
@@
 \end{mathpar}
+which is slightly simplified (by dropping per-address-space resources) for clarity in making the following point.
 This rule captures the intuitive change of address space in a Hoare triple.
@@
 }\\
-\inferrule[WriteToRegCtlFromReg]{
+\inferrule[WriteToRegCtlFromRegModal]{
   \{[\textsf{rtv}](P \ast \mathcal{I}\textsf{ASpace})\ast \mathcal{I}\textsf{ASpace} \ast R\ast r_s \mapsto_{r}\{q\}  \textsf{ rvs}  \}_{\textsf{rvs}} \overline{is}
@@
 \lstinline|mov| to $\textsf{rvs}$ afterwards.
-\todo[inline]{Colin says: the cr3 update rule should have IASpace (x2) and modalities...?}
+From that rule, with one step outside the \textsf{vProp} logic (so it is not quite a derived rule within
+our embedded logic), we can derive 
+\TirNameStyle{WriteToRegCtlFromRegModal},
+which is the (full) Hoare double version of the address space switch rule proposed in Section \ref{sec:issues}.
 

### 2023-04-15 intro.tex
clean up earlier todos

@@
 \item We verify simplified versions of several critical virtual-memory-related pieces of OS functionality, 
-      including mapping  pages, the assembly code for switching address spaces, and code which manipulates page
-      tables for \emph{currently mapped memory} without disturbing mappings.\todo[inline]{Colin says: I have optimistically
-      added this last example to the text here, because I think we can probably do it. Leave this TODO
-      here until we either do or cut it.}
-      The latter in particular goes beyond the capabilities of prior work.
+      including mapping  pages, and the assembly code for switching address spaces. 
+      The latter in particular goes beyond the capabilities of prior work, and reveals unexpected
+	subtleties.
 \end{itemize}

### 2023-04-15 examples.tex
bus draft

@@
 \subsection{Change of Address Space}
-A critical piece of \emph{trusted} code in verified OS kernels is the assembly code to change the current address space; current verified OS kernels currently lack effective ways to specify and reason about this low-level operation, for resaons outlined in Section \ref{sec:relwork}.
-\todo[inline]{probably not much more than just the cr3 update, but could be extended to a full context switch since the instruction pointer isn't updated in a swtch}
+A critical piece of \emph{trusted} code in verified OS kernels is the assembly code to change the current address space; current verified OS kernels currently lack effective ways to specify and reason about this low-level operation, for reasons outlined in Section \ref{sec:relwork}.
 \begin{figure}\footnotesize
@@
 in the specfication. 
-\todo{Colin: could you associate the rest of this paragraph with spec?}Thus the precondition of this code is that the then-current stack pointer has a return address expecting the then-current callee-save register values and suitablly updated (i.e., post-return) stack in the \emph{current} (initial) address space, while the stack pointer saved in the context to restore expects the same of the saved registers and stack \emph{in the other address space}.
-And the postcondition is analagous, but interpreted \emph{in a different address space}: the then-current (updated) stack has a return address expecting the new (restored) register values, and the saved context's invariant captures the precondition for restoring its execution \emph{in the previous address space}.
-At that point, it is safe to execute a return (or in a variant on this for user-mode and/or preemptive scheduling, a return-from-interrupt instruction).
+Thus the precondition of this code,
+in context, would include that the initial stack pointer (before \lstinline|rsp| is updated)
+has a return address expecting the then-current callee-save register values and 
+suitably updated (i.e., post-return) stack in the \emph{current} (initial) address space;
+this would be part of \textsf{P} in the precondition.
+The specification also requires that
+the stack pointer saved in the context to restore expects the same of the saved registers and stack 
+\emph{in the other address space}. 
+The other-space modality plays a critical role here; \textsf{Pother} would contain these assumptions in the other
+address space.
 
-In addition, immediately after the page table switch, the points-to information about the saved and restored contexts is \emph{also} guarded by a modality (Specification Line 52)
+Lines 10--16 save the current context into memory (in the current address space).
+Line 22 saves the initial page table root.
+Lines 33--38 begin restoring the target context, including the stack pointer (line 33),
+which may not be mapped in the address space at that time: it is the stack for the context being
+loaded into the CPU.
+The actual address switch occurs on line 45, which is verified with our modal rule for updating \lstinline|cr3|,
+and thus shifts resources in and out of other-space modalities as appropriate.
+
+The postcondition is analagous to the precondition, but interpreted \emph{in the new address space}: the then-current (updated) stack would have a return address expecting the new (restored) register values (again, in \textsf{Pother}),
+and the saved context's invariant captures the precondition for restoring its execution \emph{in the previous address space} (as part of \textsf{P}). 
+%At that point, it is safe to execute a return (or in a variant on this for user-mode and/or preemptive scheduling, a return-from-interrupt instruction).
+
+Note that immediately after the page table switch, the points-to information about the saved and restored contexts is \emph{also} guarded by a modality (Specification Line 52)
 \[ [\rtv](\texttt{Pcurrent}) \qquad \textsf{ where Pcurrent is all available logical assertions under addrress-space} \rtv\]
@@
 Thus many modern kernels have reverted to the older kernel design where the kernel inhabits its own unique address space, and user processes have only enough extra material mapped in their address spaces to switch into the kernel (CPUs do not speculate past updates to \texttt{cr3}).
-
-\todo[inline]{technically we weren't planning to prove reads of cr3...}
-\todo[inline]{need to figure out the right pre/post condition to highlight the change in address space, beyond the vpoints-to of the two structures}

### 2023-04-15 programlogic.tex
reg ptsto fix

@@
 \paragraph{Register points-to} 
-The assertion $\pfpointsto\rg\rv\rpts\qfrac$ ensures the ownership of the register $\rg$ naming the register value $\rv$. 
+The assertion $\textsf{r}\;\mapsto_{\textsf{r}}\{q\}\;\textsf{rv}$ ensures the ownership of the register $\rg$ naming the register value $\rv$. 
 The fraction $\qfrac$ with value 1 asserts the unique ownership of the register mapping, and grants update permission on it, 

### 2023-05-09 examples.tex
+
fix minor typographical weirdness
fix figures after submission-after-proof-fix

@@
 $\specline{\texttt{pte\_addr} \mapsto_{\texttt{vpte}} \paddr\; (\texttt{wzero 64})  \ast \ulcorner \texttt{entry\_present } \entryf \land \texttt{entry\_present } \entrytr \land  \texttt{entry\_present } \entrytw \urcorner}_{\rtv}$
-$\specline{\exists \texttt{ fpaddr} \ldotp \ulcorner \texttt{aligned fpaddr} \urcorner \ast \texttt{rax}\mapsto_{\textsf{r}} \texttt{fpaddr} \ast \texttt{fpaddr} \mapsto_{\textsf{p}} (\texttt{wzero 64}) \ast \ulcorner \texttt{entry\_present fpaddr}\urcorner}_{\rtv}$
+$\specline{\exists \texttt{ fpaddr} \ldotp \ulcorner \texttt{aligned fpaddr} \urcorner \ast \texttt{rax}\mapsto_{\textsf{r}} \texttt{fpaddr+3} \ast \texttt{fpaddr} \mapsto_{\textsf{p}} (\texttt{wzero 64}) \ast \ulcorner \texttt{entry\_present (fpaddr+3)}\urcorner}_{\rtv}$
 ;; Calculate new L1 entry
@@
 $\specline{  \nfpointsto{\mask\vaddr\maskthree\entryf}{\mask\vaddr\maskthreeoff\entryf}\entrytr\qtwo\naddr \ast \nfpointsto{\mask\vaddr\masktwo\entrytr}{\mask\vaddr\masktwooff\entrytr}\paddr\qthree\entryo \;\ast}_{\rtv}$
-$\specline{\texttt{pte\_addr} \mapsto_{\texttt{vpte}} \paddr \;(\texttt{fpaddr}) \; \ast \ulcorner \texttt{entry\_present } \entryf \land \texttt{entry\_present } \entrytr \land  \texttt{entry\_present } \entrytw \urcorner }_{\rtv}$
-$\specline{\exists \texttt{ fpaddr} \ldotp \ulcorner \texttt{aligned fpaddr} \urcorner \ast \texttt{rax}\mapsto_{r} \texttt{fpaddr} \ast \texttt{fpaddr} \mapsto_{p} (\texttt{wzero 64}) \ast \ulcorner \texttt{entry\_present fpaddr}\urcorner}_{\rtv}$
+$\specline{\texttt{pte\_addr} \mapsto_{\texttt{vpte}} \paddr \;(\texttt{fpaddr+3}) \; \ast \ulcorner \texttt{entry\_present } \entryf \land \texttt{entry\_present } \entrytr \land  \texttt{entry\_present } \entrytw \urcorner }_{\rtv}$
+$\specline{\ulcorner \texttt{aligned fpaddr} \urcorner \ast \texttt{rax}\mapsto_{\textsf{r}} \texttt{fpaddr+3} \ast \texttt{fpaddr} \mapsto_{\textsf{p}} (\texttt{wzero 64}) \ast \ulcorner \texttt{entry\_present fpaddr+3}\urcorner}_{\rtv}$
 $\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\; \sqsubseteq $
 $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast  \texttt{r14}\mapsto_{\textsf{r}} \texttt{pte\_addr} \ast \texttt{rdi}\mapsto_{\textsf{r}} \vaddr \ast }_{\rtv}$
-$\specline{\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\vaddr,\entryf,\entrytr,\entrytw,\fpaddr) \ast \ulcorner \theta \;!!\;\vaddr = \texttt{None}\urcorner \; \ast}_{\rtv}$
-$\specline{\ulcorner \texttt{aligned fpaddr} \urcorner \ast \texttt{rax}\mapsto_{\textsf{r}} \texttt{fpaddr} \ast \texttt{fpaddr} \mapsto_{\textsf{p}} (\texttt{wzero 64}) }_{\rtv}$
+$\specline{\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\vaddr,\entryf,\entrytr,\entrytw,\fpaddr+3) \ast \ulcorner \theta \;!!\;\vaddr = \texttt{None}\urcorner \; \ast}_{\rtv}$
+$\specline{\ulcorner \texttt{aligned fpaddr} \urcorner \ast \texttt{rax}\mapsto_{\textsf{r}} \texttt{fpaddr+3} \ast \texttt{fpaddr} \mapsto_{\textsf{p}} (\texttt{wzero 64}) }_{\rtv}$
 $\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\; \sqsubseteq $
@@
 \end{lstlisting}
-  \caption{Specification and proof of distilled code for mapping a new page with $\textsf{wpd\_def }_{\rtv}$ with a fixed namespace $\gammaPreds$ and a global map of address-spaces $m$  }
+  \caption{Specification and proof of distilled code for mapping a new page.}
 \label{fig:mapping_code}
@@
 fpaddr)) in Line 30.
+Reaching right at this point in the specification, we must note that we did find a trivial inconsistency in the postcondition related to conflating fpaddr as an address with the page table entry that reference fpaddr, and while the fix is straightforward we ran out of time.
 
@@
 ($\sqsubseteq$), and construct a PTE -points-to $\vaddr
-\mapsto_{\textsf{vpte}}\; \qfrac \;\fpaddr (\textsf{ wzero 64})$ (Specification
+\mapsto_{\textsf{vpte}}\; \qfrac \;\fpaddr\;(\textsf{wzero 64})$ (Specification
 Line 36) using $\sumwalkabs\vaddr\qfrac\fpaddr$ obtained from ghost

### 2023-05-09 programlogic.tex
fix figures after submission-after-proof-fix

@@
  Definition $\vaddr\mapsto_{\textsf{t}}\{\textsf{q}\}\; \vpage$ : vProp $\Sigma$ := 
-  $\exists_{\textsf{l4e l3e l2e l1}} \ldotp$  $\ulcorner$ aligned $\vaddr \urcorner \ast$ L4_L1_PointsTo(maddr l4e l3e l2e l1e) $\ast$ addr_L1$(\vaddr,\textsf{l1e}) \mapsto_{p}\{\mathsf{q}\} \vpage$.
- Definition L4_L1_PointsTo (maddr l4e l3e l2e l1e :word 64) : vProp $\Sigma$ :=
+  $\exists_{\textsf{l4e l3e l2e l1}} \ldotp$  $\ulcorner$ aligned $\vaddr \urcorner \ast$ L4_L1_PointsTo(maddr l4e l3e l2e l1e paddr) $\ast$ \textsf{paddr} $\mapsto_{p}\{\mathsf{q}\} \vpage$.
+ Definition L4_L1_PointsTo (maddr l4e l3e l2e l1e paddr :word 64) : vProp $\Sigma$ :=
  $\lambda$ cr3val.
@@
   \mathcal{I}\textsf{ASpace}(\ptablestore,m)\stackrel{\triangle}{=} \textsf{ASpace\_Lookup}(\ptablestore,m) \ast \\
- \qquad\qquad\qquad \bigast{(\vaddr, \textsf{l1e})\in \ptablestore}{\exists\;(\textsf{l4e l3e l2e})\ldotp \textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\textsf{ maddr, l4e, l3e, l2e, l1e})} \\
+ \qquad\qquad\qquad \bigast{(\vaddr, \textsf{paddr})\in \ptablestore}{\exists\;(\textsf{l4e l3e l2e, l1e, paddr})\ldotp \textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\textsf{ maddr, l4e, l3e, l2e, l1e, paddr})} \\
   \textsf{ where } \\

### 2023-05-10 programlogic.tex
+

@@
  Definition $\vaddr\mapsto_{\textsf{t}}\{\textsf{q}\}\; \vpage$ : vProp $\Sigma$ := 
-  $\exists_{\textsf{l4e l3e l2e l1}} \ldotp$  $\ulcorner$ aligned $\vaddr \urcorner \ast$ L4_L1_PointsTo(maddr l4e l3e l2e l1e paddr) $\ast$ \textsf{paddr} $\mapsto_{p}\{\mathsf{q}\} \vpage$.
+  $\exists_{\textsf{l4e l3e l2e l1}} \ldotp$  $\ulcorner$ aligned $\vaddr \urcorner \ast$ L4_L1_PointsTo(maddr l4e l3e l2e l1e paddr) $\ast$ paddr $\mapsto_{p}\{\mathsf{q}\} \vpage$.
  Definition L4_L1_PointsTo (maddr l4e l3e l2e l1e paddr :word 64) : vProp $\Sigma$ :=

### 2023-05-12 programlogic.tex
typo
+

@@
  Definition $\vaddr\mapsto_{\textsf{t}}\{\textsf{q}\}\; \vpage$ : vProp $\Sigma$ := 
-  $\exists_{\textsf{l4e l3e l2e l1}} \ldotp$  $\ulcorner$ aligned $\vaddr \urcorner \ast$ L4_L1_PointsTo(maddr l4e l3e l2e l1e paddr) $\ast$ paddr $\mapsto_{p}\{\mathsf{q}\} \vpage$.
- Definition L4_L1_PointsTo (maddr l4e l3e l2e l1e paddr :word 64) : vProp $\Sigma$ :=
+  $\exists_{\textsf{l4e l3e l2e l1}} \ldotp$  $\ulcorner$ aligned $\vaddr \urcorner \ast$ L4_L1_PointsTo($\vaddr$ l4e l3e l2e l1e paddr) $\ast$ paddr $\mapsto_{p}\{\mathsf{q}\} \vpage$.
+ Definition L4_L1_PointsTo ($\vaddr$ l4e l3e l2e l1e paddr :word 64) : vProp $\Sigma$ :=
  $\lambda$ cr3val.

### 2023-05-16 examples.tex
+

@@
  $\texttt{fpaddr} \mapsto_{a} (\texttt{wzero 64})$ to obtain the
-complete physical table-walk assertion (L4\_L1\_PointsTo(maddr l4e l3e l2e
+complete physical table-walk assertion (L4\_L1\_PointsTo($\vaddr$ l4e l3e l2e
 fpaddr)) in Line 30.

### 2023-05-16 programlogic.tex
+

--- 2023-05-10 programlogic.tex
@@
  Definition $\vaddr\mapsto_{\textsf{t}}\{\textsf{q}\}\; \vpage$ : vProp $\Sigma$ := 
-  $\exists_{\textsf{l4e l3e l2e l1}} \ldotp$  $\ulcorner$ aligned $\vaddr \urcorner \ast$ L4_L1_PointsTo(maddr l4e l3e l2e l1e paddr) $\ast$ paddr $\mapsto_{p}\{\mathsf{q}\} \vpage$.
+  $\exists_{\textsf{l4e l3e l2e l1e}} \ldotp$  $\ulcorner$ aligned $\vaddr \urcorner \ast$ L4_L1_PointsTo($\vaddr$ l4e l3e l2e l1e paddr) $\ast$ paddr $\mapsto_{p}\{\mathsf{q}\} \vpage$.
  Definition L4_L1_PointsTo (maddr l4e l3e l2e l1e paddr :word 64) : vProp $\Sigma$ :=
@@
 \begin{array}{l}
-    \vaddr\mapsto_{\textsf{v}}\{\textsf{q}\}\vale : \mathsf{vProp}~\Sigma \stackrel{\triangle}{=} \\
+    \vaddr\mapsto_{\textsf{v}}\{\textsf{q}\}\;\vale : \mathsf{vProp}~\Sigma \stackrel{\triangle}{=} \\
     \quad
@@
   \mathcal{I}\textsf{ASpace}(\ptablestore,m)\stackrel{\triangle}{=} \textsf{ASpace\_Lookup}(\ptablestore,m) \ast \\
- \qquad\qquad\qquad \bigast{(\vaddr, \textsf{paddr})\in \ptablestore}{\exists\;(\textsf{l4e l3e l2e, l1e, paddr})\ldotp \textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\textsf{ maddr, l4e, l3e, l2e, l1e, paddr})} \\
+ \qquad\qquad\qquad \bigast{(\vaddr, \textsf{paddr})\in \ptablestore}{\exists\;(\textsf{l4e l3e l2e, l1e, paddr})\ldotp \textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\vaddr\textsf{, l4e, l3e, l2e, l1e, paddr})} \\
   \textsf{ where } \\

### 2023-06-16 response_paper_563.txt
response

OVERVIEW
Thank you so much for your valuable reviews. For clarifying, here are our answers to provide more clarification.

CHANGE LIST

Review: 563A - "There are about 11 axioms admitted about details of the physical model.
       	        I'm not an absolutist about these things, and I don't mind admitting axioms that are, say, well-known mathematical facts,
	       	but that doesn't seem to be quite what's going on here. Similarly, for one of the case studies, the paper says that
	       	the proof got stuck because actually the stated post-condition isn't quite right, and there wasn't time to fix it,
	       	so the development has to admit an axiom related to that too."

		"What is the status of the remaining axioms? Have you made progress on removing them?"

       563B  - "At the moment our proofs do rely on 11 small axioms of properties which should be provable,
	       	but are challenging to discharge due to some representation choices in our model."
	       	"Reaching right at this point in the specification, we must note that we did find a trivial inconsistency in the postcondition related to conflating
	       	fpaddr as an address with the page table entry that reference fpaddr, and while the fix is straightforward we ran out of time."
	       	Both of these should have been mentioned much earlier and explained more clearly.""

>> Response: There was one crucial axiom that we used and introducing inconsistency in the post condition of page-insert example, and we fixed
	  it. Other axioms are assertions for the equivalances of stddp gmap (used in the logic) vs FMapInterface/List (used in our amd64 model) for register and pyhsical
	  memory map update/lookups, these do not impose any importance on our reasoining principles and semantics of amd64 model.

Review: 563A - "One more minor issue is that the paper suggest that prior to this work, hybrid modal separation logics have seen "no prior application".
  	        I don't think that's quite right. For example, the iGPS weak memory logic of Kaiser et al. 2017 uses hybrid logic for reasoning about C11 weak memory:
	   	in iGPS, propositions are indexed by "views" (roughly, the subset of memory operations a thread has seen).
	       	They embed these view-indexed propositions in Iris and have a similar modality P[V] saying "under view V, P holds".
	       	In fact, this is one of the motivations for why the Iris proof mode supports such "embedded" logics with these kinds of modalities so well,
	       	as described in Krebbers et al.'s 2018 MoSeL paper."

	563B -  "You introduce the vProp universe without any reference to the existing work.
		 Iris has a monPred construction for indexed iProps that subsumes your use case.
		 This construction has been used in 2 different Iris papers, combining program logics with hybrid logic in much the same way you do,
		 but to reason about weak rather than virtual memory. Your statement that this has not been done before is hence incorrect.
		 You should reference these papers and weaken your novelty claim on this front. *.
		 Hoang-Hai Dang, Jacques-Henri Jourdan, Jan-Oliver Kaiser, and Derek Dreyer. 2019.
		 RustBelt meets relaxed memory. Proc. ACM Program. Lang. 4, POPL, Article 34 (January 2020), 29 pages. https://doi.org/10.1145/3371102 *.
		 Hoang-Hai Dang, Jaehwang Jung, Jaemin Choi, Duc-Than Nguyen, William Mansky, Jeehoon Kang, and Derek Dreyer. 2022.
		 Compass: strong and compositional library specifications in relaxed memory separation logic.
		 In Proceedings of the 43rd ACM SIGPLAN International Conference on Programming Language Design and Implementation (PLDI 2022).
		 Association for Computing Machinery, New York, NY, USA, 792–808. https://doi.org/10.1145/3519939.3523451"

>> Response: We need to cite the papers introducing propositions indexed by a certain view of a thread P[V], monPred and others. (Colin, would you like to add more on this?)

Review: 563A - "Can you please clarify rule WriteToRegCtlFromReg? It seems to me that the P above the line should be under a modality since the page register has changed,
	        as in the more elaborate rule beneath it."

>> Response: WriteToRegCtlFromReg accidentially left there, it cannot coexists with the rule below (WriteToRegCtlFromRegModel), it is fixed.

Review: 563B - "The paper is not accessible to readers without Iris knowledge. Rather than focusing on the rules/interfaces you want your ghost resources to satisfy,
	        you show the entire resource algebra construction with little explanation. This makes explanations harder to follow and more convoluted than they need to be.
		The paper contains various notational inconsistencies, and is hand-wavy at times when formal rigour matters.
		For example, Figures 11 and 12 were very difficult to read for this reason (see below)."
>> Response: We conceptualize the logical resources and their interactions in Figure 7 and explain them in a Iris-independent way.
	  We also refined them based on your suggestions accordingly.

Review 563B - "The formal machinery seems to be there, but the evaluation is lacking: the examples you verify in the paper are toy examples of code fragments;
       	       they concern no more than a few assembly instructions, and have no interaction with surrounding code.

	       + The page mapping example is unfinished, as it only maps 64 bits (which you fail to mention until the end of the section),
	       	 relies on axioms as before, and makes some strange representational choices. I discuss this in detail in the comments below."

		 >> Response: True, our page-insertion example currently relies on axiomatized call ensuring the L1 entry, and we add the call/return/branchinng instructions in our change-list to be finished.
		              However, reasoning over multiple-page vs. single page, in principle, is a trivial extension to support (big-op) over a container of pages, and does not exhibit any interesting
			      logical challenge that  we would like to discuss and explain in the page-insert example.

	       + The address space switching example would be much more convincing if you actually verified it against a few concrete processes to obtain an end-to-end result.
	       	 Currently it is quite trivial, because you leave P and Pother opaque.
		 This makes it much harder to explain as well.

		 >> Response:

	       + I expected to see the result that you mentioned in the related work:
	         "Kolanski and Klein prove very useful embedding theorems stating that code that does not modify page table entries can be verified in a VM-ignorant program logic,
	       	 and that proofs in that logic can be embedded". This seems like a must-have if you intend to verify interesting examples.

		 >> Response:

	       + You should have at least one example that implements what you describe in section 4.3, and supports many address spaces.
	       	 You currently have no proof that this will actually work.

		 >> Response: (Colin, would you put some stuff here?)

	       + Does your paging need to be this specific to x86-64? Does it have to have four levels?
	       	 Could you somehow parameterize your implementation by the concrete paging implementation?
	       	 It is likely worthwhile to switch to a single- or two-level paging scheme for the initial examples.

		 >> Response: We build our principles on top of our amd64 model with 4-Levels paging.
		    	      There is no fundamental restriction to extend our reasoning principles to more than four-levels.
		              Right now, our reasoning principles are not shaped to be parameterized with paging-policy, but
			      since our reasoning principles are not paging-policy dependent, we can parameterize them on the paging-policy.

Review 563B - "Describe the exact subset of x86-64 that you support in more detail.
       	       It seems very limited right now. Motivate why this is still an interesting subset.
	       This would also avoid the aforementioned issue where the call instruction suddenly pops up without prior introduction."

>> Response: For us, "interesting" implies the necessity to model the address virtualization, i.e. pieces needed for address-translation.
   	     In this submission, particularly in the page-insertion example, we intended to show how we abstract away the page-table walk from the memory points-to relation
	     with our reasoning principles, and this was our primary intention to share in the paper.
	     It is true that, by the time we submitted this paper, we miss the "call" instruction, and relied on its axiomatization, then it is put in our change-list to be finished as the next step.

### 2023-06-16 response_paper_563.txt
+

@@
 	     It is true that, by the time we submitted this paper, we miss the "call" instruction, and relied on its axiomatization, then it is put in our change-list to be finished as the next step.
+
+Review 563B - "section 4.5: I am unconvinced by the justification for dropping the frame rule, both the argument that it is necessary and the argument that it is unproblematic..."
+
+>> Response: (Colin: I really don'y understand this text)
+
+Question 563B: "(related to remarks about section 4.5 above) Would it be possible to support (1) a frame rule that applies to Facts only together with
+	  	(2) a way to convert "P * cr3 = v" into "[v] P" and to convert "[v] P * cr3 = v" back into "P" and
+		(3) the assumption that [v] P is always a Fact."
+
+>> Response: All trivial by definition
+
+Questions 536B: "How would you share the kernel page tables in multiple processes?
+	  	 It seems you have hard-coded the fractions for the physical points-tos in Figure 4
+		 (although you only mention this on line 1016) under the assumption that no page table sharing occurs,
+		 so the sum of all fractions would be n in the case of n-way sharing.
+		 This is why I think having more complicated examples is required for evaluation."
+
+>> Response: Figure 4 has not been used in any of the reasoning. We have just mentioned it to motivate our virtual pointsto definition in Figure 5.
+   	     However, altough they are misleadingly used as hard-coded fractions in the code-base, we can define a fraction at a certain-level in terms of the previous level.
+	     At this moment, we do not support kernel-page-table management, and it is also becoming a recent attitude of OSes to be restrictive on the kernel page-table accesses (more isolation)
+	     after  "spectre meltdown".
+
+Question 563B: "Your virtual points-to chunks require full ownership of physical memory in Fig.5.
+	        It seems like in most cases (when you're not editing page tables) you do not care about the fraction of the page table mapping you possess,
+		but you do care about the fraction of physical memory.
+		-> It seems strange that the entailment va |->_v {q1 + q2} val |- va |->_v {q1} val \ast va |->_v {q2} val does not hold? ->
+		How would you go about mapping physical memory in multiple virtualized processes (e.g. when forking a process or sharing memory)?
+		It seems like you get into trouble because multiple processes need to own the same physical memory, which is not susceptible to the [r]-modality."
+
+>> Response:

### 2023-06-16 response_paper_563.txt
+

@@
 
->> Response:
+>> Response: First, there is a typo fix on the Figure 5 where the fraction has to appear on ->p, it is fixed .
+   	     On the other hand, we would introduce another level of abstraction to bookkeep the accesses to physical memory locations which are shared across the processes.
+	     This would look like and be on of the one we introduced of abstracting the physical-page-table walk away from virtual points-to, whereas, this abstraction would
+	     just hold the mappings from root addresses of address-spaces to the physical-memory locations accesses.

### 2023-06-16 response_paper_563.txt
+

@@
 	     just hold the mappings from root addresses of address-spaces to the physical-memory locations accesses.
+
+Question 563B: "Could you support the root of the page table being moved to a different physical address? Is it possible to change the index in ASpaces to make IAAspace hold again?"
+
+>> Response: Just making the IASpace to non-permanent assertion (which is in-fact something we used in the mechanized proof) introduces this capability.

### 2023-06-16 response_paper_563.txt
+

@@
 Thank you so much for your valuable reviews. For clarifying, here are our answers to provide more clarification.
-
-CHANGE LIST
+In general, we incorporated the structural changes and small comments as part our paper-presentation refinement.
+We fixed the inconsistency in the page-insertion example, and removed the axiom we used to introduce it after the fix.
 
@@
 
->> Response: (Colin: I really don'y understand this text)
+>> Response: (Colin: I am aware of the discussion we had but I really don'y understand this text)
+
+Review 563B - "Fig. 11 is frustratingly difficult to read in combination with the text, for various reasons: ..."
+
+>> Response: (Colin, I should speak about these with Colin)
 

### 2023-06-16 response_paper_563.txt
+

@@
 	       	Both of these should have been mentioned much earlier and explained more clearly.""
+
+      563C  -  "However the semantics of page tables is quite fiddly (as you appear to have discovered with the final axiom and associated comment), so I think care does need to be taken here!"
 
@@
 
-Review 563B - "Fig. 11 is frustratingly difficult to read in combination with the text, for various reasons: ..."
+Review 563B - "Fig. 11 is frustratingly difficult to read in combination with the text, for various reasons: "
 
->> Response: (Colin, I should speak about these with Colin)
+       	      + All of the interesting work required in Figure 11 happens inside the 2 opaque calls, which you do not show any implementation of.
+	      	I went to check your development, and your auxiliary methods are simply Axioms! This is misleading and should be signposted clearly in the paper.
+		Actually implementing and verifying these calls could have prevented many of the issues popping up in the text and the figure."
+
+		>> Response:
+
+	      + "Many details that you show in the proof are only partly explained (and in a convoluted way) in the text; e.g., is pte_addr just an alias pointing to pa for convenience reasons?
+	      	 Why does it point to wzero 64 and not some existentially quantified value (a function called ensure_L1_page does not sound like it will unmap the previous entry)? etc."
+
+		 >> Response: There aliasing to the ensured entry happens but then, as a widely adopted policy, OSes drops these aliases (not all). For the sake clarity in the example, we wanted
+		    get the entry already zeroed, but technically, it is same with initialization of an existentially quantified entry.
+
+	       + "Does aligned mean word-aligned or page aligned? From fig.11 I am getting the impression that it is the latter but Fig.4 gave the opposite impression.
+	       	  753 Where does this +3 come from? The text mentions +1, which I don't get either; fpaddr is said to be aligned, not fpaddr + 3. Additionally, 811 directly contradicts this.
+
+		 >> Response: (Colin, ask Colin for a polite explanation)
 
@@
 
->> Response: All trivial by definition
+>> Response: All trivial by definition.
 
@@
 >> Response: Just making the IASpace to non-permanent assertion (which is in-fact something we used in the mechanized proof) introduces this capability.
+
+Some Small Comments 536B: "The black later means something different than what you think in Iris; it is an endofunctor on the category of OFEs! What you want is the white triangle."
+
+>> Response: Yes, we are well-aware of this, thanks.

### 2023-06-16 response_paper_563.txt
some sketching

@@
 
->> Response: We need to cite the papers introducing propositions indexed by a certain view of a thread P[V], monPred and others. (Colin, would you like to add more on this?)
+>> Response: Thank you for the pointers, we will cite these and adjust our claims accordingly (and at some point shift to using monPred rather than our own pointwise lifting).
 
@@
 >> Response: WriteToRegCtlFromReg accidentially left there, it cannot coexists with the rule below (WriteToRegCtlFromRegModel), it is fixed.
+^^ Colin says: this is unclear to me
 
@@
 
-		 >> Response:
+		 >> Response: This is actually the correct form for actual context switching code.
+We can better explain the intended context of that example, but most kernels have a single function which implements a context switch.
+That particular function would be called with the current thread of the current address space (so it is callable in any address space), and with some arbitrary other target process address space (whichever the scheduler decided should run next).
+P and Pother are opaque because they are and should be parameters to the specification of the context switch function.
+Consider, e.g., OpenBSD's ```cpu_switchto``` (https://github.com/openbsd/src/blob/master/sys/arch/amd64/amd64/locore.S#LL273C20-L273C20 with the CR3 update on line 415); this is used to switch from any process that is blocking (or terminating) to any other process that is supposed to run, and an analogous assembly fragment exists in Linux, MacOS, Windows, FreeBSD....
+(that code snipped is dealing with additional scheduling metadata and more novel stack management than our example, which is focused on the core of the context switch.
 
@@
 
-		 >> Response: (Colin, would you put some stuff here?)
+		 >> Response: As mentioned in Section 4.3, the details of packaging all of the address spaces together is highly dependent on other kernel policies we are not attempting to model, and we do not believe a strawman will be informative.
+The simplest version, where all address spaces are simply disjoint, would simply involve another layer of ghost maps from process identifiers to $\Theta$ and $m$ for the particular address space.
 

### 2023-06-16 response_paper_563.txt
colin updates to response

@@
 Thank you so much for your valuable reviews. For clarifying, here are our answers to provide more clarification.
-In general, we incorporated the structural changes and small comments as part our paper-presentation refinement.
+We will incorporate the structural changes suggested to improve the exposition, as well as the smaller comments as part our paper-presentation refinement.
 We fixed the inconsistency in the page-insertion example, and removed the axiom we used to introduce it after the fix.
@@
 
->> Response: There was one crucial axiom that we used and introducing inconsistency in the post condition of page-insert example, and we fixed
-	  it. Other axioms are assertions for the equivalances of stddp gmap (used in the logic) vs FMapInterface/List (used in our amd64 model) for register and pyhsical
-	  memory map update/lookups, these do not impose any importance on our reasoining principles and semantics of amd64 model.
+>> Response: There was one crucial axiom that we used which allowed an inconsistency in the post condition of page-insert example.
+We have fixed it.
+The remaining axioms are assertions for the equivalances of stddp gmap (used in the logic) vs FMapInterface/List (used in our amd64 model) for register and pyhsical memory map update/lookups.
 
@@
 
->> Response: WriteToRegCtlFromReg accidentially left there, it cannot coexists with the rule below (WriteToRegCtlFromRegModel), it is fixed.
-^^ Colin says: this is unclear to me
+>> Response: WriteToRegCtlFromReg was accidentially left there, it is a pre-modal rule for updates (its assertions are iProps rather than vProps) used to build up the modal rule, and should have been omitted from the paper.
 
@@
 >> Response: We conceptualize the logical resources and their interactions in Figure 7 and explain them in a Iris-independent way.
-	  We also refined them based on your suggestions accordingly.
+	  We will also refine them based on your suggestions.
 
@@
 
-		 >> Response: True, our page-insertion example currently relies on axiomatized call ensuring the L1 entry, and we add the call/return/branchinng instructions in our change-list to be finished.
-		              However, reasoning over multiple-page vs. single page, in principle, is a trivial extension to support (big-op) over a container of pages, and does not exhibit any interesting
-			      logical challenge that  we would like to discuss and explain in the page-insert example.
+		 >> Response: The example does map a full page, but only proves the single entry is mapped; verifying the full-page mapping requires plumbing a big-op through the proof, which is longer but not fundementally difficult.
+We did axiomatize the preparation of the L1 entry; this is not unlike assumptions made in Kolanski and Klein's papers, where the actual preparation of the page is not modeled (but the retrieval of a page to map from a simple free-page-list is included in their work).
 
@@
 
-		 >> Response: This is actually the correct form for actual context switching code.
-We can better explain the intended context of that example, but most kernels have a single function which implements a context switch.
-That particular function would be called with the current thread of the current address space (so it is callable in any address space), and with some arbitrary other target process address space (whichever the scheduler decided should run next).
+		 >> Response: This is actually the correct form for actual context switching code for a general-purpose OS.
+We can better explain the intended context of that example, but most kernels have a single function which implements a context switch, which is used for every single context switch the OS performs, regardless of which processes are being switched from/to.
+That particular function would be called with the current thread of the current address space (so it is callable in any address space, and usually from multiple call sites, so even the caller context is not fixed), and with some arbitrary other target process address space (whichever the scheduler decided should run next, including switching to processes which were preempted).
 P and Pother are opaque because they are and should be parameters to the specification of the context switch function.
 Consider, e.g., OpenBSD's ```cpu_switchto``` (https://github.com/openbsd/src/blob/master/sys/arch/amd64/amd64/locore.S#LL273C20-L273C20 with the CR3 update on line 415); this is used to switch from any process that is blocking (or terminating) to any other process that is supposed to run, and an analogous assembly fragment exists in Linux, MacOS, Windows, FreeBSD....
-(that code snipped is dealing with additional scheduling metadata and more novel stack management than our example, which is focused on the core of the context switch.
+(that OpenBSD code snippet is dealing with additional scheduling metadata and more novel stack management than our example, which is focused on the core of the context switch and only integer/program registers).
 
@@
 
-		 >> Response:
+		 >> Response: Strictly speaking, Kolanski and Klein can prove that because their assertions and logical rules are defined inductively, making it possible to prove the embedding by induction over the rules in the VM-ignorant logic.
+In Iris neither assertions nor proof rules are closed, so such a result cannot be directly proven.
+It does not even make sense to talk about ideas like assertions that don't use certain constructs, since the user can add new vProps (or iProps in Iris in general).
+Instead we would highlight that our rules that do not manipulate page tables or cr3 are the same rules one would see for any other Hoare double logic, just passing IASpace and the cr3 value through without change.
 
@@
 		 >> Response: We build our principles on top of our amd64 model with 4-Levels paging.
-		    	      There is no fundamental restriction to extend our reasoning principles to more than four-levels.
+		    	      This is a current implementation limitation, but not a fundamental limitation.
 		              Right now, our reasoning principles are not shaped to be parameterized with paging-policy, but
-			      since our reasoning principles are not paging-policy dependent, we can parameterize them on the paging-policy.
+			      since our reasoning principles are not paging-policy dependent, we can parameterize them on the paging-policy; this amounts to
+                              making L4_L1_PointsTo and some associated properties regarding its accurate modeling of page table walks into a parameter.
 
@@
 
->> Response: For us, "interesting" implies the necessity to model the address virtualization, i.e. pieces needed for address-translation.
-   	     In this submission, particularly in the page-insertion example, we intended to show how we abstract away the page-table walk from the memory points-to relation
-	     with our reasoning principles, and this was our primary intention to share in the paper.
-	     It is true that, by the time we submitted this paper, we miss the "call" instruction, and relied on its axiomatization, then it is put in our change-list to be finished as the next step.
-
-Review 563B - "section 4.5: I am unconvinced by the justification for dropping the frame rule, both the argument that it is necessary and the argument that it is unproblematic..."
-
->> Response: (Colin: I am aware of the discussion we had but I really don'y understand this text)
+>> Response: For us, the subset is interesting because it is the subset needed to model the core challenges of reasoning about address translation.
+             The subset supported by the logic includes 64-bit register and memory mov instructions, mov with a 32-bit immediate to set a register, and adding a 32-bit immediate value to a 64-bit register.
+We can clarify this very early in the paper.
+The hardware model supports more (see ```external/amd64coq/src/Instructions.v```).
+             Our priority for this paper was the principles for reasoning with virtual memory, which can be explored with only the instructions we have rules for (though more convincing contextual examples, as noted, to benefit from use of call, hence our axiomatization of ensure_L1).
 
@@
 
-		>> Response:
+		>> Response: We can (and should) definitely signpost these aspects earlier to avoid surprising readers.
+However, we disagree that all of the interesting work happens in the auxilliary functions.
+The focus of our work is on manipulating the active page table mappings, and while ensure_L1 does abstract some of this (no more than prior work on logics for virtual memory) it retains both the final mapping operation and critically retains the creation of a new virtual points-to fact from constituent pieces.
 
@@
 
-		 >> Response: There aliasing to the ensured entry happens but then, as a widely adopted policy, OSes drops these aliases (not all). For the sake clarity in the example, we wanted
-		    get the entry already zeroed, but technically, it is same with initialization of an existentially quantified entry.
+		 >> Response: For the sake clarity in the example, we wanted to have the entry already zeroed.
+If it were not, the code snippet would also have to deal with deallocating the page that entry previously pointed to.
 
@@
 
-		 >> Response: (Colin, ask Colin for a polite explanation)
+		 >> Response: This word is unfortunately overloaded; our mechanization has separate predicates for being page-aligned and word-aligned.
+Figure 4 & 5 are talking about word alignment, but Figure 11 is dealing with page alignment.
+We can clarify this in the paper.
+                    You also noticed hints of a mistake in our specification of Figure 11 which we found and fixed shortly after submission: the code there is inconsistent about whether fpaddr is page aligned or has the two low-order bits set (i.e., whether it's a page-aligned address plus 0x3) to indicate a valid (bit 0) and read-write (bit 1, as opposed to read-only) page table entry.
+But the +1 and +3 are about setting the validity and read-write bits in the x86-64 page table entry.
 
@@
 
->> Response: All trivial by definition.
+>> Response: Yes, a frame rule that only frames out Facts would be possible for (1).
+(3) is trivially true by the definitions of Fact and the modality in Figure 8.
+(2) isn't quite right as written, as it is inconsistent about where the register resource for cr3 goes.
+But ```[v](P) * cr3=v -||- P * cr3=v``` holds for our model.
 
@@
 
->> Response: Figure 4 has not been used in any of the reasoning. We have just mentioned it to motivate our virtual pointsto definition in Figure 5.
-   	     However, altough they are misleadingly used as hard-coded fractions in the code-base, we can define a fraction at a certain-level in terms of the previous level.
-	     At this moment, we do not support kernel-page-table management, and it is also becoming a recent attitude of OSes to be restrictive on the kernel page-table accesses (more isolation)
-	     after  "spectre meltdown".
+>> Response: Figure 4 has not been used in any of the reasoning, it is a simpler baseline that we improve upon in Figure 5, which is what we actually use.
+             The typesetting there is slightly misleading, as q4 is actually defined in our implementation as 1/512 (since the L1 PTE maps 512 8-byte words),
+             q2 is 1/(512*512) since an L2 entry maps 512 L1 tables, and so on. You are correct that supporting page table sharing would require more flexible fractions;
+             as with other decisions, we have actively avoided committing to artificially specialized models of sharing that may not generalize to the actual sharing
+             that occurs in real kernels, the treatment of which is our longer-term goal. Our current hardware model has only a single core, so separating implication
+             (e.g., <shared page table portion> -* IASpace(T,m)) could be used to borrow the shared portion while running in another address space.
 
@@
 
->> Response: First, there is a typo fix on the Figure 5 where the fraction has to appear on ->p, it is fixed .
-   	     On the other hand, we would introduce another level of abstraction to bookkeep the accesses to physical memory locations which are shared across the processes.
-	     This would look like and be on of the one we introduced of abstracting the physical-page-table walk away from virtual points-to, whereas, this abstraction would
-	     just hold the mappings from root addresses of address-spaces to the physical-memory locations accesses.
+>> Response: First, there is a typo fix on the Figure 5 where the fraction q bound in the first line should appear in "pa ->p{q} val" on the second line, which is what our mechanization uses.
+So it is possible to have two virtual points-to assertions with shared fractional ownership of the data but which correspond to virtual address aliasing (whether within or across address spaces).
+The definition of our virtual points-to already supports that, though we have not proven rules to allow the construction of such a situation without unfolding the definition, and it is not obvious what those rules should be for cross-process sharing.
+More precisely, it is not clear what the appropriate logical modeling of transferring ownership between the processes should be.
+We consider this future work; our model is already more flexible than prior work and asking the paper to present a full solution to cross-process memory sharing is asking for a solution to a different (though complementary and related) problem than we set out to solve (logical foundations of address space management with a focus on task switching).
 
@@
 
->> Response: Just making the IASpace to non-permanent assertion (which is in-fact something we used in the mechanized proof) introduces this capability.
+>> Response: Making the IASpace a non-permanent assertion (which is in-fact something we used in the mechanized proof) makes this feasible.
 

### 2023-07-26 background.tex
crank on section 4, nearly done, along the way adding some hybrid logic comparison stuff
clean up earlier todos
try to clarify what makes a hybrid logic, and also talk about the weak memory papers

@@
     \includegraphics[width=\columnwidth]{pagetables.png}
-  \todo[inline]{this image got \emph{very} pixelated. is this possible to fix?}
     \caption{x86-64 page table lookups.}
@@
 with assertions of the form $[\ell](P)$ indicating that $P$ is true in the specific alternate circumstance (Kripke world)
- named by the \emph{nominal} $\ell$.
+ named by the \emph{nominal} $\ell$. Note that a distinctive property of hybrid logics is that, rather than hiding
+the points at which a modal assertion is evaluated inside the modality's definition, the choice of what world a modalized
+assertion should be true in is chosen \emph{in the assertion itself}. This allows assertions to talk about not simply whether some other assertion
+is true in some possible future or past world related in a fixed way to the current world, but to talk about \emph{arbitrary}
+other worlds. This explicit naming of alternate worlds increases the power of modal logics~\cite{blackburn1995hybrid}, and is actually
+necessary for completeness in classical separation logics~\cite{brotherston2014parametric}.
 
@@
 %Another pillar of our work is hybrid logic, a branch of modal logic with \emph{names} (called \emph{nominals} for states in Kripke models~\cite{blackburn1995hybrid,goranko1996hierarchies,areces2001hybrid,gargov1993modal}, in contrast to the typical flavor of modality which refers to an \emph{unspecified} (in the assertion) set of alternative circumstances.
-We are aware of only two prior works combining hybrid logics with program logics specifically.
-Brotherston and Villard~\cite{brotherston2014parametric} demonstrated that may properties true of various separation logics are not definable in boolean \BI (\BBI), and showed that a hybrid extension \HyBBI allows most such properties to be defined (e.g., the fact that separating conjunction is cancellative is unprovable in boolean \BI, but provable in \HyBBI).
-There, nominals named resources (roughly, but not exactly, heap fragments).
-Gordon~\cite{gordon2019modal} described a use of hybrid logic in the verification of actor programs, where nominals named the local state of individual actors (with such assertions stabilized with a rely/guarantee approach).
-Beyond these, there is limited work on the interaction of specifically \emph{hybrid logic} with substructural logics.
-Primarily there is a line of work on hybrid linear logic (\HyLL)~\cite{despeyroux2014hybrid}, originally used as a way to more conveniently express aspects of transition systems in linear logic.
-However, \HyLL's proof rules offer no non-trivial interactions with multiplicative connectives (every \HyLL proof can in fact be embedded into regular linear logic~\cite{chaudhuri2019hybrid}, unlike Brotherston and Villard's \HyBBI, which demonstrably increases expressive power over its base \BBI.\colinsays{some of this is too detailed for this point in the paper, needs to go to related work.
-But I'll split it later.}
-%
-%In both \HyLL and \HyBBI, nominals denote worlds with monadic structure (as worlds in Kripke semantics for either LL or \BBI necessarily have monoidal structure).
-Our nominals, by contrast, do not name worlds in the same sense with respect to Iris's CMRAs, but in fact \emph{classes} of worlds, because the names are locations (a means of \emph{selecting} resources) rather than resources.
-A key difference is that the use of nominals in those logics corresponds specifically to hypothetical reasoning about resources (until a nominal is connected to a current resource, in which case conclusions can be drawn about the current resource), which means the modalities themselves do not ``own'' resources.
-Instead, our hybrid modalities directly own portions of the data structure...\colinsays{also need to move to related work}
 %

### 2023-07-26 vmm_arxiv.tex
tweaks so latexpand can generate a working tex file for arxiv

\pdfoutput=1
%
%
%
%
%
%
%
%
%
%
%
%
%
%
%
%
%
%
%
%
%
%
%
%
%
%
%
%
%
%
%
%
%
%
%
%
%
%
%
%
%
%
%
%
%
%
%
\documentclass[acmsmall,screen,nonacm]{acmart}
\settopmatter{printacmref=false}
%
%
\AtBeginDocument{%
  \providecommand\BibTeX{{%
    Bib\TeX}}}

%
%
%
%
%
\setcopyright{none}
%
\copyrightyear{2018}
\acmYear{2018}
\acmDOI{XXXXXXX.XXXXXXX}

%
\acmConference[Conference acronym 'XX]{Make sure to enter the correct
  conference title from your rights confirmation emai}{June 03--05,
  2018}{Woodstock, NY}
%
%
%
%
%
%
\acmPrice{15.00}
\acmISBN{978-1-4503-XXXX-X/18/06}

%
%
%
%
%
%

%
%
%
%
%
%
%
%
%
%
%
%

%
%
%
%
%
%
%
%
%
%
%
%
%
%
%

\usepackage{hyperref}
\usepackage{array}
%
\usepackage{proof}
\usepackage{stmaryrd}
\usepackage{xspace}
%
\usepackage{graphicx}

\usepackage{lipsum}
\usepackage[utf8]{inputenc}
%
\usepackage{blindtext}
\usepackage{listings}
\lstset{
  columns=fullflexible,
  numbers=left,
  basicstyle=\ttfamily,
  keywordstyle=\color{blue}\bfseries,
  morekeywords={mov,add,call},
  emph={rsp,rdx,rax,rbx,rbp,rsi,rdi,rcx,r8,r9,r10,r11,r12,r13,r14,r15},
  emphstyle=\color{green},
  emph={[2]cr3},
  emphstyle={[2]\color{violet}},
  morecomment=[l]{;;}
}
%
%
\newcommand{\setbackgroundcolour}{\pagecolor[rgb]{0.19,0.19,0.19}}  
\newcommand{\settextcolour}{\color[rgb]{0.77,0.77,0.77}}    
\newcommand{\invertbackgroundtext}{\setbackgroundcolour\settextcolour}
%
%
%
\usepackage{wrapfig}
\newtheorem*{remark}{Remark}

%
\usepackage{amsmath}

%
\usepackage{cleveref}
\crefformat{footnote}{#2\footnotemark[#1]#3}
\usepackage[T1]{fontenc}
\usepackage{fontawesome}
\usepackage[utf8]{inputenc}
\usepackage{iris}
\usepackage{mathpartir}
\renewcommand{\TirNameStyle}[1]{\hypertarget{#1}{\textsc{#1}}}
\newcommand{\RULE}[1]{\hyperlink{#1}{\textsc{#1}}\xspace}
\usepackage{xspace}
%
%
\newcommand{\sref}[1]{\S\ref{#1}}
\newcommand{\fref}[1]{Figure~\ref{#1}}
%
\def\etal.{\emph{et al.}}
\newcommand{\SL}{Separation Logic\xspace}
\newcommand{\spacelang}{SpaceLang\xspace}
\let\ar\rightarrow
\newcommand{\iProp}{\mathit{iProp}}
\newcommand{\lc}{$\lambda$-calculus\xspace}

%
%
\newcommand{\logical}{logical\xspace}
\newcommand{\logically}{logically\xspace}
\newcommand{\Logical}{Logical\xspace}
%
\newcommand{\tlfoff}[1]{\kw{pml4off}:#1}
\newcommand{\tltoff}[1]{\kw{pdpoff}:#1}
\newcommand{\tltwoff}[1]{\kw{pdoff}:#1}
\newcommand{\tlooff}[1]{\kw{ptoff}:#1}
\newcommand{\tpgoff}[1]{\kw{pageoff}:#1}
\newcommand{\lfoff}{\kw{pml4off}}
\newcommand{\ltoff}{\kw{pdpoff}}
\newcommand{\ltwoff}{\kw{pdoff}}
\newcommand{\looff}{\kw{ptoff}}
\newcommand{\pgoff}{\kw{pageoff}}
\newcommand{\crt}{\kw{cr3}}
\newcommand{\lt}{\kw{pdp}}
\newcommand{\ltw}{\kw{pd}}
\newcommand{\lo}{\kw{pt}}
\newcommand{\pg}{\kw{page}}
\newcommand{\tlf}[1]{\kw{cr3}:#1}
\newcommand{\tlt}[1]{\kw{pdp}:#1}
\newcommand{\tltw}[1]{\kw{pd}:#1}
\newcommand{\tlo}[1]{\kw{pt}:#1}
\newcommand{\tpg}[1]{\kw{page}:#1}
%
\newcommand{\modalPs}[1]{\vec#1}
\newcommand{\modalP}{P}
\newcommand{\modalQ}{Q}
\newcommand{\modaldefper}[2]{@[#1]\;\square\;#2}
\newcommand{\outsidemodaldefper}[2]{\square \; @[#1]\; #2}
\newcommand{\modaldef}[2]{@[#1]\; #2 }
\newcommand{\modaldefunfold}[2]{ #2 \; #1}
\newcommand{\modalstar}[2]{#1 \star #2}

%
\newcommand{\qfracone}{\kw{1}}
\newcommand{\qfracfot}{\kw{q}/512}
\newcommand{\qfracfots}{\kw{q}/512^{2}}
\newcommand{\qfracfotss}{\kw{q}/512^{4}}
\newcommand{\qfracfotsss}{\kw{q}/512^{8}}

\newcommand{\qfrac}{\kw{q}}
\newcommand{\qfractw}{\kw{q}/2}
\newcommand{\qfract}{\kw{q}/3}
\newcommand{\qfracf}{\kw{q}/4}
\newcommand{\naddr}{\kw{a}}
\newcommand{\vaddr}{\kw{va}}
\newcommand{\vpts}{\kw{v}}
\newcommand{\ppts}{\kw{p}}
\newcommand{\rpts}{\kw{r}}
\newcommand{\pfpointsto}[4]{#1\mapsto_{#4}\{#3\}\;#2}
\newcommand{\ppointsto}[3]{#1\mapsto_{#3}\;#2}
\newcommand{\npointsto}[4]{#1 / #2 \mapsto_{#4}\;#3}

\newcommand{\nfpointsto}[5]{#1\sim#2 \mapsto_{#5}\{#4\}\;#3}
\newcommand{\maskfour}{\textsf{l4M52}}
\newcommand{\maskfouroff}{\textsf{l4off}}
\newcommand{\maskthree}{\textsf{l3M52}}
\newcommand{\maskthreeoff}{\textsf{l3off}}
\newcommand{\masktwo}{\textsf{l2M52}}
\newcommand{\masktwooff}{\textsf{l2off}}
\newcommand{\maskone}{\textsf{l1M52}}
\newcommand{\maskoneoff}{\textsf{l1off}}
%
\newcommand{\mask}[3]{(#2\; #1 \; #3)}
\newcommand{\ft}{52}
\newcommand{\tw}{12}
\newcommand{\pointsto}{\mapsto}
\newcommand{\mystackrel}[2]{#2_{#1}}
\newcommand{\fpointsto}[1]{\mathrel{\mystackrel{#1}{\mapsto}}}
\newcommand{\pointedby}{\mapsfrom}
\newcommand{\vfpointedby}[1]{\mathrel{\mystackrel{#1}{\mapsfrom}}}
\newcommand{\vpointedby}{\mathrel{\mapsfrom}}
\newcommand{\mydagger}{\mathord{\dagger}}
\renewcommand{\ddag}{\mydagger\kern-0.8mm\mydagger}
\newcommand{\ddagsingleton}[1]{\ddag\{#1\}}
%
\newcommand{\ocloud}[3]{#3\;\text{\faCloud}^{#2}\,#1}
\newcommand{\cloud}[2]{\ocloud{#1}{#2}{#1}}
\newcommand{\pure}[1]{#1}
%
\newcommand{\rcpointedby}{\mathrel{\mapsfrom}}
\newcommand{\rcvpointedby}{\mathrel{\mapsfrom}}
\newcommand{\rcvfpointedby}[1]{\mathrel{\mapsfrom}_{#1}}
\newcommand{\nmaster}{m}
\newcommand{\nv}{n}
%
\newcommand{\centralinvariant}{I}
\newcommand{\generalentry}{\textsf{ent}}
\newcommand{\entryf}{\textsf{l4e}}
\newcommand{\entrytr}{\textsf{l3e}}
\newcommand{\entrytw}{\textsf{l2e}}
\newcommand{\entryo}{\textsf{l1e}}
\newcommand{\crthree}{\textsf{CR3}}
\newcommand{\paddr}{\textsf{pa}}
\newcommand{\vsome}{\_}
\newcommand{\vpage}{\textsf{v}}
%
\newcommand{\kw}[1]{\mathsf{#1}}

%

%
\newcommand{\pageptstosum}[2]{(#1+#2)}
\newcommand{\lvlsum}[2]{(#1+8*#2)}
\newcommand{\lvlbor}[1]{(#1 | \kw{w0}_{3})}
\newcommand{\Locsx}{\mathcal{W}_{16}}
\newcommand{\locsx}{\kw{w}_{16}}
\newcommand{\Locsz}{\mathcal{W0}_{16}}
\newcommand{\locsz}{\kw{w0}_{16}}

\newcommand{\Locn}{\mathcal{W}_{9}}
\newcommand{\locn}{\kw{w}_{9}}

\newcommand{\Loctw}{\mathcal{W}_{12}}
\newcommand{\loctw}{\kw{w}_{12}}
\newcommand{\Locsf}{\mathcal{W}_{64}}
\newcommand{\locsf}{\kw{w}_{64}}
\newcommand{\Locft}{\mathcal{W}_{52}}
\newcommand{\locft}{\kw{w}_{52}}

\newcommand{\rvsrc}{\kw{rv}_{\textsf{src}}}
\newcommand{\rgsrc}{\kw{r}_{\textsf{src}}}
\newcommand{\rv}{\kw{rv}}
\newcommand{\rg}{\kw{r}}
\newcommand{\rvdst}{\kw{rv}_{\textsf{dst}}}
\newcommand{\rgdst}{\kw{r}_{\textsf{dst}}}

\newcommand{\mvsrc}{\kw{rv}_{\textsf{src}}}
\newcommand{\mgsrc}{\kw{r}_{\textsf{src}}}
\newcommand{\mv}{\kw{rv}}
\newcommand{\mg}{\kw{r}}
\newcommand{\mvdst}{\kw{rv}_{\textsf{dst}}}
\newcommand{\mgdst}{\kw{r}_{\textsf{dst}}}

\newcommand{\Loc}{\mathcal{W}_n}
\newcommand{\loc}{\kw{w}_n}
\newcommand{\regset}{\kw{greg}}
\newcommand{\regvaltype}{\kw{regval}}
\newcommand{\reg}{\kw{r}}
\newcommand{\regval}{\kw{rv}}
%
\newcommand{\xs}{\mathit{xs}}
%
\newcommand{\pre}{\Phi}
\newcommand{\post}{\Psi}
\newcommand{\midpoint}{\chi}
%
\newcommand{\ql}{p}
\newcommand{\qv}{q}
\newcommand{\qw}{q'}
%
\newcommand{\lsv}{L}
\newcommand{\lsw}{L'}
%
\newcommand{\locs}{D}
\newcommand{\antecedents}{P}

%
%
\newcommand{\metaif}[3]{#1 \mathrel{?} #2 : #3}

%
\newcommand{\val}{v}
\newcommand{\vals}{\vec\val}
\newcommand{\wal}{v'}
\newcommand{\vunit}{()}
\newcommand{\vk}{k}
\newcommand{\vcode}[2]{\lambda#1.#2}
\newcommand{\vars}{\vec\var}

%
\newcommand{\lval}{\varrho}
\newcommand{\var}{x}
\newcommand{\sloc}{c}
\newcommand{\src}{r}
\newcommand{\dst}{s}
\newcommand{\dsts}{\vec\dst}
\newcommand{\lvals}{\vec\lval}

\newcommand{\ofs}{\mathit{o}}
\newcommand{\maddr}{\kw{maddr}}
\newcommand{\offs}{\kw{offset}}
%
\newcommand{\deref}{\mathord{*}}
\newcommand{\assign}{=}
\newcommand{\instr}{i}
\newcommand{\mvrr}{\kw{mvrr}}
\newcommand{\mvrmb}{\kw{mvrmb}}
\newcommand{\mvrmo}{\kw{mvrmo}}
\newcommand{\mvmrb}{\kw{mvmrb}}
\newcommand{\mvmro}{\kw{mvmro}}
\newcommand{\instrexpr}{\kw{ie}}
\newcommand{\instrs}{\ensuremath{\vec\instr}}
\newcommand{\iskip}{\ensuremath{\kw{skip}}}
\newcommand{\iseq}[2]{\ensuremath{#1; #2}}
\newcommand{\ising}[1]{\kw{sexec} (#1)}
\newcommand{\icall}[2]{\deref#1(#2)}
\newcommand{\ialloc}[2]{\deref#1 \assign \kw{alloc}\;#2}
\newcommand{\iload}[3]{\deref#1 \assign [\deref#2+#3]}
\newcommand{\istore}[3]{[\deref#1+#2] \assign \deref#3}
\newcommand{\iloceq}[3]{\deref#1 \assign (\deref#2 == \deref#3)}
\newcommand{\iconst}[2]{\deref#1 \assign #2}
\newcommand{\imove}[2]{\deref#1 \assign \deref#2}
\newcommand{\ialloca}[2]{\kw{alloca}\,#1\,\kw{in}\,#2}
\let\iallocaactive\ialloca
\newcommand{\ifork}[3]{\kw{fork}\,\deref#1\,\kw{as}\,#2\,\kw{in}\,#3}

%

\newcommand{\movctl}{\kw{movctl}}
\newcommand{\imreg}{\kw{imreg}}
\newcommand{\amode}{\kw{amode}}
\newcommand{\amodeb}[1]{\kw{amode}  #1}
\newcommand{\amodeo}[2]{\kw{amode} = #1 + #2}
\newcommand{\imov}[3]{#1 \; #2 \; #3}
%
\newcommand{\blk}{b}
\newcommand{\btuple}[1]{#1}
\newcommand{\bcell}[1]{\langle#1\rangle}
\newcommand{\bdeallocated}{\text{\normalfont\footnotesize\faBolt}}
\newcommand{\sz}[1]{\mathit{size}(#1)}
\newcommand{\replicate}[2]{#2^{#1}}
\newcommand{\pointers}[1]{\mathit{pointers}(#1)}

%
\newcommand{\storereg}{\sigma.\mathcal{R}}
\newcommand{\storemem}{\sigma.\mathcal{M}}
\newcommand{\storememstar}[1]{\sigma.\mathcal{M}^{*[#1]}}
\newcommand{\store}{\sigma}
\newcommand{\storeregprime}{\sigma'.\mathcal{R}}
\newcommand{\storememprime}{\sigma'.\mathcal{M}}
\newcommand{\storememprimestar}[1]{\sigma'.\mathcal{M}^{*[#1]}}

\newcommand{\storeprime}{\sigma'}

\newcommand{\logicalstore}{\theta}
\newcommand{\maxsize}{S}
%
\newcommand{\valid}[1]{\sz{#1}\leqmaxsize}
\newcommand{\available}[1]{\mathit{available}(#1)}
\newcommand{\closed}[1]{\text{$#1$ is closed}}
\newcommand{\related}[2]{#1\approx#2}

%
\newcommand{\predstore}{\pi}
\newcommand{\phinvariant}[2]{#1\bowtie#2}

%
\newcommand{\ectx}{K}
\newcommand{\hole}{[]}
\newcommand{\efill}[2]{#1[#2]}

%
\newcommand{\subst}[2]{[#2/#1]}
\newcommand{\readlval}[3]{#2(#1)=\bcell{#3}}
\newcommand{\writelvalsidecondition}[4]{
  #3(#1) = \bcell{\_}
}
\newcommand{\writelvalmaincondition}[4]{
  #4 = \mupd{#1}{\bcell{#2}}{#3}
}
%
%
%
%
\newcommand{\writelval}[4]{
  #4 = \langle #1 := #2 \rangle #3
}

\newcommand{\sexec}[4]{\kw{sexec}\(#1 #2 #3 #4\)}
\newcommand{\headstep}{\longrightarrow}
\newcommand{\cfg}[2]{#1\;/\;#2}
\newcommand{\hs}[4]{\cfg{#1}{#2}\headstep\cfg{#3}{#4}}
\newcommand{\hsfork}[5]{\cfg{#1}{#2}\headstep\cfg{#3}{#4}\mathrel{\textit{spawning}}#5}
\newcommand{\hsforknl}[5]{\begin{array}{@{}r@{}}\cfg{#1}{#2}\headstep\cfg{#3}{#4}\\\textit{spawning}\;#5\end{array}}
\newcommand{\hsnostore}[2]{\hs{#1}\store{#2}\store}
%
\newcommand{\gcSymbol}{\text{\normalfont\footnotesize\faTrashO}}
\newcommand{\gc}[2]{#1\mathrel{\gcSymbol}#2}
\newcommand{\headstepgc}{\mathrel{\;\gcSymbol\!\!\longrightarrow\,}}
\newcommand{\hsgc}[4]{\cfg{#1}{#2}\headstepgc\cfg{#3}{#4}}
\newcommand{\hsgcfork}[5]{\cfg{#1}{#2}\headstepgc\cfg{#3}{#4}\mathrel{\textit{spawning}}#5}

%
\newcommand{\alen}[1]{\mathord\mid#1\mathord\mid}

%
\newcommand{\varray}[1]{\begin{array}{@{}c@{}}#1\end{array}}
\newcommand{\starvarray}[1]{\bgroup\renewcommand{\star}{\\}\varray{#1}\egroup}
\newcommand{\BRACE}[1]{\left\{#1\right\}}
\newcommand{\BRACEvarray}[1]{\BRACE{\varray{#1}}}
\newcommand{\BRACEstarvarray}[1]{\BRACE{\starvarray{#1}}}
\newcommand{\triple}[3]{\{#1\}\;#2\;\{#3\}}
\newcommand{\atriple}[3]{\triple{#1}{&#2&}{#3}}
\newcommand{\vtriple}[3]{\varray{\{#1\}\\#2\\\{#3\}}}
\newcommand{\bigvtriple}[3]{\varray{\BRACEvarray{#1}\\#2\\\BRACEvarray{#3}}}
\newcommand{\bightriplehskip}{\;}
\newcommand{\bightriple}[3]{\BRACEstarvarray{#1}\bightriplehskip#2\bightriplehskip\BRACEstarvarray{#3}}
\newcommand{\bightriplex}[4]{\BRACEstarvarray{#1}\bightriplehskip#2\bightriplehskip\BRACE{\exists#3.\;\starvarray{#4}}}
\newcommand{\bigsupd}[2]{\BRACEstarvarray{#1}\supd\BRACEstarvarray{#2}}
%
\renewcommand{\wp}[2]{\mathit{wp}\;#1\;#2}

%
%
\let\ordinarystar\star
\DeclareMathOperator*{\Sep}{\scalerel*{\ast}{\sum}}
\newcommand{\bigast}[2]{\mathop{\textstyle\Sep}\limits_{#1}\,#2}
\let\ordinarybigast\bigast
\newcommand{\later}{\triangleright\;}

\newcommand{\logequiv}{\;\equiv\;}
\newcommand{\supd}{\Rrightarrow_\centralinvariant}
\newcommand{\fupd}{\Rrightarrow}
\let\ordinarysupd\supd
\newcommand{\iFalse}{\mathit{False}}
\newcommand{\iTrue}{\mathit{True}}

%
\newcommand{\singleton}[1]{\{#1\}}
\newcommand{\dom}[1]{\mathit{dom}(#1)}

%
\newcommand{\multiplicity}[2]{#1 \mathop{\text{\small\$}} #2}
\newcommand{\cardinal}[1]{|#1|}
%
\newcommand{\valoccs}[2]{#1 \mathop{\$} #2}

%
\newcommand{\singletonMap}[2]{[#1 := #2]}
\newcommand{\mupd}[2]{[#1 := #2]} %
\newcommand{\mext}[2]{[#1 \mathbin{+\!\!=} #2]} %

%
\newcommand{\JoinPointsto}{Join$\mapsto$}
\newcommand{\JoinPointedBy}{Join$\mapsfrom$}
\newcommand{\CovPointedBy}{Covariance$\mapsfrom$}
\newcommand{\ConfrontMapstoMapsfrom}{Confront$\mapsto\mapsfrom$}
\newcommand{\ConfrontDagMapsfrom}{Confront$\ddag\mathord{\mapsfrom}$}
\newcommand{\ZeroDiams}{Zero$\diamond$}
\newcommand{\JoinDiams}{Join$\diamond$}
\newcommand{\ZeroDag}{Zero$\ddag{}$}
\newcommand{\JoinDag}{Join$\ddag{}$}

\newcommand{\mathsmash}[1]{\ensuremath{\smash{#1}}}

%

%
\newcommand{\crval}{\kw{cr3val}}
\newcommand{\plusaddr}[2]{#1 + #2}
%

%

\newcommand{\listcopy}{\textit{copy}}
\newcommand{\self}{\textit{self}\,}
\newcommand{\destination}{\textit{dst}}
\newcommand{\source}{\textit{src}}
\newcommand{\etiquette}{\textit{tag}}
\newcommand{\head}{\textit{head}}
\newcommand{\tail}{\textit{tail}}
\newcommand{\res}{\destination'}
\newcommand{\callee}{\textit{f}\,}
%
\newcommand{\logval}{\mathit{v}}
\newcommand{\loglist}{\mathit{vs}}
%
\newcommand{\isListName}{\textit{isList}}
\newcommand{\isList}[2]{\isListName\;#1\;#2}
%
\newcommand{\lognil}{[]}
\newcommand{\logcons}[2]{#1 :: #2}
\newcommand{\loglistbrackets}[1]{[#1]}

%

%

\newcommand{\stackcreate}{\textit{create}}
\newcommand{\stackpush}{\textit{push}}
\newcommand{\stackpop}{\textit{pop}}
\newcommand{\isStack}[2]{\textit{isStack}\;#1\;#2}
\newcommand{\stack}{\textit{stack}}
\newcommand{\elem}{\textit{elem}}
\newcommand{\vns}{\mathit{vns}}
\newcommand{\nonreccallee}{\ensuremath{f}}

\newcommand{\coq}{\textsc{Coq}\xspace}
\newcommand{\HyLL}{\textsc{HyLL}\xspace}
\newcommand{\BI}{\textsc{BI}\xspace}
\newcommand{\BBI}{\textsc{BBI}\xspace}
\newcommand{\HyBBI}{\textsc{HyBBI}\xspace}
\newcommand{\iris}{\textsc{Iris}\xspace}
 \newcommand{\mytodo}[1]{%
  \todo[linecolor=white, bordercolor=white, textcolor=white]{#1}%
}

%
%
%
%
%
%

%
%
%
%
%
%
%
%
%
%
%
%

%
%
%
%
%
%
%
%
%
%
%

%
%
\begin{document}

%
%
%
\title{Modal Abstractions for Virtualizing Memory Addresses}
%
%
%
%
%
%
\author{Ismail Kuru}
\email{ik335@drexel.edu}
\author{Colin S. Gordon}
\email{csgordon@drexel.edu}
\affiliation{%
  \institution{Drexel University}
  \city{Philadelphia, PA}
  \country{USA}
}

%
%
%
%
%
%
\renewcommand{\shortauthors}{Kuru and Gordon}

%
%
%
\begin{abstract}
Operating system kernels employ virtual memory subsystems, which use a CPU's memory management units (MMUs) to virtualize the addresses of memory regions:
a logical (virtual) address is translated to a physical address in memory by the MMU based on
kernel-controlled page tables -- a hardware-defined sparse tree-map structure -- stored in memory, which itself is 
accessed by the kernel through virtual addresses.
Operating systems manipulate these virtualized memory mappings to isolate untrusted processes,
 restrict which memory is accessible to different processes, 
hide memory limits from user programs, 
ensure process isolation, implement demand-paging and copy-on-write behaviors for performance
and resource controls.
At the same time, misuse of MMU hardware can lead to kernel crashes.

Virtual memory management (VMM) code is a critical piece of general-purpose OS kernels, but verification of this functionality
is challenging due to the complexity of the hardware interface (the page tables are updated via writes to those
memory locations, using addresses which are themselves virtualized).
Prior work on verification of VMM code has either only handled a single address space, trusted significant
pieces of assembly code, or resorted to direct reasoning over machine semantics rather than
exposing a clean logical interface.

In this paper, we introduce a modal abstraction to describe
the truth of assertions relative to a specific virtual address space: [r]P indicating that P holds in the
virtual address space rooted at r. Such modal assertions 
allow different address spaces to refer to each other, enabling complete verification of instruction sequences
manipulating multiple address spaces. Using them effectively requires working with other assertions,
such as points-to assertions in our separation logic, as relative to a given address space.
We therefore define virtual points-to relations, which mimic hardware address translation,
relative to a page table root.
We demonstrate our approach with challenging fragments of VMM code showing that our approach
handles examples beyond what prior work can address, including reasoning about
a sequence of instructions as it changes address spaces.
All definitions and theorems mentioned in this paper including the operational model of a RISC-like fragment of x86-64, 
a simple language run on this operational model, and a logic as an instantiation of the Iris framework are mechanized 
inside Coq.
\end{abstract}

%
%
%
%
%
%
%
%
%
%
%
%
%
%
%
%
%
%
%
%
%
%
%
%
%
%
%
%

%
%
%
%

%
%
%
%

\maketitle

{
\theoremstyle{acmdefinition}
\newtheorem{assumption}[theorem]{Assumption}
}

%
\section{Introduction}
\label{sec:intro}
Virtual memory management lies at the core of modern OS kernel implementation. It is deeply intertwined with most other parts of a typical general-purpose OS kernel design, including scheduling, hardware drivers, and even the filesystem buffer cache. In writing the authoritative reference on the internals of the Solaris kernel, McDougall and Mauro went so far as to claim that ``\emph{the virtual memory sub-system can be considered the core of a Solaris instance, and the implementation of Solaris virtual memory affects just about every other subsystem in the operating system}''~\cite{mcdougall2006solaris}.
This makes rich support for verification the virtual memory management subsystem of an OS kernel critical to the correctness of every other piece of an OS or any software running atop it.

At its core, the virtual memory functionality of modern CPUs is about \emph{location virtualization}: the memory locations
(addresses) seen by most code are not, in fact, the exact location in physical memory where data reside. Instead these 
are \emph{virtual} addresses, which are mapped to actual physical resources by the cooperation of the hardware and OS. 
This is what enables separation of process memory resources:
the OS manipulates hardware functionality to ensure that any attempt by a process to access memory not explicitly granted 
to it by the kernel will fail. But this is complicated by the fact that 
%
%
control over these mappings of virtual to physical addresses is itself mediated by \emph{in-memory data structures}, 
which the kernel still accesses via virtual address, leading to indirect cycles.

Further complicating matters is that addresses themselves bear no information about which address space they originate 
from. For user processes this is of little concern, as these have access to only their own address space. But the kernel has
(or can grant itself) access to all address spaces. Mixing up addresses from different address spaces leads to severe bugs.
More concerning, keeping track of which \emph{assertions} hold in different address spaces during kernel verification is 
difficult: some assertions should hold across all address spaces, while others hold in only one, and others may hold in 
multiple but still not all.

This kind of context-dependent assertion, where a fact may be true in one address space but not others, has a modal flavor. 
We propose tackling the verification of virtual memory subsystems (and kernels more broadly) by adapting ideas from hybrid
modal logic, which can label assertions true under \emph{other, named} circumstances (i.e., in another address space) with a 
modality indexed by a name for that space (in our case, the root of the page tables for an address space). This offers a 
\textit{convenient} and \textit{powerful} way to \emph{modularly}
\begin{itemize}
\item isolate assertions specific to a particular address space,
\item explicitly state when an assertion is true across address spaces,
\item manipulate address spaces from within other address spaces, and
\item reason about change in address spaces.
\end{itemize}
These advantages make this approach to reasoning about virtual memory more flexible than prior program logic techniques~\cite{kolanski08vstte,kolanski09tphols}, 
which were only able to work with a single address space (the current address space on the CPU) because they were unable
to speak directly \emph{within the logic} about other address spaces.
%
%
%
%
%
%

\begin{itemize}
\item We develop these ideas in the form of a logic for working with virtual-address-space-relative assertions,
      implemented as an embedded separation logic within Iris~\cite{jung2018iris} separation logic.
\item We prove the soundness of our \textsf{vProp} logic with respect to a RISC-like fragment of \textsf{AMD64} instructions.
\item We verify simplified versions of several critical virtual-memory-related pieces of OS functionality, 
      including mapping  pages, and the assembly code for switching address spaces. 
      The latter in particular goes beyond the capabilities of prior work, and reveals unexpected
	subtleties.
\end{itemize}
 %
%
\section{Background}
\label{sec:background}
\subsection{Machine Model}
\label{sec:backgroundonmachinemodel}

In typical system configurations, all memory addresses seen by programs running on modern computers are \emph{virtualized}: the address observed by a running program generally will not correspond directly to the physical location in memory, and may not even correspond to a physical location that \emph{exists} in the machine. Instead, these \emph{virtual} addresses are translated to \emph{physical} addresses that correspond directly to locations in RAM. On most modern architectures, this translation is performed through cooperation of the hardware and OS kernel: while executing an instruction that dereferences a (virtual) address, the CPU's \emph{memory management unit} (\textsc{MMU}) implements an architecture-specific process of \emph{address translation}, resulting in a physical address used to access the cache\footnote{Technically, for performance reasons most caches are indexed with parts of the virtual address, but tagged with the physical data addresses, so cache lookups and address translations can proceed in parallel.} and/or memory-bus.

\begin{figure}
    \includegraphics[width=\columnwidth]{pagetables.png}
    \caption{x86-64 page table lookups.}
    \label{fig:pagetables}
\end{figure}

On the x86-64 architecture, the input to the \textsc{MMU}'s address translation is a sparse hierarchical set of tables 
called \emph{page tables} (referring to pages of memory). As Figure \ref{fig:pagetables} (based on Figure 5-17 of the 
AMD64 architecture manual~\cite{amd64_manual_vol2}\footnote{While x86 up through its 32-bit incarnation were due to Intel,
the x86-64 architecture as a 64-bit extension to x86 was originally due to AMD. As a result, it is sometimes also referred to as the \texttt{amd64} architecture.}) shows, 
address translation proceeds by repeatedly taking designated slices of the virtual address and indexing into a table.
The final lookup in the page tables gives the base physical address of a page of physical memory, to which the low-order
bits of the accessed virtual address are added to determine the actual physical address retrieved. 
On x86-64, standard configurations use 4 levels of page tables, labelled levels 4 through 1, with lookups in the level 
1 page table resulting in the actual page of physical memory holding the requested data, and the low-order 12 bits 
being used to index into this page.\footnote{Technically levels 1--3 have designated names originating when there were 
fewer levels. We omit those names to avoid confusion for those not already steeped in the arcana of x86 history,
 and for greater consistency with the recent addition of a 5th level 
on the most recent processors, where levels 4 and 5 are simply known by their numbers. Our formalization
only deals with 4-level page tables, but is straightforwardly extensible to 5.}  
Because address translation 
traverses a series of tables, the translation process or algorithm is sometimes referred to as a \emph{page-table walk}. 
While Figure \ref{fig:pagetables} and most of our details are specific to the x86-64 architecture, ARMv8 (a.k.a.\ 
\texttt{aarch64}), RISC-V, and PowerPC use similar hierarchical page tables for address translation.

The entries of each table are 64 bits wide, but each points to a physical address aligned to 4KB (4096 byte) boundaries, which leaves 12 bits to spare to control a validity bit (called the \emph{present} bit), a read-write bit (which permits write access through the entry if and only if it is set), and a range of additional bits which can be used to control caching, write-through, and more. This paper will only consider the present bit (0), the read-write bit (1), and the accessed bit (5), which is set to 1 by the hardware every time an entry is used, allowing the OS to determine which pieces of memory are used or unused.

The page tables mapping virtual addresses to physical addresses are managed by the OS. Typically each process has its own 
virtual-to-physical mappings represented by its own page table, which the OS registers with the CPU by storing the (page-aligned)
physical address of the root of the page table tree (in our context, the start of the L4 table) in a 
specific register (\texttt{cr3}) as part of switching to a new process. 
Using different mappings, which map only disjoint portions of physical memory (with some exceptions in the next section) 
is how the OS ensures memory isolation between processes.

If an instruction is executed that accesses a virtual address that either has no mapping, or does not have a mapping permitting the kind of access that was performed (e.g., the instruction was a memory write, but the relevant address range was marked read-only in the relevant page table entry), the hardware triggers a \emph{page fault}, transferring control to a \emph{page fault handler} registered with the hardware by the OS, allowing it to take corrective action (described in the next section).

\subsection{Virtual Memory Managers}
\label{sec:backgroundonvmm}
Most OS kernels have a component called a \emph{Virtual Memory Manager} (VMM)\footnote{Not to be confused with Virtual 
Machine Monitor. In this paper we focus on non-hypervisor scenarios, but hardware virtualization extensions for both 
x86-64 and ARM make use of an additional set of page tables translating what a \emph{guest} considers to be its 
(virtualized) physical memory to actual physical memory. Our contributions should offer value in this scenario as well.}
which is responsible for setting up page table mappings and for taking action when a page fault occurs. Often, a page 
fault indicates a bug in a program --- for example, a null pointer dereference crashes a program not because the hardware
 designates it an error, but because most OSes refuse to map the first 4K worth of virtual addresses, so dereferencing 
\texttt{NULL} results in a page fault. In such cases, the OS will terminate the program.

In other cases the OS uses page faults to implement specialized functionality and optimizations. A key example of this is 
(confusingly) called \emph{paging}: saving room in physical RAM and avoiding unnecessary IO 
by waiting until memory addresses for code are accessed 
before copying them to RAM (in particular, saving process startup time); 
or copying memory that has not recently been accessed to disk 
(i.e., in a \emph{swap} file or partition) and marking those page table entries invalid so a future page fault will give 
the OS the chance to copy the relevant data back into memory when the program tries to access it again.

The key pieces of functionality a VMM must implement itself, or have available in another OS component (depending on 
design), are
adding a new page mapping (whether the mapped page contains zeros, file data, or swap data), and removing an existing 
page mapping.
While this initially sounds like relatively modest functionality whose implementation may be complicated by hardware 
subtleties, even these basic operations are actually quite intrictate.
Most notably, updates to the page tables are performed as writes to memory --- which are themselves subject to address translation.
In the case of changing the mappings for the currently-active set of page tables, the OS kernel is modifying the tables involved in its
own access of the tables.

%
%
%
%
%
%
%
%
%
%
%
%
%
%
%
%
%
%
%
%
%
%

%
%
%
%
%
%
%
%
%

Virtual memory concerns propagate to the OS scheduler, which is responsible for dealing with multiple 
address spaces, so must keep track of which virtual addresses are valid (and in what way) in which address spaces. 
Some virtual addresses are valid in only a single address space (e.g., a code address for a particular usermode 
process), while others are valid in all address spaces (e.g., kernel data structure pointers). 
The VMM must maintain some of these assumptions on behalf of the rest of the kernel, for example by guaranteeing that 
a certain range of virtual addresses (corresponding to the kernel's code and data) are valid in every address space.

\subsection{Modal Logic}
\label{sec:backgroundonmodallogic}
The problem of needing to keep track of things being true in some contexts and not in others is hardly unique to virtual 
memory management, and is the general insight behind most flavors of modal logic.

Most modal logics use unary operators on propositions to express that a logical claim $P$ is \emph{contingently} true 
in certain other circumstances, such as in other times~\cite{pnueli1977temporal} or places~\cite{gordon2019modal}. That is, for 
different modal operators $M$, $M(P)$ may mean that $P$ is true in another time or place (which may be a physical place 
like another network node~\cite{murphy2008type,gordon2019modal} or a conceptual place like a person's 
mind~\cite{hintikka1962knowledge}). A unifying concept across any modality is that they behave as applicative functors, 
typically satisfying (directly, or as a derived law, depending on the modality):
\[ (P\rightarrow Q) \rightarrow M(P) \rightarrow M(Q)\]
%
%
%
%
%
Many modalities, so-called \emph{normal} modalities also possess introduction rules of the form $P\rightarrow M(P)$, 
the classic example being that if $P$ is true, then $P$ is \emph{necessarily} true with the contingency picked up.%

Of particular interest for reasoning about virtual memory are modalities that permit \emph{naming} the alternate 
circumstances, prominently \emph{hybrid} modal logics~\cite{blackburn1995hybrid,areces2001hybrid}, which come equipped 
with assertions of the form $[\ell](P)$ indicating that $P$ is true in the specific alternate circumstance (Kripke world)
 named by the \emph{nominal} $\ell$. Note that a distinctive property of hybrid logics is that, rather than hiding
the points at which a modal assertion is evaluated inside the modality's definition, the choice of what world a modalized
assertion should be true in is chosen \emph{in the assertion itself}. This allows assertions to talk about not simply whether some other assertion
is true in some possible future or past world related in a fixed way to the current world, but to talk about \emph{arbitrary}
other worlds. This explicit naming of alternate worlds increases the power of modal logics~\cite{blackburn1995hybrid}, and is actually
necessary for completeness in classical separation logics~\cite{brotherston2014parametric}.

For our purposes, these are natural candidates to adapt for virtual memory management. We can reinterpret the notion of 
naming an alternate world slightly more loosely, and instead name \emph{address spaces} by the physical address of the 
page table root, since these structures are the physical representations of page tables. Thus in this paper we develop 
the notion that we can represent contingent truth of an assertion via $[r](P)$ indicating that $P$ holds in the address 
space rooted at physical address $r$.
The typical hybrid modality introduction rule, that $P$ and $\ell$ (indicating the current world is $\ell$) imply 
$[\ell](P)$, has a natural analogue: knowing $P$ and that the current address space is $r$ (i.e., that $r$ is the 
current \texttt{cr3} value) suggests a way to construct $[r](P)$. By the choice of \texttt{cr3} identifying the contingency, we can identify a fact
\begin{itemize}
  \item as a \textit{context} fact if its validity depends on \texttt{cr3} when it is outside the modality (i.e. modal context), and made \texttt{cr3} independent as a part of custom-tailored modal logic for virtual-memory: a truth representing a virtual-memory addressing depends \texttt{cr3} due to address-translation operation, is dependent on \texttt{cr3} in the ambient-logic (e.g. separation-logic), and is made \textit{independent} of the facts related to \texttt{cr3} by being introduced into the modal context under the assumption that it exhibits the knowledge on its validity with respect to \texttt{cr3} in the ambient logic.
  \item as a \textit{pure} fact as long as it does not \textit{necessarily} depend on any fact related to \texttt{cr3}: a truth representing a raw physical memory addressing, unlike virtual-address translation, does not need the \textit{knowledge} of \texttt{cr3}, therefore it can be introduce into the modal context as a pure fact
\end{itemize}

A relatively under-explored space of modal logics is the interaction of modal and substructural logics, 
in particular hybrid-style modalities in substructural logics, which has seen only minimal exploration~\cite{dovsen1992modal,restall1993modalities,d1997grafting,kamide2002kripke,licata2017fibrational} and no prior 
application. We develop our ideas in the Iris program logic~\cite{jung2018iris}, as a result exploring additional subtleties 
that arise where the modality itself may entail ownership of resources, as well as interactions between our hybrid-style 
modality and substructural rules.  For example, Iris contains a number of modalities that distribute over separating 
conjunction, or for which resources can freely move into the modality 
(e.g., $\blacktriangleright(P)\ast Q \wand \blacktriangleright(P\ast Q)$). In our setting some of these rules 
apply while others do not. For example, in our setting, an assertion that involves no modalities is interpreted as 
holding in the current (active-on-the-CPU) address space, so clearly cannot move into arbitrary other address spaces 
--- unless guarded by another address space modality.
%
%
%
%
%
%
%
%
%
%
%
 %
\definecolor{dkgreen}{rgb}{0,0.6,0}
\definecolor{ltblue}{rgb}{0,0.4,0.4}
\definecolor{dkviolet}{rgb}{0.3,0,0.5}

%
\lstdefinelanguage{Coq}{ 
    %
    mathescape=true,
    %
    texcl=false, 
    %
    morekeywords=[1]{Section, Module, End, Require, Import, Export,
        Variable, Variables, Parameter, Parameters, Axiom, Hypothesis,
        Hypotheses, Notation, Local, Tactic, Reserved, Scope, Open, Close,
        Bind, Delimit, Definition, Let, Ltac, Fixpoint, CoFixpoint, Add,
        Morphism, Relation, Implicit, Arguments, Unset, Contextual,
        Strict, Prenex, Implicits, Inductive, CoInductive, Record,
        Structure, Canonical, Coercion, Context, Class, Global, Instance,
        Program, Infix, Theorem, Lemma, Corollary, Proposition, Fact,
        Remark, Example, Proof, Goal, Save, Qed, Defined, Hint, Resolve,
        Rewrite, View, Search, Show, Print, Printing, All, Eval, Check,
        Projections, inside, outside, Def},
    %
    morekeywords=[2]{forall, exists, exists2, fun, fix, cofix, struct,
        match, with, end, as, in, return, let, if, is, then, else, for, of,
        nosimpl, when},
    %
    morekeywords=[3]{Type, Prop, Set, true, false, option},
    %
    morekeywords=[4]{pose, set, move, case, elim, apply, clear, hnf,
        intro, intros, generalize, rename, pattern, after, destruct,
        induction, using, refine, inversion, injection, rewrite, congr,
        unlock, compute, ring, field, fourier, replace, fold, unfold,
        change, cutrewrite, simpl, have, suff, wlog, suffices, without,
        loss, nat_norm, assert, cut, trivial, revert, bool_congr, nat_congr,
        symmetry, transitivity, auto, split, left, right, autorewrite},
    %
    morekeywords=[5]{by, done, exact, reflexivity, tauto, romega, omega,
        assumption, solve, contradiction, discriminate},
    %
    morekeywords=[6]{do, last, first, try, idtac, repeat},
    %
    morecomment=[s]{(*}{*)},
    %
    showstringspaces=false,
    %
    morestring=[b]",
    morestring=[d],
    %
    tabsize=3,
    %
    extendedchars=false,
    %
    sensitive=true,
    %
    breaklines=false,
    %
    basicstyle=\small,
    %
    captionpos=b,
    %
    columns=[l]flexible,
    %
    identifierstyle={\ttfamily\color{black}},
    %
    keywordstyle=[1]{\ttfamily\color{dkviolet}},
    %
    keywordstyle=[2]{\ttfamily\color{dkgreen}},
    %
    keywordstyle=[3]{\ttfamily\color{ltblue}},
    %
    keywordstyle=[4]{\ttfamily\color{dkblue}},
    %
    keywordstyle=[5]{\ttfamily\color{dkred}},
    %
    %
    %
    stringstyle=\ttfamily,
    %
    commentstyle={\ttfamily\color{dkgreen}},
    %
    literate=
    {\\forall}{{\color{dkgreen}{$\forall\;$}}}1
    {\\exists}{{$\exists\;$}}1
    {<-}{{$\leftarrow\;$}}1
    {=>}{{$\Rightarrow\;$}}1
    {==}{{\code{==}\;}}1
    {==>}{{\code{==>}\;}}1
    %
    {->}{{$\rightarrow\;$}}1
    {<->}{{$\leftrightarrow\;$}}1
    {<==}{{$\leq\;$}}1
    {\#}{{$^\star$}}1 
    {\\o}{{$\circ\;$}}1 
    {\@}{{$\cdot$}}1 
    {\/\\}{{$\wedge\;$}}1
    {\\\/}{{$\vee\;$}}1
    {++}{{\code{++}}}1
    {~}{{\ }}1
    {\@\@}{{$@$}}1
    {\\mapsto}{{$\mapsto\;$}}1
    {\\hline}{{\rule{\linewidth}{0.5pt}}}1
    %
}[keywords,comments,strings]

\section{Machine State \& Syntax}
\label{sec:syntax}
To develop our core logical ideas, we instantiate \textsf{Iris} with a simple language for streams of instructions, 
and a logical machine model corresponding to execution of x86-64 assembly instructions with virtual memory enabled on the 
CPU.

\subsection{Registers and Memory}
Programs we demonstrate in this paper requires accessing two types of computer resource: registers and memory.
A register identifier, $\reg$, is chosen from a fixed finite set of register identifiers, $\regset$. 
We use these identifiers $\reg$ for register names such as \texttt{rax}, \texttt{r8}, or \texttt{cr3}.
%
%
%
%
%
%
%
%
%
%
%
%
%
%
%
%
%
%
%
%
%
%
%
%
%
%
%
%
%
%
%
%
%
%
%
%
%
%
%
%
%
%
For clarity and ease of representation, we use machine words, $\loc \in \Loc$, with the subscripts showing the number of bits in a word,
for memory addresses, values, and offsets, rather than distinct location types that wrap machine words.
For example, $\kw{w}_{12}$ is a 12-bit word, which can be obtained for example truncating away 52 bits of a 64-bit word ($\kw{w}_{64}$).
\subsection{State}
\label{sec:state}
We represent the machine state mainly as a finite map of registers to register values and a map of word-aligned physical memory addresses 
to 64-bit physical memory values. 
Thus our states $\sigma$ include:
%
\begin{itemize}
%
\item A register map: $\sigma.\mathcal{R}: \kw{greg} \rightarrow_{\textrm{fin}} \kw{regval} $
\item A memory map: $\sigma.\mathcal{M}: \Locft \rightharpoonup_{\textrm{fin}} (\Loctw \rightharpoonup_{\textrm{fin}} \Locsf )$
\end{itemize}
Of particular note, \texttt{cr3}, the page table register, is included in the machine state.
%
\subsection{Instructions}
\label{sec:instructions}

\begin{figure}[t]
\begin{lstlisting}[language=Coq,mathescape]
 Definition translate (rtv: word 64) ($\sigma$: state) (w: word 64): word 64 + MemFail :=
      l4e $\leftarrow$ translate_top_level $\sigma.\mathcal{M}$ rtv w; 
      l3e $\leftarrow$ translate_from_l4e $\sigma.\mathcal{M}$ l4e w; 
      l2e $\leftarrow$ translate_from_l3e $\sigma.\mathcal{M}$ l3e w;
      l1e $\leftarrow$ translate_from_l2e $\sigma.\mathcal{M}$ l2e w;
      addr $\leftarrow$ translate_from_l1e $\sigma.\mathcal{M}$ l1e w;
      return concat (shift_dropn addr 3 three_lt64) low3.
\end{lstlisting}
\vspace{-1em}
\caption{\coq implementation of address translation (slightly simplified).}
\label{fig:coq_addr_translation}
\end{figure}

Programs in our logic are instruction sequences \instrs, which are formed by either a basic instruction \iskip, or prefixing an existing instruction
sequence with an additional instruction (\iseq\instr\instrs).
We model (and later, give program logic rules for) instructions for basic register moves, and reading and writing memory.
The latter require page table walks.
Figure \ref{fig:coq_addr_translation} gives a slightly simplified version of our address translation code in monadic style:
starting from the root page table address \textsf{rtv}, the virtual address \textsf{w} is either resolved to a physical address or an error is returned.
The full set of possible errors is not relevant to this paper; our logic is tailored for kernel code which should not fault,
so our proof rules guarantee no memory failures occur --- that the page table walk for any dereferenced address will succeed.

%
%
%
%
%

The main instructions we implement are variants of the x86-64 \lstinline|mov| instruction, which we format in Intel syntax (destination on the left, source on the right).
As expected \textsf{store} would be the update on the memory map entry
%
For memory operations, we treat
\[
\begin{array}{l}
  \hbox{(\TirNameStyle{WriteToMemFromReg})} \qquad
  \textsf{mov}~[r_m]~r_r \\
  \hbox{(\TirNameStyle{WriteToRegFromMem})} \qquad 
  \textsf{mov}~r_r~[r_m]
  \end{array}
  \]
which are realized by first translating the virtual-memory address of a memory location stored in $r_m$ to a physical memory location,
then updating those memory contents with the contents of register $r_r$ (\TirNameStyle{WriteToMemFromReg}) 
or loading the value at that physical memory location into the register $r_r$
(\TirNameStyle{WriteToRegFromMem}). 
%
Additional variants (such as instructions which access memory at a constant offset from the virtual address stored in a register) are also treated.

For register operations, we support moves back and forth between any pair of registers, though the \coq formalization splits these according to whether
the source or destination registers are standard data registers (e.g., \lstinline|rax|) or control registers (e.g., \lstinline|cr3|).
%
%
%
%
%
%
%
%
%
%
%
%
%
%
%
%
%
%
%
%
%
 %
\section{Program Logic for Location Virtualization}
\label{sec:logic}
%
\newcommand{\gammaPred}{\delta}
\newcommand{\gammaPreds}{\delta\textsf{s}}
\newcommand{\rtv}{\textsf{rtv}}
\newcommand{\qone}{\texttt{q1}}
\newcommand{\qtwo}{\texttt{q2}}
\newcommand{\qthree}{\texttt{q3}}
\newcommand{\qfour}{\texttt{q4}}

\newcommand{\sumwalkabs}[3]{
  \ownGhost\gammaPred{\authfrag{\singletonMap{#1}{(#2, #3)}}}
}

\newcommand{\sumapaces}[2]{
  \ownGhost\gammaPreds{\authfrag{\singletonMap{#1}{#2}}}
}
\newcommand{\ptableabswalk}[1]{\mathcal{A}\textsf{bsPTableWalk}(#1)}
\newcommand{\ptablestore}{\theta}

We derive a program logic (a separation logic) supporting the following stances and constraints:
\begin{enumerate}
\item \textit{address-spaces as modal contexts}: Assertions in our logic are context-dependent,
  in the sense that their truth depends on which address space they are used in, due to the need to support virtual points-to assertions.
\item \textit{sharing}: The physical location backing a virtual address's storage is located (during a page table walk) through a 
      set of physical page-table (L4-L1 page-tables) acceses that are shared amongst different virtual addresses (specifically,
      those on the same page of memory\footnote{of in the case of L2 or higher levels of tables, within a given broader region.}).
      This sharing imposes constraints on definiting points-to assertions
      in terms of physical (L4-L1) page-table memory accesses
\item \textit{context-agnostic-resources}: each virtual address is valid under a certain address-space, 
      but it does not represent this \textit{knowledge} of its address-space. 
      That is, assertions are not explicitly tagged with their address space validity
\item \textit{updating address-space mappings}: We present logical abstractions to enable 
      updating not only pages of typical data in memory, but also page tables themselves.\footnote{Prior work relied on unfolding operational semantics
      to verify page table updates.}
\item \textit{explicitly-modal assertions}: Our logic includes a means to talk about facts being true
      in another address space
\item \textit{address-space switch as changing the "World" of truth}: Switching from one address-space to another logically
      becomes a simultaneous introduction-and-elimination of a pair of modal assertions (for different address spaces)
\end{enumerate}

The first viewpoint, that the truth of an assertion is relative to an address space, has far-reaching consequences.
Therefore, as a first step, we work entirely in a pointwise lifting of Iris's base BI logic,
essentially working with separation logic assertions indexed by a page table root as a $\mathcal{W}_{64}$
(Figure \ref{fig:vprop}), which we call $\textsf{vProp }\Sigma$; this is the (\textsc{Coq}) type of assertions in our logic.
Most constructs in Iris's base logic are defined with respect to any BI-algebra (of \textsc{Coq} type \lstinline|bi|), so automatically
carry over to our derived logic.
However, be must still build up from existing Iris primitives to provide new primitives that depend on the address space --- primarily the notion
of virtual points-to.

\begin{figure}[t]
\begin{lstlisting}[language=Coq]
  Definition vProp  $\Sigma$ : bi := word 64 -b> iPropI  $\Sigma$.
  \end{lstlisting}
\caption{$\textsf{vProp }\Sigma$: Root-Address Indexed Address-Space Proposition}
  \label{fig:vprop}
\end{figure}

\subsection{Base Points-To Assertions}
\label{sec:pointsto}
We build up our context-dependent assertions from two basic types of assertions that do not care about address spaces.

\paragraph{Register points-to} 
The assertion $\textsf{r}\;\mapsto_{\textsf{r}}\{q\}\;\textsf{rv}$ ensures the ownership of the register $\rg$ naming the register value $\rv$. 
The fraction $\qfrac$ with value 1 asserts the unique ownership of the register mapping, and grants update permission on it, 
otherwise, any value $0 < \qfrac <1$ represents partial ownership granting readonly permission on the mapping.

\paragraph{Physical memory  points-to} The soundness proofs for our logic's rules largely center around page-table-walk accesses as in Figure \ref{fig:coq_addr_translation},
 where (inside the helper functions) we need to split a 64-bit word representing entries in these tables into 52-bit segments
which specify the physical page number of another page in the walk, and 12-bit segments containing various control bits.
For this reason we have two notions of physical points-to facts. The primitive notion closest to our machine model is:
\[ \textsf{w1} \ \sim \ \textsf{w2} \mapsto_{\textsf{a}} \; \{\textsf{q}\} \; \textsf{v} \]
In this form, we see that \textsf{w1} (a $\mathcal{W}_{52}$) essentially selects a 4KB page of physical memory,
and \textsf{w2} (a $\mathcal{W}_{12}$) is an offset within that page.
%
%
%
From this we can derive a more concise physical points-to when the split is unimportant:
%
\[\textsf{w} \mapsto_{\textsf{p}} \{q\} \textsf{ v} \stackrel{\triangle}{=} (\textsf{drop 12}~w) \ \sim \ (\textsf{bottom 12}~w)\mapsto_{\textsf{a}} \textsf{ v} \]

\begin{figure*}
  \begin{lstlisting}[language=Coq]
 Definition $\vaddr\mapsto_{\textsf{t}}\{\textsf{q}\}\; \vpage$ : vProp $\Sigma$ := 
  $\exists_{\textsf{l4e l3e l2e l1e}} \ldotp$  $\ulcorner$ aligned $\vaddr \urcorner \ast$ L4_L1_PointsTo($\vaddr$ l4e l3e l2e l1e paddr) $\ast$ paddr $\mapsto_{p}\{\mathsf{q}\} \vpage$.
 Definition L4_L1_PointsTo (maddr l4e l3e l2e l1e paddr :word 64) : vProp $\Sigma$ :=
 $\lambda$ cr3val.
  $\ulcorner$ entry_present l4e $\land$ entry_present l3e $\land$ entry_present l2e $\land$ entry_present l1e$\urcorner$ $\ast$
  (l4M52 maddr cr3val) $\sim$ (l4off maddr cr3val) $\mapsto_{a}$ {q1}  l4e  $\ast$
  (l3M52 maddr l4e) $\sim$ (l3off maddr l4e)  $\mapsto_{a}$ {q2}  l3e $\ast$ 
  (l2M52 maddr l3e) $\sim$ (l2off maddr l3e) $\mapsto_{a}$ {q3}  l2e $\ast$
  (l1M52 maddr l2e) $\sim$ (l1off maddr l2e) $\mapsto_{a}$ {q4}  l1e.
\end{lstlisting}
\vspace{-1em}
\caption{A Strong Virtual Points-to Relation}
  \label{fig:strongvirtualpointsto}
\end{figure*}

\subsection{An Overly-Restrictive Definition for Virtual Memory Addressing}
\label{sec:overly-restrictive}
A natural definition for a virtual points-to that depends on the current address space is to
require that in order for a virtual address \textsf{va} to point to a value \textsf{v}, the assertion contains
partial ownership of the physical memory involved in the page table walk that would translate \textsf{va} to
its backing physical location --- with locations existentially quantified since a virtual points-to should not assert
\emph{which} locations are accessed in a page table walk.
Figure \ref{figure:strongvirtualpointsto} gives such a definition.
It asserts the existence of for page table entries, one at each translation level, and via \lstinline|L4_L1_PointsTo|
asserts from the current \lstinline|cr3| value, taking the offsets from \textsf{va} (as in Figure \ref{fig:pagetables})
to index each level, finding the existentially quantified entries in each level, 
ending with the L1 entry, whose physical page base is added to the page offset of \textsf{va} to obtain the physical address
 --- and there (\lstinline|addr_L1(va,l1e)|) in physical memory, is the value.
Most of the definition lives directly in \textsf{vProp}, using the separation logic structure lifted from Iris's \textsf{iProp}.

\lstinline|L4_L1_PointsTo| works by
chaining together the entries for each level, using the sequence of table offsets from the address being translated to index
each table level, and using the physical page address embedded in each entry.
For example, 
  the first level address translation to get the L4 entry (\lstinline|l4e|) 
  uses the masks \textsf{l4M52} with the current \lstinline|cr3| to get the 
  physical address of the start of the L4 table,
  and \textsf{l4off} with the virtual address being translated to compute the correct offset within that table.
  Then at that physical location is the appropriate entry in the L4 table \textsf{l4e}.
  Subsequent levels of the page table walk work similarly.
The statement of these assertions is simplified by the use of our split physical points-to assertions, since
each level of tables is page-sized.\footnote{We do not address superpages and hugepages in this paper.}
This helper definition is also a more explicit \textsf{vProp}, explicitly binding a value for \lstinline|cr3| and using it
to start the translation process.

%
%
%
%
%
%
%
%
%
%
%
%
%
%
%
%
%
%
%
%
%
%
%
%

This solution is in fact very close to that of \citet{kolanski08vstte}, who define a separation logic from scratch in \textsc{Isabelle/HOL},
where the semantics of all assertions are functions from pairs of heaps and page table root values to booleans.\footnote{
  This was a typical explicit construction at the time; their work significantly predates Iris.
}
Our solution removes some restrictions in this definition by further abstracting the handling of address translation.

\subsection{Aliasing/Sharing Physical Pages}
  \label{sec:sharingpages}  
  One might have already observed that the virtual points-to definition shown in Figure \ref{fig:strongvirtualpointsto} 
  is too strong to specify some operations a virtual memory manager may need to do, such as
  move one level of the page table to a different physical location while preserving all virtual-to-physical mappings.\footnote{
    x86-64 hardware, like other architectures, includes a feature (which we do not formalize assertions for) to
    replace an L1 page table address in an L2 entry with a pointer to a \emph{larger} 2MB page (called super-pages), 
    or replace an L2 page table address in an L3 entry with a pointer to a 1GB page (called huge-pages).
  }
   To speak more concretely, the fractional permissions on the page-table-walk assertion ($\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$ in Figure \ref{fig:strongvirtualpointsto}) 
  ensure fractional ownership of the four physical page table entries accessed by the hardware during address translation,
  as well as the physical memory holding the data of the virtual points-to.
  However, this formulation stores knowledge of the page table walk's details with ownership of the backing physical
  memory.
  Updating any of these mappings (e.g., moving the page tables in physical memory, as in coallescing for
  superpages or hugepages)
  would require explicitly collecting all virtual points-to facts that traverse affected entries.
  It is preferable to permit the page tables themselves to be updated independently of the virtual points-to assertions,
  so long as those updates preserve the same virtual-to-physical translations.
\newcommand{\vale}{\textsf{val}}
\begin{figure*}
\[
\begin{array}{l}
    \vaddr\mapsto_{\textsf{v}}\{\textsf{q}\}\;\vale : \mathsf{vProp}~\Sigma \stackrel{\triangle}{=} \\
    \quad
    \exists \paddr\ldotp
    \exists \delta\ldotp
    \underbrace{\lambda\mathit{cr3val}\ldotp\sumapaces{\mathit{cr3val}}\delta}_\text{\begin{tabular}{c}Find ghost map for this\\address space\end{tabular}} \ast 
  \underbrace{\sumwalkabs\vaddr\qfrac\paddr }_\text{\begin{tabular}{c}Fractional ownership of \\virt-to-phys translation\end{tabular}}\ast 
  \underbrace{\paddr \mapsto_{\mathsf{p}} \vale}_\text{\begin{tabular}{c}Value at\\physical location\end{tabular}}
\end{array}
\]
\caption{Virtual-Pointsto for Sharing Pages}
  \label{fig:virtualpointstosharing}
\end{figure*}

  Intuitively, the definition in Figure \ref{fig:strongvirtualpointsto} is too strong because the virtual points-to
  assertion there tracks too much information: when writing programs that access memory via virtual addresses,
  most code does not care \emph{which physical memory locations are involved in address translation}: it only cares
  that virtual address translation would succeed. The necessary information about the physical page table walk
  must still be tracked, but can be tracked separately from the virtual points-to assertion itself.

  We take the approach of separating the physical page-table-walk pointsto ($\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$ in Figure \ref{fig:strongvirtualpointsto}) 
  from the virtual pointsto relation. 
  To do so, we take advantage of an Iris construction which permits tracking a map in ghost state that maps
  virtual addresses to the physical addresses they translate to, and has for each key in the map (i.e.,
  each virtual address) a token. Updating the virtual-to-physical (ghost) mapping for a virtual address $va$
  requires full ownership of the token. In addition, we keep an invariant that for each $va\mapsto pa$ mapping
  in this map, there are \emph{physical} resources sufficient to ensure that address translation for $va$
  will resolve in the hardware to $pa$. But these physical resources, which specify \emph{which} physical locations
  are involved in the page table walk, are now separated from but consistent with
  the knowledge that such resources exist (which is embodied by the token for $va$, which tracks that $va$ maps to $pa$
  in the ghost map). Thus we can store the token which summarizes the translation and ensures it exists in the virtual
  points-to, and keep the ghost map and the invariant that every mapping in the ghost map has corresponding physical resources
  for translation in a separate global invariant.

  %
  %
  In Iris this is realized by using an authoritative resource algebra: there is a single \emph{authoritative} global copy of the (ghost)
  map caching virtual-to-physical address translations, and for each entry a read only \emph{partial} ownership of that key-value pair.
  The resource algebra itself is instantiated as:
  \[\mathcal{A}\textsf{bsPTableWalk} \stackrel{\triangle}{=} \textsc{Auth} (\; \mathcal{W}_{64} \;\rightarrow_{\textrm{fin}} \;  ( (\textsc{Frac }, \mathord{+}) \times (\textsc{Agree } \Loc,\mathord{=}) ))\]
We replace the physical
 $\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$
with
fragmental ownership of the ghost-map ($ \sumwalkabs\vaddr\qfrac\paddr$),
 whose ownership just ensures that the virtual address ($\vaddr$) is mapped to a page (referenced by $\paddr$)

We place the authorative ownership $\mathcal{A}\textsf{PTableWalk}$ of the ghost-map in a per-address-space invariant
($\mathcal{I}$\textsf{ASpace} in Figure \ref{fig:peraspaceinvariant}) whose ownership, when combined with the
fraction stored in the virtual points-to, permits changes in the virtual-to-physical translations.
By itself, it allows changes to the page tables as long as they do not change the overall mappings from virtual
to physical addresses.

  \begin{figure*}
\[
\begin{array}{l}
  \mathcal{I}\textsf{ASpace}(\ptablestore,m)\stackrel{\triangle}{=} \textsf{ASpace\_Lookup}(\ptablestore,m) \ast \\
 \qquad\qquad\qquad \bigast{(\vaddr, \textsf{paddr})\in \ptablestore}{\exists\;(\textsf{l4e l3e l2e, l1e, paddr})\ldotp \textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\vaddr\textsf{, l4e, l3e, l2e, l1e, paddr})} \\
  \textsf{ where } \\
   \textsf{ASpace\_Lookup} (\ptablestore,m) \stackrel{\triangle}{=} \lambda\textsf{ cr3val} \ldotp \; \exists \gammaPred \; \ldotp \ulcorner m \; !!\; \textsf{cr3val} = \textsf{Some } \gammaPred \urcorner \ast
    \ownGhost\gammaPred{\authfull{\ptableabswalk\ptablestore}}

\end{array}
\]
\caption{Global Address-Space Invariant with a fixed global map of address-space names $m$}
  \label{fig:peraspaceinvariant}
  \end{figure*}

Another major change is that, because our logic is the first to explicitly address the existence of multiple address spaces,
the ghost map of virtual-to-physical translations (identified by the ghost name $\delta$ in Figure \ref{fig:virtualpointstosharing}
 \emph{for each address space}. This means we also require a way to locate the map for the current address space
from the current root page table pointer stored in \lstinline|cr3|.

Figure \ref{fig:peraspaceinvariant} describes the resources for each address space, which are collected
for the full address space, but separated from virtual points-to assertions.
$\Theta$ is the logical map from virtual addresses to the physical addresses they should translate to, 
for the \emph{current} address space, corresponding to the currently installed page tables (the ones
indicated by \lstinline|cr3|). In Figure \ref{fig:virtualpointstosharing}, the existentially-quantified $\delta$
is the ghost name for the authoritative copy of this map (again, for the current address space).
$m$ is the (full) logical map from the various page table bases to ghost names for the 
per-address-space mappings like $\Theta$.
It is the full version of the map named by the ghost name $\delta{}s$ in
Figure \ref{fig:virtualpointstosharing}; there is only one such map in the system.
This is why Figure \ref{fig:virtualpointstosharing} asserts there is some ghost name in that (global) map
for the current address space.
The global map itself is tracked by a resource algebra 
\[\mathcal{A}\textsf{spaces} \stackrel{\triangle}{=} \textsc{Auth} (\; \mathcal{W}_{64} \;\rightarrow_{\textrm{fin}} \;  ( (\textsc{Frac }, \mathord{+}) \times (\textsc{Agree } \textsc{GhostName},\mathord{=}) )) \]

$\mathcal{I}\textsf{ASpace}(\Theta,m)$ is then the resources for the current address space:
It ensures that the current address space is valid (it has an entry in $m$, whose corresponding
authoritative map matches $\Theta$), and asserts ownership of the appropriate fractional ownership
of the physical page table walk for each virtual-to-physical translation in $\Theta$ (for the current address space).

Figure \ref{fig:logicaladdrspace} depicts the relationships between these pieces, for a single
address space.
The invariant for the currently active address space (solid round square labelled as 
$\mathcal{I}$\textsf{ASpace} in Figure \ref{fig:logicaladdrspace}) asserts authorative ownership (i.e. 
update capability), $\ownGhost\gammaPred{\authfull{\ptableabswalk\ptablestore}}$, of the ghost map $\Theta$
describing virtual-to-physical address translations
-- shown as a solid-round-head arrow to $\mathcal{A}\textsf{PTableWalk}$.
Each virtual points-to assertion (dashed round square labelled as \textsf{Virtual PointsTo}
 in Figure \ref{fig:peraspaceinvariant})
asserts fragmental ownership of the ghost map $\Theta$, $\sumwalkabs\vaddr\qfrac\paddr$,
shown via the dashed-diamond-head arrow to $\mathcal{A}\textsf{PTableWalk}$. Specifically, it asserts that
the virtual address the virtual points-to assertion concerns corresponds to some physical address.
There exists only one persistent authorative ownership 
($\ownGhost\gammaPred{\authfull{\ptableabswalk\ptablestore}}$) which collects the fragmental 
ghost page-table-walk mapping ($\sumwalkabs\vaddr\qfrac\paddr$),
 shown as double-ended solid arrow 1 to N imposing the update protocol.

\begin{figure}
   \includegraphics[width=0.75\columnwidth]{logical_addr_space.png}
  \caption{Logical Resources of Address-Space Abstraction (1 Address Space)}
  \label{fig:logicaladdrspace}
  \end{figure}

\paragraph{From A Single Address Space to Many}
In principle the resources from each individual address space should be collected into a single
shared invariant, and for each memory access,
a fragment of this corresponding to $\mathcal{I}\textsf{ASpace}(\Theta,m)$
for the current address space should be extracted from this global resource, used to prove correctness of the
memory access, and put back.
In this paper we focus on the middle section, explicity identifying resources for each
address space. The reason for this is that in practice, this global resource is also
the place where kernel-specific assumptions, such as guaranteeing that certain virtual address range
was mapped in all address spaces, would be enforced. This paper focuses on the
reasoning principles behind the virtual memory access, and we leave the use of this within a larger
kernel to future work.

\subsection{Address-Space Management}
\label{sec:aspacemanagement}
So far, we have introduced logical abstractions for a single address space. However, virtual memory management
implementations typically handle more than one address-space,
and doing this requires assertions which talk about other address spaces, and means to switch address spaces.
%

%

%
%
%

%
\begin{figure}
\[
  \begin{array}{r@{\qquad}l}
    \hbox{(\TirNameStyle{ModalAddressSpace})} &
         [r](\mathsf{P}) \;: \mathsf{vProp} \; \Sigma \; \stackrel{\triangle}{=} \; \lambda \_,\; \mathsf{P}~\mathsf{r}.\\
    \hbox{(\TirNameStyle{ModalAddressSpaceMono})} & (\textsf{P} \vdash \textsf{Q}) \vdash  ([r](\textsf{P}) \vdash  [r](\textsf{Q}))\\
    \hbox{(\TirNameStyle{ModalAddressSpaceSep})} & [r](\textsf{P} \ast \textsf{Q}) \dashv\vdash ([r](\textsf{P}) \ast [r](\textsf{Q})\\
    \hbox{(\TirNameStyle{ModalAddressSpaceAnd})} &   [r](\textsf{P}\land \textsf{Q}) \dashv\vdash [r](\textsf{P})\land [r](\textsf{Q})\\    
    \hbox{(\TirNameStyle{ModalAddressSpaceOr})} &    [r](\textsf{P}\lor \textsf{Q}) \dashv\vdash [r](\textsf{P})\lor [r](\textsf{Q})\\   
    %
    \hbox{(\TirNameStyle{ModalAddressSpaceFact})} &  \textsf{Fact P} \stackrel{\triangle}{=} \;\forall \; ,r \; r' \ldotp  \textsf{P r} \dashv\vdash \textsf{P r'}\\
    %
    %
    \hbox{(\TirNameStyle{ModalAddressSpaceFactIntroElim})} 
    &\textsf{Fact P} \vdash  ([r](\textsf{P}) \dashv\vdash (\textsf{P}) %
  \end{array}
  \]
  \caption{Other-space Modality and Its Laws}
  \label{fig:modaldef}
  \end{figure}
Figure \ref{fig:modaldef} gives the definition of our modal operator for asserting truth of a modal
(address-space-contingent) assertion \emph{in another address space}, which we call
the \emph{other-space} modality. The definition itself is not
particularly surprising --- as our modal assertions are semantically predicates on a page table root (physical)
address, the assertion $[r](P)$ is a modal assertion which ignores the (implicit) current page table root,
and evaluates the truth of $P$ as if $r$ were the page table root. This simple definition is 
what we need in order for assertions in one address space to talk about assertions in another.

The inspiration for this assertion comes from hybrid logic~\cite{blackburn1995hybrid,areces2001hybrid,goranko1996hierarchies,gargov1993modal},
which as mentioned earlier uses as modalities the names of worlds in a Kripke model to state
that an assertion is true in that specific world.
Conceptually the address space rooted at some physical address (intended as the start of an L4 table)
is a distinct world, and it makes sense to name it by referring to the root of the
supporting page table structure. However, rather than explicitly building a Kripke model for our
assertions, we inherit them by lifting the model of our ambient logic, Iris.

This assertion is also useful in proofs: we can prove that this modality follows certain useful laws, and that it interacts with
a class of assertions which are true regardless of which address space they are evaluated in.
The other-space modality distributes over the various separation logic connectives, and follows the rule of
consequence.
We also define a class of \textsf{vProp}s we call \textsf{Fact}s, which are those assertions
which ignore their address space entirely.
Physical points-tos and register value assertions, when treated as \textsf{vProp}s, are \textsf{Fact}s.
\textsf{Fact}s can move in and out of other-space modalities freely.

%
%
%
%
%

%
%
%
%
%

%
%
%

%

%

%
%
%
%
%
%

\subsection{The Impact of Changing Address Spaces on Modal Assertions}
%
%
\label{sec:issues}
An important subtlety arises with supporting \lstinline|mov|s into \lstinline|cr3|. Consider the hypothetical rule
\begin{mathpar}
\inferrule[Broken]{ }{
  \{P \ast \textsf{cr3}\mapsto_{r} \;r_1 \ast r \mapsto_{r} \;r_2 \ast [r_2](Q)\}
  \texttt{mov}~\texttt{\%cr3},~r%
  \{[r_1](P) \ast \textsf{cr3} \mapsto_{r} r_2 \ast r \mapsto_{r} r_2 \ast Q\}
}
\end{mathpar}
which is slightly simplified (by dropping per-address-space resources) for clarity in making the following point.
This rule captures the intuitive change of address space in a Hoare triple.
The problem with this is that it interacts quite poorly with the traditional frame
rule and the modal flavor of virtual points-to assertions.
If we abbreviate the pre- and post-conditions above as \textsf{Pre} and \textsf{Post}:
\begin{mathpar}
  \inferrule*[right=Frame]{
    \inferrule*[right=Broken]{ }{
    \{\textsf{Pre}\}
    \texttt{mov}~\texttt{\%cr3},~r%
    \{\textsf{Post}\}
    }
  }{
    \{a\mapsto_\mathsf{v} x \ast \textsf{Pre}\}
    \texttt{mov}~\texttt{\%cr3},~r%
    \{a\mapsto_\mathsf{v} x \ast \textsf{Post}\}
  }
\end{mathpar}
Notice that both the precondition and postcondition assert that $a\mapsto_\mathsf{v} x$ in the current address space, but we have no basis for concluding that address translation is preserved by the change of address space. So this derivation clearly leads to an unsound conclusion. 
This suggestss that the traditional frame rule and the Hoare triple presentation of the change-of-address-space rule cannot soundly coexist in the same system.
The heart of the problem is that while updating \lstinline|cr3| is \emph{physically} local, it globally changes the interpretation of virtual addresses. So it is simply unsound to frame around \lstinline|cr3| updates.
In hindsight this is not surprising, because we have essentially defined
our assertions to depend on resources they do not explicitly mention --- the value of \lstinline|cr3|.
However, it has important impacts on our formal development.

Switching to Hoare doubles resolves this problem because an under-appreciated subtlety of Hoare doubles 
is that typically \emph{there is no primitive frame rule}. 
Instead each verification essentially includes a local frame that it passes to the next instructions
(think continuation-passing style), giving each overall rule a \emph{global} (rather than local) precondition. For most rules this is not that important, but it does permit rules that have global effects on their preconditions.

This is then how we justify our actual rule for \lstinline|cr3| updates:
\begin{mathpar}
\inferrule[ChangeAddressSpace]{
  \{[r_1](P) \ast cr3 \mapsto_{r} r_2 \ast r \mapsto_{r} r_2 \ast Q\}\overline{is}
}{
  \{P \ast cr3 \mapsto_{r} r_1 \ast r \mapsto_{r} r_2 \ast [r_2](Q)\}
  \texttt{mov}~\texttt{\%cr3},~r;\;\overline{is}
  %
}
\end{mathpar}
Because the precondition on this rule is global, we avoid issues with framing.

If we wanted to consider a frame rule that would work for this logic, we could consider:
\begin{mathpar}
  \inferrule[Cr3Frame]{
    \{P\ast cr3=v\}\;C\;\{ Q \ast cr3=v\}
  }{
    \{R\ast P\ast cr3=v\}\;C\;\{ R \ast Q \ast cr3=v\}
  }
\end{mathpar}
By demanding that \lstinline|cr3| be held constant (or rather, at least restored to its original value) we could frame almost traditionally. In particular, this rule would work with framing around calls that might lead to address space switches, such as calling blocking operations in the kernel.
Attempting to derive a frame rule as some work using Hoare doubles does~\cite{Chlipala2013Bedrock}
would enforce this restriction.

Readers familiar with dynamic frames~\cite{parkinson2011relationship} might find it useful to notice that a different perspective on this matter is that virtual points-to assertions are self-stable \emph{except} for changes in \lstinline|cr3|, so framing would then naturally require other means of holding \lstinline|cr3| constant (or saving and restoring it).
Virtual points-to assertions could be made self-stable by also giving them partial ownership over \lstinline|cr3| assertions, but this would require explicitly plumbing that ownership from \emph{all} assertions back to any place an address space change might occur; this would seem to be a far graver loss of modularity than this extra quirk in framing discussions.

\subsection{Selected Logical Rules}
\label{sec:selected_rules}
To address the issues described in Section \ref{sec:issues}, we give our logic using
the Hoare-Double definition in Figure \ref{fig:wpddefinition}.\footnote{The figure does
omit some low-level Iris details related to stuckness and observations, which play no meaningful
role in our development.}
\begin{figure} 
  \[
  \begin{array}{l}
    %
    \{ \Phi \}_\mathsf{ rtv }\;\textsf{e} : \textsf{iProp }\Sigma := 
   %
   ((\textsf{cr3} \mapsto_{\textsf{r}} \textsf{rtv} \ast \Phi) \textsf{ rtv}) \wand \textsf{WP e } \{\_, \textsf{True} \}
    \end{array}
  \]
\caption{Unfolded Hoare-Double Definition for \textsf{vProp} Logic }
\label{fig:wpddefinition}
\end{figure}
Our Hoare doubles $\{\Phi\}_\textsf{rtv}\;\textsf{e}$ state that the expression (i.e., sequence of instructions)
\textsf{e} are safe to execute (will not fault)
when executed with precondition $\Phi\ast\textsf{cr3}\mapsto_{\textsf{r}} \textsf{rtv}$.
\textsf{WP} is Iris's own weakest precondition modality, unmodified~\cite{jung2018iris}.
Making \textsf{rtv} an explicit parameter to the double, rather than simply using register assertions
directly solves a technical problem with ensuring that the page table root used to evaluate
the \textsf{vProp} (i.e., evaluating the assertion in the \emph{current}) address space
is feasible.\footnote{Consider the difficulty of selecting the correct page table root value from an arbitrary
opaque $\Phi$, which may even existentially quantify the page table root. An alternative is to
require $\Phi$ to have a syntactic form where we can directly extract the value of \lstinline|cr3|,
but this makes using Iris Proof Mode (IPM)~\cite{Krebbers:2017:IPH:3009837.3009855} with \textsf{vProps}
  difficult; IPM works for any type matching the signature of an Iris \lstinline|bi|, which includes
  \textsf{vProp}s, but manually guiding IPM to put an assertion in a specific position over and over adds
  significant proof burden.
}

The rest of this section describes specifications of some selected \textsf{AMD64} instructions 
in our logic. 
These rules and others (e.g., including accessing memory at an instruction-specified offset from a register
value, which is common in most ISAs)
can be found in our artifact.
Each rule in Figure \ref{fig:wpdamd}, 
is annotated with an address value (\textsf{rtv}), 
i.e. a root address of an address-space, under which the resources mentioned in the specification are valid,
or equivalently, the assumed value of \lstinline|cr3| prior to executing the instructions.
In general, we use $\textsf{r}_s$ and $\textsf{r}_d$ to specify source and destination registers
for each instruction, and prefix various register value variables with \textsf{rv}.
We sometimes use $\textsf{r}_a$ to emphasize when a register is expected to hold an address used
for memory access, though the figure also uses typical assembler conventions of specifying
memory access operands by bracketing the register holding the memory address.
Standard for Hoare doubles, there is a frame resource $P$ in each rule for passing resources
not used by the first instruction in sequence through and on to subsequent instructions.

\begin{figure}
\begin{mathpar}
  \inferrule[WriteToRegFromReg]{
  \{P \ast r_d \mapsto_{r} \textsf{rvs} \ast r_s \mapsto_{r}\{q\} \textsf{ rvs} \}_{\textsf{rtv}}\;\overline{ is}
}{
  \{P \ast r_d \mapsto_{r} \textsf{rvd} \ast r_s \mapsto_{r}\{q\} \textsf{ rvs} \}_{\textsf{rtv}}
  \textsf{ mov}~\textsf{r}_d,~\textsf{r}_s;\;\overline{is}
  %
}
\\
%
%
%
%
%
%
%
%
%
%
%
%
%
%
\inferrule[WriteToRegFromVirtMem]{
  \{P \ast \mathcal{I}\textsf{ASpace}\ast \mathcal{I}\textsf{ASpace} \ast r_d \mapsto_{r}  \textsf{v} \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{v,rtv}} \textsf{v} \}_{\textsf{rtv}}\;\overline{is}
}{
  \{P \ast \mathcal{I}\textsf{ASpace}\ast r_d \mapsto_{r}  \textsf{rvd} \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{v}} \textsf{v} \}_{\textsf{rtv}}
  \textsf{mov}~\textsf{r}_d,~[\textsf{r}_a];\;\overline{is}
}
\\
\inferrule[WriteToVirtMemFromReg]{
  \{P \ast \mathcal{I}\textsf{ASpace}\ast r_s \mapsto_{r}\{q\}  \textsf{rvs}  \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{v}} \textsf{v} \}_{\textsf{rtv}}\;\overline{ is}  
}{
  \{P \ast \mathcal{I}\textsf{ASpace}\ast r_s \mapsto_{r}\{q\}  \textsf{rvs}   \ast r_a \mapsto_{r}\{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{v}} \textsf{rvs} \}_{\textsf{rtv}}
  \textsf{ mov}~[\textsf{r}_a],~\textsf{r}_s;\;\overline{is}
}
\\
\inferrule[WriteToRegCtlFromReg]{
  \{P \ast r_s \mapsto_{r}\{q\}  \textsf{ rvs}  \}_{\textsf{rvs}} \overline{is}
}{
  \{P \ast r_s \mapsto_{r}\{q\}  \textsf{ rvs}   \}_{\textsf{rtv}}
  \textsf{mov}~\textsf{cr3},~r_s;\;\overline{is}
  %
}\\
\inferrule[WriteToRegCtlFromRegModal]{
  \{[\textsf{rtv}](P \ast \mathcal{I}\textsf{ASpace})\ast \mathcal{I}\textsf{ASpace} \ast R\ast r_s \mapsto_{r}\{q\}  \textsf{ rvs}  \}_{\textsf{rvs}} \overline{is}
}{
  \{P \ast \mathcal{I}\textsf{ASpace} \ast [\textsf{rvs}](R \ast\mathcal{I}\textsf{ASpace})\ast r_s \mapsto_{r}\{q\}  \textsf{ rvs}   \}_{\textsf{rtv}}
  \textsf{mov}~\textsf{cr3},~r_s;\;\overline{is}
  %
}\\
\inferrule[WriteToRegFromRegCtl]{
  \{P \ast r_d \mapsto_{r} \textsf{rvs} \ast r_s \mapsto_{r}\{q\} \textsf{ rvs} \}_{\textsf{rtv}}\;\overline{ is}
}{
  \{P \ast r_d \mapsto_{r} \textsf{rvd} \ast r_s \mapsto_{r}\{q\} \textsf{ rvs} \}_{\textsf{rtv}}
  \textsf{ mov}~\textsf{r}_d,~\textsf{r}_s;\;\overline{is}
  %
}
\end{mathpar}
\caption{Reasoning Rules for Selected \textsf{AMD64} Instructions under}
\label{fig:wpdamd}
\end{figure}

We briefly describe several key rules representative of the system.

Figure \ref{fig:wpdamd}, includes two  rules for accessing memory at an address in a register $r_a$. 
Ultimately, any memory access needs to ensure the relevant address translation would succeed,
which can be ensured by what we informally call the page-table-walk points-to collection
($\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$ in Figure \ref{fig:strongvirtualpointsto}).

%
%
%
%
%
%
%
%
%

\TirNameStyle{WriteToRegFromVirtMem} and \TirNameStyle{WriteToVirtMemFromReg}
each use a virtual-pointsto assertion ($\textsf{vaddr} \mapsto_{\textsf{v},\textsf{rtv}}$),
and are nearly-standard (assembly) separation logic rules for memory accesses.
However, because we split the physical resources for the page table walk from the
virtual points-to itself (in order to permit the physical page tables to be freely modified
as long as they preserve virtual-to-physical translations), the rule requires $\mathcal{I}\textsf{ASpace}$
for the current address space to be carried through.
While doing the proof of soundness for this rule,
the token ($\sumwalkabs\vaddr\qfrac\paddr$) from the virtual points-to
is used to extract to extract page-table-traversal points-to collection
for the relevant address from that resource,
and then those  physical points-to assertions describing the page table walk
(from $\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$) are used to prove
the operational semantics succeed and have the desired effect.

\paragraph{Updating \lstinline|cr3|} 
Unlike other rules, \TirNameStyle{WriteToRegCtlFromReg} updates the root address of the 
address-space determining the validity of resources, from $\rtv$ before the
\lstinline|mov| to $\textsf{rvs}$ afterwards.
From that rule, with one step outside the \textsf{vProp} logic (so it is not quite a derived rule within
our embedded logic), we can derive 
\TirNameStyle{WriteToRegCtlFromRegModal},
which is the (full) Hoare double version of the address space switch rule proposed in Section \ref{sec:issues}.

\subsection{Soundness}
Our rules from Figure \ref{fig:wpdamd} are proven sound in Iris against an assembly-level hardware model
implementing a fragment of x86-64 including 64-bit address translation with 4-level page tables.

At the moment our proofs do rely on 11 small axioms of properties which should be provable, but
are challenging to discharge due to some representation choices in our model.\footnote{See \lstinline|srx/x64/machine/current_axioms.v|}
 %
%
%
%
\section{Experiment}
\label{sec:experiment}
To both validate and demonstrate the value of the modal approach to reasoning about virtual memory management, we study several
distillations of real concerns of virtual memory managers.
Recall from Section \ref{sec:logic} that virtual points-to assertions work just like regular points-to assertions, by design.

%
%
%
%
%
%
%
%
%
%
\subsection{Mapping a New Page}
One of the key tasks of a page fault handler in a general-purpose OS kernel is
to map new pages into an address space by writing into an existing page table.
To do so, we first allocate a fresh page, then calculate the appropriate
known-valid page table walks and update the appropriate L1 page table entry.
\lstset{
  columns=fullflexible,
  numbers=left,
  basicstyle=\ttfamily,
  keywordstyle=\color{blue}\bfseries,
  morekeywords={mov,add,call},
  emph={rsp,rdx,rax,rbx,rbp,rsi,rdi,rcx,r8,r9,r10,r11,r12,r13,r14,r15},
  emphstyle=\color{green},
  emph={[2]cr3},
  emphstyle={[2]\color{violet}},
  morecomment=[l]{;;},
  mathescape
}
\newcommand{\fpaddr}{\texttt{fpaddr}}
\newcommand{\specline}[1]{{\color{blue}\left\{#1\right\}}}
\begin{figure}\footnotesize
  \begin{lstlisting}
$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m)  \ast  \texttt{r14}\mapsto_{\textsf{r}} \_ \ast \texttt{rdi}\mapsto_{r} \vaddr \ast \texttt{rax}\mapsto_{\textsf{r}} \_ \ast \ulcorner \texttt{aligned va} \land  \theta \; !!\; \vaddr = \texttt{None}\urcorner}_{\rtv}$
call ensure_L1_page
$\specline{\exists (\entryf ,\;\entrytr,\; \entrytw,\; \entryo,\;\textsf{pte\_addr },\paddr) \; \ldotp\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast  \texttt{r14}\mapsto_{\textsf{r}} \_ \ast \texttt{rdi}\mapsto_{r} \vaddr \ast \texttt{rax}\mapsto_{\textsf{r}} \textsf{ pte\_addr} \; \ast }_{\rtv}$
$\specline{ \ulcorner  \texttt{addr\_L1 }(\vaddr, \entryo) = \paddr \urcorner \ast \ulcorner \texttt{entry\_present } \entryf \land \texttt{entry\_present } \entrytr \land  \texttt{entry\_present } \entrytw \urcorner \; \ast}_{\rtv}$
$\specline{\nfpointsto{\mask\vaddr\maskfour\rtv}{\mask\vaddr\maskfouroff\rtv}\entryf\qone\naddr \; \ast}_{\rtv}$ 
$\specline{  \nfpointsto{\mask\vaddr\maskthree\entryf}{\mask\vaddr\maskthreeoff\entryf}\entrytr\qtwo\naddr \ast \nfpointsto{\mask\vaddr\masktwo\entrytr}{\mask\vaddr\masktwooff\entrytr}\paddr\qthree\entryo \;\ast}_{\rtv}$
$\specline{\texttt{pte\_addr} \mapsto_{\texttt{vpte}} \paddr \;(\texttt{wzero 64}) \ast \texttt{rax}\mapsto_{\textsf{r}} \texttt{pte\_addr}  }_{\rtv}$
;; Returns the virtual address of the L1 entry in rax
mov %
$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast  \texttt{r14}\mapsto_{\textsf{r}} \texttt{pte\_addr} \ast \texttt{rdi}\mapsto_{\textsf{r}} \vaddr \ast \texttt{rax}\mapsto_{\textsf{r}} \textsf{ pte\_addr} \; \ast }_{\rtv}$
$\specline{ \nfpointsto{\mask\vaddr\maskfour\rtv}{\mask\vaddr\maskfouroff\rtv}\entryf\qone\naddr \ast \ulcorner \texttt{entry\_present } \entryf \land \texttt{entry\_present } \entrytr \land  \texttt{entry\_present } \entrytw \urcorner \ast}_{\rtv}$ 
$\specline{  \nfpointsto{\mask\vaddr\maskthree\entryf}{\mask\vaddr\maskthreeoff\entryf}\entrytr\qtwo\naddr \ast \nfpointsto{\mask\vaddr\masktwo\entrytr}{\mask\vaddr\masktwooff\entrytr}\paddr\qthree\entryo \;\ast}_{\rtv}$
$\specline{\texttt{pte\_addr} \mapsto_{\texttt{vpte}} \paddr \;(\texttt{wzero 64}) \ast \texttt{rax}\mapsto_{\textsf{r}} \texttt{pte\_addr}  }_{\rtv}$
call alloc_phys_page_or_panic
$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast  \texttt{r14}\mapsto_{\textsf{r}} \texttt{pte\_addr} \ast \texttt{rdi}\mapsto_{\textsf{r}} \vaddr \;\ast}_{\rtv}$
$\specline{ \nfpointsto{\mask\vaddr\maskfour\rtv}{\mask\vaddr\maskfouroff\rtv}\entryf\qone\naddr \ast}_{\rtv}$ 
$\specline{  \nfpointsto{\mask\vaddr\maskthree\entryf}{\mask\vaddr\maskthreeoff\entryf}\entrytr\qtwo\naddr \ast \nfpointsto{\mask\vaddr\masktwo\entrytr}{\mask\vaddr\masktwooff\entrytr}\paddr\qthree\naddr \;\ast}_{\rtv}$
$\specline{\texttt{pte\_addr} \mapsto_{\texttt{vpte}} \paddr\; (\texttt{wzero 64})  \ast \ulcorner \texttt{entry\_present } \entryf \land \texttt{entry\_present } \entrytr \land  \texttt{entry\_present } \entrytw \urcorner}_{\rtv}$
$\specline{\exists \texttt{ fpaddr} \ldotp \ulcorner \texttt{aligned fpaddr} \urcorner \ast \texttt{rax}\mapsto_{\textsf{r}} \texttt{fpaddr+3} \ast \texttt{fpaddr} \mapsto_{\textsf{p}} (\texttt{wzero 64}) \ast \ulcorner \texttt{entry\_present (fpaddr+3)}\urcorner}_{\rtv}$
;; Calculate new L1 entry
;; update the page table entry, mapping the page
mov (%
$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast  \texttt{r14}\mapsto_{\textsf{r}} \texttt{pte\_addr} \ast \texttt{rdi}\mapsto_{\textsf{r}} \vaddr \;\ast}_{\rtv}$
$\specline{ \nfpointsto{\mask\vaddr\maskfour\rtv}{\mask\vaddr\maskfouroff\rtv}\entryf\qone\naddr \ast}_{\rtv}$ 
$\specline{  \nfpointsto{\mask\vaddr\maskthree\entryf}{\mask\vaddr\maskthreeoff\entryf}\entrytr\qtwo\naddr \ast \nfpointsto{\mask\vaddr\masktwo\entrytr}{\mask\vaddr\masktwooff\entrytr}\paddr\qthree\entryo \;\ast}_{\rtv}$
$\specline{\texttt{pte\_addr} \mapsto_{\texttt{vpte}} \paddr \;(\texttt{fpaddr+3}) \; \ast \ulcorner \texttt{entry\_present } \entryf \land \texttt{entry\_present } \entrytr \land  \texttt{entry\_present } \entrytw \urcorner }_{\rtv}$
$\specline{\ulcorner \texttt{aligned fpaddr} \urcorner \ast \texttt{rax}\mapsto_{\textsf{r}} \texttt{fpaddr+3} \ast \texttt{fpaddr} \mapsto_{\textsf{p}} (\texttt{wzero 64}) \ast \ulcorner \texttt{entry\_present fpaddr+3}\urcorner}_{\rtv}$
$\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\; \sqsubseteq $
$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast  \texttt{r14}\mapsto_{\textsf{r}} \texttt{pte\_addr} \ast \texttt{rdi}\mapsto_{\textsf{r}} \vaddr \ast }_{\rtv}$
$\specline{\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\vaddr,\entryf,\entrytr,\entrytw,\fpaddr+3) \ast \ulcorner \theta \;!!\;\vaddr = \texttt{None}\urcorner \; \ast}_{\rtv}$
$\specline{\ulcorner \texttt{aligned fpaddr} \urcorner \ast \texttt{rax}\mapsto_{\textsf{r}} \texttt{fpaddr+3} \ast \texttt{fpaddr} \mapsto_{\textsf{p}} (\texttt{wzero 64}) }_{\rtv}$
$\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\; \sqsubseteq $
$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace} (<[\vaddr:=\texttt{fpaddr}]> \theta,m) \ast}_{\rtv}$
$\specline{\ulcorner \texttt{aligned fpaddr} \urcorner \ast \texttt{fpaddr} \mapsto_{\textsf{p}} \textsf{ wzero 64} \ast \sumapaces\rtv\delta  \ast\sumwalkabs\vaddr\qfrac\fpaddr}_{\rtv}$
$\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\; \sqsubseteq $
  $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace} (<[\vaddr:=\texttt{fpaddr}]> \theta,m) \ast \vaddr \mapsto_{\textsf{vpte}}\; \{\qfrac\} \;\fpaddr \textsf{ wzero 64}}_{\rtv}$
$\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\; \sqsubseteq $
  $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace} (<[\vaddr:=\texttt{fpaddr}]> \theta,m) \ast \vaddr \mapsto_{\textsf{v}}\; \{\qfrac\} \textsf{wzero 64}}_{\rtv}$
\end{lstlisting}
  \caption{Specification and proof of distilled code for mapping a new page.}
\label{fig:mapping_code}
\end{figure}
In Figure \ref{fig:mapping_code}, we see an address ($\vaddr$) currently not
mapped to a page ($\theta \; !!\; \vaddr = \texttt{None}$). Mapping a fresh
phyiscal page to back the desired virtual page first requires ensuring
the existence of a memory location of an appropriate L1 table entry,
which is realized in two pieces by the post-condition of \lstinline|ensure_L1|:
\begin{itemize}
\item physical pointsto assertions on the tables L4, L3, and L2 reaching to the
	L1 level entry (l1e) ensure higher levels of the page table exist, 
	are marked present (\textsf{entry\_present} for the relevant
		entries \textsf{l4e l3e l2e}) (Specification Lines 4-6) 
	\item a virtual \emph{PTE}
		pointsto ($\mapsto_{\textsf{vpte}}$) allows access to the memory of the L1 entry
		via virtual address \lstinline|pte_pointsto|.
		A PTE points-to is defined just like the normal virtual points-to of Figure \ref{fig:virtualpointstosharing}, except the physical address is explicit in the assertion (here, \textsf{pa})
		rather than existentially quantified:
 \[
\begin{array}{l}
    \vaddr\mapsto_{\textsf{vpte}} \;\{\textsf{q}\} \; \paddr \; \vpage : \mathsf{vProp}~\Sigma \stackrel{\triangle}{=} 
    \exists \delta\ldotp
	(\lambda \mathit{cr3val}\ldotp
	\sumapaces{\mathit{cr3val}}\delta) \ast 
  \sumwalkabs\vaddr\qfrac\paddr \ast \paddr \mapsto_{\mathsf{p}} \vpage
\end{array}
\]
		This supports rules for accessing memory
		at that virtual address, but exposing the physical location being modified
		makes this convenient to use for page table modifications, since we must ensure
		the modified address is the correct physical address that will be used as the L1 entry.
		This is also guaranteed ($ \ulcorner
		\texttt{addr\_L1}(\vaddr,\entryo) = \paddr\urcorner$).
\end{itemize}
After obtaining a virtual address \textsf{pte\_addr} in \textsf{rax} backed 
by the physical memory for the L1 entry that will be used to translate the virtual addresses
we are mapping, we save it to \textsf{r14} to be updated later in Line 9.

Then we allocate a fresh page initialized (i.e. zeroed) in Line 14 which return
an \textsf{aligned} address of fresh page (\textsf{fpaddr}) in \textsf{rax}
(Specification Line 19), plus 1 (so the return value is already a valid page table entry).

The crucial point in the proof is then actually updating the L1 entry,
via the virtual address
(\textsf{pte\_addr}) known to translate to the appropriate physical address, in our example the L1
table entry address ($\textsf{addr\_L1}(\textsf{va, l1e})$), to hold the freshly
allocated physical page address (\textsf{fpaddr}) in Line 22.
After this write, we have (Specification Line 26)
\[\texttt{pte\_addr} \mapsto_{\texttt{vpte}}  \; \paddr\; \textsf{fpaddr}  \qquad\qquad (\textsf{by the PTE variant of Rule}~\TirNameStyle{WriteToVirtMemFromReg})\]

At this point in the proof,  $\texttt{pte\_addr} \mapsto_{\texttt{vpte}}  \; \paddr\; \textsf{fpaddr})$ 
contains  full ownerhsip of the L1 table entry physical pointsto $\paddr
\mapsto_{\mathsf{p}} \textsf{fpaddr}$ . Then, we can split the full-ownership of that
physical pointsto
($\textsf{fpaddr}\mapsto_{\textsf{p}} \;\textsf{wzero 64}$) to obtain
fractional ownership on it appropriate for a single address's share of an L1 entry (q4).
Since we know $\texttt{addr\_L1 }(\vaddr,
\entryo) \mapsto_{\mathsf{p}} \{q4\} \;\textsf{fpaddr}$) from
$\mapsto_{\textsf{vpte}}$, after the split of the PTE points-to, we can
connect L4-L2 points-to assertions (in Specification Lines 24 and 25) with 
 $\texttt{fpaddr} \mapsto_{a} (\texttt{wzero 64})$ to obtain the
complete physical table-walk assertion (L4\_L1\_PointsTo($\vaddr$ l4e l3e l2e
fpaddr)) in Line 30.

Finally, we can insert $\vaddr$ into the ghost page-table-walk summarization
map ($\theta$) to change its state from unmapped ($\ulcorner \theta
\;!!\;\vaddr = \texttt{None}\urcorner$ in Specification Line 30 and the
precondition) to mapped (Specification Line 33) using Iris' ghost-map update
($\sqsubseteq$), and construct a PTE -points-to $\vaddr
\mapsto_{\textsf{vpte}}\; \qfrac \;\fpaddr\;(\textsf{wzero 64})$ (Specification
Line 36) using $\sumwalkabs\vaddr\qfrac\fpaddr$ obtained from ghost
page-table-walk insertions, $\sumapaces\rtv\delta$ from unfolding the
definition of $\mapsto_{\textsf{vpte}}$ in Specification Line 18, and $\fpaddr
\mapsto_{\textsf{p}} \textsf{wzero 64}$ in Specification Line 34.
Finally, by existentially quantifying away \textsf{fpaddr},
we can shift the PTE points-to into a normal virtual points-to.

This proof only maps the first word of the allocated page; the generalization to the full page
is straightforward use of iterated separating conjunction to manage collections of clusters of 512 things
(slices of the L1 entry physical ownership, virtual addresses at 8 byte increments, etc.).

\subsection{Unmapping a Page}
The reverse operation, unmapping a designated page that is currently mapped,
would essentially be the reverse of
the reasoning around line 22 above: given the virtual points-to assertions for all 512
machine words of memory that the L1 entry would map,
and information about the physical location, 
full permission on the L1 entry could be obtained, allowing the construction of a
full virtual PTE pointer for it, setting to 0, and reclaiming the now-unmapped physical memory.

\subsection{Change of Address Space}
A critical piece of \emph{trusted} code in verified OS kernels is the assembly code to change the current address space; current verified OS kernels currently lack effective ways to specify and reason about this low-level operation, for reasons outlined in Section \ref{sec:relwork}.
\begin{figure}\footnotesize
\begin{lstlisting}
;; Assume the save-space is in rdi,
;; load-space in rsi
;; No need to save/restore caller-save regs
;; Save yielding context
$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast [\rtv'+56](\mathcal{I}\texttt{ASpace}(\theta,m) \ast \texttt{Pother})}_{\rtv}$
$\specline{ \texttt{rsi}\mapsto_{\textsf{r}} \rtv' \ast \texttt{rdi}\mapsto_{\textsf{r}} \_ \ast \texttt{rbx}\mapsto_{\textsf{r}} \texttt{rbxv} \ast  \texttt{rsp}\mapsto_{\textsf{r}} \texttt{rspv} \ast \texttt{rbp}\mapsto_{\textsf{r}} \texttt{rbpv} }_{\rtv}$
$\specline{\texttt{r12}\mapsto_{\textsf{r}} \texttt{r12v} \ast \texttt{r13}\mapsto_{\textsf{r}} \texttt{r13v} \ast \texttt{r14}\mapsto_{\textsf{r}} \texttt{r14v} \ast \texttt{r15}\mapsto_{\textsf{r}} \texttt{r15v}}_{\rtv}$
$\specline{\texttt{rdi+0} \mapsto_{\textsf{v}} \_ \ast \texttt{rdi+8} \mapsto_{\textsf{v}} \_ \ast \texttt{rdi+16} \mapsto_{\textsf{v}} \_ \ast \texttt{rdi+24} \mapsto_{\textsf{v}} \_ \ast \texttt{rdi+32} \mapsto_{\textsf{v}} \_}_{\rtv}$
$\specline{\texttt{rdi+40} \mapsto_{\textsf{v}} \_ \ast \texttt{rdi+48} \mapsto_{\textsf{v}} \_\ast \texttt{rdi+56} \mapsto_{\textsf{v}} \_}_{\rtv}$
mov 0[%
mov 8[%
mov 16[%
mov 24[%
mov 32[%
mov 40[%
mov 48[%
$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast [\rtv'+56](\mathcal{I}\texttt{ASpace}(\theta,m) \ast \texttt{Pother})}_{\rtv}$
$\specline{ \texttt{rdi}\mapsto_{\textsf{r}} \_ \ast \texttt{rbx}\mapsto_{\textsf{r}} \texttt{rbxv} \ast  \texttt{rsp}\mapsto_{\textsf{r}} \texttt{rspv} \ast \texttt{rbp}\mapsto_{\textsf{r}} \texttt{rbpv} }_{\rtv}$
$\specline{\texttt{r12}\mapsto_{\textsf{r}} \texttt{r12v} \ast \texttt{r13}\mapsto_{\textsf{r}} \texttt{r13v} \ast \texttt{r14}\mapsto_{\textsf{r}} \texttt{r14v} \ast \texttt{r15}\mapsto_{\textsf{r}} \texttt{r15v}}_{\rtv}$
$\specline{\texttt{rdi+0} \mapsto_{\textsf{v}} \texttt{rbxv} \ast \texttt{rdi+8} \mapsto_{\textsf{v}} \texttt{rspv} \ast \texttt{rdi+16} \mapsto_{\textsf{v}} \texttt{rbpv} \ast \texttt{rdi+24} \mapsto_{\textsf{v}} \texttt{r12v} \ast}_{\rtv}$
$\specline{ \texttt{rdi+32} \mapsto_{\textsf{v}} \texttt{r13v} \ast \texttt{rdi+40} \mapsto_{v} \texttt{r14v} \ast \texttt{rdi+48} \mapsto_{\textsf{v}} \texttt{r15v}\ast \texttt{rdi+56} \mapsto_{\textsf{v}} \_}_{\rtv}$
mov 56[%
$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast [\rtv'+56](\mathcal{I}\texttt{ASpace}(\theta,m) \ast \texttt{Pother})}_{\rtv}$
$\specline{ \texttt{rdi}\mapsto_{\textsf{r}} \_ \ast \texttt{rbx}\mapsto_{\textsf{r}} \texttt{rbxv} \ast  \texttt{rsp}\mapsto_{\textsf{r}} \texttt{rspv} \ast \texttt{rbp}\mapsto_{\textsf{r}} \texttt{rbpv} }_{\rtv}$
$\specline{\texttt{r12}\mapsto_{\textsf{r}} \texttt{r12v} \ast \texttt{r13}\mapsto_{\textsf{r}} \texttt{r13v} \ast \texttt{r14}\mapsto_{\textsf{r}} \texttt{r14v} \ast \texttt{r15}\mapsto_{\textsf{r}} \texttt{r15v}}_{\rtv}$
$\specline{\texttt{rdi+0} \mapsto_{\textsf{v}} \texttt{rbxv} \ast \texttt{rdi+8} \mapsto_{\textsf{v}} \texttt{rspv} \ast \texttt{rdi+16} \mapsto_{\textsf{v}} \texttt{rbpv} \ast \texttt{rdi+24} \mapsto_{\textsf{v}} \texttt{r12v} \ast }_{\rtv}$
$\specline{\texttt{rdi+32} \mapsto_{\textsf{v}} \texttt{r13v} \ast \texttt{rdi+40} \mapsto_{\textsf{v}} \texttt{r14v} \ast \texttt{rdi+48} \mapsto_{\textsf{v}} \texttt{r15v}\ast \texttt{rdi+56} \mapsto_{\textsf{v}} \rtv}_{\rtv}$    
;; Restore target context
mov %
;; This switches to a new stack
;; This stack may not be mapped in the
;; current address space!
mov %
mov %
mov %
mov %
mov %
mov %
$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast [\rtv'+56](\mathcal{I}\texttt{ASpace}(\theta,m) \ast \texttt{Pother})}_{\rtv}$
$\specline{ \texttt{rdi}\mapsto_{\textsf{r}} \_ \ast \texttt{rbx}\mapsto_{\textsf{r}} \rtv'+0 \ast  \texttt{rsp}\mapsto_{\textsf{r}} \rtv'+8 \ast \texttt{rbp}\mapsto_{\textsf{r}} \rtv'+16 }_{\rtv}$
$\specline{\texttt{r12}\mapsto_{\textsf{r}} \rtv'+24 \ast \texttt{r13}\mapsto_{\textsf{r}} \rtv'+32 \ast \texttt{r14}\mapsto_{\textsf{r}} \rtv'+40 \ast \texttt{r15}\mapsto_{\textsf{r}} \rtv'+48}_{\rtv}$
$\specline{\texttt{rdi+0} \mapsto_{\textsf{v}} \texttt{rbxv} \ast \texttt{rdi+8} \mapsto_{\textsf{v}} \texttt{rspv} \ast \texttt{rdi+16} \mapsto_{\textsf{v}} \texttt{rbpv} \ast \texttt{rdi+24} \mapsto_{\textsf{v}} \texttt{r12v} \ast }_{\rtv}$
$\specline{\texttt{rdi+32} \mapsto_{\textsf{v}} \texttt{r13v} \ast \texttt{rdi+40} \mapsto_{\textsf{v}} \texttt{r14v} \ast \texttt{rdi+48} \mapsto_{\textsf{v}} \texttt{r15v}\ast \texttt{rdi+56} \mapsto_{\textsf{v}} \rtv}_{\rtv}$
;; Switch to the new address space
mov %
$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast [\rtv'+56](\mathcal{I}\texttt{ASpace}(\theta,m) \ast \texttt{Pother}) }_{\rtv}$
$\specline{ \texttt{rdi}\mapsto_{\textsf{r}} \_ \ast \texttt{rbx}\mapsto_{r} \rtv'+0 \ast  \texttt{rsp}\mapsto_{\textsf{r}} \rtv'+8 \ast \texttt{rbp}\mapsto_{\textsf{r}} \rtv'+16 }_{\rtv}$
$\specline{\texttt{r12}\mapsto_{\textsf{r}} \rtv'+24 \ast \texttt{r13}\mapsto_{\textsf{r}} \rtv'+32 \ast \texttt{r14}\mapsto_{\textsf{r}} \rtv'+40 \ast \texttt{r15}\mapsto_{\textsf{r}} \rtv'+48}_{\rtv}$
$\specline{\texttt{rdi+0} \mapsto_{\textsf{v}} \texttt{rbxv} \ast \texttt{rdi+8} \mapsto_{\textsf{v}} \texttt{rspv} \ast \texttt{rdi+16} \mapsto_{\textsf{v}} \texttt{rbpv} \ast  \texttt{rdi+24} \mapsto_{\textsf{v}} \texttt{r12v} \ast }_{\rtv}$
$\specline{\texttt{rdi+32} \mapsto_{\textsf{v}} \texttt{r13v} \ast \texttt{rdi+40} \mapsto_{\textsf{v}} \texttt{r14v} \ast \texttt{rdi+48} \mapsto_{\textsf{v}} \texttt{r15v}\ast \texttt{rdi+56} \mapsto_{\textsf{v}} \rtv}_{\rtv}$
$\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\; \sqsubseteq $
$\specline{ [\rtv](\texttt{Pcurrent}) \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast \texttt{Pother}}_{\rtv'+56}$
\end{lstlisting}
\caption{Basic task switch code that switches address spaces.}
\label{fig:swtch}
\end{figure}

Figure \ref{fig:swtch} gives simplified code for a basic task switch, the heart of an OS scheduler implementation. This is code that saves the context (registers and stack) of the running thread (here in a structure pointed to by \lstinline|rdi|'s value shown in Lines 10-16 and Line 23 of Figure \ref{fig:swtch}) and restores the context of an existing thread (from \lstinline|rsi| shown in Lines 28-38 and Line 45), including the corresponding change of address space for a target thread in another process.
This code assumes the System V AMD64 ABI calling convention, where the normal registers not mentioned are caller-save, and therefore saved on the stack of the thread that calls this code, as well as on the new stack of the thread that is restored, thus only the callee-save registers and \texttt{cr3} must be restored.\footnote{We are simplifying in a couple basic ways. First, we are ignoring non-integer registers (e.g., floating point, vector registers) entirely. Second, we are ignoring that the caller-save registers should still be initialized to 0 to avoid leaking information across processes. We focus on the core logical requirements.}
With the addition of a return instruction, this code would satisfy the C function signature\footnote{The name comes from the UNIX 6th Edition \lstinline|swtch| function, the source of the infamous ``You are not expected to understand this'' comment~\cite{lions1996lions}.}
\begin{lstlisting}[language=C]
void swtch(context_t* save, context_t* restore);
\end{lstlisting}
A call to this code begins executing one thread (Lines 10-16) in one address space ($\rtv$ in Figure \ref{fig:swtch}) (whose information will be saved in \lstinline[language=C]|save| shown in Line 22) and finishes execution executing a different thread in a different address space (whose information is initially in \lstinline[language=C]|restore| Line 29).

Because this code does not directly update the instruction pointer, it is worth explaining \emph{how} this switches threads: by switching stacks. This is meant to be called with a return address for the current thread stored on the current stack when called --- which must be reflected in the calling convention. In particular, the precondition of the return address on the initial stack requires the callee-save register values at the time of the call: those stored in the first half of the code.
Likewise, part of the invariant of the stack of the second thread, the one being restored, is that the return address on \emph{that} stack requires the saved callee-save registers stored in that context to be in registers as its precondition.

The wrinkle, and the importance of the modal treatment of assertions, is that the target thread's precondition is \emph{relative to its address space}, not the address space of the calling thread shown as 
\[[\rtv'+56](\mathcal{I}\texttt{ASpace}(\theta,m) \ast \texttt{Pother})\]
in the specfication. 
Thus the precondition of this code,
in context, would include that the initial stack pointer (before \lstinline|rsp| is updated)
has a return address expecting the then-current callee-save register values and 
suitably updated (i.e., post-return) stack in the \emph{current} (initial) address space;
this would be part of \textsf{P} in the precondition.
The specification also requires that
the stack pointer saved in the context to restore expects the same of the saved registers and stack 
\emph{in the other address space}. 
The other-space modality plays a critical role here; \textsf{Pother} would contain these assumptions in the other
address space.

Lines 10--16 save the current context into memory (in the current address space).
Line 22 saves the initial page table root.
Lines 33--38 begin restoring the target context, including the stack pointer (line 33),
which may not be mapped in the address space at that time: it is the stack for the context being
loaded into the CPU.
The actual address switch occurs on line 45, which is verified with our modal rule for updating \lstinline|cr3|,
and thus shifts resources in and out of other-space modalities as appropriate.

The postcondition is analagous to the precondition, but interpreted \emph{in the new address space}: the then-current (updated) stack would have a return address expecting the new (restored) register values (again, in \textsf{Pother}),
and the saved context's invariant captures the precondition for restoring its execution \emph{in the previous address space} (as part of \textsf{P}). 
%

Note that immediately after the page table switch, the points-to information about the saved and restored contexts is \emph{also} guarded by a modality (Specification Line 52)
\[ [\rtv](\texttt{Pcurrent}) \qquad \textsf{ where Pcurrent is all available logical assertions under addrress-space} \rtv\]
for the old address space, since there is no automatic guarantee that that memory is mapped in the new address space.  The ability to transfer that points-to information out of that modality is specific to a given kernel's design. Kernels that map kernel memory into all address spaces would need to ensure and specify enough specific details about memory mappings to allow a proof of an elimination rule for specific modally-constrained points-to assertions.
Following Spectre and Meltdown, this kernel design became less prevalent because speculative execution of accesses to kernel addresses could leak information even if the access did eventually cause a fault (the user/kernel mode permission check was done after fetching data from memory). Thus many modern kernels have reverted to the older kernel design where the kernel inhabits its own unique address space, and user processes have only enough extra material mapped in their address spaces to switch into the kernel (CPUs do not speculate past updates to \texttt{cr3}).

 %
%
\section{Related Work}
\label{sec:relwork}

There has been relatively little prior work on formal verification of virtual memory.
Instead, much OS verification work has focused on minimizing reasoning about virtual memory management.
The original Verisoft project~\cite{alkassar2008verisoft,alkassar2010pervasive,alkassar2008formal,dalinger2005verification,hillebrand2005address,alkassar2008formal,starostin2010formal} relied on custom hardware which, among other things, always ran kernel code with virtual memory disabled, removing the circularity that is a key challenge of verifying actual virtual memory code: at that point page tables become a basic partial map data structure to represent user program address translations.
Other work on OS verification either never progressed far enough to address VMM verification (Verisoft XT~\cite{cohen2009vcc,cohen2010local,dahlweid2009vcc,cohen2013SOFSEM}), or uses memory-safe languages to enable safe co-habitation of a single address space by all processes (Singularity~\cite{Fahndrich2006language,Hunt2007singularity,Hunt2007sealing,Barnett2011specsharp}, Verve~\cite{Yang2010Verve}, and Tock~\cite{levy2017multiprogramming}).

The work that does address the core challenges of VMM verification is all associated with either \textsc{seL4} or \textsc{CertiKOS}.

\textsc{CertiKOS}~\cite{gu15,gu2016certikos,gu2018certikos,chen2016interrupts} is a microkernel intended for use as a hypervisor,
and its papers do not explicitly detail verification of the VMM, so we do not know the full space of which VMM functionality 
is verified, but we do know it includes the ability to map or unmap pages.
The work is clear, however, that it trusts low-level assembly fragments such as the instruction sequence which actually
switches address spaces, rather than verifying them.
The overall approach in that body of work is many layers of refinement proofs, using a
 proliferation of layers with small differences to keep most individual refinements tractable. In keeping with precursor work 
on the project from the same group~\cite{vaynberg2012compositional}, the purpose of some layers is to abstract away from 
virtual memory, so the proof is essentially a simulation proof covering for example a proof that execution with page-in on 
page faults is a valid refinement of an execution model where no paging occurs.
%
%
%
%
%
%
%
%
%
%

\textsc{seL4}~\cite{Klein2009seL4,seL4TOCS,Sewell2013translation} is a formally verified L4 microkernel~\cite{Liedtke1995,Liedtke1996} (and the first verified OS kernel to run on real-world hardware), verified with a mix of refinement proofs and program logic reasoning down to the assembly level.
Because \textsc{seL4} is a microkernel, most VMM functionality actually lives in usermode and is unverified, and moreover, their hardware model omits address translation entirely and the MMU entirely~\cite{Klein2009seL4,seL4TOCS}. As a result, the limited page table management present in the microkernel treats page tables as idiosyncratic tree-maps, ignoring the risks posed by even transient inconsistencies that would crash the kernel on real hardware (like ``temporarily'' unmapping the kernel). This is mitigated primarily by manually identifying some trusted invariants (e.g., that the address range designated for the kernel is appropriately mapped) and setting up the proof to ensure those invariants are maintained (i.e., as an extra proof obligation not required by their hardware model).

One important outgrowth of the \textsc{seL4} project, not integrated into the main project's proof, was work by 
Kolanski and Klein which studied verification of code against a hardware model that \emph{did} include address translation
 --- the only work aside from ours to do so --- initially in terms of basic memory~\cite{kolanski08vstte} and subsequently 
integrating source-level types into the interpretation~\cite{kolanski09tphols}. 
They were the first work to model physical and virtual points-to assertions separately, defining virtual points-to assertions
in terms of physical points-to assertions mimicking page table walks, and defining all of their assertions as predicates on a
pair of (physical) machine memory and a page table root, an approach we improve on.

Their work has a number of significant limitations which our work addresses.
They also define their virtual points-to assertions such that a virtual points-to $p\mapsto_\mathsf{v} a$ owns the full 
lookup path to virtual address $p$. This means that given two virtual points-to assertions at the same time, such as 
$p\mapsto_\mathsf{v}a \ast p'\mapsto_\mathsf{v}b$, the memory locations traversed to translate $p$ and $p'$ must be disjoint. 
This means the logic has a peculiar limit on how many virtual points-to assertions can coexist in a proof. Since page tables 
fan out, the bottleneck is the number of entries in the root table. For their 32-bit ARMv6 example, the top-level address is 
still 4Kb (4096 bytes), and each entry (consumed entirely by a virtual points-to in their scheme) is 4 bytes, so they have a 
maximum of 1024 virtual points-tos in their ARMv6 configuration. Any assertion which implies more than that number
of virtual addresses are mapped implies false in their logic.
(They do formulate their logic over an abstract model, but every architecture would incur a similar limitation;
Na\"ively transferring their model to x86-64 4-level tables would yield a limit of 512 assertions (also a 4Kb root page, 
but 8-byte entries).

%
%
%
%
%
%
Our definitions make use of fractional permissions throughout; Figure \ref{fig:strongvirtualpointsto}'s definition
of \lstinline|L4_L1_PointsTo| ellides the specific fractions used, but it in fact asserts 1/512 ownership of
the L1 entry, 1/($512^2$) of the L2 entry, and so on, so each entry may map the appropriate number of machine words.

As noted earlier, Kolanski and Klein's logics, by collocating both the physical ownership of the page table walk
as part of the virtual points-to itself, preempt support for changes to page tables which do not actually affect 
address translation.

The other major distinction is that Kolanski and Klein have no accounting for other address spaces.
Their logic does not deal with change of address space, and has no way to assert that certain facts hold
in another address space.
They verify only one address space manipulation: mapping a single unmapped page into the current address space (in both papers).
We verify this, as well as a change-of-address-space, which requires us to introduce assertions for talking
about other address spaces (we must know, for example, that the precondition of the code after the change must be true
in the \emph{other} address space), and to deal with the fact that the standard frame rule
for separation logic is unsound in the presence of address space changes and address-space-contingent assertions.
%
%

Our approach in this paper uses modalities to distinguish virtual-address-based assertions that hold only in specific 
address spaces, making it possible to manipulate other address spaces, and equally critically, to \emph{change} address 
spaces while reasoning about correctness.

Unlike our work, Kolanski and Klein prove very useful embedding theorems stating that code that does not modify page table 
entries can be verified in a VM-ignorant program logic, and that proofs in that logic can be embedded into the VM-aware logic 
(essentially by interpreting ``normal'' points-to relations as virtual points-to facts). While we have not proven such a result,
an analagous result {should} hold of our work: consider that the doubles for the \texttt{mov} instructions
that access memory behave just as one would expect for a VM-ignorant logic~\cite{Chlipala2013Bedrock}.
With our general approach to virtual points-to assertions being inspired by Kolanski and Klein, \emph{both}
 our approach and theirs could in principle be extended to account for pageable points-to assertions by adding additional 
disjunctions to an extended points-to definition; embedding ``regular'' separation logic into such a variant
is the appropriate next step to extend reasoning to usermode programs running with a kernel that may demand-page the program's
memory.

As noted throughout the paper, the inspiration for our other-space modality comes from hybrid logic~\cite{areces2001hybrid,blackburn1995hybrid,gargov1993modal,goranko1996hierarchies},
where modalities are indexed by \emph{nominals} which are names for specific individual states in a Kripke model.
We are aware of only two prior works combining hybrid logics with program logics specifically. 
Brotherston and Villard~\cite{brotherston2014parametric} demonstrated that may properties true of various 
separation logics are not definable in boolean \BI (\BBI), and showed that a hybrid extension \HyBBI allows
most such properties to be defined (e.g., the fact that separating conjunction is cancellative is unprovable 
in boolean \BI, but provable in \HyBBI). There, nominals named resources 
(roughly, but not exactly, heap fragments). 
Gordon~\cite{gordon2019modal} described a use of hybrid logic in the verification of actor programs, 
where nominals named the local state of individual actors (with such assertions stabilized with a 
rely/guarantee approach). Beyond these, there is limited work on the interaction of specifically 
\emph{hybrid logic} with substructural logics. 
Primarily there is a line of work on hybrid linear logic (\HyLL)~\cite{despeyroux2014hybrid}, 
originally used as a way to more conveniently express aspects of transition systems in linear logic. 
However, \HyLL's proof rules offer no non-trivial interactions with multiplicative connectives 
(every \HyLL proof can in fact be embedded into regular linear logic~\cite{chaudhuri2019hybrid}, 
unlike Brotherston and Villard's \HyBBI, which demonstrably increases expressive power over its base \BBI.

In both \HyLL and \HyBBI, nominals denote worlds with monoidal structure (as worlds in Kripke semantics
for either LL or \BBI necessarily have monoidal structure). Our nominals, by contrast, 
do not name worlds in the same sense with respect to Iris's CMRAs, 
but in fact \emph{classes} of worlds, because the names are locations 
(a means of \emph{selecting} resources) rather than resources.  
A key difference is that the use of nominals in those logics corresponds specifically to hypothetical 
reasoning about resources (until a nominal is connected to a current resource, in which case conclusions 
can be drawn about the current resource), which means the modalities themselves do not ``own'' resources. 
Instead, assertions under our other-space modality can and do
have resource footprints.
Pleasantly, we sidestep most of the metatheoretical complexity of those other substructural hybrid
systems by building our logic within a substructural metatheory (\iris).

\iris has been used to build other logics through pointwise lifting, notably logics that deal with weak
memory models~\cite{dang2019rustbelt,dang2022compass}. Those systems build a derived logic
whose lifting consists of functions from thread-local views of events (an operationalization of the release-acquire + nonatomic
portion of the repaired C11 memory model~\cite{lahav2017repairing}): there modalities $\Delta_\pi(P)$ and $\nabla_\pi(P)$
represent that $P$ held before or will hold after certain memory fence operations by thread $\pi$.
The definitions of those specific modalities existentially quantify over other views, related to the ``current'' view (the one where
the current thread's assertions are evaluated), and evaluate $P$ with respect to those other views. This approach to parameterizing
assertion semantics by a point of evaluation, and evaluating modalized assertions at other points, is what it means
to have a modality at all.
It is \emph{not}, however, an instance of hybrid logic, which is specifically demarcated by an assertion language where
\emph{assertions}, not their semantics, choose and name the evaluation points for modal assertions.
A hybrid extension of the aforementioned logics would include assertions which named specific views at which to evaluate
$P$, in the syntax of the assertion (e.g., $\Delta_\pi^v(P):=\lambda\_\ldotp (P\;v)$) rather than the 
$\Delta_\pi(P):= \lambda v\ldotp (\exists v_{rel}\ldotp \ownGhost{\pi}{\mathsf{RelV}(v_{rel})\;v} \ast (P\;v_{rel})))$ actually used.
Note the hybrid version takes the place to evaluate $P$ as a parameter, and therefore allows the \emph{derived} (modal) logic to explicitly
reason in terms of evaluation points, rather than hiding all points of evaluation in the internal definitions of modalities.

 \section{Conclusions}
This paper advances the state of the art in formal verification of programs subject to address translation
on hardware using virtual memory. We proposed to treat assertions about virtual
memory locations explicitly as assertions in a modal logic, where the notion of context
is the choice of current address space based on the page table root installed on the CPU.
We improved the modularity of our virtual address translation to allow page table modifications that
preserve mappings without collecting all affected virtual points-to assertions.
To make specifications of code involving other address spaces cleaner, we borrowed the idea of
modalities which explicitly name the conditions under which they are true from hybrid logics.
We implemented these ideas in a derived separation logic within Iris, and proved soundness of
the rules for essential memory- and address-space-change-related x86-64 instructions 
sound against a hardware model of 64-bit 4-level address translation.
Finally, we used our rules to verify the correctness of key VMM instruction sequences,
including giving the first assembly-level proof of correctness in a program logic for a change
of address space, against a specification which clearly identifies which assertions hold in which address space.
%
%
%
%

%
%
\bibliographystyle{ACM-Reference-Format}
\bibliography{vmm}

\end{document}

### 2023-08-13 examples.tex
remove line about inconsistency with fpaddr since that was fixed
+

@@
 known-valid page table walks and update the appropriate L1 page table entry.
-\lstset{
-  columns=fullflexible,
-  numbers=left,
-  basicstyle=\ttfamily,
-  keywordstyle=\color{blue}\bfseries,
-  morekeywords={mov,add,call},
-  emph={rsp,rdx,rax,rbx,rbp,rsi,rdi,rcx,r8,r9,r10,r11,r12,r13,r14,r15},
-  emphstyle=\color{green},
-  emph={[2]cr3},
-  emphstyle={[2]\color{violet}},
-  morecomment=[l]{;;},
-  mathescape
-}
+%\lstset{
+%  columns=fullflexible,
+%  numbers=left,
+%  basicstyle=\ttfamily,
+%  keywordstyle=\color{blue}\bfseries,
+%  morekeywords={mov,add,call},
+%  emph={rsp,rdx,rax,rbx,rbp,rsi,rdi,rcx,r8,r9,r10,r11,r12,r13,r14,r15},
+%  emphstyle=\color{green},
+%  emph={[2]cr3},
+%  emphstyle={[2]\color{violet}},
+%  morecomment=[l]{;;},
+%  mathescape
+%}
 \newcommand{\fpaddr}{\texttt{fpaddr}}
@@
 ;; Returns the virtual address of the L1 entry in rax
-mov %r14, %rax ;; Save that before another call
+mov r14, rax ;; Save that before another call
 $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast  \texttt{r14}\mapsto_{\textsf{r}} \texttt{pte\_addr} \ast \texttt{rdi}\mapsto_{\textsf{r}} \vaddr \ast \texttt{rax}\mapsto_{\textsf{r}} \textsf{ pte\_addr} \; \ast }_{\rtv}$
@@
 ;; update the page table entry, mapping the page
-mov (%r14), %rax
+mov [r14], rax
 $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast  \texttt{r14}\mapsto_{\textsf{r}} \texttt{pte\_addr} \ast \texttt{rdi}\mapsto_{\textsf{r}} \vaddr \;\ast}_{\rtv}$
@@
 fpaddr)) in Line 30.
-Reaching right at this point in the specification, we must note that we did find a trivial inconsistency in the postcondition related to conflating fpaddr as an address with the page table entry that reference fpaddr, and while the fix is straightforward we ran out of time.
 
@@
 $\specline{\texttt{rdi+40} \mapsto_{\textsf{v}} \_ \ast \texttt{rdi+48} \mapsto_{\textsf{v}} \_\ast \texttt{rdi+56} \mapsto_{\textsf{v}} \_}_{\rtv}$
-mov 0[%rdi], %rbx
-mov 8[%rdi], %rsp
-mov 16[%rdi], %rbp
-mov 24[%rdi], %r12
-mov 32[%rdi], %r13
-mov 40[%rdi], %r14
-mov 48[%rdi], %r15
+mov 0[rdi], rbx
+mov 8[rdi], rsp
+mov 16[rdi], rbp
+mov 24[rdi], r12
+mov 32[rdi], r13
+mov 40[rdi], r14
+mov 48[rdi], r15
 $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast [\rtv'+56](\mathcal{I}\texttt{ASpace}(\theta,m) \ast \texttt{Pother})}_{\rtv}$
@@
 ;; Restore target context
-mov %rbx, 0[%rsi] 
+mov rbx, 0[rsi] 
 ;; This switches to a new stack
@@
 ;; current address space!
-mov %rsp, 8[%rsi] 
-mov %rbp, 16[%rsi]
-mov %r12, 24[%rsi]
-mov %r13, 32[%rsi]
-mov %r14, 40[%rsi]
-mov %r15, 48[%rsi]
+mov rsp, 8[rsi] 
+mov rbp, 16[rsi]
+mov r12, 24[rsi]
+mov r13, 32[rsi]
+mov r14, 40[rsi]
+mov r15, 48[rsi]
 $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast [\rtv'+56](\mathcal{I}\texttt{ASpace}(\theta,m) \ast \texttt{Pother})}_{\rtv}$
@@
 ;; Switch to the new address space
-mov %cr3, 56[%rsi]
+mov cr3, 56[rsi]
 $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast [\rtv'+56](\mathcal{I}\texttt{ASpace}(\theta,m) \ast \texttt{Pother}) }_{\rtv}$

### 2023-11-03 background.tex
edits through middle of section 4

@@
 
-In typical system configurations, all memory addresses seen by programs running on modern computers are \emph{virtualized}: the address observed by a running program generally will not correspond directly to the physical location in memory, and may not even correspond to a physical location that \emph{exists} in the machine.
-Instead, these \emph{virtual} addresses are translated to \emph{physical} addresses that correspond directly to locations in RAM.
-On most modern architectures, this translation is performed through cooperation of the hardware and OS kernel: while executing an instruction that dereferences a (virtual) address, the CPU's \emph{memory management unit} (\textsc{MMU}) implements an architecture-specific process of \emph{address translation}, resulting in a physical address used to access the cache\footnote{Technically, for performance reasons most caches are indexed with parts of the virtual address, but tagged with the physical data addresses, so cache lookups and address translations can proceed in parallel.} and/or memory-bus.
+In typical system configurations, all memory addresses seen by programs running on modern computers are \emph{virtualized}: the address observed by a running
+program generally will not correspond directly to the physical location in memory, and may not even correspond to a physical location that \emph{exists} 
+in the machine. Instead, these \emph{virtual} addresses are translated to \emph{physical} addresses that correspond directly to locations in RAM. On most 
+modern architectures, this translation is performed through cooperation of the hardware and OS kernel: while executing an instruction that dereferences a 
+(virtual) address, the CPU's \emph{memory management unit} (\textsc{MMU}) hardware performs \emph{address translation},
+resulting in a physical address used to access the cache\footnote{Technically, for performance reasons most caches are indexed with parts of the virtual 
+address, but tagged with the physical data addresses, so cache lookups and address translations can proceed in parallel.} and/or memory-bus.
 
@@
 traverses a series of tables, the translation process or algorithm is sometimes referred to as a \emph{page-table walk}. 
-While Figure \ref{fig:pagetables} and most of our details are specific to the x86-64 architecture, ARMv8 (a.k.a.\ 
+While Figure \ref{fig:pagetables} and most of our constants (how many levels, which virtual address bits index which
+table levels) are specific to the x86-64 architecture, ARMv8 (a.k.a.\ 
 \texttt{aarch64}), RISC-V, and PowerPC use similar hierarchical page tables for address translation.
 
-The entries of each table are 64 bits wide, but each points to a physical address aligned to 4KB (4096 byte) boundaries, which leaves 12 bits to spare to control a validity bit (called the \emph{present} bit), a read-write bit (which permits write access through the entry if and only if it is set), and a range of additional bits which can be used to control caching, write-through, and more.
-This paper will only consider the present bit (0), the read-write bit (1), and the accessed bit (5), which is set to 1 by the hardware every time an entry is used, allowing the OS to determine which pieces of memory are used or unused.
+The entries of each table are 64 bits wide, but each points to a physical address aligned to 4KB (4096 byte) boundaries, which leaves 12 bits to spare to 
+control a validity bit (called the \emph{present} bit), a read-write bit (which permits write access through the entry if and only if it is set), 
+and a range of additional bits which can be used to control caching, write-through, and more. This paper will only consider the present bit (0), 
+the read-write bit (1), and the accessed bit (5), which is set to 1 by the hardware every time an entry is used, allowing the OS to determine which 
+pieces of memory are used or unused.
 
@@
 
-If an instruction is executed that accesses a virtual address that either has no mapping, or does not have a mapping permitting the kind of access that was performed (e.g., the instruction was a memory write, but the relevant address range was marked read-only in the relevant page table entry), the hardware triggers a \emph{page fault}, transferring control to a \emph{page fault handler} registered with the hardware by the OS, allowing it to take corrective action (described in the next section).
+If an instruction is executed that accesses a virtual address that either has no mapping, or does not have a mapping permitting 
+the kind of access that was performed (e.g., the instruction was a memory write, but the relevant address range was marked read-only in
+the relevant page table entry), the hardware triggers a \emph{page fault}, transferring control to a \emph{page fault handler} registered 
+with the hardware by the OS, allowing it to take corrective action (described in the next section).
+
+\paragraph{Out of Scope: Translation Lookaside Buffers (TLBs)}
+CPUs with MMUs typically have an additional \emph{translation lookaside buffer}
+(TLB), to cache the (successful) results of page table walks, rather than merely transforming every virtual
+memory access into 5 physical accesses. 
+Any time a virtual address that was accessible becomes inaccessible (or has downgraded permissions),
+the TLB (or at least entries in affected virtual address ranges) should be flushed.
+In most kernels, this occurs in only a few well-known places, which is why existing verified kernels
+\textsc{seL4}~\cite{Klein2009seL4,seL4TOCS} and \textsc{CertiKOS}~\cite{gu15,gu2016certikos} currently
+trust that TLB flushes are handled correctly rather than actually modeling TLB hardware and verifying.
+Eventual verification of TLB handling is a worthwhile long-term goal, but it is a challenging pursuit in its
+own right. Based on others' early progress on verifying TLB operations in isolation~\cite{syeda2020formal}, we expect
+it to be possible to combine this paper's insights with that support. Even without TLB modeling,
+our logic already enables verification of virtual memory management functionality that prior verified kernel work
+either trusts completely (\textsc{seL4}, \textsc{CertiKOS}) or is incapable of reasoning about.
+Appendix \ref{apdx:tlb} gives more details on the challenges involved for the interested reader.
 
@@
 \begin{itemize}
-  \item as a \textit{context} fact if its validity depends on \texttt{cr3} when it is outside the modality (i.e.
+  \item as a \textit{contextual} fact if its validity depends on \texttt{cr3} when it is outside the modality (i.e.
 modal context), and made \texttt{cr3} independent as a part of custom-tailored modal logic for virtual-memory: a truth representing a virtual-memory addressing depends \texttt{cr3} due to address-translation operation, is dependent on \texttt{cr3} in the ambient-logic (e.g.
@@
 --- unless guarded by another address space modality.
+Some prior \iris-based work~\cite{dang2019rustbelt,dang2022compass} has constructed derived modalities in the style we propose, indexed
+by thread IDs. However in that setting, the intepretation of those modalities was fully fixed ahead of time (to refer to essentially buffers in operationalized versions of C11
+concurrency). In this setting, while our modalities will be indexed by page table roots, it is possible to modify the address translation for an address
+space with root $r$ --- thus changing the interpretation of a modality --- \emph{while assertions with that modality are active}.
 %

### 2023-11-03 intro.tex
edits through middle of section 4

@@
 which were only able to work with a single address space (the current address space on the CPU) because they were unable
-to speak directly \emph{within the logic} about other address spaces.
+to speak directly \emph{within the logic} about other address spaces, in addition to handling
+the \emph{non-local} effects of page table updates whether within the current address space or across address spaces.
 %\begin{figure}
@@
 \item We develop these ideas in the form of a logic for working with virtual-address-space-relative assertions,
-      implemented as an embedded separation logic within Iris~\cite{jung2018iris} separation logic.
+      implemented as an embedded separation logic within the Iris~\cite{jung2018iris} separation logic.
+      The result is a separation logic that lifts a number of major semantic restrictions present in the few
+      prior logics tackling virtual address translation.
 \item We prove the soundness of our \textsf{vProp} logic with respect to a RISC-like fragment of \textsf{AMD64} instructions.
 \item We verify simplified versions of several critical virtual-memory-related pieces of OS functionality, 
-      including mapping  pages, and the assembly code for switching address spaces. 
+      including software page table walks, mapping  pages, and the assembly code for switching address spaces. 
       The latter in particular goes beyond the capabilities of prior work, and reveals unexpected
@@
 \end{itemize}
+
+The logic we develop here covers core reasoning principles for reasoning about memory configurations and code
+reliant upon or manipulating those memory configurations in the presence of in-memory page tables, the primary
+memory protection mechanism across Intel/AMD's x86-64 processors, ARM's application class processors including
+aarch64 CPUs, POWER, RISC-V, and other architectures.
+% One aspect of virtual memory management we do not address in this paper is updates to the translation lookaside
+% buffer (TLB) present in most CPUs. Like CertiKOS~\cite{?}, seL4~\cite{sel4tocs}, and most other OS verification
+% efforts we currently trust insertion of TLB flush operations, whose locations are critical but also few and
+% well-known (after unmapping memory, or changing address spaces). Tackling correctness of TLB invalidation
+% rigorously is an additional challenging project unto itself~\cite{...} with similarities to but
+% significant divergences from reasoning about weak memory models, which future work should combine
+% with the present work.

### 2023-11-03 programlogic.tex
tweaks so latexpand can generate a working tex file for arxiv
edits through middle of section 4

@@
       those on the same page of memory\footnote{of in the case of L2 or higher levels of tables, within a given broader region.}).
-      This sharing imposes constraints on definiting points-to assertions
+      This sharing imposes constraints on defining points-to assertions
       in terms of physical (L4-L1) page-table memory accesses
@@
 \paragraph{Physical memory  points-to} The soundness proofs for our logic's rules largely center around page-table-walk accesses as in Figure \ref{fig:coq_addr_translation},
- where (inside the helper functions) we need to split a 64-bit word representing entries in these tables into 52-bit segments
-which specify the physical page number of another page in the walk, and 12-bit segments containing various control bits.
+ where (inside the helper functions) we need to split a 64-bit address into 52-bits' worth of segments
+that specify the table indices used to locate physical page numbers for successive pages in the walk, and 12-bit segments containing various control bits.
 For this reason we have two notions of physical points-to facts. The primitive notion closest to our machine model is:
@@
 \emph{which} locations are accessed in a page table walk.
-Figure \ref{figure:strongvirtualpointsto} gives such a definition.
+Figure \ref{fig:strongvirtualpointsto} gives such a definition.
 It asserts the existence of for page table entries, one at each translation level, and via \lstinline|L4_L1_PointsTo|
@@
 
+\todo[inline,color=cyan]{Colin edit frontier}
+
 \subsection{Aliasing/Sharing Physical Pages}
@@
     \exists \delta\ldotp
-    \underbrace{\lambda\mathit{cr3val}\ldotp\sumapaces{\mathit{cr3val}}\delta}_\text{\begin{tabular}{c}Find ghost map for this\\address space\end{tabular}} \ast 
+    \underbrace{(\lambda\mathit{cr3val}\ldotp\sumapaces{\mathit{cr3val}}\delta)}_\text{\begin{tabular}{c}Find ghost map for this\\address space\end{tabular}} \ast 
   \underbrace{\sumwalkabs\vaddr\qfrac\paddr }_\text{\begin{tabular}{c}Fractional ownership of \\virt-to-phys translation\end{tabular}}\ast 
-  \underbrace{\paddr \mapsto_{\mathsf{p}} \vale}_\text{\begin{tabular}{c}Value at\\physical location\end{tabular}}
+  \underbrace{\paddr \mapsto_{\mathsf{p}}\{\textsf{q}\}\; \vale}_\text{\begin{tabular}{c}Value at\\physical location\end{tabular}}
 \end{array}
@@
   points-to, and keep the ghost map and the invariant that every mapping in the ghost map has corresponding physical resources
-  for translation in a separate global invariant.
+  for translation in a separate global invariant for each address space.
 
@@
 \label{sec:issues}
-An important subtlety arises with supporting \lstinline|mov|s into \lstinline|%cr3|. Consider the hypothetical rule
+An important subtlety arises with supporting \lstinline|mov|s into \lstinline|cr3|. Consider the hypothetical rule
 \begin{mathpar}
@@
 \end{mathpar}
+\todo[inline]{update for rip tracking}
 \caption{Reasoning Rules for Selected \textsf{AMD64} Instructions under}

### 2023-11-03 relwork.tex
edits through middle of section 4

@@
 Pleasantly, we sidestep most of the metatheoretical complexity of those other substructural hybrid
-systems by building our logic within a substructural metatheory (Iris).
+systems by building our logic within a substructural metatheory (\iris).
+
+\iris has been used to build other logics through pointwise lifting, notably logics that deal with weak
+memory models~\cite{dang2019rustbelt,dang2022compass}. Those systems build a derived logic
+whose lifting consists of functions from thread-local views of events (an operationalization of the release-acquire + nonatomic
+portion of the repaired C11 memory model~\cite{lahav2017repairing}): there modalities $\Delta_\pi(P)$ and $\nabla_\pi(P)$
+represent that $P$ held before or will hold after certain memory fence operations by thread $\pi$.
+The definitions of those specific modalities existentially quantify over other views, related to the ``current'' view (the one where
+the current thread's assertions are evaluated), and evaluate $P$ with respect to those other views. This approach to parameterizing
+assertion semantics by a point of evaluation, and evaluating modalized assertions at other points, is what it means
+to have a modality at all.
+It is \emph{not}, however, an instance of hybrid logic, which is specifically demarcated by an assertion language where
+\emph{assertions}, not their semantics, choose and name the evaluation points for modal assertions.
+A hybrid extension of the aforementioned logics would include assertions which named specific views at which to evaluate
+$P$, in the syntax of the assertion (e.g., $\Delta_\pi^v(P):=\lambda\_\ldotp (P\;v)$) rather than the 
+$\Delta_\pi(P):= \lambda v\ldotp (\exists v_{rel}\ldotp \ownGhost{\pi}{\mathsf{RelV}(v_{rel})\;v} \ast (P\;v_{rel})))$ actually used.
+Note the hybrid version takes the place to evaluate $P$ as a parameter, and therefore allows the \emph{derived} (modal) logic to explicitly
+reason in terms of evaluation points, rather than hiding all points of evaluation in the internal definitions of modalities.
+This prior \iris-based work also uses modalities where the interpretations are fixed a priori by logic designers. In contrast, our
+address space modalities' interpretations can be changed by program behaviors via page table updates; the equivalent in the prior work would
+be if programs could directly manipulate the buffers used to model weak memory behaviors,
+which they cannot.\footnote{Note that the views modelled modally in this prior work are abstractions of a wide range of hardware, so such
+manipulation of corresponding hardware resources is in fact impossible,
+while the address space mappings in our work reflect known hardware components present in a variety of CPUs.}

### 2023-11-03 syntaxsemantics.tex
edits through middle of section 4

@@
 A register identifier, $\reg$, is chosen from a fixed finite set of register identifiers, $\regset$. 
-We use these identifiers $\reg$ for register names such as \texttt{rax}, \texttt{r8}, or \texttt{cr3}.
+We use these identifiers $\reg$ for register names such as \texttt{rax}, \texttt{r8}, or \texttt{cr3}. Our model includes
+all x86-64 integer registers (including stack and instruction pointers), as well as \texttt{cr3} (for page table roots) and \texttt{rflags} (for
+flags set by comparison operations and inspected by conditional jumps).
 % \begin{figure}[t]
@@
 %\end{itemize}
+
+In addition to \lstinline|mov| instructions involving 64-bit registers, our formalization includes basic integer and bitwise operations (\lstinline|add|,
+\lstinline|and|, bit shifts, etc.) with their effects on \lstinline|rflags|, jumps and some (not all) conditional jumps, \lstinline|call|, \lstinline|ret|, \lstinline|push|,
+and \lstinline|pop|.

### 2023-11-03 examples.tex
fix rendering of figures 11 and 12 (mathescape)
outlining updates for page table walk example

@@
 \begin{figure}\footnotesize
-  \begin{lstlisting}
+  \begin{lstlisting}[mathescape]
 $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m)  \ast  \texttt{r14}\mapsto_{\textsf{r}} \_ \ast \texttt{rdi}\mapsto_{r} \vaddr \ast \texttt{rax}\mapsto_{\textsf{r}} \_ \ast \ulcorner \texttt{aligned va} \land  \theta \; !!\; \vaddr = \texttt{None}\urcorner}_{\rtv}$
@@
 
+\subsection{Traversing Live Page Tables}
+The mapping operation of Figure \ref{fig:mapping_code} assumes an operation \lstinline|ensure_L1| which must traverse the page tables
+in order to locate the address of the L1 entry to update --- possibly allocating pages at levels 3 and 2 in the process.
+Traversing the page tables is itself challenging functionality to verify: loading the current table root from \lstinline|cr3| is straightforward
+(a \lstinline|mov| instruction), however this produces the physical address of \lstinline|cr3|, not the virtual address the kernel code would use to access that memory.
+This problem repeats at each level of the page table: assuming the code has \emph{somehow} read the appropriate L4 (or L3, or L2) entry, those entries again
+yield physical addresses, not virtual.
+Most kernels maintain an invariant on their page tables that the virtual address of any page used for a page table lives at a virtual address
+whose value is a constant offset from the physical address a practice sometimes referred to as \emph{identity mapping} (even though the physical-to-virtual translation
+is typically not literally the identity function).\footnote{Some kernels do this for all physical memory on the machine, simplifying interaction
+with DMA devices.}
+We have verified\todo{todo!} the page table walk corresponding to \lstinline|ensure_L1| under this constant-offset assumption (including allocating missing page table levels).
+
+The critical challenges of verifying this code relate to both proving memory safety of the traversal, and peeling out enough resources from the per-address-space
+invariant to construct the pte-points-to used in Figure \ref{fig:mapping_code}.
+This is not possible to do modularly (without knowledge of other address spaces' details) with prior work.
+\citet{kolanski08vstte} could reason about the walk, but would need to explicitly prove that all other invariants
+of the kernel, the current address space, and all other address spaces of interest were preserved by each update, because their model
+only supports separation within a single address space. In our model, this follows for free from making
+our separation logic directly aware of address translation and internalizing assumptions about other address spaces as further separable assertions.
+\textsc{seL4} currently still trusts address translations; it models page tables as a data structure in regular memory, thus not capturing the possibility that even
+temporarily destroying the mappings and restoring them can actually crash the OS. \textsc{CertiKOS} \todo{hard to word since they intentionally don't publicize any
+details}
+
+A detailed proof outline and discussion of this code is given in Appendix \ref{apdx:tablewalk}.\todo{todo}
+
 \subsection{Change of Address Space}
@@
 \begin{figure}\footnotesize
-\begin{lstlisting}
+\begin{lstlisting}[mathescape]
 ;; Assume the save-space is in rdi,

### 2023-11-04 examples.tex
more outlining of distinctions over prior work

@@
 
+Unlike the only prior work verifying analogous code for mapping a new page~\cite{kolanski08vstte,kolanski09tphols}, our proof above
+does \emph{not} need to unfold the definitions of derived logic (therefore reasoning directly over the operational semantics),
+but instead uses derived logic rules throughout to reason about the effect of the mapping. By incorporating verification of the
+\lstinline|ensure_L1| function (see Section \ref{sec:traversing}), our verification also directly handles several subtle aspects which
+were axiomatized in prior work.
+
 \subsection{Unmapping a Page}
@@
 \subsection{Traversing Live Page Tables}
+\label{sec:traversing}
 The mapping operation of Figure \ref{fig:mapping_code} assumes an operation \lstinline|ensure_L1| which must traverse the page tables
@@
 We have verified\todo{todo!} the page table walk corresponding to \lstinline|ensure_L1| under this constant-offset assumption (including allocating missing page table levels).
+A detailed proof outline and discussion of this code is given in Appendix \ref{apdx:tablewalk}.\todo{todo}
 
-The critical challenges of verifying this code relate to both proving memory safety of the traversal, and peeling out enough resources from the per-address-space
-invariant to construct the pte-points-to used in Figure \ref{fig:mapping_code}.
-This is not possible to do modularly (without knowledge of other address spaces' details) with prior work.
-\citet{kolanski08vstte} could reason about the walk, but would need to explicitly prove that all other invariants
+The critical challenges of verifying this code relate to both proving memory safety of the traversal --- which requires reasoning accurately about 
+computing a virtual address for accessing a page table from its physical address --- and peeling out enough resources from the per-address-space
+invariant to construct the pte-points-to used in Figure \ref{fig:mapping_code}. Prior work cannot do this.
+
+\citet{kolanski08vstte,kolanski09tphols} verified a single code block with their logic which was roughly Figure \ref{fig:mapping_code} for a 2-level ARM
+page table, but several critical complexities our work deals with were not addressed.
+First, beyond the limitations discussed in Section \ref{sec:overly-restrictive}, Kolanski and Klein assumed that virtual addresses
+for page tables at each level were given as parameters rather than verifying the conversion from physical addresses to virtual addresses (or even axiomatizing their lookup).
+In contrast, our verification articulates the address space invariant from which the physical-to-virtual translation can be implemented.
+Second, our proof deals with the construction of a valid virtual points-to \emph{to the PTE to update} in mapping, which Kolanski and Klein also
+assumed was given.
+
+Reasoning about the page table walk in their logic would have required 
+could reason about the walk, but would need to explicitly prove that all other invariants
 of the kernel, the current address space, and all other address spaces of interest were preserved by each update, because their model
@@
 our separation logic directly aware of address translation and internalizing assumptions about other address spaces as further separable assertions.
+Kolanski and Klein did address part of the walk information for a 2-level page table (a possible ARM configuration), but
+
 \textsc{seL4} currently still trusts address translations; it models page tables as a data structure in regular memory, thus not capturing the possibility that even
@@
 details}
-
-A detailed proof outline and discussion of this code is given in Appendix \ref{apdx:tablewalk}.\todo{todo}
 

### 2023-11-10 examples.tex
adjust for focusing on P2V

@@
 %$\assert{\ulcorner \ptablestore !! \textsf{maddr} = \textsf{None} \urcorner \ast \ownGhost\gammaPred{\authfull{\ptableabswalk\ptablestore}} \ast \textsf{Pf}} $
-\section{Experiment}
+\section{Experiments}
 \label{sec:experiment}
@@
 yield physical addresses, not virtual.
+
 Most kernels maintain an invariant on their page tables that the virtual address of any page used for a page table lives at a virtual address
 whose value is a constant offset from the physical address a practice sometimes referred to as \emph{identity mapping} (even though the physical-to-virtual translation
-is typically not literally the identity function).\footnote{Some kernels do this for all physical memory on the machine, simplifying interaction
+is typically not literally the identity function, but adding a constant offset).\footnote{Some kernels do this for all physical memory on the machine, simplifying interaction
 with DMA devices.}
-We have verified\todo{todo!} the page table walk corresponding to \lstinline|ensure_L1| under this constant-offset assumption (including allocating missing page table levels).
-A detailed proof outline and discussion of this code is given in Appendix \ref{apdx:tablewalk}.\todo{todo}
 
-The critical challenges of verifying this code relate to both proving memory safety of the traversal --- which requires reasoning accurately about 
-computing a virtual address for accessing a page table from its physical address --- and peeling out enough resources from the per-address-space
-invariant to construct the pte-points-to used in Figure \ref{fig:mapping_code}. Prior work cannot do this.
+Work to verify the full page table walk is ongoing, but we have verified the key step for a small x86-64 kernel, which is the physical-to-virtual
+conversion (often appearing as a macro \texttt{P2V} in C source code). In our small kernel, as in larger kernels, \texttt{P2V} is actually just addition,
+though the Hoare triple we verify for this add instruction is rather complex:
+
+\todo[inline]{typeset as a triple}
 
@@
 First, beyond the limitations discussed in Section \ref{sec:overly-restrictive}, Kolanski and Klein assumed that virtual addresses
-for page tables at each level were given as parameters rather than verifying the conversion from physical addresses to virtual addresses (or even axiomatizing their lookup).
+for page tables at each level were given as parameters rather than verifying any conversion from physical addresses to virtual addresses (or even axiomatizing their lookup).
 In contrast, our verification articulates the address space invariant from which the physical-to-virtual translation can be implemented.
@@
 assumed was given.
+\todo{some of this is really an argument for our verification being more thorough, rather than being about our logic}
 
@@
 \textsc{seL4} currently still trusts address translations; it models page tables as a data structure in regular memory, thus not capturing the possibility that even
-temporarily destroying the mappings and restoring them can actually crash the OS. \textsc{CertiKOS} \todo{hard to word since they intentionally don't publicize any
-details}
+temporarily destroying the mappings and restoring them can actually crash the OS. \textsc{CertiKOS} papers share little in the way of precise details about
+their virtual memory management, but because their core technology is based on a fork of \textsc{CompCert}, whose model of memory is
+a set of unordered block allocations, we can infer their proofs must also trust these translations.
 

### 2023-11-13 background.tex
revisions / compression up through end of 4.1

@@
     \includegraphics[width=\columnwidth]{pagetables.png}
+    \todo[inline]{We should try to produce a more compact version of this diagram}
     \caption{x86-64 page table lookups.}
@@
 address translation proceeds by repeatedly taking designated slices of the virtual address and indexing into a table.
-The final lookup in the page tables gives the base physical address of a page of physical memory, to which the low-order
+The final lookup in the page tables gives the base physical address of a 4KB page of physical memory, to which the low-order
 bits of the accessed virtual address are added to determine the actual physical address retrieved. 
@@
 \texttt{aarch64}), RISC-V, and PowerPC use similar hierarchical page tables for address translation.
+RISC-V's \texttt{sv48} paging configuration~\cite[\S 4.5]{riscv-privileged} and AArch64's 4-level paging configuration
+both split 64-bit virtual addresses at the same points to index into the respective tables, so most of what follows
+is equally applicable to those architectures (the only difference is that page table entries place control bits of write access, etc.,
+in different orders).
 
@@
 control a validity bit (called the \emph{present} bit), a read-write bit (which permits write access through the entry if and only if it is set), 
-and a range of additional bits which can be used to control caching, write-through, and more. This paper will only consider the present bit (0), 
-the read-write bit (1), and the accessed bit (5), which is set to 1 by the hardware every time an entry is used, allowing the OS to determine which 
-pieces of memory are used or unused.
+and a range of additional bits which can be used to control caching, write-through, and more. This paper will only consider the present bit (0).
 
@@
 other worlds. This explicit naming of alternate worlds increases the power of modal logics~\cite{blackburn1995hybrid}, and is actually
-necessary for completeness in classical separation logics~\cite{brotherston2014parametric}.
+necessary for completeness in classical separation logics~\cite{brotherston2014parametric}.  A key distinguishing feature of
+hybrid logic is that the logic itself does not fix the set of modalities, but is defined relative to any set of sensible worlds.
 
@@
 the notion that we can represent contingent truth of an assertion via $[r](P)$ indicating that $P$ holds in the address 
-space rooted at physical address $r$.
+space rooted at physical address $r$. Because OS kernels create and destroy address spaces, it is sensible to use
+a hybrid-style logic that does not care about the exact set of modalities.
+
 The typical hybrid modality introduction rule, that $P$ and $\ell$ (indicating the current world is $\ell$) imply 
 $[\ell](P)$, has a natural analogue: knowing $P$ and that the current address space is $r$ (i.e., that $r$ is the 
-current \texttt{cr3} value) suggests a way to construct $[r](P)$. By the choice of \texttt{cr3} identifying the contingency, we can identify a fact
-\begin{itemize}
-  \item as a \textit{contextual} fact if its validity depends on \texttt{cr3} when it is outside the modality (i.e.
-modal context), and made \texttt{cr3} independent as a part of custom-tailored modal logic for virtual-memory: a truth representing a virtual-memory addressing depends \texttt{cr3} due to address-translation operation, is dependent on \texttt{cr3} in the ambient-logic (e.g.
-separation-logic), and is made \textit{independent} of the facts related to \texttt{cr3} by being introduced into the modal context under the assumption that it exhibits the knowledge on its validity with respect to \texttt{cr3} in the ambient logic.
-  \item as a \textit{pure} fact as long as it does not \textit{necessarily} depend on any fact related to \texttt{cr3}: a truth representing a raw physical memory addressing, unlike virtual-address translation, does not need the \textit{knowledge} of \texttt{cr3}, therefore it can be introduce into the modal context as a pure fact
-\end{itemize}
+current \texttt{cr3} value) suggests a way to construct $[r](P)$. 
+We identify an assertion as \emph{contextual} if its validity depends on the choice of address space. Used outside an explicit
+modality, the truth of this assertion depends on the current \texttt{cr3} value. Used under an explicit address space
+modality, its truth depends on both the modality-chosen address space and associated physical resources.
+% \begin{itemize}
+%   \item as a \textit{contextual} fact if its validity depends on \texttt{cr3} when it is outside the modality (i.e. modal context), and made 
+%   \texttt{cr3} independent as a part of custom-tailored modal logic for virtual-memory: a truth representing a virtual-memory addressing depends
+%    \texttt{cr3} due to address-translation operation, is dependent on \texttt{cr3} in the ambient-logic (e.g. separation-logic),
+%     and is made \textit{independent} of the facts related to \texttt{cr3} by being introduced into the modal context under the assumption that
+%     it exhibits the knowledge on its validity with respect to \texttt{cr3} in the ambient logic.
+%   \item as a \textit{pure} fact as long as it does not \textit{necessarily} depend on any fact related to \texttt{cr3}: a truth representing a raw physical memory addressing, unlike virtual-address translation, does not need the \textit{knowledge} of \texttt{cr3}, therefore it can be introduce into the modal context as a pure fact
+% \end{itemize}
 
@@
 concurrency). In this setting, while our modalities will be indexed by page table roots, it is possible to modify the address translation for an address
-space with root $r$ --- thus changing the interpretation of a modality --- \emph{while assertions with that modality are active}.
+space with root $r$ --- thus changing the interpretation of a modality, and even whether a modality is valid --- \emph{while assertions with that modality are active}.
 %

### 2023-11-13 programlogic.tex
revisions / compression up through end of 4.1

@@
 Therefore, as a first step, we work entirely in a pointwise lifting of Iris's base BI logic,
-essentially working with separation logic assertions indexed by a page table root as a $\mathcal{W}_{64}$
-(Figure \ref{fig:vprop}), which we call $\textsf{vProp }\Sigma$; this is the (\textsc{Coq}) type of assertions in our logic.
+essentially working with separation logic assertions indexed by a choice of page table root as a $\mathcal{W}_{64}$, which we call $\textsf{vProp }\Sigma$:\footnote{
+  \iris experts may notice our \lstinline|-b>| resembles another variety of pointwise lifting that already exists in the main \iris
+  distribution~\cite{dang2019rustbelt,dang2022compass}. This is essentially true, but the standard formalization does not work for index types which
+  are indexed in the sense of \coq's metatheory, as our \lstinline|word n| types are.
+}
+\begin{lstlisting}[language=Coq]
+  Definition vProp  $\Sigma$ : bi := word 64 -b> iPropI  $\Sigma$.
+\end{lstlisting}
+This is the (\textsc{Coq}) type of assertions in our logic.
 Most constructs in Iris's base logic are defined with respect to any BI-algebra (of \textsc{Coq} type \lstinline|bi|), so automatically
@@
 
-\begin{figure}[t]
-\begin{lstlisting}[language=Coq]
-  Definition vProp  $\Sigma$ : bi := word 64 -b> iPropI  $\Sigma$.
-  \end{lstlisting}
-\caption{$\textsf{vProp }\Sigma$: Root-Address Indexed Address-Space Proposition}
-  \label{fig:vprop}
-\end{figure}
+% \begin{figure}[t]
+% \begin{lstlisting}[language=Coq]
+%   Definition vProp  $\Sigma$ : bi := word 64 -b> iPropI  $\Sigma$.
+%   \end{lstlisting}
+% \caption{$\textsf{vProp }\Sigma$: Root-Address Indexed Address-Space Proposition}
+%   \label{fig:vprop}
+% \end{figure}
 
@@
 that specify the table indices used to locate physical page numbers for successive pages in the walk, and 12-bit segments containing various control bits.
-For this reason we have two notions of physical points-to facts. The primitive notion closest to our machine model is:
-\[ \textsf{w1} \ \sim \ \textsf{w2} \mapsto_{\textsf{a}} \; \{\textsf{q}\} \; \textsf{v} \]
-In this form, we see that \textsf{w1} (a $\mathcal{W}_{52}$) essentially selects a 4KB page of physical memory,
-and \textsf{w2} (a $\mathcal{W}_{12}$) is an offset within that page.
+For this reason we have two notions of physical points-to facts. The primitive notion closest to our machine model is captured by an assertion
+$ \textsf{pfn} \ \sim \ \textsf{pageoff} \mapsto_{\textsf{a}} \; \{\textsf{q}\} \; \textsf{v} $, where \textsf{pfn} (a $\mathcal{W}_{52}$ \emph{page frame number}) essentially selects a 4KB page of physical memory,
+and \textsf{pageoff} (a $\mathcal{W}_{12}$) is an offset within that page.
 % could be an 52-bits masked address to level 4 table 
@@
 % Giving a raw 64-bits memory pointsto assertion becomes
-\[\textsf{w} \mapsto_{\textsf{p}} \{q\} \textsf{ v} \stackrel{\triangle}{=} (\textsf{drop 12}~w) \ \sim \ (\textsf{bottom 12}~w)\mapsto_{\textsf{a}} \textsf{ v} \]
+\[\textsf{w} \mapsto_{\textsf{p}} \{q\} \textsf{ v} \stackrel{\triangle}{=} (\textsf{drop 12}~w) \ \sim \ (\textsf{bottom 12}~w)\mapsto_{\textsf{a}} \; \{\textsf{q}\} \textsf{ v} \]
 
@@
   $\exists_{\textsf{l4e l3e l2e l1e}} \ldotp$  $\ulcorner$ aligned $\vaddr \urcorner \ast$ L4_L1_PointsTo($\vaddr$ l4e l3e l2e l1e paddr) $\ast$ paddr $\mapsto_{p}\{\mathsf{q}\} \vpage$.
- Definition L4_L1_PointsTo (maddr l4e l3e l2e l1e paddr :word 64) : vProp $\Sigma$ :=
- $\lambda$ cr3val.
+ Definition L4_L1_PointsTo (maddr l4e l3e l2e l1e paddr :word 64) : vProp $\Sigma$ := $\lambda$ cr3val.
   $\ulcorner$ entry_present l4e $\land$ entry_present l3e $\land$ entry_present l2e $\land$ entry_present l1e$\urcorner$ $\ast$

### 2023-11-13 syntaxsemantics.tex
revisions / compression up through end of 4.1

@@
 For example, $\kw{w}_{12}$ is a 12-bit word, which can be obtained for example truncating away 52 bits of a 64-bit word ($\kw{w}_{64}$).
+
 \subsection{State}
@@
 to 64-bit physical memory values. 
-Thus our states $\sigma$ include:
-%Together with default \textsf{CPU} instantiation, we end up having our state with following pieces:
-\begin{itemize}
-%\item A constant cpu initialized with default value: $\sigma.\mathcal{C}$
-\item A register map: $\sigma.\mathcal{R}: \kw{greg} \rightarrow_{\textrm{fin}} \kw{regval} $
-\item A memory map: $\sigma.\mathcal{M}: \Locft \rightharpoonup_{\textrm{fin}} (\Loctw \rightharpoonup_{\textrm{fin}} \Locsf )$
-\end{itemize}
+Thus our states $\sigma$ include register maps $\sigma.\mathcal{R}: \kw{greg} \rightarrow_{\textrm{fin}} \kw{regval} $ and
+memory maps $\sigma.\mathcal{M}: \Locft \rightharpoonup_{\textrm{fin}} (\Loctw \rightharpoonup_{\textrm{fin}} \Locsf )$.
 Of particular note, \texttt{cr3}, the page table register, is included in the machine state.
@@
 \begin{lstlisting}[language=Coq,mathescape]
- Definition translate (rtv: word 64) ($\sigma$: state) (w: word 64): word 64 + MemFail :=
-      l4e $\leftarrow$ translate_top_level $\sigma.\mathcal{M}$ rtv w; 
-      l3e $\leftarrow$ translate_from_l4e $\sigma.\mathcal{M}$ l4e w; 
-      l2e $\leftarrow$ translate_from_l3e $\sigma.\mathcal{M}$ l3e w;
-      l1e $\leftarrow$ translate_from_l2e $\sigma.\mathcal{M}$ l2e w;
-      addr $\leftarrow$ translate_from_l1e $\sigma.\mathcal{M}$ l1e w;
+ Definition translate (rtv: word 64) ($\sigma$: state) (vaddr: word 64): word 64 + MemFail :=
+      l4e $\leftarrow$ translate_top_level $\sigma.\mathcal{M}$ rtv vaddr;
+      l3e $\leftarrow$ translate_from_l4e $\sigma.\mathcal{M}$ l4e vaddr;
+      l2e $\leftarrow$ translate_from_l3e $\sigma.\mathcal{M}$ l3e vaddr;
+      l1e $\leftarrow$ translate_from_l2e $\sigma.\mathcal{M}$ l2e vaddr;
+      addr $\leftarrow$ translate_from_l1e $\sigma.\mathcal{M}$ l1e vaddr;
       return concat (shift_dropn addr 3 three_lt64) low3.
@@
 
-The main instructions we implement are variants of the x86-64 \lstinline|mov| instruction, which we format in Intel syntax (destination on the left, source on the right).
-As expected \textsf{store} would be the update on the memory map entry
-% After defining the translation in terms of physical load and store action, now we can define our memory \lstinline|mov| instructions
-For memory operations, we treat
+The most important instructions we model are memory-accessing variants of the x86-64 \lstinline|mov| instruction, which we format in Intel syntax
+(destination on the left, source on the right):
 \[
@@
   \]
-which are realized by first translating the virtual-memory address of a memory location stored in $r_m$ to a physical memory location,
-then updating those memory contents with the contents of register $r_r$ (\TirNameStyle{WriteToMemFromReg}) 
+The semantic of each is realized by first translating the virtual-memory address of a memory location stored in $r_m$ to a physical memory location,
+then either updating those memory contents with the contents of register $r_r$ (\TirNameStyle{WriteToMemFromReg}) 
 or loading the value at that physical memory location into the register $r_r$
 (\TirNameStyle{WriteToRegFromMem}). 
-% More \lstinline|mov| instructions operating with memory addresses can be found in the submission artifact for different addressing schemes and data sizes.
-Additional variants (such as instructions which access memory at a constant offset from the virtual address stored in a register) are also treated.
-
-For register operations, we support moves back and forth between any pair of registers, though the \coq formalization splits these according to whether
-the source or destination registers are standard data registers (e.g., \lstinline|rax|) or control registers (e.g., \lstinline|cr3|).
-% \[
-% \begin{array}{l}
-%   \hbox{(\TirNameStyle{WriteToRegFromReg})} \qquad 
-%    \textsf{mov ireg64 ireg64} \\
-%   \hbox{(\TirNameStyle{WriteToCtlRegFromReg})} \qquad 
-%   \textsf{mov creg64 ireg64}\\
-%     \hbox{(\TirNameStyle{WriteToRegFromCtlReg})} \qquad 
-%   \textsf{mov ireg64 creg64}
-%   \end{array}
-% \]
-% for different types of registers (e.g. control and general).
-%
-%\subsection{Machine State under Address Translation}
-%\label{sec:selectedinstrsemantics}
-%Althought we give the complete set of operational semantics rule in \sref{appendix:movops}, It is worth building the intuition on the the way some of these rules bahave in the context of address translation.
-In fact $\readlval\maddr{\storememstar\crval}\locsf$ is unfoled into multiple physical memory lookups for the final page address retrieval -- i.e.
-address translation traversal as shown in Figure \ref{fig:pagetables}:
-%\begin{itemize}
-%\item top-level-address translation: with the given root address (64-bit $\crval$) of the address space, performs the address translation, handling the first level (to get the PML4 entry) itself.
-The next level table address is computed with the fetched PML4 offset value which exhibits itself as 9-bit offset in $\kw{maddr}$.
-\mytodo{iso: put move}
-%\item translating from PML4 entry: performs the second level of address translation, to retrieve starting at the PML4 table entry, and interprets the PML4 entry that references a Page-Directory-Pointer Table (PDPT).
-We obtain the PDPT offset which exhibits itself as 9 bit offset in $\kw{maddr}$ to obtain the address of the next level page directory table (PDT) \mytodo{iso: put move}
-%\item translating from PD entry: performs the third level of address translation, to retrieve starting at the PDP table entry, and interprets a PDPTE that references a PD table.
-Likewise, we obtain the PD offset which exhibits itself as 9 bit offset in $\kw{maddr}$ to obtain the address of the next level page directory table (PT) \mytodo{iso:put move }
-%\item translate from PT entry: performs final level of address translation, starting from the PT entry, with a given 12 bit page offset, we can compute the physical address referencing $\locsf$ \mytodo{ismo: put mov}
-%\end{itemize}
-
-In addition to \lstinline|mov| instructions involving 64-bit registers, our formalization includes basic integer and bitwise operations (\lstinline|add|,
+Additional \lstinline|mov| variants (e.g., accessing memory at constant offsets from the base register, or moves between registers) are also treated.
+Our formalization also includes basic integer and bitwise operations (\lstinline|add|,
 \lstinline|and|, bit shifts, etc.) with their effects on \lstinline|rflags|, jumps and some (not all) conditional jumps, \lstinline|call|, \lstinline|ret|, \lstinline|push|,

### 2023-11-14 programlogic.tex
more compression and editing, through middle of section 4

@@
 
-We derive a program logic (a separation logic) supporting the following stances and constraints:
-\begin{enumerate}
-\item \textit{address-spaces as modal contexts}: Assertions in our logic are context-dependent,
-  in the sense that their truth depends on which address space they are used in, due to the need to support virtual points-to assertions.
-\item \textit{sharing}: The physical location backing a virtual address's storage is located (during a page table walk) through a 
-      set of physical page-table (L4-L1 page-tables) acceses that are shared amongst different virtual addresses (specifically,
-      those on the same page of memory\footnote{of in the case of L2 or higher levels of tables, within a given broader region.}).
-      This sharing imposes constraints on defining points-to assertions
-      in terms of physical (L4-L1) page-table memory accesses
-\item \textit{context-agnostic-resources}: each virtual address is valid under a certain address-space, 
-      but it does not represent this \textit{knowledge} of its address-space. 
-      That is, assertions are not explicitly tagged with their address space validity
-\item \textit{updating address-space mappings}: We present logical abstractions to enable 
-      updating not only pages of typical data in memory, but also page tables themselves.\footnote{Prior work relied on unfolding operational semantics
-      to verify page table updates.}
-\item \textit{explicitly-modal assertions}: Our logic includes a means to talk about facts being true
-      in another address space
-\item \textit{address-space switch as changing the "World" of truth}: Switching from one address-space to another logically
-      becomes a simultaneous introduction-and-elimination of a pair of modal assertions (for different address spaces)
-\end{enumerate}
+We describe a program logic (a separation logic) along the lines suggested earlier, where every assertion is relative
+to an address space in which it is interpreted, allowing us to define \emph{virtual points-to} assertions that make claims
+about memory locations in a particular address space. Virtual addresses, and even virtual points-to assertions, 
+are not tagged with their address spaces in any way. Memory access in this logic is validated through the use 
+of virtual points-to assertions in pre-conditions, which guarantee that address translations succeed.
+This supports rules for updating not only typical data in memory that happens to be subject to address translation, but
+manipulation of the page tables themselves via virtual addresses (as demanded by all modern hardware), also
+via virtual points-to assertions.
+To support specifications that deal with multiple address spaces, our logic incorporates a hybrid-style modality
+$[r](P)$ to state that an assertion is true in another (assertion-specified) address space rather than the address space
+currently active in hardware, which is not only useful for virtual memory manager invariants but critical to reasoning
+about change-of-address-space.
+By developing this within the \iris framework, we obtain additional features (e.g., fractional permissions) that allow us to verify
+some of the most subtle and technically challenging instruction sequences in an OS kernel (Section \ref{sec:experiment}).
 
-The first viewpoint, that the truth of an assertion is relative to an address space, has far-reaching consequences.
-Therefore, as a first step, we work entirely in a pointwise lifting of Iris's base BI logic,
+% We derive a program logic (a separation logic) supporting the following stances and constraints:
+% \begin{enumerate}
+% \item \textit{address-spaces as modal contexts}: Assertions in our logic are context-dependent,
+%   in the sense that their truth depends on which address space they are used in, due to the need to support virtual points-to assertions.
+% \item \textit{sharing}: The physical location backing a virtual address's storage is located (during a page table walk) through a 
+%       set of physical page-table (L4-L1 page-tables) acceses that are shared amongst different virtual addresses (specifically,
+%       those on the same page of memory\footnote{of in the case of L2 or higher levels of tables, within a given broader region.}).
+%       This sharing imposes constraints on defining points-to assertions
+%       in terms of physical (L4-L1) page-table memory accesses
+% \item \textit{context-agnostic-resources}: each virtual address is valid under a certain address-space, 
+%       but it does not represent this \textit{knowledge} of its address-space. 
+%       That is, assertions are not explicitly tagged with their address space validity
+% \item \textit{updating address-space mappings}: We present logical abstractions to enable 
+%       updating not only pages of typical data in memory, but also page tables themselves.\footnote{Prior work relied on unfolding operational semantics
+%       to verify page table updates.}
+% \item \textit{explicitly-modal assertions}: Our logic includes a means to talk about facts being true
+%       in another address space
+% \item \textit{address-space switch as changing the "World" of truth}: Switching from one address-space to another logically
+%       becomes a simultaneous introduction-and-elimination of a pair of modal assertions (for different address spaces)
+% \end{enumerate}
+
+% The idea that the truth of an assertion is relative to an address space has far-reaching consequences.
+To support making assertions depend on a choice of address space, we work entirely in a pointwise lifting of Iris's base BI logic,
 essentially working with separation logic assertions indexed by a choice of page table root as a $\mathcal{W}_{64}$, which we call $\textsf{vProp }\Sigma$:\footnote{
@@
 
-\paragraph{Physical memory  points-to} The soundness proofs for our logic's rules largely center around page-table-walk accesses as in Figure \ref{fig:coq_addr_translation},
- where (inside the helper functions) we need to split a 64-bit address into 52-bits' worth of segments
-that specify the table indices used to locate physical page numbers for successive pages in the walk, and 12-bit segments containing various control bits.
-For this reason we have two notions of physical points-to facts. The primitive notion closest to our machine model is captured by an assertion
+\paragraph{Physical memory  points-to} The soundness proofs for our logic's rules largely center around 
+proving that page-table-walk accesses as in Figure \ref{fig:pagetables} succeed, which requires assertions
+dealing with physical memory locations.
+We have two notions of physical points-to facts. The primitive notion closest to our machine model is captured by an assertion
 $ \textsf{pfn} \ \sim \ \textsf{pageoff} \mapsto_{\textsf{a}} \; \{\textsf{q}\} \; \textsf{v} $, where \textsf{pfn} (a $\mathcal{W}_{52}$ \emph{page frame number}) essentially selects a 4KB page of physical memory,
@@
 
-Figure \ref{fig:logicaladdrspace} depicts the relationships between these pieces, for a single
-address space.
-The invariant for the currently active address space (solid round square labelled as 
-$\mathcal{I}$\textsf{ASpace} in Figure \ref{fig:logicaladdrspace}) asserts authorative ownership (i.e. 
-update capability), $\ownGhost\gammaPred{\authfull{\ptableabswalk\ptablestore}}$, of the ghost map $\Theta$
-describing virtual-to-physical address translations
--- shown as a solid-round-head arrow to $\mathcal{A}\textsf{PTableWalk}$.
-Each virtual points-to assertion (dashed round square labelled as \textsf{Virtual PointsTo}
- in Figure \ref{fig:peraspaceinvariant})
-asserts fragmental ownership of the ghost map $\Theta$, $\sumwalkabs\vaddr\qfrac\paddr$,
-shown via the dashed-diamond-head arrow to $\mathcal{A}\textsf{PTableWalk}$. Specifically, it asserts that
-the virtual address the virtual points-to assertion concerns corresponds to some physical address.
-There exists only one persistent authorative ownership 
-($\ownGhost\gammaPred{\authfull{\ptableabswalk\ptablestore}}$) which collects the fragmental 
-ghost page-table-walk mapping ($\sumwalkabs\vaddr\qfrac\paddr$),
- shown as double-ended solid arrow 1 to N imposing the update protocol.
+% Figure \ref{fig:logicaladdrspace} depicts the relationships between these pieces, for a single
+% address space.
+% The invariant for the currently active address space (solid round square labelled as 
+% $\mathcal{I}$\textsf{ASpace} in Figure \ref{fig:logicaladdrspace}) asserts authorative ownership (i.e. 
+% update capability), $\ownGhost\gammaPred{\authfull{\ptableabswalk\ptablestore}}$, of the ghost map $\Theta$
+% describing virtual-to-physical address translations
+% -- shown as a solid-round-head arrow to $\mathcal{A}\textsf{PTableWalk}$.
+% Each virtual points-to assertion (dashed round square labelled as \textsf{Virtual PointsTo}
+%  in Figure \ref{fig:peraspaceinvariant})
+% asserts fragmental ownership of the ghost map $\Theta$, $\sumwalkabs\vaddr\qfrac\paddr$,
+% shown via the dashed-diamond-head arrow to $\mathcal{A}\textsf{PTableWalk}$. Specifically, it asserts that
+% the virtual address the virtual points-to assertion concerns corresponds to some physical address.
+% There exists only one persistent authorative ownership 
+% ($\ownGhost\gammaPred{\authfull{\ptableabswalk\ptablestore}}$) which collects the fragmental 
+% ghost page-table-walk mapping ($\sumwalkabs\vaddr\qfrac\paddr$),
+%  shown as double-ended solid arrow 1 to N imposing the update protocol.
 
-\begin{figure}
-   \includegraphics[width=0.75\columnwidth]{logical_addr_space.png}
-  \caption{Logical Resources of Address-Space Abstraction (1 Address Space)}
-  \label{fig:logicaladdrspace}
-  \end{figure}
+% \begin{figure}
+%    \includegraphics[width=0.75\columnwidth]{logical_addr_space.png}
+%   \caption{Logical Resources of Address-Space Abstraction (1 Address Space)}
+%   \label{fig:logicaladdrspace}
+%   \end{figure}
 
@@
 \subsection{The Impact of Changing Address Spaces on Modal Assertions}
+\todo[inline]{Colin must update this to address the frame rule comments from the last submission}
 % \subsection{Considering Hoare Doubles in the Context of Switching Address-Space}

### 2023-11-14 syntaxsemantics.tex
more compression and editing, through middle of section 4

@@
 }[keywords,comments,strings]
-
-\section{Machine State \& Syntax}
+\section{Machine State and Semantics}
+% \section{Machine State \& Syntax}
 \label{sec:syntax}
@@
 
-\subsection{Registers and Memory}
+% \subsection{Registers and Memory}
 Programs we demonstrate in this paper requires accessing two types of computer resource: registers and memory.
@@
 
-\subsection{State}
+% \subsection{State}
 \label{sec:state}
@@
 $\kw{r}\in\kw{greg}$) -- to the global maps.
-\subsection{Instructions}
+
+% \subsection{Instructions}
 \label{sec:instructions}
 
-\begin{figure}[t]
-\begin{lstlisting}[language=Coq,mathescape]
- Definition translate (rtv: word 64) ($\sigma$: state) (vaddr: word 64): word 64 + MemFail :=
-      l4e $\leftarrow$ translate_top_level $\sigma.\mathcal{M}$ rtv vaddr;
-      l3e $\leftarrow$ translate_from_l4e $\sigma.\mathcal{M}$ l4e vaddr;
-      l2e $\leftarrow$ translate_from_l3e $\sigma.\mathcal{M}$ l3e vaddr;
-      l1e $\leftarrow$ translate_from_l2e $\sigma.\mathcal{M}$ l2e vaddr;
-      addr $\leftarrow$ translate_from_l1e $\sigma.\mathcal{M}$ l1e vaddr;
-      return concat (shift_dropn addr 3 three_lt64) low3.
-\end{lstlisting}
-\vspace{-1em}
-\caption{\coq implementation of address translation (slightly simplified).}
-\label{fig:coq_addr_translation}
-\end{figure}
+% \begin{figure}[t]
+% \begin{lstlisting}[language=Coq,mathescape]
+%  Definition translate (rtv: word 64) ($\sigma$: state) (vaddr: word 64): word 64 + MemFail :=
+%       l4e $\leftarrow$ translate_top_level $\sigma.\mathcal{M}$ rtv vaddr;
+%       l3e $\leftarrow$ translate_from_l4e $\sigma.\mathcal{M}$ l4e vaddr;
+%       l2e $\leftarrow$ translate_from_l3e $\sigma.\mathcal{M}$ l3e vaddr;
+%       l1e $\leftarrow$ translate_from_l2e $\sigma.\mathcal{M}$ l2e vaddr;
+%       addr $\leftarrow$ translate_from_l1e $\sigma.\mathcal{M}$ l1e vaddr;
+%       return concat (shift_dropn addr 3 three_lt64) low3.
+% \end{lstlisting}
+% \todo[inline]{I don't think we need this given a more compact version of Figure \ref{fig:pagetables}}
+% \vspace{-1em}
+% \caption{\coq implementation of address translation (slightly simplified).}
+% \label{fig:coq_addr_translation}
+% \end{figure}
 
@@
 The latter require page table walks.
-Figure \ref{fig:coq_addr_translation} gives a slightly simplified version of our address translation code in monadic style:
-starting from the root page table address \textsf{rtv}, the virtual address \textsf{w} is either resolved to a physical address or an error is returned.
-The full set of possible errors is not relevant to this paper; our logic is tailored for kernel code which should not fault,
-so our proof rules guarantee no memory failures occur --- that the page table walk for any dereferenced address will succeed.
+% Figure \ref{fig:coq_addr_translation} gives a slightly simplified version of our address translation code in monadic style:
+% starting from the root page table address \textsf{rtv}, the virtual address \textsf{w} is either resolved to a physical address or an error is returned.
+% The full set of possible errors is not relevant to this paper; our logic is tailored for kernel code which should not fault,
+% so our proof rules guarantee no memory failures occur --- that the page table walk for any dereferenced address will succeed.
 
@@
   \]
-The semantic of each is realized by first translating the virtual-memory address of a memory location stored in $r_m$ to a physical memory location,
+The semantics of each is realized by first translating the virtual-memory address of a memory location stored in $r_m$ to a physical memory location
+per the description in Section \ref{sec:backgroundonmachinemodel},
 then either updating those memory contents with the contents of register $r_r$ (\TirNameStyle{WriteToMemFromReg}) 

### 2023-11-15 examples.tex
clarification
continuing to edit/compress/clarify, make figure 9 slightly more compact

@@
 
+We can see that this specification neatly encapsulates reasoning about invariants on the page tables in terms of general modal assumptions.\todo{speculative
+but ought to turn out this way}
+
 \citet{kolanski08vstte,kolanski09tphols} verified a single code block with their logic which was roughly Figure \ref{fig:mapping_code} for a 2-level ARM
@@
 \begin{lstlisting}[mathescape]
-;; Assume the save-space is in rdi,
-;; load-space in rsi
-;; No need to save/restore caller-save regs
-;; Save yielding context
+;; Assume the save-space is in rdi, load-space in rsi
+;; First, save yielding context
 $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast [\rtv'+56](\mathcal{I}\texttt{ASpace}(\theta,m) \ast \texttt{Pother})}_{\rtv}$
@@
 mov 0[rdi], rbx
-mov 8[rdi], rsp
-mov 16[rdi], rbp
-mov 24[rdi], r12
-mov 32[rdi], r13
-mov 40[rdi], r14
+... ;; mov rsp, rbp, r12, r13, r14, saved to offsets 8, 16, 24, 32, and 40 from rdi
 mov 48[rdi], r15
@@
 mov rbx, 0[rsi] 
-;; This switches to a new stack
-;; This stack may not be mapped in the
-;; current address space!
+;; This switches to a new stack, which may not be mapped in the current address space!
 mov rsp, 8[rsi] 
-mov rbp, 16[rsi]
-mov r12, 24[rsi]
-mov r13, 32[rsi]
-mov r14, 40[rsi]
+... ;; mov rbp, r12, r13, r14, to offsets 16, 24, 32, and 40 from rsi
 mov r15, 48[rsi]

### 2023-11-15 programlogic.tex
continuing to edit/compress/clarify, make figure 9 slightly more compact

@@
 chaining together the entries for each level, using the sequence of table offsets from the address being translated to index
-each table level, and using the physical page address embedded in each entry.
+each table level, and using the physical page address embedded in each entry.\footnote{
+  The fractions \lstinline|q1| through \lstinline|q4| represent the fractional ownership of each entry based on how many
+  word-aligned addresses might need to share the entry.
+}
 For example, 
@@
 This helper definition is also a more explicit \textsf{vProp}, explicitly binding a value for \lstinline|cr3| and using it
-to start the translation process.
+to start the translation process.\todo{Ismail: aren't phys ptstos vProps? Or iProps?}
 
@@
 
-\todo[inline,color=cyan]{Colin edit frontier}
-
 \subsection{Aliasing/Sharing Physical Pages}
   \label{sec:sharingpages}  
-  One might have already observed that the virtual points-to definition shown in Figure \ref{fig:strongvirtualpointsto} 
+  The virtual points-to definition shown in Figure \ref{fig:strongvirtualpointsto} 
   is too strong to specify some operations a virtual memory manager may need to do, such as
@@
   }
-   To speak more concretely, the fractional permissions on the page-table-walk assertion ($\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$ in Figure \ref{fig:strongvirtualpointsto}) 
-  ensure fractional ownership of the four physical page table entries accessed by the hardware during address translation,
-  as well as the physical memory holding the data of the virtual points-to.
-  However, this formulation stores knowledge of the page table walk's details with ownership of the backing physical
+  The use of $\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$ in Figure \ref{fig:strongvirtualpointsto}'s
+  virtual points-to definition'
+   stores knowledge of the page table walk's details with ownership of the backing physical
   memory.
@@
     \exists \delta\ldotp
-    \underbrace{(\lambda\mathit{cr3val}\ldotp\sumapaces{\mathit{cr3val}}\delta)}_\text{\begin{tabular}{c}Find ghost map for this\\address space\end{tabular}} \ast 
-  \underbrace{\sumwalkabs\vaddr\qfrac\paddr }_\text{\begin{tabular}{c}Fractional ownership of \\virt-to-phys translation\end{tabular}}\ast 
-  \underbrace{\paddr \mapsto_{\mathsf{p}}\{\textsf{q}\}\; \vale}_\text{\begin{tabular}{c}Value at\\physical location\end{tabular}}
+    \underbrace{(\lambda\mathit{cr3val}\ldotp\sumapaces{\mathit{cr3val}}\delta)}_\text{Find addr.\;space invariant} \ast 
+  \underbrace{\sumwalkabs\vaddr\qfrac\paddr }_\text{Ghost translation}\ast 
+  \underbrace{\paddr \mapsto_{\mathsf{p}}\{\textsf{q}\}\; \vale}_\text{Physical location}
 \end{array}
 \]
+\vspace{-1em}
 \caption{Virtual-Pointsto for Sharing Pages}
@@
   must still be tracked, but can be tracked separately from the virtual points-to assertion itself.
+  In practice the decisions about which virtual addresses are valid rest not with code posessing a virtual address, but with
+  the virtual memory manager --- and its invariants.
 
-  We take the approach of separating the physical page-table-walk pointsto ($\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$ in Figure \ref{fig:strongvirtualpointsto}) 
-  from the virtual pointsto relation. 
-  To do so, we take advantage of an Iris construction which permits tracking a map in ghost state that maps
-  virtual addresses to the physical addresses they translate to, and has for each key in the map (i.e.,
+  We take separate the physical page-table-walk  
+  from the virtual pointsto relation, replacing it with ghost state that merely guarantees the address translation would succeed.
+  Iris includes a construction which permits tracking a ghost map 
+  from virtual addresses to the physical addresses they translate to, and has for each key in the map (i.e.,
   each virtual address) a token. Updating the virtual-to-physical (ghost) mapping for a virtual address $va$
-  requires full ownership of the token. In addition, we keep an invariant that for each $va\mapsto pa$ mapping
+  requires full ownership of the token. Then the \emph{virtual memory manager's invariant} ensures that for each $va\mapsto pa$ mapping
   in this map, there are \emph{physical} resources sufficient to ensure that address translation for $va$
-  will resolve in the hardware to $pa$. But these physical resources, which specify \emph{which} physical locations
-  are involved in the page table walk, are now separated from but consistent with
-  the knowledge that such resources exist (which is embodied by the token for $va$, which tracks that $va$ maps to $pa$
-  in the ghost map). Thus we can store the token which summarizes the translation and ensures it exists in the virtual
-  points-to, and keep the ghost map and the invariant that every mapping in the ghost map has corresponding physical resources
-  for translation in a separate global invariant for each address space.
+  will resolve in the hardware to $pa$ --- $\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$.
+  This allows modification of the page tables themselves as long as the changes do not invalidate overall virtual-to-physical
+  translation.
+  Consistency follows from replacing the page table walk with ownership of the ghost map token for the relevant ghost map
+  entry, preventing the overall translation from being changed separately from the virtual points-to.
+
+  %Thus the specification of \emph{which} physical addresses support translation is separated from the virtual points-to.
+  % But these physical resources, which specify \emph{which} physical locations
+  %are involved in the page table walk, are now separated from but consistent with
+  %the knowledge that such resources exist (which is embodied by the token for $va$, which tracks that $va$ maps to $pa$
+  %in the ghost map). Thus we can store the token which summarizes the translation and ensures it exists in the virtual
+  %points-to, and keep the ghost map and the invariant that every mapping in the ghost map has corresponding physical resources
+  %for translation in a separate global invariant for each address space.
 
@@
 We replace the physical
- $\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$
+ $\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$ in the virtual points-to
 with
 fragmental ownership of the ghost-map ($ \sumwalkabs\vaddr\qfrac\paddr$),
- whose ownership just ensures that the virtual address ($\vaddr$) is mapped to a page (referenced by $\paddr$)
+guaranteeing that ghost map contains an mapping from the virtual address ($\vaddr$) to a physical address ($\paddr$).
 
-We place the authorative ownership $\mathcal{A}\textsf{PTableWalk}$ of the ghost-map in a per-address-space invariant
-($\mathcal{I}$\textsf{ASpace} in Figure \ref{fig:peraspaceinvariant}) whose ownership, when combined with the
-fraction stored in the virtual points-to, permits changes in the virtual-to-physical translations.
-By itself, it allows changes to the page tables as long as they do not change the overall mappings from virtual
-to physical addresses.
+We place the authorative ownership of the ghost translation $\mathcal{A}\textsf{PTableWalk}$ in a per-address-space invariant
+$\mathcal{I}$\textsf{ASpace} (Figure \ref{fig:peraspaceinvariant}). 
+$\mathcal{I}$\textsf{ASpace} alone allows changes to the page tables 
+that preserve overall virtual-to-physical translations.
+When combined with the
+fragment stored in the virtual points-to (Figure \ref{fig:virtualpointstosharing}),
+changes in the virtual-to-physical translations become possible.
 
@@
 
-Figure \ref{fig:peraspaceinvariant} describes the resources for each address space, which are collected
-for the full address space, but separated from virtual points-to assertions.
+Figure \ref{fig:peraspaceinvariant} describes the resources for each address space.
 $\Theta$ is the logical map from virtual addresses to the physical addresses they should translate to, 
@@
 
-$\mathcal{I}\textsf{ASpace}(\Theta,m)$ is then the resources for the current address space:
-It ensures that the current address space is valid (it has an entry in $m$, whose corresponding
-authoritative map matches $\Theta$), and asserts ownership of the appropriate fractional ownership
+$\mathcal{I}\textsf{ASpace}(\Theta,m)$ is then the resources for the current address space,
+ensuring that the current address space is valid (has an entry in $m$ whose corresponding
+authoritative map matches $\Theta$), and asserting ownership of the appropriate fractional ownership
 of the physical page table walk for each virtual-to-physical translation in $\Theta$ (for the current address space).
@@
 reasoning principles behind the virtual memory access, and we leave the use of this within a larger
-kernel to future work.
+kernel to future work. We do, however, deal with the existence of multiple address spaces.
 
@@
 \label{sec:aspacemanagement}
-So far, we have introduced logical abstractions for a single address space. However, virtual memory management
-implementations typically handle more than one address-space,
-and doing this requires assertions which talk about other address spaces, and means to switch address spaces.
+So far, we have introduced logical abstractions for a single address space, but VMMs
+ handle more than one address-space,
+and doing so requires a way to talk about other address spaces, and means to switch address spaces.
 % requiring a means to  which we abstract as a ghost map from root address of an address-space to a namespace labelling the address space
@@
 what we need in order for assertions in one address space to talk about assertions in another.
+\looseness=-1
+
+\todo[inline,color=cyan]{Colin revision frontier}
 

### 2023-11-15 programlogic.tex
clean up some framing discussion

@@
 \begin{figure}
-\[
-  \begin{array}{r@{\qquad}l}
-    \hbox{(\TirNameStyle{ModalAddressSpace})} &
-         [r](\mathsf{P}) \;: \mathsf{vProp} \; \Sigma \; \stackrel{\triangle}{=} \; \lambda \_,\; \mathsf{P}~\mathsf{r}.\\
-    \hbox{(\TirNameStyle{ModalAddressSpaceMono})} & (\textsf{P} \vdash \textsf{Q}) \vdash  ([r](\textsf{P}) \vdash  [r](\textsf{Q}))\\
-    \hbox{(\TirNameStyle{ModalAddressSpaceSep})} & [r](\textsf{P} \ast \textsf{Q}) \dashv\vdash ([r](\textsf{P}) \ast [r](\textsf{Q})\\
-    \hbox{(\TirNameStyle{ModalAddressSpaceAnd})} &   [r](\textsf{P}\land \textsf{Q}) \dashv\vdash [r](\textsf{P})\land [r](\textsf{Q})\\    
-    \hbox{(\TirNameStyle{ModalAddressSpaceOr})} &    [r](\textsf{P}\lor \textsf{Q}) \dashv\vdash [r](\textsf{P})\lor [r](\textsf{Q})\\   
-    % \hbox{(\TirNameStyle{ModalAddressSpacePure})} &  \textsf{pure\_vprop P}(\textsf{iPropI }\Sigma)\textsf{ : vProp }\Sigma \stackrel{\triangle}{=} \; \lambda \_ \ldotp \textsf{P}\\
-    \hbox{(\TirNameStyle{ModalAddressSpaceFact})} &  \textsf{Fact P} \stackrel{\triangle}{=} \;\forall \; ,r \; r' \ldotp  \textsf{P r} \dashv\vdash \textsf{P r'}\\
-    % \hbox{(\TirNameStyle{ModalAddressSpaceFactIntroElim})} \\
-    % \qquad\qquad \textsf{Fact pure\_vprop P} \vdash  ([r](\textsf{pure\_vprop P}) \dashv\vdash (\textsf{pure\_vprop P}) (\textsf{vProp }\Sigma))\\
-    \hbox{(\TirNameStyle{ModalAddressSpaceFactIntroElim})} 
-    &\textsf{Fact P} \vdash  ([r](\textsf{P}) \dashv\vdash (\textsf{P}) %(\textsf{vProp }\Sigma))\\
-  \end{array}
-  \]
+\begin{mathpar}
+    [r](\mathsf{P}) \;: \mathsf{vProp} \; \Sigma \; \stackrel{\triangle}{=} \; \lambda \_,\; \mathsf{P}~\mathsf{r} \and
+    \textsf{Fact P} \stackrel{\triangle}{=} \;\forall \; ,r \; r' \ldotp  \textsf{P r} \dashv\vdash \textsf{P r'}\\
+    \inferrule{}{(\textsf{P} \vdash \textsf{Q}) \vdash  ([r](\textsf{P}) \vdash  [r](\textsf{Q}))}\and
+    \inferrule{}{[r](\textsf{P} \ast \textsf{Q}) \dashv\vdash ([r](\textsf{P}) \ast [r](\textsf{Q})}\and
+    \inferrule{}{[r](\textsf{P}\land \textsf{Q}) \dashv\vdash [r](\textsf{P})\land [r](\textsf{Q})}\and
+    \inferrule{}{[r](\textsf{P}\lor \textsf{Q}) \dashv\vdash [r](\textsf{P})\lor [r](\textsf{Q})}\and
+    \inferrule{}{\textsf{Fact P} \vdash  ([r](\textsf{P}) \dashv\vdash (\textsf{P})} %(\textsf{vProp }\Sigma))\\
+\end{mathpar}
   \caption{Other-space Modality and Its Laws}
@@
 
-\todo[inline,color=cyan]{Colin revision frontier}
+% The inspiration for this assertion comes from hybrid logic~\cite{blackburn1995hybrid,areces2001hybrid,goranko1996hierarchies,gargov1993modal},
+% which as mentioned earlier uses as modalities the names of worlds in a Kripke model to state
+% that an assertion is true in that specific world.
+% Conceptually the address space rooted at some physical address (intended as the start of an L4 table)
+% is a distinct world, and it makes sense to name it by referring to the root of the
+% supporting page table structure. However, rather than explicitly building a Kripke model for our
+% assertions, we inherit them by lifting the model of our ambient logic, Iris.
 
-The inspiration for this assertion comes from hybrid logic~\cite{blackburn1995hybrid,areces2001hybrid,goranko1996hierarchies,gargov1993modal},
-which as mentioned earlier uses as modalities the names of worlds in a Kripke model to state
-that an assertion is true in that specific world.
-Conceptually the address space rooted at some physical address (intended as the start of an L4 table)
-is a distinct world, and it makes sense to name it by referring to the root of the
-supporting page table structure. However, rather than explicitly building a Kripke model for our
-assertions, we inherit them by lifting the model of our ambient logic, Iris.
-
-This assertion is also useful in proofs: we can prove that this modality follows certain useful laws, and that it interacts with
-a class of assertions which are true regardless of which address space they are evaluated in.
-The other-space modality distributes over the various separation logic connectives, and follows the rule of
+We can prove that this modality follows certain useful laws: that its truth is independent of the address
+space in which it is considered,  it distributes over various logical connectives, and it follows the rule of
 consequence.
-We also define a class of \textsf{vProp}s we call \textsf{Fact}s, which are those assertions
-which ignore their address space entirely.
-Physical points-tos and register value assertions, when treated as \textsf{vProp}s, are \textsf{Fact}s.
+We call \textsf{vProp} assertions whose truth is independent of the current address space
+\textsf{Fact}s; these include other-space assertions, physical memory points-tos, and register assertions.
 \textsf{Fact}s can move in and out of other-space modalities freely.
@@
 \subsection{The Impact of Changing Address Spaces on Modal Assertions}
-\todo[inline]{Colin must update this to address the frame rule comments from the last submission}
 % \subsection{Considering Hoare Doubles in the Context of Switching Address-Space}
@@
 \begin{mathpar}
-\inferrule[Broken]{ }{
+\inferrule*[left=Broken]{ }{
   \{P \ast \textsf{cr3}\mapsto_{r} \;r_1 \ast r \mapsto_{r} \;r_2 \ast [r_2](Q)\}
@@
 This rule captures the intuitive change of address space in a Hoare triple.
-The problem with this is that it interacts quite poorly with the traditional frame
+This interacts quite poorly with the traditional frame
 rule and the modal flavor of virtual points-to assertions.
@@
 So this derivation clearly leads to an unsound conclusion.
-This suggestss that the traditional frame rule and the Hoare triple presentation of the change-of-address-space rule cannot soundly coexist in the same system.
-The heart of the problem is that while updating \lstinline|cr3| is \emph{physically} local, it globally changes the interpretation of virtual addresses.
-So it is simply unsound to frame around \lstinline|cr3| updates.
-In hindsight this is not surprising, because we have essentially defined
-our assertions to depend on resources they do not explicitly mention --- the value of \lstinline|cr3|.
-However, it has important impacts on our formal development.
+The essential problem is that the frame rule is motivated by local reasoning about local updates, but
+a switch of address space is a \emph{global} change, since it may invalidate information about virtual addresses.
+Thus framing around \lstinline|cr3| updates is unsound, which has important impacts on our formal development.
 
@@
 This is then how we justify our actual rule for \lstinline|cr3| updates:
+\todo[inline]{Ismail: Can we chat Wednesday about this rule, which has cr3 register pointstos instead of the cr3 subscripts?
+The subscript form might make the derivation of a frame rule harder.}
 \begin{mathpar}
-\inferrule[ChangeAddressSpace]{
+\inferrule*[left=ChangeAddressSpace]{
   \{[r_1](P) \ast cr3 \mapsto_{r} r_2 \ast r \mapsto_{r} r_2 \ast Q\}\overline{is}
@@
 \begin{mathpar}
-  \inferrule[Cr3Frame]{
+  \inferrule*[left=Cr3Frame]{
     \{P\ast cr3=v\}\;C\;\{ Q \ast cr3=v\}
@@
 
-Readers familiar with dynamic frames~\cite{parkinson2011relationship} might find it useful to notice that a different perspective on this matter is that virtual points-to assertions are self-stable \emph{except} for changes in \lstinline|cr3|, so framing would then naturally require other means of holding \lstinline|cr3| constant (or saving and restoring it).
-Virtual points-to assertions could be made self-stable by also giving them partial ownership over \lstinline|cr3| assertions, but this would require explicitly plumbing that ownership from \emph{all} assertions back to any place an address space change might occur; this would seem to be a far graver loss of modularity than this extra quirk in framing discussions.
+Another useful perspective on this might be the dynamic frames~\cite{parkinson2011relationship} 
+view: virtual points-to assertions are self-stable \emph{except} for changes in \lstinline|cr3|, so framing would then naturally require other means of holding \lstinline|cr3| constant (or saving and restoring it).
+% Virtual points-to assertions could be made self-stable by also giving them partial ownership over \lstinline|cr3| assertions, but this would require explicitly plumbing that ownership from \emph{all} assertions back to any place an address space change might occur; this would seem to be a far graver loss of modularity than this extra quirk in framing discussions.
 
@@
 \label{sec:selected_rules}
+\todo[inline]{Ismail: Please update this section with rip stuff}
 To address the issues described in Section \ref{sec:issues}, we give our logic using
@@
 
-At the moment our proofs do rely on 11 small axioms of properties which should be provable, but
+At the moment our proofs do rely on 11\todo{update!} small axioms of properties which should be provable, but
 are challenging to discharge due to some representation choices in our model.\footnote{See \lstinline|srx/x64/machine/current_axioms.v|}

### 2023-11-15 relwork.tex
make related work slightly shorter, but also cover related work limitations on examples there in a single pass, so that information can be dropped from earlier sections (examples)

--- 2023-04-14 relwork.tex
@@
 There has been relatively little prior work on formal verification of virtual memory.
-Instead, much OS verification work has focused on minimizing reasoning about virtual memory management.
-The original Verisoft project~\cite{alkassar2008verisoft,alkassar2010pervasive,alkassar2008formal,dalinger2005verification,hillebrand2005address,alkassar2008formal,starostin2010formal} relied on custom hardware which, among other things, always ran kernel code with virtual memory disabled, removing the circularity that is a key challenge of verifying actual virtual memory code: at that point page tables become a basic partial map data structure to represent user program address translations.
-Other work on OS verification either never progressed far enough to address VMM verification (Verisoft XT~\cite{cohen2009vcc,cohen2010local,dahlweid2009vcc,cohen2013SOFSEM}), or uses memory-safe languages to enable safe co-habitation of a single address space by all processes (Singularity~\cite{Fahndrich2006language,Hunt2007singularity,Hunt2007sealing,Barnett2011specsharp}, Verve~\cite{Yang2010Verve}, and Tock~\cite{levy2017multiprogramming}).
+Instead, most OS verification work has focused on minimizing reasoning about virtual memory management.
+The original \textsc{Verisoft} project~\cite{alkassar2008verisoft,alkassar2010pervasive,alkassar2008formal,dalinger2005verification,hillebrand2005address,alkassar2008formal,starostin2010formal} 
+relied on custom hardware which, among other things, always ran kernel code with virtual memory disabled, removing the circularity that is a key challenge of verifying 
+VMM code for real hardware: at that point page tables become a basic partial map data structure to represent user program address translations,
+with an idiosyncratic format. It turns out that subsequent OS verification work also treats page tables this way, but
+unsoundly given that other projects target hardware that \emph{does} run the kernel with address translation: they \emph{trust}
+that the particular page table manipulations do not, for example, unmap kernel code (which can crash the machine even if done ``temporarily''\footnote{A bug
+one author has personally encountered in a research kernel, though not one discussed here.}).
 
-The work that does address the core challenges of VMM verification is all associated with either \textsc{seL4} or \textsc{CertiKOS}.
+\textsc{seL4}~\cite{Klein2009seL4,seL4TOCS,Sewell2013translation} is a formally verified L4 microkernel~\cite{Liedtke1995,Liedtke1996} 
+(and the first verified OS kernel to run on real-world hardware), verified with a mix of refinement proofs and program logic reasoning down to the assembly level.
+Because \textsc{seL4} is a microkernel, most VMM functionality actually lives in usermode and is unverified, and moreover, 
+their hardware model omits address translation entirely and the MMU entirely~\cite{Klein2009seL4,seL4TOCS}. 
+As a result, the limited page table management present in the microkernel treats page tables as idiosyncratic tree-maps as in
+\textsc{Verisoft}, despite actually running with address translation.
+This is partly mitigated by manually identifying some trusted invariants (e.g., that the address range designated for the kernel is appropriately mapped)
+and setting up the proof to ensure those invariants are maintained (i.e., as an extra proof obligation not required by their hardware model).
 
-\textsc{CertiKOS}~\cite{gu15,gu2016certikos,gu2018certikos,chen2016interrupts} is a microkernel intended for use as a hypervisor,
-and its papers do not explicitly detail verification of the VMM, so we do not know the full space of which VMM functionality 
-is verified, but we do know it includes the ability to map or unmap pages.
-The work is clear, however, that it trusts low-level assembly fragments such as the instruction sequence which actually
-switches address spaces, rather than verifying them.
+\textsc{CertiKOS}~\cite{gu15,gu2016certikos,gu2018certikos,chen2016interrupts} is a microkernel intended for use as a hypervisor.
 The overall approach in that body of work is many layers of refinement proofs, using a
@@
 on the project from the same group~\cite{vaynberg2012compositional}, the purpose of some layers is to abstract away from 
-virtual memory, so the proof is essentially a simulation proof covering for example a proof that execution with page-in on 
-page faults is a valid refinement of an execution model where no paging occurs.
+virtual memory (as early as possible).
+The papers on \textsc{CertiKOS} do not explicitly detail the VMM beyond highlighting its existence and referencing
+that it performs mapping operations for user code.
+The work is clear, however, that it fully trusts low-level assembly fragments such as the instruction sequence which actually
+switches address spaces, rather than verifying them.
+We can also infer other limitations from the project's pervasive reliance on a modified \textsc{CompCert},
+which fixes an abstract memory model aimed at \emph{user-mode} C programs~\cite{leroy2008formal,leroy2009formally}. 
+\textsc{CompCert}'s model of memory cannot handle non-local updates, so we can conclude \textsc{CertiKOS}'s proofs cannot
+address impact of virtual memory mapping changes on the kernel itself.
 % Another key aspect of their approach is that the OS is written in Clight and compiled with \textsc{CompCert}~\cite{blazy2006formal,leroy2009formally,leroy2008formal}.
@@
 
-\textsc{seL4}~\cite{Klein2009seL4,seL4TOCS,Sewell2013translation} is a formally verified L4 microkernel~\cite{Liedtke1995,Liedtke1996} (and the first verified OS kernel to run on real-world hardware), verified with a mix of refinement proofs and program logic reasoning down to the assembly level.
-Because \textsc{seL4} is a microkernel, most VMM functionality actually lives in usermode and is unverified, and moreover, their hardware model omits address translation entirely and the MMU entirely~\cite{Klein2009seL4,seL4TOCS}.
-As a result, the limited page table management present in the microkernel treats page tables as idiosyncratic tree-maps, ignoring the risks posed by even transient inconsistencies that would crash the kernel on real hardware (like ``temporarily'' unmapping the kernel).
-This is mitigated primarily by manually identifying some trusted invariants (e.g., that the address range designated for the kernel is appropriately mapped) and setting up the proof to ensure those invariants are maintained (i.e., as an extra proof obligation not required by their hardware model).
+Other work on OS verification either never progressed far enough to address VMM verification 
+(\textsc{Verisoft XT}~\cite{cohen2009vcc,cohen2010local,dahlweid2009vcc,cohen2013SOFSEM}), or uses memory-safe languages to enable safe co-habitation 
+of a single address space by all processes (\textsc{Singularity}~\cite{Fahndrich2006language,Hunt2007singularity,Hunt2007sealing,Barnett2011specsharp}, \textsc{Verve}~\cite{Yang2010Verve},
+and \textsc{Tock}~\cite{levy2017multiprogramming}).
 
+Some work has been done like ours, studying VMM verification separately from the rest of the kernel.
 One important outgrowth of the \textsc{seL4} project, not integrated into the main project's proof, was work by 
@@
 
-Their work has a number of significant limitations which our work addresses.
 They also define their virtual points-to assertions such that a virtual points-to $p\mapsto_\mathsf{v} a$ owns the full 
@@
 but 8-byte entries).
-
-% Kolanski and Klein's points-to assertions do not model that page table entries for nearby addresses typically 
-% \emph{share} entries in higher layers of the page tables --- a single L1 entry maps 4KB of memory on many architectures,
-% but their logics avoid fractional ownership, so they can in fact only use a single memory location per page of memory.
-% In fact, because their virtual points-to assertions contain \emph{full} ownership of all entries, even in the highest-level
-% page table (L4 in our case), each entry can contribute to only a single mapped address. Thus any assertion
-% in their logic that implies there are more virtual addresses mapped than entries in the top-level page table implies false.
 Our definitions make use of fractional permissions throughout; Figure \ref{fig:strongvirtualpointsto}'s definition
@@
 as part of the virtual points-to itself, preempt support for changes to page tables which do not actually affect 
-address translation.
+address translation. We address this by moving from Figure \ref{fig:strongvirtualpointsto}'s definition (essentially Kolanski and Klein's
+with fractional sharing of intermediate table entries) to Figure \ref{fig:virtualpointstosharing}'s.
 
-The other major distinction is that Kolanski and Klein have no accounting for other address spaces.
+Kolanski and Klein do verify code to map in a new page by installing an L1 page table entry, akin to our Figure \ref{fig:mapping_code}.
+However, our logic treatment of mapping goes beyond theirs. 
+Kolanski and Klein must unfold machine semantics at one point in their proof of mapping correctness~\cite[p.~27]{kolanski08vstte}, while
+our logic permits us to conduct the proof entirely in separation logic.
+Their proofs also do not address
+converting physical addresses of page table entries to virtual addresses that can be used to access them.
+Their original low-level (pseudo-assembly) proof assumes the correct virtual address already exists~\cite{kolanski08vstte},
+and their subsequent C-level proof~\cite{kolanski09tphols} axiomatizes a function akin to our \lstinline|ensure_L1|.
+While full proof of \lstinline|ensure_L1|'s page table walk is ongoing work, we have verified that virtual-to-physical
+translation can be done in our logic with appropriate kernel invariants (Section \ref{sec:traversing}), while
+Kolanski and Klein's model may not be able to do so --- because the virtual points-to owns the physical resources associated with
+intermediate page table entries, it appears impossible to have a virtual points-to for a page table entry that is part of the
+page table walk for another virtual address.
+
+The other major distinction is that Kolanski and Klein have no explicit accounting for other address spaces.
 Their logic does not deal with change of address space, and has no way to assert that certain facts hold
@@
 for separation logic is unsound in the presence of address space changes and address-space-contingent assertions.
-% The mapping write is verified in an ad hoc way by unfolding the machine semantics, because the logic lacks proper reusable rules for
-% updating page tables.
-
 Our approach in this paper uses modalities to distinguish virtual-address-based assertions that hold only in specific 
@@
 memory.
+
+As noted earlier, our other-space modality comes from ideas in hybrid logic~\cite{areces2001hybrid,blackburn1995hybrid,gargov1993modal,goranko1996hierarchies},
+where modalities are indexed by \emph{nominals} which are names for specific individual states in a Kripke model.
+We are aware of only two prior works combining these ideas with program logics, one showing that traditional
+nominals extends the expressive power of separation logic~\cite{brotherston2014parametric}, another
+using nominals to refer to states of other actors in an actor language~\cite{gordon2019modal}.
+
+% As noted throughout the paper, the inspiration for our other-space modality comes from hybrid logic~\cite{areces2001hybrid,blackburn1995hybrid,gargov1993modal,goranko1996hierarchies},
+% where modalities are indexed by \emph{nominals} which are names for specific individual states in a Kripke model.
+% We are aware of only two prior works combining hybrid logics with program logics specifically. 
+% Brotherston and Villard~\cite{brotherston2014parametric} demonstrated that may properties true of various 
+% separation logics are not definable in boolean \BI (\BBI), and showed that a hybrid extension \HyBBI allows
+% most such properties to be defined (e.g., the fact that separating conjunction is cancellative is unprovable 
+% in boolean \BI, but provable in \HyBBI). There, nominals named resources 
+% (roughly, but not exactly, heap fragments). 
+% Gordon~\cite{gordon2019modal} described a use of hybrid logic in the verification of actor programs, 
+% where nominals named the local state of individual actors (with such assertions stabilized with a 
+% rely/guarantee approach). 
+Beyond these, there is limited work on the interaction of hybrid logic with substructural logics, in restricted forms
+that do not affect expressivity~\cite{despeyroux2014hybrid,chaudhuri2019hybrid}.
+% Primarily there is a line of work on hybrid linear logic (\HyLL)~\cite{despeyroux2014hybrid}, 
+% originally used as a way to more conveniently express aspects of transition systems in linear logic. 
+% However, \HyLL's proof rules offer no non-trivial interactions with multiplicative connectives 
+% (every \HyLL proof can in fact be embedded into regular linear logic~\cite{chaudhuri2019hybrid}, 
+% unlike Brotherston and Villard's \HyBBI, which demonstrably increases expressive power over its base \BBI.
+
+% In both \HyLL and \HyBBI, nominals denote worlds with monoidal structure (as worlds in Kripke semantics
+% for either LL or \BBI necessarily have monoidal structure). Our nominals, by contrast, 
+% do not name worlds in the same sense with respect to Iris's CMRAs, 
+% but in fact \emph{classes} of worlds, because the names are locations 
+% (a means of \emph{selecting} resources) rather than resources.  
+% A key difference is that the use of nominals in those logics corresponds specifically to hypothetical 
+% reasoning about resources (until a nominal is connected to a current resource, in which case conclusions 
+% can be drawn about the current resource), which means the modalities themselves do not ``own'' resources. 
+% Instead, assertions under our other-space modality can and do
+% have resource footprints.
+% Pleasantly, we sidestep most of the metatheoretical complexity of those other substructural hybrid
+% systems by building our logic within a substructural metatheory (\iris).
+
+\iris has been used to build other logics through pointwise lifting, notably logics that deal with weak
+memory models~\cite{dang2019rustbelt,dang2022compass}. Those systems build a derived logic
+whose lifting consists of functions from thread-local views of events (an operationalization of the release-acquire + nonatomic
+portion of the repaired C11 memory model~\cite{lahav2017repairing}): there modalities $\Delta_\pi(P)$ and $\nabla_\pi(P)$
+represent that $P$ held before or will hold after certain memory fence operations by thread $\pi$.
+The definitions of those specific modalities existentially quantify over other views, related to the ``current'' view (the one where
+the current thread's assertions are evaluated), and evaluate $P$ with respect to those other views. This approach to parameterizing
+assertion semantics by a point of evaluation, and evaluating modalized assertions at other points quanfied in the definition of a modality,
+is the classic notion of modal assertions, whereas hybrid logics expose the choice of evaluation point in assertions,
+allowing statements of more properties. In these weak memory examples this additional expressive power would not be useful,
+because any relevant points of evaluation (thread views) are intimately tied to memory fences performed by the program, whereas
+for virtual memory management the kernel must be able to choose or construct arbitrary other address spaces.
+
+%   is what it means
+% to have a modality at all.
+% It is \emph{not}, however, an instance of hybrid logic, which is specifically demarcated by an assertion language where
+% \emph{assertions}, not their semantics, choose and name the evaluation points for modal assertions.
+% A hybrid extension of the aforementioned logics would include assertions which named specific views at which to evaluate
+% $P$, in the syntax of the assertion (e.g., $\Delta_\pi^v(P):=\lambda\_\ldotp (P\;v)$) rather than the 
+% $\Delta_\pi(P):= \lambda v\ldotp (\exists v_{rel}\ldotp \ownGhost{\pi}{\mathsf{RelV}(v_{rel})\;v} \ast (P\;v_{rel})))$ actually used.
+% Note the hybrid version takes the place to evaluate $P$ as a parameter, and therefore allows the \emph{derived} (modal) logic to explicitly
+% reason in terms of evaluation points, rather than hiding all points of evaluation in the internal definitions of modalities.
+% This prior \iris-based work also uses modalities where the interpretations are fixed a priori by logic designers. In contrast, our
+% address space modalities' interpretations can be changed by program behaviors via page table updates; the equivalent in the prior work would
+% be if programs could directly manipulate the buffers used to model weak memory behaviors,
+% which they cannot.\footnote{Note that the views modelled modally in this prior work are abstractions of a wide range of hardware, so such
+% manipulation of corresponding hardware resources is in fact impossible,
+% while the address space mappings in our work reflect known hardware components present in a variety of CPUs.}

### 2023-11-15 examples.tex

@@
 
+\begin{figure*}
+\[
+\begin{array}{l}
+  \mathcal{I}\textsf{ASpace}(\ptablestore,m)\stackrel{\triangle}{=} \textsf{ASpace\_Lookup}(\ptablestore,\Xi,m) \ast \\
+  \qquad\qquad\qquad \bigast{(\vaddr, \textsf{paddr})\in \ptablestore}{\exists\;(\textsf{l4e, l3e, l2e, l1e, paddr})\ldotp \textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\vaddr\textsf{, l4e, l3e, l2e, l1e, paddr})} \\
+  \qquad\qquad\qquad \bigast{(\vaddr, \textsf{paddr})\in \Xi}{\exists\; (\textsf{qfrac, val, vaddr}) \ldotp \ulcorner \textsf{vaddr} = \textsf{paddr} + \textsf{KERNBASE} \urcorner \ast  \underbrace{\sumwalkabs\vaddr\qfrac\paddr }_\text{Ghost translation} \ast \\
+\qquad\qquad\qquad\qquad\qquad\qquad\qquad\qquad    \underbrace{\paddr \mapsto_{\mathsf{p}}\{\textsf{qfrac}\}\; \vale}_\text{Physical location} \ast \ulcorner \textsf{qfrac} = 1 \leftrightarrow \; \lnot\textsf{entry\_present }(\vale) \urcorner }\\
+  \textsf{ where } \\
+   \textsf{ASpace\_Lookup} (\ptablestore,m) \stackrel{\triangle}{=} \lambda\textsf{ cr3val} \ldotp \; \exists \gammaPred \; \ldotp \ulcorner m \; !!\; \textsf{cr3val} = \textsf{Some } \gammaPred \urcorner \ast
+   \ownGhost\gammaPred{\authfull{\ptableabswalk\ptablestore}} \ast \\ \qquad\qquad\qquad\qquad\qquad\qquad\qquad\qquad \ownGhost\gammaPred{\authfull{\ptableabswalk\Xi}}
+
+\end{array}
+\]
+\caption{Global Address-Space Invariant in Figure \ref{fig:peraspaceinvariant} extended with a ghost map bookkeeping identity mappings }
+  \label{fig:peraspaceinvariant_with_p2v_extension}
+  \end{figure*}
+
 Work to verify the full page table walk is ongoing, but we have verified the key step for a small x86-64 kernel, which is the physical-to-virtual
@@
 
-\todo[inline]{typeset as a triple}
+\begin{figure}\footnotesize
+\begin{lstlisting}[mathescape]
+;; Assume the save-space is in rdi, load-space in rsi
+;; First, save yielding context
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast [\rtv'+56](\mathcal{I}\texttt{ASpace}(\theta,m) \ast \texttt{Pother})}_{\rtv}$
+$\specline{ \texttt{rsi}\mapsto_{\textsf{r}} \rtv' \ast \texttt{rdi}\mapsto_{\textsf{r}} \_ \ast \texttt{rbx}\mapsto_{\textsf{r}} \texttt{rbxv} \ast  \texttt{rsp}\mapsto_{\textsf{r}} \texttt{rspv} \ast \texttt{rbp}\mapsto_{\textsf{r}} \texttt{rbpv} }_{\rtv}$
+$\specline{\texttt{r12}\mapsto_{\textsf{r}} \texttt{r12v} \ast \texttt{r13}\mapsto_{\textsf{r}} \texttt{r13v} \ast \texttt{r14}\mapsto_{\textsf{r}} \texttt{r14v} \ast \texttt{r15}\mapsto_{\textsf{r}} \texttt{r15v}}_{\rtv}$
+$\specline{\texttt{rdi+0} \mapsto_{\textsf{v}} \_ \ast \texttt{rdi+8} \mapsto_{\textsf{v}} \_ \ast \texttt{rdi+16} \mapsto_{\textsf{v}} \_ \ast \texttt{rdi+24} \mapsto_{\textsf{v}} \_ \ast \texttt{rdi+32} \mapsto_{\textsf{v}} \_}_{\rtv}$
+$\specline{\texttt{rdi+40} \mapsto_{\textsf{v}} \_ \ast \texttt{rdi+48} \mapsto_{\textsf{v}} \_\ast \texttt{rdi+56} \mapsto_{\textsf{v}} \_}_{\rtv}$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast [\rtv'+56](\mathcal{I}\texttt{ASpace}(\theta,m) \ast \texttt{Pother})}_{\rtv}$
+$\specline{ \texttt{rdi}\mapsto_{\textsf{r}} \_ \ast \texttt{rbx}\mapsto_{\textsf{r}} \texttt{rbxv} \ast  \texttt{rsp}\mapsto_{\textsf{r}} \texttt{rspv} \ast \texttt{rbp}\mapsto_{\textsf{r}} \texttt{rbpv} }_{\rtv}$
+$\specline{\texttt{r12}\mapsto_{\textsf{r}} \texttt{r12v} \ast \texttt{r13}\mapsto_{\textsf{r}} \texttt{r13v} \ast \texttt{r14}\mapsto_{\textsf{r}} \texttt{r14v} \ast \texttt{r15}\mapsto_{\textsf{r}} \texttt{r15v}}_{\rtv}$
+$\specline{\texttt{rdi+0} \mapsto_{\textsf{v}} \texttt{rbxv} \ast \texttt{rdi+8} \mapsto_{\textsf{v}} \texttt{rspv} \ast \texttt{rdi+16} \mapsto_{\textsf{v}} \texttt{rbpv} \ast \texttt{rdi+24} \mapsto_{\textsf{v}} \texttt{r12v} \ast }_{\rtv}$
+$\specline{\texttt{rdi+32} \mapsto_{\textsf{v}} \texttt{r13v} \ast \texttt{rdi+40} \mapsto_{\textsf{v}} \texttt{r14v} \ast \texttt{rdi+48} \mapsto_{\textsf{v}} \texttt{r15v}\ast \texttt{rdi+56} \mapsto_{\textsf{v}} \rtv}_{\rtv}$
+;; This offsetst physical address at KERNBASE!
+add r,  
+... ;; mov rbp, r12, r13, r14, to offsets 16, 24, 32, and 40 from rsi
+mov r15, 48[rsi]
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast [\rtv'+56](\mathcal{I}\texttt{ASpace}(\theta,m) \ast \texttt{Pother})}_{\rtv}$
+$\specline{ \texttt{rdi}\mapsto_{\textsf{r}} \_ \ast \texttt{rbx}\mapsto_{\textsf{r}} \rtv'+0 \ast  \texttt{rsp}\mapsto_{\textsf{r}} \rtv'+8 \ast \texttt{rbp}\mapsto_{\textsf{r}} \rtv'+16 }_{\rtv}$
+$\specline{\texttt{r12}\mapsto_{\textsf{r}} \rtv'+24 \ast \texttt{r13}\mapsto_{\textsf{r}} \rtv'+32 \ast \texttt{r14}\mapsto_{\textsf{r}} \rtv'+40 \ast \texttt{r15}\mapsto_{\textsf{r}} \rtv'+48}_{\rtv}$
+$\specline{\texttt{rdi+0} \mapsto_{\textsf{v}} \texttt{rbxv} \ast \texttt{rdi+8} \mapsto_{\textsf{v}} \texttt{rspv} \ast \texttt{rdi+16} \mapsto_{\textsf{v}} \texttt{rbpv} \ast \texttt{rdi+24} \mapsto_{\textsf{v}} \texttt{r12v} \ast }_{\rtv}$
+$\specline{\texttt{rdi+32} \mapsto_{\textsf{v}} \texttt{r13v} \ast \texttt{rdi+40} \mapsto_{\textsf{v}} \texttt{r14v} \ast \texttt{rdi+48} \mapsto_{\textsf{v}} \texttt{r15v}\ast \texttt{rdi+56} \mapsto_{\textsf{v}} \rtv}_{\rtv}$
+;; Switch to the new address space
+mov cr3, 56[rsi]
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast [\rtv'+56](\mathcal{I}\texttt{ASpace}(\theta,m) \ast \texttt{Pother}) }_{\rtv}$
+$\specline{ \texttt{rdi}\mapsto_{\textsf{r}} \_ \ast \texttt{rbx}\mapsto_{r} \rtv'+0 \ast  \texttt{rsp}\mapsto_{\textsf{r}} \rtv'+8 \ast \texttt{rbp}\mapsto_{\textsf{r}} \rtv'+16 }_{\rtv}$
+$\specline{\texttt{r12}\mapsto_{\textsf{r}} \rtv'+24 \ast \texttt{r13}\mapsto_{\textsf{r}} \rtv'+32 \ast \texttt{r14}\mapsto_{\textsf{r}} \rtv'+40 \ast \texttt{r15}\mapsto_{\textsf{r}} \rtv'+48}_{\rtv}$
+$\specline{\texttt{rdi+0} \mapsto_{\textsf{v}} \texttt{rbxv} \ast \texttt{rdi+8} \mapsto_{\textsf{v}} \texttt{rspv} \ast \texttt{rdi+16} \mapsto_{\textsf{v}} \texttt{rbpv} \ast  \texttt{rdi+24} \mapsto_{\textsf{v}} \texttt{r12v} \ast }_{\rtv}$
+$\specline{\texttt{rdi+32} \mapsto_{\textsf{v}} \texttt{r13v} \ast \texttt{rdi+40} \mapsto_{\textsf{v}} \texttt{r14v} \ast \texttt{rdi+48} \mapsto_{\textsf{v}} \texttt{r15v}\ast \texttt{rdi+56} \mapsto_{\textsf{v}} \rtv}_{\rtv}$
+$\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\; \sqsubseteq $
+$\specline{ [\rtv](\texttt{Pcurrent}) \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast \texttt{Pother}}_{\rtv'+56}$
+\end{lstlisting}
+\caption{Basic task switch code that switches address spaces.}
+\label{fig:mapping_code}
+\end{figure}
 

### 2023-11-15 examples.tex
p2v stuff

@@
 
+\newcommand{\true}{\textsf{true}}
+\tikzstyle{boxedassert_border} = [sharp corners,line width=0.2pt]
+\NewDocumentCommand \boxedassertpv {O{} m o}{%
+	\tikz[baseline=(m.base)]{
+		%	  \node[rectangle, draw,inner sep=0.8pt,anchor=base,#1] (m) {${#2}\mathstrut$};
+		\node[rectangle,inner sep=1.5pt,outer sep=0.2pt,anchor=base] (m) {${\,#2\,}\mathstrut$};
+		\draw[#1,boxedassert_border] ($(m.south west)$) rectangle ($(m.north east)$);
+	}\IfNoValueF{#3}{^{\,#3}}%
+}
+\newcommand*{\knowInvpv}[2]{\boxedassertpv{#2}[#1]}
+\newcommand*{\ownGhostpv}[2]{\boxedassertpv[dash dot]{#2}[#1]}
+
+\newcommand{\sumpv}[3]{
+  \ownGhostpv\gammaPred{\authfrag{\singletonMap{#1}{(#2, #3)}}}
+}
+
+\newcommand{\pvmapping}[1]{\mathcal{A}\textsf{PVMappings}(#1)}
+\newcommand{\sumapacesfull}[2]{
+  \ownGhost\gammaPreds{\authfull{\singletonMap{#1}{#2}}}
+}
 \begin{figure*}
@@
 \begin{array}{l}
-  \mathcal{I}\textsf{ASpace}(\ptablestore,m)\stackrel{\triangle}{=} \textsf{ASpace\_Lookup}(\ptablestore,\Xi,m) \ast \\
+  \mathcal{I}\textsf{ASpace}(\ptablestore,\Xi,m)\stackrel{\triangle}{=} \textsf{ASpace\_Lookup}(\ptablestore,\Xi,m) \ast \\
   \qquad\qquad\qquad \bigast{(\vaddr, \textsf{paddr})\in \ptablestore}{\exists\;(\textsf{l4e, l3e, l2e, l1e, paddr})\ldotp \textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\vaddr\textsf{, l4e, l3e, l2e, l1e, paddr})} \\
-  \qquad\qquad\qquad \bigast{(\vaddr, \textsf{paddr})\in \Xi}{\exists\; (\textsf{qfrac, val, vaddr}) \ldotp \ulcorner \textsf{vaddr} = \textsf{paddr} + \textsf{KERNBASE} \urcorner \ast  \underbrace{\sumwalkabs\vaddr\qfrac\paddr }_\text{Ghost translation} \ast \\
+  \qquad\qquad\qquad \bigast{(\textsf{paddr}) \in \Xi}{\exists\; (\textsf{qfrac, val, vaddr}) \ldotp \ulcorner \textsf{vaddr} = \textsf{paddr} + \textsf{KERNELBASE} \urcorner \ast  \underbrace{\sumwalkabs\vaddr\qfrac\paddr }_\text{Ghost translation} \ast \\
 \qquad\qquad\qquad\qquad\qquad\qquad\qquad\qquad    \underbrace{\paddr \mapsto_{\mathsf{p}}\{\textsf{qfrac}\}\; \vale}_\text{Physical location} \ast \ulcorner \textsf{qfrac} = 1 \leftrightarrow \; \lnot\textsf{entry\_present }(\vale) \urcorner }\\
@@
    \textsf{ASpace\_Lookup} (\ptablestore,m) \stackrel{\triangle}{=} \lambda\textsf{ cr3val} \ldotp \; \exists \gammaPred \; \ldotp \ulcorner m \; !!\; \textsf{cr3val} = \textsf{Some } \gammaPred \urcorner \ast
-   \ownGhost\gammaPred{\authfull{\ptableabswalk\ptablestore}} \ast \\ \qquad\qquad\qquad\qquad\qquad\qquad\qquad\qquad \ownGhost\gammaPred{\authfull{\ptableabswalk\Xi}}
+   \ownGhost\gammaPred{\authfull{\ptableabswalk\ptablestore}} \ast \\ \qquad\qquad\qquad\qquad\qquad\qquad\qquad\qquad \ownGhostpv\gammaPred{\authfull{\pvmapping\Xi}}
 
@@
 \begin{lstlisting}[mathescape]
-;; Assume the save-space is in rdi, load-space in rsi
-;; First, save yielding context
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast [\rtv'+56](\mathcal{I}\texttt{ASpace}(\theta,m) \ast \texttt{Pother})}_{\rtv}$
-$\specline{ \texttt{rsi}\mapsto_{\textsf{r}} \rtv' \ast \texttt{rdi}\mapsto_{\textsf{r}} \_ \ast \texttt{rbx}\mapsto_{\textsf{r}} \texttt{rbxv} \ast  \texttt{rsp}\mapsto_{\textsf{r}} \texttt{rspv} \ast \texttt{rbp}\mapsto_{\textsf{r}} \texttt{rbpv} }_{\rtv}$
-$\specline{\texttt{r12}\mapsto_{\textsf{r}} \texttt{r12v} \ast \texttt{r13}\mapsto_{\textsf{r}} \texttt{r13v} \ast \texttt{r14}\mapsto_{\textsf{r}} \texttt{r14v} \ast \texttt{r15}\mapsto_{\textsf{r}} \texttt{r15v}}_{\rtv}$
-$\specline{\texttt{rdi+0} \mapsto_{\textsf{v}} \_ \ast \texttt{rdi+8} \mapsto_{\textsf{v}} \_ \ast \texttt{rdi+16} \mapsto_{\textsf{v}} \_ \ast \texttt{rdi+24} \mapsto_{\textsf{v}} \_ \ast \texttt{rdi+32} \mapsto_{\textsf{v}} \_}_{\rtv}$
-$\specline{\texttt{rdi+40} \mapsto_{\textsf{v}} \_ \ast \texttt{rdi+48} \mapsto_{\textsf{v}} \_\ast \texttt{rdi+56} \mapsto_{\textsf{v}} \_}_{\rtv}$
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast [\rtv'+56](\mathcal{I}\texttt{ASpace}(\theta,m) \ast \texttt{Pother})}_{\rtv}$
-$\specline{ \texttt{rdi}\mapsto_{\textsf{r}} \_ \ast \texttt{rbx}\mapsto_{\textsf{r}} \texttt{rbxv} \ast  \texttt{rsp}\mapsto_{\textsf{r}} \texttt{rspv} \ast \texttt{rbp}\mapsto_{\textsf{r}} \texttt{rbpv} }_{\rtv}$
-$\specline{\texttt{r12}\mapsto_{\textsf{r}} \texttt{r12v} \ast \texttt{r13}\mapsto_{\textsf{r}} \texttt{r13v} \ast \texttt{r14}\mapsto_{\textsf{r}} \texttt{r14v} \ast \texttt{r15}\mapsto_{\textsf{r}} \texttt{r15v}}_{\rtv}$
-$\specline{\texttt{rdi+0} \mapsto_{\textsf{v}} \texttt{rbxv} \ast \texttt{rdi+8} \mapsto_{\textsf{v}} \texttt{rspv} \ast \texttt{rdi+16} \mapsto_{\textsf{v}} \texttt{rbpv} \ast \texttt{rdi+24} \mapsto_{\textsf{v}} \texttt{r12v} \ast }_{\rtv}$
-$\specline{\texttt{rdi+32} \mapsto_{\textsf{v}} \texttt{r13v} \ast \texttt{rdi+40} \mapsto_{\textsf{v}} \texttt{r14v} \ast \texttt{rdi+48} \mapsto_{\textsf{v}} \texttt{r15v}\ast \texttt{rdi+56} \mapsto_{\textsf{v}} \rtv}_{\rtv}$
-;; This offsetst physical address at KERNBASE!
-add r,  
-... ;; mov rbp, r12, r13, r14, to offsets 16, 24, 32, and 40 from rsi
-mov r15, 48[rsi]
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast [\rtv'+56](\mathcal{I}\texttt{ASpace}(\theta,m) \ast \texttt{Pother})}_{\rtv}$
-$\specline{ \texttt{rdi}\mapsto_{\textsf{r}} \_ \ast \texttt{rbx}\mapsto_{\textsf{r}} \rtv'+0 \ast  \texttt{rsp}\mapsto_{\textsf{r}} \rtv'+8 \ast \texttt{rbp}\mapsto_{\textsf{r}} \rtv'+16 }_{\rtv}$
-$\specline{\texttt{r12}\mapsto_{\textsf{r}} \rtv'+24 \ast \texttt{r13}\mapsto_{\textsf{r}} \rtv'+32 \ast \texttt{r14}\mapsto_{\textsf{r}} \rtv'+40 \ast \texttt{r15}\mapsto_{\textsf{r}} \rtv'+48}_{\rtv}$
-$\specline{\texttt{rdi+0} \mapsto_{\textsf{v}} \texttt{rbxv} \ast \texttt{rdi+8} \mapsto_{\textsf{v}} \texttt{rspv} \ast \texttt{rdi+16} \mapsto_{\textsf{v}} \texttt{rbpv} \ast \texttt{rdi+24} \mapsto_{\textsf{v}} \texttt{r12v} \ast }_{\rtv}$
-$\specline{\texttt{rdi+32} \mapsto_{\textsf{v}} \texttt{r13v} \ast \texttt{rdi+40} \mapsto_{\textsf{v}} \texttt{r14v} \ast \texttt{rdi+48} \mapsto_{\textsf{v}} \texttt{r15v}\ast \texttt{rdi+56} \mapsto_{\textsf{v}} \rtv}_{\rtv}$
-;; Switch to the new address space
-mov cr3, 56[rsi]
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast [\rtv'+56](\mathcal{I}\texttt{ASpace}(\theta,m) \ast \texttt{Pother}) }_{\rtv}$
-$\specline{ \texttt{rdi}\mapsto_{\textsf{r}} \_ \ast \texttt{rbx}\mapsto_{r} \rtv'+0 \ast  \texttt{rsp}\mapsto_{\textsf{r}} \rtv'+8 \ast \texttt{rbp}\mapsto_{\textsf{r}} \rtv'+16 }_{\rtv}$
-$\specline{\texttt{r12}\mapsto_{\textsf{r}} \rtv'+24 \ast \texttt{r13}\mapsto_{\textsf{r}} \rtv'+32 \ast \texttt{r14}\mapsto_{\textsf{r}} \rtv'+40 \ast \texttt{r15}\mapsto_{\textsf{r}} \rtv'+48}_{\rtv}$
-$\specline{\texttt{rdi+0} \mapsto_{\textsf{v}} \texttt{rbxv} \ast \texttt{rdi+8} \mapsto_{\textsf{v}} \texttt{rspv} \ast \texttt{rdi+16} \mapsto_{\textsf{v}} \texttt{rbpv} \ast  \texttt{rdi+24} \mapsto_{\textsf{v}} \texttt{r12v} \ast }_{\rtv}$
-$\specline{\texttt{rdi+32} \mapsto_{\textsf{v}} \texttt{r13v} \ast \texttt{rdi+40} \mapsto_{\textsf{v}} \texttt{r14v} \ast \texttt{rdi+48} \mapsto_{\textsf{v}} \texttt{r15v}\ast \texttt{rdi+56} \mapsto_{\textsf{v}} \rtv}_{\rtv}$
-$\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\; \sqsubseteq $
-$\specline{ [\rtv](\texttt{Pcurrent}) \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast \texttt{Pother}}_{\rtv'+56}$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,\Xi,m) \ast \ulcorner \textsf{KERNELBASE} + \fpaddr > 0 \urcorner }_{\rtv}$
+$\specline{ \texttt{rip}\mapsto_{\textsf{r}} \textsf{iv} \ast \texttt{rflags}\mapsto_{\textsf{r}} \textsf{flaglist} \ast \texttt{rbx}\mapsto_{\textsf{r}} \fpaddr  }_{\rtv}$
+$\specline{ \sumapaces\rtv\delta  \ast \sumwalkabs\vaddr\qfrac\fpaddr \ast \sumpv\fpaddr\qfrac\true  }_{\rtv}$
+;; This offsetst physical address at KERNELBASE!
+add rbx, KERNELBASE
+$\specline{\exists \textsf{flaglistadd, qphysical, q0, w, flaglistadd, val}, \textsf{P} \ast \texttt{rip}\mapsto_{\textsf{r}} \textsf{iv+7} \ast \texttt{rflags}\mapsto_{\textsf{r}} \textsf{flaglistadd} \ast \texttt{rbx}\mapsto_{\textsf{r}} \fpaddr + \textsf{KERNELBASE} }_{\rtv}$
+$\specline{\bigast{(\vaddr, \textsf{paddr})\in \ptablestore}{\exists\;(\textsf{l4e, l3e, l2e, l1e, } \fpaddr)\ldotp \textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\vaddr\textsf{, l4e, l3e, l2e, l1e, \fpaddr})}}_{\rtv}$
+$\specline{\mathcal{R} -\ast \bigast{(\fpaddr) \in \Xi}{\exists\; (\textsf{qfrac, val, vaddr}) \ldotp \ulcorner \textsf{vaddr} = \fpaddr + \textsf{KERNELBASE} \urcorner \ast \sumwalkabs\vaddr\qfrac\fpaddr  \ast \paddr \mapsto_{\mathsf{p}}\{\textsf{qfrac}\} \; \vale\ast  \ulcorner \textsf{qfrac} = 1 \leftrightarrow \; \lnot\textsf{entry\_present }(\vale) \urcorner}}_{\rtv}$
+$\specline{ \ownGhost\gammaPred{\authfull{\ptableabswalk\ptablestore}} \ast \ownGhostpv\gammaPred{\authfull{\pvmapping\Xi}}}$
+$\specline{ \sumapaces\rtv\delta  \ast\sumwalkabs\vaddr\qfrac\fpaddr \ast \sumpv\fpaddr\qfrac\true \ast \textsf{P}}_{\rtv}$
+$\specline{\vaddr \mapsto_{\texttt{vpte}} \{\textsf{q}, \textsf{qfrac}\} \paddr \;(\texttt{val})}$
 \end{lstlisting}
-\caption{Basic task switch code that switches address spaces.}
+\caption{Identity mapping of a physical address to a virtual one.}
 \label{fig:mapping_code}

### 2023-11-15 examples.tex
+

--- 2023-11-15 examples.tex
@@
 with DMA devices.}
+\newcommand{\qfraczero}{\textsf{qfrac}}
+\newcommand{\true}{\textsf{true}}
+\tikzstyle{boxedassert_border} = [sharp corners,line width=0.2pt]
+\NewDocumentCommand \boxedassertpv {O{} m o}{%
+	\tikz[baseline=(m.base)]{
+		%	  \node[rectangle, draw,inner sep=0.8pt,anchor=base,#1] (m) {${#2}\mathstrut$};
+		\node[rectangle,inner sep=1.5pt,outer sep=0.2pt,anchor=base] (m) {${\,#2\,}\mathstrut$};
+		\draw[#1,boxedassert_border] ($(m.south west)$) rectangle ($(m.north east)$);
+	}\IfNoValueF{#3}{^{\,#3}}%
+}
+\newcommand*{\knowInvpv}[2]{\boxedassertpv{#2}[#1]}
+\newcommand*{\ownGhostpv}[2]{\boxedassertpv[dash dot]{#2}[#1]}
+
+\newcommand{\sumpv}[3]{
+  \ownGhostpv\gammaPred{\authfrag{\singletonMap{#1}{(#2, #3)}}}
+}
+
+\newcommand{\pvmapping}[1]{\mathcal{A}\textsf{PVMappings}(#1)}
+\newcommand{\sumapacesfull}[2]{
+  \ownGhost\gammaPreds{\authfull{\singletonMap{#1}{#2}}}
+}
+\begin{figure*}
+\[
+\begin{array}{l}
+  \mathcal{I}\textsf{ASpace}_{\textsf{id}}(\ptablestore,\Xi,m)\stackrel{\triangle}{=} \textsf{ASpace\_Lookup}(\ptablestore,\Xi,m) \ast \\
+  \qquad\qquad\qquad \bigast{(\vaddr, \textsf{paddr})\in \ptablestore}{\exists\;(\textsf{l4e, l3e, l2e, l1e, paddr})\ldotp \textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\vaddr\textsf{, l4e, l3e, l2e, l1e, paddr})} \\
+  \qquad\qquad\qquad \bigast{(\paddr) \in \Xi}{\exists\; (\textsf{qfrac, q, val,}\vaddr) \ldotp \ulcorner \vaddr = \paddr + \textsf{KERNELBASE} \urcorner \ast  \underbrace{\sumwalkabs\vaddr\qfrac\paddr }_\text{Ghost translation} \ast \\
+\qquad\qquad\qquad\qquad\qquad\qquad\qquad\qquad    \underbrace{\paddr \mapsto_{\mathsf{p}}\{\textsf{qfrac}\}\; \vale}_\text{Physical location} \ast \ulcorner \textsf{qfrac} = 1 \leftrightarrow \; \lnot\textsf{entry\_present }(\vale) \urcorner }\\
+  \textsf{ where } \\
+   \textsf{ASpace\_Lookup}_{\textsf{id}}(\ptablestore,m) \stackrel{\triangle}{=} \lambda\textsf{ cr3val} \ldotp \; \exists \gammaPred \; \ldotp \ulcorner m \; !!\; \textsf{cr3val} = \textsf{Some } \gammaPred \urcorner \ast
+   \ownGhost\gammaPred{\authfull{\ptableabswalk\ptablestore}} \ast \\ \qquad\qquad\qquad\qquad\qquad\qquad\qquad\qquad \ownGhostpv\gammaPred{\authfull{\pvmapping\Xi}}
+
+\end{array}
+\]
+\caption{Global Address-Space Invariant in Figure \ref{fig:peraspaceinvariant} extended with a ghost map bookkeeping identity mappings }
+  \label{fig:peraspaceinvariant_with_p2v_extension}
+  \end{figure*}
 
@@
 
-\todo[inline]{typeset as a triple}
+\begin{figure}\footnotesize
+\begin{lstlisting}[mathescape]
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \ulcorner \textsf{KERNELBASE} + \paddr > 0 \urcorner }_{\rtv}$
+$\specline{ \texttt{rip}\mapsto_{\textsf{r}} \textsf{iv} \ast \texttt{rflags}\mapsto_{\textsf{r}} \textsf{flaglist} \ast \texttt{rbx}\mapsto_{\textsf{r}} \paddr  }_{\rtv}$
+$\specline{ \sumapaces\rtv\delta  \ast \sumwalkabs\vaddr\qfrac\fpaddr \ast \sumpv\fpaddr\qfrac\true  }_{\rtv}$
+;; This offsetst physical address at KERNELBASE!
+add rbx, KERNELBASE
+$\specline{\exists \qfraczero; \qfrac,\textsf{ flaglistadd, w, flaglistadd, val}, \textsf{P} \ast \texttt{rip}\mapsto_{\textsf{r}} \textsf{iv+7} \ast \texttt{rflags}\mapsto_{\textsf{r}} \textsf{flaglistadd} \ast \texttt{rbx}\mapsto_{\textsf{r}} \paddr + \textsf{KERNELBASE} }_{\rtv}$
+$\specline{\bigast{(\vaddr, \paddr)\in \ptablestore}{\exists\;(\textsf{l4e, l3e, l2e, l1e, } \paddr)\ldotp \textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\vaddr\textsf{, l4e, l3e, l2e, l1e,} \paddr)}}_{\rtv}$
+$\specline{\mathcal{R} -\ast \bigast{(\paddr) \in \Xi}{\exists\; (\textsf{qfrac, q, val, va}) \ldotp \ulcorner \textsf{va} = \paddr + \textsf{KERNELBASE} \urcorner \ast \sumwalkabs\vaddr\qfrac\paddr  \ast \paddr \mapsto_{\mathsf{p}}\{\textsf{qfrac}\} \; \vale\ast  \ulcorner \textsf{qfrac} = 1 \leftrightarrow \; \lnot\textsf{entry\_present }(\vale) \urcorner}}_{\rtv}$
+$\specline{ \ownGhost\gammaPred{\authfull{\ptableabswalk\ptablestore}} \ast \ownGhostpv\gammaPred{\authfull{\pvmapping\Xi}}}$
+$\specline{ \sumapaces\rtv\delta  \ast\sumwalkabs\vaddr\qfrac\paddr \ast \sumpv\paddr\qfraczero\true \ast \textsf{P}}_{\rtv}$
+$\specline{\vaddr \mapsto_{\texttt{vpte}} \{\textsf{q}, \textsf{qfrac}\} \paddr \;(\texttt{val})}$
+\end{lstlisting}
+\caption{Identity mapping of a physical address to a virtual one.}
+\label{fig:mapping_code}
+\end{figure}
 

### 2023-11-15 examples.tex
+

@@
 $\specline{ \sumapaces\rtv\delta  \ast \sumwalkabs\vaddr\qfrac\fpaddr \ast \sumpv\fpaddr\qfrac\true  }_{\rtv}$
-;; This offsetst physical address at KERNELBASE!
+;; This offsets the physical address bt KERNELBASE!
 add rbx, KERNELBASE

### 2023-11-16 programlogic.tex
+

@@
 \begin{mathpar}
-  \inferrule[WriteToRegFromReg]{
-  \{P \ast r_d \mapsto_{r} \textsf{rvs} \ast r_s \mapsto_{r}\{q\} \textsf{ rvs} \}_{\textsf{rtv}}\;\overline{ is}
-}{
-  \{P \ast r_d \mapsto_{r} \textsf{rvd} \ast r_s \mapsto_{r}\{q\} \textsf{ rvs} \}_{\textsf{rtv}}
-  \textsf{ mov}~\textsf{r}_d,~\textsf{r}_s;\;\overline{is}
-  %\lstinline|mov %cr3, r| 
-}
-\\
+%  \inferrule[AddRegImm]{
+%  \{P \ast r_d \mapsto_{r} \textsf{rvs} \ast r_s \mapsto_{r}\{q\} \textsf{ rvs} \}_{\textsf{rtv}}\;\overline{ is}
+%}{
+%  \{P \ast r_d \mapsto_{r} \textsf{rvd} \ast r_s \mapsto_{r}\{q\} \textsf{ rvs} \}_{\textsf{rtv}}
+%  \textsf{ mov}~\textsf{r}_d,~\textsf{r}_s;\;\overline{is}
+%  \lstinline|mov %cr3, r| 
+%}
+%\\
 % \inferrule[WriteToPhysMemFromReg]{
@@
 \inferrule[WriteToRegFromVirtMem]{
-  \{P \ast \mathcal{I}\textsf{ASpace}\ast \mathcal{I}\textsf{ASpace} \ast r_d \mapsto_{r}  \textsf{v} \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{v,rtv}} \textsf{v} \}_{\textsf{rtv}}\;\overline{is}
+  \{P \ast \mathcal{I}\textsf{ASpace}\ast \mathcal{I}\textsf{ASpace} \ast \textsf{rip} \mapsto_{\textsf{r}} \textsf{iv + x} \ast \ulcorner x = 3 \lor x =4 \urcorner \ast r_d \mapsto_{r}  \textsf{v} \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{v,rtv}} \textsf{v} \}_{\textsf{rtv}}\;\overline{is}
 }{
-  \{P \ast \mathcal{I}\textsf{ASpace}\ast r_d \mapsto_{r}  \textsf{rvd} \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{v}} \textsf{v} \}_{\textsf{rtv}}
+  \{P \ast \mathcal{I}\textsf{ASpace}\ast \textsf{rip} \mapsto_{\textsf{r}} \textsf{iv} \ast r_d \mapsto_{r}  \textsf{rvd} \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{v}} \textsf{v} \}_{\textsf{rtv}}\;
   \textsf{mov}~\textsf{r}_d,~[\textsf{r}_a];\;\overline{is}
@@
 \inferrule[WriteToVirtMemFromReg]{
-  \{P \ast \mathcal{I}\textsf{ASpace}\ast r_s \mapsto_{r}\{q\}  \textsf{rvs}  \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{v}} \textsf{v} \}_{\textsf{rtv}}\;\overline{ is}
+  \{P \ast \mathcal{I}\textsf{ASpace}\ast \textsf{rip} \mapsto_{\textsf{r}} \textsf{iv + x} \ast \ulcorner x = 3 \lor x =4 \urcorner \ast r_s \mapsto_{r}\{q\}  \textsf{rvs}  \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{v}} \textsf{v} \}_{\textsf{rtv}}\;\overline{ is}
 }{
-  \{P \ast \mathcal{I}\textsf{ASpace}\ast r_s \mapsto_{r}\{q\}  \textsf{rvs}   \ast r_a \mapsto_{r}\{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{v}} \textsf{rvs} \}_{\textsf{rtv}}
+  \{P \ast \mathcal{I}\textsf{ASpace}\ast \textsf{rip} \mapsto_{\textsf{r}} \textsf{iv} \ast r_s \mapsto_{r}\{q\}  \textsf{rvs}   \ast r_a \mapsto_{r}\{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{v}} \textsf{rvs} \}_{\textsf{rtv}}
   \textsf{ mov}~[\textsf{r}_a],~\textsf{r}_s;\;\overline{is}
@@
 \\
-\inferrule[WriteToRegCtlFromReg]{
-  \{P \ast r_s \mapsto_{r}\{q\}  \textsf{ rvs}  \}_{\textsf{rvs}} \overline{is}
+%\inferrule[WriteToRegCtlFromReg]{
+%  \{P \ast r_s \mapsto_{r}\{q\}  \textsf{ rvs}  \}_{\textsf{rvs}} \overline{is}
+%}{
+%  \{P \ast r_s \mapsto_{r}\{q\}  \textsf{ rvs}   \}_{\textsf{rtv}}
+%  \textsf{mov}~\textsf{cr3},~r_s;\;\overline{is}
+  %\lstinline|mov %cr3, r| 
+%}\\
+\inferrule[WriteToRegCtlFromRegModal]{
+  \{[\textsf{rtv}](P \ast \mathcal{I}\textsf{ASpace})\ast \textsf{rip} \mapsto_{\textsf{r}} \textsf{iv + 4} \ast \mathcal{I}\textsf{ASpace} \ast R\ast r_s \mapsto_{r}\{q\}  \textsf{ rvs}  \}_{\textsf{rvs}} \;\overline{is}
 }{
-  \{P \ast r_s \mapsto_{r}\{q\}  \textsf{ rvs}   \}_{\textsf{rtv}}
-  \textsf{mov}~\textsf{cr3},~r_s;\;\overline{is}
-  %\lstinline|mov %cr3, r| 
-}\\
-\inferrule[WriteToRegCtlFromRegModal]{
-  \{[\textsf{rtv}](P \ast \mathcal{I}\textsf{ASpace})\ast \mathcal{I}\textsf{ASpace} \ast R\ast r_s \mapsto_{r}\{q\}  \textsf{ rvs}  \}_{\textsf{rvs}} \overline{is}
-}{
-  \{P \ast \mathcal{I}\textsf{ASpace} \ast [\textsf{rvs}](R \ast\mathcal{I}\textsf{ASpace})\ast r_s \mapsto_{r}\{q\}  \textsf{ rvs}   \}_{\textsf{rtv}}
+  \{P \ast \mathcal{I}\textsf{ASpace} \ast \textsf{rip} \mapsto_{\textsf{r}} \textsf{iv} \ast [\textsf{rvs}](R \ast\mathcal{I}\textsf{ASpace})\ast r_s \mapsto_{r}\{q\}  \textsf{ rvs}   \}_{\textsf{rtv}} \;
   \textsf{mov}~\textsf{cr3},~r_s;\;\overline{is}
@@
 \inferrule[WriteToRegFromRegCtl]{
-  \{P \ast r_d \mapsto_{r} \textsf{rvs} \ast r_s \mapsto_{r}\{q\} \textsf{ rvs} \}_{\textsf{rtv}}\;\overline{ is}
+  \{P \ast  \textsf{rip} \mapsto_{\textsf{r}} \textsf{iv + 4} \ast r_d \mapsto_{r} \textsf{rvs} \ast r_s \mapsto_{r}\{q\} \textsf{ rvs} \}_{\textsf{rtv}}\;\overline{ is}
 }{
-  \{P \ast r_d \mapsto_{r} \textsf{rvd} \ast r_s \mapsto_{r}\{q\} \textsf{ rvs} \}_{\textsf{rtv}}
+  \{P \ast  \textsf{rip} \mapsto_{\textsf{r}} \textsf{iv} \ast r_d \mapsto_{r} \textsf{rvd} \ast r_s \mapsto_{r}\{q\} \textsf{ rvs} \}_{\textsf{rtv}}
   \textsf{ mov}~\textsf{r}_d,~\textsf{r}_s;\;\overline{is}

### 2023-11-16 programlogic.tex
+

@@
 % \\
-% \inferrule[WriteToRegFromPhysMem]{
-%   \{P \ast r_d \mapsto_{r}  \textsf{v} \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{t}} \textsf{v} \}_{\textsf{rtv}}\;\overline{is}
-% }{
-%   \{P \ast r_d \mapsto_{r}  \textsf{rvd} \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{t}} \textsf{v} \}_{\textsf{rtv}}
-%   \textsf{mov}~\textsf{r}_d,~[\textsf{r}_a];\;\overline{is}
-% }
-% \\
+ \inferrule[AddRegImm]{
+   \{P \ast (r_d \mapsto_{r}  \textsf{rvd+imm} \lor  r_d \mapsto_{r} 0 ) \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{t}} \textsf{v} \}_{\textsf{rtv}}\;\overline{is}
+ }{
+   \{P \ast \textsf{rip} \mapsto_{\textsf{r}} \textsf{iv} \ast \textsf{rflags} \mapsto_{\textsf{r}} \textsf{flgs} \ast \ulcorner (\textsf{rvd + imm} > 0 \lor \textsf{rvd + imm} = 0 ) \urcorner \ast r_d \mapsto_{r}  \textsf{rvd} \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{t}} \textsf{v} \}_{\textsf{rtv}}
+   \textsf{add}~\textsf{r}_d,~[\textsf{imm}];\;\overline{is}
+ }
+ \\
 \inferrule[WriteToRegFromVirtMem]{

### 2023-11-16 programlogic.tex
+

@@
  \inferrule[AddRegImm]{
-   \{P \ast (r_d \mapsto_{r}  \textsf{rvd+imm} \lor  r_d \mapsto_{r} 0 ) \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{t}} \textsf{v} \}_{\textsf{rtv}}\;\overline{is}
+   \{P \ast\textsf{rip} \mapsto_{\textsf{r}} \textsf{iv+7} \ast  \ast (r_d \mapsto_{r}  \textsf{rvd+imm} \lor  r_d \mapsto_{r} 0 ) \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{t}} \textsf{v} \}_{\textsf{rtv}}\;\overline{is}
  }{

### 2023-11-16 programlogic.tex
sync

@@
  }{
-   \{P \ast \textsf{rip} \mapsto_{\textsf{r}} \textsf{iv} \ast \textsf{rflags} \mapsto_{\textsf{r}} \textsf{flgs} \ast \ulcorner (\textsf{rvd + imm} > 0 \lor \textsf{rvd + imm} = 0 ) \urcorner \ast r_d \mapsto_{r}  \textsf{rvd} \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{t}} \textsf{v} \}_{\textsf{rtv}}
+   \{\begin{array}{l}
+   P \ast \textsf{rip} \mapsto_{\textsf{r}} \textsf{iv} \ast \textsf{rflags} \mapsto_{\textsf{r}} \textsf{flgs} \ast \ulcorner (\textsf{rvd + imm} > 0 \lor \textsf{rvd + imm} = 0 ) \urcorner \ast \\
+   r_d \mapsto_{r}  \textsf{rvd} \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{t}} \textsf{v}
+   \end{array} \}_{\textsf{rtv}}
    \textsf{add}~\textsf{r}_d,~[\textsf{imm}];\;\overline{is}

### 2023-11-16 programlogic.tex
fig 7 fixed

@@
 \begin{figure}
+  \makeatletter % allow us to mention @-commands
+\def\arcr{\@arraycr}
+\makeatother
 \begin{mathpar}
@@
  }{
-   \{\begin{array}{l}
-   P \ast \textsf{rip} \mapsto_{\textsf{r}} \textsf{iv} \ast \textsf{rflags} \mapsto_{\textsf{r}} \textsf{flgs} \ast \ulcorner (\textsf{rvd + imm} > 0 \lor \textsf{rvd + imm} = 0 ) \urcorner \ast \\
+   \left\{
+   \begin{array}{l}
+   P \ast \textsf{rip} \mapsto_{\textsf{r}} \textsf{iv} \ast \textsf{rflags} \mapsto_{\textsf{r}} \textsf{flgs} \ast \ulcorner (\textsf{rvd + imm} > 0 \lor \textsf{rvd + imm} = 0 ) \urcorner \ast \arcr
    r_d \mapsto_{r}  \textsf{rvd} \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{t}} \textsf{v}
-   \end{array} \}_{\textsf{rtv}}
+ \end{array}\right\}_{\textsf{rtv}}  \;
    \textsf{add}~\textsf{r}_d,~[\textsf{imm}];\;\overline{is}
@@
 \inferrule[WriteToRegFromVirtMem]{
-  \{P \ast \mathcal{I}\textsf{ASpace}\ast \mathcal{I}\textsf{ASpace} \ast \textsf{rip} \mapsto_{\textsf{r}} \textsf{iv + x} \ast \ulcorner x = 3 \lor x =4 \urcorner \ast r_d \mapsto_{r}  \textsf{v} \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{v,rtv}} \textsf{v} \}_{\textsf{rtv}}\;\overline{is}
+  \left\{ \begin{array}{l}
+    P \ast \mathcal{I}\textsf{ASpace}\ast \mathcal{I}\textsf{ASpace} \ast \textsf{rip} \mapsto_{\textsf{r}} \textsf{iv + x} \ast \ulcorner x = 3 \lor x =4 \urcorner \ast \arcr
+    r_d \mapsto_{r}  \textsf{v} \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{v,rtv}} \textsf{v}
+  \end{array} \right\}_{\textsf{rtv}}\;\overline{is}
 }{

### 2023-11-16 examples.tex
drop text that is now redundant with relwork

@@
 but ought to turn out this way}
+This is a simple piece of code whose functionality is critical and whose correctness is highly non-trivial. No prior work engages with this problem.
 
-\citet{kolanski08vstte,kolanski09tphols} verified a single code block with their logic which was roughly Figure \ref{fig:mapping_code} for a 2-level ARM
-page table, but several critical complexities our work deals with were not addressed.
-First, beyond the limitations discussed in Section \ref{sec:overly-restrictive}, Kolanski and Klein assumed that virtual addresses
-for page tables at each level were given as parameters rather than verifying any conversion from physical addresses to virtual addresses (or even axiomatizing their lookup).
-In contrast, our verification articulates the address space invariant from which the physical-to-virtual translation can be implemented.
-Second, our proof deals with the construction of a valid virtual points-to \emph{to the PTE to update} in mapping, which Kolanski and Klein also
-assumed was given.
-\todo{some of this is really an argument for our verification being more thorough, rather than being about our logic}
+% \citet{kolanski08vstte,kolanski09tphols} verified a single code block with their logic which was roughly Figure \ref{fig:mapping_code} for a 2-level ARM
+% page table, but several critical complexities our work deals with were not addressed.
+% First, beyond the limitations discussed in Section \ref{sec:overly-restrictive}, Kolanski and Klein assumed that virtual addresses
+% for page tables at each level were given as parameters rather than verifying any conversion from physical addresses to virtual addresses (or even axiomatizing their lookup).
+% In contrast, our verification articulates the address space invariant from which the physical-to-virtual translation can be implemented.
+% Second, our proof deals with the construction of a valid virtual points-to \emph{to the PTE to update} in mapping, which Kolanski and Klein also
+% assumed was given.
+% \todo{some of this is really an argument for our verification being more thorough, rather than being about our logic}
 
-Reasoning about the page table walk in their logic would have required 
-could reason about the walk, but would need to explicitly prove that all other invariants
-of the kernel, the current address space, and all other address spaces of interest were preserved by each update, because their model
-only supports separation within a single address space. In our model, this follows for free from making
-our separation logic directly aware of address translation and internalizing assumptions about other address spaces as further separable assertions.
-Kolanski and Klein did address part of the walk information for a 2-level page table (a possible ARM configuration), but
+% Reasoning about the page table walk in their logic would have required 
+% could reason about the walk, but would need to explicitly prove that all other invariants
+% of the kernel, the current address space, and all other address spaces of interest were preserved by each update, because their model
+% only supports separation within a single address space. In our model, this follows for free from making
+% our separation logic directly aware of address translation and internalizing assumptions about other address spaces as further separable assertions.
+% Kolanski and Klein did address part of the walk information for a 2-level page table (a possible ARM configuration), but
 
-\textsc{seL4} currently still trusts address translations; it models page tables as a data structure in regular memory, thus not capturing the possibility that even
-temporarily destroying the mappings and restoring them can actually crash the OS. \textsc{CertiKOS} papers share little in the way of precise details about
-their virtual memory management, but because their core technology is based on a fork of \textsc{CompCert}, whose model of memory is
-a set of unordered block allocations, we can infer their proofs must also trust these translations.
+% \textsc{seL4} currently still trusts address translations; it models page tables as a data structure in regular memory, thus not capturing the possibility that even
+% temporarily destroying the mappings and restoring them can actually crash the OS. \textsc{CertiKOS} papers share little in the way of precise details about
+% their virtual memory management, but because their core technology is based on a fork of \textsc{CompCert}, whose model of memory is
+% a set of unordered block allocations, we can infer their proofs must also trust these translations.
 

### 2023-11-16 examples.tex
+

--- 2023-11-15 examples.tex
@@
 $\specline{ \texttt{rip}\mapsto_{\textsf{r}} \textsf{iv} \ast \texttt{rflags}\mapsto_{\textsf{r}} \textsf{flaglist} \ast \texttt{rbx}\mapsto_{\textsf{r}} \paddr  }_{\rtv}$
-$\specline{ \sumapaces\rtv\delta  \ast \sumwalkabs\vaddr\qfrac\fpaddr \ast \sumpv\fpaddr\qfrac\true  }_{\rtv}$
-;; This offsetst physical address at KERNELBASE!
+$\specline{ \sumapaces\rtv\delta  \ast \sumwalkabs\vaddr\qfrac\paddr \ast \sumpv\paddr\qfrac\true  }_{\rtv}$
+;; This offsets the physical address bt KERNELBASE!
 add rbx, KERNELBASE

### 2023-11-16 examples.tex
+

@@
 $\specline{ \texttt{rip}\mapsto_{\textsf{r}} \textsf{iv} \ast \texttt{rflags}\mapsto_{\textsf{r}} \textsf{flaglist} \ast \texttt{rbx}\mapsto_{\textsf{r}} \paddr  }_{\rtv}$
-$\specline{ \sumapaces\rtv\delta  \ast \sumwalkabs\vaddr\qfrac\paddr \ast \sumpv\paddr\qfrac\true  }_{\rtv}$
+$\specline{ \sumapaces\rtv\delta  \ast \sumpv\paddr\qfrac\true  }_{\rtv}$
 ;; This offsets the physical address bt KERNELBASE!
@@
 $\specline{ \ownGhost\gammaPred{\authfull{\ptableabswalk\ptablestore}} \ast \ownGhostpv\gammaPred{\authfull{\pvmapping\Xi}}}$
-$\specline{ \sumapaces\rtv\delta  \ast\sumwalkabs\vaddr\qfrac\paddr \ast \sumpv\paddr\qfraczero\true \ast \textsf{P}}_{\rtv}$
+$\specline{ \sumapaces\rtv\delta  \ast\sumwalkabs\vaddr\qfrac\paddr \ast \sumpv\paddr\qfraczero\true \ast  \paddr \mapsto_{\mathsf{p}}\{\textsf{qfrac}\} \; \vale \ast \textsf{P}}_{\rtv}$
+$\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\; \sqsubseteq $
+$\specline{\exists \qfraczero; \qfrac,\textsf{ flaglistadd, w, flaglistadd, val}, \textsf{P} \ast \texttt{rip}\mapsto_{\textsf{r}} \textsf{iv+7} \ast \texttt{rflags}\mapsto_{\textsf{r}} \textsf{flaglistadd} \ast \texttt{rbx}\mapsto_{\textsf{r}} \paddr + \textsf{KERNELBASE} }_{\rtv}$
+$\specline{\bigast{(\vaddr, \paddr)\in \ptablestore}{\exists\;(\textsf{l4e, l3e, l2e, l1e, } \paddr)\ldotp \textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\vaddr\textsf{, l4e, l3e, l2e, l1e,} \paddr)}}_{\rtv}$
+$\specline{\mathcal{R} -\ast \bigast{(\paddr) \in \Xi}{\exists\; (\textsf{qfrac, q, val, va}) \ldotp \ulcorner \textsf{va} = \paddr + \textsf{KERNELBASE} \urcorner \ast \sumwalkabs\vaddr\qfrac\paddr  \ast \paddr \mapsto_{\mathsf{p}}\{\textsf{qfrac}\} \; \vale\ast  \ulcorner \textsf{qfrac} = 1 \leftrightarrow \; \lnot\textsf{entry\_present }(\vale) \urcorner}}_{\rtv}$
+$\specline{ \ownGhost\gammaPred{\authfull{\ptableabswalk\ptablestore}} \ast \ownGhostpv\gammaPred{\authfull{\pvmapping\Xi}}}$
+$\specline{ \sumapaces\rtv\delta \ast \sumpv\paddr\qfraczero\true \ast \textsf{P}}_{\rtv}$
 $\specline{\vaddr \mapsto_{\texttt{vpte}} \{\textsf{q}, \textsf{qfrac}\} \paddr \;(\texttt{val})}$

### 2023-11-16 programlogic.tex
clean up framing discussion

--- 2023-11-16 programlogic.tex
@@
 
-This is then how we justify our actual rule for \lstinline|cr3| updates:
-\todo[inline]{Ismail: Can we chat Wednesday about this rule, which has cr3 register pointstos instead of the cr3 subscripts?
-The subscript form might make the derivation of a frame rule harder.}
+This is then how we justify our actual rule for \lstinline|cr3| updates, which is \emph{almost}
 \begin{mathpar}
 \inferrule*[left=ChangeAddressSpace]{
-  \{[r_1](P) \ast cr3 \mapsto_{r} r_2 \ast r \mapsto_{r} r_2 \ast Q\}\overline{is}
+  \{[r_1](P) \ast cr3 \mapsto_{r} r_2 \ast r \mapsto_{r} r_2 \ast Q\}\;\overline{is}
 }{
-  \{P \ast cr3 \mapsto_{r} r_1 \ast r \mapsto_{r} r_2 \ast [r_2](Q)\}
+  \{P \ast cr3 \mapsto_{r} r_1 \ast r \mapsto_{r} r_2 \ast [r_2](Q)\}\;
   \texttt{mov}~\texttt{\%cr3},~r;\;\overline{is}
@@
 Because the precondition on this rule is global, we avoid issues with framing.
+A more typical frame rule can be recovered from Hoare doubles in the typical way~\cite{Chlipala2013Bedrock} A more typical frame rule can be recovered from Hoare doubles in the typical way~\cite{Chlipala2013Bedrock}
+with the added restriction that \lstinline|cr3| is restored to its pre-framing value before un-framing (in this case, the framed-out assertion is effectively
+stashed under an other-space assertion for the initial address space, and pulled out as part of restoring).
 
-If we wanted to consider a frame rule that would work for this logic, we could consider:
-\begin{mathpar}
-  \inferrule*[left=Cr3Frame]{
-    \{P\ast cr3=v\}\;C\;\{ Q \ast cr3=v\}
-  }{
-    \{R\ast P\ast cr3=v\}\;C\;\{ R \ast Q \ast cr3=v\}
-  }
-\end{mathpar}
-By demanding that \lstinline|cr3| be held constant (or rather, at least restored to its original value) we could frame almost traditionally.
+% If we wanted to consider a frame rule that would work for this logic, we could consider:
+% \begin{mathpar}
+%   \inferrule*[left=Cr3Frame]{
+%     \{P\ast cr3=v\}\;C\;\{ Q \ast cr3=v\}
+%   }{
+%     \{R\ast P\ast cr3=v\}\;C\;\{ R \ast Q \ast cr3=v\}
+%   }
+% \end{mathpar}
+% By demanding that \lstinline|cr3| be held constant (or rather, at least restored to its original value) we could frame almost traditionally.
 In particular, this rule would work with framing around calls that might lead to address space switches, such as calling blocking operations in the kernel.
-Attempting to derive a frame rule as some work using Hoare doubles does~\cite{Chlipala2013Bedrock}
-would enforce this restriction.
+% Attempting to derive a frame rule as some work using Hoare doubles does~\cite{Chlipala2013Bedrock}
+% would enforce this restriction.
 
-Another useful perspective on this might be the dynamic frames~\cite{parkinson2011relationship} 
-view: virtual points-to assertions are self-stable \emph{except} for changes in \lstinline|cr3|, so framing would then naturally require other means of holding \lstinline|cr3| constant (or saving and restoring it).
+% Another useful perspective on this might be the dynamic frames~\cite{parkinson2011relationship} 
+% view: virtual points-to assertions are self-stable \emph{except} for changes in \lstinline|cr3|, so framing would then naturally require other means of holding \lstinline|cr3| constant (or saving and restoring it).
 % Virtual points-to assertions could be made self-stable by also giving them partial ownership over \lstinline|cr3| assertions, but this would require explicitly plumbing that ownership from \emph{all} assertions back to any place an address space change might occur; this would seem to be a far graver loss of modularity than this extra quirk in framing discussions.

### 2023-11-16 examples.tex

--- 2023-11-16 examples.tex
@@
 $\specline{ \texttt{rip}\mapsto_{\textsf{r}} \textsf{iv} \ast \texttt{rflags}\mapsto_{\textsf{r}} \textsf{flaglist} \ast \texttt{rbx}\mapsto_{\textsf{r}} \paddr  }_{\rtv}$
-$\specline{ \sumapaces\rtv\delta  \ast \sumwalkabs\vaddr\qfrac\fpaddr \ast \sumpv\fpaddr\qfrac\true  }_{\rtv}$
+$\specline{ \sumapaces\rtv\delta  \ast \sumpv\paddr\qfrac\true  }_{\rtv}$
 ;; This offsets the physical address bt KERNELBASE!

### 2023-11-16 programlogic.tex

@@
  }{
-   \{P \ast \textsf{rip} \mapsto_{\textsf{r}} \textsf{iv} \ast \textsf{rflags} \mapsto_{\textsf{r}} \textsf{flgs} \ast \ulcorner (\textsf{rvd + imm} > 0 \lor \textsf{rvd + imm} = 0 ) \urcorner \ast r_d \mapsto_{r}  \textsf{rvd} \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{t}} \textsf{v} \}_{\textsf{rtv}}
+   \left\{
+   \begin{array}{l}
+   P \ast \textsf{rip} \mapsto_{\textsf{r}} \textsf{iv} \ast \textsf{rflags} \mapsto_{\textsf{r}} \textsf{flgs} \ast \ulcorner (\textsf{rvd + imm} > 0 \lor \textsf{rvd + imm} = 0 ) \urcorner \ast \arcr
+   r_d \mapsto_{r}  \textsf{rvd} \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{t}} \textsf{v}
+ \end{array}\right\}_{\textsf{rtv}}  \;
    \textsf{add}~\textsf{r}_d,~[\textsf{imm}];\;\overline{is}

### 2023-11-16 examples.tex
significantly condense task switch code, slightly condense its text

@@
 \subsection{Change of Address Space}
-A critical piece of \emph{trusted} code in verified OS kernels is the assembly code to change the current address space; current verified OS kernels currently lack effective ways to specify and reason about this low-level operation, for reasons outlined in Section \ref{sec:relwork}.
+A critical piece of \emph{trusted} code in current verified OS kernels is the assembly code to change the current address space; current verified OS kernels currently 
+lack effective ways to specify and reason about this low-level operation, for reasons outlined in Section \ref{sec:relwork}.
 \begin{figure}\footnotesize
 \begin{lstlisting}[mathescape]
-;; Assume the save-space is in rdi, load-space in rsi
-;; First, save yielding context
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast [\rtv'+56](\mathcal{I}\texttt{ASpace}(\theta,m) \ast \texttt{Pother})}_{\rtv}$
-$\specline{ \texttt{rsi}\mapsto_{\textsf{r}} \rtv' \ast \texttt{rdi}\mapsto_{\textsf{r}} \_ \ast \texttt{rbx}\mapsto_{\textsf{r}} \texttt{rbxv} \ast  \texttt{rsp}\mapsto_{\textsf{r}} \texttt{rspv} \ast \texttt{rbp}\mapsto_{\textsf{r}} \texttt{rbpv} }_{\rtv}$
-$\specline{\texttt{r12}\mapsto_{\textsf{r}} \texttt{r12v} \ast \texttt{r13}\mapsto_{\textsf{r}} \texttt{r13v} \ast \texttt{r14}\mapsto_{\textsf{r}} \texttt{r14v} \ast \texttt{r15}\mapsto_{\textsf{r}} \texttt{r15v}}_{\rtv}$
-$\specline{\texttt{rdi+0} \mapsto_{\textsf{v}} \_ \ast \texttt{rdi+8} \mapsto_{\textsf{v}} \_ \ast \texttt{rdi+16} \mapsto_{\textsf{v}} \_ \ast \texttt{rdi+24} \mapsto_{\textsf{v}} \_ \ast \texttt{rdi+32} \mapsto_{\textsf{v}} \_}_{\rtv}$
-$\specline{\texttt{rdi+40} \mapsto_{\textsf{v}} \_ \ast \texttt{rdi+48} \mapsto_{\textsf{v}} \_\ast \texttt{rdi+56} \mapsto_{\textsf{v}} \_}_{\rtv}$
+;; Assume the save-space is in rdi, load-space in rsi. First, save the yielding context
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast [\rtv'](\mathcal{I}\texttt{ASpace}(\theta',m') \ast \texttt{Pother}) \ast \texttt{rsi}\mapsto_{\textsf{r}} old \ast \texttt{rdi}\mapsto_{\tnewoldr}} old \ast \texttt{rbx}\mapsto_{\textsf{r}} \texttt{rbxv} }_{\rtv}$
+$\specline{\texttt{rsp}\mapsto_{\textsf{r}} \texttt{rspv} \ast \texttt{rbp}\mapsto_{\textsf{r}} \texttt{rbpv} \ast \texttt{r12}\mapsto_{\textsf{r}} \texttt{r12v} \ast \texttt{r13}\mapsto_{\textsf{r}} \texttt{r13v} \ast \texttt{r14}\mapsto_{\textsf{r}} \texttt{r14v} \ast \texttt{r15}\mapsto_{\textsf{r}} \texttt{r15v}}_{\rtv}$
+$\specline{\mathsf{ContextAt}(old,\_) \ast \mathsf{ContextAt}(new,[\mathsf{rbxv}',\ldots,\rtv']) }_{\rtv}$
 mov 0[rdi], rbx
@@
 mov 48[rdi], r15
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast [\rtv'+56](\mathcal{I}\texttt{ASpace}(\theta,m) \ast \texttt{Pother})}_{\rtv}$
-$\specline{ \texttt{rdi}\mapsto_{\textsf{r}} \_ \ast \texttt{rbx}\mapsto_{\textsf{r}} \texttt{rbxv} \ast  \texttt{rsp}\mapsto_{\textsf{r}} \texttt{rspv} \ast \texttt{rbp}\mapsto_{\textsf{r}} \texttt{rbpv} }_{\rtv}$
-$\specline{\texttt{r12}\mapsto_{\textsf{r}} \texttt{r12v} \ast \texttt{r13}\mapsto_{\textsf{r}} \texttt{r13v} \ast \texttt{r14}\mapsto_{\textsf{r}} \texttt{r14v} \ast \texttt{r15}\mapsto_{\textsf{r}} \texttt{r15v}}_{\rtv}$
-$\specline{\texttt{rdi+0} \mapsto_{\textsf{v}} \texttt{rbxv} \ast \texttt{rdi+8} \mapsto_{\textsf{v}} \texttt{rspv} \ast \texttt{rdi+16} \mapsto_{\textsf{v}} \texttt{rbpv} \ast \texttt{rdi+24} \mapsto_{\textsf{v}} \texttt{r12v} \ast}_{\rtv}$
-$\specline{ \texttt{rdi+32} \mapsto_{\textsf{v}} \texttt{r13v} \ast \texttt{rdi+40} \mapsto_{v} \texttt{r14v} \ast \texttt{rdi+48} \mapsto_{\textsf{v}} \texttt{r15v}\ast \texttt{rdi+56} \mapsto_{\textsf{v}} \_}_{\rtv}$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast [\rtv'](\mathcal{I}\texttt{ASpace}(\theta',m') \ast \texttt{Pother})  \ast  \texttt{rsi}\mapsto_{\textsf{r}} new \ast \texttt{rdi}\mapsto_{\textsf{r}} old \ast \texttt{rbx}\mapsto_{\textsf{r}} \texttt{rbxv} }_{\rtv}$
+$\specline{ \texttt{rsp}\mapsto_{\textsf{r}} \texttt{rspv} \ast \texttt{rbp}\mapsto_{\textsf{r}} \texttt{rbpv} \ast \texttt{r12}\mapsto_{\textsf{r}} \texttt{r12v} \ast \texttt{r13}\mapsto_{\textsf{r}} \texttt{r13v} \ast \texttt{r14}\mapsto_{\textsf{r}} \texttt{r14v} \ast \texttt{r15}\mapsto_{\textsf{r}} \texttt{r15v}}_{\rtv}$
+$\specline{\mathsf{ContextAt}(old,[\mathsf{rbxv,\ldots,\rtv}]) \ast \mathsf{ContextAt}(new,[\mathsf{rbxv}',\ldots,\rtv']) }_{\rtv}$
 mov 56[%rdi], %cr3
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast [\rtv'+56](\mathcal{I}\texttt{ASpace}(\theta,m) \ast \texttt{Pother})}_{\rtv}$
-$\specline{ \texttt{rdi}\mapsto_{\textsf{r}} \_ \ast \texttt{rbx}\mapsto_{\textsf{r}} \texttt{rbxv} \ast  \texttt{rsp}\mapsto_{\textsf{r}} \texttt{rspv} \ast \texttt{rbp}\mapsto_{\textsf{r}} \texttt{rbpv} }_{\rtv}$
-$\specline{\texttt{r12}\mapsto_{\textsf{r}} \texttt{r12v} \ast \texttt{r13}\mapsto_{\textsf{r}} \texttt{r13v} \ast \texttt{r14}\mapsto_{\textsf{r}} \texttt{r14v} \ast \texttt{r15}\mapsto_{\textsf{r}} \texttt{r15v}}_{\rtv}$
-$\specline{\texttt{rdi+0} \mapsto_{\textsf{v}} \texttt{rbxv} \ast \texttt{rdi+8} \mapsto_{\textsf{v}} \texttt{rspv} \ast \texttt{rdi+16} \mapsto_{\textsf{v}} \texttt{rbpv} \ast \texttt{rdi+24} \mapsto_{\textsf{v}} \texttt{r12v} \ast }_{\rtv}$
-$\specline{\texttt{rdi+32} \mapsto_{\textsf{v}} \texttt{r13v} \ast \texttt{rdi+40} \mapsto_{\textsf{v}} \texttt{r14v} \ast \texttt{rdi+48} \mapsto_{\textsf{v}} \texttt{r15v}\ast \texttt{rdi+56} \mapsto_{\textsf{v}} \rtv}_{\rtv}$
+$\specline{\ldots \ast \texttt{rdi+56} \mapsto_{\textsf{v}} \rtv}_{\rtv}$    
 ;; Restore target context
 mov rbx, 0[rsi] 
-;; This switches to a new stack, which may not be mapped in the current address space!
-mov rsp, 8[rsi] 
-... ;; mov rbp, r12, r13, r14, to offsets 16, 24, 32, and 40 from rsi
+mov rsp, 8[rsi] ;; Switch to new stack, which may not be mapped in the current address space!
+... ;; load rbp, r12, r13, r14, from offsets 16, 24, 32, and 40 from rsi
 mov r15, 48[rsi]
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast [\rtv'+56](\mathcal{I}\texttt{ASpace}(\theta,m) \ast \texttt{Pother})}_{\rtv}$
-$\specline{ \texttt{rdi}\mapsto_{\textsf{r}} \_ \ast \texttt{rbx}\mapsto_{\textsf{r}} \rtv'+0 \ast  \texttt{rsp}\mapsto_{\textsf{r}} \rtv'+8 \ast \texttt{rbp}\mapsto_{\textsf{r}} \rtv'+16 }_{\rtv}$
-$\specline{\texttt{r12}\mapsto_{\textsf{r}} \rtv'+24 \ast \texttt{r13}\mapsto_{\textsf{r}} \rtv'+32 \ast \texttt{r14}\mapsto_{\textsf{r}} \rtv'+40 \ast \texttt{r15}\mapsto_{\textsf{r}} \rtv'+48}_{\rtv}$
-$\specline{\texttt{rdi+0} \mapsto_{\textsf{v}} \texttt{rbxv} \ast \texttt{rdi+8} \mapsto_{\textsf{v}} \texttt{rspv} \ast \texttt{rdi+16} \mapsto_{\textsf{v}} \texttt{rbpv} \ast \texttt{rdi+24} \mapsto_{\textsf{v}} \texttt{r12v} \ast }_{\rtv}$
-$\specline{\texttt{rdi+32} \mapsto_{\textsf{v}} \texttt{r13v} \ast \texttt{rdi+40} \mapsto_{\textsf{v}} \texttt{r14v} \ast \texttt{rdi+48} \mapsto_{\textsf{v}} \texttt{r15v}\ast \texttt{rdi+56} \mapsto_{\textsf{v}} \rtv}_{\rtv}$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast [\rtv'](\mathcal{I}\texttt{ASpace}(\theta',m') \ast \texttt{Pother})  \ast  \texttt{rsi}\mapsto_{\textsf{r}} new \ast \texttt{rdi}\mapsto_{\textsf{r}} old \ast \texttt{rbx}\mapsto_{\textsf{r}} \texttt{rbxv}' }_{\rtv}$
+$\specline{\texttt{rsp}\mapsto_{\textsf{r}} \texttt{rspv}' \ast \texttt{rbp}\mapsto_{\textsf{r}} \texttt{rbpv}' \ast \texttt{r12}\mapsto_{\textsf{r}} \texttt{r12v}' \ast \texttt{r13}\mapsto_{\textsf{r}} \texttt{r13v}' \ast \texttt{r14}\mapsto_{\textsf{r}} \texttt{r14v}' \ast \texttt{r15}\mapsto_{\textsf{r}} \texttt{r15v}'}_{\rtv}$
+$\specline{\mathsf{ContextAt}(old,[\mathsf{rbxv,\ldots,\rtv}]) \ast \mathsf{ContextAt}(new,[\mathsf{rbxv}',\ldots,\rtv']) }_{\rtv}$
 ;; Switch to the new address space
 mov cr3, 56[rsi]
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast [\rtv'+56](\mathcal{I}\texttt{ASpace}(\theta,m) \ast \texttt{Pother}) }_{\rtv}$
-$\specline{ \texttt{rdi}\mapsto_{\textsf{r}} \_ \ast \texttt{rbx}\mapsto_{r} \rtv'+0 \ast  \texttt{rsp}\mapsto_{\textsf{r}} \rtv'+8 \ast \texttt{rbp}\mapsto_{\textsf{r}} \rtv'+16 }_{\rtv}$
-$\specline{\texttt{r12}\mapsto_{\textsf{r}} \rtv'+24 \ast \texttt{r13}\mapsto_{\textsf{r}} \rtv'+32 \ast \texttt{r14}\mapsto_{\textsf{r}} \rtv'+40 \ast \texttt{r15}\mapsto_{\textsf{r}} \rtv'+48}_{\rtv}$
-$\specline{\texttt{rdi+0} \mapsto_{\textsf{v}} \texttt{rbxv} \ast \texttt{rdi+8} \mapsto_{\textsf{v}} \texttt{rspv} \ast \texttt{rdi+16} \mapsto_{\textsf{v}} \texttt{rbpv} \ast  \texttt{rdi+24} \mapsto_{\textsf{v}} \texttt{r12v} \ast }_{\rtv}$
-$\specline{\texttt{rdi+32} \mapsto_{\textsf{v}} \texttt{r13v} \ast \texttt{rdi+40} \mapsto_{\textsf{v}} \texttt{r14v} \ast \texttt{rdi+48} \mapsto_{\textsf{v}} \texttt{r15v}\ast \texttt{rdi+56} \mapsto_{\textsf{v}} \rtv}_{\rtv}$
-$\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\; \sqsubseteq $
-$\specline{ [\rtv](\texttt{Pcurrent}) \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast \texttt{Pother}}_{\rtv'+56}$
+$\specline{ [\rtv](\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast \mathsf{ContextAt}(old,[\mathsf{rbxv,\ldots,\rtv}]) \ast \mathsf{ContextAt}(new,[\mathsf{rbxv}',\ldots,\rtv']))  }_{\rtv}$
+$\specline{ \mathcal{I}\texttt{ASpace}(\theta',m') \ast \texttt{Pother} \ast  \texttt{rsi}\mapsto_{\textsf{r}} new \ast \texttt{rdi}\mapsto_{\textsf{r}} old \ast \ldots }_{\rtv'}$
 \end{lstlisting}
+\vspace{-1em}
 \caption{Basic task switch code that switches address spaces.}
@@
 
-Figure \ref{fig:swtch} gives simplified code for a basic task switch, the heart of an OS scheduler implementation.
-This is code that saves the context (registers and stack) of the running thread (here in a structure pointed to by \lstinline|rdi|'s value shown in Lines 10-16 and Line 23 of Figure \ref{fig:swtch}) and restores the context of an existing thread (from \lstinline|rsi| shown in Lines 28-38 and Line 45), including the corresponding change of address space for a target thread in another process.
-This code assumes the System V AMD64 ABI calling convention, where the normal registers not mentioned are caller-save, and therefore saved on the stack of the thread that calls this code, as well as on the new stack of the thread that is restored, thus only the callee-save registers and \texttt{cr3} must be restored.\footnote{We are simplifying in a couple basic ways.
+Figure \ref{fig:swtch} gives simplified code for a basic task switch, the heart of an OS scheduler implementation. This is code that saves the context (registers and stack)
+of the running thread (here in a structure pointed to by \lstinline|rdi|'s value shown in Lines 5--7 and Line 11 of Figure \ref{fig:swtch}) and restores the context of 
+an existing thread (from \lstinline|rsi| shown in abbreviated Lines 14--17 and Line 22), including the corresponding change of address space for a target thread in another process.
+This code assumes the System V AMD64 ABI calling convention, where the normal registers not mentioned are caller-save, and therefore saved on the stack of the thread
+that calls this code, as well as on the new stack of the thread that is restored, thus only the callee-save registers and \texttt{cr3} must be 
+restored.\footnote{We are simplifying in a couple basic ways.
 First, we are ignoring non-integer registers (e.g., floating point, vector registers) entirely.
@@
 \end{lstlisting}
-A call to this code begins executing one thread (Lines 10-16) in one address space ($\rtv$ in Figure \ref{fig:swtch}) (whose information will be saved in \lstinline[language=C]|save| shown in Line 22) and finishes execution executing a different thread in a different address space (whose information is initially in \lstinline[language=C]|restore| Line 29).
+A call to this code begins executing one thread (up through Line 17) in one address space ($\rtv$), whose information will be saved in a structure at address $old$,
+and finishes execution executing a different thread in a different address space (whose information is initially in $new$).
 
-Because this code does not directly update the instruction pointer, it is worth explaining \emph{how} this switches threads: by switching stacks.
+Because this code does not directly update the instruction pointer, it is worth explaining \emph{how} this switches threads: by switching address spaces and stacks. 
 This is meant to be called with a return address for the current thread stored on the current stack when called --- which must be reflected in the calling convention. 
-In particular, the precondition of the return address on the initial stack requires the callee-save register values at the time of the call: those stored in the first half of the code.
-Likewise, part of the invariant of the stack of the second thread, the one being restored, is that the return address on \emph{that} stack requires the saved callee-save registers stored in that context to be in registers as its precondition.
+In particular, the precondition of the return address on the initial stack requires the callee-save register values at the time of the call: those stored in the first 
+half of the code.
+Likewise, part of the invariant of the stack of the second thread, the one being restored, is that the return address on \emph{that} stack requires the saved 
+callee-save registers stored in that context to be in registers as its precondition.
 
-The wrinkle, and the importance of the modal treatment of assertions, is that the target thread's precondition is \emph{relative to its address space}, not the address space of the calling thread shown as
-\[[\rtv'+56](\mathcal{I}\texttt{ASpace}(\theta,m) \ast \texttt{Pother})\]
+The wrinkle, and the importance of the modal treatment of assertions, is that the target thread's precondition is \emph{relative to its address space}, 
+not the address space of the calling thread shown as 
+\[[\rtv'](\mathcal{I}\texttt{ASpace}(\theta,m) \ast \texttt{Pother})\]
 in the specfication. 
@@
 
-Lines 10--16 save the current context into memory (in the current address space).
-Line 22 saves the initial page table root.
-Lines 33--38 begin restoring the target context, including the stack pointer (line 33),
-which may not be mapped in the address space at that time: it is the stack for the context being
-loaded into the CPU.
-The actual address switch occurs on line 45, which is verified with our modal rule for updating \lstinline|cr3|,
-and thus shifts resources in and out of other-space modalities as appropriate.
+% Lines 10--16 save the current context into memory (in the current address space).
+% Line 22 saves the initial page table root.
+% Lines 33--38 begin restoring the target context, including the stack pointer (line 33),
+% which may not be mapped in the address space at that time: it is the stack for the context being
+% loaded into the CPU.
+% The actual address switch occurs on line 45, which is verified with our modal rule for updating \lstinline|cr3|,
+% and thus shifts resources in and out of other-space modalities as appropriate.
 
@@
 and the saved context's invariant captures the precondition for restoring its execution \emph{in the previous address space} (as part of \textsf{P}).
-%At that point, it is safe to execute a return (or in a variant on this for user-mode and/or preemptive scheduling, a return-from-interrupt instruction).
 
-Note that immediately after the page table switch, the points-to information about the saved and restored contexts is \emph{also} guarded by a modality (Specification Line 52)
-\[ [\rtv](\texttt{Pcurrent}) \qquad \textsf{ where Pcurrent is all available logical assertions under addrress-space} \rtv\]
-for the old address space, since there is no automatic guarantee that that memory is mapped in the new address space.
+Note that immediately after the page table switch, the points-to information about the saved and restored contexts is guarded by a modality for the retiring
+address space \rtv (Line 23). This is enforced by \textsc{WriteToRegCtlFromRegModal} (Figure \ref{fig:wpdamd}), and is sensible because
+there is no general guarantee that the data structures of the previous addres space are mapped in the new address space.
 The ability to transfer that points-to information out of that modality is specific to a given kernel's design. 
-Kernels that map kernel memory into all address spaces would need to ensure and specify enough specific details about memory mappings to allow a proof of an elimination rule for specific modally-constrained points-to assertions.
-Following Spectre and Meltdown, this kernel design became less prevalent because speculative execution of accesses to kernel addresses could leak information even if the access did eventually cause a fault (the user/kernel mode permission check was done after fetching data from memory).
+Kernels that map kernel memory into all address spaces would need to ensure and specify enough specific details about memory mappings to allow a 
+proof of an elimination rule for specific modally-constrained points-to assertions.
+% Following Spectre and Meltdown, this kernel design became less prevalent because speculative execution of accesses to kernel addresses could leak information even if the access did eventually cause a fault (the user/kernel mode permission check was done after fetching data from memory).
 Thus many modern kernels have reverted to the older kernel design where the kernel inhabits its own unique address space, and user processes have only enough extra material mapped in their address spaces to switch into the kernel (CPUs do not speculate past updates to \texttt{cr3}).

### 2023-11-16 examples.tex
condensing new p2v example

@@
 
-Work to verify the full page table walk is ongoing, but we have verified the key step for a small x86-64 kernel, which is the physical-to-virtual
-conversion (often appearing as a macro \texttt{P2V} in C source code). In our small kernel, as in larger kernels, \texttt{P2V} is actually just addition,
-though the Hoare triple we verify for this add instruction is rather complex:
-
 \begin{figure}\footnotesize
 \begin{lstlisting}[mathescape]
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \ulcorner \textsf{KERNELBASE} + \paddr > 0 \urcorner }_{\rtv}$
-$\specline{ \texttt{rip}\mapsto_{\textsf{r}} \textsf{iv} \ast \texttt{rflags}\mapsto_{\textsf{r}} \textsf{flaglist} \ast \texttt{rbx}\mapsto_{\textsf{r}} \paddr  }_{\rtv}$
-$\specline{ \sumapaces\rtv\delta  \ast \sumpv\paddr\qfrac\true  }_{\rtv}$
-;; This offsets the physical address bt KERNELBASE!
-add rbx, KERNELBASE
-$\specline{\exists \qfraczero; \qfrac,\textsf{ flaglistadd, w, flaglistadd, val}, \textsf{P} \ast \texttt{rip}\mapsto_{\textsf{r}} \textsf{iv+7} \ast \texttt{rflags}\mapsto_{\textsf{r}} \textsf{flaglistadd} \ast \texttt{rbx}\mapsto_{\textsf{r}} \paddr + \textsf{KERNELBASE} }_{\rtv}$
-$\specline{\bigast{(\vaddr, \paddr)\in \ptablestore}{\exists\;(\textsf{l4e, l3e, l2e, l1e, } \paddr)\ldotp \textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\vaddr\textsf{, l4e, l3e, l2e, l1e,} \paddr)}}_{\rtv}$
-$\specline{\mathcal{R} -\ast \bigast{(\paddr) \in \Xi}{\exists\; (\textsf{qfrac, q, val, va}) \ldotp \ulcorner \textsf{va} = \paddr + \textsf{KERNELBASE} \urcorner \ast \sumwalkabs\vaddr\qfrac\paddr  \ast \paddr \mapsto_{\mathsf{p}}\{\textsf{qfrac}\} \; \vale\ast  \ulcorner \textsf{qfrac} = 1 \leftrightarrow \; \lnot\textsf{entry\_present }(\vale) \urcorner}}_{\rtv}$
-$\specline{ \ownGhost\gammaPred{\authfull{\ptableabswalk\ptablestore}} \ast \ownGhostpv\gammaPred{\authfull{\pvmapping\Xi}}}$
-$\specline{ \sumapaces\rtv\delta  \ast\sumwalkabs\vaddr\qfrac\paddr \ast \sumpv\paddr\qfraczero\true \ast  \paddr \mapsto_{\mathsf{p}}\{\textsf{qfrac}\} \; \vale \ast \textsf{P}}_{\rtv}$
-$\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\; \sqsubseteq $
-$\specline{\exists \qfraczero; \qfrac,\textsf{ flaglistadd, w, flaglistadd, val}, \textsf{P} \ast \texttt{rip}\mapsto_{\textsf{r}} \textsf{iv+7} \ast \texttt{rflags}\mapsto_{\textsf{r}} \textsf{flaglistadd} \ast \texttt{rbx}\mapsto_{\textsf{r}} \paddr + \textsf{KERNELBASE} }_{\rtv}$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \ulcorner \textsf{KERNELBASE} + \paddr > 0 \urcorner \ast \texttt{rbx}\mapsto_{\textsf{r}} \paddr \ast \sumapaces\rtv\delta  \ast \sumpv\paddr\qfrac\true  }_{\rtv}$
+add rbx, KERNELBASE ;; shift physical address by identity mapping constant
+$\specline{\exists \qfraczero; \qfrac,\textsf{ w, val}, \textsf{P} \ast \texttt{rbx}\mapsto_{\textsf{r}} \paddr + \textsf{KERNELBASE} }_{\rtv}$
 $\specline{\bigast{(\vaddr, \paddr)\in \ptablestore}{\exists\;(\textsf{l4e, l3e, l2e, l1e, } \paddr)\ldotp \textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\vaddr\textsf{, l4e, l3e, l2e, l1e,} \paddr)}}_{\rtv}$
@@
 \end{lstlisting}
-\caption{Identity mapping of a physical address to a virtual one.}
-\label{fig:mapping_code}
+\todo[inline]{Ismail: Should \delta be something else? $\delta$ was used for $\theta$, we need a different gname for $\Xi$.}
+\caption{Converting a physical address of a PTE to a virtual address (without instruction pointer or flag updates).}
+\label{fig:p2v}
 \end{figure}
 
-We can see that this specification neatly encapsulates reasoning about invariants on the page tables in terms of general modal assumptions.\todo{speculative
-but ought to turn out this way}
+Work to verify the full page table walk is ongoing, but we have verified the key step for a small x86-64 kernel, which is the physical-to-virtual
+conversion (often appearing as a macro \texttt{P2V} in C source code). In our small kernel, as in larger kernels, \texttt{P2V} is actually just addition,
+though the Hoare triple we verify for this add instruction is rather complex, with the specification shown in Figure \ref{fig:p2v}.
+Beyond a frame ($P$), the precondition requires the current address space invariant, a check that the addition will not overflow, the physical address to convert
+(in \lstinline|rbx|), knowledge that the current address space is valid, and knowledge that the physical address is in a range dedicated
+to page table entries (as this physical-to-virtual conversion is only guaranteed for page table memory). The latter uses an additional ghost set added to the per-address-space
+invariant per Figure \ref{fig:peraspaceinvariant_with_p2v_extension}. The extended invariant owns, for each memory location dedicated to page tables, some fraction of
+that physical memory, and also guarantees the offset is the correct calculation.\todo{Colin needs a second pass here}
 This is a simple piece of code whose functionality is critical and whose correctness is highly non-trivial. No prior work engages with this problem.

### 2023-11-16 examples.tex
fix build failures

@@
 \end{lstlisting}
-\todo[inline]{Ismail: Should \delta be something else? $\delta$ was used for $\theta$, we need a different gname for $\Xi$.}
+\todo[inline]{Ismail: Should $\delta$ be something else? $\delta$ was used for $\theta$, we need a different gname for $\Xi$.}
 \caption{Converting a physical address of a PTE to a virtual address (without instruction pointer or flag updates).}
@@
 ;; Assume the save-space is in rdi, load-space in rsi. First, save the yielding context
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast [\rtv'](\mathcal{I}\texttt{ASpace}(\theta',m') \ast \texttt{Pother}) \ast \texttt{rsi}\mapsto_{\textsf{r}} old \ast \texttt{rdi}\mapsto_{\tnewoldr}} old \ast \texttt{rbx}\mapsto_{\textsf{r}} \texttt{rbxv} }_{\rtv}$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast [\rtv'](\mathcal{I}\texttt{ASpace}(\theta',m') \ast \texttt{Pother}) \ast \texttt{rsi}\mapsto_{\textsf{r}} new \ast \texttt{rdi}\mapsto_{\textsf{r}} old \ast \texttt{rbx}\mapsto_{\textsf{r}} \texttt{rbxv} }_{\rtv}$
 $\specline{\texttt{rsp}\mapsto_{\textsf{r}} \texttt{rspv} \ast \texttt{rbp}\mapsto_{\textsf{r}} \texttt{rbpv} \ast \texttt{r12}\mapsto_{\textsf{r}} \texttt{r12v} \ast \texttt{r13}\mapsto_{\textsf{r}} \texttt{r13v} \ast \texttt{r14}\mapsto_{\textsf{r}} \texttt{r14v} \ast \texttt{r15}\mapsto_{\textsf{r}} \texttt{r15v}}_{\rtv}$

### 2023-11-16 programlogic.tex
round out and clean up rule discussion, be more selective about rules covered

@@
 \label{sec:selected_rules}
-\todo[inline]{Ismail: Please update this section with rip stuff}
-To address the issues described in Section \ref{sec:issues}, we give our logic using
-the Hoare-Double definition in Figure \ref{fig:wpddefinition}.\footnote{The figure does
+Per the discussion in Section \ref{sec:issues}, we define our logic using Hoare doubles:\footnote{This does
 omit some low-level Iris details related to stuckness and observations, which play no meaningful
 role in our development.}
-\begin{figure} 
   \[
@@
   \]
-\caption{Unfolded Hoare-Double Definition for \textsf{vProp} Logic }
-\label{fig:wpddefinition}
-\end{figure}
 Our Hoare doubles $\{\Phi\}_\textsf{rtv}\;\textsf{e}$ state that the expression (i.e., sequence of instructions)
 \textsf{e} are safe to execute (will not fault)
-when executed with precondition $\Phi\ast\textsf{cr3}\mapsto_{\textsf{r}} \textsf{rtv}$.
+when executed with \textsf{vProp} precondition $\Phi\ast\textsf{cr3}\mapsto_{\textsf{r}} \textsf{rtv}$.
 \textsf{WP} is Iris's own weakest precondition modality, unmodified~\cite{jung2018iris}.
@@
 Each rule in Figure \ref{fig:wpdamd}, 
-is annotated with an address value (\textsf{rtv}), 
-i.e. a root address of an address-space, under which the resources mentioned in the specification are valid,
-or equivalently, the assumed value of \lstinline|cr3| prior to executing the instructions.
-In general, we use $\textsf{r}_s$ and $\textsf{r}_d$ to specify source and destination registers
+is annotated with a root (i.e., \lstinline|cr3|) address value (\textsf{rtv}), 
+under which the resources mentioned in the specification are valid.
+In general, we use metavariables $\textsf{r}_s$ and $\textsf{r}_d$ to specify source and destination registers
 for each instruction, and prefix various register value variables with \textsf{rv}.
@@
 Standard for Hoare doubles, there is a frame resource $P$ in each rule for passing resources
-not used by the first instruction in sequence through and on to subsequent instructions.
+not used by the first instruction in sequence through to subsequent instructions.
+Our rules include tracking of each instruction's memory address to track \lstinline|rip| updates, which is critical
+for control transfer instructions. Our development also includes handling of the \lstinline|rflags| register updates from arithmetic instructions.
+Most rules are otherwise standard (e.g., \lstinline|mov| between register, etc.), with Figure \ref{fig:wpdamd} showing the rules
+most unique to our development.
+\looseness=-1
 
@@
 % \\
- \inferrule[AddRegImm]{
-   \{P \ast\textsf{rip} \mapsto_{\textsf{r}} \textsf{iv+7} \ast  \ast (r_d \mapsto_{r}  \textsf{rvd+imm} \lor  r_d \mapsto_{r} 0 ) \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{t}} \textsf{v} \}_{\textsf{rtv}}\;\overline{is}
- }{
-   \left\{
-   \begin{array}{l}
-   P \ast \textsf{rip} \mapsto_{\textsf{r}} \textsf{iv} \ast \textsf{rflags} \mapsto_{\textsf{r}} \textsf{flgs} \ast \ulcorner (\textsf{rvd + imm} > 0 \lor \textsf{rvd + imm} = 0 ) \urcorner \ast \arcr
-   r_d \mapsto_{r}  \textsf{rvd} \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{t}} \textsf{v}
- \end{array}\right\}_{\textsf{rtv}}  \;
-   \textsf{add}~\textsf{r}_d,~[\textsf{imm}];\;\overline{is}
- }
- \\
+ % \inferrule[AddRegImm]{
+ %   \{P \ast\textsf{rip} \mapsto_{\textsf{r}} \textsf{iv+7} \ast  \ast (r_d \mapsto_{r}  \textsf{rvd+imm} \lor  r_d \mapsto_{r} 0 ) \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{t}} \textsf{v} \}_{\textsf{rtv}}\;\overline{is}
+ % }{
+ %   \left\{
+ %   \begin{array}{l}
+ %   P \ast \textsf{rip} \mapsto_{\textsf{r}} \textsf{iv} \ast \textsf{rflags} \mapsto_{\textsf{r}} \textsf{flgs} \ast \ulcorner (\textsf{rvd + imm} > 0 \lor \textsf{rvd + imm} = 0 ) \urcorner \ast \arcr
+ %   r_d \mapsto_{r}  \textsf{rvd} \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{t}} \textsf{v}
+ % \end{array}\right\}_{\textsf{rtv}}  \;
+ %   \textsf{add}~\textsf{r}_d,~[\textsf{imm}];\;\overline{is}
+ % }
+ % \\
 \inferrule[WriteToRegFromVirtMem]{
   \left\{ \begin{array}{l}
-    P \ast \mathcal{I}\textsf{ASpace}\ast \mathcal{I}\textsf{ASpace} \ast \textsf{rip} \mapsto_{\textsf{r}} \textsf{iv + x} \ast \ulcorner x = 3 \lor x =4 \urcorner \ast \arcr
+    P \ast \mathcal{I}\textsf{ASpace}\ast \mathcal{I}\textsf{ASpace} \ast \textsf{rip} \mapsto_{\textsf{r}} {\textsf{iv} + \textsf{MovLen}(\textsf{r}_d,[\textsf{r}_a])} \ast \arcr
     r_d \mapsto_{r}  \textsf{v} \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{v,rtv}} \textsf{v}
@@
 \inferrule[WriteToVirtMemFromReg]{
-  \{P \ast \mathcal{I}\textsf{ASpace}\ast \textsf{rip} \mapsto_{\textsf{r}} \textsf{iv + x} \ast \ulcorner x = 3 \lor x =4 \urcorner \ast r_s \mapsto_{r}\{q\}  \textsf{rvs}  \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{v}} \textsf{v} \}_{\textsf{rtv}}\;\overline{ is}
+  \{P \ast \mathcal{I}\textsf{ASpace}\ast \textsf{rip} \mapsto_{\textsf{r}} \textsf{iv} + \textsf{MovLen}(\textsf{r}_d,[\textsf{r}_a])  \ast r_s \mapsto_{r}\{q\}  \textsf{rvs}  \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{v}} \textsf{v} \}_{\textsf{rtv}}\;\overline{ is}
 }{
@@
   %\lstinline|mov %cr3, r| 
-}\\
-\inferrule[WriteToRegFromRegCtl]{
-  \{P \ast  \textsf{rip} \mapsto_{\textsf{r}} \textsf{iv + 4} \ast r_d \mapsto_{r} \textsf{rvs} \ast r_s \mapsto_{r}\{q\} \textsf{ rvs} \}_{\textsf{rtv}}\;\overline{ is}
-}{
-  \{P \ast  \textsf{rip} \mapsto_{\textsf{r}} \textsf{iv} \ast r_d \mapsto_{r} \textsf{rvd} \ast r_s \mapsto_{r}\{q\} \textsf{ rvs} \}_{\textsf{rtv}}
-  \textsf{ mov}~\textsf{r}_d,~\textsf{r}_s;\;\overline{is}
-  %\lstinline|mov %cr3, r| 
 }
+% \\
+% \inferrule[WriteToRegFromRegCtl]{
+%   \{P \ast  \textsf{rip} \mapsto_{\textsf{r}} \textsf{iv + 4} \ast r_d \mapsto_{r} \textsf{rvs} \ast r_s \mapsto_{r}\{q\} \textsf{ rvs} \}_{\textsf{rtv}}\;\overline{ is}
+% }{
+%   \{P \ast  \textsf{rip} \mapsto_{\textsf{r}} \textsf{iv} \ast r_d \mapsto_{r} \textsf{rvd} \ast r_s \mapsto_{r}\{q\} \textsf{ rvs} \}_{\textsf{rtv}}
+%   \textsf{ mov}~\textsf{r}_d,~\textsf{r}_s;\;\overline{is}
+%   %\lstinline|mov %cr3, r| 
+% }
 \end{mathpar}
-\todo[inline]{update for rip tracking}
 \caption{Reasoning Rules for Selected \textsf{AMD64} Instructions under}
@@
 
-We briefly describe several key rules representative of the system.
-
-Figure \ref{fig:wpdamd}, includes two  rules for accessing memory at an address in a register $r_a$. 
+Figure \ref{fig:wpdamd} includes two  rules for accessing memory at an address stored in a register $r_a$. 
 Ultimately, any memory access needs to ensure the relevant address translation would succeed,
@@
 ($\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$ in Figure \ref{fig:strongvirtualpointsto}).
-
-% We mainly have two options to achieve this. 
-% First one is to use  page-table-walk pointsto 
-% ($\textsf{vaddr} \mapsto_{\textsf{t}} \textsf{v}$), 
-% which is consists of all-physical page-table memory assertions, 
-% to prove the page-table traversal \textsf{AMD64} implements, update the register $r_d$ the value obtained 
-% ($\textsf{v}$) as shown in Figure (\TirNameStyle{WriteToRegFromPhysMem}). 
-% By doing so, we expose the resources ($\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$) needed only 
-% for the proof of \textsf{AMD64} page-table traversal to the client of the rule. 
-% This breaks the modularity of reasoning we want to maintain in location virtualization.
-
 \TirNameStyle{WriteToRegFromVirtMem} and \TirNameStyle{WriteToVirtMemFromReg}
-each use a virtual-pointsto assertion ($\textsf{vaddr} \mapsto_{\textsf{v},\textsf{rtv}}$),
+each use a virtual-pointsto assertion ($\textsf{vaddr} \mapsto_{\textsf{v}} \textsf{v}$),
 and are nearly-standard (assembly) separation logic rules for memory accesses.
@@
 for the current address space to be carried through.
-While doing the proof of soundness for this rule,
+While doing the proof of soundness for these rules,
 the token ($\sumwalkabs\vaddr\qfrac\paddr$) from the virtual points-to
@@
 \paragraph{Updating \lstinline|cr3|} 
-Unlike other rules, \TirNameStyle{WriteToRegCtlFromReg} updates the root address of the 
+Unlike other rules, \TirNameStyle{WriteToRegCtlFromRegModal} updates the root address of the 
 address-space determining the validity of resources, from $\rtv$ before the
-\lstinline|mov| to $\textsf{rvs}$ afterwards.
-From that rule, with one step outside the \textsf{vProp} logic (so it is not quite a derived rule within
-our embedded logic), we can derive 
-\TirNameStyle{WriteToRegCtlFromRegModal},
-which is the (full) Hoare double version of the address space switch rule proposed in Section \ref{sec:issues}.
+\lstinline|mov| to $\textsf{rvs}$ afterwards. The global effects of this rule are reflected in moving
+bare assertions under an other-space modality for $\rtv$, and moving the new address space's assertions out of
+the corresponding modality.
+% From that rule, with one step outside the \textsf{vProp} logic (so it is not quite a derived rule within
+% our embedded logic), we can derive 
+% \TirNameStyle{WriteToRegCtlFromRegModal},
+% which is the (full) Hoare double version of the address space switch rule proposed in Section \ref{sec:issues}.
 

### 2023-11-16 examples.tex
space hacks, rounding out p2v explanation

@@
 \begin{figure*}
+\footnotesize
 \[
 \begin{array}{l}
-  \mathcal{I}\textsf{ASpace}_{\textsf{id}}(\ptablestore,\Xi,m)\stackrel{\triangle}{=} \textsf{ASpace\_Lookup}(\ptablestore,\Xi,m) \ast \\
-  \qquad\qquad\qquad \bigast{(\vaddr, \textsf{paddr})\in \ptablestore}{\exists\;(\textsf{l4e, l3e, l2e, l1e, paddr})\ldotp \textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\vaddr\textsf{, l4e, l3e, l2e, l1e, paddr})} \\
-  \qquad\qquad\qquad \bigast{(\paddr) \in \Xi}{\exists\; (\textsf{qfrac, q, val,}\vaddr) \ldotp \ulcorner \vaddr = \paddr + \textsf{KERNELBASE} \urcorner \ast  \underbrace{\sumwalkabs\vaddr\qfrac\paddr }_\text{Ghost translation} \ast \\
-\qquad\qquad\qquad\qquad\qquad\qquad\qquad\qquad    \underbrace{\paddr \mapsto_{\mathsf{p}}\{\textsf{qfrac}\}\; \vale}_\text{Physical location} \ast \ulcorner \textsf{qfrac} = 1 \leftrightarrow \; \lnot\textsf{entry\_present }(\vale) \urcorner }\\
+  \mathcal{I}\textsf{ASpace}_{\textsf{id}}(\ptablestore,\Xi,m)\stackrel{\triangle}{=}\\ \textsf{ASpace\_Lookup}(\ptablestore,\Xi,m) \ast 
+  \bigast{(\vaddr, \textsf{paddr})\in \ptablestore}{\exists\;(\textsf{l4e, l3e, l2e, l1e, paddr})\ldotp \textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\vaddr\textsf{, l4e, l3e, l2e, l1e, paddr})} \\
+  \bigast{(\paddr) \in \Xi}{\exists\; (\textsf{qfrac, q, val,}\vaddr) \ldotp \ulcorner \vaddr = \paddr + \textsf{OFF} \urcorner \ast  \underbrace{\sumwalkabs\vaddr\qfrac\paddr }_\text{Ghost translation} \ast
+% \qquad\qquad\qquad\qquad\qquad\qquad\qquad\qquad    
+\underbrace{\paddr \mapsto_{\mathsf{p}}\{\textsf{qfrac}\}\; \vale}_\text{Physical location} \ast \ulcorner \textsf{qfrac} = 1 \leftrightarrow \; \lnot\textsf{entry\_present }(\vale) \urcorner }\\
   \textsf{ where } \\
    \textsf{ASpace\_Lookup}_{\textsf{id}}(\ptablestore,m) \stackrel{\triangle}{=} \lambda\textsf{ cr3val} \ldotp \; \exists \gammaPred \; \ldotp \ulcorner m \; !!\; \textsf{cr3val} = \textsf{Some } \gammaPred \urcorner \ast
-   \ownGhost\gammaPred{\authfull{\ptableabswalk\ptablestore}} \ast \\ \qquad\qquad\qquad\qquad\qquad\qquad\qquad\qquad \ownGhostpv\gammaPred{\authfull{\pvmapping\Xi}}
+   \ownGhost\gammaPred{\authfull{\ptableabswalk\ptablestore}} \ast  \ownGhostpv\gammaPred{\authfull{\pvmapping\Xi}}
 
@@
 \]
+\vspace{-1em}
 \caption{Global Address-Space Invariant in Figure \ref{fig:peraspaceinvariant} extended with a ghost map bookkeeping identity mappings }
@@
 \begin{lstlisting}[mathescape]
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \ulcorner \textsf{KERNELBASE} + \paddr > 0 \urcorner \ast \texttt{rbx}\mapsto_{\textsf{r}} \paddr \ast \sumapaces\rtv\delta  \ast \sumpv\paddr\qfrac\true  }_{\rtv}$
-add rbx, KERNELBASE ;; shift physical address by identity mapping constant
-$\specline{\exists \qfraczero; \qfrac,\textsf{ w, val}, \textsf{P} \ast \texttt{rbx}\mapsto_{\textsf{r}} \paddr + \textsf{KERNELBASE} }_{\rtv}$
-$\specline{\bigast{(\vaddr, \paddr)\in \ptablestore}{\exists\;(\textsf{l4e, l3e, l2e, l1e, } \paddr)\ldotp \textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\vaddr\textsf{, l4e, l3e, l2e, l1e,} \paddr)}}_{\rtv}$
-$\specline{\mathcal{R} -\ast \bigast{(\paddr) \in \Xi}{\exists\; (\textsf{qfrac, q, val, va}) \ldotp \ulcorner \textsf{va} = \paddr + \textsf{KERNELBASE} \urcorner \ast \sumwalkabs\vaddr\qfrac\paddr  \ast \paddr \mapsto_{\mathsf{p}}\{\textsf{qfrac}\} \; \vale\ast  \ulcorner \textsf{qfrac} = 1 \leftrightarrow \; \lnot\textsf{entry\_present }(\vale) \urcorner}}_{\rtv}$
-$\specline{ \ownGhost\gammaPred{\authfull{\ptableabswalk\ptablestore}} \ast \ownGhostpv\gammaPred{\authfull{\pvmapping\Xi}}}$
-$\specline{ \sumapaces\rtv\delta \ast \sumpv\paddr\qfraczero\true \ast \textsf{P}}_{\rtv}$
-$\specline{\vaddr \mapsto_{\texttt{vpte}} \{\textsf{q}, \textsf{qfrac}\} \paddr \;(\texttt{val})}$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \ulcorner \textsf{OFF} + \paddr > 0 \urcorner \ast \texttt{rbx}\mapsto_{\textsf{r}} \paddr \ast \sumapaces\rtv\delta  \ast \sumpv\paddr\qfrac\true  }_{\rtv}$
+add rbx, OFF ;; shift physical address by identity mapping constant OFF
+$\specline{\textsf{P} \ast \exists \qfraczero; \qfrac,\textsf{ w, val},  \texttt{rbx}\mapsto_{\textsf{r}} \paddr + \textsf{OFF} \ast \sumapaces\rtv\delta \ast \sumpv\paddr\qfrac\true \ast \vaddr \mapsto_{\texttt{vpte}} \{\textsf{q}, \textsf{qfrac}\} \paddr \;(\texttt{val})}$
+$\specline{\ulcorner \vaddr = \paddr + \textsf{OFF} \urcorner \ast \ulcorner \textsf{qfrac} = 1 \leftrightarrow \; \lnot\textsf{entry\_present }(\vale) \urcorner \ast \sumwalkabs\vaddr\qfrac\paddr \ast (R \wand \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)) }_{\rtv}$
 \end{lstlisting}
-\todo[inline]{Ismail: Should $\delta$ be something else? $\delta$ was used for $\theta$, we need a different gname for $\Xi$.}
-\caption{Converting a physical address of a PTE to a virtual address (without instruction pointer or flag updates).}
+\vspace{-1em}
+\caption{Converting a physical address of a PTE to a virtual address (w/o instruction pointer or flag updates). $R$ abbreviates the resources
+associated with \paddr\ by $\Xi$ in Figure \ref{fig:peraspaceinvariant_with_p2v_extension}.}
 \label{fig:p2v}
@@
 
-Work to verify the full page table walk is ongoing, but we have verified the key step for a small x86-64 kernel, which is the physical-to-virtual
+Work to verify the full page table walk is ongoing, but we have verified the critical step for a small x86-64 kernel, which is the physical-to-virtual
 conversion (often appearing as a macro \texttt{P2V} in C source code). In our small kernel, as in larger kernels, \texttt{P2V} is actually just addition,
-though the Hoare triple we verify for this add instruction is rather complex, with the specification shown in Figure \ref{fig:p2v}.
+but the correctness of this simple instruction is quite subtle, as shown in Figure \ref{fig:p2v}.
 Beyond a frame ($P$), the precondition requires the current address space invariant, a check that the addition will not overflow, the physical address to convert
 (in \lstinline|rbx|), knowledge that the current address space is valid, and knowledge that the physical address is in a range dedicated
-to page table entries (as this physical-to-virtual conversion is only guaranteed for page table memory). The latter uses an additional ghost set added to the per-address-space
+to page table entries (as this physical-to-virtual conversion is only guaranteed for page table memory). 
+The latter uses an additional ghost set added to the per-address-space
 invariant per Figure \ref{fig:peraspaceinvariant_with_p2v_extension}. The extended invariant owns, for each memory location dedicated to page tables, some fraction of
-that physical memory, and also guarantees the offset is the correct calculation.\todo{Colin needs a second pass here}
+that physical memory, and also guarantees the offset is the correct calculation.
+Subtly, the operation ensures the relevant table entry is readable, but the exact portion of ownership returned must be determined by inspecting the valid bit
+of the value in memory --- so full ownership is returned only for unused entries.
 This is a simple piece of code whose functionality is critical and whose correctness is highly non-trivial. No prior work engages with this problem.

### 2023-11-16 programlogic.tex
space hacks
update axiom count

@@
 \begin{figure*}
+\footnotesize
 \[
 \begin{array}{l}
-    \vaddr\mapsto_{\textsf{v}}\{\textsf{q}\}\;\vale : \mathsf{vProp}~\Sigma \stackrel{\triangle}{=} \\
-    \quad
+    \vaddr\mapsto_{\textsf{v}}\{\textsf{q}\}\;\vale : \mathsf{vProp}~\Sigma \stackrel{\triangle}{=} 
     \exists \paddr\ldotp
@@
   \begin{figure*}
+  \footnotesize
 \[
@@
 \begin{figure}
+\footnotesize
 \begin{mathpar}
@@
 \begin{figure}
+  \footnotesize
   \makeatletter % allow us to mention @-commands
@@
 
-At the moment our proofs do rely on 11\todo{update!} small axioms of properties which should be provable, but
+At the moment our proofs do rely on 13 small axioms of properties which should be provable, but
 are challenging to discharge due to some representation choices in our model.\footnote{See \lstinline|srx/x64/machine/current_axioms.v|}

### 2023-11-16 examples.tex
compress the mapping example, now at 20 pages + 5 lines

@@
 $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m)  \ast  \texttt{r14}\mapsto_{\textsf{r}} \_ \ast \texttt{rdi}\mapsto_{r} \vaddr \ast \texttt{rax}\mapsto_{\textsf{r}} \_ \ast \ulcorner \texttt{aligned va} \land  \theta \; !!\; \vaddr = \texttt{None}\urcorner}_{\rtv}$
-call ensure_L1_page
+call ensure_L1_page ;; Returns the virtual address of the L1 entry in rax
 $\specline{\exists (\entryf ,\;\entrytr,\; \entrytw,\; \entryo,\;\textsf{pte\_addr },\paddr) \; \ldotp\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast  \texttt{r14}\mapsto_{\textsf{r}} \_ \ast \texttt{rdi}\mapsto_{r} \vaddr \ast \texttt{rax}\mapsto_{\textsf{r}} \textsf{ pte\_addr} \; \ast }_{\rtv}$
 $\specline{ \ulcorner  \texttt{addr\_L1 }(\vaddr, \entryo) = \paddr \urcorner \ast \ulcorner \texttt{entry\_present } \entryf \land \texttt{entry\_present } \entrytr \land  \texttt{entry\_present } \entrytw \urcorner \; \ast}_{\rtv}$
-$\specline{\nfpointsto{\mask\vaddr\maskfour\rtv}{\mask\vaddr\maskfouroff\rtv}\entryf\qone\naddr \; \ast}_{\rtv}$ 
-$\specline{  \nfpointsto{\mask\vaddr\maskthree\entryf}{\mask\vaddr\maskthreeoff\entryf}\entrytr\qtwo\naddr \ast \nfpointsto{\mask\vaddr\masktwo\entrytr}{\mask\vaddr\masktwooff\entrytr}\paddr\qthree\entryo \;\ast}_{\rtv}$
-$\specline{\texttt{pte\_addr} \mapsto_{\texttt{vpte}} \paddr \;(\texttt{wzero 64}) \ast \texttt{rax}\mapsto_{\textsf{r}} \texttt{pte\_addr}  }_{\rtv}$
-;; Returns the virtual address of the L1 entry in rax
+$\specline{\nfpointsto{\mask\vaddr\maskfour\rtv}{\mask\vaddr\maskfouroff\rtv}\entryf\qone\naddr \; \ast \nfpointsto{\mask\vaddr\maskthree\entryf}{\mask\vaddr\maskthreeoff\entryf}\entrytr\qtwo\naddr \ast}_{\rtv}$
+$\specline{  \nfpointsto{\mask\vaddr\masktwo\entrytr}{\mask\vaddr\masktwooff\entrytr}\paddr\qthree\entryo \;\ast \texttt{pte\_addr} \mapsto_{\texttt{vpte}} \paddr \;(\texttt{wzero 64}) \ast \texttt{rax}\mapsto_{\textsf{r}} \texttt{pte\_addr}  }_{\rtv}$
 mov r14, rax ;; Save that before another call
@@
 call alloc_phys_page_or_panic
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast  \texttt{r14}\mapsto_{\textsf{r}} \texttt{pte\_addr} \ast \texttt{rdi}\mapsto_{\textsf{r}} \vaddr \;\ast}_{\rtv}$
-$\specline{ \nfpointsto{\mask\vaddr\maskfour\rtv}{\mask\vaddr\maskfouroff\rtv}\entryf\qone\naddr \ast}_{\rtv}$ 
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast  \texttt{r14}\mapsto_{\textsf{r}} \texttt{pte\_addr} \ast \texttt{rdi}\mapsto_{\textsf{r}} \vaddr \;\ast \nfpointsto{\mask\vaddr\maskfour\rtv}{\mask\vaddr\maskfouroff\rtv}\entryf\qone\naddr \ast}_{\rtv}$
 $\specline{  \nfpointsto{\mask\vaddr\maskthree\entryf}{\mask\vaddr\maskthreeoff\entryf}\entrytr\qtwo\naddr \ast \nfpointsto{\mask\vaddr\masktwo\entrytr}{\mask\vaddr\masktwooff\entrytr}\paddr\qthree\naddr \;\ast}_{\rtv}$
@@
 ;; Calculate new L1 entry
-;; update the page table entry, mapping the page
-mov [r14], rax
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast  \texttt{r14}\mapsto_{\textsf{r}} \texttt{pte\_addr} \ast \texttt{rdi}\mapsto_{\textsf{r}} \vaddr \;\ast}_{\rtv}$
-$\specline{ \nfpointsto{\mask\vaddr\maskfour\rtv}{\mask\vaddr\maskfouroff\rtv}\entryf\qone\naddr \ast}_{\rtv}$ 
+mov [r14], rax ;; store the page table entry, mapping the page
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast  \texttt{r14}\mapsto_{\textsf{r}} \texttt{pte\_addr} \ast \texttt{rdi}\mapsto_{\textsf{r}} \vaddr \;\ast \nfpointsto{\mask\vaddr\maskfour\rtv}{\mask\vaddr\maskfouroff\rtv}\entryf\qone\naddr \ast}_{\rtv}$
 $\specline{  \nfpointsto{\mask\vaddr\maskthree\entryf}{\mask\vaddr\maskthreeoff\entryf}\entrytr\qtwo\naddr \ast \nfpointsto{\mask\vaddr\masktwo\entrytr}{\mask\vaddr\masktwooff\entrytr}\paddr\qthree\entryo \;\ast}_{\rtv}$
@@
 \end{figure}
+
 In Figure \ref{fig:mapping_code}, we see an address ($\vaddr$) currently not
@@
 phyiscal page to back the desired virtual page first requires ensuring
-the existence of a memory location of an appropriate L1 table entry,
-which is realized in two pieces by the post-condition of \lstinline|ensure_L1|:
-\begin{itemize}
-\item physical pointsto assertions on the tables L4, L3, and L2 reaching to the
-	L1 level entry (l1e) ensure higher levels of the page table exist, 
-	are marked present (\textsf{entry\_present} for the relevant
-		entries \textsf{l4e l3e l2e}) (Specification Lines 4-6) 
-	\item a virtual \emph{PTE}
+the existence of a memory location for an appropriate L1 table entry.
+The code uses a (currently axiomatized) helper function \lstinline|ensure_L1| (discussed again in Section \ref{sec:traversing}).
+\lstinline|ensure_L1|'s postcondition contains physical pointsto assertions for a partial page table walk reaching the
+L1 entry (l1e), ensuring higher levels of the page table exist. 
+The postcondition also includes a virtual \emph{PTE}
 pointsto ($\mapsto_{\textsf{vpte}}$) allows access to the memory of the L1 entry
 via virtual address \lstinline|pte_pointsto|.
-		A PTE points-to is defined just like the normal virtual points-to of Figure \ref{fig:virtualpointstosharing}, except the physical address is explicit in the assertion (here, \textsf{pa})
+A PTE points-to is defined just like the normal virtual points-to of Figure \ref{fig:virtualpointstosharing}, except the physical address (here, \textsf{pa}) is explicit in the assertion 
 rather than existentially quantified:
@@
 makes this convenient to use for page table modifications, since we must ensure
-		the modified address is the correct physical address that will be used as the L1 entry.
+the modified data is at the specific physical location that will be used as the L1 entry.
 This is also guaranteed ($ \ulcorner
 \texttt{addr\_L1}(\vaddr,\entryo) = \paddr\urcorner$).
-\end{itemize}
-After obtaining a virtual address \textsf{pte\_addr} in \textsf{rax} backed 
-by the physical memory for the L1 entry that will be used to translate the virtual addresses
-we are mapping, we save it to \textsf{r14} to be updated later in Line 9.
+% After obtaining a virtual address \textsf{pte\_addr} in \textsf{rax} backed 
+% by the physical memory for the L1 entry that will be used to translate the virtual addresses
+% we are mapping, we save it to \textsf{r14} to be updated later in Line 9.
 
-Then we allocate a fresh page initialized (i.e. zeroed) in Line 14 which return
-an \textsf{aligned} address of fresh page (\textsf{fpaddr}) in \textsf{rax}
-(Specification Line 19), plus 1 (so the return value is already a valid page table entry).
+Line 12 allocates a fresh page-aligned, zero-initialized page  (at \textsf{fpaddr}),
+returning a pre-filled PTE entry in \textsf{rax} ($+3$ sets the lower 2 bits).
 
-The crucial point in the proof is then actually updating the L1 entry,
+The crucial step is actually updating the L1 entry,
 via the virtual address
@@
 table entry address ($\textsf{addr\_L1}(\textsf{va, l1e})$), to hold the freshly
-allocated physical page address (\textsf{fpaddr}) in Line 22.
-After this write, we have (Specification Line 26)
+allocated physical page address (\textsf{fpaddr}) in Line 18.
+After this write, we have (Specification Line 21)
 \[\texttt{pte\_addr} \mapsto_{\texttt{vpte}}  \; \paddr\; \textsf{fpaddr}  \qquad\qquad (\textsf{by the PTE variant of Rule}~\TirNameStyle{WriteToVirtMemFromReg})\]
@@
 $\mapsto_{\textsf{vpte}}$, after the split of the PTE points-to, we can
-connect L4-L2 points-to assertions (in Specification Lines 24 and 25) with 
+connect L4-L2 points-to assertions with 
  $\texttt{fpaddr} \mapsto_{a} (\texttt{wzero 64})$ to obtain the
 complete physical table-walk assertion (L4\_L1\_PointsTo($\vaddr$ l4e l3e l2e
-fpaddr)) in Line 30.
+fpaddr)) in Line 25.
 
@@
 map ($\theta$) to change its state from unmapped ($\ulcorner \theta
-\;!!\;\vaddr = \texttt{None}\urcorner$ in Specification Line 30 and the
-precondition) to mapped (Specification Line 33) using Iris' ghost-map update
+\;!!\;\vaddr = \texttt{None}\urcorner$ (Line 28) and the
+precondition) to mapped (Line 29) using Iris' ghost-map update
 ($\sqsubseteq$), and construct a PTE -points-to $\vaddr
-\mapsto_{\textsf{vpte}}\; \qfrac \;\fpaddr\;(\textsf{wzero 64})$ (Specification
-Line 36) using $\sumwalkabs\vaddr\qfrac\fpaddr$ obtained from ghost
+\mapsto_{\textsf{vpte}}\; \qfrac \;\fpaddr\;(\textsf{wzero 64})$ (Line 31) using $\sumwalkabs\vaddr\qfrac\fpaddr$ obtained from ghost
 page-table-walk insertions, $\sumapaces\rtv\delta$ from unfolding the
-definition of $\mapsto_{\textsf{vpte}}$ in Specification Line 18, and $\fpaddr
-\mapsto_{\textsf{p}} \textsf{wzero 64}$ in Specification Line 34.
+definition of $\mapsto_{\textsf{vpte}}$, and $\fpaddr
+\mapsto_{\textsf{p}} \textsf{wzero 64}$.
 Finally, by existentially quantifying away \textsf{fpaddr},
@@
 
-This proof only maps the first word of the allocated page; the generalization to the full page
+For brevity this proof only maps the first word of the allocated page; the generalization to the full page
 is straightforward use of iterated separating conjunction to manage collections of clusters of 512 things
 (slices of the L1 entry physical ownership, virtual addresses at 8 byte increments, etc.).
+\looseness=-1
 
 Unlike the only prior work verifying analogous code for mapping a new page~\cite{kolanski08vstte,kolanski09tphols}, our proof above
-does \emph{not} need to unfold the definitions of derived logic (therefore reasoning directly over the operational semantics),
-but instead uses derived logic rules throughout to reason about the effect of the mapping. By incorporating verification of the
-\lstinline|ensure_L1| function (see Section \ref{sec:traversing}), our verification also directly handles several subtle aspects which
-were axiomatized in prior work.
+does \emph{not} need to unfold the definitions of the logic,
+but instead uses logic rules throughout to reason about the effect of the mapping. 
+This is the first proof we know of for mapping a virtual memory page that does not resort to direct reasoning over the operational semantics.
+% By incorporating verification of the
+% \lstinline|ensure_L1| function (see Section \ref{sec:traversing}), our verification also directly handles several subtle aspects which
+% were axiomatized in prior work.
 

### 2023-11-16 programlogic.tex
space hacks!
fix math mode mixup
a line in between two Definitions in Figure 2
removed the comment on vProp iProp physical pointsto

@@
 \[\textsf{w} \mapsto_{\textsf{p}} \{q\} \textsf{ v} \stackrel{\triangle}{=} (\textsf{drop 12}~w) \ \sim \ (\textsf{bottom 12}~w)\mapsto_{\textsf{a}} \; \{\textsf{q}\} \textsf{ v} \]
-
+ %I put a newline in between the following two Definitions as the second Definition seems not proper without the newline
 \begin{figure*}
@@
  $\exists_{\textsf{l4e l3e l2e l1e}} \ldotp$  $\ulcorner$ aligned $\vaddr \urcorner \ast$ L4_L1_PointsTo($\vaddr$ l4e l3e l2e l1e paddr) $\ast$ paddr $\mapsto_{p}\{\mathsf{q}\} \vpage$.
+
  Definition L4_L1_PointsTo (maddr l4e l3e l2e l1e paddr :word 64) : vProp $\Sigma$ := $\lambda$ cr3val.
@@
 This helper definition is also a more explicit \textsf{vProp}, explicitly binding a value for \lstinline|cr3| and using it
-to start the translation process.\todo{Ismail: aren't phys ptstos vProps? Or iProps?}
+to start the translation process.
 
@@
 \footnotesize
-\begin{mathpar}
-    [r](\mathsf{P}) \;: \mathsf{vProp} \; \Sigma \; \stackrel{\triangle}{=} \; \lambda \_,\; \mathsf{P}~\mathsf{r} \and
+\[\begin{array}{c}
+    [r](\mathsf{P}) \;: \mathsf{vProp} \; \Sigma \; \stackrel{\triangle}{=} \; \lambda \_,\; \mathsf{P}~\mathsf{r} \vspace{0.5em}\qquad
     \textsf{Fact P} \stackrel{\triangle}{=} \;\forall \; ,r \; r' \ldotp  \textsf{P r} \dashv\vdash \textsf{P r'}\\
-    \inferrule{}{(\textsf{P} \vdash \textsf{Q}) \vdash  ([r](\textsf{P}) \vdash  [r](\textsf{Q}))}\and
-    \inferrule{}{[r](\textsf{P} \ast \textsf{Q}) \dashv\vdash ([r](\textsf{P}) \ast [r](\textsf{Q})}\and
-    \inferrule{}{[r](\textsf{P}\land \textsf{Q}) \dashv\vdash [r](\textsf{P})\land [r](\textsf{Q})}\and
-    \inferrule{}{[r](\textsf{P}\lor \textsf{Q}) \dashv\vdash [r](\textsf{P})\lor [r](\textsf{Q})}\and
-    \inferrule{}{\textsf{Fact P} \vdash  ([r](\textsf{P}) \dashv\vdash (\textsf{P})} %(\textsf{vProp }\Sigma))\\
-\end{mathpar}
+    \inferrule{}{(\textsf{P} \vdash \textsf{Q}) \vdash  ([r](\textsf{P}) \vdash  [r](\textsf{Q}))}\qquad
+    \inferrule{}{[r](\textsf{P} \ast \textsf{Q}) \dashv\vdash ([r](\textsf{P}) \ast [r](\textsf{Q})}\vspace{0.5em}\qquad
+    \inferrule{}{[r](\textsf{P}\land \textsf{Q}) \dashv\vdash [r](\textsf{P})\land [r](\textsf{Q})}\\
+    \inferrule{}{[r](\textsf{P}\lor \textsf{Q}) \dashv\vdash [r](\textsf{P})\lor [r](\textsf{Q})}\qquad
+    \inferrule{}{\textsf{Fact P} \vdash  [r](\textsf{P}) \dashv\vdash (\textsf{P})} %(\textsf{vProp }\Sigma))\\
+\end{array}\]
+\vspace{-1em}
   \caption{Other-space Modality and Its Laws}

### 2023-11-17 examples.tex
+

@@
 \begin{array}{l}
-  \mathcal{I}\textsf{ASpace}_{\textsf{id}}(\ptablestore,\Xi,m)\stackrel{\triangle}{=}\\ \textsf{ASpace\_Lookup}(\ptablestore,\Xi,m) \ast 
+  \mathcal{I}\textsf{ASpace}_{\textsf{id}}(\ptablestore,\Xi,m)\stackrel{\triangle}{=}\\ \textsf{ASpace\_Lookup}_{\textsf{id}}(\ptablestore,\Xi,m) \ast 
   \bigast{(\vaddr, \textsf{paddr})\in \ptablestore}{\exists\;(\textsf{l4e, l3e, l2e, l1e, paddr})\ldotp \textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\vaddr\textsf{, l4e, l3e, l2e, l1e, paddr})} \\
@@
   \textsf{ where } \\
-   \textsf{ASpace\_Lookup}_{\textsf{id}}(\ptablestore,m) \stackrel{\triangle}{=} \lambda\textsf{ cr3val} \ldotp \; \exists \gammaPred \; \ldotp \ulcorner m \; !!\; \textsf{cr3val} = \textsf{Some } \gammaPred \urcorner \ast
+   \textsf{ASpace\_Lookup}_{\textsf{id}}(\ptablestore,\Xi,m) \stackrel{\triangle}{=} \lambda\textsf{ cr3val} \ldotp \; \exists \gammaPred \; \ldotp \ulcorner m \; !!\; \textsf{cr3val} = \textsf{Some } \gammaPred \urcorner \ast
    \ownGhost\gammaPred{\authfull{\ptableabswalk\ptablestore}} \ast  \ownGhostpv\gammaPred{\authfull{\pvmapping\Xi}}

### 2024-06-28 examples.tex
+

@@
 $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m)  \ast  \texttt{r14}\mapsto_{\textsf{r}} \_ \ast \texttt{rdi}\mapsto_{r} \vaddr \ast \texttt{rax}\mapsto_{\textsf{r}} \_ \ast \ulcorner \texttt{aligned va} \land  \theta \; !!\; \vaddr = \texttt{None}\urcorner}_{\rtv}$
-call ensure_L1_page ;; Returns the virtual address of the L1 entry in rax
+call pte_walkpgdir ;; Returns the virtual address of the L1 entry in rax
 $\specline{\exists (\entryf ,\;\entrytr,\; \entrytw,\; \entryo,\;\textsf{pte\_addr },\paddr) \; \ldotp\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast  \texttt{r14}\mapsto_{\textsf{r}} \_ \ast \texttt{rdi}\mapsto_{r} \vaddr \ast \texttt{rax}\mapsto_{\textsf{r}} \textsf{ pte\_addr} \; \ast }_{\rtv}$
@@
 the existence of a memory location for an appropriate L1 table entry.
-The code uses a (currently axiomatized) helper function \lstinline|ensure_L1| (discussed again in Section \ref{sec:traversing}).
-\lstinline|ensure_L1|'s postcondition contains physical pointsto assertions for a partial page table walk reaching the
+The code uses a helper function \lstinline|pte\_walkpgdir (discussed again in Section \ref{sec:traversing}).
+\lstinline|pte\_walkpgdir|'s postcondition contains physical pointsto assertions for a partial page table walk reaching the
 L1 entry (l1e), ensuring higher levels of the page table exist. 
@@
 \label{sec:traversing}
-The mapping operation of Figure \ref{fig:mapping_code} assumes an operation \lstinline|ensure_L1| which must traverse the page tables
+The mapping operation of Figure \ref{fig:mapping_code} assumes an operation \lstinline|pte\_walkpgdir| which must traverse the page tables
 in order to locate the address of the L1 entry to update --- possibly allocating pages at levels 3 and 2 in the process.

### 2024-06-28 examples.tex
+

@@
 
+\subsubsection{Identity Mappings}
 Most kernels maintain an invariant on their page tables that the virtual address of any page used for a page table lives at a virtual address
@@
 
-Work to verify the full page table walk is ongoing, but we have verified the critical step for a small x86-64 kernel, which is the physical-to-virtual
+As part of the full page table walk verification, we have verified the critical step for a small x86-64 kernel, which is the physical-to-virtual
 conversion (often appearing as a macro \texttt{P2V} in C source code). In our small kernel, as in larger kernels, \texttt{P2V} is actually just addition,
@@
 The latter uses an additional ghost set added to the per-address-space
-invariant per Figure \ref{fig:peraspaceinvariant_with_p2v_extension}. The extended invariant owns, for each memory location dedicated to page tables, some fraction of
+invariant per Figure \ref{fig:peraspaceinvariant_with_p2v_extension}.\todo{should we also discuss the following before while introducing the invariant?} The extended invariant owns, for each memory location dedicated to page tables, some fraction of
 that physical memory, and also guarantees the offset is the correct calculation.
@@
 This is a simple piece of code whose functionality is critical and whose correctness is highly non-trivial. No prior work engages with this problem.
+\subsubsection{Traversing Page Tables}
+
+\begin{figure}\footnotesize
+\begin{lstlisting}[mathescape]
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \ulcorner \textsf{OFF} + \paddr > 0 \urcorner \ast \texttt{rbx}\mapsto_{\textsf{r}} \paddr \ast \sumapaces\rtv\delta  \ast \sumpv\paddr\qfrac\true  }_{\rtv}$
+add rbx, OFF ;; shift physical address by identity mapping constant OFF
+$\specline{\textsf{P} \ast \exists \qfraczero; \qfrac,\textsf{ w, val},  \texttt{rbx}\mapsto_{\textsf{r}} \paddr + \textsf{OFF} \ast \sumapaces\rtv\delta \ast \sumpv\paddr\qfrac\true \ast \vaddr \mapsto_{\texttt{vpte}} \{\textsf{q}, \textsf{qfrac}\} \paddr \;(\texttt{val})}$
+$\specline{\ulcorner \vaddr = \paddr + \textsf{OFF} \urcorner \ast \ulcorner \textsf{qfrac} = 1 \leftrightarrow \; \lnot\textsf{entry\_present }(\vale) \urcorner \ast \sumwalkabs\vaddr\qfrac\paddr \ast (R \wand \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)) }_{\rtv}$
+\end{lstlisting}
+\vspace{-1em}
+\caption{Traversing page-tables, and allocating entries as needed while mapping-a-page in Figure \ref{fig:mappingcode}.}
+\label{fig:walkpgdir}
+\end{figure}
 

### 2024-06-28 examples.tex
compiles

@@
 the existence of a memory location for an appropriate L1 table entry.
-The code uses a helper function \lstinline|pte\_walkpgdir (discussed again in Section \ref{sec:traversing}).
+The code uses a helper function \lstinline|pte\_walkpgdir| (discussed again in Section \ref{sec:traversing}).
 \lstinline|pte\_walkpgdir|'s postcondition contains physical pointsto assertions for a partial page table walk reaching the
@@
 The latter uses an additional ghost set added to the per-address-space
-invariant per Figure \ref{fig:peraspaceinvariant_with_p2v_extension}.\todo{should we also discuss the following before while introducing the invariant?} The extended invariant owns, for each memory location dedicated to page tables, some fraction of
+invariant per Figure \ref{fig:peraspaceinvariant_with_p2v_extension}. The extended invariant owns, for each memory location dedicated to page tables, some fraction of
 that physical memory, and also guarantees the offset is the correct calculation.
@@
 \begin{lstlisting}[mathescape]
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \ulcorner \textsf{OFF} + \paddr > 0 \urcorner \ast \texttt{rbx}\mapsto_{\textsf{r}} \paddr \ast \sumapaces\rtv\delta  \ast \sumpv\paddr\qfrac\true  }_{\rtv}$
-add rbx, OFF ;; shift physical address by identity mapping constant OFF
-$\specline{\textsf{P} \ast \exists \qfraczero; \qfrac,\textsf{ w, val},  \texttt{rbx}\mapsto_{\textsf{r}} \paddr + \textsf{OFF} \ast \sumapaces\rtv\delta \ast \sumpv\paddr\qfrac\true \ast \vaddr \mapsto_{\texttt{vpte}} \{\textsf{q}, \textsf{qfrac}\} \paddr \;(\texttt{val})}$
-$\specline{\ulcorner \vaddr = \paddr + \textsf{OFF} \urcorner \ast \ulcorner \textsf{qfrac} = 1 \leftrightarrow \; \lnot\textsf{entry\_present }(\vale) \urcorner \ast \sumwalkabs\vaddr\qfrac\paddr \ast (R \wand \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)) }_{\rtv}$
+push   rbp
+mov    rsp,rbp
+sub    0x20,rsp
+mov    rdi,-0x8[rbp]
+
+mov    -0x8[rbp],rdi
+mov    [rdi],rdi
+and    0x1,rdi
+mov    rdi,rax
+cmp    0x0,rax
+jne    161 <pte_get_next_table_succ+0xa1>
+
+lea    -0x10[rbp],rdi
+callq  70 <pte_initialize>
+mov    -0x8[rbp],rdi
+mov    [rdi],rax
+and    0xfffffffffffffffd,rax
+or     0x2,rax
+mov    rax,[rdi]
+
+mov    -0x8[rbp],rax
+mov    [rax],rdi
+and    0xfffffffffffffffb,%rdi
+or     0x4,rdi
+mov    rdi,[rax]
+
+mov    -0x10[rbp],rax
+movabs 0x0,rdi	115: R_X86_64_64	KERN_BASE
+sub    rdi,rax
+shr    0xc,rax
+mov    -0x8[rbp],rdi
+mov    [rdi],rcx
+movabs 0xffffffffff,rdx
+and    rdx,rax
+shl    0xc,rax
+movabs 0xfff0000000000fff,rdx
+and    rdx,rcx
+or     rax,rcx
+mov    rcx,[rdi]
+
+mov    -0x8[rbp],rax
+mov    [rax],rcx
+and    0xfffffffffffffffe,rcx
+or     0x1,rcx
+mov    rcx,[rax]
+
+mov    -0x8[rbp],rax
+mov    [rax],rax
+shr    0xc,rax
+movabs 0xffffffffff,rcx
+and    rcx,rax
+shl    0xc,rax
+mov    rax,-0x18[rbp]
+
+mov    -0x18[rbp],rax
+movabs 0x0,rcx	187: R_X86_64_64	KERN_BASE
+add    rcx,rax
+mov    rax,-0x20[rbp]
+mov    -0x20[rbp],rax
+mov    rax,-0x10[rbp]
+
+ mov    -0x10[rbp],rax
+ add    0x20,rsp
+ pop    rbp
+ ret
 \end{lstlisting}

### 2024-06-28 examples.tex
+

--- 2024-06-28 examples.tex
@@
 \subsubsection{Traversing Page Tables}
+As we explain in Section \ref{}, a traversal over a page-table tree can be done through the indices into the tables at each level followed by obtaining the address of the next table.
+In Figure \ref{fig:getnexttable}, we see how to obtain a
+\begin{figure}\footnotesize
+ \begin{lstlisting}[mathescape]
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \ulcorner \textsf{OFF} + \paddr > 0 \urcorner \ast \texttt{rbx}\mapsto_{\textsf{r}} \paddr \ast \sumapaces\rtv\delta  \ast \sumpv\paddr\qfrac\true  }_{\rtv}$
+push   rbp
+mov    rsp,rbp
+sub    0x20,rsp
+mov    rdi,-0x8[rbp]
+;;  pte_t *next;
+
+mov    -0x8[rbp],rdi
+mov    [rdi],rdi
+and    0x1,rdi
+mov    rdi,rax
+
+;;  if (!entry->present) {
+cmp    0x0,rax
+jne    161 <pte_get_next_table_succ+0xa1>
+
+lea    -0x10[rbp],rdi
+
+;;pte_initialize(&next);
+callq  70 <pte_initialize>
+mov    -0x8[rbp],rdi
+mov    [rdi],rax
+
+;; entry->writable = 1
+;;entry->user_acc = 1;
+;;entry->pfn = PTE_ADDR_TO_PHYS_PFN((uintptr_t) next);
+;;entry->present = 1;
+and    0xfffffffffffffffd,rax
+or     0x2,rax
+mov    rax,[rdi]
+...
+;;}
+
+;;uintptr_t next_virt_addr = (uintptr_t) P2V(next_phys_addr);
+mov    -0x8[rbp],rax
+mov    [rax],rax
+shr    0xc,rax
+movabs 0xffffffffff,rcx
+and    rcx,rax
+shl    0xc,rax
+mov    rax,-0x18[rbp]
+
+;;uintptr_t next_virt_addr = (uintptr_t) P2V(next_phys_addr);
+mov    -0x18[rbp],rax
+movabs 0x0,rcx	187: R_X86_64_64	KERN_BASE
+add    rcx,rax
+
+;;next = (pte_t *) next_virt_addr;
+mov    rax,-0x20[rbp]
+mov    -0x20[rbp],rax
+mov    rax,-0x10[rbp]
+
+;;return next
+mov    -0x10[rbp],rax
+add    0x20,rsp
+pop    rbp
+ret
+\end{lstlisting}
+\vspace{-1em}
+\caption{ in Figure \ref{fig:walkpgdir}.}
+\label{fig:getnexttable}
+\end{figure}
 
@@
   \begin{lstlisting}[mathescape]
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \ulcorner \textsf{OFF} + \paddr > 0 \urcorner \ast \texttt{rbx}\mapsto_{\textsf{r}} \paddr \ast \sumapaces\rtv\delta  \ast \sumpv\paddr\qfrac\true  }_{\rtv}$
-add rbx, OFF ;; shift physical address by identity mapping constant OFF
-$\specline{\textsf{P} \ast \exists \qfraczero; \qfrac,\textsf{ w, val},  \texttt{rbx}\mapsto_{\textsf{r}} \paddr + \textsf{OFF} \ast \sumapaces\rtv\delta \ast \sumpv\paddr\qfrac\true \ast \vaddr \mapsto_{\texttt{vpte}} \{\textsf{q}, \textsf{qfrac}\} \paddr \;(\texttt{val})}$
-$\specline{\ulcorner \vaddr = \paddr + \textsf{OFF} \urcorner \ast \ulcorner \textsf{qfrac} = 1 \leftrightarrow \; \lnot\textsf{entry\_present }(\vale) \urcorner \ast \sumwalkabs\vaddr\qfrac\paddr \ast (R \wand \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)) }_{\rtv}$
+;;pte_t *walkpgdir_succ(pte_t *pml4, const void *va) {
+push   rbp
+mov    rsp,rbp
+sub    0x40,rsp
+mov    rdi,-0x8[rbp]
+mov    rsi,-0x10[rbp]
+
+;;pte_t *pml4_entry = &pml4[PML4EX(va)];
+mov    -0x8[rbp],[rsi]
+mov    -0x10[rbp],rdi
+shr    0x27,rdi
+and    0x1ff,rdi
+shl    0x3,rdi
+add    rdi,rsi
+mov    rsi,-0x18[rbp]
+
+;;pte_t *pdp = pte_get_next_table_succ(pml4_entry);
+mov    -0x18[rbp],rdi
+callq  <pte_get_next_table_succ>
+mov    rax,-0x20[rbp]
+
+;;pte_t *pdp_entry = &pdp[PDPEX(va)];
+mov    -0x20[rbp],rax
+mov    -0x10[rbp],rsi
+shr    0x1e,rsi
+and    0x1ff,rsi
+shl    0x3,rsi
+add    rsi,rax
+mov    rax,-0x28[rbp]
+
+;;pte_t *pd = pte_get_next_table_succ(pdp_entry);
+mov    -0x28[rbp],rdi
+callq  <pte_get_next_table_succ>
+mov    rax,-0x30[rbp]
+
+;;pte_t *pd_entry = &pd[PDEX(va)];
+mov    -0x30[rbp],rax
+mov    -0x10[rbp],rsi
+shr    0x15,rsi
+and    0x1ff,rsi
+shl    0x3,rsi
+add    rsi,rax
+mov    rax,-0x38[rbp]
+
+;;pte_t *pt = pte_get_next_table_succ(pd_entry);
+mov    -0x38[rbp],rdi
+callq  <pte_get_next_table_succ>
+mov    rax,-0x40[rbp]
+
+;;return &pt[PTEX(va)];
+mov    -0x40[rbp],rax
+mov    -0x10[rbp],rsi
+shr    0xc,rsi
+and    0x1ff,rsi
+shl    0x3,rsi
+add    rsi,rax
+add    0x40,rsp
+pop    rbp
+retq   
 \end{lstlisting}
@@
 \end{figure}
-
 % \citet{kolanski08vstte,kolanski09tphols} verified a single code block with their logic which was roughly Figure \ref{fig:mapping_code} for a 2-level ARM

### 2024-06-29 examples.tex
sync with desktop

@@
 As we explain in Section \ref{}, a traversal over a page-table tree can be done through the indices into the tables at each level followed by obtaining the address of the next table.
-In Figure \ref{fig:getnexttable}, we see how to obtain a
+In Figure \ref{fig:getnexttable}, we see how to obtain a physicall address of the next level page-table.
 \begin{figure}\footnotesize
@@
 
-;; entry->writable = 1
+;;setting the fields of the entry with a fresh allocated page
+;;entry->writable = 1
 ;;entry->user_acc = 1;

### 2024-06-29 examples.tex
+

@@
 As we explain in Section \ref{}, a traversal over a page-table tree can be done through the indices into the tables at each level followed by obtaining the address of the next table.
-In Figure \ref{fig:getnexttable}, we see how to obtain a physicall address of the next level page-table.
+In Figure \ref{fig:getnexttable}, we see how to obtain the physicall address of the next level page-table.
+
+\begin{figure}\footnotesize
+ \begin{lstlisting}[mathescape]
+;;void pte_initialize(pte_t **table) {
+push   rbp
+mov    rsp,rbp
+sub    0x10,rsp
+mov    rdi,-0x8[rbp]
+
+;;Allocate a full page for 512 8-byte entries
+;;pte_t *local = kalloc();
+;;clear the entire table. Ensures present (and reserved, etc.) bits are 0
+callq  81 <pte_initialize+0x11>	7d: R_X86_64_PLT32	kalloc-0x4
+mov    rax,-0x10[rbp]
+
+;;*table = local;
+mov    -0x10[rbp],rax
+mov    -0x8[rbp],rdi
+mov    rax,[rdi]
+;;return;
+add    0x10,rsp
+pop    rbp
+retq   
+  \end{lstlisting}
+\vspace{-1em}
+\caption{ in Figure \ref{fig:getnexttable}.}
+\label{fig:pteinitialize}
+\end{figure}
+
 \begin{figure}\footnotesize
@@
    $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \ulcorner \textsf{OFF} + \paddr > 0 \urcorner \ast \texttt{rbx}\mapsto_{\textsf{r}} \paddr \ast \sumapaces\rtv\delta  \ast \sumpv\paddr\qfrac\true  }_{\rtv}$
+;; pte_t *pte_get_next_table_succ(pte_t *entry) {
 push   rbp

### 2024-06-29 examples.tex
+

@@
 In Figure \ref{fig:getnexttable}, we see how to obtain the physicall address of the next level page-table.
-
-\begin{figure}\footnotesize
- \begin{lstlisting}[mathescape]
+traver
+\begin{figure*}[t!]
+\begin{tabular}{p{0.47\textwidth}p{0.47\textwidth}}
+  \begin{lstlisting}[mathescape,basicstyle=\scriptsize\ttfamily]
 ;;void pte_initialize(pte_t **table) {
-push   rbp
-mov    rsp,rbp
-sub    0x10,rsp
-mov    rdi,-0x8[rbp]
+...
 
@@
 mov    rax,[rdi]
+
 ;;return;
-add    0x10,rsp
-pop    rbp
-retq   
-  \end{lstlisting}
-\vspace{-1em}
-\caption{ in Figure \ref{fig:getnexttable}.}
-\label{fig:pteinitialize}
-\end{figure}
+...
 
-\begin{figure}\footnotesize
- \begin{lstlisting}[mathescape]
+  \end{lstlisting}&
+  \begin{lstlisting}[mathescape,basicstyle=\scriptsize\ttfamily]
    $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \ulcorner \textsf{OFF} + \paddr > 0 \urcorner \ast \texttt{rbx}\mapsto_{\textsf{r}} \paddr \ast \sumapaces\rtv\delta  \ast \sumpv\paddr\qfrac\true  }_{\rtv}$
@@
 ret
+
   \end{lstlisting}
+  \end{tabular}
 \vspace{-1em}
@@
 \label{fig:getnexttable}
-\end{figure}
+\end{figure*}
 
 \begin{figure}\footnotesize
-  \begin{lstlisting}[mathescape]
-;;pte_t *walkpgdir_succ(pte_t *pml4, const void *va) {
+  \begin{lstlisting}[mathescape,basicstyle=\scriptsize\ttfamily]
+;;pte_t *walkpgdir_succ(pte_t *pml4, const void *va) 
 push   rbp

### 2024-06-29 examples.tex
+

@@
 In Figure \ref{fig:getnexttable}, we see how to obtain the physicall address of the next level page-table.
-traver
-\begin{figure*}[t!]
+Traversal start from the root address of the L4 table (\textsf{pte\_t *pml4} as a parameter to \textsf{walkpgdir}).
+Then, we obtain the address of the entry (\textsf{pml4[PML4EX(va)]}) via the L4 index in the virtual address (\textsf{va}).
+Finally, we call \textsf{get\_next\_table} to get the virtual address to the next level page-table -- L3.
+\begin{figure*}[ht]
 \begin{tabular}{p{0.47\textwidth}p{0.47\textwidth}}
@@
 
+;;pte_t *walkpgdir_succ(pte_t *pml4, const void *va) 
+push   rbp
+mov    rsp,rbp
+sub    0x40,rsp
+mov    rdi,-0x8[rbp]
+mov    rsi,-0x10[rbp]
+;;pte_t *pml4_entry = &pml4[PML4EX(va)];
+mov    -0x8[rbp],[rsi]
+mov    -0x10[rbp],rdi
+shr    0x27,rdi
+and    0x1ff,rdi
+shl    0x3,rdi
+add    rdi,rsi
+mov    rsi,-0x18[rbp]
+;;pte_t *pdp = pte_get_next_table_succ(pml4_entry);
+mov    -0x18[rbp],rdi
+callq  <pte_get_next_table_succ>
+mov    rax,-0x20[rbp]
+;;pte_t *pdp_entry = &pdp[PDPEX(va)];
+mov    -0x20[rbp],rax
+mov    -0x10[rbp],rsi
+shr    0x1e,rsi
+and    0x1ff,rsi
+shl    0x3,rsi
+add    rsi,rax
+mov    rax,-0x28[rbp]
+;;pte_t *pd = pte_get_next_table_succ(pdp_entry);
+mov    -0x28[rbp],rdi
+callq  <pte_get_next_table_succ>
+mov    rax,-0x30[rbp]
+;;pte_t *pd_entry = &pd[PDEX(va)];
+mov    -0x30[rbp],rax
+mov    -0x10[rbp],rsi
+shr    0x15,rsi
+and    0x1ff,rsi
+shl    0x3,rsi
+add    rsi,rax
+mov    rax,-0x38[rbp]
+;;pte_t *pt = pte_get_next_table_succ(pd_entry);
+mov    -0x38[rbp],rdi
+callq  <pte_get_next_table_succ>
+mov    rax,-0x40[rbp]
+;;return &pt[PTEX(va)];
+mov    -0x40[rbp],rax
+mov    -0x10[rbp],rsi
+shr    0xc,rsi
+and    0x1ff,rsi
+shl    0x3,rsi
+add    rsi,rax
+add    0x40,rsp
+pop    rbp
+retq   
   \end{lstlisting}&
@@
 ;;  pte_t *next;
-
 mov    -0x8[rbp],rdi
@@
 mov    rdi,rax
-
 ;;  if (!entry->present) {
@@
 jne    161 <pte_get_next_table_succ+0xa1>
-
 lea    -0x10[rbp],rdi
-
 ;;pte_initialize(&next);
@@
 mov    [rdi],rax
-
 ;;setting the fields of the entry with a fresh allocated page
@@
 ;;}
-
 ;;uintptr_t next_virt_addr = (uintptr_t) P2V(next_phys_addr);
@@
 mov    rax,-0x18[rbp]
-
 ;;uintptr_t next_virt_addr = (uintptr_t) P2V(next_phys_addr);
@@
 add    rcx,rax
-
 ;;next = (pte_t *) next_virt_addr;
@@
 mov    rax,-0x10[rbp]
-
 ;;return next
@@
 ret
-
   \end{lstlisting}
@@
 
-\begin{figure}\footnotesize
-  \begin{lstlisting}[mathescape,basicstyle=\scriptsize\ttfamily]
-;;pte_t *walkpgdir_succ(pte_t *pml4, const void *va) 
-push   rbp
-mov    rsp,rbp
-sub    0x40,rsp
-mov    rdi,-0x8[rbp]
-mov    rsi,-0x10[rbp]
-
-;;pte_t *pml4_entry = &pml4[PML4EX(va)];
-mov    -0x8[rbp],[rsi]
-mov    -0x10[rbp],rdi
-shr    0x27,rdi
-and    0x1ff,rdi
-shl    0x3,rdi
-add    rdi,rsi
-mov    rsi,-0x18[rbp]
-
-;;pte_t *pdp = pte_get_next_table_succ(pml4_entry);
-mov    -0x18[rbp],rdi
-callq  <pte_get_next_table_succ>
-mov    rax,-0x20[rbp]
-
-;;pte_t *pdp_entry = &pdp[PDPEX(va)];
-mov    -0x20[rbp],rax
-mov    -0x10[rbp],rsi
-shr    0x1e,rsi
-and    0x1ff,rsi
-shl    0x3,rsi
-add    rsi,rax
-mov    rax,-0x28[rbp]
-
-;;pte_t *pd = pte_get_next_table_succ(pdp_entry);
-mov    -0x28[rbp],rdi
-callq  <pte_get_next_table_succ>
-mov    rax,-0x30[rbp]
-
-;;pte_t *pd_entry = &pd[PDEX(va)];
-mov    -0x30[rbp],rax
-mov    -0x10[rbp],rsi
-shr    0x15,rsi
-and    0x1ff,rsi
-shl    0x3,rsi
-add    rsi,rax
-mov    rax,-0x38[rbp]
-
-;;pte_t *pt = pte_get_next_table_succ(pd_entry);
-mov    -0x38[rbp],rdi
-callq  <pte_get_next_table_succ>
-mov    rax,-0x40[rbp]
-
-;;return &pt[PTEX(va)];
-mov    -0x40[rbp],rax
-mov    -0x10[rbp],rsi
-shr    0xc,rsi
-and    0x1ff,rsi
-shl    0x3,rsi
-add    rsi,rax
-add    0x40,rsp
-pop    rbp
-retq   
-\end{lstlisting}
-\vspace{-1em}
-\caption{Traversing page-tables, and allocating entries as needed while mapping-a-page in Figure \ref{fig:mappingcode}.}
-\label{fig:walkpgdir}
-\end{figure}
+%\caption{Traversing page-tables, and allocating entries as needed while mapping-a-page in Figure \ref{fig:mappingcode}.}
 % \citet{kolanski08vstte,kolanski09tphols} verified a single code block with their logic which was roughly Figure \ref{fig:mapping_code} for a 2-level ARM

### 2024-06-29 examples.tex
+

@@
 ;;pte_t *pdp_entry = &pdp[PDPEX(va)];
-mov    -0x20[rbp],rax
-mov    -0x10[rbp],rsi
-shr    0x1e,rsi
-and    0x1ff,rsi
-shl    0x3,rsi
-add    rsi,rax
-mov    rax,-0x28[rbp]
-;;pte_t *pd = pte_get_next_table_succ(pdp_entry);
-mov    -0x28[rbp],rdi
-callq  <pte_get_next_table_succ>
-mov    rax,-0x30[rbp]
-;;pte_t *pd_entry = &pd[PDEX(va)];
-mov    -0x30[rbp],rax
-mov    -0x10[rbp],rsi
-shr    0x15,rsi
-and    0x1ff,rsi
-shl    0x3,rsi
-add    rsi,rax
-mov    rax,-0x38[rbp]
-;;pte_t *pt = pte_get_next_table_succ(pd_entry);
-mov    -0x38[rbp],rdi
+...
+...
+...
+;;last call at L1 for virtual address to physical page 
 callq  <pte_get_next_table_succ>

### 2024-06-29 examples.tex
+

@@
 Traversal start from the root address of the L4 table (\textsf{pte\_t *pml4} as a parameter to \textsf{walkpgdir}).
-Then, we obtain the address of the entry (\textsf{pml4[PML4EX(va)]}) via the L4 index in the virtual address (\textsf{va}).
+Then, we obtain the address of the entry in the L4 table via the index in the virtual address (\textsf{va}).
 Finally, we call \textsf{get\_next\_table} to get the virtual address to the next level page-table -- L3.
+Withing \textsf{get\_next\_table} first checks if the entry whose address is passed as a parameter -- e.g., for L4 \textsf{pml4[PML4EX(va)]}.
+In case of absence, \textsf{pte\_initialize} is called for a physical-zeroed page (\textsf{kalloc} -- currently an axiom as it is not part of \textsf{pte} module) to be backing the physical frame address of the entry --  \textsf{entry->pfn = PTE\_ADDR\_TO\_PHYS\_PFN((uintptr\_t) next)}.
 \begin{figure*}[ht]

### 2024-06-29 examples.tex
+

--- 2024-06-29 examples.tex
@@
 Traversal start from the root address of the L4 table (\textsf{pte\_t *pml4} as a parameter to \textsf{walkpgdir}).
-Then, we obtain the address of the entry (\textsf{pml4[PML4EX(va)]}) via the L4 index in the virtual address (\textsf{va}).
+Then, we obtain the address of the entry in the L4 table via the index in the virtual address (\textsf{va}).
+\[
+\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \ulcorner \textsf{OFF} + \paddr > 0 \urcorner \ast \texttt{rbx}\mapsto_{\textsf{r}} \paddr \ast \sumapaces\rtv\delta  \ast \sumpv\paddr\qfrac\true  }_{\rtv}
+\]
 Finally, we call \textsf{get\_next\_table} to get the virtual address to the next level page-table -- L3.
+\[
+\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \ulcorner \textsf{OFF} + \paddr > 0 \urcorner \ast \texttt{rbx}\mapsto_{\textsf{r}} \paddr \ast \sumapaces\rtv\delta  \ast \sumpv\paddr\qfrac\true  }_{\rtv}
+\]
+Within \textsf{get\_next\_table} first checks if the entry whose address is passed as a parameter -- e.g., for L4 \textsf{pml4[PML4EX(va)]}.
+\[
+\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \ulcorner \textsf{OFF} + \paddr > 0 \urcorner \ast \texttt{rbx}\mapsto_{\textsf{r}} \paddr \ast \sumapaces\rtv\delta  \ast \sumpv\paddr\qfrac\true  }_{\rtv}
+\]
+In case of absence, \textsf{pte\_initialize} is called for a physical-zeroed page (\textsf{kalloc} -- currently an axiom as it is not part of \textsf{pte} module) to be backing the physical frame address of the entry --  \textsf{entry->pfn = PTE\_ADDR\_TO\_PHYS\_PFN((uintptr\_t) next)}.
+Once we make sure that the frame of the page-table entry is backed with an allocated physical page,
 \begin{figure*}[ht]

### 2024-06-30 examples.tex
+

--- 2024-06-29 examples.tex
@@
 
+As we explain in Section \ref{}, a traversal over a page-table tree can be done through the indices into the tables at each level followed by obtaining the address of the next table.
+In Figure \ref{fig:getnexttable}, we see how to obtain the physicall address of the next level page-table.
+Traversal start from the root address of the L4 table (\textsf{pte\_t *pml4} given as a parameter to \textsf{walkpgdir}).
+Then, we obtain the address of the entry in the L4 table via the index in the virtual address (\textsf{va}).
+
+Finally, we call \textsf{get\_next\_table} to get the virtual address to the next level page-table -- L3.
+Within \textsf{get\_next\_table} first checks if the entry whose address is passed as a parameter -- e.g., for L4 \textsf{pml4[PML4EX(va)]}.
+In case of absence, \textsf{pte\_initialize} is called for a physical-zeroed page (\textsf{kalloc} -- currently an axiom as it is not part of \textsf{pte} module) to be backing the physical frame address of the entry.
+
+Once we make sure that the frame of the page-table entry is backed with an allocated physical page, we can obtain the virtual address of the next level-page table (\textsf{next} of function \textsf{get\_next\_table} in Figure \ref{}) via looking up the \textit{identity mapping} from the physical address (the address value referenced by \textsf{entry->pfn}) to virtual address value to be returned by \textsf{get\_next\_table}.
 \subsubsection{Identity Mappings}
@@
 \begin{lstlisting}[mathescape]
+\[
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \ulcorner\urcorner \ast \texttt{rbx}\mapsto_{\textsf{r}} \paddr \ast \sumapaces\rtv\delta  \ast \sumpv\paddr\qfrac\true  }_{\rtv}$
+\mathsf{pte\_initialize}(\mathsf{next})
+...
   $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \ulcorner \textsf{OFF} + \paddr > 0 \urcorner \ast \texttt{rbx}\mapsto_{\textsf{r}} \paddr \ast \sumapaces\rtv\delta  \ast \sumpv\paddr\qfrac\true  }_{\rtv}$
@@
 \vspace{-1em}
-\caption{Converting a physical address of a PTE to a virtual address (w/o instruction pointer or flag updates). $R$ abbreviates the resources
+\caption{Converting a physical address, whose exsitence is made certain, of a PTE to a virtual address (w/o instruction pointer or flag updates). $R$ abbreviates the resources
 associated with \paddr\ by $\Xi$ in Figure \ref{fig:peraspaceinvariant_with_p2v_extension}.}
@@
 This is a simple piece of code whose functionality is critical and whose correctness is highly non-trivial. No prior work engages with this problem.
-\subsubsection{Traversing Page Tables}
-As we explain in Section \ref{}, a traversal over a page-table tree can be done through the indices into the tables at each level followed by obtaining the address of the next table.
-In Figure \ref{fig:getnexttable}, we see how to obtain the physicall address of the next level page-table.
-Traversal start from the root address of the L4 table (\textsf{pte\_t *pml4} as a parameter to \textsf{walkpgdir}).
-Then, we obtain the address of the entry (\textsf{pml4[PML4EX(va)]}) via the L4 index in the virtual address (\textsf{va}).
-Finally, we call \textsf{get\_next\_table} to get the virtual address to the next level page-table -- L3.
+
 \begin{figure*}[ht]

### 2024-06-30 examples.tex
+
sync with desktop

@@
 In case of absence, \textsf{pte\_initialize} is called for a physical-zeroed page (\textsf{kalloc} -- currently an axiom as it is not part of \textsf{pte} module) to be backing the physical frame address of the entry.
+\begin{figure}\footnotesize
+\begin{lstlisting}[mathescape]
+\[
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \ulcorner\urcorner \ast \texttt{rbx}\mapsto_{\textsf{r}} \paddr \ast \sumapaces\rtv\delta  \ast \sumpv\paddr\qfrac\true  }_{\rtv}$
+
+;;  if (!entry->present) {
+cmp    0x0,rax
+jne    161 <pte_get_next_table_succ+0xa1>
+lea    -0x10[rbp],rdi
+;;pte_initialize(&next);
+callq  70 <pte_initialize>
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \ulcorner\urcorner \ast \texttt{rbx}\mapsto_{\textsf{r}} \paddr \ast \sumapaces\rtv\delta  \ast \sumpv\paddr\qfrac\true  }_{\rtv}$
+...
+
+;;}
+\end{lstlisting}
+\vspace{-1em}
+\caption{Making sure that \textsf{entry}'s frame is referencing to an address backed up by a physical page.}
+\label{fig:calltopteinitialize}
+\end{figure}
 
@@
 $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \ulcorner\urcorner \ast \texttt{rbx}\mapsto_{\textsf{r}} \paddr \ast \sumapaces\rtv\delta  \ast \sumpv\paddr\qfrac\true  }_{\rtv}$
-\mathsf{pte\_initialize}(\mathsf{next})
+
+;;if (!entry->present) {
   ...
+
+;;}
+;; uintptr_t next_phys_addr = PTE_PFN_TO_ADDR(entry->pfn);
+;;uintptr_t next_virt_addr = (uintptr_t) P2V(next_phys_addr);
   $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \ulcorner \textsf{OFF} + \paddr > 0 \urcorner \ast \texttt{rbx}\mapsto_{\textsf{r}} \paddr \ast \sumapaces\rtv\delta  \ast \sumpv\paddr\qfrac\true  }_{\rtv}$
-add rbx, OFF ;; shift physical address by identity mapping constant OFF
+add rbx, KERNBASE ;; shift physical address by identity mapping constant OFF
 $\specline{\textsf{P} \ast \exists \qfraczero; \qfrac,\textsf{ w, val},  \texttt{rbx}\mapsto_{\textsf{r}} \paddr + \textsf{OFF} \ast \sumapaces\rtv\delta \ast \sumpv\paddr\qfrac\true \ast \vaddr \mapsto_{\texttt{vpte}} \{\textsf{q}, \textsf{qfrac}\} \paddr \;(\texttt{val})}$
@@
 As part of the full page table walk verification, we have verified the critical step for a small x86-64 kernel, which is the physical-to-virtual
-conversion (often appearing as a macro \texttt{P2V} in C source code). In our small kernel, as in larger kernels, \texttt{P2V} is actually just addition,
+conversion (often appearing as a macro \texttt{P2V} inside the macro \textsf{ PTE\_PFN\_TO\_ADDR} in C source code). In our small kernel, as in larger kernels, \texttt{P2V} is actually just addition,
 but the correctness of this simple instruction is quite subtle, as shown in Figure \ref{fig:p2v}.

### 2024-06-30 examples.tex
+

@@
 
-As we explain in Section \ref{}, a traversal over a page-table tree can be done through the indices into the tables at each level followed by obtaining the address of the next table.
-In Figure \ref{fig:getnexttable}, we see how to obtain the physicall address of the next level page-table.
-Traversal start from the root address of the L4 table (\textsf{pte\_t *pml4} given as a parameter to \textsf{walkpgdir}).
-Then, we obtain the address of the entry in the L4 table via the index in the virtual address (\textsf{va}).
+\subsubsection{Loading Page-Table Address Value}
+In Figure \ref{fig:getnexttable}, we see how to obtain the physical address of the next level page-table.
+Traversal start from the root address of the page-table tree (the address of L4 table \textsf{pte\_t *pml4}) given as a parameter to \textsf{pte\_walkpgdir}).
+Then, we obtain the address of the entry (\textsf{pml4[PML4EX(va)]}) in the L4 table via the index in the virtual address (\textsf{va}).
 
-Finally, we call \textsf{get\_next\_table} to get the virtual address to the next level page-table -- L3.
-Within \textsf{get\_next\_table} first checks if the entry whose address is passed as a parameter -- e.g., for L4 \textsf{pml4[PML4EX(va)]}.
-In case of absence, \textsf{pte\_initialize} is called for a physical-zeroed page (\textsf{kalloc} -- currently an axiom as it is not part of \textsf{pte} module) to be backing the physical frame address of the entry.
+Finally, we call \textsf{get\_next\_table} to get the virtual address to the next level page-table -- at this point the address of L3 table.
+Within \textsf{get\_next\_table} first checks if the entry whose address is passed as a parameter ($\textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry}$).
+To be able to do so, it needs to load the entry value (a 64-bits word divided into bit-fields for corresponding fileds shown in Section {}) and check that \textit{presence} bit of the entry is set or not.
 \begin{figure}\footnotesize
   \begin{lstlisting}[mathescape]
-\[
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \ulcorner\urcorner \ast \texttt{rbx}\mapsto_{\textsf{r}} \paddr \ast \sumapaces\rtv\delta  \ast \sumpv\paddr\qfrac\true  }_{\rtv}$
-
-;;  if (!entry->present) {
-cmp    0x0,rax
-jne    161 <pte_get_next_table_succ+0xa1>
-lea    -0x10[rbp],rdi
-;;pte_initialize(&next);
-callq  70 <pte_initialize>
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \ulcorner\urcorner \ast \texttt{rbx}\mapsto_{\textsf{r}} \paddr \ast \sumapaces\rtv\delta  \ast \sumpv\paddr\qfrac\true  }_{\rtv}$
-...
-
-;;}
+;;if(entry->present)
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast r8  \mapsto_{\textsf{r}} \textsf{\_} \ast rdi  \mapsto_{\textsf{r}} \textsf{\_} \ast \sumapaces\rtv\delta  \ast \sumpv\paddr\qfrac\true  }_{\rtv}$
+mov    -0x8[rbp],rdi
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast r8  \mapsto_{\textsf{r}} \textsf{\_} \ast rdi  \mapsto_{\textsf{r}} \textsf{entry} \ast \textsf{entry} \ast \sumapaces\rtv\delta  \ast \sumpv\paddr\qfrac\true  }_{\rtv}$
+mov     rdi, r8
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast r8  \mapsto_{\textsf{r}} \textsf{entry} \textsf{\_} \ast rdi  \mapsto_{\textsf{r}} \textsf{entry} \ast \sumapaces\rtv\delta  \ast \sumpv\paddr\qfrac\true  }_{\rtv}$
+mov    [r8],rdi
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast r8  \mapsto_{\textsf{r}} \textsf{entry} \textsf{\_} \ast rdi  \mapsto_{\textsf{r}} \textsf{entry} \ast \sumapaces\rtv\delta  \ast \sumpv\paddr\qfrac\true  }_{\rtv}$
   \end{lstlisting}
@@
 \caption{Making sure that \textsf{entry}'s frame is referencing to an address backed up by a physical page.}
-\label{fig:calltopteinitialize}
+\label{fig:loadentry}
 \end{figure}
-
-Once we make sure that the frame of the page-table entry is backed with an allocated physical page, we can obtain the virtual address of the next level-page table (\textsf{next} of function \textsf{get\_next\_table} in Figure \ref{}) via looking up the \textit{identity mapping} from the physical address (the address value referenced by \textsf{entry->pfn}) to virtual address value to be returned by \textsf{get\_next\_table}.
+As we see in Figure \ref{fig:loadentry}, we have the physical address (\textsf{entry}) of the page table entry kept in the register $r8$, but not the memory mapping to load the entry value referenced by the physical address \textsf{entry}.
 \subsubsection{Identity Mappings}
@@
   \end{figure*}
+Now, to read the value referenced by physical address \textsf{entry} in Figure \ref{fig:loadentry}, first, we open our the addresss-space invariant ($\mathcal{I}\textsf{ASpace}_{\textsf{id}}(\ptablestore,\Xi,m)$), and obtain the virtual mapping out of the identity mapping $\Xi$ with the side condition
+\[
+ \ulcorner \textsf{qfrac} = 1 \leftrightarrow \; \lnot\textsf{entry\_present }(\vale) \urcorner
+ \]
+ assuring that looking up identity mappings are safe, because the full ownership asserts the non-presence of the entry which requires allocation a new physical page and setting the presence bit.
+By obtaining the identity mapping for \textsf{entry} in Figure \ref{fig:loadentry} -- ($\textsf{entry-KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry\_val}$ with page addr \textsf{entry}),
+ \[
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{entry\},m) \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast r8  \mapsto_{\textsf{r}} \textsf{entry} \textsf{\_} \ast rdi  \mapsto_{\textsf{r}} \textsf{entry} \ast \textsf{entry-KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry\_val} \ast \sumapaces\rtv\delta  \ast \sumpv\paddr\qfrac\true  }_{\rtv}$
+mov    [r8],rdi
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast r8  \mapsto_{\textsf{r}} \textsf{entry} \textsf{\_} \ast rdi  \mapsto_{\textsf{r}} \textsf{entry\_val} \ast  \textsf{entry-KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry\_val} \ast \sumapaces\rtv\delta  \ast \sumpv\paddr\qfrac\true  }_{\rtv}$
+ \]
+ we are able to load the \textsf{entry\_val} into \textsf{rdi}, and check the presence bit via $and 0x1,rdi$.
 
+ \subsubsection{Ensuring Presence of Entry}
+In case of absence, \textsf{pte\_initialize} is called for a physical-zeroed page (\textsf{kalloc} -- currently an axiom as it is not part of \textsf{pte} module but the allocator's) to be backing the physical frame address of the entry.
+
+\begin{figure}\footnotesize
+\begin{lstlisting}[mathescape]
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \ulcorner\urcorner \ast \texttt{rbx}\mapsto_{\textsf{r}} \paddr \ast \sumapaces\rtv\delta  \ast \sumpv\paddr\qfrac\true  }_{\rtv}$
+
+;;  if (!entry->present) {
+cmp    0x0,rax
+jne    161 <pte_get_next_table_succ+0xa1>
+lea    -0x10[rbp],rdi
+;;pte_initialize(&next);
+callq  70 <pte_initialize>
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \ulcorner\urcorner \ast \texttt{rbx}\mapsto_{\textsf{r}} \paddr \ast \sumapaces\rtv\delta  \ast \sumpv\paddr\qfrac\true  }_{\rtv}$
+...
+
+;;}
+\end{lstlisting}
+\vspace{-1em}
+\caption{Making sure that \textsf{entry}'s frame is referencing to an address backed up by a physical page.}
+\label{fig:calltopteinitialize}
+\end{figure}
+
+Once we make sure that the frame of the page-table entry is backed with an allocated physical page, we can obtain the virtual address of the next level-page table (\textsf{next} of function \textsf{get\_next\_table} in Figure \ref{}) via looking up the \textit{identity mapping} from the physical address (the address value referenced by \textsf{entry->pfn}) to virtual address value to be returned by \textsf{get\_next\_table}.
 \begin{figure}\footnotesize
@@
 
+\subsubsection{}
+
 %\caption{Traversing page-tables, and allocating entries as needed while mapping-a-page in Figure \ref{fig:mappingcode}.}

### 2024-06-30 examples.tex
+

--- 2024-06-30 examples.tex
@@
 %}
+
+\newcommand{\qfraczero}{\textsf{qfrac}}
+\newcommand{\true}{\textsf{true}}
+\tikzstyle{boxedassert_border} = [sharp corners,line width=0.2pt]
+\NewDocumentCommand \boxedassertpv {O{} m o}{%
+	\tikz[baseline=(m.base)]{
+		%	  \node[rectangle, draw,inner sep=0.8pt,anchor=base,#1] (m) {${#2}\mathstrut$};
+		\node[rectangle,inner sep=1.5pt,outer sep=0.2pt,anchor=base] (m) {${\,#2\,}\mathstrut$};
+		\draw[#1,boxedassert_border] ($(m.south west)$) rectangle ($(m.north east)$);
+	}\IfNoValueF{#3}{^{\,#3}}%
+}
+\newcommand*{\knowInvpv}[2]{\boxedassertpv{#2}[#1]}
+\newcommand*{\ownGhostpv}[2]{\boxedassertpv[dash dot]{#2}[#1]}
+
+\newcommand{\sumpv}[3]{
+  \ownGhostpv\gammaPred{\authfrag{\singletonMap{#1}{(#2, #3)}}}
+}
+
+\newcommand{\pvmapping}[1]{\mathcal{A}\textsf{PVMappings}(#1)}
+
 \newcommand{\fpaddr}{\texttt{fpaddr}}
 \newcommand{\specline}[1]{{\color{blue}\left\{#1\right\}}}
+\newcommand{\sumapacesfull}[2]{
+  \ownGhost\gammaPreds{\authfull{\singletonMap{#1}{#2}}}
+}
+
 \begin{figure}\footnotesize
@@
 
-As we explain in Section \ref{}, a traversal over a page-table tree can be done through the indices into the tables at each level followed by obtaining the address of the next table.
-In Figure \ref{fig:getnexttable}, we see how to obtain the physicall address of the next level page-table.
-Traversal start from the root address of the L4 table (\textsf{pte\_t *pml4} given as a parameter to \textsf{walkpgdir}).
-Then, we obtain the address of the entry in the L4 table via the index in the virtual address (\textsf{va}).
+\subsubsection{Loading Page-Table Address Value}
+In Figure \ref{fig:getnexttable}, we see how to obtain the physical address of the next level page-table.
+Traversal start from the root address of the page-table tree (the address of L4 table \textsf{pte\_t *pml4}) given as a parameter to \textsf{pte\_walkpgdir}).
+Then, we obtain the address of the entry (\textsf{pml4[PML4EX(va)]}) in the L4 table via the index in the virtual address (\textsf{va}).
 
-Finally, we call \textsf{get\_next\_table} to get the virtual address to the next level page-table -- L3.
-Within \textsf{get\_next\_table} first checks if the entry whose address is passed as a parameter -- e.g., for L4 \textsf{pml4[PML4EX(va)]}.
-In case of absence, \textsf{pte\_initialize} is called for a physical-zeroed page (\textsf{kalloc} -- currently an axiom as it is not part of \textsf{pte} module) to be backing the physical frame address of the entry.
+Finally, we call \textsf{get\_next\_table} to get the virtual address to the next level page-table -- at this point the address of L3 table.
+Within \textsf{get\_next\_table} first checks if the entry whose address is passed as a parameter ($\textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry}$).
+To be able to do so, it needs to load the entry value (a 64-bits word divided into bit-fields for corresponding fileds shown in Section {}) and check that \textit{presence} bit of the entry is set or not.
+
 \begin{figure}\footnotesize
   \begin{lstlisting}[mathescape]
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)  \ast \texttt{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{\_} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{\_} \ast \sumpv\paddr\qfrac\true  \sumapaces\rtv\delta  }_{\rtv}$
+mov    -0x8[rbp],rdi
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{\_} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry} \ast \sumpv\paddr\qfrac\true  \sumapaces\rtv\delta   }_{\rtv}$
+mov     rdi, r8
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry} \ast \textsf{entry-KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \sumpv\paddr\qfrac\true  \sumapaces\rtv\delta  }_{\rtv}$
+mov    [r8],rdi
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry} \ast rdi  \mapsto_{\textsf{r}} \textsf{entry\_val} \ast \textsf{entry-KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \sumpv\paddr\qfrac\true  \sumapaces\rtv\delta  }_{\rtv}$
+and    0x1,rdi
+mov    rdi,rax
+;;  if (!entry->present) {
+cmp    0x0,rax
+jne    161 <pte_get_next_table_succ+0xa1>
+lea    -0x10[rbp],rdi
+;;pte_initialize(&next);
+callq  70 <pte_initialize>
+...
+;;}
+\end{lstlisting}
+\vspace{-1em}
+\caption{Making sure that \textsf{entry}'s frame is referencing to an address backed up by a physical page.}
+\label{fig:calltopteinitialize}
+\end{figure}
+
+As we see in Figure \ref{fig:loadentry}, we have the physical address (\textsf{entry}) of the page table entry kept in the register $r8$, but not the memory mapping to load the entry value referenced by the physical address \textsf{entry}.
+\subsubsection{Identity Mappings}
+Most kernels maintain an invariant on their page tables that the virtual address of any page used for a page table lives at a virtual address
+whose value is a constant offset from the physical address a practice sometimes referred to as \emph{identity mapping} (even though the physical-to-virtual translation
+is typically not literally the identity function, but adding a constant offset).\footnote{Some kernels do this for all physical memory on the machine, simplifying interaction
+with DMA devices.}
+\begin{figure*}
+\footnotesize
 \[
+\begin{array}{l}
+  \mathcal{I}\textsf{ASpace}_{\textsf{id}}(\ptablestore,\Xi,m)\stackrel{\triangle}{=}\\ \textsf{ASpace\_Lookup}_{\textsf{id}}(\ptablestore,\Xi,m) \ast 
+  \bigast{(\vaddr, \textsf{paddr})\in \ptablestore}{\exists\;(\textsf{l4e, l3e, l2e, l1e, paddr})\ldotp \textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\vaddr\textsf{, l4e, l3e, l2e, l1e, paddr})} \\
+  \bigast{(\paddr) \in \Xi}{\exists\; (\textsf{qfrac, q, val,}\vaddr) \ldotp \ulcorner \vaddr = \paddr + \textsf{OFF} \urcorner \ast  \underbrace{\sumwalkabs\vaddr\qfrac\paddr }_\text{Ghost translation} \ast
+% \qquad\qquad\qquad\qquad\qquad\qquad\qquad\qquad    
+\underbrace{\paddr \mapsto_{\mathsf{p}}\{\textsf{qfrac}\}\; \vale}_\text{Physical location} \ast \ulcorner \textsf{qfrac} = 1 \leftrightarrow \; \lnot\textsf{entry\_present }(\vale) \urcorner }\\
+  \textsf{ where } \\
+   \textsf{ASpace\_Lookup}_{\textsf{id}}(\ptablestore,\Xi,m) \stackrel{\triangle}{=} \lambda\textsf{ cr3val} \ldotp \; \exists \gammaPred \; \ldotp \ulcorner m \; !!\; \textsf{cr3val} = \textsf{Some } \gammaPred \urcorner \ast
+   \ownGhost\gammaPred{\authfull{\ptableabswalk\ptablestore}} \ast  \ownGhostpv\gammaPred{\authfull{\pvmapping\Xi}}
+
+\end{array}
+\]
+\vspace{-1em}
+\caption{Global Address-Space Invariant in Figure \ref{fig:peraspaceinvariant} extended with a ghost map bookkeeping identity mappings }
+  \label{fig:peraspaceinvariant_with_p2v_extension}
+  \end{figure*}
+Now, to read the value referenced by physical address \textsf{entry} in Figure \ref{fig:loadentry}, first, we open our the addresss-space invariant ($\mathcal{I}\textsf{ASpace}_{\textsf{id}}(\ptablestore,\Xi,m)$), and obtain the virtual mapping out of the identity mapping $\Xi$ with the side condition
+
+\begin{equation*}
+ \ulcorner \textsf{qfrac} = 1 \leftrightarrow \; \lnot\textsf{entry\_present }(\vale) \urcorner \tag{*}
+\end{equation*}
+
+ assuring that looking up identity mappings are safe, because the full ownership asserts the non-presence of the entry which requires allocation a new physical page and setting the presence bit.
+By obtaining the identity mapping for \textsf{entry} in Figure \ref{fig:loadentry} -- ($\textsf{entry-KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry\_val}$ with page addr \textsf{entry}),
+ \[
+ \begin{array}{l}
+\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{entry\},m) \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast r8  \mapsto_{\textsf{r}} \textsf{entry} \textsf{\_} \ast rdi  \mapsto_{\textsf{r}} \textsf{entry} \ast \textsf{entry-KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry\_val} \ast \sumapaces\rtv\delta  \ast \sumpv\paddr\qfrac\true  }_{\rtv}
+mov    [r8],rdi
+\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast r8  \mapsto_{\textsf{r}} \textsf{entry} \textsf{\_} \ast rdi  \mapsto_{\textsf{r}} \textsf{entry\_val} \ast  \textsf{entry-KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry\_val} \ast \sumapaces\rtv\delta  \ast \sumpv\paddr\qfrac\true  }_{\rtv}
+\end{array}
+ \]
+ we are able to load the \textsf{entry\_val} into \textsf{rdi}, and check the presence bit via $and 0x1,rdi$.
+
+ \subsubsection{Ensuring Presence of Entry}
+ Accessing the presence bit and checking the value allow us to exploit the fact in the side condition ()such that we can own twe introduced for the  
+In case of absence, \textsf{pte\_initialize} is called for a physical-zeroed page (\textsf{kalloc} -- currently an axiom as it is not part of \textsf{pte} module but the allocator's) to be backing the physical frame address of the entry.
+
+\begin{figure}\footnotesize
+\begin{lstlisting}[mathescape]
 $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \ulcorner\urcorner \ast \texttt{rbx}\mapsto_{\textsf{r}} \paddr \ast \sumapaces\rtv\delta  \ast \sumpv\paddr\qfrac\true  }_{\rtv}$
@@
 Once we make sure that the frame of the page-table entry is backed with an allocated physical page, we can obtain the virtual address of the next level-page table (\textsf{next} of function \textsf{get\_next\_table} in Figure \ref{}) via looking up the \textit{identity mapping} from the physical address (the address value referenced by \textsf{entry->pfn}) to virtual address value to be returned by \textsf{get\_next\_table}.
-\subsubsection{Identity Mappings}
-Most kernels maintain an invariant on their page tables that the virtual address of any page used for a page table lives at a virtual address
-whose value is a constant offset from the physical address a practice sometimes referred to as \emph{identity mapping} (even though the physical-to-virtual translation
-is typically not literally the identity function, but adding a constant offset).\footnote{Some kernels do this for all physical memory on the machine, simplifying interaction
-with DMA devices.}
-\newcommand{\qfraczero}{\textsf{qfrac}}
-\newcommand{\true}{\textsf{true}}
-\tikzstyle{boxedassert_border} = [sharp corners,line width=0.2pt]
-\NewDocumentCommand \boxedassertpv {O{} m o}{%
-	\tikz[baseline=(m.base)]{
-		%	  \node[rectangle, draw,inner sep=0.8pt,anchor=base,#1] (m) {${#2}\mathstrut$};
-		\node[rectangle,inner sep=1.5pt,outer sep=0.2pt,anchor=base] (m) {${\,#2\,}\mathstrut$};
-		\draw[#1,boxedassert_border] ($(m.south west)$) rectangle ($(m.north east)$);
-	}\IfNoValueF{#3}{^{\,#3}}%
-}
-\newcommand*{\knowInvpv}[2]{\boxedassertpv{#2}[#1]}
-\newcommand*{\ownGhostpv}[2]{\boxedassertpv[dash dot]{#2}[#1]}
-
-\newcommand{\sumpv}[3]{
-  \ownGhostpv\gammaPred{\authfrag{\singletonMap{#1}{(#2, #3)}}}
-}
-
-\newcommand{\pvmapping}[1]{\mathcal{A}\textsf{PVMappings}(#1)}
-\newcommand{\sumapacesfull}[2]{
-  \ownGhost\gammaPreds{\authfull{\singletonMap{#1}{#2}}}
-}
-\begin{figure*}
-\footnotesize
-\[
-\begin{array}{l}
-  \mathcal{I}\textsf{ASpace}_{\textsf{id}}(\ptablestore,\Xi,m)\stackrel{\triangle}{=}\\ \textsf{ASpace\_Lookup}_{\textsf{id}}(\ptablestore,\Xi,m) \ast 
-  \bigast{(\vaddr, \textsf{paddr})\in \ptablestore}{\exists\;(\textsf{l4e, l3e, l2e, l1e, paddr})\ldotp \textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\vaddr\textsf{, l4e, l3e, l2e, l1e, paddr})} \\
-  \bigast{(\paddr) \in \Xi}{\exists\; (\textsf{qfrac, q, val,}\vaddr) \ldotp \ulcorner \vaddr = \paddr + \textsf{OFF} \urcorner \ast  \underbrace{\sumwalkabs\vaddr\qfrac\paddr }_\text{Ghost translation} \ast
-% \qquad\qquad\qquad\qquad\qquad\qquad\qquad\qquad    
-\underbrace{\paddr \mapsto_{\mathsf{p}}\{\textsf{qfrac}\}\; \vale}_\text{Physical location} \ast \ulcorner \textsf{qfrac} = 1 \leftrightarrow \; \lnot\textsf{entry\_present }(\vale) \urcorner }\\
-  \textsf{ where } \\
-   \textsf{ASpace\_Lookup}_{\textsf{id}}(\ptablestore,\Xi,m) \stackrel{\triangle}{=} \lambda\textsf{ cr3val} \ldotp \; \exists \gammaPred \; \ldotp \ulcorner m \; !!\; \textsf{cr3val} = \textsf{Some } \gammaPred \urcorner \ast
-   \ownGhost\gammaPred{\authfull{\ptableabswalk\ptablestore}} \ast  \ownGhostpv\gammaPred{\authfull{\pvmapping\Xi}}
-
-\end{array}
-\]
-\vspace{-1em}
-\caption{Global Address-Space Invariant in Figure \ref{fig:peraspaceinvariant} extended with a ghost map bookkeeping identity mappings }
-  \label{fig:peraspaceinvariant_with_p2v_extension}
-  \end{figure*}
-
 \begin{figure}\footnotesize

### 2024-06-30 examples.tex
+

@@
 mov     rdi, r8
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry} \ast \textsf{entry-KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \sumpv\paddr\qfrac\true  \sumapaces\rtv\delta  }_{\rtv}$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry}  \ast \sumpv\paddr\qfrac\true  \sumapaces\rtv\delta  }_{\rtv}$
+$\specline{ \textsf{entry-KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner \textsf{qfrac}=1 \leftrightarrow \textsf{not\_present entry} \urcorner}$
 mov    [r8],rdi
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry} \ast rdi  \mapsto_{\textsf{r}} \textsf{entry\_val} \ast \textsf{entry-KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \sumpv\paddr\qfrac\true  \sumapaces\rtv\delta  }_{\rtv}$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry} \ast rdi  \mapsto_{\textsf{r}} \textsf{entry\_val} \ast \sumpv\paddr\qfrac\true  \sumapaces\rtv\delta  }_{\rtv}$
+$\specline{ \textsf{entry-KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner \textsf{qfrac}=1 \leftrightarrow \textsf{not\_present entry} \urcorner }$
 and    0x1,rdi
@@
 ...
-;;}
 \end{lstlisting}

### 2024-06-30 examples.tex
+

--- 2024-06-30 examples.tex
@@
 mov     rdi, r8
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry} \ast \textsf{entry-KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \sumpv\paddr\qfrac\true  \sumapaces\rtv\delta  }_{\rtv}$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry}  \ast \sumpv\paddr\qfrac\true  \sumapaces\rtv\delta  }_{\rtv}$
+$\specline{ \textsf{entry-KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot\textsf{entry\_present }\urcorner}$
 mov    [r8],rdi
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry} \ast rdi  \mapsto_{\textsf{r}} \textsf{entry\_val} \ast \textsf{entry-KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \sumpv\paddr\qfrac\true  \sumapaces\rtv\delta  }_{\rtv}$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry} \ast rdi  \mapsto_{\textsf{r}} \textsf{entry\_val} \ast \sumpv\paddr\qfrac\true  \sumapaces\rtv\delta  }_{\rtv}$
+$\specline{ \textsf{entry-KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot\textsf{entry\_present }  \urcorner }$
 and    0x1,rdi
@@
 
-As we see in Figure \ref{fig:loadentry}, we have the physical address (\textsf{entry}) of the page table entry kept in the register $r8$, but not the memory mapping to load the entry value referenced by the physical address \textsf{entry}.
+As we see in Figure \ref{fig:calltopteinitialize}, we have the physical address (\textsf{entry}) of the page table entry kept in the register $\texttt{r8}$, but not the memory mapping to load the entry value referenced by the physical address \textsf{entry}.
+We need to construct this virtual mappings for \textsf{entry}.
 \subsubsection{Identity Mappings}
@@
   \end{figure*}
-Now, to read the value referenced by physical address \textsf{entry} in Figure \ref{fig:loadentry}, first, we open our the addresss-space invariant ($\mathcal{I}\textsf{ASpace}_{\textsf{id}}(\ptablestore,\Xi,m)$), and obtain the virtual mapping out of the identity mapping $\Xi$ with the side condition
-
+Now, to read the value referenced by physical address \textsf{entry} in Figure \ref{fig:calltopteinitialize}, first, we open our the addresss-space invariant ($\mathcal{I}\textsf{ASpace}_{\textsf{id}}(\ptablestore,\Xi,m)$), and obtain the virtual mapping out of the identity mapping $\Xi$ with the side condition
 \begin{equation*}
@@
 \end{equation*}
-
  assuring that looking up identity mappings are safe, because the full ownership asserts the non-presence of the entry which requires allocation a new physical page and setting the presence bit.
-By obtaining the identity mapping for \textsf{entry} in Figure \ref{fig:loadentry} -- ($\textsf{entry-KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry\_val}$ with page addr \textsf{entry}),
+By obtaining the identity mapping for \textsf{entry} in Figure \ref{fig:calltopteinitialize} -- ($\textsf{entry-KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val}$ with page addr \textsf{entry}),
  \[

### 2024-07-01 examples.tex
+

@@
 \subsubsection{Loading Page-Table Address Value}
-In Figure \ref{fig:getnexttable}, we see how to obtain the physical address of the next level page-table.
-Traversal start from the root address of the page-table tree (the address of L4 table \textsf{pte\_t *pml4}) given as a parameter to \textsf{pte\_walkpgdir}).
-Then, we obtain the address of the entry (\textsf{pml4[PML4EX(va)]}) in the L4 table via the index in the virtual address (\textsf{va}).
-
-Finally, we call \textsf{get\_next\_table} to get the virtual address to the next level page-table -- at this point the address of L3 table.
+We call \textsf{get\_next\_table} to get the virtual address to the next level page-table.
+In Figure \ref{fig:calltopteinitialize}, we present the critical piece of \textsf{get\_next\_table}, and construct the specification of it
 Within \textsf{get\_next\_table} first checks if the entry whose address is passed as a parameter ($\textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry}$).
@@
 ...
+;;}
+;; uintptr_t next_phys_addr = PTE_PFN_TO_ADDR(entry->pfn);
+...
+;;uintptr_t next_virt_addr = (uintptr_t) P2V(next_phys_addr);
+mov    -0x18[rbp],rax
+movabs 0x0,rcx	                187: R_X86_64_64	KERN_BASE
+add    rcx,rax
+...
 \end{lstlisting}
@@
 
-As we see in Figure \ref{fig:calltopteinitialize}, we have the physical address (\textsf{entry}) of the page table entry kept in the register $\texttt{r8}$, but not the memory mapping to load the entry value referenced by the physical address \textsf{entry}.
+As we see in Figure \ref{fig:calltopteinitialize}, we have the physical address (\textsf{entry}) of the page table entry kept in the register $\texttt{r8}$, but not the memory mapping to load the entry value referenced by the physical address \textsf{entry} (Line 7 in Figure \ref{fig:calltopteinitialize}).
 We need to construct this virtual mappings for \textsf{entry}.
@@
   \end{figure*}
-Now, to read the value referenced by physical address \textsf{entry} in Figure \ref{fig:calltopteinitialize}, first, we open our the addresss-space invariant ($\mathcal{I}\textsf{ASpace}_{\textsf{id}}(\ptablestore,\Xi,m)$), and obtain the virtual mapping out of the identity mapping $\Xi$ with the side condition
+Before Line 7 in Figure \ref{fig:calltopteinitialize}, to read the value referenced by physical address \textsf{entry}, first we open our the addresss-space invariant ($\mathcal{I}\textsf{ASpace}_{\textsf{id}}(\ptablestore,\Xi,m)$), and then obtain the virtual mapping out of the identity mapping $\Xi$ with the side condition
 \begin{equation*}
@@
  \subsubsection{Ensuring Presence of Entry}
- Accessing the presence bit and checking the value allow us to exploit the fact in the side condition ()such that we can own twe introduced for the  
+ Accessing the presence bit and checking the value allow (Lines 7-13 in Figure \ref{fig:calltopteinitialize}) us to exploit the fact in the side condition (*) such that we can own twe introduced for the
 In case of absence, \textsf{pte\_initialize} is called for a physical-zeroed page (\textsf{kalloc} -- currently an axiom as it is not part of \textsf{pte} module but the allocator's) to be backing the physical frame address of the entry.
-
-\begin{figure}\footnotesize
-\begin{lstlisting}[mathescape]
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \ulcorner\urcorner \ast \texttt{rbx}\mapsto_{\textsf{r}} \paddr \ast \sumapaces\rtv\delta  \ast \sumpv\paddr\qfrac\true  }_{\rtv}$
-
-;;  if (!entry->present) {
-cmp    0x0,rax
-jne    161 <pte_get_next_table_succ+0xa1>
-lea    -0x10[rbp],rdi
-;;pte_initialize(&next);
-callq  70 <pte_initialize>
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \ulcorner\urcorner \ast \texttt{rbx}\mapsto_{\textsf{r}} \paddr \ast \sumapaces\rtv\delta  \ast \sumpv\paddr\qfrac\true  }_{\rtv}$
-...
-
-;;}
-\end{lstlisting}
-\vspace{-1em}
-\caption{Making sure that \textsf{entry}'s frame is referencing to an address backed up by a physical page.}
-\label{fig:calltopteinitialize}
-\end{figure}
 

### 2024-07-01 examples.tex
+

@@
  \subsubsection{Ensuring Presence of Entry}
- Accessing the presence bit and checking the value allow (Lines 7-13 in Figure \ref{fig:calltopteinitialize}) us to exploit the fact in the side condition (*) such that we can own twe introduced for the
+ Accessing the presence bit and checking the value allow (Lines 7-13 in Figure \ref{fig:calltopteinitialize}) us to exploit the fact in the side condition (*) in such a way that the condition ($\textsf{entry}\mapsto\textsf{present}$ determines whether we take the non-presence path (Lines 12-19 in Figure \ref{calltopteinitialize}) with full ownership ($\textsf{qfrac} = 1$) to set the entry with proper values for the bitfields, e.g., frame showing freshly allocated page and presence bit.
+
 In case of absence, \textsf{pte\_initialize} is called for a physical-zeroed page (\textsf{kalloc} -- currently an axiom as it is not part of \textsf{pte} module but the allocator's) to be backing the physical frame address of the entry.
-
+\mytodo{Discuss with Colin: v2p and spec on pteinitialize.
+After clarification expand this piece a little bit more.
+Right now we have physical pointsto returned from kalloc so pfn is directly set not through v2p}
 Once we make sure that the frame of the page-table entry is backed with an allocated physical page, we can obtain the virtual address of the next level-page table (\textsf{next} of function \textsf{get\_next\_table} in Figure \ref{}) via looking up the \textit{identity mapping} from the physical address (the address value referenced by \textsf{entry->pfn}) to virtual address value to be returned by \textsf{get\_next\_table}.
@@
 \begin{lstlisting}[mathescape]
-\[
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \ulcorner\urcorner \ast \texttt{rbx}\mapsto_{\textsf{r}} \paddr \ast \sumapaces\rtv\delta  \ast \sumpv\paddr\qfrac\true  }_{\rtv}$
-
 ;;if (!entry->present) {
@@
 ;;}
-;; uintptr_t next_phys_addr = PTE_PFN_TO_ADDR(entry->pfn);
+...
 ;;uintptr_t next_virt_addr = (uintptr_t) P2V(next_phys_addr);
@@
 This is a simple piece of code whose functionality is critical and whose correctness is highly non-trivial. No prior work engages with this problem.
-
-\begin{figure*}[ht]
-\begin{tabular}{p{0.47\textwidth}p{0.47\textwidth}}
-  \begin{lstlisting}[mathescape,basicstyle=\scriptsize\ttfamily]
-;;void pte_initialize(pte_t **table) {
-...
-
-;;Allocate a full page for 512 8-byte entries
-;;pte_t *local = kalloc();
-;;clear the entire table. Ensures present (and reserved, etc.) bits are 0
-callq  81 <pte_initialize+0x11>	7d: R_X86_64_PLT32	kalloc-0x4
-mov    rax,-0x10[rbp]
-
-;;*table = local;
-mov    -0x10[rbp],rax
-mov    -0x8[rbp],rdi
-mov    rax,[rdi]
-
-;;return;
-...
-
-;;pte_t *walkpgdir_succ(pte_t *pml4, const void *va) 
-push   rbp
-mov    rsp,rbp
-sub    0x40,rsp
-mov    rdi,-0x8[rbp]
-mov    rsi,-0x10[rbp]
-;;pte_t *pml4_entry = &pml4[PML4EX(va)];
-mov    -0x8[rbp],[rsi]
-mov    -0x10[rbp],rdi
-shr    0x27,rdi
-and    0x1ff,rdi
-shl    0x3,rdi
-add    rdi,rsi
-mov    rsi,-0x18[rbp]
-;;pte_t *pdp = pte_get_next_table_succ(pml4_entry);
-mov    -0x18[rbp],rdi
-callq  <pte_get_next_table_succ>
-mov    rax,-0x20[rbp]
-;;pte_t *pdp_entry = &pdp[PDPEX(va)];
-...
-...
-...
-;;last call at L1 for virtual address to physical page 
-callq  <pte_get_next_table_succ>
-mov    rax,-0x40[rbp]
-;;return &pt[PTEX(va)];
-mov    -0x40[rbp],rax
-mov    -0x10[rbp],rsi
-shr    0xc,rsi
-and    0x1ff,rsi
-shl    0x3,rsi
-add    rsi,rax
-add    0x40,rsp
-pop    rbp
-retq   
-  \end{lstlisting}&
-  \begin{lstlisting}[mathescape,basicstyle=\scriptsize\ttfamily]
-   $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \ulcorner \textsf{OFF} + \paddr > 0 \urcorner \ast \texttt{rbx}\mapsto_{\textsf{r}} \paddr \ast \sumapaces\rtv\delta  \ast \sumpv\paddr\qfrac\true  }_{\rtv}$
-;; pte_t *pte_get_next_table_succ(pte_t *entry) {
-push   rbp
-mov    rsp,rbp
-sub    0x20,rsp
-mov    rdi,-0x8[rbp]
-;;  pte_t *next;
-mov    -0x8[rbp],rdi
-mov    [rdi],rdi
-and    0x1,rdi
-mov    rdi,rax
-;;  if (!entry->present) {
-cmp    0x0,rax
-jne    161 <pte_get_next_table_succ+0xa1>
-lea    -0x10[rbp],rdi
-;;pte_initialize(&next);
-callq  70 <pte_initialize>
-mov    -0x8[rbp],rdi
-mov    [rdi],rax
-;;setting the fields of the entry with a fresh allocated page
-;;entry->writable = 1
-;;entry->user_acc = 1;
-;;entry->pfn = PTE_ADDR_TO_PHYS_PFN((uintptr_t) next);
-;;entry->present = 1;
-and    0xfffffffffffffffd,rax
-or     0x2,rax
-mov    rax,[rdi]
-...
-;;}
-;;uintptr_t next_virt_addr = (uintptr_t) P2V(next_phys_addr);
-mov    -0x8[rbp],rax
-mov    [rax],rax
-shr    0xc,rax
-movabs 0xffffffffff,rcx
-and    rcx,rax
-shl    0xc,rax
-mov    rax,-0x18[rbp]
-;;uintptr_t next_virt_addr = (uintptr_t) P2V(next_phys_addr);
-mov    -0x18[rbp],rax
-movabs 0x0,rcx	187: R_X86_64_64	KERN_BASE
-add    rcx,rax
-;;next = (pte_t *) next_virt_addr;
-mov    rax,-0x20[rbp]
-mov    -0x20[rbp],rax
-mov    rax,-0x10[rbp]
-;;return next
-mov    -0x10[rbp],rax
-add    0x20,rsp
-pop    rbp
-ret
-  \end{lstlisting}
-  \end{tabular}
-\vspace{-1em}
-\caption{ in Figure \ref{fig:walkpgdir}.}
-\label{fig:getnexttable}
-\end{figure*}
-
-\subsubsection{}
 

### 2024-07-01 examples.tex
+

@@
 \end{figure*}
-Before Line 7 in Figure \ref{fig:calltopteinitialize}, to read the value referenced by physical address \textsf{entry}, first we open our the addresss-space invariant ($\mathcal{I}\textsf{ASpace}_{\textsf{id}}(\ptablestore,\Xi,m)$), and then obtain the virtual mapping out of the identity mapping $\Xi$ with the side condition
+
+As part of the full page table walk verification, we have verified the critical step for a small x86-64 kernel, which is the physical-to-virtual
+conversion (often appearing as a macro \texttt{P2V} inside the macro \textsf{ PTE\_PFN\_TO\_ADDR} in C source code). In our small kernel, as in larger kernels, \texttt{P2V} is actually just addition,
+but the correctness of this simple instruction is quite subtle, as shown in Figure \ref{fig:p2v} -- showing the section of the \textsf{get\_next\_table} in Figure \ref{calltopteinitialize} after the entry presence ensured (Lines 1-19 in Figure \ref{fig:calltopteinitialize}).
+Beyond a frame, the precondition before Line 7 requires the current address space invariant, a check that the addition will not overflow, the physical address to convert
+(in \lstinline|rax|), knowledge that the current address space is valid, and knowledge that the physical address is in a range dedicated
+to page table entries (as this physical-to-virtual conversion is only guaranteed for page table memory). 
+The latter uses an additional ghost set added to the per-address-space
+invariant per Figure \ref{fig:peraspaceinvariant_with_p2v_extension}. The extended invariant owns, for each memory location dedicated to page tables, some fraction of
+that physical memory, and also guarantees the offset is the correct calculation.
+Subtly, the operation ensures the relevant table entry is readable, but the exact portion of ownership returned must be determined by inspecting the valid bit
+of the value in memory --- so full ownership is returned only for unused entries.
+This is a simple piece of code whose functionality is critical and whose correctness is highly non-trivial. No prior work engages with this problem.
+
+Consequently, before Line 7 in Figure \ref{fig:calltopteinitialize}, to read the value referenced by physical address \textsf{entry}, first we open our the addresss-space invariant ($\mathcal{I}\textsf{ASpace}_{\textsf{id}}(\ptablestore,\Xi,m)$), and then obtain the virtual mapping out of the identity mapping $\Xi$ with the side condition
 \begin{equation*}
@@
 Right now we have physical pointsto returned from kalloc so pfn is directly set not through v2p}
-Once we make sure that the frame of the page-table entry is backed with an allocated physical page, we can obtain the virtual address of the next level-page table (\textsf{next} of function \textsf{get\_next\_table} in Figure \ref{}) via looking up the \textit{identity mapping} from the physical address (the address value referenced by \textsf{entry->pfn}) to virtual address value to be returned by \textsf{get\_next\_table}.
+Once we make sure that the frame of the page-table entry is backed with an allocated physical page, we can obtain the virtual address of the next level-page table (\textsf{next} of function \textsf{get\_next\_table} in Figure \ref{}) via looking up the \textit{identity mapping} from the physical address (the address value referenced by $\textsf{entry}\mapsto\textsf{pfn}$) to virtual address value to be returned by \textsf{get\_next\_table}.
 \begin{figure}\footnotesize
@@
 ...
+mov    rax,-0x18[rbp]
 ;;uintptr_t next_virt_addr = (uintptr_t) P2V(next_phys_addr);
-  $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \ulcorner \textsf{OFF} + \paddr > 0 \urcorner \ast \texttt{rbx}\mapsto_{\textsf{r}} \paddr \ast \sumapaces\rtv\delta  \ast \sumpv\paddr\qfrac\true  }_{\rtv}$
-add rbx, KERNBASE ;; shift physical address by identity mapping constant OFF
-$\specline{\textsf{P} \ast \exists \qfraczero; \qfrac,\textsf{ w, val},  \texttt{rbx}\mapsto_{\textsf{r}} \paddr + \textsf{OFF} \ast \sumapaces\rtv\delta \ast \sumpv\paddr\qfrac\true \ast \vaddr \mapsto_{\texttt{vpte}} \{\textsf{q}, \textsf{qfrac}\} \paddr \;(\texttt{val})}$
-$\specline{\ulcorner \vaddr = \paddr + \textsf{OFF} \urcorner \ast \ulcorner \textsf{qfrac} = 1 \leftrightarrow \; \lnot\textsf{entry\_present }(\vale) \urcorner \ast \sumwalkabs\vaddr\qfrac\paddr \ast (R \wand \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)) }_{\rtv}$
-\end{lstlisting}
+mov    -0x18[rbp],rax
+movabs $0x0,rcx            187: R_X86_64_64	KERN_BASE
+add    rcx,rax
+mov    rax,-0x20[rbp]
 \vspace{-1em}
@@
 \end{figure}
-
-As part of the full page table walk verification, we have verified the critical step for a small x86-64 kernel, which is the physical-to-virtual
-conversion (often appearing as a macro \texttt{P2V} inside the macro \textsf{ PTE\_PFN\_TO\_ADDR} in C source code). In our small kernel, as in larger kernels, \texttt{P2V} is actually just addition,
-but the correctness of this simple instruction is quite subtle, as shown in Figure \ref{fig:p2v}.
-Beyond a frame ($P$), the precondition requires the current address space invariant, a check that the addition will not overflow, the physical address to convert
-(in \lstinline|rbx|), knowledge that the current address space is valid, and knowledge that the physical address is in a range dedicated
-to page table entries (as this physical-to-virtual conversion is only guaranteed for page table memory). 
-The latter uses an additional ghost set added to the per-address-space
-invariant per Figure \ref{fig:peraspaceinvariant_with_p2v_extension}. The extended invariant owns, for each memory location dedicated to page tables, some fraction of
-that physical memory, and also guarantees the offset is the correct calculation.
-Subtly, the operation ensures the relevant table entry is readable, but the exact portion of ownership returned must be determined by inspecting the valid bit
-of the value in memory --- so full ownership is returned only for unused entries.
-This is a simple piece of code whose functionality is critical and whose correctness is highly non-trivial. No prior work engages with this problem.
 

### 2024-07-01 examples.tex
+

@@
   \begin{lstlisting}[mathescape]
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)  \ast \texttt{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{\_} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{\_} \ast \sumpv\paddr\qfrac\true  \sumapaces\rtv\delta  }_{\rtv}$
+;;pte_t *pte_get_next_table_succ(pte_t *entry) {
+push   rbp
+mov    rsp,rbp
+sub    0x20,rsp
+mov    rdi,-0x8[rbp]
+;; pte_t *next;
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)  \ast \texttt{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast  \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{next} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{\_} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{\_}  }_{\rtv}$
+$\specline{\sumpv\paddr\qfrac\true \ast \sumapaces\rtv\delta  }_{\rtv}$
 mov    -0x8[rbp],rdi
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{\_} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry} \ast \sumpv\paddr\qfrac\true  \sumapaces\rtv\delta   }_{\rtv}$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{\_}  \ast  \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{next} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry}  }_{\rtv}$
+$\specline{\sumpv\paddr\qfrac\true \ast \sumapaces\rtv\delta  }_{\rtv}$
 mov     rdi, r8
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry}  \ast \sumpv\paddr\qfrac\true  \sumapaces\rtv\delta  }_{\rtv}$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast  \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{next} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry} }_{\rtv}$
+$\specline{\sumpv\paddr\qfrac\true \ast \sumapaces\rtv\delta  }_{\rtv}$
 $\specline{ \textsf{entry-KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot\textsf{entry\_present }\urcorner}$
 mov    [r8],rdi
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry} \ast rdi  \mapsto_{\textsf{r}} \textsf{entry\_val} \ast \sumpv\paddr\qfrac\true  \sumapaces\rtv\delta  }_{\rtv}$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast  \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{next} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val}   }_{\rtv}$
+$\specline{\sumpv\paddr\qfrac\true \ast \sumapaces\rtv\delta  }_{\rtv}$
 $\specline{ \textsf{entry-KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot\textsf{entry\_present }  \urcorner }$
@@
 jne    161 <pte_get_next_table_succ+0xa1>
-lea    -0x10[rbp],rdi
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry} \ast \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{next} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val} }_{\rtv}$
+$\specline{\sumpv\paddr\qfrac\true \ast \sumapaces\rtv\delta  }_{\rtv}$
+$\specline{ \textsf{entry-KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot\textsf{entry\_present }  \urcorner }$
+mov    rbp,rdi
+sub    0x10, rdi
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry} \ast \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{next} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val}  }_{\rtv}$
+$\specline{\sumpv\paddr\qfrac\true \ast \sumapaces\rtv\delta  }_{\rtv}$
+$\specline{ \textsf{entry-KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \land \lnot\textsf{entry\_present }  \urcorner }$
 ;;pte_initialize(&next);
 callq  70 <pte_initialize>
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val}  }_{\rtv}$
+$\specline{\sumpv\paddr\qfrac\true \ast \sumapaces\rtv\delta  }_{\rtv}$
+$\specline{ \textsf{entry-KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \land \lnot\textsf{entry\_present }  \urcorner }$
+$\specline{ \texttt{rbp-16} \mapsto_{\textsf{vpte,1}} \textsf{nextpaddr wzero}  \ast \texttt{rax} \mapsto_{\textsf{r}} \textsf{nextpaddr}  }$
+;;entry->present = 1;
 ...
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val}  }_{\rtv}$
+$\specline{\sumpv\paddr\qfrac\true \ast \sumapaces\rtv\delta  }_{\rtv}$
+$\specline{ \textsf{entry-KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \land \lnot\textsf{entry\_present }  \urcorner }$
+$\specline{ \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{nextpaddr}  \ast \texttt{rax} \mapsto_{\textsf{r}} \textsf{nextpaddr}  }$
 ;;}
-;; uintptr_t next_phys_addr = PTE_PFN_TO_ADDR(entry->pfn);
 ...
@@
  \]
- we are able to load the \textsf{entry\_val} into \textsf{rdi}, and check the presence bit via $and 0x1,rdi$.
+ we are able to load the \textsf{entry\_val} into \textsf{rdi}, and check the presence bit via $and \;0x1,rdi$.
 
- \subsubsection{Ensuring Presence of Entry}
+ \subsubsection{Ensuring Presence of an Entry}
  Accessing the presence bit and checking the value allow (Lines 7-13 in Figure \ref{fig:calltopteinitialize}) us to exploit the fact in the side condition (*) in such a way that the condition ($\textsf{entry}\mapsto\textsf{present}$ determines whether we take the non-presence path (Lines 12-19 in Figure \ref{calltopteinitialize}) with full ownership ($\textsf{qfrac} = 1$) to set the entry with proper values for the bitfields, e.g., frame showing freshly allocated page and presence bit.
 
-In case of absence, \textsf{pte\_initialize} is called for a physical-zeroed page (\textsf{kalloc} -- currently an axiom as it is not part of \textsf{pte} module but the allocator's) to be backing the physical frame address of the entry.
+In case of absence, \textsf{pte\_initialize} is called for a physical-zeroed page (utilizing the page-allocator's \textsf{kalloc} -- currently the only axiom as it is not part of \textsf{pte} (virtual-memory-management) module) to be backing the physical frame address of the entry.
 \mytodo{Discuss with Colin: v2p and spec on pteinitialize.
@@
 Right now we have physical pointsto returned from kalloc so pfn is directly set not through v2p}
-Once we make sure that the frame of the page-table entry is backed with an allocated physical page, we can obtain the virtual address of the next level-page table (\textsf{next} of function \textsf{get\_next\_table} in Figure \ref{}) via looking up the \textit{identity mapping} from the physical address (the address value referenced by $\textsf{entry}\mapsto\textsf{pfn}$) to virtual address value to be returned by \textsf{get\_next\_table}.
+Once we make sure that the frame of the page-table entry ($\textsf{entry}\mapsto\textsf{pfn}$)is backed with an allocated physical page, we can obtain the virtual address of the next level-page table.
+To do so, again, we refer to the  (\textsf{next} of function \textsf{get\_next\_table} in Figure \ref{}) via looking up the \textit{identity mapping} from the physical address (the address value referenced by $\textsf{entry}\mapsto\textsf{pfn}$) to virtual address value to be returned by \textsf{get\_next\_table}.
 \begin{figure}\footnotesize
@@
 ...
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)  \ast \texttt{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{\_} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{\_} \ast \sumpv\paddr\qfrac\true  \sumapaces\rtv\delta  }_{\rtv}$
 mov    rax,-0x18[rbp]
+
 ;;uintptr_t next_virt_addr = (uintptr_t) P2V(next_phys_addr);
 mov    -0x18[rbp],rax
-movabs $0x0,rcx            187: R_X86_64_64	KERN_BASE
+movabs 0x0,rcx            187: R_X86_64_64	KERN_BASE
 add    rcx,rax
 mov    rax,-0x20[rbp]
+\end{lstlisting}
 \vspace{-1em}

### 2024-07-01 examples.tex
+

@@
 ;;pte_t *pte_get_next_table_succ(pte_t *entry) {
-push   rbp
-mov    rsp,rbp
-sub    0x20,rsp
-mov    rdi,-0x8[rbp]
+;; setting up the stack
 ;; pte_t *next;
@@
 $\specline{\sumpv\paddr\qfrac\true \ast \sumapaces\rtv\delta  }_{\rtv}$
-$\specline{ \textsf{entry-KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot\textsf{entry\_present }\urcorner}$
+$\specline{ \textsf{entry-KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present entry\_val})\urcorner}$
 mov    [r8],rdi
@@
 $\specline{\sumpv\paddr\qfrac\true \ast \sumapaces\rtv\delta  }_{\rtv}$
-$\specline{ \textsf{entry-KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot\textsf{entry\_present }  \urcorner }$
+$\specline{ \textsf{entry-KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present entry\_val})  \urcorner }$
 and    0x1,rdi
@@
 $\specline{\sumpv\paddr\qfrac\true \ast \sumapaces\rtv\delta  }_{\rtv}$
-$\specline{ \textsf{entry-KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot\textsf{entry\_present }  \urcorner }$
+$\specline{ \textsf{entry-KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present entry\_val})  \urcorner }$
 mov    rbp,rdi
@@
 $\specline{\sumpv\paddr\qfrac\true \ast \sumapaces\rtv\delta  }_{\rtv}$
-$\specline{ \textsf{entry-KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \land \lnot\textsf{entry\_present }  \urcorner }$
+$\specline{ \textsf{entry-KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \land \lnot(\textsf{entry\_present entry\_val})  \urcorner }$
 ;;pte_initialize(&next);
@@
 $\specline{\sumpv\paddr\qfrac\true \ast \sumapaces\rtv\delta  }_{\rtv}$
-$\specline{ \textsf{entry-KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \land \lnot\textsf{entry\_present }  \urcorner }$
-$\specline{ \texttt{rbp-16} \mapsto_{\textsf{vpte,1}} \textsf{nextpaddr wzero}  \ast \texttt{rax} \mapsto_{\textsf{r}} \textsf{nextpaddr}  }$
+$\specline{ \textsf{entry-KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \land \lnot(\textsf{entry\_present entry\_val})  \urcorner }$
+$\specline{ \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{nextpaddr} \ast \textsf{nextpaddr} \mapsto_{\textsf{p}} \textsf{w64\_zero}  \ast \texttt{rax} \mapsto_{\textsf{r}} \textsf{nextpaddr}  }$
 ;;entry->present = 1;
-...
+;;entry->pfn = next
 $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val}  }_{\rtv}$
 $\specline{\sumpv\paddr\qfrac\true \ast \sumapaces\rtv\delta  }_{\rtv}$
-$\specline{ \textsf{entry-KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \land \lnot\textsf{entry\_present }  \urcorner }$
+$\specline{ \textsf{entry-KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \land (\textsf{entry\_present entry\_val}) \urcorner }$
 $\specline{ \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{nextpaddr}  \ast \texttt{rax} \mapsto_{\textsf{r}} \textsf{nextpaddr}  }$
@@
 
-As we see in Figure \ref{fig:calltopteinitialize}, we have the physical address (\textsf{entry}) of the page table entry kept in the register $\texttt{r8}$, but not the memory mapping to load the entry value referenced by the physical address \textsf{entry} (Line 7 in Figure \ref{fig:calltopteinitialize}).
-We need to construct this virtual mappings for \textsf{entry}.
+As we see in Figure \ref{fig:calltopteinitialize}, we have the physical address (\textsf{entry}) of the page table entry kept in the register $\texttt{r8}$, but not the virtual memory mapping to load the entry value referenced by the physical address \textsf{entry} (Line 7 in Figure \ref{fig:calltopteinitialize}).
+We need to construct this virtual mapping for \textsf{entry}.
 \subsubsection{Identity Mappings}
@@
 
-As part of the full page table walk verification, we have verified the critical step for a small x86-64 kernel, which is the physical-to-virtual
-conversion (often appearing as a macro \texttt{P2V} inside the macro \textsf{ PTE\_PFN\_TO\_ADDR} in C source code). In our small kernel, as in larger kernels, \texttt{P2V} is actually just addition,
-but the correctness of this simple instruction is quite subtle, as shown in Figure \ref{fig:p2v} -- showing the section of the \textsf{get\_next\_table} in Figure \ref{calltopteinitialize} after the entry presence ensured (Lines 1-19 in Figure \ref{fig:calltopteinitialize}).
 Beyond a frame, the precondition before Line 7 requires the current address space invariant, a check that the addition will not overflow, the physical address to convert
@@
 
-Consequently, before Line 7 in Figure \ref{fig:calltopteinitialize}, to read the value referenced by physical address \textsf{entry}, first we open our the addresss-space invariant ($\mathcal{I}\textsf{ASpace}_{\textsf{id}}(\ptablestore,\Xi,m)$), and then obtain the virtual mapping out of the identity mapping $\Xi$ with the side condition
+Concretely speaking, before Line 7 in Figure \ref{fig:calltopteinitialize}, to read the value referenced by physical address \textsf{entry}, first we open our the addresss-space invariant ($\mathcal{I}\textsf{ASpace}_{\textsf{id}}(\ptablestore,\Xi,m)$), and then obtain the virtual mapping out of the identity mapping $\Xi$ with the side condition (\tag{*})
 \begin{equation*}
@@
  In case of absence, \textsf{pte\_initialize} is called for a physical-zeroed page (utilizing the page-allocator's \textsf{kalloc} -- currently the only axiom as it is not part of \textsf{pte} (virtual-memory-management) module) to be backing the physical frame address of the entry.
-\mytodo{Discuss with Colin: v2p and spec on pteinitialize.
-After clarification expand this piece a little bit more.
-Right now we have physical pointsto returned from kalloc so pfn is directly set not through v2p}
+
  Once we make sure that the frame of the page-table entry ($\textsf{entry}\mapsto\textsf{pfn}$)is backed with an allocated physical page, we can obtain the virtual address of the next level-page table.
 To do so, again, we refer to the  (\textsf{next} of function \textsf{get\_next\_table} in Figure \ref{}) via looking up the \textit{identity mapping} from the physical address (the address value referenced by $\textsf{entry}\mapsto\textsf{pfn}$) to virtual address value to be returned by \textsf{get\_next\_table}.
+
+ As part of the full page table walk verification, we have verified the critical step for a small x86-64 kernel, which is the physical-to-virtual
+conversion (often appearing as a macro \texttt{P2V} inside the macro \textsf{ PTE\_PFN\_TO\_ADDR} in C source code). In our small kernel, as in larger kernels, \texttt{P2V} is actually just addition,
+but the correctness of this simple instruction is quite subtle, as shown in Figure \ref{fig:p2v} -- showing the section of the \textsf{get\_next\_table} in Figure \ref{calltopteinitialize} after the entry presence ensured (Lines 1-19 in Figure \ref{fig:calltopteinitialize}).
+
 \begin{figure}\footnotesize

### 2024-07-01 examples.tex
+

@@
 $\specline{\sumpv\paddr\qfrac\true \ast \sumapaces\rtv\delta  }_{\rtv}$
-$\specline{ \textsf{entry-KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present entry\_val})\urcorner}$
+$\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present entry\_val})\urcorner}_{\rtv}$
 mov    [r8],rdi
@@
 $\specline{\sumpv\paddr\qfrac\true \ast \sumapaces\rtv\delta  }_{\rtv}$
-$\specline{ \textsf{entry-KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present entry\_val})  \urcorner }$
+$\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present entry\_val})  \urcorner }_{\rtv}$
 and    0x1,rdi
@@
 $\specline{\sumpv\paddr\qfrac\true \ast \sumapaces\rtv\delta  }_{\rtv}$
-$\specline{ \textsf{entry-KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present entry\_val})  \urcorner }$
+$\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present entry\_val})  \urcorner }_{\rtv}$
 mov    rbp,rdi
@@
 $\specline{\sumpv\paddr\qfrac\true \ast \sumapaces\rtv\delta  }_{\rtv}$
-$\specline{ \textsf{entry-KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \land \lnot(\textsf{entry\_present entry\_val})  \urcorner }$
+$\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \land \lnot(\textsf{entry\_present entry\_val})  \urcorner }_{\rtv}$
 ;;pte_initialize(&next);
@@
 $\specline{\sumpv\paddr\qfrac\true \ast \sumapaces\rtv\delta  }_{\rtv}$
-$\specline{ \textsf{entry-KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \land \lnot(\textsf{entry\_present entry\_val})  \urcorner }$
-$\specline{ \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{nextpaddr} \ast \textsf{nextpaddr} \mapsto_{\textsf{p}} \textsf{w64\_zero}  \ast \texttt{rax} \mapsto_{\textsf{r}} \textsf{nextpaddr}  }$
+$\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \land v\lnot(\textsf{entry\_present entry\_val})  \urcorner }_{\rtv}$
+$\specline{ \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{nextpaddr} \ast \textsf{nextpaddr} \mapsto_{\textsf{p}} \textsf{w64\_zero}  \ast \texttt{rax} \mapsto_{\textsf{r}} \textsf{nextpaddr}  }_{\rtv}$
 ;;entry->present = 1;
@@
 $\specline{\sumpv\paddr\qfrac\true \ast \sumapaces\rtv\delta  }_{\rtv}$
-$\specline{ \textsf{entry-KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \land (\textsf{entry\_present entry\_val}) \urcorner }$
-$\specline{ \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{nextpaddr}  \ast \texttt{rax} \mapsto_{\textsf{r}} \textsf{nextpaddr}  }$
+$\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \land (\textsf{entry\_present entry\_val}) \urcorner }_{\rtv}$
+$\specline{ \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{nextpaddr}  \ast \texttt{rax} \mapsto_{\textsf{r}} \textsf{nextpaddr}  }_{\rtv}$
 ;;}
@@
 ;;uintptr_t next_virt_addr = (uintptr_t) P2V(next_phys_addr);
-mov    -0x18[rbp],rax
 movabs 0x0,rcx	                187: R_X86_64_64	KERN_BASE
@@
 ...
+;;next = (pte_t *) next_virt_addr;
+;;clean up the stack and return next
 \end{lstlisting}
@@
 \subsubsection{Identity Mappings}
+\label{subsec:identitymappings}
 Most kernels maintain an invariant on their page tables that the virtual address of any page used for a page table lives at a virtual address
@@
 
-Concretely speaking, before Line 7 in Figure \ref{fig:calltopteinitialize}, to read the value referenced by physical address \textsf{entry}, first we open our the addresss-space invariant ($\mathcal{I}\textsf{ASpace}_{\textsf{id}}(\ptablestore,\Xi,m)$), and then obtain the virtual mapping out of the identity mapping $\Xi$ with the side condition (\tag{*})
+Concretely speaking, before Line 7 in Figure \ref{fig:calltopteinitialize}, to read the value referenced by physical address \textsf{entry}, first we open our the addresss-space invariant ($\mathcal{I}\textsf{ASpace}_{\textsf{id}}(\ptablestore,\Xi,m)$), and then obtain the virtual mapping out of the identity mapping $\Xi$ with the side condition (*)
 \begin{equation*}
@@
  assuring that looking up identity mappings are safe, because the full ownership asserts the non-presence of the entry which requires allocation a new physical page and setting the presence bit.
-By obtaining the identity mapping for \textsf{entry} in Figure \ref{fig:calltopteinitialize} -- ($\textsf{entry-KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val}$ with page addr \textsf{entry}),
+By obtaining the identity mapping for \textsf{entry} in Figure \ref{fig:calltopteinitialize} -- ($\textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val}$ with page addr \textsf{entry}),
  \[
  \begin{array}{l}
-\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{entry\},m) \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast r8  \mapsto_{\textsf{r}} \textsf{entry} \textsf{\_} \ast rdi  \mapsto_{\textsf{r}} \textsf{entry} \ast \textsf{entry-KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry\_val} \ast \sumapaces\rtv\delta  \ast \sumpv\paddr\qfrac\true  }_{\rtv}
+\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{entry\},m) \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast r8  \mapsto_{\textsf{r}} \textsf{entry} \textsf{\_} \ast rdi  \mapsto_{\textsf{r}} \textsf{entry} \ast \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry\_val} \ast \sumapaces\rtv\delta  \ast \sumpv\paddr\qfrac\true  }_{\rtv}
 mov    [r8],rdi
-\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast r8  \mapsto_{\textsf{r}} \textsf{entry} \textsf{\_} \ast rdi  \mapsto_{\textsf{r}} \textsf{entry\_val} \ast  \textsf{entry-KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry\_val} \ast \sumapaces\rtv\delta  \ast \sumpv\paddr\qfrac\true  }_{\rtv}
+\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast r8  \mapsto_{\textsf{r}} \textsf{entry} \textsf{\_} \ast rdi  \mapsto_{\textsf{r}} \textsf{entry\_val} \ast  \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry\_val} \ast \sumapaces\rtv\delta  \ast \sumpv\paddr\qfrac\true  }_{\rtv}
 \end{array}
@@
  \subsubsection{Ensuring Presence of an Entry}
- Accessing the presence bit and checking the value allow (Lines 7-13 in Figure \ref{fig:calltopteinitialize}) us to exploit the fact in the side condition (*) in such a way that the condition ($\textsf{entry}\mapsto\textsf{present}$ determines whether we take the non-presence path (Lines 12-19 in Figure \ref{calltopteinitialize}) with full ownership ($\textsf{qfrac} = 1$) to set the entry with proper values for the bitfields, e.g., frame showing freshly allocated page and presence bit.
+ Accessing the presence bit and checking the value allow (Lines 7-13 in Figure \ref{fig:calltopteinitialize}) us to exploit the fact in the side condition (*) in such a way that the condition ($\textsf{entry}\mapsto\textsf{present}$ determines whether we take the non-presence path (Lines 12-19 in Figure \ref{fig:calltopteinitialize}) with full ownership ($\textsf{qfrac} = 1$) to set the entry with proper values for the bitfields, e.g., frame showing freshly allocated page and presence bit.
+  In case of absence, \textsf{pte\_initialize} is called for a physical-zeroed page (utilizing the page-allocator's \textsf{kalloc} -- currently the only axiom as it is not part of \textsf{pte} (virtual-memory-management) module) to be backing the physical frame address of the entry -- Specification Line 35 in Figure \ref{fig:calltopteinitialize}.
 
- In case of absence, \textsf{pte\_initialize} is called for a physical-zeroed page (utilizing the page-allocator's \textsf{kalloc} -- currently the only axiom as it is not part of \textsf{pte} (virtual-memory-management) module) to be backing the physical frame address of the entry.
+  \subsubsection{Kernels Utilizing \textsf{P2V}}
+  \label{sec:p2v}
+ Once we make sure that the frame of the page-table entry ($\textsf{entry}\mapsto\textsf{pfn}$)is backed with an allocated physical page, we must obtain the virtual address of the next level-page table (Line 46 in \ref{calltopteinitialize}).
+To do so, as for obtaining any virtual page-table address, we, again, need to open the invariant look up the \textit{identity mapping} for the physical address ($\textsf{entry}\mapsto \textsf{pfn}$).
 
- Once we make sure that the frame of the page-table entry ($\textsf{entry}\mapsto\textsf{pfn}$)is backed with an allocated physical page, we can obtain the virtual address of the next level-page table.
-To do so, again, we refer to the  (\textsf{next} of function \textsf{get\_next\_table} in Figure \ref{}) via looking up the \textit{identity mapping} from the physical address (the address value referenced by $\textsf{entry}\mapsto\textsf{pfn}$) to virtual address value to be returned by \textsf{get\_next\_table}.
-
- As part of the full page table walk verification, we have verified the critical step for a small x86-64 kernel, which is the physical-to-virtual
-conversion (often appearing as a macro \texttt{P2V} inside the macro \textsf{ PTE\_PFN\_TO\_ADDR} in C source code). In our small kernel, as in larger kernels, \texttt{P2V} is actually just addition,
-but the correctness of this simple instruction is quite subtle, as shown in Figure \ref{fig:p2v} -- showing the section of the \textsf{get\_next\_table} in Figure \ref{calltopteinitialize} after the entry presence ensured (Lines 1-19 in Figure \ref{fig:calltopteinitialize}).
-
+This actually constitutes a very critical piece of the full page table walk verification, and we have verified the critical step for a small x86-64 kernel, which is the physical-to-virtual conversion (often appearing as a macro \texttt{P2V} inside the macro \textsf{ PTE\_PFN\_TO\_ADDR} in C source code).
+In our small kernel, as in larger kernels, \texttt{P2V} is actually just addition, but the correctness of this simple instruction is quite subtle and relies on identity mappings explained in the previous section (Section \ref{subsec:identitymappings}) as a part of address-space invariant.
+As as separate Lemma shown in Figure \ref{fig:p2v},
 \begin{figure}\footnotesize
 \begin{lstlisting}[mathescape]
-;;if (!entry->present) {
 ...
-
-;;}
+;;uintptr_t next_virt_addr = (uintptr_t) P2V(next_phys_addr);
+movabs 0x0,rcx	                187: R_X86_64_64	KERN_BASE
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val}  }_{\rtv}$
+$\specline{\sumpv\paddr\qfrac\true \ast \sumapaces\rtv\delta  }_{\rtv}$
+$\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \land (\textsf{entry\_present entry\_val}) \urcorner }$
+$\specline{ \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{nextpaddr}  \ast \texttt{rax} \mapsto_{\textsf{r}} \textsf{nextpaddr} \ast  \texttt{rcx} \mapsto_{\textsf{r}} \textsf{KERNBASE}  }_{\rtv}$
+add    rcx,rax
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val}  }_{\rtv}$
+$\specline{\sumpv\paddr\qfrac\true \ast \sumapaces\rtv\delta  }_{\rtv}$
+$\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \land (\textsf{entry\_present entry\_val}) \urcorner }$
+$\specline{ \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{nextpaddr}  \ast \texttt{rax} \mapsto_{\textsf{r}} \textsf{nextpaddr} \ast \texttt{rcx} \mapsto_{\textsf{r}} \textsf{KERNBASE} }_{\rtv}$
+$\specline{ \texttt{nextpaddr+KERNBASE} \mapsto_{\textsf{vpte,qfracn}} \textsf{nextpaddr w64\_zero} \ }_{\rtv}$
 ...
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)  \ast \texttt{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{\_} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{\_} \ast \sumpv\paddr\qfrac\true  \sumapaces\rtv\delta  }_{\rtv}$
-mov    rax,-0x18[rbp]
-
-;;uintptr_t next_virt_addr = (uintptr_t) P2V(next_phys_addr);
-mov    -0x18[rbp],rax
-movabs 0x0,rcx            187: R_X86_64_64	KERN_BASE
-add    rcx,rax
-mov    rax,-0x20[rbp]
+;;next = (pte_t *) next_virt_addr;
+;;clean up the stack and return next
 \end{lstlisting}
 \vspace{-1em}
-\caption{Converting a physical address, whose exsitence is made certain, of a PTE to a virtual address (w/o instruction pointer or flag updates). $R$ abbreviates the resources
+\caption{Converting a physical address of a PTE to a virtual address (w/o instruction pointer or flag updates). $R$ abbreviates the resources
 associated with \paddr\ by $\Xi$ in Figure \ref{fig:peraspaceinvariant_with_p2v_extension}.}
@@
 \end{figure}
+
+%\begin{figure}\footnotesize
+%\begin{lstlisting}[mathescape]
+%;;if (!entry->present) {
+%  ...
+%
+%;;}
+%...
+%$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)  \ast \texttt{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{\_} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{\_} \ast \sumpv\paddr\qfrac\true  \sumapaces\rtv\delta  }_{\rtv}$
+%mov    rax,-0x18[rbp]
+ % 
+%;;uintptr_t next_virt_addr = (uintptr_t) P2V(next_phys_addr);
+%mov    -0x18[rbp],rax
+%movabs 0x0,rcx            187: R_X86_64_64	KERN_BASE
+%add    rcx,rax
+%mov    rax,-0x20[rbp]
+%\end{lstlisting}
+%\vspace{-1em}
+%\caption{Converting a physical address, whose exsitence is made certain, of a PTE to a virtual address (w/o instruction pointer or flag updates). $R$ abbreviates the resources
+%associated with \paddr\ by $\Xi$ in Figure \ref{fig:peraspaceinvariant_with_p2v_extension}.}
+%\label{fig:p2v}
+%\end{figure}
 

### 2024-07-01 examples.tex
+

@@
 $\specline{\sumpv\paddr\qfrac\true \ast \sumapaces\rtv\delta  }_{\rtv}$
-$\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \land (\textsf{entry\_present entry\_val}) \urcorner }_{\rtv}$
+$\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val\_present} \ast \ulcorner  \textsf{qfrac} = 1 \land (\textsf{entry\_present entry\_val}) \urcorner }_{\rtv}$
 $\specline{ \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{nextpaddr}  \ast \texttt{rax} \mapsto_{\textsf{r}} \textsf{nextpaddr}  }_{\rtv}$
@@
  \begin{array}{l}
-\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{entry\},m) \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast r8  \mapsto_{\textsf{r}} \textsf{entry} \textsf{\_} \ast rdi  \mapsto_{\textsf{r}} \textsf{entry} \ast \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry\_val} \ast \sumapaces\rtv\delta  \ast \sumpv\paddr\qfrac\true  }_{\rtv}
-mov    [r8],rdi
+\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{entry\},m) \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast r8  \mapsto_{\textsf{r}} \textsf{entry} \textsf{\_} \ast rdi  \mapsto_{\textsf{r}} \textsf{entry} \ast \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry\_val} \ast \sumapaces\rtv\delta  \ast \sumpv\paddr\qfrac\true  }_{\rtv} \\
+mov    [r8],rdi\\
 \specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast r8  \mapsto_{\textsf{r}} \textsf{entry} \textsf{\_} \ast rdi  \mapsto_{\textsf{r}} \textsf{entry\_val} \ast  \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry\_val} \ast \sumapaces\rtv\delta  \ast \sumpv\paddr\qfrac\true  }_{\rtv}
@@
 This actually constitutes a very critical piece of the full page table walk verification, and we have verified the critical step for a small x86-64 kernel, which is the physical-to-virtual conversion (often appearing as a macro \texttt{P2V} inside the macro \textsf{ PTE\_PFN\_TO\_ADDR} in C source code).
-In our small kernel, as in larger kernels, \texttt{P2V} is actually just addition, but the correctness of this simple instruction is quite subtle and relies on identity mappings explained in the previous section (Section \ref{subsec:identitymappings}) as a part of address-space invariant.
-As as separate Lemma shown in Figure \ref{fig:p2v},
+In our small kernel (Line 46 in Figure \ref{calltopteinitialize}), as in larger kernels, \texttt{P2V} is actually just addition, but the correctness of this simple instruction is quite subtle and relies on identity mappings explained in the previous section (Section \ref{subsec:identitymappings}) as a part of understanding our address-space invariant in Figure \ref{fig:peraspaceinvariant_with_p2v_extension}.
 \begin{figure}\footnotesize
@@
 $\specline{ \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{nextpaddr}  \ast \texttt{rax} \mapsto_{\textsf{r}} \textsf{nextpaddr} \ast \texttt{rcx} \mapsto_{\textsf{r}} \textsf{KERNBASE} }_{\rtv}$
-$\specline{ \texttt{nextpaddr+KERNBASE} \mapsto_{\textsf{vpte,qfracn}} \textsf{nextpaddr w64\_zero} \ }_{\rtv}$
+$\specline{ \texttt{nextpaddr+KERNBASE} \mapsto_{\textsf{vpte,qfracn}} \textsf{nextpaddr w64\_zero} }_{\rtv}$
 ...
@@
 \vspace{-1em}
-\caption{Converting a physical address of a PTE to a virtual address (w/o instruction pointer or flag updates). $R$ abbreviates the resources
-associated with \paddr\ by $\Xi$ in Figure \ref{fig:peraspaceinvariant_with_p2v_extension}.}
+\caption{Converting a physical address of a PTE to a virtual address (w/o instruction pointer or flag updates).}
 \label{fig:p2v}
 \end{figure}
+As shown in Figure \ref{fig:p2v}, \lstinline|rcx| holds the offset value \textsf{KERNBASE} which gives us the virtual address (\textsf{nextpaddr+KERNBASE}) to look-up in the identity map ($\Xi$) of the kernel invariant.
+Assuming the path for \textsf{pte\_initialize} is taken, we can obtain the virtual pte-pointsto for the physical address \textsf{nextpaddr} (Line 13 in Figure \ref{fig:p2v}).
 

### 2024-07-02 examples.tex
+

@@
 $\specline{\sumpv\paddr\qfrac\true \ast \sumapaces\rtv\delta  }_{\rtv}$
-$\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present entry\_val})  \urcorner }_{\rtv}$
+$\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \land \lnot(\textsf{entry\_present entry\_val})  \urcorner }_{\rtv}$
 mov    rbp,rdi
@@
 $\specline{\sumpv\paddr\qfrac\true \ast \sumapaces\rtv\delta  }_{\rtv}$
-$\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \land v\lnot(\textsf{entry\_present entry\_val})  \urcorner }_{\rtv}$
+$\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \land \lnot(\textsf{entry\_present entry\_val})  \urcorner }_{\rtv}$
 $\specline{ \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{nextpaddr} \ast \textsf{nextpaddr} \mapsto_{\textsf{p}} \textsf{w64\_zero}  \ast \texttt{rax} \mapsto_{\textsf{r}} \textsf{nextpaddr}  }_{\rtv}$
@@
 ;;entry->pfn = next
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val}  }_{\rtv}$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val'}  }_{\rtv}$
 $\specline{\sumpv\paddr\qfrac\true \ast \sumapaces\rtv\delta  }_{\rtv}$
-$\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val\_present} \ast \ulcorner  \textsf{qfrac} = 1 \land (\textsf{entry\_present entry\_val}) \urcorner }_{\rtv}$
+$\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val\_present} \ast \ulcorner  \textsf{qfrac} = 1 \land (\textsf{entry\_present entry\_val'}) \urcorner }_{\rtv}$
 $\specline{ \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{nextpaddr}  \ast \texttt{rax} \mapsto_{\textsf{r}} \textsf{nextpaddr}  }_{\rtv}$
@@
   \bigast{(\vaddr, \textsf{paddr})\in \ptablestore}{\exists\;(\textsf{l4e, l3e, l2e, l1e, paddr})\ldotp \textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\vaddr\textsf{, l4e, l3e, l2e, l1e, paddr})} \\
-  \bigast{(\paddr) \in \Xi}{\exists\; (\textsf{qfrac, q, val,}\vaddr) \ldotp \ulcorner \vaddr = \paddr + \textsf{OFF} \urcorner \ast  \underbrace{\sumwalkabs\vaddr\qfrac\paddr }_\text{Ghost translation} \ast
+  \bigast{(\paddr) \in \Xi}{\exists\; (\textsf{qfrac, q, val,}\vaddr) \ldotp \ulcorner \vaddr = \paddr + \textsf{KERNBASE} \urcorner \ast \ulcorner \textsf{qfrac} = 1 \leftrightarrow \; \lnot\textsf{entry\_present }(\vale) \urcorner \\
+                           \qquad\qquad\qquad\qquad\qquad\qquad          \underbrace{\sumwalkabs\vaddr\qfrac\paddr }_\text{Ghost translation} \ast 
 % \qquad\qquad\qquad\qquad\qquad\qquad\qquad\qquad    
-\underbrace{\paddr \mapsto_{\mathsf{p}}\{\textsf{qfrac}\}\; \vale}_\text{Physical location} \ast \ulcorner \textsf{qfrac} = 1 \leftrightarrow \; \lnot\textsf{entry\_present }(\vale) \urcorner }\\
+\underbrace{\paddr \mapsto_{\mathsf{p}}\{\textsf{qfrac}\}\; \vale}_\text{Physical location}}\\
   \textsf{ where } \\
@@
 
-Beyond a frame, the precondition before Line 7 requires the current address space invariant, a check that the addition will not overflow, the physical address to convert
-(in \lstinline|rax|), knowledge that the current address space is valid, and knowledge that the physical address is in a range dedicated
-to page table entries (as this physical-to-virtual conversion is only guaranteed for page table memory). 
+Beyond a frame, the precondition before Line 7 requires the current address space invariant, and knowledge that \textsf{entry} is in the identity-map ($\Xi$) -- $\textsf{paddr} \in \Xi$.
+By knowing this, we can also refer to the facts in the invariant that are related to a physical page-table addresses is in a range dedicated to page-table entries (as this physical-to-virtual conversion is only guaranteed for page table memory), and check that the addition of the offset (i.e.
+\textsf{KERNBASE}) to a physical page-table address do not overflow (in \lstinline|rax|).
+
 The latter uses an additional ghost set added to the per-address-space
@@
 
-Concretely speaking, before Line 7 in Figure \ref{fig:calltopteinitialize}, to read the value referenced by physical address \textsf{entry}, first we open our the addresss-space invariant ($\mathcal{I}\textsf{ASpace}_{\textsf{id}}(\ptablestore,\Xi,m)$), and then obtain the virtual mapping out of the identity mapping $\Xi$ with the side condition (*)
+Concretely speaking, before Line 7 in Figure \ref{fig:calltopteinitialize}, to read the value referenced by physical address \textsf{entry}, first, we open our the addresss-space invariant ($\mathcal{I}\textsf{ASpace}_{\textsf{id}}(\ptablestore,\Xi,m)$), and then obtain the virtual mapping out of the identity mapping $\Xi$ with the side condition (*)
 \begin{equation*}
@@
 \end{equation*}
- assuring that looking up identity mappings are safe, because the full ownership asserts the non-presence of the entry which requires allocation a new physical page and setting the presence bit.
-By obtaining the identity mapping for \textsf{entry} in Figure \ref{fig:calltopteinitialize} -- ($\textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val}$ with page addr \textsf{entry}),
+ assuring that looking up identity mappings are safe.
+The subtle justification relies on the fact that the full ownership asserts the non-presence of the entry, accessing to which requires allocation a new physical page and having the presence bit set.
+By obtaining the identity mapping for \textsf{entry} needed before Line 7 in Figure \ref{fig:calltopteinitialize}, we have to get it from the $Xi$ with the specification shown in \ref{fig:lookingupidpampping}, relying on the
+
+ \begin{figure*}
+   \footnotesize
    \[
  \begin{array}{l}
-\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{entry\},m) \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast r8  \mapsto_{\textsf{r}} \textsf{entry} \textsf{\_} \ast rdi  \mapsto_{\textsf{r}} \textsf{entry} \ast \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry\_val} \ast \sumapaces\rtv\delta  \ast \sumpv\paddr\qfrac\true  }_{\rtv} \\
-mov    [r8],rdi\\
-\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast r8  \mapsto_{\textsf{r}} \textsf{entry} \textsf{\_} \ast rdi  \mapsto_{\textsf{r}} \textsf{entry\_val} \ast  \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry\_val} \ast \sumapaces\rtv\delta  \ast \sumpv\paddr\qfrac\true  }_{\rtv}
+   \specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast r8  \mapsto_{\textsf{r}} \textsf{entry} \textsf{\_} \ast }_{\rtv} \\
+\specline{  rdi  \mapsto_{\textsf{r}} \textsf{entry\_val} \ast  \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry\_val} \ast \sumapaces\rtv\delta  \ast \sumpv\paddr\qfrac\true  }_{\rtv} \\
+\sqsubseteq \\
+   \specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{entry\},m) \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast r8  \mapsto_{\textsf{r}} \textsf{entry} \textsf{\_} \ast }_{\rtv} \\
+ \specline{     rdi  \mapsto_{\textsf{r}} \textsf{entry} \ast \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry\_val} \ast \sumapaces\rtv\delta  \ast \sumpv\paddr\qfrac\true  }_{\rtv} \\
+\mathsf{mov} \;   [ \mathsf{r8} ],\mathsf{rdi}\\
+\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{\textsf{entry}\},m) \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast r8  \mapsto_{\textsf{r}} \textsf{entry} \textsf{\_} \ast }_{\rtv} \\
+\specline{  rdi  \mapsto_{\textsf{r}} \textsf{entry\_val} \ast  \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry\_val} \ast \sumapaces\rtv\delta  \ast \sumpv\paddr\qfrac\true  }_{\rtv}
  \end{array}
  \]
+ \vspace{-1em}
+\caption{Global Address-Space Invariant in Figure \ref{fig:peraspaceinvariant} extended with a ghost map bookkeeping identity mappings }
+  \label{fig:lookingupidpampping}
+ \end{figure*}
  we are able to load the \textsf{entry\_val} into \textsf{rdi}, and check the presence bit via $and \;0x1,rdi$.

### 2024-07-02 examples.tex
+

@@
 The subtle justification relies on the fact that the full ownership asserts the non-presence of the entry, accessing to which requires allocation a new physical page and having the presence bit set.
-By obtaining the identity mapping for \textsf{entry} needed before Line 7 in Figure \ref{fig:calltopteinitialize}, we have to get it from the $Xi$ with the specification shown in \ref{fig:lookingupidpampping}, relying on the
-
+Acquiring the identity mapping for \textsf{entry} needed before Line 7 in Figure \ref{fig:calltopteinitialize}, we need the knowledge $\textsf{entry} \in \Xi$ to peel-out the fractional token representing the ghost address translation from the offsetted virtual address (\textsf{entry+KERNBASE}) to the physical page address (\textsf{entry}) and physical pointsto relation from the physical address \textsf{entry} to an existentially quantified value (\textsf{entry\_val}) and the side condition (*) we explained above.
+In in Figure \ref{fig:lookingupidpampping}, we see that all the logical resources needed to construct a virtual-pte pointsto
+\[ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \]
+to execute Line 7 in \ref{fig:calltopteinitialize} that can either be mutated or not based on the side condition (*).
  \begin{figure*}
@@
  \begin{array}{l}
-   \specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast r8  \mapsto_{\textsf{r}} \textsf{entry} \textsf{\_} \ast }_{\rtv} \\
-\specline{  rdi  \mapsto_{\textsf{r}} \textsf{entry\_val} \ast  \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry\_val} \ast \sumapaces\rtv\delta  \ast \sumpv\paddr\qfrac\true  }_{\rtv} \\
+   \specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry} \textsf{\_} \ast \textsf{entry} \in \Xi }_{\rtv}  \\
 \sqsubseteq \\
-   \specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{entry\},m) \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast r8  \mapsto_{\textsf{r}} \textsf{entry} \textsf{\_} \ast }_{\rtv} \\
- \specline{     rdi  \mapsto_{\textsf{r}} \textsf{entry} \ast \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry\_val} \ast \sumapaces\rtv\delta  \ast \sumpv\paddr\qfrac\true  }_{\rtv} \\
+   \specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{entry\},m) \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry} \textsf{\_} \ast  \textsf{entry} \notin \Xi }_{\rtv} \\
+ \specline{     \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry} \ast \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \sumapaces\rtv\delta  \ast \sumpv\paddr\qfrac\true  }_{\rtv} \\
 \mathsf{mov} \;   [ \mathsf{r8} ],\mathsf{rdi}\\
-\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{\textsf{entry}\},m) \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast r8  \mapsto_{\textsf{r}} \textsf{entry} \textsf{\_} \ast }_{\rtv} \\
-\specline{  rdi  \mapsto_{\textsf{r}} \textsf{entry\_val} \ast  \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry\_val} \ast \sumapaces\rtv\delta  \ast \sumpv\paddr\qfrac\true  }_{\rtv}
+\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{\textsf{entry}\},m) \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry} \textsf{\_} \ast  \textsf{entry} \notin \Xi}_{\rtv} \\
+\specline{  \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val} \ast  \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \sumapaces\rtv\delta  \ast \sumpv\paddr\qfrac\true  }_{\rtv}
  \end{array}
@@
  \vspace{-1em}
-\caption{Global Address-Space Invariant in Figure \ref{fig:peraspaceinvariant} extended with a ghost map bookkeeping identity mappings }
+\caption{Peel-out the resources needed for the virtual-pte pointsto relation for physical address \textsf{entry} to be accessed in Line 7 in Figure \ref{fig:calltopteinitialize}. }
   \label{fig:lookingupidpampping}
  \end{figure*}
- we are able to load the \textsf{entry\_val} into \textsf{rdi}, and check the presence bit via $and \;0x1,rdi$.
 
  \subsubsection{Ensuring Presence of an Entry}
+After obtaining the identity mapping for \textsf{entry}, we are able to load the \textsf{entry\_val} into \textsf{rdi}, and check the presence bit via $and \;0x1,rdi$.
  Accessing the presence bit and checking the value allow (Lines 7-13 in Figure \ref{fig:calltopteinitialize}) us to exploit the fact in the side condition (*) in such a way that the condition ($\textsf{entry}\mapsto\textsf{present}$ determines whether we take the non-presence path (Lines 12-19 in Figure \ref{fig:calltopteinitialize}) with full ownership ($\textsf{qfrac} = 1$) to set the entry with proper values for the bitfields, e.g., frame showing freshly allocated page and presence bit.

### 2024-07-02 examples.tex
+

@@
 To be able to do so, it needs to load the entry value (a 64-bits word divided into bit-fields for corresponding fileds shown in Section {}) and check that \textit{presence} bit of the entry is set or not.
-
+%wshiftll (wshiftll (natToWord 64 entry) (WordImpl.concat (WordImpl.zero 56) (WordImpl.from_nat 8 12 ^& WordImpl.concat (WordImpl.zero 2) WO~1~1~1~1~1~1)) ^& constf)
+%(WordImpl.concat (WordImpl.zero 56) (natToWord 8 12 ^& WordImpl.concat (WordImpl.zero 2) WO~1~1~1~1~1~1))
+%
+%wshiftll
+ %      (wshiftll
+%          ((((natToWord 64 entry ^& WordImpl.concat (WordImpl.zero 32) consta ^| WordImpl.concat (WordImpl.zero 32) (natToWord 32 2))
+%             ^& WordImpl.concat (WordImpl.zero 32) constb ^| WordImpl.concat (WordImpl.zero 32) (natToWord 32 4)) ^& constd
+%            ^| wshiftll
+%                 (wshiftll (nextpaddr ^+ ^~ (natToWord 64 KERNBASE))
+%                    (WordImpl.concat (WordImpl.zero 56) (WordImpl.from_nat 8 12 ^& WordImpl.concat (WordImpl.zero 2) WO~1~1~1~1~1~1))
+%                  ^& constf)
+%                 (WordImpl.concat (WordImpl.zero 56) (WordImpl.from_nat 8 12 ^& WordImpl.concat (WordImpl.zero 2) WO~1~1~1~1~1~1)))
+%           ^& WordImpl.concat (WordImpl.zero 32) conste ^| wone 64)
+%          (WordImpl.concat (WordImpl.zero 56) (WordImpl.from_nat 8 12 ^& WordImpl.concat (WordImpl.zero 2) WO~1~1~1~1~1~1)) ^& constf)
+%       (WordImpl.concat (WordImpl.zero 56) (natToWord 8 12 ^& WordImpl.concat (WordImpl.zero 2) WO~1~1~1~1~1~1)) 
 \begin{figure}\footnotesize
@@
 $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)  \ast \texttt{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast  \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{next} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{\_} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{\_}  }_{\rtv}$
-$\specline{\sumpv\paddr\qfrac\true \ast \sumapaces\rtv\delta  }_{\rtv}$
+$\specline{\mathsf{entry} \in \Xi \ast \sumapaces\rtv\delta  }_{\rtv}$
 mov    -0x8[rbp],rdi
 $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{\_}  \ast  \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{next} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry}  }_{\rtv}$
-$\specline{\sumpv\paddr\qfrac\true \ast \sumapaces\rtv\delta  }_{\rtv}$
+$\specline{ \mathsf{entry} \in \Xi \ast \sumapaces\rtv\delta  }_{\rtv}$
 mov     rdi, r8
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast  \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{next} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry} }_{\rtv}$
-$\specline{\sumpv\paddr\qfrac\true \ast \sumapaces\rtv\delta  }_{\rtv}$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ entry \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast  \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{next} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry} }_{\rtv}$
+$\specline{ \sumapaces\rtv\delta  }_{\rtv}$
 $\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present entry\_val})\urcorner}_{\rtv}$
 mov    [r8],rdi
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast  \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{next} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val}   }_{\rtv}$
-$\specline{\sumpv\paddr\qfrac\true \ast \sumapaces\rtv\delta  }_{\rtv}$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ entry \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast  \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{next} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val}   }_{\rtv}$
+$\specline{ \sumapaces\rtv\delta  }_{\rtv}$
 $\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present entry\_val})  \urcorner }_{\rtv}$
@@
 jne    161 <pte_get_next_table_succ+0xa1>
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry} \ast \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{next} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val} }_{\rtv}$
-$\specline{\sumpv\paddr\qfrac\true \ast \sumapaces\rtv\delta  }_{\rtv}$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ entry \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry} \ast \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{next} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val} }_{\rtv}$
+$\specline{ \sumapaces\rtv\delta  }_{\rtv}$
 $\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \land \lnot(\textsf{entry\_present entry\_val})  \urcorner }_{\rtv}$
@@
 sub    0x10, rdi
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry} \ast \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{next} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val}  }_{\rtv}$
-$\specline{\sumpv\paddr\qfrac\true \ast \sumapaces\rtv\delta  }_{\rtv}$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ entry \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry} \ast \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{next} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val}  }_{\rtv}$
+$\specline{ \sumapaces\rtv\delta  }_{\rtv}$
 $\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \land \lnot(\textsf{entry\_present entry\_val})  \urcorner }_{\rtv}$
@@
 callq  70 <pte_initialize>
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val}  }_{\rtv}$
-$\specline{\sumpv\paddr\qfrac\true \ast \sumapaces\rtv\delta  }_{\rtv}$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ entry \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val}  }_{\rtv}$
+$\specline{\sumapaces\rtv\delta }_{\rtv}$
 $\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \land \lnot(\textsf{entry\_present entry\_val})  \urcorner }_{\rtv}$
@@
 ;;entry->pfn = next
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val'}  }_{\rtv}$
-$\specline{\sumpv\paddr\qfrac\true \ast \sumapaces\rtv\delta  }_{\rtv}$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ entry \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val'}  }_{\rtv}$
+$\specline{ \sumapaces\rtv\delta  \ast \mathsf{entry\_val.pfn} \in \Xi \setminus \{ entry \} }_{\rtv}$
 $\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val\_present} \ast \ulcorner  \textsf{qfrac} = 1 \land (\textsf{entry\_present entry\_val'}) \urcorner }_{\rtv}$

### 2024-07-02 examples.tex
+

@@
 mov     rdi, r8
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ entry \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast  \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{next} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry} }_{\rtv}$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast  \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{next} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry} }_{\rtv}$
 $\specline{ \sumapaces\rtv\delta  }_{\rtv}$
@@
 mov    [r8],rdi
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ entry \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast  \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{next} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val}   }_{\rtv}$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast  \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{next} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val}   }_{\rtv}$
 $\specline{ \sumapaces\rtv\delta  }_{\rtv}$
@@
 jne    161 <pte_get_next_table_succ+0xa1>
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ entry \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry} \ast \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{next} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val} }_{\rtv}$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry} \ast \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{next} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val} }_{\rtv}$
 $\specline{ \sumapaces\rtv\delta  }_{\rtv}$
@@
 sub    0x10, rdi
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ entry \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry} \ast \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{next} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val}  }_{\rtv}$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry} \ast \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{next} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val}  }_{\rtv}$
 $\specline{ \sumapaces\rtv\delta  }_{\rtv}$
@@
 callq  70 <pte_initialize>
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ entry \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val}  }_{\rtv}$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val}  }_{\rtv}$
 $\specline{\sumapaces\rtv\delta }_{\rtv}$
-$\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \land \lnot(\textsf{entry\_present entry\_val})  \urcorner }_{\rtv}$
+$\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \mathsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \land \lnot(\textsf{entry\_present entry\_val})  \urcorner }_{\rtv}$
 $\specline{ \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{nextpaddr} \ast \textsf{nextpaddr} \mapsto_{\textsf{p}} \textsf{w64\_zero}  \ast \texttt{rax} \mapsto_{\textsf{r}} \textsf{nextpaddr}  }_{\rtv}$
@@
 ;;entry->pfn = next
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ entry \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val'}  }_{\rtv}$
-$\specline{ \sumapaces\rtv\delta  \ast \mathsf{entry\_val.pfn} \in \Xi \setminus \{ entry \} }_{\rtv}$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val'}  }_{\rtv}$
+$\specline{ \sumapaces\rtv\delta   \} }_{\rtv}$
 $\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val\_present} \ast \ulcorner  \textsf{qfrac} = 1 \land (\textsf{entry\_present entry\_val'}) \urcorner }_{\rtv}$
@@
 \sqsubseteq \\
-   \specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{entry\},m) \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry} \textsf{\_} \ast  \textsf{entry} \notin \Xi }_{\rtv} \\
- \specline{     \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry} \ast \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \sumapaces\rtv\delta  \ast \sumpv\paddr\qfrac\true  }_{\rtv} \\
+   \specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast  \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{next} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry} }_{\rtv}\\
+   \specline{ \sumapaces\rtv\delta  }_{\rtv} \\
+\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present entry\_val})\urcorner}_{\rtv} \\
 \mathsf{mov} \;   [ \mathsf{r8} ],\mathsf{rdi}\\
-\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{\textsf{entry}\},m) \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry} \textsf{\_} \ast  \textsf{entry} \notin \Xi}_{\rtv} \\
-\specline{  \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val} \ast  \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \sumapaces\rtv\delta  \ast \sumpv\paddr\qfrac\true  }_{\rtv}
+ \specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast  \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{next} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry} }_{\rtv}\\
+   \specline{ \sumapaces\rtv\delta  }_{\rtv} \\
+\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present entry\_val})\urcorner}_{\rtv} \\
  \end{array}
@@
 ...
-;;uintptr_t next_virt_addr = (uintptr_t) P2V(next_phys_addr);
+;;uintptr_t next_virt_addr = (uintptr_t) P2V(next<<12);
 movabs 0x0,rcx	                187: R_X86_64_64	KERN_BASE
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val}  }_{\rtv}$
-$\specline{\sumpv\paddr\qfrac\true \ast \sumapaces\rtv\delta  }_{\rtv}$
-$\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \land (\textsf{entry\_present entry\_val}) \urcorner }$
-$\specline{ \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{nextpaddr}  \ast \texttt{rax} \mapsto_{\textsf{r}} \textsf{nextpaddr} \ast  \texttt{rcx} \mapsto_{\textsf{r}} \textsf{KERNBASE}  }_{\rtv}$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val'}  }_{\rtv}$
+$\specline{ \sumapaces\rtv\delta  \ast \mathsf{entry\_val'.pfn}<<12 \in \Xi \setminus \{ \mathsf{entry} \} }_{\rtv}$
+$\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val\_present} \ast \ulcorner  \textsf{qfrac} = 1 \land (\textsf{entry\_present entry\_val'}) \urcorner }_{\rtv}$
+$\specline{ \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{nextpaddr}  \ast \texttt{rax} \mapsto_{\textsf{r}} \textsf{nextpaddr}  }_{\rtv}$
 add    rcx,rax
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val}  }_{\rtv}$
-$\specline{\sumpv\paddr\qfrac\true \ast \sumapaces\rtv\delta  }_{\rtv}$
-$\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \land (\textsf{entry\_present entry\_val}) \urcorner }$
-$\specline{ \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{nextpaddr}  \ast \texttt{rax} \mapsto_{\textsf{r}} \textsf{nextpaddr} \ast \texttt{rcx} \mapsto_{\textsf{r}} \textsf{KERNBASE} }_{\rtv}$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} , \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val'}  }_{\rtv}$
+$\specline{ \sumapaces\rtv\delta  \ast \mathsf{entry\_val'.pfn}<<12 \in \Xi \setminus \{ \mathsf{entry} \} }_{\rtv}$
+$\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val\_present} \ast \ulcorner  \textsf{qfrac} = 1 \land (\textsf{entry\_present entry\_val'}) \urcorner }_{\rtv}$
+$\specline{ \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{nextpaddr}  \ast \texttt{rax} \mapsto_{\textsf{r}} \textsf{nextpaddr}  }_{\rtv}$
 $\specline{ \texttt{nextpaddr+KERNBASE} \mapsto_{\textsf{vpte,qfracn}} \textsf{nextpaddr w64\_zero} }_{\rtv}$

### 2024-07-02 examples.tex
+

@@
 add    rcx,rax
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} , \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val'}  }_{\rtv}$
-$\specline{ \sumapaces\rtv\delta  \ast \mathsf{entry\_val'.pfn}<<12 \in \Xi \setminus \{ \mathsf{entry} \} }_{\rtv}$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} ,\mathsf{entry\_val'.pfn}<<12 \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val'}  }_{\rtv}$
+$\specline{ \sumapaces\rtv\delta  }_{\rtv}$
 $\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val\_present} \ast \ulcorner  \textsf{qfrac} = 1 \land (\textsf{entry\_present entry\_val'}) \urcorner }_{\rtv}$

### 2024-07-02 examples.tex
+

@@
 $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val}  }_{\rtv}$
-$\specline{\sumapaces\rtv\delta }_{\rtv}$
+$\specline{\sumapaces\rtv\delta \ast \mathsf{nextpaddr} \in \Xi\setminus \{ \mathsf{entry} \} \cup \{ \mathsf{nextpaddr} \} }_{\rtv}$
 $\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \mathsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \land \lnot(\textsf{entry\_present entry\_val})  \urcorner }_{\rtv}$
@@
 $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val'}  }_{\rtv}$
-$\specline{ \sumapaces\rtv\delta   \} }_{\rtv}$
-$\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val\_present} \ast \ulcorner  \textsf{qfrac} = 1 \land (\textsf{entry\_present entry\_val'}) \urcorner }_{\rtv}$
-$\specline{ \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{nextpaddr}  \ast \texttt{rax} \mapsto_{\textsf{r}} \textsf{nextpaddr}  }_{\rtv}$
+$\specline{\sumapaces\rtv\delta \ast \mathsf{nextpaddr} \in \Xi\setminus \{ \mathsf{entry} \} \cup \{ \mathsf{nextpaddr} \} }_{\rtv}$
+$\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \mathsf{entry entry\_val'} \ast \ulcorner  \textsf{qfrac} = 1 \land (\textsf{entry\_present entry\_val'})  \urcorner }_{\rtv}$
+$\specline{ \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{nextpaddr} \ast \textsf{nextpaddr} \mapsto_{\textsf{p}} \textsf{w64\_zero}  \ast \texttt{rax} \mapsto_{\textsf{r}} \textsf{nextpaddr}  }_{\rtv}$
 ;;}
@@
  In case of absence, \textsf{pte\_initialize} is called for a physical-zeroed page (utilizing the page-allocator's \textsf{kalloc} -- currently the only axiom as it is not part of \textsf{pte} (virtual-memory-management) module) to be backing the physical frame address of the entry -- Specification Line 35 in Figure \ref{fig:calltopteinitialize}.
-
+Since we are using \textsf{pte\_initialize} for page-table address allocation, it also ensures us that newly allocated physicall address is in the identity mapping map $\Xi$ -- Line X in Figure \ref{fig:calltopteinitialize}.
   \subsubsection{Kernels Utilizing \textsf{P2V}}
@@
 movabs 0x0,rcx	                187: R_X86_64_64	KERN_BASE
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val'}  }_{\rtv}$
-$\specline{ \sumapaces\rtv\delta  \ast \mathsf{entry\_val'.pfn}<<12 \in \Xi \setminus \{ \mathsf{entry} \} }_{\rtv}$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \} \cup \{\mathsf{nextpaddr}\},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val'}  }_{\rtv}$
+$\specline{ \sumapaces\rtv\delta  \ast \mathsf{nextpaddr}<<12 \in \Xi \setminus \{ \mathsf{entry} \} }_{\rtv}$
 $\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val\_present} \ast \ulcorner  \textsf{qfrac} = 1 \land (\textsf{entry\_present entry\_val'}) \urcorner }_{\rtv}$
@@
 add    rcx,rax
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} ,\mathsf{entry\_val'.pfn}<<12 \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val'}  }_{\rtv}$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} ,\mathsf{nextpaddr} \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val'}  }_{\rtv}$
 $\specline{ \sumapaces\rtv\delta  }_{\rtv}$
@@
 \end{figure}
-As shown in Figure \ref{fig:p2v}, \lstinline|rcx| holds the offset value \textsf{KERNBASE} which gives us the virtual address (\textsf{nextpaddr+KERNBASE}) to look-up in the identity map ($\Xi$) of the kernel invariant.
-Assuming the path for \textsf{pte\_initialize} is taken, we can obtain the virtual pte-pointsto for the physical address \textsf{nextpaddr} (Line 13 in Figure \ref{fig:p2v}).
+As shown in Figure \ref{fig:p2v}, \lstinline|rcx| holds the offset value \textsf{KERNBASE} which gives us the virtual address (\textsf{nextpaddr+KERNBASE}) to look-up in the identity map ($\Xi\setminus\{entry\}$) of the kernel invariant.
+Assuming the path for \textsf{pte\_initialize} is taken, we are assured that the freshly allocated physical address of the next page table (i.e.
+\textsf{nextpaddr}) is in the identity map ($\Xi$) -- from the post-condition of \textsf{pte\_initialize} in Figure \ref{fig:calltopteinitialize}, so that we can obtain the virtual pte-pointsto for the physical address \textsf{nextpaddr} from the  (Line 13 in Figure \ref{fig:p2v}).
 

### 2024-07-02 examples.tex
+

@@
 $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{\_}  \ast  \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{next} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry}  }_{\rtv}$
-$\specline{ \mathsf{entry} \in \Xi \ast \sumapaces\rtv\delta  }_{\rtv}$
+$\specline{ \ulcorner \Xi !! \mathsf{entry} = \mathsf{Some V} \urcorner \ast \sumapaces\rtv\delta  }_{\rtv}$
 mov     rdi, r8
@@
 callq  70 <pte_initialize>
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val}  }_{\rtv}$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \}\cup\{\mathsf{nextpaddr}\},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val}  }_{\rtv}$
 $\specline{\sumapaces\rtv\delta \ast \mathsf{nextpaddr} \in \Xi\setminus \{ \mathsf{entry} \} \cup \{ \mathsf{nextpaddr} \} }_{\rtv}$
 $\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \mathsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \land \lnot(\textsf{entry\_present entry\_val})  \urcorner }_{\rtv}$
-$\specline{ \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{nextpaddr} \ast \textsf{nextpaddr} \mapsto_{\textsf{p}} \textsf{w64\_zero}  \ast \texttt{rax} \mapsto_{\textsf{r}} \textsf{nextpaddr}  }_{\rtv}$
+$\specline{ \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{nextpaddr} \ast \ulcorner \Xi\setminus \{ \mathsf{entry} \}\cup\{\mathsf{nextpaddr}\} !!
+\mathsf{entry} = \mathsf{Some Z} \urcorner \ast \texttt{rax} \mapsto_{\textsf{r}} \textsf{nextpaddr}  }_{\rtv}$
+;;entry value updates
+...
 ;;entry->present = 1;
 ;;entry->pfn = next
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val'}  }_{\rtv}$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \}\cup\{\mathsf{nextpaddr}\},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val'}  }_{\rtv}$
 $\specline{\sumapaces\rtv\delta \ast \mathsf{nextpaddr} \in \Xi\setminus \{ \mathsf{entry} \} \cup \{ \mathsf{nextpaddr} \} }_{\rtv}$
-$\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \mathsf{entry entry\_val'} \ast \ulcorner  \textsf{qfrac} = 1 \land (\textsf{entry\_present entry\_val'})  \urcorner }_{\rtv}$
-$\specline{ \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{nextpaddr} \ast \textsf{nextpaddr} \mapsto_{\textsf{p}} \textsf{w64\_zero}  \ast \texttt{rax} \mapsto_{\textsf{r}} \textsf{nextpaddr}  }_{\rtv}$
+$\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \mathsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \land (\textsf{entry\_present entry\_val'})  \urcorner }_{\rtv}$
+$\specline{ \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{nextpaddr} \ast \ulcorner \Xi\setminus \{ \mathsf{entry} \}\cup\{\mathsf{nextpaddr}\} !!
+\mathsf{entry} = \mathsf{Some Z} \urcorner \ast \texttt{rax} \mapsto_{\textsf{r}} \textsf{nextpaddr}  }_{\rtv}$
 ;;}
@@
  \begin{array}{l}
-   \specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry} \textsf{\_} \ast \textsf{entry} \in \Xi }_{\rtv}  \\
+   \specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry} \textsf{\_} \ast \ulcorner \Xi !!
+\textsf{entry} = \textsf{Some V} \urcorner  }_{\rtv}  \\
 \sqsubseteq \\
@@
 Since we are using \textsf{pte\_initialize} for page-table address allocation, it also ensures us that newly allocated physicall address is in the identity mapping map $\Xi$ -- Line X in Figure \ref{fig:calltopteinitialize}.
+One might have expected to see a physical pointsto for the fresly page-table address  ($\mathsf{nextpaddr} \mapsto_{\mathsf{p}} \mathsf{w64\_0}$) after the call to \textsf{pte\_initialize} but we only see $\ulcorner \Xi !!
+\textsf{nextpaddr} = \textsf{ Some Z} \urcorner$ which exhibits itself with a different value (\textsf{Z}) than the identity mapping for the physical address \textsf{entry} has (\textsf{V}).
   \subsubsection{Kernels Utilizing \textsf{P2V}}
@@
 movabs 0x0,rcx	                187: R_X86_64_64	KERN_BASE
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \} \cup \{\mathsf{nextpaddr}\},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val'}  }_{\rtv}$
-$\specline{ \sumapaces\rtv\delta  \ast \mathsf{nextpaddr}<<12 \in \Xi \setminus \{ \mathsf{entry} \} }_{\rtv}$
-$\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val\_present} \ast \ulcorner  \textsf{qfrac} = 1 \land (\textsf{entry\_present entry\_val'}) \urcorner }_{\rtv}$
-$\specline{ \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{nextpaddr}  \ast \texttt{rax} \mapsto_{\textsf{r}} \textsf{nextpaddr}  }_{\rtv}$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \}\cup\{\mathsf{nextpaddr}\},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val'}  }_{\rtv}$
+$\specline{\sumapaces\rtv\delta \ast \mathsf{nextpaddr} \in \Xi\setminus \{ \mathsf{entry} \} \cup \{ \mathsf{nextpaddr} \} }_{\rtv}$
+$\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \mathsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \land (\textsf{entry\_present entry\_val'})  \urcorner }_{\rtv}$
+$\specline{ \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{nextpaddr} \ast \ulcorner \Xi\setminus \{ \mathsf{entry} \}\cup\{\mathsf{nextpaddr}\} !!
+\mathsf{entry} = \mathsf{Some Z} \urcorner \ast \texttt{rax} \mapsto_{\textsf{r}} \textsf{nextpaddr}  }_{\rtv}$
 add    rcx,rax
@@
 \textsf{nextpaddr}) is in the identity map ($\Xi$) -- from the post-condition of \textsf{pte\_initialize} in Figure \ref{fig:calltopteinitialize}, so that we can obtain the virtual pte-pointsto for the physical address \textsf{nextpaddr} from the (Line 13 in Figure \ref{fig:p2v}).
-
+\subsubsection{Walking Page-Table Tree: Calling \textsf{pte\_get\_nexttable} for Each Level}
+\label{wlkpgdir}
+Regarding the realization of 
 %\begin{figure}\footnotesize

### 2024-07-02 examples.tex
+

@@
 Since we are using \textsf{pte\_initialize} for page-table address allocation, it also ensures us that newly allocated physicall address is in the identity mapping map $\Xi$ -- Line X in Figure \ref{fig:calltopteinitialize}.
-One might have expected to see a physical pointsto for the fresly page-table address  ($\mathsf{nextpaddr} \mapsto_{\mathsf{p}} \mathsf{w64\_0}$) after the call to \textsf{pte\_initialize} but we only see $\ulcorner \Xi !!
-\textsf{nextpaddr} = \textsf{ Some Z} \urcorner$ which exhibits itself with a different value (\textsf{Z}) than the identity mapping for the physical address \textsf{entry} has (\textsf{V}).
+
+ There two important aspects on the specification after the call to \textsf{pte\_initialize}.
+First, one might have expected to see a physical pointsto for the fresly page-table address  ($\mathsf{nextpaddr} \mapsto_{\mathsf{p}} \mathsf{w64\_0}$) after the call to \textsf{pte\_initialize} but we do not have this physical pointsto resource.
+Second, we see the identity mapping for \textsf{nextpaddr} -- $\ulcorner \Xi !!
+\textsf{nextpaddr} = \textsf{ Some Z} \urcorner$ -- mapped to a different value (\textsf{Z}) than the identity mapping for the physical address \textsf{entry} is (\textsf{V}).
+The first aspect is important for not invalidating the soundness: \textsf{pte\_initialize} extends the identity map with the freshly allocated \textsf{nextpaddr}, and must not reveal the ownership of the physical pointsto in the post-condition (Line X in Figure \ref{}) because it would create a duplication of the full-ownership of the physical resource in the post-condition would potentially coexists with the one that can be obtained from the side-condition (*) inside the extended invariant in Figure \ref{fig:peraspaceinvariant_with_p2v_extension} with the capability token $\ulcorner \Xi !!
+\textsf{nextpaddr} = \textsf{ Some Z} \urcorner$.
+Although, the first aspect recovers the soundness while preserving the existence of the identity mapping for the fresly allocated physical page-table address, unlike  tas the  resource  inside  the  two subtle aspects are important f but quite but quite important for the soundness and
   \subsubsection{Kernels Utilizing \textsf{P2V}}

### 2024-07-02 examples.tex
+

@@
 Since we are using \textsf{pte\_initialize} for page-table address allocation, it also ensures us that newly allocated physicall address is in the identity mapping map $\Xi$ -- Line X in Figure \ref{fig:calltopteinitialize}.
-
- There two important aspects on the specification after the call to \textsf{pte\_initialize}.
+\begin{figure}\footnotesize
+\begin{lstlisting}[mathescape]
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \}\cup\{\mathsf{nextpaddr}\},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val'}  }_{\rtv}$
+$\specline{\sumapaces\rtv\delta \ast \mathsf{nextpaddr} \in \Xi\setminus \{ \mathsf{entry} \} \cup \{ \mathsf{nextpaddr} \} }_{\rtv}$
+$\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \mathsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \land (\textsf{entry\_present entry\_val'})  \urcorner }_{\rtv}$
+$\specline{ \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{nextpaddr} \ast \ulcorner \Xi\setminus \{ \mathsf{entry} \}\cup\{\mathsf{nextpaddr}\} !!
+\mathsf{entry} = \mathsf{Some Z} \urcorner \ast \texttt{rax} \mapsto_{\textsf{r}} \textsf{nextpaddr}  }_{\rtv}$
+;;void pte_initialize(pte_t **table) {
+...
+;;set up the stack    
+;;allocate a full zeroed page for 512 8-byte entries 
+;;pte_t *local = kalloc();
+callq  81 <pte_initialize+0x11>	7d: R_X86_64_PLT32	kalloc-0x4
+mov    rax,-0x10[rbp] 
+;;*table = local;
+mov    -0x10[rbp],rax
+mov    -0x8[rbp],rdi
+mov    rax,[rdi]
+;;return;
+;;..
+;;clean up the stack
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} ,\mathsf{nextpaddr} \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val'}  }_{\rtv}$
+$\specline{ \sumapaces\rtv\delta  }_{\rtv}$
+$\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val\_present} \ast \ulcorner  \textsf{qfrac} = 1 \land (\textsf{entry\_present entry\_val'}) \urcorner }_{\rtv}$
+$\specline{ \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{nextpaddr}  \ast \texttt{rax} \mapsto_{\textsf{r}} \textsf{nextpaddr}  }_{\rtv}$
+$\specline{ \texttt{nextpaddr+KERNBASE} \mapsto_{\textsf{vpte,qfracn}} \textsf{nextpaddr w64\_zero} }_{\rtv}$
+\end{lstlisting}
+\vspace{-1em}
+\caption{Allocating a physical page }
+\label{pteinitializespec}
+\end{figure}
+ There two important aspects on the specification after the call to \textsf{pte\_initialize} -- Specification in Figure \ref{fig:pteinitializespec}.
 First, one might have expected to see a physical pointsto for the fresly page-table address  ($\mathsf{nextpaddr} \mapsto_{\mathsf{p}} \mathsf{w64\_0}$) after the call to \textsf{pte\_initialize} but we do not have this physical pointsto resource.
@@
 \textsf{nextpaddr} = \textsf{ Some Z} \urcorner$.
-Although, the first aspect recovers the soundness while preserving the existence of the identity mapping for the fresly allocated physical page-table address, unlike  tas the  resource  inside  the  two subtle aspects are important f but quite but quite important for the soundness and
+Although, the first aspect recovers the soundness while preserving the knowledge of existence of the identity mapping for the fresly allocated physical page-table address, we still need the knowledge that the fresly allocated address references to a \textit{zeroed} page that an explicitly revealed physical pointsto would give -- $\mathsf{nextpaddr} \mapsto_{\mathsf{p}} \mathsf{w64\_0}$.
+Since a value of an identity mapping inside the invariant is existentially quantified (\textsf{val} in Figure \ref{fig:peraspaceinvariant_with_p2v_extension}), we would loose the knowledge of identity mapping with zeroed value.
+To carry this knowledge, we distinguish a random value of an identity mapping -- capability representing the existence of the random value (\textsf{V}) inside the invariant for the address \textsf{entry} at Line X in Figure \ref{} -- from the capability representing the zeroed identity-mapping value zeroed (\textsf{Z}).
+As we do for the address \textsf{nextpaddr}, when we use the capability with value \textsf{Z} for the identity-map lookup ($\Xi !!
+\textsf{nextpaddr} = \textsf{ Some Z}$), we know that the value existentially quantified inside the invariant is zeroed -- \textsf{val == w64\_0}.
   \subsubsection{Kernels Utilizing \textsf{P2V}}

### 2024-07-03 examples.tex
+

@@
   \begin{lstlisting}[mathescape]
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \}\cup\{\mathsf{nextpaddr}\},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val'}  }_{\rtv}$
-$\specline{\sumapaces\rtv\delta \ast \mathsf{nextpaddr} \in \Xi\setminus \{ \mathsf{entry} \} \cup \{ \mathsf{nextpaddr} \} }_{\rtv}$
-$\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \mathsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \land (\textsf{entry\_present entry\_val'})  \urcorner }_{\rtv}$
-$\specline{ \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{nextpaddr} \ast \ulcorner \Xi\setminus \{ \mathsf{entry} \}\cup\{\mathsf{nextpaddr}\} !!
-\mathsf{entry} = \mathsf{Some Z} \urcorner \ast \texttt{rax} \mapsto_{\textsf{r}} \textsf{nextpaddr}  }_{\rtv}$
+
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)   \ast \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{next} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val}  }_{\rtv}$
+$\specline{\texttt{rax} \mapsto_{\textsf{v}} \textsf{\_} \ast \sumapaces\rtv\delta }_{\rtv}$
 ;;void pte_initialize(pte_t **table) {
@@
 callq  81 <pte_initialize+0x11>	7d: R_X86_64_PLT32	kalloc-0x4
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)   \ast \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{next} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val}  }_{\rtv}$
+$\specline{\texttt{rax} \mapsto_{\textsf{v}} \textsf{nextpaddr} \ast \ulcorner \Xi !! \mathsf{nextpaddr} = None \urcorner \ast \sumapaces\rtv\delta }_{\rtv}$
+$\sqsubseteq$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\cup\{\mathsf{nextpaddr}\},m)   \ast \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{next} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val}  }_{\rtv}$
+$\specline{\texttt{rax} \mapsto_{\textsf{v}} \textsf{nextpaddr} \ast \ulcorner \Xi !! \mathsf{nextpaddr} = Z \urcorner \sumapaces\rtv\delta }_{\rtv}$  
 mov    rax,-0x10[rbp] 
@@
 mov    -0x10[rbp],rax
+
 mov    -0x8[rbp],rdi

### 2024-07-03 examples.tex
+

--- 2024-07-02 examples.tex
@@
   \begin{lstlisting}[mathescape]
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \}\cup\{\mathsf{nextpaddr}\},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val'}  }_{\rtv}$
-$\specline{\sumapaces\rtv\delta \ast \mathsf{nextpaddr} \in \Xi\setminus \{ \mathsf{entry} \} \cup \{ \mathsf{nextpaddr} \} }_{\rtv}$
-$\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \mathsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \land (\textsf{entry\_present entry\_val'})  \urcorner }_{\rtv}$
-$\specline{ \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{nextpaddr} \ast \ulcorner \Xi\setminus \{ \mathsf{entry} \}\cup\{\mathsf{nextpaddr}\} !!
-\mathsf{entry} = \mathsf{Some Z} \urcorner \ast \texttt{rax} \mapsto_{\textsf{r}} \textsf{nextpaddr}  }_{\rtv}$
+
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)   \ast \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{next} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{table\_addr}  }_{\rtv}$
+$\specline{\texttt{rax} \mapsto_{\textsf{v}} \textsf{\_} \ast \textsf{table\_addr} \mapsto_{\mathsf{v}} \mathsf{\_}\ast\sumapaces\rtv\delta }_{\rtv}$
 ;;void pte_initialize(pte_t **table) {
@@
 callq  81 <pte_initialize+0x11>	7d: R_X86_64_PLT32	kalloc-0x4
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)   \ast \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{next} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{table\_addr}  }_{\rtv}$
+$\specline{\texttt{rax} \mapsto_{\textsf{v}} \textsf{nextpaddr} \ast \ulcorner \Xi !!
+\mathsf{nextpaddr} = None \urcorner \ast \textsf{table\_addr} \mapsto_{\mathsf{v}} \mathsf{\_} \ast \sumapaces\rtv\delta }_{\rtv}$
+$\sqsubseteq$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\cup\{\mathsf{nextpaddr}\},m)   \ast \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{next} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val}  }_{\rtv}$
+$\specline{\texttt{rax} \mapsto_{\textsf{v}} \textsf{nextpaddr} \ast \ulcorner \Xi !!
+\mathsf{nextpaddr} = Z \urcorner \ast \textsf{table\_addr} \mapsto_{\mathsf{v}} \mathsf{\_} \ast \sumapaces\rtv\delta }_{\rtv}$
 mov    rax,-0x10[rbp] 
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\cup\{\mathsf{nextpaddr}\},m)   \ast \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{nextpaddr} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{table\_addr}  }_{\rtv}$
+$\specline{\texttt{rax} \mapsto_{\textsf{v}} \textsf{nextpaddr} \ast \ulcorner \Xi !!
+\mathsf{nextpaddr} = Z \urcorner \ast \textsf{table\_addr} \mapsto_{\mathsf{v}} \mathsf{\_}\ast \sumapaces\rtv\delta }_{\rtv}$
 ;;*table = local;
@@
 mov    rax,[rdi]
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\cup\{\mathsf{nextpaddr}\},m)   \ast \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{nextpaddr} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{table\_addr}  }_{\rtv}$
+$\specline{\texttt{rax} \mapsto_{\textsf{v}} \textsf{nextpaddr} \ast \ulcorner \Xi !!
+\mathsf{nextpaddr} = Z \urcorner \ast \textsf{table\_addr} \mapsto_{\mathsf{v}} \mathsf{\mathsf{nextpaddr}} \ast \sumapaces\rtv\delta }_{\rtv}$
 ;;return;
@@
 ;;clean up the stack
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} ,\mathsf{nextpaddr} \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val'}  }_{\rtv}$
-$\specline{ \sumapaces\rtv\delta  }_{\rtv}$
-$\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val\_present} \ast \ulcorner  \textsf{qfrac} = 1 \land (\textsf{entry\_present entry\_val'}) \urcorner }_{\rtv}$
-$\specline{ \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{nextpaddr}  \ast \texttt{rax} \mapsto_{\textsf{r}} \textsf{nextpaddr}  }_{\rtv}$
-$\specline{ \texttt{nextpaddr+KERNBASE} \mapsto_{\textsf{vpte,qfracn}} \textsf{nextpaddr w64\_zero} }_{\rtv}$
 \end{lstlisting}
@@
 \end{figure}
+\mytodo{give line numbers andrelated post here and pteinitcal }
  There two important aspects on the specification after the call to \textsf{pte\_initialize} -- Specification in Figure \ref{fig:pteinitializespec}.
@@
 \label{wlkpgdir}
-Regarding the realization of 
+Realizing page-table directory walk amounts to calling \textsf{pte\_get\_next\_table} for each level as shown in Figure \ref{walkpgdir}.
+Since there is not much interesting piece additional to the specification we discuss for \textsf{pte\_get\_next\_table} in Section \ref{}, in Figure \ref{}, we present the expected specification after the calls made for each level entry.
+At each specification Line (X,..) in Figure \ref{}, we see the accumulation of virtual-pte pointsto resources returned from \textsf{pte\_get\_next\_tabel} for each level (L4-L1), and finally return L1 level entry that is needed by \textsf{walkpgdir} at Line X in Figure \ref{}.
+\begin{figure}\footnotesize
+\begin{lstlisting}[mathescape]
+;;pte_t *walkpgdir_succ(pte_t *pml4, const void *va) {
+...
+;;set up the stack for root address and virtual address    
+;;pte_t *pml4_entry = &pml4[PML4EX(va)];
+mov    -0x8[rbp],rsi
+mov    -0x10[rbp],rdi
+shr    0x27,rdi
+and    0x1ff,rdi
+shl    0x3,rdi
+add    rdi,rsi
+mov    rsi,-0x18[rbp]
+;;pte_t *pdp = pte_get_next_table_succ(pml4_entry);
+mov    -0x18[rbp],rdi
+
+$\specline{}$
+callq  c0 <pte_get_next_table_succ>
+;;save the physical next table address in rax
+mov    rax,-0x20[rbp]
+$\specline{}$
+;;pte_t *pdp_entry = &pdp[PDPEX(va)];
+;;pte_t *pd = pte_get_next_table_succ(pdp_entry);
+$\specline{}$
+...
+;;pte_t *pd_entry = &pd[PDEX(va)];
+;;pte_t *pt = pte_get_next_table_succ(pd_entry);
+ $\specline{}$
+;;return &pt[PTEX(va)];
+...
+;; clean up the stack 
+\end{lstlisting}
+\vspace{-1em}
+\caption{Walking page-table directory via calls to \textsf{pte\_get\_next\_table} in Figure \ref{fig:calltopteinitialize}}
+\label{walkpgdir}
+\end{figure}
 %\begin{figure}\footnotesize

### 2024-07-03 examples.tex
+

@@
 mov    -0x18[rbp],rdi
-
-$\specline{}$
+$\specline{P_0 }_{\rtv}$
+...
 callq  c0 <pte_get_next_table_succ>
 ;;save the physical next table address in rax
-mov    rax,-0x20[rbp]
-$\specline{}$
+$\specline{P_1 \ast \texttt{pdp} \mapsto_{\textsf{vpte,qfracn}} \textsf{pdppaddr \_} }_{\rtv}$
 ;;pte_t *pdp_entry = &pdp[PDPEX(va)];
 ;;pte_t *pd = pte_get_next_table_succ(pdp_entry);
-$\specline{}$
+$\specline{P_2 \ast \texttt{pdp} \mapsto_{\textsf{vpte,qfracn}} \textsf{pdppaddr w64\_zero} \ast \texttt{pd} \mapsto_{\textsf{vpte,qfracn}} \textsf{pdpaddr \_} }_{\rtv}$
 ...
@@
 ;;pte_t *pt = pte_get_next_table_succ(pd_entry);
- $\specline{}$
+$\specline{P_3 \ast \texttt{pdp} \mapsto_{\textsf{vpte,qfracn}} \textsf{pdppaddr w64\_zero} \ast \texttt{pd} \mapsto_{\textsf{vpte,qfracn}} \textsf{pdpaddr \_} \ast \texttt{pt} \mapsto_{\textsf{vpte,qfracn}} \textsf{ptpaddr \_} }_{\rtv}$
+;;access and return L1 entry
 ;;return &pt[PTEX(va)];

### 2024-07-03 examples.tex
+

@@
 One of the key tasks of a page fault handler in a general-purpose OS kernel is
-to map new pages into an address space by writing into an existing page table.
-To do so, we first allocate a fresh page, then calculate the appropriate
-known-valid page table walks and update the appropriate L1 page table entry.
+to map new pages into an address space by writing into an existing page table (textsf{vaspace\_mappage(pte_t *pml4, void *va,uintptr_t pa, uint64_t perm )} in Figure \ref{fig:mapping_code}).
+To do so, with a given allocated a fresh page (\textsf{pa}), then calculate the appropriate
+known-valid page table walks (\textsf{pte\_walkpgdir} Line X in Figure \ref{fig:mapping_code})  and update the appropriate L1 page table entry (permissions and frame address Lines in Figure \ref{fig:mapping_code}).
 %\lstset{
@@
   \begin{lstlisting}[mathescape]
+;;complstatus_t vaspace_mappage(pte_t *pml4, void *va,uintptr_t pa, uint64_t perm ) {
+...
+;;setting up the stack      
+;;char *a; pte_t *pte;
+;;a = (char *)PGROUNDUP((uintptr_t)va);
  $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m)  \ast  \texttt{r14}\mapsto_{\textsf{r}} \_ \ast \texttt{rdi}\mapsto_{r} \vaddr \ast \texttt{rax}\mapsto_{\textsf{r}} \_ \ast \ulcorner \texttt{aligned va} \land  \theta \; !!\; \vaddr = \texttt{None}\urcorner}_{\rtv}$
+;;pte = walkpgdir_succ(pml4, a);
 call pte_walkpgdir ;; Returns the virtual address of the L1 entry in rax

### 2024-07-03 examples.tex
+

@@
 \end{lstlisting}
-  \caption{Specification and proof of distilled code for mapping a new page.}
+  \caption{Specification and proof of code for mapping a new page.}
 \label{fig:mapping_code}
@@
 The code uses a helper function \lstinline|pte\_walkpgdir| (discussed again in Section \ref{sec:traversing}).
-\lstinline|pte\_walkpgdir|'s postcondition contains physical pointsto assertions for a partial page table walk reaching the
-L1 entry (l1e), ensuring higher levels of the page table exist. 
-The postcondition also includes a virtual \emph{PTE}
-pointsto ($\mapsto_{\textsf{vpte}}$) allows access to the memory of the L1 entry
-via virtual address \lstinline|pte_pointsto|.
+\textsf{pte\_walkpgdir}'s postcondition contains virtual \emph{PTE} pointsto assertions ($\mapsto_{\textsf{vpte}}$) \todo{colin, these are no more plain physical pointstos}  both for ensuring partial page table walk reaching the
+L1 entry (l1e) with the justification of higher levels of the page table exist, and for allowing access to the memory of the L1 entry via virtual address.
 A PTE points-to is defined just like the normal virtual points-to of Figure \ref{fig:virtualpointstosharing}, except the physical address (here, \textsf{pa}) is explicit in the assertion 
@@
 \begin{array}{l}
-    \vaddr\mapsto_{\textsf{vpte}} \;\{\textsf{q}\} \; \paddr \; \vpage : \mathsf{vProp}~\Sigma \stackrel{\triangle}{=} 
+    \vaddr\mapsto_{\textsf{vpte,q}} \; \paddr \; \vpage : \mathsf{vProp}~\Sigma \stackrel{\triangle}{=} 
     \exists \delta\ldotp

### 2024-07-03 examples.tex
+

@@
 }
-
+\todo{ismail change the following spec base on the changes you have in the code}
 \begin{figure}\footnotesize
@@
 
-Line 12 allocates a fresh page-aligned, zero-initialized page  (at \textsf{fpaddr}),
-returning a pre-filled PTE entry in \textsf{rax} ($+3$ sets the lower 2 bits).
+%In the precondition, we see Line 12 allocates a fresh page-aligned, zero-initialized page  (at \textsf{fpaddr}),
+%returning a pre-filled PTE entry in \textsf{rax} ($+3$ sets the lower 2 bits).
 
@@
 table entry address ($\textsf{addr\_L1}(\textsf{va, l1e})$), to hold the freshly
-allocated physical page address (\textsf{fpaddr}) in Line 18.
-After this write, we have (Specification Line 21)
+allocated physical page address (\textsf{fpaddr}) in Line X.
+After this write, we have (Specification Line Y)
 \[\texttt{pte\_addr} \mapsto_{\texttt{vpte}}  \; \paddr\; \textsf{fpaddr}  \qquad\qquad (\textsf{by the PTE variant of Rule}~\TirNameStyle{WriteToVirtMemFromReg})\]
@@
 \entryo) \mapsto_{\mathsf{p}} \{q4\} \;\textsf{fpaddr}$) from
-$\mapsto_{\textsf{vpte}}$, after the split of the PTE points-to, we can
-connect L4-L2 points-to assertions with 
- $\texttt{fpaddr} \mapsto_{a} (\texttt{wzero 64})$ to obtain the
-complete physical table-walk assertion (L4\_L1\_PointsTo($\vaddr$ l4e l3e l2e
-fpaddr)) in Line 25.
+$\mapsto_{\textsf{vpte}}$, after the split of the PTE points-to, and, thanks to the post-condition of the page-table walk (Specification of \textsf{pte\_walkpgdir} in Figure \ref{}), we know the existence of L4-L2 mappings reaching to the L1 exist, so that we can add the last mapping $\texttt{fpaddr} \mapsto_{p} (\texttt{wzero 64})$ to obtain the complete table-walk.
+\todo{with the virtual-pte-pointsto assertions (L4\_L1\_PointsTo($\vaddr$ l4e l3e l2e
+fpaddr)) in Line 25}
 
@@
 In Figure \ref{fig:calltopteinitialize}, we present the critical piece of \textsf{get\_next\_table}, and construct the specification of it
-Within \textsf{get\_next\_table} first checks if the entry whose address is passed as a parameter ($\textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry}$).
-To be able to do so, it needs to load the entry value (a 64-bits word divided into bit-fields for corresponding fileds shown in Section {}) and check that \textit{presence} bit of the entry is set or not.
+Within \textsf{get\_next\_table}, first, we check if the entry whose address is passed as a parameter ($\textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry}$) is present or not.
+To be able to do so, it needs to load the entry value (a 64-bit word divided into bit-fields for corresponding fileds shown in Section {}) and check that \textit{presence} bit of the entry is set or not.
 %wshiftll (wshiftll (natToWord 64 entry) (WordImpl.concat (WordImpl.zero 56) (WordImpl.from_nat 8 12 ^& WordImpl.concat (WordImpl.zero 2) WO~1~1~1~1~1~1)) ^& constf)
@@
 $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{\_}  \ast  \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{next} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry}  }_{\rtv}$
-$\specline{ \ulcorner \Xi !! \mathsf{entry} = \mathsf{Some V} \urcorner \ast \sumapaces\rtv\delta  }_{\rtv}$
+$\specline{ \ulcorner \Xi \;!!\; \mathsf{entry} = \mathsf{Some V} \urcorner \ast \sumapaces\rtv\delta  }_{\rtv}$
 mov     rdi, r8
@@
 $\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \mathsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \land \lnot(\textsf{entry\_present entry\_val})  \urcorner }_{\rtv}$
-$\specline{ \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{nextpaddr} \ast \ulcorner \Xi\setminus \{ \mathsf{entry} \}\cup\{\mathsf{nextpaddr}\} !!
-\mathsf{entry} = \mathsf{Some Z} \urcorner \ast \texttt{rax} \mapsto_{\textsf{r}} \textsf{nextpaddr}  }_{\rtv}$
+$\specline{ \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{nextpaddr} \ast \ulcorner \Xi\setminus \{ \mathsf{entry} \}\cup\{\mathsf{nextpaddr}\} \;!!\; \mathsf{entry} = \mathsf{Some Z} \urcorner \ast \texttt{rax} \mapsto_{\textsf{r}} \textsf{nextpaddr}  }_{\rtv}$
 ;;entry value updates
@@
 $\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \mathsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \land (\textsf{entry\_present entry\_val'})  \urcorner }_{\rtv}$
-$\specline{ \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{nextpaddr} \ast \ulcorner \Xi\setminus \{ \mathsf{entry} \}\cup\{\mathsf{nextpaddr}\} !!
-\mathsf{entry} = \mathsf{Some Z} \urcorner \ast \texttt{rax} \mapsto_{\textsf{r}} \textsf{nextpaddr}  }_{\rtv}$
+$\specline{ \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{nextpaddr} \ast \ulcorner \Xi\setminus \{ \mathsf{entry} \}\cup\{\mathsf{nextpaddr}\} \;!!\; \mathsf{entry} = \mathsf{Some Z} \urcorner \ast \texttt{rax} \mapsto_{\textsf{r}} \textsf{nextpaddr}  }_{\rtv}$
 ;;}
@@
 
-As we see in Figure \ref{fig:calltopteinitialize}, we have the physical address (\textsf{entry}) of the page table entry kept in the register $\texttt{r8}$, but not the virtual memory mapping to load the entry value referenced by the physical address \textsf{entry} (Line 7 in Figure \ref{fig:calltopteinitialize}).
-We need to construct this virtual mapping for \textsf{entry}.
+As we see in Figure \ref{fig:calltopteinitialize}, we have the physical address (\textsf{entry}) of the page table entry kept in the register $\texttt{r8}$, but not the virtual memory mapping to load the entry value referenced by the physical address \textsf{entry} (Line 13 in Figure \ref{fig:calltopteinitialize}).
+We need to construct this virtual mapping for the physical address \textsf{entry}.
 \subsubsection{Identity Mappings}
@@
   \bigast{(\vaddr, \textsf{paddr})\in \ptablestore}{\exists\;(\textsf{l4e, l3e, l2e, l1e, paddr})\ldotp \textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\vaddr\textsf{, l4e, l3e, l2e, l1e, paddr})} \\
-  \bigast{(\paddr) \in \Xi}{\exists\; (\textsf{qfrac, q, val,}\vaddr) \ldotp \ulcorner \vaddr = \paddr + \textsf{KERNBASE} \urcorner \ast \ulcorner \textsf{qfrac} = 1 \leftrightarrow \; \lnot\textsf{entry\_present }(\vale) \urcorner \\
-                           \qquad\qquad\qquad\qquad\qquad\qquad          \underbrace{\sumwalkabs\vaddr\qfrac\paddr }_\text{Ghost translation} \ast 
+  \bigast{(\paddr) \in \Xi}{\exists\; (\textsf{qfrac, q, val,}\vaddr) \ldotp \ulcorner \vaddr = \paddr + \textsf{KERNBASE} \urcorner \ast \ulcorner \textsf{qfrac} = 1 \leftrightarrow \; \lnot\textsf{entry\_present }(\vale) \urcorner \ast \underbrace{\sumwalkabs\vaddr\qfrac\paddr }_\text{Ghost translation} \ast
 % \qquad\qquad\qquad\qquad\qquad\qquad\qquad\qquad    

### 2024-07-03 examples.tex
+

@@
 One of the key tasks of a page fault handler in a general-purpose OS kernel is
-to map new pages into an address space by writing into an existing page table (textsf{vaspace\_mappage(pte_t *pml4, void *va,uintptr_t pa, uint64_t perm )} in Figure \ref{fig:mapping_code}).
+to map new pages into an address space by writing into an existing page table (\textsf{vaspace\_mappage(pte\_t *pml4, void *va,uintptr\_t pa, uint64\_t perm )} in Figure \ref{fig:mapping_code}).
 To do so, with a given allocated a fresh page (\textsf{pa}), then calculate the appropriate

### 2024-07-03 examples.tex
+

@@
   \bigast{(\vaddr, \textsf{paddr})\in \ptablestore}{\exists\;(\textsf{l4e, l3e, l2e, l1e, paddr})\ldotp \textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\vaddr\textsf{, l4e, l3e, l2e, l1e, paddr})} \\
-  \bigast{(\paddr) \in \Xi}{\exists\; (\textsf{qfrac, q, val,}\vaddr) \ldotp \ulcorner \vaddr = \paddr + \textsf{KERNBASE} \urcorner \ast \ulcorner \textsf{qfrac} = 1 \leftrightarrow \; \lnot\textsf{entry\_present }(\vale) \urcorner \ast \underbrace{\sumwalkabs\vaddr\qfrac\paddr }_\text{Ghost translation} \ast
+  \bigast{(\paddr,\mathsf{T}) \in \Xi}{\exists\; (\textsf{qfrac, q, val,}\vaddr) \ldotp \ulcorner \vaddr = \paddr + \textsf{KERNBASE} \urcorner \ast \ulcorner \textsf{qfrac} = 1 \leftrightarrow \; \lnot\textsf{entry\_present }(\vale) \urcorner \ast
+    \ulcorner \mathsf{T} = \mathsf{Z} \leftrightarrow \mathsf{val} = \mathsf{w0\_64}\urcorner \; \ast\\
+   \qquad\qquad\qquad\qquad\qquad\qquad\qquad\qquad   \underbrace{\sumwalkabs\vaddr\qfrac\paddr }_\text{Ghost translation} \ast 
 % \qquad\qquad\qquad\qquad\qquad\qquad\qquad\qquad    

### 2024-07-03 examples.tex
+

@@
 
-Beyond a frame, the precondition before Line 7 requires the current address space invariant, and knowledge that \textsf{entry} is in the identity-map ($\Xi$) -- $\textsf{paddr} \in \Xi$.
-By knowing this, we can also refer to the facts in the invariant that are related to a physical page-table addresses is in a range dedicated to page-table entries (as this physical-to-virtual conversion is only guaranteed for page table memory), and check that the addition of the offset (i.e.
-\textsf{KERNBASE}) to a physical page-table address do not overflow (in \lstinline|rax|).
-
-The latter uses an additional ghost set added to the per-address-space
+Beyond a frame, the precondition before Line 13 requires the current address space invariant, and knowledge that \textsf{entry} is mapped to a random entry value ($\Xi \;!!\; \mathsf{entry} = Some V$ in Figure \ref{fig:peraspaceinvariant_with_p2v_extension}).
+By owning this capability, we can also refer to the facts in the invariant that there exists an virtual address from a range set dedicated to page-table entries, as this physical-to-virtual conversion is only guaranteed for page table memory, and obtained by offsetting the physical address ($\textsf{va} = \textsf{pa} + \textsf{KERNBASE}$.
+We see that to realize identity mappings, an additional ghost map for keeping identity mappings ($\Xi$) added to the per-address-space
 invariant per Figure \ref{fig:peraspaceinvariant_with_p2v_extension}. The extended invariant owns, for each memory location dedicated to page tables, some fraction of
@@
 
-Concretely speaking, before Line 7 in Figure \ref{fig:calltopteinitialize}, to read the value referenced by physical address \textsf{entry}, first, we open our the addresss-space invariant ($\mathcal{I}\textsf{ASpace}_{\textsf{id}}(\ptablestore,\Xi,m)$), and then obtain the virtual mapping out of the identity mapping $\Xi$ with the side condition (*)
+Concretely speaking, before Line 13 in Figure \ref{fig:calltopteinitialize}, to read the value referenced by physical address \textsf{entry}, first, we open our the addresss-space invariant ($\mathcal{I}\textsf{ASpace}_{\textsf{id}}(\ptablestore,\Xi,m)$), and then obtain the virtual mapping out of the identity mapping $\Xi$ with the side condition (*)
 \begin{equation*}

### 2024-07-03 examples.tex
+

@@
 invariant per Figure \ref{fig:peraspaceinvariant_with_p2v_extension}. The extended invariant owns, for each memory location dedicated to page tables, some fraction of
-that physical memory, and also guarantees the offset is the correct calculation.
+that physical memory, for the sake of constructing the virtual-mapping, also the fractional ownership abstracting translation (i.e.
+ghost translation in Figure \ref{fig:peraspaceinvariant_with_p2v_extension}), and guarantees the offset is the correct calculation.
 Subtly, the operation ensures the relevant table entry is readable, but the exact portion of ownership returned must be determined by inspecting the valid bit
@@
 
-Concretely speaking, before Line 13 in Figure \ref{fig:calltopteinitialize}, to read the value referenced by physical address \textsf{entry}, first, we open our the addresss-space invariant ($\mathcal{I}\textsf{ASpace}_{\textsf{id}}(\ptablestore,\Xi,m)$), and then obtain the virtual mapping out of the identity mapping $\Xi$ with the side condition (*)
+Concretely speaking, before Line 13 in Figure \ref{fig:calltopteinitialize}, to read the value referenced by physical address \textsf{entry}, first, we open our the addresss-space invariant ($\mathcal{I}\textsf{ASpace}_{\textsf{id}}(\ptablestore,\Xi,m)$), and then obtain the virtual mapping out of the identity mapping $\Xi$
+\[ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \]
+via using the capability $\Xi \;!!\; \textsf{entry} = Some \textsf{ V}$ under the side condition (*)
 \begin{equation*}
@@
 \end{equation*}
-assuring that looking up identity mappings are safe.
-The subtle justification relies on the fact that the full ownership asserts the non-presence of the entry, accessing to which requires allocation a new physical page and having the presence bit set.
-Acquiring the identity mapping for \textsf{entry} needed before Line 7 in Figure \ref{fig:calltopteinitialize}, we need the knowledge $\textsf{entry} \in \Xi$ to peel-out the fractional token representing the ghost address translation from the offsetted virtual address (\textsf{entry+KERNBASE}) to the physical page address (\textsf{entry}) and physical pointsto relation from the physical address \textsf{entry} to an existentially quantified value (\textsf{entry\_val}) and the side condition (*) we explained above.
-In in Figure \ref{fig:lookingupidpampping}, we see that all the logical resources needed to construct a virtual-pte pointsto
-\[ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \]
-to execute Line 7 in \ref{fig:calltopteinitialize} that can either be mutated or not based on the side condition (*).
+assuring that looking up an identity mapping is safe under the subtle justification which equates the fact of full ownership to the non-presence of the entry, accessing to which requires allocation a new physical page and having the presence bit set: \textsf{qfrac = 1} is not known for the address \textsf{entry} before the line before Line 13 in Figure \ref{fig:calltopteinitialize} (which makes the entry value accessible not mutable), and can only be known when the entry presence is inspected at Line 19.
  \begin{figure*}

### 2024-07-03 examples.tex
+

@@
 After obtaining the identity mapping for \textsf{entry}, we are able to load the \textsf{entry\_val} into \textsf{rdi}, and check the presence bit via $and \;0x1,rdi$.
- Accessing the presence bit and checking the value allow (Lines 7-13 in Figure \ref{fig:calltopteinitialize}) us to exploit the fact in the side condition (*) in such a way that the condition ($\textsf{entry}\mapsto\textsf{present}$ determines whether we take the non-presence path (Lines 12-19 in Figure \ref{fig:calltopteinitialize}) with full ownership ($\textsf{qfrac} = 1$) to set the entry with proper values for the bitfields, e.g., frame showing freshly allocated page and presence bit.
+ Accessing the presence bit and checking the value (Line 19 in Figure \ref{fig:calltopteinitialize}) allow  us to exploit the fact in the side condition (*) in such a way that the condition ($\textsf{entry}\mapsto\textsf{present}$ determines whether we take the non-presence path (Lines 19-44 in Figure \ref{fig:calltopteinitialize}) with full ownership ($\textsf{qfrac} = 1$) to set the entry with proper values for the bitfields, e.g., frame showing freshly allocated page and presence bit.
  In case of absence, \textsf{pte\_initialize} is called for a physical-zeroed page (utilizing the page-allocator's \textsf{kalloc} -- currently the only axiom as it is not part of \textsf{pte} (virtual-memory-management) module) to be backing the physical frame address of the entry -- Specification Line 35 in Figure \ref{fig:calltopteinitialize}.

### 2024-07-03 examples.tex
+

@@
 We need to construct this virtual mapping for the physical address \textsf{entry}.
+\newcommand{\ventry}[1]{\mathsf{#1}+\mathsf{KERNBASE}}
+\newcommand{\entry}[1]{\mathsf{#1}}
 \subsubsection{Identity Mappings}
@@
 
-Concretely speaking, before Line 13 in Figure \ref{fig:calltopteinitialize}, to read the value referenced by physical address \textsf{entry}, first, we open our the addresss-space invariant ($\mathcal{I}\textsf{ASpace}_{\textsf{id}}(\ptablestore,\Xi,m)$), and then obtain the virtual mapping out of the identity mapping $\Xi$
+Concretely speaking, when we focus on the specfication of Line 13 in Figure \ref{fig:calltopteinitialize} in Figure \ref{fig:lookingupidpampping}, to read the value referenced by physical address \textsf{entry}, as shown in the first specification Line in Figure \ref{fig:lookingupidpampping}, we open our the addresss-space invariant ($\mathcal{I}\textsf{ASpace}_{\textsf{id}}(\ptablestore,\Xi,m)$), and then obtain the fractional ownership of the ghost translation and physical pointsto
+\[\sumwalkabs\ventry\qfrac\entry\ \ast \mathsf{entry}\mapsto_{\mathsf{p,qfrac}}\; \vale \]
+via using the capability $\Xi \;!!\; \textsf{entry} = Some \textsf{ V}$ and left the invariant as in the state $\mathcal{I}\textsf{ASpace}_{\textsf{id}}(\ptablestore,\Xi\setminus\{\mathsf{entry}\},m)$ , and construct the virtual mapping with the resources peeled-out of the identity mapping $\Xi$
 \[ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \]
-via using the capability $\Xi \;!!\; \textsf{entry} = Some \textsf{ V}$ under the side condition (*)
+under the side condition (*)
 \begin{equation*}
@@
 \end{equation*}
-assuring that looking up an identity mapping is safe under the subtle justification which equates the fact of full ownership to the non-presence of the entry, accessing to which requires allocation a new physical page and having the presence bit set: \textsf{qfrac = 1} is not known for the address \textsf{entry} before the line before Line 13 in Figure \ref{fig:calltopteinitialize} (which makes the entry value accessible not mutable), and can only be known when the entry presence is inspected at Line 19.
+assuring that looking the identity mapping for \textsf{entry} is safe under the subtle justification which equates the full ownership to the non/presence of the entry which can only be known when investigated ad Line 19in Figure \ref{fig:calltopteinitialize}.
  \begin{figure*}

### 2024-07-03 examples.tex
+

@@
 %}
+\newcommand{\sumwalkabsent}{
+  \ownGhost\gammaPred{\authfrag{\singletonMap{\texttt{entry+KERNBASE}}{(\textsf{qfrac}, \textsf{entry})}}}
+}
 
+\newcommand{\ventry}{\texttt{entry + KERNBASE}}
+\newcommand{\entry}{\texttt{entry}}
 \newcommand{\qfraczero}{\textsf{qfrac}}
@@
 We need to construct this virtual mapping for the physical address \textsf{entry}.
-\newcommand{\ventry}[1]{\mathsf{#1}+\mathsf{KERNBASE}}
-\newcommand{\entry}[1]{\mathsf{#1}}
 \subsubsection{Identity Mappings}
@@
 
-Beyond a frame, the precondition before Line 13 requires the current address space invariant, and knowledge that \textsf{entry} is mapped to a random entry value ($\Xi \;!!\; \mathsf{entry} = Some V$ in Figure \ref{fig:peraspaceinvariant_with_p2v_extension}).
+Beyond a frame, the precondition before Line 13 requires the current address space invariant, and knowledge that \textsf{entry} is mapped to a random entry value ($\Xi \;!!\; \mathsf{entry} = \mathsf{Some V}$ in Figure \ref{fig:peraspaceinvariant_with_p2v_extension}).
 By owning this capability, we can also refer to the facts in the invariant that there exists an virtual address from a range set dedicated to page-table entries, as this physical-to-virtual conversion is only guaranteed for page table memory, and obtained by offsetting the physical address ($\textsf{va} = \textsf{pa} + \textsf{KERNBASE}$.
@@
 Concretely speaking, when we focus on the specfication of Line 13 in Figure \ref{fig:calltopteinitialize} in Figure \ref{fig:lookingupidpampping}, to read the value referenced by physical address \textsf{entry}, as shown in the first specification Line in Figure \ref{fig:lookingupidpampping}, we open our the addresss-space invariant ($\mathcal{I}\textsf{ASpace}_{\textsf{id}}(\ptablestore,\Xi,m)$), and then obtain the fractional ownership of the ghost translation and physical pointsto
-\[\sumwalkabs\ventry\qfrac\entry\ \ast \mathsf{entry}\mapsto_{\mathsf{p,qfrac}}\; \vale \]
+\[\sumwalkabsent \ast \mathsf{entry}\mapsto_{\mathsf{p,qfrac}}\; \vale \]
 via using the capability $\Xi \;!!\; \textsf{entry} = Some \textsf{ V}$ and left the invariant as in the state $\mathcal{I}\textsf{ASpace}_{\textsf{id}}(\ptablestore,\Xi\setminus\{\mathsf{entry}\},m)$ , and construct the virtual mapping with the resources peeled-out of the identity mapping $\Xi$
@@
  \subsubsection{Ensuring Presence of an Entry}
-After obtaining the identity mapping for \textsf{entry}, we are able to load the \textsf{entry\_val} into \textsf{rdi}, and check the presence bit via $and \;0x1,rdi$.
- Accessing the presence bit and checking the value (Line 19 in Figure \ref{fig:calltopteinitialize}) allow  us to exploit the fact in the side condition (*) in such a way that the condition ($\textsf{entry}\mapsto\textsf{present}$ determines whether we take the non-presence path (Lines 19-44 in Figure \ref{fig:calltopteinitialize}) with full ownership ($\textsf{qfrac} = 1$) to set the entry with proper values for the bitfields, e.g., frame showing freshly allocated page and presence bit.
+ After obtaining the identity mapping for \textsf{entry}, we are able to load the \textsf{entry\_val} into \textsf{rdi}, and check the presence bit via
+ \[ \mathsf{and} \;0x1,rdi\] Line 17 in Figure \ref{fig:calltopteinitialize}.
+ Accessing the presence bit and checking the value allow  us to exploit the fact in the side condition (*) in such a way that the condition ($\textsf{entry}\mapsto\textsf{present}$ determines whether we take the non-presence path (Lines 19-44 in Figure \ref{fig:calltopteinitialize}) with full ownership ($\textsf{qfrac} = 1$) to set the entry with proper values for the bitfields, e.g., frame showing freshly allocated page and presence bit.
  In case of absence, \textsf{pte\_initialize} is called for a physical-zeroed page (utilizing the page-allocator's \textsf{kalloc} -- currently the only axiom as it is not part of \textsf{pte} (virtual-memory-management) module) to be backing the physical frame address of the entry -- Specification Line 35 in Figure \ref{fig:calltopteinitialize}.

### 2024-07-04 examples.tex
+

@@
 
-Beyond a frame, the precondition before Line 13 requires the current address space invariant, and knowledge that \textsf{entry} is mapped to a random entry value ($\Xi \;!!\; \mathsf{entry} = \mathsf{Some V}$ in Figure \ref{fig:peraspaceinvariant_with_p2v_extension}).
+Beyond a frame, the precondition before Line 13 requires the current address space invariant, and knowledge that \textsf{entry} is mapped to a random entry value ($\Xi \;!!\; \mathsf{entry} = \textsf{Some V}$ in Figure \ref{fig:peraspaceinvariant_with_p2v_extension}).
 By owning this capability, we can also refer to the facts in the invariant that there exists an virtual address from a range set dedicated to page-table entries, as this physical-to-virtual conversion is only guaranteed for page table memory, and obtained by offsetting the physical address ($\textsf{va} = \textsf{pa} + \textsf{KERNBASE}$.
@@
 \[\sumwalkabsent \ast \mathsf{entry}\mapsto_{\mathsf{p,qfrac}}\; \vale \]
-via using the capability $\Xi \;!!\; \textsf{entry} = Some \textsf{ V}$ and left the invariant as in the state $\mathcal{I}\textsf{ASpace}_{\textsf{id}}(\ptablestore,\Xi\setminus\{\mathsf{entry}\},m)$ , and construct the virtual mapping with the resources peeled-out of the identity mapping $\Xi$
+via using the capability $\Xi \;!!\; \textsf{entry} = \textsf{Some V}$ and left the invariant as in the state $\mathcal{I}\textsf{ASpace}_{\textsf{id}}(\ptablestore,\Xi\setminus\{\mathsf{entry}\},m)$ , and construct the virtual mapping with the resources peeled-out of the identity mapping $\Xi$
 \[ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \]
@@
  \subsubsection{Ensuring Presence of an Entry}
- After obtaining the identity mapping for \textsf{entry}, we are able to load the \textsf{entry\_val} into \textsf{rdi}, and check the presence bit via
- \[ \mathsf{and} \;0x1,rdi\] Line 17 in Figure \ref{fig:calltopteinitialize}.
- Accessing the presence bit and checking the value allow  us to exploit the fact in the side condition (*) in such a way that the condition ($\textsf{entry}\mapsto\textsf{present}$ determines whether we take the non-presence path (Lines 19-44 in Figure \ref{fig:calltopteinitialize}) with full ownership ($\textsf{qfrac} = 1$) to set the entry with proper values for the bitfields, e.g., frame showing freshly allocated page and presence bit.
+ After obtaining the identity mapping for \textsf{entry}, we are able to load the \textsf{entry\_val} into \textsf{rdi}, and check the presence bit through
+ \[ \begin{array}{l}
+   \mathsf{and }\; 0x1, \;rdi \\
+   \mathsf{mov } \;\mathsf{rdi}, \; \mathsf{rax} \\
+   \mathsf{cmp } \; 0x0, \;\mathsf{rax}
+ \end{array}
+ \] Lines (17-20) in Figure \ref{fig:calltopteinitialize}.
+ Accessing the presence bit and checking the value allow  us to exploit the fact in the side condition (*) in such a way that we either take the allocation path in the non-presence case (Lines 19-44 in Figure \ref{fig:calltopteinitialize}) with the knowledge of full ownership ($\textsf{qfrac} = 1$) to set the entry with proper values for the bitfields, most imporantly frame referencing freshly allocated page and presence bit, or, take non-allocation path and jump to Line 45 with fractional permission for looking up the entry values.
+
  In case of absence, \textsf{pte\_initialize} is called for a physical-zeroed page (utilizing the page-allocator's \textsf{kalloc} -- currently the only axiom as it is not part of \textsf{pte} (virtual-memory-management) module) to be backing the physical frame address of the entry -- Specification Line 35 in Figure \ref{fig:calltopteinitialize}.

### 2024-07-04 examples.tex
+

@@
 
- In case of absence, \textsf{pte\_initialize} is called for a physical-zeroed page (utilizing the page-allocator's \textsf{kalloc} -- currently the only axiom as it is not part of \textsf{pte} (virtual-memory-management) module) to be backing the physical frame address of the entry -- Specification Line 35 in Figure \ref{fig:calltopteinitialize}.
-Since we are using \textsf{pte\_initialize} for page-table address allocation, it also ensures us that newly allocated physicall address is in the identity mapping map $\Xi$ -- Line X in Figure \ref{fig:calltopteinitialize}.
+ In case of absence, \textsf{pte\_initialize} is called (Line 31 in Figure \ref{fig:calltopteinitialize}) for a physical-zeroed page (utilizing the page-allocator's \textsf{kalloc} -- currently the only axiomatized call in the the proof of \textsf{pte\_initialize} (Line 9 in Figure \ref{pteinitializespec}, it is not part of \textsf{pte} (virtual-memory-management) module).
 \begin{figure}\footnotesize
@@
 \end{figure}
-\mytodo{give line numbers andrelated post here and pteinitcal }
- There two important aspects on the specification after the call to \textsf{pte\_initialize} -- Specification in Figure \ref{fig:pteinitializespec}.
+
+Since we are using \textsf{pte\_initialize} for page-table address allocation, we must relate this newly allocated physical address to the identity mapping map $\Xi$ -- Specification Lines 33 and 35 in Figure \ref{fig:calltopteinitialize}.
+There are two important aspects on relating a freshly allocated page-table physical address to the identity-mapping which we focus in more detail separately in the specification of \textsf{pte\_initialize} in Figure \ref{fig:pteinitializespec}.
 First, one might have expected to see a physical pointsto for the fresly page-table address  ($\mathsf{nextpaddr} \mapsto_{\mathsf{p}} \mathsf{w64\_0}$) after the call to \textsf{pte\_initialize} but we do not have this physical pointsto resource.
-Second, we see the identity mapping for \textsf{nextpaddr} -- $\ulcorner \Xi !!
-\textsf{nextpaddr} = \textsf{ Some Z} \urcorner$ -- mapped to a different value (\textsf{Z}) than the identity mapping for the physical address \textsf{entry} is (\textsf{V}).
-The first aspect is important for not invalidating the soundness: \textsf{pte\_initialize} extends the identity map with the freshly allocated \textsf{nextpaddr}, and must not reveal the ownership of the physical pointsto in the post-condition (Line X in Figure \ref{}) because it would create a duplication of the full-ownership of the physical resource in the post-condition would potentially coexists with the one that can be obtained from the side-condition (*) inside the extended invariant in Figure \ref{fig:peraspaceinvariant_with_p2v_extension} with the capability token $\ulcorner \Xi !!
+Second, we see the identity mapping for \textsf{nextpaddr}, first not appearing in the identity map (Specification Line 11 in Figure \ref{fig:pteinitializespec}), then as inserted into the identity-map with a value \textsf{Z} ($\Xi \cup \{\mathsf{nextpaddr}\}$) together with the fragmental capability  -- $\ulcorner \Xi \; !!\; \textsf{nextpaddr} = \textsf{ Some Z} \urcorner$ -- at Lines 13 and 14 in Figure \ref{fig:pteinitializespec} .
+The value type \textsf{Z}, unlike the value type \textsf{V} denoting the random value in Line 8 in Figure \ref{fig:calltopteinitialize}, denotes \textsf{zeroed} value.
+
+The first aspect is important for not invalidating the soundness: \textsf{pte\_initialize} extends the identity map with the freshly allocated \textsf{nextpaddr}, and must not reveal the ownership of the physical pointsto in the post-condition (Line 33 and 35 in Figure \ref{fig:calltopteinitialize}, Lines 22-23 in Figure \ref{fig:pteinitializespec}) because it would create unsound coexistence of the full-ownership of the physical pointsto resource (a hypothetical $\mathsf{nextpaddr} \mapsto_{\mathsf{p,1}} w0_64$) and fragmental (potentially full based on the side-condition *) physical pointsto resource that can be obtained from the identity map inside the extended invariant in Figure \ref{fig:peraspaceinvariant_with_p2v_extension} with the capability token $\ulcorner \Xi !!
 \textsf{nextpaddr} = \textsf{ Some Z} \urcorner$.
-Although, the first aspect recovers the soundness while preserving the knowledge of existence of the identity mapping for the fresly allocated physical page-table address, we still need the knowledge that the fresly allocated address references to a \textit{zeroed} page that an explicitly revealed physical pointsto would give -- $\mathsf{nextpaddr} \mapsto_{\mathsf{p}} \mathsf{w64\_0}$.
-Since a value of an identity mapping inside the invariant is existentially quantified (\textsf{val} in Figure \ref{fig:peraspaceinvariant_with_p2v_extension}), we would loose the knowledge of identity mapping with zeroed value.
-To carry this knowledge, we distinguish a random value of an identity mapping -- capability representing the existence of the random value (\textsf{V}) inside the invariant for the address \textsf{entry} at Line X in Figure \ref{} -- from the capability representing the zeroed identity-mapping value zeroed (\textsf{Z}).
-As we do for the address \textsf{nextpaddr}, when we use the capability with value \textsf{Z} for the identity-map lookup ($\Xi !!
-\textsf{nextpaddr} = \textsf{ Some Z}$), we know that the value existentially quantified inside the invariant is zeroed -- \textsf{val == w64\_0}.
+Although, the first aspect -- not explicitly revealing the physical pointsto resource for the freshly allocated page-table address but providing access through the identity map inside the invariant with access-capability tokens -- recovers the soundness argument expalined, we still need the knowledge that the fresly allocated address references to a \textit{zeroed} page.
+The invariant preserves this information through the constraint on relating the type (\textsf{T}) of mapped data -- \textsf{T = Z} for zeroed or \textsf{T = V} for random page value -- to the value of existentially quantified value -- \textsf{val} -- in Figure \ref{fig:peraspaceinvariant_with_p2v_extension}.
+So, for accessing the physical pointsto resource to the zeroed value referenced by \textsf{nextpaddr}, we need to have the fragmental ownership of the identity-map $\Xi !!
+\textsf{nextpaddr} = \textsf{ Some Z}$ which ensures and carries the \textit{being-zeroed} knowledge when we open and access the the existentially quantified value inside the invariant.
   \subsubsection{Kernels Utilizing \textsf{P2V}}

### 2024-07-04 examples.tex
+

@@
 call pte_walkpgdir ;; Returns the virtual address of the L1 entry in rax
+
+;;char *a; pte_t *pte;
+;;a = (char *)PGROUNDUP((uintptr_t)va);
+mov    -0x10[rbp],rcx
+add    0x1000,rcx
+sub    0x1,rcx
+and    0xfffffffffffff000,%rcx
+mov    rcx,-0x28[rbp]
+
+;;pte = walkpgdir_succ(pml4, a);
+mov    -0x8[rbp],rdi
+mov    -0x28[rbp],rsi
+mov    0x0,rax
+callq  291 <vaspace_mappage+0x41>	28d: R_X86_64_PLT32	walkpgdir_succ-0x4
+xor    r8,r8
+movslq eax,rcx
+mov    rcx,-0x30[rbp]
+
+;;pte->writable     = PTE_WRITEABLE_SET(perm);
+...
+pte->user_acc     = PTE_USER_ACC_SET(perm);
+...
+;;pte->writethrough = PTE_WRITETHRU_SET(perm);
+...
+;;pte->nocache      = PTE_NOCACHE_SET(perm);
+pte->pfn          = PTE_ADDR_TO_PFN(pa);
+mov    -0x18[rbp],rcx
+shr    0xc,rcx
+mov    -0x30[rbp],rdx
+mov    [rdx],rsi
+movabs 0xffffffffff,rdi
+and    rdi,rcx
+shl    0xc,rcx
+movabs 0xfff0000000000fff,rdi
+and    rdi,rsi
+or     rcx,rsi
+mov    rsi,[rdx]
+;;pte->nx           = PTE_NX_SET(perm);
+;;pte->present      = 1;
+;;clean up the stack and return
+...
+;;}
 $\specline{\exists (\entryf ,\;\entrytr,\; \entrytw,\; \entryo,\;\textsf{pte\_addr },\paddr) \; \ldotp\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast  \texttt{r14}\mapsto_{\textsf{r}} \_ \ast \texttt{rdi}\mapsto_{r} \vaddr \ast \texttt{rax}\mapsto_{\textsf{r}} \textsf{ pte\_addr} \; \ast }_{\rtv}$
@@
   \label{sec:p2v}
- Once we make sure that the frame of the page-table entry ($\textsf{entry}\mapsto\textsf{pfn}$)is backed with an allocated physical page, we must obtain the virtual address of the next level-page table (Line 46 in \ref{calltopteinitialize}).
+ Once we make sure that the frame of the page-table entry ($\textsf{entry}\mapsto\textsf{pfn}$)is backed with an allocated physical page, we must obtain the virtual address of the next level-page table (Line 48 in \ref{calltopteinitialize}).
 To do so, as for obtaining any virtual page-table address, we, again, need to open the invariant look up the \textit{identity mapping} for the physical address ($\textsf{entry}\mapsto \textsf{pfn}$).
@@
 This actually constitutes a very critical piece of the full page table walk verification, and we have verified the critical step for a small x86-64 kernel, which is the physical-to-virtual conversion (often appearing as a macro \texttt{P2V} inside the macro \textsf{ PTE\_PFN\_TO\_ADDR} in C source code).
-In our small kernel (Line 46 in Figure \ref{calltopteinitialize}), as in larger kernels, \texttt{P2V} is actually just addition, but the correctness of this simple instruction is quite subtle and relies on identity mappings explained in the previous section (Section \ref{subsec:identitymappings}) as a part of understanding our address-space invariant in Figure \ref{fig:peraspaceinvariant_with_p2v_extension}.
+In our small kernel (Line 48 in Figure \ref{calltopteinitialize}), as in larger kernels, \texttt{P2V} is actually just addition, but the correctness of this simple instruction is quite subtle and relies on identity mappings explained in the previous section (Section \ref{subsec:identitymappings}) as a part of understanding our address-space invariant in Figure \ref{fig:peraspaceinvariant_with_p2v_extension}.
 \begin{figure}\footnotesize

### 2024-07-04 examples.tex
before the last check on mapapage

@@
  $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m)  \ast  \texttt{r14}\mapsto_{\textsf{r}} \_ \ast \texttt{rdi}\mapsto_{r} \vaddr \ast \texttt{rax}\mapsto_{\textsf{r}} \_ \ast \ulcorner \texttt{aligned va} \land  \theta \; !!\; \vaddr = \texttt{None}\urcorner}_{\rtv}$
-;;pte = walkpgdir_succ(pml4, a);
-call pte_walkpgdir ;; Returns the virtual address of the L1 entry in rax
-
 ;;char *a; pte_t *pte;
@@
 callq  291 <vaspace_mappage+0x41>	28d: R_X86_64_PLT32	walkpgdir_succ-0x4
-xor    r8,r8
+mov    0x0,r8
 movslq eax,rcx
 mov    rcx,-0x30[rbp]
-
 ;;pte->writable     = PTE_WRITEABLE_SET(perm);
@@
 ;;}
-$\specline{\exists (\entryf ,\;\entrytr,\; \entrytw,\; \entryo,\;\textsf{pte\_addr },\paddr) \; \ldotp\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast  \texttt{r14}\mapsto_{\textsf{r}} \_ \ast \texttt{rdi}\mapsto_{r} \vaddr \ast \texttt{rax}\mapsto_{\textsf{r}} \textsf{ pte\_addr} \; \ast }_{\rtv}$
-$\specline{ \ulcorner  \texttt{addr\_L1 }(\vaddr, \entryo) = \paddr \urcorner \ast \ulcorner \texttt{entry\_present } \entryf \land \texttt{entry\_present } \entrytr \land  \texttt{entry\_present } \entrytw \urcorner \; \ast}_{\rtv}$
-$\specline{\nfpointsto{\mask\vaddr\maskfour\rtv}{\mask\vaddr\maskfouroff\rtv}\entryf\qone\naddr \; \ast \nfpointsto{\mask\vaddr\maskthree\entryf}{\mask\vaddr\maskthreeoff\entryf}\entrytr\qtwo\naddr \ast}_{\rtv}$
-$\specline{  \nfpointsto{\mask\vaddr\masktwo\entrytr}{\mask\vaddr\masktwooff\entrytr}\paddr\qthree\entryo \;\ast \texttt{pte\_addr} \mapsto_{\texttt{vpte}} \paddr \;(\texttt{wzero 64}) \ast \texttt{rax}\mapsto_{\textsf{r}} \texttt{pte\_addr}  }_{\rtv}$
-mov r14, rax ;; Save that before another call
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast  \texttt{r14}\mapsto_{\textsf{r}} \texttt{pte\_addr} \ast \texttt{rdi}\mapsto_{\textsf{r}} \vaddr \ast \texttt{rax}\mapsto_{\textsf{r}} \textsf{ pte\_addr} \; \ast }_{\rtv}$
-$\specline{ \nfpointsto{\mask\vaddr\maskfour\rtv}{\mask\vaddr\maskfouroff\rtv}\entryf\qone\naddr \ast \ulcorner \texttt{entry\_present } \entryf \land \texttt{entry\_present } \entrytr \land  \texttt{entry\_present } \entrytw \urcorner \ast}_{\rtv}$
-$\specline{  \nfpointsto{\mask\vaddr\maskthree\entryf}{\mask\vaddr\maskthreeoff\entryf}\entrytr\qtwo\naddr \ast \nfpointsto{\mask\vaddr\masktwo\entrytr}{\mask\vaddr\masktwooff\entrytr}\paddr\qthree\entryo \;\ast}_{\rtv}$
-$\specline{\texttt{pte\_addr} \mapsto_{\texttt{vpte}} \paddr \;(\texttt{wzero 64}) \ast \texttt{rax}\mapsto_{\textsf{r}} \texttt{pte\_addr}  }_{\rtv}$
-call alloc_phys_page_or_panic
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast  \texttt{r14}\mapsto_{\textsf{r}} \texttt{pte\_addr} \ast \texttt{rdi}\mapsto_{\textsf{r}} \vaddr \;\ast \nfpointsto{\mask\vaddr\maskfour\rtv}{\mask\vaddr\maskfouroff\rtv}\entryf\qone\naddr \ast}_{\rtv}$
-$\specline{  \nfpointsto{\mask\vaddr\maskthree\entryf}{\mask\vaddr\maskthreeoff\entryf}\entrytr\qtwo\naddr \ast \nfpointsto{\mask\vaddr\masktwo\entrytr}{\mask\vaddr\masktwooff\entrytr}\paddr\qthree\naddr \;\ast}_{\rtv}$
-$\specline{\texttt{pte\_addr} \mapsto_{\texttt{vpte}} \paddr\; (\texttt{wzero 64})  \ast \ulcorner \texttt{entry\_present } \entryf \land \texttt{entry\_present } \entrytr \land  \texttt{entry\_present } \entrytw \urcorner}_{\rtv}$
-$\specline{\exists \texttt{ fpaddr} \ldotp \ulcorner \texttt{aligned fpaddr} \urcorner \ast \texttt{rax}\mapsto_{\textsf{r}} \texttt{fpaddr+3} \ast \texttt{fpaddr} \mapsto_{\textsf{p}} (\texttt{wzero 64}) \ast \ulcorner \texttt{entry\_present (fpaddr+3)}\urcorner}_{\rtv}$
-;; Calculate new L1 entry
-mov [r14], rax ;; store the page table entry, mapping the page
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast  \texttt{r14}\mapsto_{\textsf{r}} \texttt{pte\_addr} \ast \texttt{rdi}\mapsto_{\textsf{r}} \vaddr \;\ast \nfpointsto{\mask\vaddr\maskfour\rtv}{\mask\vaddr\maskfouroff\rtv}\entryf\qone\naddr \ast}_{\rtv}$
-$\specline{  \nfpointsto{\mask\vaddr\maskthree\entryf}{\mask\vaddr\maskthreeoff\entryf}\entrytr\qtwo\naddr \ast \nfpointsto{\mask\vaddr\masktwo\entrytr}{\mask\vaddr\masktwooff\entrytr}\paddr\qthree\entryo \;\ast}_{\rtv}$
-$\specline{\texttt{pte\_addr} \mapsto_{\texttt{vpte}} \paddr \;(\texttt{fpaddr+3}) \; \ast \ulcorner \texttt{entry\_present } \entryf \land \texttt{entry\_present } \entrytr \land  \texttt{entry\_present } \entrytw \urcorner }_{\rtv}$
-$\specline{\ulcorner \texttt{aligned fpaddr} \urcorner \ast \texttt{rax}\mapsto_{\textsf{r}} \texttt{fpaddr+3} \ast \texttt{fpaddr} \mapsto_{\textsf{p}} (\texttt{wzero 64}) \ast \ulcorner \texttt{entry\_present fpaddr+3}\urcorner}_{\rtv}$
-$\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\; \sqsubseteq $
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast  \texttt{r14}\mapsto_{\textsf{r}} \texttt{pte\_addr} \ast \texttt{rdi}\mapsto_{\textsf{r}} \vaddr \ast }_{\rtv}$
-$\specline{\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\vaddr,\entryf,\entrytr,\entrytw,\fpaddr+3) \ast \ulcorner \theta \;!!\;\vaddr = \texttt{None}\urcorner \; \ast}_{\rtv}$
-$\specline{\ulcorner \texttt{aligned fpaddr} \urcorner \ast \texttt{rax}\mapsto_{\textsf{r}} \texttt{fpaddr+3} \ast \texttt{fpaddr} \mapsto_{\textsf{p}} (\texttt{wzero 64}) }_{\rtv}$
-$\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\; \sqsubseteq $
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace} (<[\vaddr:=\texttt{fpaddr}]> \theta,m) \ast}_{\rtv}$
-$\specline{\ulcorner \texttt{aligned fpaddr} \urcorner \ast \texttt{fpaddr} \mapsto_{\textsf{p}} \textsf{ wzero 64} \ast \sumapaces\rtv\delta  \ast\sumwalkabs\vaddr\qfrac\fpaddr}_{\rtv}$
-$\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\; \sqsubseteq $
-  $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace} (<[\vaddr:=\texttt{fpaddr}]> \theta,m) \ast \vaddr \mapsto_{\textsf{vpte}}\; \{\qfrac\} \;\fpaddr \textsf{ wzero 64}}_{\rtv}$
-$\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\; \sqsubseteq $
-  $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace} (<[\vaddr:=\texttt{fpaddr}]> \theta,m) \ast \vaddr \mapsto_{\textsf{v}}\; \{\qfrac\} \textsf{wzero 64}}_{\rtv}$
 \end{lstlisting}
@@
 Thus many modern kernels have reverted to the older kernel design where the kernel inhabits its own unique address space, and user processes have only enough extra material mapped in their address spaces to switch into the kernel (CPUs do not speculate past updates to \texttt{cr3}).
+
+\begin{comment}    
+$\specline{\exists (\entryf ,\;\entrytr,\; \entrytw,\; \entryo,\;\textsf{pte\_addr },\paddr) \; \ldotp\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast  \texttt{r14}\mapsto_{\textsf{r}} \_ \ast \texttt{rdi}\mapsto_{r} \vaddr \ast \texttt{rax}\mapsto_{\textsf{r}} \textsf{ pte\_addr} \; \ast }_{\rtv}$
+$\specline{ \ulcorner  \texttt{addr\_L1 }(\vaddr, \entryo) = \paddr \urcorner \ast \ulcorner \texttt{entry\_present } \entryf \land \texttt{entry\_present } \entrytr \land  \texttt{entry\_present } \entrytw \urcorner \; \ast}_{\rtv}$
+$\specline{\nfpointsto{\mask\vaddr\maskfour\rtv}{\mask\vaddr\maskfouroff\rtv}\entryf\qone\naddr \; \ast \nfpointsto{\mask\vaddr\maskthree\entryf}{\mask\vaddr\maskthreeoff\entryf}\entrytr\qtwo\naddr \ast}_{\rtv}$
+$\specline{  \nfpointsto{\mask\vaddr\masktwo\entrytr}{\mask\vaddr\masktwooff\entrytr}\paddr\qthree\entryo \;\ast \texttt{pte\_addr} \mapsto_{\texttt{vpte}} \paddr \;(\texttt{wzero 64}) \ast \texttt{rax}\mapsto_{\textsf{r}} \texttt{pte\_addr}  }_{\rtv}$
+mov r14, rax ;; Save that before another call
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast  \texttt{r14}\mapsto_{\textsf{r}} \texttt{pte\_addr} \ast \texttt{rdi}\mapsto_{\textsf{r}} \vaddr \ast \texttt{rax}\mapsto_{\textsf{r}} \textsf{ pte\_addr} \; \ast }_{\rtv}$
+$\specline{ \nfpointsto{\mask\vaddr\maskfour\rtv}{\mask\vaddr\maskfouroff\rtv}\entryf\qone\naddr \ast \ulcorner \texttt{entry\_present } \entryf \land \texttt{entry\_present } \entrytr \land  \texttt{entry\_present } \entrytw \urcorner \ast}_{\rtv}$
+$\specline{  \nfpointsto{\mask\vaddr\maskthree\entryf}{\mask\vaddr\maskthreeoff\entryf}\entrytr\qtwo\naddr \ast \nfpointsto{\mask\vaddr\masktwo\entrytr}{\mask\vaddr\masktwooff\entrytr}\paddr\qthree\entryo \;\ast}_{\rtv}$
+$\specline{\texttt{pte\_addr} \mapsto_{\texttt{vpte}} \paddr \;(\texttt{wzero 64}) \ast \texttt{rax}\mapsto_{\textsf{r}} \texttt{pte\_addr}  }_{\rtv}$
+call alloc_phys_page_or_panic
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast  \texttt{r14}\mapsto_{\textsf{r}} \texttt{pte\_addr} \ast \texttt{rdi}\mapsto_{\textsf{r}} \vaddr \;\ast \nfpointsto{\mask\vaddr\maskfour\rtv}{\mask\vaddr\maskfouroff\rtv}\entryf\qone\naddr \ast}_{\rtv}$
+$\specline{  \nfpointsto{\mask\vaddr\maskthree\entryf}{\mask\vaddr\maskthreeoff\entryf}\entrytr\qtwo\naddr \ast \nfpointsto{\mask\vaddr\masktwo\entrytr}{\mask\vaddr\masktwooff\entrytr}\paddr\qthree\naddr \;\ast}_{\rtv}$
+$\specline{\texttt{pte\_addr} \mapsto_{\texttt{vpte}} \paddr\; (\texttt{wzero 64})  \ast \ulcorner \texttt{entry\_present } \entryf \land \texttt{entry\_present } \entrytr \land  \texttt{entry\_present } \entrytw \urcorner}_{\rtv}$
+$\specline{\exists \texttt{ fpaddr} \ldotp \ulcorner \texttt{aligned fpaddr} \urcorner \ast \texttt{rax}\mapsto_{\textsf{r}} \texttt{fpaddr+3} \ast \texttt{fpaddr} \mapsto_{\textsf{p}} (\texttt{wzero 64}) \ast \ulcorner \texttt{entry\_present (fpaddr+3)}\urcorner}_{\rtv}$
+;; Calculate new L1 entry
+mov [r14], rax ;; store the page table entry, mapping the page
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast  \texttt{r14}\mapsto_{\textsf{r}} \texttt{pte\_addr} \ast \texttt{rdi}\mapsto_{\textsf{r}} \vaddr \;\ast \nfpointsto{\mask\vaddr\maskfour\rtv}{\mask\vaddr\maskfouroff\rtv}\entryf\qone\naddr \ast}_{\rtv}$
+$\specline{  \nfpointsto{\mask\vaddr\maskthree\entryf}{\mask\vaddr\maskthreeoff\entryf}\entrytr\qtwo\naddr \ast \nfpointsto{\mask\vaddr\masktwo\entrytr}{\mask\vaddr\masktwooff\entrytr}\paddr\qthree\entryo \;\ast}_{\rtv}$
+$\specline{\texttt{pte\_addr} \mapsto_{\texttt{vpte}} \paddr \;(\texttt{fpaddr+3}) \; \ast \ulcorner \texttt{entry\_present } \entryf \land \texttt{entry\_present } \entrytr \land  \texttt{entry\_present } \entrytw \urcorner }_{\rtv}$
+$\specline{\ulcorner \texttt{aligned fpaddr} \urcorner \ast \texttt{rax}\mapsto_{\textsf{r}} \texttt{fpaddr+3} \ast \texttt{fpaddr} \mapsto_{\textsf{p}} (\texttt{wzero 64}) \ast \ulcorner \texttt{entry\_present fpaddr+3}\urcorner}_{\rtv}$
+$\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\; \sqsubseteq $
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast  \texttt{r14}\mapsto_{\textsf{r}} \texttt{pte\_addr} \ast \texttt{rdi}\mapsto_{\textsf{r}} \vaddr \ast }_{\rtv}$
+$\specline{\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\vaddr,\entryf,\entrytr,\entrytw,\fpaddr+3) \ast \ulcorner \theta \;!!\;\vaddr = \texttt{None}\urcorner \; \ast}_{\rtv}$
+$\specline{\ulcorner \texttt{aligned fpaddr} \urcorner \ast \texttt{rax}\mapsto_{\textsf{r}} \texttt{fpaddr+3} \ast \texttt{fpaddr} \mapsto_{\textsf{p}} (\texttt{wzero 64}) }_{\rtv}$
+$\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\; \sqsubseteq $
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace} (<[\vaddr:=\texttt{fpaddr}]> \theta,m) \ast}_{\rtv}$
+$\specline{\ulcorner \texttt{aligned fpaddr} \urcorner \ast \texttt{fpaddr} \mapsto_{\textsf{p}} \textsf{ wzero 64} \ast \sumapaces\rtv\delta  \ast\sumwalkabs\vaddr\qfrac\fpaddr}_{\rtv}$
+$\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\; \sqsubseteq $
+  $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace} (<[\vaddr:=\texttt{fpaddr}]> \theta,m) \ast \vaddr \mapsto_{\textsf{vpte}}\; \{\qfrac\} \;\fpaddr \textsf{ wzero 64}}_{\rtv}$
+$\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\; \sqsubseteq $
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace} (<[\vaddr:=\texttt{fpaddr}]> \theta,m) \ast \vaddr \mapsto_{\textsf{v}}\; \{\qfrac\} \textsf{wzero 64}}_{\rtv}$
+\end{comment}

### 2024-07-04 examples.tex
+

@@
 
-The first aspect is important for not invalidating the soundness: \textsf{pte\_initialize} extends the identity map with the freshly allocated \textsf{nextpaddr}, and must not reveal the ownership of the physical pointsto in the post-condition (Line 33 and 35 in Figure \ref{fig:calltopteinitialize}, Lines 22-23 in Figure \ref{fig:pteinitializespec}) because it would create unsound coexistence of the full-ownership of the physical pointsto resource (a hypothetical $\mathsf{nextpaddr} \mapsto_{\mathsf{p,1}} w0_64$) and fragmental (potentially full based on the side-condition *) physical pointsto resource that can be obtained from the identity map inside the extended invariant in Figure \ref{fig:peraspaceinvariant_with_p2v_extension} with the capability token $\ulcorner \Xi !!
-\textsf{nextpaddr} = \textsf{ Some Z} \urcorner$.
+The first aspect is important for not invalidating the soundness: \textsf{pte\_initialize} extends the identity map with the freshly allocated \textsf{nextpaddr}, and must not reveal the ownership of the physical pointsto in the post-condition (Line 33 and 35 in Figure \ref{fig:calltopteinitialize}, Lines 22-23 in Figure \ref{fig:pteinitializespec}) because it would create unsound coexistence of the full-ownership of the physical pointsto resource (a hypothetical $\mathsf{nextpaddr} \mapsto_{\mathsf{p,1}} w0_64$) and fragmental (potentially full based on the side-condition *) physical pointsto resource that can be obtained from the identity map inside the extended invariant in Figure \ref{fig:peraspaceinvariant_with_p2v_extension} with the capability token $\ulcorner \Xi \;!!;\ \textsf{nextpaddr} = \textsf{ Some Z} \urcorner$.
 Although, the first aspect -- not explicitly revealing the physical pointsto resource for the freshly allocated page-table address but providing access through the identity map inside the invariant with access-capability tokens -- recovers the soundness argument expalined, we still need the knowledge that the fresly allocated address references to a \textit{zeroed} page.
 The invariant preserves this information through the constraint on relating the type (\textsf{T}) of mapped data -- \textsf{T = Z} for zeroed or \textsf{T = V} for random page value -- to the value of existentially quantified value -- \textsf{val} -- in Figure \ref{fig:peraspaceinvariant_with_p2v_extension}.
-So, for accessing the physical pointsto resource to the zeroed value referenced by \textsf{nextpaddr}, we need to have the fragmental ownership of the identity-map $\Xi !!
-\textsf{nextpaddr} = \textsf{ Some Z}$ which ensures and carries the \textit{being-zeroed} knowledge when we open and access the the existentially quantified value inside the invariant.
+So, for accessing the physical pointsto resource to the zeroed value referenced by \textsf{nextpaddr}, we need to have the fragmental ownership of the identity-map $\Xi \; !!
+\; \textsf{nextpaddr} = \textsf{ Some Z}$ which ensures and carries the \textit{being-zeroed} knowledge when we open and access the the existentially quantified value inside the invariant.
   \subsubsection{Kernels Utilizing \textsf{P2V}}
@@
 ;; Assume the save-space is in rdi, load-space in rsi. First, save the yielding context
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast [\rtv'](\mathcal{I}\texttt{ASpace}(\theta',m') \ast \texttt{Pother}) \ast \texttt{rsi}\mapsto_{\textsf{r}} new \ast \texttt{rdi}\mapsto_{\textsf{r}} old \ast \texttt{rbx}\mapsto_{\textsf{r}} \texttt{rbxv} }_{\rtv}$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,\Xi,m) \ast [\rtv'](\mathcal{I}\texttt{ASpace}(\theta',\Xi',m') \ast \texttt{Pother}) \ast \texttt{rsi}\mapsto_{\textsf{r}} new \ast \texttt{rdi}\mapsto_{\textsf{r}} old \ast \texttt{rbx}\mapsto_{\textsf{r}} \texttt{rbxv} }_{\rtv}$
 $\specline{\texttt{rsp}\mapsto_{\textsf{r}} \texttt{rspv} \ast \texttt{rbp}\mapsto_{\textsf{r}} \texttt{rbpv} \ast \texttt{r12}\mapsto_{\textsf{r}} \texttt{r12v} \ast \texttt{r13}\mapsto_{\textsf{r}} \texttt{r13v} \ast \texttt{r14}\mapsto_{\textsf{r}} \texttt{r14v} \ast \texttt{r15}\mapsto_{\textsf{r}} \texttt{r15v}}_{\rtv}$
@@
 mov 48[rdi], r15
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast [\rtv'](\mathcal{I}\texttt{ASpace}(\theta',m') \ast \texttt{Pother})  \ast  \texttt{rsi}\mapsto_{\textsf{r}} new \ast \texttt{rdi}\mapsto_{\textsf{r}} old \ast \texttt{rbx}\mapsto_{\textsf{r}} \texttt{rbxv} }_{\rtv}$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,\Xi,m) \ast [\rtv'](\mathcal{I}\texttt{ASpace}(\theta',\Xi',m') \ast \texttt{Pother})  \ast  \texttt{rsi}\mapsto_{\textsf{r}} new \ast \texttt{rdi}\mapsto_{\textsf{r}} old \ast \texttt{rbx}\mapsto_{\textsf{r}} \texttt{rbxv} }_{\rtv}$
 $\specline{ \texttt{rsp}\mapsto_{\textsf{r}} \texttt{rspv} \ast \texttt{rbp}\mapsto_{\textsf{r}} \texttt{rbpv} \ast \texttt{r12}\mapsto_{\textsf{r}} \texttt{r12v} \ast \texttt{r13}\mapsto_{\textsf{r}} \texttt{r13v} \ast \texttt{r14}\mapsto_{\textsf{r}} \texttt{r14v} \ast \texttt{r15}\mapsto_{\textsf{r}} \texttt{r15v}}_{\rtv}$
@@
 mov r15, 48[rsi]
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast [\rtv'](\mathcal{I}\texttt{ASpace}(\theta',m') \ast \texttt{Pother})  \ast  \texttt{rsi}\mapsto_{\textsf{r}} new \ast \texttt{rdi}\mapsto_{\textsf{r}} old \ast \texttt{rbx}\mapsto_{\textsf{r}} \texttt{rbxv}' }_{\rtv}$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,\Xi,m) \ast [\rtv'](\mathcal{I}\texttt{ASpace}(\theta',\Xi',m') \ast \texttt{Pother})  \ast  \texttt{rsi}\mapsto_{\textsf{r}} new \ast \texttt{rdi}\mapsto_{\textsf{r}} old \ast \texttt{rbx}\mapsto_{\textsf{r}} \texttt{rbxv}' }_{\rtv}$
 $\specline{\texttt{rsp}\mapsto_{\textsf{r}} \texttt{rspv}' \ast \texttt{rbp}\mapsto_{\textsf{r}} \texttt{rbpv}' \ast \texttt{r12}\mapsto_{\textsf{r}} \texttt{r12v}' \ast \texttt{r13}\mapsto_{\textsf{r}} \texttt{r13v}' \ast \texttt{r14}\mapsto_{\textsf{r}} \texttt{r14v}' \ast \texttt{r15}\mapsto_{\textsf{r}} \texttt{r15v}'}_{\rtv}$
@@
 mov cr3, 56[rsi]
-$\specline{ [\rtv](\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast \mathsf{ContextAt}(old,[\mathsf{rbxv,\ldots,\rtv}]) \ast \mathsf{ContextAt}(new,[\mathsf{rbxv}',\ldots,\rtv']))  }_{\rtv}$
-$\specline{ \mathcal{I}\texttt{ASpace}(\theta',m') \ast \texttt{Pother} \ast  \texttt{rsi}\mapsto_{\textsf{r}} new \ast \texttt{rdi}\mapsto_{\textsf{r}} old \ast \ldots }_{\rtv'}$
+$\specline{ [\rtv](\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,\Xi,m) \ast \mathsf{ContextAt}(old,[\mathsf{rbxv,\ldots,\rtv}]) \ast \mathsf{ContextAt}(new,[\mathsf{rbxv}',\ldots,\rtv']))  }_{\rtv}$
+$\specline{ \mathcal{I}\texttt{ASpace}(\theta',\Xi',m') \ast \texttt{Pother} \ast  \texttt{rsi}\mapsto_{\textsf{r}} new \ast \texttt{rdi}\mapsto_{\textsf{r}} old \ast \ldots }_{\rtv'}$
 \end{lstlisting}
@@
 not the address space of the calling thread shown as 
-\[[\rtv'](\mathcal{I}\texttt{ASpace}(\theta,m) \ast \texttt{Pother})\]
+\[[\rtv'](\mathcal{I}\texttt{ASpace}(\theta,\Xi,m) \ast \texttt{Pother})\]
 in the specfication. 

### 2024-07-04 examples.tex
+

@@
 One of the key tasks of a page fault handler in a general-purpose OS kernel is
-to map new pages into an address space by writing into an existing page table (\textsf{vaspace\_mappage(pte\_t *pml4, void *va,uintptr\_t pa, uint64\_t perm )} in Figure \ref{fig:mapping_code}).
+to map new pages into an address space by writing into an existing page table via a call
+\[\textsf{vaspace\_mappage(pte\_t *pml4, void *va,uintptr\_t pa, uint64\_t perm )}\] in Figure \ref{fig:mapping_code}).
 To do so, with a given allocated a fresh page (\textsf{pa}), then calculate the appropriate
@@
 ;;a = (char *)PGROUNDUP((uintptr_t)va);
- $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m)  \ast  \texttt{r14}\mapsto_{\textsf{r}} \_ \ast \texttt{rdi}\mapsto_{r} \vaddr \ast \texttt{rax}\mapsto_{\textsf{r}} \_ \ast \ulcorner \texttt{aligned va} \land  \theta \; !!\; \vaddr = \texttt{None}\urcorner}_{\rtv}$
+      $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m)  \ast \texttt{rdi}\mapsto_{r} \textsf{plm4} \ast \texttt{rsi}\mapsto_{r} \textsf{va} \ast  \texttt{rdx}\mapsto_{r} \textsf{pa} \ast \texttt{rcx}\mapsto_{r} \textsf{perm} \ast \texttt{rax}\mapsto_{\textsf{r}} \_ \ast \ulcorner \texttt{aligned va} \land  \theta \; !!\; \vaddr = \texttt{None}\urcorner\ast }_{\rtv}$
+      $\specline{\texttt{rbp-8}\mapsto_{v} \textsf{plm4} \ast \texttt{rbp-16}\mapsto_{v} \textsf{va}  \ast \texttt{rbp-24}\mapsto_{v} \textsf{pa} \ast \texttt{rbp-32}\mapsto_{v} \textsf{perm} }_{\rtv}$
 ;;char *a; pte_t *pte;
 ;;a = (char *)PGROUNDUP((uintptr_t)va);
-mov    -0x10[rbp],rcx
-add    0x1000,rcx
-sub    0x1,rcx
-and    0xfffffffffffff000,%rcx
-mov    rcx,-0x28[rbp]
-
+...
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast  \texttt{r14}\mapsto_{\textsf{r}} \texttt{pte\_addr} \ast \texttt{rdi}\mapsto_{\textsf{r}} \vaddr \;\ast \nfpointsto{\mask\vaddr\maskfour\rtv}{\mask\vaddr\maskfouroff\rtv}\entryf\qone\naddr \ast}_{\rtv}$
+$\specline{  \nfpointsto{\mask\vaddr\maskthree\entryf}{\mask\vaddr\maskthreeoff\entryf}\entrytr\qtwo\naddr \ast \nfpointsto{\mask\vaddr\masktwo\entrytr}{\mask\vaddr\masktwooff\entrytr}\paddr\qthree\naddr \;\ast}_{\rtv}$
+$\specline{\texttt{pte\_addr} \mapsto_{\texttt{vpte}} \paddr\; (\texttt{wzero 64})  \ast \ulcorner \texttt{entry\_present } \entryf \land \texttt{entry\_present } \entrytr \land  \texttt{entry\_present } \entrytw \urcorner}_{\rtv}$
+$\specline{\exists \texttt{ fpaddr} \ldotp \ulcorner \texttt{aligned fpaddr} \urcorner \ast \texttt{rax}\mapsto_{\textsf{r}} \texttt{fpaddr+3} \ast \texttt{fpaddr} \mapsto_{\textsf{p}} (\texttt{wzero 64}) \ast \ulcorner \texttt{entry\_present (fpaddr+3)}\urcorner}_{\rtv}$
 ;;pte = walkpgdir_succ(pml4, a);
@@
 mov    0x0,r8
-movslq eax,rcx
+mov    rax,rcx
 mov    rcx,-0x30[rbp]
@@
 the existence of a memory location for an appropriate L1 table entry.
-The code uses a helper function \lstinline|pte\_walkpgdir| (discussed again in Section \ref{sec:traversing}).
+The code uses a helper function \lstinline{pte\_walkpgdir} (discussed again in Section \ref{sec:traversing}).
 \textsf{pte\_walkpgdir}'s postcondition contains virtual \emph{PTE} pointsto assertions ($\mapsto_{\textsf{vpte}}$) \todo{colin, these are no more plain physical pointstos}  both for ensuring partial page table walk reaching the
@@
 
-The crucial step is actually updating the L1 entry,
+The crucial step is actually updating the L1 entry (Line X\todo{X} in Figure \ref{fig:mapping_code}),
 via the virtual address

### 2024-07-04 examples.tex
+

--- 2024-07-04 examples.tex
@@
 %\end{comment}
-\subsection{Mapping a New Page}
-One of the key tasks of a page fault handler in a general-purpose OS kernel is
-to map new pages into an address space by writing into an existing page table (\textsf{vaspace\_mappage(pte\_t *pml4, void *va,uintptr\_t pa, uint64\_t perm )} in Figure \ref{fig:mapping_code}).
-To do so, with a given allocated a fresh page (\textsf{pa}), then calculate the appropriate
-known-valid page table walks (\textsf{pte\_walkpgdir} Line X in Figure \ref{fig:mapping_code})  and update the appropriate L1 page table entry (permissions and frame address Lines in Figure \ref{fig:mapping_code}).
-%\lstset{
-%  columns=fullflexible,
-%  numbers=left,
-%  basicstyle=\ttfamily,
-%  keywordstyle=\color{blue}\bfseries,
-%  morekeywords={mov,add,call},
-%  emph={rsp,rdx,rax,rbx,rbp,rsi,rdi,rcx,r8,r9,r10,r11,r12,r13,r14,r15},
-%  emphstyle=\color{green},
-%  emph={[2]cr3},
-%  emphstyle={[2]\color{violet}},
-%  morecomment=[l]{;;},
-%  mathescape
-%}
-\newcommand{\sumwalkabsent}{
-  \ownGhost\gammaPred{\authfrag{\singletonMap{\texttt{entry+KERNBASE}}{(\textsf{qfrac}, \textsf{entry})}}}
-}
-
-\newcommand{\ventry}{\texttt{entry + KERNBASE}}
-\newcommand{\entry}{\texttt{entry}}
-\newcommand{\qfraczero}{\textsf{qfrac}}
-\newcommand{\true}{\textsf{true}}
-\tikzstyle{boxedassert_border} = [sharp corners,line width=0.2pt]
-\NewDocumentCommand \boxedassertpv {O{} m o}{%
-	\tikz[baseline=(m.base)]{
-		%	  \node[rectangle, draw,inner sep=0.8pt,anchor=base,#1] (m) {${#2}\mathstrut$};
-		\node[rectangle,inner sep=1.5pt,outer sep=0.2pt,anchor=base] (m) {${\,#2\,}\mathstrut$};
-		\draw[#1,boxedassert_border] ($(m.south west)$) rectangle ($(m.north east)$);
-	}\IfNoValueF{#3}{^{\,#3}}%
-}
-\newcommand*{\knowInvpv}[2]{\boxedassertpv{#2}[#1]}
-\newcommand*{\ownGhostpv}[2]{\boxedassertpv[dash dot]{#2}[#1]}
-
-\newcommand{\sumpv}[3]{
-  \ownGhostpv\gammaPred{\authfrag{\singletonMap{#1}{(#2, #3)}}}
-}
-
-\newcommand{\pvmapping}[1]{\mathcal{A}\textsf{PVMappings}(#1)}
-
-\newcommand{\fpaddr}{\texttt{fpaddr}}
-\newcommand{\specline}[1]{{\color{blue}\left\{#1\right\}}}
-\newcommand{\sumapacesfull}[2]{
-  \ownGhost\gammaPreds{\authfull{\singletonMap{#1}{#2}}}
-}
-\todo{ismail change the following spec base on the changes you have in the code}
-\begin{figure}\footnotesize
-  \begin{lstlisting}[mathescape]
-;;complstatus_t vaspace_mappage(pte_t *pml4, void *va,uintptr_t pa, uint64_t perm ) {
-...
-;;setting up the stack      
-;;char *a; pte_t *pte;
-;;a = (char *)PGROUNDUP((uintptr_t)va);
- $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m)  \ast  \texttt{r14}\mapsto_{\textsf{r}} \_ \ast \texttt{rdi}\mapsto_{r} \vaddr \ast \texttt{rax}\mapsto_{\textsf{r}} \_ \ast \ulcorner \texttt{aligned va} \land  \theta \; !!\; \vaddr = \texttt{None}\urcorner}_{\rtv}$
-;;char *a; pte_t *pte;
-;;a = (char *)PGROUNDUP((uintptr_t)va);
-mov    -0x10[rbp],rcx
-add    0x1000,rcx
-sub    0x1,rcx
-and    0xfffffffffffff000,%rcx
-mov    rcx,-0x28[rbp]
-
-;;pte = walkpgdir_succ(pml4, a);
-mov    -0x8[rbp],rdi
-mov    -0x28[rbp],rsi
-mov    0x0,rax
-callq  291 <vaspace_mappage+0x41>	28d: R_X86_64_PLT32	walkpgdir_succ-0x4
-mov    0x0,r8
-movslq eax,rcx
-mov    rcx,-0x30[rbp]
-;;pte->writable     = PTE_WRITEABLE_SET(perm);
-...
-pte->user_acc     = PTE_USER_ACC_SET(perm);
-...
-;;pte->writethrough = PTE_WRITETHRU_SET(perm);
-...
-;;pte->nocache      = PTE_NOCACHE_SET(perm);
-pte->pfn          = PTE_ADDR_TO_PFN(pa);
-mov    -0x18[rbp],rcx
-shr    0xc,rcx
-mov    -0x30[rbp],rdx
-mov    [rdx],rsi
-movabs 0xffffffffff,rdi
-and    rdi,rcx
-shl    0xc,rcx
-movabs 0xfff0000000000fff,rdi
-and    rdi,rsi
-or     rcx,rsi
-mov    rsi,[rdx]
-;;pte->nx           = PTE_NX_SET(perm);
-;;pte->present      = 1;
-;;clean up the stack and return
-...
-;;}
-\end{lstlisting}
-  \caption{Specification and proof of code for mapping a new page.}
-\label{fig:mapping_code}
-\end{figure}
-
-In Figure \ref{fig:mapping_code}, we see an address ($\vaddr$) currently not
-mapped to a page ($\theta \; !!\; \vaddr = \texttt{None}$). Mapping a fresh
-phyiscal page to back the desired virtual page first requires ensuring
-the existence of a memory location for an appropriate L1 table entry.
-The code uses a helper function \lstinline|pte\_walkpgdir| (discussed again in Section \ref{sec:traversing}).
-\textsf{pte\_walkpgdir}'s postcondition contains virtual \emph{PTE} pointsto assertions ($\mapsto_{\textsf{vpte}}$) \todo{colin, these are no more plain physical pointstos}  both for ensuring partial page table walk reaching the
-L1 entry (l1e) with the justification of higher levels of the page table exist, and for allowing access to the memory of the L1 entry via virtual address.
-A PTE points-to is defined just like the normal virtual points-to of Figure \ref{fig:virtualpointstosharing}, except the physical address (here, \textsf{pa}) is explicit in the assertion 
-rather than existentially quantified:
- \[
-\begin{array}{l}
-    \vaddr\mapsto_{\textsf{vpte,q}} \; \paddr \; \vpage : \mathsf{vProp}~\Sigma \stackrel{\triangle}{=} 
-    \exists \delta\ldotp
-	(\lambda \mathit{cr3val}\ldotp
-	\sumapaces{\mathit{cr3val}}\delta) \ast 
-  \sumwalkabs\vaddr\qfrac\paddr \ast \paddr \mapsto_{\mathsf{p}} \vpage
-\end{array}
-\]
-This supports rules for accessing memory
-at that virtual address, but exposing the physical location being modified
-makes this convenient to use for page table modifications, since we must ensure
-the modified data is at the specific physical location that will be used as the L1 entry.
-This is also guaranteed ($ \ulcorner
-\texttt{addr\_L1}(\vaddr,\entryo) = \paddr\urcorner$).
-% After obtaining a virtual address \textsf{pte\_addr} in \textsf{rax} backed 
-% by the physical memory for the L1 entry that will be used to translate the virtual addresses
-% we are mapping, we save it to \textsf{r14} to be updated later in Line 9.
-
-%In the precondition, we see Line 12 allocates a fresh page-aligned, zero-initialized page  (at \textsf{fpaddr}),
-%returning a pre-filled PTE entry in \textsf{rax} ($+3$ sets the lower 2 bits).
-
-The crucial step is actually updating the L1 entry,
-via the virtual address
-(\textsf{pte\_addr}) known to translate to the appropriate physical address, in our example the L1
-table entry address ($\textsf{addr\_L1}(\textsf{va, l1e})$), to hold the freshly
-allocated physical page address (\textsf{fpaddr}) in Line X.
-After this write, we have (Specification Line Y)
-\[\texttt{pte\_addr} \mapsto_{\texttt{vpte}}  \; \paddr\; \textsf{fpaddr}  \qquad\qquad (\textsf{by the PTE variant of Rule}~\TirNameStyle{WriteToVirtMemFromReg})\]
-
-At this point in the proof,  $\texttt{pte\_addr} \mapsto_{\texttt{vpte}}  \; \paddr\; \textsf{fpaddr})$ 
-contains  full ownerhsip of the L1 table entry physical pointsto $\paddr
-\mapsto_{\mathsf{p}} \textsf{fpaddr}$ . Then, we can split the full-ownership of that
-physical pointsto
-($\textsf{fpaddr}\mapsto_{\textsf{p}} \;\textsf{wzero 64}$) to obtain
-fractional ownership on it appropriate for a single address's share of an L1 entry (q4).
-Since we know $\texttt{addr\_L1 }(\vaddr,
-\entryo) \mapsto_{\mathsf{p}} \{q4\} \;\textsf{fpaddr}$) from
-$\mapsto_{\textsf{vpte}}$, after the split of the PTE points-to, and, thanks to the post-condition of the page-table walk (Specification of \textsf{pte\_walkpgdir} in Figure \ref{}), we know the existence of L4-L2 mappings reaching to the L1 exist, so that we can add the last mapping $\texttt{fpaddr} \mapsto_{p} (\texttt{wzero 64})$ to obtain the complete table-walk.
-\todo{with the virtual-pte-pointsto assertions (L4\_L1\_PointsTo($\vaddr$ l4e l3e l2e
-fpaddr)) in Line 25}
-
-Finally, we can insert $\vaddr$ into the ghost page-table-walk summarization
-map ($\theta$) to change its state from unmapped ($\ulcorner \theta
-\;!!\;\vaddr = \texttt{None}\urcorner$ (Line 28) and the
-precondition) to mapped (Line 29) using Iris' ghost-map update
-($\sqsubseteq$), and construct a PTE -points-to $\vaddr
-\mapsto_{\textsf{vpte}}\; \qfrac \;\fpaddr\;(\textsf{wzero 64})$ (Line 31) using $\sumwalkabs\vaddr\qfrac\fpaddr$ obtained from ghost
-page-table-walk insertions, $\sumapaces\rtv\delta$ from unfolding the
-definition of $\mapsto_{\textsf{vpte}}$, and $\fpaddr
-\mapsto_{\textsf{p}} \textsf{wzero 64}$.
-Finally, by existentially quantifying away \textsf{fpaddr},
-we can shift the PTE points-to into a normal virtual points-to.
-
-For brevity this proof only maps the first word of the allocated page; the generalization to the full page
-is straightforward use of iterated separating conjunction to manage collections of clusters of 512 things
-(slices of the L1 entry physical ownership, virtual addresses at 8 byte increments, etc.).
-\looseness=-1
-
-Unlike the only prior work verifying analogous code for mapping a new page~\cite{kolanski08vstte,kolanski09tphols}, our proof above
-does \emph{not} need to unfold the definitions of the logic,
-but instead uses logic rules throughout to reason about the effect of the mapping. 
-This is the first proof we know of for mapping a virtual memory page that does not resort to direct reasoning over the operational semantics.
-% By incorporating verification of the
-% \lstinline|ensure_L1| function (see Section \ref{sec:traversing}), our verification also directly handles several subtle aspects which
-% were axiomatized in prior work.
-
 \subsection{Unmapping a Page}
@@
 % a set of unordered block allocations, we can infer their proofs must also trust these translations.
+\subsection{Mapping a New Page}
+One of the key tasks of a page fault handler in a general-purpose OS kernel is
+to map new pages into an address space by writing into an existing page table via a call
+\[\textsf{vaspace\_mappage(pte\_t *pml4, void *va,uintptr\_t pa, uint64\_t perm )}\] in Figure \ref{fig:mapping_code}).
+To do so, with a given allocated a fresh page (\textsf{pa}), then calculate the appropriate
+known-valid page table walks (\textsf{pte\_walkpgdir} Line X in Figure \ref{fig:mapping_code})  and update the appropriate L1 page table entry (permissions and frame address Lines in Figure \ref{fig:mapping_code}).
+%\lstset{
+%  columns=fullflexible,
+%  numbers=left,
+%  basicstyle=\ttfamily,
+%  keywordstyle=\color{blue}\bfseries,
+%  morekeywords={mov,add,call},
+%  emph={rsp,rdx,rax,rbx,rbp,rsi,rdi,rcx,r8,r9,r10,r11,r12,r13,r14,r15},
+%  emphstyle=\color{green},
+%  emph={[2]cr3},
+%  emphstyle={[2]\color{violet}},
+%  morecomment=[l]{;;},
+%  mathescape
+%}
+\newcommand{\sumwalkabsent}{
+  \ownGhost\gammaPred{\authfrag{\singletonMap{\texttt{entry+KERNBASE}}{(\textsf{qfrac}, \textsf{entry})}}}
+}
+
+\newcommand{\ventry}{\texttt{entry + KERNBASE}}
+\newcommand{\entry}{\texttt{entry}}
+\newcommand{\qfraczero}{\textsf{qfrac}}
+\newcommand{\true}{\textsf{true}}
+\tikzstyle{boxedassert_border} = [sharp corners,line width=0.2pt]
+\NewDocumentCommand \boxedassertpv {O{} m o}{%
+	\tikz[baseline=(m.base)]{
+		%	  \node[rectangle, draw,inner sep=0.8pt,anchor=base,#1] (m) {${#2}\mathstrut$};
+		\node[rectangle,inner sep=1.5pt,outer sep=0.2pt,anchor=base] (m) {${\,#2\,}\mathstrut$};
+		\draw[#1,boxedassert_border] ($(m.south west)$) rectangle ($(m.north east)$);
+	}\IfNoValueF{#3}{^{\,#3}}%
+}
+\newcommand*{\knowInvpv}[2]{\boxedassertpv{#2}[#1]}
+\newcommand*{\ownGhostpv}[2]{\boxedassertpv[dash dot]{#2}[#1]}
+
+\newcommand{\sumpv}[3]{
+  \ownGhostpv\gammaPred{\authfrag{\singletonMap{#1}{(#2, #3)}}}
+}
+
+\newcommand{\pvmapping}[1]{\mathcal{A}\textsf{PVMappings}(#1)}
+
+\newcommand{\fpaddr}{\texttt{fpaddr}}
+\newcommand{\specline}[1]{{\color{blue}\left\{#1\right\}}}
+\newcommand{\sumapacesfull}[2]{
+  \ownGhost\gammaPreds{\authfull{\singletonMap{#1}{#2}}}
+}
+\todo{ismail change the following spec base on the changes you have in the code}
+\begin{figure}\footnotesize
+  \begin{lstlisting}[mathescape]
+;;complstatus_t vaspace_mappage(pte_t *pml4, void *va,uintptr_t pa, uint64_t perm ) {
+...
+;;setting up the stack      
+;;char *a; pte_t *pte;
+;;a = (char *)PGROUNDUP((uintptr_t)va);
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m)  \ast \texttt{rdi}\mapsto_{r} \textsf{plm4} \ast \texttt{rsi}\mapsto_{r} \textsf{va} \ast  \texttt{rdx}\mapsto_{r} \textsf{pa} \ast \texttt{rcx}\mapsto_{r} \textsf{perm} \ast \texttt{rax}\mapsto_{\textsf{r}} \_ \ast \ulcorner \texttt{aligned va} \land  \theta \; !!\; \vaddr = \texttt{None}\urcorner\ast }_{\rtv}$
+$\specline{\texttt{rbp-8}\mapsto_{v} \textsf{plm4} \ast \texttt{rbp-16}\mapsto_{v} \textsf{va}  \ast \texttt{rbp-24}\mapsto_{v} \textsf{pa} \ast \texttt{rbp-32}\mapsto_{v} \textsf{perm} \ast \textsf{r8} \mapsto_{\textsf{r}} \textsf{\_}}_{\rtv}$
+;;char *a; pte_t *pte;
+;;a = (char *)PGROUNDUP((uintptr_t)va);
+...
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast  \texttt{rdi}\mapsto_{\textsf{r}} \vaddr  \;\ast}_{\rtv}$
+$\specline{\texttt{pte\_addr} \mapsto_{\texttt{vpte}} \paddr\; (\texttt{wzero 64})  \ast \ulcorner \texttt{entry\_present } \entryf \land \texttt{entry\_present } \entrytr \land  \texttt{entry\_present } \entrytw \urcorner}_{\rtv}$
+$\specline{\exists \texttt{ fpaddr} \ldotp \ulcorner \texttt{aligned fpaddr} \urcorner \ast \texttt{rax}\mapsto_{\textsf{r}} \texttt{fpaddr+3} \ast \texttt{fpaddr} \mapsto_{\textsf{p}} (\texttt{wzero 64}) \ast \ulcorner \texttt{entry\_present (fpaddr+3)}\urcorner}_{\rtv}$
+;;pte = walkpgdir_succ(pml4, a);
+mov    -0x8[rbp],rdi
+mov    -0x28[rbp],rsi
+mov    0x0,rax
+callq  291 <vaspace_mappage+0x41>	28d: R_X86_64_PLT32	walkpgdir_succ-0x4
+mov    0x0,r8
+mov    rax,rcx
+mov    rcx,-0x30[rbp]
+;;pte->writable     = PTE_WRITEABLE_SET(perm);
+...
+pte->user_acc     = PTE_USER_ACC_SET(perm);
+...
+;;pte->writethrough = PTE_WRITETHRU_SET(perm);
+...
+;;pte->nocache      = PTE_NOCACHE_SET(perm);
+pte->pfn          = PTE_ADDR_TO_PFN(pa);
+mov    -0x18[rbp],rcx
+shr    0xc,rcx
+mov    -0x30[rbp],rdx
+mov    [rdx],rsi
+movabs 0xffffffffff,rdi
+and    rdi,rcx
+shl    0xc,rcx
+movabs 0xfff0000000000fff,rdi
+and    rdi,rsi
+or     rcx,rsi
+mov    rsi,[rdx]
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast  \texttt{r14}\mapsto_{\textsf{r}} \texttt{pte\_addr} \ast \texttt{rdi}\mapsto_{\textsf{r}} \vaddr \;\ast \nfpointsto{\mask\vaddr\maskfour\rtv}{\mask\vaddr\maskfouroff\rtv}\entryf\qone\naddr \ast}_{\rtv}$
+$\specline{  \nfpointsto{\mask\vaddr\maskthree\entryf}{\mask\vaddr\maskthreeoff\entryf}\entrytr\qtwo\naddr \ast \nfpointsto{\mask\vaddr\masktwo\entrytr}{\mask\vaddr\masktwooff\entrytr}\paddr\qthree\entryo \;\ast}_{\rtv}$
+$\specline{\texttt{pte\_addr} \mapsto_{\texttt{vpte}} \paddr \;(\texttt{fpaddr+3}) \; \ast \ulcorner \texttt{entry\_present } \entryf \land \texttt{entry\_present } \entrytr \land  \texttt{entry\_present } \entrytw \urcorner }_{\rtv}$
+$\specline{\ulcorner \texttt{aligned fpaddr} \urcorner \ast \texttt{rax}\mapsto_{\textsf{r}} \texttt{fpaddr+3} \ast \texttt{fpaddr} \mapsto_{\textsf{p}} (\texttt{wzero 64}) \ast \ulcorner \texttt{entry\_present fpaddr+3}\urcorner}_{\rtv}$
+$\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\; \sqsubseteq $
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast  \texttt{r14}\mapsto_{\textsf{r}} \texttt{pte\_addr} \ast \texttt{rdi}\mapsto_{\textsf{r}} \vaddr \ast }_{\rtv}$
+$\specline{\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\vaddr,\entryf,\entrytr,\entrytw,\fpaddr+3) \ast \ulcorner \theta \;!!\;\vaddr = \texttt{None}\urcorner \; \ast}_{\rtv}$
+$\specline{\ulcorner \texttt{aligned fpaddr} \urcorner \ast \texttt{rax}\mapsto_{\textsf{r}} \texttt{fpaddr+3} \ast \texttt{fpaddr} \mapsto_{\textsf{p}} (\texttt{wzero 64}) }_{\rtv}$
+$\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\; \sqsubseteq $
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace} (<[\vaddr:=\texttt{fpaddr}]> \theta,m) \ast}_{\rtv}$
+$\specline{\ulcorner \texttt{aligned fpaddr} \urcorner \ast \texttt{fpaddr} \mapsto_{\textsf{p}} \textsf{ wzero 64} \ast \sumapaces\rtv\delta  \ast\sumwalkabs\vaddr\qfrac\fpaddr}_{\rtv}$
+$\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\; \sqsubseteq $
+  $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace} (<[\vaddr:=\texttt{fpaddr}]> \theta,m) \ast \vaddr \mapsto_{\textsf{vpte}}\; \{\qfrac\} \;\fpaddr \textsf{ wzero 64}}_{\rtv}$
+$\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\; \sqsubseteq $
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace} (<[\vaddr:=\texttt{fpaddr}]> \theta,m) \ast \vaddr \mapsto_{\textsf{v}}\; \{\qfrac\} \textsf{wzero 64}}_{\rtv}$
+;;pte->nx           = PTE_NX_SET(perm);
+;;pte->present      = 1;
+;;clean up the stack and return
+...
+;;}
+\end{lstlisting}
+  \caption{Specification and proof of code for mapping a new page.}
+\label{fig:mapping_code}
+\end{figure}
+
+In Figure \ref{fig:mapping_code}, we see an address ($\vaddr$) currently not
+mapped to a page ($\theta \; !!\; \vaddr = \texttt{None}$). Mapping a fresh
+phyiscal page to back the desired virtual page first requires ensuring
+the existence of a memory location for an appropriate L1 table entry.
+The code uses a helper function \lstinline{pte\_walkpgdir} (discussed again in Section \ref{sec:traversing}).
+\textsf{pte\_walkpgdir}'s postcondition contains virtual \emph{PTE} pointsto assertions ($\mapsto_{\textsf{vpte}}$) \todo{colin, these are no more plain physical pointstos}  both for ensuring partial page table walk reaching the
+L1 entry (l1e) with the justification of higher levels of the page table exist, and for allowing access to the memory of the L1 entry via virtual address.
+A PTE points-to is defined just like the normal virtual points-to of Figure \ref{fig:virtualpointstosharing}, except the physical address (here, \textsf{pa}) is explicit in the assertion 
+rather than existentially quantified:
+ \[
+\begin{array}{l}
+    \vaddr\mapsto_{\textsf{vpte,q}} \; \paddr \; \vpage : \mathsf{vProp}~\Sigma \stackrel{\triangle}{=} 
+    \exists \delta\ldotp
+	(\lambda \mathit{cr3val}\ldotp
+	\sumapaces{\mathit{cr3val}}\delta) \ast 
+  \sumwalkabs\vaddr\qfrac\paddr \ast \paddr \mapsto_{\mathsf{p}} \vpage
+\end{array}
+\]
+This supports rules for accessing memory
+at that virtual address, but exposing the physical location being modified
+makes this convenient to use for page table modifications, since we must ensure
+the modified data is at the specific physical location that will be used as the L1 entry.
+This is also guaranteed ($ \ulcorner
+\texttt{addr\_L1}(\vaddr,\entryo) = \paddr\urcorner$).
+% After obtaining a virtual address \textsf{pte\_addr} in \textsf{rax} backed 
+% by the physical memory for the L1 entry that will be used to translate the virtual addresses
+% we are mapping, we save it to \textsf{r14} to be updated later in Line 9.
+
+%In the precondition, we see Line 12 allocates a fresh page-aligned, zero-initialized page  (at \textsf{fpaddr}),
+%returning a pre-filled PTE entry in \textsf{rax} ($+3$ sets the lower 2 bits).
+
+The crucial step is actually updating the L1 entry (Line X\todo{X} in Figure \ref{fig:mapping_code}),
+via the virtual address
+(\textsf{pte\_addr}) known to translate to the appropriate physical address, in our example the L1
+table entry address ($\textsf{addr\_L1}(\textsf{va, l1e})$), to hold the freshly
+allocated physical page address (\textsf{fpaddr}) in Line X.
+After this write, we have (Specification Line Y)
+\[\texttt{pte\_addr} \mapsto_{\texttt{vpte}}  \; \paddr\; \textsf{fpaddr}  \qquad\qquad (\textsf{by the PTE variant of Rule}~\TirNameStyle{WriteToVirtMemFromReg})\]
+
+At this point in the proof,  $\texttt{pte\_addr} \mapsto_{\texttt{vpte}}  \; \paddr\; \textsf{fpaddr})$ 
+contains  full ownerhsip of the L1 table entry physical pointsto $\paddr
+\mapsto_{\mathsf{p}} \textsf{fpaddr}$ . Then, we can split the full-ownership of that
+physical pointsto
+($\textsf{fpaddr}\mapsto_{\textsf{p}} \;\textsf{wzero 64}$) to obtain
+fractional ownership on it appropriate for a single address's share of an L1 entry (q4).
+Since we know $\texttt{addr\_L1 }(\vaddr,
+\entryo) \mapsto_{\mathsf{p}} \{q4\} \;\textsf{fpaddr}$) from
+$\mapsto_{\textsf{vpte}}$, after the split of the PTE points-to, and, thanks to the post-condition of the page-table walk (Specification of \textsf{pte\_walkpgdir} in Figure \ref{}), we know the existence of L4-L2 mappings reaching to the L1 exist, so that we can add the last mapping $\texttt{fpaddr} \mapsto_{p} (\texttt{wzero 64})$ to obtain the complete table-walk.
+\todo{with the virtual-pte-pointsto assertions (L4\_L1\_PointsTo($\vaddr$ l4e l3e l2e
+fpaddr)) in Line 25}
+
+Finally, we can insert $\vaddr$ into the ghost page-table-walk summarization
+map ($\theta$) to change its state from unmapped ($\ulcorner \theta
+\;!!\;\vaddr = \texttt{None}\urcorner$ (Line 28) and the
+precondition) to mapped (Line 29) using Iris' ghost-map update
+($\sqsubseteq$), and construct a PTE -points-to $\vaddr
+\mapsto_{\textsf{vpte}}\; \qfrac \;\fpaddr\;(\textsf{wzero 64})$ (Line 31) using $\sumwalkabs\vaddr\qfrac\fpaddr$ obtained from ghost
+page-table-walk insertions, $\sumapaces\rtv\delta$ from unfolding the
+definition of $\mapsto_{\textsf{vpte}}$, and $\fpaddr
+\mapsto_{\textsf{p}} \textsf{wzero 64}$.
+Finally, by existentially quantifying away \textsf{fpaddr},
+we can shift the PTE points-to into a normal virtual points-to.
+
+For brevity this proof only maps the first word of the allocated page; the generalization to the full page
+is straightforward use of iterated separating conjunction to manage collections of clusters of 512 things
+(slices of the L1 entry physical ownership, virtual addresses at 8 byte increments, etc.).
+\looseness=-1
+
+Unlike the only prior work verifying analogous code for mapping a new page~\cite{kolanski08vstte,kolanski09tphols}, our proof above
+does \emph{not} need to unfold the definitions of the logic,
+but instead uses logic rules throughout to reason about the effect of the mapping. 
+This is the first proof we know of for mapping a virtual memory page that does not resort to direct reasoning over the operational semantics.
+% By incorporating verification of the
+% \lstinline|ensure_L1| function (see Section \ref{sec:traversing}), our verification also directly handles several subtle aspects which
+% were axiomatized in prior work.
 

### 2024-07-04 examples.tex
a final discussion on vaspace\_mapapage entries vs physicall addrs

@@
 %$\assert{\ulcorner \ptablestore !! \textsf{maddr} = \textsf{None} \urcorner \ast \ownGhost\gammaPred{\authfull{\ptableabswalk\ptablestore}} \ast \textsf{Pf}} $
+\newcommand{\sumwalkabsent}{
+  \ownGhost\gammaPred{\authfrag{\singletonMap{\texttt{entry+KERNBASE}}{(\textsf{qfrac}, \textsf{entry})}}}
+}
+
+\newcommand{\ventry}{\texttt{entry + KERNBASE}}
+\newcommand{\entry}{\texttt{entry}}
+\newcommand{\qfraczero}{\textsf{qfrac}}
+\newcommand{\true}{\textsf{true}}
+\tikzstyle{boxedassert_border} = [sharp corners,line width=0.2pt]
+\NewDocumentCommand \boxedassertpv {O{} m o}{%
+	\tikz[baseline=(m.base)]{
+		%	  \node[rectangle, draw,inner sep=0.8pt,anchor=base,#1] (m) {${#2}\mathstrut$};
+		\node[rectangle,inner sep=1.5pt,outer sep=0.2pt,anchor=base] (m) {${\,#2\,}\mathstrut$};
+		\draw[#1,boxedassert_border] ($(m.south west)$) rectangle ($(m.north east)$);
+	}\IfNoValueF{#3}{^{\,#3}}%
+}
+\newcommand*{\knowInvpv}[2]{\boxedassertpv{#2}[#1]}
+\newcommand*{\ownGhostpv}[2]{\boxedassertpv[dash dot]{#2}[#1]}
+
+\newcommand{\sumpv}[3]{
+  \ownGhostpv\gammaPred{\authfrag{\singletonMap{#1}{(#2, #3)}}}
+}
+
+\newcommand{\pvmapping}[1]{\mathcal{A}\textsf{PVMappings}(#1)}
+
+\newcommand{\fpaddr}{\texttt{fpaddr}}
+\newcommand{\specline}[1]{{\color{blue}\left\{#1\right\}}}
+\newcommand{\sumapacesfull}[2]{
+  \ownGhost\gammaPreds{\authfull{\singletonMap{#1}{#2}}}
+}
 \section{Experiments}
@@
 $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)   \ast \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{next} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{table\_addr}  }_{\rtv}$
-$\specline{\texttt{rax} \mapsto_{\textsf{v}} \textsf{nextpaddr} \ast \ulcorner \Xi !!
-\mathsf{nextpaddr} = None \urcorner \ast \textsf{table\_addr} \mapsto_{\mathsf{v}} \mathsf{\_} \ast \sumapaces\rtv\delta }_{\rtv}$
+$\specline{\texttt{rax} \mapsto_{\textsf{v}} \textsf{nextpaddr} \ast \ulcorner \Xi \;!!\; \mathsf{nextpaddr} = \textsf{None} \urcorner \ast \textsf{table\_addr} \mapsto_{\mathsf{v}} \mathsf{\_} \ast \sumapaces\rtv\delta }_{\rtv}$
 $\sqsubseteq$
 $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\cup\{\mathsf{nextpaddr}\},m)   \ast \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{next} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val}  }_{\rtv}$
-$\specline{\texttt{rax} \mapsto_{\textsf{v}} \textsf{nextpaddr} \ast \ulcorner \Xi !!
-\mathsf{nextpaddr} = Z \urcorner \ast \textsf{table\_addr} \mapsto_{\mathsf{v}} \mathsf{\_} \ast \sumapaces\rtv\delta }_{\rtv}$
+$\specline{\texttt{rax} \mapsto_{\textsf{v}} \textsf{nextpaddr} \ast \ulcorner \Xi \;!!\; \mathsf{nextpaddr} = Z \urcorner \ast \textsf{table\_addr} \mapsto_{\mathsf{v}} \mathsf{\_} \ast \sumapaces\rtv\delta }_{\rtv}$
 mov    rax,-0x10[rbp] 
 $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\cup\{\mathsf{nextpaddr}\},m)   \ast \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{nextpaddr} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{table\_addr}  }_{\rtv}$
-$\specline{\texttt{rax} \mapsto_{\textsf{v}} \textsf{nextpaddr} \ast \ulcorner \Xi !!
-\mathsf{nextpaddr} = Z \urcorner \ast \textsf{table\_addr} \mapsto_{\mathsf{v}} \mathsf{\_}\ast \sumapaces\rtv\delta }_{\rtv}$
+$\specline{\texttt{rax} \mapsto_{\textsf{v}} \textsf{nextpaddr} \ast \ulcorner \Xi \;!!\; \mathsf{nextpaddr} = Z \urcorner \ast \textsf{table\_addr} \mapsto_{\mathsf{v}} \mathsf{\_}\ast \sumapaces\rtv\delta }_{\rtv}$
 ;;*table = local;
@@
 $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\cup\{\mathsf{nextpaddr}\},m)   \ast \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{nextpaddr} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{table\_addr}  }_{\rtv}$
-$\specline{\texttt{rax} \mapsto_{\textsf{v}} \textsf{nextpaddr} \ast \ulcorner \Xi !!
-\mathsf{nextpaddr} = Z \urcorner \ast \textsf{table\_addr} \mapsto_{\mathsf{v}} \mathsf{\mathsf{nextpaddr}} \ast \sumapaces\rtv\delta }_{\rtv}$
+$\specline{\texttt{rax} \mapsto_{\textsf{v}} \textsf{nextpaddr} \ast \ulcorner \Xi \;!!\; \mathsf{nextpaddr} = Z \urcorner \ast \textsf{table\_addr} \mapsto_{\mathsf{v}} \mathsf{\mathsf{nextpaddr}} \ast \sumapaces\rtv\delta }_{\rtv}$
 ;;return;
@@
 to map new pages into an address space by writing into an existing page table via a call
-\[\textsf{vaspace\_mappage(pte\_t *pml4, void *va,uintptr\_t pa, uint64\_t perm )}\] in Figure \ref{fig:mapping_code}).
-To do so, with a given allocated a fresh page (\textsf{pa}), then calculate the appropriate
+\[\textsf{vaspace\_mappage(pte\_t *pml4, void *va,uintptr\_t fpaddr, uint64\_t perm )}\] in Figure \ref{fig:mapping_code}).
+To do so, with a given allocated a fresh page (\textsf{fpaddr}), then calculate the appropriate
 known-valid page table walks (\textsf{pte\_walkpgdir} Line X in Figure \ref{fig:mapping_code})  and update the appropriate L1 page table entry (permissions and frame address Lines in Figure \ref{fig:mapping_code}).
@@
 %}
-\newcommand{\sumwalkabsent}{
-  \ownGhost\gammaPred{\authfrag{\singletonMap{\texttt{entry+KERNBASE}}{(\textsf{qfrac}, \textsf{entry})}}}
-}
-
-\newcommand{\ventry}{\texttt{entry + KERNBASE}}
-\newcommand{\entry}{\texttt{entry}}
-\newcommand{\qfraczero}{\textsf{qfrac}}
-\newcommand{\true}{\textsf{true}}
-\tikzstyle{boxedassert_border} = [sharp corners,line width=0.2pt]
-\NewDocumentCommand \boxedassertpv {O{} m o}{%
-	\tikz[baseline=(m.base)]{
-		%	  \node[rectangle, draw,inner sep=0.8pt,anchor=base,#1] (m) {${#2}\mathstrut$};
-		\node[rectangle,inner sep=1.5pt,outer sep=0.2pt,anchor=base] (m) {${\,#2\,}\mathstrut$};
-		\draw[#1,boxedassert_border] ($(m.south west)$) rectangle ($(m.north east)$);
-	}\IfNoValueF{#3}{^{\,#3}}%
-}
-\newcommand*{\knowInvpv}[2]{\boxedassertpv{#2}[#1]}
-\newcommand*{\ownGhostpv}[2]{\boxedassertpv[dash dot]{#2}[#1]}
-
-\newcommand{\sumpv}[3]{
-  \ownGhostpv\gammaPred{\authfrag{\singletonMap{#1}{(#2, #3)}}}
-}
-
-\newcommand{\pvmapping}[1]{\mathcal{A}\textsf{PVMappings}(#1)}
-
-\newcommand{\fpaddr}{\texttt{fpaddr}}
-\newcommand{\specline}[1]{{\color{blue}\left\{#1\right\}}}
-\newcommand{\sumapacesfull}[2]{
-  \ownGhost\gammaPreds{\authfull{\singletonMap{#1}{#2}}}
-}
 \todo{ismail change the following spec base on the changes you have in the code}
@@
   \begin{lstlisting}[mathescape]
-;;complstatus_t vaspace_mappage(pte_t *pml4, void *va,uintptr_t pa, uint64_t perm ) {
+;;complstatus_t vaspace_mappage(pte_t *pml4, void *va,uintptr_t fpaddr, uint64_t perm ) {
 ...
@@
 ;;a = (char *)PGROUNDUP((uintptr_t)va);
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m)  \ast \texttt{rdi}\mapsto_{r} \textsf{plm4} \ast \texttt{rsi}\mapsto_{r} \textsf{va} \ast  \texttt{rdx}\mapsto_{r} \textsf{pa} \ast \texttt{rcx}\mapsto_{r} \textsf{perm} \ast \texttt{rax}\mapsto_{\textsf{r}} \_ \ast \ulcorner \texttt{aligned va} \land  \theta \; !!\; \vaddr = \texttt{None}\urcorner\ast }_{\rtv}$
-$\specline{\texttt{rbp-8}\mapsto_{v} \textsf{plm4} \ast \texttt{rbp-16}\mapsto_{v} \textsf{va}  \ast \texttt{rbp-24}\mapsto_{v} \textsf{pa} \ast \texttt{rbp-32}\mapsto_{v} \textsf{perm} \ast \textsf{r8} \mapsto_{\textsf{r}} \textsf{\_}}_{\rtv}$
 ;;char *a; pte_t *pte;
@@
 ...
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast  \texttt{rdi}\mapsto_{\textsf{r}} \vaddr  \;\ast}_{\rtv}$
-$\specline{\texttt{pte\_addr} \mapsto_{\texttt{vpte}} \paddr\; (\texttt{wzero 64})  \ast \ulcorner \texttt{entry\_present } \entryf \land \texttt{entry\_present } \entrytr \land  \texttt{entry\_present } \entrytw \urcorner}_{\rtv}$
-$\specline{\exists \texttt{ fpaddr} \ldotp \ulcorner \texttt{aligned fpaddr} \urcorner \ast \texttt{rax}\mapsto_{\textsf{r}} \texttt{fpaddr+3} \ast \texttt{fpaddr} \mapsto_{\textsf{p}} (\texttt{wzero 64}) \ast \ulcorner \texttt{entry\_present (fpaddr+3)}\urcorner}_{\rtv}$
 ;;pte = walkpgdir_succ(pml4, a);
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m)  \ast \texttt{rdi}\mapsto_{r} \textsf{plm4} \ast \texttt{rsi}\mapsto_{r} \textsf{va} \ast  \texttt{rdx}\mapsto_{r} \textsf{pa} \ast \texttt{rcx}\mapsto_{r} \textsf{perm} \ast \texttt{rax}\mapsto_{\textsf{r}} \textsf{pte\_addr} \ast }_{\rtv}$
+$\specline{\ulcorner \texttt{aligned va} \land  \theta \; !!\; \vaddr = \texttt{None}\urcorner\ast \textsf{vl4} \mapsto_{\textsf{vpte,qfrac}} \textsf{l4e l4} \ast \textsf{vl3} \mapsto_{\textsf{vpte,qfrac}} \textsf{l3e l3} \ast \textsf{vl2} \mapsto_{\textsf{vpte,qfrac}} \textsf{l2e pa}  \ast }_{\rtv}$
+$\specline{\textsf{pte\_addr} \mapsto_{\textsf{vpte,qfrac}} \textsf{pa} \textsf{pte\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \lnot (\textsf{entry\_present pte\_val}) \urcorner}_{\rtv}$
+$\specline{\texttt{rbp-8}\mapsto_{v} \textsf{plm4} \ast \texttt{rbp-16}\mapsto_{v} \textsf{va}  \ast \texttt{rbp-24}\mapsto_{v} \textsf{pa} \ast \texttt{rbp-32}\mapsto_{v} \textsf{perm} \ast \textsf{r8} \mapsto_{\textsf{r}} \textsf{\_}}_{\rtv}$
 mov    -0x8[rbp],rdi
@@
 callq  291 <vaspace_mappage+0x41>	28d: R_X86_64_PLT32	walkpgdir_succ-0x4
+;;if(pte->present != 0) {
+...        
 mov    0x0,r8
@@
 ...
-pte->user_acc     = PTE_USER_ACC_SET(perm);
+;;pte->user_acc     = PTE_USER_ACC_SET(perm);
 ...
@@
 ;;pte->nocache      = PTE_NOCACHE_SET(perm);
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m)  \ast \texttt{rdi}\mapsto_{r} \textsf{plm4} \ast \texttt{rsi}\mapsto_{r} \textsf{va} \ast  \texttt{rdx}\mapsto_{r} \textsf{pa} \ast \texttt{rcx}\mapsto_{r} \textsf{perm} \ast \texttt{rax}\mapsto_{\textsf{r}} \textsf{pte\_addr} \ast \ulcorner \texttt{aligned va} \land  \theta \; !!\; \vaddr = \texttt{None}\urcorner\ast }_{\rtv}$
+$\specline{\textsf{pte\_addr} \mapsto_{\textsf{vpte,qfrac}} \textsf{pa} \textsf{pte\_val} \ast \ulcorner \lnot \textsf{qfrac} = 1 \land \lnot(\textsf{entry\_present pte\_val}) \urcorner}$     
+$\specline{\texttt{rbp-8}\mapsto_{v} \textsf{plm4} \ast \texttt{rbp-16}\mapsto_{v} \textsf{va}  \ast \texttt{rbp-24}\mapsto_{v} \textsf{pa} \ast \texttt{rbp-32}\mapsto_{v} \textsf{perm} \ast \textsf{r8} \mapsto_{\textsf{r}} \textsf{\_}}_{\rtv}$
 pte->pfn          = PTE_ADDR_TO_PFN(pa);
@@
 mov    [rdx],rsi
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m)  \ast \texttt{rdi}\mapsto_{r} \textsf{plm4} \ast \texttt{rsi}\mapsto_{r} \textsf{va} \ast  \texttt{rdx}\mapsto_{r} \textsf{pa} \ast \texttt{rcx}\mapsto_{r} \textsf{perm} \ast \texttt{rax}\mapsto_{\textsf{r}} \textsf{pte\_addr} \ast \ulcorner \texttt{aligned va} \land  \theta \; !!\; \vaddr = \texttt{None}\urcorner\ast }_{\rtv}$
+$\specline{\textsf{pte\_addr} \mapsto_{\textsf{vpte,qfrac}} \textsf{pa} (\textsf{pa}>>12) \ast \ulcorner \lnot \textsf{qfrac} = 1 \land \lnot(\textsf{entry\_present pte\_val}) \urcorner}$     
+$\specline{\texttt{rbp-8}\mapsto_{v} \textsf{plm4} \ast \texttt{rbp-16}\mapsto_{v} \textsf{va}  \ast \texttt{rbp-24}\mapsto_{v} \textsf{pa} \ast \texttt{rbp-32}\mapsto_{v} \textsf{perm} \ast \textsf{r8} \mapsto_{\textsf{r}} \textsf{\_}}_{\rtv}$
 movabs 0xffffffffff,rdi
@@
 mov    rsi,[rdx]
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast  \texttt{r14}\mapsto_{\textsf{r}} \texttt{pte\_addr} \ast \texttt{rdi}\mapsto_{\textsf{r}} \vaddr \;\ast \nfpointsto{\mask\vaddr\maskfour\rtv}{\mask\vaddr\maskfouroff\rtv}\entryf\qone\naddr \ast}_{\rtv}$
-$\specline{  \nfpointsto{\mask\vaddr\maskthree\entryf}{\mask\vaddr\maskthreeoff\entryf}\entrytr\qtwo\naddr \ast \nfpointsto{\mask\vaddr\masktwo\entrytr}{\mask\vaddr\masktwooff\entrytr}\paddr\qthree\entryo \;\ast}_{\rtv}$
-$\specline{\texttt{pte\_addr} \mapsto_{\texttt{vpte}} \paddr \;(\texttt{fpaddr+3}) \; \ast \ulcorner \texttt{entry\_present } \entryf \land \texttt{entry\_present } \entrytr \land  \texttt{entry\_present } \entrytw \urcorner }_{\rtv}$
-$\specline{\ulcorner \texttt{aligned fpaddr} \urcorner \ast \texttt{rax}\mapsto_{\textsf{r}} \texttt{fpaddr+3} \ast \texttt{fpaddr} \mapsto_{\textsf{p}} (\texttt{wzero 64}) \ast \ulcorner \texttt{entry\_present fpaddr+3}\urcorner}_{\rtv}$
-$\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\; \sqsubseteq $
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast  \texttt{r14}\mapsto_{\textsf{r}} \texttt{pte\_addr} \ast \texttt{rdi}\mapsto_{\textsf{r}} \vaddr \ast }_{\rtv}$
-$\specline{\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\vaddr,\entryf,\entrytr,\entrytw,\fpaddr+3) \ast \ulcorner \theta \;!!\;\vaddr = \texttt{None}\urcorner \; \ast}_{\rtv}$
-$\specline{\ulcorner \texttt{aligned fpaddr} \urcorner \ast \texttt{rax}\mapsto_{\textsf{r}} \texttt{fpaddr+3} \ast \texttt{fpaddr} \mapsto_{\textsf{p}} (\texttt{wzero 64}) }_{\rtv}$
-$\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\; \sqsubseteq $
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace} (<[\vaddr:=\texttt{fpaddr}]> \theta,m) \ast}_{\rtv}$
-$\specline{\ulcorner \texttt{aligned fpaddr} \urcorner \ast \texttt{fpaddr} \mapsto_{\textsf{p}} \textsf{ wzero 64} \ast \sumapaces\rtv\delta  \ast\sumwalkabs\vaddr\qfrac\fpaddr}_{\rtv}$
-$\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\; \sqsubseteq $
-  $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace} (<[\vaddr:=\texttt{fpaddr}]> \theta,m) \ast \vaddr \mapsto_{\textsf{vpte}}\; \{\qfrac\} \;\fpaddr \textsf{ wzero 64}}_{\rtv}$
-$\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\; \sqsubseteq $
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace} (<[\vaddr:=\texttt{fpaddr}]> \theta,m) \ast \vaddr \mapsto_{\textsf{v}}\; \{\qfrac\} \textsf{wzero 64}}_{\rtv}$
-;;pte->nx           = PTE_NX_SET(perm);
+;;}
 ;;pte->present      = 1;
@@
 fractional ownership on it appropriate for a single address's share of an L1 entry (q4).
+\todo{ismail comment out the rest of this subsection and fix according to the new map-a-page}
+\begin{comment}
 Since we know $\texttt{addr\_L1 }(\vaddr,
@@
 \mapsto_{\textsf{p}} \textsf{wzero 64}$.
+\end{comment}
 Finally, by existentially quantifying away \textsf{fpaddr},
@@
 \begin{comment}
+\[  
 $\specline{\exists (\entryf ,\;\entrytr,\; \entrytw,\; \entryo,\;\textsf{pte\_addr },\paddr) \; \ldotp\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast  \texttt{r14}\mapsto_{\textsf{r}} \_ \ast \texttt{rdi}\mapsto_{r} \vaddr \ast \texttt{rax}\mapsto_{\textsf{r}} \textsf{ pte\_addr} \; \ast }_{\rtv}$

### 2024-07-08 examples.tex
+

@@
 callq  70 <pte_initialize>
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \}\cup\{\mathsf{nextpaddr}\},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val}  }_{\rtv}$
-$\specline{\sumapaces\rtv\delta \ast \mathsf{nextpaddr} \in \Xi\setminus \{ \mathsf{entry} \} \cup \{ \mathsf{nextpaddr} \} }_{\rtv}$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val}  }_{\rtv}$
+$\specline{\sumapaces\rtv\delta \ast \mathsf{nextpaddr} \in \Xi\setminus \{ \mathsf{entry} \} \} }_{\rtv}$
 $\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \mathsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \land \lnot(\textsf{entry\_present entry\_val})  \urcorner }_{\rtv}$
-$\specline{ \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{nextpaddr} \ast \ulcorner \Xi\setminus \{ \mathsf{entry} \}\cup\{\mathsf{nextpaddr}\} \;!!\; \mathsf{entry} = \mathsf{Some Z} \urcorner \ast \texttt{rax} \mapsto_{\textsf{r}} \textsf{nextpaddr}  }_{\rtv}$
+$\specline{ \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{nextpaddr} \ast  \mathsf{nextpaddr} \mapsto_{\textsf{p}} \textsf{w0\_64} \ast \texttt{rax} \mapsto_{\textsf{r}} \textsf{nextpaddr}  }_{\rtv}$
 ;;entry value updates
@@
 ;;entry->pfn = next
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \}\cup\{\mathsf{nextpaddr}\},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val'}  }_{\rtv}$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val'}  }_{\rtv}$
 $\specline{\sumapaces\rtv\delta \ast \mathsf{nextpaddr} \in \Xi\setminus \{ \mathsf{entry} \} \cup \{ \mathsf{nextpaddr} \} }_{\rtv}$
 $\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \mathsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \land (\textsf{entry\_present entry\_val'})  \urcorner }_{\rtv}$
-$\specline{ \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{nextpaddr} \ast \ulcorner \Xi\setminus \{ \mathsf{entry} \}\cup\{\mathsf{nextpaddr}\} \;!!\; \mathsf{entry} = \mathsf{Some Z} \urcorner \ast \texttt{rax} \mapsto_{\textsf{r}} \textsf{nextpaddr}  }_{\rtv}$
+$\specline{ \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{nextpaddr} \ast  \mathsf{nextpaddr} \mapsto_{\textsf{p}} \textsf{w0\_64} \ast \texttt{rax} \mapsto_{\textsf{r}} \textsf{nextpaddr}  }_{\rtv}$
 ;;}
@@
 with DMA devices.}
+\todo{ismail: CHnage the text!  it is not as simple as relating paddrs existence type, more concrete relation on pfn and fresly allocated addr }
 \begin{figure*}
@@
   \bigast{(\paddr,\mathsf{T}) \in \Xi}{\exists\; (\textsf{qfrac, q, val,}\vaddr) \ldotp \ulcorner \vaddr = \paddr + \textsf{KERNBASE} \urcorner \ast \ulcorner \textsf{qfrac} = 1 \leftrightarrow \; \lnot\textsf{entry\_present }(\vale) \urcorner \ast
-    \ulcorner \mathsf{T} = \mathsf{Z} \leftrightarrow \mathsf{val} = \mathsf{w0\_64}\urcorner \; \ast\\
+   \exists \textsf{freshaddr, entry} \textsf{entry\_pfn} entry freshaddr = paddr \ast freshaddr \mapsto_{\textsf{p}} \textsf{w0\_64}  \leftrightarrow \mathsf{val} = \mathsf{w0\_64}\urcorner \; \ast\\
    \qquad\qquad\qquad\qquad\qquad\qquad\qquad\qquad   \underbrace{\sumwalkabs\vaddr\qfrac\paddr }_\text{Ghost translation} \ast 
@@
 
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)   \ast \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{next} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{table\_addr}  }_{\rtv}$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)   \ast \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{\_} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{table\_addr}  }_{\rtv}$
 $\specline{\texttt{rax} \mapsto_{\textsf{v}} \textsf{\_} \ast \textsf{table\_addr} \mapsto_{\mathsf{v}} \mathsf{\_}\ast\sumapaces\rtv\delta }_{\rtv}$
@@
 callq  81 <pte_initialize+0x11>	7d: R_X86_64_PLT32	kalloc-0x4
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)   \ast \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{next} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{table\_addr}  }_{\rtv}$
-$\specline{\texttt{rax} \mapsto_{\textsf{v}} \textsf{nextpaddr} \ast \ulcorner \Xi \;!!\; \mathsf{nextpaddr} = \textsf{None} \urcorner \ast \textsf{table\_addr} \mapsto_{\mathsf{v}} \mathsf{\_} \ast \sumapaces\rtv\delta }_{\rtv}$
-$\sqsubseteq$
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\cup\{\mathsf{nextpaddr}\},m)   \ast \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{next} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val}  }_{\rtv}$
-$\specline{\texttt{rax} \mapsto_{\textsf{v}} \textsf{nextpaddr} \ast \ulcorner \Xi \;!!\; \mathsf{nextpaddr} = Z \urcorner \ast \textsf{table\_addr} \mapsto_{\mathsf{v}} \mathsf{\_} \ast \sumapaces\rtv\delta }_{\rtv}$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)   \ast \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{\_} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{table\_addr}  }_{\rtv}$
+$\specline{\texttt{rax} \mapsto_{\textsf{v}} \textsf{nextpaddr} \ast \textsf{table\_addr} \mapsto_{\mathsf{v}} \mathsf{\_} \ast \textsf{nextpaddr} \mapsto_{\textsf{p}} \textsf{w0\_64}\ast\sumapaces\rtv\delta }_{\rtv}$
 mov    rax,-0x10[rbp] 
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\cup\{\mathsf{nextpaddr}\},m)   \ast \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{nextpaddr} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{table\_addr}  }_{\rtv}$
-$\specline{\texttt{rax} \mapsto_{\textsf{v}} \textsf{nextpaddr} \ast \ulcorner \Xi \;!!\; \mathsf{nextpaddr} = Z \urcorner \ast \textsf{table\_addr} \mapsto_{\mathsf{v}} \mathsf{\_}\ast \sumapaces\rtv\delta }_{\rtv}$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)   \ast \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{nextpaddr} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{table\_addr}  }_{\rtv}$
+$\specline{\texttt{rax} \mapsto_{\textsf{v}} \textsf{nextpaddr} \ast \ulcorner \Xi \;!!\; \mathsf{nextpaddr} = Z \urcorner \ast \textsf{table\_addr} \mapsto_{\mathsf{v}} \mathsf{\_} \ast \textsf{nextpaddr} \mapsto_{\textsf{p}} \textsf{w0\_64} \ast \sumapaces\rtv\delta }_{\rtv}$
 ;;*table = local;
@@
 mov    rax,[rdi]
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\cup\{\mathsf{nextpaddr}\},m)   \ast \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{nextpaddr} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{table\_addr}  }_{\rtv}$
-$\specline{\texttt{rax} \mapsto_{\textsf{v}} \textsf{nextpaddr} \ast \ulcorner \Xi \;!!\; \mathsf{nextpaddr} = Z \urcorner \ast \textsf{table\_addr} \mapsto_{\mathsf{v}} \mathsf{\mathsf{nextpaddr}} \ast \sumapaces\rtv\delta }_{\rtv}$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)   \ast \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{nextpaddr} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{table\_addr}  }_{\rtv}$
+$\specline{\texttt{rax} \mapsto_{\textsf{v}} \textsf{nextpaddr}  \ast \textsf{table\_addr} \mapsto_{\mathsf{v}} \mathsf{\mathsf{nextpaddr}} \ast \textsf{nextpaddr} \mapsto_{\textsf{p}} \textsf{w0\_64} \ast \sumapaces\rtv\delta }_{\rtv}$
 ;;return;
@@
 \end{figure}
-
+\todo{ISMAIL: Change the following paragraphs and LInes 33 35  in \ref{fig:calltopteinitialize}, because wpdp2V is where we only know the relation between entry->pfn nextpaddr , consequently zeroed.
+Not here }
 Since we are using \textsf{pte\_initialize} for page-table address allocation, we must relate this newly allocated physical address to the identity mapping map $\Xi$ -- Specification Lines 33 and 35 in Figure \ref{fig:calltopteinitialize}.

### 2024-07-09 programlogic.tex
fix a few typos, note critical missing content in figure 2?

@@
 carry over to our derived logic.
-However, be must still build up from existing Iris primitives to provide new primitives that depend on the address space --- primarily the notion
+However, we must still build up from existing Iris primitives to provide new primitives that depend on the address space --- primarily the notion
 of virtual points-to.
@@
 Figure \ref{fig:strongvirtualpointsto} gives such a definition.
-It asserts the existence of for page table entries, one at each translation level, and via \lstinline|L4_L1_PointsTo|
+It asserts the existence of four page table entries, one at each translation level, and via \lstinline|L4_L1_PointsTo|
 asserts from the current \lstinline|cr3| value, taking the offsets from \textsf{va} (as in Figure \ref{fig:pagetables})
@@
 Most of the definition lives directly in \textsf{vProp}, using the separation logic structure lifted from Iris's \textsf{iProp}.
+\todo[inline]{Colin notes: the \lstinline|addr_L1| mentioned in the paragraph above doesn't appear in Figure 2. paddr isn't mentioned
+in the definition of \lstinline|L4_L1_PointsTo| in the figure. Shouldn't there be a pure \lstinline|addr_L1(va,l1e)=paddr| in there?
+}
 
@@
 
-  We take separate the physical page-table-walk  
+  We separate the physical page-table-walk  
   from the virtual pointsto relation, replacing it with ghost state that merely guarantees the address translation would succeed.

### 2024-07-09 programlogic.tex
revise handling of ghost maps in Section 4; needs a second pass to make it more non-Iris-expert-friendly, but it is better than all the references to auth and excl

@@
   so long as those updates preserve the same virtual-to-physical translations.
+
+  % iris.sty lacks nice syntax for the ghost maps
+  \newcommand{\ghostmaptoken}[3]{\ensuremath{#2\hookrightarrow^{#1}#3}}
+  \newcommand{\fracghostmaptoken}[4]{\ensuremath{#2\hookrightarrow^{#1}_{#4}#3}}
+
 \newcommand{\vale}{\textsf{val}}
@@
     \exists \delta\ldotp
-    \underbrace{(\lambda\mathit{cr3val}\ldotp\sumapaces{\mathit{cr3val}}\delta)}_\text{Find addr.\;space invariant} \ast 
-  \underbrace{\sumwalkabs\vaddr\qfrac\paddr }_\text{Ghost translation}\ast 
+    % \underbrace{(\lambda\mathit{cr3val}\ldotp\sumapaces{\mathit{cr3val}}\delta)}_\text{Find addr.\;space invariant} \ast 
+    \underbrace{(\lambda\mathit{cr3val}\ldotp\ghostmaptoken{\delta{}s}{cr3val}{\delta})}_\text{Find addr.\;space invariant} \ast 
+  % \underbrace{\sumwalkabs\vaddr\qfrac\paddr }_\text{Ghost translation}\ast 
+  \underbrace{\fracghostmaptoken{\delta}{\mathsf{va}}{\mathsf{pa}}{\mathsf{q}} }_\text{Ghost translation}\ast 
   \underbrace{\paddr \mapsto_{\mathsf{p}}\{\textsf{q}\}\; \vale}_\text{Physical location}
@@
   from the virtual pointsto relation, replacing it with ghost state that merely guarantees the address translation would succeed.
-  Iris includes a construction which permits tracking a ghost map 
-  from virtual addresses to the physical addresses they translate to, and has for each key in the map (i.e.,
-  each virtual address) a token. Updating the virtual-to-physical (ghost) mapping for a virtual address $va$
-  requires full ownership of the token. Then the \emph{virtual memory manager's invariant} ensures that for each $va\mapsto pa$ mapping
+  Iris includes a \emph{ghost map} construction which we use to track mappings from
+  virtual addresses to the physical addresses they translate to.
+  The implementation details in terms of named invariants and various resource algebras are not important to our presentation, so
+  we treat the maps abstractly.
+   The map includes for each key in the map (i.e.,
+  each virtual address) a token $\ghostmaptoken{\gamma}{k}{v}$ whose ownership is required to update that key-value pair in the ghost map
+  named $\gamma$. The existence of such a token implies that the actual map $\theta$ tracked by a corresponding $\mathsf{GhostMap}(\gamma,\theta)$
+  resource indeed maps $k$ to $v$. These properties are captured by the Iris rules:\footnote{Iris ghost maps lack established notation in the literature,
+  but the syntax we use captures the details of \texttt{iris.base\_logic.lib.ghost\_map}.}
+
+  \begin{mathpar}
+  \inferrule{ }{\mathsf{GhostMap}(\gamma,\theta) \ast \ghostmaptoken{\gamma}{\mathsf{pa}}{\mathsf{va}} \vs 
+  \mathsf{GhostMap}(\gamma,\theta[\mathsf{pa}\mapsto\mathsf{va'}]) \ast \ghostmaptoken{\gamma}{\mathsf{pa}}{\mathsf{va}'} 
+  }
+  \and
+  \inferrule{ }{\mathsf{GhostMap}(\gamma,\theta) \ast \ghostmaptoken{\gamma}{\mathsf{pa}}{\mathsf{va}} \wand \ulcorner \theta(\mathsf{pa})=\mathsf{Some}(\matsf{va}) \urcorner }
+  \end{mathpar}
+
+  % Updating the virtual-to-physical (ghost) mapping for a virtual address $va$
+  % requires full ownership of the token. 
+  Then the \emph{virtual memory manager's invariant} ensures that for each $\ghostmaptoken{\gamma}{va}{pa}$ mapping
   in this map, there are \emph{physical} resources sufficient to ensure that address translation for $va$
-  will resolve in the hardware to $pa$ --- $\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$.
+  will resolve in the hardware to $pa$ via --- $\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$.
   This allows modification of the page tables themselves as long as the changes do not invalidate overall virtual-to-physical
@@
 
-  % To do so, we introduce the protocol separating read-only (fragmental) and authorative (full ownership) accesses to the page tables, specifically $\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$.
-  % We realize this protocol by embedding it into an already existing \textsf{Iris}'s ghost-map (or view\_maps) construction
-  In Iris this is realized by using an authoritative resource algebra: there is a single \emph{authoritative} global copy of the (ghost)
-  map caching virtual-to-physical address translations, and for each entry a read only \emph{partial} ownership of that key-value pair.
-  The resource algebra itself is instantiated as:
-  \[\mathcal{A}\textsf{bsPTableWalk} \stackrel{\triangle}{=} \textsc{Auth} (\; \mathcal{W}_{64} \;\rightarrow_{\textrm{fin}} \;  ( (\textsc{Frac }, \mathord{+}) \times (\textsc{Agree } \Loc,\mathord{=}) ))\]
-We replace the physical
- $\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$ in the virtual points-to
-with
-fragmental ownership of the ghost-map ($ \sumwalkabs\vaddr\qfrac\paddr$),
-guaranteeing that ghost map contains an mapping from the virtual address ($\vaddr$) to a physical address ($\paddr$).
+  % In Iris this is realized by using an authoritative resource algebra: there is a single \emph{authoritative} global copy of the (ghost)
+  % map caching virtual-to-physical address translations, and for each entry a read only \emph{partial} ownership of that key-value pair.
+  % The resource algebra itself is instantiated as:
+  % \[\mathcal{A}\textsf{bsPTableWalk} \stackrel{\triangle}{=} \textsc{Auth} (\; \mathcal{W}_{64} \;\rightarrow_{\textrm{fin}} \;  ( (\textsc{Frac }, \mathord{+}) \times (\textsc{Agree } \Loc,\mathord{=}) ))\]
+For clarity we refer to the ghost map summarizing virtual-to-physical translations by 
+\[\mathcal{A}\textsf{bsPTableWalk}(\delta,\theta) \stackrel{\triangle}{=} \mathsf{GhostMap}(\delta,\theta)\]
+(omitting $\delta$ for brevity when only one is in scope)
+and keep this in a per-address-space invariant described shortly.
+We then replace the physical
+ $\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$ in the virtual points-to definition
+with ownership of the token \ghostmaptoken{\delta}{\mathsf{va}}{\mathsf{paddr}}, %ghost-map ($ \sumwalkabs\vaddr\qfrac\paddr$),
+guaranteeing that ghost map contains an mapping from the virtual address ($\vaddr$) to a physical address ($\paddr$),
+and thus that the per-address-space invariant contains the physical resources that guarantee the hardware resolves the translation.
 
@@
    \textsf{ASpace\_Lookup} (\ptablestore,m) \stackrel{\triangle}{=} \lambda\textsf{ cr3val} \ldotp \; \exists \gammaPred \; \ldotp \ulcorner m \; !!\; \textsf{cr3val} = \textsf{Some } \gammaPred \urcorner \ast
-    \ownGhost\gammaPred{\authfull{\ptableabswalk\ptablestore}}
+    %\ownGhost\gammaPred{\authfull{\ptableabswalk\ptablestore}}
+    \ptableabswalk{\delta,\theta}
 
@@
 
-Another major change is that, because our logic is the first to explicitly address the existence of multiple address spaces,
+Because our logic is the first to address the existence of multiple address spaces,
 the ghost map of virtual-to-physical translations (identified by the ghost name $\delta$ in Figure \ref{fig:virtualpointstosharing}
- \emph{for each address space}. This means we also require a way to locate the map for the current address space
-from the current root page table pointer stored in \lstinline|cr3|.
+ \emph{for each address space}) must be locateable from a representation of the specific address space:
+% This means we also require a way to locate the map for the current address space from 
+the current root page table pointer stored in \lstinline|cr3|.
 
 Figure \ref{fig:peraspaceinvariant} describes the resources for each address space.
-$\Theta$ is the logical map from virtual addresses to the physical addresses they should translate to, 
+$\theta$ is the logical map from virtual addresses to the physical addresses they should translate to, 
 for the \emph{current} address space, corresponding to the currently installed page tables (the ones
 indicated by \lstinline|cr3|). In Figure \ref{fig:virtualpointstosharing}, the existentially-quantified $\delta$
-is the ghost name for the authoritative copy of this map (again, for the current address space).
-$m$ is the (full) logical map from the various page table bases to ghost names for the 
-per-address-space mappings like $\Theta$.
-It is the full version of the map named by the ghost name $\delta{}s$ in
+is the ghost name for this map (again, for the current address space).
+$m$ is the (full) logical map from the various page table roots to ghost names for the 
+per-address-space mappings like $\theta$.
+It is tracked by a ghost map named by the ghost name $\delta{}s$ in
 Figure \ref{fig:virtualpointstosharing}; there is only one such map in the system.
-This is why Figure \ref{fig:virtualpointstosharing} asserts there is some ghost name in that (global) map
-for the current address space.
-The global map itself is tracked by a resource algebra 
-\[\mathcal{A}\textsf{spaces} \stackrel{\triangle}{=} \textsc{Auth} (\; \mathcal{W}_{64} \;\rightarrow_{\textrm{fin}} \;  ( (\textsc{Frac }, \mathord{+}) \times (\textsc{Agree } \textsc{GhostName},\mathord{=}) )) \]
+\looseness=-1
+% This is why Figure \ref{fig:virtualpointstosharing} asserts there is some ghost name in that (global) map
+% for the current address space.
+% The global map itself is tracked by a resource algebra 
+% \[\mathcal{A}\textsf{spaces} \stackrel{\triangle}{=} \textsc{Auth} (\; \mathcal{W}_{64} \;\rightarrow_{\textrm{fin}} \;  ( (\textsc{Frac }, \mathord{+}) \times (\textsc{Agree } \textsc{GhostName},\mathord{=}) )) \]
 
@@
 reasoning principles behind the virtual memory access, and we leave the use of this within a larger
-kernel to future work. We do, however, deal with the existence of multiple address spaces.
+kernel to future work. We do, however, deal with the existence of multiple address spaces in examples (Section \ref{sec:experiment}).
 
@@
 \label{sec:selected_rules}
-Per the discussion in Section \ref{sec:issues}, we define our logic using Hoare doubles:\footnote{This does
-omit some low-level Iris details related to stuckness and observations, which play no meaningful
+Per the discussion in Section \ref{sec:issues}, we define our logic using Hoare doubles:\footnote{This
+omits some low-level Iris details (stuckness, observations) that play no meaningful
 role in our development.}

### 2024-07-09 programlogic.tex
finish pass on section 4, including tentative text on the control transfer axioms

@@
   (l2M52 maddr l3e) $\sim$ (l2off maddr l3e) $\mapsto_{a}$ {q3}  l2e $\ast$
-  (l1M52 maddr l2e) $\sim$ (l1off maddr l2e) $\mapsto_{a}$ {q4}  l1e.
+  (l1M52 maddr l2e) $\sim$ (l1off maddr l2e) $\mapsto_{a}$ {q4}  l1e $\ast$ $\ulcorner$ addr_L1(va,l1e) = paddr $\urcorner$.
 \end{lstlisting}
@@
 Most of the definition lives directly in \textsf{vProp}, using the separation logic structure lifted from Iris's \textsf{iProp}.
-\todo[inline]{Colin notes: the \lstinline|addr_L1| mentioned in the paragraph above doesn't appear in Figure 2. paddr isn't mentioned
-in the definition of \lstinline|L4_L1_PointsTo| in the figure. Shouldn't there be a pure \lstinline|addr_L1(va,l1e)=paddr| in there?
-}
 
@@
   \and
-  \inferrule{ }{\mathsf{GhostMap}(\gamma,\theta) \ast \ghostmaptoken{\gamma}{\mathsf{pa}}{\mathsf{va}} \wand \ulcorner \theta(\mathsf{pa})=\mathsf{Some}(\matsf{va}) \urcorner }
+  \inferrule{ }{\mathsf{GhostMap}(\gamma,\theta) \ast \ghostmaptoken{\gamma}{\mathsf{pa}}{\mathsf{va}} \wand \ulcorner \theta(\mathsf{pa})=\mathsf{Some}(\mathsf{va}) \urcorner }
   \end{mathpar}
@@
   entry, preventing the overall translation from being changed separately from the virtual points-to.
+  \looseness=-1
 
@@
  \qquad\qquad\qquad \bigast{(\vaddr, \textsf{paddr})\in \ptablestore}{\exists\;(\textsf{l4e l3e l2e, l1e, paddr})\ldotp \textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\vaddr\textsf{, l4e, l3e, l2e, l1e, paddr})} \\
-  \textsf{ where } \\
+  \textsf{ where } 
    \textsf{ASpace\_Lookup} (\ptablestore,m) \stackrel{\triangle}{=} \lambda\textsf{ cr3val} \ldotp \; \exists \gammaPred \; \ldotp \ulcorner m \; !!\; \textsf{cr3val} = \textsf{Some } \gammaPred \urcorner \ast
@@
 \]
+\vspace{-1em}
 \caption{Global Address-Space Invariant with a fixed global map of address-space names $m$}
@@
 \footnotesize
-\[\begin{array}{c}
-    [r](\mathsf{P}) \;: \mathsf{vProp} \; \Sigma \; \stackrel{\triangle}{=} \; \lambda \_,\; \mathsf{P}~\mathsf{r} \vspace{0.5em}\qquad
-    \textsf{Fact P} \stackrel{\triangle}{=} \;\forall \; ,r \; r' \ldotp  \textsf{P r} \dashv\vdash \textsf{P r'}\\
-    \inferrule{}{(\textsf{P} \vdash \textsf{Q}) \vdash  ([r](\textsf{P}) \vdash  [r](\textsf{Q}))}\qquad
-    \inferrule{}{[r](\textsf{P} \ast \textsf{Q}) \dashv\vdash ([r](\textsf{P}) \ast [r](\textsf{Q})}\vspace{0.5em}\qquad
-    \inferrule{}{[r](\textsf{P}\land \textsf{Q}) \dashv\vdash [r](\textsf{P})\land [r](\textsf{Q})}\\
-    \inferrule{}{[r](\textsf{P}\lor \textsf{Q}) \dashv\vdash [r](\textsf{P})\lor [r](\textsf{Q})}\qquad
-    \inferrule{}{\textsf{Fact P} \vdash  [r](\textsf{P}) \dashv\vdash (\textsf{P})} %(\textsf{vProp }\Sigma))\\
-\end{array}\]
+\begin{mathpar}
+    \mbox{$[r](\mathsf{P}) \;: \mathsf{vProp} \; \Sigma \; \stackrel{\triangle}{=} \; \lambda \_,\; \mathsf{P}~\mathsf{r} \vspace{0.5em}$}
+    \and
+    \mbox{$\textsf{Fact P} \stackrel{\triangle}{=} \;\forall \; ,r \; r' \ldotp  \textsf{P r} \dashv\vdash \textsf{P r'}$}
+    \and
+    \inferrule*{ }{\textsf{Fact}\;[r](\textsf{P}) }
+    \and
+    \inferrule*{ }{\textsf{Fact}\;(\textsf{r}\;\mapsto_{\textsf{r}}\{q\}\;\textsf{rv})}
+    \and
+    \inferrule{ }{\textsf{Fact}\;(\textsf{w} \mapsto_{\textsf{p}} \{q\} \textsf{ v} )}
+    \and
+    \inferrule{ }{(\textsf{P} \vdash \textsf{Q}) \vdash  ([r](\textsf{P}) \vdash  [r](\textsf{Q}))}
+    \and
+    \inferrule{ }{[r](\textsf{P} \ast \textsf{Q}) \dashv\vdash ([r](\textsf{P}) \ast [r](\textsf{Q})}\vspace{0.5em}
+    \and
+    \inferrule{ }{[r](\textsf{P}\land \textsf{Q}) \dashv\vdash [r](\textsf{P})\land [r](\textsf{Q})}
+    \and
+    \inferrule{ }{[r](\textsf{P}\lor \textsf{Q}) \dashv\vdash [r](\textsf{P})\lor [r](\textsf{Q})}
+    \and
+    \inferrule{ }{\textsf{Fact P} \vdash  [r](\textsf{P}) \dashv\vdash (\textsf{P})} %(\textsf{vProp }\Sigma))\\
+\end{mathpar}
 \vspace{-1em}
@@
 address, the assertion $[r](P)$ is a modal assertion which ignores the (implicit) current page table root,
-and evaluates the truth of $P$ as if $r$ were the page table root. This simple definition is 
-what we need in order for assertions in one address space to talk about assertions in another.
+and evaluates the truth of $P$ as if $r$ were the page table root. 
+The novelty here is not in the details of the definition, but in recognizing that this is the right way to deal with
+multiple address spaces, and working out how to support interaction of multiple address spaces (discussed in the next section).
+% This simple definition is 
+% what we need in order for assertions in one address space to talk about assertions in another.
 \looseness=-1
@@
 
-We can prove that this modality follows certain useful laws: that its truth is independent of the address
+We can prove that this modality follows certain basic laws: that its truth is independent of the address
 space in which it is considered,  it distributes over various logical connectives, and it follows the rule of
@@
 
-\subsection{The Impact of Changing Address Spaces on Modal Assertions}
+\subsection{Subtleties of Changing Address Spaces Using Modalities}
 % \subsection{Considering Hoare Doubles in the Context of Switching Address-Space}
@@
 \label{sec:issues}
-An important subtlety arises with supporting \lstinline|mov|s into \lstinline|cr3|. Consider the hypothetical rule
+Operating systems on x86-64 change address spaces by updating \lstinline|cr3|.
+Conceptually, we expect a proof rule for such an update to take move resources in and out of modalities for
+the old and new address spaces. This suggests a rule like the following, which captures the intuition of the
+address space change, but is subtly unsound under common assumptions.
+% An important subtlety arises with supporting \lstinline|mov|s into \lstinline|cr3|. Consider the hypothetical rule
 \begin{mathpar}
@@
 \end{mathpar}
-which is slightly simplified (by dropping per-address-space resources) for clarity in making the following point.
-This rule captures the intuitive change of address space in a Hoare triple.
-This interacts quite poorly with the traditional frame
-rule and the modal flavor of virtual points-to assertions.
+This rule takes all assertions not under a modality ($P$), and moves them under a modality for the old address space
+(which is where they were valid), while pulling assertions for the new, target address space out of a corresponding modality.
+(For clarity, we also omit the address space invariants, which must also be present.)
+% which is slightly simplified (by dropping per-address-space resources) for clarity in making the following point.
+% This rule captures the intuitive change of address space in a Hoare triple.
+Unfortunately, as stated, this interacts quite poorly with the traditional frame
+rule.
 If we abbreviate the pre- and post-conditions above as \textsf{Pre} and \textsf{Post}:
@@
 The essential problem is that the frame rule is motivated by local reasoning about local updates, but
-a switch of address space is a \emph{global} change, since it may invalidate information about virtual addresses.
+a switch of address space is a \emph{global} change that may invalidate information about virtual addresses.
 Thus framing around \lstinline|cr3| updates is unsound, which has important impacts on our formal development.
@@
 
-This is then how we justify our actual rule for \lstinline|cr3| updates, which is \emph{almost}
+% This is then how we justify our actual rule for \lstinline|cr3| updates, which is \emph{almost}
+This leads to our actual rule for \lstinline|cr3| updates, which is \emph{almost}
 \begin{mathpar}
@@
 \end{mathpar}
+(Again, the actual rule in Section \ref{sec:selected_rules} includes the address space invariants.)
 Because the precondition on this rule is global, we avoid issues with framing. 
-A more typical frame rule can be recovered from Hoare doubles in the typical way~\cite{Chlipala2013Bedrock} A more typical frame rule can be recovered from Hoare doubles in the typical way~\cite{Chlipala2013Bedrock}
+A more typical frame rule can be recovered from Hoare doubles in the typical way~\cite{Chlipala2013Bedrock} 
 with the added restriction that \lstinline|cr3| is restored to its pre-framing value before un-framing (in this case, the framed-out assertion is effectively
@@
  % \\
-\inferrule[WriteToRegFromVirtMem]{
+\inferrule[\footnotesize WriteToRegFromVirtMem]{
   \left\{ \begin{array}{l}
@@
 \\
-\inferrule[WriteToVirtMemFromReg]{
+\inferrule[\footnotesize WriteToVirtMemFromReg]{
   \{P \ast \mathcal{I}\textsf{ASpace}\ast \textsf{rip} \mapsto_{\textsf{r}} \textsf{iv} + \textsf{MovLen}(\textsf{r}_d,[\textsf{r}_a])  \ast r_s \mapsto_{r}\{q\}  \textsf{rvs}  \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{v}} \textsf{v} \}_{\textsf{rtv}}\;\overline{ is}
@@
 %}\\
-\inferrule[WriteToRegCtlFromRegModal]{
+\inferrule[\footnotesize WriteToRegCtlFromRegModal]{
   \{[\textsf{rtv}](P \ast \mathcal{I}\textsf{ASpace})\ast \textsf{rip} \mapsto_{\textsf{r}} \textsf{iv + 4} \ast \mathcal{I}\textsf{ASpace} \ast R\ast r_s \mapsto_{r}\{q\}  \textsf{ rvs}  \}_{\textsf{rvs}} \;\overline{is}
@@
 
+\paragraph{Accessing Virtual Addresses}
 Figure \ref{fig:wpdamd} includes two  rules for accessing memory at an address stored in a register $r_a$. 
@@
 for the current address space to be carried through.
-While doing the proof of soundness for these rules,
-the token ($\sumwalkabs\vaddr\qfrac\paddr$) from the virtual points-to
-is used to extract to extract page-table-traversal points-to collection
-for the relevant address from that resource,
-and then those  physical points-to assertions describing the page table walk
-(from $\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$) are used to prove
-the operational semantics succeed and have the desired effect.
+The soundness proofs for these rules extract
+% the token ($\sumwalkabs\vaddr\qfrac\paddr$) from the virtual points-to
+the token ($\fracghostmaptoken{\delta}{\vaddr}{\paddr}{\qfrac}$) from the virtual points-to,
+use that to extract the physical page-table-traversal points-to collection describing
+the page table walk for the relevant address ($\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$)
+from the invariant ($\mathcal{I}\mathsf{ASpace}$), prove that the page table walk succeeds and
+memory or registers are updated appropriately, before re-packing the invariant and virtual points-to resources.
 
@@
 implementing a fragment of x86-64 including 64-bit address translation with 4-level page tables.
-
-At the moment our proofs do rely on 13 small axioms of properties which should be provable, but
-are challenging to discharge due to some representation choices in our model.\footnote{See \lstinline|srx/x64/machine/current_axioms.v|}
+\todo[inline]{
+Our rules for control transfers (\lstinline|jmp|, \lstinline|call|, and \lstinline|ret|) are currently
+axiomatized (with completely standard specifications~\cite{ni2007contexts,Chlipala2013Bedrock}) because Iris's built-in machinery does not provide
+convenient ways to discard the current continuation. We are working to adapt existing machinery~\cite{de2023type} to prove their soundness.
+Our soundness proofs for all other instructions (including the instructions above, arithmetic, other heap-accessing operations, push, and pop)
+are axiom-free.
+}
+% At the moment our proofs do rely on 13 small axioms of properties which should be provable, but
+% are challenging to discharge due to some representation choices in our model.\footnote{See \lstinline|srx/x64/machine/current_axioms.v|}

### 2024-07-09 examples.tex
convert examples to new ghost map token syntax

@@
 
+\todo[inline]{need new text explaining both mapping and the page table walk}
+
 %\begin{comment}
@@
 \subsection{Unmapping a Page}
+\todo[inline]{this must move, must come after mapping}
 The reverse operation, unmapping a designated page that is currently mapped,
@@
 $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)  \ast \texttt{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast  \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{next} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{\_} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{\_}  }_{\rtv}$
-$\specline{\mathsf{entry} \in \Xi \ast \sumapaces\rtv\delta  }_{\rtv}$
+$\specline{\mathsf{entry} \in \Xi \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
 mov    -0x8[rbp],rdi
 $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{\_}  \ast  \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{next} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry}  }_{\rtv}$
-$\specline{ \ulcorner \Xi \;!!\; \mathsf{entry} = \mathsf{Some V} \urcorner \ast \sumapaces\rtv\delta  }_{\rtv}$
+$\specline{ \ulcorner \Xi \;!!\; \mathsf{entry} = \mathsf{Some V} \urcorner \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
 mov     rdi, r8
 $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast  \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{next} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry} }_{\rtv}$
-$\specline{ \sumapaces\rtv\delta  }_{\rtv}$
+$\specline{ \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
 $\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present entry\_val})\urcorner}_{\rtv}$
@@
 $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast  \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{next} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val}   }_{\rtv}$
-$\specline{ \sumapaces\rtv\delta  }_{\rtv}$
+$\specline{ \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
 $\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present entry\_val})  \urcorner }_{\rtv}$
@@
 $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry} \ast \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{next} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val} }_{\rtv}$
-$\specline{ \sumapaces\rtv\delta  }_{\rtv}$
+$\specline{ \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
 $\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \land \lnot(\textsf{entry\_present entry\_val})  \urcorner }_{\rtv}$
@@
 $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry} \ast \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{next} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val}  }_{\rtv}$
-$\specline{ \sumapaces\rtv\delta }_{\rtv}$
+$\specline{ \ghostmaptoken{\delta{}s}{\rtv}{\delta} }_{\rtv}$
 $\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \land \lnot(\textsf{entry\_present entry\_val})  \urcorner }_{\rtv}$
@@
 $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val}  }_{\rtv}$
-$\specline{\sumapaces\rtv\delta \ast \mathsf{nextpaddr} \in \Xi\setminus \{ \mathsf{entry} \} \} }_{\rtv}$
+$\specline{\ghostmaptoken{\delta{}s}{\rtv}{\delta} \ast \mathsf{nextpaddr} \in \Xi\setminus \{ \mathsf{entry} \} \} }_{\rtv}$
 $\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \mathsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \land \lnot(\textsf{entry\_present entry\_val})  \urcorner }_{\rtv}$
@@
 $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val'}  }_{\rtv}$
-$\specline{\sumapaces\rtv\delta \ast \mathsf{nextpaddr} \in \Xi\setminus \{ \mathsf{entry} \} \cup \{ \mathsf{nextpaddr} \} }_{\rtv}$
+$\specline{\ghostmaptoken{\delta{}s}{\rtv}{\delta} \ast \mathsf{nextpaddr} \in \Xi\setminus \{ \mathsf{entry} \} \cup \{ \mathsf{nextpaddr} \} }_{\rtv}$
 $\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \mathsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \land (\textsf{entry\_present entry\_val'})  \urcorner }_{\rtv}$
@@
 \vspace{-1em}
-\caption{Making sure that \textsf{entry}'s frame is referencing to an address backed up by a physical page.}
+\caption{Ensuring \textsf{entry} points to a valid next table, allocating if necessary.}
 \label{fig:calltopteinitialize}
@@
 \footnotesize
+\todo[inline]{Ismail, look at this figure, something went weird (was already like this, I didn't add this)}
 \[
@@
    \exists \textsf{freshaddr, entry} \textsf{entry\_pfn} entry freshaddr = paddr \ast freshaddr \mapsto_{\textsf{p}} \textsf{w0\_64}  \leftrightarrow \mathsf{val} = \mathsf{w0\_64}\urcorner \; \ast\\
-   \qquad\qquad\qquad\qquad\qquad\qquad\qquad\qquad   \underbrace{\sumwalkabs\vaddr\qfrac\paddr }_\text{Ghost translation} \ast 
+   \qquad\qquad\qquad\qquad\qquad\qquad\qquad\qquad   \underbrace{\fracghostmaptoken{\delta}{\vaddr}{\paddr}{\qfrac} }_\text{Ghost translation} \ast 
 % \qquad\qquad\qquad\qquad\qquad\qquad\qquad\qquad    
@@
    \textsf{ASpace\_Lookup}_{\textsf{id}}(\ptablestore,\Xi,m) \stackrel{\triangle}{=} \lambda\textsf{ cr3val} \ldotp \; \exists \gammaPred \; \ldotp \ulcorner m \; !!\; \textsf{cr3val} = \textsf{Some } \gammaPred \urcorner \ast
-   \ownGhost\gammaPred{\authfull{\ptableabswalk\ptablestore}} \ast  \ownGhostpv\gammaPred{\authfull{\pvmapping\Xi}}
+   % \ownGhost\gammaPred{\authfull{\ptableabswalk\ptablestore}} \ast  \ownGhostpv\gammaPred{\authfull{\pvmapping\Xi}}
+   \ptableabswalk{\delta,\ptablestore} \ast \pvmapping{\delta,\Xi}
 
@@
 Concretely speaking, when we focus on the specfication of Line 13 in Figure \ref{fig:calltopteinitialize} in Figure \ref{fig:lookingupidpampping}, to read the value referenced by physical address \textsf{entry}, as shown in the first specification Line in Figure \ref{fig:lookingupidpampping}, we open our the addresss-space invariant ($\mathcal{I}\textsf{ASpace}_{\textsf{id}}(\ptablestore,\Xi,m)$), and then obtain the fractional ownership of the ghost translation and physical pointsto
-\[\sumwalkabsent \ast \mathsf{entry}\mapsto_{\mathsf{p,qfrac}}\; \vale \]
-via using the capability $\Xi \;!!\; \textsf{entry} = \textsf{Some V}$ and left the invariant as in the state $\mathcal{I}\textsf{ASpace}_{\textsf{id}}(\ptablestore,\Xi\setminus\{\mathsf{entry}\},m)$ , and construct the virtual mapping with the resources peeled-out of the identity mapping $\Xi$
+% \[\sumwalkabsent \ast \mathsf{entry}\mapsto_{\mathsf{p,qfrac}}\; \vale \]
+\[\fracghostmaptoken{\delta}{(\texttt{entry + KERNBASE})}{\textsf{entry}}{\textsf{qfrac}} \ast \mathsf{entry}\mapsto_{\mathsf{p,qfrac}}\; \vale \]
+by using the capability $\Xi \;!!\; \textsf{entry} = \textsf{Some V}$ and left the invariant as in the state $\mathcal{I}\textsf{ASpace}_{\textsf{id}}(\ptablestore,\Xi\setminus\{\mathsf{entry}\},m)$ , and construct the virtual mapping with the resources peeled-out of the identity mapping $\Xi$
 \[ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \]
@@
 \sqsubseteq \\
-   \specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast  \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{next} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry} }_{\rtv}\\
-   \specline{ \sumapaces\rtv\delta  }_{\rtv} \\
+   \specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast  \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{next} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv} \\
 \specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present entry\_val})\urcorner}_{\rtv} \\
 \mathsf{mov} \;   [ \mathsf{r8} ],\mathsf{rdi}\\
- \specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast  \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{next} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry} }_{\rtv}\\
-   \specline{ \sumapaces\rtv\delta  }_{\rtv} \\
+ \specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast  \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{next} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv} \\
 \specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present entry\_val})\urcorner}_{\rtv} \\
@@
 $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)   \ast \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{\_} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{table\_addr}  }_{\rtv}$
-$\specline{\texttt{rax} \mapsto_{\textsf{v}} \textsf{\_} \ast \textsf{table\_addr} \mapsto_{\mathsf{v}} \mathsf{\_}\ast\sumapaces\rtv\delta }_{\rtv}$
+$\specline{\texttt{rax} \mapsto_{\textsf{v}} \textsf{\_} \ast \textsf{table\_addr} \mapsto_{\mathsf{v}} \mathsf{\_}\ast\ghostmaptoken{\delta{}s}{\rtv}{\delta} }_{\rtv}$
 ;;void pte_initialize(pte_t **table) {
@@
 $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)   \ast \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{\_} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{table\_addr}  }_{\rtv}$
-$\specline{\texttt{rax} \mapsto_{\textsf{v}} \textsf{nextpaddr} \ast \textsf{table\_addr} \mapsto_{\mathsf{v}} \mathsf{\_} \ast \textsf{nextpaddr} \mapsto_{\textsf{p}} \textsf{w0\_64}\ast\sumapaces\rtv\delta }_{\rtv}$
+$\specline{\texttt{rax} \mapsto_{\textsf{v}} \textsf{nextpaddr} \ast \textsf{table\_addr} \mapsto_{\mathsf{v}} \mathsf{\_} \ast \textsf{nextpaddr} \mapsto_{\textsf{p}} \textsf{w0\_64}\ast\ghostmaptoken{\delta{}s}{\rtv}{\delta} }_{\rtv}$
 mov    rax,-0x10[rbp] 
 $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)   \ast \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{nextpaddr} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{table\_addr}  }_{\rtv}$
-$\specline{\texttt{rax} \mapsto_{\textsf{v}} \textsf{nextpaddr} \ast \ulcorner \Xi \;!!\; \mathsf{nextpaddr} = Z \urcorner \ast \textsf{table\_addr} \mapsto_{\mathsf{v}} \mathsf{\_} \ast \textsf{nextpaddr} \mapsto_{\textsf{p}} \textsf{w0\_64} \ast \sumapaces\rtv\delta }_{\rtv}$
+$\specline{\texttt{rax} \mapsto_{\textsf{v}} \textsf{nextpaddr} \ast \ulcorner \Xi \;!!\; \mathsf{nextpaddr} = Z \urcorner \ast \textsf{table\_addr} \mapsto_{\mathsf{v}} \mathsf{\_} \ast \textsf{nextpaddr} \mapsto_{\textsf{p}} \textsf{w0\_64} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta} }_{\rtv}$
 ;;*table = local;
@@
 $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)   \ast \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{nextpaddr} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{table\_addr}  }_{\rtv}$
-$\specline{\texttt{rax} \mapsto_{\textsf{v}} \textsf{nextpaddr}  \ast \textsf{table\_addr} \mapsto_{\mathsf{v}} \mathsf{\mathsf{nextpaddr}} \ast \textsf{nextpaddr} \mapsto_{\textsf{p}} \textsf{w0\_64} \ast \sumapaces\rtv\delta }_{\rtv}$
+$\specline{\texttt{rax} \mapsto_{\textsf{v}} \textsf{nextpaddr}  \ast \textsf{table\_addr} \mapsto_{\mathsf{v}} \mathsf{\mathsf{nextpaddr}} \ast \textsf{nextpaddr} \mapsto_{\textsf{p}} \textsf{w0\_64} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta} }_{\rtv}$
 ;;return;
@@
 $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \}\cup\{\mathsf{nextpaddr}\},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val'}  }_{\rtv}$
-$\specline{\sumapaces\rtv\delta \ast \mathsf{nextpaddr} \in \Xi\setminus \{ \mathsf{entry} \} \cup \{ \mathsf{nextpaddr} \} }_{\rtv}$
+$\specline{\ghostmaptoken{\delta{}s}{\rtv}{\delta} \ast \mathsf{nextpaddr} \in \Xi\setminus \{ \mathsf{entry} \} \cup \{ \mathsf{nextpaddr} \} }_{\rtv}$
 $\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \mathsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \land (\textsf{entry\_present entry\_val'})  \urcorner }_{\rtv}$
@@
 $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} ,\mathsf{nextpaddr} \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val'}  }_{\rtv}$
-$\specline{ \sumapaces\rtv\delta  }_{\rtv}$
+$\specline{ \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
 $\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val\_present} \ast \ulcorner  \textsf{qfrac} = 1 \land (\textsf{entry\_present entry\_val'}) \urcorner }_{\rtv}$
@@
 %...
-%$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)  \ast \texttt{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{\_} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{\_} \ast \sumpv\paddr\qfrac\true  \sumapaces\rtv\delta  }_{\rtv}$
+%$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)  \ast \texttt{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{\_} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{\_} \ast \sumpv\paddr\qfrac\true  \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
 %mov    rax,-0x18[rbp]
@@
 	(\lambda \mathit{cr3val}\ldotp
-	\sumapaces{\mathit{cr3val}}\delta) \ast 
-  \sumwalkabs\vaddr\qfrac\paddr \ast \paddr \mapsto_{\mathsf{p}} \vpage
+	\ghostmaptoken{\delta{}s}{\mathit{cr3val}}{\delta}) \ast 
+  \fracghostmaptoken{\delta}{\vaddr}{\paddr}{\qfrac} \ast \paddr \mapsto_{\mathsf{p}} \vpage
 \end{array}
@@
 \mapsto_{\textsf{vpte}}\; \qfrac \;\fpaddr\;(\textsf{wzero 64})$ (Line 31) using $\sumwalkabs\vaddr\qfrac\fpaddr$ obtained from ghost
-page-table-walk insertions, $\sumapaces\rtv\delta$ from unfolding the
+page-table-walk insertions, $\ghostmaptoken{\delta{}s}{\rtv}{\delta}$ from unfolding the
 definition of $\mapsto_{\textsf{vpte}}$, and $\fpaddr
@@
 $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace} (<[\vaddr:=\texttt{fpaddr}]> \theta,m) \ast}_{\rtv}$
-$\specline{\ulcorner \texttt{aligned fpaddr} \urcorner \ast \texttt{fpaddr} \mapsto_{\textsf{p}} \textsf{ wzero 64} \ast \sumapaces\rtv\delta  \ast\sumwalkabs\vaddr\qfrac\fpaddr}_{\rtv}$
+$\specline{\ulcorner \texttt{aligned fpaddr} \urcorner \ast \texttt{fpaddr} \mapsto_{\textsf{p}} \textsf{ wzero 64} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  \ast\sumwalkabs\vaddr\qfrac\fpaddr}_{\rtv}$
 $\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\; \sqsubseteq $

### 2024-07-10 examples.tex
start changing specs
+

@@
 ;; pte_t *next;
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)  \ast \texttt{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast  \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{next} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{\_} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{\_}  }_{\rtv}$
-$\specline{\mathsf{entry} \in \Xi \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{entry}\},m)  \ast \texttt{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast  \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{next} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{\_} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{\_}  }_{\rtv}$
+$\specline{ \textsf{entry} \mapsto_{\textsf{id}} \textsf{v} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
+$\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present entry\_val})\urcorner}_{\rtv}$
 mov    -0x8[rbp],rdi
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{\_}  \ast  \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{next} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry}  }_{\rtv}$
-$\specline{ \ulcorner \Xi \;!!\; \mathsf{entry} = \mathsf{Some V} \urcorner \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{entry\},m)  \ast \textsf{entry}\mapsto_{\textsf{pte,q}}\textsf{entryv} \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{\_}  \ast  \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{next} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry}  }_{\rtv}$
+$\specline{ \textsf{entry} \mapsto_{\textsf{id}} \textsf{v} \ast\urcorner \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
+$\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present entry\_val})\urcorner}_{\rtv}$
 mov     rdi, r8
 $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast  \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{next} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry} }_{\rtv}$
-$\specline{ \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
+$\specline{ \textsf{entry} \mapsto_{\textsf{id}} \textsf{v} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
 $\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present entry\_val})\urcorner}_{\rtv}$
@@
 $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast  \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{next} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val}   }_{\rtv}$
-$\specline{ \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
+$\specline{  \textsf{entry} \mapsto_{\textsf{id}} \textsf{v} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
 $\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present entry\_val})  \urcorner }_{\rtv}$
@@
 $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry} \ast \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{next} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val} }_{\rtv}$
-$\specline{ \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
+$\specline{  \textsf{entry} \mapsto_{\textsf{id}} \textsf{v} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
 $\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \land \lnot(\textsf{entry\_present entry\_val})  \urcorner }_{\rtv}$
@@
 $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry} \ast \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{next} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val}  }_{\rtv}$
-$\specline{ \ghostmaptoken{\delta{}s}{\rtv}{\delta} }_{\rtv}$
+$\specline{  \textsf{entry} \mapsto_{\textsf{id}} \textsf{v} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta} }_{\rtv}$
 $\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \land \lnot(\textsf{entry\_present entry\_val})  \urcorner }_{\rtv}$
@@
 $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val}  }_{\rtv}$
-$\specline{\ghostmaptoken{\delta{}s}{\rtv}{\delta} \ast \mathsf{nextpaddr} \in \Xi\setminus \{ \mathsf{entry} \} \} }_{\rtv}$
+$\specline{ \textsf{entry} \mapsto_{\textsf{id}} \textsf{v} \ast \textsf{entry\_val}_{\textsf{pfn}} \mapsto_{\textsf{id}} \textsf{\_} \ast \forall_{i\in \textssf{0 ...
+511} } \ldotp  \textsf{entry\_val }_{\textsf{pfn}} + \texstf{i * 8} \mapsto_{\textsf{id}} \textsf{\_}   \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta} \ast \mathsf{nextpaddr} \in \Xi\setminus \{ \mathsf{entry} \} \} }_{\rtv}$
 $\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \mathsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \land \lnot(\textsf{entry\_present entry\_val})  \urcorner }_{\rtv}$
-$\specline{ \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{nextpaddr} \ast  \mathsf{nextpaddr} \mapsto_{\textsf{p}} \textsf{w0\_64} \ast \texttt{rax} \mapsto_{\textsf{r}} \textsf{nextpaddr}  }_{\rtv}$
+$\specline{ \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{entry\_val}_{\textsf{pfn}}  \ast \texttt{rax} \mapsto_{\textsf{r}} \textsf{nextpaddr}  }_{\rtv}$
 ;;entry value updates

### 2024-07-10 examples.tex
+

@@
 $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val}  }_{\rtv}$
-$\specline{ \textsf{entry} \mapsto_{\textsf{id}} \textsf{v} \ast \textsf{entry\_val}_{\textsf{pfn}} \mapsto_{\textsf{id}} \textsf{\_} \ast \forall_{i\in \textssf{0 ...
-511} } \ldotp  \textsf{entry\_val }_{\textsf{pfn}} + \texstf{i * 8} \mapsto_{\textsf{id}} \textsf{\_}   \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta} \ast \mathsf{nextpaddr} \in \Xi\setminus \{ \mathsf{entry} \} \} }_{\rtv}$
+$\specline{ \textsf{entry} \mapsto_{\textsf{id}} \textsf{_} \ast \textsf{entry\_val}_{\textsf{pfn}} \mapsto_{\textsf{id}} \textsf{\_} \ast \forall_{i\in \textssf{0 ...
+511} } \ldotp  \textsf{entry\_val }_{\textsf{pfn}} + \texstf{i * 8} \mapsto_{\textsf{id}} \textsf{\_}   \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta} \ast \mathsf{nextpaddr} \in \Xi\setminus \{ \mathsf{entry} \}) }_{\rtv}$
 $\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \mathsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \land \lnot(\textsf{entry\_present entry\_val})  \urcorner }_{\rtv}$
@@
 $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val'}  }_{\rtv}$
-$\specline{\ghostmaptoken{\delta{}s}{\rtv}{\delta} \ast \mathsf{nextpaddr} \in \Xi\setminus \{ \mathsf{entry} \} \cup \{ \mathsf{nextpaddr} \} }_{\rtv}$
+$\specline{\ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
 $\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \mathsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \land (\textsf{entry\_present entry\_val'})  \urcorner }_{\rtv}$

### 2024-07-10 examples.tex
+

--- 2024-07-10 examples.tex
@@
 $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val}  }_{\rtv}$
-$\specline{ \textsf{entry} \mapsto_{\textsf{id}} \textsf{v} \ast \textsf{entry\_val}_{\textsf{pfn}} \mapsto_{\textsf{id}} \textsf{\_} \ast \forall_{i\in \textssf{0 ...
-511} } \ldotp  \textsf{entry\_val }_{\textsf{pfn}} + \texstf{i * 8} \mapsto_{\textsf{id}} \textsf{\_}   \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta} \ast \mathsf{nextpaddr} \in \Xi\setminus \{ \mathsf{entry} \} \} }_{\rtv}$
+$\specline{ \textsf{entry} \mapsto_{\textsf{id}} \textsf{\_} \ast \textsf{entry\_val}_{\textsf{pfn}} \mapsto_{\textsf{id}} \textsf{\_} \ast \ulcorner  \textsf{qfrac} = 1 \land \lnot(\textsf{entry\_present entry\_val}_{\textsf{pfn}})  \urcorner  \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
 $\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \mathsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \land \lnot(\textsf{entry\_present entry\_val})  \urcorner }_{\rtv}$
 $\specline{ \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{entry\_val}_{\textsf{pfn}}  \ast \texttt{rax} \mapsto_{\textsf{r}} \textsf{nextpaddr}  }_{\rtv}$
+$\specline{ \forall_{i\in \textssf{0 ...
+511} } \ldotp  \textsf{entry\_val }_{\textsf{pfn}} + \texstf{i * 8} \mapsto_{\textsf{id}} \textsf{\_} \ast  \textsf{qfrac} = 1 \land \lnot(\textsf{entry\_present entry\_val }_{\textsf{pfn}} + \texstf{i * 8})  }$
 ;;entry value updates
 ...
+;;entry->pfn = next
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val}  }_{\rtv}$
+$\specline{ \textsf{entry} \mapsto_{\textsf{id}} \textsf{\_} \ast \textsf{entry\_val}_{\textsf{pfn}} \mapsto_{\textsf{id}} \textsf{\_} \ast \ulcorner  \textsf{qfrac} < 1 \land (\textsf{entry\_present entry\_val}_{\textsf{pfn}})  \urcorner  \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
+$\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \mathsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \land \lnot(\textsf{entry\_present entry\_val})  \urcorner }_{\rtv}$
+$\specline{ \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{entry\_val}_{\textsf{pfn}}  \ast \texttt{rax} \mapsto_{\textsf{r}} \textsf{nextpaddr}  }_{\rtv}$
+$\specline{ \forall_{i\in \textssf{0 ...
+511} } \ldotp  \textsf{entry\_val }_{\textsf{pfn}} + \texstf{i * 8} \mapsto_{\textsf{id}} \textsf{\_} \ast  \textsf{qfrac} = 1 \land \lnot(\textsf{entry\_present entry\_val }_{\textsf{pfn}} + \texstf{i * 8})  }$
 ;;entry->present = 1;
-;;entry->pfn = next
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val'}  }_{\rtv}$
-$\specline{\ghostmaptoken{\delta{}s}{\rtv}{\delta} \ast \mathsf{nextpaddr} \in \Xi\setminus \{ \mathsf{entry} \} \cup \{ \mathsf{nextpaddr} \} }_{\rtv}$
-$\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \mathsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \land (\textsf{entry\_present entry\_val'})  \urcorner }_{\rtv}$
-$\specline{ \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{nextpaddr} \ast  \mathsf{nextpaddr} \mapsto_{\textsf{p}} \textsf{w0\_64} \ast \texttt{rax} \mapsto_{\textsf{r}} \textsf{nextpaddr}  }_{\rtv}$
 ;;}

### 2024-07-10 examples.tex
+

@@
 $\specline{ \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{entry\_val}_{\textsf{pfn}}  \ast \texttt{rax} \mapsto_{\textsf{r}} \textsf{nextpaddr}  }_{\rtv}$
-$\specline{ \forall_{i\in \textssf{0 ...
-511} } \ldotp  \textsf{entry\_val }_{\textsf{pfn}} + \texstf{i * 8} \mapsto_{\textsf{id}} \textsf{\_} \ast  \textsf{qfrac} = 1 \land \lnot(\textsf{entry\_present entry\_val }_{\textsf{pfn}} + \texstf{i * 8})  }$
+$\specline{ \forall_{i\in \textsf{0 ...
+511} } \ldotp  \textsf{entry\_val }_{\textsf{pfn}} + \textsf{i * 8} \mapsto_{\textsf{id}} \textsf{\_} \ast  \textsf{qfrac} = 1 \land \lnot(\textsf{entry\_present entry\_val }_{\textsf{pfn}} + \textsf{i * 8})  }$
 ;;entry value updates
@@
 $\specline{ \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{entry\_val}_{\textsf{pfn}}  \ast \texttt{rax} \mapsto_{\textsf{r}} \textsf{nextpaddr}  }_{\rtv}$
-$\specline{ \forall_{i\in \textssf{0 ...
-511} } \ldotp  \textsf{entry\_val }_{\textsf{pfn}} + \texstf{i * 8} \mapsto_{\textsf{id}} \textsf{\_} \ast  \textsf{qfrac} = 1 \land \lnot(\textsf{entry\_present entry\_val }_{\textsf{pfn}} + \texstf{i * 8})  }$
+$\specline{ \forall_{i\in \textsf{0 ...
+511} } \ldotp  \textsf{entry\_val }_{\textsf{pfn}} + \textsf{i * 8} \mapsto_{\textsf{id}} \textsf{\_} \ast  \textsf{qfrac} = 1 \land \lnot(\textsf{entry\_present entry\_val }_{\textsf{pfn}} + \textsf{i * 8})  }$
 ;;entry->present = 1;

### 2024-07-10 examples.tex
+

@@
 \begin{figure}\footnotesize
-  \begin{lstlisting}[mathescape]
+  \begin{lstlisting}[mathescape, basicstyle=\tiny]
 ;;pte_t *pte_get_next_table_succ(pte_t *entry) {
@@
 \end{equation*}
-assuring that looking the identity mapping for \textsf{entry} is safe under the subtle justification which equates the full ownership to the non/presence of the entry which can only be known when investigated ad Line 19in Figure \ref{fig:calltopteinitialize}.
+assuring that looking the identity mapping for \textsf{entry} is safe under the subtle justification which equates the full ownership to the non/presence of the entry which can only be known when investigated in Line 19in Figure \ref{fig:calltopteinitialize}.
+\begin{comment}
  \begin{figure*}
@@
  \end{figure*}
-
+\end{comment}
  \subsubsection{Ensuring Presence of an Entry}
@@
 
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)   \ast \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{\_} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{table\_addr}  }_{\rtv}$
-$\specline{\texttt{rax} \mapsto_{\textsf{v}} \textsf{\_} \ast \textsf{table\_addr} \mapsto_{\mathsf{v}} \mathsf{\_}\ast\ghostmaptoken{\delta{}s}{\rtv}{\delta} }_{\rtv}$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)   \ast \texttt{rbp-16} \mapsto_{\textsf{r}} \textsf{\_} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{table\_addr}  }_{\rtv}$
+$\specline{\texttt{rax} \mapsto_{\textsf{r}} \textsf{\_} \ast \textsf{table\_addr} \mapsto_{\mathsf{v}} \mathsf{\_}\ast\ghostmaptoken{\delta{}s}{\rtv}{\delta} }_{\rtv}$
 ;;void pte_initialize(pte_t **table) {
@@
 callq  81 <pte_initialize+0x11>	7d: R_X86_64_PLT32	kalloc-0x4
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)   \ast \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{\_} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{table\_addr}  }_{\rtv}$
-$\specline{\texttt{rax} \mapsto_{\textsf{v}} \textsf{nextpaddr} \ast \textsf{table\_addr} \mapsto_{\mathsf{v}} \mathsf{\_} \ast \textsf{nextpaddr} \mapsto_{\textsf{p}} \textsf{w0\_64}\ast\ghostmaptoken{\delta{}s}{\rtv}{\delta} }_{\rtv}$
 mov    rax,-0x10[rbp] 
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)   \ast \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{nextpaddr} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{table\_addr}  }_{\rtv}$
-$\specline{\texttt{rax} \mapsto_{\textsf{v}} \textsf{nextpaddr} \ast \ulcorner \Xi \;!!\; \mathsf{nextpaddr} = Z \urcorner \ast \textsf{table\_addr} \mapsto_{\mathsf{v}} \mathsf{\_} \ast \textsf{nextpaddr} \mapsto_{\textsf{p}} \textsf{w0\_64} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta} }_{\rtv}$
 ;;*table = local;
@@
 mov    rax,[rdi]
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)   \ast \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{nextpaddr} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{table\_addr}  }_{\rtv}$
-$\specline{\texttt{rax} \mapsto_{\textsf{v}} \textsf{nextpaddr}  \ast \textsf{table\_addr} \mapsto_{\mathsf{v}} \mathsf{\mathsf{nextpaddr}} \ast \textsf{nextpaddr} \mapsto_{\textsf{p}} \textsf{w0\_64} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta} }_{\rtv}$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)   \ast \texttt{rbp-16} \mapsto_{\textsf{r}} \textsf{\_} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{table\_addr}  }_{\rtv}$
+$\specline{\texttt{rax} \mapsto_{\textsf{v}} \textsf{nextpaddr} \ast \textsf{table\_addr} \mapsto_{\mathsf{v}} \mathsf{\mathsf{nextpaddr}} \ast \textsf{nextpaddr} \mapsto_{\textsf{id}} \textsf{\_}\ast \forall_{i\in\textsf{0..511}} \textsf{nextpaddr + i * 8} \mapsto_{\textsf{id}} \textsf{\_} \ast\ghostmaptoken{\delta{}s}{\rtv}{\delta} }_{\rtv}$
 ;;return;

### 2024-07-10 examples.tex
+

@@
 ;;entry->pfn = next
+;;entry->present = 1;
 $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val}  }_{\rtv}$
-$\specline{ \textsf{entry} \mapsto_{\textsf{id}} \textsf{\_} \ast \textsf{entry\_val}_{\textsf{pfn}} \mapsto_{\textsf{id}} \textsf{\_} \ast \ulcorner  \textsf{qfrac} < 1 \land (\textsf{entry\_present entry\_val}_{\textsf{pfn}})  \urcorner  \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
+$\specline{ \textsf{entry} \mapsto_{\textsf{id}} \textsf{\_} \ast \textsf{entry\_val}_{\textsf{pfn}} \mapsto_{\textsf{id}} \textsf{\_} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \lnot(\textsf{entry\_present entry\_val}_{\textsf{pfn}})  \urcorner  \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
 $\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \mathsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \land \lnot(\textsf{entry\_present entry\_val})  \urcorner }_{\rtv}$
@@
 $\specline{ \forall_{i\in \textsf{0 ...
-511} } \ldotp  \textsf{entry\_val }_{\textsf{pfn}} + \textsf{i * 8} \mapsto_{\textsf{id}} \textsf{\_} \ast  \textsf{qfrac} = 1 \land \lnot(\textsf{entry\_present entry\_val }_{\textsf{pfn}} + \textsf{i * 8})  }$
-;;entry->present = 1;
+511} } \ldotp  \textsf{entry\_val }_{\textsf{pfn}} + \textsf{i * 8} \mapsto_{\textsf{id}} \textsf{\_} \ast  \textsf{qfrac} = 1 \leftrightarrow \lnot(\textsf{entry\_present entry\_val }_{\textsf{pfn}} + \textsf{i * 8})  }$
+\subseteq
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \} \cup \{\textsf{entry\_val}_{\textsf{pfn}}) \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val}  }_{\rtv}$
+$\specline{ \textsf{entry} \mapsto_{\textsf{id}} \textsf{\_} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
+$\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \mathsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \land \lnot(\textsf{entry\_present entry\_val})  \urcorner }_{\rtv}$
+$\specline{ \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{entry\_val}_{\textsf{pfn}}  \ast \texttt{rax} \mapsto_{\textsf{r}} \textsf{entry\_val}_{\textsf{pfn}}  }_{\rtv}$
 ;;}
@@
 $\specline{\texttt{rax} \mapsto_{\textsf{v}} \textsf{nextpaddr} \ast \textsf{table\_addr} \mapsto_{\mathsf{v}} \mathsf{\mathsf{nextpaddr}} \ast \textsf{nextpaddr} \mapsto_{\textsf{id}} \textsf{\_}\ast \forall_{i\in\textsf{0..511}} \textsf{nextpaddr + i * 8} \mapsto_{\textsf{id}} \textsf{\_} \ast\ghostmaptoken{\delta{}s}{\rtv}{\delta} }_{\rtv}$
-;;return;
-;;..
 ;;clean up the stack
+..
 \end{lstlisting}
@@
  movabs 0x0,rcx	                187: R_X86_64_64	KERN_BASE
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \}\cup\{\mathsf{nextpaddr}\},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val'}  }_{\rtv}$
-$\specline{\ghostmaptoken{\delta{}s}{\rtv}{\delta} \ast \mathsf{nextpaddr} \in \Xi\setminus \{ \mathsf{entry} \} \cup \{ \mathsf{nextpaddr} \} }_{\rtv}$
-$\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \mathsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \land (\textsf{entry\_present entry\_val'})  \urcorner }_{\rtv}$
-$\specline{ \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{nextpaddr} \ast \ulcorner \Xi\setminus \{ \mathsf{entry} \}\cup\{\mathsf{nextpaddr}\} !!
-\mathsf{entry} = \mathsf{Some Z} \urcorner \ast \texttt{rax} \mapsto_{\textsf{r}} \textsf{nextpaddr}  }_{\rtv}$
  add    rcx,rax
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} ,\mathsf{nextpaddr} \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val'}  }_{\rtv}$
-$\specline{ \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
-$\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val\_present} \ast \ulcorner  \textsf{qfrac} = 1 \land (\textsf{entry\_present entry\_val'}) \urcorner }_{\rtv}$
-$\specline{ \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{nextpaddr}  \ast \texttt{rax} \mapsto_{\textsf{r}} \textsf{nextpaddr}  }_{\rtv}$
-$\specline{ \texttt{nextpaddr+KERNBASE} \mapsto_{\textsf{vpte,qfracn}} \textsf{nextpaddr w64\_zero} }_{\rtv}$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \} \cup \{\textsf{entry\_val}_{\textsf{pfn}}) \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val}  }_{\rtv}$
+$\specline{ \textsf{entry} \mapsto_{\textsf{id}} \textsf{\_} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
+$\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \mathsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \land \lnot(\textsf{entry\_present entry\_val})  \urcorner }_{\rtv}$
+$\specline{ \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{entry\_val}_{\textsf{pfn}}  \ast \texttt{rax} \mapsto_{\textsf{r}} \textsf{entry\_val}_{\textsf{pfn}}  }_{\rtv}$
+\subseteq
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val}  }_{\rtv}$
+$\specline{ \textsf{entry} \mapsto_{\textsf{id}} \textsf{\_} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
+$\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \mathsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \land \lnot(\textsf{entry\_present entry\_val})  \urcorner }_{\rtv}$
+$\specline{  \textsf{entry\_val + KERNBASE}_{\textsf{pfn}}  \mapsto_{\textsf{vpte,qfrac}} \mathsf{entry entry\_val}  \textsf{entry\_val}_{\textsf{pfn}}} \ast R_{\textsf{children}}$
+$\specline{ \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{entry\_val}_{\textsf{pfn}}  \ast \texttt{rax} \mapsto_{\textsf{r}} \textsf{entry\_val}_{\textsf{pfn}}  }_{\rtv}$
 ...
@@
 \vspace{-1em}
-\caption{Converting a physical address of a PTE to a virtual address (w/o instruction pointer or flag updates).}
+\caption{Converting a physical address of a PTE to a virtual address (w/o instruction pointer or flag updates).
+Abreviating the other relevant resources (including the lower entries) peeled-out of the invariant as $\textsf{R}_{\textsf{children}}$ }
 \label{fig:p2v}

### 2024-07-10 examples.tex
+
fix

@@
 %\end{comment}
-\subsection{Unmapping a Page}
-\todo[inline]{this must move, must come after mapping}
-The reverse operation, unmapping a designated page that is currently mapped,
-would essentially be the reverse of
-the reasoning around line 22 above: given the virtual points-to assertions for all 512
-machine words of memory that the L1 entry would map,
-and information about the physical location, 
-full permission on the L1 entry could be obtained, allowing the construction of a
-full virtual PTE pointer for it, setting to 0, and reclaiming the now-unmapped physical memory.
-
 \subsection{Traversing Live Page Tables}
@@
 511} } \ldotp  \textsf{entry\_val }_{\textsf{pfn}} + \textsf{i * 8} \mapsto_{\textsf{id}} \textsf{\_} \ast  \textsf{qfrac} = 1 \leftrightarrow \lnot(\textsf{entry\_present entry\_val }_{\textsf{pfn}} + \textsf{i * 8})  }$
-\subseteq
+$\subseteq$
 $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \} \cup \{\textsf{entry\_val}_{\textsf{pfn}}) \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val}  }_{\rtv}$
@@
 $\specline{ \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{entry\_val}_{\textsf{pfn}}  \ast \texttt{rax} \mapsto_{\textsf{r}} \textsf{entry\_val}_{\textsf{pfn}}  }_{\rtv}$
-\subseteq
+$\subseteq$
 $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val}  }_{\rtv}$
@@
  ...
+ $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}}_{\rtv}$
 ;;set up the stack for root address and virtual address    
  ;;pte_t *pml4_entry = &pml4[PML4EX(va)];
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{entry}\},m)  }_{\rtv}$
+$\specline{ \textsf{entry} \mapsto_{\textsf{id}} \textsf{v} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
+$\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present entry\_val})\urcorner}_{\rtv}$
 mov    -0x8[rbp],rsi
@@
 mov    -0x18[rbp],rdi
-$\specline{P_0 }_{\rtv}$
 ...
@@
 ;;save the physical next table address in rax
-$\specline{P_1 \ast \texttt{pdp} \mapsto_{\textsf{vpte,qfracn}} \textsf{pdppaddr \_} }_{\rtv}$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{entry}\},m)  }_{\rtv}$
+$\specline{ \textsf{entry} \mapsto_{\textsf{id}} \textsf{v} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
+$\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present entry\_val})\urcorner}_{\rtv}$
+$\specline{ \textsf{entry}_{\textsf{pfn}} \mapsto_{\textsf{v,qf}} \textsf{pdp\_val} \ast \textsf{rax} \mapsto_{\textsf{r}} \textsf{entry\_val}_{\textsf{pfn}}}_{\rtv}$    
 ;;pte_t *pdp_entry = &pdp[PDPEX(va)];
 ;;pte_t *pd = pte_get_next_table_succ(pdp_entry);
-$\specline{P_2 \ast \texttt{pdp} \mapsto_{\textsf{vpte,qfracn}} \textsf{pdppaddr w64\_zero} \ast \texttt{pd} \mapsto_{\textsf{vpte,qfracn}} \textsf{pdpaddr \_} }_{\rtv}$
 ...
@@
 ;;pte_t *pt = pte_get_next_table_succ(pd_entry);
-$\specline{P_3 \ast \texttt{pdp} \mapsto_{\textsf{vpte,qfracn}} \textsf{pdppaddr w64\_zero} \ast \texttt{pd} \mapsto_{\textsf{vpte,qfracn}} \textsf{pdpaddr \_} \ast \texttt{pt} \mapsto_{\textsf{vpte,qfracn}} \textsf{ptpaddr \_} }_{\rtv}$
+$\specline{\textsf{R}_{\textsf{walk\_acc}}}_{\rtv}$
 ;;access and return L1 entry
@@
 ;;pte->writable     = PTE_WRITEABLE_SET(perm);
-...
 ;;pte->user_acc     = PTE_USER_ACC_SET(perm);
-...
 ;;pte->writethrough = PTE_WRITETHRU_SET(perm);
-...
 ;;pte->nocache      = PTE_NOCACHE_SET(perm);
@@
 % were axiomatized in prior work.
-
+\subsection{Unmapping a Page}
+\todo[inline]{this must move, must come after mapping}
+The reverse operation, unmapping a designated page that is currently mapped,
+would essentially be the reverse of
+the reasoning around line 22 above: given the virtual points-to assertions for all 512
+machine words of memory that the L1 entry would map,
+and information about the physical location, 
+full permission on the L1 entry could be obtained, allowing the construction of a
+full virtual PTE pointer for it, setting to 0, and reclaiming the now-unmapped physical memory.
 \subsection{Change of Address Space}

### 2024-07-10 examples.tex
+

@@
 \label{sec:traversing}
-The mapping operation of Figure \ref{fig:mapping_code} assumes an operation \lstinline|pte\_walkpgdir| which must traverse the page tables
+The mapping operation of Figure \ref{fig:mapping_code} assumes an operation \textsf{pte\_walkpgdir} which must traverse the page tables
 in order to locate the address of the L1 entry to update --- possibly allocating pages at levels 3 and 2 in the process.
@@
 We call \textsf{get\_next\_table} to get the virtual address to the next level page-table.
-In Figure \ref{fig:calltopteinitialize}, we present the critical piece of \textsf{get\_next\_table}, and construct the specification of it
-Within \textsf{get\_next\_table}, first, we check if the entry whose address is passed as a parameter ($\textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry}$) is present or not.
-To be able to do so, it needs to load the entry value (a 64-bit word divided into bit-fields for corresponding fileds shown in Section {}) and check that \textit{presence} bit of the entry is set or not.
+In Figure \ref{fig:calltopteinitialize}, we present the most critical piece of \textsf{get\_next\_table}, and construct the specification for it.
+Within \textsf{get\_next\_table}, first, we check if the entry whose address is passed as a parameter ($\textsf{entry}$) is present or not (Line 21 in Figure \ref{fig:callptetoinitialize}).
+To be able to do so, it needs to load the entry value (a 64-bit word divided into bit-fields for corresponding fileds shown in Section {}), and then,check if \textit{presence} bit of the entry is set or not.
+We see in Specification Line 6 that \textsf{entry} has \textsf{pte-pointsto} that is constructed on the client of \textfsf{pte\_get\_next\_table} -- e.g.\textsf{pte\_walkpgdir} in Figure \ref{walkpgdir}.
 %wshiftll (wshiftll (natToWord 64 entry) (WordImpl.concat (WordImpl.zero 56) (WordImpl.from_nat 8 12 ^& WordImpl.concat (WordImpl.zero 2) WO~1~1~1~1~1~1)) ^& constf)
@@
 511} } \ldotp  \textsf{entry\_val }_{\textsf{pfn}} + \textsf{i * 8} \mapsto_{\textsf{id}} \textsf{\_} \ast  \textsf{qfrac} = 1 \leftrightarrow \lnot(\textsf{entry\_present entry\_val }_{\textsf{pfn}} + \textsf{i * 8})  }$
-$\subseteq$
+$\sqsubseteq$
 $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \} \cup \{\textsf{entry\_val}_{\textsf{pfn}}) \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val}  }_{\rtv}$
@@
 \end{figure}
-
-As we see in Figure \ref{fig:calltopteinitialize}, we have the physical address (\textsf{entry}) of the page table entry kept in the register $\texttt{r8}$, but not the virtual memory mapping to load the entry value referenced by the physical address \textsf{entry} (Line 13 in Figure \ref{fig:calltopteinitialize}).
-We need to construct this virtual mapping for the physical address \textsf{entry}.
+Looking at the update from Specfication Lines 5-10 to 10-14 in Figure \ref{walkpgdir}, we see that the removal of \textsf{entry} from the map $\Xi$ in the invariant to construct the \texstsf{pte-ptsto} (Specfication Line 13 in Figure \ref{walkpgdir}) needed for executing the load instruction (Line 15 in Figure \ref{fig:calltopteinitialize}) to obtain the value of a page entry whose physical address (\textsf{entry}) kept in the register $\texttt{r8}$.
+This magical abstraction -- represented as $\Xi$ in the invariant definition -- provides us to construct this \textsf{pte-ptsto} out of an physical address is called \textit{identity-mappings}.
 \subsubsection{Identity Mappings}
@@
 with DMA devices.}
-\todo{ismail: CHnage the text!  it is not as simple as relating paddrs existence type, more concrete relation on pfn and fresly allocated addr }
 \begin{figure*}
@@
 \end{figure*}
-
-Beyond a frame, the precondition before Line 13 requires the current address space invariant, and knowledge that \textsf{entry} is mapped to a random entry value ($\Xi \;!!\; \mathsf{entry} = \textsf{Some V}$ in Figure \ref{fig:peraspaceinvariant_with_p2v_extension}).
+\textsf{pte\_walkpgdir}, as a client, holds the knowledge that there exists an identity mapping for the physical entry address (\textsf{entry}) in the root page table ($\textsf{L}_{4}$):  $\mathsf{entry} \mapsto_{\textsf{id}} \textsf{\_}$ in Specification Line 3 is a partially owned token for accessing and looking up the resources in the identity map $\Xi$ to construct the \ref\textit{virtual-to-physical} pointsto relation $\textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry \entry\_val}$ with the virtual address (\textsf{entry+KERNBASE}) obtained by offsetting the physical address (\textsf{entry}).
+With this knowledge on the root-page-table-entry, we start traversing the page-table tree which requires locating the address of the next table shown in Figure \ref{fig:calltoptepteinitialize}.
+Beyond a frame, the precondition before Line 13 requires the current address space invariant, and knowledge that \textsf{entry} is mapped to a random entry value in Figure \ref{fig:peraspaceinvariant_with_p2v_extension}).
 By owning this capability, we can also refer to the facts in the invariant that there exists an virtual address from a range set dedicated to page-table entries, as this physical-to-virtual conversion is only guaranteed for page table memory, and obtained by offsetting the physical address ($\textsf{va} = \textsf{pa} + \textsf{KERNBASE}$.

### 2024-07-10 examples.tex
+

@@
 With this knowledge on the root-page-table-entry, we start traversing the page-table tree which requires locating the address of the next table shown in Figure \ref{fig:calltoptepteinitialize}.
-Beyond a frame, the precondition before Line 13 requires the current address space invariant, and knowledge that \textsf{entry} is mapped to a random entry value in Figure \ref{fig:peraspaceinvariant_with_p2v_extension}).
-By owning this capability, we can also refer to the facts in the invariant that there exists an virtual address from a range set dedicated to page-table entries, as this physical-to-virtual conversion is only guaranteed for page table memory, and obtained by offsetting the physical address ($\textsf{va} = \textsf{pa} + \textsf{KERNBASE}$.
-We see that to realize identity mappings, an additional ghost map for keeping identity mappings ($\Xi$) added to the per-address-space
-invariant per Figure \ref{fig:peraspaceinvariant_with_p2v_extension}. The extended invariant owns, for each memory location dedicated to page tables, some fraction of
-that physical memory, for the sake of constructing the virtual-mapping, also the fractional ownership abstracting translation (i.e.
-ghost translation in Figure \ref{fig:peraspaceinvariant_with_p2v_extension}), and guarantees the offset is the correct calculation.
-Subtly, the operation ensures the relevant table entry is readable, but the exact portion of ownership returned must be determined by inspecting the valid bit
+Beyond a frame, the precondition before Line 15 requires the current address space invariant, and knowledge that \textsf{entry} is mapped to a random entry value, subtly, the operation also, at least, requires that the relevant table entry is readable, but the exact portion of ownership returned must be determined by inspecting the valid bit
 of the value in memory --- so full ownership is returned only for unused entries.
@@
 
-Concretely speaking, when we focus on the specfication of Line 13 in Figure \ref{fig:calltopteinitialize} in Figure \ref{fig:lookingupidpampping}, to read the value referenced by physical address \textsf{entry}, as shown in the first specification Line in Figure \ref{fig:lookingupidpampping}, we open our the addresss-space invariant ($\mathcal{I}\textsf{ASpace}_{\textsf{id}}(\ptablestore,\Xi,m)$), and then obtain the fractional ownership of the ghost translation and physical pointsto
-% \[\sumwalkabsent \ast \mathsf{entry}\mapsto_{\mathsf{p,qfrac}}\; \vale \]
-\[\fracghostmaptoken{\delta}{(\texttt{entry + KERNBASE})}{\textsf{entry}}{\textsf{qfrac}} \ast \mathsf{entry}\mapsto_{\mathsf{p,qfrac}}\; \vale \]
-by using the capability $\Xi \;!!\; \textsf{entry} = \textsf{Some V}$ and left the invariant as in the state $\mathcal{I}\textsf{ASpace}_{\textsf{id}}(\ptablestore,\Xi\setminus\{\mathsf{entry}\},m)$ , and construct the virtual mapping with the resources peeled-out of the identity mapping $\Xi$
-\[ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \]
+Concretely speaking, when we focus on the specfication of Line 15 in Figure \ref{fig:calltopteinitialize} in Figure \ref{fig:lookingupidpampping}, to read the value referenced by physical address \textsf{entry}, we need the guarantee
 under the side condition (*)
@@
 \end{equation*}
-assuring that looking the identity mapping for \textsf{entry} is safe under the subtle justification which equates the full ownership to the non/presence of the entry which can only be known when investigated in Line 19in Figure \ref{fig:calltopteinitialize}.
+assuring that looking the identity mapping for \textsf{entry} is safe under the subtle justification which equates the full ownership to the non/presence of the entry which can only be known when investigated in Line 21 in Figure \ref{fig:calltopteinitialize}.
 \begin{comment}

### 2024-07-10 examples.tex
+

@@
 \textsf{pte\_walkpgdir}, as a client, holds the knowledge that there exists an identity mapping for the physical entry address (\textsf{entry}) in the root page table ($\textsf{L}_{4}$):  $\mathsf{entry} \mapsto_{\textsf{id}} \textsf{\_}$ in Specification Line 3 is a partially owned token for accessing and looking up the resources in the identity map $\Xi$ to construct the \ref\textit{virtual-to-physical} pointsto relation $\textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry \entry\_val}$ with the virtual address (\textsf{entry+KERNBASE}) obtained by offsetting the physical address (\textsf{entry}).
-With this knowledge on the root-page-table-entry, we start traversing the page-table tree which requires locating the address of the next table shown in Figure \ref{fig:calltoptepteinitialize}.
+With this knowledge on the root-page-table-entry, we can start traversing the page-table tree which requires locating the address of the next table -- a call to \textsf{pte\_get\_next\_table} shown in Figure \ref{fig:calltoptepteinitialize}.
 Beyond a frame, the precondition before Line 15 requires the current address space invariant, and knowledge that \textsf{entry} is mapped to a random entry value, subtly, the operation also, at least, requires that the relevant table entry is readable, but the exact portion of ownership returned must be determined by inspecting the valid bit
@@
  \end{array}
- \] Lines (17-20) in Figure \ref{fig:calltopteinitialize}.
+ \] Lines (19-22) in Figure \ref{fig:calltopteinitialize}.
  Accessing the presence bit and checking the value allow  us to exploit the fact in the side condition (*) in such a way that we either take the allocation path in the non-presence case (Lines 19-44 in Figure \ref{fig:calltopteinitialize}) with the knowledge of full ownership ($\textsf{qfrac} = 1$) to set the entry with proper values for the bitfields, most imporantly frame referencing freshly allocated page and presence bit, or, take non-allocation path and jump to Line 45 with fractional permission for looking up the entry values.
 
- In case of absence, \textsf{pte\_initialize} is called (Line 31 in Figure \ref{fig:calltopteinitialize}) for a physical-zeroed page (utilizing the page-allocator's \textsf{kalloc} -- currently the only axiomatized call in the the proof of \textsf{pte\_initialize} (Line 9 in Figure \ref{pteinitializespec}, it is not part of \textsf{pte} (virtual-memory-management) module).
+ In case of absence, \textsf{pte\_initialize} is called (Line 33 in Figure \ref{fig:calltopteinitialize}) for a physical-zeroed page (utilizing the page-allocator's \textsf{kalloc} -- currently the only axiomatized call in the the proof of \textsf{pte\_initialize} (Line 9 in Figure \ref{pteinitializespec}, it is not part of \textsf{pte} (virtual-memory-management) module).
 \begin{figure}\footnotesize

### 2024-07-10 examples.tex
+

@@
 \end{figure}
-\todo{ISMAIL: Change the following paragraphs and LInes 33 35  in \ref{fig:calltopteinitialize}, because wpdp2V is where we only know the relation between entry->pfn nextpaddr , consequently zeroed.
-Not here }
-Since we are using \textsf{pte\_initialize} for page-table address allocation, we must relate this newly allocated physical address to the identity mapping map $\Xi$ -- Specification Lines 33 and 35 in Figure \ref{fig:calltopteinitialize}.
+Since we are using \textsf{pte\_initialize} for page-table address allocation, we must relate this newly allocated physical address to the identity mapping map $\Xi$ -- Specification Lines 34 and 38 in Figure \ref{fig:calltopteinitialize}.
 There are two important aspects on relating a freshly allocated page-table physical address to the identity-mapping which we focus in more detail separately in the specification of \textsf{pte\_initialize} in Figure \ref{fig:pteinitializespec}.
 First, one might have expected to see a physical pointsto for the fresly page-table address  ($\mathsf{nextpaddr} \mapsto_{\mathsf{p}} \mathsf{w64\_0}$) after the call to \textsf{pte\_initialize} but we do not have this physical pointsto resource.
-Second, we see the identity mapping for \textsf{nextpaddr}, first not appearing in the identity map (Specification Line 11 in Figure \ref{fig:pteinitializespec}), then as inserted into the identity-map with a value \textsf{Z} ($\Xi \cup \{\mathsf{nextpaddr}\}$) together with the fragmental capability  -- $\ulcorner \Xi \; !!\; \textsf{nextpaddr} = \textsf{ Some Z} \urcorner$ -- at Lines 13 and 14 in Figure \ref{fig:pteinitializespec} .
+Second, we see the identity mapping for the entry frame ($\textsf{entry\_val}_{\textsf{pfn}}$), and all the knowledge on the existence of the identity mappings for both the frame and the page entries indexed from the frame (Line 38 in Figure \ref{fig:calltopteinitialize}) identity mappings exists, which are not part of the invariant yet, which could occur once the presence bit of the entry is set (Line 42 in Figure \ref{fig:calltopteinitialize}).
+Presence bit guard in the extended invariant shown in Figure \ref{fig:peraspaceinvariant_with_p2v_extension} \todo{we need to say something about the improtance of being present and bein inside the invariant.} allows us to preserve the \textit{soundness} of the table-traversal it eliminates all the entries (and tables) that are being under mutation for any other access other than mutation itself.
+%%%%deprecated 
+\begin{comment}
+once the presence bit of the entry for which the invariant  the entry}first not appearing in the identity map (Specification Line 11 in Figure \ref{fig:pteinitializespec}), then as inserted into the identity-map with a value \textsf{Z} ($\Xi \cup \{\mathsf{nextpaddr}\}$) together with the fragmental capability  -- $\ulcorner \Xi \; !!\; \textsf{nextpaddr} = \textsf{ Some Z} \urcorner$ -- at Lines 13 and 14 in Figure \ref{fig:pteinitializespec} .
 The value type \textsf{Z}, unlike the value type \textsf{V} denoting the random value in Line 8 in Figure \ref{fig:calltopteinitialize}, denotes \textsf{zeroed} value.
@@
 \; \textsf{nextpaddr} = \textsf{ Some Z}$ which ensures and carries the \textit{being-zeroed} knowledge when we open and access the the existentially quantified value inside the invariant.
+\end{comment}
+%%%%%%%%%%
   \subsubsection{Kernels Utilizing \textsf{P2V}}

### 2024-07-10 examples.tex
+

@@
 To be able to do so, it needs to load the entry value (a 64-bit word divided into bit-fields for corresponding fileds shown in Section {}), and then,check if \textit{presence} bit of the entry is set or not.
-We see in Specification Line 6 that \textsf{entry} has \textsf{pte-pointsto} that is constructed on the client of \textfsf{pte\_get\_next\_table} -- e.g.\textsf{pte\_walkpgdir} in Figure \ref{walkpgdir}.
+We see in Specification Line 6 that \textsf{entry} has \textsf{pte-pointsto} that is constructed on the client of \textsf{pte\_get\_next\_table} -- e.g.\textsf{pte\_walkpgdir} in Figure \ref{walkpgdir}.
 %wshiftll (wshiftll (natToWord 64 entry) (WordImpl.concat (WordImpl.zero 56) (WordImpl.from_nat 8 12 ^& WordImpl.concat (WordImpl.zero 2) WO~1~1~1~1~1~1)) ^& constf)
@@
 \end{figure}
-Looking at the update from Specfication Lines 5-10 to 10-14 in Figure \ref{walkpgdir}, we see that the removal of \textsf{entry} from the map $\Xi$ in the invariant to construct the \texstsf{pte-ptsto} (Specfication Line 13 in Figure \ref{walkpgdir}) needed for executing the load instruction (Line 15 in Figure \ref{fig:calltopteinitialize}) to obtain the value of a page entry whose physical address (\textsf{entry}) kept in the register $\texttt{r8}$.
+Looking at the update from Specfication Lines 5-10 to 10-14 in Figure \ref{walkpgdir}, we see that the removal of \textsf{entry} from the map $\Xi$ in the invariant to construct the \textsf{pte-ptsto} (Specfication Line 13 in Figure \ref{walkpgdir}) needed for executing the load instruction (Line 15 in Figure \ref{fig:calltopteinitialize}) to obtain the value of a page entry whose physical address (\textsf{entry}) kept in the register $\texttt{r8}$.
 This magical abstraction -- represented as $\Xi$ in the invariant definition -- provides us to construct this \textsf{pte-ptsto} out of an physical address is called \textit{identity-mappings}.
@@
 \end{figure*}
-\textsf{pte\_walkpgdir}, as a client, holds the knowledge that there exists an identity mapping for the physical entry address (\textsf{entry}) in the root page table ($\textsf{L}_{4}$):  $\mathsf{entry} \mapsto_{\textsf{id}} \textsf{\_}$ in Specification Line 3 is a partially owned token for accessing and looking up the resources in the identity map $\Xi$ to construct the \ref\textit{virtual-to-physical} pointsto relation $\textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry \entry\_val}$ with the virtual address (\textsf{entry+KERNBASE}) obtained by offsetting the physical address (\textsf{entry}).
+\textsf{pte\_walkpgdir}, as a client, holds the knowledge that there exists an identity mapping for the physical entry address (\textsf{entry}) in the root page table ($\textsf{L}_{4}$):  $\mathsf{entry} \mapsto_{\textsf{id}} \textsf{\_}$ in Specification Line 3 is a partially owned token for accessing and looking up the resources in the identity map, $\Xi$, to construct the \textit{virtual-to-physical} pointsto relation $\textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry \entry\_val}$ with the virtual address (\textsf{entry+KERNBASE}) obtained by offsetting the physical address (\textsf{entry}).
 With this knowledge on the root-page-table-entry, we can start traversing the page-table tree which requires locating the address of the next table -- a call to \textsf{pte\_get\_next\_table} shown in Figure \ref{fig:calltoptepteinitialize}.

### 2024-07-10 examples.tex
+

@@
 \end{figure}
-Looking at the update from Specfication Lines 5-10 to 10-14 in Figure \ref{walkpgdir}, we see that the removal of \textsf{entry} from the map $\Xi$ in the invariant to construct the \textsf{pte-ptsto} (Specfication Line 13 in Figure \ref{walkpgdir}) needed for executing the load instruction (Line 15 in Figure \ref{fig:calltopteinitialize}) to obtain the value of a page entry whose physical address (\textsf{entry}) kept in the register $\texttt{r8}$.
+Looking at the update from Specfication Lines 3 to 15 in Figure \ref{walkpgdir}, we see that the removal of \textsf{entry} from the map $\Xi$ in the invariant to construct the \textsf{pte-ptsto} (Specfication Line 15 in Figure \ref{walkpgdir}) needed for executing the load instruction (Line 15 in Figure \ref{fig:calltopteinitialize}) to obtain the value of a page entry whose physical address (\textsf{entry}) kept in the register $\texttt{r8}$.
 This magical abstraction -- represented as $\Xi$ in the invariant definition -- provides us to construct this \textsf{pte-ptsto} out of an physical address is called \textit{identity-mappings}.
@@
   \bigast{(\vaddr, \textsf{paddr})\in \ptablestore}{\exists\;(\textsf{l4e, l3e, l2e, l1e, paddr})\ldotp \textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\vaddr\textsf{, l4e, l3e, l2e, l1e, paddr})} \\
-  \bigast{(\paddr,\mathsf{T}) \in \Xi}{\exists\; (\textsf{qfrac, q, val,}\vaddr) \ldotp \ulcorner \vaddr = \paddr + \textsf{KERNBASE} \urcorner \ast \ulcorner \textsf{qfrac} = 1 \leftrightarrow \; \lnot\textsf{entry\_present }(\vale) \urcorner \ast
-   \exists \textsf{freshaddr, entry} \textsf{entry\_pfn} entry freshaddr = paddr \ast freshaddr \mapsto_{\textsf{p}} \textsf{w0\_64}  \leftrightarrow \mathsf{val} = \mathsf{w0\_64}\urcorner \; \ast\\
-   \qquad\qquad\qquad\qquad\qquad\qquad\qquad\qquad   \underbrace{\fracghostmaptoken{\delta}{\vaddr}{\paddr}{\qfrac} }_\text{Ghost translation} \ast 
-% \qquad\qquad\qquad\qquad\qquad\qquad\qquad\qquad    
-\underbrace{\paddr \mapsto_{\mathsf{p}}\{\textsf{qfrac}\}\; \vale}_\text{Physical location}}\\
+  \bigast{(\paddr,\mathsf{T}) \in \Xi}{\exists\; (\textsf{qfrac, q, val,}\vaddr) \ldotp \ulcorner \vaddr = \paddr + \textsf{KERNBASE} \urcorner \ast  \underbrace{\fracghostmaptoken{\delta}{\vaddr}{\paddr}{\qfrac} }_\text{Ghost translation} \ast \underbrace{\paddr \mapsto_{\mathsf{p}}\{\textsf{qfrac}\}\; \vale}_\text{Physical location}}\\
+   \forall_{\textsf{i}\in\textsf{0..511}} \ldotp \paddr + \textsf{i * 8} \mapsto_{\textsf{id,qfacc}} \textsf{mvalidc} \ast \ulcorner \textsf{qfrac} = 1 \leftrightarrow \; \lnot\textsf{entry\_present }(\vale) \urcorner \ast \\
   \textsf{ where } \\
@@
 ;;clean up the stack
-..
+...
 \end{lstlisting}
@@
 $\specline{  \textsf{entry\_val + KERNBASE}_{\textsf{pfn}}  \mapsto_{\textsf{vpte,qfrac}} \mathsf{entry entry\_val}  \textsf{entry\_val}_{\textsf{pfn}}} \ast R_{\textsf{children}}$
-$\specline{ \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{entry\_val}_{\textsf{pfn}}  \ast \texttt{rax} \mapsto_{\textsf{r}} \textsf{entry\_val}_{\textsf{pfn}}  }_{\rtv}$
+$\specline{ \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{entry\_val}_{\textsf{pfn}}  \ast \texttt{rax} \mapsto_{\textsf{r}} \textsf{entry\_val}_{\textsf{pfn}} \ast \textsf{entry\_val}_{\textsf{pfn}} \mapsto_{\textsf{id}} \textsf{\_}}_{\rtv}$
 ...
@@
  ...
- $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}}_{\rtv}$
+ $\specline{\textsf{P} \ast \textsf{entry} \mapsto_{\textsf{id}} \textsf{\_} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}}_{\rtv}$
 ;;set up the stack for root address and virtual address    
  ;;pte_t *pml4_entry = &pml4[PML4EX(va)];
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{entry}\},m)  }_{\rtv}$
-$\specline{ \textsf{entry} \mapsto_{\textsf{id}} \textsf{v} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
-$\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present entry\_val})\urcorner}_{\rtv}$
 mov    -0x8[rbp],rsi
@@
 mov    rsi,-0x18[rbp]
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{entry}\},m)  }_{\rtv}$
+$\specline{ \textsf{entry} \mapsto_{\textsf{id}} \textsf{v} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
+$\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present entry\_val})\urcorner}_{\rtv}$
 ;;pte_t *pdp = pte_get_next_table_succ(pml4_entry);

### 2024-07-10 examples.tex
+

@@
   \bigast{(\paddr,\mathsf{T}) \in \Xi}{\exists\; (\textsf{qfrac, q, val,}\vaddr) \ldotp \ulcorner \vaddr = \paddr + \textsf{KERNBASE} \urcorner \ast  \underbrace{\fracghostmaptoken{\delta}{\vaddr}{\paddr}{\qfrac} }_\text{Ghost translation} \ast \underbrace{\paddr \mapsto_{\mathsf{p}}\{\textsf{qfrac}\}\; \vale}_\text{Physical location}}\\
-   \forall_{\textsf{i}\in\textsf{0..511}} \ldotp \paddr + \textsf{i * 8} \mapsto_{\textsf{id,qfacc}} \textsf{mvalidc} \ast \ulcorner \textsf{qfrac} = 1 \leftrightarrow \; \lnot\textsf{entry\_present }(\vale) \urcorner \ast \\
+   \forall_{\textsf{i}\in\textsf{0..511}} \ldotp \paddr + \textsf{i * 8} \mapsto_{\textsf{id,qfacc}} \textsf{mvalidc} \ast  \underbrace{ \ulcorner \textsf{qfrac} = 1 \leftrightarrow \; \lnot\textsf{entry\_present }(\vale) \urcorner}_\text{Entry validity}\ast \\
   \textsf{ where } \\

### 2024-07-10 examples.tex
+

@@
 To be able to do so, it needs to load the entry value (a 64-bit word divided into bit-fields for corresponding fileds shown in Section {}), and then,check if \textit{presence} bit of the entry is set or not.
-We see in Specification Line 6 that \textsf{entry} has \textsf{pte-pointsto} that is constructed on the client of \textsf{pte\_get\_next\_table} -- e.g.\textsf{pte\_walkpgdir} in Figure \ref{walkpgdir}.
+In the very first precondition on the client side -- e.g.\textsf{pte\_walkpgdir} in Figure \ref{walkpgdir} -- , we see that \textsf{vpte-ptsto} for the entry being accessed (\textsf{entry}) is constructed -- in Specification Line 6 that \textsf{entry}.
+Intuitively speaking, we start traversal from a certain entry address at the current level for which we know ($\textsf{entry} \mapsto_{\textsf{id}} \textsf{\_}$ in Line 3 in Figure \ref{walkpgdir}) that there is a memory mapping ($\textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val}$  Line 15 in Figure \ref{walkpgdir}) that can be used to explore deeper in the page-tree.
 %wshiftll (wshiftll (natToWord 64 entry) (WordImpl.concat (WordImpl.zero 56) (WordImpl.from_nat 8 12 ^& WordImpl.concat (WordImpl.zero 2) WO~1~1~1~1~1~1)) ^& constf)

### 2024-07-10 examples.tex
+

@@
 $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val}  }_{\rtv}$
-$\specline{ \textsf{entry} \mapsto_{\textsf{id}} \textsf{\_} \ast \textsf{entry\_val}_{\textsf{pfn}} \mapsto_{\textsf{id}} \textsf{\_} \ast \ulcorner  \textsf{qfrac} = 1 \land \lnot(\textsf{entry\_present entry\_val}_{\textsf{pfn}})  \urcorner  \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
+$\specline{ \textsf{entry} \mapsto_{\textsf{id}} \textsf{\_} \ast \textsf{entry}_{\textsf{pfn}} \mapsto_{\textsf{id}} \textsf{\_} \ast \ulcorner  \textsf{qfrac} = 1 \land \lnot(\textsf{entry\_present entry\_val}_{\textsf{pfn}})  \urcorner  \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
 $\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \mathsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \land \lnot(\textsf{entry\_present entry\_val})  \urcorner }_{\rtv}$
-$\specline{ \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{entry\_val}_{\textsf{pfn}}  \ast \texttt{rax} \mapsto_{\textsf{r}} \textsf{nextpaddr}  }_{\rtv}$
+$\specline{ \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{entry}_{\textsf{pfn}}  \ast \texttt{rax} \mapsto_{\textsf{r}} \textsf{nextpaddr}  }_{\rtv}$
 $\specline{ \forall_{i\in \textsf{0 ...
-511} } \ldotp  \textsf{entry\_val }_{\textsf{pfn}} + \textsf{i * 8} \mapsto_{\textsf{id}} \textsf{\_} \ast  \textsf{qfrac} = 1 \land \lnot(\textsf{entry\_present entry\_val }_{\textsf{pfn}} + \textsf{i * 8})  }$
+511} } \ldotp  \textsf{entry }_{\textsf{pfn}} + \textsf{i * 8} \mapsto_{\textsf{id}} \textsf{\_} \ast  \textsf{qfrac} = 1 \land \lnot(\textsf{entry\_present entry\_val }_{\textsf{pfn}} + \textsf{i * 8})  }$
 ;;entry value updates
@@
 $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val}  }_{\rtv}$
-$\specline{ \textsf{entry} \mapsto_{\textsf{id}} \textsf{\_} \ast \textsf{entry\_val}_{\textsf{pfn}} \mapsto_{\textsf{id}} \textsf{\_} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \lnot(\textsf{entry\_present entry\_val}_{\textsf{pfn}})  \urcorner  \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
-$\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \mathsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \land \lnot(\textsf{entry\_present entry\_val})  \urcorner }_{\rtv}$
-$\specline{ \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{entry\_val}_{\textsf{pfn}}  \ast \texttt{rax} \mapsto_{\textsf{r}} \textsf{nextpaddr}  }_{\rtv}$
+$\specline{ \textsf{entry} \mapsto_{\textsf{id}} \textsf{\_} \ast \textsf{entry}_{\textsf{pfn}} \mapsto_{\textsf{id}} \textsf{\_} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \lnot(\textsf{entry\_present entry\_val}_{\textsf{pfn}})  \urcorner  \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
+$\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \mathsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \lnot(\textsf{entry\_present entry\_val})  \urcorner }_{\rtv}$
+$\specline{ \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{entry}_{\textsf{pfn}}  \ast \texttt{rax} \mapsto_{\textsf{r}} \textsf{nextpaddr}  }_{\rtv}$
 $\specline{ \forall_{i\in \textsf{0 ...
-511} } \ldotp  \textsf{entry\_val }_{\textsf{pfn}} + \textsf{i * 8} \mapsto_{\textsf{id}} \textsf{\_} \ast  \textsf{qfrac} = 1 \leftrightarrow \lnot(\textsf{entry\_present entry\_val }_{\textsf{pfn}} + \textsf{i * 8})  }$
+511} } \ldotp  \textsf{entry }_{\textsf{pfn}} + \textsf{i * 8} \mapsto_{\textsf{id}} \textsf{\_} \ast  \textsf{qfrac} = 1 \leftrightarrow \lnot(\textsf{entry\_present entry\_val }_{\textsf{pfn}} + \textsf{i * 8})  }$
 $\sqsubseteq$
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \} \cup \{\textsf{entry\_val}_{\textsf{pfn}}) \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val}  }_{\rtv}$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \} \cup \{\textsf{entry}_{\textsf{pfn}}) \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val}  }_{\rtv}$
 $\specline{ \textsf{entry} \mapsto_{\textsf{id}} \textsf{\_} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
-$\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \mathsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \land \lnot(\textsf{entry\_present entry\_val})  \urcorner }_{\rtv}$
-$\specline{ \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{entry\_val}_{\textsf{pfn}}  \ast \texttt{rax} \mapsto_{\textsf{r}} \textsf{entry\_val}_{\textsf{pfn}}  }_{\rtv}$
+$\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \mathsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \lnot(\textsf{entry\_present entry\_val})  \urcorner }_{\rtv}$
+$\specline{ \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{entry}_{\textsf{pfn}}  \ast \texttt{rax} \mapsto_{\textsf{r}} \textsf{entry}_{\textsf{pfn}}  }_{\rtv}$
 ;;}
@@
 \end{figure}
-Looking at the update from Specfication Lines 3 to 15 in Figure \ref{walkpgdir}, we see that the removal of \textsf{entry} from the map $\Xi$ in the invariant to construct the \textsf{pte-ptsto} (Specfication Line 15 in Figure \ref{walkpgdir}) needed for executing the load instruction (Line 15 in Figure \ref{fig:calltopteinitialize}) to obtain the value of a page entry whose physical address (\textsf{entry}) kept in the register $\texttt{r8}$.
+
+The most crucial aspect of page-table entry access relies on the subtle interplay between updating the shared page tables, and the not violating the memory mappings.
+Concretely speaking, accessing to the entry -- whose address is kept in the register $\texttt{r8}$ -- in Line 15 in Figure \ref{fig:calltopteinitialize} requires a memory mapping that is logically provided through the update from Specfication Lines 3 to 15 in Figure \ref{walkpgdir} on the client side by the removal of \textsf{entry} from the map $\Xi$ in the invariant to construct the \textsf{vpte-ptsto} (Specfication Line 15 in Figure \ref{walkpgdir}).
 This magical abstraction -- represented as $\Xi$ in the invariant definition -- provides us to construct this \textsf{pte-ptsto} out of an physical address is called \textit{identity-mappings}.
@@
 \footnotesize
-\todo[inline]{Ismail, look at this figure, something went weird (was already like this, I didn't add this)}
 \[
@@
 
-Concretely speaking, when we focus on the specfication of Line 15 in Figure \ref{fig:calltopteinitialize} in Figure \ref{fig:lookingupidpampping}, to read the value referenced by physical address \textsf{entry}, we need the guarantee
-under the side condition (*)
+Concretely speaking, going back to Line 15 in Figure \ref{fig:calltopteinitialize}, to read the value referenced by physical address \textsf{entry} while preserving the soundness of memory mappings, our extended invariant introduces the side condition (*)
 \begin{equation*}
@@
 
- In case of absence, \textsf{pte\_initialize} is called (Line 33 in Figure \ref{fig:calltopteinitialize}) for a physical-zeroed page (utilizing the page-allocator's \textsf{kalloc} -- currently the only axiomatized call in the the proof of \textsf{pte\_initialize} (Line 9 in Figure \ref{pteinitializespec}, it is not part of \textsf{pte} (virtual-memory-management) module).
+ In case of absence, \textsf{pte\_initialize} is called (Line 33 in Figure \ref{fig:calltopteinitialize}) for a physical page (utilizing the page-allocator's \textsf{kalloc} -- currently the only axiomatized call in the the proof of \textsf{pte\_initialize} (Line 9 in Figure \ref{pteinitializespec}, it is not part of \textsf{pte} (virtual-memory-management) module).
 \begin{figure}\footnotesize
@@
 Since we are using \textsf{pte\_initialize} for page-table address allocation, we must relate this newly allocated physical address to the identity mapping map $\Xi$ -- Specification Lines 34 and 38 in Figure \ref{fig:calltopteinitialize}.
-There are two important aspects on relating a freshly allocated page-table physical address to the identity-mapping which we focus in more detail separately in the specification of \textsf{pte\_initialize} in Figure \ref{fig:pteinitializespec}.
-First, one might have expected to see a physical pointsto for the fresly page-table address  ($\mathsf{nextpaddr} \mapsto_{\mathsf{p}} \mathsf{w64\_0}$) after the call to \textsf{pte\_initialize} but we do not have this physical pointsto resource.
-Second, we see the identity mapping for the entry frame ($\textsf{entry\_val}_{\textsf{pfn}}$), and all the knowledge on the existence of the identity mappings for both the frame and the page entries indexed from the frame (Line 38 in Figure \ref{fig:calltopteinitialize}) identity mappings exists, which are not part of the invariant yet, which could occur once the presence bit of the entry is set (Line 42 in Figure \ref{fig:calltopteinitialize}).
-Presence bit guard in the extended invariant shown in Figure \ref{fig:peraspaceinvariant_with_p2v_extension} \todo{we need to say something about the improtance of being present and bein inside the invariant.} allows us to preserve the \textit{soundness} of the table-traversal it eliminates all the entries (and tables) that are being under mutation for any other access other than mutation itself.
-%%%%deprecated 
-\begin{comment}
-once the presence bit of the entry for which the invariant  the entry}first not appearing in the identity map (Specification Line 11 in Figure \ref{fig:pteinitializespec}), then as inserted into the identity-map with a value \textsf{Z} ($\Xi \cup \{\mathsf{nextpaddr}\}$) together with the fragmental capability  -- $\ulcorner \Xi \; !!\; \textsf{nextpaddr} = \textsf{ Some Z} \urcorner$ -- at Lines 13 and 14 in Figure \ref{fig:pteinitializespec} .
-The value type \textsf{Z}, unlike the value type \textsf{V} denoting the random value in Line 8 in Figure \ref{fig:calltopteinitialize}, denotes \textsf{zeroed} value.
-
-    The first aspect is important for not invalidating the soundness: \textsf{pte\_initialize} extends the identity map with the freshly allocated \textsf{nextpaddr}, and must not reveal the ownership of the physical pointsto in the post-condition (Line 33 and 35 in Figure \ref{fig:calltopteinitialize}, Lines 22-23 in Figure \ref{fig:pteinitializespec}) because it would create unsound coexistence of the full-ownership of the physical pointsto resource (a hypothetical $\mathsf{nextpaddr} \mapsto_{\mathsf{p,1}} w0_64$) and fragmental (potentially full based on the side-condition *) physical pointsto resource that can be obtained from the identity map inside the extended invariant in Figure \ref{fig:peraspaceinvariant_with_p2v_extension} with the capability token $\ulcorner \Xi \;!!;\ \textsf{nextpaddr} = \textsf{ Some Z} \urcorner$.
-Although, the first aspect -- not explicitly revealing the physical pointsto resource for the freshly allocated page-table address but providing access through the identity map inside the invariant with access-capability tokens -- recovers the soundness argument expalined, we still need the knowledge that the fresly allocated address references to a \textit{zeroed} page.
-The invariant preserves this information through the constraint on relating the type (\textsf{T}) of mapped data -- \textsf{T = Z} for zeroed or \textsf{T = V} for random page value -- to the value of existentially quantified value -- \textsf{val} -- in Figure \ref{fig:peraspaceinvariant_with_p2v_extension}.
-So, for accessing the physical pointsto resource to the zeroed value referenced by \textsf{nextpaddr}, we need to have the fragmental ownership of the identity-map $\Xi \; !!
-\; \textsf{nextpaddr} = \textsf{ Some Z}$ which ensures and carries the \textit{being-zeroed} knowledge when we open and access the the existentially quantified value inside the invariant.
-\end{comment}
-%%%%%%%%%%
+Focusing on the specification of \textsf{pte\_initialize} separately in Figure \ref{fig:pteinitializespec}, we right immediately realize that instead of seeing see a physical pointsto for the fresly page-table address (e.g.
+$\mathsf{nextpaddr} \mapsto_{\mathsf{p}} \mathsf{w64\_0}$) deliberately in the post-conditoin in Lines 15-16, we observe a full-ownership token representing the knowledge that a frame and all the entries indexed from this frame are freshly allocated with full-ownership to be a part of the identity map, $\Xi$.
+The soundness argument of this specification relies on the fact that these freshly allocated resources is part of an entry construction that has not been completed yet: the presence bit is set (Line 42 in Figure \ref{fig:calltopteinitialize}) after these freshly allocated resources are incorporated to the entry construction (Line 41 in Figure \ref{fig:calltopteinitialize}).
+In other words, the side condition, (*), acknowledges any access to the entry with these resources as \textit{invalid}, as the entry is revealed to shared-accesses when the presence bit is set.
   \subsubsection{Kernels Utilizing \textsf{P2V}}
   \label{sec:p2v}
- Once we make sure that the frame of the page-table entry ($\textsf{entry}\mapsto\textsf{pfn}$)is backed with an allocated physical page, we must obtain the virtual address of the next level-page table (Line 48 in \ref{calltopteinitialize}).
-To do so, as for obtaining any virtual page-table address, we, again, need to open the invariant look up the \textit{identity mapping} for the physical address ($\textsf{entry}\mapsto \textsf{pfn}$).
+ Once we reach to the certain knowledge of having an entry with a frame referencing to an allocated resource resides inside the identity mappings which can already be known (not taking the allocation path in (Lines 21-53 in Figure \ref{fig:calltopteinitialize}) or by ensuring the allocated frame is made present to shared accesses, then put into identity mappings -- logical update from Specification Lines 43-47 to 49-52 putting $\textsf{entry}_{\textsf{pfn}}$ into $\Xi$ in Figure \ref{fig:calltopteinitialize}), no matther how it is constructed, we can utilize this knowledge to convert this frame address into an virtual address of the next page table through, again, identity mappings -- Line 57 in Figure \ref{fig:calltopteinitialize} specified in Figure \ref{fig:p2v}.
 
 This actually constitutes a very critical piece of the full page table walk verification, and we have verified the critical step for a small x86-64 kernel, which is the physical-to-virtual conversion (often appearing as a macro \texttt{P2V} inside the macro \textsf{ PTE\_PFN\_TO\_ADDR} in C source code).
-In our small kernel (Line 48 in Figure \ref{calltopteinitialize}), as in larger kernels, \texttt{P2V} is actually just addition, but the correctness of this simple instruction is quite subtle and relies on identity mappings explained in the previous section (Section \ref{subsec:identitymappings}) as a part of understanding our address-space invariant in Figure \ref{fig:peraspaceinvariant_with_p2v_extension}.
+In our small kernel (Line 57 in Figure \ref{calltopteinitialize}), as in larger kernels, \texttt{P2V} is actually just addition, but the correctness of this simple instruction is quite subtle and relies on identity mappings explained in the previous section (Section \ref{subsec:identitymappings}) as a part of understanding our address-space invariant in Figure \ref{fig:peraspaceinvariant_with_p2v_extension}.
 \begin{figure}\footnotesize
@@
  add    rcx,rax
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \} \cup \{\textsf{entry\_val}_{\textsf{pfn}}) \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val}  }_{\rtv}$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \} \cup \{\textsf{entry}_{\textsf{pfn}}) \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val}  }_{\rtv}$
 $\specline{ \textsf{entry} \mapsto_{\textsf{id}} \textsf{\_} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
-$\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \mathsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \land \lnot(\textsf{entry\_present entry\_val})  \urcorner }_{\rtv}$
-$\specline{ \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{entry\_val}_{\textsf{pfn}}  \ast \texttt{rax} \mapsto_{\textsf{r}} \textsf{entry\_val}_{\textsf{pfn}}  }_{\rtv}$
+$\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val}  \urcorner }_{\rtv}$
+$\specline{ \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{entry}_{\textsf{pfn}}  \ast \texttt{rax} \mapsto_{\textsf{r}} \textsf{entry}_{\textsf{pfn}} \ast \textsf{entry}_{\textsf{pfn}} \mapsto_{\textsf{id}} \textsf{\_}  }_{\rtv}$
 $\subseteq$
@@
 $\specline{ \textsf{entry} \mapsto_{\textsf{id}} \textsf{\_} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
-$\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \mathsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \land \lnot(\textsf{entry\_present entry\_val})  \urcorner }_{\rtv}$
-$\specline{  \textsf{entry\_val + KERNBASE}_{\textsf{pfn}}  \mapsto_{\textsf{vpte,qfrac}} \mathsf{entry entry\_val}  \textsf{entry\_val}_{\textsf{pfn}}} \ast R_{\textsf{children}}$
-$\specline{ \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{entry\_val}_{\textsf{pfn}}  \ast \texttt{rax} \mapsto_{\textsf{r}} \textsf{entry\_val}_{\textsf{pfn}} \ast \textsf{entry\_val}_{\textsf{pfn}} \mapsto_{\textsf{id}} \textsf{\_}}_{\rtv}$
+$\specline{ \textsf{entry}_{\textsf{pfn}}\textsf{+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val}  }_{\rtv}$
+$\specline{  \textsf{entry}_{\textsf{pfn}} \textsf{+KERNBASE}  \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entrypnf\_val}  \textsf{entry}_{\textsf{pfn}}} \ast R_{\textsf{children}}$
+$\specline{ \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{entry}_{\textsf{pfn}}  \ast \texttt{rax} \mapsto_{\textsf{r}} \textsf{entry}_{\textsf{pfn}} \ast \textsf{entry}_{\textsf{pfn}} \mapsto_{\textsf{id}} \textsf{\_}}_{\rtv}$
 ...
@@
 \end{figure}
-As shown in Figure \ref{fig:p2v}, \lstinline|rcx| holds the offset value \textsf{KERNBASE} which gives us the virtual address (\textsf{nextpaddr+KERNBASE}) to look-up in the identity map ($\Xi\setminus\{entry\}$) of the kernel invariant.
-Assuming the path for \textsf{pte\_initialize} is taken, we are assured that the freshly allocated physical address of the next page table (i.e.
-\textsf{nextpaddr}) is in the identity map ($\Xi$) -- from the post-condition of \textsf{pte\_initialize} in Figure \ref{fig:calltopteinitialize}, so that we can obtain the virtual pte-pointsto for the physical address \textsf{nextpaddr} from the (Line 13 in Figure \ref{fig:p2v}).
+As shown in Figure \ref{fig:p2v}, \lstinline|rcx| holds the offset value \textsf{KERNBASE} which gives us the value of the virtual address ($\textsf{entry}_{\textsf{pfn}}$ \textsf{+KERNBASE}) in the identity map ($\Xi\setminus\{entry\}\cup \{\textsf{entry}_{\textsf{pfn}}) \}$) of the kernel invariant: the logical update in Specification  Lines 5-10 to 10-14 for obtaining virtual-pointsto resource for the frame ($\textsf{entry}_{\textsf{pfn}}$) by removing it from the ghost map ($\Xi\setminus\{entry\}\cup \{\textsf{entry}_{\textsf{pfn}}) \}$) in Line 5 and compute the identity mapping for this physical frame address in Line 13 in Figure \ref{fig:p2v}).
 \subsubsection{Walking Page-Table Tree: Calling \textsf{pte\_get\_nexttable} for Each Level}
@@
 $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{entry}\},m)  }_{\rtv}$
-$\specline{ \textsf{entry} \mapsto_{\textsf{id}} \textsf{v} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
+$\specline{ \textsf{entry} \mapsto_{\textsf{id}} \textsf{\_} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
 $\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present entry\_val})\urcorner}_{\rtv}$
@@
 $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{entry}\},m)  }_{\rtv}$
-$\specline{ \textsf{entry} \mapsto_{\textsf{id}} \textsf{v} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
+$\specline{ \textsf{entry} \mapsto_{\textsf{id}} \textsf{\_} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
 $\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present entry\_val})\urcorner}_{\rtv}$
-$\specline{ \textsf{entry}_{\textsf{pfn}} \mapsto_{\textsf{v,qf}} \textsf{pdp\_val} \ast \textsf{rax} \mapsto_{\textsf{r}} \textsf{entry\_val}_{\textsf{pfn}}}_{\rtv}$    
+$\specline{ \textsf{entry}_{\textsf{pfn}} \mapsto_{\textsf{v,qf}} \textsf{pdp\_val} \ast \textsf{rax} \mapsto_{\textsf{r}} \textsf{entry}_{\textsf{pfn}}}_{\rtv}$    
 ;;pte_t *pdp_entry = &pdp[PDPEX(va)];

### 2024-07-10 examples.tex
a pass over section 5

@@
 Realizing page-table directory walk amounts to calling \textsf{pte\_get\_next\_table} for each level as shown in Figure \ref{walkpgdir}.
-Since there is not much interesting piece additional to the specification we discuss for \textsf{pte\_get\_next\_table} in Section \ref{}, in Figure \ref{}, we present the expected specification after the calls made for each level entry.
-At each specification Line (X,..) in Figure \ref{}, we see the accumulation of virtual-pte pointsto resources returned from \textsf{pte\_get\_next\_tabel} for each level (L4-L1), and finally return L1 level entry that is needed by \textsf{walkpgdir} at Line X in Figure \ref{}.
+The special part of the specficication for a page table walk can be considered as accumulation of memory mappings for the page-table entries visited and frame addresses for page-tables.
+For example, Line 23-24 in Figure \ref{walkpgdir} shows the virtual-pointsto for a L3 page table and the virtual-page-table-entry-pointsto for an L4 entry.
+In the final post-condition, we expect the accumulation of these resources from each level -- $\textsf{R}_{\textsf{walk\_succ}}$ -- which allows us to construct the path to the L1 entry in the tree to insert a new page.
 \begin{figure}\footnotesize
@@
 ;;save the physical next table address in rax
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{entry}\},m)  }_{\rtv}$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{entry},\textsf{entry}_{\textsf{pfn}}\},m)  }_{\rtv}$
 $\specline{ \textsf{entry} \mapsto_{\textsf{id}} \textsf{\_} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$

### 2024-07-11 examples.tex
starting a pass on section 5

@@
 Recall from Section \ref{sec:logic} that virtual points-to assertions work just like regular points-to assertions, by design.
-
-\todo[inline]{need new text explaining both mapping and the page table walk}
+In this section we work through two critical and challenging aspects of virtual memory management.
+First, in several stages, we work up to mapping a new page in the current address space.
+This requires a number of challenging substeps: dynamically traversing a page table to find
+the appropriate L1 entry to update; inserting additional levels of the page table if necessary (updating
+the VMM invariants along the way);
+converting the physical addresses found in intermediate entries into the corresponding virtual addresses
+that can be used for access;
+installing the new mapping;
+and collecting sufficient resources to form a virtual points-to assertion.
+Of these, only the second-to-last step (installing the correct mapping into the
+current address space) has previously been formally verified with respect to a machine model with address translation.
+Second, we formally verify a switch into a new address space as part of a task switch.\todo{Colin will double-check the TLB work's details before claiming this is the first}
 
@@
 The mapping operation of Figure \ref{fig:mapping_code} assumes an operation \textsf{pte\_walkpgdir} which must traverse the page tables
-in order to locate the address of the L1 entry to update --- possibly allocating pages at levels 3 and 2 in the process.
+in order to locate the address of the L1 entry to update --- possibly allocating tables at levels 3, 2, and 1 in the process.
 Traversing the page tables is itself challenging functionality to verify: loading the current table root from \lstinline|cr3| is straightforward
@@
 \subsubsection{Loading Page-Table Address Value}
-We call \textsf{get\_next\_table} to get the virtual address to the next level page-table.
-In Figure \ref{fig:calltopteinitialize}, we present the most critical piece of \textsf{get\_next\_table}, and construct the specification for it.
-Within \textsf{get\_next\_table}, first, we check if the entry whose address is passed as a parameter ($\textsf{entry}$) is present or not (Line 21 in Figure \ref{fig:callptetoinitialize}).
-To be able to do so, it needs to load the entry value (a 64-bit word divided into bit-fields for corresponding fileds shown in Section {}), and then,check if \textit{presence} bit of the entry is set or not.
+We call \textsf{get\_next\_table}\todo{Ismail, this name differs from the figure. Intentional?} to get 
+the virtual address to the next level page-table.
+This section describes the verification of this function, with
+Figure \ref{fig:calltopteinitialize} presenting the most critical piece for discussion.
+Within \textsf{get\_next\_table}, first, we check if the entry whose address is passed as a 
+parameter ($\textsf{entry}$) is present or not (Line 21 in Figure \ref{fig:calltopteinitialize}).
+To be able to do so, it needs to load the entry value (a 64-bit word divided into bit-fields for
+the physical address of the next table, and control bits like the valid bit, as discussed in 
+Section \ref{sec:backgroundonmachinemodel}).
+Next the code checks whether or not the \textit{present} bit of the entry is set.
+\todo[inline,color=blue]{Colin revision frontier}
 In the very first precondition on the client side -- e.g.\textsf{pte\_walkpgdir} in Figure \ref{walkpgdir} -- , we see that \textsf{vpte-ptsto} for the entry being accessed (\textsf{entry}) is constructed -- in Specification Line 6 that \textsf{entry}.

### 2024-07-11 examples.tex
some writing, some reformatting, some addition of proper line number references so reformatting stops breaking line numbers

@@
 \label{sec:traversing}
+We build up to the main 
 The mapping operation of Figure \ref{fig:mapping_code} assumes an operation \textsf{pte\_walkpgdir} which must traverse the page tables
@@
 \subsubsection{Loading Page-Table Address Value}
-We call \textsf{get\_next\_table}\todo{Ismail, this name differs from the figure. Intentional?} to get 
-the virtual address to the next level page-table.
-This section describes the verification of this function, with
-Figure \ref{fig:calltopteinitialize} presenting the most critical piece for discussion.
-Within \textsf{get\_next\_table}, first, we check if the entry whose address is passed as a 
-parameter ($\textsf{entry}$) is present or not (Line 21 in Figure \ref{fig:calltopteinitialize}).
-To be able to do so, it needs to load the entry value (a 64-bit word divided into bit-fields for
+We will discuss access to the level 4 table later. But for subsequent levels, the base address of level $n$ must be
+fetched from the appropriate entry in the level $n+1$ table. This is the role of \lstinline|pte_get_next_table| (Figures \ref{fig:calltopteinitialize} and \ref{fig:p2v}):
+it is passed the virtual address of the page table entry in level $n+1$, and should return the \emph{virtual} address of the \emph{base} of the level $n$ table
+indicated by that entry. If the entry is empty (i.e., this is a sparse part of the page table representation),
+the code also allocates a page for the level $n$ table, installs it in the level $n+1$ entry, and establishes appropriate invariants.
+Figure \ref{fig:calltopteinitialize} presents the initial part of the function, which performs the allocation if necessary.
+Figure \ref{fig:p2v} (discussed in Section \ref{sec:p2v}) deals with the cases where no allocation is necessary \emph{or} the allocation has already
+been performed by the code in this figure.
+Note that none of the verification for this function assumes specific page table levels (it is not even a parameter to the verification);
+it is used for all three level transitions when traversing page tables (4 to 3, 3 to 2, 2 to 1).
+
+Within \textsf{get\_next\_table}, after a standard function prologue, the code 
+loads the entry \textsf{entry} pointed to by the argument.
+This is a page table entry: a 64-bit word divided into bit-fields for
 the physical address of the next table, and control bits like the valid bit, as discussed in 
-Section \ref{sec:backgroundonmachinemodel}).
-Next the code checks whether or not the \textit{present} bit of the entry is set.
+Section \ref{sec:backgroundonmachinemodel}.
+
+Lines \ref{line:mask_present}--\ref{line:check_entry_present} check if the entry's ``present'' bit is set.
+If it is zero, a new page must be allocated for the next level of the table --- which is done by the fall-through
+from Line \ref{line:check_entry_present_jump}'s conditional jump. Otherwise the code jumps ahead to
+the case for the next level already existing, which is discussed in Section \ref{sec:p2v} and Figure \ref{fig:p2v}.
 \todo[inline,color=blue]{Colin revision frontier}
-In the very first precondition on the client side -- e.g.\textsf{pte\_walkpgdir} in Figure \ref{walkpgdir} -- , we see that \textsf{vpte-ptsto} for the entry being accessed (\textsf{entry}) is constructed -- in Specification Line 6 that \textsf{entry}.
+In the very first precondition on the client side -- e.g.\textsf{pte\_walkpgdir} in Figure \ref{walkpgdir} -- 
+we see that \textsf{vpte-ptsto} for the entry being accessed (\textsf{entry}) is constructed -- in Specification Line \ref{line:get_next_vpte_precondition} of Figure \ref{fig:calltopteinitialize} for \textsf{entry}.
 Intuitively speaking, we start traversal from a certain entry address at the current level for which we know ($\textsf{entry} \mapsto_{\textsf{id}} \textsf{\_}$ in Line 3 in Figure \ref{walkpgdir}) that there is a memory mapping ($\textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val}$  Line 15 in Figure \ref{walkpgdir}) that can be used to explore deeper in the page-tree.
@@
 %       (WordImpl.concat (WordImpl.zero 56) (natToWord 8 12 ^& WordImpl.concat (WordImpl.zero 2) WO~1~1~1~1~1~1)) 
-\begin{figure}\footnotesize
-  \begin{lstlisting}[mathescape, basicstyle=\tiny]
+\begin{figure}%\footnotesize
+  \begin{lstlisting}[mathescape, basicstyle=\footnotesize,escapeinside={(*}{*)}]
 ;;pte_t *pte_get_next_table_succ(pte_t *entry) {
@@
 ;; pte_t *next;
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{entry}\},m)  \ast \texttt{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast  \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{next} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{\_} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{\_}  }_{\rtv}$
-$\specline{ \textsf{entry} \mapsto_{\textsf{id}} \textsf{v} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
-$\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present entry\_val})\urcorner}_{\rtv}$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{entry}\},m)  \ast \texttt{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast  \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{next} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{\_} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{\_}  \ast \textsf{entry} \mapsto_{\textsf{id}} \textsf{v} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
+$\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present entry\_val})\urcorner}_{\rtv}$ (*\label{line:get_next_vpte_precondition}*)
 mov    -0x8[rbp],rdi
@@
 $\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present entry\_val})  \urcorner }_{\rtv}$
-and    0x1,rdi
+and    0x1,rdia (* \label{line:mask_present} *)
 mov    rdi,rax
 ;;  if (!entry->present) {
-cmp    0x0,rax
-jne    161 <pte_get_next_table_succ+0xa1>
+cmp    0x0,rax (* \label{line:check_entry_present} *)
+jne    161 <pte_get_next_table_succ+0xa1> (* \label{line:check_entry_present_jump} *) ; Jump if the present bit is not zero
 $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry} \ast \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{next} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val} }_{\rtv}$
@@
 ;;pte_initialize(&next);
-callq  70 <pte_initialize>
+callq  70 <pte_initialize> (* \label{line:call_to_pte_initialize} *)
 $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val}  }_{\rtv}$
@@
 511} } \ldotp  \textsf{entry }_{\textsf{pfn}} + \textsf{i * 8} \mapsto_{\textsf{id}} \textsf{\_} \ast  \textsf{qfrac} = 1 \land \lnot(\textsf{entry\_present entry\_val }_{\textsf{pfn}} + \textsf{i * 8})  }$
-;;entry value updates
-...
-;;entry->pfn = next
-;;entry->present = 1;
+... ;;entry value updates: entry->pfn = next; entry->present = 1;
 $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val}  }_{\rtv}$
@@
 ;;}
-...
-;;uintptr_t next_virt_addr = (uintptr_t) P2V(next_phys_addr);
-movabs 0x0,rcx	                187: R_X86_64_64	KERN_BASE
-add    rcx,rax
-...
-;;next = (pte_t *) next_virt_addr;
-;;clean up the stack and return next
+... ;; Continued in Figure (*\ref{fig:p2v}*)
 \end{lstlisting}
+% ;;uintptr_t next_virt_addr = (uintptr_t) P2V(next_phys_addr);
+% movabs 0x0,rcx	                187: R_X86_64_64	KERN_BASE
+% add    rcx,rax
+% ...
+% ;;next = (pte_t *) next_virt_addr;
+% ;;clean up the stack and return next
+% \end{lstlisting}
 \vspace{-1em}
@@
 The most crucial aspect of page-table entry access relies on the subtle interplay between updating the shared page tables, and the not violating the memory mappings. 
-Concretely speaking, accessing to the entry -- whose address is kept in the register $\texttt{r8}$ -- in Line 15 in Figure \ref{fig:calltopteinitialize} requires a memory mapping that is logically provided through the update from Specfication Lines 3 to 15 in Figure \ref{walkpgdir} on the client side by the removal of \textsf{entry} from the map $\Xi$ in the invariant to construct the \textsf{vpte-ptsto} (Specfication Line 15 in Figure \ref{walkpgdir}).
+Concretely speaking, accessing to the entry -- whose address is kept in the register $\texttt{r8}$ -- in Line 15 in Figure \ref{fig:calltopteinitialize} requires a 
+memory mapping that is logically provided through the update from Specfication Lines 3 to 15 in Figure \ref{walkpgdir} on the client side by the removal of \textsf{entry}
+from the map $\Xi$ in the invariant to construct the \textsf{vpte-ptsto} (Specfication Line 15 in Figure \ref{walkpgdir}). 
 This magical abstraction -- represented as $\Xi$ in the invariant definition -- provides us to construct this \textsf{pte-ptsto} out of an physical address is called \textit{identity-mappings}.
+
 \subsubsection{Identity Mappings}
@@
  After obtaining the identity mapping for \textsf{entry}, we are able to load the \textsf{entry\_val} into \textsf{rdi}, and check the presence bit through
- \[ \begin{array}{l}
-   \mathsf{and }\; 0x1, \;rdi \\
-   \mathsf{mov } \;\mathsf{rdi}, \; \mathsf{rax} \\
-   \mathsf{cmp } \; 0x0, \;\mathsf{rax}
- \end{array}
- \] Lines (19-22) in Figure \ref{fig:calltopteinitialize}.
+ % \[ \begin{array}{l}
+ %   \mathsf{and }\; 0x1, \;rdi \\
+ %   \mathsf{mov } \;\mathsf{rdi}, \; \mathsf{rax} \\
+ %   \mathsf{cmp } \; 0x0, \;\mathsf{rax}
+ % \end{array}
+ % \] 
+Lines \ref{line:mask_present}--\ref{line:check_entry_present} in Figure \ref{fig:calltopteinitialize}.
  Accessing the presence bit and checking the value allow  us to exploit the fact in the side condition (*) in such a way that we either take the allocation path in the non-presence case (Lines 19-44 in Figure \ref{fig:calltopteinitialize}) with the knowledge of full ownership ($\textsf{qfrac} = 1$) to set the entry with proper values for the bitfields, most imporantly frame referencing freshly allocated page and presence bit, or, take non-allocation path and jump to Line 45 with fractional permission for looking up the entry values.
@@
 In other words, the side condition, (*), acknowledges any access to the entry with these resources as \textit{invalid}, as the entry is revealed to shared-accesses when the presence bit is set.
+
   \subsubsection{Kernels Utilizing \textsf{P2V}}
@@
 \begin{figure}\footnotesize
-\begin{lstlisting}[mathescape]
-...
+\begin{lstlisting}[mathescape],escapeinside={(*}{*)}
+... ;; Continued from Figure (*\ref{fig:calltopteinitialize}*)
 ;;uintptr_t next_virt_addr = (uintptr_t) P2V(next<<12);

### 2024-07-11 examples.tex
revisions for explaining the p2v invariant section, with questions

@@
 \begin{figure}%\footnotesize
+\todo[inline]{TODO: resolve entry+KERNBASE question from chat, add information about rax holding \lstinline|entry_val & 0x1| and the conditional
+outcome that that is 0 in the then branch}
   \begin{lstlisting}[mathescape, basicstyle=\footnotesize,escapeinside={(*}{*)}]
@@
 $\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present entry\_val})\urcorner}_{\rtv}$
-mov    [r8],rdi
+mov    [r8],rdi (*\label{line:read_entry_contents}*)
 $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast  \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{next} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val}   }_{\rtv}$
@@
 511} } \ldotp  \textsf{entry }_{\textsf{pfn}} + \textsf{i * 8} \mapsto_{\textsf{id}} \textsf{\_} \ast  \textsf{qfrac} = 1 \land \lnot(\textsf{entry\_present entry\_val }_{\textsf{pfn}} + \textsf{i * 8})  }$
-... ;;entry value updates: entry->pfn = next; entry->present = 1;
+... ;;entry value updates: entry->pfn = next; entry->present = 1; (*\label{line:install_new_entry}*)
 $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val}  }_{\rtv}$
@@
 \label{subsec:identitymappings}
+It is typical for kernels to need to convert between physical and virtual addresses, in both directions.
+Traversing the page tables in software is the simplest way to convert a virtual address to a physical address; this is the context we are working up to.
+However, implementing this virtual-to-physical (V2P) translation ironically requires physical-to-virtual (P2V) translation,
+because the addresses stored in page table entries are physical, but memory accesses issued by the OS code use virtual addresses.
+There is no universal way to convert physical addresses to virtual --- doing so relies on the kernel maintaining careful invariants or
+additional data structures to enable P2V translation.
+
+In practice, VMM operations are performance-critical for many workloads, so most kernels opt for using invariants to make P2V conversion very fast,
+rather than maintaining yet another data structure.
 Most kernels maintain an invariant on their page tables that the virtual address of any page used for a page table lives at a virtual address
-whose value is a constant offset from the physical address a practice sometimes referred to as \emph{identity mapping} (even though the physical-to-virtual translation
-is typically not literally the identity function, but adding a constant offset).\footnote{Some kernels do this for all physical memory on the machine, simplifying interaction
+whose value is \emph{a constant offset from the physical address} --- a practice sometimes referred to as \emph{identity mapping} 
+(even though the physical-to-virtual translation
+is typically not literally the identity function, but adding a non-zero constant offset).\footnote{Some kernels do this for all physical memory on the machine, simplifying interaction
 with DMA devices.}
+
+For this reason we extend the per-address-space invariant as in Figure \ref{fig:peraspaceinvariant_with_p2v_extension}, to also track which
+addresses we can perform a P2V conversion on by a adding a constant offset.
+$\Xi$ is another ghost map, from physical addresses to the unit type (so, effectively a ghost set). 
+\emph{Only} physical addresses in $\Xi$ can undergo P2V conversion. We describe proof of such a conversion in Section \ref{sec:p2v},
+but describe the invariant here because installing a new level 3/2/1 table requires maintaining that invariant.
+
 \begin{figure*}
@@
 \end{figure*}
+
+For each $\paddr\in\Xi$, the invariant tracks a virtual points-to justifying that virtual address $\paddr+\textsf{KERNBASE}$ maps to physical address $\paddr$
+(the ``Ghost translation'' in Figure \ref{fig:peraspaceinvariant_with_p2v_extension}),
+fractional ownership of the physical memory for that page table entry,
+and 
+\todo[inline,color=red]{
+  Another Figure 8 issue: last line provide membership in the id map / $\Xi$ not for pa plus whatever, but for \emph{the table base in val} plus whatever,
+  also only when the entry is present. I know we whiteboarded this way and sketched it in Zulip this way, and I'm pretty sure I saw something when
+  working on issue 19, but need to check w/ Ismail.
+}
+The fractional ownership of the entry's physical memory is subtle. Recall that $\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$ retains some physical
+ownership of each page table entry that is traversed (proportional to how many virtual addresses share the entry).
+So in general the invariant cannot keep full permission to the memory in this part of the invariant, or it would overlap the page table walk for virtual points-to
+assertions. But in the case where the entry is invalid, we may need to write to it (e.g., to install a reference to a next-level table, as we do in Figure \ref{fig:calltopteinitialize}),
+which requires full permission. Fortunately, the entry can only be in use if its valid bit is set; if the valid bit is not set we know
+that no virtual points-to assertions in $\delta$/$\theta$ have any partial ownership.
+Thus we use the invariant portion annotated as ``Entry validity'' in Figure \ref{fig:peraspaceinvariant_with_p2v_extension} to capture this:
+if the entry is invalid the invariant holds full ownership of the entry, so it can be updated; while if the entry is valid,
+the invariant owns only a constant non-zero fragment sufficient to read the entry, but not modify it (which would invalidate some virtual points-to assertions).
+
+Thus the fractional ownership of the physical location is enough for Line \ref{line:read_entry_contents} to access the entry, though in \lstinline|get_next_table|
+the caller has pulled that piece of information out of the invariant and passed it for the entry at hand (which is why this appears explicitly in assertions).
+Line \ref{line:check_entry_present_jump}'s conditional then determines in the fall-through case that the bit is not set, which resolves $\textsf{qfrac} = 1$,
+and permits the new entry to be installed (in ellided code around Line \ref{line:install_new_entry})
+
+This seemingly-simple piece of code has a highly non-trivial correctness argument, which depends critically on detailed invariants on how access to page table
+entries is shared between parts of the kernel. No prior work has engaged with this problem.
+
+\todo[inline,color=blue]{Colin revision frontier: need to resolve entry+KERNBASE issue to finish pulling this into prose above}
+
 \textsf{pte\_walkpgdir}, as a client, holds the knowledge that there exists an identity mapping for the physical entry address (\textsf{entry}) in the root page table ($\textsf{L}_{4}$):  $\mathsf{entry} \mapsto_{\textsf{id}} \textsf{\_}$ in Specification Line 3 is a partially owned token for accessing and looking up the resources in the identity map, $\Xi$, to construct the \textit{virtual-to-physical} pointsto relation $\textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry \entry\_val}$ with the virtual address (\textsf{entry+KERNBASE}) obtained by offsetting the physical address (\textsf{entry}).
@@
 \begin{figure}\footnotesize
-  \begin{lstlisting}[mathescape]
+  \begin{lstlisting}[mathescape,escapeinside={(*}{*)}]
 
@@
 \begin{figure}\footnotesize
-\begin{lstlisting}[mathescape],escapeinside={(*}{*)}
+\begin{lstlisting}[mathescape,escapeinside={(*}{*)}]
 ... ;; Continued from Figure (*\ref{fig:calltopteinitialize}*)
@@
 \begin{figure}\footnotesize
-\begin{lstlisting}[mathescape]
+\begin{lstlisting}[mathescape,escapeinside={(*}{*)}]
 ;;pte_t *walkpgdir_succ(pte_t *pml4, const void *va) {
@@
 \begin{figure}\footnotesize
-  \begin{lstlisting}[mathescape]
+  \begin{lstlisting}[mathescape,escapeinside={(*}{*)}]
 ;;complstatus_t vaspace_mappage(pte_t *pml4, void *va,uintptr_t fpaddr, uint64_t perm ) {
@@
 \begin{figure}\footnotesize
-\begin{lstlisting}[mathescape]
+\begin{lstlisting}[mathescape,escapeinside={(*}{*)}]
 ;; Assume the save-space is in rdi, load-space in rsi. First, save the yielding context

### 2024-07-11 examples.tex
..

@@
 \footnotesize
+\todo[inline]{C wants to add authoritative ownership of the ghost map of $\Xi$ so talking about the tokens makes sense}
 \[
@@
   \bigast{(\paddr,\mathsf{T}) \in \Xi}{\exists\; (\textsf{qfrac, q, val,}\vaddr) \ldotp \ulcorner \vaddr = \paddr + \textsf{KERNBASE} \urcorner \ast  \underbrace{\fracghostmaptoken{\delta}{\vaddr}{\paddr}{\qfrac} }_\text{Ghost translation} \ast \underbrace{\paddr \mapsto_{\mathsf{p}}\{\textsf{qfrac}\}\; \vale}_\text{Physical location}}\\
-   \forall_{\textsf{i}\in\textsf{0..511}} \ldotp \paddr + \textsf{i * 8} \mapsto_{\textsf{id,qfacc}} \textsf{mvalidc} \ast  \underbrace{ \ulcorner \textsf{qfrac} = 1 \leftrightarrow \; \lnot\textsf{entry\_present }(\vale) \urcorner}_\text{Entry validity}\ast \\
+   \underbrace{ \ulcorner \textsf{qfrac} = 1 \leftrightarrow \; \lnot\textsf{entry\_present }(\vale) \urcorner}_\text{Entry validity}\ast (\ulcorner\textsf{entry\_present}(val)\urcorner \wand \forall_{\textsf{i}\in\textsf{0..511}} \ldotp ((\paddr + \textsf{i * 8}) \mapsto_{\textsf{id,qfacc}} \_)) \\
   \textsf{ where } \\
@@
 Thus the fractional ownership of the physical location is enough for Line \ref{line:read_entry_contents} to access the entry, though in \lstinline|get_next_table|
-the caller has pulled that piece of information out of the invariant and passed it for the entry at hand (which is why this appears explicitly in assertions).
+the caller has pulled that piece of information out of the invariant and passed it for the entry at hand, which is why this appears explicitly in assertions,
+and the argument to the invariant is $\Xi\setminus\{\mathsf{entry}\}$\todo{update pending entry+KB} (indexing by the set $\Xi$ allows us to borrow the physical resources
+for a specific page table entry out of the invariant, and later put them back).
 Line \ref{line:check_entry_present_jump}'s conditional then determines in the fall-through case that the bit is not set, which resolves $\textsf{qfrac} = 1$,

### 2024-07-12 examples.tex
more figure updates, scaffolding prose smoothing in 5.1

@@
 \begin{figure}%\footnotesize
-\todo[inline]{TODO: resolve entry+KERNBASE question from chat, add information about rax holding \lstinline|entry_val & 0x1| and the conditional
-outcome that that is 0 in the then branch}
+\todo[inline]{TODO: resolve entry+KERNBASE question from chat, update all rdi points-tos after line \ref{line:pass_addrof_next}}
   \begin{lstlisting}[mathescape, basicstyle=\footnotesize,escapeinside={(*}{*)}]
 ;;pte_t *pte_get_next_table_succ(pte_t *entry) {
-;; setting up the stack
+... ;; setting up the stack
 ;; pte_t *next;
@@
 cmp    0x0,rax (* \label{line:check_entry_present} *)
-jne    161 <pte_get_next_table_succ+0xa1> (* \label{line:check_entry_present_jump} *) ; Jump if the present bit is not zero
+jne    161 <pte_get_next_table_succ+0xa1> (* \label{line:check_entry_present_jump} *) ;; Jump if the present bit is not zero
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry} \ast \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{next} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val \& 0x1} }_{\rtv}$
+$\specline{  \textsf{entry} \mapsto_{\textsf{id}} \textsf{v} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  \ast \texttt{rax}  \mapsto_{\textsf{r}} \textsf{entry\_val \& 0x1} \ast \ulcorner\textsf{entry\_val \& 0x1}=\textsf{0x0}\urcorner}_{\rtv}$
+$\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow  \lnot(\textsf{entry\_val \& 0x1}=\textsf{0x1})  \urcorner }_{\rtv}$ (*\label{line:before_concluding_qfrac1}*)
+mov    rbp,rdi
+sub    0x10, rdi ;; Store the value of rbp minus 16 bytes (address of (*\textsf{next}*)) into rdi (*\label{line:pass_addrof_next}*)
 $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry} \ast \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{next} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val}  }_{\rtv}$
 $\specline{  \textsf{entry} \mapsto_{\textsf{id}} \textsf{v} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta} }_{\rtv}$
-$\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \land \lnot(\textsf{entry\_present entry\_val})  \urcorner }_{\rtv}$
-mov    rbp,rdi
-sub    0x10, rdi
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry} \ast \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{next} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val}  }_{\rtv}$
-$\specline{  \textsf{entry} \mapsto_{\textsf{id}} \textsf{v} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta} }_{\rtv}$
-$\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \land \lnot(\textsf{entry\_present entry\_val})  \urcorner }_{\rtv}$
+$\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \land \lnot(\textsf{entry\_present entry\_val})  \urcorner }_{\rtv}$ (*\label{line:after_concluding_qfrac1}*)
 ;;pte_initialize(&next);
@@
 for a specific page table entry out of the invariant, and later put them back).
-Line \ref{line:check_entry_present_jump}'s conditional then determines in the fall-through case that the bit is not set, which resolves $\textsf{qfrac} = 1$,
-and permits the new entry to be installed (in ellided code around Line \ref{line:install_new_entry})
+Line \ref{line:check_entry_present_jump}'s conditional then determines in the fall-through case that the bit is not set, which 
+together with other facts entails $\textsf{qfrac} = 1$ at Line \ref{line:after_concluding_qfrac1},
+and permits storing a new entry (in ellided code around Line \ref{line:install_new_entry})
 
@@
 \end{comment}
+
  \subsubsection{Ensuring Presence of an Entry}
+\todo[inline]{some of this content is now earlier, rename to ``installing a new entry'' and just talk about the invariant update from installing the newly-allocated
+entry. include explanation of the resources returned from initialization call.
+Carefully check speck and then explain code of pte initialize, which currently has not prose (though this is a candidate for appendix in an oopsla submission)
+}
  After obtaining the identity mapping for \textsf{entry}, we are able to load the \textsf{entry\_val} into \textsf{rdi}, and check the presence bit through

### 2024-07-18 examples.tex
more work on examples text, figures

@@
 $\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow  \lnot(\textsf{entry\_val \& 0x1}=\textsf{0x1})  \urcorner }_{\rtv}$ (*\label{line:before_concluding_qfrac1}*)
-mov    rbp,rdi
+mov    rbp,rdi (*\label{line:alloc_path_start}*)
 sub    0x10, rdi ;; Store the value of rbp minus 16 bytes (address of (*\textsf{next}*)) into rdi (*\label{line:pass_addrof_next}*)
@@
 $\specline{ \forall_{i\in \textsf{0 ...
-511} } \ldotp  \textsf{entry }_{\textsf{pfn}} + \textsf{i * 8} \mapsto_{\textsf{id}} \textsf{\_} \ast  \textsf{qfrac} = 1 \land \lnot(\textsf{entry\_present entry\_val }_{\textsf{pfn}} + \textsf{i * 8})  }$
+511} } \ldotp  \ghostmaptoken{\textsf{id}}{((\mathsf{entry\_page}\;\textsf{entry}) + \textsf{i * 8})}{\textsf{\_}} \ast  \textsf{qfrac} = 1 \land \lnot(\textsf{entry\_present entry\_val }_{\textsf{pfn}} + \textsf{i * 8})  }$
 ... ;;entry value updates: entry->pfn = next; entry->present = 1; (*\label{line:install_new_entry}*)
@@
 $\specline{ \forall_{i\in \textsf{0 ...
-511} } \ldotp  \textsf{entry }_{\textsf{pfn}} + \textsf{i * 8} \mapsto_{\textsf{id}} \textsf{\_} \ast  \textsf{qfrac} = 1 \leftrightarrow \lnot(\textsf{entry\_present entry\_val }_{\textsf{pfn}} + \textsf{i * 8})  }$
+511} } \ldotp \ghostmaptoken{\textsf{id}}{((\mathsf{entry\_page}\;\textsf{entry}) + \textsf{i * 8})}{\textsf{\_}} \ast  \textsf{qfrac} = 1 \leftrightarrow \lnot(\textsf{entry\_present entry\_val }_{\textsf{pfn}} + \textsf{i * 8})  }$
 $\sqsubseteq$
@@
 $\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \mathsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \lnot(\textsf{entry\_present entry\_val})  \urcorner }_{\rtv}$
-$\specline{ \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{entry}_{\textsf{pfn}}  \ast \texttt{rax} \mapsto_{\textsf{r}} \textsf{entry}_{\textsf{pfn}}  }_{\rtv}$
+$\specline{ \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{entry}_{\textsf{pfn}}  \ast \texttt{rax} \mapsto_{\textsf{r}} \textsf{entry}_{\textsf{pfn}}  }_{\rtv}$ (*\label{line:alloc_path_end}*)
 ;;}
@@
 \footnotesize
-\todo[inline]{C wants to add authoritative ownership of the ghost map of $\Xi$ so talking about the tokens makes sense}
 \[
 \begin{array}{l}
-  \mathcal{I}\textsf{ASpace}_{\textsf{id}}(\ptablestore,\Xi,m)\stackrel{\triangle}{=}\\ \textsf{ASpace\_Lookup}_{\textsf{id}}(\ptablestore,\Xi,m) \ast 
-  \bigast{(\vaddr, \textsf{paddr})\in \ptablestore}{\exists\;(\textsf{l4e, l3e, l2e, l1e, paddr})\ldotp \textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\vaddr\textsf{, l4e, l3e, l2e, l1e, paddr})} \\
-  \bigast{(\paddr,\mathsf{T}) \in \Xi}{\exists\; (\textsf{qfrac, q, val,}\vaddr) \ldotp \ulcorner \vaddr = \paddr + \textsf{KERNBASE} \urcorner \ast  \underbrace{\fracghostmaptoken{\delta}{\vaddr}{\paddr}{\qfrac} }_\text{Ghost translation} \ast \underbrace{\paddr \mapsto_{\mathsf{p}}\{\textsf{qfrac}\}\; \vale}_\text{Physical location}}\\
-   \underbrace{ \ulcorner \textsf{qfrac} = 1 \leftrightarrow \; \lnot\textsf{entry\_present }(\vale) \urcorner}_\text{Entry validity}\ast (\ulcorner\textsf{entry\_present}(val)\urcorner \wand \forall_{\textsf{i}\in\textsf{0..511}} \ldotp ((\paddr + \textsf{i * 8}) \mapsto_{\textsf{id,qfacc}} \_)) \\
+  \mathcal{I}\textsf{ASpace}_{\textsf{id}}(\ptablestore,\Xi,m)\stackrel{\triangle}{=} \textsf{ASpace\_Lookup}_{\textsf{id}}(\ptablestore,\Xi,m) \ast \mathsf{GhostMap}(\mathsf{id},\Xi)\ast\\
+  \left(\bigast{(\vaddr, \textsf{paddr})\in \ptablestore}{\exists\;(\textsf{l4e, l3e, l2e, l1e, paddr})\ldotp \textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\vaddr\textsf{, l4e, l3e, l2e, l1e, paddr})}\right)\ast \\
+  \bigast{(\paddr,\mathsf{T}) \in \Xi}{\exists\; (\textsf{qfrac, q, val,}\vaddr) \ldotp \ulcorner \vaddr = \paddr + \textsf{KERNBASE} \urcorner \ast  \underbrace{\fracghostmaptoken{\delta}{\vaddr}{\paddr}{\qfrac} }_\text{Ghost translation} \ast \underbrace{\paddr \mapsto_{\mathsf{p}}\{\textsf{qfrac}\}\; \vale}_\text{Physical location}} \ast\\
+   \qquad\underbrace{ \ulcorner \textsf{qfrac} = 1 \leftrightarrow \; \lnot\textsf{entry\_present }(\vale) \urcorner}_\text{Entry validity}\ast 
+    \underbrace{\left(\ulcorner\textsf{entry\_present}(\vale)\urcorner \wand \forall_{\textsf{i}\in\textsf{0..511}} \ldotp \ghostmaptoken{\textsf{id}}{((\mathsf{entry\_page}\;\vale) + \textsf{i * 8})}{\_}\right)}_{\text{Indexing into next level of tables}} \\
   \textsf{ where } \\
@@
 For each $\paddr\in\Xi$, the invariant tracks a virtual points-to justifying that virtual address $\paddr+\textsf{KERNBASE}$ maps to physical address $\paddr$
-(the ``Ghost translation'' in Figure \ref{fig:peraspaceinvariant_with_p2v_extension}),
-fractional ownership of the physical memory for that page table entry,
-and 
-\todo[inline,color=red]{
-  Another Figure 8 issue: last line provide membership in the id map / $\Xi$ not for pa plus whatever, but for \emph{the table base in val} plus whatever,
-  also only when the entry is present. I know we whiteboarded this way and sketched it in Zulip this way, and I'm pretty sure I saw something when
-  working on issue 19, but need to check w/ Ismail.
+(the ``Ghost translation'' in Figure \ref{fig:peraspaceinvariant_with_p2v_extension});
+fractional ownership of the physical memory for that page table entry;
+and for valid entries (with the present bit set), ghost map tokens for every entry in the table pointed to by the entry.
+
+\todo[inline,color=green]{Colin says: 
+The indexing into the next level of tables is a little weird at the L1 level. The page pointed to by an L1 entry will be a data
+page. We don't actually need to perform P2V on data pages in general. But also problematic is that this invariant implies
+partial ownership of the data (by taking the $\Xi$ tokens for the data page back to this invariant). So we need a way
+to make this ``bottom out'' after the L1 entries. Something like switching $\Xi$ to be a map from a physical address to the corresponding level of tables
+(as a natural number; we'll always start out knowing that $\ghostmaptoken{\textsf{id}}{\textsf{cr3val}}{4}$ for a traversal),
+and then switching the indexing into the next level to be
+\[
+\bigast{(\paddr,\mathsf{level})}{\ldots\left(\ulcorner\textsf{entry\_present}(\vale)\land \mathsf{level} > 1 \urcorner \wand \forall_{\textsf{i}\in\textsf{0..511}} \ldotp \ghostmaptoken{\textsf{id}}{((\mathsf{page\_addr}\;\vale) + \textsf{i * 8})}{(\mathsf{level}-1)}\right)}\]
+I'm surprised this hasn't come up during page table setup
 }
+
 The fractional ownership of the entry's physical memory is subtle. Recall that $\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$ retains some physical
@@
 
- \subsubsection{Ensuring Presence of an Entry}
-\todo[inline]{some of this content is now earlier, rename to ``installing a new entry'' and just talk about the invariant update from installing the newly-allocated
-entry. include explanation of the resources returned from initialization call.
-Carefully check speck and then explain code of pte initialize, which currently has not prose (though this is a candidate for appendix in an oopsla submission)
-}
+ \subsubsection{Installing a New Table}
  After obtaining the identity mapping for \textsf{entry}, we are able to load the \textsf{entry\_val} into \textsf{rdi}, and check the presence bit through
@@
 Lines \ref{line:mask_present}--\ref{line:check_entry_present} in Figure \ref{fig:calltopteinitialize}.
- Accessing the presence bit and checking the value allow  us to exploit the fact in the side condition (*) in such a way that we either take the allocation path in the non-presence case (Lines 19-44 in Figure \ref{fig:calltopteinitialize}) with the knowledge of full ownership ($\textsf{qfrac} = 1$) to set the entry with proper values for the bitfields, most imporantly frame referencing freshly allocated page and presence bit, or, take non-allocation path and jump to Line 45 with fractional permission for looking up the entry values.
+Accessing the presence bit and checking the value allows us to exploit the side condition (*) when verifying the allocation
+path for when the entry is invalid (Lines \ref{line:alloc_path_start}--\ref{line:alloc_path_end} in Figure \ref{fig:calltopteinitialize}),
+as it entails full ownership of the entry's memory ($\textsf{qfrac} = 1$) and justifies writing to that memory.
+Otherwise, the code jumps past the end of this listing, to the following code at the top of Figure \ref{fig:p2v} (which is also the
+continuation of this code).
 
- In case of absence, \textsf{pte\_initialize} is called (Line 33 in Figure \ref{fig:calltopteinitialize}) for a physical page (utilizing the page-allocator's \textsf{kalloc} -- currently the only axiomatized call in the the proof of \textsf{pte\_initialize} (Line 9 in Figure \ref{pteinitializespec}, it is not part of \textsf{pte} (virtual-memory-management) module).
+If the entry is not set, \textsf{pte\_initialize} is called (Line \ref{line:call_to_pte_initialize} in Figure \ref{fig:calltopteinitialize}) 
+for a physical page (utilizing the page-allocator's \textsf{kalloc} -- currently the only axiomatized call in the the proof of \textsf{pte\_initialize} 
+(Line \ref{line:call_to_kalloc} in Figure \ref{pteinitializespec}). 
 \begin{figure}\footnotesize
@@
 ;;pte_t *local = kalloc();
-callq  81 <pte_initialize+0x11>	7d: R_X86_64_PLT32	kalloc-0x4
+callq  81 <pte_initialize+0x11>	7d: R_X86_64_PLT32	kalloc-0x4 (*\label{line:call_to_kalloc}*)
 mov    rax,-0x10[rbp] 
@@
 \end{figure}
+\todo[inline,color=blue]{Colin revision frontier}
 Since we are using \textsf{pte\_initialize} for page-table address allocation, we must relate this newly allocated physical address to the identity mapping map $\Xi$ -- Specification Lines 34 and 38 in Figure \ref{fig:calltopteinitialize}.

### 2024-07-24 examples.tex
added level to the extended inv

@@
   \left(\bigast{(\vaddr, \textsf{paddr})\in \ptablestore}{\exists\;(\textsf{l4e, l3e, l2e, l1e, paddr})\ldotp \textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\vaddr\textsf{, l4e, l3e, l2e, l1e, paddr})}\right)\ast \\
-  \bigast{(\paddr,\mathsf{T}) \in \Xi}{\exists\; (\textsf{qfrac, q, val,}\vaddr) \ldotp \ulcorner \vaddr = \paddr + \textsf{KERNBASE} \urcorner \ast  \underbrace{\fracghostmaptoken{\delta}{\vaddr}{\paddr}{\qfrac} }_\text{Ghost translation} \ast \underbrace{\paddr \mapsto_{\mathsf{p}}\{\textsf{qfrac}\}\; \vale}_\text{Physical location}} \ast\\
+  \bigast{(\paddr,\mathsf{level}) \in \Xi}{\exists\; (\textsf{qfrac, q, val,}\vaddr) \ldotp \ulcorner \vaddr = \paddr + \textsf{KERNBASE} \; \textsf{level} > 1\urcorner \ast  \underbrace{\fracghostmaptoken{\delta}{\vaddr}{\paddr}{\qfrac} }_\text{Ghost translation} \ast \underbrace{\paddr \mapsto_{\mathsf{p}}\{\textsf{qfrac}\}\; \vale}_\text{Physical location}} \ast\\
    \qquad\underbrace{ \ulcorner \textsf{qfrac} = 1 \leftrightarrow \; \lnot\textsf{entry\_present }(\vale) \urcorner}_\text{Entry validity}\ast 
-    \underbrace{\left(\ulcorner\textsf{entry\_present}(\vale)\urcorner \wand \forall_{\textsf{i}\in\textsf{0..511}} \ldotp \ghostmaptoken{\textsf{id}}{((\mathsf{entry\_page}\;\vale) + \textsf{i * 8})}{\_}\right)}_{\text{Indexing into next level of tables}} \\
+    \underbrace{\left(\ulcorner\textsf{entry\_present}(\vale)\urcorner \wand \forall_{\textsf{i}\in\textsf{0..511}} \ldotp \ghostmaptoken{\textsf{id}}{((\mathsf{entry\_page}\;\vale) + \textsf{i * 8})}{\textsf{level-1}}\right)}_{\text{Indexing into next level of tables}} \\
   \textsf{ where } \\

### 2024-07-24 examples.tex
spec of initialize fixed

@@
 
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)   \ast \texttt{rbp-16} \mapsto_{\textsf{r}} \textsf{\_} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{table\_addr}  }_{\rtv}$
-$\specline{\texttt{rax} \mapsto_{\textsf{r}} \textsf{\_} \ast \textsf{table\_addr} \mapsto_{\mathsf{v}} \mathsf{\_}\ast\ghostmaptoken{\delta{}s}{\rtv}{\delta} }_{\rtv}$
-;;void pte_initialize(pte_t **table) {
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)   \ast \texttt{rbp-16} \mapsto_{\textsf{r}} \textsf{\_} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry+KERNBASE}  }_{\rtv}$
+$\specline{\texttt{rax} \mapsto_{\textsf{r}} \textsf{\_} \ast \lnot(\textsf{entry\_present entry\_val} }_{\rtv}$
+$\specline{\textsf{entry+KERNBASE} \mapsto_{\textsf{pte,1}} \textsf{entry entryv} \ast \ghostmaptoken{\textsf{id}}{(\mathsf{entry})}{\textsf{level}} }$
+;;void pte_initialize(pte_t *entry) {
 ...
@@
 mov    rax,-0x10[rbp] 
-;;*table = local;
+;;entry->pfn = PTE_ADDR_TO_PFN((uintptr_t) local);
 mov    -0x10[rbp],rax
@@
 mov    rax,[rdi]
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)   \ast \texttt{rbp-16} \mapsto_{\textsf{r}} \textsf{\_} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{table\_addr}  }_{\rtv}$
-$\specline{\texttt{rax} \mapsto_{\textsf{v}} \textsf{nextpaddr} \ast \textsf{table\_addr} \mapsto_{\mathsf{v}} \mathsf{\mathsf{nextpaddr}} \ast \textsf{nextpaddr} \mapsto_{\textsf{id}} \textsf{\_}\ast \forall_{i\in\textsf{0..511}} \textsf{nextpaddr + i * 8} \mapsto_{\textsf{id}} \textsf{\_} \ast\ghostmaptoken{\delta{}s}{\rtv}{\delta} }_{\rtv}$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)   \ast \texttt{rbp-16} \mapsto_{\textsf{r}} \textsf{\_} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry+KERNBASE} \ast \texttt{rax} \mapsto_{\textsf{v}} \textsf{nextpaddr}  }_{\rtv}$
+$\specline{\textsf{entry+KERNBASE} \mapsto_{\textsf{pte,q}} \textsf{entry entryv} \ast \ghostmaptoken{\textsf{id}}{(\mathsf{entry})}{\textsf{level}} }$
+$\specline{\texttt{rax} \mapsto_{\textsf{v}} \textsf{nextpaddr} \ast \lnot(\textsf{entry\_present entry\_val}) }$
+$\specline{ \textsf{entry+KERNBASE} \mapsto_{\mathsf{pte,1}} \mathsf{entry} \; \textsf{pfn\_set(entryv nextpaddr)}}$
+$\specline{\ghostmaptoken{\delta{}s}{\rtv}{\delta} \ast \begin{array}{l} \textsf{entry\_present(pte\_initialize(pfn\_set(entryv nextpaddr)))} -* \\  \;\;\;\;\;\;\; \forall_{i\in\textsf{0..511} \; \textsf{table\_root(pte\_initialized (pfn\_set (entryv nextpaddr))) + i * 8} \mapsto_{\textsf{id}} \textsf{level-1}    \end{array}  }_{\rtv}$
 ;;clean up the stack

### 2024-07-24 examples.tex
+

--- 2024-07-24 examples.tex
@@
 
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)   \ast \texttt{rbp-16} \mapsto_{\textsf{r}} \textsf{\_} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{table\_addr}  }_{\rtv}$
-$\specline{\texttt{rax} \mapsto_{\textsf{r}} \textsf{\_} \ast \textsf{table\_addr} \mapsto_{\mathsf{v}} \mathsf{\_}\ast\ghostmaptoken{\delta{}s}{\rtv}{\delta} }_{\rtv}$
-;;void pte_initialize(pte_t **table) {
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)   \ast \texttt{rbp-16} \mapsto_{\textsf{r}} \textsf{\_} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry+KERNBASE}  }_{\rtv}$
+$\specline{\texttt{rax} \mapsto_{\textsf{r}} \textsf{\_} \ast \lnot(\textsf{entry\_present entry\_val} }_{\rtv}$
+$\specline{\textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,1}} \textsf{entry entryv} \ast \ghostmaptoken{\textsf{id}}{(\mathsf{entry})}{\textsf{level}} }$
+;;void pte_initialize(pte_t *entry) {
 ...
@@
 mov    rax,[rdi]
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)   \ast \texttt{rbp-16} \mapsto_{\textsf{r}} \textsf{\_} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{table\_addr}  }_{\rtv}$
-$\specline{\texttt{rax} \mapsto_{\textsf{v}} \textsf{nextpaddr} \ast \textsf{table\_addr} \mapsto_{\mathsf{v}} \mathsf{\mathsf{nextpaddr}} \ast \textsf{nextpaddr} \mapsto_{\textsf{id}} \textsf{\_}\ast \forall_{i\in\textsf{0..511}} \textsf{nextpaddr + i * 8} \mapsto_{\textsf{id}} \textsf{\_} \ast\ghostmaptoken{\delta{}s}{\rtv}{\delta} }_{\rtv}$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)   \ast \texttt{rbp-16} \mapsto_{\textsf{r}} \textsf{\_} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry+KERNBASE} \ast \texttt{rax} \mapsto_{\textsf{v}} \textsf{nextpaddr}  }_{\rtv}$
+$\specline{\textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,}} \textsf{entry entryv} \ast \ghostmaptoken{\textsf{id}}{(\mathsf{entry})}{\textsf{level}} }$
+$\specline{\texttt{rax} \mapsto_{\textsf{v}} \textsf{nextpaddr} \ast \lnot(\textsf{entry\_present entry\_val}) }$
+$\specline{ \textsf{entry+KERNBASE} \mapsto_{\mathsf{pte,1}} \mathsf{entry} \; \textsf{pfn\_set(entryv nextpaddr)}}$
+$\specline{\ghostmaptoken{\delta{}s}{\rtv}{\delta} \ast \begin{array}{l} \textsf{entry\_present(pte\_initialize(pfn\_set(entryv nextpaddr)))} -* \\  \;\;\;\;\;\;\; \forall_{i\in\textsf{0..511} \; \textsf{table\_root(pte\_initialized (pfn\_set (entryv nextpaddr))) + i * 8} \mapsto_{\textsf{id}} \textsf{level-1}    \end{array}  }_{\rtv}$
 ;;clean up the stack

### 2024-07-24 examples.tex
figure 10

@@
 $\specline{ \textsf{entry+KERNBASE} \mapsto_{\mathsf{pte,1}} \mathsf{entry} \; \textsf{pfn\_set(entryv nextpaddr)}}$
-$\specline{\ghostmaptoken{\delta{}s}{\rtv}{\delta} \ast \begin{array}{l} \textsf{entry\_present(pte\_initialize(pfn\_set(entryv nextpaddr)))} -* \\  \;\;\;\;\;\;\; \forall_{i\in\textsf{0..511} \; \textsf{table\_root(pte\_initialized (pfn\_set (entryv nextpaddr))) + i * 8} \mapsto_{\textsf{id}} \textsf{level-1}    \end{array}  }_{\rtv}$
+$\specline{\ghostmaptoken{\delta{}s}{\rtv}{\delta} \ast \begin{array}{l} \textsf{entry\_present(pte\_initialize(pfn\_set(entryv nextpaddr)))} -* \\  \;\;\;\;\;\;\; \forall_{i\in\textsf{0..511}} \; \textsf{table\_root(pte\_initialized (pfn\_set (entryv nextpaddr))) + i * 8} \mapsto_{\textsf{id}} \textsf{level-1}    \end{array}  }_{\rtv}$
 ;;clean up the stack
@@
 $\specline{ \textsf{entry} \mapsto_{\textsf{id}} \textsf{\_} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
-$\specline{ \textsf{entry}_{\textsf{pfn}}\textsf{+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val}  }_{\rtv}$
-$\specline{  \textsf{entry}_{\textsf{pfn}} \textsf{+KERNBASE}  \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entrypnf\_val}  \textsf{entry}_{\textsf{pfn}}} \ast R_{\textsf{children}}$
+$\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val}  }_{\rtv}$
+$\specline{  \textsf{entry}_{\textsf{pfn}} \textsf{+KERNBASE}  \mapsto_{\textsf{vpte,qfrac}}   \textsf{entry}_{\textsf{pfn}} \; \textsf{entry\_pfn\_val} \ast R_{\textsf{children}}}_{\rtv}$
 $\specline{ \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{entry}_{\textsf{pfn}}  \ast \texttt{rax} \mapsto_{\textsf{r}} \textsf{entry}_{\textsf{pfn}} \ast \textsf{entry}_{\textsf{pfn}} \mapsto_{\textsf{id}} \textsf{\_}}_{\rtv}$

### 2024-07-24 examples.tex
walkpgdir
+

@@
  ...
- $\specline{\textsf{P} \ast \textsf{entry} \mapsto_{\textsf{id}} \textsf{\_} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}}_{\rtv}$
+ $\specline{\textsf{P} \ast \textsf{plm4} \mapsto_{\textsf{id}} \textsf{\_} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}}_{\rtv}$
 ;;set up the stack for root address and virtual address    
@@
 mov    rsi,-0x18[rbp]
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{entry}\},m)  }_{\rtv}$
-$\specline{ \textsf{entry} \mapsto_{\textsf{id}} \textsf{\_} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
-$\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present entry\_val})\urcorner}_{\rtv}$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{plm4\_entry}\},m)  }_{\rtv}$
+$\specline{ \ghostmaptoken{\textsf{id}}{(\mathsf{plm4\_entry})}{\textsf{4}} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
+$\specline{\begin{array}{l} \textsf{entry\_present(l4e\_val)} -* \\  \;\;\;\;\;\;\; \forall_{i\in\textsf{0..511}} \; \textsf{table\_root(plm4)}_{pfn} \textsf{ + i * 8} \mapsto_{\textsf{id}} \textsf{3}    \end{array}}_{\rtv} $
+$\specline{ \textsf{plm4\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{plm4 l4e\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l4e\_val})\urcorner}_{\rtv}$
 ;;pte_t *pdp = pte_get_next_table_succ(pml4_entry);
@@
 ;;save the physical next table address in rax
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{entry},\textsf{entry}_{\textsf{pfn}}\},m)  }_{\rtv}$
-$\specline{ \textsf{entry} \mapsto_{\textsf{id}} \textsf{\_} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
-$\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present entry\_val})\urcorner}_{\rtv}$
-$\specline{ \textsf{entry}_{\textsf{pfn}} \mapsto_{\textsf{v,qf}} \textsf{pdp\_val} \ast \textsf{rax} \mapsto_{\textsf{r}} \textsf{entry}_{\textsf{pfn}}}_{\rtv}$    
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{plm4\_entry}\},m)  }_{\rtv}$
+$\specline{ \ghostmaptoken{\textsf{id}}{(\mathsf{plm4\_entry})}{\textsf{4}}  \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
+$\specline{\forall_{i\in\textsf{0..511}} \; \textsf{table\_root(plm4\_entry}_{\textsf{pfn}}) \textsf{ + i * 8} \mapsto_{\textsf{id}} \textsf{3} } $
+$\specline{ \textsf{plm4\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{plm4\_entry l4e\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l4e\_val})\urcorner}_{\rtv}$
+$\specline{\texttt{rax}  \mapsto_{\textsf{r}} \textsf{table\_root(pml4\_entry}_{\textsf{pfn}})}_{\rtv}$
 ;;pte_t *pdp_entry = &pdp[PDPEX(va)];
+$\specline{\textsf{pdp}+\textsf{PDPEX(va)} = \textsf{pdp\_entry} \land \textsf{table\_root(pml4\_entry}_{\textsf{pfn}})  = \textsf{pdp}}_{\rtv}$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{plm4\_entry},\textsf{pdp\_entry}\},m)  }_{\rtv}$
+$\specline{\ghostmaptoken{\textsf{id}}{(\mathsf{plm4\_entry})}{\textsf{4}} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
+$\specline{\forall_{i\in\textsf{0..511}} \; \textsf{table\_root(plm4\_entry}_{\textsf{pfn}}) \textsf{ + i * 8} \mapsto_{\textsf{id}} \textsf{3} }_{\rtv}$
+$\specline{ \textsf{plm4\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l4e\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present entry\_val})\urcorner}_{\rtv}$
+$\specline{ \textsf{pdp\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l3e\_val} \ast \ghostmaptoken{\textsf{id}}{(\mathsf{pdp\_entry})}{\textsf{3}}}_{\rtv}$
+$\specline{\begin{array}{l} \textsf{entry\_present(l3e\_val)} -* \\  \;\;\;\;\;\;\; \forall_{i\in\textsf{0..511}} \; \textsf{table\_root(pdp\_entry}_{\textsf{pfn}}) \textsf{ + i * 8} \mapsto_{\textsf{id}} \textsf{2}    \end{array}}_{\rtv} $
+$\specline{\ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l3e\_val})\urcorner}_{\rtv}$
 ;;pte_t *pd = pte_get_next_table_succ(pdp_entry);
@@
 ...
+ $\specline{\textsf{P} \ast \textsf{plm4} \mapsto_{\textsf{id}} \textsf{\_} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}}_{\rtv}$      
 ;;pte = walkpgdir_succ(pml4, a);
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m)  \ast \texttt{rdi}\mapsto_{r} \textsf{plm4} \ast \texttt{rsi}\mapsto_{r} \textsf{va} \ast  \texttt{rdx}\mapsto_{r} \textsf{pa} \ast \texttt{rcx}\mapsto_{r} \textsf{perm} \ast \texttt{rax}\mapsto_{\textsf{r}} \textsf{pte\_addr} \ast }_{\rtv}$
-$\specline{\ulcorner \texttt{aligned va} \land  \theta \; !!\; \vaddr = \texttt{None}\urcorner\ast \textsf{vl4} \mapsto_{\textsf{vpte,qfrac}} \textsf{l4e l4} \ast \textsf{vl3} \mapsto_{\textsf{vpte,qfrac}} \textsf{l3e l3} \ast \textsf{vl2} \mapsto_{\textsf{vpte,qfrac}} \textsf{l2e pa}  \ast }_{\rtv}$
-$\specline{\textsf{pte\_addr} \mapsto_{\textsf{vpte,qfrac}} \textsf{pa} \textsf{pte\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \lnot (\textsf{entry\_present pte\_val}) \urcorner}_{\rtv}$
-$\specline{\texttt{rbp-8}\mapsto_{v} \textsf{plm4} \ast \texttt{rbp-16}\mapsto_{v} \textsf{va}  \ast \texttt{rbp-24}\mapsto_{v} \textsf{pa} \ast \texttt{rbp-32}\mapsto_{v} \textsf{perm} \ast \textsf{r8} \mapsto_{\textsf{r}} \textsf{\_}}_{\rtv}$
 mov    -0x8[rbp],rdi
@@
 ;;pte->nocache      = PTE_NOCACHE_SET(perm);
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m)  \ast \texttt{rdi}\mapsto_{r} \textsf{plm4} \ast \texttt{rsi}\mapsto_{r} \textsf{va} \ast  \texttt{rdx}\mapsto_{r} \textsf{pa} \ast \texttt{rcx}\mapsto_{r} \textsf{perm} \ast \texttt{rax}\mapsto_{\textsf{r}} \textsf{pte\_addr} \ast \ulcorner \texttt{aligned va} \land  \theta \; !!\; \vaddr = \texttt{None}\urcorner\ast }_{\rtv}$
-$\specline{\textsf{pte\_addr} \mapsto_{\textsf{vpte,qfrac}} \textsf{pa} \textsf{pte\_val} \ast \ulcorner \lnot \textsf{qfrac} = 1 \land \lnot(\textsf{entry\_present pte\_val}) \urcorner}$     
-$\specline{\texttt{rbp-8}\mapsto_{v} \textsf{plm4} \ast \texttt{rbp-16}\mapsto_{v} \textsf{va}  \ast \texttt{rbp-24}\mapsto_{v} \textsf{pa} \ast \texttt{rbp-32}\mapsto_{v} \textsf{perm} \ast \textsf{r8} \mapsto_{\textsf{r}} \textsf{\_}}_{\rtv}$
+$\specline{ \left( \begin{array}{l} \textsf{pdp}+\textsf{PDPEX(va)} = \textsf{pdp\_entry} \land \textsf{table\_root(pml4\_entry}_{\textsf{pfn}})  = \textsf{pdp} \\ \textsf{pd}+\textsf{PDEX(va)} = \textsf{pd\_entry} \land \textsf{table\_root(pml3\_entry}_{\textsf{pfn}})  = \textsf{pd} \\ \textsf{pt}+\textsf{PDPEX(va)} = \textsf{pt\_entry} \land \textsf{table\_root(pml2\_entry}_{\textsf{pfn}})  = \textsf{pt} \\  \textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{plm4\_entry},\textsf{pdp\_entry},\textsf{pd\_entry},\textsf{pt\_entry}\},m)  \\ \ghostmaptoken{\delta{}s}{\rtv}{\delta} \\
+\ghostmaptoken{\textsf{id}}{(\mathsf{plm4\_entry})}{\textsf{4}}\ast \ghostmaptoken{\textsf{id}}{(\mathsf{plm3\_entry})}{\textsf{3}} \ast \ghostmaptoken{\textsf{id}}{(\mathsf{plm2\_entry})}{\textsf{2}}  \\ \forall_{i\in\textsf{0..511}} \; \textsf{table\_root(plm4\_entry}_{\textsf{pfn}}) \textsf{ + i * 8} \mapsto_{\textsf{id}} \textsf{3}  \\ \forall_{i\in\textsf{0..511}} \; \textsf{table\_root(pdp\_entry}_{\textsf{pfn}}) \textsf{ + i * 8} \mapsto_{\textsf{id}} \textsf{2}  \\ \forall_{i\in\textsf{0..511}} \; \textsf{table\_root(pd\_entry}_{\textsf{pfn}}) \textsf{ + i * 8} \mapsto_{\textsf{id}} \textsf{1}  \\  \textsf{plm4\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l4e\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l4e\_val})\urcorner \\  \textsf{pdp\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l3e\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l3e\_val})\urcorner \\  \textsf{pd\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l2e\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l2e\_val})\urcorner \\   \textsf{pt\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l1e\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l1e\_val})\urcorner    \end{array} \right)  =\textsf{R}_{\textsf{walk\_acc}}}$
 pte->pfn          = PTE_ADDR_TO_PFN(pa);
@@
 mov    [rdx],rsi
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m)  \ast \texttt{rdi}\mapsto_{r} \textsf{plm4} \ast \texttt{rsi}\mapsto_{r} \textsf{va} \ast  \texttt{rdx}\mapsto_{r} \textsf{pa} \ast \texttt{rcx}\mapsto_{r} \textsf{perm} \ast \texttt{rax}\mapsto_{\textsf{r}} \textsf{pte\_addr} \ast \ulcorner \texttt{aligned va} \land  \theta \; !!\; \vaddr = \texttt{None}\urcorner\ast }_{\rtv}$
-$\specline{\textsf{pte\_addr} \mapsto_{\textsf{vpte,qfrac}} \textsf{pa} (\textsf{pa}>>12) \ast \ulcorner \lnot \textsf{qfrac} = 1 \land \lnot(\textsf{entry\_present pte\_val}) \urcorner}$     
-$\specline{\texttt{rbp-8}\mapsto_{v} \textsf{plm4} \ast \texttt{rbp-16}\mapsto_{v} \textsf{va}  \ast \texttt{rbp-24}\mapsto_{v} \textsf{pa} \ast \texttt{rbp-32}\mapsto_{v} \textsf{perm} \ast \textsf{r8} \mapsto_{\textsf{r}} \textsf{\_}}_{\rtv}$
+$\specline{\textsf{pt\_entry}_{\textsf{pfn}} \mapsto_{p} \textsf{addr\_to\_pfn}(\textsf{pa})}_{\rtv}$
 movabs 0xffffffffff,rdi
@@
 \end{lstlisting}
-  \caption{Specification and proof of code for mapping a new page.}
+  \caption{Specification and proof of code for mapping a new page with $\textsf{R}_{\textsf{walk\_acc}}$ in Figure \ref{walkpgdir} expanded.}
 \label{fig:mapping_code}

### 2024-07-26 examples.tex
done with figures
fix up alignment, rdi in Figures 7 and 10

@@
 \begin{figure}%\footnotesize
-\todo[inline]{TODO: resolve entry+KERNBASE question from chat, update all rdi points-tos after line \ref{line:pass_addrof_next}}
 \begin{lstlisting}[mathescape, basicstyle=\footnotesize,escapeinside={(*}{*)}]
@@
 $\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present entry\_val})  \urcorner }_{\rtv}$
-and    0x1,rdia (* \label{line:mask_present} *)
+and    0x1,rdi (* \label{line:mask_present} *)
 mov    rdi,rax
-;;  if (!entry->present) {
-cmp    0x0,rax (* \label{line:check_entry_present} *)
+cmp    0x0,rax ;;  if (!entry->present) {(* \label{line:check_entry_present} *)
 jne    161 <pte_get_next_table_succ+0xa1> (* \label{line:check_entry_present_jump} *) ;; Jump if the present bit is not zero
@@
 sub    0x10, rdi ;; Store the value of rbp minus 16 bytes (address of (*\textsf{next}*)) into rdi (*\label{line:pass_addrof_next}*)
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry} \ast \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{next} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val}  }_{\rtv}$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry} \ast \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{next} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{rbp - 16} }_{\rtv}$
 $\specline{  \textsf{entry} \mapsto_{\textsf{id}} \textsf{v} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta} }_{\rtv}$
 $\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \land \lnot(\textsf{entry\_present entry\_val})  \urcorner }_{\rtv}$ (*\label{line:after_concluding_qfrac1}*)
-;;pte_initialize(&next);
-callq  70 <pte_initialize> (* \label{line:call_to_pte_initialize} *)
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val}  }_{\rtv}$
+callq  70 <pte_initialize> ;;pte_initialize(&next);(* \label{line:call_to_pte_initialize} *)
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{\_}  }_{\rtv}$
 $\specline{ \textsf{entry} \mapsto_{\textsf{id}} \textsf{\_} \ast \textsf{entry}_{\textsf{pfn}} \mapsto_{\textsf{id}} \textsf{\_} \ast \ulcorner  \textsf{qfrac} = 1 \land \lnot(\textsf{entry\_present entry\_val}_{\textsf{pfn}})  \urcorner  \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
@@
 ... ;;entry value updates: entry->pfn = next; entry->present = 1; (*\label{line:install_new_entry}*)
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val}  }_{\rtv}$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{\_}  }_{\rtv}$
 $\specline{ \textsf{entry} \mapsto_{\textsf{id}} \textsf{\_} \ast \textsf{entry}_{\textsf{pfn}} \mapsto_{\textsf{id}} \textsf{\_} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \lnot(\textsf{entry\_present entry\_val}_{\textsf{pfn}})  \urcorner  \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
@@
 $\sqsubseteq$
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \} \cup \{\textsf{entry}_{\textsf{pfn}}) \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val}  }_{\rtv}$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \} \cup \{\textsf{entry}_{\textsf{pfn}}) \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{\_} }_{\rtv}$
 $\specline{ \textsf{entry} \mapsto_{\textsf{id}} \textsf{\_} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
@@
 ;;}
-... ;; Continued in Figure (*\ref{fig:p2v}*)
+... ;; Code after conditional continued in Figure (*\ref{fig:p2v}*)
 \end{lstlisting}
 % ;;uintptr_t next_virt_addr = (uintptr_t) P2V(next_phys_addr);
-% movabs 0x0,rcx	                187: R_X86_64_64	KERN_BASE
+% movabs KERNBASE,rcx
 % add    rcx,rax
@@
    \qquad\underbrace{ \ulcorner \textsf{qfrac} = 1 \leftrightarrow \; \lnot\textsf{entry\_present }(\vale) \urcorner}_\text{Entry validity}\ast 
-    \underbrace{\left(\ulcorner\textsf{entry\_present}(\vale)\urcorner \wand \forall_{\textsf{i}\in\textsf{0..511}} \ldotp \ghostmaptoken{\textsf{id}}{((\mathsf{entry\_page}\;\vale) + \textsf{i * 8})}{\textsf{level-1}}\right)}_{\text{Indexing into next level of tables}} \\
+    \underbrace{\left(\ulcorner\textsf{present\_L}(\vale,\mathsf{level})\urcorner \wand \forall_{\textsf{i}\in\textsf{0..511}} \ldotp \ghostmaptoken{\textsf{id}}{((\mathsf{entry\_page}\;\vale) + \textsf{i * 8})}{\textsf{level-1}}\right)}_{\text{Indexing into next level of tables}} \\
   \textsf{ where } \\
@@
    % \ownGhost\gammaPred{\authfull{\ptableabswalk\ptablestore}} \ast  \ownGhostpv\gammaPred{\authfull{\pvmapping\Xi}}
-   \ptableabswalk{\delta,\ptablestore} \ast \pvmapping{\delta,\Xi}
+   \ptableabswalk{\delta,\ptablestore} \ast \pvmapping{\delta,\Xi}\\
+  \textsf{present\_L}(\vale,\mathsf{level})\stackrel{\triangle}{=} \mathsf{entry\_present}(\vale)\land \mathsf{level} > 0
 
@@
 ;;uintptr_t next_virt_addr = (uintptr_t) P2V(next<<12);
- movabs 0x0,rcx	                187: R_X86_64_64	KERN_BASE
+ movabs KERNBASE,rcx
  add    rcx,rax
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \} \cup \{\textsf{entry}_{\textsf{pfn}}) \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val}  }_{\rtv}$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \} \cup \{\textsf{entry}_{\textsf{pfn}}) \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{\_}  }_{\rtv}$
 $\specline{ \textsf{entry} \mapsto_{\textsf{id}} \textsf{\_} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
@@
 $\subseteq$
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val}  }_{\rtv}$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{\_}  }_{\rtv}$
 $\specline{ \textsf{entry} \mapsto_{\textsf{id}} \textsf{\_} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
@@
 %mov    -0x18[rbp],rax
-%movabs 0x0,rcx            187: R_X86_64_64	KERN_BASE
+%movabs KERN_BASE,rcx
 %add    rcx,rax
@@
 mov    [rdx],rsi
-$\specline{\textsf{pt\_entry}_{\textsf{pfn}} \mapsto_{p} \textsf{addr\_to\_pfn}(\textsf{pa})}_{\rtv}$
+$\specline{\textsf{pt\_entry}_{\textsf{pfn}} \mapsto_{p} \textsf{addr\_to\_pfn}(\textsf{pa}) \ast \textsf{R}_{\textsf{walk\_acc}}}_{\rtv}$
+$\sqsubseteq$
+$\specline{\textsf{va} \mapsto_{\textsf{v,q}} \textsf{pa} \ast \textsf{P}}$
 movabs 0xffffffffff,rdi

### 2024-07-26 examples.tex
more writing, more fiddly details

@@
 We will discuss access to the level 4 table later. But for subsequent levels, the base address of level $n$ must be
-fetched from the appropriate entry in the level $n+1$ table. This is the role of \lstinline|pte_get_next_table| (Figures \ref{fig:calltopteinitialize} and \ref{fig:p2v}):
-it is passed the virtual address of the page table entry in level $n+1$, and should return the \emph{virtual} address of the \emph{base} of the level $n$ table
-indicated by that entry. If the entry is empty (i.e., this is a sparse part of the page table representation),
+fetched from the appropriate entry in the level $n+1$ table.
+This is the role of \lstinline|pte_get_next_table| (Figures \ref{fig:calltopteinitialize} and \ref{fig:p2v}):
+it is passed the virtual address of the page table entry in level $n+1$, and should return the \emph{virtual} 
+address of the \emph{base} of the level $n$ table
+indicated by that entry.
+If the entry is empty (i.e., this is a sparse part of the page table representation),
 the code also allocates a page for the level $n$ table, installs it in the level $n+1$ entry, and establishes appropriate invariants.
@@
 been performed by the code in this figure.
+
 Note that none of the verification for this function assumes specific page table levels (it is not even a parameter to the verification);
 it is used for all three level transitions when traversing page tables (4 to 3, 3 to 2, 2 to 1).
+This comes into play with a subtlety of the specification of \lstinline|pte_get_next_table| that we will
+revisit several times: \lstinline|pte_get_next_table|'s specification
+assumes it is given a virtual \emph{vpte-pointsto}\footnote{A \emph{vpte-pointsto}
+$\vaddr\mapsto_{\textsf{vpte,qfrac}} \textsf{\paddr~\val}$ is a virtual points-to granting
+access to virtual address $\vaddr$, which is known to hold \textsf{\val}, but additionally exposing the physical address
+$\paddr$ that the virtual address translates to. This is commonly used, as the name suggests, when updating PTEs,
+where we require assurance that we are accessing the intended PTE.
+Its definition is just like the regular virtual points-to in Figure \ref{fig:virtualpointstosharing},
+except taking the physical address as a parameter rather than existentially quantifying it.
+}
+ granting access to the specified entry,
+but its postcondition does not yield new virtual points-to assertions!
+Instead it merely computes the base address of the next table, and returns adequate capabilities (discussed in Section \ref{subsec:identitymappings})
+for the \emph{caller} to construct a vpte-pointsto for the next table level (if this is not an L1 entry ---
+the caller knows which level of the table this is for).
+\looseness=-1
 
@@
 the case for the next level already existing, which is discussed in Section \ref{sec:p2v} and Figure \ref{fig:p2v}.
-\todo[inline,color=blue]{Colin revision frontier}
-In the very first precondition on the client side -- e.g.\textsf{pte\_walkpgdir} in Figure \ref{walkpgdir} -- 
-we see that \textsf{vpte-ptsto} for the entry being accessed (\textsf{entry}) is constructed -- in Specification Line \ref{line:get_next_vpte_precondition} of Figure \ref{fig:calltopteinitialize} for \textsf{entry}.
-Intuitively speaking, we start traversal from a certain entry address at the current level for which we know ($\textsf{entry} \mapsto_{\textsf{id}} \textsf{\_}$ in Line 3 in Figure \ref{walkpgdir}) that there is a memory mapping ($\textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val}$  Line 15 in Figure \ref{walkpgdir}) that can be used to explore deeper in the page-tree.
+First, we must discuss another refinement of the address space invariant, establishing
+enough structure on the page tables themselves to allow the traversal.
+The code for allocating a new level of the page table must establish this extended invariant.
+
 %wshiftll (wshiftll (natToWord 64 entry) (WordImpl.concat (WordImpl.zero 56) (WordImpl.from_nat 8 12 ^& WordImpl.concat (WordImpl.zero 2) WO~1~1~1~1~1~1)) ^& constf)
@@
 $\specline{ \forall_{i\in \textsf{0 ...
-511} } \ldotp  \ghostmaptoken{\textsf{id}}{((\mathsf{entry\_page}\;\textsf{entry}) + \textsf{i * 8})}{\textsf{\_}} \ast  \textsf{qfrac} = 1 \land \lnot(\textsf{entry\_present entry\_val }_{\textsf{pfn}} + \textsf{i * 8})  }$
+511} } \ldotp  \ghostmaptoken{\textsf{id}}{((\mathsf{entry\_page}\;\textsf{entry}) + \textsf{i * 8})}{\textsf{\_}} \ast  \textsf{qfrac} = 1 \land \lnot(\textsf{entry\_present entry\_val }_{\textsf{pfn}} + \textsf{i * 8})  }$ (*\label{line:page_of_caps}*)
 ... ;;entry value updates: entry->pfn = next; entry->present = 1; (*\label{line:install_new_entry}*)
@@
 
-The most crucial aspect of page-table entry access relies on the subtle interplay between updating the shared page tables, and the not violating the memory mappings. 
-Concretely speaking, accessing to the entry -- whose address is kept in the register $\texttt{r8}$ -- in Line 15 in Figure \ref{fig:calltopteinitialize} requires a 
-memory mapping that is logically provided through the update from Specfication Lines 3 to 15 in Figure \ref{walkpgdir} on the client side by the removal of \textsf{entry}
-from the map $\Xi$ in the invariant to construct the \textsf{vpte-ptsto} (Specfication Line 15 in Figure \ref{walkpgdir}). 
-This magical abstraction -- represented as $\Xi$ in the invariant definition -- provides us to construct this \textsf{pte-ptsto} out of an physical address is called \textit{identity-mappings}.
-
 \subsubsection{Identity Mappings}
@@
 Traversing the page tables in software is the simplest way to convert a virtual address to a physical address; this is the context we are working up to.
-However, implementing this virtual-to-physical (V2P) translation ironically requires physical-to-virtual (P2V) translation,
+However, implementing this virtual-to-physical (V2P) translation in this way ironically requires physical-to-virtual (P2V) translation,
 because the addresses stored in page table entries are physical, but memory accesses issued by the OS code use virtual addresses.
@@
 is typically not literally the identity function, but adding a non-zero constant offset).\footnote{Some kernels do this for all physical memory on the machine, simplifying interaction
-with DMA devices.}
+with DMA devices.
+On newer platforms like RISC-V, this sometimes truly is an identity mapping ---
+x86-64 machines are forced into offsets by backwards compatibility with bootloaders that cannot access the full memory space of the
+machine.
+}
 
@@
 addresses we can perform a P2V conversion on by a adding a constant offset.
-$\Xi$ is another ghost map, from physical addresses to the unit type (so, effectively a ghost set). 
+$\Xi$ is another ghost map, from physical addresses to the level of the page table they represent (1--4).
 \emph{Only} physical addresses in $\Xi$ can undergo P2V conversion. We describe proof of such a conversion in Section \ref{sec:p2v},
@@
 fractional ownership of the physical memory for that page table entry;
-and for valid entries (with the present bit set), ghost map tokens for every entry in the table pointed to by the entry.
-
-\todo[inline,color=green]{Colin says: 
-The indexing into the next level of tables is a little weird at the L1 level. The page pointed to by an L1 entry will be a data
-page. We don't actually need to perform P2V on data pages in general. But also problematic is that this invariant implies
-partial ownership of the data (by taking the $\Xi$ tokens for the data page back to this invariant). So we need a way
-to make this ``bottom out'' after the L1 entries. Something like switching $\Xi$ to be a map from a physical address to the corresponding level of tables
-(as a natural number; we'll always start out knowing that $\ghostmaptoken{\textsf{id}}{\textsf{cr3val}}{4}$ for a traversal),
-and then switching the indexing into the next level to be
-\[
-\bigast{(\paddr,\mathsf{level})}{\ldots\left(\ulcorner\textsf{entry\_present}(\vale)\land \mathsf{level} > 1 \urcorner \wand \forall_{\textsf{i}\in\textsf{0..511}} \ldotp \ghostmaptoken{\textsf{id}}{((\mathsf{page\_addr}\;\vale) + \textsf{i * 8})}{(\mathsf{level}-1)}\right)}\]
-I'm surprised this hasn't come up during page table setup
-}
+and for valid entries (with the present bit set) above L1, ghost map tokens for every entry in the table pointed to by the entry.
+(L1 entries point to data pages, whose physical memory ownership resides in some virtual points-to).
 
@@
 
-Thus the fractional ownership of the physical location is enough for Line \ref{line:read_entry_contents} to access the entry, though in \lstinline|get_next_table|
-the caller has pulled that piece of information out of the invariant and passed it for the entry at hand, which is why this appears explicitly in assertions,
-and the argument to the invariant is $\Xi\setminus\{\mathsf{entry}\}$\todo{update pending entry+KB} (indexing by the set $\Xi$ allows us to borrow the physical resources
+Thus the fractional ownership of the physical location is enough for Line \ref{line:read_entry_contents} in Figure \ref{fig:calltopteinitialize} to access the entry, though in \lstinline|get_next_table|
+the caller has pulled that piece of information out of the invariant and passed it for the entry at hand.
+This removal appears explicitly in assertions,
+as the argument to the invariant is $\Xi\setminus\{\mathsf{entry}\}$ (indexing by the set $\Xi$ allows us to borrow the physical resources
 for a specific page table entry out of the invariant, and later put them back).
@@
 and permits storing a new entry (in ellided code around Line \ref{line:install_new_entry})
+\looseness=-1
 
@@
 \textsf{pte\_walkpgdir}, as a client, holds the knowledge that there exists an identity mapping for the physical entry address (\textsf{entry}) in the root page table ($\textsf{L}_{4}$):  $\mathsf{entry} \mapsto_{\textsf{id}} \textsf{\_}$ in Specification Line 3 is a partially owned token for accessing and looking up the resources in the identity map, $\Xi$, to construct the \textit{virtual-to-physical} pointsto relation $\textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry \entry\_val}$ with the virtual address (\textsf{entry+KERNBASE}) obtained by offsetting the physical address (\textsf{entry}).
-With this knowledge on the root-page-table-entry, we can start traversing the page-table tree which requires locating the address of the next table -- a call to \textsf{pte\_get\_next\_table} shown in Figure \ref{fig:calltoptepteinitialize}.
+With this knowledge on the root-page-table-entry, we can start traversing the page-table tree which requires locating the address of the next table -- a call to \textsf{pte\_get\_next\_table} shown in Figure \ref{fig:calltopteinitialize}.
 Beyond a frame, the precondition before Line 15 requires the current address space invariant, and knowledge that \textsf{entry} is mapped to a random entry value, subtly, the operation also, at least, requires that the relevant table entry is readable, but the exact portion of ownership returned must be determined by inspecting the valid bit
@@
   \begin{lstlisting}[mathescape,escapeinside={(*}{*)}]
-
 $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)   \ast \texttt{rbp-16} \mapsto_{\textsf{r}} \textsf{\_} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry+KERNBASE}  }_{\rtv}$
@@
 ;;void pte_initialize(pte_t *entry) {
-...
-;;set up the stack    
+... ;;set up the stack    
 ;;allocate a full zeroed page for 512 8-byte entries 
-;;pte_t *local = kalloc();
-callq  81 <pte_initialize+0x11>	7d: R_X86_64_PLT32	kalloc-0x4 (*\label{line:call_to_kalloc}*)
-mov    rax,-0x10[rbp] 
+callq  81 <kalloc>	;;pte_t *local = kalloc();(*\label{line:call_to_kalloc}*)
+mov    rax,-0x10[rbp] ;; Store into 'local'
 ;;entry->pfn = PTE_ADDR_TO_PFN((uintptr_t) local);
@@
 $\specline{ \textsf{entry+KERNBASE} \mapsto_{\mathsf{pte,1}} \mathsf{entry} \; \textsf{pfn\_set(entryv nextpaddr)}}$
-$\specline{\ghostmaptoken{\delta{}s}{\rtv}{\delta} \ast \begin{array}{l} \textsf{entry\_present(pte\_initialize(pfn\_set(entryv nextpaddr)))} -* \\  \;\;\;\;\;\;\; \forall_{i\in\textsf{0..511}} \; \textsf{table\_root(pte\_initialized (pfn\_set (entryv nextpaddr))) + i * 8} \mapsto_{\textsf{id}} \textsf{level-1}    \end{array}  }_{\rtv}$
-;;clean up the stack
-...
+$\specline{\ghostmaptoken{\delta{}s}{\rtv}{\delta} \ast \begin{array}{l} \ulcorner\textsf{present\_L(pte\_initialize(pfn\_set(entryv nextpaddr)),level)}\urcorner \wand \\  \;\;\;\;\;\;\; \forall_{i\in\textsf{0..511}} \; \textsf{table\_root(pte\_initialized (pfn\_set (entryv nextpaddr))) + i * 8} \mapsto_{\textsf{id}} \textsf{level-1}    \end{array}  }_{\rtv}$
+... ;;clean up the stack, return
 \end{lstlisting}
@@
 \end{figure}
-\todo[inline,color=blue]{Colin revision frontier}
-Since we are using \textsf{pte\_initialize} for page-table address allocation, we must relate this newly allocated physical address to the identity mapping map $\Xi$ -- Specification Lines 34 and 38 in Figure \ref{fig:calltopteinitialize}.
-Focusing on the specification of \textsf{pte\_initialize} separately in Figure \ref{fig:pteinitializespec}, we right immediately realize that instead of seeing see a physical pointsto for the fresly page-table address (e.g.
-$\mathsf{nextpaddr} \mapsto_{\mathsf{p}} \mathsf{w64\_0}$) deliberately in the post-conditoin in Lines 15-16, we observe a full-ownership token representing the knowledge that a frame and all the entries indexed from this frame are freshly allocated with full-ownership to be a part of the identity map, $\Xi$.
-The soundness argument of this specification relies on the fact that these freshly allocated resources is part of an entry construction that has not been completed yet: the presence bit is set (Line 42 in Figure \ref{fig:calltopteinitialize}) after these freshly allocated resources are incorporated to the entry construction (Line 41 in Figure \ref{fig:calltopteinitialize}).
-In other words, the side condition, (*), acknowledges any access to the entry with these resources as \textit{invalid}, as the entry is revealed to shared-accesses when the presence bit is set.
+
+Since we are using \textsf{pte\_initialize} for page-table address allocation, we must relate this newly
+allocated physical address to the identity mapping map $\Xi$ --- 
+see Line \ref{line:page_of_caps} in Figure \ref{fig:calltopteinitialize}, where
+\texttt{kalloc}'s specification guarantees it has returned memory from a designated memory
+pool that is already mapped\footnote{A reasonable reader might wonder where this pool
+initially comes from, and how it might grow when needed. Typically an initial mapping subject to this identity mapping
+constraint is set up prior to transition to 64-bit kernel code (notably,
+a page table must exist \emph{before} virtual memory is enabled during boot, as part of enabling it is setting
+a page table root).
+Growing this pool later requires cooperation of physical memory range allocation and virtual memory range allocation,
+typically by starting general virtual address allocation at the highest physical memory address plus the identity mapping offset.
+This reserves the virtual addresses corresponding to all physical addresses plus the offset for later use in this pool,
+as needed.
+} and satisfies the offset invariants.
+\todo[inline,color=blue]{colin frontier.
+Stuck with line 31 onwards in Figure 7. rax holds nextpaddr, but I think that should be entrypfn, and 
+the explicit entrypfn id token assertion should go away, as its covered by the forall assertion.
+then the postcondition for pte-initialize should have a specific level now for the entries,
+like 0, which can be updated in the view shift on line 42.
+}
+% Focusing on the specification of \textsf{pte\_initialize} separately in Figure \ref{fig:pteinitializespec}, 
+% we right immediately realize that instead of seeing see a physical pointsto for the fresly page-table address 
+% (e.g. $\mathsf{nextpaddr} \mapsto_{\mathsf{p}} \mathsf{w64\_0}$) deliberately in the post-conditoin in Lines 15-16,
+%  we observe a full-ownership token representing the knowledge that a frame and all the entries indexed from this 
+% frame are freshly allocated with full-ownership to be a part of the identity map, $\Xi$. 
+The soundness argument of this specification relies on the fact that these freshly allocated resources is part 
+of an entry construction that has not been completed yet: the presence bit is set 
+(Line 42 in Figure \ref{fig:calltopteinitialize}) after these freshly allocated resources are incorporated to the 
+entry construction (Line 41 in Figure \ref{fig:calltopteinitialize}). In other words, the side condition, (*),
+ acknowledges any access to the entry with these resources as \textit{invalid}, as the entry is revealed to 
+shared-accesses when the presence bit is set.
 
@@
 This actually constitutes a very critical piece of the full page table walk verification, and we have verified the critical step for a small x86-64 kernel, which is the physical-to-virtual conversion (often appearing as a macro \texttt{P2V} inside the macro \textsf{ PTE\_PFN\_TO\_ADDR} in C source code).
-In our small kernel (Line 57 in Figure \ref{calltopteinitialize}), as in larger kernels, \texttt{P2V} is actually just addition, but the correctness of this simple instruction is quite subtle and relies on identity mappings explained in the previous section (Section \ref{subsec:identitymappings}) as a part of understanding our address-space invariant in Figure \ref{fig:peraspaceinvariant_with_p2v_extension}.
+In our small kernel (Line 57 in Figure \ref{fig:calltopteinitialize}), as in larger kernels, \texttt{P2V} is actually just addition, but the correctness of this simple instruction is quite subtle and relies on identity mappings explained in the previous section (Section \ref{subsec:identitymappings}) as a part of understanding our address-space invariant in Figure \ref{fig:peraspaceinvariant_with_p2v_extension}.
 \begin{figure}\footnotesize
@@
 As shown in Figure \ref{fig:p2v}, \lstinline|rcx| holds the offset value \textsf{KERNBASE} which gives us the value of the virtual address ($\textsf{entry}_{\textsf{pfn}}$ \textsf{+KERNBASE}) in the identity map ($\Xi\setminus\{entry\}\cup \{\textsf{entry}_{\textsf{pfn}}) \}$) of the kernel invariant: the logical update in Specification  Lines 5-10 to 10-14 for obtaining virtual-pointsto resource for the frame ($\textsf{entry}_{\textsf{pfn}}$) by removing it from the ghost map ($\Xi\setminus\{entry\}\cup \{\textsf{entry}_{\textsf{pfn}}) \}$) in Line 5 and compute the identity mapping for this physical frame address in Line 13 in Figure \ref{fig:p2v}).
+
 \subsubsection{Walking Page-Table Tree: Calling \textsf{pte\_get\_nexttable} for Each Level}
 \label{wlkpgdir}
+
+\todo[inline,color=blue]{TAKEN FROM get-next-table DISCUSSION:
+In the very first precondition on the client side -- e.g.\textsf{pte\_walkpgdir} in Figure \ref{walkpgdir} -- 
+we see that \textsf{vpte-ptsto} for the entry being accessed (\textsf{entry}) is constructed -- in Specification Line \ref{line:get_next_vpte_precondition} of Figure \ref{fig:calltopteinitialize} for \textsf{entry}.
+Intuitively speaking, we start traversal from a certain entry address at the current level for which we know ($\textsf{entry} \mapsto_{\textsf{id}} \textsf{\_}$ in Line 3 in Figure \ref{walkpgdir}) that there is a memory mapping ($\textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val}$  Line 15 in Figure \ref{walkpgdir}) that can be used to explore deeper in the page-tree.
+
+The most crucial aspect of page-table entry access relies on the subtle interplay between updating the shared page tables, and the not violating the memory mappings. 
+Concretely speaking, accessing to the entry -- whose address is kept in the register $\texttt{r8}$ -- in Line 15 in Figure \ref{fig:calltopteinitialize} requires a 
+memory mapping that is logically provided through the update from Specfication Lines 3 to 15 in Figure \ref{walkpgdir} on the client side by the removal of \textsf{entry}
+from the map $\Xi$ in the invariant to construct the \textsf{vpte-ptsto} (Specfication Line 15 in Figure \ref{walkpgdir}). 
+This magical abstraction -- represented as $\Xi$ in the invariant definition -- provides us to construct this \textsf{pte-ptsto} out of an physical address is called \textit{identity-mappings}.
+}
+
 Realizing page-table directory walk amounts to calling \textsf{pte\_get\_next\_table} for each level as shown in Figure \ref{walkpgdir}.
@@
 ;;complstatus_t vaspace_mappage(pte_t *pml4, void *va,uintptr_t fpaddr, uint64_t perm ) {
-...
-;;setting up the stack      
-;;char *a; pte_t *pte;
-;;a = (char *)PGROUNDUP((uintptr_t)va);
+... ;;setting up the stack      
 ;;char *a; pte_t *pte;
@@
 \subsection{Unmapping a Page}
-\todo[inline]{this must move, must come after mapping}
+\todo[inline]{update (esp. line refs) for new mapping code}
 The reverse operation, unmapping a designated page that is currently mapped,
@@
 full virtual PTE pointer for it, setting to 0, and reclaiming the now-unmapped physical memory.
+
 \subsection{Change of Address Space}

### 2024-07-26 examples.tex
fix C version of call to pte_initialize for our tweaks
redundant rax in fig9

@@
 $\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \land \lnot(\textsf{entry\_present entry\_val})  \urcorner }_{\rtv}$ (*\label{line:after_concluding_qfrac1}*)
-callq  70 <pte_initialize> ;;pte_initialize(&next);(* \label{line:call_to_pte_initialize} *)
+callq  70 <pte_initialize> ;;pte_initialize(entry);(* \label{line:call_to_pte_initialize} *)
 $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{\_}  }_{\rtv}$
@@
 mov    rax,[rdi]
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)   \ast \texttt{rbp-16} \mapsto_{\textsf{r}} \textsf{\_} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry+KERNBASE} \ast \texttt{rax} \mapsto_{\textsf{v}} \textsf{nextpaddr}  }_{\rtv}$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)   \ast \texttt{rbp-16} \mapsto_{\textsf{r}} \textsf{\_} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry+KERNBASE} }_{\rtv}$
 $\specline{\textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,}} \textsf{entry entryv} \ast \ghostmaptoken{\textsf{id}}{(\mathsf{entry})}{\textsf{level}} }$
-$\specline{\texttt{rax} \mapsto_{\textsf{v}} \textsf{nextpaddr} \ast \lnot(\textsf{entry\_present entry\_val}) }$
+$\specline{\texttt{rax} \mapsto_{\textsf{v}} \textsf{nextpaddr} \ast \lnot(\textsf{entry\_present entry\_v}) }$
 $\specline{ \textsf{entry+KERNBASE} \mapsto_{\mathsf{pte,1}} \mathsf{entry} \; \textsf{pfn\_set(entryv nextpaddr)}}$

### 2024-07-26 examples.tex
+

@@
 callq  70 <pte_initialize> ;;pte_initialize(entry);(* \label{line:call_to_pte_initialize} *)
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{\_}  }_{\rtv}$
-$\specline{ \textsf{entry} \mapsto_{\textsf{id}} \textsf{\_} \ast \textsf{entry}_{\textsf{pfn}} \mapsto_{\textsf{id}} \textsf{\_} \ast \ulcorner  \textsf{qfrac} = 1 \land \lnot(\textsf{entry\_present entry\_val}_{\textsf{pfn}})  \urcorner  \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
-$\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \mathsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \land \lnot(\textsf{entry\_present entry\_val})  \urcorner }_{\rtv}$
-$\specline{ \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{entry}_{\textsf{pfn}}  \ast \texttt{rax} \mapsto_{\textsf{r}} \textsf{nextpaddr}  }_{\rtv}$
-$\specline{ \forall_{i\in \textsf{0 ...
-511} } \ldotp  \ghostmaptoken{\textsf{id}}{((\mathsf{entry\_page}\;\textsf{entry}) + \textsf{i * 8})}{\textsf{\_}} \ast  \textsf{qfrac} = 1 \land \lnot(\textsf{entry\_present entry\_val }_{\textsf{pfn}} + \textsf{i * 8})  }$ (*\label{line:page_of_caps}*)
-... ;;entry value updates: entry->pfn = next; entry->present = 1; (*\label{line:install_new_entry}*)
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{\_}  }_{\rtv}$
-$\specline{ \textsf{entry} \mapsto_{\textsf{id}} \textsf{\_} \ast \textsf{entry}_{\textsf{pfn}} \mapsto_{\textsf{id}} \textsf{\_} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \lnot(\textsf{entry\_present entry\_val}_{\textsf{pfn}})  \urcorner  \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
-$\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \mathsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \lnot(\textsf{entry\_present entry\_val})  \urcorner }_{\rtv}$
-$\specline{ \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{entry}_{\textsf{pfn}}  \ast \texttt{rax} \mapsto_{\textsf{r}} \textsf{nextpaddr}  }_{\rtv}$
-$\specline{ \forall_{i\in \textsf{0 ...
-511} } \ldotp \ghostmaptoken{\textsf{id}}{((\mathsf{entry\_page}\;\textsf{entry}) + \textsf{i * 8})}{\textsf{\_}} \ast  \textsf{qfrac} = 1 \leftrightarrow \lnot(\textsf{entry\_present entry\_val }_{\textsf{pfn}} + \textsf{i * 8})  }$
-$\sqsubseteq$
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \} \cup \{\textsf{entry}_{\textsf{pfn}}) \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{\_} }_{\rtv}$
-$\specline{ \textsf{entry} \mapsto_{\textsf{id}} \textsf{\_} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
-$\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \mathsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \lnot(\textsf{entry\_present entry\_val})  \urcorner }_{\rtv}$
-$\specline{ \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{entry}_{\textsf{pfn}}  \ast \texttt{rax} \mapsto_{\textsf{r}} \textsf{entry}_{\textsf{pfn}}  }_{\rtv}$ (*\label{line:alloc_path_end}*)
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  }_{\rtv}$
+$\specline{  \textsf{entry} \mapsto_{\textsf{id}} \textsf{v} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta} }_{\rtv}$
+$\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \land \lnot(\textsf{entry\_present entry\_val})  \urcorner }_{\rtv}$
+$\specline{\ulcorner  \textsf{qfrac} = 1 \land \lnot(\textsf{entry\_present (pfn\_set (entry\_val nextpaddr))}  \urcorner   }_{\rtv}$
+$\specline{ \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{pfn\_set(entry\_val nextpaddr)}   }_{\rtv}$
+$\specline{ \begin{array}{l}\textsf{entry\_present(pfn\_set(entryv nextpaddr))} -*  \\ \;\;\;\;\;\;\;\;\  \forall_{i\in \textsf{0 ...
+511} } \ldotp  \ghostmaptoken{\textsf{id}}{((\textsf{table\_root entry\_present(pfn\_set(entry\_val nextpaddr))}) + \textsf{i * 8})}{\textsf{v-1}} \end{array}  }$ (*\label{line:page_of_caps}*)
+... ;;entry value updates: entry->pfn = nextpaddr; entry->present = 1; (*\label{line:install_new_entry}*)
+... ;;now we know that entry is initialized, so we satisfy the condition to access children list
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  }_{\rtv}$
+$\specline{  \textsf{entry} \mapsto_{\textsf{id}} \textsf{v} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta} }_{\rtv}$
+$\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \land \lnot(\textsf{entry\_present entry\_val})  \urcorner }_{\rtv}$
+$\specline{ \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{pfn\_set(entry\_val nextpaddr)}   }_{\rtv}$
+$\specline{\forall_{i\in \textsf{0 ...
+511} } \ldotp  \ghostmaptoken{\textsf{id}}{((\textsf{table\_root (pfn\_set(entry\_val nextpaddr))}) + \textsf{i * 8})}{\textsf{v-1}}  }$ (*\label{line:alloc_path_end}*)
 ;;}
@@
 $\specline{\texttt{rax} \mapsto_{\textsf{v}} \textsf{nextpaddr} \ast \lnot(\textsf{entry\_present entry\_v}) }$
-$\specline{ \textsf{entry+KERNBASE} \mapsto_{\mathsf{pte,1}} \mathsf{entry} \; \textsf{pfn\_set(entryv nextpaddr)}}$
+$\specline{ \textsf{entry+KERNBASE} \mapsto_{\mathsf{vpte,1}} \mathsf{entry} \; \textsf{pfn\_set(entryv nextpaddr)}}$
 $\specline{\ghostmaptoken{\delta{}s}{\rtv}{\delta} \ast \begin{array}{l} \ulcorner\textsf{present\_L(pte\_initialize(pfn\_set(entryv nextpaddr)),level)}\urcorner \wand \\  \;\;\;\;\;\;\; \forall_{i\in\textsf{0..511}} \; \textsf{table\_root(pte\_initialized (pfn\_set (entryv nextpaddr))) + i * 8} \mapsto_{\textsf{id}} \textsf{level-1}    \end{array}  }_{\rtv}$

### 2024-07-26 examples.tex
triggered fixes

@@
 mov    -0x8[rbp],rdi
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{entry\},m)  \ast \textsf{entry}\mapsto_{\textsf{pte,q}}\textsf{entryv} \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{\_}  \ast  \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{next} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry}  }_{\rtv}$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{entry\},m)  \ast \textsf{entry}\mapsto_{\textsf{pte,q}}\textsf{entry\_val} \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{\_}  \ast  \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{next} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry}  }_{\rtv}$
 $\specline{ \textsf{entry} \mapsto_{\textsf{id}} \textsf{v} \ast\urcorner \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
@@
 $\specline{ \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{pfn\_set(entry\_val nextpaddr)}   }_{\rtv}$
-$\specline{ \begin{array}{l}\textsf{entry\_present(pfn\_set(entryv nextpaddr))} -*  \\ \;\;\;\;\;\;\;\;\  \forall_{i\in \textsf{0 ...
-511} } \ldotp  \ghostmaptoken{\textsf{id}}{((\textsf{table\_root entry\_present(pfn\_set(entry\_val nextpaddr))}) + \textsf{i * 8})}{\textsf{v-1}} \end{array}  }$ (*\label{line:page_of_caps}*)
+$\specline{ \begin{array}{l}\textsf{entry\_present (pte\_initialized (pfn\_set(entryv nextpaddr)))} \wand  \\ \;\;\;\;\;\;\;\;\  \forall_{i\in \textsf{0 ...
+511} } \ldotp  \ghostmaptoken{\textsf{id}}{((\textsf{table\_root (pte\initialized (pfn\_set(entry\_val nextpaddr)))}) + \textsf{i * 8})}{\textsf{v-1}} \end{array}  }$ (*\label{line:page_of_caps}*)
 ... ;;entry value updates: entry->pfn = nextpaddr; entry->present = 1; (*\label{line:install_new_entry}*)
@@
 $\specline{  \textsf{entry} \mapsto_{\textsf{id}} \textsf{v} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta} }_{\rtv}$
-$\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \land \lnot(\textsf{entry\_present entry\_val})  \urcorner }_{\rtv}$
-$\specline{ \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{pfn\_set(entry\_val nextpaddr)}   }_{\rtv}$
+$\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}}  \textsf{pte\_initialized(pfn\_set(entryv nextpaddr))} \ast \ulcorner  \textsf{qfrac} = 1 \land \lnot(\textsf{entry\_present entry\_val})  \urcorner }_{\rtv}$
+$\specline{ \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{pte\_initialized(pfn\_set(entry\_val nextpaddr))}    }_{\rtv}$
 $\specline{\forall_{i\in \textsf{0 ...
-511} } \ldotp  \ghostmaptoken{\textsf{id}}{((\textsf{table\_root (pfn\_set(entry\_val nextpaddr))}) + \textsf{i * 8})}{\textsf{v-1}}  }$ (*\label{line:alloc_path_end}*)
+511} } \ldotp  \ghostmaptoken{\textsf{id}}{((\textsf{table\_root (pte\_initialized (pfn\_set(entry\_val nextpaddr)))}) + \textsf{i * 8})}{\textsf{v-1}}  }$ (*\label{line:alloc_path_end}*)
 ;;}
@@
 $\specline{\textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,}} \textsf{entry entryv} \ast \ghostmaptoken{\textsf{id}}{(\mathsf{entry})}{\textsf{level}} }$
-$\specline{\texttt{rax} \mapsto_{\textsf{v}} \textsf{nextpaddr} \ast \lnot(\textsf{entry\_present entry\_v}) }$
+$\specline{\texttt{rax} \mapsto_{\textsf{v}} \textsf{nextpaddr} \ast \lnot(\textsf{entry\_present entry\_val}) }$
 $\specline{ \textsf{entry+KERNBASE} \mapsto_{\mathsf{vpte,1}} \mathsf{entry} \; \textsf{pfn\_set(entryv nextpaddr)}}$
-$\specline{\ghostmaptoken{\delta{}s}{\rtv}{\delta} \ast \begin{array}{l} \ulcorner\textsf{present\_L(pte\_initialize(pfn\_set(entryv nextpaddr)),level)}\urcorner \wand \\  \;\;\;\;\;\;\; \forall_{i\in\textsf{0..511}} \; \textsf{table\_root(pte\_initialized (pfn\_set (entryv nextpaddr))) + i * 8} \mapsto_{\textsf{id}} \textsf{level-1}    \end{array}  }_{\rtv}$
+$\specline{\ghostmaptoken{\delta{}s}{\rtv}{\delta} \ast \begin{array}{l} \ulcorner\textsf{entry\_present (pte\_initialize(pfn\_set(entry\_val nextpaddr)),level)}\urcorner \wand \\  \;\;\;\;\;\;\; \forall_{i\in\textsf{0..511}} \; \textsf{table\_root(pte\_initialized (pfn\_set (entry\_val nextpaddr))) + i * 8} \mapsto_{\textsf{id}} \textsf{level-1}    \end{array}  }_{\rtv}$
 ... ;;clean up the stack, return
@@
  add    rcx,rax
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \} \cup \{\textsf{entry}_{\textsf{pfn}}) \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{\_}  }_{\rtv}$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \}),m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  }_{\rtv}$
 $\specline{ \textsf{entry} \mapsto_{\textsf{id}} \textsf{\_} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
-$\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val}  \urcorner }_{\rtv}$
-$\specline{ \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{entry}_{\textsf{pfn}}  \ast \texttt{rax} \mapsto_{\textsf{r}} \textsf{entry}_{\textsf{pfn}} \ast \textsf{entry}_{\textsf{pfn}} \mapsto_{\textsf{id}} \textsf{\_}  }_{\rtv}$
-$\subseteq$
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{\_}  }_{\rtv}$
-$\specline{ \textsf{entry} \mapsto_{\textsf{id}} \textsf{\_} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
-$\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val}  }_{\rtv}$
-$\specline{  \textsf{entry}_{\textsf{pfn}} \textsf{+KERNBASE}  \mapsto_{\textsf{vpte,qfrac}}   \textsf{entry}_{\textsf{pfn}} \; \textsf{entry\_pfn\_val} \ast R_{\textsf{children}}}_{\rtv}$
-$\specline{ \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{entry}_{\textsf{pfn}}  \ast \texttt{rax} \mapsto_{\textsf{r}} \textsf{entry}_{\textsf{pfn}} \ast \textsf{entry}_{\textsf{pfn}} \mapsto_{\textsf{id}} \textsf{\_}}_{\rtv}$
+$\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{pte\_initialize(pfn\_set(entry\_val nextpaddr))} \urcorner }_{\rtv}$
+$\specline{ \texttt{rbp-16} \mapsto_{\textsf{v}}  \textsf{pte\_initialize(pfn\_set(entry\_val nextpaddr))}  \ast \texttt{rax} \mapsto_{\textsf{r}} \textsf{ table\_root (pte\_initialize(pfn\_set(entry\_val nextpaddr))} \ast \textsf{entry}_{\textsf{pfn}} \mapsto_{\textsf{id}} \textsf{\_}  }_{\rtv}$
 ...

### 2024-07-26 examples.tex
+

@@
 $\specline{ \begin{array}{l}\textsf{entry\_present (pte\_initialized (pfn\_set(entryv nextpaddr)))} \wand  \\ \;\;\;\;\;\;\;\;\  \forall_{i\in \textsf{0 ...
-511} } \ldotp  \ghostmaptoken{\textsf{id}}{((\textsf{table\_root (pte\initialized (pfn\_set(entry\_val nextpaddr)))}) + \textsf{i * 8})}{\textsf{v-1}} \end{array}  }$ (*\label{line:page_of_caps}*)
+511} } \ldotp  \ghostmaptoken{\textsf{id}}{((\textsf{table\_root (pte\_initialized (pfn\_set(entry\_val nextpaddr)))}) + \textsf{i * 8})}{\textsf{v-1}} \end{array}  }$ (*\label{line:page_of_caps}*)
 ... ;;entry value updates: entry->pfn = nextpaddr; entry->present = 1; (*\label{line:install_new_entry}*)
@@
 $\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}}  \textsf{pte\_initialized(pfn\_set(entryv nextpaddr))} \ast \ulcorner  \textsf{qfrac} = 1 \land \lnot(\textsf{entry\_present entry\_val})  \urcorner }_{\rtv}$
-$\specline{ \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{pte\_initialized(pfn\_set(entry\_val nextpaddr))}    }_{\rtv}$
+$\specline{ \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{pte\_initialized(pfn\_set(entry\_val nextpaddr))}  \ast rax \mapsto_{r} \textsf{table\_root (pte\_initialized (pfn\_set(entry\_val nextpaddr)))}  }_{\rtv}$
 $\specline{\forall_{i\in \textsf{0 ...
@@
 $\specline{\textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,}} \textsf{entry entryv} \ast \ghostmaptoken{\textsf{id}}{(\mathsf{entry})}{\textsf{level}} }$
-$\specline{\texttt{rax} \mapsto_{\textsf{v}} \textsf{nextpaddr} \ast \lnot(\textsf{entry\_present entry\_val}) }$
+$\specline{\texttt{rax} \mapsto_{\textsf{r}} \textsf{nextpaddr} \ast \lnot(\textsf{entry\_present entry\_val}) }$
 $\specline{ \textsf{entry+KERNBASE} \mapsto_{\mathsf{vpte,1}} \mathsf{entry} \; \textsf{pfn\_set(entryv nextpaddr)}}$
-$\specline{\ghostmaptoken{\delta{}s}{\rtv}{\delta} \ast \begin{array}{l} \ulcorner\textsf{entry\_present (pte\_initialize(pfn\_set(entry\_val nextpaddr)),level)}\urcorner \wand \\  \;\;\;\;\;\;\; \forall_{i\in\textsf{0..511}} \; \textsf{table\_root(pte\_initialized (pfn\_set (entry\_val nextpaddr))) + i * 8} \mapsto_{\textsf{id}} \textsf{level-1}    \end{array}  }_{\rtv}$
+$\specline{\ghostmaptoken{\delta{}s}{\rtv}{\delta} \ast \begin{array}{l} \ulcorner\textsf{entry\_present (pte\_initialize(pfn\_set(entry\_val nextpaddr)),level)}\urcorner \wand \\  \;\;\;\;\;\;\; \forall_{i\in\textsf{0..511}} \; \textsf{table\_root (pte\_initialized (pfn\_set (entry\_val nextpaddr))) + i * 8} \mapsto_{\textsf{id}} \textsf{level-1}    \end{array}  }_{\rtv}$
 ... ;;clean up the stack, return
@@
 \begin{lstlisting}[mathescape,escapeinside={(*}{*)}]
-... ;; Continued from Figure (*\ref{fig:calltopteinitialize}*)
-;;uintptr_t next_virt_addr = (uintptr_t) P2V(next<<12);
+... ;; Continued from Figure (*\ref{fig:calltopteinitialize}*) either by following the pte allocation path or not
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \}),m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  }_{\rtv}$
+$\specline{ \textsf{entry} \mapsto_{\textsf{id}} \textsf{\_} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
+$\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{(pte\_initialized (entry\_val.pfn))} \urcorner }_{\rtv}$
+$\specline{ \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{(pte\_initialized (entry\_val.pfn)))} \ast \texttt{rax} \mapsto_{\textsf{r}} \textsf{ table\_root (pte\_initialize(entry\_val.pfn))} }_{\rtv}$
+;;uintptr_t next_virt_addr = (uintptr_t) P2V(entry.pfn<<12);  
  movabs KERNBASE,rcx
@@
 $\specline{ \textsf{entry} \mapsto_{\textsf{id}} \textsf{\_} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
-$\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{pte\_initialize(pfn\_set(entry\_val nextpaddr))} \urcorner }_{\rtv}$
-$\specline{ \texttt{rbp-16} \mapsto_{\textsf{v}}  \textsf{pte\_initialize(pfn\_set(entry\_val nextpaddr))}  \ast \texttt{rax} \mapsto_{\textsf{r}} \textsf{ table\_root (pte\_initialize(pfn\_set(entry\_val nextpaddr))} \ast \textsf{entry}_{\textsf{pfn}} \mapsto_{\textsf{id}} \textsf{\_}  }_{\rtv}$
+$\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{(pte\_initialized (entry\_val.pfn))} \urcorner }_{\rtv}$
+$\specline{ \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{(pte\_initialized (entry\_val.pfn)))} \ast \texttt{rax} \mapsto_{\textsf{r}} \textsf{ table\_root (pte\_initialize(entry\_val.pfn)) + KERNBASE} }_{\rtv}$
 ...

### 2024-07-27 examples.tex
+

@@
 ... ;; Continued from Figure (*\ref{fig:calltopteinitialize}*) either by following the pte allocation path or not
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \}),m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  }_{\rtv}$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \}),m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{rcx}  \mapsto_{\textsf{r}} \textsf{\_}  }_{\rtv}$
 $\specline{ \textsf{entry} \mapsto_{\textsf{id}} \textsf{\_} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
@@
  add    rcx,rax
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \}),m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  }_{\rtv}$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \}),m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{rcx}  \mapsto_{\textsf{r}} \textsf{KERNBASE} }_{\rtv}$
 $\specline{ \textsf{entry} \mapsto_{\textsf{id}} \textsf{\_} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$

### 2024-08-05 examples.tex
cleaning up some unfinished polishing

@@
 511} } \ldotp  \ghostmaptoken{\textsf{id}}{((\textsf{table\_root (pte\_initialized (pfn\_set(entry\_val nextpaddr)))}) + \textsf{i * 8})}{\textsf{v-1}}  }$ (*\label{line:alloc_path_end}*)
-;;}
+;;} (*\label{line:end_of_allocation_path}*)
 ... ;; Code after conditional continued in Figure (*\ref{fig:p2v}*)
@@
 if the entry is invalid the invariant holds full ownership of the entry, so it can be updated; while if the entry is valid,
-the invariant owns only a constant non-zero fragment sufficient to read the entry, but not modify it (which would invalidate some virtual points-to assertions).
-
+the invariant owns only a constant non-zero fragment sufficient to read the entry, but not modify it (which would invalidate some virtual points-to assertions):
+\begin{equation*}
+ \ulcorner \textsf{qfrac} = 1 \leftrightarrow \; \lnot\textsf{entry\_present }(\vale) \urcorner \tag{*}
+\end{equation*}
 Thus the fractional ownership of the physical location is enough for Line \ref{line:read_entry_contents} in Figure \ref{fig:calltopteinitialize} to access the entry, though in \lstinline|get_next_table|
@@
 
-\todo[inline,color=blue]{Colin revision frontier: need to resolve entry+KERNBASE issue to finish pulling this into prose above}
-
-\textsf{pte\_walkpgdir}, as a client, holds the knowledge that there exists an identity mapping for the physical entry address (\textsf{entry}) in the root page table ($\textsf{L}_{4}$):  $\mathsf{entry} \mapsto_{\textsf{id}} \textsf{\_}$ in Specification Line 3 is a partially owned token for accessing and looking up the resources in the identity map, $\Xi$, to construct the \textit{virtual-to-physical} pointsto relation $\textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry \entry\_val}$ with the virtual address (\textsf{entry+KERNBASE}) obtained by offsetting the physical address (\textsf{entry}).
-With this knowledge on the root-page-table-entry, we can start traversing the page-table tree which requires locating the address of the next table -- a call to \textsf{pte\_get\_next\_table} shown in Figure \ref{fig:calltopteinitialize}.
-Beyond a frame, the precondition before Line 15 requires the current address space invariant, and knowledge that \textsf{entry} is mapped to a random entry value, subtly, the operation also, at least, requires that the relevant table entry is readable, but the exact portion of ownership returned must be determined by inspecting the valid bit
-of the value in memory --- so full ownership is returned only for unused entries.
-This is a simple piece of code whose functionality is critical and whose correctness is highly non-trivial. No prior work engages with this problem.
-
-Concretely speaking, going back to Line 15 in Figure \ref{fig:calltopteinitialize}, to read the value referenced by physical address \textsf{entry} while preserving the soundness of memory mappings, our extended invariant introduces the side condition (*)
-\begin{equation*}
- \ulcorner \textsf{qfrac} = 1 \leftrightarrow \; \lnot\textsf{entry\_present }(\vale) \urcorner \tag{*}
-\end{equation*}
-assuring that looking the identity mapping for \textsf{entry} is safe under the subtle justification which equates the full ownership to the non/presence of the entry which can only be known when investigated in Line 21 in Figure \ref{fig:calltopteinitialize}.
+% Concretely speaking, going back to Line 15 in Figure \ref{fig:calltopteinitialize}, to read the value referenced by physical address \textsf{entry} while preserving the soundness of memory mappings, our extended invariant introduces the side condition (*)
+% \begin{equation*}
+%  \ulcorner \textsf{qfrac} = 1 \leftrightarrow \; \lnot\textsf{entry\_present }(\vale) \urcorner \tag{*}
+% \end{equation*}
+% assuring that looking the identity mapping for \textsf{entry} is safe under the subtle justification which equates the full ownership to the non/presence of the entry which can only be known when investigated in Line 21 in Figure \ref{fig:calltopteinitialize}.
 \begin{comment}
@@
 } and satisfies the offset invariants.
-\todo[inline,color=blue]{colin frontier.
-Stuck with line 31 onwards in Figure 7. rax holds nextpaddr, but I think that should be entrypfn, and 
-the explicit entrypfn id token assertion should go away, as its covered by the forall assertion.
-then the postcondition for pte-initialize should have a specific level now for the entries,
-like 0, which can be updated in the view shift on line 42.
-}
+% \todo[inline,color=blue]{colin frontier.
+% Stuck with line 31 onwards in Figure 7. rax holds nextpaddr, but I think that should be entrypfn, and 
+% the explicit entrypfn id token assertion should go away, as its covered by the forall assertion.
+% then the postcondition for pte-initialize should have a specific level now for the entries,
+% like 0, which can be updated in the view shift on line 42.
+% }
 % Focusing on the specification of \textsf{pte\_initialize} separately in Figure \ref{fig:pteinitializespec}, 
@@
 of an entry construction that has not been completed yet: the presence bit is set 
-(Line 42 in Figure \ref{fig:calltopteinitialize}) after these freshly allocated resources are incorporated to the 
-entry construction (Line 41 in Figure \ref{fig:calltopteinitialize}). In other words, the side condition, (*),
- acknowledges any access to the entry with these resources as \textit{invalid}, as the entry is revealed to 
-shared-accesses when the presence bit is set.
+(Line \ref{line:install_new_entry} in Figure \ref{fig:calltopteinitialize}) after these freshly allocated resources are incorporated to the 
+entry construction via the page-frame portion of the PTE. In other words, the side condition, (*),
+ formalizes that any access to the entry with these resources as \textit{invalid}, until the entry is revealed to 
+shared accesses when the presence bit is set.
 
-  \subsubsection{Kernels Utilizing \textsf{P2V}}
+\subsubsection{Physical-to-Virtual Conversion with \textsf{P2V}}
 \label{sec:p2v}
- Once we reach to the certain knowledge of having an entry with a frame referencing to an allocated resource resides inside the identity mappings which can already be known (not taking the allocation path in (Lines 21-53 in Figure \ref{fig:calltopteinitialize}) or by ensuring the allocated frame is made present to shared accesses, then put into identity mappings -- logical update from Specification Lines 43-47 to 49-52 putting $\textsf{entry}_{\textsf{pfn}}$ into $\Xi$ in Figure \ref{fig:calltopteinitialize}), no matther how it is constructed, we can utilize this knowledge to convert this frame address into an virtual address of the next page table through, again, identity mappings -- Line 57 in Figure \ref{fig:calltopteinitialize} specified in Figure \ref{fig:p2v}.
+Once we reach to the certain knowledge of having an entry with a frame referencing to an allocated resource resides inside the identity mappings
+(which can already be known if the branch at Line \ref{line:check_entry_present_jump} is taken, or ensured by allocating and installing a new entry
+as just discussed for Lines \ref{line:check_entry_present_jump}--\ref{line:end_of_allocation_path}), 
+we can utilize this knowledge 
+to convert this frame address into an virtual address of the next page table through, again, identity mappings -- Line 57 in Figure \ref{fig:calltopteinitialize} 
+specified in Figure \ref{fig:p2v}.
 
@@
 This magical abstraction -- represented as $\Xi$ in the invariant definition -- provides us to construct this \textsf{pte-ptsto} out of an physical address is called \textit{identity-mappings}.
+}
+
+\todo[inline,color=blue]{MORE TO INTEGRATE: 
+\textsf{pte\_walkpgdir}, as a client, holds the knowledge that there exists an identity mapping for the physical entry address (\textsf{entry}) in the root page table ($\textsf{L}_{4}$):  $\mathsf{entry} \mapsto_{\textsf{id}} \textsf{\_}$ in Specification Line 3 is a partially owned token for accessing and looking up the resources in the identity map, $\Xi$, to construct the \textit{virtual-to-physical} pointsto relation $\textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry \entry\_val}$ with the virtual address (\textsf{entry+KERNBASE}) obtained by offsetting the physical address (\textsf{entry}).
+With this knowledge on the root-page-table-entry, we can start traversing the page-table tree which requires locating the address of the next table -- a call to \textsf{pte\_get\_next\_table} shown in Figure \ref{fig:calltopteinitialize}.
+Beyond a frame, the precondition before Line 15 requires the current address space invariant, and knowledge that \textsf{entry} is mapped to a random entry value, subtly, the operation also, at least, requires that the relevant table entry is readable, but the exact portion of ownership returned must be determined by inspecting the valid bit
+of the value in memory --- so full ownership is returned only for unused entries.
+This is a simple piece of code whose functionality is critical and whose correctness is highly non-trivial. No prior work engages with this problem.
 }

### 2024-08-05 examples.tex
finish text for 5.1.4

@@
 
-This actually constitutes a very critical piece of the full page table walk verification, and we have verified the critical step for a small x86-64 kernel, which is the physical-to-virtual conversion (often appearing as a macro \texttt{P2V} inside the macro \textsf{ PTE\_PFN\_TO\_ADDR} in C source code).
-In our small kernel (Line 57 in Figure \ref{fig:calltopteinitialize}), as in larger kernels, \texttt{P2V} is actually just addition, but the correctness of this simple instruction is quite subtle and relies on identity mappings explained in the previous section (Section \ref{subsec:identitymappings}) as a part of understanding our address-space invariant in Figure \ref{fig:peraspaceinvariant_with_p2v_extension}.
+This actually constitutes a very critical piece of the full page table walk verification, and we have verified the critical step for a small x86-64 kernel, 
+which is the physical-to-virtual conversion (often appearing as a macro \texttt{P2V} in C source code). 
+In our small kernel (Line \ref{line:p2v} in Figure \ref{fig:p2v}), as in larger kernels, \texttt{P2V} is actually just addition by the
+constant offset mentioned in Section \ref{subsec:identitymappings}, but the correctness 
+of this simple instruction is quite subtle and relies on the extended invariant (Figure \ref{fig:peraspaceinvariant_with_p2v_extension})
+explained in that section.
+
 \begin{figure}\footnotesize
@@
 $\specline{ \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{(pte\_initialized (entry\_val.pfn)))} \ast \texttt{rax} \mapsto_{\textsf{r}} \textsf{ table\_root (pte\_initialize(entry\_val.pfn))} }_{\rtv}$
-;;uintptr_t next_virt_addr = (uintptr_t) P2V(entry.pfn<<12);  
+;;uintptr_t next_virt_addr = (uintptr_t) P2V(entry.pfn<<12); (*\label{line:p2v}*) 
  movabs KERNBASE,rcx
@@
 \end{figure}
-As shown in Figure \ref{fig:p2v}, \lstinline|rcx| holds the offset value \textsf{KERNBASE} which gives us the value of the virtual address ($\textsf{entry}_{\textsf{pfn}}$ \textsf{+KERNBASE}) in the identity map ($\Xi\setminus\{entry\}\cup \{\textsf{entry}_{\textsf{pfn}}) \}$) of the kernel invariant: the logical update in Specification  Lines 5-10 to 10-14 for obtaining virtual-pointsto resource for the frame ($\textsf{entry}_{\textsf{pfn}}$) by removing it from the ghost map ($\Xi\setminus\{entry\}\cup \{\textsf{entry}_{\textsf{pfn}}) \}$) in Line 5 and compute the identity mapping for this physical frame address in Line 13 in Figure \ref{fig:p2v}).
+Figure \ref{fig:p2v} shows the verification of the end of \lstinline|pte_get_next_table| specialized to the case where 
+where no allocation was necessary (i.e., the conditional on Line \ref{line:check_entry_present} of Figure \ref{fig:calltopteinitialize} was taken).
+The code loads \lstinline|rcx| with the offset value \textsf{KERNBASE}, which gives us the value of the virtual address ($\textsf{entry}_{\textsf{pfn}}$ \textsf{+KERNBASE})
+of the \emph{base} of the next level of the page table.
+\todo[inline]{the next sentence depends on having figure 10 updated to reflect the page-worth of tokens}
+While we could now convert this address to a virtual points-to, this is not necessarily the correct thing to do.
+The caller \lstinline|walkpgdir| (discussed next) uses \lstinline|pte_get_next_table| to retrieve just the base address,
+because only the caller knows which entry in the subsequent table will be accessed (it depends on the corresponding bits from the virtual
+address being translated). So instead we pass back the per-address-space invariant with the identity mapping resources for \lstinline|entry|
+pulled out. It is up to the caller to determine which entry in that table must actually
+be accessed --- by selecting the appropriate index into the 512 ghost map tokens returned in the postcondition,
+and using the ghost translation and physical location portions of the invariant to assemble a vpte-pointsto.
+% in the identity map ($\Xi\setminus\{entry\}$) of the kernel invariant.
+% the logical update in Specification  Lines 5-10 to 10-14 for obtaining virtual-pointsto resource for the frame 
+% ($\textsf{entry}_{\textsf{pfn}}$) by removing it from the ghost map ($\Xi\setminus\{entry\}\cup \{\textsf{entry}_{\textsf{pfn}}) \}$) 
+% in Line 5 and compute the identity mapping for this physical frame address in Line 13 in Figure \ref{fig:p2v}).
 

### 2024-08-05 examples.tex
working into 5.1.5, draft fix to first line of fig 11

@@
 \begin{lstlisting}[mathescape, basicstyle=\footnotesize,escapeinside={(*}{*)}]
-;;pte_t *pte_get_next_table_succ(pte_t *entry) {
+;;pte_t *pte_get_next_table(pte_t *entry) {
 ... ;; setting up the stack
@@
 cmp    0x0,rax ;;  if (!entry->present) {(* \label{line:check_entry_present} *)
-jne    161 <pte_get_next_table_succ+0xa1> (* \label{line:check_entry_present_jump} *) ;; Jump if the present bit is not zero
+jne    161 <pte_get_next_table+0xa1> (* \label{line:check_entry_present_jump} *) ;; Jump if the present bit is not zero
 $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry} \ast \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{next} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val \& 0x1} }_{\rtv}$
@@
 \begin{lstlisting}[mathescape,escapeinside={(*}{*)}]
-... ;; Continued from Figure (*\ref{fig:calltopteinitialize}*) either by following the pte allocation path or not
+... ;; Continued from Figure (*\ref{fig:calltopteinitialize}*); assertions below specialized to non-allocating path for clarity
 $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \}),m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{rcx}  \mapsto_{\textsf{r}} \textsf{\_}  }_{\rtv}$
@@
 $\specline{ \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{(pte\_initialized (entry\_val.pfn)))} \ast \texttt{rax} \mapsto_{\textsf{r}} \textsf{ table\_root (pte\_initialize(entry\_val.pfn))} }_{\rtv}$
+$\specline{\forall_{i\in \textsf{0 ... 511} } \ldotp  \ghostmaptoken{\textsf{id}}{((\textsf{table\_root (pte\_initialized (entry\_val.pfn)))}) + \textsf{i * 8})}{\textsf{v-1}}  }$
 ;;uintptr_t next_virt_addr = (uintptr_t) P2V(entry.pfn<<12); (*\label{line:p2v}*) 
@@
 $\specline{ \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{(pte\_initialized (entry\_val.pfn)))} \ast \texttt{rax} \mapsto_{\textsf{r}} \textsf{ table\_root (pte\_initialize(entry\_val.pfn)) + KERNBASE} }_{\rtv}$
+$\specline{\forall_{i\in \textsf{0 ... 511} } \ldotp  \ghostmaptoken{\textsf{id}}{((\textsf{table\_root (pte\_initialized (entry\_val.pfn)))}) + \textsf{i * 8})}{\textsf{v-1}}  }$
 ...
@@
 \caption{Converting a physical address of a PTE to a virtual address (w/o instruction pointer or flag updates).
-Abreviating the other relevant resources (including the lower entries) peeled-out of the invariant as $\textsf{R}_{\textsf{children}}$ }
+ % Abreviating the other relevant resources (including the lower entries) peeled-out of the invariant as $\textsf{R}_{\textsf{children}}$
+}
 \label{fig:p2v}
@@
 of the \emph{base} of the next level of the page table.
-\todo[inline]{the next sentence depends on having figure 10 updated to reflect the page-worth of tokens}
+% \todo[inline]{the next sentence depends on having figure 10 updated to reflect the page-worth of tokens}
 While we could now convert this address to a virtual points-to, this is not necessarily the correct thing to do.
@@
 
-\subsubsection{Walking Page-Table Tree: Calling \textsf{pte\_get\_nexttable} for Each Level}
+\subsubsection{Walking Page-Table Tree: Calling \textsf{pte\_get\_next\_table} for Each Level}
 \label{wlkpgdir}
 
-\todo[inline,color=blue]{TAKEN FROM get-next-table DISCUSSION:
-In the very first precondition on the client side -- e.g.\textsf{pte\_walkpgdir} in Figure \ref{walkpgdir} -- 
-we see that \textsf{vpte-ptsto} for the entry being accessed (\textsf{entry}) is constructed -- in Specification Line \ref{line:get_next_vpte_precondition} of Figure \ref{fig:calltopteinitialize} for \textsf{entry}.
-Intuitively speaking, we start traversal from a certain entry address at the current level for which we know ($\textsf{entry} \mapsto_{\textsf{id}} \textsf{\_}$ in Line 3 in Figure \ref{walkpgdir}) that there is a memory mapping ($\textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val}$  Line 15 in Figure \ref{walkpgdir}) that can be used to explore deeper in the page-tree.
-
-The most crucial aspect of page-table entry access relies on the subtle interplay between updating the shared page tables, and the not violating the memory mappings. 
-Concretely speaking, accessing to the entry -- whose address is kept in the register $\texttt{r8}$ -- in Line 15 in Figure \ref{fig:calltopteinitialize} requires a 
-memory mapping that is logically provided through the update from Specfication Lines 3 to 15 in Figure \ref{walkpgdir} on the client side by the removal of \textsf{entry}
-from the map $\Xi$ in the invariant to construct the \textsf{vpte-ptsto} (Specfication Line 15 in Figure \ref{walkpgdir}). 
-This magical abstraction -- represented as $\Xi$ in the invariant definition -- provides us to construct this \textsf{pte-ptsto} out of an physical address is called \textit{identity-mappings}.
-}
-
-\todo[inline,color=blue]{MORE TO INTEGRATE: 
-\textsf{pte\_walkpgdir}, as a client, holds the knowledge that there exists an identity mapping for the physical entry address (\textsf{entry}) in the root page table ($\textsf{L}_{4}$):  $\mathsf{entry} \mapsto_{\textsf{id}} \textsf{\_}$ in Specification Line 3 is a partially owned token for accessing and looking up the resources in the identity map, $\Xi$, to construct the \textit{virtual-to-physical} pointsto relation $\textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry \entry\_val}$ with the virtual address (\textsf{entry+KERNBASE}) obtained by offsetting the physical address (\textsf{entry}).
-With this knowledge on the root-page-table-entry, we can start traversing the page-table tree which requires locating the address of the next table -- a call to \textsf{pte\_get\_next\_table} shown in Figure \ref{fig:calltopteinitialize}.
-Beyond a frame, the precondition before Line 15 requires the current address space invariant, and knowledge that \textsf{entry} is mapped to a random entry value, subtly, the operation also, at least, requires that the relevant table entry is readable, but the exact portion of ownership returned must be determined by inspecting the valid bit
-of the value in memory --- so full ownership is returned only for unused entries.
-This is a simple piece of code whose functionality is critical and whose correctness is highly non-trivial. No prior work engages with this problem.
-}
-
-Realizing page-table directory walk amounts to calling \textsf{pte\_get\_next\_table} for each level as shown in Figure \ref{walkpgdir}.
-The special part of the specficication for a page table walk can be considered as accumulation of memory mappings for the page-table entries visited and frame addresses for page-tables.
-For example, Line 23-24 in Figure \ref{walkpgdir} shows the virtual-pointsto for a L3 page table and the virtual-page-table-entry-pointsto for an L4 entry.
-In the final post-condition, we expect the accumulation of these resources from each level -- $\textsf{R}_{\textsf{walk\_succ}}$ -- which allows us to construct the path to the L1 entry in the tree to insert a new page.
 \begin{figure}\footnotesize
 \begin{lstlisting}[mathescape,escapeinside={(*}{*)}]
-;;pte_t *walkpgdir_succ(pte_t *pml4, const void *va) {
+;;pte_t *walkpgdir(pte_t *pml4, const void *va) {
  ...
- $\specline{\textsf{P} \ast \textsf{plm4} \mapsto_{\textsf{id}} \textsf{\_} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}}_{\rtv}$
+ $\specline{\textsf{P} \ast \ulcorner\textsf{rtv+KERNBASE}=\textsf{pml4}\urcorner \ast \forall_{i\in\textsf{0..511}}\ghostmaptoken{\textsf{id}}{(\textsf{rtv}+i*8)}{\textsf{4}} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}}_{\rtv}$
 ;;set up the stack for root address and virtual address    
@@
 $\specline{ \textsf{plm4\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{plm4 l4e\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l4e\_val})\urcorner}_{\rtv}$
-;;pte_t *pdp = pte_get_next_table_succ(pml4_entry);
+;;pte_t *pdp = pte_get_next_table(pml4_entry);
 mov    -0x18[rbp],rdi
 ...
-callq  c0 <pte_get_next_table_succ>
+callq  c0 <pte_get_next_table>
 ;;save the physical next table address in rax
@@
 $\specline{\forall_{i\in\textsf{0..511}} \; \textsf{table\_root(plm4\_entry}_{\textsf{pfn}}) \textsf{ + i * 8} \mapsto_{\textsf{id}} \textsf{3} }_{\rtv}$
-$\specline{ \textsf{plm4\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l4e\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present entry\_val})\urcorner}_{\rtv}$
-$\specline{ \textsf{pdp\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l3e\_val} \ast \ghostmaptoken{\textsf{id}}{(\mathsf{pdp\_entry})}{\textsf{3}}}_{\rtv}$
+$\specline{ \textsf{plm4\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l4e\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present entry\_val})\urcorner}_{\rtv}$(*\label{line:ex_l4_vpte}*)
+$\specline{ \textsf{pdp\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l3e\_val} \ast \ghostmaptoken{\textsf{id}}{(\mathsf{pdp\_entry})}{\textsf{3}}}_{\rtv}$(*\label{line:ex_l3_vpte}*)
 $\specline{\begin{array}{l} \textsf{entry\_present(l3e\_val)} -* \\  \;\;\;\;\;\;\; \forall_{i\in\textsf{0..511}} \; \textsf{table\_root(pdp\_entry}_{\textsf{pfn}}) \textsf{ + i * 8} \mapsto_{\textsf{id}} \textsf{2}    \end{array}}_{\rtv} $
 $\specline{\ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l3e\_val})\urcorner}_{\rtv}$
-;;pte_t *pd = pte_get_next_table_succ(pdp_entry);
+;;pte_t *pd = pte_get_next_table(pdp_entry);
 ...
 ;;pte_t *pd_entry = &pd[PDEX(va)];
-;;pte_t *pt = pte_get_next_table_succ(pd_entry);
-$\specline{\textsf{R}_{\textsf{walk\_acc}}}_{\rtv}$
+;;pte_t *pt = pte_get_next_table(pd_entry);
+$\specline{\textsf{R}_{\textsf{walk}}}_{\rtv}$
 ;;access and return L1 entry
@@
 \end{figure}
+
+Realizing a software page-table walk amounts to calling \textsf{pte\_get\_next\_table} for each level as shown in Figure \ref{walkpgdir}. 
+The special part of the specficication for a page table walk can be considered as accumulation of memory mappings for the page-table entries 
+visited and frame addresses for page-tables. 
+For example, Lines \ref{line:ex_l4_vpte} and \ref{line:ex_l3_vpte} in Figure \ref{walkpgdir} show the virtual pte-pointsto assertions for L4 and L3 entries.
+In the final post-condition, we expect the accumulation of these resources from each level -- $\textsf{R}_{\textsf{walk}}$ -- 
+which allows us to construct and return the path to the L1 entry in the tree to insert a new page.
+
+This is the code which performs the actual physical-to-virtual conversions using the identity mapping portion of the per-address-space invariant.
+
+\todo[inline]{colin frontier, need to incorporate these next 3 (2+1) paragraphs taken from earlier}
+\todo[inline,color=green]{need to finish updating ghost token assertions in figure}
+
+In the very first precondition on the client side -- e.g.\textsf{pte\_walkpgdir} in Figure \ref{walkpgdir} -- 
+we see that \textsf{vpte-ptsto} for the entry being accessed (\textsf{entry}) is constructed -- in S
+pecification Line \ref{line:get_next_vpte_precondition} of Figure \ref{fig:calltopteinitialize} for \textsf{entry}.
+Intuitively speaking, we start traversal from a certain entry address at the current level for which we know ($\textsf{entry} \mapsto_{\textsf{id}} \textsf{\_}$ in Line 3 in Figure \ref{walkpgdir}) that there is a memory mapping ($\textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val}$  Line 15 in Figure \ref{walkpgdir}) that can be used to explore deeper in the page-tree.
+
+The most crucial aspect of page-table entry access relies on the subtle interplay between updating the shared page tables, and the not violating the memory mappings. 
+Concretely speaking, accessing to the entry -- whose address is kept in the register $\texttt{r8}$ -- in Line 15 in Figure \ref{fig:calltopteinitialize} requires a 
+memory mapping that is logically provided through the update from Specfication Lines 3 to 15 in Figure \ref{walkpgdir} on the client side by the removal of \textsf{entry}
+from the map $\Xi$ in the invariant to construct the \textsf{vpte-ptsto} (Specfication Line 15 in Figure \ref{walkpgdir}). 
+This magical abstraction -- represented as $\Xi$ in the invariant definition -- provides us to construct this \textsf{pte-ptsto} out of an physical address is called \textit{identity-mappings}.
+
+\textsf{pte\_walkpgdir}, as a client, holds the knowledge that there exists an identity mapping for the physical entry address (\textsf{entry}) in the root page table ($\textsf{L}_{4}$):  $\mathsf{entry} \mapsto_{\textsf{id}} \textsf{\_}$ in Specification Line 3 is a partially owned token for accessing and looking up the resources in the identity map, $\Xi$, to construct the \textit{virtual-to-physical} pointsto relation $\textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry \entry\_val}$ with the virtual address (\textsf{entry+KERNBASE}) obtained by offsetting the physical address (\textsf{entry}).
+With this knowledge on the root-page-table-entry, we can start traversing the page-table tree which requires locating the address of the next table -- a call to \textsf{pte\_get\_next\_table} shown in Figure \ref{fig:calltopteinitialize}.
+Beyond a frame, the precondition before Line 15 requires the current address space invariant, and knowledge that \textsf{entry} is mapped to a random entry value, subtly, the operation also, at least, requires that the relevant table entry is readable, but the exact portion of ownership returned must be determined by inspecting the valid bit
+of the value in memory --- so full ownership is returned only for unused entries.
+This is a simple piece of code whose functionality is critical and whose correctness is highly non-trivial. No prior work engages with this problem.
+
 %\begin{figure}\footnotesize
@@
  $\specline{\textsf{P} \ast \textsf{plm4} \mapsto_{\textsf{id}} \textsf{\_} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}}_{\rtv}$      
-;;pte = walkpgdir_succ(pml4, a);
+;;pte = walkpgdir(pml4, a);
 mov    -0x8[rbp],rdi
@@
 mov    0x0,rax
-callq  291 <vaspace_mappage+0x41>	28d: R_X86_64_PLT32	walkpgdir_succ-0x4
+callq  291 <vaspace_mappage+0x41>	28d: R_X86_64_PLT32	walkpgdir-0x4
 ;;if(pte->present != 0) {

### 2024-08-05 examples.tex
halfway through updating figure 11

@@
 mov    -0x10[rbp],rdi
-shr    0x27,rdi
-and    0x1ff,rdi
-shl    0x3,rdi
-add    rdi,rsi
-mov    rsi,-0x18[rbp]
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{plm4\_entry}\},m)  }_{\rtv}$
-$\specline{ \ghostmaptoken{\textsf{id}}{(\mathsf{plm4\_entry})}{\textsf{4}} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
-$\specline{\begin{array}{l} \textsf{entry\_present(l4e\_val)} -* \\  \;\;\;\;\;\;\; \forall_{i\in\textsf{0..511}} \; \textsf{table\_root(plm4)}_{pfn} \textsf{ + i * 8} \mapsto_{\textsf{id}} \textsf{3}    \end{array}}_{\rtv} $
-$\specline{ \textsf{plm4\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{plm4 l4e\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l4e\_val})\urcorner}_{\rtv}$
+shr    0x27,rdi       ;; Shift L4 index to lowest bits
+and    0x1ff,rdi      ;; Mask to just lower 9 bits (0x1ff=511)
+shl    0x3,rdi        ;; Multiply by 8
+add    rdi,rsi        ;; Add to pml4 (virtual) table base
+mov    rsi,-0x18[rbp] ;; Store to local variable pml4_entry; logical pml4_entry is physical, program variable is virtual
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{pml4\_entry}\},m) \ast \forall_{i\in\textsf{0..511}\setminus\{\textsf{PML4EX(va)}\}}\ghostmaptoken{\textsf{id}}{(\textsf{rtv}+i*8)}{\textsf{4}}\ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
+$\specline{\textsf{entry\_present(l4e\_val)} \wand \forall_{i\in\textsf{0..511}} \; \textsf{table\_root(pml4)}_{pfn} \textsf{ + i * 8} \mapsto_{\textsf{id}} \textsf{3} }_{\rtv} $
+$\specline{ \textsf{pml4\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{pml4\_entry l4e\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l4e\_val})\urcorner}_{\rtv}$
 ;;pte_t *pdp = pte_get_next_table(pml4_entry);
@@
 ;;save the physical next table address in rax
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{plm4\_entry}\},m)  }_{\rtv}$
-$\specline{ \ghostmaptoken{\textsf{id}}{(\mathsf{plm4\_entry})}{\textsf{4}}  \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
-$\specline{\forall_{i\in\textsf{0..511}} \; \textsf{table\_root(plm4\_entry}_{\textsf{pfn}}) \textsf{ + i * 8} \mapsto_{\textsf{id}} \textsf{3} } $
-$\specline{ \textsf{plm4\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{plm4\_entry l4e\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l4e\_val})\urcorner}_{\rtv}$
-$\specline{\texttt{rax}  \mapsto_{\textsf{r}} \textsf{table\_root(pml4\_entry}_{\textsf{pfn}})}_{\rtv}$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{pml4\_entry}\},m)  \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
+$\specline{\forall_{i\in\textsf{0..511}} \; \textsf{table\_root(l4e\_val'.pfn) + i * 8} \mapsto_{\textsf{id}} \textsf{3}  \ast \ulcorner\textsf{pdp-KERNBASE}=\textsf{table\_root(l4e\_val'.pfn)}\urcorner} $
+$\specline{ \textsf{pml4\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{pml4\_entry l4e\_val'} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l4e\_val})\urcorner}_{\rtv}$
+$\specline{\texttt{rax}  \mapsto_{\textsf{r}} \textsf{table\_root(l4e\_val'.pfn)} }_{\rtv}$ ;; pte_get_next_table may have allocated a new page, updating entry
 ;;pte_t *pdp_entry = &pdp[PDPEX(va)];
 $\specline{\textsf{pdp}+\textsf{PDPEX(va)} = \textsf{pdp\_entry} \land \textsf{table\_root(pml4\_entry}_{\textsf{pfn}})  = \textsf{pdp}}_{\rtv}$
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{plm4\_entry},\textsf{pdp\_entry}\},m)  }_{\rtv}$
-$\specline{\ghostmaptoken{\textsf{id}}{(\mathsf{plm4\_entry})}{\textsf{4}} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
-$\specline{\forall_{i\in\textsf{0..511}} \; \textsf{table\_root(plm4\_entry}_{\textsf{pfn}}) \textsf{ + i * 8} \mapsto_{\textsf{id}} \textsf{3} }_{\rtv}$
-$\specline{ \textsf{plm4\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l4e\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present entry\_val})\urcorner}_{\rtv}$(*\label{line:ex_l4_vpte}*)
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{pml4\_entry},\textsf{pdp\_entry}\},m)  }_{\rtv}$
+$\specline{\ghostmaptoken{\textsf{id}}{(\mathsf{pml4\_entry})}{\textsf{4}} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
+$\specline{\forall_{i\in\textsf{0..511}} \; \textsf{table\_root(pml4\_entry}_{\textsf{pfn}}) \textsf{ + i * 8} \mapsto_{\textsf{id}} \textsf{3} }_{\rtv}$
+$\specline{ \textsf{pml4\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l4e\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present entry\_val})\urcorner}_{\rtv}$(*\label{line:ex_l4_vpte}*)
 $\specline{ \textsf{pdp\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l3e\_val} \ast \ghostmaptoken{\textsf{id}}{(\mathsf{pdp\_entry})}{\textsf{3}}}_{\rtv}$(*\label{line:ex_l3_vpte}*)
@@
 \todo[inline]{colin frontier, need to incorporate these next 3 (2+1) paragraphs taken from earlier}
-\todo[inline,color=green]{need to finish updating ghost token assertions in figure}
+\todo[inline,color=green]{need to finish updating ghost token assertions in figure; I think they're correct now up to the pdp_entry assignment line}
 
 In the very first precondition on the client side -- e.g.\textsf{pte\_walkpgdir} in Figure \ref{walkpgdir} -- 
-we see that \textsf{vpte-ptsto} for the entry being accessed (\textsf{entry}) is constructed -- in S
-pecification Line \ref{line:get_next_vpte_precondition} of Figure \ref{fig:calltopteinitialize} for \textsf{entry}.
-Intuitively speaking, we start traversal from a certain entry address at the current level for which we know ($\textsf{entry} \mapsto_{\textsf{id}} \textsf{\_}$ in Line 3 in Figure \ref{walkpgdir}) that there is a memory mapping ($\textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val}$  Line 15 in Figure \ref{walkpgdir}) that can be used to explore deeper in the page-tree.
+we see that \textsf{vpte-ptsto} for the entry being accessed (\textsf{entry}) is constructed -- in 
+Specification Line \ref{line:get_next_vpte_precondition} of Figure \ref{fig:calltopteinitialize} for \textsf{entry}. 
+Intuitively speaking, we start traversal from a certain entry address at the current level for which we know 
+($\textsf{entry} \mapsto_{\textsf{id}} \textsf{\_}$ in Line 3 in Figure \ref{walkpgdir}) that 
+there is a memory mapping ($\textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val}$ 
+Line 15 in Figure \ref{walkpgdir}) that can be used to explore deeper in the page-tree.
 
@@
 
-\textsf{pte\_walkpgdir}, as a client, holds the knowledge that there exists an identity mapping for the physical entry address (\textsf{entry}) in the root page table ($\textsf{L}_{4}$):  $\mathsf{entry} \mapsto_{\textsf{id}} \textsf{\_}$ in Specification Line 3 is a partially owned token for accessing and looking up the resources in the identity map, $\Xi$, to construct the \textit{virtual-to-physical} pointsto relation $\textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry \entry\_val}$ with the virtual address (\textsf{entry+KERNBASE}) obtained by offsetting the physical address (\textsf{entry}).
-With this knowledge on the root-page-table-entry, we can start traversing the page-table tree which requires locating the address of the next table -- a call to \textsf{pte\_get\_next\_table} shown in Figure \ref{fig:calltopteinitialize}.
-Beyond a frame, the precondition before Line 15 requires the current address space invariant, and knowledge that \textsf{entry} is mapped to a random entry value, subtly, the operation also, at least, requires that the relevant table entry is readable, but the exact portion of ownership returned must be determined by inspecting the valid bit
+\textsf{walkpgdir}, as a client, holds the knowledge that there exists an identity mapping for the physical entry address (\textsf{entry})
+ in the root page table ($\textsf{L}_{4}$):  $\mathsf{entry} \mapsto_{\textsf{id}} \textsf{\_}$ in Specification Line 3 is a partially owned
+ token for accessing and looking up the resources in the identity map, $\Xi$, to construct the \textit{virtual-to-physical} pointsto relation 
+$\textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry \entry\_val}$ with the virtual address (\textsf{entry+KERNBASE}) obtained 
+by offsetting the physical address (\textsf{entry}). With this knowledge on the root-page-table-entry, we can start traversing the page-table 
+tree which requires locating the address of the next table -- a call to \textsf{pte\_get\_next\_table} shown in Figure \ref{fig:calltopteinitialize}. 
+Beyond a frame, the precondition before Line 15 requires the current address space invariant, and knowledge that \textsf{entry} is mapped to a 
+random entry value, subtly, the operation also, at least, requires that the relevant table entry is readable, but the exact portion of ownership 
+returned must be determined by inspecting the valid bit
 of the value in memory --- so full ownership is returned only for unused entries.
@@
 ...
- $\specline{\textsf{P} \ast \textsf{plm4} \mapsto_{\textsf{id}} \textsf{\_} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}}_{\rtv}$      
+ $\specline{\textsf{P} \ast \textsf{pml4} \mapsto_{\textsf{id}} \textsf{\_} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}}_{\rtv}$      
 ;;pte = walkpgdir(pml4, a);
@@
 ;;pte->nocache      = PTE_NOCACHE_SET(perm);
-$\specline{ \left( \begin{array}{l} \textsf{pdp}+\textsf{PDPEX(va)} = \textsf{pdp\_entry} \land \textsf{table\_root(pml4\_entry}_{\textsf{pfn}})  = \textsf{pdp} \\ \textsf{pd}+\textsf{PDEX(va)} = \textsf{pd\_entry} \land \textsf{table\_root(pml3\_entry}_{\textsf{pfn}})  = \textsf{pd} \\ \textsf{pt}+\textsf{PDPEX(va)} = \textsf{pt\_entry} \land \textsf{table\_root(pml2\_entry}_{\textsf{pfn}})  = \textsf{pt} \\  \textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{plm4\_entry},\textsf{pdp\_entry},\textsf{pd\_entry},\textsf{pt\_entry}\},m)  \\ \ghostmaptoken{\delta{}s}{\rtv}{\delta} \\
-\ghostmaptoken{\textsf{id}}{(\mathsf{plm4\_entry})}{\textsf{4}}\ast \ghostmaptoken{\textsf{id}}{(\mathsf{plm3\_entry})}{\textsf{3}} \ast \ghostmaptoken{\textsf{id}}{(\mathsf{plm2\_entry})}{\textsf{2}}  \\ \forall_{i\in\textsf{0..511}} \; \textsf{table\_root(plm4\_entry}_{\textsf{pfn}}) \textsf{ + i * 8} \mapsto_{\textsf{id}} \textsf{3}  \\ \forall_{i\in\textsf{0..511}} \; \textsf{table\_root(pdp\_entry}_{\textsf{pfn}}) \textsf{ + i * 8} \mapsto_{\textsf{id}} \textsf{2}  \\ \forall_{i\in\textsf{0..511}} \; \textsf{table\_root(pd\_entry}_{\textsf{pfn}}) \textsf{ + i * 8} \mapsto_{\textsf{id}} \textsf{1}  \\  \textsf{plm4\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l4e\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l4e\_val})\urcorner \\  \textsf{pdp\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l3e\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l3e\_val})\urcorner \\  \textsf{pd\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l2e\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l2e\_val})\urcorner \\   \textsf{pt\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l1e\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l1e\_val})\urcorner    \end{array} \right)  =\textsf{R}_{\textsf{walk\_acc}}}$
+$\specline{ \left( \begin{array}{l} \textsf{pdp}+\textsf{PDPEX(va)} = \textsf{pdp\_entry} \land \textsf{table\_root(pml4\_entry}_{\textsf{pfn}})  = \textsf{pdp} \\ \textsf{pd}+\textsf{PDEX(va)} = \textsf{pd\_entry} \land \textsf{table\_root(pml3\_entry}_{\textsf{pfn}})  = \textsf{pd} \\ \textsf{pt}+\textsf{PDPEX(va)} = \textsf{pt\_entry} \land \textsf{table\_root(pml2\_entry}_{\textsf{pfn}})  = \textsf{pt} \\  \textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{pml4\_entry},\textsf{pdp\_entry},\textsf{pd\_entry},\textsf{pt\_entry}\},m)  \\ \ghostmaptoken{\delta{}s}{\rtv}{\delta} \\
+\ghostmaptoken{\textsf{id}}{(\mathsf{pml4\_entry})}{\textsf{4}}\ast \ghostmaptoken{\textsf{id}}{(\mathsf{plm3\_entry})}{\textsf{3}} \ast \ghostmaptoken{\textsf{id}}{(\mathsf{plm2\_entry})}{\textsf{2}}  \\ \forall_{i\in\textsf{0..511}} \; \textsf{table\_root(pml4\_entry}_{\textsf{pfn}}) \textsf{ + i * 8} \mapsto_{\textsf{id}} \textsf{3}  \\ \forall_{i\in\textsf{0..511}} \; \textsf{table\_root(pdp\_entry}_{\textsf{pfn}}) \textsf{ + i * 8} \mapsto_{\textsf{id}} \textsf{2}  \\ \forall_{i\in\textsf{0..511}} \; \textsf{table\_root(pd\_entry}_{\textsf{pfn}}) \textsf{ + i * 8} \mapsto_{\textsf{id}} \textsf{1}  \\  \textsf{pml4\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l4e\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l4e\_val})\urcorner \\  \textsf{pdp\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l3e\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l3e\_val})\urcorner \\  \textsf{pd\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l2e\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l2e\_val})\urcorner \\   \textsf{pt\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l1e\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l1e\_val})\urcorner    \end{array} \right)  =\textsf{R}_{\textsf{walk\_acc}}}$
 pte->pfn          = PTE_ADDR_TO_PFN(pa);

### 2024-08-05 examples.tex
finished revising figure 11 assertions, found maybe inconsistency between figures 11 and 7?

@@
 ;; pte_t *next;
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{entry}\},m)  \ast \texttt{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast  \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{next} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{\_} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{\_}  \ast \textsf{entry} \mapsto_{\textsf{id}} \textsf{v} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{entry}\},m)  \ast \texttt{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast  \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{next} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{\_} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{\_}  \ast \ghostmaptoken{\textsf{id}}{\textsf{entry}}{\textsf{v}} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
 $\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present entry\_val})\urcorner}_{\rtv}$ (*\label{line:get_next_vpte_precondition}*)
@@
 ;;pte_t *walkpgdir(pte_t *pml4, const void *va) {
- ...
+... ;; Stack setup
 $\specline{\textsf{P} \ast \ulcorner\textsf{rtv+KERNBASE}=\textsf{pml4}\urcorner \ast \forall_{i\in\textsf{0..511}}\ghostmaptoken{\textsf{id}}{(\textsf{rtv}+i*8)}{\textsf{4}} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}}_{\rtv}$
 ;;set up the stack for root address and virtual address    
- ;;pte_t *pml4_entry = &pml4[PML4EX(va)];
+;;pte_t *pml4_entry = &pml4[PML4EX(va)]; // Virtual address of L4 entry(*\label{line:start_pml4_calc}*)
 mov    -0x8[rbp],rsi
@@
 shl    0x3,rdi        ;; Multiply by 8
-add    rdi,rsi        ;; Add to pml4 (virtual) table base
+add    rdi,rsi        ;; Add to pml4 (virtual) table base(*\label{line:end_pml4_calc}*)
 mov    rsi,-0x18[rbp] ;; Store to local variable pml4_entry; logical pml4_entry is physical, program variable is virtual
 $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{pml4\_entry}\},m) \ast \forall_{i\in\textsf{0..511}\setminus\{\textsf{PML4EX(va)}\}}\ghostmaptoken{\textsf{id}}{(\textsf{rtv}+i*8)}{\textsf{4}}\ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
-$\specline{\textsf{entry\_present(l4e\_val)} \wand \forall_{i\in\textsf{0..511}} \; \textsf{table\_root(pml4)}_{pfn} \textsf{ + i * 8} \mapsto_{\textsf{id}} \textsf{3} }_{\rtv} $
+$\specline{\textsf{entry\_present(l4e\_val)} \wand \forall_{i\in\textsf{0..511}} \; \ghostmaptoken{\textsf{id}}{\textsf{table\_root(pml4.pfn)+ i * 8}}{\textsf{3}} }_{\rtv} $
 $\specline{ \textsf{pml4\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{pml4\_entry l4e\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l4e\_val})\urcorner}_{\rtv}$
@@
 $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{pml4\_entry}\},m)  \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
-$\specline{\forall_{i\in\textsf{0..511}} \; \textsf{table\_root(l4e\_val'.pfn) + i * 8} \mapsto_{\textsf{id}} \textsf{3}  \ast \ulcorner\textsf{pdp-KERNBASE}=\textsf{table\_root(l4e\_val'.pfn)}\urcorner} $
+$\specline{\forall_{i\in\textsf{0..511}} \ghostmaptoken{\textsf{id}}{\textsf{table\_root(l4e\_val'.pfn) + i * 8}}{\textsf{3}}  \ast \ulcorner\textsf{pdp-KERNBASE}=\textsf{table\_root(l4e\_val'.pfn)}\urcorner} $
 $\specline{ \textsf{pml4\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{pml4\_entry l4e\_val'} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l4e\_val})\urcorner}_{\rtv}$
 $\specline{\texttt{rax}  \mapsto_{\textsf{r}} \textsf{table\_root(l4e\_val'.pfn)} }_{\rtv}$ ;; pte_get_next_table may have allocated a new page, updating entry
-;;pte_t *pdp_entry = &pdp[PDPEX(va)];
-$\specline{\textsf{pdp}+\textsf{PDPEX(va)} = \textsf{pdp\_entry} \land \textsf{table\_root(pml4\_entry}_{\textsf{pfn}})  = \textsf{pdp}}_{\rtv}$
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{pml4\_entry},\textsf{pdp\_entry}\},m)  }_{\rtv}$
-$\specline{\ghostmaptoken{\textsf{id}}{(\mathsf{pml4\_entry})}{\textsf{4}} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
-$\specline{\forall_{i\in\textsf{0..511}} \; \textsf{table\_root(pml4\_entry}_{\textsf{pfn}}) \textsf{ + i * 8} \mapsto_{\textsf{id}} \textsf{3} }_{\rtv}$
-$\specline{ \textsf{pml4\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l4e\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present entry\_val})\urcorner}_{\rtv}$(*\label{line:ex_l4_vpte}*)
-$\specline{ \textsf{pdp\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l3e\_val} \ast \ghostmaptoken{\textsf{id}}{(\mathsf{pdp\_entry})}{\textsf{3}}}_{\rtv}$(*\label{line:ex_l3_vpte}*)
-$\specline{\begin{array}{l} \textsf{entry\_present(l3e\_val)} -* \\  \;\;\;\;\;\;\; \forall_{i\in\textsf{0..511}} \; \textsf{table\_root(pdp\_entry}_{\textsf{pfn}}) \textsf{ + i * 8} \mapsto_{\textsf{id}} \textsf{2}    \end{array}}_{\rtv} $
-$\specline{\ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l3e\_val})\urcorner}_{\rtv}$
+;;pte_t *pdp_entry = &pdp[PDPEX(va)]; // Virtual address of L3 entry
+$\specline{\ulcorner\textsf{pdp}+\textsf{PDPEX(va)*8} = \textsf{pdp\_entry} \land \textsf{table\_root(l4e\_val'.pfn)}  = \textsf{pdp} \urcorner }_{\rtv}$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{pml4\_entry},\textsf{pdp\_entry}\},m)  \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
+$\specline{\forall_{i\in\textsf{0..511}\setminus\{\textsf{PDPEX(va)}\}} \; \textsf{table\_root(pml4\_entry}_{\textsf{pfn}}) \textsf{ + i * 8} \mapsto_{\textsf{id}} \textsf{3} }_{\rtv}$
+$\specline{ \textsf{pml4\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac4}}\textsf{pml4\_entry}\; \textsf{l4e\_val'} \ast \ulcorner  \textsf{qfrac4} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l4e\_val'})\urcorner}_{\rtv}$(*\label{line:ex_l4_vpte}*)
+$\specline{ \textsf{pdp\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac3}}\textsf{pdp\_entry}\; \textsf{l3e\_val} \ast \ulcorner  \textsf{qfrac3} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l3e\_val})\urcorner}_{\rtv}$(*\label{line:ex_l3_vpte}*)
+$\specline{\textsf{entry\_present(l3e\_val)} \wand \forall_{i\in\textsf{0..511}} \;\ghostmaptoken{\textsf{id}}{\textsf{table\_root(pdp\_entry.pfn) + i * 8}}{ \textsf{2}} } _{\rtv} $
 ;;pte_t *pd = pte_get_next_table(pdp_entry);
-...
-;;pte_t *pd_entry = &pd[PDEX(va)];
+... ;; Similar assembly to reach next level
+;;pte_t *pd_entry = &pd[PDEX(va)]; // Virtual address of L2 entry
 ;;pte_t *pt = pte_get_next_table(pd_entry);
@@
 ;;access and return L1 entry
-;;return &pt[PTEX(va)];
+;;return &pt[PTEX(va)]; // Virtual address of L1 entry
 ...
@@
 This is the code which performs the actual physical-to-virtual conversions using the identity mapping portion of the per-address-space invariant.
+\lstinline|walkpgdir| accepts a \emph{virtual} pointer to the base of the L4 table, and the address to translate.
+The precondition provides knowledge that the virtual base of the L4 is at the appropriate offset from the current \lstineline|cr3| value,
+but does not provide a virtual points-to assertion --- because the function must calculate (Lines \ref{line:start_pml4_calc}--\ref{line:end_pml4_calc})
+which entry it needs access to.
+Instead the precondition has 512 identity map tokens, guaranteeing that every entry on the page is subject to the identity mapping invariant.
+Line \ref{line:end_pml4_calc} calculates the virtual address of the relevant entry, and the subsequent view shift
+pulls that entry out of the identity mapping ($\Xi$) and fetches its
 
-\todo[inline]{colin frontier, need to incorporate these next 3 (2+1) paragraphs taken from earlier}
-\todo[inline,color=green]{need to finish updating ghost token assertions in figure; I think they're correct now up to the pdp_entry assignment line}
+\todo[inline]{Colin is unsure:
+It looks like the call to pte-get-next-table on line 19 is done after using the idmap ($\Xi$) token for the entry's physical address.
+But the verification of pte-get-next-table starts by assuming it has a vpte-pointsto \emph{and} the token!
+Except to build the vpte-pointsto, we had to remove that from the invariant around $\Xi$; 2/3 of the resources membership in $\Xi$
+entails were already consumed to build it. That's why we have the entry-present-implies-l3-tokens
+assertion on line 14 of Figure 11.
+So there's a misalignment of these specs. It seems like pte-get-next-table should not get the token for the physical address, but should instead get
+the conditional assertion at line 14 directly?
+}
 

### 2024-08-06 examples.tex
finish updating walkpgdir discussion, need to loop back and update get-next discussion and figure again per zulip discussion

@@
 Accessing the presence bit and checking the value allows us to exploit the side condition (*) when verifying the allocation
-path for when the entry is invalid (Lines \ref{line:alloc_path_start}--\ref{line:alloc_path_end} in Figure \ref{fig:calltopteinitialize}),
-as it entails full ownership of the entry's memory ($\textsf{qfrac} = 1$) and justifies writing to that memory.
+path for when the entry is invalid (Lines \ref{line:alloc_path_start}--\ref{line:alloc_path_end} in Figure \ref{fig:calltopteinitialize}).
+This operation is subtle: the operation requires that the relevant table entry is readable, but the exact portion of ownership 
+returned must be determined by inspecting the valid bit of the value in memory --- so full ownership is returned only for unused entries.
+When the bit is not set, that entails full ownership of the entry's memory ($\textsf{qfrac} = 1$) and justifies writing to that memory.
 Otherwise, the code jumps past the end of this listing, to the following code at the top of Figure \ref{fig:p2v} (which is also the
@@
 % frame are freshly allocated with full-ownership to be a part of the identity map, $\Xi$. 
-The soundness argument of this specification relies on the fact that these freshly allocated resources is part 
+The soundness argument of this specification relies on the fact that these freshly allocated resources are part 
 of an entry construction that has not been completed yet: the presence bit is set 
@@
 ...
-callq  c0 <pte_get_next_table>
+callq  c0 <pte_get_next_table>(*\label{line:first_getnext_call}*)
 ;;save the physical next table address in rax
@@
 $\specline{\forall_{i\in\textsf{0..511}} \ghostmaptoken{\textsf{id}}{\textsf{table\_root(l4e\_val'.pfn) + i * 8}}{\textsf{3}}  \ast \ulcorner\textsf{pdp-KERNBASE}=\textsf{table\_root(l4e\_val'.pfn)}\urcorner} $
-$\specline{ \textsf{pml4\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{pml4\_entry l4e\_val'} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l4e\_val})\urcorner}_{\rtv}$
+$\specline{ \textsf{pml4\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{pml4\_entry l4e\_val'} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l4e\_val})\urcorner}_{\rtv}$(*\label{line:first_pte_pointsto}*)
 $\specline{\texttt{rax}  \mapsto_{\textsf{r}} \textsf{table\_root(l4e\_val'.pfn)} }_{\rtv}$ ;; pte_get_next_table may have allocated a new page, updating entry
@@
 
-This is the code which performs the actual physical-to-virtual conversions using the identity mapping portion of the per-address-space invariant.
+This is the code which performs most actual physical-to-virtual conversions using the identity mapping portion of the per-address-space invariant.
 \lstinline|walkpgdir| accepts a \emph{virtual} pointer to the base of the L4 table, and the address to translate.
-The precondition provides knowledge that the virtual base of the L4 is at the appropriate offset from the current \lstineline|cr3| value,
+The precondition provides knowledge that the virtual base of the L4 is at the appropriate offset from the current \lstinline|cr3| value,
 but does not provide a virtual points-to assertion --- because the function must calculate (Lines \ref{line:start_pml4_calc}--\ref{line:end_pml4_calc})
@@
 Line \ref{line:end_pml4_calc} calculates the virtual address of the relevant entry, and the subsequent view shift
-pulls that entry out of the identity mapping ($\Xi$) and fetches its
+pulls that entry out of the identity mapping ($\Xi$) and fetches its corresponding resources as
+described by Figure \ref{fig:peraspaceinvariant_with_p2v_extension} and Section \ref{subsec:identitymappings}.
+The ghost translation and physical location are used to form the virtual pte-pointsto for the L4 entry
+(Line \ref{line:first_pte_pointsto}), with the entry validity and next-level indexing
+satisfying the rest of the precondition for \lstinline|pte_get_next_table|.
+\todo[inline]{This line above assumes the fixes to the precondition of get-next-table to take the conditional child
+token map}
+\lstinline|pte_get_next_table| then, as described earlier, checks the valid bit in the indicated
+entry and either returns the (unconditional) tokens for the L3 entry physical addresses (if valid), or
+allocates into the entry and returns new (also unconditional) tokens for the L3 entry physical addresses.
+\lstinline|pte_get_next_table|'s first call (Line \ref{line:first_getnext_call}) returns
+the virtual address of the base of the L3 table (a \emph{page directory pointer}, so PDP, in official
+x86-64 terminology). Then the situation to move from that pointer to the base of the L2
+is just like the process just followed: the proof calculates the address of the relevant
+L3 entry, uses the appropriate L3 identity mapping token to construct a virtual pte-pointsto to that entry,
+and passes that along with additional resources pulled out of the invariant to another call to
+\lstinline|pte_get_next_table|. That call then returns the base of an L2 table, and the process
+repeats until the function returns the virtual address of the relevant L1 entry.
+That will then be used in the next section by the caller of \lstinline|walkpgdir|
+to install a new mapping.
 
-\todo[inline]{Colin is unsure:
-It looks like the call to pte-get-next-table on line 19 is done after using the idmap ($\Xi$) token for the entry's physical address.
-But the verification of pte-get-next-table starts by assuming it has a vpte-pointsto \emph{and} the token!
-Except to build the vpte-pointsto, we had to remove that from the invariant around $\Xi$; 2/3 of the resources membership in $\Xi$
-entails were already consumed to build it. That's why we have the entry-present-implies-l3-tokens
-assertion on line 14 of Figure 11.
-So there's a misalignment of these specs. It seems like pte-get-next-table should not get the token for the physical address, but should instead get
-the conditional assertion at line 14 directly?
-}
-
-In the very first precondition on the client side -- e.g.\textsf{pte\_walkpgdir} in Figure \ref{walkpgdir} -- 
-we see that \textsf{vpte-ptsto} for the entry being accessed (\textsf{entry}) is constructed -- in 
-Specification Line \ref{line:get_next_vpte_precondition} of Figure \ref{fig:calltopteinitialize} for \textsf{entry}. 
-Intuitively speaking, we start traversal from a certain entry address at the current level for which we know 
-($\textsf{entry} \mapsto_{\textsf{id}} \textsf{\_}$ in Line 3 in Figure \ref{walkpgdir}) that 
-there is a memory mapping ($\textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val}$ 
-Line 15 in Figure \ref{walkpgdir}) that can be used to explore deeper in the page-tree.
-
-The most crucial aspect of page-table entry access relies on the subtle interplay between updating the shared page tables, and the not violating the memory mappings. 
-Concretely speaking, accessing to the entry -- whose address is kept in the register $\texttt{r8}$ -- in Line 15 in Figure \ref{fig:calltopteinitialize} requires a 
-memory mapping that is logically provided through the update from Specfication Lines 3 to 15 in Figure \ref{walkpgdir} on the client side by the removal of \textsf{entry}
-from the map $\Xi$ in the invariant to construct the \textsf{vpte-ptsto} (Specfication Line 15 in Figure \ref{walkpgdir}). 
-This magical abstraction -- represented as $\Xi$ in the invariant definition -- provides us to construct this \textsf{pte-ptsto} out of an physical address is called \textit{identity-mappings}.
-
-\textsf{walkpgdir}, as a client, holds the knowledge that there exists an identity mapping for the physical entry address (\textsf{entry})
- in the root page table ($\textsf{L}_{4}$):  $\mathsf{entry} \mapsto_{\textsf{id}} \textsf{\_}$ in Specification Line 3 is a partially owned
- token for accessing and looking up the resources in the identity map, $\Xi$, to construct the \textit{virtual-to-physical} pointsto relation 
-$\textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry \entry\_val}$ with the virtual address (\textsf{entry+KERNBASE}) obtained 
-by offsetting the physical address (\textsf{entry}). With this knowledge on the root-page-table-entry, we can start traversing the page-table 
-tree which requires locating the address of the next table -- a call to \textsf{pte\_get\_next\_table} shown in Figure \ref{fig:calltopteinitialize}. 
-Beyond a frame, the precondition before Line 15 requires the current address space invariant, and knowledge that \textsf{entry} is mapped to a 
-random entry value, subtly, the operation also, at least, requires that the relevant table entry is readable, but the exact portion of ownership 
-returned must be determined by inspecting the valid bit
-of the value in memory --- so full ownership is returned only for unused entries.
-This is a simple piece of code whose functionality is critical and whose correctness is highly non-trivial. No prior work engages with this problem.
+% \textsf{walkpgdir}, as a client, holds the knowledge that there exists an identity mapping for the physical entry address (\textsf{entry})
+%  in the root page table ($\textsf{L}_{4}$):  $\mathsf{entry} \mapsto_{\textsf{id}} \textsf{\_}$ in Specification Line 3 is a partially owned
+%  token for accessing and looking up the resources in the identity map, $\Xi$, to construct the \textit{virtual-to-physical} pointsto relation 
+% $\textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry \entry\_val}$ with the virtual address (\textsf{entry+KERNBASE}) obtained 
+% by offsetting the physical address (\textsf{entry}). With this knowledge on the root-page-table-entry, we can start traversing the page-table 
+% tree which requires locating the address of the next table -- a call to \textsf{pte\_get\_next\_table} shown in Figure \ref{fig:calltopteinitialize}. 
+% Beyond a frame, the precondition before Line 15 requires the current address space invariant, and knowledge that \textsf{entry} is mapped to a 
+% random entry value, subtly, 
+% the operation also, at least, requires that the relevant table entry is readable, but the exact portion of ownership 
+% returned must be determined by inspecting the valid bit
+% of the value in memory --- so full ownership is returned only for unused entries.
+% This is a simple piece of code whose functionality is critical and whose correctness is highly non-trivial. No prior work engages with this problem.
 

### 2024-08-06 examples.tex
fixed figure 11, tied together passing of guarded children

@@
 \label{sec:traversing}
-We build up to the main 
-The mapping operation of Figure \ref{fig:mapping_code} assumes an operation \textsf{pte\_walkpgdir} which must traverse the page tables
-in order to locate the address of the L1 entry to update --- possibly allocating tables at levels 3, 2, and 1 in the process.
+We build up to the main task of mapping a new page after traversing page tables in software.
+The mapping operation of Figure \ref{fig:mapping_code} assumes an operation \textsf{walkpgdir} which must traverse the page tables
+in order to locate the address of the L1 entry to update --- possibly allocating tables for levels 3, 2, and 1 in the process,
+installing them into levels 4, 3, and 2, along the way.
 Traversing the page tables is itself challenging functionality to verify: loading the current table root from \lstinline|cr3| is straightforward
@@
 \subsubsection{Loading Page-Table Address Value}
-We will discuss access to the level 4 table later. But for subsequent levels, the base address of level $n$ must be
+We will discuss access to the level 4 table later (Section \ref{wlkpgdir}). But for subsequent levels, the base address of level $n$ must be
 fetched from the appropriate entry in the level $n+1$ table.
@@
 
-Note that none of the verification for this function assumes specific page table levels (it is not even a parameter to the verification);
-it is used for all three level transitions when traversing page tables (4 to 3, 3 to 2, 2 to 1).
+Note that none of the verification for this function assumes specific page table levels --- logical parameter \textsf{v} represents the level
+of the entry passed as an argument, and this code
+is used for all three level transitions when traversing page tables (4 to 3, 3 to 2, 2 to 1).
 This comes into play with a subtlety of the specification of \lstinline|pte_get_next_table| that we will
@@
 ;; pte_t *next;
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{entry}\},m)  \ast \texttt{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast  \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{next} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{\_} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{\_}  \ast \ghostmaptoken{\textsf{id}}{\textsf{entry}}{\textsf{v}} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{entry}\},m)  \ast \texttt{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast  \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{next} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{\_} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{\_} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
 $\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present entry\_val})\urcorner}_{\rtv}$ (*\label{line:get_next_vpte_precondition}*)
+$\specline{\ulcorner\textsf{entry\_present(entry\_val)}\urcorner \wand \forall_{i\in\textsf{0..511}} \; \ghostmaptoken{\textsf{id}}{\textsf{table\_root(entry\_val.pfn) + i * 8}}{\textsf{v-1}} }_{\rtv} $ (*\label{line:conditional_children}*)
 mov    -0x8[rbp],rdi
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{entry\},m)  \ast \textsf{entry}\mapsto_{\textsf{pte,q}}\textsf{entry\_val} \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{\_}  \ast  \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{next} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry}  }_{\rtv}$
-$\specline{ \textsf{entry} \mapsto_{\textsf{id}} \textsf{v} \ast\urcorner \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{entry\},m)   \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{\_}  \ast  \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{next} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
 $\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present entry\_val})\urcorner}_{\rtv}$
+$\specline{\ulcorner\textsf{entry\_present(entry\_val)}\urcorner \wand \forall_{i\in\textsf{0..511}} \; \ghostmaptoken{\textsf{id}}{\textsf{table\_root(entry\_val.pfn) + i * 8}}{\textsf{v-1}} }_{\rtv} $
 mov     rdi, r8
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast  \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{next} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry} }_{\rtv}$
-$\specline{ \textsf{entry} \mapsto_{\textsf{id}} \textsf{v} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast  \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{next} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
 $\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present entry\_val})\urcorner}_{\rtv}$
+$\specline{\ulcorner\textsf{entry\_present(entry\_val)}\urcorner \wand \forall_{i\in\textsf{0..511}} \; \ghostmaptoken{\textsf{id}}{\textsf{table\_root(entry\_val.pfn) + i * 8}}{\textsf{v-1}} }_{\rtv} $
 mov    [r8],rdi (*\label{line:read_entry_contents}*)
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast  \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{next} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val}   }_{\rtv}$
-$\specline{  \textsf{entry} \mapsto_{\textsf{id}} \textsf{v} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast  \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{next} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val}  \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
 $\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present entry\_val})  \urcorner }_{\rtv}$
+$\specline{\ulcorner\textsf{entry\_present(entry\_val)}\urcorner \wand \forall_{i\in\textsf{0..511}} \; \ghostmaptoken{\textsf{id}}{\textsf{table\_root(entry\_val.pfn) + i * 8}}{\textsf{v-1}} }_{\rtv} $
 and    0x1,rdi (* \label{line:mask_present} *)
@@
 
-For each $\paddr\in\Xi$, the invariant tracks a virtual points-to justifying that virtual address $\paddr+\textsf{KERNBASE}$ maps to physical address $\paddr$
+For each $\paddr\mapsto \texsf{v} \in\Xi$, the invariant tracks a virtual points-to justifying that virtual address $\paddr+\textsf{KERNBASE}$ maps to physical address $\paddr$
 (the ``Ghost translation'' in Figure \ref{fig:peraspaceinvariant_with_p2v_extension});
 fractional ownership of the physical memory for that page table entry;
-and for valid entries (with the present bit set) above L1, ghost map tokens for every entry in the table pointed to by the entry.
+and for valid entries (with the present bit set) above L1, ghost map tokens for every entry in the table pointed to by the entry, which can be used
+to repeat the process one level down. i
 (L1 entries point to data pages, whose physical memory ownership resides in some virtual points-to).
+The assertion on Line \ref{line:conditional_children} originates from the invariant one level up, and if this code determines the valid bit is set,
+it can return those child tokens without the conditional guard.
 
@@
 $\specline{ \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{(pte\_initialized (entry\_val.pfn)))} \ast \texttt{rax} \mapsto_{\textsf{r}} \textsf{ table\_root (pte\_initialize(entry\_val.pfn))} }_{\rtv}$
-$\specline{\forall_{i\in \textsf{0 ... 511} } \ldotp  \ghostmaptoken{\textsf{id}}{((\textsf{table\_root (pte\_initialized (entry\_val.pfn)))}) + \textsf{i * 8})}{\textsf{v-1}}  }$
+$\specline{\forall_{i\in \textsf{0 ...
+511} } \ldotp  \ghostmaptoken{\textsf{id}}{((\textsf{table\_root (pte\_initialized (entry\_val.pfn)))}) + \textsf{i * 8})}{\textsf{v-1}}  }$ (*\label{line:children}*)
 ;;uintptr_t next_virt_addr = (uintptr_t) P2V(entry.pfn<<12); (*\label{line:p2v}*) 
@@
 where no allocation was necessary (i.e., the conditional on Line \ref{line:check_entry_present} of Figure \ref{fig:calltopteinitialize} was taken).
+In this case, the present bit being set grants access to the child tokens from Line \ref{line:conditional_children} of Figure \ref{fig:calltopteinitialize},
+which is then refined to the assertion on Line \ref{line:children} of Figure \ref{fig:p2v}.
 The code loads \lstinline|rcx| with the offset value \textsf{KERNBASE}, which gives us the value of the virtual address ($\textsf{entry}_{\textsf{pfn}}$ \textsf{+KERNBASE})
@@
 $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{pml4\_entry}\},m) \ast \forall_{i\in\textsf{0..511}\setminus\{\textsf{PML4EX(va)}\}}\ghostmaptoken{\textsf{id}}{(\textsf{rtv}+i*8)}{\textsf{4}}\ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
-$\specline{\textsf{entry\_present(l4e\_val)} \wand \forall_{i\in\textsf{0..511}} \; \ghostmaptoken{\textsf{id}}{\textsf{table\_root(pml4.pfn)+ i * 8}}{\textsf{3}} }_{\rtv} $
+$\specline{\textsf{entry\_present(l4e\_val)} \wand \forall_{i\in\textsf{0..511}} \; \ghostmaptoken{\textsf{id}}{\textsf{table\_root(l4e\_val)+ i * 8}}{\textsf{3}} }_{\rtv} $
 $\specline{ \textsf{pml4\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{pml4\_entry l4e\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l4e\_val})\urcorner}_{\rtv}$
@@
 satisfying the rest of the precondition for \lstinline|pte_get_next_table|.
-\todo[inline]{This line above assumes the fixes to the precondition of get-next-table to take the conditional child
-token map}
 \lstinline|pte_get_next_table| then, as described earlier, checks the valid bit in the indicated
@@
 % a set of unordered block allocations, we can infer their proofs must also trust these translations.
+
 \subsection{Mapping a New Page}

### 2024-08-06 examples.tex
helps when you spell command names correctly

@@
 
-For each $\paddr\mapsto \texsf{v} \in\Xi$, the invariant tracks a virtual points-to justifying that virtual address $\paddr+\textsf{KERNBASE}$ maps to physical address $\paddr$
+For each $\paddr\mapsto \textsf{v} \in\Xi$, the invariant tracks a virtual points-to justifying that virtual address $\paddr+\textsf{KERNBASE}$ maps to physical address $\paddr$
 (the ``Ghost translation'' in Figure \ref{fig:peraspaceinvariant_with_p2v_extension});

### 2024-09-16 examples.tex
arxiv2

@@
 current address space) has previously been formally verified with respect to a machine model with address translation.
-Second, we formally verify a switch into a new address space as part of a task switch.\todo{Colin will double-check the TLB work's details before claiming this is the first}
+Second, we formally verify a switch into a new address space as part of a task switch.%\todo{Colin will double-check the TLB work's details before claiming this is the first}
 
@@
 To do so, with a given allocated a fresh page (\textsf{fpaddr}), then calculate the appropriate
-known-valid page table walks (\textsf{pte\_walkpgdir} Line X in Figure \ref{fig:mapping_code})  and update the appropriate L1 page table entry (permissions and frame address Lines in Figure \ref{fig:mapping_code}).
+known-valid page table walks ($\textsf{pte\_walkpgdir}$ Line 11 in Figure \ref{fig:mapping_code})  and update the appropriate L1 page table entry (permissions and frame address Lines 17-22 in Figure \ref{fig:mapping_code}).
 %\lstset{
@@
 %}
-\todo{ismail change the following spec base on the changes you have in the code}
+%\todo{ismail change the following spec base on the changes you have in the code}
 \begin{figure}\footnotesize
@@
 In Figure \ref{fig:mapping_code}, we see an address ($\vaddr$) currently not
-mapped to a page ($\theta \; !!\; \vaddr = \texttt{None}$). Mapping a fresh
+mapped to a page (assuming $\theta \; !!\; \vaddr = \texttt{None}$). Mapping a fresh
 phyiscal page to back the desired virtual page first requires ensuring
@@
 The code uses a helper function \lstinline{pte\_walkpgdir} (discussed again in Section \ref{sec:traversing}).
-\textsf{pte\_walkpgdir}'s postcondition contains virtual \emph{PTE} pointsto assertions ($\mapsto_{\textsf{vpte}}$) \todo{colin, these are no more plain physical pointstos}  both for ensuring partial page table walk reaching the
+\textsf{pte\_walkpgdir}'s postcondition contains virtual \emph{PTE} pointsto assertions ($\mapsto_{\textsf{vpte}}$) %\todo{colin, these are no more plain physical pointstos}
+both for ensuring partial page table walk reaching the
 L1 entry (l1e) with the justification of higher levels of the page table exist, and for allowing access to the memory of the L1 entry via virtual address.
@@
 the modified data is at the specific physical location that will be used as the L1 entry.
-This is also guaranteed ($ \ulcorner
-\texttt{addr\_L1}(\vaddr,\entryo) = \paddr\urcorner$).
+By traversing the page-table tree, we obtain virtual page-table entry pointsto relations (Line 21 in Figure  \ref{fig:mapping_code}) for each level down to the L1. 
 % After obtaining a virtual address \textsf{pte\_addr} in \textsf{rax} backed 
@@
 
-The crucial step is actually updating the L1 entry (Line X\todo{X} in Figure \ref{fig:mapping_code}),
-via the virtual address
-(\textsf{pte\_addr}) known to translate to the appropriate physical address, in our example the L1
-table entry address ($\textsf{addr\_L1}(\textsf{va, l1e})$), to hold the freshly
-allocated physical page address (\textsf{fpaddr}) in Line X.
-After this write, we have (Specification Line Y)
-\[\texttt{pte\_addr} \mapsto_{\texttt{vpte}}  \; \paddr\; \textsf{fpaddr}  \qquad\qquad (\textsf{by the PTE variant of Rule}~\TirNameStyle{WriteToVirtMemFromReg})\]
+The crucial step is actually updating the L1 entry (Line 22 in Figure \ref{fig:mapping_code}),
+via the virtual address known to translate to the appropriate physical address to hold the freshly
+allocated physical page address (\textsf{pa}): from Specification Line 27 to 29 by (\textsf{by the PTE variant of Rule}~\TirNameStyle{WriteToVirtMemFromReg}).
 
-At this point in the proof,  $\texttt{pte\_addr} \mapsto_{\texttt{vpte}}  \; \paddr\; \textsf{fpaddr})$ 
-contains  full ownerhsip of the L1 table entry physical pointsto $\paddr
-\mapsto_{\mathsf{p}} \textsf{fpaddr}$ . Then, we can split the full-ownership of that
-physical pointsto
-($\textsf{fpaddr}\mapsto_{\textsf{p}} \;\textsf{wzero 64}$) to obtain
-fractional ownership on it appropriate for a single address's share of an L1 entry (q4).
-\todo{ismail comment out the rest of this subsection and fix according to the new map-a-page}
+In fact, by the application of Rule ~\TirNameStyle{WriteToVirtMemFromReg}, we obtain full ownerhsip of the L1 table entry physical pointsto. Then, we can split the full-ownership of that
+physical pointsto to obtain
+fractional ownership on it appropriate for a single address's share of an L1 entry (q in Specification Line 29).
+%\todo{ismail comment out the rest of this subsection and fix according to the new map-a-page}
 \begin{comment}
@@
 \mapsto_{\textsf{p}} \textsf{wzero 64}$.
-\end{comment}
+
 Finally, by existentially quantifying away \textsf{fpaddr},
 we can shift the PTE points-to into a normal virtual points-to.
-
+\end{comment}
 For brevity this proof only maps the first word of the allocated page; the generalization to the full page
@@
 but instead uses logic rules throughout to reason about the effect of the mapping. 
-This is the first proof we know of for mapping a virtual memory page that does not resort to direct reasoning over the operational semantics.
+This is the first proof we know of for mapping a virtual memory page via using virtual memory handles that does not resort to direct reasoning over the operational semantics.
 % By incorporating verification of the
@@
 \subsection{Unmapping a Page}
-\todo[inline]{update (esp. line refs) for new mapping code}
+%\todo[inline]{update (esp. line refs) for new mapping code}
 The reverse operation, unmapping a designated page that is currently mapped,

### 2024-09-16 programlogic.tex
arxiv2

@@
 implementing a fragment of x86-64 including 64-bit address translation with 4-level page tables.
-\todo[inline]{
+
+%\todo[inline]{
 Our rules for control transfers (\lstinline|jmp|, \lstinline|call|, and \lstinline|ret|) are currently
-axiomatized (with completely standard specifications~\cite{ni2007contexts,Chlipala2013Bedrock}) because Iris's built-in machinery does not provide
-convenient ways to discard the current continuation. We are working to adapt existing machinery~\cite{de2023type} to prove their soundness.
+axiomatized (with completely standard specifications~\cite{ni2007contexts,Chlipala2013Bedrock}).% because Iris's built-in machinery does not provide
+%convenient ways to discard the current continuation.
+We are working to adapt existing machinery~\cite{de2023type} to prove their soundness.
 Our soundness proofs for all other instructions (including the instructions above, arithmetic, other heap-accessing operations, push, and pop)
 are axiom-free.
-}
+%}
 % At the moment our proofs do rely on 13 small axioms of properties which should be provable, but

### 2024-11-11 background.tex
more compression and editing, through middle of section 4
remove unnecessary todos
condense the heck out of bg section, shaving off nearly a page

@@
 \begin{figure}
-    \includegraphics[width=\columnwidth]{pagetables.png}
-    \todo[inline]{We should try to produce a more compact version of this diagram}
+    \includegraphics[width=\columnwidth]{pagetables2.png}
+    \vspace{-2.5em}
     \caption{x86-64 page table lookups.}
     \label{fig:pagetables}
+    \vspace{-1em}
 \end{figure}
 
-On the x86-64 architecture, the input to the \textsc{MMU}'s address translation is a sparse hierarchical set of tables 
-called \emph{page tables} (referring to pages of memory). As Figure \ref{fig:pagetables} (based on Figure 5-17 of the 
+On the x86-64 architecture, the \textsc{MMU}'s address translation uses a sparse hierarchical set of tables:
+\emph{page tables} (referring to pages of memory). As Figure \ref{fig:pagetables} (based on Figure 5-17 of the 
 AMD64 architecture manual~\cite{amd64_manual_vol2}\footnote{While x86 up through its 32-bit incarnation were due to Intel,
@@
 only deals with 4-level page tables, but is straightforwardly extensible to 5.}  
-Because address translation 
-traverses a series of tables, the translation process or algorithm is sometimes referred to as a \emph{page-table walk}. 
+The translation process or algorithm is sometimes referred to as a \emph{page-table walk}. 
 While Figure \ref{fig:pagetables} and most of our constants (how many levels, which virtual address bits index which
@@
 
-The page tables mapping virtual addresses to physical addresses are managed by the OS. Typically each process has its own 
-virtual-to-physical mappings represented by its own page table, which the OS registers with the CPU by storing the (page-aligned)
-physical address of the root of the page table tree (in our context, the start of the L4 table) in a 
+The page tables are managed by the OS, typically by a \emph{virtual memory manager} (VMM).\footnote{Not to be confused with Virtual 
+Machine Monitor. In this paper we focus on non-hypervisor scenarios, but hardware virtualization extensions for both 
+x86-64 and ARM make use of an additional set of page tables translating what a \emph{guest} considers to be its 
+(virtualized) physical memory to actual physical memory. Our contributions should offer value in this scenario as well.}
+Typically each process has its own 
+page table, which the OS registers with the CPU by storing the (page-aligned)
+physical address of the root of the page table treethe start of the L4 table) in a 
 specific register (\texttt{cr3}) as part of switching to a new process. 
@@
 the relevant page table entry), the hardware triggers a \emph{page fault}, transferring control to a \emph{page fault handler} registered 
-with the hardware by the OS, allowing it to take corrective action (described in the next section).
+with the hardware by the OS, allowing it to take corrective action.
+If no mapping was supposed to exist, this is a program bug (e.g., dereferencing virtual address 0 / NULL)
+and the faulting program should be terminated. But this can also be used for
+specialized functionality and optimizations, such as \emph{paging} (saving room in physical RAM and deferring
+unnecessary IO by only reading program code from disk when it is accessed, or even swapping memory that has not
+recently been accessed to disk, to read back in when a page fault indicates access).
+
+The key pieces of VMM functionality are
+adding a new page mapping (whether the mapped page contains zeros, file data, or swap data), and removing an existing 
+page mapping.
+While this initially sounds like relatively modest functionality whose implementation may be complicated by hardware 
+subtleties, correctness of even these basic operations are actually quite intrictate.
+Notably, updates to the page tables are performed as writes to memory --- \emph{which are themselves subject to address translation},
+and finding the correct page table to update requires converting between physical and virtual addresses.
+In the case of changing the mappings for the currently-active set of page tables, 
+\emph{the OS kernel is modifying the tables involved in its
+own access of the tables}.
+
+Virtual memory concerns propagate to the OS scheduler, which deals with multiple 
+address spaces, so must keep track of which virtual addresses are valid (and in what way) in which address spaces. 
+Some virtual addresses are valid in only a single address space (e.g., a code address for a particular usermode 
+process), while others are valid in all address spaces (e.g., kernel data structure pointers). 
+The VMM must maintain some of these assumptions on behalf of the rest of the kernel, for example by guaranteeing that 
+a certain range of virtual addresses (corresponding to the kernel's code and data) are valid in every address space.
 
@@
 CPUs with MMUs typically have an additional \emph{translation lookaside buffer}
-(TLB), to cache the (successful) results of page table walks, rather than merely transforming every virtual
+(TLB), to cache the (successful) results of page table walks, rather than transforming every virtual
 memory access into 5 physical accesses. 
@@
 Appendix \ref{apdx:tlb} gives more details on the challenges involved for the interested reader.
+\looseness=-1
 
-\subsection{Virtual Memory Managers}
-\label{sec:backgroundonvmm}
-Most OS kernels have a component called a \emph{Virtual Memory Manager} (VMM)\footnote{Not to be confused with Virtual 
-Machine Monitor. In this paper we focus on non-hypervisor scenarios, but hardware virtualization extensions for both 
-x86-64 and ARM make use of an additional set of page tables translating what a \emph{guest} considers to be its 
-(virtualized) physical memory to actual physical memory. Our contributions should offer value in this scenario as well.}
-which is responsible for setting up page table mappings and for taking action when a page fault occurs. Often, a page 
-fault indicates a bug in a program --- for example, a null pointer dereference crashes a program not because the hardware
- designates it an error, but because most OSes refuse to map the first 4K worth of virtual addresses, so dereferencing 
-\texttt{NULL} results in a page fault. In such cases, the OS will terminate the program.
+% \subsection{Virtual Memory Managers}
+% \label{sec:backgroundonvmm}
+% Most OS kernels have a component called a \emph{Virtual Memory Manager} (VMM)
+% which is responsible for setting up page table mappings and for taking action when a page fault occurs. Often, a page 
+% fault indicates a bug in a program --- for example, a null pointer dereference crashes a program not because the hardware
+%  designates it an error, but because most OSes refuse to map the first 4K worth of virtual addresses, so dereferencing 
+% \texttt{NULL} results in a page fault. In such cases, the OS will terminate the program.
 
-In other cases the OS uses page faults to implement specialized functionality and optimizations. A key example of this is 
-(confusingly) called \emph{paging}: saving room in physical RAM and avoiding unnecessary IO 
-by waiting until memory addresses for code are accessed 
-before copying them to RAM (in particular, saving process startup time); 
-or copying memory that has not recently been accessed to disk 
-(i.e., in a \emph{swap} file or partition) and marking those page table entries invalid so a future page fault will give 
-the OS the chance to copy the relevant data back into memory when the program tries to access it again.
-
-The key pieces of functionality a VMM must implement itself, or have available in another OS component (depending on 
-design), are
-adding a new page mapping (whether the mapped page contains zeros, file data, or swap data), and removing an existing 
-page mapping.
-While this initially sounds like relatively modest functionality whose implementation may be complicated by hardware 
-subtleties, even these basic operations are actually quite intrictate.
-Most notably, updates to the page tables are performed as writes to memory --- which are themselves subject to address translation.
-In the case of changing the mappings for the currently-active set of page tables, the OS kernel is modifying the tables involved in its
-own access of the tables.
+% In other cases the OS uses page faults to implement specialized functionality and optimizations. A key example of this is 
+% (confusingly) called \emph{paging}: saving room in physical RAM and avoiding unnecessary IO 
+% by waiting until memory addresses for code are accessed 
+% before copying them to RAM (in particular, saving process startup time); 
+% or copying memory that has not recently been accessed to disk 
+% (i.e., in a \emph{swap} file or partition) and marking those page table entries invalid so a future page fault will give 
+% the OS the chance to copy the relevant data back into memory when the program tries to access it again.
 
@@
 
-Virtual memory concerns propagate to the OS scheduler, which is responsible for dealing with multiple 
-address spaces, so must keep track of which virtual addresses are valid (and in what way) in which address spaces. 
-Some virtual addresses are valid in only a single address space (e.g., a code address for a particular usermode 
-process), while others are valid in all address spaces (e.g., kernel data structure pointers). 
-The VMM must maintain some of these assumptions on behalf of the rest of the kernel, for example by guaranteeing that 
-a certain range of virtual addresses (corresponding to the kernel's code and data) are valid in every address space.
-
 \subsection{Modal Logic}
@@
 The problem of needing to keep track of things being true in some contexts and not in others is hardly unique to virtual 
-memory management, and is the general insight behind most flavors of modal logic.
+memory management, and is the general insight behind most flavors of modal logic, which use
+unary operators to express that a logical claim $P$ is \emph{contingently} true 
+in certain other circumstances, such as in other times~\cite{pnueli1977temporal} or places~\cite{murphy2008type,gordon2019modal}.
 
-Most modal logics use unary operators on propositions to express that a logical claim $P$ is \emph{contingently} true 
-in certain other circumstances, such as in other times~\cite{pnueli1977temporal} or places~\cite{gordon2019modal}. That is, for 
-different modal operators $M$, $M(P)$ may mean that $P$ is true in another time or place (which may be a physical place 
-like another network node~\cite{murphy2008type,gordon2019modal} or a conceptual place like a person's 
-mind~\cite{hintikka1962knowledge}). A unifying concept across any modality is that they behave as applicative functors, 
-typically satisfying (directly, or as a derived law, depending on the modality):
-\[ (P\rightarrow Q) \rightarrow M(P) \rightarrow M(Q)\]
-%\todo[inline]{Ismail, I think this means we have pure intro, $\square P \mathrel{-\ast} [r](\square P)$ if I'm using the right symbol for pure assertions }
-%\todo[inline,color=red]{The above todo comment isn't quite right.
-%Purity is about being able to duplicate. You had defined another typeclass/property-of-assertions that meant it didn't
-%care what address space it was in (like physical pointstos). That's orthogonal to purity.
-%}
-Many modalities, so-called \emph{normal} modalities also possess introduction rules of the form $P\rightarrow M(P)$, 
-the classic example being that if $P$ is true, then $P$ is \emph{necessarily} true with the contingency picked up.%($\square P$).
+%   A unifying concept across any modality is that they behave as applicative functors, 
+% typically satisfying (directly, or as a derived law, depending on the modality):
+% \[ (P\rightarrow Q) \rightarrow M(P) \rightarrow M(Q)\]
+% %\todo[inline]{Ismail, I think this means we have pure intro, $\square P \mathrel{-\ast} [r](\square P)$ if I'm using the right symbol for pure assertions }
+% %\todo[inline,color=red]{The above todo comment isn't quite right.
+% %Purity is about being able to duplicate. You had defined another typeclass/property-of-assertions that meant it didn't
+% %care what address space it was in (like physical pointstos). That's orthogonal to purity.
+% %}
+% Many modalities, so-called \emph{normal} modalities also possess introduction rules of the form $P\rightarrow M(P)$, 
+% the classic example being that if $P$ is true, then $P$ is \emph{necessarily} true with the contingency picked up.%($\square P$).
 
@@
 other worlds. This explicit naming of alternate worlds increases the power of modal logics~\cite{blackburn1995hybrid}, and is actually
-necessary for completeness in classical separation logics~\cite{brotherston2014parametric}.  A key distinguishing feature of
-hybrid logic is that the logic itself does not fix the set of modalities, but is defined relative to any set of sensible worlds.
+necessary for completeness in classical separation logics~\cite{brotherston2014parametric}.
+%   A key distinguishing feature of
+% hybrid logic is that the logic itself does not fix the set of modalities, but is defined relative to any set of sensible worlds.
 
@@
 space rooted at physical address $r$. Because OS kernels create and destroy address spaces, it is sensible to use
-a hybrid-style logic that does not care about the exact set of modalities.
+a hybrid-style logic that is not specialized to a fixed set of modalities, but this introduces
+some subtleties from the fact that the existence of certain modalities (address spaces) can change.
 
-The typical hybrid modality introduction rule, that $P$ and $\ell$ (indicating the current world is $\ell$) imply 
-$[\ell](P)$, has a natural analogue: knowing $P$ and that the current address space is $r$ (i.e., that $r$ is the 
-current \texttt{cr3} value) suggests a way to construct $[r](P)$. 
-We identify an assertion as \emph{contextual} if its validity depends on the choice of address space. Used outside an explicit
-modality, the truth of this assertion depends on the current \texttt{cr3} value. Used under an explicit address space
-modality, its truth depends on both the modality-chosen address space and associated physical resources.
+% The typical hybrid modality introduction rule, that $P$ and $\ell$ (indicating the current world is $\ell$) imply 
+% $[\ell](P)$, has a natural analogue: knowing $P$ and that the current address space is $r$ (i.e., that $r$ is the 
+% current \texttt{cr3} value) suggests a way to construct $[r](P)$. 
+% We identify an assertion as \emph{contextual} if its validity depends on the choice of address space. Used outside an explicit
+% modality, the truth of this assertion depends on the current \texttt{cr3} value. Used under an explicit address space
+% modality, its truth depends on both the modality-chosen address space and associated physical resources.
 % \begin{itemize}
@@
 
-A relatively under-explored space of modal logics is the interaction of modal and substructural logics, 
-in particular hybrid-style modalities in substructural logics, which has seen only minimal exploration~\cite{dovsen1992modal,restall1993modalities,d1997grafting,kamide2002kripke,licata2017fibrational} and no prior
-application. We develop our ideas in the Iris program logic~\cite{jung2018iris}, as a result exploring additional subtleties 
+Interaction of hybrid modalities and substructural reasoning is relatively unexplored (see Section \ref{sec:relwork}).
+% A relatively under-explored space of modal logics is the interaction of modal and substructural logics, 
+% in particular hybrid-style modalities in substructural logics, which has seen only minimal exploration~\cite{dovsen1992modal,restall1993modalities,d1997grafting,kamide2002kripke,licata2017fibrational} and no prior
+% application. 
+We develop our ideas in the Iris program logic~\cite{jung2018iris}, as a result exploring additional subtleties 
 that arise where the modality itself may entail ownership of resources, as well as interactions between our hybrid-style 
-modality and substructural rules.  For example, Iris contains a number of modalities that distribute over separating 
-conjunction, or for which resources can freely move into the modality 
-(e.g., $\blacktriangleright(P)\ast Q \wand \blacktriangleright(P\ast Q)$). In our setting some of these rules 
-apply while others do not. For example, in our setting, an assertion that involves no modalities is interpreted as 
-holding in the current (active-on-the-CPU) address space, so clearly cannot move into arbitrary other address spaces 
---- unless guarded by another address space modality.
+modality and substructural rules.  
+% For example, Iris contains a number of modalities that distribute over separating 
+% conjunction, or for which resources can freely move into the modality 
+% (e.g., $\blacktriangleright(P)\ast Q \wand \blacktriangleright(P\ast Q)$). In our setting some of these rules 
+% apply while others do not. For example, 
+% In our setting, an assertion that involves no modalities is interpreted as 
+% holding in the current (active-on-the-CPU) address space, so clearly cannot move into arbitrary other address spaces,
+% while address-space-relative assertions 
+% --- unless guarded by another address space modality.
 Some prior \iris-based work~\cite{dang2019rustbelt,dang2022compass} has constructed derived modalities in the style we propose, indexed

### 2024-11-12 intro.tex
compressing stuff up through (including) the strong virtual pointsto def

@@
 which the kernel still accesses via virtual address, leading to indirect cycles.
+\looseness=-1
 
-Further complicating matters is that addresses themselves bear no information about which address space they originate 
-from. For user processes this is of little concern, as these have access to only their own address space. But the kernel has
+Further complicating matters, addresses themselves bear no information about which address space they originate 
+from. 
+%For user processes this is of little concern, as these have access to only their own address space. 
+This is not a problem for usermode code, which only accesses its own address space.
+But the kernel has
 (or can grant itself) access to all address spaces. Mixing up addresses from different address spaces leads to severe bugs.
-More concerning, keeping track of which \emph{assertions} hold in different address spaces during kernel verification is 
-difficult: some assertions should hold across all address spaces, while others hold in only one, and others may hold in 
+Proofs of kernel correctness must also track which \emph{assertions} hold in different address spaces:
+some assertions should hold across all address spaces, while others hold in only one, and others may hold in 
 multiple but still not all.
@@
 \textit{convenient} and \textit{powerful} way to \emph{modularly}
-\begin{itemize}
-\item isolate assertions specific to a particular address space,
-\item explicitly state when an assertion is true across address spaces,
-\item manipulate address spaces from within other address spaces, and
-\item reason about change in address spaces.
-\end{itemize}
+isolate assertions specific to a particular address space,
+explicitly state when an assertion is true across address spaces,
+manipulate address spaces from within other address spaces, and
+reason about change in address spaces.
 These advantages make this approach to reasoning about virtual memory more flexible than prior program logic techniques~\cite{kolanski08vstte,kolanski09tphols}, 
@@
 the \emph{non-local} effects of page table updates whether within the current address space or across address spaces.
-%\begin{figure}
-%   \includegraphics[width=0.5\columnwidth]{architecture.png}
-%  \caption{Layers of Implementation}
-%  \todo[inline]{Colin says: I don't think this figure helps readers; it is too abstract}
-%  \label{fig:architecture}
-%\end{figure}
 
-\begin{itemize}
-\item We develop these ideas in the form of a logic for working with virtual-address-space-relative assertions,
-      implemented as an embedded separation logic within the Iris~\cite{jung2018iris} separation logic.
+\paragraph{Contributions}
+We develop these ideas in the form of a logic for working with virtual-address-space-relative assertions,
+implemented as an embedded separation logic within Iris~\cite{jung2018iris}.
 The result is a separation logic that lifts a number of major semantic restrictions present in the few
 prior logics tackling virtual address translation.
-\item We prove the soundness of our \textsf{vProp} logic with respect to a RISC-like fragment of \textsf{AMD64} instructions.
-\item We verify simplified versions of several critical virtual-memory-related pieces of OS functionality, 
-      including software page table walks, mapping  pages, and the assembly code for switching address spaces. 
-      The latter in particular goes beyond the capabilities of prior work, and reveals unexpected
-	subtleties.
-\end{itemize}
-
-The logic we develop here covers core reasoning principles for reasoning about memory configurations and code
+The logic we develop covers core reasoning principles for reasoning about memory configurations and code
 reliant upon or manipulating those memory configurations in the presence of in-memory page tables, the primary
 memory protection mechanism across Intel/AMD's x86-64 processors, ARM's application class processors including
-aarch64 CPUs, POWER, RISC-V, and other architectures.
+AArch64 CPUs, POWER, RISC-V, and other architectures.
+We prove the soundness of our \textsf{vProp} logic with respect to a RISC-like fragment of \textsf{AMD64} instructions.
+We verify simplified versions of several critical virtual-memory-related pieces of OS functionality, 
+including mapping  pages, switching address spaces, and converting physical addresses to virtual addresses for
+software page table walks (as when mapping pages). These examples demonstrate the suitability and flexibility of
+a modal treatment of address spaces: each has a concise specification, each example either goes beyond the technical
+capabilities of prior logics, or revisits an example from prior work with more details (treating parts of the kernel
+invariant never addressed by prior work) and fewer assumptions (e.g., verifying virtual-to-physical conversion rather
+than assuming it has already occurred).
+
 % One aspect of virtual memory management we do not address in this paper is updates to the translation lookaside

### 2024-11-12 programlogic.tex
compressing stuff up through (including) the strong virtual pointsto def

--- 2024-07-09 programlogic.tex
@@
 }
-\begin{lstlisting}[language=Coq]
-  Definition vProp  $\Sigma$ : bi := word 64 -b> iPropI  $\Sigma$.
-\end{lstlisting}
+\lstinline[language=Coq]|Definition vProp  $\Sigma$ : bi := word 64 -b> iPropI  $\Sigma$|.
 This is the (\textsc{Coq}) type of assertions in our logic.
@@
 
-\subsection{Base Points-To Assertions}
-\label{sec:pointsto}
-We build up our context-dependent assertions from two basic types of assertions that do not care about address spaces.
+% \subsection{Base Points-To Assertions}
+% \label{sec:pointsto}
+% We build up our context-dependent assertions from two basic types of assertions that do not care about address spaces.
+To define and use virtual points-to assertions, we require two more basic assertions:
 
@@
 %and, expectedly \textsf{w2} is an address computed by page-offset computation (e.g. $\textsf{l4off maddr cr3val}$). 
-From this we can derive a more concise physical points-to when the split is unimportant:
+From this we can derive a more concise physical points-to when the split is unimportant:\\
 % Giving a raw 64-bits memory pointsto assertion becomes
-\[\textsf{w} \mapsto_{\textsf{p}} \{q\} \textsf{ v} \stackrel{\triangle}{=} (\textsf{drop 12}~w) \ \sim \ (\textsf{bottom 12}~w)\mapsto_{\textsf{a}} \; \{\textsf{q}\} \textsf{ v} \]
+\centerline{$\textsf{w} \mapsto_{\textsf{p}} \{q\} \textsf{ v} \stackrel{\triangle}{=} (\textsf{drop 12}~w) \ \sim \ (\textsf{bottom 12}~w)\mapsto_{\textsf{a}} \; \{\textsf{q}\} \textsf{ v} $}
+
  %I put a newline in between the following two Definitions as the second Definition seems not proper without the newline
@@
  $\exists_{\textsf{l4e l3e l2e l1e}} \ldotp$  $\ulcorner$ aligned $\vaddr \urcorner \ast$ L4_L1_PointsTo($\vaddr$ l4e l3e l2e l1e paddr) $\ast$ paddr $\mapsto_{p}\{\mathsf{q}\} \vpage$.
-
  Definition L4_L1_PointsTo (maddr l4e l3e l2e l1e paddr :word 64) : vProp $\Sigma$ := $\lambda$ cr3val.
@@
   \label{fig:strongvirtualpointsto}
+\vspace{-1em}
 \end{figure*}
@@
 \label{sec:overly-restrictive}
-A natural definition for a virtual points-to that depends on the current address space is to
-require that in order for a virtual address \textsf{va} to point to a value \textsf{v}, the assertion contains
+A natural definition for a virtual points-to
+asserting that virtual address \textsf{va} points to a value \textsf{v}
+would 
+% require that in order for a virtual address \textsf{va} to point to a value \textsf{v}, the assertion 
+contain
 partial ownership of the physical memory involved in the page table walk that would translate \textsf{va} to
 its backing physical location --- with locations existentially quantified since a virtual points-to should not assert
-\emph{which} locations are accessed in a page table walk.
-Figure \ref{fig:strongvirtualpointsto} gives such a definition.
+\emph{which} locations are accessed in a page table walk, as in Figure \ref{fig:strongvirtualpointsto}.
+% Figure \ref{fig:strongvirtualpointsto} gives such a definition.
 It asserts the existence of four page table entries, one at each translation level, and via \lstinline|L4_L1_PointsTo|
-asserts from the current \lstinline|cr3| value, taking the offsets from \textsf{va} (as in Figure \ref{fig:pagetables})
-to index each level, finding the existentially quantified entries in each level, 
-ending with the L1 entry, whose physical page base is added to the page offset of \textsf{va} to obtain the physical address
- --- and there (\lstinline|addr_L1(va,l1e)|) in physical memory, is the value.
+asserts that the physical page table walk (per Figure \ref{fig:pagetables}) succeeds in reaching the L1 entry,
+% asserts from the current \lstinline|cr3| value, taking the offsets from \textsf{va} (as in Figure \ref{fig:pagetables})
+% to index each level, finding the existentially quantified entries in each level, 
+% ending with the L1 entry, 
+% whose physical page base is added to the page offset of \textsf{va} to obtain the physical address
+ % --- and there (\lstinline|addr_L1(va,l1e)|) in physical memory, is the value.
+which points to the page holding the physical memory backing the virtual address, which contains value \textsf{v}.
 Most of the definition lives directly in \textsf{vProp}, using the separation logic structure lifted from Iris's \textsf{iProp}.
+\looseness=-1
 
@@
 to start the translation process.
+\looseness=-1
 

### 2024-11-12 syntaxsemantics.tex
compressing stuff up through (including) the strong virtual pointsto def

@@
 CPU.
-
+%
 % \subsection{Registers and Memory}
-Programs we demonstrate in this paper requires accessing two types of computer resource: registers and memory.
+% Programs we demonstrate in this paper requires accessing two types of computer resource: registers and memory.
 A register identifier, $\reg$, is chosen from a fixed finite set of register identifiers, $\regset$. 
-We use these identifiers $\reg$ for register names such as \texttt{rax}, \texttt{r8}, or \texttt{cr3}. Our model includes
-all x86-64 integer registers (including stack and instruction pointers), as well as \texttt{cr3} (for page table roots) and \texttt{rflags} (for
+We use these identifiers $\reg$ for register names such as \lstinline|rax|, \lstinline|r8|, or \lstinline|cr3|. Our model includes
+all x86-64 integer registers (including stack and instruction pointers), as well as \lstinline|cr3| (for page table roots) and \lstinline|rflags| (for
 flags set by comparison operations and inspected by conditional jumps).
@@
 memory maps $\sigma.\mathcal{M}: \Locft \rightharpoonup_{\textrm{fin}} (\Loctw \rightharpoonup_{\textrm{fin}} \Locsf )$.
-Of particular note, \texttt{cr3}, the page table register, is included in the machine state.
+Of particular note, \lstinline|cr3|, the page table register, is included in the machine state.
 % As one might have already anticipated from the syntax we introduce, we do not bind any value of an evaluated expression.
@@
 The most important instructions we model are memory-accessing variants of the x86-64 \lstinline|mov| instruction, which we format in Intel syntax
-(destination on the left, source on the right):
-\[
-\begin{array}{l}
-  \hbox{(\TirNameStyle{WriteToMemFromReg})} \qquad
-  \textsf{mov}~[r_m]~r_r \\
-  \hbox{(\TirNameStyle{WriteToRegFromMem})} \qquad 
-  \textsf{mov}~r_r~[r_m]
-  \end{array}
-  \]
-The semantics of each is realized by first translating the virtual-memory address of a memory location stored in $r_m$ to a physical memory location
-per the description in Section \ref{sec:backgroundonmachinemodel},
-then either updating those memory contents with the contents of register $r_r$ (\TirNameStyle{WriteToMemFromReg}) 
-or loading the value at that physical memory location into the register $r_r$
-(\TirNameStyle{WriteToRegFromMem}). 
-Additional \lstinline|mov| variants (e.g., accessing memory at constant offsets from the base register, or moves between registers) are also treated.
-Our formalization also includes basic integer and bitwise operations (\lstinline|add|,
+(destination on the left, source on the right).
+Thus a store to memory is $\textsf{mov}~[r_m]~r_r$, and a load from memory is $\textsf{mov}~r_r~[r_m]$.
+% \[
+% \begin{array}{l}
+%   \hbox{(\TirNameStyle{WriteToMemFromReg})} \qquad
+%   \textsf{mov}~[r_m]~r_r \\
+%   \hbox{(\TirNameStyle{WriteToRegFromMem})} \qquad 
+%   \textsf{mov}~r_r~[r_m]
+%   \end{array}
+%   \]
+Operationally, each first translates the virtual address stored in register $r_m$ to a physical address,
+then either updates the memory at that physical location with the contents of $r_r$ (for store) or
+or updates $r_r$ with the contents of that memory (for load).
+% The semantics of each is realized by first translating the virtual-memory address of a memory location stored in $r_m$ to a physical memory location
+% per the description in Section \ref{sec:backgroundonmachinemodel},
+% then either updating those memory contents with the contents of register $r_r$ (\TirNameStyle{WriteToMemFromReg}) 
+% or loading the value at that physical memory location into the register $r_r$
+% (\TirNameStyle{WriteToRegFromMem}). 
+Our formalization includes
+additional \lstinline|mov| variants (e.g., accessing memory at constant offsets from the base register, or moves between registers),
+basic integer and bitwise operations (\lstinline|add|,
 \lstinline|and|, bit shifts, etc.) with their effects on \lstinline|rflags|, jumps and some (not all) conditional jumps, \lstinline|call|, \lstinline|ret|, \lstinline|push|,
 and \lstinline|pop|.
+The semantics model updates to the instruction pointer, though we elide that register (\lstinline|rip|)
+from specifications in the paper for clarity.

### 2024-11-12 appendix.tex
appendix c

\definecolor{mGreen}{rgb}{0,0.6,0}
\definecolor{mGray}{rgb}{0.5,0.5,0.5}
\definecolor{mPurple}{rgb}{0.58,0,0.82}
\definecolor{backgroundColour}{rgb}{0.95,0.95,0.92}

\lstdefinestyle{CStyle}{
    backgroundcolor=\color{backgroundColour},   
    commentstyle=\color{mGreen},
    keywordstyle=\color{magenta},
    numberstyle=\tiny\color{mGray},
    stringstyle=\color{mPurple},
    basicstyle=\footnotesize,
    breakatwhitespace=false,         
    breaklines=true,                 
    captionpos=b,                    
    keepspaces=true,                 
    numbers=left,                    
    numbersep=5pt,                  
    showspaces=false,                
    showstringspaces=false,
    showtabs=false,                  
    tabsize=2,
    language=C
}

\section{C Implementation of \textsf{PTE} Library}
\label{sec:cimpl}

\subsection{Page-Table-Walk}
\label{sec:ctablewalk}

\begin{lstlisting}[style=CStyle]
void pte_initialize(pte_t *entry) {

    // Allocate a full page for 512 8-byte entries
    pte_t *local = kalloc();

    // Clear the entire table. Ensures present (and reserved, etc.) bits are 0
    //memset(local, 0, PAGE_SIZE);
    entry->pfn = PTE_ADDR_TO_PFN((uintptr_t) local);

    return;
}

/*
 * pte_get_next_table - given a page table entry, returns a pointer to the base
 * of the next level of page tables, allocating a new one if necessary and if
 * told to do so
 * @param entry - the entry from which to get the next level of tables
 * @param alloc - if non-zero, if the next table is not present, allocates it
 * @return - pointer to the next page table level or, if not present and alloc
 * is not set or memory is not sufficient, NULL
 */
 pte_t *pte_get_next_table_succ(pte_t *entry) {
    pte_t *next;

    // If not already present, try to allocate
    if (!entry->present) {
      // If it shouldn't be or cannot be allocated, indicate failure
      pte_initialize(entry);

      // Update the entry to indicate that a new table has now been allocated.
      // Make it writable and user-accesible - if necessary, the lowest level
      // of page table can fix these access rights. Indicate the *physical*
      // address of the newly allocated structure in pfn.
      entry->writable = 1;
      entry->user_acc = 1;
      // entry->pfn = PTE_ADDR_TO_PFN((uintptr_t) next);
      entry->present = 1;
    }
    uintptr_t next_phys_addr = PTE_PFN_TO_ADDR(entry->pfn);
    uintptr_t next_virt_addr = (uintptr_t) P2V(next_phys_addr);
    next = (pte_t *) next_virt_addr;

    return next;
}

pte_t *walkpgdir_succ(pte_t *pml4, const void *va) {
  pte_t *pdp;
  pte_t *pd;
  pte_t *pt;
  uintptr_t next_phys_addr;
  uintptr_t next_virt_addr;  
  pte_t *local;

  pte_t *pml4_entry = &pml4[PML4EX(va)];

    // If not already present, try to allocate
    if (!pml4_entry->present) {
      // If it shouldn't be or cannot be allocated, indicate failure
      // Allocate a full page for 512 8-byte entries
      local = kalloc();

      // Clear the entire table. Ensures present (and reserved, etc.) bits are 0
      //memset(local, 0, PAGE_SIZE);

      *(&pdp) = local;

      // Update the entry to indicate that a new table has now been allocated.
      // Make it writable and user-accesible - if necessary, the lowest level
      // of page table can fix these access rights. Indicate the *physical*
      // address of the newly allocated structure in pfn.
      pml4_entry->writable = 1;
      pml4_entry->user_acc = 1;
      pml4_entry->pfn = PTE_ADDR_TO_PFN((uintptr_t) pdp);
      pml4_entry->present = 1;
    }
    next_phys_addr = PTE_PFN_TO_ADDR(pml4_entry->pfn);
    next_virt_addr = (uintptr_t) P2V(next_phys_addr);
    pdp = (pte_t *) next_virt_addr;

    //= pte_get_next_table_succ(pml4_entry);

    pte_t *pdp_entry = &pdp[PDPEX(va)];

     // If not already present, try to allocate
    if (!pdp_entry->present) {
      // If it shouldn't be or cannot be allocated, indicate failure
      // Allocate a full page for 512 8-byte entries
      local = kalloc();

      // Clear the entire table. Ensures present (and reserved, etc.) bits are 0
      //memset(local, 0, PAGE_SIZE);

      *(&pd) = local;

      // Update the entry to indicate that a new table has now been allocated.
      // Make it writable and user-accesible - if necessary, the lowest level
      // of page table can fix these access rights. Indicate the *physical*
      // address of the newly allocated structure in pfn.
      pdp_entry->writable = 1;
      pdp_entry->user_acc = 1;
      pdp_entry->pfn = PTE_ADDR_TO_PFN((uintptr_t) pd);
      pdp_entry->present = 1;
    }
    next_phys_addr = PTE_PFN_TO_ADDR(pdp_entry->pfn);
    next_virt_addr = (uintptr_t) P2V(next_phys_addr);
    pd = (pte_t *) next_virt_addr;

    pte_t *pd_entry = &pd[PDEX(va)];
  // If not already present, try to allocate
    if (!pd_entry->present) {
      // If it shouldn't be or cannot be allocated, indicate failure
      // Allocate a full page for 512 8-byte entries
      local = kalloc();

      // Clear the entire table. Ensures present (and reserved, etc.) bits are 0
      //memset(local, 0, PAGE_SIZE);

      *(&pt) = local;

      // Update the entry to indicate that a new table has now been allocated.
      // Make it writable and user-accesible - if necessary, the lowest level
      // of page table can fix these access rights. Indicate the *physical*
      // address of the newly allocated structure in pfn.
      pd_entry->writable = 1;
      pd_entry->user_acc = 1;
      pd_entry->pfn = PTE_ADDR_TO_PFN((uintptr_t) pt);
      pd_entry->present = 1;
    }
    next_phys_addr = PTE_PFN_TO_ADDR(pd_entry->pfn);
    next_virt_addr = (uintptr_t) P2V(next_phys_addr);
    pt = (pte_t *) next_virt_addr;

    pt = pte_get_next_table_succ(pd_entry);

    return &pt[PTEX(va)];
}
\end{lstlisting}

### 2024-11-12 appendix.tex
appendix of c code

@@
 
-\begin{lstlisting}[style=CStyle]
+\begin{lstlisting}[caption={Page-Table-Entry Library for Page-Table Traversal},label={cpagetablewalk},style=CStyle]
 void pte_initialize(pte_t *entry) {
@@
 \end{lstlisting}
+
+\subsection{Mapping a Page}
+\label{sec:cmapapage}
+
+\begin{lstlisting}[caption={Mapping a Single Page},label={cmapapage},style=CStyle]
+complstatus_t vaspace_mappage(
+    pte_t *pml4,
+    void *va,
+    uintptr_t pa,
+    uint64_t perm
+) {
+    char *a;
+    pte_t *pte;
+
+    a = (char *)PGROUNDUP((uintptr_t)va);
+
+    pte = walkpgdir_succ(pml4, a);
+    if (!pte->present){
+      pte->writable     = PTE_WRITEABLE_SET(perm);
+      pte->user_acc     = PTE_USER_ACC_SET(perm);
+      pte->writethrough = PTE_WRITETHRU_SET(perm);
+      pte->nocache      = PTE_NOCACHE_SET(perm);
+      pte->pfn          = PTE_ADDR_TO_PFN(pa);
+      pte->nx           = PTE_NX_SET(perm);
+      pte->present      = 1;
+    }
+
+    return COMPL_SUCCESS;
+}
+\end{lstlisting}
+
+\subsection{Address-Space-Switch}
+\label{sec:caspaceswitch}
+
+\begin{lstlisting}[caption={Loading \textsf{cr3} for Address-Space Switch},label={caspace},style=CStyle]
+void pte_load(pte_t *pml4) {
+    lcr3(V2P(pml4));
+}
+\end{lstlisting}

### 2024-11-12 programlogic.tex
compressing...

@@
   is too strong to specify some operations a virtual memory manager may need to do, such as
-  move one level of the page table to a different physical location while preserving all virtual-to-physical mappings.\footnote{
-    x86-64 hardware, like other architectures, includes a feature (which we do not formalize assertions for) to
-    replace an L1 page table address in an L2 entry with a pointer to a \emph{larger} 2MB page (called super-pages), 
-    or replace an L2 page table address in an L3 entry with a pointer to a 1GB page (called huge-pages).
-  }
+  move one level of the page table to a different physical location while preserving all virtual-to-physical mappings. %\footnote{
+  %   x86-64 hardware, like other architectures, includes a feature (which we do not formalize assertions for) to
+  %   replace an L1 page table address in an L2 entry with a pointer to a \emph{larger} 2MB page (called super-pages), 
+  %   or replace an L2 page table address in an L3 entry with a pointer to a 1GB page (called huge-pages).
+  % }
   The use of $\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$ in Figure \ref{fig:strongvirtualpointsto}'s
@@
   so long as those updates preserve the same virtual-to-physical translations.
+  But this is not possible with Figure \ref{fig:strongvirtualpointsto}'s definition, which ties ownership of particular
+  pieces of page table memory to the virtual points-to.
 
@@
 \footnotesize
-\[
+\centerline{$
 \begin{array}{l}
@@
 \end{array}
-\]
+$}
 \vspace{-1em}
@@
 
-  Intuitively, the definition in Figure \ref{fig:strongvirtualpointsto} is too strong because the virtual points-to
-  assertion there tracks too much information: when writing programs that access memory via virtual addresses,
-  most code does not care \emph{which physical memory locations are involved in address translation}: it only cares
-  that virtual address translation would succeed. The necessary information about the physical page table walk
-  must still be tracked, but can be tracked separately from the virtual points-to assertion itself.
-  In practice the decisions about which virtual addresses are valid rest not with code posessing a virtual address, but with
-  the virtual memory manager --- and its invariants.
+  % Intuitively, the definition in Figure \ref{fig:strongvirtualpointsto} is too strong because the virtual points-to
+  % assertion there tracks too much information: when writing programs that access memory via virtual addresses,
+  % most code does not care \emph{which physical memory locations are involved in address translation}: it only cares
+  % that virtual address translation would succeed. The necessary information about the physical page table walk
+  % must still be tracked, but can be tracked separately from the virtual points-to assertion itself.
+  % In practice the decisions about which virtual addresses are valid rest not with code posessing a virtual address, but with
+  % the virtual memory manager --- and its invariants.
 
@@
   named $\gamma$. The existence of such a token implies that the actual map $\theta$ tracked by a corresponding $\mathsf{GhostMap}(\gamma,\theta)$
-  resource indeed maps $k$ to $v$. These properties are captured by the Iris rules:\footnote{Iris ghost maps lack established notation in the literature,
+  resource indeed maps $k$ to $v$.
+These properties are captured by Iris rules using the token to update or draw conclusions about the logical map:\footnote{Iris ghost maps lack established notation in the literature,
   but the syntax we use captures the details of \texttt{iris.base\_logic.lib.ghost\_map}.}
-
   \begin{mathpar}
@@
   \footnotesize
-\[
+\centerline{$
 \begin{array}{l}
@@
     \ptableabswalk{\delta,\theta}
-
 \end{array}
-\]
+$}
 \vspace{-1em}
@@
   \label{fig:peraspaceinvariant}
+\vspace{-1em}
   \end{figure*}
 
-Because our logic is the first to address the existence of multiple address spaces,
-the ghost map of virtual-to-physical translations (identified by the ghost name $\delta$ in Figure \ref{fig:virtualpointstosharing}
- \emph{for each address space}) must be locateable from a representation of the specific address space:
+Our logic is the first to address the existence of multiple address spaces, so requires
+a way to look up
+the ghost map of virtual-to-physical translations for an address space (the ghost name $\delta$ in Figure \ref{fig:virtualpointstosharing}).
+% for each address space must be locateable from a representation of the specific address space:
 % This means we also require a way to locate the map for the current address space from 
-the current root page table pointer stored in \lstinline|cr3|.
+We support lookup by the current root page table pointer stored in \lstinline|cr3|.
+\looseness=-1
 
@@
 and doing so requires a way to talk about other address spaces, and means to switch address spaces.
-% requiring a means to  which we abstract as a ghost map from root address of an address-space to a namespace labelling the address space
-
-% \paragraph{Valid Context-Resources} Fragmental ownership of $\mathcal{A}\textsf{spaces}$ ($\sumapaces\rtv\gammaPred$) conveys the fact that a current address-space ($\gammaPred$) is the one selected, i.e.
-loaded into the $\rtv$.
-This simple fact becomes essential when we observe it inside virtual-pointsto definition in Figure \ref{fig:virtualpointstosharing} and pictorially Figure \ref{logicaladdrspace}.
-Existence of this fragmental ownership ensures that the contextual resource, i.e.
-a virtual pointsto assertion, resides under the correct address-space, i.e.
-modal-context, that makes it valid.
-
-% \paragraph{Modal Resource Context\&Interaction with Ambient Logic \textsf{vProp}}
-% \label{sec:resourcecontext}
-% The relation between the address-space and and the validity of virtual-pointsto assertions exhibits itself as a contingency, i.e.
-\textit{"a certain virtual-pointsto facts whose validity is indexed by the root address of an address space"}.
-But, what sort of convenience this pattern provides when we do reasoning about more than one address-space.
-
-% In Figure \ref{fig:addrswitch}, we see two different address-spaces rooted at R0 and R1 in the pre/post condition.
-To be put into words, while switching an address-space by executing \lstinline|mov_ctl|) updating the  \lstinline|cr3|, it is expected to mention resources of two different address-spaces in the specification.
-In the precondition,  \lstinline|cr3| is loaded with R0 and all the virtual-pointsto assertions $a_0 \mapsto_{\textsf{v}}p_0$ and $a_1 \mapsto_{\textsf{v}}p_1$ are in the current-view of the memory and valid in the ambient logic \textsf{vProp}.
-The other address-space with the root address R1 has its own resources ($b_0 \mapsto_{\textsf{v}}p_3$ and $b_1 \mapsto_{\textsf{v}}p_4$) which are invalid under the current address space rooted at R0.
-This pattern of specification requires resources being mentioned under the \textit{world} they are valid, i.e.
-a modal-framing within the specification.
-Once the current address-space becomes the one with root address R1, i.e.
-\lstinline|cr3| is loaded with R1, the context-resources $b_0 \mapsto_{\textsf{v}}p_3$ and $b_1 \mapsto_{\textsf{v}}p_4$ are loaded into the current-view of the memory, and all the other resources that are valid under R0 are introduced into the modal context rooted at R0
 \begin{figure}
@@
   \label{fig:modaldef}
+\vspace{-1em}
   \end{figure}
@@
 % what we need in order for assertions in one address space to talk about assertions in another.
-\looseness=-1
-
 % The inspiration for this assertion comes from hybrid logic~\cite{blackburn1995hybrid,areces2001hybrid,goranko1996hierarchies,gargov1993modal},
@@
 % assertions, we inherit them by lifting the model of our ambient logic, Iris.
-
 We can prove that this modality follows certain basic laws: that its truth is independent of the address
@@
 \textsf{Fact}s can move in and out of other-space modalities freely.
+\looseness=-1
 
@@
 \subsection{Subtleties of Changing Address Spaces Using Modalities}
+\todo[inline]{Colin will push this to an appendix, and leave just a short comment of a few lines here with a forward ref}
 % \subsection{Considering Hoare Doubles in the Context of Switching Address-Space}

### 2024-11-12 programlogic.tex
compression, clarifying relation to syeda et al

@@
 \label{sec:aspacemanagement}
-So far, we have introduced logical abstractions for a single address space, but VMMs
- handle more than one address-space,
-and doing so requires a way to talk about other address spaces, and means to switch address spaces.
+% So far, we have introduced logical abstractions for a single address space, but VMMs
+Real VMMs must
+ handle more than one address-space.
+Doing so requires a way to talk about other address spaces, and means to switch address spaces.
 \begin{figure}
@@
 
-\subsection{Subtleties of Changing Address Spaces Using Modalities}
-\todo[inline]{Colin will push this to an appendix, and leave just a short comment of a few lines here with a forward ref}
-% \subsection{Considering Hoare Doubles in the Context of Switching Address-Space}
-% \subsubsection{Issues with Using Hoare Triples in Address-Space Specification}
-\label{sec:issues}
-Operating systems on x86-64 change address spaces by updating \lstinline|cr3|.
-Conceptually, we expect a proof rule for such an update to take move resources in and out of modalities for
-the old and new address spaces. This suggests a rule like the following, which captures the intuition of the
-address space change, but is subtly unsound under common assumptions.
-% An important subtlety arises with supporting \lstinline|mov|s into \lstinline|cr3|. Consider the hypothetical rule
-\begin{mathpar}
-\inferrule*[left=Broken]{ }{
-  \{P \ast \textsf{cr3}\mapsto_{r} \;r_1 \ast r \mapsto_{r} \;r_2 \ast [r_2](Q)\}
-  \texttt{mov}~\texttt{\%cr3},~r%\lstinline|mov %cr3, r| 
-  \{[r_1](P) \ast \textsf{cr3} \mapsto_{r} r_2 \ast r \mapsto_{r} r_2 \ast Q\}
-}
-\end{mathpar}
-This rule takes all assertions not under a modality ($P$), and moves them under a modality for the old address space
-(which is where they were valid), while pulling assertions for the new, target address space out of a corresponding modality.
-(For clarity, we also omit the address space invariants, which must also be present.)
-% which is slightly simplified (by dropping per-address-space resources) for clarity in making the following point.
-% This rule captures the intuitive change of address space in a Hoare triple.
-Unfortunately, as stated, this interacts quite poorly with the traditional frame
-rule.
-If we abbreviate the pre- and post-conditions above as \textsf{Pre} and \textsf{Post}:
-\begin{mathpar}
-  \inferrule*[right=Frame]{
-    \inferrule*[right=Broken]{ }{
-    \{\textsf{Pre}\}
-    \texttt{mov}~\texttt{\%cr3},~r%\lstinline|mov %cr3, r| 
-    \{\textsf{Post}\}
-    }
-  }{
-    \{a\mapsto_\mathsf{v} x \ast \textsf{Pre}\}
-    \texttt{mov}~\texttt{\%cr3},~r%\lstinline|mov %cr3, r| 
-    \{a\mapsto_\mathsf{v} x \ast \textsf{Post}\}
-  }
-\end{mathpar}
-Notice that both the precondition and postcondition assert that $a\mapsto_\mathsf{v} x$ in the current address space, but we have no basis for concluding that address translation is preserved by the change of address space.
-So this derivation clearly leads to an unsound conclusion.
-The essential problem is that the frame rule is motivated by local reasoning about local updates, but
-a switch of address space is a \emph{global} change that may invalidate information about virtual addresses.
-Thus framing around \lstinline|cr3| updates is unsound, which has important impacts on our formal development.
-
-Switching to Hoare doubles resolves this problem because an under-appreciated subtlety of Hoare doubles 
-is that typically \emph{there is no primitive frame rule}. 
-Instead each verification essentially includes a local frame that it passes to the next instructions
-(think continuation-passing style), giving each overall rule a \emph{global} (rather than local) precondition.
-For most rules this is not that important, but it does permit rules that have global effects on their preconditions.
-
-% This is then how we justify our actual rule for \lstinline|cr3| updates, which is \emph{almost}
-This leads to our actual rule for \lstinline|cr3| updates, which is \emph{almost}
-\begin{mathpar}
-\inferrule*[left=ChangeAddressSpace]{
-  \{[r_1](P) \ast cr3 \mapsto_{r} r_2 \ast r \mapsto_{r} r_2 \ast Q\}\;\overline{is}
-}{
-  \{P \ast cr3 \mapsto_{r} r_1 \ast r \mapsto_{r} r_2 \ast [r_2](Q)\}\;
-  \texttt{mov}~\texttt{\%cr3},~r;\;\overline{is}
-  %\lstinline|mov %cr3, r| 
-}
-\end{mathpar}
-(Again, the actual rule in Section \ref{sec:selected_rules} includes the address space invariants.)
-Because the precondition on this rule is global, we avoid issues with framing. 
-A more typical frame rule can be recovered from Hoare doubles in the typical way~\cite{Chlipala2013Bedrock} 
-with the added restriction that \lstinline|cr3| is restored to its pre-framing value before un-framing (in this case, the framed-out assertion is effectively
-stashed under an other-space assertion for the initial address space, and pulled out as part of restoring).
-
-% If we wanted to consider a frame rule that would work for this logic, we could consider:
-% \begin{mathpar}
-%   \inferrule*[left=Cr3Frame]{
-%     \{P\ast cr3=v\}\;C\;\{ Q \ast cr3=v\}
-%   }{
-%     \{R\ast P\ast cr3=v\}\;C\;\{ R \ast Q \ast cr3=v\}
-%   }
-% \end{mathpar}
-% By demanding that \lstinline|cr3| be held constant (or rather, at least restored to its original value) we could frame almost traditionally.
-In particular, this rule would work with framing around calls that might lead to address space switches, such as calling blocking operations in the kernel.
-% Attempting to derive a frame rule as some work using Hoare doubles does~\cite{Chlipala2013Bedrock}
-% would enforce this restriction.
-
-% Another useful perspective on this might be the dynamic frames~\cite{parkinson2011relationship} 
-% view: virtual points-to assertions are self-stable \emph{except} for changes in \lstinline|cr3|, so framing would then naturally require other means of holding \lstinline|cr3| constant (or saving and restoring it).
-% Virtual points-to assertions could be made self-stable by also giving them partial ownership over \lstinline|cr3| assertions, but this would require explicitly plumbing that ownership from \emph{all} assertions back to any place an address space change might occur; this would seem to be a far graver loss of modularity than this extra quirk in framing discussions.
-
 \subsection{Selected Logical Rules}
 \label{sec:selected_rules}
-Per the discussion in Section \ref{sec:issues}, we define our logic using Hoare doubles:\footnote{This
+% Per the discussion in Section \ref{sec:issues}, w
+As common for assembly-level verification, we define our logic using Hoare doubles:\footnote{This
 omits some low-level Iris details (stuckness, observations) that play no meaningful
@@
 Making \textsf{rtv} an explicit parameter to the double, rather than simply using register assertions
-directly solves a technical problem with ensuring that the page table root used to evaluate
-the \textsf{vProp} (i.e., evaluating the assertion in the \emph{current}) address space
-is feasible.\footnote{Consider the difficulty of selecting the correct page table root value from an arbitrary
-opaque $\Phi$, which may even existentially quantify the page table root. An alternative is to
-require $\Phi$ to have a syntactic form where we can directly extract the value of \lstinline|cr3|,
-but this makes using Iris Proof Mode (IPM)~\cite{Krebbers:2017:IPH:3009837.3009855} with \textsf{vProps}
-  difficult; IPM works for any type matching the signature of an Iris \lstinline|bi|, which includes
-  \textsf{vProp}s, but manually guiding IPM to put an assertion in a specific position over and over adds
-  significant proof burden.
-}
+directly 
+makes it possible to ensure ownership of the \lstinline|cr3| register and its value is accounted for
+while avoiding some technical headaches with trying to enforce that $\Phi$ itself contains that.
+% solves a technical problem with ensuring that the page table root used to evaluate
+% the \textsf{vProp} (i.e., evaluating the assertion in the \emph{current}) address space
+% is feasible.
+% \footnote{Consider the difficulty of selecting the correct page table root value from an arbitrary
+% opaque $\Phi$, which may even existentially quantify the page table root. An alternative is to
+% require $\Phi$ to have a syntactic form where we can directly extract the value of \lstinline|cr3|,
+% but this makes using Iris Proof Mode (IPM)~\cite{Krebbers:2017:IPH:3009837.3009855} with \textsf{vProps}
+%   difficult; IPM works for any type matching the signature of an Iris \lstinline|bi|, which includes
+%   \textsf{vProp}s, but manually guiding IPM to put an assertion in a specific position over and over adds
+%   significant proof burden.
+% }
 
-The rest of this section describes specifications of some selected \textsf{AMD64} instructions 
+The rest of this section describes specifications of three key \textsf{AMD64} instructions 
 in our logic. 
@@
 \end{mathpar}
+\vspace{-1em}
 \caption{Reasoning Rules for Selected \textsf{AMD64} Instructions under}
@@
 the corresponding modality.
+Note that the \emph{global} aspect is important, as a na\"ive frame rule would be unsound for updates to \lstinline|cr3|,
+because one could frame out assertions in one address space, switch address spaces, and bring those assertions from the \emph{old}
+address space back into the \emph{new} address space, where they may not hold. Appendix \ref{sec:issues} gives more details.
 % From that rule, with one step outside the \textsf{vProp} logic (so it is not quite a derived rule within

### 2024-11-12 relwork.tex
compression, clarifying relation to syeda et al

@@
 of virtual addresses are mapped implies false in their logic.
-(They do formulate their logic over an abstract model, but every architecture would incur a similar limitation;
-Na\"ively transferring their model to x86-64 4-level tables would yield a limit of 512 assertions (also a 4Kb root page, 
-but 8-byte entries).
+(They do formulate their logic over an abstract model, but every architecture would incur a similar limitation.)
+% Na\"ively transferring their model to x86-64 4-level tables would yield a limit of 512 assertions (also a 4Kb root page, 
+% but 8-byte entries).
 Our definitions make use of fractional permissions throughout; Figure \ref{fig:strongvirtualpointsto}'s definition
@@
 the L1 entry, 1/($512^2$) of the L2 entry, and so on, so each entry may map the appropriate number of machine words.
+\looseness=-1
 
-As noted earlier, Kolanski and Klein's logics, by collocating both the physical ownership of the page table walk
-as part of the virtual points-to itself, preempt support for changes to page tables which do not actually affect 
+As noted earlier, by collocating both the physical ownership of the page table walk
+as part of the virtual points-to itself these logics preempt support for changes to page tables which do not actually affect 
 address translation. We address this by moving from Figure \ref{fig:strongvirtualpointsto}'s definition (essentially Kolanski and Klein's
 with fractional sharing of intermediate table entries) to Figure \ref{fig:virtualpointstosharing}'s.
+\looseness=-1
 
@@
 page table walk for another virtual address.
+\looseness=-1
 
@@
 spaces while reasoning about correctness.
+The only other attempt to verify a change of address space was by \citet{syeda2018program,syeda2020formal}, whose focus was on TLB maintenance (which we
+do not model). However, they verified only the 4 instructions to switch address spaces and update the TLB on an ARM processor, in isolation.
+The specification they proved for those instructions did not address program invariants that may be valid in one address space and not the other,
+so is not flexible enough to extend directly to a full context switching primitive as in Figure \ref{fig:swtch}.
+Of course as we do not address TLB management (see Appendix \ref{apdx:tlb}), the right eventual solution would be to combine our work
+and theirs, which we leave to future work.
 
@@
 
-As noted earlier, our other-space modality comes from ideas in hybrid logic~\cite{areces2001hybrid,blackburn1995hybrid,gargov1993modal,goranko1996hierarchies},
+As noted earlier, our other-space modality derives from hybrid logic~\cite{areces2001hybrid,blackburn1995hybrid,gargov1993modal,goranko1996hierarchies},
 where modalities are indexed by \emph{nominals} which are names for specific individual states in a Kripke model.
-We are aware of only two prior works combining these ideas with program logics, one showing that traditional
-nominals extends the expressive power of separation logic~\cite{brotherston2014parametric}, another
-using nominals to refer to states of other actors in an actor language~\cite{gordon2019modal}.
-
+Little prior work combines these ideas with program logics: \citet{brotherston2014parametric} show that traditional
+nominals extends the expressive power of separation logic, and \citet{gordon2019modal}
+uses nominals to refer to states of other actors in an actor language.
 % As noted throughout the paper, the inspiration for our other-space modality comes from hybrid logic~\cite{areces2001hybrid,blackburn1995hybrid,gargov1993modal,goranko1996hierarchies},
@@
 Beyond these, there is limited work on the interaction of hybrid logic with substructural logics, in restricted forms
-that do not affect expressivity~\cite{despeyroux2014hybrid,chaudhuri2019hybrid}.
+that do not affect expressivity~\cite{despeyroux2014hybrid,chaudhuri2019hybrid}.\looseness=-1
 % Primarily there is a line of work on hybrid linear logic (\HyLL)~\cite{despeyroux2014hybrid}, 

### 2024-11-13 examples.tex
compression, clarifying relation to syeda et al
more edits, adjusting presentation of some code figures, and removing some redundant text now has the experiments nearly fitting in 20 pages, but then there is related work after that still

--- 2024-08-06 examples.tex
@@
 \label{sec:experiment}
-To both validate and demonstrate the value of the modal approach to reasoning about virtual memory management, we study several
-distillations of real concerns of virtual memory managers.
-Recall from Section \ref{sec:logic} that virtual points-to assertions work just like regular points-to assertions, by design.
-In this section we work through two critical and challenging aspects of virtual memory management.
+%To both validate and demonstrate the value of the modal approach to reasoning about virtual memory management, 
+% we study several
+% We validate our logic by studying
+% distillations of key VMM functionality.
+% real concerns of virtual memory managers.
+% Recall from Section \ref{sec:logic} that virtual points-to assertions work just like regular points-to assertions, by design.
+In this section we verify several critical and challenging pieces of VMM code.
 First, in several stages, we work up to mapping a new page in the current address space.
-This requires a number of challenging substeps: dynamically traversing a page table to find
+This requires a number of independently challenging substeps: dynamically traversing a page table to find
 the appropriate L1 entry to update; inserting additional levels of the page table if necessary (updating
@@
 current address space) has previously been formally verified with respect to a machine model with address translation.
-Second, we formally verify a switch into a new address space as part of a task switch.\todo{Colin will double-check the TLB work's details before claiming this is the first}
+Second, we formally verify a switch into a new address space as part of a task switch,
+the first such verification handling both old and new processes' assertions at the time of the switch.
 
@@
 revisit several times: \lstinline|pte_get_next_table|'s specification
-assumes it is given a virtual \emph{vpte-pointsto}\footnote{A \emph{vpte-pointsto}
-$\vaddr\mapsto_{\textsf{vpte,qfrac}} \textsf{\paddr~\val}$ is a virtual points-to granting
-access to virtual address $\vaddr$, which is known to hold \textsf{\val}, but additionally exposing the physical address
-$\paddr$ that the virtual address translates to. This is commonly used, as the name suggests, when updating PTEs,
-where we require assurance that we are accessing the intended PTE.
-Its definition is just like the regular virtual points-to in Figure \ref{fig:virtualpointstosharing},
-except taking the physical address as a parameter rather than existentially quantifying it.
-}
- granting access to the specified entry,
+assumes it is given a virtual \emph{vpte-pointsto}
+% \footnote{A \emph{vpte-pointsto}
+% $\vaddr\mapsto_{\textsf{vpte,qfrac}} \textsf{\paddr~\val}$ is a virtual points-to granting
+% access to virtual address $\vaddr$, which is known to hold \textsf{\val}, but additionally exposing the physical address
+% $\paddr$ that the virtual address translates to. This is commonly used, as the name suggests, when updating PTEs,
+% where we require assurance that we are accessing the intended PTE.
+% Its definition is just like the regular virtual points-to in Figure \ref{fig:virtualpointstosharing},
+% except taking the physical address as a parameter rather than existentially quantifying it.
+% }
+(a virtual points-to exposing the underlying physical address; see Section \ref{sec:mapnew}) granting access to the specified entry,
 but its postcondition does not yield new virtual points-to assertions!
@@
 \begin{figure}%\footnotesize
+\ifPLDI
+\begin{lstlisting}[mathescape, basicstyle=\footnotesize,escapeinside={(*}{*)}]
+;;pte_t *pte_get_next_table(pte_t *entry) { ... // setting up the stack for a pte_t *next elided
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{entry}\},m)  \ast \texttt{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast  \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{next} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{\_} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{\_} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
+$\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present entry\_val})\urcorner}_{\rtv}$ (*\label{line:get_next_vpte_precondition}*)
+$\specline{\ulcorner\textsf{entry\_present(entry\_val)}\urcorner \wand \forall_{i\in\textsf{0..511}} \; \ghostmaptoken{\textsf{id}}{\textsf{table\_root(entry\_val.pfn) + i * 8}}{\textsf{v-1}} }_{\rtv} $ (*\label{line:conditional_children}*)
+mov    -0x8[rbp],rdi $\specline{\ldots \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry} \ast \ldots }$
+mov     rdi, r8 $\specline{\ldots \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry} \ast \ldots }$
+mov    [r8],rdi (*\label{line:read_entry_contents}*) $\specline{\ldots \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val}  \ast\ldots}$
+and    0x1,rdi (* \label{line:mask_present} *)
+mov    rdi,rax
+cmp    0x0,rax ;;  if (!entry->present) {(* \label{line:check_entry_present} *)
+jne    161 <pte_get_next_table+0xa1> (* \label{line:check_entry_present_jump} *) ;; Jump if the present bit is not zero
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry} \ast \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{next} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val \& 0x1} }_{\rtv}$
+$\specline{  \textsf{entry} \mapsto_{\textsf{id}} \textsf{v} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  \ast \texttt{rax}  \mapsto_{\textsf{r}} \textsf{entry\_val \& 0x1} \ast \ulcorner\textsf{entry\_val \& 0x1}=\textsf{0x0}\urcorner}_{\rtv}$
+$\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow  \lnot(\textsf{entry\_val \& 0x1}=\textsf{0x1})  \urcorner }_{\rtv}$ (*\label{line:before_concluding_qfrac1}*)
+mov    rbp,rdi (*\label{line:alloc_path_start}*)
+sub    0x10, rdi ;; Store the value of rbp minus 16 bytes (address of (*\textsf{next}*)) into rdi (*\label{line:pass_addrof_next}*)
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry} \ast \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{next} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{rbp - 16} }_{\rtv}$
+$\specline{  \textsf{entry} \mapsto_{\textsf{id}} \textsf{v} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta} }_{\rtv}$
+$\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \land \lnot(\textsf{entry\_present entry\_val})  \urcorner }_{\rtv}$ (*\label{line:after_concluding_qfrac1}*)
+callq  70 <pte_initialize> ;;pte_initialize(entry);(* \label{line:call_to_pte_initialize} *)
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  }_{\rtv}$
+$\specline{  \textsf{entry} \mapsto_{\textsf{id}} \textsf{v} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta} }_{\rtv}$
+$\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \land \lnot(\textsf{entry\_present entry\_val})  \urcorner }_{\rtv}$
+$\specline{\ulcorner  \textsf{qfrac} = 1 \land \lnot(\textsf{entry\_present (pfn\_set (entry\_val nextpaddr))}  \urcorner   }_{\rtv}$
+$\specline{ \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{pfn\_set(entry\_val nextpaddr)}   }_{\rtv}$
+$\specline{ \begin{array}{l}\textsf{entry\_present (pte\_initialized (pfn\_set(entryv nextpaddr)))} \wand  \\ \;\;\;\;\;\;\;\;\  \forall_{i\in \textsf{0 ...
+511} } \ldotp  \ghostmaptoken{\textsf{id}}{((\textsf{table\_root (pte\_initialized (pfn\_set(entry\_val nextpaddr)))}) + \textsf{i * 8})}{\textsf{v-1}} \end{array}  }$ (*\label{line:page_of_caps}*)
+... ;;entry value updates: entry->pfn = nextpaddr; entry->present = 1; (*\label{line:install_new_entry}*)
+... ;;now we know that entry is initialized, so we satisfy the condition to access children list
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  }_{\rtv}$
+$\specline{  \textsf{entry} \mapsto_{\textsf{id}} \textsf{v} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta} }_{\rtv}$
+$\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}}  \textsf{pte\_initialized(pfn\_set(entryv nextpaddr))} \ast \ulcorner  \textsf{qfrac} = 1 \land \lnot(\textsf{entry\_present entry\_val})  \urcorner }_{\rtv}$
+$\specline{ \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{pte\_initialized(pfn\_set(entry\_val nextpaddr))}  \ast rax \mapsto_{r} \textsf{table\_root (pte\_initialized (pfn\_set(entry\_val nextpaddr)))}  }_{\rtv}$
+$\specline{\forall_{i\in \textsf{0 ...
+511} } \ldotp  \ghostmaptoken{\textsf{id}}{((\textsf{table\_root (pte\_initialized (pfn\_set(entry\_val nextpaddr)))}) + \textsf{i * 8})}{\textsf{v-1}}  }$ (*\label{line:alloc_path_end}*)
+;;} (*\label{line:end_of_allocation_path}*)
+... ;; Code after conditional continued in Figure (*\ref{fig:p2v}*)
+\end{lstlisting}
+\else
 \begin{lstlisting}[mathescape, basicstyle=\footnotesize,escapeinside={(*}{*)}]
@@
 \end{lstlisting}
+\fi
 % ;;uintptr_t next_virt_addr = (uintptr_t) P2V(next_phys_addr);
@@
 ;;entry->pfn = PTE_ADDR_TO_PFN((uintptr_t) local);
-mov    -0x10[rbp],rax
-mov    -0x8[rbp],rdi
-mov    rax,[rdi]
+mov    -0x10[rbp],rax; mov    -0x8[rbp],rdi; mov    rax,[rdi]
 $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)   \ast \texttt{rbp-16} \mapsto_{\textsf{r}} \textsf{\_} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry+KERNBASE} }_{\rtv}$
@@
 \subsection{Mapping a New Page}
+\label{sec:mapnew}
 One of the key tasks of a page fault handler in a general-purpose OS kernel is
@@
 $\specline{\mathsf{ContextAt}(old,\_) \ast \mathsf{ContextAt}(new,[\mathsf{rbxv}',\ldots,\rtv']) }_{\rtv}$
-mov 0[rdi], rbx
-... ;; mov rsp, rbp, r12, r13, r14, saved to offsets 8, 16, 24, 32, and 40 from rdi
+mov 0[rdi], rbx ... ;; also mov rsp, rbp, r12, r13, r14, saved to offsets 8, 16, 24, 32, and 40 from rdi
 mov 48[rdi], r15
@@
 $\specline{\mathsf{ContextAt}(old,[\mathsf{rbxv,\ldots,\rtv}]) \ast \mathsf{ContextAt}(new,[\mathsf{rbxv}',\ldots,\rtv']) }_{\rtv}$
-mov 56[%rdi], %cr3
-$\specline{\ldots \ast \texttt{rdi+56} \mapsto_{\textsf{v}} \rtv}_{\rtv}$    
+mov 56[%rdi], %cr3 $\specline{\ldots \ast \texttt{rdi+56} \mapsto_{\textsf{v}} \rtv \ast \ldots }_{\rtv}$    
 ;; Restore target context
@@
 $\specline{\mathsf{ContextAt}(old,[\mathsf{rbxv,\ldots,\rtv}]) \ast \mathsf{ContextAt}(new,[\mathsf{rbxv}',\ldots,\rtv']) }_{\rtv}$
-;; Switch to the new address space
-mov cr3, 56[rsi]
+mov cr3, 56[rsi] ;; <-- Switch to the new address space
 $\specline{ [\rtv](\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,\Xi,m) \ast \mathsf{ContextAt}(old,[\mathsf{rbxv,\ldots,\rtv}]) \ast \mathsf{ContextAt}(new,[\mathsf{rbxv}',\ldots,\rtv']))  }_{\rtv}$

### 2024-11-13 programlogic.tex
more edits, adjusting presentation of some code figures, and removing some redundant text now has the experiments nearly fitting in 20 pages, but then there is related work after that still

@@
 For clarity we refer to the ghost map summarizing virtual-to-physical translations by 
-\[\mathcal{A}\textsf{bsPTableWalk}(\delta,\theta) \stackrel{\triangle}{=} \mathsf{GhostMap}(\delta,\theta)\]
+\centerline{$\mathcal{A}\textsf{bsPTableWalk}(\delta,\theta) \stackrel{\triangle}{=} \mathsf{GhostMap}(\delta,\theta)$}\\
 (omitting $\delta$ for brevity when only one is in scope)
@@
 
-Our logic is the first to address the existence of multiple address spaces, so requires
+Our logic is the first to address the existence of multiple address spaces, so also requires
 a way to look up
@@
 
-\paragraph{From A Single Address Space to Many}
-In principle the resources from each individual address space should be collected into a single
-shared invariant, and for each memory access,
-a fragment of this corresponding to $\mathcal{I}\textsf{ASpace}(\Theta,m)$
-for the current address space should be extracted from this global resource, used to prove correctness of the
-memory access, and put back.
-In this paper we focus on the middle section, explicity identifying resources for each
-address space. The reason for this is that in practice, this global resource is also
-the place where kernel-specific assumptions, such as guaranteeing that certain virtual address range
-was mapped in all address spaces, would be enforced. This paper focuses on the
-reasoning principles behind the virtual memory access, and we leave the use of this within a larger
-kernel to future work. We do, however, deal with the existence of multiple address spaces in examples (Section \ref{sec:experiment}).
+% \paragraph{From A Single Address Space to Many}
+% In principle the resources from each individual address space should be collected into a single
+% shared invariant, and for each memory access,
+% a fragment of this corresponding to $\mathcal{I}\textsf{ASpace}(\Theta,m)$
+% for the current address space should be extracted from this global resource, used to prove correctness of the
+% memory access, and put back.
+% In this paper we focus on the middle section, explicity identifying resources for each
+% address space. The reason for this is that in practice, this global resource is also
+% the place where kernel-specific assumptions, such as guaranteeing that certain virtual address range
+% was mapped in all address spaces, would be enforced. This paper focuses on the
+% reasoning principles behind the virtual memory access, and we leave the use of this within a larger
+% kernel to future work. We do, however, deal with the existence of multiple address spaces in examples (Section \ref{sec:experiment}).
 
@@
 
+In general, per-address-space invariants should be collected in a larger
+VMM invariant, with individual address spaces' invariants pulled out as needed, such as when proving
+soundness of an individual virtual memory access.
+But such larger invariants would contain many kernel-specific properties that are orthogonal
+to the fundamental reasoning principles that are the focus of this paper.
+We leave such kernel-specific reasoning to future work, but our verification of task switching
+(Section \ref{sec:experiment}) demonstrates support for managing multiple address spaces.
+\looseness=-1
+
 % \begin{figure}
@@
 % Per the discussion in Section \ref{sec:issues}, w
-As common for assembly-level verification, we define our logic using Hoare doubles:\footnote{This
-omits some low-level Iris details (stuckness, observations) that play no meaningful
-role in our development.}
+As common for assembly-level verification, we define our logic using Hoare doubles:%\footnote{This
+% omits some low-level Iris details (stuckness, observations) that play no meaningful
+% role in our development.}
   \[
@@
 can be found in our artifact.
-Each rule in Figure \ref{fig:wpdamd}, 
-is annotated with a root (i.e., \lstinline|cr3|) address value (\textsf{rtv}), 
-under which the resources mentioned in the specification are valid.
+% Each rule in Figure \ref{fig:wpdamd}, 
+% is annotated with a root (i.e., \lstinline|cr3|) address value (\textsf{rtv}), 
+% under which the resources mentioned in the specification are valid.
 In general, we use metavariables $\textsf{r}_s$ and $\textsf{r}_d$ to specify source and destination registers
@@
 \vspace{-1em}
-\caption{Reasoning Rules for Selected \textsf{AMD64} Instructions under}
+\caption{Proof rules for selected \textsf{AMD64} instructions}
 \label{fig:wpdamd}
+\vspace{-1em}
 \end{figure}
@@
 implementing a fragment of x86-64 including 64-bit address translation with 4-level page tables.
-\todo[inline]{
-Our rules for control transfers (\lstinline|jmp|, \lstinline|call|, and \lstinline|ret|) are currently
-axiomatized (with completely standard specifications~\cite{ni2007contexts,Chlipala2013Bedrock}) because Iris's built-in machinery does not provide
-convenient ways to discard the current continuation. We are working to adapt existing machinery~\cite{de2023type} to prove their soundness.
-Our soundness proofs for all other instructions (including the instructions above, arithmetic, other heap-accessing operations, push, and pop)
+% \todo[inline]{
+Our rules for control transfers (\lstinline|jne|, \lstinline|call|, and \lstinline|ret|) are currently
+axiomatized (with completely standard specifications~\cite{ni2007contexts,Chlipala2013Bedrock}) 
+because Iris's built-in machinery does not provide
+convenient ways to discard the current continuation; adaptation of others'
+approaches~\cite{de2023type} is ongoing work.
+Our soundness proofs for all other instructions (including, critically, all memory accesses)
 are axiom-free.
-}
+% }
 % At the moment our proofs do rely on 13 small axioms of properties which should be provable, but
 % are challenging to discharge due to some representation choices in our model.\footnote{See \lstinline|srx/x64/machine/current_axioms.v|}
+\looseness=-1

### 2024-11-13 relwork.tex
took an axe to related work, and now we are down to 22.5 pages overall

@@
 one author has personally encountered in a research kernel, though not one discussed here.}).
+This is true of \textsc{seL4}~\cite{Klein2009seL4,seL4TOCS,Sewell2013translation}, whose formal machine model omits address translation;
+and \textsc{CertiKOS}~\cite{gu15,gu2016certikos,gu2018certikos,chen2016interrupts}, whose refinement proofs rely on
+\textsc{CompCert}'s usermode-oriented memory model~\cite{leroy2008formal,leroy2009formally} which fundamentally assumes
+all memory addresses are totally independent, which is not the case when a single page table write can make regions of memory inaccessible.
+% \textsc{seL4}~\cite{Klein2009seL4,seL4TOCS,Sewell2013translation} is a formally verified L4 microkernel~\cite{Liedtke1995,Liedtke1996} 
+% (and the first verified OS kernel to run on real-world hardware), verified with a mix of refinement proofs and program logic reasoning down to the assembly level.
+% Because \textsc{seL4} is a microkernel, most VMM functionality actually lives in usermode and is unverified, and moreover, 
+% their hardware model omits address translation entirely and the MMU entirely~\cite{Klein2009seL4,seL4TOCS}. 
+% As a result, the limited page table management present in the microkernel treats page tables as idiosyncratic tree-maps as in
+% \textsc{Verisoft}, despite actually running with address translation.
+% This is partly mitigated by manually identifying some trusted invariants (e.g., that the address range designated for the kernel is appropriately mapped)
+% and setting up the proof to ensure those invariants are maintained (i.e., as an extra proof obligation not required by their hardware model).
 
-\textsc{seL4}~\cite{Klein2009seL4,seL4TOCS,Sewell2013translation} is a formally verified L4 microkernel~\cite{Liedtke1995,Liedtke1996} 
-(and the first verified OS kernel to run on real-world hardware), verified with a mix of refinement proofs and program logic reasoning down to the assembly level.
-Because \textsc{seL4} is a microkernel, most VMM functionality actually lives in usermode and is unverified, and moreover, 
-their hardware model omits address translation entirely and the MMU entirely~\cite{Klein2009seL4,seL4TOCS}. 
-As a result, the limited page table management present in the microkernel treats page tables as idiosyncratic tree-maps as in
-\textsc{Verisoft}, despite actually running with address translation.
-This is partly mitigated by manually identifying some trusted invariants (e.g., that the address range designated for the kernel is appropriately mapped)
-and setting up the proof to ensure those invariants are maintained (i.e., as an extra proof obligation not required by their hardware model).
-
-\textsc{CertiKOS}~\cite{gu15,gu2016certikos,gu2018certikos,chen2016interrupts} is a microkernel intended for use as a hypervisor.
-The overall approach in that body of work is many layers of refinement proofs, using a
- proliferation of layers with small differences to keep most individual refinements tractable. In keeping with precursor work 
-on the project from the same group~\cite{vaynberg2012compositional}, the purpose of some layers is to abstract away from 
-virtual memory (as early as possible).
-The papers on \textsc{CertiKOS} do not explicitly detail the VMM beyond highlighting its existence and referencing
-that it performs mapping operations for user code.
-The work is clear, however, that it fully trusts low-level assembly fragments such as the instruction sequence which actually
-switches address spaces, rather than verifying them.
-We can also infer other limitations from the project's pervasive reliance on a modified \textsc{CompCert},
-which fixes an abstract memory model aimed at \emph{user-mode} C programs~\cite{leroy2008formal,leroy2009formally}. 
-\textsc{CompCert}'s model of memory cannot handle non-local updates, so we can conclude \textsc{CertiKOS}'s proofs cannot
-address impact of virtual memory mapping changes on the kernel itself.
-% Another key aspect of their approach is that the OS is written in Clight and compiled with \textsc{CompCert}~\cite{blazy2006formal,leroy2009formally,leroy2008formal}.
-% CompCert's memory abstraction~\cite{leroy2008formal} assumes
-% memory is a set of disjoint chunks of bytes with no overlap, so the lowest levels of CertiKOS must provide a matching 
-% machine model as a layer. This prohibits virtual address aliasing, so CertiKOS cannot support simultaneous memory-mapped 
-% (\texttt{mmap}) and stream-oriented (\texttt{read}/\texttt{write}) IO to a single file\todo{should we go into this detail?}, 
-% and cannot use
-% the common kernel design choice of mapping all physical memory into the bottom of the kernel's address space for direct access i
-% while the kernel code is simultaneously mapped (and executed) at higher virtual addresses.
-% This is not necessary for \textsc{CertiKOS}'s intended primary use case (a hypervisor), but means that \textsc{CertiKOS}'s
-% approach cannot be used to support this functionality in other systems, without major surgery to \textsc{CompCert}.
-
+% \textsc{CertiKOS}~\cite{gu15,gu2016certikos,gu2018certikos,chen2016interrupts} is a microkernel intended for use as a hypervisor.
+% The overall approach in that body of work is many layers of refinement proofs, using a
+%  proliferation of layers with small differences to keep most individual refinements tractable. In keeping with precursor work 
+% on the project from the same group~\cite{vaynberg2012compositional}, the purpose of some layers is to abstract away from 
+% virtual memory (as early as possible).
+% The papers on \textsc{CertiKOS} do not explicitly detail the VMM beyond highlighting its existence and referencing
+% that it performs mapping operations for user code.
+% The work is clear, however, that it fully trusts low-level assembly fragments such as the instruction sequence which actually
+% switches address spaces, rather than verifying them.
+% We can also infer other limitations from the project's pervasive reliance on a modified \textsc{CompCert},
+% which fixes an abstract memory model aimed at \emph{user-mode} C programs~\cite{leroy2008formal,leroy2009formally}. 
+% \textsc{CompCert}'s model of memory cannot handle non-local updates, so we can conclude \textsc{CertiKOS}'s proofs cannot
+% address impact of virtual memory mapping changes on the kernel itself.
 Other work on OS verification either never progressed far enough to address VMM verification 
@@
 of a single address space by all processes (\textsc{Singularity}~\cite{Fahndrich2006language,Hunt2007singularity,Hunt2007sealing,Barnett2011specsharp}, \textsc{Verve}~\cite{Yang2010Verve},
-and \textsc{Tock}~\cite{levy2017multiprogramming}).
+and \textsc{Tock}~\cite{levy2017multiprogramming}). The latter ensures memory safety, but ignores other functional uses of virtual memory
+hardware like swapping~\cite{Denning1970VM} or exploiting copy-on-write techniques for dynamic migration of virtual machines~\cite{clark2005live}.
 
-Some work has been done like ours, studying VMM verification separately from the rest of the kernel.
-One important outgrowth of the \textsc{seL4} project, not integrated into the main project's proof, was work by 
-Kolanski and Klein which studied verification of code against a hardware model that \emph{did} include address translation
- --- the only work aside from ours to do so --- initially in terms of basic memory~\cite{kolanski08vstte} and subsequently 
-integrating source-level types into the interpretation~\cite{kolanski09tphols}. 
-They were the first work to model physical and virtual points-to assertions separately, defining virtual points-to assertions
-in terms of physical points-to assertions mimicking page table walks, and defining all of their assertions as predicates on a
-pair of (physical) machine memory and a page table root, an approach we improve on.
+\citet{kolanski08vstte} are the only other researchers to study VMM verification against a realistic hardware
+model, where page table updates are performed through virtual memory accesses (later adding C-level type modeling~\cite{kolanski09tphols}).
+As noted in Section \ref{sec:overly-restrictive}, Kolanski and Klein's virtual points-to is similar to that in Figure \ref{fig:strongvirtualpointsto},
+with the attendant problems discussed earlier and lifted by our model. Their approach had modal elements,
+but the work never addressed the issues of managing multiple address spaces, switching address spaces, or
+invariants referring to other address spaces, which our explicitly modal approach and case studies do address.
+They verified a fraction of an ARM version of our Figure \ref{fig:mapping_code}, which assumed they already had the virtual address
+of the L1 entry (which they assumed existed), thereby skipping the physical-to-virtual conversion issues we treat in Section \ref{sec:p2v}.
+Due to lack of fractional permissions in their formalism, they incidentally pick up other limitations orthogonal to their
+foundational focus: by requiring
+a virtual points-to have \emph{full} ownership of the page table walk memory, they limit their system to having only as many
+virtual points-to assertions as there are entries in the top-level table (4096) because they cannot share access to entries.
+We inherit fractional permission support form Iris, and use it extensively (the overly-restrictive
+Figure \ref{fig:strongvirtualpointsto} is already an improvement in this way).
 
-They also define their virtual points-to assertions such that a virtual points-to $p\mapsto_\mathsf{v} a$ owns the full 
-lookup path to virtual address $p$. This means that given two virtual points-to assertions at the same time, such as 
-$p\mapsto_\mathsf{v}a \ast p'\mapsto_\mathsf{v}b$, the memory locations traversed to translate $p$ and $p'$ must be disjoint. 
-This means the logic has a peculiar limit on how many virtual points-to assertions can coexist in a proof. Since page tables 
-fan out, the bottleneck is the number of entries in the root table. For their 32-bit ARMv6 example, the top-level address is 
-still 4Kb (4096 bytes), and each entry (consumed entirely by a virtual points-to in their scheme) is 4 bytes, so they have a 
-maximum of 1024 virtual points-tos in their ARMv6 configuration. Any assertion which implies more than that number
-of virtual addresses are mapped implies false in their logic.
-(They do formulate their logic over an abstract model, but every architecture would incur a similar limitation.)
-% Na\"ively transferring their model to x86-64 4-level tables would yield a limit of 512 assertions (also a 4Kb root page, 
-% but 8-byte entries).
-Our definitions make use of fractional permissions throughout; Figure \ref{fig:strongvirtualpointsto}'s definition
-of \lstinline|L4_L1_PointsTo| ellides the specific fractions used, but it in fact asserts 1/512 ownership of
-the L1 entry, 1/($512^2$) of the L2 entry, and so on, so each entry may map the appropriate number of machine words.
-\looseness=-1
+% Some work has been done like ours, studying VMM verification separately from the rest of the kernel.
+% One important outgrowth of the \textsc{seL4} project, not integrated into the main project's proof, was work by 
+% Kolanski and Klein which studied verification of code against a hardware model that \emph{did} include address translation
+%  --- the only work aside from ours to do so --- initially in terms of basic memory~\cite{kolanski08vstte} and subsequently 
+% integrating source-level types into the interpretation~\cite{kolanski09tphols}. 
+% They were the first work to model physical and virtual points-to assertions separately, defining virtual points-to assertions
+% in terms of physical points-to assertions mimicking page table walks, and defining all of their assertions as predicates on a
+% pair of (physical) machine memory and a page table root, an approach we improve on.
 
-As noted earlier, by collocating both the physical ownership of the page table walk
-as part of the virtual points-to itself these logics preempt support for changes to page tables which do not actually affect 
-address translation. We address this by moving from Figure \ref{fig:strongvirtualpointsto}'s definition (essentially Kolanski and Klein's
-with fractional sharing of intermediate table entries) to Figure \ref{fig:virtualpointstosharing}'s.
-\looseness=-1
+% They also define their virtual points-to assertions such that a virtual points-to $p\mapsto_\mathsf{v} a$ owns the full 
+% lookup path to virtual address $p$. This means that given two virtual points-to assertions at the same time, such as 
+% $p\mapsto_\mathsf{v}a \ast p'\mapsto_\mathsf{v}b$, the memory locations traversed to translate $p$ and $p'$ must be disjoint. 
+% This means the logic has a peculiar limit on how many virtual points-to assertions can coexist in a proof. Since page tables 
+% fan out, the bottleneck is the number of entries in the root table. For their 32-bit ARMv6 example, the top-level address is 
+% still 4Kb (4096 bytes), and each entry (consumed entirely by a virtual points-to in their scheme) is 4 bytes, so they have a 
+% maximum of 1024 virtual points-tos in their ARMv6 configuration. Any assertion which implies more than that number
+% of virtual addresses are mapped implies false in their logic.
+% (They do formulate their logic over an abstract model, but every architecture would incur a similar limitation.)
+% % Na\"ively transferring their model to x86-64 4-level tables would yield a limit of 512 assertions (also a 4Kb root page, 
+% % but 8-byte entries).
+% Our definitions make use of fractional permissions throughout; Figure \ref{fig:strongvirtualpointsto}'s definition
+% of \lstinline|L4_L1_PointsTo| ellides the specific fractions used, but it in fact asserts 1/512 ownership of
+% the L1 entry, 1/($512^2$) of the L2 entry, and so on, so each entry may map the appropriate number of machine words.
+% \looseness=-1
 
-Kolanski and Klein do verify code to map in a new page by installing an L1 page table entry, akin to our Figure \ref{fig:mapping_code}.
-However, our logic treatment of mapping goes beyond theirs. 
-Kolanski and Klein must unfold machine semantics at one point in their proof of mapping correctness~\cite[p.~27]{kolanski08vstte}, while
-our logic permits us to conduct the proof entirely in separation logic.
-Their proofs also do not address
-converting physical addresses of page table entries to virtual addresses that can be used to access them.
-Their original low-level (pseudo-assembly) proof assumes the correct virtual address already exists~\cite{kolanski08vstte},
-and their subsequent C-level proof~\cite{kolanski09tphols} axiomatizes a function akin to our \lstinline|ensure_L1|.
-While full proof of \lstinline|ensure_L1|'s page table walk is ongoing work, we have verified that virtual-to-physical
-translation can be done in our logic with appropriate kernel invariants (Section \ref{sec:traversing}), while
-Kolanski and Klein's model may not be able to do so --- because the virtual points-to owns the physical resources associated with
-intermediate page table entries, it appears impossible to have a virtual points-to for a page table entry that is part of the
-page table walk for another virtual address.
-\looseness=-1
+% As noted earlier, by collocating both the physical ownership of the page table walk
+% as part of the virtual points-to itself these logics preempt support for changes to page tables which do not actually affect 
+% address translation. We address this by moving from Figure \ref{fig:strongvirtualpointsto}'s definition (essentially Kolanski and Klein's
+% with fractional sharing of intermediate table entries) to Figure \ref{fig:virtualpointstosharing}'s.
+% \looseness=-1
 
-The other major distinction is that Kolanski and Klein have no explicit accounting for other address spaces.
-Their logic does not deal with change of address space, and has no way to assert that certain facts hold
-in another address space.
-They verify only one address space manipulation: mapping a single unmapped page into the current address space (in both papers).
-We verify this, as well as a change-of-address-space, which requires us to introduce assertions for talking
-about other address spaces (we must know, for example, that the precondition of the code after the change must be true
-in the \emph{other} address space), and to deal with the fact that the standard frame rule
-for separation logic is unsound in the presence of address space changes and address-space-contingent assertions.
-Our approach in this paper uses modalities to distinguish virtual-address-based assertions that hold only in specific 
-address spaces, making it possible to manipulate other address spaces, and equally critically, to \emph{change} address 
-spaces while reasoning about correctness.
+% Kolanski and Klein do verify code to map in a new page by installing an L1 page table entry, akin to our Figure \ref{fig:mapping_code}.
+% However, our logic treatment of mapping goes beyond theirs. 
+% Kolanski and Klein must unfold machine semantics at one point in their proof of mapping correctness~\cite[p.~27]{kolanski08vstte}, while
+% our logic permits us to conduct the proof entirely in separation logic.
+% Their proofs also do not address
+% converting physical addresses of page table entries to virtual addresses that can be used to access them.
+% Their original low-level (pseudo-assembly) proof assumes the correct virtual address already exists~\cite{kolanski08vstte},
+% and their subsequent C-level proof~\cite{kolanski09tphols} axiomatizes a function akin to our \lstinline|ensure_L1|.
+% While full proof of \lstinline|ensure_L1|'s page table walk is ongoing work, we have verified that virtual-to-physical
+% translation can be done in our logic with appropriate kernel invariants (Section \ref{sec:traversing}), while
+% Kolanski and Klein's model may not be able to do so --- because the virtual points-to owns the physical resources associated with
+% intermediate page table entries, it appears impossible to have a virtual points-to for a page table entry that is part of the
+% page table walk for another virtual address.
+% \looseness=-1
+
+% The other major distinction is that Kolanski and Klein have no explicit accounting for other address spaces.
+% Their logic does not deal with change of address space, and has no way to assert that certain facts hold
+% in another address space.
+% They verify only one address space manipulation: mapping a single unmapped page into the current address space (in both papers).
+% We verify this, as well as a change-of-address-space, which requires us to introduce assertions for talking
+% about other address spaces (we must know, for example, that the precondition of the code after the change must be true
+% in the \emph{other} address space), and to deal with the fact that the standard frame rule
+% for separation logic is unsound in the presence of address space changes and address-space-contingent assertions.
+% Our approach in this paper uses modalities to distinguish virtual-address-based assertions that hold only in specific 
+% address spaces, making it possible to manipulate other address spaces, and equally critically, to \emph{change} address 
+% spaces while reasoning about correctness.
 The only other attempt to verify a change of address space was by \citet{syeda2018program,syeda2020formal}, whose focus was on TLB maintenance (which we
-do not model). However, they verified only the 4 instructions to switch address spaces and update the TLB on an ARM processor, in isolation.
+do not model). However, they verified only the 4 instructions to switch address spaces and update the TLB on an ARM processor, in isolation
+(i.e., not the full context switch including changing stacks with address spaces).
 The specification they proved for those instructions did not address program invariants that may be valid in one address space and not the other,
@@
 
+\ifPLDI
+\else
 Unlike our work, Kolanski and Klein prove very useful embedding theorems stating that code that does not modify page table 
@@
 memory.
+\fi
 

### 2024-11-14 examples.tex
more compressing

@@
 mov     rdi, r8 $\specline{\ldots \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry} \ast \ldots }$
-mov    [r8],rdi (*\label{line:read_entry_contents}*) $\specline{\ldots \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val}  \ast\ldots}$
+mov    [r8],rdi (*\label{line:read_entry_contents}*) $\specline{\ldots \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val}  \ast\ldots}$ ;; Justified by virtual pte-pointsto
 and    0x1,rdi (* \label{line:mask_present} *)
@@
 ... ;; Continued from Figure (*\ref{fig:calltopteinitialize}*); assertions below specialized to non-allocating path for clarity
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \}),m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{rcx}  \mapsto_{\textsf{r}} \textsf{\_}  }_{\rtv}$
-$\specline{ \textsf{entry} \mapsto_{\textsf{id}} \textsf{\_} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \}),m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{rcx}  \mapsto_{\textsf{r}} \textsf{\_}  \ast \textsf{entry} \mapsto_{\textsf{id}} \textsf{\_} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
 $\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{(pte\_initialized (entry\_val.pfn))} \urcorner }_{\rtv}$
@@
 ;;uintptr_t next_virt_addr = (uintptr_t) P2V(entry.pfn<<12); (*\label{line:p2v}*) 
- movabs KERNBASE,rcx
- add    rcx,rax
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \}),m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{rcx}  \mapsto_{\textsf{r}} \textsf{KERNBASE} }_{\rtv}$
-$\specline{ \textsf{entry} \mapsto_{\textsf{id}} \textsf{\_} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
-$\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{(pte\_initialized (entry\_val.pfn))} \urcorner }_{\rtv}$
-$\specline{ \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{(pte\_initialized (entry\_val.pfn)))} \ast \texttt{rax} \mapsto_{\textsf{r}} \textsf{ table\_root (pte\_initialize(entry\_val.pfn)) + KERNBASE} }_{\rtv}$
-$\specline{\forall_{i\in \textsf{0 ... 511} } \ldotp  \ghostmaptoken{\textsf{id}}{((\textsf{table\_root (pte\_initialized (entry\_val.pfn)))}) + \textsf{i * 8})}{\textsf{v-1}}  }$
-...
-;;next = (pte_t *) next_virt_addr;
-;;clean up the stack and return next
+movabs KERNBASE,rcx $\specline{\ldots \ast  \texttt{rcx}  \mapsto_{\textsf{r}} \textsf{KERNBASE}  \ast \ldots}_{\rtv}$
+add    rcx,rax $\specline{ \ldots \ast \texttt{rax} \mapsto_{\textsf{r}} \textsf{ table\_root (pte\_initialize(entry\_val.pfn)) + KERNBASE}  \ast \ldots}_{\rtv}$
+... ;;clean up the stack and return the rax value
 \end{lstlisting}
@@
 To do so, with a given allocated a fresh page (\textsf{fpaddr}), then calculate the appropriate
-known-valid page table walks (\textsf{pte\_walkpgdir} Line X in Figure \ref{fig:mapping_code})  and update the appropriate L1 page table entry (permissions and frame address Lines in Figure \ref{fig:mapping_code}).
+known-valid page table walks (\textsf{pte\_walkpgdir} Line X\todo{!!!} in Figure \ref{fig:mapping_code})  and update the appropriate L1 page table entry (permissions and frame address Lines in Figure \ref{fig:mapping_code});
+unmapping is the reverse of the logic we discuss here.
+\looseness=-1
 %\lstset{
@@
 ... ;;setting up the stack      
-;;char *a; pte_t *pte;
-;;a = (char *)PGROUNDUP((uintptr_t)va);
-...
  $\specline{\textsf{P} \ast \textsf{pml4} \mapsto_{\textsf{id}} \textsf{\_} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}}_{\rtv}$      
-;;pte = walkpgdir(pml4, a);
-mov    -0x8[rbp],rdi
-mov    -0x28[rbp],rsi
-mov    0x0,rax
-callq  291 <vaspace_mappage+0x41>	28d: R_X86_64_PLT32	walkpgdir-0x4
+;; pte_t *pte = walkpgdir(pml4, a);
+mov    -0x8[rbp],rdi; mov    -0x28[rbp],rsi; mov    0x0,rax; callq  291 <vaspace_mappage+0x41>
 ;;if(pte->present != 0) {
 ...        
-mov    0x0,r8
-mov    rax,rcx
-mov    rcx,-0x30[rbp]
-;;pte->writable     = PTE_WRITEABLE_SET(perm);
-;;pte->user_acc     = PTE_USER_ACC_SET(perm);
-;;pte->writethrough = PTE_WRITETHRU_SET(perm);
-;;pte->nocache      = PTE_NOCACHE_SET(perm);
+mov    0x0,r8; mov    rax,rcx; mov    rcx,-0x30[rbp]
+... ;; Copy user access bit, writethrough, etc. (not part of our semantics) into pte from perm
 $\specline{ \left( \begin{array}{l} \textsf{pdp}+\textsf{PDPEX(va)} = \textsf{pdp\_entry} \land \textsf{table\_root(pml4\_entry}_{\textsf{pfn}})  = \textsf{pdp} \\ \textsf{pd}+\textsf{PDEX(va)} = \textsf{pd\_entry} \land \textsf{table\_root(pml3\_entry}_{\textsf{pfn}})  = \textsf{pd} \\ \textsf{pt}+\textsf{PDPEX(va)} = \textsf{pt\_entry} \land \textsf{table\_root(pml2\_entry}_{\textsf{pfn}})  = \textsf{pt} \\  \textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{pml4\_entry},\textsf{pdp\_entry},\textsf{pd\_entry},\textsf{pt\_entry}\},m)  \\ \ghostmaptoken{\delta{}s}{\rtv}{\delta} \\
 \ghostmaptoken{\textsf{id}}{(\mathsf{pml4\_entry})}{\textsf{4}}\ast \ghostmaptoken{\textsf{id}}{(\mathsf{plm3\_entry})}{\textsf{3}} \ast \ghostmaptoken{\textsf{id}}{(\mathsf{plm2\_entry})}{\textsf{2}}  \\ \forall_{i\in\textsf{0..511}} \; \textsf{table\_root(pml4\_entry}_{\textsf{pfn}}) \textsf{ + i * 8} \mapsto_{\textsf{id}} \textsf{3}  \\ \forall_{i\in\textsf{0..511}} \; \textsf{table\_root(pdp\_entry}_{\textsf{pfn}}) \textsf{ + i * 8} \mapsto_{\textsf{id}} \textsf{2}  \\ \forall_{i\in\textsf{0..511}} \; \textsf{table\_root(pd\_entry}_{\textsf{pfn}}) \textsf{ + i * 8} \mapsto_{\textsf{id}} \textsf{1}  \\  \textsf{pml4\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l4e\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l4e\_val})\urcorner \\  \textsf{pdp\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l3e\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l3e\_val})\urcorner \\  \textsf{pd\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l2e\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l2e\_val})\urcorner \\   \textsf{pt\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l1e\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l1e\_val})\urcorner    \end{array} \right)  =\textsf{R}_{\textsf{walk\_acc}}}$
-pte->pfn          = PTE_ADDR_TO_PFN(pa);
-mov    -0x18[rbp],rcx
-shr    0xc,rcx
-mov    -0x30[rbp],rdx
-mov    [rdx],rsi
-$\specline{\textsf{pt\_entry}_{\textsf{pfn}} \mapsto_{p} \textsf{addr\_to\_pfn}(\textsf{pa}) \ast \textsf{R}_{\textsf{walk\_acc}}}_{\rtv}$
-$\sqsubseteq$
-$\specline{\textsf{va} \mapsto_{\textsf{v,q}} \textsf{pa} \ast \textsf{P}}$
-movabs 0xffffffffff,rdi
-and    rdi,rcx
-shl    0xc,rcx
-movabs 0xfff0000000000fff,rdi
-and    rdi,rsi
-or     rcx,rsi
-mov    rsi,[rdx]
+;;pte->pfn          = PTE_ADDR_TO_PFN(pa);
+mov    -0x18[rbp],rcx; shr    0xc,rcx; mov    -0x30[rbp],rdx; mov    [rdx],rsi ;; Load current entry in rsi
+$\specline{\textsf{pt\_entry}_{\textsf{pfn}} \mapsto_{p} \textsf{addr\_to\_pfn}(\textsf{pa}) \ast \textsf{R}_{\textsf{walk\_acc}}}_{\rtv} \Rrightarrow \specline{\textsf{va} \mapsto_{\textsf{v,q}} \textsf{pa} \ast \textsf{P}}$
+movabs 0xffffffffff,rdi; and    rdi,rcx; shl    0xc,rcx; movabs 0xfff0000000000fff,rdi
+and    rdi,rsi; or     rcx,rsi; mov    rsi,[rdx] ;; Store updated entry back to L1 entry
 ;;}
-;;pte->present      = 1;
-;;clean up the stack and return
-...
+... ;; set present bit in entry, then clean up the stack and return (proof re-packs invariants)
 ;;}
 \end{lstlisting}
+\vspace{-1em}
   \caption{Specification and proof of code for mapping a new page with $\textsf{R}_{\textsf{walk\_acc}}$ in Figure \ref{walkpgdir} expanded.}
@@
 % were axiomatized in prior work.
+\ifPLDI
+\else
 \subsection{Unmapping a Page}
@@
 full virtual PTE pointer for it, setting to 0, and reclaiming the now-unmapped physical memory.
+\fi
 
@@
 $\specline{\mathsf{ContextAt}(old,\_) \ast \mathsf{ContextAt}(new,[\mathsf{rbxv}',\ldots,\rtv']) }_{\rtv}$
-mov 0[rdi], rbx ... ;; also mov rsp, rbp, r12, r13, r14, saved to offsets 8, 16, 24, 32, and 40 from rdi
+mov 0[rdi], rbx ... ;; also save rsp, rbp, r12, r13, r14, to offsets 8, 16, 24, 32, and 40 from rdi
 mov 48[rdi], r15

### 2024-11-14 examples.tex
make  large footnote non-PLDI only, which unblocks TeXs layout algorithm

@@
 \texttt{kalloc}'s specification guarantees it has returned memory from a designated memory
-pool that is already mapped\footnote{A reasonable reader might wonder where this pool
+pool that is already mapped
+\ifPLDI
+\else
+\footnote{A reasonable reader might wonder where this pool
 initially comes from, and how it might grow when needed. Typically an initial mapping subject to this identity mapping
@@
 as needed.
-} and satisfies the offset invariants.
+} 
+\fi
+and satisfies the offset invariants.
 % \todo[inline,color=blue]{colin frontier.

### 2024-11-14 examples.tex
compress figure 11

@@
 \begin{figure}\footnotesize
+\ifPLDI
+\begin{lstlisting}[mathescape,escapeinside={(*}{*)}]
+;;pte_t *walkpgdir(pte_t *pml4, const void *va) {
+... ;; Stack setup
+$\specline{\textsf{P} \ast \ulcorner\textsf{rtv+KERNBASE}=\textsf{pml4}\urcorner \ast \forall_{i\in\textsf{0..511}}\ghostmaptoken{\textsf{id}}{(\textsf{rtv}+i*8)}{\textsf{4}} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}}_{\rtv}$
+;;pte_t *pml4_entry = &pml4[PML4EX(va)]; // Virtual address of L4 entry(*\label{line:start_pml4_calc}*)
+mov    -0x8[rbp],rsi; mov    -0x10[rbp],rdi; ;; Load pml4 and va
+;; Extract 9-bit (0x1ff=511) L4 index from va, multiply by 8 (byte index), add to pml4 for entry address
+shr    0x27,rdi; and    0x1ff,rdi; shl    0x3,rdi; add    rdi,rsi(*\label{line:end_pml4_calc}*)
+mov    rsi,-0x18[rbp] ;; Store virt. addr. to local var pml4_entry; logical pml4_entry is physical!
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{pml4\_entry}\},m) \ast \forall_{i\in\textsf{0..511}\setminus\{\textsf{PML4EX(va)}\}}\ghostmaptoken{\textsf{id}}{(\textsf{rtv}+i*8)}{\textsf{4}}\ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
+$\specline{\textsf{entry\_present(l4e\_val)} \wand \forall_{i\in\textsf{0..511}} \; \ghostmaptoken{\textsf{id}}{\textsf{table\_root(l4e\_val)+ i * 8}}{\textsf{3}} }_{\rtv} $
+$\specline{ \textsf{pml4\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{pml4\_entry l4e\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l4e\_val})\urcorner}_{\rtv}$
+;;pte_t *pdp = pte_get_next_table(pml4_entry);
+mov    -0x18[rbp],rdi; ...; callq  c0 <pte_get_next_table>(*\label{line:first_getnext_call}*)
+;;save the physical next table address in rax
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{pml4\_entry}\},m)  \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
+$\specline{\forall_{i\in\textsf{0..511}} \ghostmaptoken{\textsf{id}}{\textsf{table\_root(l4e\_val'.pfn) + i * 8}}{\textsf{3}}  \ast \ulcorner\textsf{pdp-KERNBASE}=\textsf{table\_root(l4e\_val'.pfn)}\urcorner} $
+$\specline{ \textsf{pml4\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{pml4\_entry l4e\_val'} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l4e\_val})\urcorner}_{\rtv}$(*\label{line:first_pte_pointsto}*)
+$\specline{\texttt{rax}  \mapsto_{\textsf{r}} \textsf{table\_root(l4e\_val'.pfn)} }_{\rtv}$ ;; pte_get_next_table may have allocated a new page, updating entry
+;;pte_t *pdp_entry = &pdp[PDPEX(va)]; // Virtual address of L3 entry
+$\specline{\ulcorner\textsf{pdp}+\textsf{PDPEX(va)*8} = \textsf{pdp\_entry} \land \textsf{table\_root(l4e\_val'.pfn)}  = \textsf{pdp} \urcorner }_{\rtv}$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{pml4\_entry},\textsf{pdp\_entry}\},m)  \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
+$\specline{\forall_{i\in\textsf{0..511}\setminus\{\textsf{PDPEX(va)}\}} \; \textsf{table\_root(pml4\_entry}_{\textsf{pfn}}) \textsf{ + i * 8} \mapsto_{\textsf{id}} \textsf{3} }_{\rtv}$
+$\specline{ \textsf{pml4\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac4}}\textsf{pml4\_entry}\; \textsf{l4e\_val'} \ast \ulcorner  \textsf{qfrac4} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l4e\_val'})\urcorner}_{\rtv}$(*\label{line:ex_l4_vpte}*)
+$\specline{ \textsf{pdp\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac3}}\textsf{pdp\_entry}\; \textsf{l3e\_val} \ast \ulcorner  \textsf{qfrac3} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l3e\_val})\urcorner}_{\rtv}$(*\label{line:ex_l3_vpte}*)
+$\specline{\textsf{entry\_present(l3e\_val)} \wand \forall_{i\in\textsf{0..511}} \;\ghostmaptoken{\textsf{id}}{\textsf{table\_root(pdp\_entry.pfn) + i * 8}}{ \textsf{2}} } _{\rtv} $
+... ;; Similar assembly & proofs for levels 3 and 2
+$\specline{\textsf{R}_{\textsf{walk}}}_{\rtv}$
+... ;;return &pt[PTEX(va)]; // return address of L1 entry, similar index computation
+\end{lstlisting}
+
+\else
 \begin{lstlisting}[mathescape,escapeinside={(*}{*)}]
@@
 \end{lstlisting}
+\fi
 \vspace{-1em}

### 2024-11-14 examples.tex
compressing figure 7 more

@@
 mov     rdi, r8 $\specline{\ldots \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry} \ast \ldots }$
-mov    [r8],rdi (*\label{line:read_entry_contents}*) $\specline{\ldots \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val}  \ast\ldots}$ ;; Justified by virtual pte-pointsto
-and    0x1,rdi (* \label{line:mask_present} *)
-mov    rdi,rax
-cmp    0x0,rax ;;  if (!entry->present) {(* \label{line:check_entry_present} *)
+mov    [r8],rdi (*\label{line:read_entry_contents}*) $\specline{\ldots \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val}  \ast\ldots}$ ;; reading page table entry Justified by virtual pte-pointsto
+and    0x1,rdi;  mov    rdi,rax; cmp    0x0,rax ;;  if (!entry->present) {(* \label{line:check_entry_present} *)(* \label{line:mask_present} *)
 jne    161 <pte_get_next_table+0xa1> (* \label{line:check_entry_present_jump} *) ;; Jump if the present bit is not zero
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry} \ast \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{next} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val \& 0x1} }_{\rtv}$
-$\specline{  \textsf{entry} \mapsto_{\textsf{id}} \textsf{v} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  \ast \texttt{rax}  \mapsto_{\textsf{r}} \textsf{entry\_val \& 0x1} \ast \ulcorner\textsf{entry\_val \& 0x1}=\textsf{0x0}\urcorner}_{\rtv}$
-$\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow  \lnot(\textsf{entry\_val \& 0x1}=\textsf{0x1})  \urcorner }_{\rtv}$ (*\label{line:before_concluding_qfrac1}*)
-mov    rbp,rdi (*\label{line:alloc_path_start}*)
-sub    0x10, rdi ;; Store the value of rbp minus 16 bytes (address of (*\textsf{next}*)) into rdi (*\label{line:pass_addrof_next}*)
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry} \ast \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{next} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{rbp - 16} }_{\rtv}$
-$\specline{  \textsf{entry} \mapsto_{\textsf{id}} \textsf{v} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta} }_{\rtv}$
+$\specline{\ldots \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val \& 0x1} }_{\rtv}$
+mov    rbp,rdi; sub    0x10, rdi ;; Store the value of rbp minus 16 bytes (address of (*\textsf{next}*)) into rdi (*\label{line:pass_addrof_next}*)(*\label{line:alloc_path_start}*)
 $\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \land \lnot(\textsf{entry\_present entry\_val})  \urcorner }_{\rtv}$ (*\label{line:after_concluding_qfrac1}*)
 callq  70 <pte_initialize> ;;pte_initialize(entry);(* \label{line:call_to_pte_initialize} *)
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  }_{\rtv}$
-$\specline{  \textsf{entry} \mapsto_{\textsf{id}} \textsf{v} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta} }_{\rtv}$
-$\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \land \lnot(\textsf{entry\_present entry\_val})  \urcorner }_{\rtv}$
-$\specline{\ulcorner  \textsf{qfrac} = 1 \land \lnot(\textsf{entry\_present (pfn\_set (entry\_val nextpaddr))}  \urcorner   }_{\rtv}$
+$\specline{ \ldots \ast \ulcorner  \textsf{qfrac} = 1 \land \lnot(\textsf{entry\_present entry\_val})  \urcorner \ast \ulcorner  \textsf{qfrac} = 1 \land \lnot(\textsf{entry\_present (pfn\_set (entry\_val nextpaddr))}  \urcorner   }_{\rtv}$
 $\specline{ \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{pfn\_set(entry\_val nextpaddr)}   }_{\rtv}$
@@
 ... ;;now we know that entry is initialized, so we satisfy the condition to access children list
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  }_{\rtv}$
-$\specline{  \textsf{entry} \mapsto_{\textsf{id}} \textsf{v} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta} }_{\rtv}$
 $\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}}  \textsf{pte\_initialized(pfn\_set(entryv nextpaddr))} \ast \ulcorner  \textsf{qfrac} = 1 \land \lnot(\textsf{entry\_present entry\_val})  \urcorner }_{\rtv}$

### 2024-11-14 examples.tex
a little more figure condensing...

@@
   \begin{lstlisting}[mathescape,escapeinside={(*}{*)}]
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)   \ast \texttt{rbp-16} \mapsto_{\textsf{r}} \textsf{\_} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry+KERNBASE}  }_{\rtv}$
-$\specline{\texttt{rax} \mapsto_{\textsf{r}} \textsf{\_} \ast \lnot(\textsf{entry\_present entry\_val} }_{\rtv}$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)   \ast \texttt{rbp-16} \mapsto_{\textsf{r}} \textsf{\_} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry+KERNBASE}  \ast \texttt{rax} \mapsto_{\textsf{r}} \textsf{\_} \ast \lnot(\textsf{entry\_present entry\_val} }_{\rtv}$
 $\specline{\textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,1}} \textsf{entry entryv} \ast \ghostmaptoken{\textsf{id}}{(\mathsf{entry})}{\textsf{level}} }$
-;;void pte_initialize(pte_t *entry) {
-... ;;set up the stack    
-;;allocate a full zeroed page for 512 8-byte entries 
-callq  81 <kalloc>	;;pte_t *local = kalloc();(*\label{line:call_to_kalloc}*)
+;;void pte_initialize(pte_t *entry) { ... Stack frame setup elided    
+callq  81 <kalloc> ;;allocate a full zeroed page for 512 8-byte entries(*\label{line:call_to_kalloc}*)
 mov    rax,-0x10[rbp] ;; Store into 'local'
@@
 mov    -0x10[rbp],rax; mov    -0x8[rbp],rdi; mov    rax,[rdi]
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)   \ast \texttt{rbp-16} \mapsto_{\textsf{r}} \textsf{\_} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry+KERNBASE} }_{\rtv}$
-$\specline{\textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,}} \textsf{entry entryv} \ast \ghostmaptoken{\textsf{id}}{(\mathsf{entry})}{\textsf{level}} }$
-$\specline{\texttt{rax} \mapsto_{\textsf{r}} \textsf{nextpaddr} \ast \lnot(\textsf{entry\_present entry\_val}) }$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)   \ast \texttt{rbp-16} \mapsto_{\textsf{r}} \textsf{\_} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry+KERNBASE} \ast \texttt{rax} \mapsto_{\textsf{r}} \textsf{nextpaddr} \ast \lnot(\textsf{entry\_present entry\_val}) }$
 $\specline{ \textsf{entry+KERNBASE} \mapsto_{\mathsf{vpte,1}} \mathsf{entry} \; \textsf{pfn\_set(entryv nextpaddr)}}$

### 2024-11-15 background.tex
various space tweaks...

@@
 1 page table resulting in the actual page of physical memory holding the requested data, and the low-order 12 bits 
-being used to index into this page.\footnote{Technically levels 1--3 have designated names originating when there were 
-fewer levels. We omit those names to avoid confusion for those not already steeped in the arcana of x86 history,
- and for greater consistency with the recent addition of a 5th level 
-on the most recent processors, where levels 4 and 5 are simply known by their numbers. Our formalization
+being used to index into this page.\footnote{Technically levels 1--3 have explicit historical names,
+but for brevity and consistency, we simply number them, in keeping with the newer 5th level.
+Our formalization
 only deals with 4-level page tables, but is straightforwardly extensible to 5.}  
@@
 The page tables are managed by the OS, typically by a \emph{virtual memory manager} (VMM).\footnote{Not to be confused with Virtual 
-Machine Monitor. In this paper we focus on non-hypervisor scenarios, but hardware virtualization extensions for both 
+Machine Monitor. We focus on non-hypervisor scenarios, but hardware virtualization extensions for both 
 x86-64 and ARM make use of an additional set of page tables translating what a \emph{guest} considers to be its 

### 2024-11-15 examples.tex
various space tweaks...

@@
 been performed by the code in this figure.
+\looseness=-1
 
@@
 \label{fig:calltopteinitialize}
+\vspace{-1em}
 \end{figure}
@@
 $\Xi$ is another ghost map, from physical addresses to the level of the page table they represent (1--4).
-\emph{Only} physical addresses in $\Xi$ can undergo P2V conversion. We describe proof of such a conversion in Section \ref{sec:p2v},
-but describe the invariant here because installing a new level 3/2/1 table requires maintaining that invariant.
+\emph{Only} physical addresses in $\Xi$ can undergo P2V conversion. 
+Section \ref{sec:p2v} describes the actual conversion,
+but we describe the invariant here 
+because adding new level 3/2/1 tables must maintain the invariant.
 
@@
 \label{pteinitializespec}
+\vspace{-1em}
 \end{figure}
@@
  % Abreviating the other relevant resources (including the lower entries) peeled-out of the invariant as $\textsf{R}_{\textsf{children}}$
+\vspace{-1em}
 }
@@
 \label{walkpgdir}
+\vspace{-1em}
 \end{figure}
@@
 rather than existentially quantified:
+\looseness=-1
  \[

### 2024-11-15 intro.tex
various space tweaks...

@@
 reason about change in address spaces.
-These advantages make this approach to reasoning about virtual memory more flexible than prior program logic techniques~\cite{kolanski08vstte,kolanski09tphols}, 
+This approach to reasoning about virtual memory is more flexible than prior program logic techniques~\cite{kolanski08vstte,kolanski09tphols}, 
 which were only able to work with a single address space (the current address space on the CPU) because they were unable
@@
 the \emph{non-local} effects of page table updates whether within the current address space or across address spaces.
+\looseness=-1
 

### 2024-11-15 programlogic.tex
various space tweaks...

@@
 essentially working with separation logic assertions indexed by a choice of page table root as a $\mathcal{W}_{64}$, which we call $\textsf{vProp }\Sigma$:\footnote{
-  \iris experts may notice our \lstinline|-b>| resembles another variety of pointwise lifting that already exists in the main \iris
-  distribution~\cite{dang2019rustbelt,dang2022compass}. This is essentially true, but the standard formalization does not work for index types which
-  are indexed in the sense of \coq's metatheory, as our \lstinline|word n| types are.
+  \iris experts may notice our \lstinline|-b>| resembles another pointwise lifting already in  \iris~\cite{dang2019rustbelt,dang2022compass}. 
+  This similarity is real, but the existing lifting does not work with indexed \coq types
+  like our \lstinline|word n| as a domain.
 }
@@
 \textsf{WP} is Iris's own weakest precondition modality, unmodified~\cite{jung2018iris}.
-Making \textsf{rtv} an explicit parameter to the double, rather than simply using register assertions
-directly 
+Making \textsf{rtv} a parameter to the double (vs. a simple register assertion)
 makes it possible to ensure ownership of the \lstinline|cr3| register and its value is accounted for
 while avoiding some technical headaches with trying to enforce that $\Phi$ itself contains that.
+\looseness=-1
 % solves a technical problem with ensuring that the page table root used to evaluate

### 2024-11-15 examples.tex
compress some text enough that only the conclusions are over the page limit

@@
 The mapping operation of Figure \ref{fig:mapping_code} assumes an operation \textsf{walkpgdir} which must traverse the page tables
-in order to locate the address of the L1 entry to update --- possibly allocating tables for levels 3, 2, and 1 in the process,
-installing them into levels 4, 3, and 2, along the way.
+in order to locate the address of the L1 entry to update --- 
+% possibly allocating tables for levels 3, 2, and 1 in the process,
+% installing them into levels 4, 3, and 2, along the way.
+possibly allocating new L3, L2, and L1 tables as necessary.
 Traversing the page tables is itself challenging functionality to verify: loading the current table root from \lstinline|cr3| is straightforward
@@
 
-Note that none of the verification for this function assumes specific page table levels --- logical parameter \textsf{v} represents the level
+Note that the specification does \emph{not} assume a specific page table --- logical parameter \textsf{v} represents the level
 of the entry passed as an argument, and this code
@@
 \label{subsec:identitymappings}
-It is typical for kernels to need to convert between physical and virtual addresses, in both directions.
+Kernels need to convert between physical and virtual addresses, in both directions.
 Traversing the page tables in software is the simplest way to convert a virtual address to a physical address; this is the context we are working up to.
@@
 because the addresses stored in page table entries are physical, but memory accesses issued by the OS code use virtual addresses.
-There is no universal way to convert physical addresses to virtual --- doing so relies on the kernel maintaining careful invariants or
-additional data structures to enable P2V translation.
+% There is no universal way to convert physical addresses to virtual --- doing so relies on the kernel maintaining careful invariants or
+% additional data structures to enable P2V translation.
+\looseness=-1
 
-In practice, VMM operations are performance-critical for many workloads, so most kernels opt for using invariants to make P2V conversion very fast,
-rather than maintaining yet another data structure.
+Because VMM operations are performance-critical for many workloads, most kernels 
+maintain invariants that enable very fast P2V conversions (rather than adding another data structure).
 Most kernels maintain an invariant on their page tables that the virtual address of any page used for a page table lives at a virtual address
@@
   \label{fig:peraspaceinvariant_with_p2v_extension}
+\vspace{-1em}
 \end{figure*}
@@
 
-If the entry is not set, \textsf{pte\_initialize} is called (Line \ref{line:call_to_pte_initialize} in Figure \ref{fig:calltopteinitialize}) 
-for a physical page (utilizing the page-allocator's \textsf{kalloc} -- currently the only axiomatized call in the the proof of \textsf{pte\_initialize} 
-(Line \ref{line:call_to_kalloc} in Figure \ref{pteinitializespec}). 
+If the entry is not set, \textsf{pte\_initialize} (Line \ref{line:call_to_pte_initialize} in Figure \ref{fig:calltopteinitialize}) 
+allocates a physical page (internally utilizing the only axiomatized call in our case studies, the page-allocator's \textsf{kalloc}, 
+on Line \ref{line:call_to_kalloc} in Figure \ref{pteinitializespec}). 
+\looseness=-1
 \begin{figure}\footnotesize
@@
 \label{sec:p2v}
-Once we reach to the certain knowledge of having an entry with a frame referencing to an allocated resource resides inside the identity mappings
-(which can already be known if the branch at Line \ref{line:check_entry_present_jump} is taken, or ensured by allocating and installing a new entry
+Once we know the entry refers to an allocated resource inside the identity mapping range ($\Xi$)
+% Once we reach to the certain knowledge of having an entry with a frame referencing to an allocated resource resides inside the identity mappings
+(via the branch at Line \ref{line:check_entry_present_jump}, or  by allocating and installing a new entry
 as just discussed for Lines \ref{line:check_entry_present_jump}--\ref{line:end_of_allocation_path}), 
-we can utilize this knowledge 
+we can %utilize this knowledge 
 to convert this frame address into an virtual address of the next page table through, again, identity mappings -- Line 57 in Figure \ref{fig:calltopteinitialize} 
 specified in Figure \ref{fig:p2v}.
-
 This actually constitutes a very critical piece of the full page table walk verification, and we have verified the critical step for a small x86-64 kernel, 

### 2024-11-15 examples.tex
3 lines over...

@@
 (L1 entries point to data pages, whose physical memory ownership resides in some virtual points-to).
-The assertion on Line \ref{line:conditional_children} originates from the invariant one level up, and if this code determines the valid bit is set,
-it can return those child tokens without the conditional guard.
+The assertion on Line \ref{line:conditional_children} comes from the invariant one level up; 
+if the valid bit is set,
+the code can return those child tokens without the conditional guard.
+\looseness=-1
 
@@
 \label{sec:p2v}
-Once we know the entry refers to an allocated resource inside the identity mapping range ($\Xi$)
+Once we know the entry refers to a physical address the identity mapping range ($\Xi$)
 % Once we reach to the certain knowledge of having an entry with a frame referencing to an allocated resource resides inside the identity mappings
@@
 we can %utilize this knowledge 
-to convert this frame address into an virtual address of the next page table through, again, identity mappings -- Line 57 in Figure \ref{fig:calltopteinitialize} 
+to convert this frame address to a corresponding virtual address via identity mappings -- Line 57 in Figure \ref{fig:calltopteinitialize} 
 specified in Figure \ref{fig:p2v}.
-This actually constitutes a very critical piece of the full page table walk verification, and we have verified the critical step for a small x86-64 kernel, 
+This is a critical piece of the full page table walk verification, and we have verified the critical step for a small x86-64 kernel, 
 which is the physical-to-virtual conversion (often appearing as a macro \texttt{P2V} in C source code). 
@@
 where no allocation was necessary (i.e., the conditional on Line \ref{line:check_entry_present} of Figure \ref{fig:calltopteinitialize} was taken).
-In this case, the present bit being set grants access to the child tokens from Line \ref{line:conditional_children} of Figure \ref{fig:calltopteinitialize},
+In this case, the true present bit allows access to the child tokens from Line \ref{line:conditional_children} of Figure \ref{fig:calltopteinitialize},
 which is then refined to the assertion on Line \ref{line:children} of Figure \ref{fig:p2v}.
@@
 address being translated). So instead we pass back the per-address-space invariant with the identity mapping resources for \lstinline|entry|
-pulled out. It is up to the caller to determine which entry in that table must actually
+pulled out. The caller determines which entry in that table must actually
 be accessed --- by selecting the appropriate index into the 512 ghost map tokens returned in the postcondition,
@@
 Unlike the only prior work verifying analogous code for mapping a new page~\cite{kolanski08vstte,kolanski09tphols}, our proof above
-does \emph{not} need to unfold the definitions of the logic,
-but instead uses logic rules throughout to reason about the effect of the mapping. 
-This is the first proof we know of for mapping a virtual memory page that does not resort to direct reasoning over the operational semantics.
+does \emph{not} need to reason directly over the operational semantics,
+making this the first verification we know of for mapping a virtual memory page that 
+stays entirely at the program logic level.
+\looseness=-1
 % By incorporating verification of the
@@
 
-\subsection{Change of Address Space}
-A critical piece of \emph{trusted} code in current verified OS kernels is the assembly code to change the current address space; current verified OS kernels currently 
-lack effective ways to specify and reason about this low-level operation, for reasons outlined in Section \ref{sec:relwork}.
 \begin{figure}\footnotesize
 \begin{lstlisting}[mathescape,escapeinside={(*}{*)}]
-;; Assume the save-space is in rdi, load-space in rsi. First, save the yielding context
 $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,\Xi,m) \ast [\rtv'](\mathcal{I}\texttt{ASpace}(\theta',\Xi',m') \ast \texttt{Pother}) \ast \texttt{rsi}\mapsto_{\textsf{r}} new \ast \texttt{rdi}\mapsto_{\textsf{r}} old \ast \texttt{rbx}\mapsto_{\textsf{r}} \texttt{rbxv} }_{\rtv}$
@@
 
+\subsection{Change of Address Space}
+A critical piece of \emph{trusted} code in current verified OS kernels is the assembly code to change the current address space; current verified OS kernels currently 
+lack effective ways to specify and reason about this low-level operation, for reasons outlined in Section \ref{sec:relwork}.
+
 Figure \ref{fig:swtch} gives simplified code for a basic task switch, the heart of an OS scheduler implementation. This is code that saves the context (registers and stack)
@@
 that calls this code, as well as on the new stack of the thread that is restored, thus only the callee-save registers and \texttt{cr3} must be 
-restored.\footnote{We are simplifying in a couple basic ways.
-First, we are ignoring non-integer registers (e.g., floating point, vector registers) entirely.
-Second, we are ignoring that the caller-save registers should still be initialized to 0 to avoid leaking information across processes.
-We focus on the core logical requirements.}
-With the addition of a return instruction, this code would satisfy the C function signature\footnote{The name comes from the UNIX 6th Edition \lstinline|swtch| function, the source of the infamous ``You are not expected to understand this'' comment~\cite{lions1996lions}.}
+restored.\footnote{We are simplifying by ignoring non-integer registers (e.g., floating point, vector registers),
+and the caller-save registers should be initialized to 0 to avoid leaking information across processes, but this captures the key challenges.}
+With the addition of a return instruction, this code would satisfy the C function signature\footnote{This is the function in UNIX 6th Edition 
+with the infamous ``You are not expected to understand this'' comment~\cite{lions1996lions}.}
 \begin{lstlisting}[language=C]
@@
 Because this code does not directly update the instruction pointer, it is worth explaining \emph{how} this switches threads: by switching address spaces and stacks. 
-This is meant to be called with a return address for the current thread stored on the current stack when called --- which must be reflected in the calling convention. 
-In particular, the precondition of the return address on the initial stack requires the callee-save register values at the time of the call: those stored in the first 
+This is meant to be called with a return address for the current thread stored on the current stack when called. 
+The precondition of the return address on the initial stack requires the callee-save register values at the time of the call: those stored in the first 
 half of the code.
@@
 not the address space of the calling thread shown as 
-\[[\rtv'](\mathcal{I}\texttt{ASpace}(\theta,\Xi,m) \ast \texttt{Pother})\]
+$[\rtv'](\mathcal{I}\texttt{ASpace}(\theta,\Xi,m) \ast \texttt{Pother})$
 in the specfication. 
-Thus the precondition of this code,
+The precondition of this code,
 in context, would include that the initial stack pointer (before \lstinline|rsp| is updated)

### 2024-11-15 relwork.tex
3 lines over...

@@
 unsoundly given that other projects target hardware that \emph{does} run the kernel with address translation: they \emph{trust}
-that the particular page table manipulations do not, for example, unmap kernel code (which can crash the machine even if done ``temporarily''\footnote{A bug
-one author has personally encountered in a research kernel, though not one discussed here.}).
+that the particular page table manipulations do not, for example, unmap kernel code (which can crash the machine even if done ``temporarily'').
+% \footnote{A bug
+% one author has personally encountered in a research kernel, though not one discussed here.}).
 This is true of \textsc{seL4}~\cite{Klein2009seL4,seL4TOCS,Sewell2013translation}, whose formal machine model omits address translation;
 and \textsc{CertiKOS}~\cite{gu15,gu2016certikos,gu2018certikos,chen2016interrupts}, whose refinement proofs rely on
-\textsc{CompCert}'s usermode-oriented memory model~\cite{leroy2008formal,leroy2009formally} which fundamentally assumes
-all memory addresses are totally independent, which is not the case when a single page table write can make regions of memory inaccessible.
+\textsc{CompCert}'s usermode-oriented memory model~\cite{leroy2008formal,leroy2009formally} which assumes
+updates to one memory address are independent of updates to another ---
+which is not true of page table updates.
+ % which is not the case when a single page table write can make regions of memory inaccessible.
 % \textsc{seL4}~\cite{Klein2009seL4,seL4TOCS,Sewell2013translation} is a formally verified L4 microkernel~\cite{Liedtke1995,Liedtke1996} 
@@
 
-\iris has been used to build other logics through pointwise lifting, notably logics that deal with weak
-memory models~\cite{dang2019rustbelt,dang2022compass}. Those systems build a derived logic
-whose lifting consists of functions from thread-local views of events (an operationalization of the release-acquire + nonatomic
-portion of the repaired C11 memory model~\cite{lahav2017repairing}): there modalities $\Delta_\pi(P)$ and $\nabla_\pi(P)$
-represent that $P$ held before or will hold after certain memory fence operations by thread $\pi$.
+Some logics for weak memory models~\cite{dang2019rustbelt,dang2022compass} have been formalized
+in \iris using pointwise lifting, parameterizing by thread-local views of events (an operationalization of the release-acquire + nonatomic
+portion of the repaired C11 memory model~\cite{lahav2017repairing}). 
+There modalities $\Delta_\pi(P)$ and $\nabla_\pi(P)$
+represent that $P$ held before or will hold after certain memory fences by thread $\pi$.
 The definitions of those specific modalities existentially quantify over other views, related to the ``current'' view (the one where
@@
 assertion semantics by a point of evaluation, and evaluating modalized assertions at other points quanfied in the definition of a modality,
-is the classic notion of modal assertions, whereas hybrid logics expose the choice of evaluation point in assertions,
-allowing statements of more properties. In these weak memory examples this additional expressive power would not be useful,
+is the classic notion of modal assertions, 
+whereas hybrid logics expose the choice of evaluation point in assertions,
+allowing statements of more properties. 
+In these weak memory examples this additional expressive power would not be useful,
 because any relevant points of evaluation (thread views) are intimately tied to memory fences performed by the program, whereas

### 2024-11-15 examples.tex
20 pages

@@
 $\specline{\mathsf{ContextAt}(old,\_) \ast \mathsf{ContextAt}(new,[\mathsf{rbxv}',\ldots,\rtv']) }_{\rtv}$
-mov 0[rdi], rbx ... ;; also save rsp, rbp, r12, r13, r14, to offsets 8, 16, 24, 32, and 40 from rdi
+mov 0[rdi], rbx ... ;; also save rsp, rbp, r12, r13, r14, to offsets 8, 16, 24, 32, and 40 from rdi (*\label{line:start_save}*)
 mov 48[rdi], r15
@@
 $\specline{\mathsf{ContextAt}(old,[\mathsf{rbxv,\ldots,\rtv}]) \ast \mathsf{ContextAt}(new,[\mathsf{rbxv}',\ldots,\rtv']) }_{\rtv}$
-mov 56[%rdi], %cr3 $\specline{\ldots \ast \texttt{rdi+56} \mapsto_{\textsf{v}} \rtv \ast \ldots }_{\rtv}$    
-;; Restore target context
-mov rbx, 0[rsi] 
+mov 56[%rdi], %cr3 $\specline{\ldots \ast \texttt{rdi+56} \mapsto_{\textsf{v}} \rtv \ast \ldots }_{\rtv}$   (*\label{line:end_save}*) 
+mov rbx, 0[rsi] (*\label{line:start_restore}*)
 mov rsp, 8[rsi] ;; Switch to new stack, which may not be mapped in the current address space!
@@
 $\specline{\mathsf{ContextAt}(old,[\mathsf{rbxv,\ldots,\rtv}]) \ast \mathsf{ContextAt}(new,[\mathsf{rbxv}',\ldots,\rtv']) }_{\rtv}$
-mov cr3, 56[rsi] ;; <-- Switch to the new address space
-$\specline{ [\rtv](\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,\Xi,m) \ast \mathsf{ContextAt}(old,[\mathsf{rbxv,\ldots,\rtv}]) \ast \mathsf{ContextAt}(new,[\mathsf{rbxv}',\ldots,\rtv']))  }_{\rtv}$
+mov cr3, 56[rsi] ;; <-- Switch to the new address space (*\label{line:end_restore}*)
+$\specline{ [\rtv](\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,\Xi,m) \ast \mathsf{ContextAt}(old,[\mathsf{rbxv,\ldots,\rtv}]) \ast \mathsf{ContextAt}(new,[\mathsf{rbxv}',\ldots,\rtv']))  }_{\rtv}$ (*\label{line:modality_switch}*)
 $\specline{ \mathcal{I}\texttt{ASpace}(\theta',\Xi',m') \ast \texttt{Pother} \ast  \texttt{rsi}\mapsto_{\textsf{r}} new \ast \texttt{rdi}\mapsto_{\textsf{r}} old \ast \ldots }_{\rtv'}$
@@
 Figure \ref{fig:swtch} gives simplified code for a basic task switch, the heart of an OS scheduler implementation. This is code that saves the context (registers and stack)
-of the running thread (here in a structure pointed to by \lstinline|rdi|'s value shown in Lines 5--7 and Line 11 of Figure \ref{fig:swtch}) and restores the context of 
-an existing thread (from \lstinline|rsi| shown in abbreviated Lines 14--17 and Line 22), including the corresponding change of address space for a target thread in another process.
+of the running thread (here in a structure pointed to by \lstinline|rdi|'s value shown in Lines \ref{line:start_save}--\ref{line:end_save} of Figure \ref{fig:swtch}) and restores the context of 
+an existing thread (from \lstinline|rsi| shown in abbreviated Lines \ref{line:start_restore}--\ref{line:end_restore}), including the corresponding change of address space for a target thread in another process.
 This code assumes the System V AMD64 ABI calling convention, where the normal registers not mentioned are caller-save, and therefore saved on the stack of the thread
@@
 with the infamous ``You are not expected to understand this'' comment~\cite{lions1996lions}.}
-\begin{lstlisting}[language=C]
-void swtch(context_t* save, context_t* restore);
-\end{lstlisting}
-A call to this code begins executing one thread (up through Line 17) in one address space ($\rtv$), whose information will be saved in a structure at address $old$,
-and finishes execution executing a different thread in a different address space (whose information is initially in $new$).
+\centerline{\lstinline[language=C]|void swtch(context_t* save, context_t* restore);|}\\
+A call to this code begins executing one thread (until just before Line \ref{line:end_save}) in one address space ($\rtv$), whose information will be saved in a structure at address $old$,
+and finishes execution executing a different thread in a different address space (Line \ref{line:end_restore} on) whose information is initially in $new$.
 
@@
 
-Note that immediately after the page table switch, the points-to information about the saved and restored contexts is guarded by a modality for the retiring
-address space \rtv (Line 23). This is enforced by \textsc{WriteToRegCtlFromRegModal} (Figure \ref{fig:wpdamd}), and is sensible because
-there is no general guarantee that the data structures of the previous addres space are mapped in the new address space.
+Immediately after the page table switch, assertions about the saved and restored contexts are
+guarded by a modality for the retiring
+address space \rtv{} (Line \ref{line:modality_switch}), per
+\textsc{WriteToRegCtlFromRegModal} (Figure \ref{fig:wpdamd}),
+because
+there is no guarantee that the data structures of the previous address space are mapped in the new address space.
 The ability to transfer that points-to information out of that modality is specific to a given kernel's design. 
-Kernels that map kernel memory into all address spaces would need to ensure and specify enough specific details about memory mappings to allow a 
-proof of an elimination rule for specific modally-constrained points-to assertions.
+Kernels that map kernel memory into all address spaces would need invariants
+that justified moving those assertions out of the other-space modality.
 % Following Spectre and Meltdown, this kernel design became less prevalent because speculative execution of accesses to kernel addresses could leak information even if the access did eventually cause a fault (the user/kernel mode permission check was done after fetching data from memory).
 Thus many modern kernels have reverted to the older kernel design where the kernel inhabits its own unique address space, and user processes have only enough extra material mapped in their address spaces to switch into the kernel (CPUs do not speculate past updates to \texttt{cr3}).
+\looseness=-1
 

### 2024-11-15 examples.tex
resolve line refs, fix some missing constant updates in 5.2

@@
 to map new pages into an address space by writing into an existing page table via a call
-\[\textsf{vaspace\_mappage(pte\_t *pml4, void *va,uintptr\_t fpaddr, uint64\_t perm )}\] in Figure \ref{fig:mapping_code}).
+\[\textsf{vaspace\_mappage(pte\_t *pml4, void *va,uintptr\_t fpaddr, uint64\_t perm )}\] in Figure \ref{fig:mapping_code}.
 To do so, with a given allocated a fresh page (\textsf{fpaddr}), then calculate the appropriate
-known-valid page table walks (\textsf{pte\_walkpgdir} Line X\todo{!!!} in Figure \ref{fig:mapping_code})  and update the appropriate L1 page table entry (permissions and frame address Lines in Figure \ref{fig:mapping_code});
+known-valid page table walks (via \textsf{walkpgdir} Line \ref{line:call_walkpgdir} in Figure \ref{fig:mapping_code})  and update 
+the appropriate L1 page table entry (Line \ref{line:l1entry_store} in Figure \ref{fig:mapping_code});
 unmapping is the reverse of the logic we discuss here.
@@
 %}
-\todo{ismail change the following spec base on the changes you have in the code}
 \begin{figure}\footnotesize
@@
 ... ;;setting up the stack      
- $\specline{\textsf{P} \ast \textsf{pml4} \mapsto_{\textsf{id}} \textsf{\_} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}}_{\rtv}$      
+ $\specline{\textsf{P} \ast \textsf{pml4} \mapsto_{\textsf{id}} \textsf{\_} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta} \ast \ulcorner\theta \; !!\; \vaddr = \texttt{None}\urcorner}_{\rtv}$
 ;; pte_t *pte = walkpgdir(pml4, a);
-mov    -0x8[rbp],rdi; mov    -0x28[rbp],rsi; mov    0x0,rax; callq  291 <vaspace_mappage+0x41>
+mov    -0x8[rbp],rdi; mov    -0x28[rbp],rsi; mov    0x0,rax; callq  291 <walkpgdir> (*\label{line:call_walkpgdir}*)
 ;;if(pte->present != 0) {
@@
 \ghostmaptoken{\textsf{id}}{(\mathsf{pml4\_entry})}{\textsf{4}}\ast \ghostmaptoken{\textsf{id}}{(\mathsf{plm3\_entry})}{\textsf{3}} \ast \ghostmaptoken{\textsf{id}}{(\mathsf{plm2\_entry})}{\textsf{2}}  \\ \forall_{i\in\textsf{0..511}} \; \textsf{table\_root(pml4\_entry}_{\textsf{pfn}}) \textsf{ + i * 8} \mapsto_{\textsf{id}} \textsf{3}  \\ \forall_{i\in\textsf{0..511}} \; \textsf{table\_root(pdp\_entry}_{\textsf{pfn}}) \textsf{ + i * 8} \mapsto_{\textsf{id}} \textsf{2}  \\ \forall_{i\in\textsf{0..511}} \; \textsf{table\_root(pd\_entry}_{\textsf{pfn}}) \textsf{ + i * 8} \mapsto_{\textsf{id}} \textsf{1}  \\  \textsf{pml4\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l4e\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l4e\_val})\urcorner \\  \textsf{pdp\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l3e\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l3e\_val})\urcorner \\  \textsf{pd\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l2e\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l2e\_val})\urcorner \\   \textsf{pt\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l1e\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l1e\_val})\urcorner    \end{array} \right)  =\textsf{R}_{\textsf{walk\_acc}}}$
-;;pte->pfn          = PTE_ADDR_TO_PFN(pa);
+;;pte->pfn          = PTE_ADDR_TO_PFN(fpaddr);
 mov    -0x18[rbp],rcx; shr    0xc,rcx; mov    -0x30[rbp],rdx; mov    [rdx],rsi ;; Load current entry in rsi
@@
 movabs 0xffffffffff,rdi; and    rdi,rcx; shl    0xc,rcx; movabs 0xfff0000000000fff,rdi
-and    rdi,rsi; or     rcx,rsi; mov    rsi,[rdx] ;; Store updated entry back to L1 entry
+and    rdi,rsi; or     rcx,rsi; mov    rsi,[rdx] ;; Store updated entry back to L1 entry (*\label{line:l1entry_store}*)
 ;;}
@@
 the existence of a memory location for an appropriate L1 table entry.
-The code uses a helper function \lstinline{pte\_walkpgdir} (discussed again in Section \ref{sec:traversing}).
-\textsf{pte\_walkpgdir}'s postcondition contains virtual \emph{PTE} pointsto assertions ($\mapsto_{\textsf{vpte}}$) \todo{colin, these are no more plain physical pointstos}  both for ensuring partial page table walk reaching the
+The code uses a helper function \lstinline{walkpgdir} (discussed again in Section \ref{sec:traversing}).
+\textsf{walkpgdir}'s postcondition contains virtual \emph{PTE} pointsto assertions ($\mapsto_{\textsf{vpte}}$)
+both for ensuring partial page table walk reaching the
 L1 entry (l1e) with the justification of higher levels of the page table exist, and for allowing access to the memory of the L1 entry via virtual address.
@@
 
-The crucial step is actually updating the L1 entry (Line X\todo{X} in Figure \ref{fig:mapping_code}),
+The crucial step is actually updating the L1 entry (Line \ref{line:l1entry_store} in Figure \ref{fig:mapping_code}),
 via the virtual address
-(\textsf{pte\_addr}) known to translate to the appropriate physical address, in our example the L1
-table entry address ($\textsf{addr\_L1}(\textsf{va, l1e})$), to hold the freshly
-allocated physical page address (\textsf{fpaddr}) in Line X.
-After this write, we have (Specification Line Y)
-\[\texttt{pte\_addr} \mapsto_{\texttt{vpte}}  \; \paddr\; \textsf{fpaddr}  \qquad\qquad (\textsf{by the PTE variant of Rule}~\TirNameStyle{WriteToVirtMemFromReg})\]
+(\textsf{pt\_entry+KERNBASE}) known to translate to the appropriate physical address, in our example the L1
+table entry address ($\textsf{addr\_L1}(\textsf{va, l1e})$).
+% , to hold the freshly
+% allocated physical page address (\textsf{fpaddr}) in Line X.
+After this write, we have
+% \[\texttt{pte\_addr} \mapsto_{\texttt{vpte}}  \; \paddr\; \textsf{fpaddr}  \qquad\qquad (\textsf{by the PTE variant of Rule}~\TirNameStyle{WriteToVirtMemFromReg})\]
+\[\texttt{pt\_entry+KERNBASE} \mapsto_{\texttt{vpte}}  \; \paddr\; (\textsf{fpaddr \& 0x1})  \quad (\textsf{by PTE variant of}~\TirNameStyle{WriteToVirtMemFromReg})\]
 
-At this point in the proof,  $\texttt{pte\_addr} \mapsto_{\texttt{vpte}}  \; \paddr\; \textsf{fpaddr})$ 
-contains  full ownerhsip of the L1 table entry physical pointsto $\paddr
-\mapsto_{\mathsf{p}} \textsf{fpaddr}$ . Then, we can split the full-ownership of that
+At this point in the proof, % $\texttt{pte\_addr} \mapsto_{\texttt{vpte}}  \; \paddr\; (\textsf{fpaddr})$ 
+the virtual PTE points-to
+contains  full ownership of the L1 table entry's backing physical memory.
+Then, we can split the full-ownership of that
 physical pointsto
-($\textsf{fpaddr}\mapsto_{\textsf{p}} \;\textsf{wzero 64}$) to obtain
-fractional ownership on it appropriate for a single address's share of an L1 entry (q4).
-\todo{ismail comment out the rest of this subsection and fix according to the new map-a-page}
+% ($\textsf{fpaddr}\mapsto_{\textsf{p}} \;\textsf{wzero 64}$) 
+into 512 pieces to obtain
+fractional ownership of the physical memory appropriate for a single virtual address's share of an L1 entry.
+% \todo{ismail comment out the rest of this subsection and fix according to the new map-a-page}
 \begin{comment}

### 2024-11-15 examples.tex
+

@@
 known-valid page table walks (via \textsf{walkpgdir} Line \ref{line:call_walkpgdir} in Figure \ref{fig:mapping_code})  and update 
-the appropriate L1 page table entry (Line \ref{line:l1entry_store} in Figure \ref{fig:mapping_code});
+the appropriate L1 page table entry (Line 11 in Figure \ref{fig:mapping_code});
 unmapping is the reverse of the logic we discuss here.
@@
 the modified data is at the specific physical location that will be used as the L1 entry.
-This is also guaranteed ($ \ulcorner
-\texttt{addr\_L1}(\vaddr,\entryo) = \paddr\urcorner$).
 % After obtaining a virtual address \textsf{pte\_addr} in \textsf{rax} backed 
@@
 
-The crucial step is actually updating the L1 entry (Line \ref{line:l1entry_store} in Figure \ref{fig:mapping_code}),
+The crucial and the only step in addition to traversing the page table in Figure \ref{walkpgdir} is actually updating the L1 entry (Line 11 in Figure \ref{fig:mapping_code}),
 via the virtual address
 (\textsf{pt\_entry+KERNBASE}) known to translate to the appropriate physical address, in our example the L1
-table entry address ($\textsf{addr\_L1}(\textsf{va, l1e})$).
+table entry address ($\textsf{addr\_to\_pfn(pa)}$).
 % , to hold the freshly
 % allocated physical page address (\textsf{fpaddr}) in Line X.
-After this write, we have
-% \[\texttt{pte\_addr} \mapsto_{\texttt{vpte}}  \; \paddr\; \textsf{fpaddr}  \qquad\qquad (\textsf{by the PTE variant of Rule}~\TirNameStyle{WriteToVirtMemFromReg})\]
-\[\texttt{pt\_entry+KERNBASE} \mapsto_{\texttt{vpte}}  \; \paddr\; (\textsf{fpaddr \& 0x1})  \quad (\textsf{by PTE variant of}~\TirNameStyle{WriteToVirtMemFromReg})\]
 
-At this point in the proof, % $\texttt{pte\_addr} \mapsto_{\texttt{vpte}}  \; \paddr\; (\textsf{fpaddr})$ 
-the virtual PTE points-to
-contains  full ownership of the L1 table entry's backing physical memory.
-Then, we can split the full-ownership of that
-physical pointsto
-% ($\textsf{fpaddr}\mapsto_{\textsf{p}} \;\textsf{wzero 64}$) 
-into 512 pieces to obtain
-fractional ownership of the physical memory appropriate for a single virtual address's share of an L1 entry.
-% \todo{ismail comment out the rest of this subsection and fix according to the new map-a-page}
-\begin{comment}
-Since we know $\texttt{addr\_L1 }(\vaddr,
-\entryo) \mapsto_{\mathsf{p}} \{q4\} \;\textsf{fpaddr}$) from
-$\mapsto_{\textsf{vpte}}$, after the split of the PTE points-to, and, thanks to the post-condition of the page-table walk (Specification of \textsf{pte\_walkpgdir} in Figure \ref{}), we know the existence of L4-L2 mappings reaching to the L1 exist, so that we can add the last mapping $\texttt{fpaddr} \mapsto_{p} (\texttt{wzero 64})$ to obtain the complete table-walk.
-\todo{with the virtual-pte-pointsto assertions (L4\_L1\_PointsTo($\vaddr$ l4e l3e l2e
-fpaddr)) in Line 25}
-
-Finally, we can insert $\vaddr$ into the ghost page-table-walk summarization
-map ($\theta$) to change its state from unmapped ($\ulcorner \theta
-\;!!\;\vaddr = \texttt{None}\urcorner$ (Line 28) and the
-precondition) to mapped (Line 29) using Iris' ghost-map update
-($\sqsubseteq$), and construct a PTE -points-to $\vaddr
-\mapsto_{\textsf{vpte}}\; \qfrac \;\fpaddr\;(\textsf{wzero 64})$ (Line 31) using $\sumwalkabs\vaddr\qfrac\fpaddr$ obtained from ghost
-page-table-walk insertions, $\ghostmaptoken{\delta{}s}{\rtv}{\delta}$ from unfolding the
-definition of $\mapsto_{\textsf{vpte}}$, and $\fpaddr
-\mapsto_{\textsf{p}} \textsf{wzero 64}$.
-\end{comment}
-Finally, by existentially quantifying away \textsf{fpaddr},
-we can shift the PTE points-to into a normal virtual points-to.
-
-For brevity this proof only maps the first word of the allocated page; the generalization to the full page
-is straightforward use of iterated separating conjunction to manage collections of clusters of 512 things
-(slices of the L1 entry physical ownership, virtual addresses at 8 byte increments, etc.).
+At this point in the proof, i.e. after the traversal to obtain the sound physical path resources ($\textsf{R}_{\textsf{walk\_acc}}$ Line 27 in Figure \ref{walkpgdir}), which coontains
+virtual PTE points-to relations along the path to the full ownership of the L1 table entry's backing physical memory, we update the L1 entry to point the fresh page, as we do for the non-existent entries in the page table traversal (Line 6 in Figure \ref{pteinitializespec}.
 \looseness=-1

### 2024-11-15 intro.tex
+

@@
 We prove the soundness of our \textsf{vProp} logic with respect to a RISC-like fragment of \textsf{AMD64} instructions.
-We verify simplified versions of several critical virtual-memory-related pieces of OS functionality, 
-including mapping  pages, switching address spaces, and converting physical addresses to virtual addresses for
-software page table walks (as when mapping pages). These examples demonstrate the suitability and flexibility of
+We verify simplified versions of several critical virtual-memory-related pieces of OS functionality, most importantly:
+switching address spaces, converting physical addresses to virtual addresses for
+software page table walks, and complete page table walk which backbones other essential functionalities of
+virtual-memory which we discuss for mapping and unmapping pages. These examples demonstrate the suitability and flexibility of
 a modal treatment of address spaces: each has a concise specification, each example either goes beyond the technical

### 2025-03-20 intro.tex
tweak abstract, intro, just slightly

--- 2024-11-15 intro.tex
@@
 This is what enables separation of process memory resources:
-the OS manipulates hardware functionality to ensure that any attempt by a process to access memory not explicitly granted 
-to it by the kernel will fail. But this is complicated by the fact that 
+the OS manipulates hardware functionality to ensure 
+program memory accesses only succeed if the kernel has granted the program access to the accessed address.
+But this is complicated by the fact that 
 %the OS and hardware can also enable, shared (overlapping) access to physical memory regions; the fact that the kernel data structures themselves are accessed via virtual memory addresses; 
@@
 control over these mappings of virtual to physical addresses is itself mediated by \emph{in-memory data structures}, 
-which the kernel still accesses via virtual address, leading to indirect cycles.
+which the kernel also accesses via virtual address, leading to indirect cycles.
 \looseness=-1
@@
 reliant upon or manipulating those memory configurations in the presence of in-memory page tables, the primary
-memory protection mechanism across Intel/AMD's x86-64 processors, ARM's application class processors including
-AArch64 CPUs, POWER, RISC-V, and other architectures.
+memory protection mechanism across Intel/AMD's x86-64 processors; ARM's application class processors including
+AArch64 CPUs; as well as POWER, RISC-V, and other architectures.
 We prove the soundness of our \textsf{vProp} logic with respect to a RISC-like fragment of \textsf{AMD64} instructions.
 We verify simplified versions of several critical virtual-memory-related pieces of OS functionality, 
-including mapping  pages, switching address spaces, and converting physical addresses to virtual addresses for
-software page table walks (as when mapping pages). These examples demonstrate the suitability and flexibility of
+including:
+switching address spaces, converting physical addresses to virtual addresses for
+software page table walks, and complete page table walk which backbones other essential functionalities of
+virtual-memory which we discuss for mapping and unmapping pages. These examples demonstrate the suitability and flexibility of
 a modal treatment of address spaces: each has a concise specification, each example either goes beyond the technical
 capabilities of prior logics, or revisits an example from prior work with more details (treating parts of the kernel
-invariant never addressed by prior work) and fewer assumptions (e.g., verifying virtual-to-physical conversion rather
-than assuming it has already occurred).
+invariant never addressed by prior work) and fewer assumptions (e.g., verifying the reverse physical-to-virtual 
+address conversion rather than assuming it has already occurred).
 

### 2025-03-21 background.tex
misc tweaks and clarifications up to early section 4, added new bg section on separation logic that feels like it is not quite right

@@
 \label{sec:background}
+This section briefly recalls both fundamentals of page-table-based address translation in
+general-puspose OS kernels, 
+high-level background on separation logics and the \iris~\cite{jung2018iris} framework
+we build on,
+and some material on modal logic that informs our development.
+
 \subsection{Machine Model}
@@
 
+\subsection{Separation Logic}
+\label{sec:seplogic}
+\todo[inline]{Colin says: I added this b/c prior systems reviewers wanted more verification background...
+not sure how I feel about it}
+Separation logic~\cite{reynolds02} is a descendant of classic Hoare logic~\cite{hoare69},
+where in addition to pre- and postcondition assertions, assertions themselves pick up a
+\emph{separating conjunction} operator $\ast$, such that an assertion $A\ast B$ means $A$ and $B$ are true
+of disjoint pieces of state. This allows for local reasoning about updates, because it articulates
+that updates to the state backing $A$ do not affect the truth of $B$. This is classically demonstrated
+through the \emph{points-to} assertion: $l\mapsto v$ asserts that the memory cell at address $l$ holds value $v$:
+knowing $x\mapsto 3\ast y\mapsto 4$ and writing through $x$ means information about $y$ is preserved:
+$\{x\mapsto 3\ast P\}\;x\mathrel{:=}5\;\{x\mapsto 5 \ast P\}$ can be derived for any $P$.
+
+We build on the \iris~\cite{jung2018iris} separation logic framework,
+an abstract separation logic embedded in Rocq, which is useful for both metatheoretical work
+and interactive correctness proofs using the logic. Given an operational
+semantics structured a certain way (in our case, semantics for
+a fragment of x86-64 assembly including address translation),
+if a small number of ``glue'' lemmas are proven, \iris
+provides a ready-made separation logic with a number of advanced features, including
+higher-order ghost state and impredicative invariants, for no additional work.
+
+We suppress many details of \iris assertions for brevity, but briefly note
+that because \iris is an embedded framework in Rocq, proofs in \iris-derived
+logics (including ours) often encapsulate raw Rocq assertions: $\ulcorner P \urcorner$ is an embedding
+of the Rocq assertion $P$ into an \iris assertion (used for things like equality and other
+general pure logical assertions). Newly-defined \iris assertions are in fact
+Rocq terms of a particular type, rather than being drawn from a fixed vocabulary.
+We introduce other details as they arise.
+
+Of particular interest to us, \iris's machinery is largely agnostic to the particular choice of
+\emph{resource algebra}\footnote{A modern descendant of the venerable partial commutative monoid used
+to abstractly model earlier separation logics~\cite{calcagno2007local}, 
+with additions to support the step-indexing required to solve the recursive domain equations
+that arise with higher-order ghost state and impredicativity~\cite{birkedal2011step,hobor2010theory}.
+}
+used to give algebraic semantics to the connectives of the assertion language.
+This means we can drop in an alternative model that naturally supports working with respect to
+a locally-fixed context, like an address space.
+
+% Also relevant is that \textsc{Iris} naturally decomposes the traditional Hoare triple $\{P\}\;C\;\{Q\}$
+% into the precondition implying the weakest precondition of the program with respect to the postcondition ---
+% $P \wand \textsf{wp}\;C\;\{v\ldotp Q\}$.\footnote{An idea with long history~\cite{pratt1976semantical}.}
+% This is a natural fit for assembly-level verification, where the \emph{Hoare double}
+
+An important limitation here is that to date, every separation logic
+has assumed that all pointer addresses are meant for use in a single address space, which avoided the
+problem of tracking that certain points-to assertions are true only for certain address spaces, but not others.
+
 \subsection{Modal Logic}
@@
 % application. 
-We develop our ideas in the Iris program logic~\cite{jung2018iris}, as a result exploring additional subtleties 
-that arise where the modality itself may entail ownership of resources, as well as interactions between our hybrid-style 
+Our development atop \iris~\cite{jung2018iris} needs to explore some additional subtleties 
+that arise where the modality itself may entail ownership of resources, 
+as well as interactions between our hybrid-style 
 modality and substructural rules.  

### 2025-03-21 programlogic.tex
misc tweaks and clarifications up to early section 4, added new bg section on separation logic that feels like it is not quite right

@@
 % We build up our context-dependent assertions from two basic types of assertions that do not care about address spaces.
-To define and use virtual points-to assertions, we require two more basic assertions:
+To define and use virtual points-to assertions, we require two basic assertions that ignore
+the current address space:
 
 \paragraph{Register points-to} 
-The assertion $\textsf{r}\;\mapsto_{\textsf{r}}\{q\}\;\textsf{rv}$ ensures the ownership of the register $\rg$ naming the register value $\rv$. 
+The assertion $\textsf{r}\;\mapsto_{\textsf{r}}\{q\}\;\textsf{rv}$ ensures the ownership of the register $\rg$ containing the register value $\rv$. 
 The fraction $\qfrac$ with value 1 asserts the unique ownership of the register mapping, and grants update permission on it, 
@@
   The fractions \lstinline|q1| through \lstinline|q4| represent the fractional ownership of each entry based on how many
-  word-aligned addresses might need to share the entry.
+  word-aligned addresses might need to share the entry ---  $(\frac{1}{512})^n$ for each level $n$.
 }
@@
   from the virtual pointsto relation, replacing it with ghost state that merely guarantees the address translation would succeed.
-  Iris includes a \emph{ghost map} construction which we use to track mappings from
+  \iris includes a \emph{ghost map} construction which we use to track mappings from
   virtual addresses to the physical addresses they translate to.
@@
   resource indeed maps $k$ to $v$.
-These properties are captured by Iris rules using the token to update or draw conclusions about the logical map:\footnote{Iris ghost maps lack established notation in the literature,
+These properties are captured by \iris rules using the token to update or draw conclusions about the logical map:\footnote{Iris ghost maps lack established notation in the literature,
   but the syntax we use captures the details of \texttt{iris.base\_logic.lib.ghost\_map}.}

### 2025-03-21 syntaxsemantics.tex
misc tweaks and clarifications up to early section 4, added new bg section on separation logic that feels like it is not quite right

@@
 \label{sec:syntax}
-To develop our core logical ideas, we instantiate \textsf{Iris} with a simple language for streams of instructions, 
+To develop our core logical ideas, we instantiate \iris with a simple language for streams of instructions, 
 and a logical machine model corresponding to execution of x86-64 assembly instructions with virtual memory enabled on the 
@@
 Operationally, each first translates the virtual address stored in register $r_m$ to a physical address,
-then either updates the memory at that physical location with the contents of $r_r$ (for store) or
+then either updates the memory at that physical location with the contents of $r_r$ (for store)
 or updates $r_r$ with the contents of that memory (for load).

### 2025-03-21 background.tex
minor bg tweaks, significant adjustment of section 4.2, which was hard to follow because it jumped back and forth between different figures and was missing some explanation

@@
 \label{sec:seplogic}
-\todo[inline]{Colin says: I added this b/c prior systems reviewers wanted more verification background...
+\todo{Colin says: I added this b/c prior systems reviewers wanted more verification background...
 not sure how I feel about it}
@@
 We suppress many details of \iris assertions for brevity, but briefly note
-that because \iris is an embedded framework in Rocq, proofs in \iris-derived
-logics (including ours) often encapsulate raw Rocq assertions: $\ulcorner P \urcorner$ is an embedding
-of the Rocq assertion $P$ into an \iris assertion (used for things like equality and other
+a few recurring details used heavily in \iris but not necessarily in traditional
+separation logics.
+Because \iris is an embedded framework in \rocq, proofs in \iris-derived
+logics (including ours) often encapsulate raw \rocq assertions: $\ulcorner P \urcorner$ is an embedding
+of the \rocq assertion $P$ into an \iris assertion (used for things like equality and other
 general pure logical assertions). Newly-defined \iris assertions are in fact
-Rocq terms of a particular type, rather than being drawn from a fixed vocabulary.
+\rocq terms of a particular type, rather than being drawn from a fixed vocabulary.
 We introduce other details as they arise.
 
-Of particular interest to us, \iris's machinery is largely agnostic to the particular choice of
-\emph{resource algebra}\footnote{A modern descendant of the venerable partial commutative monoid used
-to abstractly model earlier separation logics~\cite{calcagno2007local}, 
-with additions to support the step-indexing required to solve the recursive domain equations
-that arise with higher-order ghost state and impredicativity~\cite{birkedal2011step,hobor2010theory}.
+\iris includes two forms of implication. The magic wand operator $\wand$ is an affine implication:
+$A\wand B$ describes a resource which, if combined with a resource satisfying $A$, will satisfy $B$.
+Notably, this implication involves no changes to ghost state. \iris, building on the Views framework~\cite{Dinsdale-Young2013Views},
+also includes a \emph{view shift} operator $\vs$ which models updates to ghost state: $A \vs B$
+means resources satisfying $A$ may be transformed into resources satisfying $B$, intuitively by updating only ghost state.\footnote{\iris
+experts may note that this is \emph{technically} mildly misleading given how \iris's update modalities and weakest precondition
+are defined, but
+this is adequate intuition for non-experts in \iris to follow their use.
 }
-used to give algebraic semantics to the connectives of the assertion language.
-This means we can drop in an alternative model that naturally supports working with respect to
-a locally-fixed context, like an address space.
+\looseness=-1
+
+% Of particular interest to us, \iris's machinery is largely agnostic to the particular choice of
+% \emph{resource algebra}\footnote{A modern descendant of the venerable partial commutative monoid used
+% to abstractly model earlier separation logics~\cite{calcagno2007local}, 
+% with additions to support the step-indexing required to solve the recursive domain equations
+% that arise with higher-order ghost state and impredicativity~\cite{birkedal2011step,hobor2010theory}.
+% }
+% used to give algebraic semantics to the connectives of the assertion language.
+% This means we can drop in an alternative model that naturally supports working with respect to
+% a locally-fixed context, like an address space.
 

### 2025-03-21 programlogic.tex
minor bg tweaks, significant adjustment of section 4.2, which was hard to follow because it jumped back and forth between different figures and was missing some explanation

@@
   \iris experts may notice our \lstinline|-b>| resembles another pointwise lifting already in  \iris~\cite{dang2019rustbelt,dang2022compass}. 
-  This similarity is real, but the existing lifting does not work with indexed \coq types
+  This similarity is real, but the existing lifting does not appear to work with indexed \coq types
   like our \lstinline|word n| as a domain.
@@
 This is the (\textsc{Coq}) type of assertions in our logic.
-Most constructs in Iris's base logic are defined with respect to any BI-algebra (of \textsc{Coq} type \lstinline|bi|), so automatically
+Most constructs in Iris's base logic are defined with respect to any BI-algebra (of \coq type \lstinline|bi|), so automatically
 carry over to our derived logic.
@@
 
+  \begin{figure}
+\footnotesize
+  \begin{mathpar}
+  \inferrule*[right=\footnotesize GhostMapUpdate]{ }{\mathsf{GhostMap}(\gamma,\theta) \ast \ghostmaptoken{\gamma}{\mathsf{pa}}{\mathsf{va}} \vs 
+  \mathsf{GhostMap}(\gamma,\theta[\mathsf{va}\mapsto\mathsf{pa'}]) \ast \ghostmaptoken{\gamma}{\mathsf{va}}{\mathsf{pa}'} 
+  }
+  \and
+  \inferrule*[right=\footnotesize GhostMapLookup]{ }{\mathsf{GhostMap}(\gamma,\theta) \ast \ghostmaptoken{\gamma}{\mathsf{va}}{\mathsf{pa}} \wand \ulcorner \theta(\mathsf{va})=\mathsf{Some}(\mathsf{pa}) \urcorner }
+  \end{mathpar}
+  \caption{\iris rules for ghost maps}
+  \label{fig:ghostmaps}
+  \end{figure}
+
   We separate the physical page-table-walk  
@@
   \iris includes a \emph{ghost map} construction which we use to track mappings from
-  virtual addresses to the physical addresses they translate to.
-  The implementation details in terms of named invariants and various resource algebras are not important to our presentation, so
-  we treat the maps abstractly.
+  virtual addresses to the physical addresses they translate to as a piece of ghost state.
   The map includes for each key in the map (i.e.,
@@
   named $\gamma$. The existence of such a token implies that the actual map $\theta$ tracked by a corresponding $\mathsf{GhostMap}(\gamma,\theta)$
-  resource indeed maps $k$ to $v$.
-These properties are captured by \iris rules using the token to update or draw conclusions about the logical map:\footnote{Iris ghost maps lack established notation in the literature,
+  resource indeed maps $k$ to $v$. These properties are captured by key \iris rules in Figure \ref{fig:ghostmaps}.\footnote{\iris ghost maps lack established notation in the literature,
   but the syntax we use captures the details of \texttt{iris.base\_logic.lib.ghost\_map}.}
-  \begin{mathpar}
-  \inferrule{ }{\mathsf{GhostMap}(\gamma,\theta) \ast \ghostmaptoken{\gamma}{\mathsf{pa}}{\mathsf{va}} \vs 
-  \mathsf{GhostMap}(\gamma,\theta[\mathsf{pa}\mapsto\mathsf{va'}]) \ast \ghostmaptoken{\gamma}{\mathsf{pa}}{\mathsf{va}'} 
-  }
-  \and
-  \inferrule{ }{\mathsf{GhostMap}(\gamma,\theta) \ast \ghostmaptoken{\gamma}{\mathsf{pa}}{\mathsf{va}} \wand \ulcorner \theta(\mathsf{pa})=\mathsf{Some}(\mathsf{va}) \urcorner }
-  \end{mathpar}
+  There are other rules, but these two are most important for explaining ghost maps.
+  \textsc{GhostMapUpdate} says that ownership of the actual ghost map with ghost name $\gamma$ and map contents $\theta$,
+  and a full (non-fractional) token witnessing that $\theta$ maps \textsf{pa} to \textsf{va} permits an update to the ghost map's state,
+  changing the map and replacing the token to represent the new value.
+  \textsc{GhostMapLookup} allows using the same information to simply conclude that the mapping indicated by the token is true.
 
@@
   % requires full ownership of the token. 
-  Then the \emph{virtual memory manager's invariant} ensures that for each $\ghostmaptoken{\gamma}{va}{pa}$ mapping
-  in this map, there are \emph{physical} resources sufficient to ensure that address translation for $va$
-  will resolve in the hardware to $pa$ via --- $\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$.
-  This allows modification of the page tables themselves as long as the changes do not invalidate overall virtual-to-physical
-  translation.
-  Consistency follows from replacing the page table walk with ownership of the ghost map token for the relevant ghost map
-  entry, preventing the overall translation from being changed separately from the virtual points-to.
+  The \emph{virtual memory manager's invariant} ensures that for each $\ghostmaptoken{\gamma}{\vaddr}{\paddr}$ mapping
+  in this map, there are \emph{physical} resources sufficient to ensure that address translation for $\vaddr$
+  will resolve in the hardware to $\paddr$ --- via $\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$.
+  % This allows modification of the page tables themselves as long as the changes do not invalidate overall virtual-to-physical
+  % translation.
+  % Consistency follows from replacing the explicit page table walk of Figure \ref{fig:strongvirtualpointsto}'s
+  % definition with ownership of the ghost map token for the relevant ghost map
+  % entry, as in Figure \ref{fig:virtualpointstosharing}.
+  % This allows the overall translation to be updated separately from the virtual points-to.
   \looseness=-1
@@
   % \[\mathcal{A}\textsf{bsPTableWalk} \stackrel{\triangle}{=} \textsc{Auth} (\; \mathcal{W}_{64} \;\rightarrow_{\textrm{fin}} \;  ( (\textsc{Frac }, \mathord{+}) \times (\textsc{Agree } \Loc,\mathord{=}) ))\]
-For clarity we refer to the ghost map summarizing virtual-to-physical translations by 
+For clarity we refer to the specific ghost map summarizing virtual-to-physical translations by \\
 \centerline{$\mathcal{A}\textsf{bsPTableWalk}(\delta,\theta) \stackrel{\triangle}{=} \mathsf{GhostMap}(\delta,\theta)$}\\
@@
 and keep this in a per-address-space invariant described shortly.
-We then replace the physical
- $\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$ in the virtual points-to definition
-with ownership of the token \ghostmaptoken{\delta}{\mathsf{va}}{\mathsf{paddr}}, %ghost-map ($ \sumwalkabs\vaddr\qfrac\paddr$),
-guaranteeing that ghost map contains an mapping from the virtual address ($\vaddr$) to a physical address ($\paddr$),
-and thus that the per-address-space invariant contains the physical resources that guarantee the hardware resolves the translation.
-
-We place the authorative ownership of the ghost translation $\mathcal{A}\textsf{PTableWalk}$ in a per-address-space invariant
-$\mathcal{I}$\textsf{ASpace} (Figure \ref{fig:peraspaceinvariant}). 
-$\mathcal{I}$\textsf{ASpace} alone allows changes to the page tables 
-that preserve overall virtual-to-physical translations.
-When combined with the
-fragment stored in the virtual points-to (Figure \ref{fig:virtualpointstosharing}),
-changes in the virtual-to-physical translations become possible.
+We then replace the physical traversal
+ $\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$ in Figure \ref{fig:strongvirtualpointsto}'s virtual points-to definition
+with ownership of the token \ghostmaptoken{\delta}{\vaddr}{\paddr}, %ghost-map ($ \sumwalkabs\vaddr\qfrac\paddr$),
+yielding Figure \ref{fig:virtualpointstosharing}'s definition.
+This new definition guarantees that the ghost map contains an mapping from the virtual address ($\vaddr$) to a physical address ($\paddr$),
+and thus that the per-address-space invariant will contain the physical resources that guarantee the hardware resolves the translation.
 
@@
 \vspace{-1em}
-\caption{Global Address-Space Invariant with a fixed global map of address-space names $m$}
+\caption{Per-address-space invariant with a fixed global map of address-space names $m$}
   \label{fig:peraspaceinvariant}
@@
 
-Our logic is the first to address the existence of multiple address spaces, so also requires
-a way to look up
-the ghost map of virtual-to-physical translations for an address space (the ghost name $\delta$ in Figure \ref{fig:virtualpointstosharing}).
-% for each address space must be locateable from a representation of the specific address space:
-% This means we also require a way to locate the map for the current address space from 
-We support lookup by the current root page table pointer stored in \lstinline|cr3|.
-\looseness=-1
+We place the authorative ownership of the ghost translation $\mathcal{A}\textsf{PTableWalk}$ in a per-address-space invariant
+$\mathcal{I}$\textsf{ASpace} (Figure \ref{fig:peraspaceinvariant}). 
+$\mathcal{I}$\textsf{ASpace} alone allows changes to the page tables 
+that preserve overall virtual-to-physical translations.
+When combined with the
+fragment stored in the virtual points-to (Figure \ref{fig:virtualpointstosharing}),
+changes in the virtual-to-physical translations become possible without needing to track all virtual points-to assertions.
 
-Figure \ref{fig:peraspaceinvariant} describes the resources for each address space.
+% Figure \ref{fig:peraspaceinvariant} describes the resources for each address space.
+In Figure \ref{fig:peraspaceinvariant},
 $\theta$ is the logical map from virtual addresses to the physical addresses they should translate to, 
@@
 is the ghost name for this map (again, for the current address space).
-$m$ is the (full) logical map from the various page table roots to ghost names for the 
-per-address-space mappings like $\theta$.
-It is tracked by a ghost map named by the ghost name $\delta{}s$ in
-Figure \ref{fig:virtualpointstosharing}; there is only one such map in the system.
+Because our logic is the first to address the existence of multiple address spaces,
+it must ensure that $\theta$ actually corresponds to the intended ghost map for the current address space.
+This requires looking up the ghost name (i.e., $\delta$) for the current address space,
+which is handled by $m$ --- another logical map (tied to ghost state shortly)
+from various page table roots to ghost names for the per-address-space ghost maps discussed above.
+$\mathcal{I}\textsf{ASpace}(\theta,m)$ then stores the physical resources for the current address space mappings,
+ensuring that the current address space is valid (has an entry in $m$ whose corresponding
+authoritative map matches $\theta$), and asserting ownership of the appropriate fractional ownership
+of the physical page table walk for each virtual-to-physical translation in $\Theta$ (for the current address space).
+
+Virtual points-to assertions must also ensure that they are tied to a valid address space, and because they
+use ghost map tokens to ensure the virtual-to-physical translation exists in a specific ghost map, they must also
+name that map. In Figure \ref{fig:virtualpointstosharing}, this is done with a ghost map token for \emph{another} ghost map,
+one shared across the kernel, holding $m$. Our virtual points-to assertion existentially quanfies the address-space-specific
+ghost map $\delta$ mapping virtual addresses to physical addresses, and ensures it exists with a token for the distinguished
+global (unique) ghost map named by $\delta{s}$.
+% the ghost map of virtual-to-physical translations for a particular address space (the ghost name $\delta$ in Figure \ref{fig:virtualpointstosharing}).
+% % for each address space must be locateable from a representation of the specific address space:
+% % This means we also require a way to locate the map for the current address space from 
+% We support lookup by the current root page table pointer stored in \lstinline|cr3|.
 \looseness=-1
-% This is why Figure \ref{fig:virtualpointstosharing} asserts there is some ghost name in that (global) map
-% for the current address space.
-% The global map itself is tracked by a resource algebra 
-% \[\mathcal{A}\textsf{spaces} \stackrel{\triangle}{=} \textsc{Auth} (\; \mathcal{W}_{64} \;\rightarrow_{\textrm{fin}} \;  ( (\textsc{Frac }, \mathord{+}) \times (\textsc{Agree } \textsc{GhostName},\mathord{=}) )) \]
-
-$\mathcal{I}\textsf{ASpace}(\Theta,m)$ is then the resources for the current address space,
-ensuring that the current address space is valid (has an entry in $m$ whose corresponding
-authoritative map matches $\Theta$), and asserting ownership of the appropriate fractional ownership
-of the physical page table walk for each virtual-to-physical translation in $\Theta$ (for the current address space).
 

### 2025-03-21 programlogic.tex
finish pass through section 4

@@
     % \underbrace{(\lambda\mathit{cr3val}\ldotp\sumapaces{\mathit{cr3val}}\delta)}_\text{Find addr.\;space invariant} \ast 
-    \underbrace{(\lambda\mathit{cr3val}\ldotp\ghostmaptoken{\delta{}s}{cr3val}{\delta})}_\text{Find addr.\;space invariant} \ast 
+    \underbrace{(\lambda\mathit{cr3val}\ldotp\exists \epsilon\ldotp\fracghostmaptoken{\delta{}s}{cr3val}{\delta}{\epsilon})}_\text{Find addr.\;space invariant} \ast 
   % \underbrace{\sumwalkabs\vaddr\qfrac\paddr }_\text{Ghost translation}\ast 
@@
 \vspace{-1em}
-\caption{Virtual-Pointsto for Sharing Pages}
+\caption{Virtual-Points-to for Sharing Pages}
   \label{fig:virtualpointstosharing}
@@
 the page table walk for the relevant address ($\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$)
-from the invariant ($\mathcal{I}\mathsf{ASpace}$), prove that the page table walk succeeds and
-memory or registers are updated appropriately, before re-packing the invariant and virtual points-to resources.
+from the invariant ($\mathcal{I}\mathsf{ASpace}$), prove that the page table walk succeeds
+and that memory or registers are updated appropriately, before re-packing the invariant and virtual points-to resources.
+
+The memory access rules model updates to the instruction pointer \lstinline|rip| using a metafunction \textsf{MovLen}.
+\textsf{MovLen} returns how long the instruction encoding for the corresponding \lstinline|mov| is;
+x86-64 instruction encodings are often longer for instructions using registers that are absent from the 32-bit
+x86 ISA that preceded x86-64.
 
@@
 \subsection{Soundness}
-Our rules from Figure \ref{fig:wpdamd} are proven sound in Iris against an assembly-level hardware model
+\label{sec:soundness}
+Our rules from Figure \ref{fig:wpdamd} are proven sound in \iris against an assembly-level hardware model
 implementing a fragment of x86-64 including 64-bit address translation with 4-level page tables.
@@
 axiomatized (with completely standard specifications~\cite{ni2007contexts,Chlipala2013Bedrock}) 
-because Iris's built-in machinery does not provide
+because \iris's built-in machinery does not provide
 convenient ways to discard the current continuation; adaptation of others'
-approaches~\cite{de2023type} is ongoing work.
+approaches~\cite{de2023type} is future work.
 Our soundness proofs for all other instructions (including, critically, all memory accesses)

### 2025-03-21 background.tex
revisions up through and including 5.1.3

@@
 of the \rocq assertion $P$ into an \iris assertion (used for things like equality and other
-general pure logical assertions). Newly-defined \iris assertions are in fact
+general pure logical assertions), similar to prior \rocq-embedded program logics~\cite{Chlipala2013Bedrock}. Newly-defined \iris assertions are in fact
 \rocq terms of a particular type, rather than being drawn from a fixed vocabulary.

### 2025-03-21 examples.tex
revisions up through and including 5.1.3

@@
 converting the physical addresses found in intermediate entries into the corresponding virtual addresses
-that can be used for access;
+that can be used for memory access;
 installing the new mapping;
@@
 Second, we formally verify a switch into a new address space as part of a task switch,
-the first such verification handling both old and new processes' assertions at the time of the switch.
+the first such verification handling both old and new processes' assertions (in different address spaces) at the time of the switch.
 
@@
 
-Note that the specification does \emph{not} assume a specific page table --- logical parameter \textsf{v} represents the level
+Note that the specification does \emph{not} assume a specific page table level --- logical parameter \textsf{v} represents the level
 of the entry passed as an argument, and this code
@@
 assumes it is given a virtual \emph{vpte-pointsto}
-% \footnote{A \emph{vpte-pointsto}
-% $\vaddr\mapsto_{\textsf{vpte,qfrac}} \textsf{\paddr~\val}$ is a virtual points-to granting
-% access to virtual address $\vaddr$, which is known to hold \textsf{\val}, but additionally exposing the physical address
-% $\paddr$ that the virtual address translates to. This is commonly used, as the name suggests, when updating PTEs,
-% where we require assurance that we are accessing the intended PTE.
-% Its definition is just like the regular virtual points-to in Figure \ref{fig:virtualpointstosharing},
-% except taking the physical address as a parameter rather than existentially quantifying it.
-% }
-(a virtual points-to exposing the underlying physical address; see Section \ref{sec:mapnew}) granting access to the specified entry,
+(a virtual points-to exposing the underlying physical address instead of existentially quantifying it;
+ see Section \ref{sec:mapnew}) granting access to the specified entry,
 but its postcondition does not yield new virtual points-to assertions!
-Instead it merely computes the base address of the next table, and returns adequate capabilities (discussed in Section \ref{subsec:identitymappings})
-for the \emph{caller} to construct a vpte-pointsto for the next table level (if this is not an L1 entry ---
+Instead it merely computes the base virtual address of the next table, and returns adequate capabilities (discussed in Section \ref{subsec:identitymappings})
+for the \emph{caller} to construct a vpte-pointsto for any entry of the next table level (if this is not an L1 entry ---
 the caller knows which level of the table this is for).
@@
 Within \textsf{get\_next\_table}, after a standard function prologue, the code 
-loads the entry \textsf{entry} pointed to by the argument.
+loads the entry pointed to by the argument (logical variable \textsf{entry} in the proof outline).
 This is a page table entry: a 64-bit word divided into bit-fields for
@@
 
-Lines \ref{line:mask_present}--\ref{line:check_entry_present} check if the entry's ``present'' bit is set.
+\ifPLDI
+Line \ref{line:mask_present} checks % In the condensed figure, it's all on one line
+\else
+Lines \ref{line:mask_present}--\ref{line:check_entry_present} check
+\fi
+if the entry's ``present'' bit is set.
 If it is zero, a new page must be allocated for the next level of the table --- which is done by the fall-through
@@
 and    0x1,rdi;  mov    rdi,rax; cmp    0x0,rax ;;  if (!entry->present) {(* \label{line:check_entry_present} *)(* \label{line:mask_present} *)
-jne    161 <pte_get_next_table+0xa1> (* \label{line:check_entry_present_jump} *) ;; Jump if the present bit is not zero
+jne    161 <pte_get_next_table+0xa1> (* \label{line:check_entry_present_jump} *) ;; Jump if the present bit is not zero; no need to allocate next level
 $\specline{\ldots \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val \& 0x1} }_{\rtv}$
@@
 maintain invariants that enable very fast P2V conversions (rather than adding another data structure).
-Most kernels maintain an invariant on their page tables that the virtual address of any page used for a page table lives at a virtual address
-whose value is \emph{a constant offset from the physical address} --- a practice sometimes referred to as \emph{identity mapping} 
+Most kernels maintain an invariant on their page tables that the virtual address of any page used for a page table 
+% lives at a virtual address whose value 
+is \emph{a constant offset from the physical address} --- a practice sometimes referred to as \emph{identity mapping} 
 (even though the physical-to-virtual translation
@@
 and for valid entries (with the present bit set) above L1, ghost map tokens for every entry in the table pointed to by the entry, which can be used
-to repeat the process one level down. i
-(L1 entries point to data pages, whose physical memory ownership resides in some virtual points-to).
-The assertion on Line \ref{line:conditional_children} comes from the invariant one level up; 
+to repeat the process one level down. 
+% (L1 entries point to data pages, whose physical memory ownership resides in some virtual points-to).
+The assertion on Line \ref{line:conditional_children} of Figure \ref{fig:calltopteinitialize} comes from the invariant one level up; 
 if the valid bit is set,
@@
 which requires full permission. Fortunately, the entry can only be in use if its valid bit is set; if the valid bit is not set we know
-that no virtual points-to assertions in $\delta$/$\theta$ have any partial ownership.
+that no virtual points-to entry in $\delta$/$\theta$ holds any partial ownership.
 Thus we use the invariant portion annotated as ``Entry validity'' in Figure \ref{fig:peraspaceinvariant_with_p2v_extension} to capture this:
@@
 together with other facts entails $\textsf{qfrac} = 1$ at Line \ref{line:after_concluding_qfrac1},
-and permits storing a new entry (in ellided code around Line \ref{line:install_new_entry})
+and permits storing a new entry (in ellided code around Line \ref{line:install_new_entry}).
 \looseness=-1
@@
  After obtaining the identity mapping for \textsf{entry}, we are able to load the \textsf{entry\_val} into \textsf{rdi}, and check the presence bit through
- % \[ \begin{array}{l}
- %   \mathsf{and }\; 0x1, \;rdi \\
- %   \mathsf{mov } \;\mathsf{rdi}, \; \mathsf{rax} \\
- %   \mathsf{cmp } \; 0x0, \;\mathsf{rax}
- % \end{array}
- % \] 
-Lines \ref{line:mask_present}--\ref{line:check_entry_present} in Figure \ref{fig:calltopteinitialize}.
-Accessing the presence bit and checking the value allows us to exploit the side condition (*) when verifying the allocation
-path for when the entry is invalid (Lines \ref{line:alloc_path_start}--\ref{line:alloc_path_end} in Figure \ref{fig:calltopteinitialize}).
-This operation is subtle: the operation requires that the relevant table entry is readable, but the exact portion of ownership 
+\ifPLDI
+Line \ref{line:mask_present} % in condensed version, all on same line
+\else
+Lines \ref{line:mask_present}--\ref{line:check_entry_present} 
+\fi
+in Figure \ref{fig:calltopteinitialize}.
+Accessing the presence bit and checking the value allows us to exploit the condition (*) that was just discussed when verifying the allocation
+path (i.e., when the entry is invalid  and Lines \ref{line:alloc_path_start}--\ref{line:alloc_path_end} in Figure \ref{fig:calltopteinitialize}
+must allocate the next level of tables).
+This operation is subtle. To reiterate: the operation requires that the relevant table entry is readable, but the exact portion of ownership 
 returned must be determined by inspecting the valid bit of the value in memory --- so full ownership is returned only for unused entries.
@@
 If the entry is not set, \textsf{pte\_initialize} (Line \ref{line:call_to_pte_initialize} in Figure \ref{fig:calltopteinitialize}) 
-allocates a physical page (internally utilizing the only axiomatized call in our case studies, the page-allocator's \textsf{kalloc}, 
+allocates a physical page (internally utilizing the only unverified (trusted) code in our case studies, the page-allocator's \textsf{kalloc},\footnote{
+  This is an allocator for regions of pre-zeroed physical memory that is mapped, but not accessed by the allocator itself,
+  as is typical for slab allocators~\cite{bonwick1994slab}.
+  Its verification would be similar to verifying a usermode \textsf{malloc} verifications~\cite{Chlipala2013Bedrock,wickerson2010explicit},
+  just with additional invariants on the memory pool.
+} 
 on Line \ref{line:call_to_kalloc} in Figure \ref{pteinitializespec}). 
@@
 entry construction via the page-frame portion of the PTE. In other words, the side condition, (*),
- formalizes that any access to the entry with these resources as \textit{invalid}, until the entry is revealed to 
-shared accesses when the presence bit is set.
+ formalizes that any access to the entry with these resources is \textit{invalid} (in the sense of not necessarily
+having accompanying resources) until the entry is marked present (and thus the memory returned from \textsf{kalloc}
+moves into the page table invariant.
 

### 2025-03-21 relwork.tex
inline tlb appendix mostly into related work, update related work

@@
 
+We will discuss related work along two axes: work on OS verification specifically, and work on 
+program logics with modalities.
+
+\subsection{OS Kernel Verification}
 There has been relatively little prior work on formal verification of virtual memory.
@@
 of a single address space by all processes (\textsc{Singularity}~\cite{Fahndrich2006language,Hunt2007singularity,Hunt2007sealing,Barnett2011specsharp}, \textsc{Verve}~\cite{Yang2010Verve},
-and \textsc{Tock}~\cite{levy2017multiprogramming}). The latter ensures memory safety, but ignores other functional uses of virtual memory
+and \textsc{Tock}~\cite{levy2017multiprogramming}). The latter approach ensures memory safety, but ignores other functional uses of virtual memory
 hardware like swapping~\cite{Denning1970VM} or exploiting copy-on-write techniques for dynamic migration of virtual machines~\cite{clark2005live}.
@@
 virtual points-to assertions as there are entries in the top-level table (4096) because they cannot share access to entries.
-We inherit fractional permission support form Iris, and use it extensively (the overly-restrictive
+We inherit fractional permission support form \iris, and use it extensively (the overly-restrictive
 Figure \ref{fig:strongvirtualpointsto} is already an improvement in this way).
@@
 % spaces while reasoning about correctness.
-The only other attempt to verify a change of address space was by \citet{syeda2018program,syeda2020formal}, whose focus was on TLB maintenance (which we
-do not model). However, they verified only the 4 instructions to switch address spaces and update the TLB on an ARM processor, in isolation
-(i.e., not the full context switch including changing stacks with address spaces).
-The specification they proved for those instructions did not address program invariants that may be valid in one address space and not the other,
-so is not flexible enough to extend directly to a full context switching primitive as in Figure \ref{fig:swtch}.
-Of course as we do not address TLB management (see Appendix \ref{apdx:tlb}), the right eventual solution would be to combine our work
-and theirs, which we leave to future work.
 
-\ifPLDI
-\else
 Unlike our work, Kolanski and Klein prove very useful embedding theorems stating that code that does not modify page table 
@@
 memory.
-\fi
+
+\paragraph{Translation Lookaside Buffers}
+As noted in Section \ref{sec:backgroundonmachinemodel}, we do not formally model or reason about TLBs.
+TLB flushes are required only when addresses are removed from a virtual address space, or when changing
+virtual address spaces. Because this occurs in few places in the kernel (in some, only 3 locations), full verified
+kernels including \textsc{seL4}~\cite{Klein2009seL4,seL4TOCS} and \textsc{CertiKOS}~\cite{gu15,gu2016certikos}
+trust TLB management. Neither of the aforementioned systems has a hardware model including a TLB, so neither is able
+to verify TLB management in any form --- they \emph{must} trust its operation.
+
+The only place TLB code becomes particularly challenging is in multiprocessor kernels, where
+not only must the running CPU flush its own TLB with special instructions, but it must send an inter-processor interrupt (IPI) to all
+other cores to ensure they also flush the relevant ranges of their own TLBs.
+Currently no formal hardware model exists with sufficient detail to reason about IPIs in full detail,
+which would require extensions to model interactions with ACPI tables populated by firmware on desktop and server machines,
+or flattened device trees for (semi-)embedded systems like single-board ARM and RISC-V machines, which tell the OS
+important information about how to control other CPU cores;
+plus the memory-mapped IO used to trigger those IPIs.
+As a result, multicore extensions of \textsc{seL4}~\cite{von2013clustered} and \textsc{CertiKOS}~\cite{gu2016certikos} also trust this functionality.
+
+% Fully grounded trust for this would require a formal model of how hardware populates ACPI tables and formal verification of ACPI
+% parsing code on general-purpose desktop and server machines, or a formalized and verified
+% parsing of flattened device trees (plus trust that a system was booted with a correct FDT) for (semi-)embedded systems like single-board
+% ARM and RISC-V machines. Either route would also require detailed treatment of memory-mapped IO triggering
+% interrupts on other CPUs. This is far beyond any formal model of computer hardware that exists today,
+% and far out of scope for this paper.
+
+% \citet{syeda2020formal,syeda2018program} are the only existing work to address formal verification of TLB management.
+% They extend the work of \citet{kolanski08ssv,kolanski09tphols,kolanski09tphols}, and therefore inherit the limitations of that work discussed
+% above. Their logic primarily tracks a set of mapped addresses and a set of
+% cached addresses, and generally preserve a global invariant (in a global Hoare logic, not a separation logic)
+% that the cached addresses are a subset of the mapped addresses. When this invariant is violated, the logic provides only
+% a Hoare-style backwards assignment type rule with unconstrained assertion $P$, rather than providing structure in the logic
+% for reasoning about how to restore the invariant. Most technical results in the paper focus on proving
+% transparency lemmas, that code that \emph{does not} modify page tables (kernel code outside the VMM, user code)
+% is unaffected by its existence. The one piece of kernel code verified is a sequence of 4 pseudo-instructions for changing address spaces
+% which include \emph{only} the installation of a new page table root and flushing of the old address space from the TLB.
+
+% Adapting Syeda and Klein's global reasoning principles to a separation logic is non-trivial, even though their approach
+% does capture some intuition about TLB reasoning. And as noted above, in the absence of support for significant additional
+% hardware functionality in the hardware model, this would be substantial work for only very limited gains in confidence compared
+% to trusting single-core TLB managment.
+% Thus we leave TLB management to our future work plans, where we believe it is best addressed in tandem with substantially
+% richer hardware models than any currently existing.
+
+The only other attempt to verify a change of address space was by \citet{syeda2018program,syeda2020formal}, whose focus was on TLB maintenance (which we
+do not model). However, they verified only the 4 instructions to switch address spaces and update the TLB on an ARM processor, in isolation
+(i.e., not the full context switch including changing stacks with address spaces).
+The specification they proved for those instructions did not address program invariants that may be valid in one address space and not the other,
+so is not flexible enough to extend directly to a full context switching primitive as in Figure \ref{fig:swtch},
+in addition to inheriting other limitations from the work they extend, discussed above~\cite{kolanski08vstte,kolanski09tphols}.
+
+The right eventual solution would be to combine our work
+and theirs, which we leave to future work, as combining both the models and reasoning principles would be highly non-trivial.
+
+\subsection{Program Logics with Modalities}
+Modalities have long been a staple of program logics, at least as far back as Dijkstra's weakest precondition calculus~\cite{dijkstra1975guarded}
+and Pratt's observation the Hoare triples
+could be decomposed using the weakest-precondition modality of dynamic logic~\cite{pratt1976semantical},
+in a form quite similar to what \iris uses today~\cite{jung2018iris}.
+Variants of Nakano's later modality~\cite{nakano2000modality} have long been used to deal with step-indexing 
+for impredicative and recursive features of logics and type systems~\cite{Appel2007,hobor2010theory,birkedal2011step,birkedal2013intensional,jung2018iris}.
 
@@
 where modalities are indexed by \emph{nominals} which are names for specific individual states in a Kripke model.
-Little prior work combines these ideas with program logics: \citet{brotherston2014parametric} show that traditional
-nominals extends the expressive power of separation logic, and \citet{gordon2019modal}
+Little prior work combines these ideas with program logics. \citet{brotherston2014parametric} show that traditional
+nominals extends the expressive power of separation logic. \citet{gordon2019modal}
 uses nominals to refer to states of other actors in an actor language.
+In parallel with our work, \citet{wagner2024realistic} use a hybrid-logic-inspired modality
+to abstract reference-counting specifics from specifications of a low-level application binary interface (ABI)
+--- their $@_l(P)$ indicates that $l$ is the location of a reference count for resources satisfying $P$.
 % As noted throughout the paper, the inspiration for our other-space modality comes from hybrid logic~\cite{areces2001hybrid,blackburn1995hybrid,gargov1993modal,goranko1996hierarchies},
@@
 % rely/guarantee approach). 
-Beyond these, there is limited work on the interaction of hybrid logic with substructural logics, in restricted forms
+Beyond these, there is limited work on the interaction of hybrid logic with general substructural logics, in restricted forms
 that do not affect expressivity~\cite{despeyroux2014hybrid,chaudhuri2019hybrid}.\looseness=-1

### 2025-03-21 programlogic.tex
fix some mentions of Coq (now Rocq) and Iris (to use the macro)

@@
 % The idea that the truth of an assertion is relative to an address space has far-reaching consequences.
-To support making assertions depend on a choice of address space, we work entirely in a pointwise lifting of Iris's base BI logic,
+To support making assertions depend on a choice of address space, we work entirely in a pointwise lifting of \iris's base BI logic,
 essentially working with separation logic assertions indexed by a choice of page table root as a $\mathcal{W}_{64}$, which we call $\textsf{vProp }\Sigma$:\footnote{
@@
 \lstinline[language=Coq]|Definition vProp  $\Sigma$ : bi := word 64 -b> iPropI  $\Sigma$|.
-This is the (\textsc{Coq}) type of assertions in our logic.
-Most constructs in Iris's base logic are defined with respect to any BI-algebra (of \coq type \lstinline|bi|), so automatically
+This is the (\rocq) type of assertions in our logic.
+Most constructs in \iris's base logic are defined with respect to any BI-algebra (of \coq type \lstinline|bi|), so automatically
 carry over to our derived logic.
-However, we must still build up from existing Iris primitives to provide new primitives that depend on the address space --- primarily the notion
+However, we must still build up from existing \iris primitives to provide new primitives that depend on the address space --- primarily the notion
 of virtual points-to.
@@
 which points to the page holding the physical memory backing the virtual address, which contains value \textsf{v}.
-Most of the definition lives directly in \textsf{vProp}, using the separation logic structure lifted from Iris's \textsf{iProp}.
+Most of the definition lives directly in \textsf{vProp}, using the separation logic structure lifted from \iris's \textsf{iProp}.
 \looseness=-1
@@
 where the semantics of all assertions are functions from pairs of heaps and page table root values to booleans.\footnote{
-  This was a typical explicit construction at the time; their work significantly predates Iris.
+  This was a typical explicit construction at the time; their work significantly predates \iris.
 }
@@
 when executed with \textsf{vProp} precondition $\Phi\ast\textsf{cr3}\mapsto_{\textsf{r}} \textsf{rtv}$.
-\textsf{WP} is Iris's own weakest precondition modality, unmodified~\cite{jung2018iris}.
+\textsf{WP} is \iris's own weakest precondition modality, unmodified~\cite{jung2018iris}.
 Making \textsf{rtv} a parameter to the double (vs. a simple register assertion)

### 2025-03-22 examples.tex
+

--- 2024-11-15 examples.tex
@@
 \label{wlkpgdir}
-
 \begin{figure}\footnotesize
@@
 ... ;; Similar assembly & proofs for levels 3 and 2
-$\specline{\textsf{R}_{\textsf{walk}}}_{\rtv}$
 ... ;;return &pt[PTEX(va)]; // return address of L1 entry, similar index computation
+ $\specline{\textsf{R}_{\textsf{walk}} \ast \textsf{R}_{\textsf{l1e}} }_{\rtv}$
 \end{lstlisting}
@@
 ;;pte_t *pt = pte_get_next_table(pd_entry);
-$\specline{\textsf{R}_{\textsf{walk}}}_{\rtv}$
+$\specline{ \left( \begin{array}{l} \textsf{pdp}+\textsf{PDPEX(va)} = \textsf{pdp\_entry} \land \textsf{table\_root(pml4\_entry}_{\textsf{pfn}})  = \textsf{pdp} \\ \textsf{pd}+\textsf{PDEX(va)} = \textsf{pd\_entry} \land \textsf{table\_root(pml3\_entry}_{\textsf{pfn}})  = \textsf{pd} \\ \textsf{pt}+\textsf{PDPEX(va)} = \textsf{pt\_entry} \land \textsf{table\_root(pml2\_entry}_{\textsf{pfn}})  = \textsf{pt} \\  \textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{pml4\_entry},\textsf{pdp\_entry},\textsf{pd\_entry},\textsf{pt\_entry}\},m)  \\ \ghostmaptoken{\delta{}s}{\rtv}{\delta} \\
+\ghostmaptoken{\textsf{id}}{(\mathsf{pml4\_entry})}{\textsf{4}}\ast \ghostmaptoken{\textsf{id}}{(\mathsf{plm3\_entry})}{\textsf{3}} \ast \ghostmaptoken{\textsf{id}}{(\mathsf{plm2\_entry})}{\textsf{2}}  \\ \forall_{i\in\textsf{0..511}} \; \textsf{table\_root(pml4\_entry}_{\textsf{pfn}}) \textsf{ + i * 8} \mapsto_{\textsf{id}} \textsf{3}  \\ \forall_{i\in\textsf{0..511}} \; \textsf{table\_root(pdp\_entry}_{\textsf{pfn}}) \textsf{ + i * 8} \mapsto_{\textsf{id}} \textsf{2}  \\ \forall_{i\in\textsf{0..511}} \; \textsf{table\_root(pd\_entry}_{\textsf{pfn}}) \textsf{ + i * 8} \mapsto_{\textsf{id}} \textsf{1}  \\  \textsf{pml4\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l4e\_val} \\  \textsf{pdp\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l3e\_val} \\  \textsf{pd\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l2e\_val} \\   \textsf{pt\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l1e\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l1e\_val})\urcorner    \end{array} \right)  =\textsf{R}_{\textsf{walk}}}$
 ;;access and return L1 entry
@@
 \caption{Walking page-table directory via calls to \textsf{pte\_get\_next\_table} in Figure \ref{fig:calltopteinitialize}}
+\label{walkpgdir}
+\vspace{-1em}
+\end{figure}
+
+\begin{figure}\footnotesize
+  \vspace{-1em}
+$\begin{array}{l}\left( \begin{array}{l} \ulcorner \textsf{pdp}+\textsf{PDPEX(va)} = \textsf{pdp\_entry} \land \textsf{table\_root(pml4\_entry}_{\textsf{pfn}})  = \textsf{pdp} \\ \land \textsf{pd}+\textsf{PDEX(va)} = \textsf{pd\_entry} \land \textsf{table\_root(pml3\_entry}_{\textsf{pfn}})  = \textsf{pd} \\ \land \textsf{pt}+\textsf{PDPEX(va)} = \textsf{pt\_entry} \land \textsf{table\_root(pml2\_entry}_{\textsf{pfn}})  = \textsf{pt} \urcorner  \ast \\  \textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{pml4\_entry},\textsf{pdp\_entry},\textsf{pd\_entry},\textsf{pt\_entry}\},m)  \\ \ghostmaptoken{\delta{}s}{\rtv}{\delta} \ast  \\
+      \ghostmaptoken{\textsf{id}}{(\mathsf{pml4\_entry})}{\textsf{4}}\ast \ghostmaptoken{\textsf{id}}{(\mathsf{plm3\_entry})}{\textsf{3}} \ast \ghostmaptoken{\textsf{id}}{(\mathsf{plm2\_entry})}{\textsf{2}}  \ast \\  \textsf{pml4\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l4e\_val} \ast  \tcbhighmath[boxrule=1pt,drop fuzzy shadow=blue]{\ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l4e\_val})\urcorner}\;\; \ast \\  \textsf{pdp\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l3e\_val} \ast  \tcbhighmath[boxrule=1pt,drop fuzzy shadow=blue]{\ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l3e\_val})\urcorner} \;\; \ast \\  \textsf{pd\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l2e\_val} \ast \tcbhighmath[boxrule=1pt,drop fuzzy shadow=blue]{\ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l2e\_val})\urcorner }  \end{array} \right)  =\textsf{R}_{\textsf{walk?} } \\ \ast \left(\begin{array}{l}   \textsf{pt\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l1e\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l1e\_val})\urcorner \end{array}\right) =\textsf{R}_{\textsf{l1e}}\end{array} $
+\caption{\textsf{R}$_{\textsf{walk?}}$: Resources needed for walking page-table directory via calls to \textsf{pte\_get\_next\_table} in Figure \ref{fig:calltopteinitialize}}
 \label{walkpgdir}
@@
 ;; pte_t *pte = walkpgdir(pml4, a);
-mov    -0x8[rbp],rdi; mov    -0x28[rbp],rsi; mov    0x0,rax; callq  291 <walkpgdir> (*\label{line:call_walkpgdir}*)
+ mov  -0x8[rbp], rdi;
+ mov  -0x16[rbp], rsi;
+ mov  -0x24[rbp], rdx;
+% mov  rdx, -0x16[rbp];
+ mov  rdx, (l4 + PML4EX(va) + natToWord 64 KERNBASE)))
+ mov  rsi,  -0x32[rbp];
+ mov  rax, 0x0;
+ callq  291 <walkpgdir> (*\label{line:call_walkpgdir}*)
+ $\specline{ \textsf{R}_{\textsf{walk}} \ast \left(\begin{array}{l}   \textsf{pte+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l1e\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l1e\_val})\urcorner \end{array}\right) =\textsf{R}_{\textsf{l1e}} }$
 ;;if(pte->present != 0) {
@@
 ... ;; Copy user access bit, writethrough, etc. (not part of our semantics) into pte from perm
-$\specline{ \left( \begin{array}{l} \textsf{pdp}+\textsf{PDPEX(va)} = \textsf{pdp\_entry} \land \textsf{table\_root(pml4\_entry}_{\textsf{pfn}})  = \textsf{pdp} \\ \textsf{pd}+\textsf{PDEX(va)} = \textsf{pd\_entry} \land \textsf{table\_root(pml3\_entry}_{\textsf{pfn}})  = \textsf{pd} \\ \textsf{pt}+\textsf{PDPEX(va)} = \textsf{pt\_entry} \land \textsf{table\_root(pml2\_entry}_{\textsf{pfn}})  = \textsf{pt} \\  \textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{pml4\_entry},\textsf{pdp\_entry},\textsf{pd\_entry},\textsf{pt\_entry}\},m)  \\ \ghostmaptoken{\delta{}s}{\rtv}{\delta} \\
-\ghostmaptoken{\textsf{id}}{(\mathsf{pml4\_entry})}{\textsf{4}}\ast \ghostmaptoken{\textsf{id}}{(\mathsf{plm3\_entry})}{\textsf{3}} \ast \ghostmaptoken{\textsf{id}}{(\mathsf{plm2\_entry})}{\textsf{2}}  \\ \forall_{i\in\textsf{0..511}} \; \textsf{table\_root(pml4\_entry}_{\textsf{pfn}}) \textsf{ + i * 8} \mapsto_{\textsf{id}} \textsf{3}  \\ \forall_{i\in\textsf{0..511}} \; \textsf{table\_root(pdp\_entry}_{\textsf{pfn}}) \textsf{ + i * 8} \mapsto_{\textsf{id}} \textsf{2}  \\ \forall_{i\in\textsf{0..511}} \; \textsf{table\_root(pd\_entry}_{\textsf{pfn}}) \textsf{ + i * 8} \mapsto_{\textsf{id}} \textsf{1}  \\  \textsf{pml4\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l4e\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l4e\_val})\urcorner \\  \textsf{pdp\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l3e\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l3e\_val})\urcorner \\  \textsf{pd\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l2e\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l2e\_val})\urcorner \\   \textsf{pt\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l1e\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l1e\_val})\urcorner    \end{array} \right)  =\textsf{R}_{\textsf{walk\_acc}}}$
 ;;pte->pfn          = PTE_ADDR_TO_PFN(fpaddr);

### 2025-03-22 examples.tex
+

@@
 $\begin{array}{l}\left( \begin{array}{l} \ulcorner \textsf{pdp}+\textsf{PDPEX(va)} = \textsf{pdp\_entry} \land \textsf{table\_root(pml4\_entry}_{\textsf{pfn}})  = \textsf{pdp} \\ \land \textsf{pd}+\textsf{PDEX(va)} = \textsf{pd\_entry} \land \textsf{table\_root(pml3\_entry}_{\textsf{pfn}})  = \textsf{pd} \\ \land \textsf{pt}+\textsf{PDPEX(va)} = \textsf{pt\_entry} \land \textsf{table\_root(pml2\_entry}_{\textsf{pfn}})  = \textsf{pt} \urcorner  \ast \\  \textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{pml4\_entry},\textsf{pdp\_entry},\textsf{pd\_entry},\textsf{pt\_entry}\},m)  \\ \ghostmaptoken{\delta{}s}{\rtv}{\delta} \ast  \\
-      \ghostmaptoken{\textsf{id}}{(\mathsf{pml4\_entry})}{\textsf{4}}\ast \ghostmaptoken{\textsf{id}}{(\mathsf{plm3\_entry})}{\textsf{3}} \ast \ghostmaptoken{\textsf{id}}{(\mathsf{plm2\_entry})}{\textsf{2}}  \ast \\  \textsf{pml4\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l4e\_val} \ast  \tcbhighmath[boxrule=1pt,drop fuzzy shadow=blue]{\ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l4e\_val})\urcorner}\;\; \ast \\  \textsf{pdp\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l3e\_val} \ast  \tcbhighmath[boxrule=1pt,drop fuzzy shadow=blue]{\ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l3e\_val})\urcorner} \;\; \ast \\  \textsf{pd\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l2e\_val} \ast \tcbhighmath[boxrule=1pt,drop fuzzy shadow=blue]{\ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l2e\_val})\urcorner }  \end{array} \right)  =\textsf{R}_{\textsf{walk?} } \\ \ast \left(\begin{array}{l}   \textsf{pt\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l1e\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l1e\_val})\urcorner \end{array}\right) =\textsf{R}_{\textsf{l1e}}\end{array} $
+      \ghostmaptoken{\textsf{id}}{(\mathsf{pml4\_entry})}{\textsf{4}}\ast \ghostmaptoken{\textsf{id}}{(\mathsf{plm3\_entry})}{\textsf{3}} \ast \ghostmaptoken{\textsf{id}}{(\mathsf{plm2\_entry})}{\textsf{2}}  \ast \\  \textsf{pml4\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l4e\_val} \ast  \tcbhighmath[boxrule=1pt,drop fuzzy shadow=blue]{\ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l4e\_val})\urcorner}\;\; \ast \\  \textsf{pdp\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l3e\_val} \ast  \tcbhighmath[boxrule=1pt,drop fuzzy shadow=blue]{\ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l3e\_val})\urcorner} \;\; \ast \\  \textsf{pd\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l2e\_val} \ast \tcbhighmath[boxrule=1pt,drop fuzzy shadow=blue]{\ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l2e\_val})\urcorner }  \end{array} \right)  =\textsf{R}_{\textsf{walk?} } \\ \ast \left(\begin{array}{l} \textsf{rax} \mapsto \textsf{pte} \ast  \textsf{pte} \mapsto_{\textsf{vpte,qfrac}} \textsf{pte\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present pte\_val})\urcorner \end{array}\right) =\textsf{R}_{\textsf{l1e}}\end{array} $
 \caption{\textsf{R}$_{\textsf{walk?}}$: Resources needed for walking page-table directory via calls to \textsf{pte\_get\_next\_table} in Figure \ref{fig:calltopteinitialize}}
@@
  callq  291 <walkpgdir> (*\label{line:call_walkpgdir}*)
- $\specline{ \textsf{R}_{\textsf{walk}} \ast \left(\begin{array}{l}   \textsf{pte+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l1e\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l1e\_val})\urcorner \end{array}\right) =\textsf{R}_{\textsf{l1e}} }$
+ $\specline{ \textsf{R}_{\textsf{walk}} \ast \left(\begin{array}{l}   \textsf{rax} \mapsto \textsf{pte} \ast  \textsf{pte} \mapsto_{\textsf{vpte,qfrac}} \textsf{pte\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present pte\_val}) \end{array}\right) =\textsf{R}_{\textsf{l1e}} }$
 ;;if(pte->present != 0) {

### 2025-03-22 examples.tex
pass through 5.1.4

--- 2025-03-21 examples.tex
@@
 \label{sec:p2v}
-Once we know the entry refers to a physical address the identity mapping range ($\Xi$)
-% Once we reach to the certain knowledge of having an entry with a frame referencing to an allocated resource resides inside the identity mappings
+Once we know the entry refers to a physical address in the identity mapping range ($\Xi$)
 (via the branch at Line \ref{line:check_entry_present_jump}, or  by allocating and installing a new entry
 as just discussed for Lines \ref{line:check_entry_present_jump}--\ref{line:end_of_allocation_path}), 
-we can %utilize this knowledge 
-to convert this frame address to a corresponding virtual address via identity mappings -- Line 57 in Figure \ref{fig:calltopteinitialize} 
-specified in Figure \ref{fig:p2v}.
-This is a critical piece of the full page table walk verification, and we have verified the critical step for a small x86-64 kernel, 
-which is the physical-to-virtual conversion (often appearing as a macro \texttt{P2V} in C source code). 
-In our small kernel (Line \ref{line:p2v} in Figure \ref{fig:p2v}), as in larger kernels, \texttt{P2V} is actually just addition by the
-constant offset mentioned in Section \ref{subsec:identitymappings}, but the correctness 
-of this simple instruction is quite subtle and relies on the extended invariant (Figure \ref{fig:peraspaceinvariant_with_p2v_extension})
-explained in that section.
+we can convert this frame address to a corresponding virtual address via the identity mappings
+discussed in Section \ref{subsec:identitymappings} and Figure \ref{fig:peraspaceinvariant_with_p2v_extension}.
+in the last lines of \lstinline|pte_get_next_table| shown in Figure \ref{fig:p2v} (the continuation of Figure \ref{fig:calltopteinitialize}).
+This is a critical piece of the full page table walk verification.
+In our small kernel (Line \ref{line:p2v} in Figure \ref{fig:p2v}), as in larger kernels, the C macro \texttt{P2V} common to many kernels
+is actually just addition by the constant offset mentioned in Section \ref{subsec:identitymappings}.
+But the correctness of this simple instruction is quite subtle.
+%  and cannot be proven 
+% without the extended invariant (Figure \ref{fig:peraspaceinvariant_with_p2v_extension})
+% worked out Section \ref{subsec:identitymappings}.
 
@@
 be accessed --- by selecting the appropriate index into the 512 ghost map tokens returned in the postcondition,
-and using the ghost translation and physical location portions of the invariant to assemble a vpte-pointsto.
+and using the ghost translation and physical location portions of the invariant to assemble a vpte-pointsto
+that justifies the caller's subsequent access to a particular entry of the returned table.
 % in the identity map ($\Xi\setminus\{entry\}$) of the kernel invariant.

### 2025-03-23 examples.tex
pass through 5.3

@@
 $\specline{\mathsf{ContextAt}(old,\_) \ast \mathsf{ContextAt}(new,[\mathsf{rbxv}',\ldots,\rtv']) }_{\rtv}$
-mov 0[rdi], rbx ... ;; also save rsp, rbp, r12, r13, r14, to offsets 8, 16, 24, 32, and 40 from rdi (*\label{line:start_save}*)
+mov 0[rdi], rbx
+... ;; also save rsp, rbp, r12, r13, r14, to offsets 8, 16, 24, 32, and 40 from rdi (*\label{line:start_save}*)
 mov 48[rdi], r15
@@
 The wrinkle, and the importance of the modal treatment of assertions, is that the target thread's precondition is \emph{relative to its address space}, 
-not the address space of the calling thread shown as 
+not the address space of the calling thread, which is reflected by
+the other-space modality 
 $[\rtv'](\mathcal{I}\texttt{ASpace}(\theta,\Xi,m) \ast \texttt{Pother})$
@@
 address space.
+\looseness=-1
 
@@
 \looseness=-1
+
+While prior work has verified context switches within a single address space~\cite{ni2007contexts}, and context switches
+without any code before or after~\cite{syeda2020formal} (i.e., not reasoning about the impact of address space change
+on what data was accessible), this is the first verification handling both.
 

### 2025-03-23 programlogic.tex
finished advisor pass, switched to author-year citations, now 23 pages with a little wiggle room

@@
 
-\paragraph{Accessing Virtual Addresses}
+\subsubsection{Accessing Virtual Addresses}
 Figure \ref{fig:wpdamd} includes two  rules for accessing memory at an address stored in a register $r_a$. 
@@
 
-\paragraph{Updating \lstinline|cr3|} 
+\subsubsection{Updating \lstinline|cr3|} 
 Unlike other rules, \TirNameStyle{WriteToRegCtlFromRegModal} updates the root address of the 
@@
 the corresponding modality.
+
 Note that the \emph{global} aspect is important, as a na\"ive frame rule would be unsound for updates to \lstinline|cr3|,
 because one could frame out assertions in one address space, switch address spaces, and bring those assertions from the \emph{old}
-address space back into the \emph{new} address space, where they may not hold. Appendix \ref{sec:issues} gives more details.
-% From that rule, with one step outside the \textsf{vProp} logic (so it is not quite a derived rule within
-% our embedded logic), we can derive 
-% \TirNameStyle{WriteToRegCtlFromRegModal},
-% which is the (full) Hoare double version of the address space switch rule proposed in Section \ref{sec:issues}.
+address space back into the \emph{new} address space, where they may not hold. 
+% Appendix \ref{sec:issues} gives more details.
+It is often said that the frame rule (below) is one of the key pieces of separation logic.
+\[
+  \mbox{$\inferrule*[right=Frame]{
+    \{P\}\;C\;\{Q\}
+  }{
+    \{P\ast R\}\;C\;\{Q\ast R\}
+  }$}
+  \qquad\textrm{(unsound with address space changes)}
+\]
+Such a rule is normally recoverable from Hoare doubles (see, e.g., \citet{Chlipala2011Bedrock,Chlipala2013Bedrock}).
+However, in the presence of address space changes the traditional frame rule is unsound.
+Consider:
+\begin{mathpar}
+  \inferrule*[right=Frame]{
+    \inferrule{\ldots }{
+    \{\textsf{Pre}\}
+    \texttt{mov}~\texttt{\%cr3},~r%\lstinline|mov %cr3, r| 
+    \{\textsf{Post}\}
+    }
+  }{
+    \{a\mapsto_\mathsf{v} x \ast \textsf{Pre}\}
+    \texttt{mov}~\texttt{\%cr3},~r%\lstinline|mov %cr3, r| 
+    \{a\mapsto_\mathsf{v} x \ast \textsf{Post}\}
+  }
+\end{mathpar}
+In this (broken!) hypothetical example,
+both the precondition and postcondition assert that $a\mapsto_\mathsf{v} x$ in the current address space, but we have no basis for concluding that address translation is preserved by the change of address space.
+So this derivation clearly leads to an unsound conclusion.
+The essential problem is that the frame rule is motivated by local reasoning about local updates, but
+a switch of address space is a \emph{global} change that may invalidate information about virtual addresses.
+Thus framing around arbitrary \lstinline|cr3| updates is unsound --- hence the \emph{global} nature of \textsc{WriteToRegCtlFromRegModal} ---
+though a variant the ensures the same \lstinline|cr3| value is installed before and after the framing
+can be recovered.
 

### 2025-03-24 examples.tex
finished advisor pass, switched to author-year citations, now 23 pages with a little wiggle room
fixes

--- 2025-03-22 examples.tex
@@
   \vspace{-1em}
-$\begin{array}{l}\left( \begin{array}{l} \ulcorner \textsf{pdp}+\textsf{PDPEX(va)} = \textsf{pdp\_entry} \land \textsf{table\_root(pml4\_entry}_{\textsf{pfn}})  = \textsf{pdp} \\ \land \textsf{pd}+\textsf{PDEX(va)} = \textsf{pd\_entry} \land \textsf{table\_root(pml3\_entry}_{\textsf{pfn}})  = \textsf{pd} \\ \land \textsf{pt}+\textsf{PDPEX(va)} = \textsf{pt\_entry} \land \textsf{table\_root(pml2\_entry}_{\textsf{pfn}})  = \textsf{pt} \urcorner  \ast \\  \textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{pml4\_entry},\textsf{pdp\_entry},\textsf{pd\_entry},\textsf{pt\_entry}\},m)  \\ \ghostmaptoken{\delta{}s}{\rtv}{\delta} \ast  \\
-      \ghostmaptoken{\textsf{id}}{(\mathsf{pml4\_entry})}{\textsf{4}}\ast \ghostmaptoken{\textsf{id}}{(\mathsf{plm3\_entry})}{\textsf{3}} \ast \ghostmaptoken{\textsf{id}}{(\mathsf{plm2\_entry})}{\textsf{2}}  \ast \\  \textsf{pml4\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l4e\_val} \ast  \tcbhighmath[boxrule=1pt,drop fuzzy shadow=blue]{\ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l4e\_val})\urcorner}\;\; \ast \\  \textsf{pdp\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l3e\_val} \ast  \tcbhighmath[boxrule=1pt,drop fuzzy shadow=blue]{\ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l3e\_val})\urcorner} \;\; \ast \\  \textsf{pd\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l2e\_val} \ast \tcbhighmath[boxrule=1pt,drop fuzzy shadow=blue]{\ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l2e\_val})\urcorner }  \end{array} \right)  =\textsf{R}_{\textsf{walk?} } \\ \ast \left(\begin{array}{l} \textsf{rax} \mapsto \textsf{pte} \ast  \textsf{pte} \mapsto_{\textsf{vpte,qfrac}} \textsf{pte\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present pte\_val})\urcorner \end{array}\right) =\textsf{R}_{\textsf{l1e}}\end{array} $
-\caption{\textsf{R}$_{\textsf{walk?}}$: Resources needed for walking page-table directory via calls to \textsf{pte\_get\_next\_table} in Figure \ref{fig:calltopteinitialize}}
+$\begin{array}{l}\left( \begin{array}{l} \ulcorner \textsf{pdp}+\textsf{PDPEX(va)} = \textsf{pdp\_entry} \land \textsf{table\_root(pml4\_entry}_{\textsf{pfn}})  = \textsf{pdp} \\ \land \textsf{pd}+\textsf{PDEX(va)} = \textsf{pd\_entry} \land \textsf{table\_root(pml3\_entry}_{\textsf{pfn}})  = \textsf{pd} \\ \land \textsf{pt}+\textsf{PDPEX(va)} = \textsf{pt\_entry} \land \textsf{table\_root(pml2\_entry}_{\textsf{pfn}})  = \textsf{pt} \urcorner  \ast \\  \textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{pml4\_entry},\textsf{pdp\_entry},\textsf{pd\_entry},\textsf{pte}\},m)  \\ \ghostmaptoken{\delta{}s}{\rtv}{\delta} \ast  \\
+      \ghostmaptoken{\textsf{id}}{(\mathsf{pml4\_entry})}{\textsf{4}}\ast \ghostmaptoken{\textsf{id}}{(\mathsf{plm3\_entry})}{\textsf{3}} \ast \ghostmaptoken{\textsf{id}}{(\mathsf{plm2\_entry})}{\textsf{2}}  \ast \\  \textsf{pml4\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l4e\_val}  \ast \\  \textsf{pdp\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l3e\_val}  \ast \\  \textsf{pd\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l2e\_val}  \end{array} \right)  =\textsf{R}_{\textsf{walk} } \\ \ast \left(\begin{array}{l} \textsf{rax} \mapsto \textsf{pte} \ast  \textsf{pt\_entry} \mapsto_{\textsf{vpte,qfrac}} \textsf{pte} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present pte\_val})\urcorner \end{array}\right) =\textsf{R}_{\textsf{l1e}}\end{array} $
+\caption{\textsf{R}$_{\textsf{walk}}$: Resources obtained from invariant for walking page-table directory with \emph{non-existing entries} via calls to \textsf{pte\_get\_next\_table} in Figure \ref{fig:calltopteinitialize}}
 \label{walkpgdir}
@@
 \end{figure}
-
 Realizing a software page-table walk amounts to calling \textsf{pte\_get\_next\_table} for each level as shown in Figure \ref{walkpgdir}. 
@@
   \begin{lstlisting}[mathescape,escapeinside={(*}{*)}]
-;;complstatus_t vaspace_mappage(pte_t *pml4, void *va,uintptr_t fpaddr, uint64_t perm ) {
+;;complstatus_t vaspace_mappage(pte_t *pml4, void *va,uintptr_t fpaddr ) {
 ... ;;setting up the stack      
@@
  mov  -0x24[rbp], rdx;
-% mov  rdx, -0x16[rbp];
- mov  rdx, (l4 + PML4EX(va) + natToWord 64 KERNBASE)))
- mov  rsi,  -0x32[rbp];
- mov  rax, 0x0;
+;; pte = walkpgdir_succ(pml4, va);
+ mov    -0x8[rbp], %rdi;
+ mov    -0x16[rbp], rsi;
+ mov    0x0, rax;
  callq  291 <walkpgdir> (*\label{line:call_walkpgdir}*)
  $\specline{ \textsf{R}_{\textsf{walk}} \ast \left(\begin{array}{l}   \textsf{rax} \mapsto \textsf{pte} \ast  \textsf{pte} \mapsto_{\textsf{vpte,qfrac}} \textsf{pte\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present pte\_val}) \end{array}\right) =\textsf{R}_{\textsf{l1e}} }$
-;;if(pte->present != 0) {
+ mov  rax, rdx;
+ mov    rdx,-0x28[rbp];
+;;if (!pte->present){ 
+ mov    -0x28[rbp], rdx
+ mov    [rdx], rcx
+ and    0x1, rcx
+ mov    rcx, r8
+ cmp    0x0, r8
+ jne    3e0 <vaspace_mappage+0x190>
  ...
-mov    0x0,r8; mov    rax,rcx; mov    rcx,-0x30[rbp]
-... ;; Copy user access bit, writethrough, etc. (not part of our semantics) into pte from perm
-;;pte->pfn          = PTE_ADDR_TO_PFN(fpaddr);
-mov    -0x18[rbp],rcx; shr    0xc,rcx; mov    -0x30[rbp],rdx; mov    [rdx],rsi ;; Load current entry in rsi
-$\specline{\textsf{pt\_entry}_{\textsf{pfn}} \mapsto_{p} \textsf{addr\_to\_pfn}(\textsf{pa}) \ast \textsf{R}_{\textsf{walk\_acc}}}_{\rtv} \Rrightarrow \specline{\textsf{va} \mapsto_{\textsf{v,q}} \textsf{pa} \ast \textsf{P}}$
-movabs 0xffffffffff,rdi; and    rdi,rcx; shl    0xc,rcx; movabs 0xfff0000000000fff,rdi
-and    rdi,rsi; or     rcx,rsi; mov    rsi,[rdx] ;; Store updated entry back to L1 entry (*\label{line:l1entry_store}*)
+;;pte->pfn          = PTE_ADDR_TO_PFN(pa);
+ mov    -0x18[rbp], rdx;
+ shr    0xc, rdx;
+ mov    -0x28[rbp], rsi;
+ mov    [rsi], rdi;
+ movabs 0xffffffffff, rcx
+ and    rcx, rdx
+ shl    0xc, rdx
+ movabs 0xfff0000000000fff, rcx
+ and    rcx, rdi
+ or     rdx, rdi
+ mov    rdi, [rsi]
+$\specline{ \textsf{pte} \mapsto_{\textsf{vpte,1}} (\textsf{set\_pfn}(\textsf{pte\_val}, \textsf{PTE\_ADDR\_TO\_PFN}(\textsf{fpaddr}))) \ast  \ulcorner \lnot(\textsf{entry\_present pte\_val}) \urcorner \ast P}$
+ ;; Store updated entry back to L1 entry (*\label{line:l1entry_store}*)
 ;;}
-... ;; set present bit in entry, then clean up the stack and return (proof re-packs invariants)
+... ;;set present bit in entry, then clean up the stack
+... ;;and return either unaccessible pte pointsto or one with fresh page (fpaddr)
+ $\specline{ \begin{array}{l}\textsf{pte} \mapsto_{\textsf{vpte,qfrac}} \textsf{pte\_val} \ast \ulcorner (\textsf{entry\_present pte\_val}) \urcorner \ast P \\
+ \lor \; \;  \textsf{pte} \mapsto_{\textsf{vpte,1}} (\textsf{set\_pfn}(\textsf{pte\_val}, \textsf{PTE\_ADDR\_TO\_PFN}(\textsf{fpaddr}))) \ast  \ulcorner \lnot(\textsf{entry\_present pte\_val}) \urcorner \ast P \end{array} }$
 ;;}

### 2025-03-24 examples.tex
+

@@
 \caption{\textsf{R}$_{\textsf{walk}}$: Resources obtained from invariant for walking page-table directory with \emph{non-existing entries} via calls to \textsf{pte\_get\_next\_table} in Figure \ref{fig:calltopteinitialize}}
-\label{walkpgdir}
+\label{fig:rwalk}
 \vspace{-1em}
@@
 to map new pages into an address space by writing into an existing page table via a call
-\[\textsf{vaspace\_mappage(pte\_t *pml4, void *va,uintptr\_t fpaddr, uint64\_t perm )}\] in Figure \ref{fig:mapping_code}.
+\[\textsf{vaspace\_mappage(pte\_t *pml4, void *va,uintptr\_t fpaddr)}\] in Figure \ref{fig:mapping_code}.
 To do so, with a given allocated a fresh page (\textsf{fpaddr}), then calculate the appropriate
 known-valid page table walks (via \textsf{walkpgdir} Line \ref{line:call_walkpgdir} in Figure \ref{fig:mapping_code})  and update 
-the appropriate L1 page table entry (Line 11 in Figure \ref{fig:mapping_code});
+the appropriate L1 page table entry (Line 35 in Figure \ref{fig:mapping_code});
 unmapping is the reverse of the logic we discuss here.
@@
  or     rdx, rdi
- mov    rdi, [rsi]
+ mov    rdi, [rsi] (*\label{line:updatepfn}*)
 $\specline{ \textsf{pte} \mapsto_{\textsf{vpte,1}} (\textsf{set\_pfn}(\textsf{pte\_val}, \textsf{PTE\_ADDR\_TO\_PFN}(\textsf{fpaddr}))) \ast  \ulcorner \lnot(\textsf{entry\_present pte\_val}) \urcorner \ast P}$
@@
 both for ensuring partial page table walk reaching the
-L1 entry (l1e) with the justification of higher levels of the page table exist, and for allowing access to the memory of the L1 entry via virtual address.
-A PTE points-to is defined just like the normal virtual points-to of Figure \ref{fig:virtualpointstosharing}, except the physical address (here, \textsf{pa}) is explicit in the assertion 
-rather than existentially quantified:
-\looseness=-1
- \[
-\begin{array}{l}
-    \vaddr\mapsto_{\textsf{vpte,q}} \; \paddr \; \vpage : \mathsf{vProp}~\Sigma \stackrel{\triangle}{=} 
-    \exists \delta\ldotp
-	(\lambda \mathit{cr3val}\ldotp
-	\ghostmaptoken{\delta{}s}{\mathit{cr3val}}{\delta}) \ast 
-  \fracghostmaptoken{\delta}{\vaddr}{\paddr}{\qfrac} \ast \paddr \mapsto_{\mathsf{p}} \vpage
-\end{array}
-\]
-This supports rules for accessing memory
-at that virtual address, but exposing the physical location being modified
-makes this convenient to use for page table modifications, since we must ensure
-the modified data is at the specific physical location that will be used as the L1 entry.
+L1 entry (l1e) with the justification of higher levels of the page table exist (R$_{\textsf{walk}}$ in Figure \ref{fig:rwalk}), and for allowing access to the memory of the L1 entry via virtual address (R$_{\textsf{l1e}}$ in Figure \ref{fig:rwalk}).
+%A PTE points-to, as we already brifly mentioned, is defined just like the normal virtual points-to of Figure \ref{fig:virtualpointstosharing}, except the physical address (here, \textsf{pa}) is explicit in the assertion
+%rather than existentially quantified:
+%\looseness=-1
+% \[
+%\begin{array}{l}
+%    \vaddr\mapsto_{\textsf{vpte,q}} \; \paddr \; \vpage : \mathsf{vProp}~\Sigma \stackrel{\triangle}{=} 
+%    \exists \delta\ldotp
+%	(\lambda \mathit{cr3val}\ldotp
+%	\ghostmaptoken{\delta{}s}{\mathit{cr3val}}{\delta}) \ast 
+%  \fracghostmaptoken{\delta}{\vaddr}{\paddr}{\qfrac} \ast \paddr \mapsto_{\mathsf{p}} \vpage
+%\end{array}
+%\]
+%This supports rules for accessing memory
+%at that virtual address, but exposing the physical location being modified
+%makes this convenient to use for page table modifications, since we must ensure
+%the modified data is at the specific physical location that will be used as the L1 entry.
 % After obtaining a virtual address \textsf{pte\_addr} in \textsf{rax} backed 
@@
 
-The crucial and the only step in addition to traversing the page table in Figure \ref{walkpgdir} is actually updating the L1 entry (Line 11 in Figure \ref{fig:mapping_code}),
-via the virtual address
-(\textsf{pt\_entry+KERNBASE}) known to translate to the appropriate physical address, in our example the L1
-table entry address ($\textsf{addr\_to\_pfn(pa)}$).
 % , to hold the freshly
@@
 
-At this point in the proof, i.e. after the traversal to obtain the sound physical path resources ($\textsf{R}_{\textsf{walk\_acc}}$ Line 27 in Figure \ref{walkpgdir}), which coontains
-virtual PTE points-to relations along the path to the full ownership of the L1 table entry's backing physical memory, we update the L1 entry to point the fresh page, as we do for the non-existent entries in the page table traversal (Line 6 in Figure \ref{pteinitializespec}.
-\looseness=-1
+We already discussed for the upper level page-tables that the entry-exists checks are properly ensured ($\textsf{R}_{\textsf{walk}}$ in Figure \ref{fig:rwalk} and Line 27 in Figure \ref{walkpgdir}).
+However, for L1 level, this check is also checked left to the caller of the traversal function \textsf{walkpgdir}.
+In other words, unlike  to what we see in R$_{\textsf{walk}}$ for the upper levels where all entry-exists checks are dropped, the specification in R$_{\textsf{l1e}}$ ensures that page table entry for L1 needs to be checked on the caller site.
+By doing so, as we see in Figure \ref{fig:mapping_code}, the page reference \textsf{fpaddr} is linked to back the virtual address \textsf{va} only if it is not already referring to a physical resource (Lines 16-23 in Figure \ref{fig:mapping_code}).
+
+The crucial and the only step in addition to traversing the page table in Figure \ref{walkpgdir} is actually updating the L1 entry (Line 35 in Figure \ref{fig:mapping_code}),
+via the virtual address
+(\textsf{pte}) known to translate to the appropriate physical address, in our example the L1
+table entry address ($\textsf{PTE\_ADDR\_TO\_PFN(fpaddr)}$).
 

### 2025-03-25 examples.tex

--- 2025-03-23 examples.tex
@@
 
-If the entry is not set, \textsf{pte\_initialize} (Line \ref{line:call_to_pte_initialize} in Figure \ref{fig:calltopteinitialize}) 
-allocates a physical page (internally utilizing the only unverified (trusted) code in our case studies, the page-allocator's \textsf{kalloc},\footnote{
-  This is an allocator for regions of pre-zeroed physical memory that is mapped, but not accessed by the allocator itself,
-  as is typical for slab allocators~\cite{bonwick1994slab}.
-  Its verification would be similar to verifying a usermode \textsf{malloc} verifications~\cite{Chlipala2013Bedrock,wickerson2010explicit},
-  just with additional invariants on the memory pool.
-} 
-on Line \ref{line:call_to_kalloc} in Figure \ref{pteinitializespec}). 
-\looseness=-1
 \begin{figure}\footnotesize
@@
 
+If the entry is not set, \textsf{pte\_initialize} (Line \ref{line:call_to_pte_initialize} in Figure \ref{fig:calltopteinitialize}) 
+allocates a physical page (internally utilizing the only unverified (trusted) code in our case studies, the page-allocator's \textsf{kalloc},\footnote{
+  This is an allocator for regions of pre-zeroed physical memory that is mapped, but not accessed by the allocator itself,
+  as is typical for slab allocators~\cite{bonwick1994slab}.
+  Its verification would be similar to verifying a usermode \textsf{malloc} verifications~\cite{Chlipala2013Bedrock,wickerson2010explicit},
+  just with additional invariants on the memory pool.
+} 
+on Line \ref{line:call_to_kalloc} in Figure \ref{pteinitializespec}). 
 Since we are using \textsf{pte\_initialize} for page-table address allocation, we must relate this newly
@@
 
+\begin{figure}\footnotesize
+  \vspace{-1em}
+$\begin{array}{l}\left( \begin{array}{l} \ulcorner \textsf{pdp}+\textsf{PDPEX(va)} = \textsf{pdp\_entry} \land \textsf{table\_root(pml4\_entry}_{\textsf{pfn}})  = \textsf{pdp} \\ \land \textsf{pd}+\textsf{PDEX(va)} = \textsf{pd\_entry} \land \textsf{table\_root(pml3\_entry}_{\textsf{pfn}})  = \textsf{pd} \\ \land \textsf{pt}+\textsf{PDPEX(va)} = \textsf{pt\_entry} \land \textsf{table\_root(pml2\_entry}_{\textsf{pfn}})  = \textsf{pt} \urcorner  \ast \\  \textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{pml4\_entry},\textsf{pdp\_entry},\textsf{pd\_entry},\textsf{pte}\},m)  \\ \ghostmaptoken{\delta{}s}{\rtv}{\delta} \ast  \\
+      \ghostmaptoken{\textsf{id}}{(\mathsf{pml4\_entry})}{\textsf{4}}\ast \ghostmaptoken{\textsf{id}}{(\mathsf{plm3\_entry})}{\textsf{3}} \ast \ghostmaptoken{\textsf{id}}{(\mathsf{plm2\_entry})}{\textsf{2}}  \ast \\  \textsf{pml4\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l4e\_val}  \ast \\  \textsf{pdp\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l3e\_val}  \ast \\  \textsf{pd\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l2e\_val}  \end{array} \right)  =\textsf{R}_{\textsf{walk} } \\ \ast \left(\begin{array}{l} \textsf{rax} \mapsto \textsf{pte} \ast  \textsf{pt\_entry} \mapsto_{\textsf{vpte,qfrac}} \textsf{pte} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present pte\_val})\urcorner \end{array}\right) =\textsf{R}_{\textsf{l1e}}\end{array} $
+\caption{\textsf{R}$_{\textsf{walk}}$: Resources obtained from invariant for walking page-table directory with \emph{non-existing entries} via calls to \textsf{pte\_get\_next\_table} in Figure \ref{fig:calltopteinitialize}}
+\label{fig:rwalk}
+\vspace{-1em}
+\end{figure}
 Realizing a software page-table walk amounts to calling \textsf{pte\_get\_next\_table} for each level as shown in Figure \ref{walkpgdir}. 
@@
 ;; pte_t *pte = walkpgdir(pml4, a);
-mov    -0x8[rbp],rdi; mov    -0x28[rbp],rsi; mov    0x0,rax; callq  291 <walkpgdir> (*\label{line:call_walkpgdir}*)
-;;if(pte->present != 0) {
+ mov  -0x8[rbp], rdi;
+ mov  -0x16[rbp], rsi;
+ mov  -0x24[rbp], rdx;
+;; pte = walkpgdir_succ(pml4, va);
+ mov    -0x8[rbp], %rdi;
+ mov    -0x16[rbp], rsi;
+ mov    0x0, rax;
+ callq  291 <walkpgdir> (*\label{line:call_walkpgdir}*)
+ $\specline{ \textsf{R}_{\textsf{walk}} \ast \left(\begin{array}{l}   \textsf{rax} \mapsto \textsf{pte} \ast  \textsf{pte} \mapsto_{\textsf{vpte,qfrac}} \textsf{pte\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present pte\_val}) \end{array}\right) =\textsf{R}_{\textsf{l1e}} }$
+ mov  rax, rdx;
+ mov    rdx,-0x28[rbp];
+;;if (!pte->present){ 
+ mov    -0x28[rbp], rdx
+ mov    [rdx], rcx
+ and    0x1, rcx
+ mov    rcx, r8
+ cmp    0x0, r8
+ jne    3e0 <vaspace_mappage+0x190>
  ...
-mov    0x0,r8; mov    rax,rcx; mov    rcx,-0x30[rbp]
-... ;; Copy user access bit, writethrough, etc. (not part of our semantics) into pte from perm
-$\specline{ \left( \begin{array}{l} \textsf{pdp}+\textsf{PDPEX(va)} = \textsf{pdp\_entry} \land \textsf{table\_root(pml4\_entry}_{\textsf{pfn}})  = \textsf{pdp} \\ \textsf{pd}+\textsf{PDEX(va)} = \textsf{pd\_entry} \land \textsf{table\_root(pml3\_entry}_{\textsf{pfn}})  = \textsf{pd} \\ \textsf{pt}+\textsf{PDPEX(va)} = \textsf{pt\_entry} \land \textsf{table\_root(pml2\_entry}_{\textsf{pfn}})  = \textsf{pt} \\  \textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{pml4\_entry},\textsf{pdp\_entry},\textsf{pd\_entry},\textsf{pt\_entry}\},m)  \\ \ghostmaptoken{\delta{}s}{\rtv}{\delta} \\
-\ghostmaptoken{\textsf{id}}{(\mathsf{pml4\_entry})}{\textsf{4}}\ast \ghostmaptoken{\textsf{id}}{(\mathsf{plm3\_entry})}{\textsf{3}} \ast \ghostmaptoken{\textsf{id}}{(\mathsf{plm2\_entry})}{\textsf{2}}  \\ \forall_{i\in\textsf{0..511}} \; \textsf{table\_root(pml4\_entry}_{\textsf{pfn}}) \textsf{ + i * 8} \mapsto_{\textsf{id}} \textsf{3}  \\ \forall_{i\in\textsf{0..511}} \; \textsf{table\_root(pdp\_entry}_{\textsf{pfn}}) \textsf{ + i * 8} \mapsto_{\textsf{id}} \textsf{2}  \\ \forall_{i\in\textsf{0..511}} \; \textsf{table\_root(pd\_entry}_{\textsf{pfn}}) \textsf{ + i * 8} \mapsto_{\textsf{id}} \textsf{1}  \\  \textsf{pml4\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l4e\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l4e\_val})\urcorner \\  \textsf{pdp\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l3e\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l3e\_val})\urcorner \\  \textsf{pd\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l2e\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l2e\_val})\urcorner \\   \textsf{pt\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l1e\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l1e\_val})\urcorner    \end{array} \right)  =\textsf{R}_{\textsf{walk\_acc}}}$
-;;pte->pfn          = PTE_ADDR_TO_PFN(fpaddr);
-mov    -0x18[rbp],rcx; shr    0xc,rcx; mov    -0x30[rbp],rdx; mov    [rdx],rsi ;; Load current entry in rsi
-$\specline{\textsf{pt\_entry}_{\textsf{pfn}} \mapsto_{p} \textsf{addr\_to\_pfn}(\textsf{pa}) \ast \textsf{R}_{\textsf{walk\_acc}}}_{\rtv} \Rrightarrow \specline{\textsf{va} \mapsto_{\textsf{v,q}} \textsf{pa} \ast \textsf{P}}$
-movabs 0xffffffffff,rdi; and    rdi,rcx; shl    0xc,rcx; movabs 0xfff0000000000fff,rdi
-and    rdi,rsi; or     rcx,rsi; mov    rsi,[rdx] ;; Store updated entry back to L1 entry (*\label{line:l1entry_store}*)
+;;pte->pfn          = PTE_ADDR_TO_PFN(pa);
+ mov    -0x18[rbp], rdx;
+ shr    0xc, rdx;
+ mov    -0x28[rbp], rsi;
+ mov    [rsi], rdi;
+ movabs 0xffffffffff, rcx
+ and    rcx, rdx
+ shl    0xc, rdx
+ movabs 0xfff0000000000fff, rcx
+ and    rcx, rdi
+ or     rdx, rdi
+ mov    rdi, [rsi] (*\label{line:updatepfn}*)
+$\specline{ \textsf{pte} \mapsto_{\textsf{vpte,1}} (\textsf{set\_pfn}(\textsf{pte\_val}, \textsf{PTE\_ADDR\_TO\_PFN}(\textsf{fpaddr}))) \ast  \ulcorner \lnot(\textsf{entry\_present pte\_val}) \urcorner \ast P}$
+ ;; Store updated entry back to L1 entry (*\label{line:l1entry_store}*)
 ;;}
-... ;; set present bit in entry, then clean up the stack and return (proof re-packs invariants)
+... ;;set present bit in entry, then clean up the stack
+... ;;and return either unaccessible pte pointsto or one with fresh page (fpaddr)
+ $\specline{ \begin{array}{l}\textsf{pte} \mapsto_{\textsf{vpte,qfrac}} \textsf{pte\_val} \ast \ulcorner (\textsf{entry\_present pte\_val}) \urcorner \ast P \\
+ \lor \; \;  \textsf{pte} \mapsto_{\textsf{vpte,1}} (\textsf{set\_pfn}(\textsf{pte\_val}, \textsf{PTE\_ADDR\_TO\_PFN}(\textsf{fpaddr}))) \ast  \ulcorner \lnot(\textsf{entry\_present pte\_val}) \urcorner \ast P \end{array} }$
 ;;}
@@
 $\specline{\mathsf{ContextAt}(old,\_) \ast \mathsf{ContextAt}(new,[\mathsf{rbxv}',\ldots,\rtv']) }_{\rtv}$
-mov 0[rdi], rbx
-... ;; also save rsp, rbp, r12, r13, r14, to offsets 8, 16, 24, 32, and 40 from rdi (*\label{line:start_save}*)
+mov 0[rdi], rbx ... ;; also save rsp, rbp, r12, r13, r14, to offsets 8, 16, 24, 32, and 40 from rdi (*\label{line:start_save}*)
 mov 48[rdi], r15
@@
 on what data was accessible), this is the first verification handling both.
+\looseness=-1
 

### 2025-03-25 examples.tex
some misc cleanup of edits to mappage, switch back to numeric citations for space

@@
   \vspace{-1em}
-$\begin{array}{l}\left( \begin{array}{l} \ulcorner \textsf{pdp}+\textsf{PDPEX(va)} = \textsf{pdp\_entry} \land \textsf{table\_root(pml4\_entry}_{\textsf{pfn}})  = \textsf{pdp} \\ \land \textsf{pd}+\textsf{PDEX(va)} = \textsf{pd\_entry} \land \textsf{table\_root(pml3\_entry}_{\textsf{pfn}})  = \textsf{pd} \\ \land \textsf{pt}+\textsf{PDPEX(va)} = \textsf{pt\_entry} \land \textsf{table\_root(pml2\_entry}_{\textsf{pfn}})  = \textsf{pt} \urcorner  \ast \\  \textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{pml4\_entry},\textsf{pdp\_entry},\textsf{pd\_entry},\textsf{pte}\},m)  \\ \ghostmaptoken{\delta{}s}{\rtv}{\delta} \ast  \\
-      \ghostmaptoken{\textsf{id}}{(\mathsf{pml4\_entry})}{\textsf{4}}\ast \ghostmaptoken{\textsf{id}}{(\mathsf{plm3\_entry})}{\textsf{3}} \ast \ghostmaptoken{\textsf{id}}{(\mathsf{plm2\_entry})}{\textsf{2}}  \ast \\  \textsf{pml4\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l4e\_val}  \ast \\  \textsf{pdp\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l3e\_val}  \ast \\  \textsf{pd\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l2e\_val}  \end{array} \right)  =\textsf{R}_{\textsf{walk} } \\ \ast \left(\begin{array}{l} \textsf{rax} \mapsto \textsf{pte} \ast  \textsf{pt\_entry} \mapsto_{\textsf{vpte,qfrac}} \textsf{pte} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present pte\_val})\urcorner \end{array}\right) =\textsf{R}_{\textsf{l1e}}\end{array} $
+$\begin{array}{l}\left( \begin{array}{l} \ulcorner \textsf{pdp}+\textsf{PDPEX(va)} = \textsf{pdp\_entry} \land \textsf{table\_root(pml4\_entry}_{\textsf{pfn}})  = \textsf{pdp} \\ \land \textsf{pd}+\textsf{PDEX(va)} = \textsf{pd\_entry} \land \textsf{table\_root(pml3\_entry}_{\textsf{pfn}})  = \textsf{pd} \\ \land \textsf{pt}+\textsf{PDPEX(va)} = \textsf{pt\_entry} \land \textsf{table\_root(pml2\_entry}_{\textsf{pfn}})  = \textsf{pt} \urcorner  \ast \\  \textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{pml4\_entry},\textsf{pdp\_entry},\textsf{pd\_entry},\textsf{pte}\},m)  \\ \ghostmaptoken{\delta{}s}{\rtv}{\delta} \ast
+      \ghostmaptoken{\textsf{id}}{(\mathsf{pml4\_entry})}{\textsf{4}}\ast \ghostmaptoken{\textsf{id}}{(\mathsf{plm3\_entry})}{\textsf{3}} \ast \ghostmaptoken{\textsf{id}}{(\mathsf{plm2\_entry})}{\textsf{2}}  \ast \\  \textsf{pml4\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{pml4\_entry},\textsf{l4e\_val}  \ast \\  \textsf{pdp\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{pdp\_entry},\textsf{l3e\_val}  \ast \\  \textsf{pd\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{pd\_entry},\textsf{l2e\_val}  \end{array} \right)  =\textsf{R}_{\textsf{walk} }
+    \\ \ast \left(\begin{array}{l} \textsf{rax} \mapsto \textsf{pte} \ast  \textsf{pt\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{pt\_entry},\textsf{pte} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present pte\_val})\urcorner \end{array}\right) =\textsf{R}_{\textsf{l1e}}\end{array} $
 \caption{\textsf{R}$_{\textsf{walk}}$: Resources obtained from invariant for walking page-table directory with \emph{non-existing entries} via calls to \textsf{pte\_get\_next\_table} in Figure \ref{fig:calltopteinitialize}}
@@
 \end{figure}
-Realizing a software page-table walk amounts to calling \textsf{pte\_get\_next\_table} for each level as shown in Figure \ref{walkpgdir}. 
-The special part of the specficication for a page table walk can be considered as accumulation of memory mappings for the page-table entries 
+Implementing a software page-table walk amounts to calling \textsf{pte\_get\_next\_table} for each level as shown in Figure \ref{walkpgdir}. 
+The key part of the specification and proof for a page table walk is accumulation of memory mappings for the page-table entries 
 visited and frame addresses for page-tables. 
@@
 ;; pte_t *pte = walkpgdir(pml4, a);
- mov  -0x8[rbp], rdi;
- mov  -0x16[rbp], rsi;
- mov  -0x24[rbp], rdx;
-;; pte = walkpgdir_succ(pml4, va);
  mov    -0x8[rbp], %rdi;

### 2025-03-25 examples.tex
bus draft

@@
 \else
-\begin{lstlisting}[mathescape,escapeinside={(*}{*)}]
-;;pte_t *walkpgdir(pte_t *pml4, const void *va) {
-... ;; Stack setup
-$\specline{\textsf{P} \ast \ulcorner\textsf{rtv+KERNBASE}=\textsf{pml4}\urcorner \ast \forall_{i\in\textsf{0..511}}\ghostmaptoken{\textsf{id}}{(\textsf{rtv}+i*8)}{\textsf{4}} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}}_{\rtv}$
-;;set up the stack for root address and virtual address    
-;;pte_t *pml4_entry = &pml4[PML4EX(va)]; // Virtual address of L4 entry(*\label{line:start_pml4_calc}*)
-mov    -0x8[rbp],rsi
-mov    -0x10[rbp],rdi
-shr    0x27,rdi       ;; Shift L4 index to lowest bits
-and    0x1ff,rdi      ;; Mask to just lower 9 bits (0x1ff=511)
-shl    0x3,rdi        ;; Multiply by 8
-add    rdi,rsi        ;; Add to pml4 (virtual) table base(*\label{line:end_pml4_calc}*)
-mov    rsi,-0x18[rbp] ;; Store to local variable pml4_entry; logical pml4_entry is physical, program variable is virtual
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{pml4\_entry}\},m) \ast \forall_{i\in\textsf{0..511}\setminus\{\textsf{PML4EX(va)}\}}\ghostmaptoken{\textsf{id}}{(\textsf{rtv}+i*8)}{\textsf{4}}\ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
-$\specline{\textsf{entry\_present(l4e\_val)} \wand \forall_{i\in\textsf{0..511}} \; \ghostmaptoken{\textsf{id}}{\textsf{table\_root(l4e\_val)+ i * 8}}{\textsf{3}} }_{\rtv} $
-$\specline{ \textsf{pml4\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{pml4\_entry l4e\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l4e\_val})\urcorner}_{\rtv}$
-;;pte_t *pdp = pte_get_next_table(pml4_entry);
-mov    -0x18[rbp],rdi
-...
-callq  c0 <pte_get_next_table>(*\label{line:first_getnext_call}*)
-;;save the physical next table address in rax
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{pml4\_entry}\},m)  \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
-$\specline{\forall_{i\in\textsf{0..511}} \ghostmaptoken{\textsf{id}}{\textsf{table\_root(l4e\_val'.pfn) + i * 8}}{\textsf{3}}  \ast \ulcorner\textsf{pdp-KERNBASE}=\textsf{table\_root(l4e\_val'.pfn)}\urcorner} $
-$\specline{ \textsf{pml4\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{pml4\_entry l4e\_val'} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l4e\_val})\urcorner}_{\rtv}$(*\label{line:first_pte_pointsto}*)
-$\specline{\texttt{rax}  \mapsto_{\textsf{r}} \textsf{table\_root(l4e\_val'.pfn)} }_{\rtv}$ ;; pte_get_next_table may have allocated a new page, updating entry
-;;pte_t *pdp_entry = &pdp[PDPEX(va)]; // Virtual address of L3 entry
-$\specline{\ulcorner\textsf{pdp}+\textsf{PDPEX(va)*8} = \textsf{pdp\_entry} \land \textsf{table\_root(l4e\_val'.pfn)}  = \textsf{pdp} \urcorner }_{\rtv}$
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{pml4\_entry},\textsf{pdp\_entry}\},m)  \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
-$\specline{\forall_{i\in\textsf{0..511}\setminus\{\textsf{PDPEX(va)}\}} \; \textsf{table\_root(pml4\_entry}_{\textsf{pfn}}) \textsf{ + i * 8} \mapsto_{\textsf{id}} \textsf{3} }_{\rtv}$
-$\specline{ \textsf{pml4\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac4}}\textsf{pml4\_entry}\; \textsf{l4e\_val'} \ast \ulcorner  \textsf{qfrac4} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l4e\_val'})\urcorner}_{\rtv}$(*\label{line:ex_l4_vpte}*)
-$\specline{ \textsf{pdp\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac3}}\textsf{pdp\_entry}\; \textsf{l3e\_val} \ast \ulcorner  \textsf{qfrac3} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l3e\_val})\urcorner}_{\rtv}$(*\label{line:ex_l3_vpte}*)
-$\specline{\textsf{entry\_present(l3e\_val)} \wand \forall_{i\in\textsf{0..511}} \;\ghostmaptoken{\textsf{id}}{\textsf{table\_root(pdp\_entry.pfn) + i * 8}}{ \textsf{2}} } _{\rtv} $
-;;pte_t *pd = pte_get_next_table(pdp_entry);
-... ;; Similar assembly to reach next level
-;;pte_t *pd_entry = &pd[PDEX(va)]; // Virtual address of L2 entry
-;;pte_t *pt = pte_get_next_table(pd_entry);
-$\specline{ \left( \begin{array}{l} \textsf{pdp}+\textsf{PDPEX(va)} = \textsf{pdp\_entry} \land \textsf{table\_root(pml4\_entry}_{\textsf{pfn}})  = \textsf{pdp} \\ \textsf{pd}+\textsf{PDEX(va)} = \textsf{pd\_entry} \land \textsf{table\_root(pml3\_entry}_{\textsf{pfn}})  = \textsf{pd} \\ \textsf{pt}+\textsf{PDPEX(va)} = \textsf{pt\_entry} \land \textsf{table\_root(pml2\_entry}_{\textsf{pfn}})  = \textsf{pt} \\  \textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{pml4\_entry},\textsf{pdp\_entry},\textsf{pd\_entry},\textsf{pt\_entry}\},m)  \\ \ghostmaptoken{\delta{}s}{\rtv}{\delta} \\
-\ghostmaptoken{\textsf{id}}{(\mathsf{pml4\_entry})}{\textsf{4}}\ast \ghostmaptoken{\textsf{id}}{(\mathsf{plm3\_entry})}{\textsf{3}} \ast \ghostmaptoken{\textsf{id}}{(\mathsf{plm2\_entry})}{\textsf{2}}  \\ \forall_{i\in\textsf{0..511}} \; \textsf{table\_root(pml4\_entry}_{\textsf{pfn}}) \textsf{ + i * 8} \mapsto_{\textsf{id}} \textsf{3}  \\ \forall_{i\in\textsf{0..511}} \; \textsf{table\_root(pdp\_entry}_{\textsf{pfn}}) \textsf{ + i * 8} \mapsto_{\textsf{id}} \textsf{2}  \\ \forall_{i\in\textsf{0..511}} \; \textsf{table\_root(pd\_entry}_{\textsf{pfn}}) \textsf{ + i * 8} \mapsto_{\textsf{id}} \textsf{1}  \\  \textsf{pml4\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l4e\_val} \\  \textsf{pdp\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l3e\_val} \\  \textsf{pd\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l2e\_val} \\   \textsf{pt\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l1e\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l1e\_val})\urcorner    \end{array} \right)  =\textsf{R}_{\textsf{walk}}}$
-;;access and return L1 entry
-;;return &pt[PTEX(va)]; // Virtual address of L1 entry
-...
-;; clean up the stack 
-\end{lstlisting}
+\todo[inline]{this is dead code, we're using the PLDI flag set to true}
+% \begin{lstlisting}[mathescape,escapeinside={(*}{*)}]
+% ;;pte_t *walkpgdir(pte_t *pml4, const void *va) {
+% ... ;; Stack setup
+% $\specline{\textsf{P} \ast \ulcorner\textsf{rtv+KERNBASE}=\textsf{pml4}\urcorner \ast \forall_{i\in\textsf{0..511}}\ghostmaptoken{\textsf{id}}{(\textsf{rtv}+i*8)}{\textsf{4}} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}}_{\rtv}$
+% ;;set up the stack for root address and virtual address    
+% ;;pte_t *pml4_entry = &pml4[PML4EX(va)]; // Virtual address of L4 entry(*\label{line:start_pml4_calc}*)
+% mov    -0x8[rbp],rsi
+% mov    -0x10[rbp],rdi
+% shr    0x27,rdi       ;; Shift L4 index to lowest bits
+% and    0x1ff,rdi      ;; Mask to just lower 9 bits (0x1ff=511)
+% shl    0x3,rdi        ;; Multiply by 8
+% add    rdi,rsi        ;; Add to pml4 (virtual) table base(*\label{line:end_pml4_calc}*)
+% mov    rsi,-0x18[rbp] ;; Store to local variable pml4_entry; logical pml4_entry is physical, program variable is virtual
+% $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{pml4\_entry}\},m) \ast \forall_{i\in\textsf{0..511}\setminus\{\textsf{PML4EX(va)}\}}\ghostmaptoken{\textsf{id}}{(\textsf{rtv}+i*8)}{\textsf{4}}\ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
+% $\specline{\textsf{entry\_present(l4e\_val)} \wand \forall_{i\in\textsf{0..511}} \; \ghostmaptoken{\textsf{id}}{\textsf{table\_root(l4e\_val)+ i * 8}}{\textsf{3}} }_{\rtv} $
+% $\specline{ \textsf{pml4\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{pml4\_entry l4e\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l4e\_val})\urcorner}_{\rtv}$
+% ;;pte_t *pdp = pte_get_next_table(pml4_entry);
+% mov    -0x18[rbp],rdi
+% ...
+% callq  c0 <pte_get_next_table>(*\label{line:first_getnext_call}*)
+% ;;save the physical next table address in rax
+% $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{pml4\_entry}\},m)  \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
+% $\specline{\forall_{i\in\textsf{0..511}} \ghostmaptoken{\textsf{id}}{\textsf{table\_root(l4e\_val'.pfn) + i * 8}}{\textsf{3}}  \ast \ulcorner\textsf{pdp-KERNBASE}=\textsf{table\_root(l4e\_val'.pfn)}\urcorner} $
+% $\specline{ \textsf{pml4\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{pml4\_entry l4e\_val'} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l4e\_val})\urcorner}_{\rtv}$(*\label{line:first_pte_pointsto}*)
+% $\specline{\texttt{rax}  \mapsto_{\textsf{r}} \textsf{table\_root(l4e\_val'.pfn)} }_{\rtv}$ ;; pte_get_next_table may have allocated a new page, updating entry
+% ;;pte_t *pdp_entry = &pdp[PDPEX(va)]; // Virtual address of L3 entry
+% $\specline{\ulcorner\textsf{pdp}+\textsf{PDPEX(va)*8} = \textsf{pdp\_entry} \land \textsf{table\_root(l4e\_val'.pfn)}  = \textsf{pdp} \urcorner }_{\rtv}$
+% $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{pml4\_entry},\textsf{pdp\_entry}\},m)  \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
+% $\specline{\forall_{i\in\textsf{0..511}\setminus\{\textsf{PDPEX(va)}\}} \; \textsf{table\_root(pml4\_entry}_{\textsf{pfn}}) \textsf{ + i * 8} \mapsto_{\textsf{id}} \textsf{3} }_{\rtv}$
+% $\specline{ \textsf{pml4\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac4}}\textsf{pml4\_entry}\; \textsf{l4e\_val'} \ast \ulcorner  \textsf{qfrac4} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l4e\_val'})\urcorner}_{\rtv}$(*\label{line:ex_l4_vpte}*)
+% $\specline{ \textsf{pdp\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac3}}\textsf{pdp\_entry}\; \textsf{l3e\_val} \ast \ulcorner  \textsf{qfrac3} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l3e\_val})\urcorner}_{\rtv}$(*\label{line:ex_l3_vpte}*)
+% $\specline{\textsf{entry\_present(l3e\_val)} \wand \forall_{i\in\textsf{0..511}} \;\ghostmaptoken{\textsf{id}}{\textsf{table\_root(pdp\_entry.pfn) + i * 8}}{ \textsf{2}} } _{\rtv} $
+% ;;pte_t *pd = pte_get_next_table(pdp_entry);
+% ... ;; Similar assembly to reach next level
+% ;;pte_t *pd_entry = &pd[PDEX(va)]; // Virtual address of L2 entry
+% ;;pte_t *pt = pte_get_next_table(pd_entry);
+% $\specline{ \left( \begin{array}{l} \textsf{pdp}+\textsf{PDPEX(va)} = \textsf{pdp\_entry} \land \textsf{table\_root(pml4\_entry}_{\textsf{pfn}})  = \textsf{pdp} \\ \textsf{pd}+\textsf{PDEX(va)} = \textsf{pd\_entry} \land \textsf{table\_root(pml3\_entry}_{\textsf{pfn}})  = \textsf{pd} \\ \textsf{pt}+\textsf{PDPEX(va)} = \textsf{pt\_entry} \land \textsf{table\_root(pml2\_entry}_{\textsf{pfn}})  = \textsf{pt} \\  \textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{pml4\_entry},\textsf{pdp\_entry},\textsf{pd\_entry},\textsf{pt\_entry}\},m)  \\ \ghostmaptoken{\delta{}s}{\rtv}{\delta} \\
+% \ghostmaptoken{\textsf{id}}{(\mathsf{pml4\_entry})}{\textsf{4}}\ast \ghostmaptoken{\textsf{id}}{(\mathsf{plm3\_entry})}{\textsf{3}} \ast \ghostmaptoken{\textsf{id}}{(\mathsf{plm2\_entry})}{\textsf{2}}  \\ \forall_{i\in\textsf{0..511}} \; \textsf{table\_root(pml4\_entry}_{\textsf{pfn}}) \textsf{ + i * 8} \mapsto_{\textsf{id}} \textsf{3}  \\ \forall_{i\in\textsf{0..511}} \; \textsf{table\_root(pdp\_entry}_{\textsf{pfn}}) \textsf{ + i * 8} \mapsto_{\textsf{id}} \textsf{2}  \\ \forall_{i\in\textsf{0..511}} \; \textsf{table\_root(pd\_entry}_{\textsf{pfn}}) \textsf{ + i * 8} \mapsto_{\textsf{id}} \textsf{1}  \\  \textsf{pml4\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l4e\_val} \\  \textsf{pdp\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l3e\_val} \\  \textsf{pd\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l2e\_val} \\   \textsf{pt\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l1e\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l1e\_val})\urcorner    \end{array} \right)  =\textsf{R}_{\textsf{walk}}}$
+% ;;access and return L1 entry
+% ;;return &pt[PTEX(va)]; // Virtual address of L1 entry
+% ...
+% ;; clean up the stack 
+% \end{lstlisting}
 \fi
@@
 One of the key tasks of a page fault handler in a general-purpose OS kernel is
-to map new pages into an address space by writing into an existing page table via a call
-\[\textsf{vaspace\_mappage(pte\_t *pml4, void *va,uintptr\_t fpaddr)}\] in Figure \ref{fig:mapping_code}.
+to map new pages into an address space by writing into an existing page table via a call\\
+\centerline{\textsf{vaspace\_mappage(pte\_t *pml4, void *va,uintptr\_t fpaddr)}}\\
+in Figure \ref{fig:mapping_code}.
 To do so, with a given allocated a fresh page (\textsf{fpaddr}), then calculate the appropriate
@@
  callq  291 <walkpgdir> (*\label{line:call_walkpgdir}*)
- $\specline{ \textsf{R}_{\textsf{walk}} \ast \left(\begin{array}{l}   \textsf{rax} \mapsto \textsf{pte} \ast  \textsf{pte} \mapsto_{\textsf{vpte,qfrac}} \textsf{pte\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present pte\_val}) \end{array}\right) =\textsf{R}_{\textsf{l1e}} }$
+ $\specline{ \textsf{R}_{\textsf{walk}} \ast \textsf{R}_{\textsf{l1e}} }$
  mov  rax, rdx;
  mov    rdx,-0x28[rbp];
-;;if (!pte->present){ 
+;;if (!pte->present){ (*\label{line:mappage_pte_present_start}*)
  mov    -0x28[rbp], rdx
@@
  cmp    0x0, r8
- jne    3e0 <vaspace_mappage+0x190>
+ jne    3e0 <vaspace_mappage+0x190> (*\label{line:mappage_pte_present_end}*)
  ...
@@
 mapped to a page ($\theta \; !!\; \vaddr = \texttt{None}$). Mapping a fresh
-phyiscal page to back the desired virtual page first requires ensuring
+physical page to back the desired virtual page first requires ensuring
 the existence of a memory location for an appropriate L1 table entry.
@@
 both for ensuring partial page table walk reaching the
-L1 entry (l1e) with the justification of higher levels of the page table exist (R$_{\textsf{walk}}$ in Figure \ref{fig:rwalk}), and for allowing access to the memory of the L1 entry via virtual address (R$_{\textsf{l1e}}$ in Figure \ref{fig:rwalk}).
-%A PTE points-to, as we already brifly mentioned, is defined just like the normal virtual points-to of Figure \ref{fig:virtualpointstosharing}, except the physical address (here, \textsf{pa}) is explicit in the assertion
-%rather than existentially quantified:
-%\looseness=-1
-% \[
-%\begin{array}{l}
-%    \vaddr\mapsto_{\textsf{vpte,q}} \; \paddr \; \vpage : \mathsf{vProp}~\Sigma \stackrel{\triangle}{=} 
-%    \exists \delta\ldotp
-%	(\lambda \mathit{cr3val}\ldotp
-%	\ghostmaptoken{\delta{}s}{\mathit{cr3val}}{\delta}) \ast 
-%  \fracghostmaptoken{\delta}{\vaddr}{\paddr}{\qfrac} \ast \paddr \mapsto_{\mathsf{p}} \vpage
-%\end{array}
-%\]
-%This supports rules for accessing memory
-%at that virtual address, but exposing the physical location being modified
-%makes this convenient to use for page table modifications, since we must ensure
-%the modified data is at the specific physical location that will be used as the L1 entry.
+L1 entry (l1e) by asserting that higher levels of the page table exist (R$_{\textsf{walk}}$ in Figure \ref{fig:rwalk}), 
+and for allowing access to the memory of the L1 entry via virtual address (R$_{\textsf{l1e}}$ in Figure \ref{fig:rwalk}).
+
+A PTE points-to, as we already brifly mentioned, is defined just like the normal virtual points-to of Figure \ref{fig:virtualpointstosharing}, except the physical address (here, \textsf{pa}) is explicit in the assertion
+rather than existentially quantified:\\
+\centerline{$
+    \vaddr\mapsto_{\textsf{vpte,q}} \; \paddr \; \vpage : \mathsf{vProp}~\Sigma \stackrel{\triangle}{=} 
+    \exists \delta\ldotp
+	(\lambda \mathit{cr3val}\ldotp
+	\ghostmaptoken{\delta{}s}{\mathit{cr3val}}{\delta}) \ast 
+  \fracghostmaptoken{\delta}{\vaddr}{\paddr}{\qfrac} \ast \paddr \mapsto_{\mathsf{p}} \vpage
+$}\\
+This supports rules for accessing memory
+at that virtual address, but exposing the physical location being modified
+makes this convenient to use for page table modifications, since we must ensure
+the modified data is at the specific physical location that will be used as the L1 entry.
 % After obtaining a virtual address \textsf{pte\_addr} in \textsf{rax} backed 
@@
 
-We already discussed for the upper level page-tables that the entry-exists checks are properly ensured ($\textsf{R}_{\textsf{walk}}$ in Figure \ref{fig:rwalk} and Line 27 in Figure \ref{walkpgdir}).
-However, for L1 level, this check is also checked left to the caller of the traversal function \textsf{walkpgdir}.
-In other words, unlike  to what we see in R$_{\textsf{walk}}$ for the upper levels where all entry-exists checks are dropped, the specification in R$_{\textsf{l1e}}$ ensures that page table entry for L1 needs to be checked on the caller site.
-By doing so, as we see in Figure \ref{fig:mapping_code}, the page reference \textsf{fpaddr} is linked to back the virtual address \textsf{va} only if it is not already referring to a physical resource (Lines 16-23 in Figure \ref{fig:mapping_code}).
+We already discussed for the upper level page-tables how the entry-present checks are handled.
+However, for L1 entries this check is left to the caller of the 
+traversal function \textsf{walkpgdir}. In other words, unlike what we see in R$_{\textsf{walk}}$ for the upper levels where all entry-present
+checks have already been performed, the specification in R$_{\textsf{l1e}}$ ensures that page table entry for L1 needs to be checked at the caller site. 
+By doing so, as we see in Figure \ref{fig:mapping_code}, the page reference \textsf{fpaddr} is linked to back the virtual address \textsf{va} 
+only if it is not already referring to a physical resource (Lines \ref{line:mappage_pte_present_start}--\ref{line:mappage_pte_present_end} in Figure \ref{fig:mapping_code}).
 
-The crucial and the only step in addition to traversing the page table in Figure \ref{walkpgdir} is actually updating the L1 entry (Line 35 in Figure \ref{fig:mapping_code}),
+The crucial step in addition to traversing the page table in Figure \ref{walkpgdir} is actually updating the L1 entry (Line \ref{line:updatepfn} in Figure \ref{fig:mapping_code}),
 via the virtual address

### 2025-03-25 examples.tex
Fixes on the figures and text

@@
   \vspace{-1em}
-$\begin{array}{l}\left( \begin{array}{l} \ulcorner \textsf{pdp}+\textsf{PDPEX(va)} = \textsf{pdp\_entry} \land \textsf{table\_root(pml4\_entry}_{\textsf{pfn}})  = \textsf{pdp} \\ \land \textsf{pd}+\textsf{PDEX(va)} = \textsf{pd\_entry} \land \textsf{table\_root(pml3\_entry}_{\textsf{pfn}})  = \textsf{pd} \\ \land \textsf{pt}+\textsf{PDPEX(va)} = \textsf{pt\_entry} \land \textsf{table\_root(pml2\_entry}_{\textsf{pfn}})  = \textsf{pt} \urcorner  \ast \\  \textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{pml4\_entry},\textsf{pdp\_entry},\textsf{pd\_entry},\textsf{pte}\},m)  \\ \ghostmaptoken{\delta{}s}{\rtv}{\delta} \ast
-      \ghostmaptoken{\textsf{id}}{(\mathsf{pml4\_entry})}{\textsf{4}}\ast \ghostmaptoken{\textsf{id}}{(\mathsf{plm3\_entry})}{\textsf{3}} \ast \ghostmaptoken{\textsf{id}}{(\mathsf{plm2\_entry})}{\textsf{2}}  \ast \\  \textsf{pml4\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{pml4\_entry},\textsf{l4e\_val}  \ast \\  \textsf{pdp\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{pdp\_entry},\textsf{l3e\_val}  \ast \\  \textsf{pd\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{pd\_entry},\textsf{l2e\_val}  \end{array} \right)  =\textsf{R}_{\textsf{walk} }
-    \\ \ast \left(\begin{array}{l} \textsf{rax} \mapsto \textsf{pte} \ast  \textsf{pt\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{pt\_entry},\textsf{pte} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present pte\_val})\urcorner \end{array}\right) =\textsf{R}_{\textsf{l1e}}\end{array} $
-\caption{\textsf{R}$_{\textsf{walk}}$: Resources obtained from invariant for walking page-table directory with \emph{non-existing entries} via calls to \textsf{pte\_get\_next\_table} in Figure \ref{fig:calltopteinitialize}}
+$\begin{array}{l}\left( \begin{array}{l} \ulcorner \textsf{pdp}+\textsf{PDPEX(va)} = \textsf{pdp\_entry} \land \textsf{table\_root(pml4\_entry}_{\textsf{pfn}})  = \textsf{pdp} \\ \land \textsf{pd}+\textsf{PDEX(va)} = \textsf{pd\_entry} \land \textsf{table\_root(pml3\_entry}_{\textsf{pfn}})  = \textsf{pd} \\ \land \textsf{pt}+\textsf{PDPEX(va)} = \textsf{pt\_entry} \land \textsf{table\_root(pml2\_entry}_{\textsf{pfn}})  = \textsf{pt} \urcorner  \ast \\  \textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{pml4\_entry},\textsf{pdp\_entry},\textsf{pd\_entry},\textsf{pt\_entry}\},m)  \\ \ghostmaptoken{\delta{}s}{\rtv}{\delta} \ast
+      \ghostmaptoken{\textsf{id}}{(\mathsf{pml4\_entry})}{\textsf{4}}\ast \ghostmaptoken{\textsf{id}}{(\mathsf{plm3\_entry})}{\textsf{3}} \ast \ghostmaptoken{\textsf{id}}{(\mathsf{plm2\_entry})}{\textsf{2}}  \ast \\  \textsf{pml4\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{pml4\_entry} \; \textsf{l4e\_val}  \ast \\  \textsf{pdp\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{pdp\_entry}\;\textsf{l3e\_val}  \ast \\  \textsf{pd\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{pd\_entry} \; \textsf{l2e\_val}  \end{array} \right)  =\textsf{R}_{\textsf{walk} }
+    \\ \ast \left(\begin{array}{l} \textsf{rax} \mapsto \textsf{pt\_entry+KERNBASE} \ast  \textsf{pt\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{pt\_entry}\;\textsf{l1e\_val} \ast \\ \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present pte\_val})\urcorner \end{array}\right) =\textsf{R}_{\textsf{l1e}}\end{array} $
+\caption{\textsf{R}$_{\textsf{walk}}$: Resources obtained from invariant for walking page-table directory with \emph{non-existing entries} via calls to \textsf{l1e\_get\_next\_table} in Figure \ref{fig:calltopteinitialize}}
 \label{fig:rwalk}
@@
  $\specline{\textsf{P} \ast \textsf{pml4} \mapsto_{\textsf{id}} \textsf{\_} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta} \ast \ulcorner\theta \; !!\; \vaddr = \texttt{None}\urcorner}_{\rtv}$
-;; pte_t *pte = walkpgdir(pml4, a);
+;; pte_t *pteaddr = walkpgdir(pml4, a);
  mov    -0x8[rbp], %rdi;
@@
  callq  291 <walkpgdir> (*\label{line:call_walkpgdir}*)
- $\specline{ \textsf{R}_{\textsf{walk}} \ast \textsf{R}_{\textsf{l1e}} }$
+ $\specline{ \textsf{R}_{\textsf{walk}} \ast \left(\begin{array}{l} \textsf{rax} \mapsto \textsf{pt\_entry+KERNBASE} \ast  \textsf{pt\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{pt\_entry}\;\textsf{l1e\_val} \ast \\ \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l1e\_val})\urcorner \end{array}\right) =\textsf{R}_{\textsf{l1e}} }$
  mov  rax, rdx;
  mov    rdx,-0x28[rbp];
-;;if (!pte->present){ (*\label{line:mappage_pte_present_start}*)
+;;if (!pteaddr->present){ (*\label{line:mappage_pte_present_start}*)
  mov    -0x28[rbp], rdx
@@
  ...
-;;pte->pfn          = PTE_ADDR_TO_PFN(pa);
+;;pteaddr->pfn          = PTE_ADDR_TO_PFN(fpaddr);
  mov    -0x18[rbp], rdx;
@@
  mov    rdi, [rsi] (*\label{line:updatepfn}*)
-$\specline{ \textsf{pte} \mapsto_{\textsf{vpte,1}} (\textsf{set\_pfn}(\textsf{pte\_val}, \textsf{PTE\_ADDR\_TO\_PFN}(\textsf{fpaddr}))) \ast  \ulcorner \lnot(\textsf{entry\_present pte\_val}) \urcorner \ast P}$
+$\specline{ \textsf{pt\_entry+KERNBASE} \mapsto_{\textsf{vpte,1}} \textsf{pt\_entry} \; (\textsf{set\_pfn}(\textsf{l1e\_val}, \textsf{PTE\_ADDR\_TO\_PFN}(\textsf{fpaddr}))) \ast  \ulcorner \lnot(\textsf{entry\_present l1e\_val}) \urcorner \ast \textsf{P}}$
  ;; Store updated entry back to L1 entry (*\label{line:l1entry_store}*)
@@
 ... ;;and return either unaccessible pte pointsto or one with fresh page (fpaddr)
- $\specline{ \begin{array}{l}\textsf{pte} \mapsto_{\textsf{vpte,qfrac}} \textsf{pte\_val} \ast \ulcorner (\textsf{entry\_present pte\_val}) \urcorner \ast P \\
- \lor \; \;  \textsf{pte} \mapsto_{\textsf{vpte,1}} (\textsf{set\_pfn}(\textsf{pte\_val}, \textsf{PTE\_ADDR\_TO\_PFN}(\textsf{fpaddr}))) \ast  \ulcorner \lnot(\textsf{entry\_present pte\_val}) \urcorner \ast P \end{array} }$
+ $\specline{ \begin{array}{l}\textsf{pt\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{pt\_entry} \; \textsf{l1e\_val} \ast \ulcorner (\textsf{entry\_present l1e\_val}) \urcorner \ast \textsf{P} \\
+ \lor \; \;  \textsf{pt\_entry+KERNBASE} \mapsto_{\textsf{vpte,1}} \textsf{pt\_entry} \; (\textsf{set\_pfn}(\textsf{l1e\_val}, \textsf{PTE\_ADDR\_TO\_PFN}(\textsf{fpaddr}))) \ast  \ulcorner \lnot(\textsf{entry\_present l1e\_val}) \urcorner \ast P \end{array} }$
 ;;}
@@
 \vspace{-1em}
-  \caption{Specification and proof of code for mapping a new page with $\textsf{R}_{\textsf{walk\_acc}}$ in Figure \ref{walkpgdir} expanded.}
+  \caption{Specification of updating L1 entry (\textsf{pt\_entry}) to reference a new page (\textsf{fpaddr}).}
 \label{fig:mapping_code}
@@
 The crucial step in addition to traversing the page table in Figure \ref{walkpgdir} is actually updating the L1 entry (Line \ref{line:updatepfn} in Figure \ref{fig:mapping_code}),
-via the virtual address
-(\textsf{pte}) known to translate to the appropriate physical address, in our example the L1
+via the virtual address (\textsf{pt\_entry+KERNBASE}) known to translate to the appropriate physical address, in our example the L1
 table entry address ($\textsf{PTE\_ADDR\_TO\_PFN(fpaddr)}$).

### 2025-03-25 examples.tex
a tiny fix

@@
       \ghostmaptoken{\textsf{id}}{(\mathsf{pml4\_entry})}{\textsf{4}}\ast \ghostmaptoken{\textsf{id}}{(\mathsf{plm3\_entry})}{\textsf{3}} \ast \ghostmaptoken{\textsf{id}}{(\mathsf{plm2\_entry})}{\textsf{2}}  \ast \\  \textsf{pml4\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{pml4\_entry} \; \textsf{l4e\_val}  \ast \\  \textsf{pdp\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{pdp\_entry}\;\textsf{l3e\_val}  \ast \\  \textsf{pd\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{pd\_entry} \; \textsf{l2e\_val}  \end{array} \right)  =\textsf{R}_{\textsf{walk} }
-    \\ \ast \left(\begin{array}{l} \textsf{rax} \mapsto \textsf{pt\_entry+KERNBASE} \ast  \textsf{pt\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{pt\_entry}\;\textsf{l1e\_val} \ast \\ \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present pte\_val})\urcorner \end{array}\right) =\textsf{R}_{\textsf{l1e}}\end{array} $
+    \\ \ast \left(\begin{array}{l} \textsf{rax} \mapsto \textsf{pt\_entry+KERNBASE} \ast  \textsf{pt\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{pt\_entry}\;\textsf{l1e\_val} \ast \\ \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l1e\_val})\urcorner \end{array}\right) =\textsf{R}_{\textsf{l1e}}\end{array} $
 \caption{\textsf{R}$_{\textsf{walk}}$: Resources obtained from invariant for walking page-table directory with \emph{non-existing entries} via calls to \textsf{l1e\_get\_next\_table} in Figure \ref{fig:calltopteinitialize}}

### 2025-03-25 examples.tex
"fold" Rl1e in Figure 14, and apparently we are down to the limit again

@@
  callq  291 <walkpgdir> (*\label{line:call_walkpgdir}*)
- $\specline{ \textsf{R}_{\textsf{walk}} \ast \left(\begin{array}{l} \textsf{rax} \mapsto \textsf{pt\_entry+KERNBASE} \ast  \textsf{pt\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{pt\_entry}\;\textsf{l1e\_val} \ast \\ \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l1e\_val})\urcorner \end{array}\right) =\textsf{R}_{\textsf{l1e}} }$
+ $\specline{ \textsf{R}_{\textsf{walk}} \ast \textsf{R}_{\textsf{l1e}} }$
  mov  rax, rdx;

### 2025-03-26 background.tex
inline tlb appendix mostly into related work, update related work
remove old todo

@@
 own right. Based on others' early progress on verifying TLB operations in isolation~\cite{syeda2020formal}, we expect
-it to be possible to combine this paper's insights with that support. Even without TLB modeling,
+it to be possible to combine this paper's insights with that support. 
+Section \ref{sec:relwork} elaborates briefly on the challenges involved.
+Even without TLB modeling,
 our logic already enables verification of virtual memory management functionality that prior verified kernel work
 either trusts completely (\textsc{seL4}, \textsc{CertiKOS}) or is incapable of reasoning about.
-Appendix \ref{apdx:tlb} gives more details on the challenges involved for the interested reader.
+% Appendix \ref{apdx:tlb} gives more details on the challenges involved for the interested reader.
 \looseness=-1
@@
 \label{sec:seplogic}
-\todo{Colin says: I added this b/c prior systems reviewers wanted more verification background...
-not sure how I feel about it}
+% \todo{Colin says: I added this b/c prior systems reviewers wanted more verification background...
+% not sure how I feel about it}
 Separation logic~\cite{reynolds02} is a descendant of classic Hoare logic~\cite{hoare69},

### 2025-07-13 examples_C.tex
some work on appendix

--- 2025-03-25 examples.tex
@@
 ;; pte_t *pteaddr = walkpgdir(pml4, a);
- mov    -0x8[rbp], %rdi;
- mov    -0x16[rbp], rsi;
- mov    0x0, rax;
+ mov    -0x8[rbp], %rdi
+ mov    -0x16[rbp], rsi
+ mov    0x0, rax
  callq  291 <walkpgdir> (*\label{line:call_walkpgdir}*)
  $\specline{ \textsf{R}_{\textsf{walk}} \ast \textsf{R}_{\textsf{l1e}} }$
- mov  rax, rdx;
- mov    rdx,-0x28[rbp];
+ mov  rax, rdx
+ mov    rdx,-0x28[rbp]
 ;;if (!pteaddr->present){ (*\label{line:mappage_pte_present_start}*)
@@
  jne    3e0 <vaspace_mappage+0x190> (*\label{line:mappage_pte_present_end}*)
- ...
+ ... ;; Not present
 ;;pteaddr->pfn          = PTE_ADDR_TO_PFN(fpaddr);
- mov    -0x18[rbp], rdx;
- shr    0xc, rdx;
- mov    -0x28[rbp], rsi;
- mov    [rsi], rdi;
+ mov    -0x18[rbp], rdx
+ shr    0xc, rdx
+ mov    -0x28[rbp], rsi
+ mov    [rsi], rdi ;; Load current entry in rdi
  movabs 0xffffffffff, rcx
@@
  or     rdx, rdi
- mov    rdi, [rsi] (*\label{line:updatepfn}*)
+ mov    rdi, [rsi] ;; Write back updated entry with new PFN (*\label{line:updatepfn}*)
 $\specline{ \textsf{pt\_entry+KERNBASE} \mapsto_{\textsf{vpte,1}} \textsf{pt\_entry} \; (\textsf{set\_pfn}(\textsf{l1e\_val}, \textsf{PTE\_ADDR\_TO\_PFN}(\textsf{fpaddr}))) \ast  \ulcorner \lnot(\textsf{entry\_present l1e\_val}) \urcorner \ast \textsf{P}}$

### 2025-07-14 examples.tex
+

@@
-%$\assert{\ulcorner \mathsf{aligned maddr} \urcorner \ast \mathsf{r14} \mapsto_{\textsf{r}} \textsf{r14v} \ast \mathsf{r13} \mapsto_{\textsf{r}} \textsf{maddr} \ast \mathsf{rdi} \mapsto_{\textsf{r}} \textsf{rdiv} \ast \mathsf{rax} \mapsto_{\textsf{r}} \textsf{raxv}}$
-%$\assert{\ulcorner \ptablestore !! \textsf{maddr} = \textsf{None} \urcorner \ast \ownGhost\gammaPred{\authfull{\ptableabswalk\ptablestore}} \ast \textsf{Pf}} $
-\newcommand{\sumwalkabsent}{
-  \ownGhost\gammaPred{\authfrag{\singletonMap{\texttt{entry+KERNBASE}}{(\textsf{qfrac}, \textsf{entry})}}}
-}
-
-\newcommand{\ventry}{\texttt{entry + KERNBASE}}
-\newcommand{\entry}{\texttt{entry}}
-\newcommand{\qfraczero}{\textsf{qfrac}}
-\newcommand{\true}{\textsf{true}}
-\tikzstyle{boxedassert_border} = [sharp corners,line width=0.2pt]
-\NewDocumentCommand \boxedassertpv {O{} m o}{%
-	\tikz[baseline=(m.base)]{
-		%	  \node[rectangle, draw,inner sep=0.8pt,anchor=base,#1] (m) {${#2}\mathstrut$};
-		\node[rectangle,inner sep=1.5pt,outer sep=0.2pt,anchor=base] (m) {${\,#2\,}\mathstrut$};
-		\draw[#1,boxedassert_border] ($(m.south west)$) rectangle ($(m.north east)$);
-	}\IfNoValueF{#3}{^{\,#3}}%
-}
-\newcommand*{\knowInvpv}[2]{\boxedassertpv{#2}[#1]}
-\newcommand*{\ownGhostpv}[2]{\boxedassertpv[dash dot]{#2}[#1]}
-
-\newcommand{\sumpv}[3]{
-  \ownGhostpv\gammaPred{\authfrag{\singletonMap{#1}{(#2, #3)}}}
-}
-
-\newcommand{\pvmapping}[1]{\mathcal{A}\textsf{PVMappings}(#1)}
-
-\newcommand{\fpaddr}{\texttt{fpaddr}}
-\newcommand{\specline}[1]{{\color{blue}\left\{#1\right\}}}
-\newcommand{\sumapacesfull}[2]{
-  \ownGhost\gammaPreds{\authfull{\singletonMap{#1}{#2}}}
-}
-\section{Experiments}
+\section{Experiments with Assembly Language}
 \label{sec:experiment}

### 2025-07-14 examples_C.tex
+

@@
 %$\assert{\ulcorner \ptablestore !! \textsf{maddr} = \textsf{None} \urcorner \ast \ownGhost\gammaPred{\authfull{\ptableabswalk\ptablestore}} \ast \textsf{Pf}} $
+\definecolor{main-color}{rgb}{0.6627, 0.7176, 0.7764}
+\definecolor{back-color}{rgb}{0.1686, 0.1686, 0.1686}
+\definecolor{string-color}{rgb}{0.3333, 0.5254, 0.345}
+\definecolor{key-color}{rgb}{0.8, 0.47, 0.196}
+
+\lstdefinestyle{mystyle}
+{
+    language = C,
+    basicstyle = {\footnotesize,\color{black}},
+    backgroundcolor = {\color{white}},
+    stringstyle = {\color{string-color}},
+    keywordstyle = {\color{key-color}},
+    keywordstyle = [2]{\color{lime}},
+    keywordstyle = [3]{\color{yellow}},
+    keywordstyle = [4]{\color{teal}},
+    otherkeywords = {;,<<,>>,++},
+    morekeywords = [2]{;},
+    morekeywords = [3]{<<, >>},
+    morekeywords = [4]{++},
+    escapeinside={(*}{*)},
+}
 \newcommand{\sumwalkabsent}{
@@
 \ifPLDI
-\begin{lstlisting}[mathescape, basicstyle=\footnotesize,escapeinside={(*}{*)}]
-;;pte_t *pte_get_next_table(pte_t *entry) { ... // setting up the stack for a pte_t *next elided
+\begin{lstlisting}[firstnumber=1, mathescape, language = C, style=mystyle]
+pte_t *pte_get_next_table(pte_t *entry) {
+  pte_t *next;
+  /*setting up the stack for a pte_t *next elided*/
   $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{entry}\},m)  \ast \texttt{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast  \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{next} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{\_} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{\_} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
@@
   $\specline{\ulcorner\textsf{entry\_present(entry\_val)}\urcorner \wand \forall_{i\in\textsf{0..511}} \; \ghostmaptoken{\textsf{id}}{\textsf{table\_root(entry\_val.pfn) + i * 8}}{\textsf{v-1}} }_{\rtv} $ (*\label{line:conditional_children}*)
-mov    -0x8[rbp],rdi $\specline{\ldots \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry} \ast \ldots }$
-mov     rdi, r8 $\specline{\ldots \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry} \ast \ldots }$
-mov    [r8],rdi (*\label{line:read_entry_contents}*) $\specline{\ldots \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val}  \ast\ldots}$ ;; reading page table entry Justified by virtual pte-pointsto
-and    0x1,rdi;  mov    rdi,rax; cmp    0x0,rax ;;  if (!entry->present) {(* \label{line:check_entry_present} *)(* \label{line:mask_present} *)
-jne    161 <pte_get_next_table+0xa1> (* \label{line:check_entry_present_jump} *) ;; Jump if the present bit is not zero; no need to allocate next level
+  $\specline{\ldots \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry} \ast \ldots }$
+  $\specline{\ldots \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry} \ast \ldots }$
+   (*\label{line:read_entry_contents}*)$\specline{\ldots \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val}  \ast\ldots}$
+   /*reading page table entry Justified by virtual pte-pointsto*/
+   if (!entry->present) {/* Jump if the present bit is not zero; no need to allocate next level (* \label{line:check_entry_present} *)(* \label{line:mask_present} *) */
      $\specline{\ldots \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val \& 0x1} }_{\rtv}$
-mov    rbp,rdi; sub    0x10, rdi ;; Store the value of rbp minus 16 bytes (address of (*\textsf{next}*)) into rdi (*\label{line:pass_addrof_next}*)(*\label{line:alloc_path_start}*)
+     /*Store the value of rbp minus 16 bytes (address of (*\textsf{next}*)) into rdi (*\label{line:pass_addrof_next}*)(*\label{line:alloc_path_start}*) */
      $\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \land \lnot(\textsf{entry\_present entry\_val})  \urcorner }_{\rtv}$ (*\label{line:after_concluding_qfrac1}*)
-callq  70 <pte_initialize> ;;pte_initialize(entry);(* \label{line:call_to_pte_initialize} *)
+     pte_initialize(entry);(* \label{line:call_to_pte_initialize} *)
      $\specline{ \ldots \ast \ulcorner  \textsf{qfrac} = 1 \land \lnot(\textsf{entry\_present entry\_val})  \urcorner \ast \ulcorner  \textsf{qfrac} = 1 \land \lnot(\textsf{entry\_present (pfn\_set (entry\_val nextpaddr))}  \urcorner   }_{\rtv}$
@@
 511} } \ldotp  \ghostmaptoken{\textsf{id}}{((\textsf{table\_root (pte\_initialized (pfn\_set(entry\_val nextpaddr)))}) + \textsf{i * 8})}{\textsf{v-1}} \end{array}  }$ (*\label{line:page_of_caps}*)
-... ;;entry value updates: entry->pfn = nextpaddr; entry->present = 1; (*\label{line:install_new_entry}*)
-... ;;now we know that entry is initialized, so we satisfy the condition to access children list
+     /*... entry value updates:*/
+     entry->pfn = nextpaddr; entry->present = 1; (*\label{line:install_new_entry}*)
+     /*... now we know that entry is initialized, so we satisfy the condition to access children list */
      $\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}}  \textsf{pte\_initialized(pfn\_set(entryv nextpaddr))} \ast \ulcorner  \textsf{qfrac} = 1 \land \lnot(\textsf{entry\_present entry\_val})  \urcorner }_{\rtv}$
@@
 511} } \ldotp  \ghostmaptoken{\textsf{id}}{((\textsf{table\_root (pte\_initialized (pfn\_set(entry\_val nextpaddr)))}) + \textsf{i * 8})}{\textsf{v-1}}  }$ (*\label{line:alloc_path_end}*)
-;;} (*\label{line:end_of_allocation_path}*)
-... ;; Code after conditional continued in Figure (*\ref{fig:p2v}*)
+    } (*\label{line:end_of_allocation_path}*)
+   /*...  Code after conditional continued in Figure (*\ref{fig:p2v}*)*/
 \end{lstlisting}

### 2025-07-14 examples_C.tex
+

@@
   /*...  Code after conditional continued in Figure (*\ref{fig:p2v}*)*/
+   $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \}),m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{rcx}  \mapsto_{\textsf{r}} \textsf{\_}  \ast \textsf{entry} \mapsto_{\textsf{id}} \textsf{\_} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
+  $\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{(pte\_initialized (entry\_val.pfn))} \urcorner }_{\rtv}$
+  $\specline{ \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{(pte\_initialized (entry\_val.pfn)))} \ast \texttt{rax} \mapsto_{\textsf{r}} \textsf{ table\_root (pte\_initialize(entry\_val.pfn))} }_{\rtv}$
+  $\specline{\forall_{i\in \textsf{0 ...
+511} } \ldotp  \ghostmaptoken{\textsf{id}}{((\textsf{table\_root (pte\_initialized (entry\_val.pfn)))}) + \textsf{i * 8})}{\textsf{v-1}}  }$ (*\label{line:children}*)
+  uintptr_t next_phys_addr = PTE_PFN_TO_ADDR(entry->pfn)
+  uintptr_t next_virt_addr = (uintptr_t) P2V(next_phys_addr); (*\label{line:p2v}*) 
+  /*... clean up the stack and return the rax value*/
+  return next;
+}
 \end{lstlisting}
@@
 
-\begin{figure}\footnotesize
-\begin{lstlisting}[mathescape,escapeinside={(*}{*)}]
-... ;; Continued from Figure (*\ref{fig:calltopteinitialize}*); assertions below specialized to non-allocating path for clarity
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \}),m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{rcx}  \mapsto_{\textsf{r}} \textsf{\_}  \ast \textsf{entry} \mapsto_{\textsf{id}} \textsf{\_} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
-$\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{(pte\_initialized (entry\_val.pfn))} \urcorner }_{\rtv}$
-$\specline{ \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{(pte\_initialized (entry\_val.pfn)))} \ast \texttt{rax} \mapsto_{\textsf{r}} \textsf{ table\_root (pte\_initialize(entry\_val.pfn))} }_{\rtv}$
-$\specline{\forall_{i\in \textsf{0 ...
+%\begin{figure}\footnotesize
+%  \begin{lstlisting}[mathescape,escapeinside={(*}{*)}]
+%  /*... Continued from Figure (*\ref{fig:calltopteinitialize}*); assertions below specialized to non-allocating path for clarity */
+%  $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \}),m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{rcx}  \mapsto_{\textsf{r}} \textsf{\_}  \ast \textsf{entry} \mapsto_{\textsf{id}} \textsf{\_} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
+%  $\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{(pte\_initialized (entry\_val.pfn))} \urcorner }_{\rtv}$
+%  $\specline{ \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{(pte\_initialized (entry\_val.pfn)))} \ast \texttt{rax} \mapsto_{\textsf{r}} \textsf{ table\_root (pte\_initialize(entry\_val.pfn))} }_{\rtv}$
+%  $\specline{\forall_{i\in \textsf{0 ...
 511} } \ldotp  \ghostmaptoken{\textsf{id}}{((\textsf{table\_root (pte\_initialized (entry\_val.pfn)))}) + \textsf{i * 8})}{\textsf{v-1}}  }$ (*\label{line:children}*)
-;;uintptr_t next_virt_addr = (uintptr_t) P2V(entry.pfn<<12); (*\label{line:p2v}*) 
-movabs KERNBASE,rcx $\specline{\ldots \ast  \texttt{rcx}  \mapsto_{\textsf{r}} \textsf{KERNBASE}  \ast \ldots}_{\rtv}$
-add    rcx,rax $\specline{ \ldots \ast \texttt{rax} \mapsto_{\textsf{r}} \textsf{ table\_root (pte\_initialize(entry\_val.pfn)) + KERNBASE}  \ast \ldots}_{\rtv}$
-... ;;clean up the stack and return the rax value
-\end{lstlisting}
-\vspace{-1em}
-\caption{Converting a physical address of a PTE to a virtual address (w/o instruction pointer or flag updates).
- % Abreviating the other relevant resources (including the lower entries) peeled-out of the invariant as $\textsf{R}_{\textsf{children}}$
-\vspace{-1em}
-}
-\label{fig:p2v}
-\end{figure}
+%  uintptr_t next_phys_addr = PTE_PFN_TO_ADDR(entry->pfn)
+%  uintptr_t next_virt_addr = (uintptr_t) P2V(next_phys_addr); (*\label{line:p2v}*) 
+%  /*... clean up the stack and return the rax value*/
+%  return next
+%}
+%\end{lstlisting}
+%\vspace{-1em}
+%\caption{Converting a physical address of a PTE to a virtual address (w/o instruction pointer or flag updates).
+% % Abreviating the other relevant resources (including the lower entries) peeled-out of the invariant as $\textsf{R}_{\textsf{children}}$
+%\vspace{-1em}
+%}
+%\label{fig:p2v}
+%\end{figure}
 Figure \ref{fig:p2v} shows the verification of the end of \lstinline|pte_get_next_table| specialized to the case where 

### 2025-07-15 examples_C.tex
+

@@
   uintptr_t next_phys_addr = PTE_PFN_TO_ADDR(entry->pfn)
-  uintptr_t next_virt_addr = (uintptr_t) P2V(next_phys_addr); (*\label{line:p2v}*) 
+  uintptr_t next_virt_addr = (uintptr_t) P2V(next_phys_addr); (*\label{line:p2v}*);
+  nezt = (pte_t *) next_virt_addr;
   /*... clean up the stack and return the rax value*/
@@
 $\specline{\textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,1}} \textsf{entry entryv} \ast \ghostmaptoken{\textsf{id}}{(\mathsf{entry})}{\textsf{level}} }$
-;;void pte_initialize(pte_t *entry) { ... Stack frame setup elided    
-callq  81 <kalloc> ;;allocate a full zeroed page for 512 8-byte entries(*\label{line:call_to_kalloc}*)
-mov    rax,-0x10[rbp] ;; Store into 'local'
-;;entry->pfn = PTE_ADDR_TO_PFN((uintptr_t) local);
-mov    -0x10[rbp],rax; mov    -0x8[rbp],rdi; mov    rax,[rdi]
+void pte_initialize(pte_t *entry) {
+  /* ... Stack frame setup elided */
+  /*allocate a full zeroed page for 512 8-byte entries(*\label{line:call_to_kalloc}*)*/
+  pte_t *local = kalloc();
+  entry->pfn = PTE_ADDR_TO_PFN((uintptr_t) local);
 $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)   \ast \texttt{rbp-16} \mapsto_{\textsf{r}} \textsf{\_} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry+KERNBASE} \ast \texttt{rax} \mapsto_{\textsf{r}} \textsf{nextpaddr} \ast \lnot(\textsf{entry\_present entry\_val}) }$
@@
 $\specline{\ghostmaptoken{\delta{}s}{\rtv}{\delta} \ast \begin{array}{l} \ulcorner\textsf{entry\_present (pte\_initialize(pfn\_set(entry\_val nextpaddr)),level)}\urcorner \wand \\  \;\;\;\;\;\;\; \forall_{i\in\textsf{0..511}} \; \textsf{table\_root (pte\_initialized (pfn\_set (entry\_val nextpaddr))) + i * 8} \mapsto_{\textsf{id}} \textsf{level-1}    \end{array}  }_{\rtv}$
-... ;;clean up the stack, return
+  /*... ;;clean up the stack,*/
+  return
+}
 \end{lstlisting}

### 2025-07-15 examples_C.tex
+

@@
 \begin{lstlisting}[mathescape,escapeinside={(*}{*)}]
-;;pte_t *walkpgdir(pte_t *pml4, const void *va) {
-... ;; Stack setup
+pte_t *walkpgdir(pte_t *pml4, const void *va) {
+  /*... ;; Stack setup*/
   $\specline{\textsf{P} \ast \ulcorner\textsf{rtv+KERNBASE}=\textsf{pml4}\urcorner \ast \forall_{i\in\textsf{0..511}}\ghostmaptoken{\textsf{id}}{(\textsf{rtv}+i*8)}{\textsf{4}} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}}_{\rtv}$
-;;pte_t *pml4_entry = &pml4[PML4EX(va)]; // Virtual address of L4 entry(*\label{line:start_pml4_calc}*)
-mov    -0x8[rbp],rsi; mov    -0x10[rbp],rdi; ;; Load pml4 and va
-;; Extract 9-bit (0x1ff=511) L4 index from va, multiply by 8 (byte index), add to pml4 for entry address
-shr    0x27,rdi; and    0x1ff,rdi; shl    0x3,rdi; add    rdi,rsi(*\label{line:end_pml4_calc}*)
-mov    rsi,-0x18[rbp] ;; Store virt. addr. to local var pml4_entry; logical pml4_entry is physical!
+  pte_t *pml4_entry = &pml4[PML4EX(va)]; // Virtual address of L4 entry(*\label{line:start_pml4_calc}*)
+  /*Load pml4 and va:*/
+  /*Extract 9-bit (0x1ff=511) L4 index from va, multiply by 8 (byte index), add to pml4 for entry address (*\label{line:end_pml4_calc}*) */
+  /*Store virt. addr. to local var pml4_entry; logical pml4_entry is physical!*/
   $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{pml4\_entry}\},m) \ast \forall_{i\in\textsf{0..511}\setminus\{\textsf{PML4EX(va)}\}}\ghostmaptoken{\textsf{id}}{(\textsf{rtv}+i*8)}{\textsf{4}}\ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
@@
   $\specline{ \textsf{pml4\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{pml4\_entry l4e\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l4e\_val})\urcorner}_{\rtv}$
-;;pte_t *pdp = pte_get_next_table(pml4_entry);
-mov    -0x18[rbp],rdi; ...; callq  c0 <pte_get_next_table>(*\label{line:first_getnext_call}*)
-;;save the physical next table address in rax
+  pte_t *pdp = pte_get_next_table(pml4_entry); /*(*\label{line:first_getnext_call}*) save the physical next table address in rax*/
   $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{pml4\_entry}\},m)  \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
@@
   $\specline{ \textsf{pml4\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{pml4\_entry l4e\_val'} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l4e\_val})\urcorner}_{\rtv}$(*\label{line:first_pte_pointsto}*)
-$\specline{\texttt{rax}  \mapsto_{\textsf{r}} \textsf{table\_root(l4e\_val'.pfn)} }_{\rtv}$ ;; pte_get_next_table may have allocated a new page, updating entry
-;;pte_t *pdp_entry = &pdp[PDPEX(va)]; // Virtual address of L3 entry
+  $\specline{\texttt{rax}  \mapsto_{\textsf{r}} \textsf{table\_root(l4e\_val'.pfn)} }_{\rtv}$
+  /*pte_get_next_table may have allocated a new page, updating entry*/
+  pte_t *pdp_entry = &pdp[PDPEX(va)]; /* Virtual address of L3 entry*/
   $\specline{\ulcorner\textsf{pdp}+\textsf{PDPEX(va)*8} = \textsf{pdp\_entry} \land \textsf{table\_root(l4e\_val'.pfn)}  = \textsf{pdp} \urcorner }_{\rtv}$
@@
   $\specline{\textsf{entry\_present(l3e\_val)} \wand \forall_{i\in\textsf{0..511}} \;\ghostmaptoken{\textsf{id}}{\textsf{table\_root(pdp\_entry.pfn) + i * 8}}{ \textsf{2}} } _{\rtv} $
-... ;; Similar assembly & proofs for levels 3 and 2
-... ;;return &pt[PTEX(va)]; // return address of L1 entry, similar index computation
+  /*... ;; Similar assembly & proofs for levels 3 and 2*/
+  return &pt[PTEX(va)]; /* return address of L1 entry, similar index computation*/
   $\specline{\textsf{R}_{\textsf{walk}} \ast \textsf{R}_{\textsf{l1e}} }_{\rtv}$

### 2025-07-15 examples_C.tex
+

@@
   return &pt[PTEX(va)]; /* return address of L1 entry, similar index computation*/
+}
 $\specline{\textsf{R}_{\textsf{walk}} \ast \textsf{R}_{\textsf{l1e}} }_{\rtv}$
@@
   \begin{lstlisting}[mathescape,escapeinside={(*}{*)}]
-;;complstatus_t vaspace_mappage(pte_t *pml4, void *va,uintptr_t fpaddr ) {
-... ;;setting up the stack      
+complstatus_t vaspace_mappage(pte_t *pml4, void *va,uintptr_t fpaddr ) {
+  /*... ;;setting up the stack */ 
   $\specline{\textsf{P} \ast \textsf{pml4} \mapsto_{\textsf{id}} \textsf{\_} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta} \ast \ulcorner\theta \; !!\; \vaddr = \texttt{None}\urcorner}_{\rtv}$
-;; pte_t *pteaddr = walkpgdir(pml4, a);
- mov    -0x8[rbp], %rdi
- mov    -0x16[rbp], rsi
- mov    0x0, rax
- callq  291 <walkpgdir> (*\label{line:call_walkpgdir}*)
+  pte_t *pteaddr = walkpgdir(pml4, a); (*\label{line:call_walkpgdir}*)
   $\specline{ \textsf{R}_{\textsf{walk}} \ast \textsf{R}_{\textsf{l1e}} }$
- mov  rax, rdx
- mov    rdx,-0x28[rbp]
-;;if (!pteaddr->present){ (*\label{line:mappage_pte_present_start}*)
- mov    -0x28[rbp], rdx
- mov    [rdx], rcx
- and    0x1, rcx
- mov    rcx, r8
- cmp    0x0, r8
- jne    3e0 <vaspace_mappage+0x190> (*\label{line:mappage_pte_present_end}*)
- ... ;; Not present
-;;pteaddr->pfn          = PTE_ADDR_TO_PFN(fpaddr);
- mov    -0x18[rbp], rdx
- shr    0xc, rdx
- mov    -0x28[rbp], rsi
- mov    [rsi], rdi ;; Load current entry in rdi
- movabs 0xffffffffff, rcx
- and    rcx, rdx
- shl    0xc, rdx
- movabs 0xfff0000000000fff, rcx
- and    rcx, rdi
- or     rdx, rdi
- mov    rdi, [rsi] ;; Write back updated entry with new PFN (*\label{line:updatepfn}*)
+  if (!pteaddr->present){ (*\label{line:mappage_pte_present_start}*)
+    /*... ;; Not present(*\label{line:mappage_pte_present_end}*)*/
+    pteaddr->pfn = PTE_ADDR_TO_PFN(fpaddr); /*Store updated entry back to L1 entry (*\label{line:l1entry_store}*)*/
     $\specline{ \textsf{pt\_entry+KERNBASE} \mapsto_{\textsf{vpte,1}} \textsf{pt\_entry} \; (\textsf{set\_pfn}(\textsf{l1e\_val}, \textsf{PTE\_ADDR\_TO\_PFN}(\textsf{fpaddr}))) \ast  \ulcorner \lnot(\textsf{entry\_present l1e\_val}) \urcorner \ast \textsf{P}}$
- ;; Store updated entry back to L1 entry (*\label{line:l1entry_store}*)
-;;}
-... ;;set present bit in entry, then clean up the stack
+  }
+  /*... ;;set present bit in entry, then clean up the stack
   ... ;;and return either unaccessible pte pointsto or one with fresh page (fpaddr)
+  */
   $\specline{ \begin{array}{l}\textsf{pt\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{pt\_entry} \; \textsf{l1e\_val} \ast \ulcorner (\textsf{entry\_present l1e\_val}) \urcorner \ast \textsf{P} \\
  \lor \; \;  \textsf{pt\_entry+KERNBASE} \mapsto_{\textsf{vpte,1}} \textsf{pt\_entry} \; (\textsf{set\_pfn}(\textsf{l1e\_val}, \textsf{PTE\_ADDR\_TO\_PFN}(\textsf{fpaddr}))) \ast  \ulcorner \lnot(\textsf{entry\_present l1e\_val}) \urcorner \ast P \end{array} }$
-;;}
+}
 \end{lstlisting}

### 2025-07-15 examples_C.tex
+

@@
   \begin{lstlisting}[mathescape,escapeinside={(*}{*)}]
-complstatus_t vaspace_mappage(pte_t *pml4, void *va,uintptr_t fpaddr ) {
+void vaspace_mappage(pte_t *pml4, void *va,uintptr_t fpaddr ) {
   /*... ;;setting up the stack */ 
@@
     $\specline{ \textsf{pt\_entry+KERNBASE} \mapsto_{\textsf{vpte,1}} \textsf{pt\_entry} \; (\textsf{set\_pfn}(\textsf{l1e\_val}, \textsf{PTE\_ADDR\_TO\_PFN}(\textsf{fpaddr}))) \ast  \ulcorner \lnot(\textsf{entry\_present l1e\_val}) \urcorner \ast \textsf{P}}$
+    /*... ;;set present bit in entry, then clean up the stack*/
+    pteaddr->presetn = 1;
   }
-  /*... ;;set present bit in entry, then clean up the stack
+  /*
   ... ;;and return either unaccessible pte pointsto or one with fresh page (fpaddr)
   */
+  return error;
   $\specline{ \begin{array}{l}\textsf{pt\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{pt\_entry} \; \textsf{l1e\_val} \ast \ulcorner (\textsf{entry\_present l1e\_val}) \urcorner \ast \textsf{P} \\

### 2025-07-15 examples_C.tex
+

@@
 void pte_initialize(pte_t *entry) {
-  /* ... Stack frame setup elided */
   /*allocate a full zeroed page for 512 8-byte entries(*\label{line:call_to_kalloc}*)*/
@@
   $\specline{\ghostmaptoken{\delta{}s}{\rtv}{\delta} \ast \begin{array}{l} \ulcorner\textsf{entry\_present (pte\_initialize(pfn\_set(entry\_val nextpaddr)),level)}\urcorner \wand \\  \;\;\;\;\;\;\; \forall_{i\in\textsf{0..511}} \; \textsf{table\_root (pte\_initialized (pfn\_set (entry\_val nextpaddr))) + i * 8} \mapsto_{\textsf{id}} \textsf{level-1}    \end{array}  }_{\rtv}$
-  /*... ;;clean up the stack,*/
-  return
+  return;
 }
@@
   /*... ;; Similar assembly & proofs for levels 3 and 2*/
-  return &pt[PTEX(va)]; /* return address of L1 entry, similar index computation*/
+  pte_t *pd = pte_get_next_table(pdp_entry);
+  pte_t pd_entry = &pd[PDEX(va)];
+  pte_t *pt = pte_get_next_table(pd_entry);
+  /* return address of L1 entry, similar index computation*/
+  return &pt[PTEX(va)]; 
 }

### 2025-07-15 examples_C.tex
right before removing regs and details from the spec
+

@@
   pte_t *next;
-  /*setting up the stack for a pte_t *next elided*/
-  $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{entry}\},m)  \ast \texttt{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast  \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{next} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{\_} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{\_} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
+  $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{entry}\},m) \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
   $\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present entry\_val})\urcorner}_{\rtv}$ (*\label{line:get_next_vpte_precondition}*)
   $\specline{\ulcorner\textsf{entry\_present(entry\_val)}\urcorner \wand \forall_{i\in\textsf{0..511}} \; \ghostmaptoken{\textsf{id}}{\textsf{table\_root(entry\_val.pfn) + i * 8}}{\textsf{v-1}} }_{\rtv} $ (*\label{line:conditional_children}*)
-  $\specline{\ldots \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry} \ast \ldots }$
-  $\specline{\ldots \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry} \ast \ldots }$
   (*\label{line:read_entry_contents}*)$\specline{\ldots \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val}  \ast\ldots}$
@@
     $\specline{ \ldots \ast \ulcorner  \textsf{qfrac} = 1 \land \lnot(\textsf{entry\_present entry\_val})  \urcorner \ast \ulcorner  \textsf{qfrac} = 1 \land \lnot(\textsf{entry\_present (pfn\_set (entry\_val nextpaddr))}  \urcorner   }_{\rtv}$
-    $\specline{ \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{pfn\_set(entry\_val nextpaddr)}   }_{\rtv}$
     $\specline{ \begin{array}{l}\textsf{entry\_present (pte\_initialized (pfn\_set(entryv nextpaddr)))} \wand  \\ \;\;\;\;\;\;\;\;\  \forall_{i\in \textsf{0 ...
@@
     $\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}}  \textsf{pte\_initialized(pfn\_set(entryv nextpaddr))} \ast \ulcorner  \textsf{qfrac} = 1 \land \lnot(\textsf{entry\_present entry\_val})  \urcorner }_{\rtv}$
-    $\specline{ \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{pte\_initialized(pfn\_set(entry\_val nextpaddr))}  \ast rax \mapsto_{r} \textsf{table\_root (pte\_initialized (pfn\_set(entry\_val nextpaddr)))}  }_{\rtv}$
     $\specline{\forall_{i\in \textsf{0 ...
@@
   /*...  Code after conditional continued in Figure (*\ref{fig:p2v}*)*/
-   $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \}),m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{rcx}  \mapsto_{\textsf{r}} \textsf{\_}  \ast \textsf{entry} \mapsto_{\textsf{id}} \textsf{\_} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
+   $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \}),m) \ast \textsf{entry} \mapsto_{\textsf{id}} \textsf{\_} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
   $\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{(pte\_initialized (entry\_val.pfn))} \urcorner }_{\rtv}$
-  $\specline{ \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{(pte\_initialized (entry\_val.pfn)))} \ast \texttt{rax} \mapsto_{\textsf{r}} \textsf{ table\_root (pte\_initialize(entry\_val.pfn))} }_{\rtv}$
   $\specline{\forall_{i\in \textsf{0 ...
@@
   uintptr_t next_virt_addr = (uintptr_t) P2V(next_phys_addr); (*\label{line:p2v}*);
-  nezt = (pte_t *) next_virt_addr;
+  next = (pte_t *) next_virt_addr;
   /*... clean up the stack and return the rax value*/
@@
   \begin{lstlisting}[mathescape,escapeinside={(*}{*)}]
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)   \ast \texttt{rbp-16} \mapsto_{\textsf{r}} \textsf{\_} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry+KERNBASE}  \ast \texttt{rax} \mapsto_{\textsf{r}} \textsf{\_} \ast \lnot(\textsf{entry\_present entry\_val} }_{\rtv}$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \lnot(\textsf{entry\_present entry\_val} }_{\rtv}$
 $\specline{\textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,1}} \textsf{entry entryv} \ast \ghostmaptoken{\textsf{id}}{(\mathsf{entry})}{\textsf{level}} }$
@@
   entry->pfn = PTE_ADDR_TO_PFN((uintptr_t) local);
-  $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)   \ast \texttt{rbp-16} \mapsto_{\textsf{r}} \textsf{\_} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry+KERNBASE} \ast \texttt{rax} \mapsto_{\textsf{r}} \textsf{nextpaddr} \ast \lnot(\textsf{entry\_present entry\_val}) }$
+  $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \lnot(\textsf{entry\_present entry\_val}) }$
 $\specline{ \textsf{entry+KERNBASE} \mapsto_{\mathsf{vpte,1}} \mathsf{entry} \; \textsf{pfn\_set(entryv nextpaddr)}}$
@@
 \begin{lstlisting}[mathescape,escapeinside={(*}{*)}]
+/*
+@param plm4: the root address of the page-table tree
+@param va: virtual address to be translated and mapped
+*/
 pte_t *walkpgdir(pte_t *pml4, const void *va) {
-  /*... ;; Stack setup*/
   $\specline{\textsf{P} \ast \ulcorner\textsf{rtv+KERNBASE}=\textsf{pml4}\urcorner \ast \forall_{i\in\textsf{0..511}}\ghostmaptoken{\textsf{id}}{(\textsf{rtv}+i*8)}{\textsf{4}} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}}_{\rtv}$

### 2025-07-15 examples_C.tex
+

@@
   $\specline{\ulcorner\textsf{entry\_present(entry\_val)}\urcorner \wand \forall_{i\in\textsf{0..511}} \; \ghostmaptoken{\textsf{id}}{\textsf{table\_root(entry\_val.pfn) + i * 8}}{\textsf{v-1}} }_{\rtv} $ (*\label{line:conditional_children}*)
-  (*\label{line:read_entry_contents}*)$\specline{\ldots \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val}  \ast\ldots}$
   /*reading page table entry Justified by virtual pte-pointsto*/
-  if (!entry->present) {/* Jump if the present bit is not zero; no need to allocate next level (* \label{line:check_entry_present} *)(* \label{line:mask_present} *) */
-    $\specline{\ldots \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val \& 0x1} }_{\rtv}$
+  if (!entry->present(*\label{line:read_entry_contents}*)) {/* Jump if the present bit is not zero; no need to allocate next level (* \label{line:check_entry_present} *)(* \label{line:mask_present} *) */
     /*Store the value of rbp minus 16 bytes (address of (*\textsf{next}*)) into rdi (*\label{line:pass_addrof_next}*)(*\label{line:alloc_path_start}*) */
@@
   next = (pte_t *) next_virt_addr;
-  /*... clean up the stack and return the rax value*/
   return next;
@@
   $\specline{ \textsf{pml4\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{pml4\_entry l4e\_val'} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l4e\_val})\urcorner}_{\rtv}$(*\label{line:first_pte_pointsto}*)
-  $\specline{\texttt{rax}  \mapsto_{\textsf{r}} \textsf{table\_root(l4e\_val'.pfn)} }_{\rtv}$
   /*pte_get_next_table may have allocated a new page, updating entry*/
@@
       \ghostmaptoken{\textsf{id}}{(\mathsf{pml4\_entry})}{\textsf{4}}\ast \ghostmaptoken{\textsf{id}}{(\mathsf{plm3\_entry})}{\textsf{3}} \ast \ghostmaptoken{\textsf{id}}{(\mathsf{plm2\_entry})}{\textsf{2}}  \ast \\  \textsf{pml4\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{pml4\_entry} \; \textsf{l4e\_val}  \ast \\  \textsf{pdp\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{pdp\_entry}\;\textsf{l3e\_val}  \ast \\  \textsf{pd\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{pd\_entry} \; \textsf{l2e\_val}  \end{array} \right)  =\textsf{R}_{\textsf{walk} }
-    \\ \ast \left(\begin{array}{l} \textsf{rax} \mapsto \textsf{pt\_entry+KERNBASE} \ast  \textsf{pt\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{pt\_entry}\;\textsf{l1e\_val} \ast \\ \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l1e\_val})\urcorner \end{array}\right) =\textsf{R}_{\textsf{l1e}}\end{array} $
+    \\ \ast \left(\begin{array}{l}  \textsf{pt\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{pt\_entry}\;\textsf{l1e\_val} \ast \\ \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l1e\_val})\urcorner \end{array}\right) =\textsf{R}_{\textsf{l1e}}\end{array} $
 \caption{\textsf{R}$_{\textsf{walk}}$: Resources obtained from invariant for walking page-table directory with \emph{non-existing entries} via calls to \textsf{l1e\_get\_next\_table} in Figure \ref{fig:calltopteinitialize}}

### 2025-07-15 examples_C.tex
+

@@
 \begin{lstlisting}[firstnumber=1, mathescape, language = C, style=mystyle]
+/*
+@param entry: entry whose frame address is used to locate the a-level-lower entry in the page-tree
+*/
 pte_t *pte_get_next_table(pte_t *entry) {
@@
   \begin{lstlisting}[mathescape,escapeinside={(*}{*)}]
+/*
+@param entry: page entry whose frame is initialized with a new physical page 
+*/
 $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \lnot(\textsf{entry\_present entry\_val} }_{\rtv}$
@@
   \begin{lstlisting}[mathescape,escapeinside={(*}{*)}]
+/*
+@param plm4: the root address of the page-table tree
+@param va: virtual address to be translated and mapped
+@param fpaddr: physical page address 
+*/
 void vaspace_mappage(pte_t *pml4, void *va,uintptr_t fpaddr ) {
@@
     /*... ;;set present bit in entry, then clean up the stack*/
-    pteaddr->presetn = 1;
+    pteaddr->present = 1;
   }

### 2025-07-15 examples_C.tex
+

@@
 @param va: virtual address to be translated and mapped
-@param fpaddr: physical page address 
+@param fpaddr: physically allocated page's address in case the virtual address va is not backed with one
 */
 void vaspace_mappage(pte_t *pml4, void *va,uintptr_t fpaddr ) {
-  /*... ;;setting up the stack */ 
   $\specline{\textsf{P} \ast \textsf{pml4} \mapsto_{\textsf{id}} \textsf{\_} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta} \ast \ulcorner\theta \; !!\; \vaddr = \texttt{None}\urcorner}_{\rtv}$

### 2025-07-15 examples_C.tex
+

@@
   pte_t *pml4_entry = &pml4[PML4EX(va)]; // Virtual address of L4 entry(*\label{line:start_pml4_calc}*)
-  /*Load pml4 and va:*/
-  /*Extract 9-bit (0x1ff=511) L4 index from va, multiply by 8 (byte index), add to pml4 for entry address (*\label{line:end_pml4_calc}*) */
-  /*Store virt. addr. to local var pml4_entry; logical pml4_entry is physical!*/
+  /*At the 4th level (and other levels ignoring the level-specific indices) an access amounts to:*/
+  /*1.Load pml4 and va*/
+  /*2.Extract 9-bit (0x1ff=511) L4 index from the virtual address va*/
+  /*3.Multiply by 8 (byte index), add to pml4 for entry address (*\label{line:end_pml4_calc}*) */
+  /*4.Store virt. addr. to local var pml4_entry; logical pml4_entry is physical!*/
   $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{pml4\_entry}\},m) \ast \forall_{i\in\textsf{0..511}\setminus\{\textsf{PML4EX(va)}\}}\ghostmaptoken{\textsf{id}}{(\textsf{rtv}+i*8)}{\textsf{4}}\ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
@@
   $\specline{\textsf{entry\_present(l3e\_val)} \wand \forall_{i\in\textsf{0..511}} \;\ghostmaptoken{\textsf{id}}{\textsf{table\_root(pdp\_entry.pfn) + i * 8}}{ \textsf{2}} } _{\rtv} $
-  /*... ;; Similar assembly & proofs for levels 3 and 2*/
+  /*... Similar assembly & proofs for levels 3 and 2*/
   pte_t *pd = pte_get_next_table(pdp_entry);
@@
     $\specline{ \textsf{pt\_entry+KERNBASE} \mapsto_{\textsf{vpte,1}} \textsf{pt\_entry} \; (\textsf{set\_pfn}(\textsf{l1e\_val}, \textsf{PTE\_ADDR\_TO\_PFN}(\textsf{fpaddr}))) \ast  \ulcorner \lnot(\textsf{entry\_present l1e\_val}) \urcorner \ast \textsf{P}}$
-    /*... ;;set present bit in entry, then clean up the stack*/
+    /*... set present bit in entry, then clean up the stack*/
     pteaddr->present = 1;
@@
   /*
-  ... ;;and return either unaccessible pte pointsto or one with fresh page (fpaddr)
+  ... and return either unaccessible pte pointsto or one with fresh page (fpaddr)
   */

### 2025-07-15 examples_C.tex
+

@@
   $\specline{\ulcorner\textsf{entry\_present(entry\_val)}\urcorner \wand \forall_{i\in\textsf{0..511}} \; \ghostmaptoken{\textsf{id}}{\textsf{table\_root(entry\_val.pfn) + i * 8}}{\textsf{v-1}} }_{\rtv} $ (*\label{line:conditional_children}*)
-  /*reading page table entry Justified by virtual pte-pointsto*/
-  if (!entry->present(*\label{line:read_entry_contents}*)) {/* Jump if the present bit is not zero; no need to allocate next level (* \label{line:check_entry_present} *)(* \label{line:mask_present} *) */
-    /*Store the value of rbp minus 16 bytes (address of (*\textsf{next}*)) into rdi (*\label{line:pass_addrof_next}*)(*\label{line:alloc_path_start}*) */
+  /*Reading page table entry justified by virtual pte-pointsto*/
+  if (!entry->present(*\label{line:read_entry_contents}*)) {
+    /* Jump if the present bit is not zero; no need to allocate next level (* \label{line:check_entry_present} *)(* \label{line:mask_present} *) */
+    /*Allocation path starts here (*\label{line:pass_addrof_next}*)(*\label{line:alloc_path_start}*)*/
     $\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \land \lnot(\textsf{entry\_present entry\_val})  \urcorner }_{\rtv}$ (*\label{line:after_concluding_qfrac1}*)

### 2025-07-15 examples_C.tex
+

@@
 @param entry: entry whose frame address is used to locate the a-level-lower entry in the page-tree
+@return next: return a virtual pointer to the descendant tree
 */
@@
 511} } \ldotp  \ghostmaptoken{\textsf{id}}{((\textsf{table\_root (pte\_initialized (pfn\_set(entry\_val nextpaddr)))}) + \textsf{i * 8})}{\textsf{v-1}} \end{array}  }$ (*\label{line:page_of_caps}*)
-    /*... entry value updates:*/
+    /*Entry value updates:*/
     entry->pfn = nextpaddr; entry->present = 1; (*\label{line:install_new_entry}*)
-    /*... now we know that entry is initialized, so we satisfy the condition to access children list */
+    /*Now we know that entry is initialized, so we satisfy the condition to access children list */
     $\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}}  \textsf{pte\_initialized(pfn\_set(entryv nextpaddr))} \ast \ulcorner  \textsf{qfrac} = 1 \land \lnot(\textsf{entry\_present entry\_val})  \urcorner }_{\rtv}$
@@
   } (*\label{line:end_of_allocation_path}*)
-  /*...  Code after conditional continued in Figure (*\ref{fig:p2v}*)*/
+  /*Physical to virtual address mapping (*\ref{fig:p2v}*)*/
   $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \}),m) \ast \textsf{entry} \mapsto_{\textsf{id}} \textsf{\_} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$

### 2025-07-15 examples_C.tex
+

@@
   $\specline{ \textsf{pml4\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{pml4\_entry l4e\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l4e\_val})\urcorner}_{\rtv}$
-  pte_t *pdp = pte_get_next_table(pml4_entry); /*(*\label{line:first_getnext_call}*) save the physical next table address in rax*/
+  pte_t *pdp = pte_get_next_table(pml4_entry); (*\label{line:first_getnext_call}*) 
   $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{pml4\_entry}\},m)  \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
@@
   $\specline{ \textsf{pml4\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{pml4\_entry l4e\_val'} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l4e\_val})\urcorner}_{\rtv}$(*\label{line:first_pte_pointsto}*)
-  /*pte_get_next_table may have allocated a new page, updating entry*/
+  /*pte_get_next_table may have allocated a new page which updates entry*/
   pte_t *pdp_entry = &pdp[PDPEX(va)]; /* Virtual address of L3 entry*/
@@
   $\specline{\textsf{entry\_present(l3e\_val)} \wand \forall_{i\in\textsf{0..511}} \;\ghostmaptoken{\textsf{id}}{\textsf{table\_root(pdp\_entry.pfn) + i * 8}}{ \textsf{2}} } _{\rtv} $
-  /*... Similar assembly & proofs for levels 3 and 2*/
+  /*Similar steps & proofs for levels 3 and 2*/
   pte_t *pd = pte_get_next_table(pdp_entry);
@@
   pte_t *pt = pte_get_next_table(pd_entry);
-  /* return address of L1 entry, similar index computation*/
+  /*Return the address of L1 entry with a similar index computation*/
   return &pt[PTEX(va)]; 

### 2025-07-15 examples_C.tex
+

@@
 %       (WordImpl.concat (WordImpl.zero 56) (natToWord 8 12 ^& WordImpl.concat (WordImpl.zero 2) WO~1~1~1~1~1~1)) 
-\begin{figure}%\footnotesize
+\begin{figure}\footnotesize
 \ifPLDI
-\begin{lstlisting}[firstnumber=1, mathescape, language = C, style=mystyle]
+\begin{lstlisting}[mathescape,escapeinside={(*}{*)}]
 /*
@@
 void pte_initialize(pte_t *entry) {
-  /*allocate a full zeroed page for 512 8-byte entries(*\label{line:call_to_kalloc}*)*/
+  /*Allocate a full zeroed page for 512 8-byte entries(*\label{line:call_to_kalloc}*)*/
   pte_t *local = kalloc();

### 2025-07-15 examples_C.tex
+

@@
   if (!pteaddr->present){ (*\label{line:mappage_pte_present_start}*)
-    /*... ;; Not present(*\label{line:mappage_pte_present_end}*)*/
+    /*The entry is not present(*\label{line:mappage_pte_present_end}*)*/
     pteaddr->pfn = PTE_ADDR_TO_PFN(fpaddr); /*Store updated entry back to L1 entry (*\label{line:l1entry_store}*)*/
     $\specline{ \textsf{pt\_entry+KERNBASE} \mapsto_{\textsf{vpte,1}} \textsf{pt\_entry} \; (\textsf{set\_pfn}(\textsf{l1e\_val}, \textsf{PTE\_ADDR\_TO\_PFN}(\textsf{fpaddr}))) \ast  \ulcorner \lnot(\textsf{entry\_present l1e\_val}) \urcorner \ast \textsf{P}}$
-    /*... set present bit in entry, then clean up the stack*/
+    /*Set present bit in entry, then clean up the stack*/
     pteaddr->present = 1;
   }
-  /*
-  ... and return either unaccessible pte pointsto or one with fresh page (fpaddr)
-  */
-  return error;
+  return;
   $\specline{ \begin{array}{l}\textsf{pt\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{pt\_entry} \; \textsf{l1e\_val} \ast \ulcorner (\textsf{entry\_present l1e\_val}) \urcorner \ast \textsf{P} \\

### 2025-07-16 background.tex
experiment w/ paragraph giving more detail on mapping a page, in bg section

@@
 a certain range of virtual addresses (corresponding to the kernel's code and data) are valid in every address space.
+
+\paragraph{\add{Mapping a New Page}}
+\add{Mapping a new page into a virtual address space is a key piece fo functionality, and also a key challenge for verification.
+This operation takes a chunk of physical memory to install, and a (page-aligned) virtual address at which to make it
+visible to the process.
+Starting from the page table root (\lstinline|cr3| in Figure \ref{fig:pagetables}), it must traverse the page tables until it locates
+the correct location of the L1 entry which should point to the physical page (installing new L3, L2, and L1 tables as needed),
+and update the L1 entry to point to the physical page. This is complicated by the fact that the kernel accesses the page
+tables through \emph{virtual} addresses, but the page table root and the addresses in the L4--L1 entries are all \emph{physical}
+addresses. So the mapping process must repeatedly perform \emph{backwards} mappings (not supported by hardware) from
+physical addresses of each level's entry to a virtual address that is then mapped back to the corresponding physical address!
+And for modularity, the translation process (and possible allocation of missing tables) is typically implemented in a function
+used for every level of translation (i.e., a level-polymorphic function).
+These compounding subtleties are one reason \textsc{CertiKOS} and \textsc{seL4} \emph{trust} their page table
+manipulations; both the reasoning to prove correctness, and the kernel invariants that ensure correctness of typical implementations,
+are remarkably subtle.
+}
+\looseness=-1
 

### 2025-07-16 background.tex
lots of disorganized edits, mostly focused on justifying use of modalities, some comments for fixing in sec 4

@@
 \rocq terms of a particular type, rather than being drawn from a fixed vocabulary.
-We introduce other details as they arise.
-
+\del{We introduce other details as they arise.}
 \iris includes two forms of implication. The magic wand operator $\wand$ is an affine implication:
@@
 }
+View shifts are essentially logical entailment plus ghost updates.
 \looseness=-1
@@
 
-An important limitation here is that to date, every separation logic
+An important limitation \del{here} is that to date, every separation logic
 has assumed that all pointer addresses are meant for use in a single address space, which avoided the
@@
 is true in some possible future or past world related in a fixed way to the current world, but to talk about \emph{arbitrary}
-other worlds. This explicit naming of alternate worlds increases the power of modal logics~\cite{blackburn1995hybrid}, and is actually
+other worlds.
+\add{This is somewhat different from the accessibility relation modalities more common in verification, but
+also well-established.}
+This explicit naming of alternate worlds increases the power of modal logics \add{over those without it}~\cite{blackburn1995hybrid}, and is actually
 necessary for completeness in classical separation logics~\cite{brotherston2014parametric}.
@@
 space with root $r$ --- thus changing the interpretation of a modality, and even whether a modality is valid --- \emph{while assertions with that modality are active}.
+\add{This is essential to updating page mappings for the current address space in use by a CPU.}
 %

### 2025-07-16 programlogic.tex
lots of disorganized edits, mostly focused on justifying use of modalities, some comments for fixing in sec 4

@@
 some of the most subtle and technically challenging instruction sequences in an OS kernel (Section \ref{sec:experiment}).
+
+\add{Since \textsf{vProp}s are ``just'' \textsc{Iris} predicates on a choice of page table root,
+it may not be obvious what the gain is over simply using \textsc{Iris} directly.
+The gain is not expressive power.
+Every modal logic can be translated straitforwardly to a logic without modalities
+(nearly always a small variation on the \emph{standard translation} from propositional modal logic
+to first-order logic~\cite{blackburn2006modal}\footnote{\add{Itself similar to verification condition generation
+for weakest precondition calculi and Hoare logics.}}) that can directly address the logic's models ---
+but these translations, in addition to making specifications far more verbose,
+generally do not preserve the appealing reasoning principles of modal logics.
+This becomes most apparent when discussing our proof rules.
+}
 
@@
 most unique to our development.
+\add{As a reminder, in systems of Hoare doubles, an instruction's precondition appears in the conclusion of the rule,
+and an instruction's ``postcondition'' appears as the precondition to subsequent instructions in the
+antecedent of the rule.}
 \looseness=-1
@@
     P \ast \mathcal{I}\textsf{ASpace}\ast \mathcal{I}\textsf{ASpace} \ast \textsf{rip} \mapsto_{\textsf{r}} {\textsf{iv} + \textsf{MovLen}(\textsf{r}_d,[\textsf{r}_a])} \ast \arcr
-    r_d \mapsto_{r}  \textsf{v} \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{v,rtv}} \textsf{v}
+    r_d \mapsto_{r}  \textsf{v} \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{v}} \textsf{v}
   \end{array} \right\}_{\textsf{rtv}}\;\overline{is}
@@
 Figure \ref{fig:wpdamd} includes two  rules for accessing memory at an address stored in a register $r_a$. 
-Ultimately, any memory access needs to ensure the relevant address translation would succeed,
+\add{Setting aside $P$, $\mathcal{I}\textsf{ASpace}$, and the instruction pointer \lstinline|rip|,}
+\del{Ultimately, any memory access needs to ensure the relevant address translation would succeed,
 which can be ensured by what we informally call the page-table-walk points-to collection
-($\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$ in Figure \ref{fig:strongvirtualpointsto}).
+($\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$ in Figure \ref{fig:strongvirtualpointsto}). }
 \TirNameStyle{WriteToRegFromVirtMem} and \TirNameStyle{WriteToVirtMemFromReg}
-each use a virtual-pointsto assertion ($\textsf{vaddr} \mapsto_{\textsf{v}} \textsf{v}$),
-and are nearly-standard (assembly) separation logic rules for memory accesses.
-However, because we split the physical resources for the page table walk from the
-virtual points-to itself (in order to permit the physical page tables to be freely modified
-as long as they preserve virtual-to-physical translations), the rule requires $\mathcal{I}\textsf{ASpace}$
+\del{each use a virtual-pointsto assertion ($\textsf{vaddr} \mapsto_{\textsf{v}} \textsf{v}$),
+and }
+are nearly-standard (assembly) separation logic rules for memory accesses~\add{\cite{Chlipala2013Bedrock,ni2007contexts}}.
+\add{For example, \textsc{WriteToRegFromVirtMem}'s specification
+reflects that it reads from the (virtual) memory address \textsf{vaddr} stored in the address
+register $\textsf{r}_a$ --- and thus requires register points-to and virtual points-to assertions describing
+that relationship and the assumed value \textsf{v} in memory in its precondition (the precondition of the rule's
+conclusion). 
+It reflects the load of that memory value into the destination register $\textsf{r}_d$, with
+the updated register points-to in the precondition for $\overline{is}$.
+$P$ describes framed resources, which are passed along to the precondition of subsequent instructions,
+as in any system of Hoare doubles~\cite{Chlipala2013Bedrock,ni2007contexts}.
+\textsc{WriteToVirtMemFromReg} is analogous for writing to memory.
+}
+
+\add{There are only two changes specific to our approach.}
+\replace{However, }{First,}
+because we split the physical resources for the page table walk from the
+virtual points-to itself (\replace{in order to permit the physical page tables to be freely modified
+as long as they preserve virtual-to-physical translations}{per the discussion of Section \ref{sec:sharingpages}}), the rule requires $\mathcal{I}\textsf{ASpace}$
 for the current address space to be carried through.
 The soundness proofs for these rules extract
-% the token ($\sumwalkabs\vaddr\qfrac\paddr$) from the virtual points-to
 the token ($\fracghostmaptoken{\delta}{\vaddr}{\paddr}{\qfrac}$) from the virtual points-to,
@@
 
-The memory access rules model updates to the instruction pointer \lstinline|rip| using a metafunction \textsf{MovLen}.
+\replace{T}{Second, t}he memory access rules --- as with all rules in our logic ---
+model updates to the instruction pointer \lstinline|rip| \add{based on the length of the encoded instruction.
+These rules} \replace{using}{use} a metafunction \textsf{MovLen}.
 \textsf{MovLen} returns how long the instruction encoding for the corresponding \lstinline|mov| is;
@@
 x86 ISA that preceded x86-64.
+
+\add{Note that the use of a modal abstraction of address space simplifies these rules.
+The antecedents of \textsc{WriteToRegFromVirtMem} and \textsc{WriteToVirtMemFromReg}
+only mention the address space in the index of the 
+Hoare double --- not in $P$, or the (virtual) points-to assertions.
+There is no extra condition to discharge that the address being accessed is from
+the current address space.
+\todo[inline,color=green]{WIP: Colin is trying to clarify the value of modal approach here}
+}
+\todo[inline,color=cyan]{Colin says:
+Ismail, Figure 7 looks broken. Why are there two IAspaces in WriteToRegFromVirtMem? Shouldn't there
+only be one?
+Also in WriteToRegFromVirtMem, why can't the virtual points-to be fractional? It is in the Rocq development,
+and we use it as such in examples.
+In WriteToVirtMemFromReg, vaddr points to a value rvs both before and after the store.
+And also in WriteToVirtMemFromReg, why are all the fractions q? They should be allowed to be different fractions.
+(Also, explaining the fraction notation was one of the questions from reviewers.)
+}
 
@@
 \lstinline|mov| to $\textsf{rvs}$ afterwards. The global effects of this rule are reflected in moving
-bare assertions under an other-space modality for $\rtv$, and moving the new address space's assertions out of
-the corresponding modality.
+\del{bare} assertions \add{for the current address space ($P$ and $\mathcal{I}\mathsf{ASpace}$)} under an other-space modality for \add{the initial
+(outgoing) address space} $\rtv$, and moving the new address space's assertions out of
+the corresponding modality \add{(since after the \lstinline|mov|, those will hold in the
+new current address space)}.
 
@@
 
+\todo[inline,color=green]{Colin needs to add more of the ``why modals'' stuff here, per the response.
+But it's a little awkward: in the response we said this was hard to specify with ghost-tagged
+pointers, but now I'm wondering if that's \emph{quite} right.
+It seems like maybe you could just plumb an implicitly-tagged P through to the following
+instructions and be okay, the instantiation of P would just be horrid.
+What we said about e.g., defining
+a generic linked list is still true, and should be raised earlier (not here).
+Here, we want to make a point about packaging all resources for a specific address space
+together. But it would be stronger if we could hone the argument in the response more.
+The bit about ``how would we talk abstractly about the fact that some unspecified resources R should
+be sufficient ... in the new address space'' is the right bit to polish...
+ghost bit discussion should either go in relwork with some Dafny(?) discussion, or earlier in motivation.
+}
+
 \subsection{Soundness}

### 2025-07-16 relwork.tex
finished advisor pass, switched to author-year citations, now 23 pages with a little wiggle room
bus draft
lots of disorganized edits, mostly focused on justifying use of modalities, some comments for fixing in sec 4

@@
 
-We will discuss related work along two axes: work on OS verification specifically, and work on 
+We will discuss two streams of related work: OS verification, and 
 program logics with modalities.
@@
 hardware like swapping~\cite{Denning1970VM} or exploiting copy-on-write techniques for dynamic migration of virtual machines~\cite{clark2005live}.
+\looseness=-1
 
@@
 to verify TLB management in any form --- they \emph{must} trust its operation.
+\looseness=-1
 
@@
 As a result, multicore extensions of \textsc{seL4}~\cite{von2013clustered} and \textsc{CertiKOS}~\cite{gu2016certikos} also trust this functionality.
+\looseness=-1
 
@@
 in addition to inheriting other limitations from the work they extend, discussed above~\cite{kolanski08vstte,kolanski09tphols}.
+\looseness=-1
 
@@
 where modalities are indexed by \emph{nominals} which are names for specific individual states in a Kripke model.
+\add{Readers mostly familiar with modalities in prominent program verification approaches~\cite{Appel2007,birkedal2013intensional,birkedal2011step,hobor2010theory,jung2018iris,nakano2000modality,pnueli1977temporal,pratt1976semantical}
+may not recognize hybrid logics, but as we discuss in Section \ref{sec:relwork}, they (like temporal logics) trace their roots back to Arthur Prior in the 1950s.
+}
 Little prior work combines these ideas with program logics. \citet{brotherston2014parametric} show that traditional
@@
 for virtual memory management the kernel must be able to choose or construct arbitrary other address spaces.
+\looseness=-1
 

### 2025-07-16 syntaxsemantics.tex
lots of disorganized edits, mostly focused on justifying use of modalities, some comments for fixing in sec 4

@@
 % Programs we demonstrate in this paper requires accessing two types of computer resource: registers and memory.
-A register identifier, $\reg$, is chosen from a fixed finite set of register identifiers, $\regset$. 
+A register identifier, $\reg$, is chosen from a fixed finite set of register identifiers, $\regset$.\footnote{\add{$\reg$ abbreviates \emph{general register}, which includes integer registers (e.g., \lstinline|rax|),
+control registers (e.g., \lstinline|cr3|), and segmentation registers (which we do not discuss in detail,
+but are still used in a limited way in 64-bit mode on modern x86-64 processors).
+}} 
 We use these identifiers $\reg$ for register names such as \lstinline|rax|, \lstinline|r8|, or \lstinline|cr3|. Our model includes
@@
 Thus our states $\sigma$ include register maps $\sigma.\mathcal{R}: \kw{greg} \rightarrow_{\textrm{fin}} \kw{regval} $ and
-memory maps $\sigma.\mathcal{M}: \Locft \rightharpoonup_{\textrm{fin}} (\Loctw \rightharpoonup_{\textrm{fin}} \Locsf )$.
+memory maps $\sigma.\mathcal{M}: \Locft \rightharpoonup_{\textrm{fin}} (\Loctw \rightharpoonup_{\textrm{fin}} \Locsf )$
+\add{which segment memory into page-sized increments}.
 Of particular note, \lstinline|cr3|, the page table register, is included in the machine state.
@@
 sequence with an additional instruction (\iseq\instr\instrs).
-We model (and later, give program logic rules for) instructions for basic register moves, and reading and writing memory.
+We model (and later, give program logic rules for) instructions for \del{basic} register \replace{moves}{loads and stores}, and reading and writing memory.
 The latter require page table walks.
@@
 %   \]
-Operationally, each first translates the virtual address stored in register $r_m$ to a physical address,
+Operationally, each first translates the virtual address stored in register $r_m$ to a physical address \add{(a page table walk)},
 then either updates the memory at that physical location with the contents of $r_r$ (for store)
@@
 from specifications in the paper for clarity.
+\looseness=-1

### 2025-07-16 examples_C.tex
first pass on refs

@@
 We build up to the main task of mapping a new page after traversing page tables in software.
-The mapping operation of Figure \ref{fig:mapping_code} assumes an operation \textsf{walkpgdir} which must traverse the page tables
+The mapping operation of Figure \ref{fig:mapping_codeC} assumes an operation \textsf{walkpgdir} which must traverse the page tables
 in order to locate the address of the L1 entry to update --- 
@@
 \subsubsection{Loading Page-Table Address Value}
-We will discuss access to the level 4 table later (Section \ref{wlkpgdir}). But for subsequent levels, the base address of level $n$ must be
+We will discuss access to the level 4 table later (Section \ref{wlkpgdirC}). But for subsequent levels, the base address of level $n$ must be
 fetched from the appropriate entry in the level $n+1$ table.
-This is the role of \lstinline|pte_get_next_table| (Figures \ref{fig:calltopteinitialize} and \ref{fig:p2v}):
+This is the role of \lstinline|pte_get_next_table| (Figures \ref{fig:calltopteinitializeC} and \ref{fig:p2vC}):
 it is passed the virtual address of the page table entry in level $n+1$, and should return the \emph{virtual} 
@@
 the code also allocates a page for the level $n$ table, installs it in the level $n+1$ entry, and establishes appropriate invariants.
-Figure \ref{fig:calltopteinitialize} presents the initial part of the function, which performs the allocation if necessary.
-Figure \ref{fig:p2v} (discussed in Section \ref{sec:p2v}) deals with the cases where no allocation is necessary \emph{or} the allocation has already
+Figure \ref{fig:calltopteinitializeC} presents the initial part of the function, which performs the allocation if necessary.
+Figure \ref{fig:p2vC} (discussed in Section \ref{sec:p2vC}) deals with the cases where no allocation is necessary \emph{or} the allocation has already
 been performed by the code in this figure.
@@
 (a virtual points-to exposing the underlying physical address instead of existentially quantifying it;
- see Section \ref{sec:mapnew}) granting access to the specified entry,
+ see Section \ref{sec:mapnewC}) granting access to the specified entry,
 but its postcondition does not yield new virtual points-to assertions!
-Instead it merely computes the base virtual address of the next table, and returns adequate capabilities (discussed in Section \ref{subsec:identitymappings})
+Instead it merely computes the base virtual address of the next table, and returns adequate capabilities (discussed in Section \ref{subsec:identitymappingsC})
 for the \emph{caller} to construct a vpte-pointsto for any entry of the next table level (if this is not an L1 entry ---
@@
   $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{entry}\},m) \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
-  $\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present entry\_val})\urcorner}_{\rtv}$ (*\label{line:get_next_vpte_precondition}*)
-  $\specline{\ulcorner\textsf{entry\_present(entry\_val)}\urcorner \wand \forall_{i\in\textsf{0..511}} \; \ghostmaptoken{\textsf{id}}{\textsf{table\_root(entry\_val.pfn) + i * 8}}{\textsf{v-1}} }_{\rtv} $ (*\label{line:conditional_children}*)
+  $\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present entry\_val})\urcorner}_{\rtv}$ (*\label{line:get_next_vpte_preconditionC}*)
+  $\specline{\ulcorner\textsf{entry\_present(entry\_val)}\urcorner \wand \forall_{i\in\textsf{0..511}} \; \ghostmaptoken{\textsf{id}}{\textsf{table\_root(entry\_val.pfn) + i * 8}}{\textsf{v-1}} }_{\rtv} $ (*\label{line:conditional_childrenC}*)
   /*Reading page table entry justified by virtual pte-pointsto*/
-  if (!entry->present(*\label{line:read_entry_contents}*)) {
-    /* Jump if the present bit is not zero; no need to allocate next level (* \label{line:check_entry_present} *)(* \label{line:mask_present} *) */
-    /*Allocation path starts here (*\label{line:pass_addrof_next}*)(*\label{line:alloc_path_start}*)*/
-    $\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \land \lnot(\textsf{entry\_present entry\_val})  \urcorner }_{\rtv}$ (*\label{line:after_concluding_qfrac1}*)
-    pte_initialize(entry);(* \label{line:call_to_pte_initialize} *)
+  if (!entry->present(*\label{line:read_entry_contentsC}*)) {
+    /* Jump if the present bit is not zero; no need to allocate next level (* \label{line:check_entry_presentC} *)(* \label{line:mask_presentC} *) */
+    /*Allocation path starts here (*\label{line:pass_addrof_nextC}*)(*\label{line:alloc_path_startC}*)*/
+    $\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \land \lnot(\textsf{entry\_present entry\_val})  \urcorner }_{\rtv}$ (*\label{line:after_concluding_qfrac1C}*)
+    pte_initialize(entry);(* \label{line:call_to_pte_initializeC} *)
     $\specline{ \ldots \ast \ulcorner  \textsf{qfrac} = 1 \land \lnot(\textsf{entry\_present entry\_val})  \urcorner \ast \ulcorner  \textsf{qfrac} = 1 \land \lnot(\textsf{entry\_present (pfn\_set (entry\_val nextpaddr))}  \urcorner   }_{\rtv}$
     $\specline{ \begin{array}{l}\textsf{entry\_present (pte\_initialized (pfn\_set(entryv nextpaddr)))} \wand  \\ \;\;\;\;\;\;\;\;\  \forall_{i\in \textsf{0 ...
-511} } \ldotp  \ghostmaptoken{\textsf{id}}{((\textsf{table\_root (pte\_initialized (pfn\_set(entry\_val nextpaddr)))}) + \textsf{i * 8})}{\textsf{v-1}} \end{array}  }$ (*\label{line:page_of_caps}*)
+511} } \ldotp  \ghostmaptoken{\textsf{id}}{((\textsf{table\_root (pte\_initialized (pfn\_set(entry\_val nextpaddr)))}) + \textsf{i * 8})}{\textsf{v-1}} \end{array}  }$ (*\label{line:page_of_capsC}*)
     /*Entry value updates:*/
-    entry->pfn = nextpaddr; entry->present = 1; (*\label{line:install_new_entry}*)
+    entry->pfn = nextpaddr; entry->present = 1; (*\label{line:install_new_entryC}*)
     /*Now we know that entry is initialized, so we satisfy the condition to access children list */
@@
     $\specline{\forall_{i\in \textsf{0 ...
-511} } \ldotp  \ghostmaptoken{\textsf{id}}{((\textsf{table\_root (pte\_initialized (pfn\_set(entry\_val nextpaddr)))}) + \textsf{i * 8})}{\textsf{v-1}}  }$ (*\label{line:alloc_path_end}*)
-  } (*\label{line:end_of_allocation_path}*)
-  /*Physical to virtual address mapping (*\ref{fig:p2v}*)*/
+511} } \ldotp  \ghostmaptoken{\textsf{id}}{((\textsf{table\_root (pte\_initialized (pfn\_set(entry\_val nextpaddr)))}) + \textsf{i * 8})}{\textsf{v-1}}  }$ (*\label{line:alloc_path_endC}*)
+  } (*\label{line:end_of_allocation_pathC}*)
+  /*Physical to virtual address mapping code segment starts here (*\ref{fig:p2vC}*)*/
   $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \}),m) \ast \textsf{entry} \mapsto_{\textsf{id}} \textsf{\_} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
@@
   $\specline{\forall_{i\in \textsf{0 ...
-511} } \ldotp  \ghostmaptoken{\textsf{id}}{((\textsf{table\_root (pte\_initialized (entry\_val.pfn)))}) + \textsf{i * 8})}{\textsf{v-1}}  }$ (*\label{line:children}*)
+511} } \ldotp  \ghostmaptoken{\textsf{id}}{((\textsf{table\_root (pte\_initialized (entry\_val.pfn)))}) + \textsf{i * 8})}{\textsf{v-1}}  }$ (*\label{line:childrenC}*)
   uintptr_t next_phys_addr = PTE_PFN_TO_ADDR(entry->pfn)
-  uintptr_t next_virt_addr = (uintptr_t) P2V(next_phys_addr); (*\label{line:p2v}*);
+  uintptr_t next_virt_addr = (uintptr_t) P2V(next_phys_addr); (*\label{line:p2vC}*);
   next = (pte_t *) next_virt_addr;
@@
 \caption{Ensuring \textsf{entry} points to a valid next table, allocating if necessary.}
-\label{fig:calltopteinitialize}
+\label{fig:calltopteinitializeC}
 \vspace{-1em}
@@
 \subsubsection{Identity Mappings}
-\label{subsec:identitymappings}
+\label{subsec:identitymappingsC}
 Kernels need to convert between physical and virtual addresses, in both directions.
@@
 
-For this reason we extend the per-address-space invariant as in Figure \ref{fig:peraspaceinvariant_with_p2v_extension}, to also track which
+For this reason we extend the per-address-space invariant as in Figure \ref{fig:peraspaceinvariant_with_p2v_extensionC}, to also track which
 addresses we can perform a P2V conversion on by a adding a constant offset.
@@
 \emph{Only} physical addresses in $\Xi$ can undergo P2V conversion. 
-Section \ref{sec:p2v} describes the actual conversion,
+Section \ref{sec:p2vC} describes the actual conversion,
 but we describe the invariant here 
@@
 \vspace{-1em}
-\caption{Global Address-Space Invariant in Figure \ref{fig:peraspaceinvariant} extended with a ghost map bookkeeping identity mappings }
-  \label{fig:peraspaceinvariant_with_p2v_extension}
+\caption{Global Address-Space Invariant in Figure \ref{fig:peraspaceinvariantC} extended with a ghost map bookkeeping identity mappings }
+  \label{fig:peraspaceinvariant_with_p2v_extensionC}
 \vspace{-1em}
@@
 For each $\paddr\mapsto \textsf{v} \in\Xi$, the invariant tracks a virtual points-to justifying that virtual address $\paddr+\textsf{KERNBASE}$ maps to physical address $\paddr$
-(the ``Ghost translation'' in Figure \ref{fig:peraspaceinvariant_with_p2v_extension});
+(the ``Ghost translation'' in Figure \ref{fig:peraspaceinvariant_with_p2v_extensionC});
 fractional ownership of the physical memory for that page table entry;
@@
 % (L1 entries point to data pages, whose physical memory ownership resides in some virtual points-to).
-The assertion on Line \ref{line:conditional_children} of Figure \ref{fig:calltopteinitialize} comes from the invariant one level up; 
+The assertion on Line \ref{line:conditional_childrenC} of Figure \ref{fig:calltopteinitializeC} comes from the invariant one level up; 
 if the valid bit is set,
@@
 So in general the invariant cannot keep full permission to the memory in this part of the invariant, or it would overlap the page table walk for virtual points-to
-assertions. But in the case where the entry is invalid, we may need to write to it (e.g., to install a reference to a next-level table, as we do in Figure \ref{fig:calltopteinitialize}),
+assertions. But in the case where the entry is invalid, we may need to write to it (e.g., to install a reference to a next-level table, as we do in Figure \ref{fig:calltopteinitializeC}),
 which requires full permission. Fortunately, the entry can only be in use if its valid bit is set; if the valid bit is not set we know
 that no virtual points-to entry in $\delta$/$\theta$ holds any partial ownership.
-Thus we use the invariant portion annotated as ``Entry validity'' in Figure \ref{fig:peraspaceinvariant_with_p2v_extension} to capture this:
+Thus we use the invariant portion annotated as ``Entry validity'' in Figure \ref{fig:peraspaceinvariant_with_p2v_extensionC} to capture this:
 if the entry is invalid the invariant holds full ownership of the entry, so it can be updated; while if the entry is valid,
@@
 \end{equation*}
-Thus the fractional ownership of the physical location is enough for Line \ref{line:read_entry_contents} in Figure \ref{fig:calltopteinitialize} to access the entry, though in \lstinline|get_next_table|
+Thus the fractional ownership of the physical location is enough for Line \ref{line:read_entry_contentsC} in Figure \ref{fig:calltopteinitializeC} to access the entry, though in \lstinline|get_next_table|
 the caller has pulled that piece of information out of the invariant and passed it for the entry at hand.
@@
 for a specific page table entry out of the invariant, and later put them back).
-Line \ref{line:check_entry_present_jump}'s conditional then determines in the fall-through case that the bit is not set, which 
-together with other facts entails $\textsf{qfrac} = 1$ at Line \ref{line:after_concluding_qfrac1},
-and permits storing a new entry (in ellided code around Line \ref{line:install_new_entry}).
+Line \ref{line:check_entry_present_jumpC}'s conditional then determines in the fall-through case that the bit is not set, which 
+together with other facts entails $\textsf{qfrac} = 1$ at Line \ref{line:after_concluding_qfrac1C},
+and permits storing a new entry (in ellided code around Line \ref{line:install_new_entryC}).
 \looseness=-1
@@
  \vspace{-1em}
-\caption{Peel-out the resources needed for the virtual-pte pointsto relation for physical address \textsf{entry} to be accessed in Line 7 in Figure \ref{fig:calltopteinitialize}. }
-  \label{fig:lookingupidpampping}
+\caption{Peel-out the resources needed for the virtual-pte pointsto relation for physical address \textsf{entry} to be accessed in Line 7 in Figure \ref{fig:calltopteinitializeC}. }
+  \label{fig:lookingupidpamppingC}
  \end{figure*}
@@
 \fi
-in Figure \ref{fig:calltopteinitialize}.
+in Figure \ref{fig:calltopteinitializeC}.
 Accessing the presence bit and checking the value allows us to exploit the condition (*) that was just discussed when verifying the allocation
-path (i.e., when the entry is invalid  and Lines \ref{line:alloc_path_start}--\ref{line:alloc_path_end} in Figure \ref{fig:calltopteinitialize}
+path (i.e., when the entry is invalid  and Lines \ref{line:alloc_path_startC}--\ref{line:alloc_path_endC} in Figure \ref{fig:calltopteinitializeC}
 must allocate the next level of tables).
@@
 void pte_initialize(pte_t *entry) {
-  /*Allocate a full zeroed page for 512 8-byte entries(*\label{line:call_to_kalloc}*)*/
+  /*Allocate a full zeroed page for 512 8-byte entries(*\label{line:call_to_kallocC}*)*/
   pte_t *local = kalloc();
@@
 \caption{Allocating a physical page }
-\label{pteinitializespec}
+\label{pteinitializespecC}
 \vspace{-1em}
@@
 
-If the entry is not set, \textsf{pte\_initialize} (Line \ref{line:call_to_pte_initialize} in Figure \ref{fig:calltopteinitialize}) 
+If the entry is not set, \textsf{pte\_initialize} (Line \ref{line:call_to_pte_initializeC} in Figure \ref{fig:calltopteinitializeC}) 
 allocates a physical page (internally utilizing the only unverified (trusted) code in our case studies, the page-allocator's \textsf{kalloc},\footnote{
@@
 } 
-on Line \ref{line:call_to_kalloc} in Figure \ref{pteinitializespec}). 
+on Line \ref{line:call_to_kallocC} in Figure \ref{pteinitializespecC}). 
 Since we are using \textsf{pte\_initialize} for page-table address allocation, we must relate this newly
 allocated physical address to the identity mapping map $\Xi$ --- 
-see Line \ref{line:page_of_caps} in Figure \ref{fig:calltopteinitialize}, where
+see Line \ref{line:page_of_capsC} in Figure \ref{fig:calltopteinitializeC}, where
 \texttt{kalloc}'s specification guarantees it has returned memory from a designated memory
@@
 of an entry construction that has not been completed yet: the presence bit is set 
-(Line \ref{line:install_new_entry} in Figure \ref{fig:calltopteinitialize}) after these freshly allocated resources are incorporated to the 
+(Line \ref{line:install_new_entryC} in Figure \ref{fig:calltopteinitializeC}) after these freshly allocated resources are incorporated to the 
 entry construction via the page-frame portion of the PTE. In other words, the side condition, (*),
@@
 \subsubsection{Physical-to-Virtual Conversion with \textsf{P2V}}
-\label{sec:p2v}
+\label{sec:p2vC}
 Once we know the entry refers to a physical address in the identity mapping range ($\Xi$)
-(via the branch at Line \ref{line:check_entry_present_jump}, or  by allocating and installing a new entry
-as just discussed for Lines \ref{line:check_entry_present_jump}--\ref{line:end_of_allocation_path}), 
+(via the branch at Line \ref{line:check_entry_present_jumpC}, or  by allocating and installing a new entry
+as just discussed for Lines \ref{line:check_entry_present_jumpC}--\ref{line:end_of_allocation_pathC}), 
 we can convert this frame address to a corresponding virtual address via the identity mappings
-discussed in Section \ref{subsec:identitymappings} and Figure \ref{fig:peraspaceinvariant_with_p2v_extension}.
-in the last lines of \lstinline|pte_get_next_table| shown in Figure \ref{fig:p2v} (the continuation of Figure \ref{fig:calltopteinitialize}).
+discussed in Section \ref{subsec:identitymappingsC} and Figure \ref{fig:peraspaceinvariant_with_p2v_extensionC}.
+in the last lines of \lstinline|pte_get_next_table| shown in Figure \ref{fig:p2v} (the continuation of Figure \ref{fig:calltopteinitializeC}).
 This is a critical piece of the full page table walk verification.
-In our small kernel (Line \ref{line:p2v} in Figure \ref{fig:p2v}), as in larger kernels, the C macro \texttt{P2V} common to many kernels
-is actually just addition by the constant offset mentioned in Section \ref{subsec:identitymappings}.
+In our small kernel (Line \ref{line:p2vC} in Figure \ref{fig:p2vC}), as in larger kernels, the C macro \texttt{P2V} common to many kernels
+is actually just addition by the constant offset mentioned in Section \ref{subsec:identitymappingsC}.
 But the correctness of this simple instruction is quite subtle.
@@
 %\end{figure}
-Figure \ref{fig:p2v} shows the verification of the end of \lstinline|pte_get_next_table| specialized to the case where 
-where no allocation was necessary (i.e., the conditional on Line \ref{line:check_entry_present} of Figure \ref{fig:calltopteinitialize} was taken).
-In this case, the true present bit allows access to the child tokens from Line \ref{line:conditional_children} of Figure \ref{fig:calltopteinitialize},
-which is then refined to the assertion on Line \ref{line:children} of Figure \ref{fig:p2v}.
+Figure \ref{fig:p2vC} shows the verification of the end of \lstinline|pte_get_next_table| specialized to the case where 
+where no allocation was necessary (i.e., the conditional on Line \ref{line:check_entry_presentC} of Figure \ref{fig:calltopteinitializeC} was taken).
+In this case, the true present bit allows access to the child tokens from Line \ref{line:conditional_childrenC} of Figure \ref{fig:calltopteinitializeC},
+which is then refined to the assertion on Line \ref{line:childrenC} of Figure \ref{fig:p2vC}.
 The code loads \lstinline|rcx| with the offset value \textsf{KERNBASE}, which gives us the value of the virtual address ($\textsf{entry}_{\textsf{pfn}}$ \textsf{+KERNBASE})
@@
 \subsubsection{Walking Page-Table Tree: Calling \textsf{pte\_get\_next\_table} for Each Level}
-\label{wlkpgdir}
+\label{wlkpgdirC}
 \begin{figure}\footnotesize
@@
   $\specline{\textsf{P} \ast \ulcorner\textsf{rtv+KERNBASE}=\textsf{pml4}\urcorner \ast \forall_{i\in\textsf{0..511}}\ghostmaptoken{\textsf{id}}{(\textsf{rtv}+i*8)}{\textsf{4}} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}}_{\rtv}$
-  pte_t *pml4_entry = &pml4[PML4EX(va)]; // Virtual address of L4 entry(*\label{line:start_pml4_calc}*)
+  pte_t *pml4_entry = &pml4[PML4EX(va)]; // Virtual address of L4 entry(*\label{line:start_pml4_calcC}*)
   /*At the 4th level (and other levels ignoring the level-specific indices) an access amounts to:*/
@@
   /*2.Extract 9-bit (0x1ff=511) L4 index from the virtual address va*/
-  /*3.Multiply by 8 (byte index), add to pml4 for entry address (*\label{line:end_pml4_calc}*) */
+  /*3.Multiply by 8 (byte index), add to pml4 for entry address (*\label{line:end_pml4_calcC}*) */
   /*4.Store virt. addr. to local var pml4_entry; logical pml4_entry is physical!*/
@@
   $\specline{ \textsf{pml4\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{pml4\_entry l4e\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l4e\_val})\urcorner}_{\rtv}$
-  pte_t *pdp = pte_get_next_table(pml4_entry); (*\label{line:first_getnext_call}*) 
+  pte_t *pdp = pte_get_next_table(pml4_entry); (*\label{line:first_getnext_callC}*) 
   $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{pml4\_entry}\},m)  \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
   $\specline{\forall_{i\in\textsf{0..511}} \ghostmaptoken{\textsf{id}}{\textsf{table\_root(l4e\_val'.pfn) + i * 8}}{\textsf{3}}  \ast \ulcorner\textsf{pdp-KERNBASE}=\textsf{table\_root(l4e\_val'.pfn)}\urcorner} $
-  $\specline{ \textsf{pml4\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{pml4\_entry l4e\_val'} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l4e\_val})\urcorner}_{\rtv}$(*\label{line:first_pte_pointsto}*)
+  $\specline{ \textsf{pml4\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{pml4\_entry l4e\_val'} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l4e\_val})\urcorner}_{\rtv}$(*\label{line:first_pte_pointstoC}*)
   /*pte_get_next_table may have allocated a new page which updates entry*/
@@
   $\specline{\forall_{i\in\textsf{0..511}\setminus\{\textsf{PDPEX(va)}\}} \; \textsf{table\_root(pml4\_entry}_{\textsf{pfn}}) \textsf{ + i * 8} \mapsto_{\textsf{id}} \textsf{3} }_{\rtv}$
-  $\specline{ \textsf{pml4\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac4}}\textsf{pml4\_entry}\; \textsf{l4e\_val'} \ast \ulcorner  \textsf{qfrac4} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l4e\_val'})\urcorner}_{\rtv}$(*\label{line:ex_l4_vpte}*)
-  $\specline{ \textsf{pdp\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac3}}\textsf{pdp\_entry}\; \textsf{l3e\_val} \ast \ulcorner  \textsf{qfrac3} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l3e\_val})\urcorner}_{\rtv}$(*\label{line:ex_l3_vpte}*)
+  $\specline{ \textsf{pml4\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac4}}\textsf{pml4\_entry}\; \textsf{l4e\_val'} \ast \ulcorner  \textsf{qfrac4} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l4e\_val'})\urcorner}_{\rtv}$(*\label{line:ex_l4_vpteC}*)
+  $\specline{ \textsf{pdp\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac3}}\textsf{pdp\_entry}\; \textsf{l3e\_val} \ast \ulcorner  \textsf{qfrac3} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l3e\_val})\urcorner}_{\rtv}$(*\label{line:ex_l3_vpteC}*)
   $\specline{\textsf{entry\_present(l3e\_val)} \wand \forall_{i\in\textsf{0..511}} \;\ghostmaptoken{\textsf{id}}{\textsf{table\_root(pdp\_entry.pfn) + i * 8}}{ \textsf{2}} } _{\rtv} $
@@
 \vspace{-1em}
-\caption{Walking page-table directory via calls to \textsf{pte\_get\_next\_table} in Figure \ref{fig:calltopteinitialize}}
-\label{walkpgdir}
+\caption{Walking page-table directory via calls to \textsf{pte\_get\_next\_table} in Figure \ref{fig:calltopteinitializeC}}
+\label{walkpgdirC}
 \vspace{-1em}
@@
     \\ \ast \left(\begin{array}{l}  \textsf{pt\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{pt\_entry}\;\textsf{l1e\_val} \ast \\ \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l1e\_val})\urcorner \end{array}\right) =\textsf{R}_{\textsf{l1e}}\end{array} $
-\caption{\textsf{R}$_{\textsf{walk}}$: Resources obtained from invariant for walking page-table directory with \emph{non-existing entries} via calls to \textsf{l1e\_get\_next\_table} in Figure \ref{fig:calltopteinitialize}}
-\label{fig:rwalk}
+\caption{\textsf{R}$_{\textsf{walk}}$: Resources obtained from invariant for walking page-table directory with \emph{non-existing entries} via calls to \textsf{l1e\_get\_next\_table} in Figure \ref{fig:calltopteinitializeC}}
+\label{fig:rwalkC}
 \vspace{-1em}
 \end{figure}
-Implementing a software page-table walk amounts to calling \textsf{pte\_get\_next\_table} for each level as shown in Figure \ref{walkpgdir}. 
+Implementing a software page-table walk amounts to calling \textsf{pte\_get\_next\_table} for each level as shown in Figure \ref{walkpgdirC}. 
 The key part of the specification and proof for a page table walk is accumulation of memory mappings for the page-table entries 
 visited and frame addresses for page-tables. 
-For example, Lines \ref{line:ex_l4_vpte} and \ref{line:ex_l3_vpte} in Figure \ref{walkpgdir} show the virtual pte-pointsto assertions for L4 and L3 entries.
+For example, Lines \ref{line:ex_l4_vpteC} and \ref{line:ex_l3_vpteC} in Figure \ref{walkpgdirC} show the virtual pte-pointsto assertions for L4 and L3 entries.
 In the final post-condition, we expect the accumulation of these resources from each level -- $\textsf{R}_{\textsf{walk}}$ -- 
@@
 The precondition provides knowledge that the virtual base of the L4 is at the appropriate offset from the current \lstinline|cr3| value,
-but does not provide a virtual points-to assertion --- because the function must calculate (Lines \ref{line:start_pml4_calc}--\ref{line:end_pml4_calc})
+but does not provide a virtual points-to assertion --- because the function must calculate (Lines \ref{line:start_pml4_calcC}--\ref{line:end_pml4_calcC})
 which entry it needs access to.
 Instead the precondition has 512 identity map tokens, guaranteeing that every entry on the page is subject to the identity mapping invariant.
-Line \ref{line:end_pml4_calc} calculates the virtual address of the relevant entry, and the subsequent view shift
+Line \ref{line:end_pml4_calcC} calculates the virtual address of the relevant entry, and the subsequent view shift
 pulls that entry out of the identity mapping ($\Xi$) and fetches its corresponding resources as
-described by Figure \ref{fig:peraspaceinvariant_with_p2v_extension} and Section \ref{subsec:identitymappings}.
+described by Figure \ref{fig:peraspaceinvariant_with_p2v_extensionC} and Section \ref{subsec:identitymappingsC}.
 The ghost translation and physical location are used to form the virtual pte-pointsto for the L4 entry
-(Line \ref{line:first_pte_pointsto}), with the entry validity and next-level indexing
+(Line \ref{line:first_pte_pointstoC}), with the entry validity and next-level indexing
 satisfying the rest of the precondition for \lstinline|pte_get_next_table|.
@@
 allocates into the entry and returns new (also unconditional) tokens for the L3 entry physical addresses.
-\lstinline|pte_get_next_table|'s first call (Line \ref{line:first_getnext_call}) returns
+\lstinline|pte_get_next_table|'s first call (Line \ref{line:first_getnext_callC}) returns
 the virtual address of the base of the L3 table (a \emph{page directory pointer}, so PDP, in official
@@
 \subsection{Mapping a New Page}
-\label{sec:mapnew}
+\label{sec:mapnewC}
 One of the key tasks of a page fault handler in a general-purpose OS kernel is
@@
 \centerline{\textsf{vaspace\_mappage(pte\_t *pml4, void *va,uintptr\_t fpaddr)}}\\
-in Figure \ref{fig:mapping_code}.
+in Figure \ref{fig:mapping_codeC}.
 To do so, with a given allocated a fresh page (\textsf{fpaddr}), then calculate the appropriate
-known-valid page table walks (via \textsf{walkpgdir} Line \ref{line:call_walkpgdir} in Figure \ref{fig:mapping_code})  and update 
-the appropriate L1 page table entry (Line 35 in Figure \ref{fig:mapping_code});
+known-valid page table walks (via \textsf{walkpgdir} Line \ref{line:call_walkpgdirC} in Figure \ref{fig:mapping_codeC})  and update 
+the appropriate L1 page table entry (Line 35 in Figure \ref{fig:mapping_codeC});
 unmapping is the reverse of the logic we discuss here.
@@
   $\specline{\textsf{P} \ast \textsf{pml4} \mapsto_{\textsf{id}} \textsf{\_} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta} \ast \ulcorner\theta \; !!\; \vaddr = \texttt{None}\urcorner}_{\rtv}$
-  pte_t *pteaddr = walkpgdir(pml4, a); (*\label{line:call_walkpgdir}*)
+  pte_t *pteaddr = walkpgdir(pml4, a); (*\label{line:call_walkpgdirC}*)
   $\specline{ \textsf{R}_{\textsf{walk}} \ast \textsf{R}_{\textsf{l1e}} }$
-  if (!pteaddr->present){ (*\label{line:mappage_pte_present_start}*)
-    /*The entry is not present(*\label{line:mappage_pte_present_end}*)*/
-    pteaddr->pfn = PTE_ADDR_TO_PFN(fpaddr); /*Store updated entry back to L1 entry (*\label{line:l1entry_store}*)*/
+  if (!pteaddr->present){ (*\label{line:mappage_pte_present_startC}*)
+    /*The entry is not present(*\label{line:mappage_pte_present_endC}*)*/
+    pteaddr->pfn = PTE_ADDR_TO_PFN(fpaddr); /*Store updated entry back to L1 entry (*\label{line:l1entry_storeC}*)*/
     $\specline{ \textsf{pt\_entry+KERNBASE} \mapsto_{\textsf{vpte,1}} \textsf{pt\_entry} \; (\textsf{set\_pfn}(\textsf{l1e\_val}, \textsf{PTE\_ADDR\_TO\_PFN}(\textsf{fpaddr}))) \ast  \ulcorner \lnot(\textsf{entry\_present l1e\_val}) \urcorner \ast \textsf{P}}$
-    /*Set present bit in entry, then clean up the stack*/
+    /*Set the present bit in entry*/
     pteaddr->present = 1;
@@
   \caption{Specification of updating L1 entry (\textsf{pt\_entry}) to reference a new page (\textsf{fpaddr}).}
-\label{fig:mapping_code}
+\label{fig:mapping_codeC}
 \end{figure}
 
-In Figure \ref{fig:mapping_code}, we see an address ($\vaddr$) currently not
+In Figure \ref{fig:mapping_codeC}, we see an address ($\vaddr$) currently not
 mapped to a page ($\theta \; !!\; \vaddr = \texttt{None}$). Mapping a fresh
@@
 the existence of a memory location for an appropriate L1 table entry.
-The code uses a helper function \lstinline{walkpgdir} (discussed again in Section \ref{sec:traversing}).
+The code uses a helper function \lstinline{walkpgdir} (discussed again in Section \ref{sec:traversingC}).
 \textsf{walkpgdir}'s postcondition contains virtual \emph{PTE} pointsto assertions ($\mapsto_{\textsf{vpte}}$)
 both for ensuring partial page table walk reaching the
-L1 entry (l1e) by asserting that higher levels of the page table exist (R$_{\textsf{walk}}$ in Figure \ref{fig:rwalk}), 
-and for allowing access to the memory of the L1 entry via virtual address (R$_{\textsf{l1e}}$ in Figure \ref{fig:rwalk}).
+L1 entry (l1e) by asserting that higher levels of the page table exist (R$_{\textsf{walk}}$ in Figure \ref{fig:rwalkC}), 
+and for allowing access to the memory of the L1 entry via virtual address (R$_{\textsf{l1e}}$ in Figure \ref{fig:rwalkC}).
 
-A PTE points-to, as we already brifly mentioned, is defined just like the normal virtual points-to of Figure \ref{fig:virtualpointstosharing}, except the physical address (here, \textsf{pa}) is explicit in the assertion
+A PTE points-to, as we already brifly mentioned, is defined just like the normal virtual points-to of Figure \ref{fig:virtualpointstosharingC}, except the physical address (here, \textsf{pa}) is explicit in the assertion
 rather than existentially quantified:\\
@@
 checks have already been performed, the specification in R$_{\textsf{l1e}}$ ensures that page table entry for L1 needs to be checked at the caller site. 
-By doing so, as we see in Figure \ref{fig:mapping_code}, the page reference \textsf{fpaddr} is linked to back the virtual address \textsf{va} 
-only if it is not already referring to a physical resource (Lines \ref{line:mappage_pte_present_start}--\ref{line:mappage_pte_present_end} in Figure \ref{fig:mapping_code}).
+By doing so, as we see in Figure \ref{fig:mapping_codeC}, the page reference \textsf{fpaddr} is linked to back the virtual address \textsf{va} 
+only if it is not already referring to a physical resource (Lines \ref{line:mappage_pte_present_startC}--\ref{line:mappage_pte_present_endC} in Figure \ref{fig:mapping_codeC}).
 
-The crucial step in addition to traversing the page table in Figure \ref{walkpgdir} is actually updating the L1 entry (Line \ref{line:updatepfn} in Figure \ref{fig:mapping_code}),
+The crucial step in addition to traversing the page table in Figure \ref{walkpgdirC} is actually updating the L1 entry (Line \ref{line:updatepfnC} in Figure \ref{fig:mapping_codeC}),
 via the virtual address (\textsf{pt\_entry+KERNBASE}) known to translate to the appropriate physical address, in our example the L1
@@
 $\specline{\mathsf{ContextAt}(old,\_) \ast \mathsf{ContextAt}(new,[\mathsf{rbxv}',\ldots,\rtv']) }_{\rtv}$
-mov 0[rdi], rbx ... ;; also save rsp, rbp, r12, r13, r14, to offsets 8, 16, 24, 32, and 40 from rdi (*\label{line:start_save}*)
+mov 0[rdi], rbx ... ;; also save rsp, rbp, r12, r13, r14, to offsets 8, 16, 24, 32, and 40 from rdi (*\label{line:start_saveC}*)
 mov 48[rdi], r15
@@
 mov 56[%rdi], %cr3 $\specline{\ldots \ast \texttt{rdi+56} \mapsto_{\textsf{v}} \rtv \ast \ldots }_{\rtv}$   (*\label{line:end_save}*) 
-mov rbx, 0[rsi] (*\label{line:start_restore}*)
+mov rbx, 0[rsi] (*\label{line:start_restoreC}*)
 mov rsp, 8[rsi] ;; Switch to new stack, which may not be mapped in the current address space!
@@
 $\specline{\mathsf{ContextAt}(old,[\mathsf{rbxv,\ldots,\rtv}]) \ast \mathsf{ContextAt}(new,[\mathsf{rbxv}',\ldots,\rtv']) }_{\rtv}$
-mov cr3, 56[rsi] ;; <-- Switch to the new address space (*\label{line:end_restore}*)
-$\specline{ [\rtv](\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,\Xi,m) \ast \mathsf{ContextAt}(old,[\mathsf{rbxv,\ldots,\rtv}]) \ast \mathsf{ContextAt}(new,[\mathsf{rbxv}',\ldots,\rtv']))  }_{\rtv}$ (*\label{line:modality_switch}*)
+mov cr3, 56[rsi] ;; <-- Switch to the new address space (*\label{line:end_restoreC}*)
+$\specline{ [\rtv](\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,\Xi,m) \ast \mathsf{ContextAt}(old,[\mathsf{rbxv,\ldots,\rtv}]) \ast \mathsf{ContextAt}(new,[\mathsf{rbxv}',\ldots,\rtv']))  }_{\rtv}$ (*\label{line:modality_switchC}*)
 $\specline{ \mathcal{I}\texttt{ASpace}(\theta',\Xi',m') \ast \texttt{Pother} \ast  \texttt{rsi}\mapsto_{\textsf{r}} new \ast \texttt{rdi}\mapsto_{\textsf{r}} old \ast \ldots }_{\rtv'}$
@@
 \caption{Basic task switch code that switches address spaces.}
-\label{fig:swtch}
+\label{fig:swtchC}
 \end{figure}
@@
 A critical piece of \emph{trusted} code in current verified OS kernels is the assembly code to change the current address space; current verified OS kernels currently 
-lack effective ways to specify and reason about this low-level operation, for reasons outlined in Section \ref{sec:relwork}.
+lack effective ways to specify and reason about this low-level operation, for reasons outlined in Section \ref{sec:relworkC}.
 
-Figure \ref{fig:swtch} gives simplified code for a basic task switch, the heart of an OS scheduler implementation. This is code that saves the context (registers and stack)
-of the running thread (here in a structure pointed to by \lstinline|rdi|'s value shown in Lines \ref{line:start_save}--\ref{line:end_save} of Figure \ref{fig:swtch}) and restores the context of 
-an existing thread (from \lstinline|rsi| shown in abbreviated Lines \ref{line:start_restore}--\ref{line:end_restore}), including the corresponding change of address space for a target thread in another process.
+Figure \ref{fig:swtchC} gives simplified code for a basic task switch, the heart of an OS scheduler implementation. This is code that saves the context (registers and stack)
+of the running thread (here in a structure pointed to by \lstinline|rdi|'s value shown in Lines \ref{line:start_saveC}--\ref{line:end_saveC} of Figure \ref{fig:swtchC}) and restores the context of 
+an existing thread (from \lstinline|rsi| shown in abbreviated Lines \ref{line:start_restoreC}--\ref{line:end_restoreC}), including the corresponding change of address space for a target thread in another process.
 This code assumes the System V AMD64 ABI calling convention, where the normal registers not mentioned are caller-save, and therefore saved on the stack of the thread
@@
 \centerline{\lstinline[language=C]|void swtch(context_t* save, context_t* restore);|}\\
-A call to this code begins executing one thread (until just before Line \ref{line:end_save}) in one address space ($\rtv$), whose information will be saved in a structure at address $old$,
-and finishes execution executing a different thread in a different address space (Line \ref{line:end_restore} on) whose information is initially in $new$.
+A call to this code begins executing one thread (until just before Line \ref{line:end_saveC}) in one address space ($\rtv$), whose information will be saved in a structure at address $old$,
+and finishes execution executing a different thread in a different address space (Line \ref{line:end_restoreC} on) whose information is initially in $new$.
 

### 2025-07-16 examples_C.tex
+

@@
 fetched from the appropriate entry in the level $n+1$ table.
-This is the role of \lstinline|pte_get_next_table| (Figures \ref{fig:calltopteinitializeC} and \ref{fig:p2vC}):
+This is the role of \lstinline|pte_get_next_table| (Figures \ref{fig:calltopteinitializeC}):
+%and \ref{fig:p2vC}):
 it is passed the virtual address of the page table entry in level $n+1$, and should return the \emph{virtual} 
@@
 Figure \ref{fig:calltopteinitializeC} presents the initial part of the function, which performs the allocation if necessary.
-Figure \ref{fig:p2vC} (discussed in Section \ref{sec:p2vC}) deals with the cases where no allocation is necessary \emph{or} the allocation has already
+Lines \ref{line:finalpieceS}-\ref{line:finalpieceE} in Figure \ref{fig:calltopteinitializeC} (discussed in Section \ref{sec:p2vC}) deals with the cases where no allocation is necessary \emph{or} the allocation has already
 been performed by the code in this figure.
@@
 \else
-Lines \ref{line:mask_present}--\ref{line:check_entry_present} check
+Lines \ref{line:mask_presentC}--\ref{line:check_entry_presentC} check
 \fi
@@
 from Line \ref{line:check_entry_present_jump}'s conditional jump. Otherwise the code jumps ahead to
-the case for the next level already existing, which is discussed in Section \ref{sec:p2v} and Figure \ref{fig:p2v}.
+the case for the next level already existing, which is discussed in Section \ref{sec:p2vC} and Figure \ref{fig:p2vC}.
 First, we must discuss another refinement of the address space invariant, establishing
@@
   } (*\label{line:end_of_allocation_pathC}*)
-  /*Physical to virtual address mapping code segment starts here (*\ref{fig:p2vC}*)*/
+  /*Physical to virtual address mapping code segment starts here (*\label{line:finalpieceS}*)*/
   $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \}),m) \ast \textsf{entry} \mapsto_{\textsf{id}} \textsf{\_} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
@@
   uintptr_t next_virt_addr = (uintptr_t) P2V(next_phys_addr); (*\label{line:p2vC}*);
-  next = (pte_t *) next_virt_addr;
+  next = (pte_t *) next_virt_addr;(*\label{line:finalpieceE}*)
   return next;
@@
 discussed in Section \ref{subsec:identitymappingsC} and Figure \ref{fig:peraspaceinvariant_with_p2v_extensionC}.
-in the last lines of \lstinline|pte_get_next_table| shown in Figure \ref{fig:p2v} (the continuation of Figure \ref{fig:calltopteinitializeC}).
+in the last lines of \lstinline|pte_get_next_table| shown in Figure \ref{fig:p2vC} (the continuation of Figure \ref{fig:calltopteinitializeC}).
 This is a critical piece of the full page table walk verification.

### 2025-07-17 background.tex
structure

--- 2025-03-26 background.tex
@@
 \begin{figure}
-    \includegraphics[width=\columnwidth]{pagetables2.png}
+%    \begin{framed}
+\begin{multicols}{2}
+%\column{.59\textwidth}    
+    \begin{figure}
+%        \centering
+\tikzset{every picture/.style={line width=0.75pt}} %set default line width to 0.75pt        
+\begin{tikzpicture}[x=0.75pt,y=0.75pt,yscale=-0.5,xscale=0.5]
+%uncomment if require: \path (0,471); %set diagram left start at 0, and has height of 471
+
+%Shape: Rectangle [id:dp6797306099675045] 
+\draw   (34,165) -- (124,165) -- (124,316) -- (34,316) -- cycle ;
+%Shape: Rectangle [id:dp6192113478200691] 
+\draw   (167,165) -- (257,165) -- (257,316) -- (167,316) -- cycle ;
+%Shape: Rectangle [id:dp046479372017499854] 
+\draw   (296,165) -- (386,165) -- (386,316) -- (296,316) -- cycle ;
+%Shape: Rectangle [id:dp1806612887251846] 
+\draw   (428,165) -- (518,165) -- (518,316) -- (428,316) -- cycle ;
+%Shape: Rectangle [id:dp16201905390239735] 
+\draw   (116,78) -- (186,78) -- (186,118) -- (116,118) -- cycle ;
+%Shape: Rectangle [id:dp46359313115432044] 
+\draw   (186,78) -- (256,78) -- (256,118) -- (186,118) -- cycle ;
+%Shape: Rectangle [id:dp8986469180554777] 
+\draw   (256,78) -- (326,78) -- (326,118) -- (256,118) -- cycle ;
+%Shape: Rectangle [id:dp3960250568381902] 
+\draw   (326,78) -- (396,78) -- (396,118) -- (326,118) -- cycle ;
+%Shape: Rectangle [id:dp76786771788994] 
+\draw   (396,78) -- (466,78) -- (466,118) -- (396,118) -- cycle ;
+%Shape: Rectangle [id:dp9070245370995538] 
+\draw   (466,78) -- (536,78) -- (536,118) -- (466,118) -- cycle ;
+%Curve Lines [id:da6951218383977624] 
+\draw [color={rgb, 255:red, 208; green, 2; blue, 27 }  ,draw opacity=1 ]   (211,121) .. controls (-36.94,123.94) and (16.69,214.28) .. (31.17,237.65) ;
+\draw [shift={(32,239)}, rotate = 238.24] [color={rgb, 255:red, 208; green, 2; blue, 27 }  ,draw opacity=1 ][line width=0.75]    (10.93,-3.29) ..
+controls (6.95,-1.4) and (3.31,-0.3) ..
+(0,0) ..
+controls (3.31,0.3) and (6.95,1.4) ..
+(10.93,3.29)   ;
+%Shape: Rectangle [id:dp48687197518810277] 
+\draw   (34,215) -- (124,215) -- (124,255) -- (34,255) -- cycle ;
+%Curve Lines [id:da7139570096435448] 
+\draw [color={rgb, 255:red, 208; green, 2; blue, 27 }  ,draw opacity=1 ]   (291,121) .. controls (117.54,136.68) and (141.92,192.7) .. (161.79,218.46) ;
+\draw [shift={(163,220)}, rotate = 231.34] [color={rgb, 255:red, 208; green, 2; blue, 27 }  ,draw opacity=1 ][line width=0.75]    (10.93,-3.29) ..
+controls (6.95,-1.4) and (3.31,-0.3) ..
+(0,0) ..
+controls (3.31,0.3) and (6.95,1.4) ..
+(10.93,3.29)   ;
+%Shape: Rectangle [id:dp32338125663089956] 
+\draw   (167,201) -- (257,201) -- (257,241) -- (167,241) -- cycle ;
+%Shape: Rectangle [id:dp6104052627034173] 
+\draw   (296,180) -- (386,180) -- (386,220) -- (296,220) -- cycle ;
+%Shape: Rectangle [id:dp49476345634718477] 
+\draw   (428,236) -- (518,236) -- (518,276) -- (428,276) -- cycle ;
+%Curve Lines [id:da3926551025530285] 
+\draw [color={rgb, 255:red, 208; green, 2; blue, 27 }  ,draw opacity=1 ]   (358,121) .. controls (263.92,134.72) and (273.56,172.45) .. (292.81,197.48) ;
+\draw [shift={(294,199)}, rotate = 231.34] [color={rgb, 255:red, 208; green, 2; blue, 27 }  ,draw opacity=1 ][line width=0.75]    (10.93,-3.29) ..
+controls (6.95,-1.4) and (3.31,-0.3) ..
+(0,0) ..
+controls (3.31,0.3) and (6.95,1.4) ..
+(10.93,3.29)   ;
+%Curve Lines [id:da38615138851194764] 
+\draw [color={rgb, 255:red, 208; green, 2; blue, 27 }  ,draw opacity=1 ]   (434,121) .. controls (385.98,135.7) and (402.31,221.47) .. (421.8,248.42) ;
+\draw [shift={(423,250)}, rotate = 231.34] [color={rgb, 255:red, 208; green, 2; blue, 27 }  ,draw opacity=1 ][line width=0.75]    (10.93,-3.29) ..
+controls (6.95,-1.4) and (3.31,-0.3) ..
+(0,0) ..
+controls (3.31,0.3) and (6.95,1.4) ..
+(10.93,3.29)   ;
+%Curve Lines [id:da8796700124764012] 
+\draw [color={rgb, 255:red, 208; green, 2; blue, 27 }  ,draw opacity=1 ]   (499,120) .. controls (574.46,142.54) and (571.17,273.61) .. (570.06,307.07) ;
+\draw [shift={(570,309)}, rotate = 271.91] [color={rgb, 255:red, 208; green, 2; blue, 27 }  ,draw opacity=1 ][line width=0.75]    (10.93,-3.29) ..
+controls (6.95,-1.4) and (3.31,-0.3) ..
+(0,0) ..
+controls (3.31,0.3) and (6.95,1.4) ..
+(10.93,3.29)   ;
+%Curve Lines [id:da26463336009584704] 
+\draw    (43,373) .. controls (-9.21,383.84) and (18.15,347.13) .. (33.32,317.36) ;
+\draw [shift={(34,316)}, rotate = 116.57] [color={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.75]    (10.93,-3.29) ..
+controls (6.95,-1.4) and (3.31,-0.3) ..
+(0,0) ..
+controls (3.31,0.3) and (6.95,1.4) ..
+(10.93,3.29)   ;
+%Curve Lines [id:da18202057955172246] 
+\draw [color={rgb, 255:red, 208; green, 2; blue, 27 }  ,draw opacity=1 ]   (123,239) .. controls (129.9,278.4) and (132.91,293.54) .. (160.71,313.1) ;
+\draw [shift={(162,314)}, rotate = 214.59] [color={rgb, 255:red, 208; green, 2; blue, 27 }  ,draw opacity=1 ][line width=0.75]    (10.93,-3.29) ..
+controls (6.95,-1.4) and (3.31,-0.3) ..
+(0,0) ..
+controls (3.31,0.3) and (6.95,1.4) ..
+(10.93,3.29)   ;
+%Curve Lines [id:da39925694981212034] 
+\draw [color={rgb, 255:red, 208; green, 2; blue, 27 }  ,draw opacity=1 ]   (258,221) .. controls (259.95,260) and (278.06,294.25) .. (294.72,314.47) ;
+\draw [shift={(296,316)}, rotate = 229.64] [color={rgb, 255:red, 208; green, 2; blue, 27 }  ,draw opacity=1 ][line width=0.75]    (10.93,-3.29) ..
+controls (6.95,-1.4) and (3.31,-0.3) ..
+(0,0) ..
+controls (3.31,0.3) and (6.95,1.4) ..
+(10.93,3.29)   ;
+%Curve Lines [id:da0025982000064483923] 
+\draw [color={rgb, 255:red, 208; green, 2; blue, 27 }  ,draw opacity=1 ]   (387,201) .. controls (393.86,250) and (409.36,294.2) .. (426.92,314.77) ;
+\draw [shift={(428,316)}, rotate = 228.01] [color={rgb, 255:red, 208; green, 2; blue, 27 }  ,draw opacity=1 ][line width=0.75]    (10.93,-3.29) ..
+controls (6.95,-1.4) and (3.31,-0.3) ..
+(0,0) ..
+controls (3.31,0.3) and (6.95,1.4) ..
+(10.93,3.29)   ;
+%Curve Lines [id:da4425456634879614] 
+\draw [color={rgb, 255:red, 208; green, 2; blue, 27 }  ,draw opacity=1 ]   (518,253) .. controls (525.76,277.25) and (544.81,300.56) .. (557.81,314.71) ;
+\draw [shift={(559,316)}, rotate = 227.12] [color={rgb, 255:red, 208; green, 2; blue, 27 }  ,draw opacity=1 ][line width=0.75]    (10.93,-3.29) ..
+controls (6.95,-1.4) and (3.31,-0.3) ..
+(0,0) ..
+controls (3.31,0.3) and (6.95,1.4) ..
+(10.93,3.29)   ;
+%Shape: Rectangle [id:dp20590678105984384] 
+\draw   (520,362) -- (590,362) -- (590,402) -- (520,402) -- cycle ;
+%Shape: Rectangle [id:dp8145206670609288] 
+\draw   (87,364) -- (128.5,364) -- (128.5,404) -- (87,404) -- cycle ;
+%Shape: Rectangle [id:dp5322999267202655] 
+\draw   (128.5,364) -- (257.5,364) -- (257.5,404) -- (128.5,404) -- cycle ;
+%Shape: Rectangle [id:dp44035388924299834] 
+\draw   (257.5,364) -- (297,364) -- (297,404) -- (257.5,404) -- cycle ;
+\draw   (561,324.5) ..
+controls (561,319.25) and (565.25,315) ..
+(570.5,315) ..
+controls (575.75,315) and (580,319.25) ..
+(580,324.5) ..
+controls (580,329.75) and (575.75,334) ..
+(570.5,334) ..
+controls (565.25,334) and (561,329.75) ..
+(561,324.5) -- cycle ; \draw   (561,324.5) -- (580,324.5) ; \draw   (570.5,315) -- (570.5,334) ;
+%Curve Lines [id:da7994290642535546] 
+\draw [color={rgb, 255:red, 208; green, 2; blue, 27 }  ,draw opacity=1 ]   (563.5,333) .. controls (513.52,328.1) and (528.85,329.92) .. (544.54,357.29) ;
+\draw [shift={(545.5,359)}, rotate = 241.11] [color={rgb, 255:red, 208; green, 2; blue, 27 }  ,draw opacity=1 ][line width=0.75]    (10.93,-3.29) ..
+controls (6.95,-1.4) and (3.31,-0.3) ..
+(0,0) ..
+controls (3.31,0.3) and (6.95,1.4) ..
+(10.93,3.29)   ;
+
+% Text Node
+\draw (12,13) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\scriptsize Virtual}};
+% Text Node
+\draw (550,442) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\scriptsize Physical}};
+% Text Node
+\draw (133,89) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\scriptsize Sign}};
+% Text Node
+\draw (196,89) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\scriptsize L4 I}};
+% Text Node
+\draw (265,89) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\scriptsize L3 I}};
+% Text Node
+\draw (336,89) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\scriptsize L2 I}};
+% Text Node
+\draw (406,89) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\scriptsize L1 I}};
+% Text Node
+\draw (482,89) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\scriptsize Pg}};
+% Text Node
+\draw (55,91) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\scriptsize va$_i$}};
+% Text Node
+\draw (34,373) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\scriptsize root}};
+% Text Node
+\draw (50,227) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\scriptsize L4 E}};
+% Text Node
+\draw (184,213) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\scriptsize L3 E}};
+% Text Node
+\draw (314,191) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\scriptsize L2 E}};
+% Text Node
+\draw (446,247) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\scriptsize L1 E}};
+% Text Node
+\draw (471,375) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\scriptsize page}};
+% Text Node
+\draw (538,372) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\scriptsize 4K}};
+% Text Node
+\draw (148,412) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\scriptsize 51}};
+% Text Node
+\draw (218,412) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\scriptsize 12}};
+% Text Node
+\draw (137,376) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\scriptsize Base Addr}};
+% Text Node
+\draw (111,49) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\scriptsize 63}};
+% Text Node
+\draw (182,49) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\scriptsize 48}};
+% Text Node
+\draw (252,49) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\scriptsize 39}};
+% Text Node
+\draw (321,49) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\scriptsize 30}};
+% Text Node
+\draw (392,49) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\scriptsize 21}};
+% Text Node
+\draw (461,49) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\scriptsize 12}};
+% Text Node
+\draw (530,49) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\scriptsize 0}};
+
+\end{tikzpicture}
+          \caption{Accessing to the Page Referenced by L1 Entry}
+        \label{fig:enter-label}
+    \end{figure}
+\columnbreak
+    \begin{figure}
+\begin{lstlisting}%[style=CStyleNum, basicstyle=\tiny]
+static pte_t *pte_nxt_table (pte_t *entry){
+ pte_t *next;
+ // If not already present, try to allocate
+ if (!entry->present){
+  if (!pte_alloc(&next)) {
+   return NULL;
+  }
+  entry->pfn = PTE_PFN((uintptr_t) next);
+  entry->present = 1;
+  } else {
+   uintptr_t next_phys_addr = PTE_PFN_TO_ADDR(entry->pfn);        
+   uintptr_t next_virt_addr = (uintptr_t) P2V(next_phys_addr);
+   next = (pte_t *) next_virt_addr;
+  }
+  return next;
+}   
+pte_t *walkpgdir(pte_t *l4, void *va){ 
+ pte_t *l4_entry = &l4[L4I(va)];
+ pte_t *l3 = pte_nxt_table(l4_entry);
+ pte_t *l3_entry = &l3[L3I(va)];
+ pte_t *l2 = pte_nxt_table(l3_entry);
+ pte_t *l2_entry = &l2[L2I(va)];  
+ pte_t *l1 = pte_nxt_table(l2_entry);
+ pte_t *l1_entry = &l1[L1I(va)];  
+ return l1_entry
+}
+\end{lstlisting}
+%          \caption{Walking page-table tree to locate L1 entry C-source code.}
+%        \label{fig:enter-label}
+    \end{figure}
+\end{multicols}
+%\end{framed}
     \vspace{-2.5em}

### 2025-07-17 background.tex
+

@@
 %    \begin{framed}
-\begin{multicols}{2}
 %\column{.59\textwidth}    
-    \begin{figure}
+    \begin{subfigure}
 %        \centering
@@
         \label{fig:enter-label}
-    \end{figure}
-\columnbreak
-    \begin{figure}
+    \end{subfigure}
+    \vfill
+    \begin{multicols}{2}
+%    \begin{figure}
 \begin{lstlisting}%[style=CStyleNum, basicstyle=\tiny]
@@
 }
+\end{lstlisting}
+\columnbreak
+\begin{lstlisting}
 pte_t *walkpgdir(pte_t *l4, void *va){ 
@@
 %        \label{fig:enter-label}
-    \end{figure}
+ %   \end{figure}
 \end{multicols}

### 2025-07-17 background.tex
+

--- 2025-07-16 background.tex
@@
 \begin{figure}
-    \includegraphics[width=\columnwidth]{pagetables2.png}
+%    \begin{framed}
+%\column{.59\textwidth}    
+\begin{subfigure}
+
+\tikzset{every picture/.style={line width=0.75pt}} %set default line width to 0.75pt
+
+\begin{tikzpicture}[x=0.75pt,y=0.75pt,yscale=-0.5,xscale=0.5]
+%uncomment if require: \path (0,430); %set diagram left start at 0, and has height of 430
+
+%Shape: Rectangle [id:dp9406932326447903] 
+\draw   (33,141) -- (123,141) -- (123,292) -- (33,292) -- cycle ;
+%Shape: Rectangle [id:dp7682116051394419] 
+\draw   (166,141) -- (256,141) -- (256,292) -- (166,292) -- cycle ;
+%Shape: Rectangle [id:dp07376733516930423] 
+\draw   (295,141) -- (385,141) -- (385,292) -- (295,292) -- cycle ;
+%Shape: Rectangle [id:dp7106629147891306] 
+\draw   (427,141) -- (517,141) -- (517,292) -- (427,292) -- cycle ;
+%Shape: Rectangle [id:dp1342906826203598] 
+\draw   (115,54) -- (185,54) -- (185,94) -- (115,94) -- cycle ;
+%Shape: Rectangle [id:dp08072439007207399] 
+\draw   (185,54) -- (255,54) -- (255,94) -- (185,94) -- cycle ;
+%Shape: Rectangle [id:dp8282933603833216] 
+\draw   (255,54) -- (325,54) -- (325,94) -- (255,94) -- cycle ;
+%Shape: Rectangle [id:dp8416255090865326] 
+\draw   (325,54) -- (395,54) -- (395,94) -- (325,94) -- cycle ;
+%Shape: Rectangle [id:dp22911958218529804] 
+\draw   (395,54) -- (465,54) -- (465,94) -- (395,94) -- cycle ;
+%Shape: Rectangle [id:dp33732955213990756] 
+\draw   (465,54) -- (535,54) -- (535,94) -- (465,94) -- cycle ;
+%Curve Lines [id:da6688730321284677] 
+\draw    (211.5,96) .. controls (-36.44,98.94) and (15.75,190.24) .. (30.17,213.65) ;
+\draw [shift={(31,215)}, rotate = 238.24] [color={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.75]    (10.93,-3.29) ..
+controls (6.95,-1.4) and (3.31,-0.3) ..
+(0,0) ..
+controls (3.31,0.3) and (6.95,1.4) ..
+(10.93,3.29)   ;
+%Shape: Rectangle [id:dp6044747978986409] 
+\draw   (33,191) -- (123,191) -- (123,231) -- (33,231) -- cycle ;
+%Curve Lines [id:da8867562998134584] 
+\draw    (289.5,95) .. controls (115.15,110.76) and (137.77,175.03) .. (163.33,197.98) ;
+\draw [shift={(164.5,199)}, rotate = 220.24] [color={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.75]    (10.93,-3.29) ..
+controls (6.95,-1.4) and (3.31,-0.3) ..
+(0,0) ..
+controls (3.31,0.3) and (6.95,1.4) ..
+(10.93,3.29)   ;
+%Shape: Rectangle [id:dp8861153167245819] 
+\draw   (166,177) -- (256,177) -- (256,217) -- (166,217) -- cycle ;
+%Shape: Rectangle [id:dp05979844200848894] 
+\draw  [dash pattern={on 4.5pt off 4.5pt}] (295,156) -- (385,156) -- (385,196) -- (295,196) -- cycle ;
+%Shape: Rectangle [id:dp17403733244247843] 
+\draw   (427,212) -- (517,212) -- (517,252) -- (427,252) -- cycle ;
+%Curve Lines [id:da20581124900573955] 
+\draw [color={rgb, 255:red, 0; green, 0; blue, 0 }  ,draw opacity=1 ] [dash pattern={on 4.5pt off 4.5pt}]  (357,97) .. controls (262.92,110.72) and (272.56,148.45) .. (291.81,173.48) ;
+\draw [shift={(293,175)}, rotate = 231.34] [color={rgb, 255:red, 0; green, 0; blue, 0 }  ,draw opacity=1 ][line width=0.75]    (10.93,-3.29) ..
+controls (6.95,-1.4) and (3.31,-0.3) ..
+(0,0) ..
+controls (3.31,0.3) and (6.95,1.4) ..
+(10.93,3.29)   ;
+%Curve Lines [id:da07589611603635316] 
+\draw    (425.5,95) .. controls (380.42,119.5) and (406.41,202.58) .. (422.53,224.72) ;
+\draw [shift={(423.5,226)}, rotate = 231.34] [color={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.75]    (10.93,-3.29) ..
+controls (6.95,-1.4) and (3.31,-0.3) ..
+(0,0) ..
+controls (3.31,0.3) and (6.95,1.4) ..
+(10.93,3.29)   ;
+%Curve Lines [id:da6852167546673456] 
+\draw    (497.5,95) .. controls (564.48,91.06) and (565,230.71) .. (567.87,269.31) ;
+\draw [shift={(568,271)}, rotate = 265.24] [color={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.75]    (10.93,-3.29) ..
+controls (6.95,-1.4) and (3.31,-0.3) ..
+(0,0) ..
+controls (3.31,0.3) and (6.95,1.4) ..
+(10.93,3.29)   ;
+%Curve Lines [id:da017521077423665377] 
+\draw    (39.5,359) .. controls (-12.71,369.84) and (17.07,323.43) .. (32.31,293.36) ;
+\draw [shift={(33,292)}, rotate = 116.57] [color={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.75]    (10.93,-3.29) ..
+controls (6.95,-1.4) and (3.31,-0.3) ..
+(0,0) ..
+controls (3.31,0.3) and (6.95,1.4) ..
+(10.93,3.29)   ;
+%Curve Lines [id:da6750042432035159] 
+\draw    (122,215) .. controls (128.9,254.4) and (131.91,269.54) .. (159.71,289.1) ;
+\draw [shift={(161,290)}, rotate = 214.59] [color={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.75]    (10.93,-3.29) ..
+controls (6.95,-1.4) and (3.31,-0.3) ..
+(0,0) ..
+controls (3.31,0.3) and (6.95,1.4) ..
+(10.93,3.29)   ;
+%Curve Lines [id:da14788158682113428] 
+\draw    (257,197) .. controls (258.95,236) and (277.06,270.25) .. (293.72,290.47) ;
+\draw [shift={(295,292)}, rotate = 229.64] [color={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.75]    (10.93,-3.29) ..
+controls (6.95,-1.4) and (3.31,-0.3) ..
+(0,0) ..
+controls (3.31,0.3) and (6.95,1.4) ..
+(10.93,3.29)   ;
+%Curve Lines [id:da3297977595767969] 
+\draw    (386,177) .. controls (392.86,226) and (408.36,270.2) .. (425.92,290.77) ;
+\draw [shift={(427,292)}, rotate = 228.01] [color={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.75]    (10.93,-3.29) ..
+controls (6.95,-1.4) and (3.31,-0.3) ..
+(0,0) ..
+controls (3.31,0.3) and (6.95,1.4) ..
+(10.93,3.29)   ;
+%Curve Lines [id:da7710391961792686] 
+\draw    (517,229) .. controls (524.76,253.25) and (540.05,269.03) .. (552.82,282.73) ;
+\draw [shift={(554,284)}, rotate = 227.12] [color={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.75]    (10.93,-3.29) ..
+controls (6.95,-1.4) and (3.31,-0.3) ..
+(0,0) ..
+controls (3.31,0.3) and (6.95,1.4) ..
+(10.93,3.29)   ;
+%Shape: Rectangle [id:dp1535611125971872] 
+\draw   (514,336) -- (584,336) -- (584,376) -- (514,376) -- cycle ;
+%Shape: Rectangle [id:dp05427947476226991] 
+\draw   (95,338) -- (136.5,338) -- (136.5,378) -- (95,378) -- cycle ;
+%Shape: Rectangle [id:dp10511138743752002] 
+\draw   (136.5,338) -- (265.5,338) -- (265.5,378) -- (136.5,378) -- cycle ;
+%Shape: Rectangle [id:dp8917680919534263] 
+\draw   (265.5,338) -- (305,338) -- (305,378) -- (265.5,378) -- cycle ;
+\draw   (558,284.5) ..
+controls (558,279.25) and (562.25,275) ..
+(567.5,275) ..
+controls (572.75,275) and (577,279.25) ..
+(577,284.5) ..
+controls (577,289.75) and (572.75,294) ..
+(567.5,294) ..
+controls (562.25,294) and (558,289.75) ..
+(558,284.5) -- cycle ; \draw   (558,284.5) -- (577,284.5) ; \draw   (567.5,275) -- (567.5,294) ;
+%Curve Lines [id:da9312068388962929] 
+\draw    (568,298) .. controls (584.66,309.76) and (586.91,318.64) .. (565.35,335.93) ;
+\draw [shift={(564,337)}, rotate = 321.95] [color={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.75]    (10.93,-3.29) ..
+controls (6.95,-1.4) and (3.31,-0.3) ..
+(0,0) ..
+controls (3.31,0.3) and (6.95,1.4) ..
+(10.93,3.29)   ;
+
+% Text Node
+\draw (11,14) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\scriptsize Virtual}};
+% Text Node
+\draw (570,395) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\scriptsize Physical}};
+% Text Node
+\draw (104,28) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\scriptsize 63}};
+% Text Node
+\draw (175,28) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\scriptsize 48}};
+% Text Node
+\draw (245,28) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\scriptsize 39}};
+% Text Node
+\draw (314,28) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\scriptsize 30}};
+% Text Node
+\draw (385,28) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\scriptsize 21}};
+% Text Node
+\draw (454,28) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\scriptsize 12}};
+% Text Node
+\draw (523,28) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\scriptsize 0}};
+% Text Node
+\draw (132,65) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\scriptsize Sign}};
+% Text Node
+\draw (195,65) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\scriptsize L4 Offset}};
+% Text Node
+\draw (264,65) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\scriptsize L3 Offset}};
+% Text Node
+\draw (336,65) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\scriptsize L2 Offset}};
+% Text Node
+\draw (406,65) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\scriptsize L1 Offset}};
+% Text Node
+\draw (472,65) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\scriptsize Page Offset}};
+% Text Node
+\draw (67,63) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\scriptsize va}};
+% Text Node
+\draw (53,203) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\scriptsize L4 Entry}};
+% Text Node
+\draw (185,189) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\scriptsize L3 Entry}};
+% Text Node
+\draw (316,167) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\scriptsize L2 Entry}};
+% Text Node
+\draw (449,223) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\scriptsize L1 Entry}};
+% Text Node
+\draw (51,347) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\scriptsize root}};
+% Text Node
+\draw (465,349) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\scriptsize page}};
+% Text Node
+\draw (532,346) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\scriptsize 4K Data}};
+% Text Node
+\draw (156,386) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\scriptsize 51}};
+% Text Node
+\draw (226,386) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\scriptsize 12}};
+% Text Node
+\draw (145,350) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\scriptsize Base Addr}};
+
+\end{tikzpicture}
+          \caption{Accessing to the Page Referenced by L1 Entry}
+        \label{fig:enter-label}
+    \end{subfigure}
+    \vfill
+    \begin{multicols}{2}
+%    \begin{figure}
+\begin{lstlisting}[mathescape,escapeinside={(*}{*)}]
+/*
+@param entry: entry whose frame address is used to locate the a-level-lower entry in the page-tree
+@return next: return a virtual pointer to the descendant tree
+*/
+pte_t *pte_get_next_table(pte_t *entry) {
+  pte_t *next;
+  /*Reading page table entry justified by virtual pte-pointsto*/
+  if (!entry->present(*\label{line:read_entry_contentsC}*) (*\label{line:conditional_childrenC}*)) {
+    /* Jump if the present bit is not zero; no need to allocate next level (* \label{line:check_entry_presentC} *)(* \label{line:mask_presentC} *)  (*\label{line:after_concluding_qfrac1C}*)*/
+    /*Allocation path starts here (*\label{line:pass_addrof_nextC}*)(*\label{line:alloc_path_startC}*)*/
+    pte_initialize(entry);(* \label{line:call_to_pte_initializeC} *)
+    /*Entry value updates:*/
+    entry->pfn = nextpaddr;
+    entry->present = 1; (*\label{line:install_new_entryC}*)
+    /*Now we know that entry is initialized, so we satisfy the condition to access children list */
+  } (*\label{line:end_of_allocation_pathCB}*)
+  /*Physical to virtual address mapping code segment starts here (*\label{line:finalpieceSB}*)*/
+  uintptr_t next_phys_addr = PTE_PFN_TO_ADDR(entry->pfn)
+  uintptr_t next_virt_addr = (uintptr_t) P2V(next_phys_addr); (*\label{line:p2vCB}*);
+  next = (pte_t *) next_virt_addr;(*\label{line:finalpieceEB}*)
+  return next;
+}
+\end{lstlisting}
+\columnbreak
+\begin{lstlisting}[mathescape,escapeinside={(*}{*)}]
+pte_t *walkpgdir(pte_t *l4, void *va){ 
+ pte_t *l4_entry = &l4[L4Offset(va)];
+ pte_t *l3 = pte_nxt_table(l4_entry);
+ pte_t *l3_entry = &l3[L3Offset(va)];
+ pte_t *l2 = pte_nxt_table(l3_entry);
+ pte_t *l2_entry = &l2[L2Offset(va)];  
+ pte_t *l1 = pte_nxt_table(l2_entry);
+ pte_t *l1_entry = &l1[L1Offset(va)];  
+ return l1_entry
+}
+\end{lstlisting}
+%          \caption{Walking page-table tree to locate L1 entry C-source code.}
+%        \label{fig:enter-label}
+ %   \end{figure}
+\end{multicols}
+%\end{framed}
     \vspace{-2.5em}
@@
 a certain range of virtual addresses (corresponding to the kernel's code and data) are valid in every address space.
-
-\paragraph{\add{Mapping a New Page}}
-\add{Mapping a new page into a virtual address space is a key piece fo functionality, and also a key challenge for verification.
-This operation takes a chunk of physical memory to install, and a (page-aligned) virtual address at which to make it
-visible to the process.
-Starting from the page table root (\lstinline|cr3| in Figure \ref{fig:pagetables}), it must traverse the page tables until it locates
-the correct location of the L1 entry which should point to the physical page (installing new L3, L2, and L1 tables as needed),
-and update the L1 entry to point to the physical page. This is complicated by the fact that the kernel accesses the page
-tables through \emph{virtual} addresses, but the page table root and the addresses in the L4--L1 entries are all \emph{physical}
-addresses. So the mapping process must repeatedly perform \emph{backwards} mappings (not supported by hardware) from
-physical addresses of each level's entry to a virtual address that is then mapped back to the corresponding physical address!
-And for modularity, the translation process (and possible allocation of missing tables) is typically implemented in a function
-used for every level of translation (i.e., a level-polymorphic function).
-These compounding subtleties are one reason \textsc{CertiKOS} and \textsc{seL4} \emph{trust} their page table
-manipulations; both the reasoning to prove correctness, and the kernel invariants that ensure correctness of typical implementations,
-are remarkably subtle.
-}
-\looseness=-1
 
@@
 \rocq terms of a particular type, rather than being drawn from a fixed vocabulary.
-\del{We introduce other details as they arise.}
+We introduce other details as they arise.
+
 \iris includes two forms of implication. The magic wand operator $\wand$ is an affine implication:
@@
 }
-View shifts are essentially logical entailment plus ghost updates.
 \looseness=-1
@@
 
-An important limitation \del{here} is that to date, every separation logic
+An important limitation here is that to date, every separation logic
 has assumed that all pointer addresses are meant for use in a single address space, which avoided the
@@
 is true in some possible future or past world related in a fixed way to the current world, but to talk about \emph{arbitrary}
-other worlds.
-\add{This is somewhat different from the accessibility relation modalities more common in verification, but
-also well-established.}
-This explicit naming of alternate worlds increases the power of modal logics \add{over those without it}~\cite{blackburn1995hybrid}, and is actually
+other worlds. This explicit naming of alternate worlds increases the power of modal logics~\cite{blackburn1995hybrid}, and is actually
 necessary for completeness in classical separation logics~\cite{brotherston2014parametric}.
@@
 space with root $r$ --- thus changing the interpretation of a modality, and even whether a modality is valid --- \emph{while assertions with that modality are active}.
-\add{This is essential to updating page mappings for the current address space in use by a CPU.}
 %

### 2025-07-17 examples_C.tex
+

@@
 
-\lstdefinestyle{mystyle}
-{
-    language = C,
-    basicstyle = {\footnotesize,\color{black}},
-    backgroundcolor = {\color{white}},
-    stringstyle = {\color{string-color}},
-    keywordstyle = {\color{key-color}},
-    keywordstyle = [2]{\color{lime}},
-    keywordstyle = [3]{\color{yellow}},
-    keywordstyle = [4]{\color{teal}},
-    otherkeywords = {;,<<,>>,++},
-    morekeywords = [2]{;},
-    morekeywords = [3]{<<, >>},
-    morekeywords = [4]{++},
-    escapeinside={(*}{*)},
-}
 \newcommand{\sumwalkabsent}{

### 2025-07-17 background.tex
+

@@
 address, but tagged with the physical data addresses, so cache lookups and address translations can proceed in parallel.} and/or memory-bus.
-
-\begin{figure}
+\setlength{\columnseprule}{4pt}
+\begin{figure}\footnotesize
 %    \begin{framed}
 %\column{.59\textwidth}    
-\begin{subfigure}
+\begin{subfigure}\footnotesize
 
@@
 
-\begin{tikzpicture}[x=0.75pt,y=0.75pt,yscale=-0.5,xscale=0.5]
-%uncomment if require: \path (0,430); %set diagram left start at 0, and has height of 430
+\begin{tikzpicture}[x=0.75pt,y=0.75pt,yscale=-0.75,xscale=0.75]
+%uncomment if require: \path (0,427); %set diagram left start at 0, and has height of 427
 
@@
 % Text Node
-\draw (53,203) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\scriptsize L4 Entry}};
+\draw (48,203) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\scriptsize L4 Entry}};
 % Text Node
-\draw (185,189) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\scriptsize L3 Entry}};
+\draw (180,189) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\scriptsize L3 Entry}};
 % Text Node
-\draw (316,167) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\scriptsize L2 Entry}};
+\draw (311,167) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\scriptsize L2 Entry}};
 % Text Node
-\draw (449,223) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\scriptsize L1 Entry}};
+\draw (444,223) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\scriptsize L1 Entry}};
 % Text Node
@@
 % Text Node
-\draw (532,346) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\scriptsize 4K Data}};
+\draw (529,346) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\scriptsize 4K Data}};
 % Text Node
-\draw (156,386) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\scriptsize 51}};
+\draw (133,386) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\scriptsize 51}};
 % Text Node
-\draw (226,386) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\scriptsize 12}};
+\draw (258,386) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\scriptsize 12}};
 % Text Node
-\draw (145,350) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\scriptsize Base Addr}};
+\draw (163,350) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\scriptsize Base Addr}};
 
@@
     \vfill
-    \begin{multicols}{2}
+\begin{minipage}{.5\textwidth}
 %    \begin{figure}
-\begin{lstlisting}[mathescape,escapeinside={(*}{*)}]
+\begin{lstlisting}[mathescape,escapeinside={(*}{*)},basicstyle=\footnotesize]
 /*
-@param entry: entry whose frame address is used to locate the a-level-lower entry in the page-tree
+@param entry: entry whose frame address is used 
+              to locate the subsequent entry in the page-tree
 @return next: return a virtual pointer to the descendant tree
@@
   pte_t *next;
-  /*Reading page table entry justified by virtual pte-pointsto*/
+  /*Reading page table entry:
+    pointed by a virtual memory address*/
+  /* Jump if the present bit is not zero:
+     no need to allocate next level (* \label{line:check_entry_presentC} *)(* \label{line:mask_presentC} *)  (*\label{line:after_concluding_qfrac1C}*)*/
   if (!entry->present(*\label{line:read_entry_contentsC}*) (*\label{line:conditional_childrenC}*)) {
-    /* Jump if the present bit is not zero; no need to allocate next level (* \label{line:check_entry_presentC} *)(* \label{line:mask_presentC} *)  (*\label{line:after_concluding_qfrac1C}*)*/
     /*Allocation path starts here (*\label{line:pass_addrof_nextC}*)(*\label{line:alloc_path_startC}*)*/
@@
     entry->present = 1; (*\label{line:install_new_entryC}*)
-    /*Now we know that entry is initialized, so we satisfy the condition to access children list */
+    /*Entry is initialized,
+      so we satisfy the condition to access children list */
   } (*\label{line:end_of_allocation_pathCB}*)
@@
 \end{lstlisting}
-\columnbreak
-\begin{lstlisting}[mathescape,escapeinside={(*}{*)}]
+\end{minipage}
+\begin{minipage}{.4\textwidth}
+\begin{lstlisting}[mathescape,escapeinside={(*}{*)}, basicstyle=\footnotesize]
 pte_t *walkpgdir(pte_t *l4, void *va){ 
@@
  %   \end{figure}
-\end{multicols}
+\end{minipage}
 %\end{framed}

### 2025-07-17 background_C.tex
+

--- 2025-07-17 background.tex
@@
 \begin{lstlisting}[mathescape,escapeinside={(*}{*)}, basicstyle=\footnotesize]
-pte_t *walkpgdir(pte_t *l4, void *va){ 
-  pte_t *l4_entry = &l4[L4Offset(va)];
-  pte_t *l3 = pte_nxt_table(l4_entry);
+pte_t *walkpgdir(pte_t *root, void *va){ 
+  pte_t *l4_entry = &root[L4Offset(va)];
+  pte_t *l3 = pte_next_table(l4_entry);
   pte_t *l3_entry = &l3[L3Offset(va)];
-  pte_t *l2 = pte_nxt_table(l3_entry);
+  pte_t *l2 = pte_next_table(l3_entry);
   pte_t *l2_entry = &l2[L2Offset(va)];  
-  pte_t *l1 = pte_nxt_table(l2_entry);
+  pte_t *l1 = pte_next_table(l2_entry);
   pte_t *l1_entry = &l1[L1Offset(va)];  

### 2025-07-17 examples_C.tex
C pseudo annotated files

@@
 /*
-@param plm4: the root address of the page-table tree
+@param l4: the root address of the page-table tree
 @param va: virtual address to be translated and mapped
 */
-pte_t *walkpgdir(pte_t *pml4, const void *va) {
-  $\specline{\textsf{P} \ast \ulcorner\textsf{rtv+KERNBASE}=\textsf{pml4}\urcorner \ast \forall_{i\in\textsf{0..511}}\ghostmaptoken{\textsf{id}}{(\textsf{rtv}+i*8)}{\textsf{4}} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}}_{\rtv}$
-  pte_t *pml4_entry = &pml4[PML4EX(va)]; // Virtual address of L4 entry(*\label{line:start_pml4_calcC}*)
+pte_t *walkpgdir(pte_t *l4, const void *va) {
+  $\specline{\textsf{P} \ast \ulcorner\textsf{rtv+KERNBASE}=\textsf{l4}\urcorner \ast \forall_{i\in\textsf{0..511}}\ghostmaptoken{\textsf{id}}{(\textsf{rtv}+i*8)}{\textsf{4}} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}}_{\rtv}$
+  pte_t *l4_entry = &l4[L4Offset(va)]; // Virtual address of L4 entry(*\label{line:start_l4_calcC}*)
   /*At the 4th level (and other levels ignoring the level-specific indices) an access amounts to:*/
-  /*1.Load pml4 and va*/
+  /*1.Load l4 and va*/
   /*2.Extract 9-bit (0x1ff=511) L4 index from the virtual address va*/
-  /*3.Multiply by 8 (byte index), add to pml4 for entry address (*\label{line:end_pml4_calcC}*) */
-  /*4.Store virt. addr. to local var pml4_entry; logical pml4_entry is physical!*/
-  $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{pml4\_entry}\},m) \ast \forall_{i\in\textsf{0..511}\setminus\{\textsf{PML4EX(va)}\}}\ghostmaptoken{\textsf{id}}{(\textsf{rtv}+i*8)}{\textsf{4}}\ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
+  /*3.Multiply by 8 (byte index), add to l4 for entry address (*\label{line:end_l4_calcC}*) */
+  /*4.Store virt. addr. to local var l4_entry; logical l4_entry is physical!*/
+  $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{l4\_entry}\},m) \ast \forall_{i\in\textsf{0..511}\setminus\{\textsf{L4Offset(va)}\}}\ghostmaptoken{\textsf{id}}{(\textsf{rtv}+i*8)}{\textsf{4}}\ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
   $\specline{\textsf{entry\_present(l4e\_val)} \wand \forall_{i\in\textsf{0..511}} \; \ghostmaptoken{\textsf{id}}{\textsf{table\_root(l4e\_val)+ i * 8}}{\textsf{3}} }_{\rtv} $
-  $\specline{ \textsf{pml4\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{pml4\_entry l4e\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l4e\_val})\urcorner}_{\rtv}$
-  pte_t *pdp = pte_get_next_table(pml4_entry); (*\label{line:first_getnext_callC}*) 
-  $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{pml4\_entry}\},m)  \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
-  $\specline{\forall_{i\in\textsf{0..511}} \ghostmaptoken{\textsf{id}}{\textsf{table\_root(l4e\_val'.pfn) + i * 8}}{\textsf{3}}  \ast \ulcorner\textsf{pdp-KERNBASE}=\textsf{table\_root(l4e\_val'.pfn)}\urcorner} $
-  $\specline{ \textsf{pml4\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{pml4\_entry l4e\_val'} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l4e\_val})\urcorner}_{\rtv}$(*\label{line:first_pte_pointstoC}*)
+  $\specline{ \textsf{l4\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l4\_entry l4e\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l4e\_val})\urcorner}_{\rtv}$
+  pte_t *l3 = pte_get_next_table(l4_entry); (*\label{line:first_getnext_callC}*) 
+  $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{l4\_entry}\},m)  \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
+  $\specline{\forall_{i\in\textsf{0..511}} \ghostmaptoken{\textsf{id}}{\textsf{table\_root(l4e\_val'.pfn) + i * 8}}{\textsf{3}}  \ast \ulcorner\textsf{l3-KERNBASE}=\textsf{table\_root(l4e\_val'.pfn)}\urcorner} $
+  $\specline{ \textsf{l4\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l4\_entry l4e\_val'} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l4e\_val})\urcorner}_{\rtv}$(*\label{line:first_pte_pointstoC}*)
   /*pte_get_next_table may have allocated a new page which updates entry*/
-  pte_t *pdp_entry = &pdp[PDPEX(va)]; /* Virtual address of L3 entry*/
-  $\specline{\ulcorner\textsf{pdp}+\textsf{PDPEX(va)*8} = \textsf{pdp\_entry} \land \textsf{table\_root(l4e\_val'.pfn)}  = \textsf{pdp} \urcorner }_{\rtv}$
-  $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{pml4\_entry},\textsf{pdp\_entry}\},m)  \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
-  $\specline{\forall_{i\in\textsf{0..511}\setminus\{\textsf{PDPEX(va)}\}} \; \textsf{table\_root(pml4\_entry}_{\textsf{pfn}}) \textsf{ + i * 8} \mapsto_{\textsf{id}} \textsf{3} }_{\rtv}$
-  $\specline{ \textsf{pml4\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac4}}\textsf{pml4\_entry}\; \textsf{l4e\_val'} \ast \ulcorner  \textsf{qfrac4} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l4e\_val'})\urcorner}_{\rtv}$(*\label{line:ex_l4_vpteC}*)
-  $\specline{ \textsf{pdp\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac3}}\textsf{pdp\_entry}\; \textsf{l3e\_val} \ast \ulcorner  \textsf{qfrac3} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l3e\_val})\urcorner}_{\rtv}$(*\label{line:ex_l3_vpteC}*)
-  $\specline{\textsf{entry\_present(l3e\_val)} \wand \forall_{i\in\textsf{0..511}} \;\ghostmaptoken{\textsf{id}}{\textsf{table\_root(pdp\_entry.pfn) + i * 8}}{ \textsf{2}} } _{\rtv} $
+  pte_t *l3_entry = &l3[L3Offset(va)]; /* Virtual address of L3 entry*/
+  $\specline{\ulcorner\textsf{l3}+\textsf{L3Offset(va)*8} = \textsf{l3\_entry} \land \textsf{table\_root(l4e\_val'.pfn)}  = \textsf{l3} \urcorner }_{\rtv}$
+  $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{l4\_entry},\textsf{l3\_entry}\},m)  \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
+  $\specline{\forall_{i\in\textsf{0..511}\setminus\{\textsf{L3Offset(va)}\}} \; \textsf{table\_root(l4\_entry}_{\textsf{pfn}}) \textsf{ + i * 8} \mapsto_{\textsf{id}} \textsf{3} }_{\rtv}$
+  $\specline{ \textsf{l4\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac4}}\textsf{l4\_entry}\; \textsf{l4e\_val'} \ast \ulcorner  \textsf{qfrac4} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l4e\_val'})\urcorner}_{\rtv}$(*\label{line:ex_l4_vpteC}*)
+  $\specline{ \textsf{l3\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac3}}\textsf{l3\_entry}\; \textsf{l3e\_val} \ast \ulcorner  \textsf{qfrac3} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l3e\_val})\urcorner}_{\rtv}$(*\label{line:ex_l3_vpteC}*)
+  $\specline{\textsf{entry\_present(l3e\_val)} \wand \forall_{i\in\textsf{0..511}} \;\ghostmaptoken{\textsf{id}}{\textsf{table\_root(l3\_entry.pfn) + i * 8}}{ \textsf{2}} } _{\rtv} $
   /*Similar steps & proofs for levels 3 and 2*/
-  pte_t *pd = pte_get_next_table(pdp_entry);
-  pte_t pd_entry = &pd[PDEX(va)];
-  pte_t *pt = pte_get_next_table(pd_entry);
+  pte_t *l2 = pte_get_next_table(l3_entry);
+  pte_t l2_entry = &l2[L2Offset(va)];
+  pte_t *l1 = pte_get_next_table(l2_entry);
   /*Return the address of L1 entry with a similar index computation*/
-  return &pt[PTEX(va)]; 
+  return &l1[L1Offset(va)]; 
 }
@@
 % \begin{lstlisting}[mathescape,escapeinside={(*}{*)}]
-% ;;pte_t *walkpgdir(pte_t *pml4, const void *va) {
+% ;;pte_t *walkpgdir(pte_t *l4, const void *va) {
 % ... ;; Stack setup
-% $\specline{\textsf{P} \ast \ulcorner\textsf{rtv+KERNBASE}=\textsf{pml4}\urcorner \ast \forall_{i\in\textsf{0..511}}\ghostmaptoken{\textsf{id}}{(\textsf{rtv}+i*8)}{\textsf{4}} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}}_{\rtv}$
+% $\specline{\textsf{P} \ast \ulcorner\textsf{rtv+KERNBASE}=\textsf{l4}\urcorner \ast \forall_{i\in\textsf{0..511}}\ghostmaptoken{\textsf{id}}{(\textsf{rtv}+i*8)}{\textsf{4}} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}}_{\rtv}$
 % ;;set up the stack for root address and virtual address    
-% ;;pte_t *pml4_entry = &pml4[PML4EX(va)]; // Virtual address of L4 entry(*\label{line:start_pml4_calc}*)
+% ;;pte_t *l4_entry = &l4[L4EX(va)]; // Virtual address of L4 entry(*\label{line:start_l4_calc}*)
 % mov    -0x8[rbp],rsi
@@
 % shl    0x3,rdi        ;; Multiply by 8
-% add    rdi,rsi        ;; Add to pml4 (virtual) table base(*\label{line:end_pml4_calc}*)
-% mov    rsi,-0x18[rbp] ;; Store to local variable pml4_entry; logical pml4_entry is physical, program variable is virtual
-% $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{pml4\_entry}\},m) \ast \forall_{i\in\textsf{0..511}\setminus\{\textsf{PML4EX(va)}\}}\ghostmaptoken{\textsf{id}}{(\textsf{rtv}+i*8)}{\textsf{4}}\ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
+% add    rdi,rsi        ;; Add to l4 (virtual) table base(*\label{line:end_l4_calc}*)
+% mov    rsi,-0x18[rbp] ;; Store to local variable l4_entry; logical l4_entry is physical, program variable is virtual
+% $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{l4\_entry}\},m) \ast \forall_{i\in\textsf{0..511}\setminus\{\textsf{L4EX(va)}\}}\ghostmaptoken{\textsf{id}}{(\textsf{rtv}+i*8)}{\textsf{4}}\ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
 % $\specline{\textsf{entry\_present(l4e\_val)} \wand \forall_{i\in\textsf{0..511}} \; \ghostmaptoken{\textsf{id}}{\textsf{table\_root(l4e\_val)+ i * 8}}{\textsf{3}} }_{\rtv} $
-% $\specline{ \textsf{pml4\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{pml4\_entry l4e\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l4e\_val})\urcorner}_{\rtv}$
-% ;;pte_t *pdp = pte_get_next_table(pml4_entry);
+% $\specline{ \textsf{l4\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l4\_entry l4e\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l4e\_val})\urcorner}_{\rtv}$
+% ;;pte_t *pdp = pte_get_next_table(l4_entry);
 % mov    -0x18[rbp],rdi
@@
 % ;;save the physical next table address in rax
-% $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{pml4\_entry}\},m)  \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
+% $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{l4\_entry}\},m)  \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
 % $\specline{\forall_{i\in\textsf{0..511}} \ghostmaptoken{\textsf{id}}{\textsf{table\_root(l4e\_val'.pfn) + i * 8}}{\textsf{3}}  \ast \ulcorner\textsf{pdp-KERNBASE}=\textsf{table\_root(l4e\_val'.pfn)}\urcorner} $
-% $\specline{ \textsf{pml4\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{pml4\_entry l4e\_val'} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l4e\_val})\urcorner}_{\rtv}$(*\label{line:first_pte_pointsto}*)
+% $\specline{ \textsf{l4\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l4\_entry l4e\_val'} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l4e\_val})\urcorner}_{\rtv}$(*\label{line:first_pte_pointsto}*)
 % $\specline{\texttt{rax}  \mapsto_{\textsf{r}} \textsf{table\_root(l4e\_val'.pfn)} }_{\rtv}$ ;; pte_get_next_table may have allocated a new page, updating entry
@@
 % $\specline{\ulcorner\textsf{pdp}+\textsf{PDPEX(va)*8} = \textsf{pdp\_entry} \land \textsf{table\_root(l4e\_val'.pfn)}  = \textsf{pdp} \urcorner }_{\rtv}$
-% $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{pml4\_entry},\textsf{pdp\_entry}\},m)  \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
-% $\specline{\forall_{i\in\textsf{0..511}\setminus\{\textsf{PDPEX(va)}\}} \; \textsf{table\_root(pml4\_entry}_{\textsf{pfn}}) \textsf{ + i * 8} \mapsto_{\textsf{id}} \textsf{3} }_{\rtv}$
-% $\specline{ \textsf{pml4\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac4}}\textsf{pml4\_entry}\; \textsf{l4e\_val'} \ast \ulcorner  \textsf{qfrac4} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l4e\_val'})\urcorner}_{\rtv}$(*\label{line:ex_l4_vpte}*)
+% $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{l4\_entry},\textsf{pdp\_entry}\},m)  \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
+% $\specline{\forall_{i\in\textsf{0..511}\setminus\{\textsf{PDPEX(va)}\}} \; \textsf{table\_root(l4\_entry}_{\textsf{pfn}}) \textsf{ + i * 8} \mapsto_{\textsf{id}} \textsf{3} }_{\rtv}$
+% $\specline{ \textsf{l4\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac4}}\textsf{l4\_entry}\; \textsf{l4e\_val'} \ast \ulcorner  \textsf{qfrac4} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l4e\_val'})\urcorner}_{\rtv}$(*\label{line:ex_l4_vpte}*)
 % $\specline{ \textsf{pdp\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac3}}\textsf{pdp\_entry}\; \textsf{l3e\_val} \ast \ulcorner  \textsf{qfrac3} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l3e\_val})\urcorner}_{\rtv}$(*\label{line:ex_l3_vpte}*)
@@
 % ;;pte_t *pt = pte_get_next_table(pd_entry);
-% $\specline{ \left( \begin{array}{l} \textsf{pdp}+\textsf{PDPEX(va)} = \textsf{pdp\_entry} \land \textsf{table\_root(pml4\_entry}_{\textsf{pfn}})  = \textsf{pdp} \\ \textsf{pd}+\textsf{PDEX(va)} = \textsf{pd\_entry} \land \textsf{table\_root(pml3\_entry}_{\textsf{pfn}})  = \textsf{pd} \\ \textsf{pt}+\textsf{PDPEX(va)} = \textsf{pt\_entry} \land \textsf{table\_root(pml2\_entry}_{\textsf{pfn}})  = \textsf{pt} \\  \textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{pml4\_entry},\textsf{pdp\_entry},\textsf{pd\_entry},\textsf{pt\_entry}\},m)  \\ \ghostmaptoken{\delta{}s}{\rtv}{\delta} \\
-% \ghostmaptoken{\textsf{id}}{(\mathsf{pml4\_entry})}{\textsf{4}}\ast \ghostmaptoken{\textsf{id}}{(\mathsf{plm3\_entry})}{\textsf{3}} \ast \ghostmaptoken{\textsf{id}}{(\mathsf{plm2\_entry})}{\textsf{2}}  \\ \forall_{i\in\textsf{0..511}} \; \textsf{table\_root(pml4\_entry}_{\textsf{pfn}}) \textsf{ + i * 8} \mapsto_{\textsf{id}} \textsf{3}  \\ \forall_{i\in\textsf{0..511}} \; \textsf{table\_root(pdp\_entry}_{\textsf{pfn}}) \textsf{ + i * 8} \mapsto_{\textsf{id}} \textsf{2}  \\ \forall_{i\in\textsf{0..511}} \; \textsf{table\_root(pd\_entry}_{\textsf{pfn}}) \textsf{ + i * 8} \mapsto_{\textsf{id}} \textsf{1}  \\  \textsf{pml4\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l4e\_val} \\  \textsf{pdp\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l3e\_val} \\  \textsf{pd\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l2e\_val} \\   \textsf{pt\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l1e\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l1e\_val})\urcorner    \end{array} \right)  =\textsf{R}_{\textsf{walk}}}$
+% $\specline{ \left( \begin{array}{l} \textsf{pdp}+\textsf{PDPEX(va)} = \textsf{pdp\_entry} \land \textsf{table\_root(l4\_entry}_{\textsf{pfn}})  = \textsf{pdp} \\ \textsf{pd}+\textsf{PDEX(va)} = \textsf{pd\_entry} \land \textsf{table\_root(pml3\_entry}_{\textsf{pfn}})  = \textsf{pd} \\ \textsf{pt}+\textsf{PDPEX(va)} = \textsf{pt\_entry} \land \textsf{table\_root(pml2\_entry}_{\textsf{pfn}})  = \textsf{pt} \\  \textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{l4\_entry},\textsf{pdp\_entry},\textsf{pd\_entry},\textsf{pt\_entry}\},m)  \\ \ghostmaptoken{\delta{}s}{\rtv}{\delta} \\
+% \ghostmaptoken{\textsf{id}}{(\mathsf{l4\_entry})}{\textsf{4}}\ast \ghostmaptoken{\textsf{id}}{(\mathsf{plm3\_entry})}{\textsf{3}} \ast \ghostmaptoken{\textsf{id}}{(\mathsf{plm2\_entry})}{\textsf{2}}  \\ \forall_{i\in\textsf{0..511}} \; \textsf{table\_root(l4\_entry}_{\textsf{pfn}}) \textsf{ + i * 8} \mapsto_{\textsf{id}} \textsf{3}  \\ \forall_{i\in\textsf{0..511}} \; \textsf{table\_root(pdp\_entry}_{\textsf{pfn}}) \textsf{ + i * 8} \mapsto_{\textsf{id}} \textsf{2}  \\ \forall_{i\in\textsf{0..511}} \; \textsf{table\_root(pd\_entry}_{\textsf{pfn}}) \textsf{ + i * 8} \mapsto_{\textsf{id}} \textsf{1}  \\  \textsf{l4\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l4e\_val} \\  \textsf{pdp\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l3e\_val} \\  \textsf{pd\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l2e\_val} \\   \textsf{pt\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l1e\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l1e\_val})\urcorner    \end{array} \right)  =\textsf{R}_{\textsf{walk}}}$
 % ;;access and return L1 entry
@@
 \fi
-\vspace{-1em}
 \caption{Walking page-table directory via calls to \textsf{pte\_get\_next\_table} in Figure \ref{fig:calltopteinitializeC}}
@@
 \begin{figure}\footnotesize
-  \vspace{-1em}
-$\begin{array}{l}\left( \begin{array}{l} \ulcorner \textsf{pdp}+\textsf{PDPEX(va)} = \textsf{pdp\_entry} \land \textsf{table\_root(pml4\_entry}_{\textsf{pfn}})  = \textsf{pdp} \\ \land \textsf{pd}+\textsf{PDEX(va)} = \textsf{pd\_entry} \land \textsf{table\_root(pml3\_entry}_{\textsf{pfn}})  = \textsf{pd} \\ \land \textsf{pt}+\textsf{PDPEX(va)} = \textsf{pt\_entry} \land \textsf{table\_root(pml2\_entry}_{\textsf{pfn}})  = \textsf{pt} \urcorner  \ast \\  \textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{pml4\_entry},\textsf{pdp\_entry},\textsf{pd\_entry},\textsf{pt\_entry}\},m)  \\ \ghostmaptoken{\delta{}s}{\rtv}{\delta} \ast
-      \ghostmaptoken{\textsf{id}}{(\mathsf{pml4\_entry})}{\textsf{4}}\ast \ghostmaptoken{\textsf{id}}{(\mathsf{plm3\_entry})}{\textsf{3}} \ast \ghostmaptoken{\textsf{id}}{(\mathsf{plm2\_entry})}{\textsf{2}}  \ast \\  \textsf{pml4\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{pml4\_entry} \; \textsf{l4e\_val}  \ast \\  \textsf{pdp\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{pdp\_entry}\;\textsf{l3e\_val}  \ast \\  \textsf{pd\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{pd\_entry} \; \textsf{l2e\_val}  \end{array} \right)  =\textsf{R}_{\textsf{walk} }
-    \\ \ast \left(\begin{array}{l}  \textsf{pt\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{pt\_entry}\;\textsf{l1e\_val} \ast \\ \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l1e\_val})\urcorner \end{array}\right) =\textsf{R}_{\textsf{l1e}}\end{array} $
+$\begin{array}{l}\left( \begin{array}{l} \ulcorner \textsf{l3}+\textsf{L3Offset(va)} = \textsf{l3\_entry} \land \textsf{table\_root(l4\_entry}_{\textsf{pfn}})  = \textsf{l3} \\ \land \textsf{l2}+\textsf{L2Offset(va)} = \textsf{l2\_entry} \land \textsf{table\_root(l3\_entry}_{\textsf{pfn}})  = \textsf{l2} \\ \land \textsf{l1}+\textsf{L1Offset(va)} = \textsf{l1\_entry} \land \textsf{table\_root(l2\_entry}_{\textsf{pfn}})  = \textsf{l1} \urcorner  \ast \\  \textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{l4\_entry},\textsf{l3\_entry},\textsf{l2\_entry},\textsf{l1\_entry}\},m)  \\ \ghostmaptoken{\delta{}s}{\rtv}{\delta} \ast
+      \ghostmaptoken{\textsf{id}}{(\mathsf{l4\_entry})}{\textsf{4}}\ast \ghostmaptoken{\textsf{id}}{(\mathsf{l3\_entry})}{\textsf{3}} \ast \ghostmaptoken{\textsf{id}}{(\mathsf{l2\_entry})}{\textsf{2}}  \ast \\  \textsf{l4\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l4\_entry} \; \textsf{l4e\_val}  \ast \\  \textsf{l3\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l3\_entry}\;\textsf{l3e\_val}  \ast \\  \textsf{l2\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l2\_entry} \; \textsf{l2e\_val}  \end{array} \right)  =\textsf{R}_{\textsf{walk} }
+    \\ \ast \left(\begin{array}{l}  \textsf{l1\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l1\_entry}\;\textsf{l1e\_val} \ast \\ \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l1e\_val})\urcorner \end{array}\right) =\textsf{R}_{\textsf{l1e}}\end{array} $
 \caption{\textsf{R}$_{\textsf{walk}}$: Resources obtained from invariant for walking page-table directory with \emph{non-existing entries} via calls to \textsf{l1e\_get\_next\_table} in Figure \ref{fig:calltopteinitializeC}}
@@
 The precondition provides knowledge that the virtual base of the L4 is at the appropriate offset from the current \lstinline|cr3| value,
-but does not provide a virtual points-to assertion --- because the function must calculate (Lines \ref{line:start_pml4_calcC}--\ref{line:end_pml4_calcC})
+but does not provide a virtual points-to assertion --- because the function must calculate (Lines \ref{line:start_l4_calcC}--\ref{line:end_l4_calcC})
 which entry it needs access to.
 Instead the precondition has 512 identity map tokens, guaranteeing that every entry on the page is subject to the identity mapping invariant.
-Line \ref{line:end_pml4_calcC} calculates the virtual address of the relevant entry, and the subsequent view shift
+Line \ref{line:end_l4_calcC} calculates the virtual address of the relevant entry, and the subsequent view shift
 pulls that entry out of the identity mapping ($\Xi$) and fetches its corresponding resources as
@@
 to map new pages into an address space by writing into an existing page table via a call\\
-\centerline{\textsf{vaspace\_mappage(pte\_t *pml4, void *va,uintptr\_t fpaddr)}}\\
+\centerline{\textsf{vaspace\_mappage(pte\_t *l4, void *va,uintptr\_t fpaddr)}}\\
 in Figure \ref{fig:mapping_codeC}.
@@
 */
-void vaspace_mappage(pte_t *pml4, void *va,uintptr_t fpaddr ) {
-  $\specline{\textsf{P} \ast \textsf{pml4} \mapsto_{\textsf{id}} \textsf{\_} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta} \ast \ulcorner\theta \; !!\; \vaddr = \texttt{None}\urcorner}_{\rtv}$
-  pte_t *pteaddr = walkpgdir(pml4, a); (*\label{line:call_walkpgdirC}*)
+void vaspace_mappage(pte_t *l4, void *va,uintptr_t fpaddr ) {
+  $\specline{\textsf{P} \ast \textsf{l4} \mapsto_{\textsf{id}} \textsf{\_} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta} \ast \ulcorner\theta \; !!\; \vaddr = \texttt{None}\urcorner}_{\rtv}$
+  pte_t *pteaddr = walkpgdir(l4, va); (*\label{line:call_walkpgdirC}*)
   $\specline{ \textsf{R}_{\textsf{walk}} \ast \textsf{R}_{\textsf{l1e}} }$
@@
     pteaddr->pfn = PTE_ADDR_TO_PFN(fpaddr); /*Store updated entry back to L1 entry (*\label{line:l1entry_storeC}*)*/
-    $\specline{ \textsf{pt\_entry+KERNBASE} \mapsto_{\textsf{vpte,1}} \textsf{pt\_entry} \; (\textsf{set\_pfn}(\textsf{l1e\_val}, \textsf{PTE\_ADDR\_TO\_PFN}(\textsf{fpaddr}))) \ast  \ulcorner \lnot(\textsf{entry\_present l1e\_val}) \urcorner \ast \textsf{P}}$
+    $\specline{ \textsf{l1\_entry+KERNBASE} \mapsto_{\textsf{vpte,1}} \textsf{l1\_entry} \; (\textsf{set\_pfn}(\textsf{l1e\_val}, \textsf{PTE\_ADDR\_TO\_PFN}(\textsf{fpaddr}))) \ast  \ulcorner \lnot(\textsf{entry\_present l1e\_val}) \urcorner \ast \textsf{P}}$
     /*Set the present bit in entry*/
@@
   return;
-  $\specline{ \begin{array}{l}\textsf{pt\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{pt\_entry} \; \textsf{l1e\_val} \ast \ulcorner (\textsf{entry\_present l1e\_val}) \urcorner \ast \textsf{P} \\
- \lor \; \;  \textsf{pt\_entry+KERNBASE} \mapsto_{\textsf{vpte,1}} \textsf{pt\_entry} \; (\textsf{set\_pfn}(\textsf{l1e\_val}, \textsf{PTE\_ADDR\_TO\_PFN}(\textsf{fpaddr}))) \ast  \ulcorner \lnot(\textsf{entry\_present l1e\_val}) \urcorner \ast P \end{array} }$
+  $\specline{ \begin{array}{l}\textsf{l1\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l1\_entry} \; \textsf{l1e\_val} \ast \ulcorner (\textsf{entry\_present l1e\_val}) \urcorner \ast \textsf{P} \\
+ \lor \; \;  \textsf{l1\_entry+KERNBASE} \mapsto_{\textsf{vpte,1}} \textsf{l1\_entry} \; (\textsf{set\_pfn}(\textsf{l1e\_val}, \textsf{PTE\_ADDR\_TO\_PFN}(\textsf{fpaddr}))) \ast  \ulcorner \lnot(\textsf{entry\_present l1e\_val}) \urcorner \ast P \end{array} }$
 }
@@
 \vspace{-1em}
-  \caption{Specification of updating L1 entry (\textsf{pt\_entry}) to reference a new page (\textsf{fpaddr}).}
+  \caption{Specification of updating L1 entry to reference a new page (\textsf{fpaddr}).}
 \label{fig:mapping_codeC}

### 2025-07-17 examples_C.tex
+

@@
 fetched from the appropriate entry in the level $n+1$ table.
-This is the role of \lstinline|pte_get_next_table| (Figures \ref{fig:calltopteinitializeC}):
+This is the role of \lstinline|pte_get_next_table| (Figure \ref{fig:calltopteinitializeC}):
 %and \ref{fig:p2vC}):
@@
 \ifPLDI
-Line \ref{line:mask_present} checks % In the condensed figure, it's all on one line
+Line \ref{line:conditional_childrenC} checks % In the condensed figure, it's all on one line
 \else
-Lines \ref{line:mask_presentC}--\ref{line:check_entry_presentC} check
+Lines \ref{line:mask_present}--\ref{line:check_entry_present} check
 \fi
 if the entry's ``present'' bit is set.
-If it is zero, a new page must be allocated for the next level of the table --- which is done by the fall-through
-from Line \ref{line:check_entry_present_jump}'s conditional jump. Otherwise the code jumps ahead to
-the case for the next level already existing, which is discussed in Section \ref{sec:p2vC} and Figure \ref{fig:p2vC}.
+If it is zero, a new page must be allocated for the next level of the table. Otherwise the code jumps ahead to
+the case for the next level already existing (Lines \ref{line:finalpieceS}-\ref{line:finalpieceE} in Figure \ref{fig:calltopteinitializeC}), which is discussed in Section \ref{sec:p2vC}.
 First, we must discuss another refinement of the address space invariant, establishing
@@
   $\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present entry\_val})\urcorner}_{\rtv}$ (*\label{line:get_next_vpte_preconditionC}*)
-  $\specline{\ulcorner\textsf{entry\_present(entry\_val)}\urcorner \wand \forall_{i\in\textsf{0..511}} \; \ghostmaptoken{\textsf{id}}{\textsf{table\_root(entry\_val.pfn) + i * 8}}{\textsf{v-1}} }_{\rtv} $ (*\label{line:conditional_childrenC}*)
+  $\specline{\ulcorner\textsf{entry\_present(entry\_val)}\urcorner \wand \forall_{i\in\textsf{0..511}} \; \ghostmaptoken{\textsf{id}}{\textsf{table\_root(entry\_val.pfn) + i * 8}}{\textsf{v-1}} }_{\rtv} $
   /*Reading page table entry justified by virtual pte-pointsto*/
-  if (!entry->present(*\label{line:read_entry_contentsC}*)) {
+  if (!entry->present(*\label{line:read_entry_contentsC}*)(*\label{line:conditional_childrenC}*)\label{line:check_entry_present_jumpC}) {
     /* Jump if the present bit is not zero; no need to allocate next level (* \label{line:check_entry_presentC} *)(* \label{line:mask_presentC} *) */
@@
     $\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \land \lnot(\textsf{entry\_present entry\_val})  \urcorner }_{\rtv}$ (*\label{line:after_concluding_qfrac1C}*)
-    pte_initialize(entry);(* \label{line:call_to_pte_initializeC} *)
+    pte_initialize(entry);(* \label{line:call_to_pte_initializeC}(*\label{line:page_of_capsC}*) *)
     $\specline{ \ldots \ast \ulcorner  \textsf{qfrac} = 1 \land \lnot(\textsf{entry\_present entry\_val})  \urcorner \ast \ulcorner  \textsf{qfrac} = 1 \land \lnot(\textsf{entry\_present (pfn\_set (entry\_val nextpaddr))}  \urcorner   }_{\rtv}$
@@
 \lstinline|pte_get_next_table|'s first call (Line \ref{line:first_getnext_callC}) returns
-the virtual address of the base of the L3 table (a \emph{page directory pointer}, so PDP, in official
-x86-64 terminology). Then the situation to move from that pointer to the base of the L2
+the virtual address of the base of the L3 table. Then the situation to move from that pointer to the base of the L2
 is just like the process just followed: the proof calculates the address of the relevant

### 2025-07-17 examples_C.tex
+

@@
   /*Reading page table entry justified by virtual pte-pointsto*/
-  if (!entry->present(*\label{line:read_entry_contentsC}*)(*\label{line:conditional_childrenC}*)\label{line:check_entry_present_jumpC}) {
+  if (!entry->present(*\label{line:read_entry_contentsC}*)(*\label{line:conditional_childrenC}*)(*\label{line:check_entry_present_jumpC}*)) {
     /* Jump if the present bit is not zero; no need to allocate next level (* \label{line:check_entry_presentC} *)(* \label{line:mask_presentC} *) */
@@
 \ifPLDI
-Line \ref{line:mask_present} % in condensed version, all on same line
+Line \ref{line:mask_presentC} % in condensed version, all on same line
 \else
@@
 %\end{figure}
-Figure \ref{fig:p2vC} shows the verification of the end of \lstinline|pte_get_next_table| specialized to the case where 
+Lines \ref{childrenC}-\ref{line:p2vC} in Figure \ref{fig:calltopteinitializeC} show the verification of the end of \lstinline|pte_get_next_table| specialized to the case where 
 where no allocation was necessary (i.e., the conditional on Line \ref{line:check_entry_presentC} of Figure \ref{fig:calltopteinitializeC} was taken).
@@
 the existence of a memory location for an appropriate L1 table entry.
-The code uses a helper function \lstinline{walkpgdir} (discussed again in Section \ref{sec:traversingC}).
+The code uses a helper function \lstinline{walkpgdir} (discussed again in Section \ref{wlkpgdirC}).
 \textsf{walkpgdir}'s postcondition contains virtual \emph{PTE} pointsto assertions ($\mapsto_{\textsf{vpte}}$)

### 2025-07-18 programlogic.tex
finish section 4 comments

@@
 
-\add{Since \textsf{vProp}s are ``just'' \textsc{Iris} predicates on a choice of page table root,
-it may not be obvious what the gain is over simply using \textsc{Iris} directly.
-The gain is not expressive power.
-Every modal logic can be translated straitforwardly to a logic without modalities
-(nearly always a small variation on the \emph{standard translation} from propositional modal logic
-to first-order logic~\cite{blackburn2006modal}\footnote{\add{Itself similar to verification condition generation
-for weakest precondition calculi and Hoare logics.}}) that can directly address the logic's models ---
-but these translations, in addition to making specifications far more verbose,
-generally do not preserve the appealing reasoning principles of modal logics.
-This becomes most apparent when discussing our proof rules.
-}
-
 % We derive a program logic (a separation logic) supporting the following stances and constraints:
@@
 carry over to our derived logic.
+\add{Since \textsf{vProp}s are ``just'' \textsc{Iris} predicates on a choice of page table root,
+it may not be obvious what the gain is over simply using \textsc{Iris} directly.
+The gain is not expressive power.
+Every modal logic can be translated straitforwardly to a logic without modalities
+(nearly always a small variation on the \emph{standard translation} from propositional modal logic
+to first-order logic~\cite{blackburn2006modal}\footnote{\add{Itself similar to verification condition generation
+for weakest precondition calculi and Hoare logics.}}) that can directly address the logic's models ---
+but these translations, in addition to making specifications far more verbose,
+generally do not preserve the appealing reasoning principles of modal logics.
+This becomes most apparent when discussing our proof rules.
+}
+
 However, we must still build up from existing \iris primitives to provide new primitives that depend on the address space --- primarily the notion
 of virtual points-to.
-
-% \begin{figure}[t]
-% \begin{lstlisting}[language=Coq]
-%   Definition vProp  $\Sigma$ : bi := word 64 -b> iPropI  $\Sigma$.
-%   \end{lstlisting}
-% \caption{$\textsf{vProp }\Sigma$: Root-Address Indexed Address-Space Proposition}
-%   \label{fig:vprop}
-% \end{figure}
-
-% \subsection{Base Points-To Assertions}
-% \label{sec:pointsto}
-% We build up our context-dependent assertions from two basic types of assertions that do not care about address spaces.
 To define and use virtual points-to assertions, we require two basic assertions that ignore
@@
 \vspace{-1em}
-\caption{A Strong Virtual Points-to Relation}
+\caption{A Strong Virtual Points-to Relation
+\todo[inline]{Ismail: Tie these lines to the steps and entries in Figure \ref{fig:pagetables}, e.g., with \textcircled{1}\ldots}
+}
   \label{fig:strongvirtualpointsto}
@@
 \emph{which} locations are accessed in a page table walk, as in Figure \ref{fig:strongvirtualpointsto}.
-% Figure \ref{fig:strongvirtualpointsto} gives such a definition.
 It asserts the existence of four page table entries, one at each translation level, and via \lstinline|L4_L1_PointsTo|
 asserts that the physical page table walk (per Figure \ref{fig:pagetables}) succeeds in reaching the L1 entry,
-% asserts from the current \lstinline|cr3| value, taking the offsets from \textsf{va} (as in Figure \ref{fig:pagetables})
-% to index each level, finding the existentially quantified entries in each level, 
-% ending with the L1 entry, 
-% whose physical page base is added to the page offset of \textsf{va} to obtain the physical address
- % --- and there (\lstinline|addr_L1(va,l1e)|) in physical memory, is the value.
 which points to the page holding the physical memory backing the virtual address, which contains value \textsf{v}.
@@
 
+\todo[inline]{This is a section where it's worth explaining the figure in a little more detail, especially as we reuse L4L1PointsTo
+in our own definitions later. I know we discussed at some point, but don't the \texttt{l$n$M52} accessors ignore the \texttt{maddr},
+and don't the \texttt{l$n$off} accessors ignore the entry arguments? If so we can drop those (simplifying the figure), and annotate 
+(e.g., with comments like \lstinline[mathescape]|(*$\textcircled{E}$*)|) each \texttt{l$n$off} in Figure \ref{fig:strongvirtualpointsto}
+with a tag matching a new tag on the corresponding offset chunk of the translated virtual addres in Figure \ref{fig:pagetables}\ldots
+}
 \lstinline|L4_L1_PointsTo| works by
@@
 }
-Our solution removes some restrictions in this definition by further abstracting the handling of address translation.
+Our solution \replace{removes}{in the next subsection improves on theirs by removing} some restrictions in this definition by further abstracting the handling of address translation.
 
@@
 
-  % Updating the virtual-to-physical (ghost) mapping for a virtual address $va$
-  % requires full ownership of the token. 
   The \emph{virtual memory manager's invariant} ensures that for each $\ghostmaptoken{\gamma}{\vaddr}{\paddr}$ mapping
@@
   will resolve in the hardware to $\paddr$ --- via $\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$.
-  % This allows modification of the page tables themselves as long as the changes do not invalidate overall virtual-to-physical
-  % translation.
-  % Consistency follows from replacing the explicit page table walk of Figure \ref{fig:strongvirtualpointsto}'s
-  % definition with ownership of the ghost map token for the relevant ghost map
-  % entry, as in Figure \ref{fig:virtualpointstosharing}.
-  % This allows the overall translation to be updated separately from the virtual points-to.
+  \add{This kernel invariant turns out to be a key ingredient in supporting proofs of VMM functionality:
+  in Section \ref{sec:experiment} we will see that separating the logical and physical virtual-to-physical mappings
+  is what allows stating the global kernel invariants needed for software page traversals, which prior work did not
+  (and could not) pursue.}
   \looseness=-1
@@
 
-% Figure \ref{fig:peraspaceinvariant} describes the resources for each address space.
+\todo[inline]{This next paragraph below explains details, but should first explain the big picture: the current address space
+is identified by a paddr, so to state the invariant for the address space named by that paddr we need to look up
+what invariants should hold, then assert that those invariants do hold.}
 In Figure \ref{fig:peraspaceinvariant},
@@
 
+\todo[inline]{What's not clear in this next paragraph is the relationship between ASpaceLookup in Figure 5 and the lookup part of Figure 3.
+I think Figure 5's $m$ and Figure 3's $\delta{s}$ are the same, aren't they? If so we should say so and use the same name.
+If not we should explain why they're different.
+}
 Virtual points-to assertions must also ensure that they are tied to a valid address space, and because they
@@
 multiple address spaces, and working out how to support interaction of multiple address spaces (discussed in the next section).
-% This simple definition is 
-% what we need in order for assertions in one address space to talk about assertions in another.
-% The inspiration for this assertion comes from hybrid logic~\cite{blackburn1995hybrid,areces2001hybrid,goranko1996hierarchies,gargov1993modal},
-% which as mentioned earlier uses as modalities the names of worlds in a Kripke model to state
-% that an assertion is true in that specific world.
-% Conceptually the address space rooted at some physical address (intended as the start of an L4 table)
-% is a distinct world, and it makes sense to name it by referring to the root of the
-% supporting page table structure. However, rather than explicitly building a Kripke model for our
-% assertions, we inherit them by lifting the model of our ambient logic, Iris.
+\add{The modal assertions, together with the other-space modality, mean we can give generic definitions
+of data structure assertions (e.g., linked lists, etc.) which do not need to track information
+about their own address space. In fact, \emph{only} assertions that explicitly deal with multiple address
+spaces need to mention address spaces at all (via the other-space modality).
+}
+
 We can prove that this modality follows certain basic laws: that its truth is independent of the address
@@
 % Per the discussion in Section \ref{sec:issues}, w
-As common for assembly-level verification, we define our logic using Hoare doubles:%\footnote{This
+As common for assembly-level verification~\add{\cite{Ni2006codeptrs,ni2007contexts}}, we define our logic using Hoare doubles:%\footnote{This
 % omits some low-level Iris details (stuckness, observations) that play no meaningful
@@
 \end{mathpar}
+\todo[inline]{Ismail, what do you think about shading the ``core'' assertions that are common to any asm program logic?
+Or shading the VMM-specific ones?
+}
 \vspace{-1em}
@@
 which can be ensured by what we informally call the page-table-walk points-to collection
-($\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$ in Figure \ref{fig:strongvirtualpointsto}). }
+($\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$ in Figure ??). }
 \TirNameStyle{WriteToRegFromVirtMem} and \TirNameStyle{WriteToVirtMemFromReg}
@@
 Our rules for control transfers (\lstinline|jne|, \lstinline|call|, and \lstinline|ret|) are currently
-axiomatized (with completely standard specifications~\cite{ni2007contexts,Chlipala2013Bedrock}) 
+axiomatized (with completely standard specifications~\cite{ni2007contexts,Chlipala2013Bedrock})\footnote{
+  \add{An assertion that code at some address is safe to call with a given precondition~\cite{Ni2006codeptrs}
+  asserts that the address is mapped and that memory at that address decodes to an instruction sequence
+  that is safe with that precondition.
+  }
+} 
 because \iris's built-in machinery does not provide

### 2025-07-18 background_C.tex

@@
 \begin{lstlisting}[mathescape,escapeinside={(*}{*)},basicstyle=\footnotesize]
-/*
-@param entry: entry whose frame address is used 
-              to locate the subsequent entry in the page-tree
-@return next: return a virtual pointer to the descendant tree
-*/
 pte_t *pte_get_next_table(pte_t *entry) {

### 2025-07-18 examples_C.tex

@@
 discussed in Section \ref{subsec:identitymappingsC} and Figure \ref{fig:peraspaceinvariant_with_p2v_extensionC}.
-in the last lines of \lstinline|pte_get_next_table| shown in Figure \ref{fig:p2vC} (the continuation of Figure \ref{fig:calltopteinitializeC}).
+in the last lines of \lstinline|pte_get_next_table| shown in Line \ref{line:p2vC} in Figure \ref{fig:calltopteinitializeC} (the continuation of Figure \ref{fig:calltopteinitializeC}).
 This is a critical piece of the full page table walk verification.
-In our small kernel (Line \ref{line:p2vC} in Figure \ref{fig:p2vC}), as in larger kernels, the C macro \texttt{P2V} common to many kernels
+In our small kernel (Line \ref{line:p2vC} in Figure \ref{fig:calltopteinitializeC}), as in larger kernels, the C macro \texttt{P2V} common to many kernels
 is actually just addition by the constant offset mentioned in Section \ref{subsec:identitymappingsC}.
@@
 %\end{figure}
-Lines \ref{childrenC}-\ref{line:p2vC} in Figure \ref{fig:calltopteinitializeC} show the verification of the end of \lstinline|pte_get_next_table| specialized to the case where 
+Lines \ref{line:childrenC}-\ref{line:p2vC} in Figure \ref{fig:calltopteinitializeC} show the verification of the end of \lstinline|pte_get_next_table| specialized to the case where 
 where no allocation was necessary (i.e., the conditional on Line \ref{line:check_entry_presentC} of Figure \ref{fig:calltopteinitializeC} was taken).
 In this case, the true present bit allows access to the child tokens from Line \ref{line:conditional_childrenC} of Figure \ref{fig:calltopteinitializeC},
-which is then refined to the assertion on Line \ref{line:childrenC} of Figure \ref{fig:p2vC}.
-The code loads \lstinline|rcx| with the offset value \textsf{KERNBASE}, which gives us the value of the virtual address ($\textsf{entry}_{\textsf{pfn}}$ \textsf{+KERNBASE})
+which is then refined to the assertion on Line \ref{line:childrenC} of Figure \ref{fig:calltopteinitializeC}.
+The code loads the value at the offseted (\textsf{KERNBASE}) virtual memory address ($\textsf{entry}_{\textsf{pfn}}$ \textsf{+KERNBASE})
 of the \emph{base} of the next level of the page table.

### 2025-07-19 background.tex
fig

@@
 address, but tagged with the physical data addresses, so cache lookups and address translations can proceed in parallel.} and/or memory-bus.
+
 \setlength{\columnseprule}{4pt}

### 2025-07-19 background.tex
background error check

@@
 address, but tagged with the physical data addresses, so cache lookups and address translations can proceed in parallel.} and/or memory-bus.
-
 \setlength{\columnseprule}{4pt}
@@
 \end{minipage}
+\hfill\hfill
 \begin{minipage}{.4\textwidth}

### 2025-07-19 syntaxsemantics.tex
synsem check

@@
 \label{sec:syntax}
-To develop our core logical ideas, we instantiate \iris with a simple language for streams of instructions, 
-and a logical machine model corresponding to execution of x86-64 assembly instructions with virtual memory enabled on the 
+To develop our core logical ideas, we instantiate \iris with a simple language for a stream of instructions
+and a logical machine model corresponding to the execution of x86-64 assembly instructions with virtual memory enabled on the
 CPU.
@@
 % Programs we demonstrate in this paper requires accessing two types of computer resource: registers and memory.
-A register identifier, $\reg$, is chosen from a fixed finite set of register identifiers, $\regset$.\footnote{\add{$\reg$ abbreviates \emph{general register}, which includes integer registers (e.g., \lstinline|rax|),
+A register identifier, $\reg$, is chosen from a fixed finite set of register identifiers, $\regset$.
+\footnote{\add{$\reg$ abbreviates \emph{general register}, which includes integer registers (e.g., \lstinline|rax|),
 control registers (e.g., \lstinline|cr3|), and segmentation registers (which we do not discuss in detail,
@@
 
-The most important instructions we model are memory-accessing variants of the x86-64 \lstinline|mov| instruction, which we format in Intel syntax
+The most important instructions that we model are memory-accessing variants of the x86-64 \lstinline|mov| instruction, which we format in Intel syntax
 (destination on the left, source on the right).
-Thus a store to memory is $\textsf{mov}~[r_m]~r_r$, and a load from memory is $\textsf{mov}~r_r~[r_m]$.
+Thus, a store to the memory is $\textsf{mov}~[r_m]~r_r$, and a load from memory is $\textsf{mov}~r_r~[r_m]$.
 % \[
@@
 %   \]
-Operationally, each first translates the virtual address stored in register $r_m$ to a physical address \add{(a page table walk)},
+Operationally, each first translates the virtual address stored in the register $r_m$ to a physical address \add{(a page table walk)},
 then either updates the memory at that physical location with the contents of $r_r$ (for store)

### 2025-07-20 programlogic.tex
+

@@
 are not tagged with their address spaces in any way. Memory access in this logic is validated through the use
-of virtual points-to assertions in pre-conditions, which guarantee that address translations succeed.
-This supports rules for updating not only typical data in memory that happens to be subject to address translation, but
-manipulation of the page tables themselves via virtual addresses (as demanded by all modern hardware), also
+of virtual points-to assertions in preconditions, which guarantee that address translations succeed.
+This supports rules for updating not only typical data in memory that happens to be subject to address translation but also
+manipulation of the page tables themselves via virtual addresses (as demanded by all modern hardware) and also
 via virtual points-to assertions.
 To support specifications that deal with multiple address spaces, our logic incorporates a hybrid-style modality
-$[r](P)$ to state that an assertion is true in another (assertion-specified) address space rather than the address space
-currently active in hardware, which is not only useful for virtual memory manager invariants but critical to reasoning
-about change-of-address-space.
+$[r](P)$ to state that an assertion is true in another address space (assertion-specified) rather than the address space
+currently active in hardware, which is not only useful for virtual memory manager invariants, but also critical to reasoning
+about change of address space.
 By developing this within the \iris framework, we obtain additional features (e.g., fractional permissions) that allow us to verify
@@
   \iris experts may notice our \lstinline|-b>| resembles another pointwise lifting already in  \iris~\cite{dang2019rustbelt,dang2022compass}. 
-  This similarity is real, but the existing lifting does not appear to work with indexed \coq types
-  like our \lstinline|word n| as a domain.
+  This similarity is real, but the existing lifting does not appear to work with indexed \coq types like our \lstinline|word n| as a domain.
 }
@@
 This is the (\rocq) type of assertions in our logic.
-Most constructs in \iris's base logic are defined with respect to any BI-algebra (of \coq type \lstinline|bi|), so automatically
+Most constructs in \iris's base logic are defined with respect to any BI-algebra (of \coq type \lstinline|bi|), so they automatically
 carry over to our derived logic.
@@
 The gain is not expressive power.
-Every modal logic can be translated straitforwardly to a logic without modalities
+Every modal logic can be translated straightforwardly to a logic without modalities
 (nearly always a small variation on the \emph{standard translation} from propositional modal logic
@@
 generally do not preserve the appealing reasoning principles of modal logics.
-This becomes most apparent when discussing our proof rules.
+This becomes more apparent when discussing our proof rules.
 }
@@
 \paragraph{Register points-to} 
-The assertion $\textsf{r}\;\mapsto_{\textsf{r}}\{q\}\;\textsf{rv}$ ensures the ownership of the register $\rg$ containing the register value $\rv$. 
-The fraction $\qfrac$ with value 1 asserts the unique ownership of the register mapping, and grants update permission on it, 
-otherwise, any value $0 < \qfrac <1$ represents partial ownership granting readonly permission on the mapping.
+The assertion $\textsf{r}\;\mapsto_{\textsf{r}}\{q\}\;\textsf{rv}$ ensures the ownership of the register $\rg$ containing the value of the register $\rv$. 
+The fraction $\qfrac$ with value 1 asserts the unique ownership of the register mapping and grants update permission to it;
+otherwise, any value $0 < \qfrac <1$ represents partial ownership, granting read-only permission on the mapping.
 
@@
 \emph{which} locations are accessed in a page table walk, as in Figure \ref{fig:strongvirtualpointsto}.
-It asserts the existence of four page table entries, one at each translation level, and via \lstinline|L4_L1_PointsTo|
+It asserts the existence of four page-table entries, one at each translation level, and via \lstinline|L4_L1_PointsTo|
 asserts that the physical page table walk (per Figure \ref{fig:pagetables}) succeeds in reaching the L1 entry,
@@
 chaining together the entries for each level, using the sequence of table offsets from the address being translated to index
-each table level, and using the physical page address embedded in each entry.\footnote{
+each table level, and using the physical page address embedded in each entry. \footnote{
   The fractions \lstinline|q1| through \lstinline|q4| represent the fractional ownership of each entry based on how many
@@
 }
-For example, 
-  the first level address translation to get the L4 entry (\lstinline|l4e|) 
-  uses the masks \textsf{l4M52} with the current \lstinline|cr3| to get the 
-  physical address of the start of the L4 table,
+For example, the first-level address translation to get the L4 entry (\lstinline|l4e|) 
+  uses the masks \textsf{l4M52} with the current \lstinline|cr3| to get the physical address of the start of the L4 table
   and \textsf{l4off} with the virtual address being translated to compute the correct offset within that table.
-  Then at that physical location is the appropriate entry in the L4 table \textsf{l4e}.
+  Then, that physical location is the appropriate entry in the L4 table \textsf{l4e}.
   Subsequent levels of the page table walk work similarly.
 The statement of these assertions is simplified by the use of our split physical points-to assertions, since
-each level of tables is page-sized.\footnote{We do not address superpages and hugepages in this paper.}
-This helper definition is also a more explicit \textsf{vProp}, explicitly binding a value for \lstinline|cr3| and using it
-to start the translation process.
+each level of tables is page-sized. \footnote{We do not address superpages and hugepages in this paper.}
+This helper definition is also more explicit in \textsf{vProp} which binds a value to \lstinline|cr3| and uses it to start the translation process.
 \looseness=-1
@@
   The virtual points-to definition shown in Figure \ref{fig:strongvirtualpointsto} 
-  is too strong to specify some operations a virtual memory manager may need to do, such as
-  move one level of the page table to a different physical location while preserving all virtual-to-physical mappings. %\footnote{
+  is too strong to specify some operations that a virtual memory manager may need to do, such as move one level of the page table to a different physical location while preserving all virtual-to-physical mappings.
+%\footnote{
   %   x86-64 hardware, like other architectures, includes a feature (which we do not formalize assertions for) to
@@
   The use of $\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$ in Figure \ref{fig:strongvirtualpointsto}'s
-  virtual points-to definition'
-   stores knowledge of the page table walk's details with ownership of the backing physical
-  memory.
-  Updating any of these mappings (e.g., moving the page tables in physical memory, as in coallescing for
-  superpages or hugepages)
+  virtual points-to definition stores knowledge of the page table walk details with ownership of the backing physical memory.
+  Updating any of these mappings (e.g., moving the page tables in physical memory, as in coalescing for superpages or hugepages)
   would require explicitly collecting all virtual points-to facts that traverse affected entries.
@@
   so long as those updates preserve the same virtual-to-physical translations.
-  But this is not possible with Figure \ref{fig:strongvirtualpointsto}'s definition, which ties ownership of particular
-  pieces of page table memory to the virtual points-to.
+  But this is not possible with Figure \ref{fig:strongvirtualpointsto}'s definition, which ties ownership of particular pieces of page table memory to the virtual points-to.
 
@@
 
-  We separate the physical page-table-walk  
-  from the virtual pointsto relation, replacing it with ghost state that merely guarantees the address translation would succeed.
-  \iris includes a \emph{ghost map} construction which we use to track mappings from
-  virtual addresses to the physical addresses they translate to as a piece of ghost state.
-  The map includes for each key in the map (i.e.,
-  each virtual address) a token $\ghostmaptoken{\gamma}{k}{v}$ whose ownership is required to update that key-value pair in the ghost map
-  named $\gamma$. The existence of such a token implies that the actual map $\theta$ tracked by a corresponding $\mathsf{GhostMap}(\gamma,\theta)$
+  We separate the physical page-table walk from the virtual points-to relation, replacing it with a ghost state that merely guarantees that the address translation would succeed.
+  \iris includes a \emph{ghost map} construction, which we use to track mappings from virtual addresses to the physical addresses they translate to as a piece of ghost state.
+  The map includes, for each key in the map (i.e.,
+  each virtual address), a token $\ghostmaptoken{\gamma}{k}{v}$ whose ownership is required to update that key-value pair in the ghost map named $\gamma$.
+The existence of such a token implies that the actual map $\theta$ tracked by a corresponding $\mathsf{GhostMap}(\gamma,\theta)$
   resource indeed maps $k$ to $v$. These properties are captured by key \iris rules in Figure \ref{fig:ghostmaps}.\footnote{\iris ghost maps lack established notation in the literature,
@@
   \textsc{GhostMapUpdate} says that ownership of the actual ghost map with ghost name $\gamma$ and map contents $\theta$,
-  and a full (non-fractional) token witnessing that $\theta$ maps \textsf{pa} to \textsf{va} permits an update to the ghost map's state,
+  and a full (nonfractional) token witnessing that $\theta$ maps \textsf{pa} to \textsf{va} permits an update to the ghost map's state,
   changing the map and replacing the token to represent the new value.
@@
 
-  The \emph{virtual memory manager's invariant} ensures that for each $\ghostmaptoken{\gamma}{\vaddr}{\paddr}$ mapping
-  in this map, there are \emph{physical} resources sufficient to ensure that address translation for $\vaddr$
-  will resolve in the hardware to $\paddr$ --- via $\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$.
+  The \emph{virtual memory manager's invariant} ensures that for each $\ghostmaptoken{\gamma}{\vaddr}{\paddr}$ mapping in this map, there are \emph{physical} resources sufficient to ensure that the address translation for $\vaddr$
+will resolve on the hardware to $\paddr$ --- via $\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$.
   \add{This kernel invariant turns out to be a key ingredient in supporting proofs of VMM functionality:
   in Section \ref{sec:experiment} we will see that separating the logical and physical virtual-to-physical mappings
-  is what allows stating the global kernel invariants needed for software page traversals, which prior work did not
-  (and could not) pursue.}
+  is what allows stating the global kernel invariants needed for software page traversals, which prior work did not (and could not) pursue.}
   \looseness=-1
@@
   % \[\mathcal{A}\textsf{bsPTableWalk} \stackrel{\triangle}{=} \textsc{Auth} (\; \mathcal{W}_{64} \;\rightarrow_{\textrm{fin}} \;  ( (\textsc{Frac }, \mathord{+}) \times (\textsc{Agree } \Loc,\mathord{=}) ))\]
-For clarity we refer to the specific ghost map summarizing virtual-to-physical translations by \\
+For clarity, we refer to the specific ghost map summarizing virtual-to-physical translations by \\
 \centerline{$\mathcal{A}\textsf{bsPTableWalk}(\delta,\theta) \stackrel{\triangle}{=} \mathsf{GhostMap}(\delta,\theta)$}\\
@@
 and keep this in a per-address-space invariant described shortly.
-We then replace the physical traversal
- $\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$ in Figure \ref{fig:strongvirtualpointsto}'s virtual points-to definition
+We then replace the physical traversal $\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$ in Figure \ref{fig:strongvirtualpointsto}'s virtual points-to definition
 with ownership of the token \ghostmaptoken{\delta}{\vaddr}{\paddr}, %ghost-map ($ \sumwalkabs\vaddr\qfrac\paddr$),
@@
 This new definition guarantees that the ghost map contains an mapping from the virtual address ($\vaddr$) to a physical address ($\paddr$),
-and thus that the per-address-space invariant will contain the physical resources that guarantee the hardware resolves the translation.
+and thus that the per-address-space invariant will contain the physical resources that guarantee that the hardware resolves the translation.
 
@@
 When combined with the
-fragment stored in the virtual points-to (Figure \ref{fig:virtualpointstosharing}),
-changes in the virtual-to-physical translations become possible without needing to track all virtual points-to assertions.
+fragment stored in virtual points-to (Figure \ref{fig:virtualpointstosharing}),
+changes in the virtual-to-physical translations become possible without the need to track all virtual points-to assertions.
 
@@
 $\theta$ is the logical map from virtual addresses to the physical addresses they should translate to, 
-for the \emph{current} address space, corresponding to the currently installed page tables (the ones
-indicated by \lstinline|cr3|). In Figure \ref{fig:virtualpointstosharing}, the existentially-quantified $\delta$
+for the \emph{current} address space, corresponding to the page tables currently installed (the ones
+indicated by \lstinline|cr3|). In Figure \ref{fig:virtualpointstosharing}, the existentially quantified $\delta$
 is the ghost name for this map (again, for the current address space).
@@
 it must ensure that $\theta$ actually corresponds to the intended ghost map for the current address space.
-This requires looking up the ghost name (i.e., $\delta$) for the current address space,
+This requires looking up the ghost name (i.e. $\delta$) for the current address space,
 which is handled by $m$ --- another logical map (tied to ghost state shortly)
@@
 \todo[inline]{What's not clear in this next paragraph is the relationship between ASpaceLookup in Figure 5 and the lookup part of Figure 3.
-I think Figure 5's $m$ and Figure 3's $\delta{s}$ are the same, aren't they? If so we should say so and use the same name.
+I think Figure 5's $m$ and Figure 3's $\delta{s}$ are the same, aren't they? If so, we should say so and use the same name.
 If not we should explain why they're different.
@@
 name that map. In Figure \ref{fig:virtualpointstosharing}, this is done with a ghost map token for \emph{another} ghost map,
-one shared across the kernel, holding $m$. Our virtual points-to assertion existentially quanfies the address-space-specific
+one shared across the kernel, holding $m$. Our virtual points-to assertion existentially quantifies the address-space-specific
 ghost map $\delta$ mapping virtual addresses to physical addresses, and ensures it exists with a token for the distinguished
@@
   \end{figure}
-Figure \ref{fig:modaldef} gives the definition of our modal operator for asserting truth of a modal
+Figure \ref{fig:modaldef} gives the definition of our modal operator for asserting the truth of a modal
 (address-space-contingent) assertion \emph{in another address space}, which we call
 the \emph{other-space} modality. The definition itself is not
-particularly surprising --- as our modal assertions are semantically predicates on a page table root (physical)
-address, the assertion $[r](P)$ is a modal assertion which ignores the (implicit) current page table root,
+particularly surprising --- as our modal assertions are semantic predicates on a page table root (physical)
+address, the assertion $[r](P)$ is a modal assertion that ignores the (implicit) current page table root,
 and evaluates the truth of $P$ as if $r$ were the page table root. 
@@
 
-We can prove that this modality follows certain basic laws: that its truth is independent of the address
-space in which it is considered,  it distributes over various logical connectives, and it follows the rule of
+We can prove that this modality follows certain basic laws, showing that its truth is independent of the address
+space in which it is considered, that it is distributed over various logical connectives, and that it follows the rule of
 consequence.
 We call \textsf{vProp} assertions whose truth is independent of the current address space
-\textsf{Fact}s; these include other-space assertions, physical memory points-tos, and register assertions.
-\textsf{Fact}s can move in and out of other-space modalities freely.
+\textsf{Fact}s; these include other address-space assertions, physical memory points-tos, and register assertions.
+\textsf{Fact}s can freely move in and out of other address-space modalities.
 \looseness=-1
@@
 soundness of an individual virtual memory access.
-But such larger invariants would contain many kernel-specific properties that are orthogonal
+However, such larger invariants would contain many kernel-specific properties that are orthogonal
 to the fundamental reasoning principles that are the focus of this paper.
@@
 \textsf{WP} is \iris's own weakest precondition modality, unmodified~\cite{jung2018iris}.
-Making \textsf{rtv} a parameter to the double (vs. a simple register assertion)
+Making \textsf{rtv} a parameter to the double (versus a simple registration assertion)
 makes it possible to ensure ownership of the \lstinline|cr3| register and its value is accounted for
@@
 for control transfer instructions. Our development also includes handling of the \lstinline|rflags| register updates from arithmetic instructions.
-Most rules are otherwise standard (e.g., \lstinline|mov| between register, etc.), with Figure \ref{fig:wpdamd} showing the rules
+Most rules are otherwise standard (e.g., \lstinline|mov| between registers, etc.), with Figure \ref{fig:wpdamd} showing the rules
 most unique to our development.
@@
 conclusion). 
-It reflects the load of that memory value into the destination register $\textsf{r}_d$, with
+It reflects the loading of that memory value into the destination register $\textsf{r}_d$, with
 the updated register points-to in the precondition for $\overline{is}$.
@@
 the current address space.
-\todo[inline,color=green]{WIP: Colin is trying to clarify the value of modal approach here}
+\todo[inline,color=green]{WIP: Colin is trying to clarify the value of the modal approach here}
 }
@@
 \lstinline|mov| to $\textsf{rvs}$ afterwards. The global effects of this rule are reflected in moving
-\del{bare} assertions \add{for the current address space ($P$ and $\mathcal{I}\mathsf{ASpace}$)} under an other-space modality for \add{the initial
+\del{bare} assertions \add{for the current address space ($P$ and $\mathcal{I}\mathsf{ASpace}$)} under an other address-space modality for \add{the initial
 (outgoing) address space} $\rtv$, and moving the new address space's assertions out of
@@
 Such a rule is normally recoverable from Hoare doubles (see, e.g., \citet{Chlipala2011Bedrock,Chlipala2013Bedrock}).
-However, in the presence of address space changes the traditional frame rule is unsound.
+However, in the presence of address space changes, the traditional frame rule is unsound.
 Consider:
@@
 In this (broken!) hypothetical example,
-both the precondition and postcondition assert that $a\mapsto_\mathsf{v} x$ in the current address space, but we have no basis for concluding that address translation is preserved by the change of address space.
-So this derivation clearly leads to an unsound conclusion.
+both the precondition and the postcondition assert that $a\mapsto_\mathsf{v} x$ in the current address space, but we have no basis for concluding that address translation is preserved by the change of address space.
+So, this derivation clearly leads to an unsound conclusion.
 The essential problem is that the frame rule is motivated by local reasoning about local updates, but
-a switch of address space is a \emph{global} change that may invalidate information about virtual addresses.
-Thus framing around arbitrary \lstinline|cr3| updates is unsound --- hence the \emph{global} nature of \textsc{WriteToRegCtlFromRegModal} ---
+a switch of address-space is a \emph{global} change that may invalidate information about virtual addresses.
+Thus, framing around arbitrary \lstinline|cr3| updates is unsound --- hence the \emph{global} nature of \textsc{WriteToRegCtlFromRegModal} ---
 though a variant the ensures the same \lstinline|cr3| value is installed before and after the framing
@@
 \label{sec:soundness}
-Our rules from Figure \ref{fig:wpdamd} are proven sound in \iris against an assembly-level hardware model
-implementing a fragment of x86-64 including 64-bit address translation with 4-level page tables.
+Our rules from Figure \ref{fig:wpdamd} are proven to be sound in \iris against an assembly-level hardware model
+implementing a fragment of x86-64, including 64-bit address translation with 4-level page tables.
 % \todo[inline]{

### 2025-07-20 examples_C.tex
Examples check

@@
 % Recall from Section \ref{sec:logic} that virtual points-to assertions work just like regular points-to assertions, by design.
-In this section we verify several critical and challenging pieces of VMM code.
+In this section, we verify several critical and challenging pieces of VMM code.
 First, in several stages, we work up to mapping a new page in the current address space.
@@
 \label{sec:traversing}
-We build up to the main task of mapping a new page after traversing page tables in software.
+We build up to the main task of mapping a new page after traversing the page tables in the software.
 The mapping operation of Figure \ref{fig:mapping_codeC} assumes an operation \textsf{walkpgdir} which must traverse the page tables
@@
 Traversing the page tables is itself challenging functionality to verify: loading the current table root from \lstinline|cr3| is straightforward
-(a \lstinline|mov| instruction), however this produces the physical address of \lstinline|cr3|, not the virtual address the kernel code would use to access that memory.
-This problem repeats at each level of the page table: assuming the code has \emph{somehow} read the appropriate L4 (or L3, or L2) entry, those entries again
+(a \lstinline|mov| instruction); however, this produces the physical address of \lstinline|cr3|, not the virtual address the kernel code would use to access that memory.
+This problem repeats at each level of the page table: assuming that the code has \emph{somehow} read the appropriate L4 (or L3, or L2) entry, those entries again
 yield physical addresses, not virtual.
@@
 \subsubsection{Loading Page-Table Address Value}
-We will discuss access to the level 4 table later (Section \ref{wlkpgdirC}). But for subsequent levels, the base address of level $n$ must be
-fetched from the appropriate entry in the level $n+1$ table.
+We will discuss access to the level 4 table later (Section \ref{wlkpgdirC}). However, for subsequent levels, the base address of level $n$ must be
+fetched from the appropriate entry in the table of level $n+1$.
 This is the role of \lstinline|pte_get_next_table| (Figure \ref{fig:calltopteinitializeC}):
@@
 Figure \ref{fig:calltopteinitializeC} presents the initial part of the function, which performs the allocation if necessary.
-Lines \ref{line:finalpieceS}-\ref{line:finalpieceE} in Figure \ref{fig:calltopteinitializeC} (discussed in Section \ref{sec:p2vC}) deals with the cases where no allocation is necessary \emph{or} the allocation has already
+Lines \ref{line:finalpieceS}-\ref{line:finalpieceE} in Figure \ref{fig:calltopteinitializeC} (discussed in Section \ref{sec:p2vC}) deal with the cases where no allocation is necessary \emph{or} the allocation has already
 been performed by the code in this figure.
@@
 
-Note that the specification does \emph{not} assume a specific page table level --- logical parameter \textsf{v} represents the level
+Note that the specification does \emph{not} assume a specific page table level --- the logical parameter \textsf{v} represents the level
 of the entry passed as an argument, and this code
@@
 the case for the next level already existing (Lines \ref{line:finalpieceS}-\ref{line:finalpieceE} in Figure \ref{fig:calltopteinitializeC}), which is discussed in Section \ref{sec:p2vC}.
-First, we must discuss another refinement of the address space invariant, establishing
+First, we must discuss another refinement of the address-space invariant, establishing
 enough structure on the page tables themselves to allow the traversal.
@@
 (even though the physical-to-virtual translation
-is typically not literally the identity function, but adding a non-zero constant offset).\footnote{Some kernels do this for all physical memory on the machine, simplifying interaction
+is typically not literally the identity function, but adding a nonzero constant offset).\footnote{Some kernels do this for all physical memory on the machine, simplifying interaction
 with DMA devices.
 On newer platforms like RISC-V, this sometimes truly is an identity mapping ---
-x86-64 machines are forced into offsets by backwards compatibility with bootloaders that cannot access the full memory space of the
+x86-64 machines are forced into offsets by backward compatibility with bootloaders that cannot access the full memory space of the
 machine.
@@
 
-For this reason we extend the per-address-space invariant as in Figure \ref{fig:peraspaceinvariant_with_p2v_extensionC}, to also track which
-addresses we can perform a P2V conversion on by a adding a constant offset.
+For this reason, we extend the per-address-space invariant as in Figure \ref{fig:peraspaceinvariant_with_p2v_extensionC}, to also track which
+addresses we can perform a P2V conversion on by adding a constant offset.
 $\Xi$ is another ghost map, from physical addresses to the level of the page table they represent (1--4).
@@
 Section \ref{sec:p2vC} describes the actual conversion,
-but we describe the invariant here 
-because adding new level 3/2/1 tables must maintain the invariant.
+but we describe the invariant here because adding new level 3/2/1 tables must maintain the invariant.
 
@@
 
-For each $\paddr\mapsto \textsf{v} \in\Xi$, the invariant tracks a virtual points-to justifying that virtual address $\paddr+\textsf{KERNBASE}$ maps to physical address $\paddr$
+For each $\paddr\mapsto \textsf{v} \in\Xi$, the invariant tracks virtual points-to justifying that virtual address $\paddr+\textsf{KERNBASE}$ maps to physical address $\paddr$
 (the ``Ghost translation'' in Figure \ref{fig:peraspaceinvariant_with_p2v_extensionC});
@@
 assertions. But in the case where the entry is invalid, we may need to write to it (e.g., to install a reference to a next-level table, as we do in Figure \ref{fig:calltopteinitializeC}),
-which requires full permission. Fortunately, the entry can only be in use if its valid bit is set; if the valid bit is not set we know
+which requires full permission. Fortunately, the entry can only be in use if its valid bit is set; if the valid bit is not set, we know
 that no virtual points-to entry in $\delta$/$\theta$ holds any partial ownership.
@@
 if the entry is invalid the invariant holds full ownership of the entry, so it can be updated; while if the entry is valid,
-the invariant owns only a constant non-zero fragment sufficient to read the entry, but not modify it (which would invalidate some virtual points-to assertions):
+the invariant owns only a constant nonzero fragment sufficient to read the entry, but not modify it (which would invalidate some virtual points-to assertions):
 \begin{equation*}
@@
 together with other facts entails $\textsf{qfrac} = 1$ at Line \ref{line:after_concluding_qfrac1C},
-and permits storing a new entry (in ellided code around Line \ref{line:install_new_entryC}).
+and permits storing a new entry (in the elided code around Line \ref{line:install_new_entryC}).
 \looseness=-1
 
-This seemingly-simple piece of code has a highly non-trivial correctness argument, which depends critically on detailed invariants on how access to page table
+This seemingly simple piece of code has a highly nontrivial correctness argument, which depends critically on detailed invariants on how access to page table
 entries is shared between parts of the kernel. No prior work has engaged with this problem.
@@
   as is typical for slab allocators~\cite{bonwick1994slab}.
-  Its verification would be similar to verifying a usermode \textsf{malloc} verifications~\cite{Chlipala2013Bedrock,wickerson2010explicit},
+  Its verification would be similar to verifying a usermode \textsf{malloc} verification ~\cite{Chlipala2013Bedrock,wickerson2010explicit},
   just with additional invariants on the memory pool.
@@
 \footnote{A reasonable reader might wonder where this pool
-initially comes from, and how it might grow when needed. Typically an initial mapping subject to this identity mapping
+initially comes from and how it might grow when needed. Typically an initial mapping subject to this identity mapping
 constraint is set up prior to transition to 64-bit kernel code (notably,
@@
 The soundness argument of this specification relies on the fact that these freshly allocated resources are part
-of an entry construction that has not been completed yet: the presence bit is set 
-(Line \ref{line:install_new_entryC} in Figure \ref{fig:calltopteinitializeC}) after these freshly allocated resources are incorporated to the 
-entry construction via the page-frame portion of the PTE. In other words, the side condition, (*),
+of an entry construction that has not been completed yet: the presence bit is set (line \ref{line:install_new_entryC} in Figure \ref{fig:calltopteinitializeC}) after these freshly allocated resources are incorporated into the
+entry construction via the page frame portion of the PTE. In other words, the side condition, (*),
  formalizes that any access to the entry with these resources is \textit{invalid} (in the sense of not necessarily
 having accompanying resources) until the entry is marked present (and thus the memory returned from \textsf{kalloc}
-moves into the page table invariant.
+moves into the page table invariant).
 
@@
 \label{sec:p2vC}
-Once we know the entry refers to a physical address in the identity mapping range ($\Xi$)
+Once we know that the entry refers to a physical address in the identity mapping range ($\Xi$)
 (via the branch at Line \ref{line:check_entry_present_jumpC}, or by allocating and installing a new entry
@@
 %\end{figure}
-Lines \ref{line:childrenC}-\ref{line:p2vC} in Figure \ref{fig:calltopteinitializeC} show the verification of the end of \lstinline|pte_get_next_table| specialized to the case where 
-where no allocation was necessary (i.e., the conditional on Line \ref{line:check_entry_presentC} of Figure \ref{fig:calltopteinitializeC} was taken).
+Lines \ref{line:childrenC}-\ref{line:p2vC} in Figure \ref{fig:calltopteinitializeC} show the verification of the end of \lstinline|pte_get_next_table| specialized to the case where no allocation was necessary (i.e., the conditional on Line \ref{line:check_entry_presentC} of Figure \ref{fig:calltopteinitializeC} was taken).
 In this case, the true present bit allows access to the child tokens from Line \ref{line:conditional_childrenC} of Figure \ref{fig:calltopteinitializeC},
@@
 but does not provide a virtual points-to assertion --- because the function must calculate (Lines \ref{line:start_l4_calcC}--\ref{line:end_l4_calcC})
-which entry it needs access to.
-Instead the precondition has 512 identity map tokens, guaranteeing that every entry on the page is subject to the identity mapping invariant.
+which entry it needs access.
+Instead, the precondition has 512 identity map tokens, guaranteeing that every entry on the page is subject to the identity mapping invariant.
 Line \ref{line:end_l4_calcC} calculates the virtual address of the relevant entry, and the subsequent view shift
@@
 The ghost translation and physical location are used to form the virtual pte-pointsto for the L4 entry
-(Line \ref{line:first_pte_pointstoC}), with the entry validity and next-level indexing
+(Line \ref{line:first_pte_pointstoC}), with entry validity and next-level indexing
 satisfying the rest of the precondition for \lstinline|pte_get_next_table|.
-\lstinline|pte_get_next_table| then, as described earlier, checks the valid bit in the indicated
+Then, as described earlier, checks the valid bit in the indicated
 entry and either returns the (unconditional) tokens for the L3 entry physical addresses (if valid), or
@@
 One of the key tasks of a page fault handler in a general-purpose OS kernel is
-to map new pages into an address space by writing into an existing page table via a call\\
+to map new pages to an address space by writing into an existing page table via a call.\\
 \centerline{\textsf{vaspace\_mappage(pte\_t *l4, void *va,uintptr\_t fpaddr)}}\\
 in Figure \ref{fig:mapping_codeC}.
-To do so, with a given allocated a fresh page (\textsf{fpaddr}), then calculate the appropriate
+To do so, with a given allocated fresh page (\textsf{fpaddr}), then calculate the appropriate
 known-valid page table walks (via \textsf{walkpgdir} Line \ref{line:call_walkpgdirC} in Figure \ref{fig:mapping_codeC}) and update
@@
 The code uses a helper function \lstinline{walkpgdir} (discussed again in Section \ref{wlkpgdirC}).
-\textsf{walkpgdir}'s postcondition contains virtual \emph{PTE} pointsto assertions ($\mapsto_{\textsf{vpte}}$)
+\textsf{walkpgdir}'s postcondition contains virtual \emph{PTE} points-to assertions ($\mapsto_{\textsf{vpte}}$)
 both for ensuring partial page table walk reaching the
@@
 
-A PTE points-to, as we already brifly mentioned, is defined just like the normal virtual points-to of Figure \ref{fig:virtualpointstosharingC}, except the physical address (here, \textsf{pa}) is explicit in the assertion
+A PTE points-to, as we already have briefly mentioned, is defined just like the normal virtual points-to of Figure \ref{fig:virtualpointstosharingC}, except the physical address (here, \textsf{pa}) is explicit in the assertion
 rather than existentially quantified:\\
@@
 However, for L1 entries this check is left to the caller of the
-traversal function \textsf{walkpgdir}. In other words, unlike what we see in R$_{\textsf{walk}}$ for the upper levels where all entry-present
-checks have already been performed, the specification in R$_{\textsf{l1e}}$ ensures that page table entry for L1 needs to be checked at the caller site. 
-By doing so, as we see in Figure \ref{fig:mapping_codeC}, the page reference \textsf{fpaddr} is linked to back the virtual address \textsf{va} 
-only if it is not already referring to a physical resource (Lines \ref{line:mappage_pte_present_startC}--\ref{line:mappage_pte_present_endC} in Figure \ref{fig:mapping_codeC}).
+traversal function \textsf{walkpgdir}. In other words, unlike what we see in R$_{\textsf{walk}}$ for the upper levels, here all entry-present
+checks have already been performed; the specification in R$_{\textsf{l1e}}$ ensures that page table entry for L1 needs to be checked at the caller site. 
+In doing so, as we see in Figure \ref{fig:mapping_codeC}, the page reference \textsf{fpaddr} is linked back to the virtual address \textsf{va} 
+only if it does not already refer to a physical resource (lines \ref{line:mappage_pte_present_startC}--\ref{line:mappage_pte_present_endC} in Figure \ref{fig:mapping_codeC}).
 
@@
 full permission on the L1 entry could be obtained, allowing the construction of a
-full virtual PTE pointer for it, setting to 0, and reclaiming the now-unmapped physical memory.
+full virtual PTE pointer for it, setting to 0, and reclaiming the now unmapped physical memory.
 \fi
@@
 This code assumes the System V AMD64 ABI calling convention, where the normal registers not mentioned are caller-save, and therefore saved on the stack of the thread
-that calls this code, as well as on the new stack of the thread that is restored, thus only the callee-save registers and \texttt{cr3} must be 
+that calls this code, as well as on the new stack of the thread that is restored; thus, only the callee-save registers and \texttt{cr3} must be
 restored.\footnote{We are simplifying by ignoring non-integer registers (e.g., floating point, vector registers),
 and the caller-save registers should be initialized to 0 to avoid leaking information across processes, but this captures the key challenges.}
-With the addition of a return instruction, this code would satisfy the C function signature\footnote{This is the function in UNIX 6th Edition 
+With the addition of a return instruction, this code would satisfy the signatures of the C functions \footnote{This is the function in UNIX 6th Edition 
 with the infamous ``You are not expected to understand this'' comment~\cite{lions1996lions}.}
 \centerline{\lstinline[language=C]|void swtch(context_t* save, context_t* restore);|}\\
-A call to this code begins executing one thread (until just before Line \ref{line:end_saveC}) in one address space ($\rtv$), whose information will be saved in a structure at address $old$,
-and finishes execution executing a different thread in a different address space (Line \ref{line:end_restoreC} on) whose information is initially in $new$.
+A call to this code begins by executing one thread (up to just before line \ref{line:end_saveC}) in an address space ($\rtv$), whose information will be saved in a structure at address $old$,
+and finishes the execution by executing a different thread in a different address-space (line \ref{line:end_restoreC} on), whose information is initially in $new$.
 
@@
 half of the code.
-Likewise, part of the invariant of the stack of the second thread, the one being restored, is that the return address on \emph{that} stack requires the saved 
-callee-save registers stored in that context to be in registers as its precondition.
+Likewise, part of the invariant of the stack of the second thread, the one being restored, is that the return address on \emph{that} stack requires the saved callee-save registers stored in that context be in registers as its precondition.
 
-The wrinkle, and the importance of the modal treatment of assertions, is that the target thread's precondition is \emph{relative to its address space}, 
+The wrinkle and the importance of the modal treatment of assertions is that the target thread's precondition is \emph{relative to its address space}, 
 not the address space of the calling thread, which is reflected by
@@
 \emph{in the other address space}. 
-The other-space modality plays a critical role here; \textsf{Pother} would contain these assumptions in the other
+The other address-space modality plays a critical role here; \textsf{Pother} would contain these assumptions in the other
 address space.
@@
 
-While prior work has verified context switches within a single address space~\cite{ni2007contexts}, and context switches
-without any code before or after~\cite{syeda2020formal} (i.e., not reasoning about the impact of address space change
-on what data was accessible), this is the first verification handling both.
+Although prior work has verified context switches within a single address space~\cite{ni2007contexts}, and context switches
+without any code before or after~\cite{syeda2020formal} (that is, not reasoning about the impact of address space change
+on what data were accessible), this is the first verification that handles both.
 \looseness=-1

### 2025-07-20 relwork.tex
whole paper except appendix proof read

@@
 The original \textsc{Verisoft} project~\cite{alkassar2008verisoft,alkassar2010pervasive,alkassar2008formal,dalinger2005verification,hillebrand2005address,alkassar2008formal,starostin2010formal} 
-relied on custom hardware which, among other things, always ran kernel code with virtual memory disabled, removing the circularity that is a key challenge of verifying 
-VMM code for real hardware: at that point page tables become a basic partial map data structure to represent user program address translations,
+relied on custom hardware, which, among other things, always ran kernel code with virtual memory disabled, removing the circularity that is a key challenge of verifying
+VMM code for real hardware: At that point page tables become a basic partial map data structure to represent user program address translations,
 with an idiosyncratic format. It turns out that subsequent OS verification work also treats page tables this way, but
@@
 % one author has personally encountered in a research kernel, though not one discussed here.}).
-This is true of \textsc{seL4}~\cite{Klein2009seL4,seL4TOCS,Sewell2013translation}, whose formal machine model omits address translation;
+This is true for \textsc{seL4}~\cite{Klein2009seL4,seL4TOCS,Sewell2013translation}, whose formal machine model omits address translation,
 and \textsc{CertiKOS}~\cite{gu15,gu2016certikos,gu2018certikos,chen2016interrupts}, whose refinement proofs rely on
@@
 Other work on OS verification either never progressed far enough to address VMM verification
-(\textsc{Verisoft XT}~\cite{cohen2009vcc,cohen2010local,dahlweid2009vcc,cohen2013SOFSEM}), or uses memory-safe languages to enable safe co-habitation 
+(\textsc{Verisoft XT}~\cite{cohen2009vcc,cohen2010local,dahlweid2009vcc,cohen2013SOFSEM}), or uses memory-safe languages to enable safe cohabitation 
 of a single address space by all processes (\textsc{Singularity}~\cite{Fahndrich2006language,Hunt2007singularity,Hunt2007sealing,Barnett2011specsharp}, \textsc{Verve}~\cite{Yang2010Verve},
 and \textsc{Tock}~\cite{levy2017multiprogramming}). The latter approach ensures memory safety, but ignores other functional uses of virtual memory
-hardware like swapping~\cite{Denning1970VM} or exploiting copy-on-write techniques for dynamic migration of virtual machines~\cite{clark2005live}.
+hardware, like swapping~\cite{Denning1970VM} or exploiting copy-on-write techniques for dynamic migration of virtual machines~\cite{clark2005live}.
 \looseness=-1
@@
 but the work never addressed the issues of managing multiple address spaces, switching address spaces, or
-invariants referring to other address spaces, which our explicitly modal approach and case studies do address.
-They verified a fraction of an ARM version of our Figure \ref{fig:mapping_code}, which assumed they already had the virtual address
+invariants referring to other address spaces, which our explicitly modal approach and case studies address.
+They verified a fraction of an ARM version of our Figure \ref{fig:mapping_code}, which assumed that they already had the virtual address
 of the L1 entry (which they assumed existed), thereby skipping the physical-to-virtual conversion issues we treat in Section \ref{sec:p2v}.
-Due to lack of fractional permissions in their formalism, they incidentally pick up other limitations orthogonal to their
+Due to a lack of fractional permissions in their formalism, they incidentally pick up other limitations orthogonal to their
 foundational focus: by requiring

### 2025-07-20 background.tex
+

@@
 \end{tikzpicture}
-          \caption{Accessing to the Page Referenced by L1 Entry}
-        \label{fig:enter-label}
+          \caption{x86-64 page table lookups.}
+        \label{fig:pagetables}
     \end{subfigure}
@@
     \vspace{-2.5em}
-    \caption{x86-64 page table lookups.}
-    \label{fig:pagetables}
+    \caption{A kernel implementing x86-64 page table lookups.}
+    \label{fig:pagetablescode}
     \vspace{-1em}
@@
 AMD64 architecture manual~\cite{amd64_manual_vol2}\footnote{While x86 up through its 32-bit incarnation were due to Intel,
-the x86-64 architecture as a 64-bit extension to x86 was originally due to AMD. As a result, it is sometimes also referred to as the \texttt{amd64} architecture.}) shows, 
-address translation proceeds by repeatedly taking designated slices of the virtual address and indexing into a table.
+the x86-64 architecture as a 64-bit extension to x86 was originally due to AMD.
+As a result, it is sometimes also referred to as the \texttt{amd64} architecture.}) shows a \emph{page-table walk} for a virtual address.
+The page tables are managed by the OS, typically by a \emph{virtual memory manager} (VMM).\footnote{Not to be confused with Virtual
+Machine Monitor. We focus on non-hypervisor scenarios, but hardware virtualization extensions for both 
+x86-64 and ARM make use of an additional set of page tables translating what a \emph{guest} considers to be its 
+(virtualized) physical memory to actual physical memory. Our contributions should offer value in this scenario as well.}
+Typically each process has its own 
+page table, which the OS registers with the CPU by storing the (page-aligned)
+physical address of the root of the page table tree (the start of the L4 table shown as \mathsf{root} in Figure \ref{fig:pagetables}) in a specific register (\texttt{cr3}) as part of switching to a new process.
+Using different mappings, which map only disjoint portions of physical memory (with some exceptions in the next section) 
+is how the OS ensures memory isolation between processes.
+
+In Figure \ref{fig:pagetablescode}, we see a simplified kernel code, \lstinline|walkpgdir|, implementing this page-table walk for translating the virtual address $\vaddr$ over the page-table tree with unique root address $\mathsf{root}$ shown in Figure \ref{fig:pagetables}.
+As shown in Figure at the top of Figure \ref{fig:pagetables}, on x86-64, standard configurations use 4 levels of page tables, labelled levels 4 through 1, with lookups in the level 
+1 page table resulting in the actual page of physical memory holding the requested data, and the low-order 12 bits 
+being used to index into this page.\footnote{Technically levels 1--3 have explicit historical names,
+but for brevity and consistency, we simply number them, in keeping with the newer 5th level. Our formalization only deals with 4-level page tables, but is straightforwardly extensible to 5.} 
+As depicted at the top of Figure \ref{fig:pagetables}, for each level of translation, address-translation implementation repeatedly takes designated slices of the virtual address and indexing into a table --- e.g., L4Offset(va) at Line 2 of \lstlinline|walkpgdir| in Figure \ref{fig:pagetablescode}.
 The final lookup in the page tables gives the base physical address of a 4KB page of physical memory, to which the low-order
 bits of the accessed virtual address are added to determine the actual physical address retrieved.
-On x86-64, standard configurations use 4 levels of page tables, labelled levels 4 through 1, with lookups in the level 
-1 page table resulting in the actual page of physical memory holding the requested data, and the low-order 12 bits 
-being used to index into this page.\footnote{Technically levels 1--3 have explicit historical names,
-but for brevity and consistency, we simply number them, in keeping with the newer 5th level.
-Our formalization
-only deals with 4-level page tables, but is straightforwardly extensible to 5.}  
-The translation process or algorithm is sometimes referred to as a \emph{page-table walk}. 
+The entries of each table are 64 bits wide, but each points to a physical address aligned to 4KB (4096 byte) boundaries, which leaves 12 bits to spare to
+control a validity bit (called the \emph{present} bit), a read-write bit (which permits write access through the entry if and only if it is set), and a range of additional bits which can be used to control caching, write-through, and more.
+This paper will only consider the present bit (0).
+
 While Figure \ref{fig:pagetables} and most of our constants (how many levels, which virtual address bits index which
@@
 in different orders).
-
-The entries of each table are 64 bits wide, but each points to a physical address aligned to 4KB (4096 byte) boundaries, which leaves 12 bits to spare to 
-control a validity bit (called the \emph{present} bit), a read-write bit (which permits write access through the entry if and only if it is set), 
-and a range of additional bits which can be used to control caching, write-through, and more. This paper will only consider the present bit (0).
-
-The page tables are managed by the OS, typically by a \emph{virtual memory manager} (VMM).\footnote{Not to be confused with Virtual 
-Machine Monitor. We focus on non-hypervisor scenarios, but hardware virtualization extensions for both 
-x86-64 and ARM make use of an additional set of page tables translating what a \emph{guest} considers to be its 
-(virtualized) physical memory to actual physical memory. Our contributions should offer value in this scenario as well.}
-Typically each process has its own 
-page table, which the OS registers with the CPU by storing the (page-aligned)
-physical address of the root of the page table treethe start of the L4 table) in a 
-specific register (\texttt{cr3}) as part of switching to a new process. 
-Using different mappings, which map only disjoint portions of physical memory (with some exceptions in the next section) 
-is how the OS ensures memory isolation between processes.
 

### 2025-07-20 background.tex
+

@@
 bits of the accessed virtual address are added to determine the actual physical address retrieved.
-The entries of each table are 64 bits wide, but each points to a physical address aligned to 4KB (4096 byte) boundaries, which leaves 12 bits to spare to
+The entries (e.g., L4 Entry in Figure \ref{fig:pagetables}) of each table are 64 bits wide, but each points to a physical address aligned to 4KB (4096 byte) boundaries, which leaves 12 bits to spare to
 control a validity bit (called the \emph{present} bit), a read-write bit (which permits write access through the entry if and only if it is set), and a range of additional bits which can be used to control caching, write-through, and more.
 This paper will only consider the present bit (0).
+As we see in Figure \ref{fig:pagetables}, the entry L2 is not be present (shown as dashed in Figure \ref{fig:pagetables}), ...
 

### 2025-07-20 background.tex
background more elaborative

@@
 This paper will only consider the present bit (0).
-As we see in Figure \ref{fig:pagetables}, the entry L2 is not be present (shown as dashed in Figure \ref{fig:pagetables}), ...
+As we see in Figure \ref{fig:pagetables}, the entry L2 is not be present (shown as dashed in Figure \ref{fig:pagetables}).
+In order to able to access to the subsequent table base address (e.g., $\mathsf{l1}$ in Line 7 of \lstinline|walkpgdir| in Figure \ref{fig:pagetablescode}).
+Accessing to the next table is realized in \lstinline|pte_get_next_table| in Figure \ref{fig:pagetablescode} to which the current entry (e.g., $\mathsf{l2_entry}$ in Line 7 of \lstinline|walkpgdir| in Figure \ref{fig:pagetablescode}), and this method first check if the entry is present (Line 7 of \lstinline|pte_get_next_table| in Figure \ref{fig:pagetablecodes}).
+For a nonexisting entry such as L2 entry, the execution goes through allocation path shown in Lines 7-15 of \lstinline|pte_get_next_table| in Figure \ref{fig:pagetablecodes} which allocates a \emph{physical addresses} (512 L1 entries) and make $\mathsf{l2}$ pointer's frame address to show the first entry of the L1 entries fresly allocated (Line 12 of \lstinline|pte_get_next_table| in Figure \ref{fig:pagetablecodes}), and set the present bit of $\mathsf{l2_entry}$ (Line 12 of \lstinline|pte_get_next_table| in Figure \ref{fig:pagetablecodes}).
+After the validity of page-table walk is restored, \lstinline|pte_get_next_table| converts the physical frame address of the entry to a \emph{virtual address} in Lines 16-20 of \lstinline|pte_get_next_table| in Figure \ref{fig:pagetablecodes} return the virtual address of the subsequent ($\mathsf{next}$ in in Figure \ref{fig:pagetablecodes}) base address (e.g., $\mathsf{l1}$ in Line 7 of \lstinline|walkpgdir| in Figure \ref{fig:pagetablescode}).
 

### 2025-07-21 background.tex
+

@@
 page table, which the OS registers with the CPU by storing the (page-aligned)
-physical address of the root of the page table tree (the start of the L4 table shown as \mathsf{root} in Figure \ref{fig:pagetables}) in a specific register (\texttt{cr3}) as part of switching to a new process.
+physical address of the root of the page table tree (the start of the L4 table shown as $\mathsf{root}$ in Figure \ref{fig:pagetables}) in a specific register (\texttt{cr3}) as part of switching to a new process.
 Using different mappings, which map only disjoint portions of physical memory (with some exceptions in the next section) 
@@
 but for brevity and consistency, we simply number them, in keeping with the newer 5th level. Our formalization only deals with 4-level page tables, but is straightforwardly extensible to 5.} 
-As depicted at the top of Figure \ref{fig:pagetables}, for each level of translation, address-translation implementation repeatedly takes designated slices of the virtual address and indexing into a table --- e.g., L4Offset(va) at Line 2 of \lstlinline|walkpgdir| in Figure \ref{fig:pagetablescode}.
+As depicted at the top of Figure \ref{fig:pagetables}, for each level of translation, address-translation implementation repeatedly takes designated slices of the virtual address and indexing into a table --- e.g., L4Offset(va) at Line 2 of \lstinline|walkpgdir| in Figure \ref{fig:pagetablescode}.
 The final lookup in the page tables gives the base physical address of a 4KB page of physical memory, to which the low-order

### 2025-07-21 programlogic.tex
pass from ismail

@@
   \todo[inline]{Ismail: Tie these lines to the steps and entries in Figure \ref{fig:pagetables}, e.g., with \textcircled{1}\ldots}
+  \todo[inline]{Ismail Addressed: earlier in the background and the current text is clearly related to \ref{fig:pagetables}}
 }
@@
 }
+\todo[inline]{Ismail Addressed: current explanation in here is fortified by the changed text in the background section smoothly}
+
 \lstinline|L4_L1_PointsTo| works by
@@
   \left\{ \begin{array}{l}
-    P \ast \mathcal{I}\textsf{ASpace}\ast \mathcal{I}\textsf{ASpace} \ast \textsf{rip} \mapsto_{\textsf{r}} {\textsf{iv} + \textsf{MovLen}(\textsf{r}_d,[\textsf{r}_a])} \ast \arcr
-    r_d \mapsto_{r}  \textsf{v} \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{v}} \textsf{v}
+    P \ast \mathcal{I}\textsf{ASpace}\ast  \textsf{rip} \mapsto_{\textsf{r}} {\textsf{iv} + \textsf{MovLen}(\textsf{r}_d,[\textsf{r}_a])} \ast \arcr
+    r_d \mapsto_{r}  \textsf{v} \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{v}} \{q\} \textsf{v}
   \end{array} \right\}_{\textsf{rtv}}\;\overline{is}
@@
 Or shading the VMM-specific ones?
+}
+\todo[inline]{Ismail Addressed: I don't think that we should do.
 }
@@
 }
-
+\todo[inline,color=cyan]{Ismail Addressed:
+Others issues are addressed.
+"(Also, explaining the fraction notation was one of the questions from reviewers.)"
+They are explained earlier in the text. No need. They should go back and read.
+}
 \subsubsection{Updating \lstinline|cr3|} 

### 2025-07-21 programlogic.tex
+

@@
   \left\{ \begin{array}{l}
-    P \ast \mathcal{I}\textsf{ASpace}\ast  \textsf{rip} \mapsto_{\textsf{r}} {\textsf{iv} + \textsf{MovLen}(\textsf{r}_d,[\textsf{r}_a])} \ast \arcr
-    r_d \mapsto_{r}  \textsf{v} \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{v}} \{q\} \textsf{v}
+    P \ast \mathcal{I}\textsf{ASpace}\ast  \textsf{rip} \mapsto_{\textsf{r}} \; {\textsf{iv} + \textsf{MovLen}(\textsf{r}_d,[\textsf{r}_a])} \ast \arcr
+    r_d \mapsto_{r}  \textsf{ v} \ast r_a \mapsto_{r} \; \{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{v}} \;\{q\} \textsf{ v}
   \end{array} \right\}_{\textsf{rtv}}\;\overline{is}
 }{
-  \{P \ast \mathcal{I}\textsf{ASpace}\ast \textsf{rip} \mapsto_{\textsf{r}} \textsf{iv} \ast r_d \mapsto_{r}  \textsf{rvd} \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{v}} \textsf{v} \}_{\textsf{rtv}}\;
+  \{P \ast \mathcal{I}\textsf{ASpace}\ast \textsf{rip} \mapsto_{\textsf{r}} \textsf{ iv} \ast r_d \mapsto_{r}  \textsf{ rvd} \ast r_a \mapsto_{r} \;\{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{v}} \textsf{ v} \}_{\textsf{rtv}}\;
   \textsf{mov}~\textsf{r}_d,~[\textsf{r}_a];\;\overline{is}
@@
 \inferrule[\footnotesize WriteToVirtMemFromReg]{
-  \{P \ast \mathcal{I}\textsf{ASpace}\ast \textsf{rip} \mapsto_{\textsf{r}} \textsf{iv} + \textsf{MovLen}(\textsf{r}_d,[\textsf{r}_a])  \ast r_s \mapsto_{r}\{q\}  \textsf{rvs}  \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{v}} \textsf{v} \}_{\textsf{rtv}}\;\overline{ is}
+  \{P \ast \mathcal{I}\textsf{ASpace}\ast \textsf{rip} \mapsto_{\textsf{r}} \textsf{ iv} + \textsf{MovLen}(\textsf{r}_d,[\textsf{r}_a])  \ast r_s \mapsto_{r}\;\{q\}  \textsf{ rvs}  \ast r_a \mapsto_{r} \; \{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{v}} \textsf{ v} \}_{\textsf{rtv}}\;\overline{ is}
 }{
-  \{P \ast \mathcal{I}\textsf{ASpace}\ast \textsf{rip} \mapsto_{\textsf{r}} \textsf{iv} \ast r_s \mapsto_{r}\{q\}  \textsf{rvs}   \ast r_a \mapsto_{r}\{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{v}} \textsf{rvs} \}_{\textsf{rtv}}
+  \{P \ast \mathcal{I}\textsf{ASpace}\ast \textsf{rip} \mapsto_{\textsf{r}} \textsf{ iv} \ast r_s \mapsto_{r} \;\{q\}  \textsf{ rvs}   \ast r_a \mapsto_{r} \; \{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{v}} \textsf{ rvs} \}_{\textsf{rtv}}
   \textsf{ mov}~[\textsf{r}_a],~\textsf{r}_s;\;\overline{is}
@@
 \inferrule[\footnotesize WriteToRegCtlFromRegModal]{
-  \{[\textsf{rtv}](P \ast \mathcal{I}\textsf{ASpace})\ast \textsf{rip} \mapsto_{\textsf{r}} \textsf{iv + 4} \ast \mathcal{I}\textsf{ASpace} \ast R\ast r_s \mapsto_{r}\{q\}  \textsf{ rvs}  \}_{\textsf{rvs}} \;\overline{is}
+  \{[\textsf{rtv}](P \ast \mathcal{I}\textsf{ASpace})\ast \textsf{rip} \mapsto_{\textsf{r}} \textsf{ iv + 4} \ast \mathcal{I}\textsf{ASpace} \ast R\ast r_s \mapsto_{r} \;\{q\}  \textsf{ rvs}  \}_{\textsf{rvs}} \;\overline{is}
 }{
-  \{P \ast \mathcal{I}\textsf{ASpace} \ast \textsf{rip} \mapsto_{\textsf{r}} \textsf{iv} \ast [\textsf{rvs}](R \ast\mathcal{I}\textsf{ASpace})\ast r_s \mapsto_{r}\{q\}  \textsf{ rvs}   \}_{\textsf{rtv}} \;
+  \{P \ast \mathcal{I}\textsf{ASpace} \ast \textsf{rip} \mapsto_{\textsf{r}} \textsf{ iv} \ast [\textsf{rvs}](R \ast\mathcal{I}\textsf{ASpace})\ast r_s \mapsto_{r}\; \{q\}  \textsf{ rvs}   \}_{\textsf{rtv}} \;
   \textsf{mov}~\textsf{cr3},~r_s;\;\overline{is}
@@
 }
-\todo[inline]{Ismail Addressed: I don't think that we should do.
+\todo[inline]{Ismail Addressed: I don't think that we should do -- general instructions' text outlines the how to read VMM related ones in the text. 
 }

### 2025-07-21 programlogic.tex
+

@@
 \inferrule[\footnotesize WriteToVirtMemFromReg]{
-  \{P \ast \mathcal{I}\textsf{ASpace}\ast \textsf{rip} \mapsto_{\textsf{r}} \textsf{ iv} + \textsf{MovLen}(\textsf{r}_d,[\textsf{r}_a])  \ast r_s \mapsto_{r}\;\{q\}  \textsf{ rvs}  \ast r_a \mapsto_{r} \; \{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{v}} \textsf{ v} \}_{\textsf{rtv}}\;\overline{ is}
+  \{P \ast \mathcal{I}\textsf{ASpace}\ast \textsf{rip} \mapsto_{\textsf{r}} \textsf{ iv} + \textsf{MovLen}(\textsf{r}_d,[\textsf{r}_a])  \ast r_s \mapsto_{r}\;\{q_1\}  \textsf{ rvs}  \ast r_a \mapsto_{r} \; \{q_2\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{v}} \textsf{ v} \}_{\textsf{rtv}}\;\overline{ is}
 }{
-  \{P \ast \mathcal{I}\textsf{ASpace}\ast \textsf{rip} \mapsto_{\textsf{r}} \textsf{ iv} \ast r_s \mapsto_{r} \;\{q\}  \textsf{ rvs}   \ast r_a \mapsto_{r} \; \{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{v}} \textsf{ rvs} \}_{\textsf{rtv}}
+  \{P \ast \mathcal{I}\textsf{ASpace}\ast \textsf{rip} \mapsto_{\textsf{r}} \textsf{ iv} \ast r_s \mapsto_{r} \;\{q_1\}  \textsf{ rvs}   \ast r_a \mapsto_{r} \; \{q_2\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{v}} \textsf{ rvs} \}_{\textsf{rtv}}
   \textsf{ mov}~[\textsf{r}_a],~\textsf{r}_s;\;\overline{is}

### 2025-07-21 background.tex
C pass through Ismail changes: adding change-tracking macros, fixing bad suggestions from grammar tool

@@
 
+\todo[inline]{Colin says: the changes in the next couple paragraphs seem to have deleted critical information about how the hardware
+translates addresses.}
 On the x86-64 architecture, the \textsc{MMU}'s address translation uses a sparse hierarchical set of tables:
@@
 AMD64 architecture manual~\cite{amd64_manual_vol2}\footnote{While x86 up through its 32-bit incarnation were due to Intel,
-the x86-64 architecture as a 64-bit extension to x86 was originally due to AMD.
-As a result, it is sometimes also referred to as the \texttt{amd64} architecture.}) shows a \emph{page-table walk} for a virtual address.
-The page tables are managed by the OS, typically by a \emph{virtual memory manager} (VMM).\footnote{Not to be confused with Virtual
+the x86-64 architecture as a 64-bit extension to x86 was originally due to AMD. As a result, it is sometimes also referred to as the \texttt{amd64} architecture.})
+\replace{shows,
+address translation proceeds by repeatedly taking designated slices of the virtual address and indexing into a table.
+The final lookup in the page tables gives the base physical address of a 4KB page of physical memory, to which the low-order
+bits of the accessed virtual address are added to determine the actual physical address retrieved.
+On x86-64, standard configurations use 4 levels of page tables, labelled levels 4 through 1, with lookups in the level 
+}{
+shows a \emph{page-table walk} for a virtual address. The page tables are managed by the OS, typically by a \emph{virtual memory manager} (VMM).\footnote{Not to be confused with Virtual 
 Machine Monitor. We focus on non-hypervisor scenarios, but hardware virtualization extensions for both 
@@
 is how the OS ensures memory isolation between processes.
+}
 
+\add{
 In Figure \ref{fig:pagetablescode}, we see a simplified kernel code, \lstinline|walkpgdir|, implementing this page-table walk for translating the virtual address $\vaddr$ over the page-table tree with unique root address $\mathsf{root}$ shown in Figure \ref{fig:pagetables}.
 As shown in Figure at the top of Figure \ref{fig:pagetables}, on x86-64, standard configurations use 4 levels of page tables, labelled levels 4 through 1, with lookups in the level 
+}
 1 page table resulting in the actual page of physical memory holding the requested data, and the low-order 12 bits 
@@
 but for brevity and consistency, we simply number them, in keeping with the newer 5th level. Our formalization only deals with 4-level page tables, but is straightforwardly extensible to 5.} 
+\replace{
+  The translation process or algorithm is sometimes referred to as a \emph{page-table walk}. 
+}{
 As depicted at the top of Figure \ref{fig:pagetables}, for each level of translation, address-translation implementation repeatedly takes designated slices of the virtual address and indexing into a table --- e.g., L4Offset(va) at Line 2 of \lstinline|walkpgdir| in Figure \ref{fig:pagetablescode}.
@@
 bits of the accessed virtual address are added to determine the actual physical address retrieved.
-The entries (e.g., L4 Entry in Figure \ref{fig:pagetables}) of each table are 64 bits wide, but each points to a physical address aligned to 4KB (4096 byte) boundaries, which leaves 12 bits to spare to
-control a validity bit (called the \emph{present} bit), a read-write bit (which permits write access through the entry if and only if it is set), and a range of additional bits which can be used to control caching, write-through, and more.
+The entries \add{(e.g., L4 Entry in Figure \ref{fig:pagetables})} of each table are 64 bits wide, but each points to a physical address aligned to 4KB (4096 byte) boundaries, which leaves 12 bits to spare to
+control a validity bit (called the \emph{present} bit), a read-write bit (which permits write access through the entry if and only if it is set),
+and a range of additional bits which can be used to control caching, write-through, and more.
 This paper will only consider the present bit (0).
-As we see in Figure \ref{fig:pagetables}, the entry L2 is not be present (shown as dashed in Figure \ref{fig:pagetables}).
+}
+\add{As we see in Figure \ref{fig:pagetables}, the entry L2 is not be present (shown as dashed in Figure \ref{fig:pagetables}).
 In order to able to access to the subsequent table base address (e.g., $\mathsf{l1}$ in Line 7 of \lstinline|walkpgdir| in Figure \ref{fig:pagetablescode}).
@@
 After the validity of page-table walk is restored, \lstinline|pte_get_next_table| converts the physical frame address of the entry to a \emph{virtual address} in Lines 16-20 of \lstinline|pte_get_next_table| in Figure \ref{fig:pagetablecodes} return the virtual address of the subsequent ($\mathsf{next}$ in in Figure \ref{fig:pagetablecodes}) base address (e.g., $\mathsf{l1}$ in Line 7 of \lstinline|walkpgdir| in Figure \ref{fig:pagetablescode}).
+}
 
@@
 in different orders).
+
+\del{
+  The entries of each table are 64 bits wide, but each points to a physical address aligned to 4KB (4096 byte) boundaries, which leaves 12 bits to spare to
+  control a validity bit (called the \emph{present} bit), a read-write bit (which permits write access through the entry if and only if it is set),
+  and a range of additional bits which can be used to control caching, write-through, and more. This paper will only consider the present bit (0).
+}
+
+\del{
+  The page tables are managed by the OS, typically by a \emph{virtual memory manager} (VMM).\footnote{Not to be confused with Virtual
+  Machine Monitor. We focus on non-hypervisor scenarios, but hardware virtualization extensions for both
+  x86-64 and ARM make use of an additional set of page tables translating what a \emph{guest} considers to be its
+  (virtualized) physical memory to actual physical memory. Our contributions should offer value in this scenario as well.}
+  Typically each process has its own
+  page table, which the OS registers with the CPU by storing the (page-aligned)
+  physical address of the root of the page table treethe start of the L4 table) in a
+  specific register (\texttt{cr3}) as part of switching to a new process.
+  Using different mappings, which map only disjoint portions of physical memory (with some exceptions in the next section)
+  is how the OS ensures memory isolation between processes.
+}
 
@@
 \rocq terms of a particular type, rather than being drawn from a fixed vocabulary.
-We introduce other details as they arise.
+\del{We introduce other details as they arise.}
 
@@
 is true in some possible future or past world related in a fixed way to the current world, but to talk about \emph{arbitrary}
-other worlds. This explicit naming of alternate worlds increases the power of modal logics~\cite{blackburn1995hybrid}, and is actually
+other worlds.
+
+\add{This is somewhat different from the accessibility relation modalities more common in verification, but also well-established.}
+This explicit naming of alternate worlds increases the power of modal logics~\cite{blackburn1995hybrid}, and is actually
 necessary for completeness in classical separation logics~\cite{brotherston2014parametric}.

### 2025-07-21 examples_C.tex
C pass through Ismail changes: adding change-tracking macros, fixing bad suggestions from grammar tool

@@
 the case for the next level already existing (Lines \ref{line:finalpieceS}-\ref{line:finalpieceE} in Figure \ref{fig:calltopteinitializeC}), which is discussed in Section \ref{sec:p2vC}.
-First, we must discuss another refinement of the address-space invariant, establishing
+First, we must discuss another refinement of the address space invariant, establishing
 enough structure on the page tables themselves to allow the traversal.
@@
 A call to this code begins by executing one thread (up to just before line \ref{line:end_saveC}) in an address space ($\rtv$), whose information will be saved in a structure at address $old$,
-and finishes the execution by executing a different thread in a different address-space (line \ref{line:end_restoreC} on), whose information is initially in $new$.
+and finishes the execution by executing a different thread in a different address space (line \ref{line:end_restoreC} on), whose information is initially in $new$.
 
@@
 \emph{in the other address space}. 
-The other address-space modality plays a critical role here; \textsf{Pother} would contain these assumptions in the other
+The other-space modality plays a critical role here; \textsf{Pother} would contain these assumptions in the other
 address space.

### 2025-07-21 programlogic.tex
C pass through Ismail changes: adding change-tracking macros, fixing bad suggestions from grammar tool

@@
 of virtual points-to assertions in preconditions, which guarantee that address translations succeed.
-This supports rules for updating not only typical data in memory that happens to be subject to address translation but also
-manipulation of the page tables themselves via virtual addresses (as demanded by all modern hardware) and also
+This supports rules for updating not only typical data in memory that happens to be subject to address translation, but \add{also}
+manipulation of the page tables themselves via virtual addresses (as demanded by all modern hardware) \replace{also}{and}
 via virtual points-to assertions.
 To support specifications that deal with multiple address spaces, our logic incorporates a hybrid-style modality
-$[r](P)$ to state that an assertion is true in another address space (assertion-specified) rather than the address space
-currently active in hardware, which is not only useful for virtual memory manager invariants, but also critical to reasoning
+$[r](P)$ to state that an assertion is true in another (assertion-specified) address space rather than the address space
+currently active in hardware, which is not only useful for virtual memory manager invariants, but \add{also} critical to reasoning
 about change of address space.
@@
 % \begin{enumerate}
-% \item \textit{address-spaces as modal contexts}: Assertions in our logic are context-dependent,
+% \item \textit{address spaces as modal contexts}: Assertions in our logic are context-dependent,
 %   in the sense that their truth depends on which address space they are used in, due to the need to support virtual points-to assertions.
@@
 %       in terms of physical (L4-L1) page-table memory accesses
-% \item \textit{context-agnostic-resources}: each virtual address is valid under a certain address-space, 
-%       but it does not represent this \textit{knowledge} of its address-space. 
+% \item \textit{context-agnostic-resources}: each virtual address is valid under a certain address space, 
+%       but it does not represent this \textit{knowledge} of its address space. 
 %       That is, assertions are not explicitly tagged with their address space validity
@@
 This is the (\rocq) type of assertions in our logic.
-Most constructs in \iris's base logic are defined with respect to any BI-algebra (of \coq type \lstinline|bi|), so they automatically
+Most constructs in \iris's base logic are defined with respect to any BI-algebra (of \coq type \lstinline|bi|), so \add{they} automatically
 carry over to our derived logic.
@@
 generally do not preserve the appealing reasoning principles of modal logics.
-This becomes more apparent when discussing our proof rules.
+This becomes \replace{most}{more} apparent when discussing our proof rules.
 }
@@
 \paragraph{Register points-to} 
-The assertion $\textsf{r}\;\mapsto_{\textsf{r}}\{q\}\;\textsf{rv}$ ensures the ownership of the register $\rg$ containing the value of the register $\rv$. 
-The fraction $\qfrac$ with value 1 asserts the unique ownership of the register mapping and grants update permission to it;
+The assertion $\textsf{r}\;\mapsto_{\textsf{r}}\{q\}\;\textsf{rv}$ ensures the ownership of the register $\rg$ containing the
+value of the register $\rv$.\todo{Ismail, you changed ``the register value $\rv$'' to ``value of the register $\rv$'' and I changed it back.
+What you wrote treats $\rv$ as the name of the register, not its value.}
+The fraction $\qfrac$ with value 1 asserts the unique ownership of the register mapping and grants update permission \replace{on}{to} it;
 otherwise, any value $0 < \qfrac <1$ represents partial ownership, granting read-only permission on the mapping.
@@
 chaining together the entries for each level, using the sequence of table offsets from the address being translated to index
-each table level, and using the physical page address embedded in each entry. \footnote{
+each table level, and using the physical page address embedded in each entry.\footnote{
   The fractions \lstinline|q1| through \lstinline|q4| represent the fractional ownership of each entry based on how many
@@
   and \textsf{l4off} with the virtual address being translated to compute the correct offset within that table.
-  Then, that physical location is the appropriate entry in the L4 table \textsf{l4e}.
+  Then \del{at} that physical location \replace{is}{holds} the appropriate entry in the L4 table \textsf{l4e}.
   Subsequent levels of the page table walk work similarly.
@@
   \textsc{GhostMapUpdate} says that ownership of the actual ghost map with ghost name $\gamma$ and map contents $\theta$,
-  and a full (nonfractional) token witnessing that $\theta$ maps \textsf{pa} to \textsf{va} permits an update to the ghost map's state,
+  and a full (non-fractional) token witnessing that $\theta$ maps \textsf{pa} to \textsf{va} permits an update to the ghost map's state,
   changing the map and replacing the token to represent the new value.
@@
 This new definition guarantees that the ghost map contains an mapping from the virtual address ($\vaddr$) to a physical address ($\paddr$),
-and thus that the per-address-space invariant will contain the physical resources that guarantee that the hardware resolves the translation.
+and thus that the per-address-space invariant will contain the physical resources that guarantee \add{that} the hardware resolves the translation.
 
@@
 \vspace{-1em}
-\caption{Per-address-space invariant with a fixed global map of address-space names $m$}
+\caption{Per-address-space invariant with a fixed global map of address space names $m$}
   \label{fig:peraspaceinvariant}
@@
 When combined with the
-fragment stored in virtual points-to (Figure \ref{fig:virtualpointstosharing}),
+fragment stored in the virtual points-to (Figure \ref{fig:virtualpointstosharing}),
 changes in the virtual-to-physical translations become possible without the need to track all virtual points-to assertions.
@@
 $\theta$ is the logical map from virtual addresses to the physical addresses they should translate to, 
-for the \emph{current} address space, corresponding to the page tables currently installed (the ones
+for the \emph{current} address space, corresponding to the  currently installed page tables (the ones
 indicated by \lstinline|cr3|). In Figure \ref{fig:virtualpointstosharing}, the existentially quantified $\delta$
@@
 it must ensure that $\theta$ actually corresponds to the intended ghost map for the current address space.
-This requires looking up the ghost name (i.e. $\delta$) for the current address space,
+This requires looking up the ghost name (i.e., $\delta$) for the current address space,
 which is handled by $m$ --- another logical map (tied to ghost state shortly)
@@
 
-\subsection{Address-Space Management}
+\subsection{Address Space Management}
 \label{sec:aspacemanagement}
@@
 Real VMMs must
- handle more than one address-space.
+ handle more than one address space.
 Doing so requires a way to talk about other address spaces, and means to switch address spaces.
@@
 the \emph{other-space} modality. The definition itself is not
-particularly surprising --- as our modal assertions are semantic predicates on a page table root (physical)
+particularly surprising --- as our modal assertions are semantically predicates on a page table root (physical)
 address, the assertion $[r](P)$ is a modal assertion that ignores the (implicit) current page table root,
@@
 We can prove that this modality follows certain basic laws, showing that its truth is independent of the address
-space in which it is considered, that it is distributed over various logical connectives, and that it follows the rule of
+space in which it is considered, \add{that} it distributes over various logical connectives, and \add{that} it follows the rule of
 consequence.
 We call \textsf{vProp} assertions whose truth is independent of the current address space
-\textsf{Fact}s; these include other address-space assertions, physical memory points-tos, and register assertions.
-\textsf{Fact}s can freely move in and out of other address-space modalities.
+\textsf{Fact}s; these include other-space assertions, physical memory points-tos, and register assertions.
+\textsf{Fact}s can \add{freely} move in and out of other address space modalities\del{ freely}.
 \looseness=-1
@@
 soundness of an individual virtual memory access.
-However, such larger invariants would contain many kernel-specific properties that are orthogonal
+\replace{But}{However,} such larger invariants would contain many kernel-specific properties that are orthogonal
 to the fundamental reasoning principles that are the focus of this paper.
@@
 \textsf{WP} is \iris's own weakest precondition modality, unmodified~\cite{jung2018iris}.
-Making \textsf{rtv} a parameter to the double (versus a simple registration assertion)
+Making \textsf{rtv} a parameter to the double (vs.\ a simple register assertion)
 makes it possible to ensure ownership of the \lstinline|cr3| register and its value is accounted for
@@
 conclusion). 
-It reflects the loading of that memory value into the destination register $\textsf{r}_d$, with
+It reflects the load (\lstinline|mov|) of that memory value into the destination register $\textsf{r}_d$, with
 the updated register points-to in the precondition for $\overline{is}$.
@@
 Unlike other rules, \TirNameStyle{WriteToRegCtlFromRegModal} updates the root address of the 
-address-space determining the validity of resources, from $\rtv$ before the
+address space determining the validity of resources, from $\rtv$ before the
 \lstinline|mov| to $\textsf{rvs}$ afterwards. The global effects of this rule are reflected in moving
-\del{bare} assertions \add{for the current address space ($P$ and $\mathcal{I}\mathsf{ASpace}$)} under an other address-space modality for \add{the initial
+\del{bare} assertions \add{for the current address space ($P$ and $\mathcal{I}\mathsf{ASpace}$)} under an other-space modality for \add{the initial
 (outgoing) address space} $\rtv$, and moving the new address space's assertions out of
@@
 The essential problem is that the frame rule is motivated by local reasoning about local updates, but
-a switch of address-space is a \emph{global} change that may invalidate information about virtual addresses.
+a switch of address space is a \emph{global} change that may invalidate information about virtual addresses.
 Thus, framing around arbitrary \lstinline|cr3| updates is unsound --- hence the \emph{global} nature of \textsc{WriteToRegCtlFromRegModal} ---

### 2025-07-21 relwork.tex
C pass through Ismail changes: adding change-tracking macros, fixing bad suggestions from grammar tool

@@
 but the work never addressed the issues of managing multiple address spaces, switching address spaces, or
-invariants referring to other address spaces, which our explicitly modal approach and case studies address.
+invariants referring to other address spaces, which our explicitly modal approach and case studies do address.
 They verified a fraction of an ARM version of our Figure \ref{fig:mapping_code}, which assumed that they already had the virtual address

### 2025-07-21 soundness.tex
+
C pass through Ismail changes: adding change-tracking macros, fixing bad suggestions from grammar tool

@@
 As a last definition, we give the head step relation required for the Iris instantiation for our simple language in Figure \ref{} to sequence instructions.
-This relation allows lifting the program expression to enable the application of changes imposed by the operational semantics on the program state (for a certain cpu, a register map, and a selected memory for an address-space) when applied for an insturction (\textsf{i}).
+This relation allows lifting the program expression to enable the application of changes imposed by the operational semantics on the program state (for a certain cpu, a register map, and a selected memory for an address space) when applied for an insturction (\textsf{i}).
 \begin{lstlisting}[language=Coq]
@@
 \inferrule{
-  \{P \ast r_d \mapsto_{r}  \textsf{v} \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{v}} \textsf{v} \}_{\textsf{rtv}}\;\overline{is}
+  \{\mathsf{P} \ast \mathsf{r}_d \mapsto_{r}  \mathsf{v} \ast \mathsf{r}_a \mapsto_{r} \{\mathsf{q}\} \; \mathsf{ vaddr} \ast \mathsf{vaddr} \mapsto_{\mathsf{v}} \mathsf{v} \}_{\mathsf{rtv}}\;\overline{is}
 }{
-  \{P \ast r_d \mapsto_{r}  \textsf{rvd} \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{v}} \textsf{v} \}_{\textsf{rtv}}
-  \textsf{ mov}~\textsf{r}_d,~\textsf{r}_a;\;\overline{is}
+  \{\mathsf{P} \ast \mathsf{r}_d \mapsto_{r}  \mathsf{rvd} \ast \mathsf{r}_a \mapsto_{r} \{\mathsf{q}\} \;\mathsf{ vaddr} \ast \mathsf{vaddr} \mapsto_{\mathsf{v}} \mathsf{v} \}_{\mathsf{rtv}}
+\; \mathsf{ mov}~\mathsf{r}_d~\mathsf{r}_a;\;\overline{is}
 }

### 2025-07-21 syntaxsemantics.tex
C pass through Ismail changes: adding change-tracking macros, fixing bad suggestions from grammar tool

@@
 \label{sec:syntax}
-To develop our core logical ideas, we instantiate \iris with a simple language for a stream of instructions
-and a logical machine model corresponding to the execution of x86-64 assembly instructions with virtual memory enabled on the
+To develop our core logical ideas, we instantiate \iris with a simple language for streams of instructions
+and a logical machine model corresponding to execution of x86-64 assembly instructions with virtual memory enabled on the
 CPU.
@@
 % Programs we demonstrate in this paper requires accessing two types of computer resource: registers and memory.
-A register identifier, $\reg$, is chosen from a fixed finite set of register identifiers, $\regset$.
-\footnote{\add{$\reg$ abbreviates \emph{general register}, which includes integer registers (e.g., \lstinline|rax|),
+A register identifier, $\reg$, is chosen from a fixed finite set of register identifiers, $\regset$.\footnote{\add{$\reg$ abbreviates \emph{general register}, which includes integer registers (e.g., \lstinline|rax|),
 control registers (e.g., \lstinline|cr3|), and segmentation registers (which we do not discuss in detail,
@@
 (destination on the left, source on the right).
-Thus, a store to the memory is $\textsf{mov}~[r_m]~r_r$, and a load from memory is $\textsf{mov}~r_r~[r_m]$.
+Thus, a store to memory is $\textsf{mov}~[r_m]~r_r$, and a load from memory is $\textsf{mov}~r_r~[r_m]$.
 % \[

### 2025-07-21 background.tex
slightly condense new figure 2

@@
   pte_t *next;
-  /*Reading page table entry:
-    pointed by a virtual memory address*/
-  /* Jump if the present bit is not zero:
-     no need to allocate next level (* \label{line:check_entry_presentC} *)(* \label{line:mask_presentC} *)  (*\label{line:after_concluding_qfrac1C}*)*/
+  /* Check if entry is populated, allocate if needed(* \label{line:check_entry_presentC} *)(* \label{line:mask_presentC} *)(*\label{line:after_concluding_qfrac1C}*)*/
   if (!entry->present(*\label{line:read_entry_contentsC}*) (*\label{line:conditional_childrenC}*)) {
-    /*Allocation path starts here (*\label{line:pass_addrof_nextC}*)(*\label{line:alloc_path_startC}*)*/
-    pte_initialize(entry);(* \label{line:call_to_pte_initializeC} *)
-    /*Entry value updates:*/
-    entry->pfn = nextpaddr;
-    entry->present = 1; (*\label{line:install_new_entryC}*)
-    /*Entry is initialized,
-      so we satisfy the condition to access children list */
+    pte_initialize(entry); // Alloc empty table for next level (*\label{line:pass_addrof_nextC}*)(*\label{line:alloc_path_startC}*)(* \label{line:call_to_pte_initializeC} *)
+    entry->pfn = nextpaddr; // Point to next table
+    entry->present = 1; // Mark valid (*\label{line:install_new_entryC}*)
+    /* Key invariant for child access now holds */
   } (*\label{line:end_of_allocation_pathCB}*)

### 2025-07-21 examples_C.tex
C WIP on examples rewrite, fresh intro, figure tidying, various TODOs

@@
 We build up to the main task of mapping a new page after traversing the page tables in the software.
-The mapping operation of Figure \ref{fig:mapping_codeC} assumes an operation \textsf{walkpgdir} which must traverse the page tables
-in order to locate the address of the L1 entry to update --- 
-% possibly allocating tables for levels 3, 2, and 1 in the process,
-% installing them into levels 4, 3, and 2, along the way.
-possibly allocating new L3, L2, and L1 tables as necessary.
-Traversing the page tables is itself challenging functionality to verify: loading the current table root from \lstinline|cr3| is straightforward
-(a \lstinline|mov| instruction); however, this produces the physical address of \lstinline|cr3|, not the virtual address the kernel code would use to access that memory.
+This algorithm is complex and corresponds to a significant amount of assembly code.
+To assist with readability, we present C code for this process, with assertions adjusted slightly to refer to
+C program variables rather than registers. The actual verification was carried out on x86-64 assembly
+\emph{generated from this source code}.
+Listings of the assembly fragments with inline assertions appear in Appendix ???. %\ref{apdx:asm}.
+\todo[inline]{Need to add the assembly appendix.}
+Whether in C or assembly,
+the page table traversal involved in mapping a new page is very challenging functionality to verify.
+Loading the current table root from \lstinline|cr3| is straightforward (a \lstinline|mov| instruction).
+However, this produces the \emph{physical} address stored in \lstinline|cr3|, not a \emph{virtual} address the kernel code can use to access that memory.
 This problem repeats at each level of the page table: assuming that the code has \emph{somehow} read the appropriate L4 (or L3, or L2) entry, those entries again
 yield physical addresses, not virtual.
+The only prior work to verify page mapping ignored the traversal and only verified mapping
+assuming code \emph{already} had an appropriate virtual address for the L1 entry, where a physical
+address could simply be stored. Our proof is the first to additionally deal with the critical code
+leading up to that point.
 
-\subsubsection{Loading Page-Table Address Value}
+\paragraph{Code Overview}
+As described in Section \ref{sec:background}, mapping a new page consists of 
+simulating the hardware address translation of Figure \ref{fig:pagetables}, but in software.
+The code for this task takes three explicit parameters:
+the root pointer (read from \lstinline|cr3| by earlier code),
+the page-aligned \emph{virtual} address (\lstinline|va|) at which to make a new piece of memory accessible,
+and the \emph{physical} address (\lstinline|fpaddr|) of the memory to map in that location.
+The function we ultimately verify, \lstinline|vaspace_mappage| (Figure \ref{fig:mapping_codeC}),
+relies primarily on a helper function already shown in Figure \ref{fig:pagetablescode}.
+\lstinline|walkpgdir| finds the (virtual) address of the the correct L1 entry to translate \lstinline|va|,
+by walking the page tables in software one level at a time.
+\lstinline|vaspace_mappage| then uses the result to install the new entry.
+\lstinline|walkpgdir| itself relies on its own helper function \lstinline|pte_get_next_table|, also shown in Figure \ref{fig:pagetablescode},
+which implements a single-level of traversal from level $n+1$ to level $n$ (and whose specification and proof are therefore
+parameterized by page table level), allocating additional levels as needed.
+
+We organize our explanation of the proofs by essentially following execution from the start of \lstinline|walkpgdir|, through
+execution of \lstinline|pte_get_next_table|, and out to its callsite in \lstinline|vaspace_mappage|.
+While slightly awkward because we start in the middle of the mapping execution,
+this ordering allows us to start with the simpler pieces of the proof, and incrementally explain the complexities
+of the proofs and kernel invariant, before concluding with the top-level verification.
+
+% The mapping operation of Figure \ref{fig:mapping_codeC} assumes an operation \textsf{walkpgdir} which must traverse the page tables
+% in order to locate the address of the L1 entry to update --- 
+% % possibly allocating tables for levels 3, 2, and 1 in the process,
+% % installing them into levels 4, 3, and 2, along the way.
+% possibly allocating new L3, L2, and L1 tables as necessary.
+
+\todo[inline]{Colin WIP note: The organization is odd here.
+We talk through access to entry, then start talking about identity mappings even though we don't need that info
+until the P2V call later. But we do need the part of the extended invariant that deals with the
+entry-present qfrac bit, which does only hold for those addresses used for page tables, so is part
+of the same invariant. Maybe we should reorg to first talk through the logic, 
+calling out the two problems the new invariant solves (permission to update the entry and p2v) along the way,
+then have a section called "extended address space invariant" or similar to talk about the solutions.
+This would roughly mean moving 5.1.3 (installing a new table) up
+}
+\todo[inline]{this and other figures need assertion updates to talk about C: see entry\_val (needs adjustment to points-to preconditions
+and explanation of entry pointing to a C bitfield so entry\_val encompasses the various bits), nextpaddr
+(was logical name in asm proof, but not a C variable)
+}
+\todo[inline]{need to narrate the main parts of assertions in prose}
+
+\subsubsection{From L$n+1$ Entries to L$n$ Tables}
+\hrule
 We will discuss access to the level 4 table later (Section \ref{wlkpgdirC}). However, for subsequent levels, the base address of level $n$ must be
 fetched from the appropriate entry in the table of level $n+1$.
-This is the role of \lstinline|pte_get_next_table| (Figure \ref{fig:calltopteinitializeC}):
-%and \ref{fig:p2vC}):
-it is passed the virtual address of the page table entry in level $n+1$, and should return the \emph{virtual} 
+This is the role of \lstinline|pte_get_next_table| (originally Figure \ref{fig:pagetablescode}, with proof details in Figure \ref{fig:calltopteinitializeC}).
+It is passed the virtual address of the page table entry in level $n+1$, and should return the \emph{virtual} 
 address of the \emph{base} of the level $n$ table
@@
 \ifPLDI
-\begin{lstlisting}[mathescape,escapeinside={(*}{*)}]
-/*
-@param entry: entry whose frame address is used to locate the a-level-lower entry in the page-tree
-@return next: return a virtual pointer to the descendant tree
-*/
+\begin{lstlisting}[language=C,mathescape,escapeinside={(*}{*)}]
+/* @param entry: virtual address of level n+1 entry which should point to the base of a level n table
+   @return next: a virtual pointer to the base of a valid level n table */
 pte_t *pte_get_next_table(pte_t *entry) {
@@
   $\specline{\ulcorner\textsf{entry\_present(entry\_val)}\urcorner \wand \forall_{i\in\textsf{0..511}} \; \ghostmaptoken{\textsf{id}}{\textsf{table\_root(entry\_val.pfn) + i * 8}}{\textsf{v-1}} }_{\rtv} $
-  /*Reading page table entry justified by virtual pte-pointsto*/
-  if (!entry->present(*\label{line:read_entry_contentsC}*)(*\label{line:conditional_childrenC}*)(*\label{line:check_entry_present_jumpC}*)) {
-    /* Jump if the present bit is not zero; no need to allocate next level (* \label{line:check_entry_presentC} *)(* \label{line:mask_presentC} *) */
-    /*Allocation path starts here (*\label{line:pass_addrof_nextC}*)(*\label{line:alloc_path_startC}*)*/
-    $\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \land \lnot(\textsf{entry\_present entry\_val})  \urcorner }_{\rtv}$ (*\label{line:after_concluding_qfrac1C}*)
-    pte_initialize(entry);(* \label{line:call_to_pte_initializeC}(*\label{line:page_of_capsC}*) *)
+  /* Reading the page table is entry justified by the virtual pte-points-to */
+  if (!entry->present(*\label{line:read_entry_contentsC}*)(*\label{line:conditional_childrenC}*)(*\label{line:check_entry_present_jumpC}*)) {// If the present bit is zero, need to allocate next level (* \label{line:check_entry_presentC} *)(* \label{line:mask_presentC} *)
+    (*\label{line:pass_addrof_nextC}*)(*\label{line:alloc_path_startC}*)$\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \land \lnot(\textsf{entry\_present entry\_val})  \urcorner }_{\rtv}$ (*\label{line:after_concluding_qfrac1C}*)
+    pte_initialize(entry); // Allocate a zeroed physical page and install in entry(*\label{line:call_to_pte_initializeC}\label{line:page_of_capsC}*)
     $\specline{ \ldots \ast \ulcorner  \textsf{qfrac} = 1 \land \lnot(\textsf{entry\_present entry\_val})  \urcorner \ast \ulcorner  \textsf{qfrac} = 1 \land \lnot(\textsf{entry\_present (pfn\_set (entry\_val nextpaddr))}  \urcorner   }_{\rtv}$
@@
 511} } \ldotp  \ghostmaptoken{\textsf{id}}{((\textsf{table\_root (pte\_initialized (pfn\_set(entry\_val nextpaddr)))}) + \textsf{i * 8})}{\textsf{v-1}} \end{array}  }$ (*\label{line:page_of_capsC}*)
-    /*Entry value updates:*/
-    entry->pfn = nextpaddr; entry->present = 1; (*\label{line:install_new_entryC}*)
+    entry->pfn = nextpaddr; entry->present = 1; // Install newly-allocated level n table, mark present(*\label{line:install_new_entryC}*)
     /*Now we know that entry is initialized, so we satisfy the condition to access children list */
@@
 % \end{lstlisting}
+\todo[inline]{this figure has an issue with entry, which is used both as a logical variable and a program variable (e.g., assertion on line 6)}
 \vspace{-1em}
-\caption{Ensuring \textsf{entry} points to a valid next table, allocating if necessary.}
+\caption{Ensuring \textsf{entry} points to a valid next-level table, allocating if necessary, returning its virtual address.}
 \label{fig:calltopteinitializeC}
@@
   \begin{lstlisting}[mathescape,escapeinside={(*}{*)}]
-/*
-@param entry: page entry whose frame is initialized with a new physical page 
-*/
+/* @param entry: virtual address of a non-present entry to initialize with a new physical page */
 $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \lnot(\textsf{entry\_present entry\_val} }_{\rtv}$
@@
 void pte_initialize(pte_t *entry) {
-  /*Allocate a full zeroed page for 512 8-byte entries(*\label{line:call_to_kallocC}*)*/
-  pte_t *local = kalloc();
+  pte_t *local = kalloc(); // Allocate a full zeroed page for 512 8-byte entries(*\label{line:call_to_kallocC}*)
   entry->pfn = PTE_ADDR_TO_PFN((uintptr_t) local);
@@
   $\specline{\ghostmaptoken{\delta{}s}{\rtv}{\delta} \ast \begin{array}{l} \ulcorner\textsf{entry\_present (pte\_initialize(pfn\_set(entry\_val nextpaddr)),level)}\urcorner \wand \\  \;\;\;\;\;\;\; \forall_{i\in\textsf{0..511}} \; \textsf{table\_root (pte\_initialized (pfn\_set (entry\_val nextpaddr))) + i * 8} \mapsto_{\textsf{id}} \textsf{level-1}    \end{array}  }_{\rtv}$
-  return;
 }

### 2025-07-21 examples_C.tex
WIP checkpoint

@@
 
-\subsubsection{From L$n+1$ Entries to L$n$ Tables}
-\hrule
-We will discuss access to the level 4 table later (Section \ref{wlkpgdirC}). However, for subsequent levels, the base address of level $n$ must be
-fetched from the appropriate entry in the table of level $n+1$.
-This is the role of \lstinline|pte_get_next_table| (originally Figure \ref{fig:pagetablescode}, with proof details in Figure \ref{fig:calltopteinitializeC}).
-It is passed the virtual address of the page table entry in level $n+1$, and should return the \emph{virtual} 
-address of the \emph{base} of the level $n$ table
-indicated by that entry.
-If the entry is empty (i.e., this is a sparse part of the page table representation),
-the code also allocates a page for the level $n$ table, installs it in the level $n+1$ entry, and establishes appropriate invariants.
-Figure \ref{fig:calltopteinitializeC} presents the initial part of the function, which performs the allocation if necessary.
-Lines \ref{line:finalpieceS}-\ref{line:finalpieceE} in Figure \ref{fig:calltopteinitializeC} (discussed in Section \ref{sec:p2vC}) deal with the cases where no allocation is necessary \emph{or} the allocation has already
-been performed by the code in this figure.
-\looseness=-1
-
-Note that the specification does \emph{not} assume a specific page table level --- the logical parameter \textsf{v} represents the level
-of the entry passed as an argument, and this code
-is used for all three level transitions when traversing page tables (4 to 3, 3 to 2, 2 to 1).
-This comes into play with a subtlety of the specification of \lstinline|pte_get_next_table| that we will
-revisit several times: \lstinline|pte_get_next_table|'s specification
-assumes it is given a virtual \emph{vpte-pointsto}
-(a virtual points-to exposing the underlying physical address instead of existentially quantifying it;
- see Section \ref{sec:mapnewC}) granting access to the specified entry,
-but its postcondition does not yield new virtual points-to assertions!
-Instead it merely computes the base virtual address of the next table, and returns adequate capabilities (discussed in Section \ref{subsec:identitymappingsC})
-for the \emph{caller} to construct a vpte-pointsto for any entry of the next table level (if this is not an L1 entry ---
-the caller knows which level of the table this is for).
-\looseness=-1
-
-Within \textsf{get\_next\_table}, after a standard function prologue, the code
-loads the entry pointed to by the argument (logical variable \textsf{entry} in the proof outline).
-This is a page table entry: a 64-bit word divided into bit-fields for
-the physical address of the next table, and control bits like the valid bit, as discussed in 
-Section \ref{sec:backgroundonmachinemodel}.
-
-\ifPLDI
-Line \ref{line:conditional_childrenC} checks % In the condensed figure, it's all on one line
-\else
-Lines \ref{line:mask_present}--\ref{line:check_entry_present} check
-\fi
-if the entry's ``present'' bit is set.
-If it is zero, a new page must be allocated for the next level of the table. Otherwise the code jumps ahead to
-the case for the next level already existing (Lines \ref{line:finalpieceS}-\ref{line:finalpieceE} in Figure \ref{fig:calltopteinitializeC}), which is discussed in Section \ref{sec:p2vC}.
-First, we must discuss another refinement of the address space invariant, establishing
-enough structure on the page tables themselves to allow the traversal.
-The code for allocating a new level of the page table must establish this extended invariant.
-
-%wshiftll (wshiftll (natToWord 64 entry) (WordImpl.concat (WordImpl.zero 56) (WordImpl.from_nat 8 12 ^& WordImpl.concat (WordImpl.zero 2) WO~1~1~1~1~1~1)) ^& constf)
-%(WordImpl.concat (WordImpl.zero 56) (natToWord 8 12 ^& WordImpl.concat (WordImpl.zero 2) WO~1~1~1~1~1~1))
-%
-%wshiftll
- %      (wshiftll
-%          ((((natToWord 64 entry ^& WordImpl.concat (WordImpl.zero 32) consta ^| WordImpl.concat (WordImpl.zero 32) (natToWord 32 2))
-%             ^& WordImpl.concat (WordImpl.zero 32) constb ^| WordImpl.concat (WordImpl.zero 32) (natToWord 32 4)) ^& constd
-%            ^| wshiftll
-%                 (wshiftll (nextpaddr ^+ ^~ (natToWord 64 KERNBASE))
-%                    (WordImpl.concat (WordImpl.zero 56) (WordImpl.from_nat 8 12 ^& WordImpl.concat (WordImpl.zero 2) WO~1~1~1~1~1~1))
-%                  ^& constf)
-%                 (WordImpl.concat (WordImpl.zero 56) (WordImpl.from_nat 8 12 ^& WordImpl.concat (WordImpl.zero 2) WO~1~1~1~1~1~1)))
-%           ^& WordImpl.concat (WordImpl.zero 32) conste ^| wone 64)
-%          (WordImpl.concat (WordImpl.zero 56) (WordImpl.from_nat 8 12 ^& WordImpl.concat (WordImpl.zero 2) WO~1~1~1~1~1~1)) ^& constf)
-%       (WordImpl.concat (WordImpl.zero 56) (natToWord 8 12 ^& WordImpl.concat (WordImpl.zero 2) WO~1~1~1~1~1~1)) 
 \begin{figure}\footnotesize
@@
   } (*\label{line:end_of_allocation_pathC}*)
-  /*Physical to virtual address mapping code segment starts here (*\label{line:finalpieceS}*)*/
   $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \}),m) \ast \textsf{entry} \mapsto_{\textsf{id}} \textsf{\_} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
@@
 511} } \ldotp  \ghostmaptoken{\textsf{id}}{((\textsf{table\_root (pte\_initialized (entry\_val.pfn)))}) + \textsf{i * 8})}{\textsf{v-1}}  }$ (*\label{line:childrenC}*)
-  uintptr_t next_phys_addr = PTE_PFN_TO_ADDR(entry->pfn)
-  uintptr_t next_virt_addr = (uintptr_t) P2V(next_phys_addr); (*\label{line:p2vC}*);
+  uintptr_t next_phys_addr = PTE_PFN_TO_ADDR(entry->pfn); // Fetch physical address of next table (*\label{line:finalpieceS}*)(*\label{line:extract_pfn}*)
+  uintptr_t next_virt_addr = (uintptr_t) P2V(next_phys_addr); // Convert to virtual address (*\label{line:p2vC}*)
   next = (pte_t *) next_virt_addr;(*\label{line:finalpieceE}*)
@@
 % \end{lstlisting}
-\todo[inline]{this figure has an issue with entry, which is used both as a logical variable and a program variable (e.g., assertion on line 6)}
+\todo[inline]{this figure has an issue with entry, which is used both as a logical variable and a program variable (e.g., assertion on line 6).
+I would also expect to see a more prominent use of \textsf{v} in the precondition
+}
 \vspace{-1em}
@@
 
-\subsubsection{Identity Mappings}
+\subsubsection{From L$n+1$ Entries to L$n$ Tables}
+We will discuss access to the level 4 table later (Section \ref{wlkpgdirC}). However, for subsequent levels, the base address of level $n$ must be
+fetched from the appropriate entry in the table of level $n+1$.
+This is the role of \lstinline|pte_get_next_table| (originally Figure \ref{fig:pagetablescode}, with proof details in Figure \ref{fig:calltopteinitializeC}).
+It is passed the virtual address of the page table entry in level $n+1$, and should return the \emph{virtual} 
+address of the \emph{base} of the level $n$ table
+indicated by that entry.
+If the entry is empty (i.e., this is a sparse part of the page table representation),
+the code also allocates a page for the level $n$ table, installs it in the level $n+1$ entry, and establishes appropriate invariants.
+Figure \ref{fig:calltopteinitializeC} presents the function with proof annotations that we will explain shortly, but we first explain the functionality.
+\lstinline|pte_get_next_table| accepts a \emph{virtual} address \lstinline|entry| which points to a level-$n+1$
+table entry. On Line \ref{line:read_entry_contentsC}, the code checks the present bit of the entry.
+If the bit is unset, there is no level-$n$ table, so one must be allocated via \lstinline|pte_initialize| (explained shortly\footnote{
+  Briefly, it allocates a fresh physical page, and initializes the memory pointed to by \lstinline|entry| with that physical address.
+}) and marked present.
+By Line \ref{line:finalpieceS} the entry is known to be valid and contain the physical address of
+the base of a level $n$ table. That address is then extracted (Line \ref{line:extract_pfn}),
+converted to a virtual address (Line \ref{line:p2vC}), and returned to the caller.
+
+We can now discuss \lstinline|pte_get_next_table|'s proof of correctness, which is surprisingly subtle for two
+local reasons, and a third related to interaction with the caller, none of which have been addressed by prior work.
+First, the update to the memory at (virtual) address \lstinline|entry| depends on subtle ownership invariants:
+if the entry is present then its fractional ownership is shared with a large number of \lstinline|L4_L1_PointsTo| assertions
+from the address space invariant (Figure \ref{fig:peraspaceinvariant}),
+but if the entry is absent the proof requires full ownership to update it. We resolve this by extending the address space invariant
+to make the owned fraction of the entry's memory \emph{dependent on its own contents}.
+Second, the conversion of physical addresses into a corresponding virtual address that can be used to modify the specific
+physical location relies on subtle, never-before-formalized kernel invariants.
+And third, these two factors percolate to the precondition (for conditional fractional ownership) and postcondition
+(for physical-to-virtual conversions), as the caller essentially passes output from one call to \lstinline|pte_get_next_table|
+as input to a subsequent call to traverse multiple levels.
+Since the key to solving these challenges is to extend the address space invariant, we
+first discuss that invariant and the kernel designs it supports, before returning to the subtle details of
+verifying lines \ref{line:install_new_entryC} and \ref{line:p2vC}.
+
+\subsubsection{An Extended Address Space Invariant: Identity Mappings and Conditional Ownership Page Table Ownership}
 \label{subsec:identitymappingsC}
+
+\ldots
+\hrule\hrule
+
 Kernels need to convert between physical and virtual addresses, in both directions.
@@
 % in Line 5 and compute the identity mapping for this physical frame address in Line 13 in Figure \ref{fig:p2v}).
+
+\subsubsection{The Specification of \lstinline|pte_get_next_table|}
+\hrule
+
+Note that the specification does \emph{not} assume a specific page table level --- the logical parameter \textsf{v} represents the level
+of the entry passed as an argument, and this code
+is used for all three level transitions when traversing page tables (4 to 3, 3 to 2, 2 to 1).
+This comes into play with a subtlety of the specification of \lstinline|pte_get_next_table| that we will
+revisit several times: \lstinline|pte_get_next_table|'s specification
+assumes it is given a virtual \emph{vpte-pointsto}
+(a virtual points-to exposing the underlying physical address instead of existentially quantifying it;
+ see Section \ref{sec:mapnewC}) granting access to the specified entry,
+but its postcondition does not yield new virtual points-to assertions!
+Instead it merely computes the base virtual address of the next table, and returns adequate capabilities (discussed in Section \ref{subsec:identitymappingsC})
+for the \emph{caller} to construct a vpte-pointsto for any entry of the next table level (if this is not an L1 entry ---
+the caller knows which level of the table this is for).
+\looseness=-1
 

### 2025-07-22 examples_C.tex
major revision of explanation up through extended invariant, with concrete plans up through (not including) walkpgdir discussion

@@
   pte_t *next;
-  $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{entry}\},m) \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
+  $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{entry}\},m) \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$(*\label{line:precondition_entry_out}*)
   $\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present entry\_val})\urcorner}_{\rtv}$ (*\label{line:get_next_vpte_preconditionC}*)
-  $\specline{\ulcorner\textsf{entry\_present(entry\_val)}\urcorner \wand \forall_{i\in\textsf{0..511}} \; \ghostmaptoken{\textsf{id}}{\textsf{table\_root(entry\_val.pfn) + i * 8}}{\textsf{v-1}} }_{\rtv} $
+  $\specline{\ulcorner\textsf{entry\_present(entry\_val)}\urcorner \wand \forall_{i\in\textsf{0..511}} \; \ghostmaptoken{\textsf{id}}{\textsf{table\_root(entry\_val.pfn) + i * 8}}{\textsf{v-1}} }_{\rtv} $(*\label{line:precondition_conditional}*)
   /* Reading the page table is entry justified by the virtual pte-points-to */
   if (!entry->present(*\label{line:read_entry_contentsC}*)(*\label{line:conditional_childrenC}*)(*\label{line:check_entry_present_jumpC}*)) {// If the present bit is zero, need to allocate next level (* \label{line:check_entry_presentC} *)(* \label{line:mask_presentC} *)
-    (*\label{line:pass_addrof_nextC}*)(*\label{line:alloc_path_startC}*)$\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \land \lnot(\textsf{entry\_present entry\_val})  \urcorner }_{\rtv}$ (*\label{line:after_concluding_qfrac1C}*)
+    (*\label{line:refined_fractional_ownership}*)(*\label{line:pass_addrof_nextC}*)(*\label{line:alloc_path_startC}*)$\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \land \lnot(\textsf{entry\_present entry\_val})  \urcorner }_{\rtv}$ (*\label{line:after_concluding_qfrac1C}*)
     pte_initialize(entry); // Allocate a zeroed physical page and install in entry(*\label{line:call_to_pte_initializeC}\label{line:page_of_capsC}*)
@@
     entry->pfn = nextpaddr; entry->present = 1; // Install newly-allocated level n table, mark present(*\label{line:install_new_entryC}*)
-    /*Now we know that entry is initialized, so we satisfy the condition to access children list */
+    /*Now we know that entry is initialized, so we satisfy the condition to access children list */(*\label{line:now_we_know}*)
     $\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}}  \textsf{pte\_initialized(pfn\_set(entryv nextpaddr))} \ast \ulcorner  \textsf{qfrac} = 1 \land \lnot(\textsf{entry\_present entry\_val})  \urcorner }_{\rtv}$
     $\specline{\forall_{i\in \textsf{0 ...
-511} } \ldotp  \ghostmaptoken{\textsf{id}}{((\textsf{table\_root (pte\_initialized (pfn\_set(entry\_val nextpaddr)))}) + \textsf{i * 8})}{\textsf{v-1}}  }$ (*\label{line:alloc_path_endC}*)
+511} } \ldotp  \ghostmaptoken{\textsf{id}}{((\textsf{table\_root (pte\_initialized (pfn\_set(entry\_val nextpaddr)))}) + \textsf{i * 8})}{\textsf{v-1}}  }$(*\label{line:available_child_tokens}*) (*\label{line:alloc_path_endC}*)
   } (*\label{line:end_of_allocation_pathC}*)
@@
 converted to a virtual address (Line \ref{line:p2vC}), and returned to the caller.
+We can now discuss \lstinline|pte_get_next_table|'s proof of correctness.
+While at first glance this code does not sound terribly complex, beyond needing some care to distinguish physical and virtual addresses,
+it has a highly nontrivial correctness argument, which depends critically on detailed invariants on how access to page table
+entries is shared between parts of the kernel. No prior work has engaged with this problem.
 
-We can now discuss \lstinline|pte_get_next_table|'s proof of correctness, which is surprisingly subtle for two
-local reasons, and a third related to interaction with the caller, none of which have been addressed by prior work.
+\subsubsection{Address Space Invariant: Identity Mappings and Conditional Ownership Page Table Ownership}
+\label{subsec:identitymappingsC}
+Assembly-level verification of compiler output from Figure \ref{fig:calltopteinitializeC} is verbose but doable
+Section \ref{sec:logic}'s logic, but only after resolving two critical challenges.\todo{Ismail, not sure if this ``verbose but doable''
+ is a good framing, we should discuss
+}
+Two key challenges stand out and end up affecting both the pre- and post-conditions, neither of which has been addressed by prior work.
 First, the update to the memory at (virtual) address \lstinline|entry| depends on subtle ownership invariants:
@@
 physical location relies on subtle, never-before-formalized kernel invariants.
-And third, these two factors percolate to the precondition (for conditional fractional ownership) and postcondition
-(for physical-to-virtual conversions), as the caller essentially passes output from one call to \lstinline|pte_get_next_table|
-as input to a subsequent call to traverse multiple levels.
+% \looseness=-1
+%
+% These two factors percolate to the precondition (for conditional fractional ownership) and postcondition
+% (for physical-to-virtual conversions), as the caller essentially passes output from one call to \lstinline|pte_get_next_table|
+% as input to a subsequent call to traverse multiple levels.
 Since the key to solving these challenges is to extend the address space invariant, we
@@
 
-\subsubsection{An Extended Address Space Invariant: Identity Mappings and Conditional Ownership Page Table Ownership}
-\label{subsec:identitymappingsC}
-
-\ldots
-\hrule\hrule
-
+\paragraph{Physical-to-Virtual Mappings}
 Kernels need to convert between physical and virtual addresses, in both directions.
-Traversing the page tables in software is the simplest way to convert a virtual address to a physical address; this is the context we are working up to.
-However, implementing this virtual-to-physical (V2P) translation in this way ironically requires physical-to-virtual (P2V) translation,
+Traversing the page tables in software is the simplest way to convert a virtual address to a physical address;
+this is the context we are working up to.
+However, implementing this virtual-to-physical (V2P) translation in software ironically requires physical-to-virtual (P2V) translation,
 because the addresses stored in page table entries are physical, but memory accesses issued by the OS code use virtual addresses.
-% There is no universal way to convert physical addresses to virtual --- doing so relies on the kernel maintaining careful invariants or
-% additional data structures to enable P2V translation.
-\looseness=-1
-
 Because VMM operations are performance-critical for many workloads, most kernels
 maintain invariants that enable very fast P2V conversions (rather than adding another data structure).
-Most kernels maintain an invariant on their page tables that the virtual address of any page used for a page table 
+Specifically, many kernels maintain an invariant on their page tables that the virtual address of any page used for a page table 
 % lives at a virtual address whose value 
@@
 }
+Thus \lstinline|P2V| on line \ref{line:p2vC} of Figure \ref{fig:calltopteinitializeC} is a macro for adding the fixed constant \lstinline|KERNBASE|.
 
-For this reason, we extend the per-address-space invariant as in Figure \ref{fig:peraspaceinvariant_with_p2v_extensionC}, to also track which
-addresses we can perform a P2V conversion on by adding a constant offset.
+Figure \ref{fig:peraspaceinvariant_with_p2v_extensionC} extends the per-address-space invariant  to also track which
+addresses we can perform a P2V conversion on by adding a constant offset (i.e., the set of physical addresses which participate in page tables).
 $\Xi$ is another ghost map, from physical addresses to the level of the page table they represent (1--4).
 \emph{Only} physical addresses in $\Xi$ can undergo P2V conversion. 
-Section \ref{sec:p2vC} describes the actual conversion,
-but we describe the invariant here because adding new level 3/2/1 tables must maintain the invariant.
+Section \ref{sec:p2vC} describes the verification of an actual conversion,
+but this invariant must be \emph{established} when adding a new page table level (notably on Line \ref{line:call_to_pte_initializeC},
+hence the comment of Line \ref{line:now_we_know}).
 
@@
   \left(\bigast{(\vaddr, \textsf{paddr})\in \ptablestore}{\exists\;(\textsf{l4e, l3e, l2e, l1e, paddr})\ldotp \textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\vaddr\textsf{, l4e, l3e, l2e, l1e, paddr})}\right)\ast \\
-  \bigast{(\paddr,\mathsf{level}) \in \Xi}{\exists\; (\textsf{qfrac, q, val,}\vaddr) \ldotp \ulcorner \vaddr = \paddr + \textsf{KERNBASE} \; \textsf{level} > 1\urcorner \ast  \underbrace{\fracghostmaptoken{\delta}{\vaddr}{\paddr}{\qfrac} }_\text{Ghost translation} \ast \underbrace{\paddr \mapsto_{\mathsf{p}}\{\textsf{qfrac}\}\; \vale}_\text{Physical location}} \ast\\
-   \qquad\underbrace{ \ulcorner \textsf{qfrac} = 1 \leftrightarrow \; \lnot\textsf{entry\_present }(\vale) \urcorner}_\text{Entry validity}\ast 
-    \underbrace{\left(\ulcorner\textsf{present\_L}(\vale,\mathsf{level})\urcorner \wand \forall_{\textsf{i}\in\textsf{0..511}} \ldotp \ghostmaptoken{\textsf{id}}{((\mathsf{entry\_page}\;\vale) + \textsf{i * 8})}{\textsf{level-1}}\right)}_{\text{Indexing into next level of tables}} \\
+  \bigast{(\paddr,\mathsf{level}) \in \Xi}{\exists\; (\textsf{qfrac, q, val,}\vaddr) \ldotp \ulcorner \vaddr = \paddr + \textsf{KERNBASE} \; \textsf{level} > 1\urcorner \ast  \underbrace{\fracghostmaptoken{\delta}{\vaddr}{\paddr}{\qfrac} }_\text{\textcircled{1} Ghost translation} \ast \underbrace{\paddr \mapsto_{\mathsf{p}}\{\textsf{qfrac}\}\; \vale}_\text{\textcircled{2} Physical location}} \ast\\
+   \qquad\underbrace{ \ulcorner \textsf{qfrac} = 1 \leftrightarrow \; \lnot\textsf{entry\_present }(\vale) \urcorner}_\text{\textcircled{3} Entry validity}\ast 
+    \underbrace{\left(\ulcorner\textsf{present\_L}(\vale,\mathsf{level})\urcorner \wand \forall_{\textsf{i}\in\textsf{0..511}} \ldotp \ghostmaptoken{\textsf{id}}{((\mathsf{entry\_page}\;\vale) + \textsf{i * 8})}{\textsf{level-1}}\right)}_{\text{\textcircled{4} Indexing into next level of tables}} \\
   \textsf{ where } \\
@@
 \]
+\todo[inline]{Ismail, what's APVMappings?}
 \vspace{-1em}
-\caption{Global Address-Space Invariant in Figure \ref{fig:peraspaceinvariantC} extended with a ghost map bookkeeping identity mappings }
+\caption{Address space invariant of Figure \ref{fig:peraspaceinvariant} extended with a ghost map bookkeeping identity mappings. }
   \label{fig:peraspaceinvariant_with_p2v_extensionC}
@@
 
-For each $\paddr\mapsto \textsf{v} \in\Xi$, the invariant tracks virtual points-to justifying that virtual address $\paddr+\textsf{KERNBASE}$ maps to physical address $\paddr$
-(the ``Ghost translation'' in Figure \ref{fig:peraspaceinvariant_with_p2v_extensionC});
-fractional ownership of the physical memory for that page table entry;
-and for valid entries (with the present bit set) above L1, ghost map tokens for every entry in the table pointed to by the entry, which can be used
-to repeat the process one level down. 
+For each $\paddr\mapsto \textsf{v} \in\Xi$, the invariant contains a virtual points-to justifying that virtual address
+$\paddr+\textsf{KERNBASE}$ maps to physical address $\paddr$
+(\textcircled{1} in Figure \ref{fig:peraspaceinvariant_with_p2v_extensionC});
+fractional ownership of the physical memory for that page table entry (\textcircled{2}, which together with \textcircled{1} is equivalent
+to a virtual points-to);
+and for valid entries (with the present bit set) above L1, ghost map tokens for $\Xi$ for every entry in the table pointed to by the entry, which can be used
+to repeat the process one level down (\textcircled{4}). 
 % (L1 entries point to data pages, whose physical memory ownership resides in some virtual points-to).
-The assertion on Line \ref{line:conditional_childrenC} of Figure \ref{fig:calltopteinitializeC} comes from the invariant one level up; 
-if the valid bit is set,
-the code can return those child tokens without the conditional guard.
+\textcircled{4} becomes part of the precondition to \lstinline|pte_get_next_table|:
+Line \ref{line:precondition_conditional}) says that if the entry is valid (points to a next-level table)
+then there are tokens for accessing $\Xi$ for every entry in the next-level table.
+By Line \ref{line:finalpieceS} the entry is guaranteed to be valid so all tokens for converting the next table level's physical addresses to virtual
+are available (Line \ref{line:available_child_tokens}).
 \looseness=-1
 
-The fractional ownership of the entry's physical memory is subtle. Recall that $\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$ retains some physical
-ownership of each page table entry that is traversed (proportional to how many virtual addresses share the entry).
-So in general the invariant cannot keep full permission to the memory in this part of the invariant, or it would overlap the page table walk for virtual points-to
-assertions. But in the case where the entry is invalid, we may need to write to it (e.g., to install a reference to a next-level table, as we do in Figure \ref{fig:calltopteinitializeC}),
-which requires full permission. Fortunately, the entry can only be in use if its valid bit is set; if the valid bit is not set, we know
-that no virtual points-to entry in $\delta$/$\theta$ holds any partial ownership.
-Thus we use the invariant portion annotated as ``Entry validity'' in Figure \ref{fig:peraspaceinvariant_with_p2v_extensionC} to capture this:
-if the entry is invalid the invariant holds full ownership of the entry, so it can be updated; while if the entry is valid,
-the invariant owns only a constant nonzero fragment sufficient to read the entry, but not modify it (which would invalidate some virtual points-to assertions):
-\begin{equation*}
- \ulcorner \textsf{qfrac} = 1 \leftrightarrow \; \lnot\textsf{entry\_present }(\vale) \urcorner \tag{*}
-\end{equation*}
-Thus the fractional ownership of the physical location is enough for Line \ref{line:read_entry_contentsC} in Figure \ref{fig:calltopteinitializeC} to access the entry, though in \lstinline|get_next_table|
-the caller has pulled that piece of information out of the invariant and passed it for the entry at hand.
-This removal appears explicitly in assertions,
-as the argument to the invariant is $\Xi\setminus\{\mathsf{entry}\}$ (indexing by the set $\Xi$ allows us to borrow the physical resources
-for a specific page table entry out of the invariant, and later put them back).
-Line \ref{line:check_entry_present_jumpC}'s conditional then determines in the fall-through case that the bit is not set, which
-together with other facts entails $\textsf{qfrac} = 1$ at Line \ref{line:after_concluding_qfrac1C},
-and permits storing a new entry (in the elided code around Line \ref{line:install_new_entryC}).
+\paragraph{Self-Conditional Fractional Ownership}
+The fractional ownership of the entry's physical memory is subtle.
+As noted above, a \emph{valid} entry must coexist with the fractional ownership from
+$\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$ and therefore have less than full ownership,
+but in the case where the entry is \emph{invalid}, Line \ref{line:call_to_pte_initializeC} must have full permissions in order
+to populate the entry (i.e., to install a reference to a next-level table).
+Fortunately, the entry can only be in use if its valid bit is set; if the valid bit is not set, we know
+that no virtual points-to entry in $\delta$ or $\theta$ holds any partial ownership.
+But determining this requires reading the very memory whose ownership is being determined.
+We use the invariant portion annotated as ``Entry validity'' (\textcircled{3}) in Figure \ref{fig:peraspaceinvariant_with_p2v_extensionC} to capture this:
+if the entry is invalid the invariant holds full ownership of the entry, so it can be updated;
+while if the entry is valid, the invariant owns only a constant nonzero fraction sufficient to read but not modify the entry.
+% \begin{equation*}
+%  \ulcorner \textsf{qfrac} = 1 \leftrightarrow \; \lnot\textsf{entry\_present }(\vale) \urcorner \tag{*}
+% \end{equation*}
+Since the fractional ownership is always non-zero, Line \ref{line:read_entry_contentsC} in Figure \ref{fig:calltopteinitializeC} can read the entry,
+and if the entry is dynamically found to be invalid, the invariant is refined (Line \ref{line:refined_fractional_ownership}) for full
+ownership, allowing updates.
+
+Note that the caller is responsible for providing this conditional ownership, having pulled it out of the invariant earlier.
+This is why the precondition (Line \ref{line:preconditoin_entry_out}) explicitly excludes the entry's physical address from the invariant ($\Xi\setminus\{\mathsf{entry}\}$) ---
+its relevant assertions have already been borrowed by the caller.
 \looseness=-1
-
-This seemingly simple piece of code has a highly nontrivial correctness argument, which depends critically on detailed invariants on how access to page table
-entries is shared between parts of the kernel. No prior work has engaged with this problem.
 
@@
  \subsubsection{Installing a New Table}
+\todo[inline]{C edit frontier. Should see if we can condense this discussion, since the C is only 2 lines and mostly follows from assumptions
+about kalloc. Maybe this should get inlined into the Self-Conditional Fractional Ownership above, and likewise the P2V subsubsection below into the Physical-to-Virtual Mappings
+bit above. Yeah, they should. I think the expository order still makes sense even if it's not program order; maybe explain we treat the
+success case first and the backup case second? This would then still permit the discussion of pte initialize to be near the proof of its call}
  After obtaining the identity mapping for \textsf{entry}, we are able to load the \textsf{entry\_val} into \textsf{rdi}, and check the presence bit through
@@
 \ifPLDI
-\begin{lstlisting}[mathescape,escapeinside={(*}{*)}]
-/*
-@param l4: the root address of the page-table tree
-@param va: virtual address to be translated and mapped
-*/
+\begin{lstlisting}[language=C,mathescape,escapeinside={(*}{*)}]
+/* @param l4: the root address of the page-table tree
+   @param va: virtual address to be translated and mapped */
 pte_t *walkpgdir(pte_t *l4, const void *va) {
@@
 \fi
+\todo[inline]{clarify virt/phys of input root and the output}
 \caption{Walking page-table directory via calls to \textsf{pte\_get\_next\_table} in Figure \ref{fig:calltopteinitializeC}}
@@
     \\ \ast \left(\begin{array}{l}  \textsf{l1\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l1\_entry}\;\textsf{l1e\_val} \ast \\ \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l1e\_val})\urcorner \end{array}\right) =\textsf{R}_{\textsf{l1e}}\end{array} $
+\todo[inline]{Colin: This is confusing: in the code \lstinline|l2_entry| is an actual entry word, but in these invariants they seem to be paddrs.}
 \caption{\textsf{R}$_{\textsf{walk}}$: Resources obtained from invariant for walking page-table directory with \emph{non-existing entries} via calls to \textsf{l1e\_get\_next\_table} in Figure \ref{fig:calltopteinitializeC}}

### 2025-07-22 examples_C.tex
WIP restructuring the get_next discussion

@@
 
-\todo[inline]{Colin WIP note: The organization is odd here.
-We talk through access to entry, then start talking about identity mappings even though we don't need that info
-until the P2V call later. But we do need the part of the extended invariant that deals with the
-entry-present qfrac bit, which does only hold for those addresses used for page tables, so is part
-of the same invariant. Maybe we should reorg to first talk through the logic, 
-calling out the two problems the new invariant solves (permission to update the entry and p2v) along the way,
-then have a section called "extended address space invariant" or similar to talk about the solutions.
-This would roughly mean moving 5.1.3 (installing a new table) up
-}
 \todo[inline]{this and other figures need assertion updates to talk about C: see entry\_val (needs adjustment to points-to preconditions
@@
 I would also expect to see a more prominent use of \textsf{v} in the precondition
+Line 20 also seems like it lost the paddr of the vpte
 }
@@
 
-\subsubsection{Address Space Invariant: Identity Mappings and Conditional Ownership Page Table Ownership}
+\subsubsection{Address Space Invariant: Identity Mappings and Conditional Page Table Ownership}
 \label{subsec:identitymappingsC}
@@
 verifying lines \ref{line:install_new_entryC} and \ref{line:p2vC}.
+The key idea is to establish extra invariants on physical addresses that are part of a page table ---
+but to do so in a way that meshes with the existing invariants (in the informal sense) already preserved in most
+unverified kernel designs.
+Each of the above problems requires its own extension to the invariant, but we will discuss
+physical-to-virtual conversion first, both because it dictates the organization of the invariant
+and because when Line \ref{line:check_entry_present_jumpC}'s
+conditional check is false that is all that is necessary for the proof; correctness of the conditional branch deals with \emph{both}
+extensions.
 
-\paragraph{Physical-to-Virtual Mappings}
+\subsubsection{Physical-to-Virtual Mappings and \textsf{P2V}}
+\label{sec:p2vC}
 Kernels need to convert between physical and virtual addresses, in both directions.
@@
 By Line \ref{line:finalpieceS} the entry is guaranteed to be valid so all tokens for converting the next table level's physical addresses to virtual
-are available (Line \ref{line:available_child_tokens}).
+are available (in the form expressed by the assertion on Line \ref{line:available_child_tokens}).
 \looseness=-1
+
+As noted above, for the code path where the conditional does not execute (there was already a valid entry), this is all we need of the new
+invariant to verify the end of the function from Line \ref{line:finalpieceS} onward.
+By that point the invariant holds and applies to the definitely-valid entry,
+so we can the physical address of the next-level table to a corresponding virtual address via the identity mappings just described.
+Line \ref{line:extract_pfn} simply retrieves the physical address.
+Line \ref{line:p2vC} is the critical piece, and arguably corresponds to the most subtle verification of an \lstinline|add| instruction
+(\lstinline|add rax, KERNBASE|)\todo{if walkpgdir returns vaddr, then KERNBASE==0, which complicates some discussion here}
+that we are aware of, and something no prior work on verified OS kernels has dealt with.
+
+After Line \ref{line:extract_pfn}, it is already known that the present bit is set in the entry,
+thus Line \ref{line:childrenC}'s assertion reflects that the tokens for $\Xi$ exist for each word-aligned
+physical address in the next-level table.
+
+\hrule
+\todo[inline]{this next (older) text is a little confusing; it talks about the conversion to virtual points-to,
+but that happens at the caller, which is the only one that knows which index is needed.
+Clean up: the end of this function returns (1) computationally, the virtual address of the base of the level $n-1$ table;
+and (2) logically, the resources for the \emph{caller} to convert any aligned virtual address in the 4K
+range after that base into a virtual points-to.
+}
+
+The code loads the value at the offseted (\textsf{KERNBASE}) virtual memory address ($\textsf{entry}_{\textsf{pfn}}$ \textsf{+KERNBASE})
+of the \emph{base} of the next level of the page table.
+While we could now convert this address to a virtual points-to, this is not necessarily the correct thing to do.
+The caller \lstinline|walkpgdir| (discussed next) uses \lstinline|pte_get_next_table| to retrieve just the base address,
+because only the caller knows which entry in the subsequent table will be accessed (it depends on the corresponding bits from the virtual
+address being translated). So instead we pass back the per-address-space invariant with the identity mapping resources for \lstinline|entry|
+pulled out. The caller determines which entry in that table must actually
+be accessed --- by selecting the appropriate index into the 512 ghost map tokens returned in the postcondition,
+and using the ghost translation and physical location portions of the invariant to assemble a vpte-pointsto
+that justifies the caller's subsequent access to a particular entry of the returned table.
 
@@
 \looseness=-1
+\\
 
@@
 
-\subsubsection{Physical-to-Virtual Conversion with \textsf{P2V}}
-\label{sec:p2vC}
-Once we know that the entry refers to a physical address in the identity mapping range ($\Xi$)
-(via the branch at Line \ref{line:check_entry_present_jumpC}, or by allocating and installing a new entry
-as just discussed for Lines \ref{line:check_entry_present_jumpC}--\ref{line:end_of_allocation_pathC}), 
-we can convert this frame address to a corresponding virtual address via the identity mappings
-discussed in Section \ref{subsec:identitymappingsC} and Figure \ref{fig:peraspaceinvariant_with_p2v_extensionC}.
-in the last lines of \lstinline|pte_get_next_table| shown in Line \ref{line:p2vC} in Figure \ref{fig:calltopteinitializeC} (the continuation of Figure \ref{fig:calltopteinitializeC}).
-This is a critical piece of the full page table walk verification.
-In our small kernel (Line \ref{line:p2vC} in Figure \ref{fig:calltopteinitializeC}), as in larger kernels, the C macro \texttt{P2V} common to many kernels
-is actually just addition by the constant offset mentioned in Section \ref{subsec:identitymappingsC}.
-But the correctness of this simple instruction is quite subtle.
-%  and cannot be proven 
-% without the extended invariant (Figure \ref{fig:peraspaceinvariant_with_p2v_extension})
-% worked out Section \ref{subsec:identitymappings}.
-
-%\begin{figure}\footnotesize
-%  \begin{lstlisting}[mathescape,escapeinside={(*}{*)}]
-%  /*... Continued from Figure (*\ref{fig:calltopteinitialize}*); assertions below specialized to non-allocating path for clarity */
-%  $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \}),m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{rcx}  \mapsto_{\textsf{r}} \textsf{\_}  \ast \textsf{entry} \mapsto_{\textsf{id}} \textsf{\_} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
-%  $\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{(pte\_initialized (entry\_val.pfn))} \urcorner }_{\rtv}$
-%  $\specline{ \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{(pte\_initialized (entry\_val.pfn)))} \ast \texttt{rax} \mapsto_{\textsf{r}} \textsf{ table\_root (pte\_initialize(entry\_val.pfn))} }_{\rtv}$
-%  $\specline{\forall_{i\in \textsf{0 ...
-511} } \ldotp  \ghostmaptoken{\textsf{id}}{((\textsf{table\_root (pte\_initialized (entry\_val.pfn)))}) + \textsf{i * 8})}{\textsf{v-1}}  }$ (*\label{line:children}*)
-%  uintptr_t next_phys_addr = PTE_PFN_TO_ADDR(entry->pfn)
-%  uintptr_t next_virt_addr = (uintptr_t) P2V(next_phys_addr); (*\label{line:p2v}*) 
-%  /*... clean up the stack and return the rax value*/
-%  return next
-%}
-%\end{lstlisting}
-%\vspace{-1em}
-%\caption{Converting a physical address of a PTE to a virtual address (w/o instruction pointer or flag updates).
-% % Abreviating the other relevant resources (including the lower entries) peeled-out of the invariant as $\textsf{R}_{\textsf{children}}$
-%\vspace{-1em}
-%}
-%\label{fig:p2v}
-%\end{figure}
-Lines \ref{line:childrenC}-\ref{line:p2vC} in Figure \ref{fig:calltopteinitializeC} show the verification of the end of \lstinline|pte_get_next_table| specialized to the case where no allocation was necessary (i.e., the conditional on Line \ref{line:check_entry_presentC} of Figure \ref{fig:calltopteinitializeC} was taken).
-In this case, the true present bit allows access to the child tokens from Line \ref{line:conditional_childrenC} of Figure \ref{fig:calltopteinitializeC},
-which is then refined to the assertion on Line \ref{line:childrenC} of Figure \ref{fig:calltopteinitializeC}.
-The code loads the value at the offseted (\textsf{KERNBASE}) virtual memory address ($\textsf{entry}_{\textsf{pfn}}$ \textsf{+KERNBASE})
-of the \emph{base} of the next level of the page table.
-% \todo[inline]{the next sentence depends on having figure 10 updated to reflect the page-worth of tokens}
-While we could now convert this address to a virtual points-to, this is not necessarily the correct thing to do.
-The caller \lstinline|walkpgdir| (discussed next) uses \lstinline|pte_get_next_table| to retrieve just the base address,
-because only the caller knows which entry in the subsequent table will be accessed (it depends on the corresponding bits from the virtual
-address being translated). So instead we pass back the per-address-space invariant with the identity mapping resources for \lstinline|entry|
-pulled out. The caller determines which entry in that table must actually
-be accessed --- by selecting the appropriate index into the 512 ghost map tokens returned in the postcondition,
-and using the ghost translation and physical location portions of the invariant to assemble a vpte-pointsto
-that justifies the caller's subsequent access to a particular entry of the returned table.
-% in the identity map ($\Xi\setminus\{entry\}$) of the kernel invariant.
-% the logical update in Specification  Lines 5-10 to 10-14 for obtaining virtual-pointsto resource for the frame 
-% ($\textsf{entry}_{\textsf{pfn}}$) by removing it from the ghost map ($\Xi\setminus\{entry\}\cup \{\textsf{entry}_{\textsf{pfn}}) \}$) 
-% in Line 5 and compute the identity mapping for this physical frame address in Line 13 in Figure \ref{fig:p2v}).
-
 \subsubsection{The Specification of \lstinline|pte_get_next_table|}
+\todo[inline]{Colin says: should make this section more concise, really just tie up earlier discussions.
+But make a point to highlight that none of the specifications need to track which address space we're dealing with!}
 \hrule

### 2025-07-22 examples_C.tex
WIP update

@@
 
-After Line \ref{line:extract_pfn}, it is already known that the present bit is set in the entry,
-thus Line \ref{line:childrenC}'s assertion reflects that the tokens for $\Xi$ exist for each word-aligned
+After Line \ref{line:extract_pfn}, it is already known that the present bit is set in the entry;
+Line \ref{line:childrenC}'s assertion reflects that the tokens for $\Xi$ exist for each word-aligned
 physical address in the next-level table.
-
-\hrule
-\todo[inline]{this next (older) text is a little confusing; it talks about the conversion to virtual points-to,
-but that happens at the caller, which is the only one that knows which index is needed.
-Clean up: the end of this function returns (1) computationally, the virtual address of the base of the level $n-1$ table;
-and (2) logically, the resources for the \emph{caller} to convert any aligned virtual address in the 4K
-range after that base into a virtual points-to.
-}
-
-The code loads the value at the offseted (\textsf{KERNBASE}) virtual memory address ($\textsf{entry}_{\textsf{pfn}}$ \textsf{+KERNBASE})
-of the \emph{base} of the next level of the page table.
-While we could now convert this address to a virtual points-to, this is not necessarily the correct thing to do.
-The caller \lstinline|walkpgdir| (discussed next) uses \lstinline|pte_get_next_table| to retrieve just the base address,
-because only the caller knows which entry in the subsequent table will be accessed (it depends on the corresponding bits from the virtual
-address being translated). So instead we pass back the per-address-space invariant with the identity mapping resources for \lstinline|entry|
-pulled out. The caller determines which entry in that table must actually
+However, note that no argument to this function specifies which virtual address is being accessed,
+so \lstinline|pte_get_next_table| does not know which entry in the next table to retrieve.
+Even if that address were passed, this function is used for each step-down, so the slice of the
+virtual address (per Figure \ref{fig:pagetables}) is not fixed.
+Thus Line \ref{line:p2vC} computes the virtual address of the \emph{base} of the next-level table,
+and the postcondition includes a renamed version of the assertion on line \ref{line:childrenC},
+for the \emph{caller} --- \lstinline|walkpgdir| (discussed next) to perform the conversion for
+The caller determines which entry in that table must actually
 be accessed --- by selecting the appropriate index into the 512 ghost map tokens returned in the postcondition,
@@
 that justifies the caller's subsequent access to a particular entry of the returned table.
+The postcondition also passes back the per-address-space invariant with the
+identity mapping resources for \lstinline|entry| still pulled out (it was removed by the caller).
 
-\paragraph{Self-Conditional Fractional Ownership}
+\subsubsection{Self-Conditional Fractional Ownership and Installing a New Table}
 The fractional ownership of the entry's physical memory is subtle.
@@
 while if the entry is valid, the invariant owns only a constant nonzero fraction sufficient to read but not modify the entry.
-% \begin{equation*}
-%  \ulcorner \textsf{qfrac} = 1 \leftrightarrow \; \lnot\textsf{entry\_present }(\vale) \urcorner \tag{*}
-% \end{equation*}
-Since the fractional ownership is always non-zero, Line \ref{line:read_entry_contentsC} in Figure \ref{fig:calltopteinitializeC} can read the entry,
-and if the entry is dynamically found to be invalid, the invariant is refined (Line \ref{line:refined_fractional_ownership}) for full
-ownership, allowing updates.
-
+Since the fractional ownership is always non-zero, Line \ref{line:read_entry_contentsC} in Figure \ref{fig:calltopteinitializeC} can read the entry
+(using a rule similar to \textsc{WriteToRegFromVirtMem}, tailored to virtual PTE-points-to assertions),
+and if the entry is dynamically found to be invalid, the invariant is refined (Line \ref{line:refined_fractional_ownership})
+to indicate full ownership, allowing updates.
 Note that the caller is responsible for providing this conditional ownership, having pulled it out of the invariant earlier.
-This is why the precondition (Line \ref{line:preconditoin_entry_out}) explicitly excludes the entry's physical address from the invariant ($\Xi\setminus\{\mathsf{entry}\}$) ---
+This is why the precondition (Line \ref{line:precondition_entry_out}) explicitly excludes the entry's physical address from the invariant ($\Xi\setminus\{\mathsf{entry}\}$) ---
 its relevant assertions have already been borrowed by the caller.
 \looseness=-1
-\\
-
-% Concretely speaking, going back to Line 15 in Figure \ref{fig:calltopteinitialize}, to read the value referenced by physical address \textsf{entry} while preserving the soundness of memory mappings, our extended invariant introduces the side condition (*)
-% \begin{equation*}
-%  \ulcorner \textsf{qfrac} = 1 \leftrightarrow \; \lnot\textsf{entry\_present }(\vale) \urcorner \tag{*}
-% \end{equation*}
-% assuring that looking the identity mapping for \textsf{entry} is safe under the subtle justification which equates the full ownership to the non/presence of the entry which can only be known when investigated in Line 21 in Figure \ref{fig:calltopteinitialize}.
-\begin{comment}
- \begin{figure*}
-   \footnotesize
-   \[
- \begin{array}{l}
-   \specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry} \textsf{\_} \ast \ulcorner \Xi !!
-\textsf{entry} = \textsf{Some V} \urcorner  }_{\rtv}  \\
-\sqsubseteq \\
-   \specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast  \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{next} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv} \\
-\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present entry\_val})\urcorner}_{\rtv} \\
-\mathsf{mov} \;   [ \mathsf{r8} ],\mathsf{rdi}\\
- \specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast  \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{next} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv} \\
-\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present entry\_val})\urcorner}_{\rtv} \\
- \end{array}
- \]
- \vspace{-1em}
-\caption{Peel-out the resources needed for the virtual-pte pointsto relation for physical address \textsf{entry} to be accessed in Line 7 in Figure \ref{fig:calltopteinitializeC}. }
-  \label{fig:lookingupidpamppingC}
- \end{figure*}
-\end{comment}
-
- \subsubsection{Installing a New Table}
-\todo[inline]{C edit frontier. Should see if we can condense this discussion, since the C is only 2 lines and mostly follows from assumptions
-about kalloc. Maybe this should get inlined into the Self-Conditional Fractional Ownership above, and likewise the P2V subsubsection below into the Physical-to-Virtual Mappings
-bit above. Yeah, they should. I think the expository order still makes sense even if it's not program order; maybe explain we treat the
-success case first and the backup case second? This would then still permit the discussion of pte initialize to be near the proof of its call}
- After obtaining the identity mapping for \textsf{entry}, we are able to load the \textsf{entry\_val} into \textsf{rdi}, and check the presence bit through
-\ifPLDI
-Line \ref{line:mask_presentC} % in condensed version, all on same line
-\else
-Lines \ref{line:mask_present}--\ref{line:check_entry_present} 
-\fi
-in Figure \ref{fig:calltopteinitializeC}.
-Accessing the presence bit and checking the value allows us to exploit the condition (*) that was just discussed when verifying the allocation
-path (i.e., when the entry is invalid and Lines \ref{line:alloc_path_startC}--\ref{line:alloc_path_endC} in Figure \ref{fig:calltopteinitializeC}
-must allocate the next level of tables).
-This operation is subtle. To reiterate: the operation requires that the relevant table entry is readable, but the exact portion of ownership
-returned must be determined by inspecting the valid bit of the value in memory --- so full ownership is returned only for unused entries.
-When the bit is not set, that entails full ownership of the entry's memory ($\textsf{qfrac} = 1$) and justifies writing to that memory.
-Otherwise, the code jumps past the end of this listing, to the following code at the top of Figure \ref{fig:p2v} (which is also the
-continuation of this code).
 
 \begin{figure}\footnotesize
-  \begin{lstlisting}[mathescape,escapeinside={(*}{*)}]
+  \begin{lstlisting}[language=C,mathescape,escapeinside={(*}{*)}]
 /* @param entry: virtual address of a non-present entry to initialize with a new physical page */
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \lnot(\textsf{entry\_present entry\_val} }_{\rtv}$
-$\specline{\textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,1}} \textsf{entry entryv} \ast \ghostmaptoken{\textsf{id}}{(\mathsf{entry})}{\textsf{level}} }$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \lnot(\textsf{entry\_present entry\_val} \ast \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,1}} \textsf{entry entryv} \ast \ghostmaptoken{\textsf{id}}{(\mathsf{entry})}{\textsf{level}} }$
 void pte_initialize(pte_t *entry) {
   pte_t *local = kalloc(); // Allocate a full zeroed page for 512 8-byte entries(*\label{line:call_to_kallocC}*)
-  entry->pfn = PTE_ADDR_TO_PFN((uintptr_t) local);
+  entry->pfn = PTE_ADDR_TO_PFN((uintptr_t) local);(*\label{line:kalloc_install}*)
   $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \lnot(\textsf{entry\_present entry\_val}) }$
   $\specline{ \textsf{entry+KERNBASE} \mapsto_{\mathsf{vpte,1}} \mathsf{entry} \; \textsf{pfn\_set(entryv nextpaddr)}}$
-  $\specline{\ghostmaptoken{\delta{}s}{\rtv}{\delta} \ast \begin{array}{l} \ulcorner\textsf{entry\_present (pte\_initialize(pfn\_set(entry\_val nextpaddr)),level)}\urcorner \wand \\  \;\;\;\;\;\;\; \forall_{i\in\textsf{0..511}} \; \textsf{table\_root (pte\_initialized (pfn\_set (entry\_val nextpaddr))) + i * 8} \mapsto_{\textsf{id}} \textsf{level-1}    \end{array}  }_{\rtv}$
+  $\specline{\ghostmaptoken{\delta{}s}{\rtv}{\delta} \ast \begin{array}{l} \ulcorner\textsf{entry\_present (pte\_initialize(pfn\_set(entry\_val nextpaddr)),level)}\urcorner \wand \\  \;\;\;\;\;\;\; \forall_{i\in\textsf{0..511}} \; \textsf{table\_root (pte\_initialized (pfn\_set (entry\_val nextpaddr))) + i * 8} \mapsto_{\textsf{id}} \textsf{level-1}    \end{array}  }_{\rtv}$(*\label{line:initialize_post}*)
 }
 \end{lstlisting}
-\vspace{-1em}
+\todo[inline]{more confusion of entry as program variable holding vaddr
+vs entry as logical variable for the paddr, nextpaddr, etc.
+update, clarifying that it returns a value for which the mapping holds.
+Also, do we need the conditional implication in this postcondition? it alloc'ed from
+a region already in Xi.
+And shouldn't we be excluding entry (the paddr) from Xi here?
+}
+\vspace{-2em}
 \caption{Allocating a physical page }
@@
 
-If the entry is not set, \textsf{pte\_initialize} (Line \ref{line:call_to_pte_initializeC} in Figure \ref{fig:calltopteinitializeC}) 
-allocates a physical page (internally utilizing the only unverified (trusted) code in our case studies, the page-allocator's \textsf{kalloc},\footnote{
+If the entry is not set, \lstinline|pte_initialize|  
+allocates a physical page for use as the next-level table.
+\lstinline|pte_initialize| is shown in Figure \ref{pteinitializespecC}: it simply calls
+\lstinline|kalloc| to allocate a physical page (Figure \ref{pteinitializespecC} Line \ref{line:call_to_kallocC}),
+and installs it into the entry (Line \ref{line:kalloc_install}).
+The page-allocator's \textsf{kalloc}
+is the only unverified (trusted) code in our case study.\footnote{
   This is an allocator for regions of pre-zeroed physical memory that is mapped, but not accessed by the allocator itself,
@@
 } 
-on Line \ref{line:call_to_kallocC} in Figure \ref{pteinitializespecC}). 
 Since we are using \textsf{pte\_initialize} for page-table address allocation, we must relate this newly
@@
 pool that is already mapped
-\ifPLDI
-\else
 \footnote{A reasonable reader might wonder where this pool
@@
 } 
-\fi
-and satisfies the offset invariants.
-% \todo[inline,color=blue]{colin frontier.
-% Stuck with line 31 onwards in Figure 7. rax holds nextpaddr, but I think that should be entrypfn, and 
-% the explicit entrypfn id token assertion should go away, as its covered by the forall assertion.
-% then the postcondition for pte-initialize should have a specific level now for the entries,
-% like 0, which can be updated in the view shift on line 42.
-% }
-% Focusing on the specification of \textsf{pte\_initialize} separately in Figure \ref{fig:pteinitializespec}, 
-% we right immediately realize that instead of seeing see a physical pointsto for the fresly page-table address 
-% (e.g. $\mathsf{nextpaddr} \mapsto_{\mathsf{p}} \mathsf{w64\_0}$) deliberately in the post-conditoin in Lines 15-16,
-%  we observe a full-ownership token representing the knowledge that a frame and all the entries indexed from this 
-% frame are freshly allocated with full-ownership to be a part of the identity map, $\Xi$. 
-The soundness argument of this specification relies on the fact that these freshly allocated resources are part
-of an entry construction that has not been completed yet: the presence bit is set (line \ref{line:install_new_entryC} in Figure \ref{fig:calltopteinitializeC}) after these freshly allocated resources are incorporated into the
-entry construction via the page frame portion of the PTE. In other words, the side condition, (*),
- formalizes that any access to the entry with these resources is \textit{invalid} (in the sense of not necessarily
-having accompanying resources) until the entry is marked present (and thus the memory returned from \textsf{kalloc}
-moves into the page table invariant).
+and satisfies the offset invariants (trivially, as the new page is zeroed).
+The presence bit of the entry is \emph{not} set during \lstinline|pte_initialize|, but upon
+return to \lstinline|pte_get_next_table|, where it will validate the conditional ownership discussed above.
+\lstinline|pte_initialize| has a full-permission virtual pte-pointsto in its precondition.
+Then the assertions that hold after Line \ref{line:now_we_know} of Figure \ref{fig:calltopteinitializeC}
+are enough to establish the same page table invariants which hold in the case where the entry was already valid,
+by updating the current address space's entry.
+\todo[inline]{still feel like this was a little too high level}
+
+\todo[inline,color=cyan]{Colin edit frontier}
 
 \subsubsection{The Specification of \lstinline|pte_get_next_table|}
-\todo[inline]{Colin says: should make this section more concise, really just tie up earlier discussions.
-But make a point to highlight that none of the specifications need to track which address space we're dealing with!}
-\hrule
-
 Note that the specification does \emph{not} assume a specific page table level --- the logical parameter \textsf{v} represents the level

### 2025-07-22 examples_C.tex
updated notes for tasks after discussion with Ismail

@@
   pte_t *next;
-  $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{entry}\},m) \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$(*\label{line:precondition_entry_out}*)
+  $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{entry}\mapsto\textsf{v}\},m) \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$(*\label{line:precondition_entry_out}*)
   $\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present entry\_val})\urcorner}_{\rtv}$ (*\label{line:get_next_vpte_preconditionC}*)
@@
 }
+\todo[inline,color=cyan]{yes update logical entry to entryp, maybe remove entry $\mapsto$ v from $\Xi$ in first line?
+and fix 20}
 \vspace{-1em}
@@
  is a good framing, we should discuss
+}\todo[color=cyan]{maybe say something about it feeling like normal asm verification \emph{because} of the virtual points-to enabling
+familiar-style proving except for specific manipulation of page tables.
+Maybe mention that it's well-beyond the capabilities of prior work because they just never tried? (said nicer...)
 }
@@
 \todo[inline]{Ismail, what's APVMappings?}
+\todo[inline,color=cyan]{APVMappings is authoritative ownership of the ghost map for the current address space's $\Xi$ (identity mappings).
+So the last line shouldn't just existentially quantify $\delta$, but also a $\delta'$ since the authoritative map for $\Xi$ should have its own ghost name
+(and does in the Rocq)
+}
 \vspace{-1em}
@@
 }
+\todo[inline,color=cyan]{no, this proof really doesn't care if entry is in or out of Xi, no need to exclude, but maybe worth explaining}
 \vspace{-2em}

### 2025-07-22 relwork.tex
updated notes for tasks after discussion with Ismail

@@
 
-Unlike our work, Kolanski and Klein prove very useful embedding theorems stating that code that does not modify page table 
+Unlike our work, Kolanski and Klein prove very useful embedding theorems\todo{no, Colin fix per response} stating that code that does not modify page table 
 entries can be verified in a VM-ignorant program logic, and that proofs in that logic can be embedded into the VM-aware logic 

### 2025-07-23 background.tex
caption cleanup
updated notes for tasks after discussion with Ismail
remove leftover unnecessary pfn update from examples

@@
     pte_initialize(entry); // Alloc empty table for next level (*\label{line:pass_addrof_nextC}*)(*\label{line:alloc_path_startC}*)(* \label{line:call_to_pte_initializeC} *)
-    entry->pfn = nextpaddr; // Point to next table
     entry->present = 1; // Mark valid (*\label{line:install_new_entryC}*)
@@
     \vspace{-2.5em}
-    \caption{A kernel implementing x86-64 page table lookups.}
+    \caption{A kernel implementation of x86-64 page table lookups.}
     \label{fig:pagetablescode}

### 2025-07-23 examples_C.tex
remove leftover unnecessary pfn update from examples
move and revise the context switch example

@@
 In this section, we verify several critical and challenging pieces of VMM code.
-First, in several stages, we work up to mapping a new page in the current address space.
+First, we formally verify a switch into a new address space as part of a task switch,
+the first such verification handling both old and new processes' assertions (in different address spaces) at the time of the switch.
+Then, in several stages, we work up to mapping a new page in the current address space, addressing significantly more of this process than prior
+work that included address translation in its hardware model.
 This requires a number of independently challenging substeps: dynamically traversing a page table to find
@@
 current address space) has previously been formally verified with respect to a machine model with address translation.
-Second, we formally verify a switch into a new address space as part of a task switch,
-the first such verification handling both old and new processes' assertions (in different address spaces) at the time of the switch.
+
+\begin{figure}\footnotesize
+\begin{lstlisting}[mathescape,escapeinside={(*}{*)}]
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,\Xi,m) \ast [\rtv'](\mathcal{I}\texttt{ASpace}(\theta',\Xi',m') \ast \texttt{Pother}) \ast \texttt{rsi}\mapsto_{\textsf{r}} restore \ast \texttt{rdi}\mapsto_{\textsf{r}} save\ast \texttt{rbx}\mapsto_{\textsf{r}} \texttt{rbxv} }_{\rtv}$(*\label{line:start_swtch_pre}*)
+$\specline{\texttt{rsp}\mapsto_{\textsf{r}} \texttt{rspv} \ast \texttt{rbp}\mapsto_{\textsf{r}} \texttt{rbpv} \ast \texttt{r12}\mapsto_{\textsf{r}} \texttt{r12v} \ast \texttt{r13}\mapsto_{\textsf{r}} \texttt{r13v} \ast \texttt{r14}\mapsto_{\textsf{r}} \texttt{r14v} \ast \texttt{r15}\mapsto_{\textsf{r}} \texttt{r15v}}_{\rtv}$
+$\specline{\mathsf{ContextAt}(save,\_) \ast \mathsf{ContextAt}(restore,[\mathsf{rbxv}',\ldots,\rtv']) }_{\rtv}$(*\label{line:end_swtch_pre}*)
+mov 0[rdi], rbx(*\label{line:start_save}*)
+... ;; also save rsp, rbp, r12, r13, r14, to offsets 8, 16, 24, 32, and 40 from rdi
+mov 48[rdi], r15
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,\Xi,m) \ast [\rtv'](\mathcal{I}\texttt{ASpace}(\theta',\Xi',m') \ast \texttt{Pother})  \ast  \texttt{rsi}\mapsto_{\textsf{r}} restore \ast \texttt{rdi}\mapsto_{\textsf{r}} save \ast \texttt{rbx}\mapsto_{\textsf{r}} \texttt{rbxv} }_{\rtv}$
+$\specline{ \texttt{rsp}\mapsto_{\textsf{r}} \texttt{rspv} \ast \texttt{rbp}\mapsto_{\textsf{r}} \texttt{rbpv} \ast \texttt{r12}\mapsto_{\textsf{r}} \texttt{r12v} \ast \texttt{r13}\mapsto_{\textsf{r}} \texttt{r13v} \ast \texttt{r14}\mapsto_{\textsf{r}} \texttt{r14v} \ast \texttt{r15}\mapsto_{\textsf{r}} \texttt{r15v}}_{\rtv}$
+$\specline{\mathsf{ContextAt}(save,[\mathsf{rbxv,\ldots,\rtv}]) \ast \mathsf{ContextAt}(restore,[\mathsf{rbxv}',\ldots,\rtv']) }_{\rtv}$
+mov 56[%rdi], %cr3 $\specline{\ldots \ast \texttt{rdi+56} \mapsto_{\textsf{v}} \rtv \ast \ldots }_{\rtv}$   (*\label{line:end_save}*) 
+mov rbx, 0[rsi] (*\label{line:start_restore}*)
+mov rsp, 8[rsi] ;; Switch to new stack, which may not be mapped in the current address space!(*\label{line:stack_switch}*)
+... ;; load rbp, r12, r13, r14, from offsets 16, 24, 32, and 40 from rsi
+mov r15, 48[rsi]
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,\Xi,m) \ast [\rtv'](\mathcal{I}\texttt{ASpace}(\theta',\Xi',m') \ast \texttt{Pother})  \ast  \texttt{rsi}\mapsto_{\textsf{r}} restore \ast \texttt{rdi}\mapsto_{\textsf{r}} save \ast \texttt{rbx}\mapsto_{\textsf{r}} \texttt{rbxv}' }_{\rtv}$
+$\specline{\texttt{rsp}\mapsto_{\textsf{r}} \texttt{rspv}' \ast \texttt{rbp}\mapsto_{\textsf{r}} \texttt{rbpv}' \ast \texttt{r12}\mapsto_{\textsf{r}} \texttt{r12v}' \ast \texttt{r13}\mapsto_{\textsf{r}} \texttt{r13v}' \ast \texttt{r14}\mapsto_{\textsf{r}} \texttt{r14v}' \ast \texttt{r15}\mapsto_{\textsf{r}} \texttt{r15v}'}_{\rtv}$
+$\specline{\mathsf{ContextAt}(save,[\mathsf{rbxv,\ldots,\rtv}]) \ast \mathsf{ContextAt}(restore,[\mathsf{rbxv}',\ldots,\rtv']) }_{\rtv}$
+mov cr3, 56[rsi] ;; <-- Switch to the new address space (*\label{line:end_restore}*)
+$\specline{ [\rtv](\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,\Xi,m) \ast \mathsf{ContextAt}(save,[\mathsf{rbxv,\ldots,\rtv}]) \ast \mathsf{ContextAt}(restore,[\mathsf{rbxv}',\ldots,\rtv']))  }_{\rtv}$ (*\label{line:modality_switchC}*)
+$\specline{ \mathcal{I}\texttt{ASpace}(\theta',\Xi',m') \ast \texttt{Pother} \ast  \texttt{rsi}\mapsto_{\textsf{r}} restore \ast \texttt{rdi}\mapsto_{\textsf{r}} save \ast \ldots }_{\rtv'}$
+\end{lstlisting}
+\vspace{-1em}
+\caption{Basic task switch code that switches address spaces.}
+\label{fig:swtchC}
+\end{figure}
+
+\subsection{Change of Address Space}
+A critical piece of \emph{trusted} code in current verified OS kernels is the assembly code to change the current address space; current verified OS kernels currently
+lack effective ways to specify and reason about this low-level operation, for reasons outlined in Section \ref{sec:relwork}.
+
+Figure \ref{fig:swtchC} gives simplified code for a basic task switch, the heart of an OS scheduler implementation. This is code that saves the context (registers and stack)
+of the running thread, and resumes execution of a previously-suspended thread of execution.
+In C this code would be given the signature
+\mbox{\lstinline[language=C]|void swtch(context_t* save, context_t* restore)|}.
+Saving the context is a straightforward matter of storing each register into the \lstinline|save| context.
+Restoring the \lstinline|restore| context is the tricky bit, because both the stack pointer and address space must be restored.
+Confusingly, a single dynamic execution of this function begins execution in one thread,
+and returns in another thread --- because the execution switches stacks, and thus returns on the second thread's stack.\footnote{This is the function in UNIX 6th Edition 
+with the infamous ``You are not expected to understand this'' comment~\cite{lions1996lions},
+though Doeppner~\cite{doeppner2010operating} and others, offer detailed explanations.}
+Hence this is used, for example, when the scheduler has chosen a new thread to execute for a voluntary (non-preempted)
+context switch, and will call
+the code with \lstinline|save| pointing to a reserved storage space for the current thread, and
+\lstinline|restore| pointing to the context of the next thread to execute.
+We will ignore non-integer registers; others are handled similarly.
+
+In Figure \ref{fig:swtchC}, Lines \ref{line:start_save}--\ref{line:end_save} store the callee-save registers (per the System V AMD64 ABI calling conventions) of the calling
+thread into the context data structure pointed to by \lstinline|rdi| (at virtual address $save$).
+This is justified by the $\textsf{ContextAt}(save,\_)$ assertion in the precondition (Line \ref{line:end_swtch_pre}), which expands
+into a full set of full-permission (thus writable) virtual points-to assertions for various offsets from $save$, one for each saved register.
+Verification up through line \ref{line:end_save} is straightforward application of \textsc{WriteToVirtMemFromReg} (Figure \ref{fig:wpdamd}).
+
+The code to restore the previously-saved context located at $restore$ (accessed via \lstinline|rsi|) in Lines
+\ref{line:start_restore}--\ref{line:end_restore}
+is where the proof becomes subtle, though our logic makes the construction of the proof feel similar to typical assembly-level verification
+because most instructions are verified with rules that work very similarly to standard proof rules while being proven
+sound against a machine model with address translation.
+Similar to the precondition for the save context, the restore context has a corresponding $\mathsf{ContextAt}(restore,[\ldots])$
+assertion expanding to virtual points-to assertions --- in the caller's adddress space.
+The \lstinline|mov| instructions prior to Line \ref{line:end_restore} are each verified with a fixed-register-offset
+variant of \textsc{WriteToRegFromVirtMem}, but
+Line \ref{line:stack_switch}'s implications are subtle because it switches stacks by updating \lstinline|rsp|.
+Because the new stack pointer may only be valid in the address space of the restored context, stack accesses at this point are unsafe.
+Prior to Line \ref{line:end_restore}, we can see in code and invariants that the local registers are updated
+to match the values populating the restore context, except for the page table root.
+Line \ref{line:end_restore} itself is verified with a rule similar to \textsc{WriteToRegCtlFromRegModal} (but obviously
+reading from a fixed offset of a register, as needed in Line \ref{line:end_restore}).
+This rule also globally moves assertions into and out-of other-space assertions, to reflect that
+assertions holding in the outgoing address space \rtv{} generally will not hold in the incoming address space $\rtv'$
+and vice versa. Thus the precondition has assertions for the new thread under an other-space modality for the new address space,
+and the postcondition has assertions for the old thread under an other-space modality for the old address space.\footnote{
+  The \textsf{ContextAt} assertions \emph{both} end up under the other-space modality for the old thread.
+  A real kernel would want to transfer both out, but justifying this is highly kernel-specific,
+  and particularly post-Spectre-and-Meltdown is quite varied.
+  \looseness=-1
+}
+\looseness=-1
+
+The specification above does not directly discuss the relationship between instruction pointers and registers --- and does not need to
+because \textsf{P} and \textsf{POther} can be instantiated to capture that relationship with additional information about stack contents.
+This code is meant to be called with a return address for the current thread stored on the current stack,
+and a return address for the target thread on the target thread's stack.
+% But the target thread's precondition is \emph{relative to its address space}, 
+% not the address space of the calling thread! This is reflected by
+% the other-space modality
+% $[\rtv'](\mathcal{I}\texttt{ASpace}(\theta,\Xi,m) \ast \texttt{Pother})$
+% in the specfication. 
+For a given call site, \textsf{P} would be instantiated to require that the initial stack pointer (before \lstinline|rsp| is updated)
+has a return address expecting the then-current callee-save register values in the \emph{current} (initial) address space
+to (together with other resources used to instantiate \textsf{P}) imply the precondition of the code at that return address.
+The situation for the target thread is similar, but using \textsf{POther}, \emph{and using the other-space modality}
+because the other thread's stack, code, and other relevant assertions may only be valid in the new address space.
+Our logic's rules for updating the page table root, and thus moving assertions into and out of other-space modalities,
+neatly manage which assertions are \emph{currently} valid, without the need to explicitly plumb address space labels through
+every assertion in the larger proof.
+
+% Immediately after the page table switch, assertions about the saved and restored contexts are
+% guarded by a modality for the retiring
+% address space \rtv{} (Line \ref{line:modality_switch}), per
+% \textsc{WriteToRegCtlFromRegModal} (Figure \ref{fig:wpdamd}),
+% because
+% there is no guarantee that the data structures of the previous address space are mapped in the new address space.
+% The ability to transfer that points-to information out of that modality is specific to a given kernel's design. 
+% Kernels that map kernel memory into all address spaces would need invariants
+% that justified moving those assertions out of the other-space modality.
+% % Following Spectre and Meltdown, this kernel design became less prevalent because speculative execution of accesses to kernel addresses could leak information even if the access did eventually cause a fault (the user/kernel mode permission check was done after fetching data from memory).
+Thus many modern kernels have reverted to the older kernel design where the kernel inhabits its own unique address space, and user processes have only enough extra material mapped in their address spaces to switch into the kernel (CPUs do not speculate past updates to \texttt{cr3}).
+% \looseness=-1
+
+Although prior work has verified context switches within a single address space~\cite{ni2007contexts}, and address space switches
+without any code before or after~\cite{syeda2020formal} (that is, not reasoning about the \emph{impact} of address space change
+on what data were accessible), this is the first verification that handles both.
+\looseness=-1
 
@@
 511} } \ldotp  \ghostmaptoken{\textsf{id}}{((\textsf{table\_root (pte\_initialized (pfn\_set(entry\_val nextpaddr)))}) + \textsf{i * 8})}{\textsf{v-1}} \end{array}  }$ (*\label{line:page_of_capsC}*)
-    entry->pfn = nextpaddr; entry->present = 1; // Install newly-allocated level n table, mark present(*\label{line:install_new_entryC}*)
+    entry->present = 1; // Install newly-allocated level n table, mark present(*\label{line:install_new_entryC}*)
     /*Now we know that entry is initialized, so we satisfy the condition to access children list */(*\label{line:now_we_know}*)
@@
 \fi
-
-\begin{figure}\footnotesize
-\begin{lstlisting}[mathescape,escapeinside={(*}{*)}]
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,\Xi,m) \ast [\rtv'](\mathcal{I}\texttt{ASpace}(\theta',\Xi',m') \ast \texttt{Pother}) \ast \texttt{rsi}\mapsto_{\textsf{r}} new \ast \texttt{rdi}\mapsto_{\textsf{r}} old \ast \texttt{rbx}\mapsto_{\textsf{r}} \texttt{rbxv} }_{\rtv}$
-$\specline{\texttt{rsp}\mapsto_{\textsf{r}} \texttt{rspv} \ast \texttt{rbp}\mapsto_{\textsf{r}} \texttt{rbpv} \ast \texttt{r12}\mapsto_{\textsf{r}} \texttt{r12v} \ast \texttt{r13}\mapsto_{\textsf{r}} \texttt{r13v} \ast \texttt{r14}\mapsto_{\textsf{r}} \texttt{r14v} \ast \texttt{r15}\mapsto_{\textsf{r}} \texttt{r15v}}_{\rtv}$
-$\specline{\mathsf{ContextAt}(old,\_) \ast \mathsf{ContextAt}(new,[\mathsf{rbxv}',\ldots,\rtv']) }_{\rtv}$
-mov 0[rdi], rbx ... ;; also save rsp, rbp, r12, r13, r14, to offsets 8, 16, 24, 32, and 40 from rdi (*\label{line:start_saveC}*)
-mov 48[rdi], r15
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,\Xi,m) \ast [\rtv'](\mathcal{I}\texttt{ASpace}(\theta',\Xi',m') \ast \texttt{Pother})  \ast  \texttt{rsi}\mapsto_{\textsf{r}} new \ast \texttt{rdi}\mapsto_{\textsf{r}} old \ast \texttt{rbx}\mapsto_{\textsf{r}} \texttt{rbxv} }_{\rtv}$
-$\specline{ \texttt{rsp}\mapsto_{\textsf{r}} \texttt{rspv} \ast \texttt{rbp}\mapsto_{\textsf{r}} \texttt{rbpv} \ast \texttt{r12}\mapsto_{\textsf{r}} \texttt{r12v} \ast \texttt{r13}\mapsto_{\textsf{r}} \texttt{r13v} \ast \texttt{r14}\mapsto_{\textsf{r}} \texttt{r14v} \ast \texttt{r15}\mapsto_{\textsf{r}} \texttt{r15v}}_{\rtv}$
-$\specline{\mathsf{ContextAt}(old,[\mathsf{rbxv,\ldots,\rtv}]) \ast \mathsf{ContextAt}(new,[\mathsf{rbxv}',\ldots,\rtv']) }_{\rtv}$
-mov 56[%rdi], %cr3 $\specline{\ldots \ast \texttt{rdi+56} \mapsto_{\textsf{v}} \rtv \ast \ldots }_{\rtv}$   (*\label{line:end_save}*) 
-mov rbx, 0[rsi] (*\label{line:start_restoreC}*)
-mov rsp, 8[rsi] ;; Switch to new stack, which may not be mapped in the current address space!
-... ;; load rbp, r12, r13, r14, from offsets 16, 24, 32, and 40 from rsi
-mov r15, 48[rsi]
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,\Xi,m) \ast [\rtv'](\mathcal{I}\texttt{ASpace}(\theta',\Xi',m') \ast \texttt{Pother})  \ast  \texttt{rsi}\mapsto_{\textsf{r}} new \ast \texttt{rdi}\mapsto_{\textsf{r}} old \ast \texttt{rbx}\mapsto_{\textsf{r}} \texttt{rbxv}' }_{\rtv}$
-$\specline{\texttt{rsp}\mapsto_{\textsf{r}} \texttt{rspv}' \ast \texttt{rbp}\mapsto_{\textsf{r}} \texttt{rbpv}' \ast \texttt{r12}\mapsto_{\textsf{r}} \texttt{r12v}' \ast \texttt{r13}\mapsto_{\textsf{r}} \texttt{r13v}' \ast \texttt{r14}\mapsto_{\textsf{r}} \texttt{r14v}' \ast \texttt{r15}\mapsto_{\textsf{r}} \texttt{r15v}'}_{\rtv}$
-$\specline{\mathsf{ContextAt}(old,[\mathsf{rbxv,\ldots,\rtv}]) \ast \mathsf{ContextAt}(new,[\mathsf{rbxv}',\ldots,\rtv']) }_{\rtv}$
-mov cr3, 56[rsi] ;; <-- Switch to the new address space (*\label{line:end_restoreC}*)
-$\specline{ [\rtv](\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,\Xi,m) \ast \mathsf{ContextAt}(old,[\mathsf{rbxv,\ldots,\rtv}]) \ast \mathsf{ContextAt}(new,[\mathsf{rbxv}',\ldots,\rtv']))  }_{\rtv}$ (*\label{line:modality_switchC}*)
-$\specline{ \mathcal{I}\texttt{ASpace}(\theta',\Xi',m') \ast \texttt{Pother} \ast  \texttt{rsi}\mapsto_{\textsf{r}} new \ast \texttt{rdi}\mapsto_{\textsf{r}} old \ast \ldots }_{\rtv'}$
-\end{lstlisting}
-\vspace{-1em}
-\caption{Basic task switch code that switches address spaces.}
-\label{fig:swtchC}
-\end{figure}
-
-\subsection{Change of Address Space}
-A critical piece of \emph{trusted} code in current verified OS kernels is the assembly code to change the current address space; current verified OS kernels currently
-lack effective ways to specify and reason about this low-level operation, for reasons outlined in Section \ref{sec:relworkC}.
-
-Figure \ref{fig:swtchC} gives simplified code for a basic task switch, the heart of an OS scheduler implementation. This is code that saves the context (registers and stack)
-of the running thread (here in a structure pointed to by \lstinline|rdi|'s value shown in Lines \ref{line:start_saveC}--\ref{line:end_saveC} of Figure \ref{fig:swtchC}) and restores the context of 
-an existing thread (from \lstinline|rsi| shown in abbreviated Lines \ref{line:start_restoreC}--\ref{line:end_restoreC}), including the corresponding change of address space for a target thread in another process.
-This code assumes the System V AMD64 ABI calling convention, where the normal registers not mentioned are caller-save, and therefore saved on the stack of the thread
-that calls this code, as well as on the new stack of the thread that is restored; thus, only the callee-save registers and \texttt{cr3} must be
-restored.\footnote{We are simplifying by ignoring non-integer registers (e.g., floating point, vector registers),
-and the caller-save registers should be initialized to 0 to avoid leaking information across processes, but this captures the key challenges.}
-With the addition of a return instruction, this code would satisfy the signatures of the C functions \footnote{This is the function in UNIX 6th Edition 
-with the infamous ``You are not expected to understand this'' comment~\cite{lions1996lions}.}
-\centerline{\lstinline[language=C]|void swtch(context_t* save, context_t* restore);|}\\
-A call to this code begins by executing one thread (up to just before line \ref{line:end_saveC}) in an address space ($\rtv$), whose information will be saved in a structure at address $old$,
-and finishes the execution by executing a different thread in a different address space (line \ref{line:end_restoreC} on), whose information is initially in $new$.
-
-Because this code does not directly update the instruction pointer, it is worth explaining \emph{how} this switches threads: by switching address spaces and stacks. 
-This is meant to be called with a return address for the current thread stored on the current stack when called. 
-The precondition of the return address on the initial stack requires the callee-save register values at the time of the call: those stored in the first
-half of the code.
-Likewise, part of the invariant of the stack of the second thread, the one being restored, is that the return address on \emph{that} stack requires the saved callee-save registers stored in that context be in registers as its precondition.
-
-The wrinkle and the importance of the modal treatment of assertions is that the target thread's precondition is \emph{relative to its address space}, 
-not the address space of the calling thread, which is reflected by
-the other-space modality
-$[\rtv'](\mathcal{I}\texttt{ASpace}(\theta,\Xi,m) \ast \texttt{Pother})$
-in the specfication. 
-The precondition of this code,
-in context, would include that the initial stack pointer (before \lstinline|rsp| is updated)
-has a return address expecting the then-current callee-save register values and 
-suitably updated (i.e., post-return) stack in the \emph{current} (initial) address space;
-this would be part of \textsf{P} in the precondition.
-The specification also requires that
-the stack pointer saved in the context to restore expects the same of the saved registers and stack
-\emph{in the other address space}. 
-The other-space modality plays a critical role here; \textsf{Pother} would contain these assumptions in the other
-address space.
-\looseness=-1
-
-% Lines 10--16 save the current context into memory (in the current address space).
-% Line 22 saves the initial page table root.
-% Lines 33--38 begin restoring the target context, including the stack pointer (line 33),
-% which may not be mapped in the address space at that time: it is the stack for the context being
-% loaded into the CPU.
-% The actual address switch occurs on line 45, which is verified with our modal rule for updating \lstinline|cr3|,
-% and thus shifts resources in and out of other-space modalities as appropriate.
-
-The postcondition is analagous to the precondition, but interpreted \emph{in the new address space}: the then-current (updated) stack would have a return address expecting the new (restored) register values (again, in \textsf{Pother}),
-and the saved context's invariant captures the precondition for restoring its execution \emph{in the previous address space} (as part of \textsf{P}).
-
-Immediately after the page table switch, assertions about the saved and restored contexts are
-guarded by a modality for the retiring
-address space \rtv{} (Line \ref{line:modality_switch}), per
-\textsc{WriteToRegCtlFromRegModal} (Figure \ref{fig:wpdamd}),
-because
-there is no guarantee that the data structures of the previous address space are mapped in the new address space.
-The ability to transfer that points-to information out of that modality is specific to a given kernel's design. 
-Kernels that map kernel memory into all address spaces would need invariants
-that justified moving those assertions out of the other-space modality.
-% Following Spectre and Meltdown, this kernel design became less prevalent because speculative execution of accesses to kernel addresses could leak information even if the access did eventually cause a fault (the user/kernel mode permission check was done after fetching data from memory).
-Thus many modern kernels have reverted to the older kernel design where the kernel inhabits its own unique address space, and user processes have only enough extra material mapped in their address spaces to switch into the kernel (CPUs do not speculate past updates to \texttt{cr3}).
-\looseness=-1
-
-Although prior work has verified context switches within a single address space~\cite{ni2007contexts}, and context switches
-without any code before or after~\cite{syeda2020formal} (that is, not reasoning about the impact of address space change
-on what data were accessible), this is the first verification that handles both.
-\looseness=-1
-
-\begin{comment}
-\[  
-$\specline{\exists (\entryf ,\;\entrytr,\; \entrytw,\; \entryo,\;\textsf{pte\_addr },\paddr) \; \ldotp\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast  \texttt{r14}\mapsto_{\textsf{r}} \_ \ast \texttt{rdi}\mapsto_{r} \vaddr \ast \texttt{rax}\mapsto_{\textsf{r}} \textsf{ pte\_addr} \; \ast }_{\rtv}$
-$\specline{ \ulcorner  \texttt{addr\_L1 }(\vaddr, \entryo) = \paddr \urcorner \ast \ulcorner \texttt{entry\_present } \entryf \land \texttt{entry\_present } \entrytr \land  \texttt{entry\_present } \entrytw \urcorner \; \ast}_{\rtv}$
-$\specline{\nfpointsto{\mask\vaddr\maskfour\rtv}{\mask\vaddr\maskfouroff\rtv}\entryf\qone\naddr \; \ast \nfpointsto{\mask\vaddr\maskthree\entryf}{\mask\vaddr\maskthreeoff\entryf}\entrytr\qtwo\naddr \ast}_{\rtv}$
-$\specline{  \nfpointsto{\mask\vaddr\masktwo\entrytr}{\mask\vaddr\masktwooff\entrytr}\paddr\qthree\entryo \;\ast \texttt{pte\_addr} \mapsto_{\texttt{vpte}} \paddr \;(\texttt{wzero 64}) \ast \texttt{rax}\mapsto_{\textsf{r}} \texttt{pte\_addr}  }_{\rtv}$
-mov r14, rax ;; Save that before another call
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast  \texttt{r14}\mapsto_{\textsf{r}} \texttt{pte\_addr} \ast \texttt{rdi}\mapsto_{\textsf{r}} \vaddr \ast \texttt{rax}\mapsto_{\textsf{r}} \textsf{ pte\_addr} \; \ast }_{\rtv}$
-$\specline{ \nfpointsto{\mask\vaddr\maskfour\rtv}{\mask\vaddr\maskfouroff\rtv}\entryf\qone\naddr \ast \ulcorner \texttt{entry\_present } \entryf \land \texttt{entry\_present } \entrytr \land  \texttt{entry\_present } \entrytw \urcorner \ast}_{\rtv}$
-$\specline{  \nfpointsto{\mask\vaddr\maskthree\entryf}{\mask\vaddr\maskthreeoff\entryf}\entrytr\qtwo\naddr \ast \nfpointsto{\mask\vaddr\masktwo\entrytr}{\mask\vaddr\masktwooff\entrytr}\paddr\qthree\entryo \;\ast}_{\rtv}$
-$\specline{\texttt{pte\_addr} \mapsto_{\texttt{vpte}} \paddr \;(\texttt{wzero 64}) \ast \texttt{rax}\mapsto_{\textsf{r}} \texttt{pte\_addr}  }_{\rtv}$
-call alloc_phys_page_or_panic
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast  \texttt{r14}\mapsto_{\textsf{r}} \texttt{pte\_addr} \ast \texttt{rdi}\mapsto_{\textsf{r}} \vaddr \;\ast \nfpointsto{\mask\vaddr\maskfour\rtv}{\mask\vaddr\maskfouroff\rtv}\entryf\qone\naddr \ast}_{\rtv}$
-$\specline{  \nfpointsto{\mask\vaddr\maskthree\entryf}{\mask\vaddr\maskthreeoff\entryf}\entrytr\qtwo\naddr \ast \nfpointsto{\mask\vaddr\masktwo\entrytr}{\mask\vaddr\masktwooff\entrytr}\paddr\qthree\naddr \;\ast}_{\rtv}$
-$\specline{\texttt{pte\_addr} \mapsto_{\texttt{vpte}} \paddr\; (\texttt{wzero 64})  \ast \ulcorner \texttt{entry\_present } \entryf \land \texttt{entry\_present } \entrytr \land  \texttt{entry\_present } \entrytw \urcorner}_{\rtv}$
-$\specline{\exists \texttt{ fpaddr} \ldotp \ulcorner \texttt{aligned fpaddr} \urcorner \ast \texttt{rax}\mapsto_{\textsf{r}} \texttt{fpaddr+3} \ast \texttt{fpaddr} \mapsto_{\textsf{p}} (\texttt{wzero 64}) \ast \ulcorner \texttt{entry\_present (fpaddr+3)}\urcorner}_{\rtv}$
-;; Calculate new L1 entry
-mov [r14], rax ;; store the page table entry, mapping the page
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast  \texttt{r14}\mapsto_{\textsf{r}} \texttt{pte\_addr} \ast \texttt{rdi}\mapsto_{\textsf{r}} \vaddr \;\ast \nfpointsto{\mask\vaddr\maskfour\rtv}{\mask\vaddr\maskfouroff\rtv}\entryf\qone\naddr \ast}_{\rtv}$
-$\specline{  \nfpointsto{\mask\vaddr\maskthree\entryf}{\mask\vaddr\maskthreeoff\entryf}\entrytr\qtwo\naddr \ast \nfpointsto{\mask\vaddr\masktwo\entrytr}{\mask\vaddr\masktwooff\entrytr}\paddr\qthree\entryo \;\ast}_{\rtv}$
-$\specline{\texttt{pte\_addr} \mapsto_{\texttt{vpte}} \paddr \;(\texttt{fpaddr+3}) \; \ast \ulcorner \texttt{entry\_present } \entryf \land \texttt{entry\_present } \entrytr \land  \texttt{entry\_present } \entrytw \urcorner }_{\rtv}$
-$\specline{\ulcorner \texttt{aligned fpaddr} \urcorner \ast \texttt{rax}\mapsto_{\textsf{r}} \texttt{fpaddr+3} \ast \texttt{fpaddr} \mapsto_{\textsf{p}} (\texttt{wzero 64}) \ast \ulcorner \texttt{entry\_present fpaddr+3}\urcorner}_{\rtv}$
-$\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\; \sqsubseteq $
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast  \texttt{r14}\mapsto_{\textsf{r}} \texttt{pte\_addr} \ast \texttt{rdi}\mapsto_{\textsf{r}} \vaddr \ast }_{\rtv}$
-$\specline{\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\vaddr,\entryf,\entrytr,\entrytw,\fpaddr+3) \ast \ulcorner \theta \;!!\;\vaddr = \texttt{None}\urcorner \; \ast}_{\rtv}$
-$\specline{\ulcorner \texttt{aligned fpaddr} \urcorner \ast \texttt{rax}\mapsto_{\textsf{r}} \texttt{fpaddr+3} \ast \texttt{fpaddr} \mapsto_{\textsf{p}} (\texttt{wzero 64}) }_{\rtv}$
-$\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\; \sqsubseteq $
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace} (<[\vaddr:=\texttt{fpaddr}]> \theta,m) \ast}_{\rtv}$
-$\specline{\ulcorner \texttt{aligned fpaddr} \urcorner \ast \texttt{fpaddr} \mapsto_{\textsf{p}} \textsf{ wzero 64} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  \ast\sumwalkabs\vaddr\qfrac\fpaddr}_{\rtv}$
-$\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\; \sqsubseteq $
-  $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace} (<[\vaddr:=\texttt{fpaddr}]> \theta,m) \ast \vaddr \mapsto_{\textsf{vpte}}\; \{\qfrac\} \;\fpaddr \textsf{ wzero 64}}_{\rtv}$
-$\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\; \sqsubseteq $
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace} (<[\vaddr:=\texttt{fpaddr}]> \theta,m) \ast \vaddr \mapsto_{\textsf{v}}\; \{\qfrac\} \textsf{wzero 64}}_{\rtv}$
-\end{comment}

### 2025-07-23 examples_C.tex
a few more edits, but out of steam for the night

@@
 
-% The mapping operation of Figure \ref{fig:mapping_codeC} assumes an operation \textsf{walkpgdir} which must traverse the page tables
-% in order to locate the address of the L1 entry to update --- 
-% % possibly allocating tables for levels 3, 2, and 1 in the process,
-% % installing them into levels 4, 3, and 2, along the way.
-% possibly allocating new L3, L2, and L1 tables as necessary.
-
-\todo[inline]{this and other figures need assertion updates to talk about C: see entry\_val (needs adjustment to points-to preconditions
-and explanation of entry pointing to a C bitfield so entry\_val encompasses the various bits), nextpaddr
-(was logical name in asm proof, but not a C variable)
-}
-\todo[inline]{need to narrate the main parts of assertions in prose}
-
 \begin{figure}\footnotesize
@@
     /*Now we know that entry is initialized, so we satisfy the condition to access children list */(*\label{line:now_we_know}*)
-    $\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}}  \textsf{pte\_initialized(pfn\_set(entryv nextpaddr))} \ast \ulcorner  \textsf{qfrac} = 1 \land \lnot(\textsf{entry\_present entry\_val})  \urcorner }_{\rtv}$
+    $\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}}  \textsf{entry} \; \textsf{pte\_initialized(pfn\_set(entryv nextpaddr))} \ast \ulcorner  \textsf{qfrac} = 1 \land \lnot(\textsf{entry\_present entry\_val})  \urcorner }_{\rtv}$
     $\specline{\forall_{i\in \textsf{0 ...
@@
   } (*\label{line:end_of_allocation_pathC}*)
-  $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \}),m) \ast \textsf{entry} \mapsto_{\textsf{id}} \textsf{\_} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
-  $\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{(pte\_initialized (entry\_val.pfn))} \urcorner }_{\rtv}$
+  $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry}\mapsto\textsf{v} \}),m) \ast \textsf{entry} \mapsto_{\textsf{id}} \textsf{\_} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
+  $\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry} \; \textsf{(pte\_initialized (entry\_val.pfn))} \urcorner }_{\rtv}$
   $\specline{\forall_{i\in \textsf{0 ...
@@
 \todo[inline]{this figure has an issue with entry, which is used both as a logical variable and a program variable (e.g., assertion on line 6).
-I would also expect to see a more prominent use of \textsf{v} in the precondition
 Line 20 also seems like it lost the paddr of the vpte
 }
-\todo[inline,color=cyan]{yes update logical entry to entryp, maybe remove entry $\mapsto$ v from $\Xi$ in first line?
-and fix 20}
+\todo[inline,color=cyan]{yes update logical entry to entryp, which requires also adding a local-variable assertion that entry the variable
+holds entryp+KERNBASE}
 \vspace{-1em}

### 2025-07-23 examples_C.tex
feeling good about revisions up through end of 5.2.5

@@
 
-\newcommand{\pvmapping}[1]{\mathcal{A}\textsf{PVMappings}(#1)}
+\newcommand{\pvmapping}[1]{\mathcal{A}\textsf{P2VMappings}(#1)}
 
@@
   pte_t *next;
-  $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{entry}\mapsto\textsf{v}\},m) \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$(*\label{line:precondition_entry_out}*)
-  $\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present entry\_val})\urcorner}_{\rtv}$ (*\label{line:get_next_vpte_preconditionC}*)
-  $\specline{\ulcorner\textsf{entry\_present(entry\_val)}\urcorner \wand \forall_{i\in\textsf{0..511}} \; \ghostmaptoken{\textsf{id}}{\textsf{table\_root(entry\_val.pfn) + i * 8}}{\textsf{v-1}} }_{\rtv} $(*\label{line:precondition_conditional}*)
+  $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{entryp}\mapsto\textsf{v}\},m) \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta} \ast \mathsf{entry} \mapsto_r \mathsf{entryp+KERNBASE} \ast \ghostmaptoken{\textsf{id}}{(\mathsf{entryp})}{\textsf{level}}}_{\rtv}$(*\label{line:precondition_entry_out}*)
+  $\specline{ \textsf{entryp+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entryp entryv} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present entryv})\urcorner}_{\rtv}$ (*\label{line:get_next_vpte_preconditionC}*)
+  $\specline{\ulcorner\textsf{entry\_present(entryv)}\urcorner \wand \forall_{i\in\textsf{0..511}} \; \ghostmaptoken{\textsf{id}}{\textsf{table\_root(entryv.pfn) + i * 8}}{\textsf{v-1}} }_{\rtv} $(*\label{line:precondition_conditional}*)
   /* Reading the page table is entry justified by the virtual pte-points-to */
   if (!entry->present(*\label{line:read_entry_contentsC}*)(*\label{line:conditional_childrenC}*)(*\label{line:check_entry_present_jumpC}*)) {// If the present bit is zero, need to allocate next level (* \label{line:check_entry_presentC} *)(* \label{line:mask_presentC} *)
-    (*\label{line:refined_fractional_ownership}*)(*\label{line:pass_addrof_nextC}*)(*\label{line:alloc_path_startC}*)$\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \land \lnot(\textsf{entry\_present entry\_val})  \urcorner }_{\rtv}$ (*\label{line:after_concluding_qfrac1C}*)
+    (*\label{line:refined_fractional_ownership}*)(*\label{line:pass_addrof_nextC}*)(*\label{line:alloc_path_startC}*)$\specline{ \textsf{entryp+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entryp entryv} \ast \ulcorner  \textsf{qfrac} = 1 \land \lnot(\textsf{entry\_present entryv})  \urcorner }_{\rtv}$ (*\label{line:after_concluding_qfrac1C}*)
     pte_initialize(entry); // Allocate a zeroed physical page and install in entry(*\label{line:call_to_pte_initializeC}\label{line:page_of_capsC}*)
-    $\specline{ \ldots \ast \ulcorner  \textsf{qfrac} = 1 \land \lnot(\textsf{entry\_present entry\_val})  \urcorner \ast \ulcorner  \textsf{qfrac} = 1 \land \lnot(\textsf{entry\_present (pfn\_set (entry\_val nextpaddr))}  \urcorner   }_{\rtv}$
+    $\specline{ \ldots \ast \ulcorner  \textsf{qfrac} = 1 \land \lnot(\textsf{entry\_present entryv})  \urcorner \ast \ulcorner  \textsf{qfrac} = 1 \land \lnot(\textsf{entry\_present (pfn\_set (entryv nextpaddr))}  \urcorner   }_{\rtv}$
     $\specline{ \begin{array}{l}\textsf{entry\_present (pte\_initialized (pfn\_set(entryv nextpaddr)))} \wand  \\ \;\;\;\;\;\;\;\;\  \forall_{i\in \textsf{0 ...
-511} } \ldotp  \ghostmaptoken{\textsf{id}}{((\textsf{table\_root (pte\_initialized (pfn\_set(entry\_val nextpaddr)))}) + \textsf{i * 8})}{\textsf{v-1}} \end{array}  }$ (*\label{line:page_of_capsC}*)
+511} } \ldotp  \ghostmaptoken{\textsf{id}}{((\textsf{table\_root (pte\_initialized (pfn\_set(entryv nextpaddr)))}) + \textsf{i * 8})}{\textsf{v-1}} \end{array}  }$ (*\label{line:page_of_capsC}*)
     entry->present = 1; // Install newly-allocated level n table, mark present(*\label{line:install_new_entryC}*)
     /*Now we know that entry is initialized, so we satisfy the condition to access children list */(*\label{line:now_we_know}*)
-    $\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}}  \textsf{entry} \; \textsf{pte\_initialized(pfn\_set(entryv nextpaddr))} \ast \ulcorner  \textsf{qfrac} = 1 \land \lnot(\textsf{entry\_present entry\_val})  \urcorner }_{\rtv}$
+    $\specline{ \textsf{entryp+KERNBASE} \mapsto_{\textsf{vpte,qfrac}}  \textsf{entryp} \; \textsf{pte\_initialized(pfn\_set(entryv nextpaddr))} \ast \ulcorner  \textsf{qfrac} = 1 \land \lnot(\textsf{entry\_present entryv})  \urcorner }_{\rtv}$
     $\specline{\forall_{i\in \textsf{0 ...
-511} } \ldotp  \ghostmaptoken{\textsf{id}}{((\textsf{table\_root (pte\_initialized (pfn\_set(entry\_val nextpaddr)))}) + \textsf{i * 8})}{\textsf{v-1}}  }$(*\label{line:available_child_tokens}*) (*\label{line:alloc_path_endC}*)
+511} } \ldotp  \ghostmaptoken{\textsf{id}}{((\textsf{table\_root (pte\_initialized (pfn\_set(entryv nextpaddr)))}) + \textsf{i * 8})}{\textsf{v-1}}  }$(*\label{line:available_child_tokens}*) (*\label{line:alloc_path_endC}*)
   } (*\label{line:end_of_allocation_pathC}*)
-  $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry}\mapsto\textsf{v} \}),m) \ast \textsf{entry} \mapsto_{\textsf{id}} \textsf{\_} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
-  $\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry} \; \textsf{(pte\_initialized (entry\_val.pfn))} \urcorner }_{\rtv}$
+  $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entryp}\mapsto\textsf{v} \}),m) \ast \textsf{entryp} \mapsto_{\textsf{id}} \textsf{\_} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
+  $\specline{ \textsf{entryp+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entryp} \; \textsf{(pte\_initialized (entryv.pfn))} \urcorner }_{\rtv}$
   $\specline{\forall_{i\in \textsf{0 ...
-511} } \ldotp  \ghostmaptoken{\textsf{id}}{((\textsf{table\_root (pte\_initialized (entry\_val.pfn)))}) + \textsf{i * 8})}{\textsf{v-1}}  }$ (*\label{line:childrenC}*)
+511} } \ldotp  \ghostmaptoken{\textsf{id}}{((\textsf{table\_root (pte\_initialized (entryv.pfn)))}) + \textsf{i * 8})}{\textsf{v-1}}  }$ (*\label{line:childrenC}*)
   uintptr_t next_phys_addr = PTE_PFN_TO_ADDR(entry->pfn); // Fetch physical address of next table (*\label{line:finalpieceS}*)(*\label{line:extract_pfn}*)
@@
 % \end{lstlisting}
-\todo[inline]{this figure has an issue with entry, which is used both as a logical variable and a program variable (e.g., assertion on line 6).
-Line 20 also seems like it lost the paddr of the vpte
+\todo[inline]{
+Ismail, why does the implication on line 13 talk about entryv? Is that really what the Rocq proof talks about, after overwriting?
+Are we missing some existential quantification of the value in the conditional branch and lines 19-21?
 }
-\todo[inline,color=cyan]{yes update logical entry to entryp, which requires also adding a local-variable assertion that entry the variable
-holds entryp+KERNBASE}
 \vspace{-1em}
@@
 \lstinline|pte_get_next_table| accepts a \emph{virtual} address \lstinline|entry| which points to a level-$n+1$
-table entry. On Line \ref{line:read_entry_contentsC}, the code checks the present bit of the entry.
+table entry.
+
+On Line \ref{line:read_entry_contentsC}, the code checks the present bit of the entry.
 If the bit is unset, there is no level-$n$ table, so one must be allocated via \lstinline|pte_initialize| (explained shortly\footnote{
@@
 
+For this C presentation of what is really an assembly-level proof, we abuse notation and
+use our register points-to for C-level program variables. So on Line \ref{line:precondition_entry_out},
+$\mathsf{entry} \mapsto_r \mathsf{entryp+KERNBASE}$ means that the \emph{register representing} the C program variable
+\lstinline|entry| (per the calling convention, \lstinline|rdi|) holds the sum on the right (a constant offset added to the physical address $\mathsf{entryp}$ of the entry).
+That particular value is one subtlety of the proof related to the aforementioned kernel invariant, and is explained in Section \ref{sec:p2vC}.
+The \emph{virtual pte-points-to} from that virtual address (Line \ref{line:get_next_vpte_preconditionC}) indicates that it points to a value
+$\mathsf{entryv}$, a (possibly-unpopulated) page table entry.
+A virtual pte-points-to is defined just like the normal virtual points-to of Figure \ref{fig:virtualpointstosharing},
+except the physical address (\textsf{entryp} on Line \ref{line:get_next_vpte_preconditionC}) is explicit in the assertion
+rather than existentially quantified:\\
+\centerline{$
+    \vaddr\mapsto_{\textsf{vpte,q}} \; \paddr \; \vpage : \mathsf{vProp}~\Sigma \stackrel{\triangle}{=} 
+    \exists \delta\ldotp
+	(\lambda \mathit{cr3val}\ldotp
+	\ghostmaptoken{\delta{}s}{\mathit{cr3val}}{\delta}) \ast 
+  \fracghostmaptoken{\delta}{\vaddr}{\paddr}{\qfrac} \ast \paddr \mapsto_{\mathsf{p}} \vpage
+$}\\
+This supports memory access rules much like Figure \ref{fig:wpdamd}'s rules (which are proven sound using soundness of virtual pte-points-to rules as lemmas!)
+while exposing the physical location being modified.
+This is useful for page table modifications, which require knowing the physical location being changed.
+They are used throughout the software page table walk because entries in any level may be initialized.
+\looseness=-1
+
 \subsubsection{Address Space Invariant: Identity Mappings and Conditional Page Table Ownership}
 \label{subsec:identitymappingsC}
-Assembly-level verification of compiler output from Figure \ref{fig:calltopteinitializeC} is verbose but doable
-Section \ref{sec:logic}'s logic, but only after resolving two critical challenges.\todo{Ismail, not sure if this ``verbose but doable''
- is a good framing, we should discuss
-}\todo[color=cyan]{maybe say something about it feeling like normal asm verification \emph{because} of the virtual points-to enabling
-familiar-style proving except for specific manipulation of page tables.
-Maybe mention that it's well-beyond the capabilities of prior work because they just never tried? (said nicer...)
-}
+Assembly-level verification of compiler output from Figure \ref{fig:calltopteinitializeC} is verbose, but largely
+similar to other assembly-level verification thanks to Section \ref{sec:logic}'s logic (including virtual pte-points-to
+assertions),\todo{figure out how to highlight advance over prior work here}
+but only after resolving two critical challenges.
 Two key challenges stand out and end up affecting both the pre- and post-conditions, neither of which has been addressed by prior work.
@@
   \textsf{ where } \\
-   \textsf{ASpace\_Lookup}_{\textsf{id}}(\ptablestore,\Xi,m) \stackrel{\triangle}{=} \lambda\textsf{ cr3val} \ldotp \; \exists \gammaPred \; \ldotp \ulcorner m \; !!\; \textsf{cr3val} = \textsf{Some } \gammaPred \urcorner \ast
+   \textsf{ASpace\_Lookup}_{\textsf{id}}(\ptablestore,\Xi,m) \stackrel{\triangle}{=} \lambda\textsf{ cr3val} \ldotp \; \exists \gammaPred,\gammaPred' \; \ldotp \ulcorner m \; !!\; \textsf{cr3val} = \textsf{Some } \gammaPred \urcorner \ast
    % \ownGhost\gammaPred{\authfull{\ptableabswalk\ptablestore}} \ast  \ownGhostpv\gammaPred{\authfull{\pvmapping\Xi}}
-   \ptableabswalk{\delta,\ptablestore} \ast \pvmapping{\delta,\Xi}\\
-  \textsf{present\_L}(\vale,\mathsf{level})\stackrel{\triangle}{=} \mathsf{entry\_present}(\vale)\land \mathsf{level} > 0
+   \ptableabswalk{\delta,\ptablestore} \ast \pvmapping{\delta',\Xi}\\
+  \textsf{present\_L}(\vale,\mathsf{level})\stackrel{\triangle}{=} \mathsf{entry\_present}(\vale)\land \mathsf{level} > 1
 
@@
 \]
-\todo[inline]{Ismail, what's APVMappings?}
-\todo[inline,color=cyan]{APVMappings is authoritative ownership of the ghost map for the current address space's $\Xi$ (identity mappings).
-So the last line shouldn't just existentially quantify $\delta$, but also a $\delta'$ since the authoritative map for $\Xi$ should have its own ghost name
-(and does in the Rocq)
-}
 \vspace{-1em}
@@
 /* @param entry: virtual address of a non-present entry to initialize with a new physical page */
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \lnot(\textsf{entry\_present entry\_val} \ast \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,1}} \textsf{entry entryv} \ast \ghostmaptoken{\textsf{id}}{(\mathsf{entry})}{\textsf{level}} }$
+$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \lnot(\textsf{entry\_present entryv}) \ast \textsf{entryp+KERNBASE} \mapsto_{\textsf{vpte,1}} \textsf{entryp entryv} \ast \ghostmaptoken{\textsf{id}}{(\mathsf{entryp})}{\textsf{level}} }$
 void pte_initialize(pte_t *entry) {
@@
   entry->pfn = PTE_ADDR_TO_PFN((uintptr_t) local);(*\label{line:kalloc_install}*)
-  $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \lnot(\textsf{entry\_present entry\_val}) }$
-  $\specline{ \textsf{entry+KERNBASE} \mapsto_{\mathsf{vpte,1}} \mathsf{entry} \; \textsf{pfn\_set(entryv nextpaddr)}}$
+  $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \lnot(\textsf{entry\_present entryv}) }$
+  $\specline{ \textsf{entryp+KERNBASE} \mapsto_{\mathsf{vpte,1}} \mathsf{entryp} \; \textsf{pfn\_set(entryv nextpaddr)}}$
   $\specline{\ghostmaptoken{\delta{}s}{\rtv}{\delta} \ast \begin{array}{l} \ulcorner\textsf{entry\_present (pte\_initialize(pfn\_set(entry\_val nextpaddr)),level)}\urcorner \wand \\  \;\;\;\;\;\;\; \forall_{i\in\textsf{0..511}} \; \textsf{table\_root (pte\_initialized (pfn\_set (entry\_val nextpaddr))) + i * 8} \mapsto_{\textsf{id}} \textsf{level-1}    \end{array}  }_{\rtv}$(*\label{line:initialize_post}*)
@@
 \end{lstlisting}
-\todo[inline]{more confusion of entry as program variable holding vaddr
-vs entry as logical variable for the paddr, nextpaddr, etc.
-update, clarifying that it returns a value for which the mapping holds.
-Also, do we need the conditional implication in this postcondition? it alloc'ed from
-a region already in Xi.
-And shouldn't we be excluding entry (the paddr) from Xi here?
-}
-\todo[inline,color=cyan]{no, this proof really doesn't care if entry is in or out of Xi, no need to exclude, but maybe worth explaining}
 \vspace{-2em}
@@
 
+\subsubsection{The Specification of \lstinline|pte_get_next_table|}
+Note that the specification does \emph{not} assume a specific page table level and is used
+for all three level transitions (4 to 3, 3 to 2, 2 to 1).
+The logical parameter \textsf{v} represents the level
+of the entry passed as an argument (c.f. the token $\ghostmaptoken{\textsf{id}}{(\mathsf{entryp})}{\textsf{v}}$ witnessing
+that \textsf{entryp} is part of the page table invariant on Line \ref{line:precondition_entry_out}).
+
+This comes into play with a key subtlety of \lstinline|pte_get_next_table|'s
+specification: its precondition
+includes a virtual {pte-points-to} (discussed earlier, Line \ref{line:get_next_vpte_preconditionC})
+but its postcondition does not yield new virtual points-to assertions!
+Instead it merely computes the base virtual address of the next table, and returns adequate capabilities
+(discussed in Section \ref{subsec:identitymappingsC}, explicit on Line \ref{line:childrenC})
+for the \emph{caller} to construct a vpte-points-to for any entry of the next table level.
+\looseness=-1
+
 \todo[inline,color=cyan]{Colin edit frontier}
-
-\subsubsection{The Specification of \lstinline|pte_get_next_table|}
-Note that the specification does \emph{not} assume a specific page table level --- the logical parameter \textsf{v} represents the level
-of the entry passed as an argument, and this code
-is used for all three level transitions when traversing page tables (4 to 3, 3 to 2, 2 to 1).
-This comes into play with a subtlety of the specification of \lstinline|pte_get_next_table| that we will
-revisit several times: \lstinline|pte_get_next_table|'s specification
-assumes it is given a virtual \emph{vpte-pointsto}
-(a virtual points-to exposing the underlying physical address instead of existentially quantifying it;
- see Section \ref{sec:mapnewC}) granting access to the specified entry,
-but its postcondition does not yield new virtual points-to assertions!
-Instead it merely computes the base virtual address of the next table, and returns adequate capabilities (discussed in Section \ref{subsec:identitymappingsC})
-for the \emph{caller} to construct a vpte-pointsto for any entry of the next table level (if this is not an L1 entry ---
-the caller knows which level of the table this is for).
-\looseness=-1
 
@@
 
+\todo[inline]{update this since we moved the definition earlier}
 A PTE points-to, as we already have briefly mentioned, is defined just like the normal virtual points-to of Figure \ref{fig:virtualpointstosharingC}, except the physical address (here, \textsf{pa}) is explicit in the assertion

### 2025-07-24 examples_C.tex
wip checkin

@@
 but its postcondition does not yield new virtual points-to assertions!
-Instead it merely computes the base virtual address of the next table, and returns adequate capabilities
+Instead it merely computes the base virtual address of the next table, and returns adequate tokens
 (discussed in Section \ref{subsec:identitymappingsC}, explicit on Line \ref{line:childrenC})
@@
 
-\todo[inline,color=cyan]{Colin edit frontier}
-
-\subsubsection{Walking Page-Table Tree: Calling \textsf{pte\_get\_next\_table} for Each Level}
+\subsubsection{Walking The Page Tables: Calling \textsf{pte\_get\_next\_table} for Each Level}
 \label{wlkpgdirC}
@@
 \begin{lstlisting}[language=C,mathescape,escapeinside={(*}{*)}]
-/* @param l4: the root address of the page-table tree
+/* @param l4: the (virtual) root address of the page-table tree
    @param va: virtual address to be translated and mapped */
@@
   /*pte_get_next_table may have allocated a new page which updates entry*/
-  pte_t *l3_entry = &l3[L3Offset(va)]; /* Virtual address of L3 entry*/
+  pte_t *l3_entry = &l3[L3Offset(va)]; /* Virtual address of L3 entry*/(*\label{line:l3offset}*)
   $\specline{\ulcorner\textsf{l3}+\textsf{L3Offset(va)*8} = \textsf{l3\_entry} \land \textsf{table\_root(l4e\_val'.pfn)}  = \textsf{l3} \urcorner }_{\rtv}$
@@
   /*Similar steps & proofs for levels 3 and 2*/
-  pte_t *l2 = pte_get_next_table(l3_entry);
-  pte_t l2_entry = &l2[L2Offset(va)];
-  pte_t *l1 = pte_get_next_table(l2_entry);
-  /*Return the address of L1 entry with a similar index computation*/
-  return &l1[L1Offset(va)]; 
+  pte_t *l2 = pte_get_next_table(l3_entry);(*\label{line:getnextl3}*)
+  pte_t l2_entry = &l2[L2Offset(va)];(*\label{line:l2offset}*)
+  pte_t *l1 = pte_get_next_table(l2_entry);(*\label{line:getnextl2}*)
+  return &l1[L1Offset(va)]; // Return the address of L1 entry with a similar index computation(*\label{line:returnl1entryaddr}*)
 }
@@
 \todo[inline]{clarify virt/phys of input root and the output}
+\todo[inline]{Colin: I think \lstinline|l2_entry| and such are also supposed to be pointers --- they hold the address of an entry.
+And they are passed as arguments to pte get next table, which accepts pointers.
+}
 \caption{Walking page-table directory via calls to \textsf{pte\_get\_next\_table} in Figure \ref{fig:calltopteinitializeC}}
@@
 \end{figure}
-Implementing a software page-table walk amounts to calling \textsf{pte\_get\_next\_table} for each level as shown in Figure \ref{walkpgdirC}. 
+
+Implementing a software page table walk amounts to calling \textsf{pte\_get\_next\_table} for each level as shown in Figure \ref{walkpgdirC}. 
+\lstinline|walkpgdir| traverses the page table anchored at \lstinline|l4| (the virtual address of the base of the L4 table)
+and returns the virtual address of the L1 entry that should map the virtual address \lstinline|va|.
+For each level, \lstinline|walkpgdir| locates the appropriate entry by using the level-specific slice of \lstinline|va| to index into
+that table (simulating the hardware translation as in Figure \ref{fig:pagetables}), and passes the virtual address of that entry to
+\lstinline|pte_get_next_table| to get the base of the next level down.
+For example, Line \ref{line:start_l4_calcC} uses \lstinline|L4Offset| (a bit shifting and masking macro) to extract bits 39--47 of \lstinline|va|,
+and uses that to find the address of the L4 entry that would map \lstinline|va| in the address space.
+That is then passed to \lstinline|pte_get_next_table| on Line \ref{line:first_getnext_callC}, which
+returns the virtual address of the base of the L3 table. This process repeats for 3-to-2 (Lines \ref{line:l3offset}--\ref{line:getnextl3}),
+and 2-to-1 (Lines \ref{line:l2offset}--\ref{line:getnextl2}), after which Line \ref{line:returnl1entryaddr} returns
+the virtual address of the appropriate L1 entry.
+
+For each of the three level transitions, the main challenges for the proof are to
+construct a virtual pte-points-to assertion for the entry in that level's table,
+and pass the conditional assertion discussed in Section \ref{sec:p2vC} that if that entry is present then there are
+identity map tokens for the the physical address of each entry in the subsequent level's table.
+\todo[inline,color=cyan]{WIP checkpoint:
+continue by highlighting parallels between assertion on line 13 and precondition of get next table,
+and between assertions on lines 4 and 16 and the postcondition.
+Then finish talking through essentially the safety argument, then loop back and talk about accumulating the info for
+constructing a virtual points-to for va
+}
+
 The key part of the specification and proof for a page table walk is accumulation of memory mappings for the page-table entries 

### 2025-07-24 examples_C.tex
almost done with examples pass

@@
 \subsubsection{Self-Conditional Fractional Ownership and Installing a New Table}
+\label{sec:selfconditional}
 The fractional ownership of the entry's physical memory is subtle.
@@
 \begin{lstlisting}[language=C,mathescape,escapeinside={(*}{*)}]
-/* @param l4: the (virtual) root address of the page-table tree
+/* @param l4: the (virtual) address of the L4 page table tree
    @param va: virtual address to be translated and mapped */
 pte_t *walkpgdir(pte_t *l4, const void *va) {
-  $\specline{\textsf{P} \ast \ulcorner\textsf{rtv+KERNBASE}=\textsf{l4}\urcorner \ast \forall_{i\in\textsf{0..511}}\ghostmaptoken{\textsf{id}}{(\textsf{rtv}+i*8)}{\textsf{4}} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}}_{\rtv}$
+  $\specline{\textsf{P} \ast \ulcorner\textsf{rtv+KERNBASE}=\textsf{l4}\urcorner \ast \forall_{i\in\textsf{0..511}}\ghostmaptoken{\textsf{id}}{(\textsf{rtv}+i*8)}{\textsf{4}} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}}_{\rtv}$(*\label{line:walkpgdir_pre}*)
   pte_t *l4_entry = &l4[L4Offset(va)]; // Virtual address of L4 entry(*\label{line:start_l4_calcC}*)
-  /*At the 4th level (and other levels ignoring the level-specific indices) an access amounts to:*/
-  /*1.Load l4 and va*/
-  /*2.Extract 9-bit (0x1ff=511) L4 index from the virtual address va*/
-  /*3.Multiply by 8 (byte index), add to l4 for entry address (*\label{line:end_l4_calcC}*) */
-  /*4.Store virt. addr. to local var l4_entry; logical l4_entry is physical!*/
   $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{l4\_entry}\},m) \ast \forall_{i\in\textsf{0..511}\setminus\{\textsf{L4Offset(va)}\}}\ghostmaptoken{\textsf{id}}{(\textsf{rtv}+i*8)}{\textsf{4}}\ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
@@
   pte_t *l3 = pte_get_next_table(l4_entry); (*\label{line:first_getnext_callC}*) 
+  /*pte_get_next_table may have allocated a new page which updates entry value*/
   $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{l4\_entry}\},m)  \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
-  $\specline{\forall_{i\in\textsf{0..511}} \ghostmaptoken{\textsf{id}}{\textsf{table\_root(l4e\_val'.pfn) + i * 8}}{\textsf{3}}  \ast \ulcorner\textsf{l3-KERNBASE}=\textsf{table\_root(l4e\_val'.pfn)}\urcorner} $
+  $\specline{\forall_{i\in\textsf{0..511}} \ghostmaptoken{\textsf{id}}{\textsf{table\_root(l4e\_val'.pfn) + i * 8}}{\textsf{3}}  \ast \ulcorner\textsf{l3-KERNBASE}=\textsf{table\_root(l4e\_val'.pfn)}\urcorner} $(*\label{line:l3tokens}*)
   $\specline{ \textsf{l4\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l4\_entry l4e\_val'} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l4e\_val})\urcorner}_{\rtv}$(*\label{line:first_pte_pointstoC}*)
-  /*pte_get_next_table may have allocated a new page which updates entry*/
   pte_t *l3_entry = &l3[L3Offset(va)]; /* Virtual address of L3 entry*/(*\label{line:l3offset}*)
@@
   pte_t *l2 = pte_get_next_table(l3_entry);(*\label{line:getnextl3}*)
-  pte_t l2_entry = &l2[L2Offset(va)];(*\label{line:l2offset}*)
+  pte_t *l2_entry = &l2[L2Offset(va)];(*\label{line:l2offset}*)
   pte_t *l1 = pte_get_next_table(l2_entry);(*\label{line:getnextl2}*)
-  return &l1[L1Offset(va)]; // Return the address of L1 entry with a similar index computation(*\label{line:returnl1entryaddr}*)
+  return &l1[L1Offset(va)]; // Return virtual address of L1 entry with a similar index computation(*\label{line:returnl1entryaddr}*)
 }
@@
 \end{lstlisting}
-
+\todo[inline]{same entry confusion of virt program and phys logical; fix and explain names}
 \else
@@
 \fi
-\todo[inline]{clarify virt/phys of input root and the output}
-\todo[inline]{Colin: I think \lstinline|l2_entry| and such are also supposed to be pointers --- they hold the address of an entry.
-And they are passed as arguments to pte get next table, which accepts pointers.
-}
 \caption{Walking page-table directory via calls to \textsf{pte\_get\_next\_table} in Figure \ref{fig:calltopteinitializeC}}
@@
 identity map tokens for the the physical address of each entry in the subsequent level's table.
-\todo[inline,color=cyan]{WIP checkpoint:
-continue by highlighting parallels between assertion on line 13 and precondition of get next table,
-and between assertions on lines 4 and 16 and the postcondition.
-Then finish talking through essentially the safety argument, then loop back and talk about accumulating the info for
-constructing a virtual points-to for va
-}
+For traversing the L4 table, this proceeds by extracting the relevant identity map token
+(the assertion on Line \ref{line:walkpgdir_pre} has one for each entry in the table)
+and pulling the resources for physical address $\mathsf{rtv}+8*\textsf{L4Offset}(\textsf{va})$
+out of the identity map invariant: 
+parts \textcircled{1} and \textcircled{2} of Figure \ref{fig:peraspaceinvariant_with_p2v_extensionC}) give a virtual pte-points-to,
+and \textcircled{3} and \textcircled{4} satisfy other parts of \lstinline|pte_get_next_table|'s precondition.
+This justifies the call on Line \ref{line:first_getnext_callC},
+which returns the virtual address of the base of the appropriate L3 table and whose postcondition
+includes 512 identity map tokens for each of those entries (anchored to the returned virtuall address minus \textsf{KERNBASE}).
+That invariant (Line \ref{line:l3tokens}) is analogous to the one that justified the 4-3 step (Line \ref{line:walkpgdir_pre}), and the next two steps proceed the same way.
 
-The key part of the specification and proof for a page table walk is accumulation of memory mappings for the page-table entries 
-visited and frame addresses for page-tables. 
-For example, Lines \ref{line:ex_l4_vpteC} and \ref{line:ex_l3_vpteC} in Figure \ref{walkpgdirC} show the virtual pte-pointsto assertions for L4 and L3 entries.
-In the final post-condition, we expect the accumulation of these resources from each level -- $\textsf{R}_{\textsf{walk}}$ -- 
-which allows us to construct and return the path to the L1 entry in the tree to insert a new page.
+Even given the slight adaptation of our assembly-level proof for the C-level presentation in Figure \ref{walkpgdirC},
+the proof outline in the figure omits may assertions that hold at various points in time.
+But by Line \ref{line:ex_l3_vpteC}, it should be clear that the proof is on track to accumulate a set of similar assertions for each level.
+Figure \ref{fig:rwalkC} expands the abbreviated postcondition to include the full set of facts that are accumulated in this way.
+$\mathsf{R}_\mathsf{walk}$ together with $\mathsf{R}_{\mathsf{l1e}}$ should look quite similar to \lstinline|L4_L1_PointsTo| (Figure \ref{fig:strongvirtualpointsto}),
+and this is quite intentional. \lstinline|walkpgdir|'s execution observes enough
+to determine that there is most of an address translation for virtual address \lstinline|va|,
+at least down to a possibly-invalid L1 entry (which \lstinline|walkpgdir|'s caller, \lstinline|vaspace_mappage|, will check). 
+Each virtual pte-points-to in $\mathsf{R}_\mathsf{walk}$ internally contains the physical points-to portion of one page table walk step for \lstinline|L4_L1_PointsTo|,
+and the pure assertions in $\mathsf{R}_\mathsf{walk}$ ensure the address arithmetic works.
+$\mathsf{R}_{\mathsf{l1e}}$ also includes the self-conditional fractional ownership of the L1 entry (per Section \ref{sec:selfconditional})
+for the caller to initialize the entry if it is empty, making it possible for the caller to complete virtual points-to assertions
+for a newly-mapped page, as we discuss next.
 
-This is the code which performs most actual physical-to-virtual conversions using the identity mapping portion of the per-address-space invariant.
-\lstinline|walkpgdir| accepts a \emph{virtual} pointer to the base of the L4 table, and the address to translate.
-The precondition provides knowledge that the virtual base of the L4 is at the appropriate offset from the current \lstinline|cr3| value,
-but does not provide a virtual points-to assertion --- because the function must calculate (Lines \ref{line:start_l4_calcC}--\ref{line:end_l4_calcC})
-which entry it needs access.
-Instead, the precondition has 512 identity map tokens, guaranteeing that every entry on the page is subject to the identity mapping invariant.
-Line \ref{line:end_l4_calcC} calculates the virtual address of the relevant entry, and the subsequent view shift
-pulls that entry out of the identity mapping ($\Xi$) and fetches its corresponding resources as
-described by Figure \ref{fig:peraspaceinvariant_with_p2v_extensionC} and Section \ref{subsec:identitymappingsC}.
-The ghost translation and physical location are used to form the virtual pte-pointsto for the L4 entry
-(Line \ref{line:first_pte_pointstoC}), with entry validity and next-level indexing
-satisfying the rest of the precondition for \lstinline|pte_get_next_table|.
-Then, as described earlier, checks the valid bit in the indicated
-entry and either returns the (unconditional) tokens for the L3 entry physical addresses (if valid), or
-allocates into the entry and returns new (also unconditional) tokens for the L3 entry physical addresses.
-\lstinline|pte_get_next_table|'s first call (Line \ref{line:first_getnext_callC}) returns
-the virtual address of the base of the L3 table. Then the situation to move from that pointer to the base of the L2
-is just like the process just followed: the proof calculates the address of the relevant
-L3 entry, uses the appropriate L3 identity mapping token to construct a virtual pte-pointsto to that entry,
-and passes that along with additional resources pulled out of the invariant to another call to
-\lstinline|pte_get_next_table|. That call then returns the base of an L2 table, and the process
-repeats until the function returns the virtual address of the relevant L1 entry.
-That will then be used in the next section by the caller of \lstinline|walkpgdir|
-to install a new mapping.
+% The key part of the specification and proof for a page table walk is accumulation of memory mappings for the page-table entries 
+% visited and frame addresses for page-tables. 
+% For example, Lines \ref{line:ex_l4_vpteC} and \ref{line:ex_l3_vpteC} in Figure \ref{walkpgdirC} show the virtual pte-pointsto assertions for L4 and L3 entries.
+% In the final post-condition, we expect the accumulation of these resources from each level -- $\textsf{R}_{\textsf{walk}}$ -- 
+% which allows us to construct and return the path to the L1 entry in the tree to insert a new page.
+
+% This is the code which performs most actual physical-to-virtual conversions using the identity mapping portion of the per-address-space invariant.
+% \lstinline|walkpgdir| accepts a \emph{virtual} pointer to the base of the L4 table, and the address to translate.
+% The precondition provides knowledge that the virtual base of the L4 is at the appropriate offset from the current \lstinline|cr3| value,
+% but does not provide a virtual points-to assertion --- because the function must calculate (Lines \ref{line:start_l4_calcC}--\ref{line:end_l4_calcC})
+% which entry it needs access.
+% Instead, the precondition has 512 identity map tokens, guaranteeing that every entry on the page is subject to the identity mapping invariant.
+% Line \ref{line:end_l4_calcC} calculates the virtual address of the relevant entry, and the subsequent view shift
+% pulls that entry out of the identity mapping ($\Xi$) and fetches its corresponding resources as
+% described by Figure \ref{fig:peraspaceinvariant_with_p2v_extensionC} and Section \ref{subsec:identitymappingsC}.
+% The ghost translation and physical location are used to form the virtual pte-pointsto for the L4 entry
+% (Line \ref{line:first_pte_pointstoC}), with entry validity and next-level indexing
+% satisfying the rest of the precondition for \lstinline|pte_get_next_table|.
+% Then, as described earlier, checks the valid bit in the indicated
+% entry and either returns the (unconditional) tokens for the L3 entry physical addresses (if valid), or
+% allocates into the entry and returns new (also unconditional) tokens for the L3 entry physical addresses.
+% \lstinline|pte_get_next_table|'s first call (Line \ref{line:first_getnext_callC}) returns
+% the virtual address of the base of the L3 table. Then the situation to move from that pointer to the base of the L2
+% is just like the process just followed: the proof calculates the address of the relevant
+% L3 entry, uses the appropriate L3 identity mapping token to construct a virtual pte-pointsto to that entry,
+% and passes that along with additional resources pulled out of the invariant to another call to
+% \lstinline|pte_get_next_table|. That call then returns the base of an L2 table, and the process
+% repeats until the function returns the virtual address of the relevant L1 entry.
+% That will then be used in the next section by the caller of \lstinline|walkpgdir|
+% to install a new mapping.
 
@@
 
+\todo[inline,color=cyan]{colin revision frontier}
+
 \subsection{Mapping a New Page}
@@
 \begin{figure}\footnotesize
-  \begin{lstlisting}[mathescape,escapeinside={(*}{*)}]
-/*
-@param plm4: the root address of the page-table tree
+  \begin{lstlisting}[language=C,mathescape,escapeinside={(*}{*)}]
+/* @param plm4: the virtual address of the root of the page table tree
    @param va: virtual address to be translated and mapped
-@param fpaddr: physically allocated page's address in case the virtual address va is not backed with one
-*/
+   @param fpaddr: physical address of a zeroed page to map if va is not already mapped */
 void vaspace_mappage(pte_t *l4, void *va,uintptr_t fpaddr ) {
@@
 and for allowing access to the memory of the L1 entry via virtual address (R$_{\textsf{l1e}}$ in Figure \ref{fig:rwalkC}).
-
-\todo[inline]{update this since we moved the definition earlier}
-A PTE points-to, as we already have briefly mentioned, is defined just like the normal virtual points-to of Figure \ref{fig:virtualpointstosharingC}, except the physical address (here, \textsf{pa}) is explicit in the assertion
-rather than existentially quantified:\\
-\centerline{$
-    \vaddr\mapsto_{\textsf{vpte,q}} \; \paddr \; \vpage : \mathsf{vProp}~\Sigma \stackrel{\triangle}{=} 
-    \exists \delta\ldotp
-	(\lambda \mathit{cr3val}\ldotp
-	\ghostmaptoken{\delta{}s}{\mathit{cr3val}}{\delta}) \ast 
-  \fracghostmaptoken{\delta}{\vaddr}{\paddr}{\qfrac} \ast \paddr \mapsto_{\mathsf{p}} \vpage
-$}\\
-This supports rules for accessing memory
-at that virtual address, but exposing the physical location being modified
-makes this convenient to use for page table modifications, since we must ensure
-the modified data is at the specific physical location that will be used as the L1 entry.
-% After obtaining a virtual address \textsf{pte\_addr} in \textsf{rax} backed 
-% by the physical memory for the L1 entry that will be used to translate the virtual addresses
-% we are mapping, we save it to \textsf{r14} to be updated later in Line 9.
-
-%In the precondition, we see Line 12 allocates a fresh page-aligned, zero-initialized page  (at \textsf{fpaddr}),
-%returning a pre-filled PTE entry in \textsf{rax} ($+3$ sets the lower 2 bits).
-
-% , to hold the freshly
-% allocated physical page address (\textsf{fpaddr}) in Line X.
 

### 2025-07-24 examples_C.tex
finish main pass through experiments

@@
   entry->pfn = PTE_ADDR_TO_PFN((uintptr_t) local);(*\label{line:kalloc_install}*)
-  $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \lnot(\textsf{entry\_present entryv}) }$
-  $\specline{ \textsf{entryp+KERNBASE} \mapsto_{\mathsf{vpte,1}} \mathsf{entryp} \; \textsf{pfn\_set(entryv nextpaddr)}}$
+  $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \lnot(\textsf{entry\_present entryv}) \ast \textsf{entryp+KERNBASE} \mapsto_{\mathsf{vpte,1}} \mathsf{entryp} \; \textsf{pfn\_set(entryv nextpaddr)}}$
   $\specline{\ghostmaptoken{\delta{}s}{\rtv}{\delta} \ast \begin{array}{l} \ulcorner\textsf{entry\_present (pte\_initialize(pfn\_set(entry\_val nextpaddr)),level)}\urcorner \wand \\  \;\;\;\;\;\;\; \forall_{i\in\textsf{0..511}} \; \textsf{table\_root (pte\_initialized (pfn\_set (entry\_val nextpaddr))) + i * 8} \mapsto_{\textsf{id}} \textsf{level-1}    \end{array}  }_{\rtv}$(*\label{line:initialize_post}*)
@@
   $\specline{\textsf{entry\_present(l3e\_val)} \wand \forall_{i\in\textsf{0..511}} \;\ghostmaptoken{\textsf{id}}{\textsf{table\_root(l3\_entry.pfn) + i * 8}}{ \textsf{2}} } _{\rtv} $
-  /*Similar steps & proofs for levels 3 and 2*/
   pte_t *l2 = pte_get_next_table(l3_entry);(*\label{line:getnextl3}*)
@@
   return &l1[L1Offset(va)]; // Return virtual address of L1 entry with a similar index computation(*\label{line:returnl1entryaddr}*)
-}
-$\specline{\textsf{R}_{\textsf{walk}} \ast \textsf{R}_{\textsf{l1e}} }_{\rtv}$
+} $\specline{\textsf{R}_{\textsf{walk}} \ast \textsf{R}_{\textsf{l1e}} }_{\rtv}$
 \end{lstlisting}
@@
     \\ \ast \left(\begin{array}{l}  \textsf{l1\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l1\_entry}\;\textsf{l1e\_val} \ast \\ \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l1e\_val})\urcorner \end{array}\right) =\textsf{R}_{\textsf{l1e}}\end{array} $
-\todo[inline]{Colin: This is confusing: in the code \lstinline|l2_entry| is an actual entry word, but in these invariants they seem to be paddrs.}
+\todo[inline]{Colin: This is confusing: in the code \lstinline|l2_entry| is an actual entry word, but in these invariants they seem to be paddrs.
+Also, these should be 4 different fractions, not all qfrac
+}
 \caption{\textsf{R}$_{\textsf{walk}}$: Resources obtained from invariant for walking page-table directory with \emph{non-existing entries} via calls to \textsf{l1e\_get\_next\_table} in Figure \ref{fig:calltopteinitializeC}}
@@
 
-\todo[inline,color=cyan]{colin revision frontier}
-
 \subsection{Mapping a New Page}
 \label{sec:mapnewC}
-One of the key tasks of a page fault handler in a general-purpose OS kernel is
-to map new pages to an address space by writing into an existing page table via a call.\\
-\centerline{\textsf{vaspace\_mappage(pte\_t *l4, void *va,uintptr\_t fpaddr)}}\\
-in Figure \ref{fig:mapping_codeC}.
-To do so, with a given allocated fresh page (\textsf{fpaddr}), then calculate the appropriate
-known-valid page table walks (via \textsf{walkpgdir} Line \ref{line:call_walkpgdirC} in Figure \ref{fig:mapping_codeC}) and update
-the appropriate L1 page table entry (Line 35 in Figure \ref{fig:mapping_codeC});
-unmapping is the reverse of the logic we discuss here.
+Finally we come to the top-level routine for mapping a new page of memory into an address space
+by updating page tables --- the \lstinline|vaspace_mappage| function in Figure \ref{fig:mapping_codeC}.
+Again, our verification was carried out at the assembly level, but presented as on the original C for readability,
+and the proof outline omits all relevant facts in favor of the most critial assertions involved in the key parts of the proof.
+\lstinline|vaspace_mappage| is typically called by a page fault handler, to map a previously-reserved but lazily-allocated page.
+It is passed the virtual address of the L4 table base (\lstinline|l4|), the virtual address to map (\lstinline|va|),
+and the \emph{physical} address of an empty memory page which should be used as backing memory for \lstinline|va| and its surrounding page.
+It begins by calling \lstinline|walkpgdir| (Line \ref{line:call_walkpgdirC}) to return the virtual address of 
+the L1 entry which corresponds to \lstinline|va| (allocating intermediate tables as needed).
+It then checks if the entry is already initialized. If not, \lstinline|fpaddr| is installed
+into the L1 entry, which is then marked valid (setting the present bit), and the page is mapped.
+% To do so, with a given allocated fresh page (\textsf{fpaddr}), then calculate the appropriate
+% known-valid page table walks (via \textsf{walkpgdir} Line \ref{line:call_walkpgdirC} in Figure \ref{fig:mapping_codeC}) and update
+% the appropriate L1 page table entry (Line 35 in Figure \ref{fig:mapping_codeC});
+Unmapping is the reverse of the logic we discuss here.
 \looseness=-1
@@
   \begin{lstlisting}[language=C,mathescape,escapeinside={(*}{*)}]
-/* @param plm4: the virtual address of the root of the page table tree
+/* @param l4: the virtual address of the root of the page table tree
    @param va: virtual address to be translated and mapped
@@
 void vaspace_mappage(pte_t *l4, void *va,uintptr_t fpaddr ) {
-  $\specline{\textsf{P} \ast \textsf{l4} \mapsto_{\textsf{id}} \textsf{\_} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta} \ast \ulcorner\theta \; !!\; \vaddr = \texttt{None}\urcorner}_{\rtv}$
+  $\specline{\textsf{P} \ast \textsf{l4} \mapsto_{\textsf{id}} \textsf{\_} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta} \ast \ulcorner\theta \; !!\; \vaddr = \texttt{None}\urcorner}_{\rtv}$ (*\label{line:mappage_pre}*)
   pte_t *pteaddr = walkpgdir(l4, va); (*\label{line:call_walkpgdirC}*)
@@
     /*Set the present bit in entry*/
-    pteaddr->present = 1;
+    pteaddr->present = 1;(*\label{line:l1entry_setpresent}*)
   }
-  return;
   $\specline{ \begin{array}{l}\textsf{l1\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l1\_entry} \; \textsf{l1e\_val} \ast \ulcorner (\textsf{entry\_present l1e\_val}) \urcorner \ast \textsf{P} \\
@@
 
-In Figure \ref{fig:mapping_codeC}, we see an address ($\vaddr$) currently not
-mapped to a page ($\theta \; !!\; \vaddr = \texttt{None}$). Mapping a fresh
-physical page to back the desired virtual page first requires ensuring
-the existence of a memory location for an appropriate L1 table entry.
-The code uses a helper function \lstinline{walkpgdir} (discussed again in Section \ref{wlkpgdirC}).
-\textsf{walkpgdir}'s postcondition contains virtual \emph{PTE} points-to assertions ($\mapsto_{\textsf{vpte}}$)
-both for ensuring partial page table walk reaching the
-L1 entry (l1e) by asserting that higher levels of the page table exist (R$_{\textsf{walk}}$ in Figure \ref{fig:rwalkC}), 
-and for allowing access to the memory of the L1 entry via virtual address (R$_{\textsf{l1e}}$ in Figure \ref{fig:rwalkC}).
+For brevity this example is specialized to the case where $\vaddr$ is known to not be mapped:
+the precondition on Line \ref{line:mappage_pre} includes $\theta \; !!\; \vaddr = \texttt{None}$.\footnote{
+  The generalization to handling either case and returning an error is straightforward.
+}
+\todo[inline,color=cyan]{
+  Ismail, shouldn't the precondition include the 512 tokens for the L4 table, and shouldn't the l4 entry be excluded from $\Xi$?
+}
+The precondition directly implies the precondition of \lstinline|walkpgdir|, which as just discussed
+returns the virtual address of an allocated L1 entry, and whose postcondition contains almost all of the information
+needed to construct a virtual points-to for \lstinline|va| --- except information about the L1 entry
+being present and pointing to a data page.
+We already discussed for the upper level page-tables how the entry-present checks are handled, and
+Line \ref{line:mappage_pte_present_startC} is similar: $\mathsf{R}_{\textsf{l1e}}$ includes the self-conditional
+fractional permission for the L1 entry, so as it is not present, by Line \ref{line:l1entry_storeC}
+it is known that full permission is held to update that entry.
+Lines \ref{line:l1entry_storeC} and \ref{line:l1entry_setpresent} are verified using the pte-points-to
+memory store rule.\footnote{In the assembly, \lstinline|fpaddr| is already known to be word-aligned, so it is simply bitwise-and-ed with 1
+(setting the present flag) and a single store occurs.}
+\looseness=-1
 
-We already discussed for the upper level page-tables how the entry-present checks are handled.
-However, for L1 entries this check is left to the caller of the
-traversal function \textsf{walkpgdir}. In other words, unlike what we see in R$_{\textsf{walk}}$ for the upper levels, here all entry-present
-checks have already been performed; the specification in R$_{\textsf{l1e}}$ ensures that page table entry for L1 needs to be checked at the caller site. 
-In doing so, as we see in Figure \ref{fig:mapping_codeC}, the page reference \textsf{fpaddr} is linked back to the virtual address \textsf{va} 
-only if it does not already refer to a physical resource (lines \ref{line:mappage_pte_present_startC}--\ref{line:mappage_pte_present_endC} in Figure \ref{fig:mapping_codeC}).
-
-The crucial step in addition to traversing the page table in Figure \ref{walkpgdirC} is actually updating the L1 entry (Line \ref{line:updatepfnC} in Figure \ref{fig:mapping_codeC}),
-via the virtual address (\textsf{pt\_entry+KERNBASE}) known to translate to the appropriate physical address, in our example the L1
-table entry address ($\textsf{PTE\_ADDR\_TO\_PFN(fpaddr)}$).
-
+\todo[inline]{Colin will fix up this paragraph: it's right but also kind of weird, will handle during related work pass}
 Unlike the only prior work verifying analogous code for mapping a new page~\cite{kolanski08vstte,kolanski09tphols}, our proof above

### 2025-07-25 examples_C.tex
add important note
...

@@
 pte_t *walkpgdir(pte_t *l4, const void *va) {
-  $\specline{\textsf{P} \ast \ulcorner\textsf{rtv+KERNBASE}=\textsf{l4}\urcorner \ast \forall_{i\in\textsf{0..511}}\ghostmaptoken{\textsf{id}}{(\textsf{rtv}+i*8)}{\textsf{4}} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}}_{\rtv}$(*\label{line:walkpgdir_pre}*)
+  $\specline{\textsf{P} \ast \ulcorner\textsf{l4p}=\textsf{rtv}\urcorner\ast\ulcorner\textsf{rtv+KERNBASE}=\textsf{l4}\urcorner \ast \forall_{i\in\textsf{0..511}}\ghostmaptoken{\textsf{id}}{(\textsf{rtv}+i*8)}{\textsf{4}} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}}_{\rtv}$(*\label{line:walkpgdir_pre}*)
   pte_t *l4_entry = &l4[L4Offset(va)]; // Virtual address of L4 entry(*\label{line:start_l4_calcC}*)
-  $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{l4\_entry}\},m) \ast \forall_{i\in\textsf{0..511}\setminus\{\textsf{L4Offset(va)}\}}\ghostmaptoken{\textsf{id}}{(\textsf{rtv}+i*8)}{\textsf{4}}\ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
+  $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{l4p}\},m) \ast \forall_{i\in\textsf{0..511}\setminus\{\textsf{L4Offset(va)}\}}\ghostmaptoken{\textsf{id}}{(\textsf{rtv}+i*8)}{\textsf{4}}\ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
   $\specline{\textsf{entry\_present(l4e\_val)} \wand \forall_{i\in\textsf{0..511}} \; \ghostmaptoken{\textsf{id}}{\textsf{table\_root(l4e\_val)+ i * 8}}{\textsf{3}} }_{\rtv} $
-  $\specline{ \textsf{l4\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l4\_entry l4e\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l4e\_val})\urcorner}_{\rtv}$
+  $\specline{ \textsf{l4p+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l4p l4e\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l4e\_val})\urcorner}_{\rtv}$
   pte_t *l3 = pte_get_next_table(l4_entry); (*\label{line:first_getnext_callC}*) 
   /*pte_get_next_table may have allocated a new page which updates entry value*/
-  $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{l4\_entry}\},m)  \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
+  $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{l4p}\},m)  \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
   $\specline{\forall_{i\in\textsf{0..511}} \ghostmaptoken{\textsf{id}}{\textsf{table\_root(l4e\_val'.pfn) + i * 8}}{\textsf{3}}  \ast \ulcorner\textsf{l3-KERNBASE}=\textsf{table\_root(l4e\_val'.pfn)}\urcorner} $(*\label{line:l3tokens}*)
-  $\specline{ \textsf{l4\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l4\_entry l4e\_val'} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l4e\_val})\urcorner}_{\rtv}$(*\label{line:first_pte_pointstoC}*)
+  $\specline{ \textsf{l4p+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l4p l4e\_val'} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l4e\_val})\urcorner}_{\rtv}$(*\label{line:first_pte_pointstoC}*)
   pte_t *l3_entry = &l3[L3Offset(va)]; /* Virtual address of L3 entry*/(*\label{line:l3offset}*)
-  $\specline{\ulcorner\textsf{l3}+\textsf{L3Offset(va)*8} = \textsf{l3\_entry} \land \textsf{table\_root(l4e\_val'.pfn)}  = \textsf{l3} \urcorner }_{\rtv}$
-  $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{l4\_entry},\textsf{l3\_entry}\},m)  \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
-  $\specline{\forall_{i\in\textsf{0..511}\setminus\{\textsf{L3Offset(va)}\}} \; \textsf{table\_root(l4\_entry}_{\textsf{pfn}}) \textsf{ + i * 8} \mapsto_{\textsf{id}} \textsf{3} }_{\rtv}$
-  $\specline{ \textsf{l4\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac4}}\textsf{l4\_entry}\; \textsf{l4e\_val'} \ast \ulcorner  \textsf{qfrac4} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l4e\_val'})\urcorner}_{\rtv}$(*\label{line:ex_l4_vpteC}*)
-  $\specline{ \textsf{l3\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac3}}\textsf{l3\_entry}\; \textsf{l3e\_val} \ast \ulcorner  \textsf{qfrac3} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l3e\_val})\urcorner}_{\rtv}$(*\label{line:ex_l3_vpteC}*)
-  $\specline{\textsf{entry\_present(l3e\_val)} \wand \forall_{i\in\textsf{0..511}} \;\ghostmaptoken{\textsf{id}}{\textsf{table\_root(l3\_entry.pfn) + i * 8}}{ \textsf{2}} } _{\rtv} $
+  $\specline{\ulcorner\textsf{l3}+\textsf{L3Offset(va)*8} = \textsf{l3p} \land \textsf{table\_root(l4e\_val'.pfn)}  = \textsf{l3} \urcorner }_{\rtv}$
+  $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{l4p},\textsf{l3p}\},m)  \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
+  $\specline{\forall_{i\in\textsf{0..511}\setminus\{\textsf{L3Offset(va)}\}} \; \ghostmaptoken{\textsf{id}}{(\textsf{table\_root(l4e\_val.pfn}) \textsf{ + i * 8})} {\textsf{3}} }_{\rtv}$
+  $\specline{ \textsf{l4p+KERNBASE} \mapsto_{\textsf{vpte,qfrac4}}\textsf{l4p}\; \textsf{l4e\_val'} \ast \ulcorner  \textsf{qfrac4} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l4e\_val'})\urcorner}_{\rtv}$(*\label{line:ex_l4_vpteC}*)
+  $\specline{ \textsf{l3p+KERNBASE} \mapsto_{\textsf{vpte,qfrac3}}\textsf{l3p}\; \textsf{l3e\_val} \ast \ulcorner  \textsf{qfrac3} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l3e\_val})\urcorner}_{\rtv}$(*\label{line:ex_l3_vpteC}*)
+  $\specline{\textsf{entry\_present(l3e\_val)} \wand \forall_{i\in\textsf{0..511}} \;\ghostmaptoken{\textsf{id}}{\textsf{table\_root(l3e\_val.pfn) + i * 8}}{ \textsf{2}} } _{\rtv} $
   pte_t *l2 = pte_get_next_table(l3_entry);(*\label{line:getnextl3}*)
@@
 \end{lstlisting}
-\todo[inline]{same entry confusion of virt program and phys logical; fix and explain names}
+\todo[inline]{Ismail proof-read this figure for Colin's changes}
 \else
@@
 \begin{figure}\footnotesize
-$\begin{array}{l}\left( \begin{array}{l} \ulcorner \textsf{l3}+\textsf{L3Offset(va)} = \textsf{l3\_entry} \land \textsf{table\_root(l4\_entry}_{\textsf{pfn}})  = \textsf{l3} \\ \land \textsf{l2}+\textsf{L2Offset(va)} = \textsf{l2\_entry} \land \textsf{table\_root(l3\_entry}_{\textsf{pfn}})  = \textsf{l2} \\ \land \textsf{l1}+\textsf{L1Offset(va)} = \textsf{l1\_entry} \land \textsf{table\_root(l2\_entry}_{\textsf{pfn}})  = \textsf{l1} \urcorner  \ast \\  \textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{l4\_entry},\textsf{l3\_entry},\textsf{l2\_entry},\textsf{l1\_entry}\},m)  \\ \ghostmaptoken{\delta{}s}{\rtv}{\delta} \ast
-      \ghostmaptoken{\textsf{id}}{(\mathsf{l4\_entry})}{\textsf{4}}\ast \ghostmaptoken{\textsf{id}}{(\mathsf{l3\_entry})}{\textsf{3}} \ast \ghostmaptoken{\textsf{id}}{(\mathsf{l2\_entry})}{\textsf{2}}  \ast \\  \textsf{l4\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l4\_entry} \; \textsf{l4e\_val}  \ast \\  \textsf{l3\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l3\_entry}\;\textsf{l3e\_val}  \ast \\  \textsf{l2\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l2\_entry} \; \textsf{l2e\_val}  \end{array} \right)  =\textsf{R}_{\textsf{walk} }
-    \\ \ast \left(\begin{array}{l}  \textsf{l1\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l1\_entry}\;\textsf{l1e\_val} \ast \\ \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l1e\_val})\urcorner \end{array}\right) =\textsf{R}_{\textsf{l1e}}\end{array} $
-\todo[inline]{Colin: This is confusing: in the code \lstinline|l2_entry| is an actual entry word, but in these invariants they seem to be paddrs.
-Also, these should be 4 different fractions, not all qfrac
+$\begin{array}{l}\left( \begin{array}{l}
+  \ulcorner
+    \textsf{l3}+\textsf{L3Offset(va)} = \textsf{l3\_entry}
+    \land \textsf{l2}+\textsf{L2Offset(va)} = \textsf{l2\_entry}
+    \land \textsf{l1}+\textsf{L1Offset(va)} = \textsf{l1\_entry} \\
+    \land \textsf{table\_root(l4e\_val}_{\textsf{pfn}})  = \textsf{l3} 
+    \land \textsf{table\_root(l3e\_val}_{\textsf{pfn}})  = \textsf{l2} 
+    \land \textsf{table\_root(l2e\_val}_{\textsf{pfn}})  = \textsf{l1} \urcorner  \ast \\
+    \textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{l4p},\textsf{l3p},\textsf{l2p},\textsf{l1p}\},m)  \ast  
+      \ghostmaptoken{\textsf{id}}{(\mathsf{l4p})}{\textsf{4}}\ast \ghostmaptoken{\textsf{id}}{(\mathsf{l3p})}{\textsf{3}} \ast \ghostmaptoken{\textsf{id}}{(\mathsf{l2p})}{\textsf{2}}  \ast \\
+      \ghostmaptoken{\delta{}s}{\rtv}{\delta} \ast \textsf{l4p+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l4p} \; \textsf{l4e\_val}  \ast \\
+      \textsf{l3p+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l3p}\;\textsf{l3e\_val}  \ast   \textsf{l2p+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l2p} \; \textsf{l2e\_val}  \end{array} \right)  =\textsf{R}_{\textsf{walk} }
+    \\ \left(\begin{array}{l}  \textsf{l1p+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l1p}\;\textsf{l1e\_val} \ast  \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l1e\_val})\urcorner \end{array}\right) =\textsf{R}_{\textsf{l1e}}\end{array} $
+\todo[inline]{Colin: These should be 4 different fractions, not all qfrac, right?
+And as a presentation question, should the vaddrs in the vpte pointstos use the physical plus KERNBASE or the l$n$\_entry vaddrs?
 }
@@
 
+In Figures \ref{walkpgdirC} and \ref{fig:rwalkC}, there are four related concepts for each level.
+\textsf{l4p} is the physical address of the L4 table base; \textsf{l4} is the corresponding virtual address
+(using the same name for the value and the program variable name for brevity, since the variable is not reassigned);
+\textsf{l4\_entry} is the virtual address of the L4 entry used to translate \textsf{va};
+and \textsf{l4e\_val} is the value of that table entry.
+Other levels are named consistently.
 For each of the three level transitions, the main challenges for the proof are to
@@
 (the assertion on Line \ref{line:walkpgdir_pre} has one for each entry in the table)
+\todo[inline]{text around this note depends on precondition to walkpgdir}
 and pulling the resources for physical address $\mathsf{rtv}+8*\textsf{L4Offset}(\textsf{va})$
@@
 void vaspace_mappage(pte_t *l4, void *va,uintptr_t fpaddr ) {
-  $\specline{\textsf{P} \ast \textsf{l4} \mapsto_{\textsf{id}} \textsf{\_} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta} \ast \ulcorner\theta \; !!\; \vaddr = \texttt{None}\urcorner}_{\rtv}$ (*\label{line:mappage_pre}*)
+  $\specline{\textsf{P} \ast \ghostmaptoken{\mathsf{\textsf{id}}}{\textsf{l4}}{4} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta} \ast \ulcorner\theta \; !!\; \vaddr = \texttt{None}\urcorner}_{\rtv}$ (*\label{line:mappage_pre}*)
   pte_t *pteaddr = walkpgdir(l4, va); (*\label{line:call_walkpgdirC}*)
@@
 }
-\todo[inline,color=cyan]{
-  Ismail, shouldn't the precondition include the 512 tokens for the L4 table, and shouldn't the l4 entry be excluded from $\Xi$?
+\todo[inline,color=cyan]{This is where Colin is stuck. I don't see how we can derive vptes for any of the 512
+L4 entries from just a single identity map token for the base of the table (i.e., entry 0).
+Surely we must need ${\bigast_{i\in 0..511}}{\ghostmaptoken{\textsf{id}}{(l4p+i*8)}{4}}$, we should discuss live.
 }
-The precondition directly implies the precondition of \lstinline|walkpgdir|, which as just discussed
-returns the virtual address of an allocated L1 entry, and whose postcondition contains almost all of the information
+From the precondition on Line \ref{line:mappage_pre}, we can derive
+the precondition of \lstinline|walkpgdir|:\\
+\centerline{$\ghostmaptoken{\mathsf{\textsf{id}}}{\textsf{l4p}}{4} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)
+\vdash
+\ghostmaptoken{\mathsf{\textsf{id}}}{\textsf{l4p}}{4} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus l4p,m)
+\ast \textsf{l4p+KERNBASE}\mapsto_{\textsf{vpte,q}}\textsf{l4p}\;
+$}\\
+\todo[inline,color=cyan]{end confusion}
+\lstinline|walkpgdir|, as just discussed,
+returns the virtual address of an allocated L1 entry and its postcondition contains almost all of the information
 needed to construct a virtual points-to for \lstinline|va| --- except information about the L1 entry

### 2025-07-25 examples_C.tex
finish pass on section 5

@@
 pte_t *walkpgdir(pte_t *l4, const void *va) {
-  $\specline{\textsf{P} \ast \ulcorner\textsf{l4p}=\textsf{rtv}\urcorner\ast\ulcorner\textsf{rtv+KERNBASE}=\textsf{l4}\urcorner \ast \forall_{i\in\textsf{0..511}}\ghostmaptoken{\textsf{id}}{(\textsf{rtv}+i*8)}{\textsf{4}} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}}_{\rtv}$(*\label{line:walkpgdir_pre}*)
+  $\specline{\textsf{P} \ast \ulcorner\textsf{l4p}=\textsf{rtv}\urcorner\ast\ulcorner\textsf{rtv+KERNBASE}=\textsf{l4}\urcorner \ast \ghostmaptoken{\textsf{id}}{(\textsf{l4p}+8*\textsf{L4Offset}(\textsf{va}))}{\textsf{4}} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}}_{\rtv}$(*\label{line:walkpgdir_pre}*)
   pte_t *l4_entry = &l4[L4Offset(va)]; // Virtual address of L4 entry(*\label{line:start_l4_calcC}*)
-  $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{l4p}\},m) \ast \forall_{i\in\textsf{0..511}\setminus\{\textsf{L4Offset(va)}\}}\ghostmaptoken{\textsf{id}}{(\textsf{rtv}+i*8)}{\textsf{4}}\ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
-  $\specline{\textsf{entry\_present(l4e\_val)} \wand \forall_{i\in\textsf{0..511}} \; \ghostmaptoken{\textsf{id}}{\textsf{table\_root(l4e\_val)+ i * 8}}{\textsf{3}} }_{\rtv} $
-  $\specline{ \textsf{l4p+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l4p l4e\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l4e\_val})\urcorner}_{\rtv}$
+  $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{l4p+8*L4Offset(va)}\},m) \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
+  $\specline{\textsf{entry\_present(l4e\_val)} \wand \forall_{i\in\textsf{0..511}} \; \ghostmaptoken{\textsf{id}}{(\textsf{table\_root(l4e\_val)+ i * 8})}{\textsf{3}} }_{\rtv} $
+  $\specline{ \textsf{l4p+KERNBASE+L4Offset(va)} \mapsto_{\textsf{vpte,qfrac4}} \textsf{(l4p+L4Offset(va)) l4e\_val} \ast \ulcorner  \textsf{qfrac4} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l4e\_val})\urcorner}_{\rtv}$
   pte_t *l3 = pte_get_next_table(l4_entry); (*\label{line:first_getnext_callC}*) 
   /*pte_get_next_table may have allocated a new page which updates entry value*/
-  $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{l4p}\},m)  \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
-  $\specline{\forall_{i\in\textsf{0..511}} \ghostmaptoken{\textsf{id}}{\textsf{table\_root(l4e\_val'.pfn) + i * 8}}{\textsf{3}}  \ast \ulcorner\textsf{l3-KERNBASE}=\textsf{table\_root(l4e\_val'.pfn)}\urcorner} $(*\label{line:l3tokens}*)
-  $\specline{ \textsf{l4p+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l4p l4e\_val'} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l4e\_val})\urcorner}_{\rtv}$(*\label{line:first_pte_pointstoC}*)
+  $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{l4p+8*L4Offset(va)}\},m)  \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
+  $\specline{\forall_{i\in\textsf{0..511}} \ghostmaptoken{\textsf{id}}{(\textsf{table\_root(l4e\_val'.pfn) + i * 8})}{\textsf{3}}  \ast \ulcorner\textsf{l3-KERNBASE}=\textsf{table\_root(l4e\_val'.pfn)}\urcorner} $(*\label{line:l3tokens}*)
+  $\specline{ \textsf{l4p+KERNBASE+L4Offset(va)} \mapsto_{\textsf{vpte,qfrac4}} \textsf{(l4p+L4Offset(va)) l4e\_val} \ast \ulcorner  \textsf{qfrac4} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l4e\_val})\urcorner}_{\rtv}$(*\label{line:first_pte_pointstoC}*)
   pte_t *l3_entry = &l3[L3Offset(va)]; /* Virtual address of L3 entry*/(*\label{line:l3offset}*)
   $\specline{\ulcorner\textsf{l3}+\textsf{L3Offset(va)*8} = \textsf{l3p} \land \textsf{table\_root(l4e\_val'.pfn)}  = \textsf{l3} \urcorner }_{\rtv}$
-  $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{l4p},\textsf{l3p}\},m)  \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
+  $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{l4p+8*L4Offset(va)},\textsf{l3p+8*L3Offset(va)}\},m)  \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
   $\specline{\forall_{i\in\textsf{0..511}\setminus\{\textsf{L3Offset(va)}\}} \; \ghostmaptoken{\textsf{id}}{(\textsf{table\_root(l4e\_val.pfn}) \textsf{ + i * 8})} {\textsf{3}} }_{\rtv}$
-  $\specline{ \textsf{l4p+KERNBASE} \mapsto_{\textsf{vpte,qfrac4}}\textsf{l4p}\; \textsf{l4e\_val'} \ast \ulcorner  \textsf{qfrac4} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l4e\_val'})\urcorner}_{\rtv}$(*\label{line:ex_l4_vpteC}*)
-  $\specline{ \textsf{l3p+KERNBASE} \mapsto_{\textsf{vpte,qfrac3}}\textsf{l3p}\; \textsf{l3e\_val} \ast \ulcorner  \textsf{qfrac3} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l3e\_val})\urcorner}_{\rtv}$(*\label{line:ex_l3_vpteC}*)
-  $\specline{\textsf{entry\_present(l3e\_val)} \wand \forall_{i\in\textsf{0..511}} \;\ghostmaptoken{\textsf{id}}{\textsf{table\_root(l3e\_val.pfn) + i * 8}}{ \textsf{2}} } _{\rtv} $
+  $\specline{ \textsf{l4p+KERNBASE+L4Offset(va)} \mapsto_{\textsf{vpte,qfrac4}} \textsf{(l4p+L4Offset(va)) l4e\_val} \ast \ulcorner  \textsf{qfrac4} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l4e\_val})\urcorner}_{\rtv}$(*\label{line:ex_l4_vpteC}*)
+  $\specline{ \textsf{l3p+KERNBASE+L3Offset(va)} \mapsto_{\textsf{vpte,qfrac3}} \textsf{(l3p+L3Offset(va)) l3e\_val} \ast \ulcorner  \textsf{qfrac3} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l3e\_val})\urcorner}_{\rtv}$(*\label{line:ex_l3_vpteC}*)
+  $\specline{\textsf{entry\_present(l3e\_val)} \wand \forall_{i\in\textsf{0..511}} \;\ghostmaptoken{\textsf{id}}{(\textsf{table\_root(l3e\_val.pfn) + i * 8})}{ \textsf{2}} } _{\rtv} $
   pte_t *l2 = pte_get_next_table(l3_entry);(*\label{line:getnextl3}*)
@@
   \ulcorner
-    \textsf{l3}+\textsf{L3Offset(va)} = \textsf{l3\_entry}
-    \land \textsf{l2}+\textsf{L2Offset(va)} = \textsf{l2\_entry}
-    \land \textsf{l1}+\textsf{L1Offset(va)} = \textsf{l1\_entry} \\
+    \textsf{l3}+8*\textsf{L3Offset(va)} = \textsf{l3\_entry}
+    \land \textsf{l2}+8*\textsf{L2Offset(va)} = \textsf{l2\_entry}
+    \land \textsf{l1}+8*\textsf{L1Offset(va)} = \textsf{l1\_entry} \\
     \land \textsf{table\_root(l4e\_val}_{\textsf{pfn}})  = \textsf{l3} 
@@
     \land \textsf{table\_root(l2e\_val}_{\textsf{pfn}})  = \textsf{l1} \urcorner  \ast \\
-    \textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{l4p},\textsf{l3p},\textsf{l2p},\textsf{l1p}\},m)  \ast  
-      \ghostmaptoken{\textsf{id}}{(\mathsf{l4p})}{\textsf{4}}\ast \ghostmaptoken{\textsf{id}}{(\mathsf{l3p})}{\textsf{3}} \ast \ghostmaptoken{\textsf{id}}{(\mathsf{l2p})}{\textsf{2}}  \ast \\
-      \ghostmaptoken{\delta{}s}{\rtv}{\delta} \ast \textsf{l4p+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l4p} \; \textsf{l4e\_val}  \ast \\
-      \textsf{l3p+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l3p}\;\textsf{l3e\_val}  \ast   \textsf{l2p+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l2p} \; \textsf{l2e\_val}  \end{array} \right)  =\textsf{R}_{\textsf{walk} }
-    \\ \left(\begin{array}{l}  \textsf{l1p+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l1p}\;\textsf{l1e\_val} \ast  \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l1e\_val})\urcorner \end{array}\right) =\textsf{R}_{\textsf{l1e}}\end{array} $
-\todo[inline]{Colin: These should be 4 different fractions, not all qfrac, right?
-And as a presentation question, should the vaddrs in the vpte pointstos use the physical plus KERNBASE or the l$n$\_entry vaddrs?
-}
+    % \textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{l4p},\textsf{l3p},\textsf{l2p},\textsf{l1p}\},m)  \ast  
+    \textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{l4p+8*L4Offset(va)},\textsf{l3p+8*L3Offset(va)},\textsf{l2p+8*L2Offset(va)},\textsf{l1p+8*L1Offset(va)}\},m)  \ast  \\
+      \ghostmaptoken{\textsf{id}}{(\mathsf{l4p+8*L4Offset(va)})}{\textsf{4}}\ast \ghostmaptoken{\textsf{id}}{(\mathsf{l3p+8*L3Offset(va)})}{\textsf{3}} \ast \ghostmaptoken{\textsf{id}}{(\mathsf{l2p+8*L2Offset(va)})}{\textsf{2}}  \ast \\
+      \ghostmaptoken{\delta{}s}{\rtv}{\delta} \ast (\textsf{l4p+KERNBASE+8*L4Offset(va)}) \mapsto_{\textsf{vpte,qfrac4}} (\textsf{l4p+8*L4Offset(va)}) \; \textsf{l4e\_val}  \ast \\
+      (\textsf{l3p+KERNBASE+8*L3Offset(va)}) \mapsto_{\textsf{vpte,qfrac3}} \textsf{l3p+8*L3Offset(va)}\;\textsf{l3e\_val}  \ast\\
+      (\textsf{l2p+KERNBASE+8*L2Offset(va)}) \mapsto_{\textsf{vpte,qfrac2}} \textsf{l2p+8*L2Offset(va)} \; \textsf{l2e\_val}
+    \end{array} \right)  =\textsf{R}_{\textsf{walk} } 
+    \\ \left(\begin{array}{l}  (\textsf{l1p+KERNBASE*8*L1Offset(va)}) \mapsto_{\textsf{vpte,qfrac1}} \textsf{l1p+8*L1Offset(va)}\;\textsf{l1e\_val} \ast  \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l1e\_val})\urcorner \end{array}\right) =\textsf{R}_{\textsf{l1e}}\end{array} $
+\todo[inline]{Ismail: Why isn't there an l1 entry identity map token in Rwalk?}
+\vspace{-1em}
 \caption{\textsf{R}$_{\textsf{walk}}$: Resources obtained from invariant for walking page-table directory with \emph{non-existing entries} via calls to \textsf{l1e\_get\_next\_table} in Figure \ref{fig:calltopteinitializeC}}
@@
 identity map tokens for the the physical address of each entry in the subsequent level's table.
-For traversing the L4 table, this proceeds by extracting the relevant identity map token
-(the assertion on Line \ref{line:walkpgdir_pre} has one for each entry in the table)
-\todo[inline]{text around this note depends on precondition to walkpgdir}
-and pulling the resources for physical address $\mathsf{rtv}+8*\textsf{L4Offset}(\textsf{va})$
+For traversing the L4 table, this proceeds by
+exchanging the relevant identity map token provided in the precondition (Line \ref{line:walkpgdir_pre})
+and pulling the resources for physical address $\mathsf{l4p}+8*\textsf{L4Offset}(\textsf{va})$
 out of the identity map invariant: 
@@
 void vaspace_mappage(pte_t *l4, void *va,uintptr_t fpaddr ) {
-  $\specline{\textsf{P} \ast \ghostmaptoken{\mathsf{\textsf{id}}}{\textsf{l4}}{4} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta} \ast \ulcorner\theta \; !!\; \vaddr = \texttt{None}\urcorner}_{\rtv}$ (*\label{line:mappage_pre}*)
+  $\specline{\textsf{P} \ast \ghostmaptoken{\mathsf{\textsf{id}}}{(\textsf{l4p+8*L4Offset(va)})}{4} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta} \ast \ulcorner\theta \; !!\; \vaddr = \texttt{None}\urcorner}_{\rtv}$ (*\label{line:mappage_pre}*)
   pte_t *pteaddr = walkpgdir(l4, va); (*\label{line:call_walkpgdirC}*)
@@
 \vspace{-1em}
+\todo[inline,color=cyan]{will switch to l4p+ l4offset of va, so caller has the 512}
   \caption{Specification of updating L1 entry to reference a new page (\textsf{fpaddr}).}
@@
 }
-\todo[inline,color=cyan]{This is where Colin is stuck. I don't see how we can derive vptes for any of the 512
-L4 entries from just a single identity map token for the base of the table (i.e., entry 0).
-Surely we must need ${\bigast_{i\in 0..511}}{\ghostmaptoken{\textsf{id}}{(l4p+i*8)}{4}}$, we should discuss live.
+The precondition on Line \ref{line:mappage_pre}, directly entails
+the precondition of the \lstinline|walkpgdir| call.\footnote{The proof of \lstinline|vaspace_mappage|'s caller
+would extract this single identity map token for the specific L4 entry from a set of 512 that are
+part of the kernel invariant, as \lstinline|walkpgdir|'s proof does for lower levels.
 }
-From the precondition on Line \ref{line:mappage_pre}, we can derive
-the precondition of \lstinline|walkpgdir|:\\
-\centerline{$\ghostmaptoken{\mathsf{\textsf{id}}}{\textsf{l4p}}{4} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)
-\vdash
-\ghostmaptoken{\mathsf{\textsf{id}}}{\textsf{l4p}}{4} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus l4p,m)
-\ast \textsf{l4p+KERNBASE}\mapsto_{\textsf{vpte,q}}\textsf{l4p}\;
-$}\\
-\todo[inline,color=cyan]{end confusion}
 \lstinline|walkpgdir|, as just discussed,
@@
 
-\todo[inline]{Colin will fix up this paragraph: it's right but also kind of weird, will handle during related work pass}
+\todo[inline]{Colin will fix up next paragraph: it's right but also kinda weird, will handle w/ related work pass}
 Unlike the only prior work verifying analogous code for mapping a new page~\cite{kolanski08vstte,kolanski09tphols}, our proof above

### 2025-07-25 examples_C.tex
l1 id mapsto

@@
     \textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{l4p+8*L4Offset(va)},\textsf{l3p+8*L3Offset(va)},\textsf{l2p+8*L2Offset(va)},\textsf{l1p+8*L1Offset(va)}\},m)  \ast  \\
-      \ghostmaptoken{\textsf{id}}{(\mathsf{l4p+8*L4Offset(va)})}{\textsf{4}}\ast \ghostmaptoken{\textsf{id}}{(\mathsf{l3p+8*L3Offset(va)})}{\textsf{3}} \ast \ghostmaptoken{\textsf{id}}{(\mathsf{l2p+8*L2Offset(va)})}{\textsf{2}}  \ast \\
+      \ghostmaptoken{\textsf{id}}{(\mathsf{l4p+8*L4Offset(va)})}{\textsf{4}}\ast \ghostmaptoken{\textsf{id}}{(\mathsf{l3p+8*L3Offset(va)})}{\textsf{3}} \ast \\ \ghostmaptoken{\textsf{id}}{(\mathsf{l2p+8*L2Offset(va)})}{\textsf{2}}  \ast \ghostmaptoken{\textsf{id}}{(\mathsf{l1p+8*L1Offset(va)})}{\textsf{1}} \ast \\
       \ghostmaptoken{\delta{}s}{\rtv}{\delta} \ast (\textsf{l4p+KERNBASE+8*L4Offset(va)}) \mapsto_{\textsf{vpte,qfrac4}} (\textsf{l4p+8*L4Offset(va)}) \; \textsf{l4e\_val}  \ast \\

### 2025-07-26 programlogic.tex
space tweak
work on related work, justifying modalities. not done with either, but progress.

@@
 
-We describe a program logic (a separation logic) along the lines suggested earlier, where every assertion is relative
+\add{
+  A program logic for reasoning about code which may work with (and possibly update) multiple address spaces requires dealing with
+  several key challenges. It must ensure that reasoning about memory accesses only depends on assertions that hold
+  in the active address space at the time of access. It must allow invariants for code or data structures to refer
+  to other address spaces. These constraints mean it must also support reasoning about when the active address space
+  changes, as this affects which memory assumptions are usable or not for memory access and thus which data structures are
+  immediately directly accessible.
+}
+\add{
+  Many approaches which could handle these problems in principle, such as tagging pointers with the relevant address space.
+  But such approaches introduce other complexities.
+  Most code, even in an operating system kernel, only works with a single address space (the current one).
+  Explicitly plumbing address space information through a simple linked list specification, simply because some other part of the program
+  may manipulate other address spaces, adds significant specification burden.
+  Specifications which need to talk about a particular invariant holding in an arbitrary other address space,
+  or with knowing a certain assertion conceptually holds in multiple specific address spaces, would need to quantify
+  over functions from address space identity to assertions rather than just assertions.
+  Using a modal approach resolves all of these technical and usability\todo{need better word, I'm always wary of using this word in PL}
+  challenges cleanly and uniformly.
+  Specifications that are not \emph{about} address space manipulation need not mention an address spaces at all:
+  one can use one's favorite separation logic data structure specifications without restructuring,
+  and every assertion can be stated for either the current or specific other address spaces.
+  In short, modalities make it possible for specifications to mention address spaces only when they deal with address spaces,
+  and not when they don't.\todo{wary of this due to the rtv in the double}
+}
+
+We describe a program logic (a separation logic) along the lines suggested \replace{earlier}{above}, where every assertion is relative
 to an address space in which it is interpreted, allowing us to define \emph{virtual points-to} assertions that make claims
@@
 %and, expectedly \textsf{w2} is an address computed by page-offset computation (e.g. $\textsf{l4off maddr cr3val}$). 
-From this we can derive a more concise physical points-to when the split is unimportant:\\
+From this we can derive a more concise physical points-to when the split is unimportant:
 % Giving a raw 64-bits memory pointsto assertion becomes
-\centerline{$\textsf{w} \mapsto_{\textsf{p}} \{q\} \textsf{ v} \stackrel{\triangle}{=} (\textsf{drop 12}~w) \ \sim \ (\textsf{bottom 12}~w)\mapsto_{\textsf{a}} \; \{\textsf{q}\} \textsf{ v} $}
+{$\textsf{w} \mapsto_{\textsf{p}} \{q\} \textsf{ v} \stackrel{\triangle}{=} (\textsf{drop 12}~w) \ \sim \ (\textsf{bottom 12}~w)\mapsto_{\textsf{a}} \; \{\textsf{q}\} \textsf{ v} $}
 
@@
 \end{lstlisting}
+\todo[inline,color=cyan]{Ismail, we can drop the maddr from the l4M52 etc, and the l4e etc. from the l3off etc., right?}
 \vspace{-1em}
 \caption{A Strong Virtual Points-to Relation
-  \todo[inline]{Ismail: Tie these lines to the steps and entries in Figure \ref{fig:pagetables}, e.g., with \textcircled{1}\ldots}
-  \todo[inline]{Ismail Addressed: earlier in the background and the current text is clearly related to \ref{fig:pagetables}}
 }
@@
 \footnotesize
+\vspace{-1em}
   \begin{mathpar}
@@
   \end{mathpar}
+\vspace{-2em}
   \caption{\iris rules for ghost maps}
@@
 \end{mathpar}
-\todo[inline]{Ismail, what do you think about shading the ``core'' assertions that are common to any asm program logic?
-Or shading the VMM-specific ones?
-}
-\todo[inline]{Ismail Addressed: I don't think that we should do -- general instructions' text outlines the how to read VMM related ones in the text. 
-}
 \vspace{-1em}
@@
 
-\todo[inline,color=green]{Colin needs to add more of the ``why modals'' stuff here, per the response.
-But it's a little awkward: in the response we said this was hard to specify with ghost-tagged
-pointers, but now I'm wondering if that's \emph{quite} right.
-It seems like maybe you could just plumb an implicitly-tagged P through to the following
-instructions and be okay, the instantiation of P would just be horrid.
-What we said about e.g., defining
-a generic linked list is still true, and should be raised earlier (not here).
-Here, we want to make a point about packaging all resources for a specific address space
-together. But it would be stronger if we could hone the argument in the response more.
-The bit about ``how would we talk abstractly about the fact that some unspecified resources R should
-be sufficient ... in the new address space'' is the right bit to polish...
-ghost bit discussion should either go in relwork with some Dafny(?) discussion, or earlier in motivation.
-}
-
 \subsection{Soundness}

### 2025-07-26 relwork.tex
a few more edits, but out of steam for the night
work on related work, justifying modalities. not done with either, but progress.

--- 2025-07-20 relwork.tex
@@
 \label{sec:relwork}
+\todo[inline]{Colin must incorporate stuff from response}
 
@@
 relied on custom hardware, which, among other things, always ran kernel code with virtual memory disabled, removing the circularity that is a key challenge of verifying
-VMM code for real hardware: At that point page tables become a basic partial map data structure to represent user program address translations,
-with an idiosyncratic format. It turns out that subsequent OS verification work also treats page tables this way, but
-unsoundly given that other projects target hardware that \emph{does} run the kernel with address translation: they \emph{trust}
-that the particular page table manipulations do not, for example, unmap kernel code (which can crash the machine even if done ``temporarily'').
+VMM code for real hardware: \replace{A}{a}t that point page tables become a basic partial map data structure to represent user program address translations,
+with an idiosyncratic format. \replace{It turns out that s}{S}ubsequent OS verification work also treats page tables this way, but
+unsoundly given that \replace{other projects}{their} target hardware \del{that} \emph{does} run the kernel with address translation: they \emph{trust}
+that the particular page table manipulations do not, for example, unmap kernel code\footnote{\add{A ``fun'' bug one author encountered in
+an industrial prototype kernel.}}
+ (which can crash the machine even if done ``temporarily'').
 % \footnote{A bug
@@
 with the attendant problems discussed earlier and lifted by our model. Their approach had modal elements,
-but the work never addressed the issues of managing multiple address spaces, switching address spaces, or
-invariants referring to other address spaces, which our explicitly modal approach and case studies address.
-They verified a fraction of an ARM version of our Figure \ref{fig:mapping_code}, which assumed that they already had the virtual address
-of the L1 entry (which they assumed existed), thereby skipping the physical-to-virtual conversion issues we treat in Section \ref{sec:p2v}.
+but
+\replace{the work never addressed the issues of managing multiple address spaces, switching address spaces, or
+invariants referring to other address spaces, which our explicitly modal approach and case studies do address.
+They verified a fraction of an ARM version of our Figure ??, which assumed that they already had the virtual address
+of the L1 entry (which they assumed existed), thereby skipping the physical-to-virtual conversion issues we treat in Section ??.
+}{
+ did not tackle evaluations that would benefit from modality.
+ We have verified significantly more VMM code than Kolanski and Klein, which pushed us towards the modal approach.
+ They verify~\cite{kolanski08vstte} roughly the (ARM) assembly corresponding to Lines \ref{line:l1entry_storeC} and \ref{line:l1entry_setpresent} of Figure
+ \ref{fig:mapping_codeC},
+ plus pulling the zeroed page to install in the L1 entry from a simple linked list.
+ They never addressed code equivalent to our \lstinline|walkpgdir| (Figure \ref{walkpgdirC}), or \lstinline|pte_get_next_table| (Figure \ref{fig:calltopteinitializeC}),
+ which are the most challenging parts of verifying a full page table walk, encompasing subtle sharing and ownership
+ specifications and delicate kernel invariants involved in converting physical to virtual addresses. Ours is the only work to date
+ to tackle verification of the \emph{entirety} of a software page table walk and mapping a new page.
+}
 Due to a lack of fractional permissions in their formalism, they incidentally pick up other limitations orthogonal to their
@@
 a virtual points-to have \emph{full} ownership of the page table walk memory, they limit their system to having only as many
-virtual points-to assertions as there are entries in the top-level table (4096) because they cannot share access to entries.
+virtual points-to assertions as there are entries in the top-level table (\replace{4096}{512}) because they cannot share access to entries.
 We inherit fractional permission support form \iris, and use it extensively (the overly-restrictive
 Figure \ref{fig:strongvirtualpointsto} is already an improvement in this way).
+\add{While we cannot claim credit for \textsc{Iris}'s extensive feature set, the fact that the model of our
+assertions is based on a classic algebraic tool (pointwise lifting) makes our approach compatible with
+other logical bases as well, such as \textsc{PulseCore}~\cite{ebner2025pulsecore} or \textsc{VST}~\cite{appel2014program}.
+}
 
-% Some work has been done like ours, studying VMM verification separately from the rest of the kernel.
-% One important outgrowth of the \textsc{seL4} project, not integrated into the main project's proof, was work by 
-% Kolanski and Klein which studied verification of code against a hardware model that \emph{did} include address translation
-%  --- the only work aside from ours to do so --- initially in terms of basic memory~\cite{kolanski08vstte} and subsequently 
-% integrating source-level types into the interpretation~\cite{kolanski09tphols}. 
-% They were the first work to model physical and virtual points-to assertions separately, defining virtual points-to assertions
-% in terms of physical points-to assertions mimicking page table walks, and defining all of their assertions as predicates on a
-% pair of (physical) machine memory and a page table root, an approach we improve on.
-
-% They also define their virtual points-to assertions such that a virtual points-to $p\mapsto_\mathsf{v} a$ owns the full 
-% lookup path to virtual address $p$. This means that given two virtual points-to assertions at the same time, such as 
-% $p\mapsto_\mathsf{v}a \ast p'\mapsto_\mathsf{v}b$, the memory locations traversed to translate $p$ and $p'$ must be disjoint. 
-% This means the logic has a peculiar limit on how many virtual points-to assertions can coexist in a proof. Since page tables 
-% fan out, the bottleneck is the number of entries in the root table. For their 32-bit ARMv6 example, the top-level address is 
-% still 4Kb (4096 bytes), and each entry (consumed entirely by a virtual points-to in their scheme) is 4 bytes, so they have a 
-% maximum of 1024 virtual points-tos in their ARMv6 configuration. Any assertion which implies more than that number
-% of virtual addresses are mapped implies false in their logic.
-% (They do formulate their logic over an abstract model, but every architecture would incur a similar limitation.)
-% % Na\"ively transferring their model to x86-64 4-level tables would yield a limit of 512 assertions (also a 4Kb root page, 
-% % but 8-byte entries).
-% Our definitions make use of fractional permissions throughout; Figure \ref{fig:strongvirtualpointsto}'s definition
-% of \lstinline|L4_L1_PointsTo| ellides the specific fractions used, but it in fact asserts 1/512 ownership of
-% the L1 entry, 1/($512^2$) of the L2 entry, and so on, so each entry may map the appropriate number of machine words.
-% \looseness=-1
-
-% As noted earlier, by collocating both the physical ownership of the page table walk
-% as part of the virtual points-to itself these logics preempt support for changes to page tables which do not actually affect 
-% address translation. We address this by moving from Figure \ref{fig:strongvirtualpointsto}'s definition (essentially Kolanski and Klein's
-% with fractional sharing of intermediate table entries) to Figure \ref{fig:virtualpointstosharing}'s.
-% \looseness=-1
-
-% Kolanski and Klein do verify code to map in a new page by installing an L1 page table entry, akin to our Figure \ref{fig:mapping_code}.
-% However, our logic treatment of mapping goes beyond theirs. 
-% Kolanski and Klein must unfold machine semantics at one point in their proof of mapping correctness~\cite[p.~27]{kolanski08vstte}, while
-% our logic permits us to conduct the proof entirely in separation logic.
-% Their proofs also do not address
-% converting physical addresses of page table entries to virtual addresses that can be used to access them.
-% Their original low-level (pseudo-assembly) proof assumes the correct virtual address already exists~\cite{kolanski08vstte},
-% and their subsequent C-level proof~\cite{kolanski09tphols} axiomatizes a function akin to our \lstinline|ensure_L1|.
-% While full proof of \lstinline|ensure_L1|'s page table walk is ongoing work, we have verified that virtual-to-physical
-% translation can be done in our logic with appropriate kernel invariants (Section \ref{sec:traversing}), while
-% Kolanski and Klein's model may not be able to do so --- because the virtual points-to owns the physical resources associated with
-% intermediate page table entries, it appears impossible to have a virtual points-to for a page table entry that is part of the
-% page table walk for another virtual address.
-% \looseness=-1
-
-% The other major distinction is that Kolanski and Klein have no explicit accounting for other address spaces.
-% Their logic does not deal with change of address space, and has no way to assert that certain facts hold
-% in another address space.
-% They verify only one address space manipulation: mapping a single unmapped page into the current address space (in both papers).
-% We verify this, as well as a change-of-address-space, which requires us to introduce assertions for talking
-% about other address spaces (we must know, for example, that the precondition of the code after the change must be true
-% in the \emph{other} address space), and to deal with the fact that the standard frame rule
-% for separation logic is unsound in the presence of address space changes and address-space-contingent assertions.
-% Our approach in this paper uses modalities to distinguish virtual-address-based assertions that hold only in specific 
-% address spaces, making it possible to manipulate other address spaces, and equally critically, to \emph{change} address 
-% spaces while reasoning about correctness.
-
-Unlike our work, Kolanski and Klein prove very useful embedding theorems stating that code that does not modify page table 
+\replace{Unlike our work, Kolanski and Klein prove very useful embedding theorems stating that code that does not modify page table 
 entries can be verified in a VM-ignorant program logic, and that proofs in that logic can be embedded into the VM-aware logic 
-(essentially by interpreting ``normal'' points-to relations as virtual points-to facts). While we have not proven such a result,
-an analagous result {should} hold of our work: consider that the doubles for the \texttt{mov} instructions
+(essentially by interpreting ``normal'' points-to relations as virtual points-to facts).}{
+ Kolanski and Klein prove that updates to any memory locations that are not part of the page tables
+ do not affect the interpretation of other memory addresses.
+ This implies that programs that do not modify memory mappings could be reasoned about without
+ concern for mappings.
+}
+While we have not \add{formally} proven such a result \add[of our machine model, it holds (just as for real hardware)],
+an analagous result \replace{{should} hold}{holds} of our work: consider that the doubles for the \texttt{mov} instructions
 that access memory behave just as one would expect for a VM-ignorant logic~\cite{Chlipala2013Bedrock}.
 With our general approach to virtual points-to assertions being inspired by Kolanski and Klein, \emph{both}
- our approach and theirs could in principle be extended to account for pageable points-to assertions by adding additional 
+ our approach and theirs could in principle be extended to account for pageable points-to assertions by adding \del{additional}
 disjunctions to an extended points-to definition; embedding ``regular'' separation logic into such a variant
@@
 
+\add{
+ No prior work has dealt with combined address space changes and context switching. \textsc{XCAP}~\cite{ni2007contexts} deals
+ only with usermode threading (in a single process). \textsc{CertiKOS} and \textsc{seL4} trust assembly primitives for
+ context switches, in addition to not modeling address translation for executing code.
+}
+\todo[inline]{author response talks about syeda and klein here briefly, and it fits, but we also want the longer discussion}
+
 \paragraph{Translation Lookaside Buffers}
 As noted in Section \ref{sec:backgroundonmachinemodel}, we do not formally model or reason about TLBs.
-TLB flushes are required only when addresses are removed from a virtual address space, or when changing
-virtual address spaces. Because this occurs in few places in the kernel (in some, only 3 locations), full verified
+\replace{TLB flushes are required only when addresses are removed from a virtual address space, or when changing
+virtual address spaces.}{
+ TLB flushes are necessary when a page is \emph{un}mapped, or when switching address spaces.
+}
+\replace{Because t}{T}his occurs in few places in \replace{the}{uniprocessor} kernel\add{s} (in some, only 3 locations), \add{so} full verified
 kernels including \textsc{seL4}~\cite{Klein2009seL4,seL4TOCS} and \textsc{CertiKOS}~\cite{gu15,gu2016certikos}
@@
 to verify TLB management in any form --- they \emph{must} trust its operation.
-\looseness=-1
-
+\replace{
 The only place TLB code becomes particularly challenging is in multiprocessor kernels, where
@@
 plus the memory-mapped IO used to trigger those IPIs.
-As a result, multicore extensions of \textsc{seL4}~\cite{von2013clustered} and \textsc{CertiKOS}~\cite{gu2016certikos} also trust this functionality.
-\looseness=-1
-
-% Fully grounded trust for this would require a formal model of how hardware populates ACPI tables and formal verification of ACPI
-% parsing code on general-purpose desktop and server machines, or a formalized and verified
-% parsing of flattened device trees (plus trust that a system was booted with a correct FDT) for (semi-)embedded systems like single-board
-% ARM and RISC-V machines. Either route would also require detailed treatment of memory-mapped IO triggering
-% interrupts on other CPUs. This is far beyond any formal model of computer hardware that exists today,
-% and far out of scope for this paper.
-
-% \citet{syeda2020formal,syeda2018program} are the only existing work to address formal verification of TLB management.
-% They extend the work of \citet{kolanski08ssv,kolanski09tphols,kolanski09tphols}, and therefore inherit the limitations of that work discussed
-% above. Their logic primarily tracks a set of mapped addresses and a set of
-% cached addresses, and generally preserve a global invariant (in a global Hoare logic, not a separation logic)
-% that the cached addresses are a subset of the mapped addresses. When this invariant is violated, the logic provides only
-% a Hoare-style backwards assignment type rule with unconstrained assertion $P$, rather than providing structure in the logic
-% for reasoning about how to restore the invariant. Most technical results in the paper focus on proving
-% transparency lemmas, that code that \emph{does not} modify page tables (kernel code outside the VMM, user code)
-% is unaffected by its existence. The one piece of kernel code verified is a sequence of 4 pseudo-instructions for changing address spaces
-% which include \emph{only} the installation of a new page table root and flushing of the old address space from the TLB.
-
-% Adapting Syeda and Klein's global reasoning principles to a separation logic is non-trivial, even though their approach
-% does capture some intuition about TLB reasoning. And as noted above, in the absence of support for significant additional
-% hardware functionality in the hardware model, this would be substantial work for only very limited gains in confidence compared
-% to trusting single-core TLB managment.
-% Thus we leave TLB management to our future work plans, where we believe it is best addressed in tandem with substantially
-% richer hardware models than any currently existing.
+As a result, multicore extensions of \textsc{seL4}~[?] and \textsc{CertiKOS}~[?] also trust this functionality.
+}{
+ This is true of multicore verified kernels as well~\cite{von2013clustered,gu2016certikos}, though there
+ the situation becomes much more challenging: when unmapping pages, the running CPU must invalidate the relevant TLB entries locally,
+ but also send an inter-procesor interrupt (IPI) to all other cores to ensure they also invalidate affected entries on their TLBs.
+ No formal hardware model currently exists for IPIs on any architecture, or even for the memory-mapped IO
+ used to trigger those IPIs.
+}
 
@@
 The specification they proved for those instructions did not address program invariants that may be valid in one address space and not the other,
-so is not flexible enough to extend directly to a full context switching primitive as in Figure \ref{fig:swtch},
-in addition to inheriting other limitations from the work they extend, discussed above~\cite{kolanski08vstte,kolanski09tphols}.
-\looseness=-1
-
-The right eventual solution would be to combine our work
-and theirs, which we leave to future work, as combining both the models and reasoning principles would be highly non-trivial.
+so is not flexible enough to extend directly to a full context switching primitive as in Figure \ref{fig:swtchC}\replace{,}{.}
+\del{in addition to inheriting other limitations from {the work they extend, discussed above}.}
+\add{Their logic adds an assertion tracking known-inconsistent addresses (i.e., recently-unmapped by a table update or change in page table root)
+and their memory access rules require accessed memory to lie outside that set.
+However their logic is proven sound against a generalized semantics where any page table change adds \emph{all}
+addresses to the inconsistent set, which is too imprecise for finer-grained unmapping.
+}
+The right \replace{eventual}{general} solution would be to combine our work
+and \add{an extension of} theirs, \add{a substantial project} which we leave to future work\replace{,}{.}\del{ as combining both the models and reasoning principles would be highly non-trivial.}
 

### 2025-07-26 examples_C.tex

@@
 \vspace{-1em}
-\caption{\textsf{R}$_{\textsf{walk}}$: Resources obtained from invariant for walking page-table directory with \emph{non-existing entries} via calls to \textsf{l1e\_get\_next\_table} in Figure \ref{fig:calltopteinitializeC}}
+\caption{\textsf{R}$_{\textsf{walk}}$: Resources obtained from invariant for walking page-table directory with \emph{non-existing entries} via calls to \textsf{pte\_get\_next\_table} in Figure \ref{fig:calltopteinitializeC}}
 \label{fig:rwalkC}

### 2025-07-27 intro.tex
lots of disorganized edits, mostly focused on justifying use of modalities, some comments for fixing in sec 4
intro tweaks

@@
 We develop these ideas in the form of a logic for working with virtual-address-space-relative assertions,
-implemented as an embedded separation logic within Iris~\cite{jung2018iris}.
+implemented as an embedded separation logic within \replace{Iris}{\textsc{Iris}}~\cite{jung2018iris}.
 The result is a separation logic that lifts a number of major semantic restrictions present in the few
@@
 switching address spaces, converting physical addresses to virtual addresses for
-software page table walks, and complete page table walk which backbones other essential functionalities of
-virtual-memory which we discuss for mapping and unmapping pages. These examples demonstrate the suitability and flexibility of
+software page table walks, and \add{a} complete \add{software} page table walk \replace{which backbones}{ --- the backbone of} other essential functionalities of
+virtual-memory which we discuss for mapping and unmapping pages. 
+These examples demonstrate the suitability and flexibility of
 a modal treatment of address spaces: each has a concise specification, each example either goes beyond the technical
-capabilities of prior logics, or revisits an example from prior work with more details (treating parts of the kernel
+capabilities of prior logics, or revisits an example from prior work
+\replace{with more details (treating parts of the kernel
 invariant never addressed by prior work) and fewer assumptions (e.g., verifying the reverse physical-to-virtual 
-address conversion rather than assuming it has already occurred).
+address conversion rather than assuming it has already occurred).}{
+  in greater detail (e.g., verifying critical and challenging supporting code that was
+  axiomatized or treated as out-of-scope in prior work).
+}
 

### 2025-07-27 background.tex
work on related work, justifying modalities. not done with either, but progress.

@@
 \setlength{\columnseprule}{4pt}
-\begin{figure}\footnotesize
+\begin{figure}\scriptsize
 %    \begin{framed}
 %\column{.59\textwidth}    
-\begin{subfigure}\footnotesize
+\begin{subfigure}\scriptsize
 
@@
 
-\begin{tikzpicture}[x=0.75pt,y=0.75pt,yscale=-0.75,xscale=0.75]
-%uncomment if require: \path (0,427); %set diagram left start at 0, and has height of 427
+\begin{tikzpicture}[x=0.75pt,y=0.75pt,yscale=-0.55,xscale=0.55]
+%uncomment if require: \path (0,403); %set diagram left start at 0, and has height of 403
 
 %Shape: Rectangle [id:dp9406932326447903] 
-\draw   (33,141) -- (123,141) -- (123,292) -- (33,292) -- cycle ;
+\draw   (33,133) -- (123,133) -- (123,284) -- (33,284) -- cycle ;
 %Shape: Rectangle [id:dp7682116051394419] 
-\draw   (166,141) -- (256,141) -- (256,292) -- (166,292) -- cycle ;
+\draw   (166,133) -- (256,133) -- (256,284) -- (166,284) -- cycle ;
 %Shape: Rectangle [id:dp07376733516930423] 
-\draw   (295,141) -- (385,141) -- (385,292) -- (295,292) -- cycle ;
+\draw   (295,133) -- (385,133) -- (385,284) -- (295,284) -- cycle ;
 %Shape: Rectangle [id:dp7106629147891306] 
-\draw   (427,141) -- (517,141) -- (517,292) -- (427,292) -- cycle ;
+\draw   (427,133) -- (517,133) -- (517,284) -- (427,284) -- cycle ;
 %Shape: Rectangle [id:dp1342906826203598] 
-\draw   (115,54) -- (185,54) -- (185,94) -- (115,94) -- cycle ;
+\draw   (115,46) -- (185,46) -- (185,86) -- (115,86) -- cycle ;
 %Shape: Rectangle [id:dp08072439007207399] 
-\draw   (185,54) -- (255,54) -- (255,94) -- (185,94) -- cycle ;
+\draw   (185,46) -- (255,46) -- (255,86) -- (185,86) -- cycle ;
 %Shape: Rectangle [id:dp8282933603833216] 
-\draw   (255,54) -- (325,54) -- (325,94) -- (255,94) -- cycle ;
+\draw   (255,46) -- (325,46) -- (325,86) -- (255,86) -- cycle ;
 %Shape: Rectangle [id:dp8416255090865326] 
-\draw   (325,54) -- (395,54) -- (395,94) -- (325,94) -- cycle ;
+\draw   (325,46) -- (395,46) -- (395,86) -- (325,86) -- cycle ;
 %Shape: Rectangle [id:dp22911958218529804] 
-\draw   (395,54) -- (465,54) -- (465,94) -- (395,94) -- cycle ;
+\draw   (395,46) -- (465,46) -- (465,86) -- (395,86) -- cycle ;
 %Shape: Rectangle [id:dp33732955213990756] 
-\draw   (465,54) -- (535,54) -- (535,94) -- (465,94) -- cycle ;
+\draw   (464.5,46) -- (555.5,46) -- (555.5,86) -- (464.5,86) -- cycle ;
 %Curve Lines [id:da6688730321284677] 
-\draw    (211.5,96) .. controls (-36.44,98.94) and (15.75,190.24) .. (30.17,213.65) ;
-\draw [shift={(31,215)}, rotate = 238.24] [color={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.75]    (10.93,-3.29) ..
+\draw    (211.5,88) .. controls (-36.44,90.94) and (15.75,182.24) .. (30.17,205.65) ;
+\draw [shift={(31,207)}, rotate = 238.24] [color={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.75]    (10.93,-3.29) ..
 controls (6.95,-1.4) and (3.31,-0.3) ..
@@
 %Shape: Rectangle [id:dp6044747978986409] 
-\draw   (33,191) -- (123,191) -- (123,231) -- (33,231) -- cycle ;
+\draw   (33,183) -- (123,183) -- (123,223) -- (33,223) -- cycle ;
 %Curve Lines [id:da8867562998134584] 
-\draw    (289.5,95) .. controls (115.15,110.76) and (137.77,175.03) .. (163.33,197.98) ;
-\draw [shift={(164.5,199)}, rotate = 220.24] [color={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.75]    (10.93,-3.29) ..
+\draw    (289.5,87) .. controls (115.15,102.76) and (137.77,167.03) .. (163.33,189.98) ;
+\draw [shift={(164.5,191)}, rotate = 220.24] [color={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.75]    (10.93,-3.29) ..
 controls (6.95,-1.4) and (3.31,-0.3) ..
@@
 %Shape: Rectangle [id:dp8861153167245819] 
-\draw   (166,177) -- (256,177) -- (256,217) -- (166,217) -- cycle ;
+\draw   (166,169) -- (256,169) -- (256,209) -- (166,209) -- cycle ;
 %Shape: Rectangle [id:dp05979844200848894] 
-\draw  [dash pattern={on 4.5pt off 4.5pt}] (295,156) -- (385,156) -- (385,196) -- (295,196) -- cycle ;
+\draw  [dash pattern={on 4.5pt off 4.5pt}] (295,148) -- (385,148) -- (385,188) -- (295,188) -- cycle ;
 %Shape: Rectangle [id:dp17403733244247843] 
-\draw   (427,212) -- (517,212) -- (517,252) -- (427,252) -- cycle ;
+\draw   (427,204) -- (517,204) -- (517,244) -- (427,244) -- cycle ;
 %Curve Lines [id:da20581124900573955] 
-\draw [color={rgb, 255:red, 0; green, 0; blue, 0 }  ,draw opacity=1 ] [dash pattern={on 4.5pt off 4.5pt}]  (357,97) .. controls (262.92,110.72) and (272.56,148.45) .. (291.81,173.48) ;
-\draw [shift={(293,175)}, rotate = 231.34] [color={rgb, 255:red, 0; green, 0; blue, 0 }  ,draw opacity=1 ][line width=0.75]    (10.93,-3.29) ..
+\draw [color={rgb, 255:red, 0; green, 0; blue, 0 }  ,draw opacity=1 ] [dash pattern={on 4.5pt off 4.5pt}]  (357,89) .. controls (262.92,102.72) and (272.56,140.45) .. (291.81,165.48) ;
+\draw [shift={(293,167)}, rotate = 231.34] [color={rgb, 255:red, 0; green, 0; blue, 0 }  ,draw opacity=1 ][line width=0.75]    (10.93,-3.29) ..
 controls (6.95,-1.4) and (3.31,-0.3) ..
@@
 %Curve Lines [id:da07589611603635316] 
-\draw    (425.5,95) .. controls (380.42,119.5) and (406.41,202.58) .. (422.53,224.72) ;
-\draw [shift={(423.5,226)}, rotate = 231.34] [color={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.75]    (10.93,-3.29) ..
+\draw    (425.5,87) .. controls (380.42,111.5) and (406.41,194.58) .. (422.53,216.72) ;
+\draw [shift={(423.5,218)}, rotate = 231.34] [color={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.75]    (10.93,-3.29) ..
 controls (6.95,-1.4) and (3.31,-0.3) ..
@@
 %Curve Lines [id:da6852167546673456] 
-\draw    (497.5,95) .. controls (564.48,91.06) and (565,230.71) .. (567.87,269.31) ;
-\draw [shift={(568,271)}, rotate = 265.24] [color={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.75]    (10.93,-3.29) ..
+\draw    (497.5,87) .. controls (564.48,83.06) and (567.43,193.61) .. (570.37,231.33) ;
+\draw [shift={(570.5,233)}, rotate = 265.24] [color={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.75]    (10.93,-3.29) ..
 controls (6.95,-1.4) and (3.31,-0.3) ..
@@
 %Curve Lines [id:da017521077423665377] 
-\draw    (39.5,359) .. controls (-12.71,369.84) and (17.07,323.43) .. (32.31,293.36) ;
-\draw [shift={(33,292)}, rotate = 116.57] [color={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.75]    (10.93,-3.29) ..
+\draw    (36.5,330) .. controls (-15.71,340.84) and (16.98,314.8) .. (32.31,285.35) ;
+\draw [shift={(33,284)}, rotate = 116.57] [color={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.75]    (10.93,-3.29) ..
 controls (6.95,-1.4) and (3.31,-0.3) ..
@@
 %Curve Lines [id:da6750042432035159] 
-\draw    (122,215) .. controls (128.9,254.4) and (131.91,269.54) .. (159.71,289.1) ;
-\draw [shift={(161,290)}, rotate = 214.59] [color={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.75]    (10.93,-3.29) ..
+\draw    (122,207) .. controls (128.9,246.4) and (131.91,261.54) .. (159.71,281.1) ;
+\draw [shift={(161,282)}, rotate = 214.59] [color={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.75]    (10.93,-3.29) ..
 controls (6.95,-1.4) and (3.31,-0.3) ..
@@
 %Curve Lines [id:da14788158682113428] 
-\draw    (257,197) .. controls (258.95,236) and (277.06,270.25) .. (293.72,290.47) ;
-\draw [shift={(295,292)}, rotate = 229.64] [color={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.75]    (10.93,-3.29) ..
+\draw    (257,189) .. controls (258.95,228) and (277.06,262.25) .. (293.72,282.47) ;
+\draw [shift={(295,284)}, rotate = 229.64] [color={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.75]    (10.93,-3.29) ..
 controls (6.95,-1.4) and (3.31,-0.3) ..
@@
 %Curve Lines [id:da3297977595767969] 
-\draw    (386,177) .. controls (392.86,226) and (408.36,270.2) .. (425.92,290.77) ;
-\draw [shift={(427,292)}, rotate = 228.01] [color={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.75]    (10.93,-3.29) ..
+\draw    (386,169) .. controls (392.86,218) and (408.36,262.2) .. (425.92,282.77) ;
+\draw [shift={(427,284)}, rotate = 228.01] [color={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.75]    (10.93,-3.29) ..
 controls (6.95,-1.4) and (3.31,-0.3) ..
@@
 %Curve Lines [id:da7710391961792686] 
-\draw    (517,229) .. controls (524.76,253.25) and (540.05,269.03) .. (552.82,282.73) ;
-\draw [shift={(554,284)}, rotate = 227.12] [color={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.75]    (10.93,-3.29) ..
+\draw    (517,221) .. controls (535.13,236.68) and (527.81,247.56) .. (555.75,240.46) ;
+\draw [shift={(557.5,240)}, rotate = 165.07] [color={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.75]    (10.93,-3.29) ..
 controls (6.95,-1.4) and (3.31,-0.3) ..
@@
 %Shape: Rectangle [id:dp1535611125971872] 
-\draw   (514,336) -- (584,336) -- (584,376) -- (514,376) -- cycle ;
+\draw   (511,313) -- (581,313) -- (581,353) -- (511,353) -- cycle ;
 %Shape: Rectangle [id:dp05427947476226991] 
-\draw   (95,338) -- (136.5,338) -- (136.5,378) -- (95,378) -- cycle ;
+\draw   (95,313) -- (136.5,313) -- (136.5,353) -- (95,353) -- cycle ;
 %Shape: Rectangle [id:dp10511138743752002] 
-\draw   (136.5,338) -- (265.5,338) -- (265.5,378) -- (136.5,378) -- cycle ;
+\draw   (136.5,313) -- (265.5,313) -- (265.5,353) -- (136.5,353) -- cycle ;
 %Shape: Rectangle [id:dp8917680919534263] 
-\draw   (265.5,338) -- (305,338) -- (305,378) -- (265.5,378) -- cycle ;
-\draw   (558,284.5) ..
-controls (558,279.25) and (562.25,275) ..
-(567.5,275) ..
-controls (572.75,275) and (577,279.25) ..
-(577,284.5) ..
-controls (577,289.75) and (572.75,294) ..
-(567.5,294) ..
-controls (562.25,294) and (558,289.75) ..
-(558,284.5) -- cycle ; \draw   (558,284.5) -- (577,284.5) ; \draw   (567.5,275) -- (567.5,294) ;
+\draw   (265.5,313) -- (305,313) -- (305,353) -- (265.5,353) -- cycle ;
+\draw   (560,243.5) ..
+controls (560,238.25) and (564.25,234) ..
+(569.5,234) ..
+controls (574.75,234) and (579,238.25) ..
+(579,243.5) ..
+controls (579,248.75) and (574.75,253) ..
+(569.5,253) ..
+controls (564.25,253) and (560,248.75) ..
+(560,243.5) -- cycle ; \draw   (560,243.5) -- (579,243.5) ; \draw   (569.5,234) -- (569.5,253) ;
 %Curve Lines [id:da9312068388962929] 
-\draw    (568,298) .. controls (584.66,309.76) and (586.91,318.64) .. (565.35,335.93) ;
-\draw [shift={(564,337)}, rotate = 321.95] [color={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.75]    (10.93,-3.29) ..
+\draw    (569,258) .. controls (585.66,269.76) and (587.91,278.64) .. (566.35,295.93) ;
+\draw [shift={(565,297)}, rotate = 321.95] [color={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.75]    (10.93,-3.29) ..
 controls (6.95,-1.4) and (3.31,-0.3) ..
@@
 % Text Node
-\draw (11,14) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\scriptsize Virtual}};
+\draw (21,20) node [anchor=north west][inner sep=0.75pt]  [font=\scriptsize] [align=left] {{\scriptsize Virtual}};
 % Text Node
-\draw (570,395) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\scriptsize Physical}};
+\draw (569,375) node [anchor=north west][inner sep=0.75pt]  [font=\scriptsize] [align=left] {{\scriptsize Physical}};
 % Text Node
-\draw (104,28) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\scriptsize 63}};
+\draw (102,17) node [anchor=north west][inner sep=0.75pt]  [font=\scriptsize] [align=left] {{\scriptsize 63}};
 % Text Node
-\draw (175,28) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\scriptsize 48}};
+\draw (173,17) node [anchor=north west][inner sep=0.75pt]  [font=\scriptsize] [align=left] {{\scriptsize 48}};
 % Text Node
-\draw (245,28) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\scriptsize 39}};
+\draw (243,17) node [anchor=north west][inner sep=0.75pt]  [font=\scriptsize] [align=left] {{\scriptsize 39}};
 % Text Node
-\draw (314,28) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\scriptsize 30}};
+\draw (312,17) node [anchor=north west][inner sep=0.75pt]  [font=\scriptsize] [align=left] {{\scriptsize 30}};
 % Text Node
-\draw (385,28) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\scriptsize 21}};
+\draw (383,17) node [anchor=north west][inner sep=0.75pt]  [font=\scriptsize] [align=left] {{\scriptsize 21}};
 % Text Node
-\draw (454,28) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\scriptsize 12}};
+\draw (452,17) node [anchor=north west][inner sep=0.75pt]  [font=\scriptsize] [align=left] {{\scriptsize 12}};
 % Text Node
-\draw (523,28) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\scriptsize 0}};
+\draw (547,16) node [anchor=north west][inner sep=0.75pt]  [font=\scriptsize] [align=left] {{\scriptsize 0}};
 % Text Node
-\draw (132,65) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\scriptsize Sign}};
+\draw (126,54) node [anchor=north west][inner sep=0.75pt]  [font=\scriptsize] [align=left] {{\scriptsize Sign}};
 % Text Node
-\draw (195,65) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\scriptsize L4 Offset}};
+\draw (188,54) node [anchor=north west][inner sep=0.75pt]  [font=\scriptsize] [align=left] {{\scriptsize L4 Offset}};
 % Text Node
-\draw (264,65) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\scriptsize L3 Offset}};
+\draw (257,54) node [anchor=north west][inner sep=0.75pt]  [font=\scriptsize] [align=left] {{\scriptsize L3 Offset}};
 % Text Node
-\draw (336,65) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\scriptsize L2 Offset}};
+\draw (328,54) node [anchor=north west][inner sep=0.75pt]  [font=\scriptsize] [align=left] {{\scriptsize L2 Offset}};
 % Text Node
-\draw (406,65) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\scriptsize L1 Offset}};
+\draw (398,55) node [anchor=north west][inner sep=0.75pt]  [font=\scriptsize] [align=left] {{\scriptsize L1 Offset}};
 % Text Node
-\draw (472,65) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\scriptsize Page Offset}};
+\draw (471,54) node [anchor=north west][inner sep=0.75pt]  [font=\scriptsize] [align=left] {{\scriptsize Page Offset}};
 % Text Node
-\draw (67,63) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\scriptsize va}};
+\draw (87,52) node [anchor=north west][inner sep=0.75pt]  [font=\scriptsize] [align=left] {{\scriptsize va}};
 % Text Node
-\draw (48,203) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\scriptsize L4 Entry}};
+\draw (46,192) node [anchor=north west][inner sep=0.75pt]  [font=\scriptsize] [align=left] {{\scriptsize L4 Entry}};
 % Text Node
-\draw (180,189) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\scriptsize L3 Entry}};
+\draw (178,178) node [anchor=north west][inner sep=0.75pt]  [font=\scriptsize] [align=left] {{\scriptsize L3 Entry}};
 % Text Node
-\draw (311,167) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\scriptsize L2 Entry}};
+\draw (309,156) node [anchor=north west][inner sep=0.75pt]  [font=\scriptsize] [align=left] {{\scriptsize L2 Entry}};
 % Text Node
-\draw (444,223) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\scriptsize L1 Entry}};
+\draw (442,212) node [anchor=north west][inner sep=0.75pt]  [font=\scriptsize] [align=left] {{\scriptsize L1 Entry}};
 % Text Node
-\draw (51,347) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\scriptsize root}};
+\draw (59,322) node [anchor=north west][inner sep=0.75pt]  [font=\scriptsize] [align=left] {{\scriptsize root}};
 % Text Node
-\draw (465,349) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\scriptsize page}};
+\draw (471,321) node [anchor=north west][inner sep=0.75pt]  [font=\scriptsize] [align=left] {{\scriptsize page}};
 % Text Node
-\draw (529,346) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\scriptsize 4K Data}};
+\draw (517,321) node [anchor=north west][inner sep=0.75pt]  [font=\scriptsize] [align=left] {{\scriptsize 4K Data}};
 % Text Node
-\draw (133,386) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\scriptsize 51}};
+\draw (131,358) node [anchor=north west][inner sep=0.75pt]  [font=\scriptsize] [align=left] {{\scriptsize 51}};
 % Text Node
-\draw (258,386) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\scriptsize 12}};
+\draw (256,358) node [anchor=north west][inner sep=0.75pt]  [font=\scriptsize] [align=left] {{\scriptsize 12}};
 % Text Node
-\draw (163,350) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\scriptsize Base Addr}};
+\draw (161,322) node [anchor=north west][inner sep=0.75pt]  [font=\scriptsize] [align=left] {{\scriptsize Base Addr}};
 
 \end{tikzpicture}
+
           \caption{x86-64 page table lookups.}
@@
 %    \begin{figure}
-\begin{lstlisting}[mathescape,escapeinside={(*}{*)},basicstyle=\footnotesize]
+\begin{lstlisting}[mathescape,escapeinside={(*}{*)},basicstyle=\scriptsize]
 pte_t *pte_get_next_table(pte_t *entry) {
@@
 \begin{minipage}{.4\textwidth}
-\begin{lstlisting}[mathescape,escapeinside={(*}{*)}, basicstyle=\footnotesize]
+\begin{lstlisting}[mathescape,escapeinside={(*}{*)}, basicstyle=\scriptsize]
 pte_t *walkpgdir(pte_t *root, void *va){ 
@@
 other worlds.
+\looseness=-1
 

### 2025-07-27 background.tex
further shrinkage to bg figures, cleanup/reorg and compression of bg text

@@
 
-\tikzset{every picture/.style={line width=0.75pt}} %set default line width to 0.75pt
+% \tikzset{every picture/.style={line width=0.75pt}} %set default line width to 0.75pt
 
-\begin{tikzpicture}[x=0.75pt,y=0.75pt,yscale=-0.55,xscale=0.55]
-%uncomment if require: \path (0,403); %set diagram left start at 0, and has height of 403
+% \begin{tikzpicture}[x=0.75pt,y=0.75pt,yscale=-0.55,xscale=0.55]
+% %uncomment if require: \path (0,403); %set diagram left start at 0, and has height of 403
 
-%Shape: Rectangle [id:dp9406932326447903] 
-\draw   (33,133) -- (123,133) -- (123,284) -- (33,284) -- cycle ;
-%Shape: Rectangle [id:dp7682116051394419] 
-\draw   (166,133) -- (256,133) -- (256,284) -- (166,284) -- cycle ;
-%Shape: Rectangle [id:dp07376733516930423] 
-\draw   (295,133) -- (385,133) -- (385,284) -- (295,284) -- cycle ;
-%Shape: Rectangle [id:dp7106629147891306] 
-\draw   (427,133) -- (517,133) -- (517,284) -- (427,284) -- cycle ;
-%Shape: Rectangle [id:dp1342906826203598] 
-\draw   (115,46) -- (185,46) -- (185,86) -- (115,86) -- cycle ;
-%Shape: Rectangle [id:dp08072439007207399] 
-\draw   (185,46) -- (255,46) -- (255,86) -- (185,86) -- cycle ;
-%Shape: Rectangle [id:dp8282933603833216] 
-\draw   (255,46) -- (325,46) -- (325,86) -- (255,86) -- cycle ;
-%Shape: Rectangle [id:dp8416255090865326] 
-\draw   (325,46) -- (395,46) -- (395,86) -- (325,86) -- cycle ;
-%Shape: Rectangle [id:dp22911958218529804] 
-\draw   (395,46) -- (465,46) -- (465,86) -- (395,86) -- cycle ;
-%Shape: Rectangle [id:dp33732955213990756] 
-\draw   (464.5,46) -- (555.5,46) -- (555.5,86) -- (464.5,86) -- cycle ;
-%Curve Lines [id:da6688730321284677] 
-\draw    (211.5,88) .. controls (-36.44,90.94) and (15.75,182.24) .. (30.17,205.65) ;
-\draw [shift={(31,207)}, rotate = 238.24] [color={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.75]    (10.93,-3.29) ..
+% %Shape: Rectangle [id:dp9406932326447903] 
+% \draw   (33,133) -- (123,133) -- (123,284) -- (33,284) -- cycle ;
+% %Shape: Rectangle [id:dp7682116051394419] 
+% \draw   (166,133) -- (256,133) -- (256,284) -- (166,284) -- cycle ;
+% %Shape: Rectangle [id:dp07376733516930423] 
+% \draw   (295,133) -- (385,133) -- (385,284) -- (295,284) -- cycle ;
+% %Shape: Rectangle [id:dp7106629147891306] 
+% \draw   (427,133) -- (517,133) -- (517,284) -- (427,284) -- cycle ;
+% %Shape: Rectangle [id:dp1342906826203598] 
+% \draw   (115,46) -- (185,46) -- (185,86) -- (115,86) -- cycle ;
+% %Shape: Rectangle [id:dp08072439007207399] 
+% \draw   (185,46) -- (255,46) -- (255,86) -- (185,86) -- cycle ;
+% %Shape: Rectangle [id:dp8282933603833216] 
+% \draw   (255,46) -- (325,46) -- (325,86) -- (255,86) -- cycle ;
+% %Shape: Rectangle [id:dp8416255090865326] 
+% \draw   (325,46) -- (395,46) -- (395,86) -- (325,86) -- cycle ;
+% %Shape: Rectangle [id:dp22911958218529804] 
+% \draw   (395,46) -- (465,46) -- (465,86) -- (395,86) -- cycle ;
+% %Shape: Rectangle [id:dp33732955213990756] 
+% \draw   (464.5,46) -- (555.5,46) -- (555.5,86) -- (464.5,86) -- cycle ;
+% %Curve Lines [id:da6688730321284677] 
+% \draw    (211.5,88) .. controls (-36.44,90.94) and (15.75,182.24) .. (30.17,205.65) ;
+% \draw [shift={(31,207)}, rotate = 238.24] [color={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.75]    (10.93,-3.29) ..
 controls (6.95,-1.4) and (3.31,-0.3) ..
@@
 (10.93,3.29)   ;
-%Shape: Rectangle [id:dp6044747978986409] 
-\draw   (33,183) -- (123,183) -- (123,223) -- (33,223) -- cycle ;
-%Curve Lines [id:da8867562998134584] 
-\draw    (289.5,87) .. controls (115.15,102.76) and (137.77,167.03) .. (163.33,189.98) ;
-\draw [shift={(164.5,191)}, rotate = 220.24] [color={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.75]    (10.93,-3.29) ..
+% %Shape: Rectangle [id:dp6044747978986409] 
+% \draw   (33,183) -- (123,183) -- (123,223) -- (33,223) -- cycle ;
+% %Curve Lines [id:da8867562998134584] 
+% \draw    (289.5,87) .. controls (115.15,102.76) and (137.77,167.03) .. (163.33,189.98) ;
+% \draw [shift={(164.5,191)}, rotate = 220.24] [color={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.75]    (10.93,-3.29) ..
 controls (6.95,-1.4) and (3.31,-0.3) ..
@@
 (10.93,3.29)   ;
-%Shape: Rectangle [id:dp8861153167245819] 
-\draw   (166,169) -- (256,169) -- (256,209) -- (166,209) -- cycle ;
-%Shape: Rectangle [id:dp05979844200848894] 
-\draw  [dash pattern={on 4.5pt off 4.5pt}] (295,148) -- (385,148) -- (385,188) -- (295,188) -- cycle ;
-%Shape: Rectangle [id:dp17403733244247843] 
-\draw   (427,204) -- (517,204) -- (517,244) -- (427,244) -- cycle ;
-%Curve Lines [id:da20581124900573955] 
-\draw [color={rgb, 255:red, 0; green, 0; blue, 0 }  ,draw opacity=1 ] [dash pattern={on 4.5pt off 4.5pt}]  (357,89) .. controls (262.92,102.72) and (272.56,140.45) .. (291.81,165.48) ;
-\draw [shift={(293,167)}, rotate = 231.34] [color={rgb, 255:red, 0; green, 0; blue, 0 }  ,draw opacity=1 ][line width=0.75]    (10.93,-3.29) ..
+% %Shape: Rectangle [id:dp8861153167245819] 
+% \draw   (166,169) -- (256,169) -- (256,209) -- (166,209) -- cycle ;
+% %Shape: Rectangle [id:dp05979844200848894] 
+% \draw  [dash pattern={on 4.5pt off 4.5pt}] (295,148) -- (385,148) -- (385,188) -- (295,188) -- cycle ;
+% %Shape: Rectangle [id:dp17403733244247843] 
+% \draw   (427,204) -- (517,204) -- (517,244) -- (427,244) -- cycle ;
+% %Curve Lines [id:da20581124900573955] 
+% \draw [color={rgb, 255:red, 0; green, 0; blue, 0 }  ,draw opacity=1 ] [dash pattern={on 4.5pt off 4.5pt}]  (357,89) .. controls (262.92,102.72) and (272.56,140.45) .. (291.81,165.48) ;
+% \draw [shift={(293,167)}, rotate = 231.34] [color={rgb, 255:red, 0; green, 0; blue, 0 }  ,draw opacity=1 ][line width=0.75]    (10.93,-3.29) ..
 controls (6.95,-1.4) and (3.31,-0.3) ..
@@
 (10.93,3.29)   ;
-%Curve Lines [id:da07589611603635316] 
-\draw    (425.5,87) .. controls (380.42,111.5) and (406.41,194.58) .. (422.53,216.72) ;
-\draw [shift={(423.5,218)}, rotate = 231.34] [color={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.75]    (10.93,-3.29) ..
+% %Curve Lines [id:da07589611603635316] 
+% \draw    (425.5,87) .. controls (380.42,111.5) and (406.41,194.58) .. (422.53,216.72) ;
+% \draw [shift={(423.5,218)}, rotate = 231.34] [color={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.75]    (10.93,-3.29) ..
 controls (6.95,-1.4) and (3.31,-0.3) ..
@@
 (10.93,3.29)   ;
-%Curve Lines [id:da6852167546673456] 
-\draw    (497.5,87) .. controls (564.48,83.06) and (567.43,193.61) .. (570.37,231.33) ;
-\draw [shift={(570.5,233)}, rotate = 265.24] [color={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.75]    (10.93,-3.29) ..
+% %Curve Lines [id:da6852167546673456] 
+% \draw    (497.5,87) .. controls (564.48,83.06) and (567.43,193.61) .. (570.37,231.33) ;
+% \draw [shift={(570.5,233)}, rotate = 265.24] [color={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.75]    (10.93,-3.29) ..
 controls (6.95,-1.4) and (3.31,-0.3) ..
@@
 (10.93,3.29)   ;
-%Curve Lines [id:da017521077423665377] 
-\draw    (36.5,330) .. controls (-15.71,340.84) and (16.98,314.8) .. (32.31,285.35) ;
-\draw [shift={(33,284)}, rotate = 116.57] [color={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.75]    (10.93,-3.29) ..
+% %Curve Lines [id:da017521077423665377] 
+% \draw    (36.5,330) .. controls (-15.71,340.84) and (16.98,314.8) .. (32.31,285.35) ;
+% \draw [shift={(33,284)}, rotate = 116.57] [color={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.75]    (10.93,-3.29) ..
 controls (6.95,-1.4) and (3.31,-0.3) ..
@@
 (10.93,3.29)   ;
-%Curve Lines [id:da6750042432035159] 
-\draw    (122,207) .. controls (128.9,246.4) and (131.91,261.54) .. (159.71,281.1) ;
-\draw [shift={(161,282)}, rotate = 214.59] [color={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.75]    (10.93,-3.29) ..
+% %Curve Lines [id:da6750042432035159] 
+% \draw    (122,207) .. controls (128.9,246.4) and (131.91,261.54) .. (159.71,281.1) ;
+% \draw [shift={(161,282)}, rotate = 214.59] [color={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.75]    (10.93,-3.29) ..
 controls (6.95,-1.4) and (3.31,-0.3) ..
@@
 (10.93,3.29)   ;
-%Curve Lines [id:da14788158682113428] 
-\draw    (257,189) .. controls (258.95,228) and (277.06,262.25) .. (293.72,282.47) ;
-\draw [shift={(295,284)}, rotate = 229.64] [color={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.75]    (10.93,-3.29) ..
+% %Curve Lines [id:da14788158682113428] 
+% \draw    (257,189) .. controls (258.95,228) and (277.06,262.25) .. (293.72,282.47) ;
+% \draw [shift={(295,284)}, rotate = 229.64] [color={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.75]    (10.93,-3.29) ..
 controls (6.95,-1.4) and (3.31,-0.3) ..
@@
 (10.93,3.29)   ;
-%Curve Lines [id:da3297977595767969] 
-\draw    (386,169) .. controls (392.86,218) and (408.36,262.2) .. (425.92,282.77) ;
-\draw [shift={(427,284)}, rotate = 228.01] [color={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.75]    (10.93,-3.29) ..
+% %Curve Lines [id:da3297977595767969] 
+% \draw    (386,169) .. controls (392.86,218) and (408.36,262.2) .. (425.92,282.77) ;
+% \draw [shift={(427,284)}, rotate = 228.01] [color={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.75]    (10.93,-3.29) ..
 controls (6.95,-1.4) and (3.31,-0.3) ..
@@
 (10.93,3.29)   ;
-%Curve Lines [id:da7710391961792686] 
-\draw    (517,221) .. controls (535.13,236.68) and (527.81,247.56) .. (555.75,240.46) ;
-\draw [shift={(557.5,240)}, rotate = 165.07] [color={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.75]    (10.93,-3.29) ..
+% %Curve Lines [id:da7710391961792686] 
+% \draw    (517,221) .. controls (535.13,236.68) and (527.81,247.56) .. (555.75,240.46) ;
+% \draw [shift={(557.5,240)}, rotate = 165.07] [color={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.75]    (10.93,-3.29) ..
 controls (6.95,-1.4) and (3.31,-0.3) ..
@@
 (10.93,3.29)   ;
-%Shape: Rectangle [id:dp1535611125971872] 
-\draw   (511,313) -- (581,313) -- (581,353) -- (511,353) -- cycle ;
-%Shape: Rectangle [id:dp05427947476226991] 
-\draw   (95,313) -- (136.5,313) -- (136.5,353) -- (95,353) -- cycle ;
-%Shape: Rectangle [id:dp10511138743752002] 
-\draw   (136.5,313) -- (265.5,313) -- (265.5,353) -- (136.5,353) -- cycle ;
-%Shape: Rectangle [id:dp8917680919534263] 
-\draw   (265.5,313) -- (305,313) -- (305,353) -- (265.5,353) -- cycle ;
-\draw   (560,243.5) ..
+% %Shape: Rectangle [id:dp1535611125971872] 
+% \draw   (511,313) -- (581,313) -- (581,353) -- (511,353) -- cycle ;
+% %Shape: Rectangle [id:dp05427947476226991] 
+% \draw   (95,313) -- (136.5,313) -- (136.5,353) -- (95,353) -- cycle ;
+% %Shape: Rectangle [id:dp10511138743752002] 
+% \draw   (136.5,313) -- (265.5,313) -- (265.5,353) -- (136.5,353) -- cycle ;
+% %Shape: Rectangle [id:dp8917680919534263] 
+% \draw   (265.5,313) -- (305,313) -- (305,353) -- (265.5,353) -- cycle ;
+% \draw   (560,243.5) ..
 controls (560,238.25) and (564.25,234) ..
@@
 (560,243.5) -- cycle ; \draw   (560,243.5) -- (579,243.5) ; \draw   (569.5,234) -- (569.5,253) ;
-%Curve Lines [id:da9312068388962929] 
-\draw    (569,258) .. controls (585.66,269.76) and (587.91,278.64) .. (566.35,295.93) ;
-\draw [shift={(565,297)}, rotate = 321.95] [color={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.75]    (10.93,-3.29) ..
+% %Curve Lines [id:da9312068388962929] 
+% \draw    (569,258) .. controls (585.66,269.76) and (587.91,278.64) .. (566.35,295.93) ;
+% \draw [shift={(565,297)}, rotate = 321.95] [color={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.75]    (10.93,-3.29) ..
 controls (6.95,-1.4) and (3.31,-0.3) ..
@@
 
-% Text Node
-\draw (21,20) node [anchor=north west][inner sep=0.75pt]  [font=\scriptsize] [align=left] {{\scriptsize Virtual}};
-% Text Node
-\draw (569,375) node [anchor=north west][inner sep=0.75pt]  [font=\scriptsize] [align=left] {{\scriptsize Physical}};
-% Text Node
-\draw (102,17) node [anchor=north west][inner sep=0.75pt]  [font=\scriptsize] [align=left] {{\scriptsize 63}};
-% Text Node
-\draw (173,17) node [anchor=north west][inner sep=0.75pt]  [font=\scriptsize] [align=left] {{\scriptsize 48}};
-% Text Node
-\draw (243,17) node [anchor=north west][inner sep=0.75pt]  [font=\scriptsize] [align=left] {{\scriptsize 39}};
-% Text Node
-\draw (312,17) node [anchor=north west][inner sep=0.75pt]  [font=\scriptsize] [align=left] {{\scriptsize 30}};
-% Text Node
-\draw (383,17) node [anchor=north west][inner sep=0.75pt]  [font=\scriptsize] [align=left] {{\scriptsize 21}};
-% Text Node
-\draw (452,17) node [anchor=north west][inner sep=0.75pt]  [font=\scriptsize] [align=left] {{\scriptsize 12}};
-% Text Node
-\draw (547,16) node [anchor=north west][inner sep=0.75pt]  [font=\scriptsize] [align=left] {{\scriptsize 0}};
-% Text Node
-\draw (126,54) node [anchor=north west][inner sep=0.75pt]  [font=\scriptsize] [align=left] {{\scriptsize Sign}};
-% Text Node
-\draw (188,54) node [anchor=north west][inner sep=0.75pt]  [font=\scriptsize] [align=left] {{\scriptsize L4 Offset}};
-% Text Node
-\draw (257,54) node [anchor=north west][inner sep=0.75pt]  [font=\scriptsize] [align=left] {{\scriptsize L3 Offset}};
-% Text Node
-\draw (328,54) node [anchor=north west][inner sep=0.75pt]  [font=\scriptsize] [align=left] {{\scriptsize L2 Offset}};
-% Text Node
-\draw (398,55) node [anchor=north west][inner sep=0.75pt]  [font=\scriptsize] [align=left] {{\scriptsize L1 Offset}};
-% Text Node
-\draw (471,54) node [anchor=north west][inner sep=0.75pt]  [font=\scriptsize] [align=left] {{\scriptsize Page Offset}};
-% Text Node
-\draw (87,52) node [anchor=north west][inner sep=0.75pt]  [font=\scriptsize] [align=left] {{\scriptsize va}};
-% Text Node
-\draw (46,192) node [anchor=north west][inner sep=0.75pt]  [font=\scriptsize] [align=left] {{\scriptsize L4 Entry}};
-% Text Node
-\draw (178,178) node [anchor=north west][inner sep=0.75pt]  [font=\scriptsize] [align=left] {{\scriptsize L3 Entry}};
-% Text Node
-\draw (309,156) node [anchor=north west][inner sep=0.75pt]  [font=\scriptsize] [align=left] {{\scriptsize L2 Entry}};
-% Text Node
-\draw (442,212) node [anchor=north west][inner sep=0.75pt]  [font=\scriptsize] [align=left] {{\scriptsize L1 Entry}};
-% Text Node
-\draw (59,322) node [anchor=north west][inner sep=0.75pt]  [font=\scriptsize] [align=left] {{\scriptsize root}};
-% Text Node
-\draw (471,321) node [anchor=north west][inner sep=0.75pt]  [font=\scriptsize] [align=left] {{\scriptsize page}};
-% Text Node
-\draw (517,321) node [anchor=north west][inner sep=0.75pt]  [font=\scriptsize] [align=left] {{\scriptsize 4K Data}};
-% Text Node
-\draw (131,358) node [anchor=north west][inner sep=0.75pt]  [font=\scriptsize] [align=left] {{\scriptsize 51}};
-% Text Node
-\draw (256,358) node [anchor=north west][inner sep=0.75pt]  [font=\scriptsize] [align=left] {{\scriptsize 12}};
-% Text Node
-\draw (161,322) node [anchor=north west][inner sep=0.75pt]  [font=\scriptsize] [align=left] {{\scriptsize Base Addr}};
+% % Text Node
+% \draw (21,20) node [anchor=north west][inner sep=0.75pt]  [font=\scriptsize] [align=left] {{\scriptsize Virtual}};
+% % Text Node
+% \draw (569,375) node [anchor=north west][inner sep=0.75pt]  [font=\scriptsize] [align=left] {{\scriptsize Physical}};
+% % Text Node
+% \draw (102,17) node [anchor=north west][inner sep=0.75pt]  [font=\scriptsize] [align=left] {{\scriptsize 63}};
+% % Text Node
+% \draw (173,17) node [anchor=north west][inner sep=0.75pt]  [font=\scriptsize] [align=left] {{\scriptsize 48}};
+% % Text Node
+% \draw (243,17) node [anchor=north west][inner sep=0.75pt]  [font=\scriptsize] [align=left] {{\scriptsize 39}};
+% % Text Node
+% \draw (312,17) node [anchor=north west][inner sep=0.75pt]  [font=\scriptsize] [align=left] {{\scriptsize 30}};
+% % Text Node
+% \draw (383,17) node [anchor=north west][inner sep=0.75pt]  [font=\scriptsize] [align=left] {{\scriptsize 21}};
+% % Text Node
+% \draw (452,17) node [anchor=north west][inner sep=0.75pt]  [font=\scriptsize] [align=left] {{\scriptsize 12}};
+% % Text Node
+% \draw (547,16) node [anchor=north west][inner sep=0.75pt]  [font=\scriptsize] [align=left] {{\scriptsize 0}};
+% % Text Node
+% \draw (126,54) node [anchor=north west][inner sep=0.75pt]  [font=\scriptsize] [align=left] {{\scriptsize Sign}};
+% % Text Node
+% \draw (188,54) node [anchor=north west][inner sep=0.75pt]  [font=\scriptsize] [align=left] {{\scriptsize L4 Offset}};
+% % Text Node
+% \draw (257,54) node [anchor=north west][inner sep=0.75pt]  [font=\scriptsize] [align=left] {{\scriptsize L3 Offset}};
+% % Text Node
+% \draw (328,54) node [anchor=north west][inner sep=0.75pt]  [font=\scriptsize] [align=left] {{\scriptsize L2 Offset}};
+% % Text Node
+% \draw (398,55) node [anchor=north west][inner sep=0.75pt]  [font=\scriptsize] [align=left] {{\scriptsize L1 Offset}};
+% % Text Node
+% \draw (471,54) node [anchor=north west][inner sep=0.75pt]  [font=\scriptsize] [align=left] {{\scriptsize Page Offset}};
+% % Text Node
+% \draw (87,52) node [anchor=north west][inner sep=0.75pt]  [font=\scriptsize] [align=left] {{\scriptsize va}};
+% % Text Node
+% \draw (46,192) node [anchor=north west][inner sep=0.75pt]  [font=\scriptsize] [align=left] {{\scriptsize L4 Entry}};
+% % Text Node
+% \draw (178,178) node [anchor=north west][inner sep=0.75pt]  [font=\scriptsize] [align=left] {{\scriptsize L3 Entry}};
+% % Text Node
+% \draw (309,156) node [anchor=north west][inner sep=0.75pt]  [font=\scriptsize] [align=left] {{\scriptsize L2 Entry}};
+% % Text Node
+% \draw (442,212) node [anchor=north west][inner sep=0.75pt]  [font=\scriptsize] [align=left] {{\scriptsize L1 Entry}};
+% % Text Node
+% \draw (59,322) node [anchor=north west][inner sep=0.75pt]  [font=\scriptsize] [align=left] {{\scriptsize root}};
+% % Text Node
+% \draw (471,321) node [anchor=north west][inner sep=0.75pt]  [font=\scriptsize] [align=left] {{\scriptsize page}};
+% % Text Node
+% \draw (517,321) node [anchor=north west][inner sep=0.75pt]  [font=\scriptsize] [align=left] {{\scriptsize 4K Data}};
+% % Text Node
+% \draw (131,358) node [anchor=north west][inner sep=0.75pt]  [font=\scriptsize] [align=left] {{\scriptsize 51}};
+% % Text Node
+% \draw (256,358) node [anchor=north west][inner sep=0.75pt]  [font=\scriptsize] [align=left] {{\scriptsize 12}};
+% % Text Node
+% \draw (161,322) node [anchor=north west][inner sep=0.75pt]  [font=\scriptsize] [align=left] {{\scriptsize Base Addr}};
 
-\end{tikzpicture}
+% \end{tikzpicture}
 
-          \caption{x86-64 page table lookups.}
+\includegraphics[width=\textwidth]{addr_translation.png}
+\vspace{-1em}
+          \caption{x86-64 \replace{page table lookups}{hardware address translation}.}
         \label{fig:pagetables}
@@
     \vfill
+\hspace{5em}
 \begin{minipage}{.5\textwidth}
@@
   pte_t *next;
-  /* Check if entry is populated, allocate if needed(* \label{line:check_entry_presentC} *)(* \label{line:mask_presentC} *)(*\label{line:after_concluding_qfrac1C}*)*/
-  if (!entry->present(*\label{line:read_entry_contentsC}*) (*\label{line:conditional_childrenC}*)) {
-    pte_initialize(entry); // Alloc empty table for next level (*\label{line:pass_addrof_nextC}*)(*\label{line:alloc_path_startC}*)(* \label{line:call_to_pte_initializeC} *)
-    entry->present = 1; // Mark valid (*\label{line:install_new_entryC}*)
-    /* Key invariant for child access now holds */
-  } (*\label{line:end_of_allocation_pathCB}*)
-  /*Physical to virtual address mapping code segment starts here (*\label{line:finalpieceSB}*)*/
+  if (!entry->present) { // Check if entry needs initialization
+    pte_initialize(entry); // Alloc empty table for next level(*\label{line:ptc_alloc}*)
+    entry->present = 1; // Mark valid(*\label{line:ptc_mark}*)
+  }
+  /* Convert phys. addr. from PTE to virt. addr. for access */
   uintptr_t next_phys_addr = PTE_PFN_TO_ADDR(entry->pfn)
-  uintptr_t next_virt_addr = (uintptr_t) P2V(next_phys_addr); (*\label{line:p2vCB}*);
-  next = (pte_t *) next_virt_addr;(*\label{line:finalpieceEB}*)
+  uintptr_t next_virt_addr = (uintptr_t) P2V(next_phys_addr);(*\label{line:ptc_p2v}*)
+  next = (pte_t *) next_virt_addr;
   return next;
@@
 \begin{minipage}{.4\textwidth}
+\hspace{3em}
 \begin{lstlisting}[mathescape,escapeinside={(*}{*)}, basicstyle=\scriptsize]
@@
     \vspace{-2.5em}
-    \caption{A kernel implementation of x86-64 page table lookups.}
+    \caption{Most of a software page table walk in C, as used to map new pages.}
     \label{fig:pagetablescode}
@@
 
-\todo[inline]{Colin says: the changes in the next couple paragraphs seem to have deleted critical information about how the hardware
-translates addresses.}
 On the x86-64 architecture, the \textsc{MMU}'s address translation uses a sparse hierarchical set of tables:
@@
 the x86-64 architecture as a 64-bit extension to x86 was originally due to AMD. As a result, it is sometimes also referred to as the \texttt{amd64} architecture.})
-\replace{shows,
-address translation proceeds by repeatedly taking designated slices of the virtual address and indexing into a table.
-The final lookup in the page tables gives the base physical address of a 4KB page of physical memory, to which the low-order
-bits of the accessed virtual address are added to determine the actual physical address retrieved.
+shows,
+address translation proceeds by repeatedly taking designated slices of the virtual address and indexing into \replace{a}{successive} table\add{s} \add{of 512 8-byte entries (making each table 4KB in size)}.
+The final lookup in the page tables gives the base physical address of a 4KB page of physical memory \add{accessible to the running
+program}, to which the low-order
+\add{12} bits of the accessed virtual address are added to determine the actual physical address retrieved.
 On x86-64, standard configurations use 4 levels of page tables, labelled levels 4 through 1, with lookups in the level 
-}{
-shows a \emph{page-table walk} for a virtual address. The page tables are managed by the OS, typically by a \emph{virtual memory manager} (VMM).\footnote{Not to be confused with Virtual 
-Machine Monitor. We focus on non-hypervisor scenarios, but hardware virtualization extensions for both 
-x86-64 and ARM make use of an additional set of page tables translating what a \emph{guest} considers to be its 
-(virtualized) physical memory to actual physical memory. Our contributions should offer value in this scenario as well.}
-Typically each process has its own 
-page table, which the OS registers with the CPU by storing the (page-aligned)
-physical address of the root of the page table tree (the start of the L4 table shown as $\mathsf{root}$ in Figure \ref{fig:pagetables}) in a specific register (\texttt{cr3}) as part of switching to a new process.
-Using different mappings, which map only disjoint portions of physical memory (with some exceptions in the next section) 
-is how the OS ensures memory isolation between processes.
-}
-
-\add{
-In Figure \ref{fig:pagetablescode}, we see a simplified kernel code, \lstinline|walkpgdir|, implementing this page-table walk for translating the virtual address $\vaddr$ over the page-table tree with unique root address $\mathsf{root}$ shown in Figure \ref{fig:pagetables}.
-As shown in Figure at the top of Figure \ref{fig:pagetables}, on x86-64, standard configurations use 4 levels of page tables, labelled levels 4 through 1, with lookups in the level 
-}
 1 page table resulting in the actual page of physical memory holding the requested data, and the low-order 12 bits 
@@
 but for brevity and consistency, we simply number them, in keeping with the newer 5th level. Our formalization only deals with 4-level page tables, but is straightforwardly extensible to 5.} 
+The translation process or algorithm is \replace{sometimes}{often} referred to as a \emph{page-table walk}.
+While Figure \ref{fig:pagetables} and most of our constants (how many levels, which virtual address bits index which
+table levels) are specific to \replace{the}{common} x86-64 \replace{architecture}{CPUs},
 \replace{
-  The translation process or algorithm is sometimes referred to as a \emph{page-table walk}. 
-}{
-As depicted at the top of Figure \ref{fig:pagetables}, for each level of translation, address-translation implementation repeatedly takes designated slices of the virtual address and indexing into a table --- e.g., L4Offset(va) at Line 2 of \lstinline|walkpgdir| in Figure \ref{fig:pagetablescode}.
-The final lookup in the page tables gives the base physical address of a 4KB page of physical memory, to which the low-order
-bits of the accessed virtual address are added to determine the actual physical address retrieved.
-The entries \add{(e.g., L4 Entry in Figure \ref{fig:pagetables})} of each table are 64 bits wide, but each points to a physical address aligned to 4KB (4096 byte) boundaries, which leaves 12 bits to spare to
-control a validity bit (called the \emph{present} bit), a read-write bit (which permits write access through the entry if and only if it is set),
-and a range of additional bits which can be used to control caching, write-through, and more.
-This paper will only consider the present bit (0).
-}
-\add{As we see in Figure \ref{fig:pagetables}, the entry L2 is not be present (shown as dashed in Figure \ref{fig:pagetables}).
-In order to able to access to the subsequent table base address (e.g., $\mathsf{l1}$ in Line 7 of \lstinline|walkpgdir| in Figure \ref{fig:pagetablescode}).
-Accessing to the next table is realized in \lstinline|pte_get_next_table| in Figure \ref{fig:pagetablescode} to which the current entry (e.g., $\mathsf{l2_entry}$ in Line 7 of \lstinline|walkpgdir| in Figure \ref{fig:pagetablescode}), and this method first check if the entry is present (Line 7 of \lstinline|pte_get_next_table| in Figure \ref{fig:pagetablecodes}).
-For a nonexisting entry such as L2 entry, the execution goes through allocation path shown in Lines 7-15 of \lstinline|pte_get_next_table| in Figure \ref{fig:pagetablecodes} which allocates a \emph{physical addresses} (512 L1 entries) and make $\mathsf{l2}$ pointer's frame address to show the first entry of the L1 entries fresly allocated (Line 12 of \lstinline|pte_get_next_table| in Figure \ref{fig:pagetablecodes}), and set the present bit of $\mathsf{l2_entry}$ (Line 12 of \lstinline|pte_get_next_table| in Figure \ref{fig:pagetablecodes}).
-After the validity of page-table walk is restored, \lstinline|pte_get_next_table| converts the physical frame address of the entry to a \emph{virtual address} in Lines 16-20 of \lstinline|pte_get_next_table| in Figure \ref{fig:pagetablecodes} return the virtual address of the subsequent ($\mathsf{next}$ in in Figure \ref{fig:pagetablecodes}) base address (e.g., $\mathsf{l1}$ in Line 7 of \lstinline|walkpgdir| in Figure \ref{fig:pagetablescode}).
-}
-
-While Figure \ref{fig:pagetables} and most of our constants (how many levels, which virtual address bits index which
-table levels) are specific to the x86-64 architecture, ARMv8 (a.k.a.\ 
-\texttt{aarch64}), RISC-V, and PowerPC use similar hierarchical page tables for address translation.
+ARMv8 (a.k.a.\ texttt{aarch64}), RISC-V, and PowerPC use similar hierarchical page tables for address translation.
 RISC-V's \texttt{sv48} paging configuration~\cite[\S 4.5]{riscv-privileged} and AArch64's 4-level paging configuration
@@
 in different orders).
-
-\del{
-  The entries of each table are 64 bits wide, but each points to a physical address aligned to 4KB (4096 byte) boundaries, which leaves 12 bits to spare to
-  control a validity bit (called the \emph{present} bit), a read-write bit (which permits write access through the entry if and only if it is set),
-  and a range of additional bits which can be used to control caching, write-through, and more. This paper will only consider the present bit (0).
+}{
+  it is straightforward to adapt our approach to ARMv8/\texttt{aarch64}, RISC-V, or PowerPC, which use nearly-identical
+  page table structures (only the order of flag bits in the entries differ),
+  or to 5-level paging present in newer x86-64 or RISC-V designs (one step would be iterated an additional time).
 }
 
-\del{
+The entries \add{(e.g., L4 Entry in Figure \ref{fig:pagetables})} of each table are 64 bits wide,
+but each points to a physical address aligned to 4KB (4096 byte) boundaries, which leaves 12 bits to spare to 
+control a validity bit (called the \emph{present} bit), a read-write bit (which permits write access through the entry if and only if it is set),
+and a range of additional bits which can be used to control caching, write-through, and more.
+This paper will only consider the present bit (0).
+
 The page tables are managed by the OS, typically by a \emph{virtual memory manager} (VMM).\footnote{Not to be confused with Virtual 
@@
 Typically each process has its own 
-  page table, which the OS registers with the CPU by storing the (page-aligned)
-  physical address of the root of the page table treethe start of the L4 table) in a
-  specific register (\texttt{cr3}) as part of switching to a new process.
+page tables, which the OS registers with the CPU by storing the (page-aligned)
+physical address of the root of the page table tree (the start of the L4 table shown as $\mathsf{root}$ in Figure \ref{fig:pagetables}) in a specific register (\texttt{cr3}) as part of switching to a new process.
 Using different mappings, which map only disjoint portions of physical memory (with some exceptions in the next section) 
 is how the OS ensures memory isolation between processes.
-}
 
@@
 and the faulting program should be terminated. But this can also be used for
+\replace{
 specialized functionality and optimizations, such as \emph{paging} (saving room in physical RAM and deferring
@@
 recently been accessed to disk, to read back in when a page fault indicates access).
+}{
+  other purposes such as demand paging to save on IO and better-manage physical memory~\cite{doeppner2010operating}.
+}
 
@@
 
+\add{
+  To get a sense of how subtle the required reasoning is, we can consider code such as Figure \ref{fig:pagetablescode},
+  used in our evaluation to locate the appropriate L1 entry to map a page into the current address space.
+  % Adding or removing mappings requires the VMM code to locate the correct page table entry to update,
+  % using code such as that in Figure \ref{fig:pagetablescode}.
+  \lstinline|walkpgdir| (right) essentially mimics the hardware address translation up to the L1 entry (its caller will modify
+  the entry to map a new page),
+  indexing into each successive table (e.g., with \lstinline|L4Offset(va)| retrieving the L4 offset of Figure \ref{fig:pagetables}),
+  with \lstinline|pte_get_next_table| (left) fetching the base address of the next table
+  from each entry.
+  If, as shown in Figure \ref{fig:pagetables}, an entry such as the L2 entry is uninitialized (the \emph{present} flag is not set),
+  \lstinline|pte_get_next_table| allocates the next table on Line \ref{line:ptc_alloc} and marks it valid on Line \ref{line:ptc_mark}.
+  But there is a critical difference from hardware: every memory access in Figure \ref{fig:pagetablescode} uses \emph{virtual} addresses, as opposed
+  to hardware's direct physical memory access.
+  Thus \lstinline|pte_get_next_table| uses the \lstinline|P2V| (physical-to-virtual) macro on line \ref{line:ptc_p2v}
+  to convert physical addresses stored in page table entries into virtual addresses which the kernel can use to access
+  the corresponding physical location.
+  Proving it correct requires proving that it yields
+  an address that is not only mapped, but known to map back to the original physical address!
+  Verifying this code is beyond the reach of prior work, which either
+  does not model address translation for the kernel~\cite{gu15,gu2018certikos,Klein2009seL4,seL4TOCS,alkassar2008formal}
+  and would thus reason unsoundly about this code,
+  or models address translation but lacks features required to reason about this code~\cite{kolanski08vstte,kolanski09tphols}.
+  We describe our proof of correctness for this code,
+  based on the first formalization of critical VMM kernel invariants in an MMU-aware logic, in Section \ref{sec:experiment}.
+}
+\looseness=-1
+
 Virtual memory concerns propagate to the OS scheduler, which deals with multiple 
@@
 
-\paragraph{\add{Mapping a New Page}}
-\add{Mapping a new page into a virtual address space is a key piece fo functionality, and also a key challenge for verification.
-This operation takes a chunk of physical memory to install, and a (page-aligned) virtual address at which to make it
-visible to the process.
-Starting from the page table root (\lstinline|cr3| in Figure \ref{fig:pagetables}), it must traverse the page tables until it locates
-the correct location of the L1 entry which should point to the physical page (installing new L3, L2, and L1 tables as needed),
-and update the L1 entry to point to the physical page. This is complicated by the fact that the kernel accesses the page
-tables through \emph{virtual} addresses, but the page table root and the addresses in the L4--L1 entries are all \emph{physical}
-addresses. So the mapping process must repeatedly perform \emph{backwards} mappings (not supported by hardware) from
-physical addresses of each level's entry to a virtual address that is then mapped back to the corresponding physical address!
-And for modularity, the translation process (and possible allocation of missing tables) is typically implemented in a function
-used for every level of translation (i.e., a level-polymorphic function).
-These compounding subtleties are one reason \textsc{CertiKOS} and \textsc{seL4} \emph{trust} their page table
-manipulations; both the reasoning to prove correctness, and the kernel invariants that ensure correctness of typical implementations,
-are remarkably subtle.
-}
-\looseness=-1
+% \paragraph{\add{Mapping a New Page}}
+% \add{Mapping a new page into a virtual address space is a key piece fo functionality, and also a key challenge for verification.
+% This operation takes a chunk of physical memory to install, and a (page-aligned) virtual address at which to make it
+% visible to the process.
+% Starting from the page table root (\lstinline|cr3| in Figure \ref{fig:pagetables}), it must traverse the page tables until it locates
+% the correct location of the L1 entry which should point to the physical page (installing new L3, L2, and L1 tables as needed),
+% and update the L1 entry to point to the physical page. This is complicated by the fact that the kernel accesses the page
+% tables through \emph{virtual} addresses, but the page table root and the addresses in the L4--L1 entries are all \emph{physical}
+% addresses. So the mapping process must repeatedly perform \emph{backwards} mappings (not supported by hardware) from
+% physical addresses of each level's entry to a virtual address that is then mapped back to the corresponding physical address!
+% And for modularity, the translation process (and possible allocation of missing tables) is typically implemented in a function
+% used for every level of translation (i.e., a level-polymorphic function).
+% These compounding subtleties are one reason \textsc{CertiKOS} and \textsc{seL4} \emph{trust} their page table
+% manipulations; both the reasoning to prove correctness, and the kernel invariants that ensure correctness of typical implementations,
+% are remarkably subtle.
+% }
+% \looseness=-1
 

### 2025-07-28 background.tex
try to further clarify modal benefits

@@
 other worlds.
-\looseness=-1
+\add{This is somewhat different from the accessibility relation modalities more common in verification, but also well-established,
+tracing its origin back to Arthur Prior~\cite{areces2006hybrid}, just as many temporal logics do.}
 
-\add{This is somewhat different from the accessibility relation modalities more common in verification, but also well-established.}
-This explicit naming of alternate worlds increases the power of modal logics~\cite{blackburn1995hybrid}, and is actually
+This explicit naming of alternate worlds increases the power of \add{propositional} modal logics~\cite{blackburn1995hybrid}, and is actually
 necessary for completeness in classical separation logics~\cite{brotherston2014parametric}.
-%   A key distinguishing feature of
-% hybrid logic is that the logic itself does not fix the set of modalities, but is defined relative to any set of sensible worlds.
+\add{However typically modal logics can always be ``compiled'' into logics with suitable quantifiers:
+the \emph{standard translation}~\cite{blackburn2006modal} of propositional modal logics into first-order logic
+has been thoroughly studied. \textsc{Iris}'s support for higher-order impredicative quantification means
+this applies to our modalities as well --- they do not strictly speaking make \textsc{Iris} more expressive.
+But while propositional modal logics are no more expressive than first-order logic (with just two variables!),
+the primary goal of a modal logic is not expressive power but intuitive readable specifications.
+Temporal logics are used because they simplify specifications (vs. having explicit time variables
+in every base asertion) and reasoning principles (because the modal correspondents of quantification over
+time is used in a highly structured way).
+Our logic offers a similar value proposition: natural specifications and reasoning principles which
+are powerful enough for reasoning about virtual memory management, but without requiring pervasive tracking of
+and explicit abstraction over address spaces.
+}
 

### 2025-07-28 programlogic.tex
more cleanup and compression, now halfway through 4.2

@@
   may manipulate other address spaces, adds significant specification burden.
-  Specifications which need to talk about a particular invariant holding in an arbitrary other address space,
-  or with knowing a certain assertion conceptually holds in multiple specific address spaces, would need to quantify
-  over functions from address space identity to assertions rather than just assertions.
-  Using a modal approach resolves all of these technical and usability\todo{need better word, I'm always wary of using this word in PL}
-  challenges cleanly and uniformly.
+  Specifications which need to talk about a particular invariant holding in a specific other address space (or multiple other address spaces) would need to quantify
+  over \emph{functions} from address space identity to assertions rather than just assertions.
+  Using a modal approach resolves all of these challenges cleanly and uniformly.
   Specifications that are not \emph{about} address space manipulation need not mention an address spaces at all:
-  one can use one's favorite separation logic data structure specifications without restructuring,
+  one can use standard separation logic data structure specifications without restructuring,
   and every assertion can be stated for either the current or specific other address spaces.
-  In short, modalities make it possible for specifications to mention address spaces only when they deal with address spaces,
-  and not when they don't.\todo{wary of this due to the rtv in the double}
+  In short, modalities make it possible for specifications to mention address spaces when necessary,
+  without requiring mention when only one address space is at hand.
 }
@@
 carry over to our derived logic.
-\add{Since \textsf{vProp}s are ``just'' \textsc{Iris} predicates on a choice of page table root,
-it may not be obvious what the gain is over simply using \textsc{Iris} directly.
-The gain is not expressive power.
-Every modal logic can be translated straightforwardly to a logic without modalities
-(nearly always a small variation on the \emph{standard translation} from propositional modal logic
-to first-order logic~\cite{blackburn2006modal}\footnote{\add{Itself similar to verification condition generation
-for weakest precondition calculi and Hoare logics.}}) that can directly address the logic's models ---
-but these translations, in addition to making specifications far more verbose,
-generally do not preserve the appealing reasoning principles of modal logics.
-This becomes \replace{most}{more} apparent when discussing our proof rules.
-}
-
+% \add{Since \textsf{vProp}s are ``just'' \textsc{Iris} predicates on a choice of page table root,\todo{this section may be subsumed under first addition to this section, don't need both}
+% it may not be obvious what the gain is over simply using \textsc{Iris} directly.
+% The gain is not expressive power.
+% Every modal logic can be translated straightforwardly to a logic without modalities
+% (nearly always a small variation on the \emph{standard translation} from propositional modal logic
+% to first-order logic~\cite{blackburn2006modal}\footnote{\add{Itself similar to verification condition generation
+% for weakest precondition calculi and Hoare logics.}}) that can directly address the logic's models ---
+% but these translations, in addition to making specifications far more verbose,
+% generally do not preserve the appealing reasoning principles of modal logics.
+% This becomes \replace{most}{more} apparent when discussing our proof rules.
+% }
 However, we must still build up from existing \iris primitives to provide new primitives that depend on the address space --- primarily the notion
@@
 The assertion $\textsf{r}\;\mapsto_{\textsf{r}}\{q\}\;\textsf{rv}$ ensures the ownership of the register $\rg$ containing the
-value of the register $\rv$.\todo{Ismail, you changed ``the register value $\rv$'' to ``value of the register $\rv$'' and I changed it back.
-What you wrote treats $\rv$ as the name of the register, not its value.}
+value of the register $\rv$.
 The fraction $\qfrac$ with value 1 asserts the unique ownership of the register mapping and grants update permission \replace{on}{to} it;
@@
 \begin{figure*}
-  \begin{lstlisting}[language=Coq]
+  \begin{lstlisting}[language=Coq,escapeinside={(*}{*)}]
  Definition $\vaddr\mapsto_{\textsf{t}}\{\textsf{q}\}\; \vpage$ : vProp $\Sigma$ := 
@@
   $\ulcorner$ entry_present l4e $\land$ entry_present l3e $\land$ entry_present l2e $\land$ entry_present l1e$\urcorner$ $\ast$
-  (l4M52 maddr cr3val) $\sim$ (l4off maddr cr3val) $\mapsto_{a}$ {q1}  l4e  $\ast$
-  (l3M52 maddr l4e) $\sim$ (l3off maddr l4e)  $\mapsto_{a}$ {q2}  l3e $\ast$ 
-  (l2M52 maddr l3e) $\sim$ (l2off maddr l3e) $\mapsto_{a}$ {q3}  l2e $\ast$
-  (l1M52 maddr l2e) $\sim$ (l1off maddr l2e) $\mapsto_{a}$ {q4}  l1e $\ast$ $\ulcorner$ addr_L1(va,l1e) = paddr $\urcorner$.
+  (l4M52 maddr cr3val) $\sim$ (l4off maddr cr3val) $\mapsto_{a}$ {q1}  l4e  $\ast$(*\label{line:l4pointsto}*)
+  (l3M52 maddr l4e) $\sim$ (l3off maddr l4e)  $\mapsto_{a}$ {q2}  l3e $\ast$(*\label{line:l3pointsto}*) 
+  (l2M52 maddr l3e) $\sim$ (l2off maddr l3e) $\mapsto_{a}$ {q3}  l2e $\ast$(*\label{line:l2pointsto}*)
+  (l1M52 maddr l2e) $\sim$ (l1off maddr l2e) $\mapsto_{a}$ {q4}  l1e $\ast$ $\ulcorner$ addr_L1(va,l1e) = paddr $\urcorner$.(*\label{line:l1pointsto}*)
 \end{lstlisting}
@@
 
-\todo[inline]{This is a section where it's worth explaining the figure in a little more detail, especially as we reuse L4L1PointsTo
-in our own definitions later. I know we discussed at some point, but don't the \texttt{l$n$M52} accessors ignore the \texttt{maddr},
-and don't the \texttt{l$n$off} accessors ignore the entry arguments? If so we can drop those (simplifying the figure), and annotate 
-(e.g., with comments like \lstinline[mathescape]|(*$\textcircled{E}$*)|) each \texttt{l$n$off} in Figure \ref{fig:strongvirtualpointsto}
-with a tag matching a new tag on the corresponding offset chunk of the translated virtual addres in Figure \ref{fig:pagetables}\ldots
-}
-\todo[inline]{Ismail Addressed: current explanation in here is fortified by the changed text in the background section smoothly}
-
 \lstinline|L4_L1_PointsTo| works by
@@
   uses the masks \textsf{l4M52} with the current \lstinline|cr3| to get the physical address of the start of the L4 table
-  and \textsf{l4off} with the virtual address being translated to compute the correct offset within that table.
-  Then \del{at} that physical location \replace{is}{holds} the appropriate entry in the L4 table \textsf{l4e}.
-  Subsequent levels of the page table walk work similarly.
+  and \textsf{l4off} with the virtual address being translated to compute the correct \add{byte} offset within that table \add{just as in the first translation
+  step of Figure \ref{fig:pagetables}}.\footnote{\add{Note the offsets mentioned in Figure \ref{fig:pagetables} are 9-bit indexes into the 512 entries; the byte offset is that times 8.}}
+  \replace{
+    Then {at} that physical location {is} the appropriate entry in the L4 table \textsf{l4e}.
+  }{
+    Thus Line \ref{line:l4pointsto} asserts that the physical address built from the table base and offset points to the L4 entry \textsf{l4e}.
+  }
+  Subsequent levels of the page table walk \add{assertion (Lines \ref{line:l3pointsto}--\ref{line:l1pointsto})} work similarly.
 The statement of these assertions is simplified by the use of our split physical points-to assertions, since
@@
 }
-Our solution \replace{removes}{in the next subsection improves on theirs by removing} some restrictions in this definition by further abstracting the handling of address translation.
+Our solution \replace{removes}{in the next subsection improves on theirs, removing} some restrictions in this definition by further abstracting the handling of address translation.
+\looseness=-1
 
@@
   \textsc{GhostMapUpdate} says that ownership of the actual ghost map with ghost name $\gamma$ and map contents $\theta$,
-  and a full (non-fractional) token witnessing that $\theta$ maps \textsf{pa} to \textsf{va} permits an update to the ghost map's state,
+  and a \del{full (non-fractional) }token witnessing that $\theta$ maps \textsf{pa} to \textsf{va} permits an update to the ghost map's state,
   changing the map and replacing the token to represent the new value.
@@
   % \[\mathcal{A}\textsf{bsPTableWalk} \stackrel{\triangle}{=} \textsc{Auth} (\; \mathcal{W}_{64} \;\rightarrow_{\textrm{fin}} \;  ( (\textsc{Frac }, \mathord{+}) \times (\textsc{Agree } \Loc,\mathord{=}) ))\]
-For clarity, we refer to the specific ghost map summarizing virtual-to-physical translations by \\
-\centerline{$\mathcal{A}\textsf{bsPTableWalk}(\delta,\theta) \stackrel{\triangle}{=} \mathsf{GhostMap}(\delta,\theta)$}\\
+For clarity, we refer to the specific ghost map summarizing virtual-to-physical translations by 
+\mbox{$\mathcal{A}\textsf{bsPTableWalk}(\delta,\theta) \stackrel{\triangle}{=} \mathsf{GhostMap}(\delta,\theta)$}
 (omitting $\delta$ for brevity when only one is in scope)
@@
 yielding Figure \ref{fig:virtualpointstosharing}'s definition.
-This new definition guarantees that the ghost map contains an mapping from the virtual address ($\vaddr$) to a physical address ($\paddr$),
-and thus that the per-address-space invariant will contain the physical resources that guarantee \add{that} the hardware resolves the translation.
+This new definition guarantees that the ghost map \add{$\theta$} \replace{contains an mapping from}{maps} the virtual address ($\vaddr$) to a physical address ($\paddr$),
+and thus that the per-address-space invariant \add{described next} will contain the physical resources that guarantee \add{that} the hardware resolves the translation.
+\looseness=-1
 
@@
   \footnotesize
+\vspace{-1em}
 \centerline{$
@@
 
-We place the authorative ownership of the ghost translation $\mathcal{A}\textsf{PTableWalk}$ in a per-address-space invariant
+We place the authorative ownership of the ghost \add{map} translation $\mathcal{A}\textsf{PTableWalk}$ in a per-address-space invariant
 $\mathcal{I}$\textsf{ASpace} (Figure \ref{fig:peraspaceinvariant}). 
@@
 
-\todo[inline]{This next paragraph below explains details, but should first explain the big picture: the current address space
-is identified by a paddr, so to state the invariant for the address space named by that paddr we need to look up
-what invariants should hold, then assert that those invariants do hold.}
+\todo[inline,color=cyan]{Explain $\delta{s}$ (Iris ghost name) vs $m$ (logical contents of ghost map) in next 2 paragraphs}
+% \todo[inline]{This next paragraph below explains details, but should first explain the big picture: the current address space
+% is identified by a paddr, so to state the invariant for the address space named by that paddr we need to look up
+% what invariants should hold, then assert that those invariants do hold.}
+\add{One remaining difficulty is ensuring that different address spaces can have independent ghost maps ---
+which we resolve with an additional global ghost map from address-space identifiers (page table roots) to
+the \textsc{Iris} ghost name for that address space's ghost map.}
 In Figure \ref{fig:peraspaceinvariant},
@@
 
-\todo[inline]{What's not clear in this next paragraph is the relationship between ASpaceLookup in Figure 5 and the lookup part of Figure 3.
-I think Figure 5's $m$ and Figure 3's $\delta{s}$ are the same, aren't they? If so, we should say so and use the same name.
-If not we should explain why they're different.
-}
 Virtual points-to assertions must also ensure that they are tied to a valid address space, and because they
@@
 \footnotesize
-\begin{mathpar}
+\[\begin{array}{c}
     \mbox{$[r](\mathsf{P}) \;: \mathsf{vProp} \; \Sigma \; \stackrel{\triangle}{=} \; \lambda \_,\; \mathsf{P}~\mathsf{r} \vspace{0.5em}$}
-    \and
+    \qquad
     \mbox{$\textsf{Fact P} \stackrel{\triangle}{=} \;\forall \; ,r \; r' \ldotp  \textsf{P r} \dashv\vdash \textsf{P r'}$}
-    \and
+    \qquad
     \inferrule*{ }{\textsf{Fact}\;[r](\textsf{P}) }
-    \and
+    \qquad
     \inferrule*{ }{\textsf{Fact}\;(\textsf{r}\;\mapsto_{\textsf{r}}\{q\}\;\textsf{rv})}
-    \and
+    \\[1.3em]
     \inferrule{ }{\textsf{Fact}\;(\textsf{w} \mapsto_{\textsf{p}} \{q\} \textsf{ v} )}
-    \and
+    \qquad
     \inferrule{ }{(\textsf{P} \vdash \textsf{Q}) \vdash  ([r](\textsf{P}) \vdash  [r](\textsf{Q}))}
-    \and
+    \qquad
     \inferrule{ }{[r](\textsf{P} \ast \textsf{Q}) \dashv\vdash ([r](\textsf{P}) \ast [r](\textsf{Q})}\vspace{0.5em}
-    \and
+    \\[1.3em]
     \inferrule{ }{[r](\textsf{P}\land \textsf{Q}) \dashv\vdash [r](\textsf{P})\land [r](\textsf{Q})}
-    \and
+    \qquad
     \inferrule{ }{[r](\textsf{P}\lor \textsf{Q}) \dashv\vdash [r](\textsf{P})\lor [r](\textsf{Q})}
-    \and
+    \qquad
     \inferrule{ }{\textsf{Fact P} \vdash  [r](\textsf{P}) \dashv\vdash (\textsf{P})} %(\textsf{vProp }\Sigma))\\
-\end{mathpar}
+\end{array}\]
 \vspace{-1em}

### 2025-07-28 programlogic.tex
more cleaning and compressing of section 4

@@
 The fraction $\qfrac$ with value 1 asserts the unique ownership of the register mapping and grants update permission \replace{on}{to} it;
-otherwise, any value $0 < \qfrac <1$ represents partial ownership, granting read-only permission on the mapping.
+otherwise, any value $0 < \qfrac <1$ represents partial ownership, granting read-only permission on the mapping.\footnote{
+\add{We adopt the standard naming convention of $\qfrac$-related names representing fractional permission, with fractions
+sometimes appearing in braces or as subscripts in various asertions.}}
 
@@
 
-\todo[inline,color=cyan]{Explain $\delta{s}$ (Iris ghost name) vs $m$ (logical contents of ghost map) in next 2 paragraphs}
+% \todo[inline,color=cyan]{Explain $\delta{s}$ (Iris ghost name) vs $m$ (logical contents of ghost map) in next 2 paragraphs}
 % \todo[inline]{This next paragraph below explains details, but should first explain the big picture: the current address space
@@
 % what invariants should hold, then assert that those invariants do hold.}
-\add{One remaining difficulty is ensuring that different address spaces can have independent ghost maps ---
-which we resolve with an additional global ghost map from address-space identifiers (page table roots) to
-the \textsc{Iris} ghost name for that address space's ghost map.}
-In Figure \ref{fig:peraspaceinvariant},
+\add{We mus also ensure that different address spaces can have independent ghost maps ---
+which we resolve with an additional unique global ghost map (with ghost name $\delta{s}$) from address-space identifiers (page table roots
+whose values are manipulated by the kernel code) to
+the \textsc{Iris} ghost name for that address space's ghost map.
+In Figure \ref{fig:virtualpointstosharing}, the extra ghost map token for $\delta{s}$ asserts that $\delta$ --- which is exisentially
+quantified --- is the correct ghost name for the current address space. That is then the ghost map named
+in the ghost virtual-to-physical translation token of Figure \ref{fig:virtualpointstosharing}.
+}
+\del{In Figure \ref{fig:peraspaceinvariant},
 $\theta$ is the logical map from virtual addresses to the physical addresses they should translate to, 
@@
 from various page table roots to ghost names for the per-address-space ghost maps discussed above.
-$\mathcal{I}\textsf{ASpace}(\theta,m)$ then stores the physical resources for the current address space mappings,
-ensuring that the current address space is valid (has an entry in $m$ whose corresponding
+}
+\add{
+Just as the ghost name $\delta$ names the ghost map with contents $\theta$,
+$\delta{s}$ names a ghost map, whose contents appear as $m$ in Figure \ref{fig:peraspaceinvariant} (the association of $\delta{s}$ to $m$
+is a global invariant not shown).
+}
+$\mathcal{I}\textsf{ASpace}(\theta,m)$ then \add{performs 3 roles:
+it associates the current address space's root with an appropriate \textsc{Iris} ghost name $\delta$;
+it tracks authoritatively that $\delta$'s logical contents match $\theta$; and it}
+stores the physical resources for the current address space mappings \add{(via the iterated $\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$)}.
+\del{ensuring that the current address space is valid (has an entry in $m$ whose corresponding
 authoritative map matches $\theta$), and asserting ownership of the appropriate fractional ownership
-of the physical page table walk for each virtual-to-physical translation in $\Theta$ (for the current address space).
+of the physical page table walk for each virtual-to-physical translation in $\Theta$ (for the current address space).}
 
+\del{
 Virtual points-to assertions must also ensure that they are tied to a valid address space, and because they
@@
 ghost map $\delta$ mapping virtual addresses to physical addresses, and ensures it exists with a token for the distinguished
-global (unique) ghost map named by $\delta{s}$.
+global (unique) ghost map named by $\delta{s}$.}
 % the ghost map of virtual-to-physical translations for a particular address space (the ghost name $\delta$ in Figure \ref{fig:virtualpointstosharing}).
@@
 and that memory or registers are updated appropriately, before re-packing the invariant and virtual points-to resources.
+\looseness=-1
 
 \replace{T}{Second, t}he memory access rules --- as with all rules in our logic ---
-model updates to the instruction pointer \lstinline|rip| \add{based on the length of the encoded instruction.
-These rules} \replace{using}{use} a metafunction \textsf{MovLen}.
+\replace{model updates to}{increment} the instruction pointer \lstinline|rip| \add{by the length of the encoded instruction.}
+\del{using a metafunction \textsf{MovLen}.}
 \textsf{MovLen} returns how long the instruction encoding for the corresponding \lstinline|mov| is;
@@
 x86 ISA that preceded x86-64.
+\looseness=-1
 
@@
 the current address space.
-\todo[inline,color=green]{WIP: Colin is trying to clarify the value of the modal approach here}
 }
-\todo[inline,color=cyan]{Colin says:
-Ismail, Figure 7 looks broken. Why are there two IAspaces in WriteToRegFromVirtMem? Shouldn't there
-only be one?
-Also in WriteToRegFromVirtMem, why can't the virtual points-to be fractional? It is in the Rocq development,
-and we use it as such in examples.
-In WriteToVirtMemFromReg, vaddr points to a value rvs both before and after the store.
-And also in WriteToVirtMemFromReg, why are all the fractions q? They should be allowed to be different fractions.
-(Also, explaining the fraction notation was one of the questions from reviewers.)
-}
-\todo[inline,color=cyan]{Ismail Addressed:
-Others issues are addressed.
-"(Also, explaining the fraction notation was one of the questions from reviewers.)"
-They are explained earlier in the text. No need. They should go back and read.
-}
+
 \subsubsection{Updating \lstinline|cr3|} 
@@
 It is often said that the frame rule (below) is one of the key pieces of separation logic.
-\[
+\centerline{$
   \mbox{$\inferrule*[right=Frame]{
@@
   }$}
-  \qquad\textrm{(unsound with address space changes)}
-\]
+  \qquad\begin{array}{c}\textrm{(unsound with address space changes)}\\\\\end{array}
+$}\\
 Such a rule is normally recoverable from Hoare doubles (see, e.g., \citet{Chlipala2011Bedrock,Chlipala2013Bedrock}).
 However, in the presence of address space changes, the traditional frame rule is unsound.
-Consider:
-\begin{mathpar}
+Consider:\\
+\centerline{$
   \inferrule*[right=Frame]{
@@
   }
-\end{mathpar}
+$}\\
 In this (broken!) hypothetical example,
-both the precondition and the postcondition assert that $a\mapsto_\mathsf{v} x$ in the current address space, but we have no basis for concluding that address translation is preserved by the change of address space.
-So, this derivation clearly leads to an unsound conclusion.
+both the precondition and the postcondition assert that $a\mapsto_\mathsf{v} x$ in the current address space, but
+\replace{we have no basis for concluding that address translation is preserved by the change of address space}{
+  the new address space may map $a$ to another value}. So, this derivation clearly leads to an unsound conclusion. 
 The essential problem is that the frame rule is motivated by local reasoning about local updates, but

### 2025-07-28 examples_C.tex
remove a couple stale TODOs

@@
 similar to other assembly-level verification thanks to Section \ref{sec:logic}'s logic (including virtual pte-points-to
-assertions),\todo{figure out how to highlight advance over prior work here}
+assertions),
 but only after resolving two critical challenges.
@@
 Line \ref{line:p2vC} is the critical piece, and arguably corresponds to the most subtle verification of an \lstinline|add| instruction
-(\lstinline|add rax, KERNBASE|)\todo{if walkpgdir returns vaddr, then KERNBASE==0, which complicates some discussion here}
+(\lstinline|add rax, KERNBASE|)
 that we are aware of, and something no prior work on verified OS kernels has dealt with.

### 2025-07-28 examples_C.tex
remove stale TODOs, add appendix and TODO after live discussion

@@
 }
+\todo[inline,color=cyan]{
+  reading carefully, it says that if entry\_present of pte\_initialize of something holds (which it always will)
+  then the 512 entries are there. So this is basically ``True implies'', but in a shape that's compatible
+  with the other branch/path's invariants
+}
 \vspace{-1em}
@@
 First, the update to the memory at (virtual) address \lstinline|entry| depends on subtle ownership invariants:
-if the entry is present then its fractional ownership is shared with a large number of \lstinline|L4_L1_PointsTo| assertions
+if the entry is present then its fractional ownership is shared with a large number of $\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$ assertions
 from the address space invariant (Figure \ref{fig:peraspaceinvariant}),
@@
 \end{lstlisting}
-\todo[inline]{Ismail proof-read this figure for Colin's changes}
 \else
@@
 \fi
+\vspace{-1em}
 \caption{Walking page-table directory via calls to \textsf{pte\_get\_next\_table} in Figure \ref{fig:calltopteinitializeC}}
@@
     \\ \left(\begin{array}{l}  (\textsf{l1p+KERNBASE*8*L1Offset(va)}) \mapsto_{\textsf{vpte,qfrac1}} \textsf{l1p+8*L1Offset(va)}\;\textsf{l1e\_val} \ast  \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l1e\_val})\urcorner \end{array}\right) =\textsf{R}_{\textsf{l1e}}\end{array} $
-\todo[inline]{Ismail: Why isn't there an l1 entry identity map token in Rwalk?}
 \vspace{-1em}
@@
 \end{lstlisting}
-\vspace{-1em}
-\todo[inline,color=cyan]{will switch to l4p+ l4offset of va, so caller has the 512}
+\vspace{-2em}
   \caption{Specification of updating L1 entry to reference a new page (\textsf{fpaddr}).}
@@
 
-\todo[inline]{Colin will fix up next paragraph: it's right but also kinda weird, will handle w/ related work pass}
-Unlike the only prior work verifying analogous code for mapping a new page~\cite{kolanski08vstte,kolanski09tphols}, our proof above
-does \emph{not} need to reason directly over the operational semantics,
-making this the first verification we know of for mapping a virtual memory page that
-stays entirely at the program logic level.
-\looseness=-1
 % By incorporating verification of the

### 2025-07-28 relwork.tex
checking in WIP on relwork

@@
 There has been relatively little prior work on formal verification of virtual memory.
-Instead, most OS verification work has focused on minimizing reasoning about virtual memory management.
+\replace{Instead, m}Most OS verification work \replace{has focused on minimizing}{minimizes} reasoning about virtual memory management.
 The original \textsc{Verisoft} project~\cite{alkassar2008verisoft,alkassar2010pervasive,alkassar2008formal,dalinger2005verification,hillebrand2005address,alkassar2008formal,starostin2010formal} 
-relied on custom hardware, which, among other things, always ran kernel code with virtual memory disabled, removing the circularity that is a key challenge of verifying
+relied on custom hardware\del{,} which\del{, among other things,} always ran kernel code with virtual memory disabled, removing the circularity that is a key challenge of verifying
 VMM code for real hardware: \replace{A}{a}t that point page tables become a basic partial map data structure to represent user program address translations,
-with an idiosyncratic format. \replace{It turns out that s}{S}ubsequent OS verification work also treats page tables this way, but
-unsoundly given that \replace{other projects}{their} target hardware \del{that} \emph{does} run the kernel with address translation: they \emph{trust}
+with an idiosyncratic format. \replace{It turns out that s}{S}ubsequent OS verification work
+\add{targets real hardware that uses virtual addressing in the kernel, but unsoundly
+uses hardware models that do not.} \del{also treats page tables this way, but
+unsoundly given that \replace{other projects}{their} target hardware \del{that} \emph{does} run the kernel with address translation:}
+\add{Thus} they \emph{trust}
 that the particular page table manipulations do not, for example, unmap kernel code\footnote{\add{A ``fun'' bug one author encountered in
@@
 which is not true of page table updates.
- % which is not the case when a single page table write can make regions of memory inaccessible.
-% \textsc{seL4}~\cite{Klein2009seL4,seL4TOCS,Sewell2013translation} is a formally verified L4 microkernel~\cite{Liedtke1995,Liedtke1996} 
-% (and the first verified OS kernel to run on real-world hardware), verified with a mix of refinement proofs and program logic reasoning down to the assembly level.
-% Because \textsc{seL4} is a microkernel, most VMM functionality actually lives in usermode and is unverified, and moreover, 
-% their hardware model omits address translation entirely and the MMU entirely~\cite{Klein2009seL4,seL4TOCS}. 
-% As a result, the limited page table management present in the microkernel treats page tables as idiosyncratic tree-maps as in
-% \textsc{Verisoft}, despite actually running with address translation.
-% This is partly mitigated by manually identifying some trusted invariants (e.g., that the address range designated for the kernel is appropriately mapped)
-% and setting up the proof to ensure those invariants are maintained (i.e., as an extra proof obligation not required by their hardware model).
-
-% \textsc{CertiKOS}~\cite{gu15,gu2016certikos,gu2018certikos,chen2016interrupts} is a microkernel intended for use as a hypervisor.
-% The overall approach in that body of work is many layers of refinement proofs, using a
-%  proliferation of layers with small differences to keep most individual refinements tractable. In keeping with precursor work 
-% on the project from the same group~\cite{vaynberg2012compositional}, the purpose of some layers is to abstract away from 
-% virtual memory (as early as possible).
-% The papers on \textsc{CertiKOS} do not explicitly detail the VMM beyond highlighting its existence and referencing
-% that it performs mapping operations for user code.
-% The work is clear, however, that it fully trusts low-level assembly fragments such as the instruction sequence which actually
-% switches address spaces, rather than verifying them.
-% We can also infer other limitations from the project's pervasive reliance on a modified \textsc{CompCert},
-% which fixes an abstract memory model aimed at \emph{user-mode} C programs~\cite{leroy2008formal,leroy2009formally}. 
-% \textsc{CompCert}'s model of memory cannot handle non-local updates, so we can conclude \textsc{CertiKOS}'s proofs cannot
-% address impact of virtual memory mapping changes on the kernel itself.
 Other work on OS verification either never progressed far enough to address VMM verification
@@
 
+\add{
+ These limitations motivate work like ours on reasoning soundly about virtual memory management.
+ As discussed earlier, 
+}
 \citet{kolanski08vstte} are the only other researchers to study VMM verification against a realistic hardware
@@
  did not tackle evaluations that would benefit from modality.
- We have verified significantly more VMM code than Kolanski and Klein, which pushed us towards the modal approach.
- They verify~\cite{kolanski08vstte} roughly the (ARM) assembly corresponding to Lines \ref{line:l1entry_storeC} and \ref{line:l1entry_setpresent} of Figure
- \ref{fig:mapping_codeC},
- plus pulling the zeroed page to install in the L1 entry from a simple linked list.
- They never addressed code equivalent to our \lstinline|walkpgdir| (Figure \ref{walkpgdirC}), or \lstinline|pte_get_next_table| (Figure \ref{fig:calltopteinitializeC}),
- which are the most challenging parts of verifying a full page table walk, encompasing subtle sharing and ownership
- specifications and delicate kernel invariants involved in converting physical to virtual addresses. Ours is the only work to date
- to tackle verification of the \emph{entirety} of a software page table walk and mapping a new page.
 }
+\looseness=-1
+
+\add{
+ Our work improves significantly on the technical capabilities of ther logic and
+ evaluates on kernel code that is more complete and more challenging than theirs.
+ Our modal approach makes it possible to specify address space changes cleanly, which their logic cannot do at all.
+ Our use of virtual pte-points-to assertions enables nearly the same proof rules\todo{if this stays, we should typeset one of the vpte-pointsto rules to justify this}
+ as standard memory accesses, and constructing virtual points-to information within the logic
+ (c.f.\ the logical entailment between $\textsf{R}_\textsf{walk}$ and $\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$)
+ rather than reasoning semantically about when the model state supports new virtual points-to assertions.
+}
+
+\add{
+While Kolanski and Klein verify the critical step of updating an already-located L1 entry to map a new page
+(ARM assembly corresponding roughly to Lines \ref{line:l1entry_storeC} and \ref{line:l1entry_set_present}
+in our Figure \ref{fig:mapping_codeC}), they ignore the essential code
+preceding that step, which as our \lstinline|walkpgdir| and \lstinline|pte_get_next_table| verifications
+demonstrate, side-steps a significant amount of complexity and critical reasoning tasks.
+We have verified the entirety of the software page table walk up to mapping a new page, aside from a trusted physical memory allocator.\footnote{
+ \add{Which works quite similarly to verified \lstinline|malloc|~\cite{Chlipala2013Bedrock,wickerson2010explicit}}
+}
+ As a consequence of tackling this larger verification challenge, we are also the first to formally specify large
+ segments of the self-referential portion of an OS kernel's virtual memory management invariants (per Sections \ref{sec:p2vC} and \ref{sec:selfconditional}),
+ and to reason about converting back from physical addresses to virtual addresses efficiently.
+}
+
 Due to a lack of fractional permissions in their formalism, they incidentally pick up other limitations orthogonal to their

### 2025-07-28 examples.tex
remove stale TODOs, add appendix and TODO after live discussion
integrate the assembly appendix by writing a brief new intro

@@
-\section{Experiments with Assembly Language}
-\label{sec:experiment}
+\section{\add{Assembly-level Verification for Page Table Traversal and Mapping}}
+\label{sec:experiment_appendix}
 %To both validate and demonstrate the value of the modal approach to reasoning about virtual memory management, 
@@
 % Recall from Section \ref{sec:logic} that virtual points-to assertions work just like regular points-to assertions, by design.
+
+\replace{
 In this section we verify several critical and challenging pieces of VMM code.
@@
 the first such verification handling both old and new processes' assertions (in different address spaces) at the time of the switch.
+}{
+While our logic was developed and proven sound for x86-64 assembly,
+Section \ref{sec:traversingC} described verification of software page table walking code (\lstinline|pte_get_next_table| and \lstinline|walkpgdir|)
+as if at the level of C for improved readability.
+This appendix describes the actual assembly-level verification carried out in Rocq.
+Careful readers of both Section \ref{sec:traversingC} and this appendix will notice
+strong similarities in the assertions and and reasoning, for good reason:
+The C code in Section \ref{sec:traversingC} was the original kernel code that was compiled
+(with no optimizations) to x86-64 assembly and verified with our logic, and the proof outlines
+in that section largely back-port the assembly proofs back to C.
+\looseness=-1
+}
+
+\add{
+ This section describes the assembly proofs without reference to the C outlines given in Section \ref{sec:traversingC}.
+ The main additional details of note at the assembly level are:
+ \begin{itemize}
+ \item Accurate treatment of register management (particularly the AMD64 System V calling convention) leading to more direct correspondence
+       with our logic
+ \item The assembly is naturally more verbose than the C, so the proof outlines are relatively more sparse, with assertions written
+       only for key updates.
+ \item Bitwise manipulations of page table entries are harder to follow than C's bitfield access support.
+       Multiple manipulations which are each explicit in C become adjacent (or sometimes non-adjacent) bitwise operations.
+       The critical ones are commented in the assembly figures.
+ \item And compared to the C-based presentation earlier, there are differences in logical variable names. For example,
+       the assembly proofs use \textsf{entry} as the name for the \emph{physical} address of the entry modified by
+       \lstinline|pte_get_next_table| in Figure \ref{fig:calltopteinitialize}, whereas to make sense of the C code
+       in Figure \ref{fig:calltopteinitializeC} we used \textsf{entry} consistently with the C variable name and introduced
+       separate logical names for physical addresses. This propagates to figures presenting larger invariants separately,
+       as they also refer to the logical names from the proofs.
+ \end{itemize}
+}
 
@@
 
+\del{
 A PTE points-to, as we already brifly mentioned, is defined just like the normal virtual points-to of Figure \ref{fig:virtualpointstosharing}, except the physical address (here, \textsf{pa}) is explicit in the assertion
@@
 the modified data is at the specific physical location that will be used as the L1 entry.
+}
 % After obtaining a virtual address \textsf{pte\_addr} in \textsf{rax} backed 
@@
 
-\begin{figure}\footnotesize
-\begin{lstlisting}[mathescape,escapeinside={(*}{*)}]
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,\Xi,m) \ast [\rtv'](\mathcal{I}\texttt{ASpace}(\theta',\Xi',m') \ast \texttt{Pother}) \ast \texttt{rsi}\mapsto_{\textsf{r}} new \ast \texttt{rdi}\mapsto_{\textsf{r}} old \ast \texttt{rbx}\mapsto_{\textsf{r}} \texttt{rbxv} }_{\rtv}$
-$\specline{\texttt{rsp}\mapsto_{\textsf{r}} \texttt{rspv} \ast \texttt{rbp}\mapsto_{\textsf{r}} \texttt{rbpv} \ast \texttt{r12}\mapsto_{\textsf{r}} \texttt{r12v} \ast \texttt{r13}\mapsto_{\textsf{r}} \texttt{r13v} \ast \texttt{r14}\mapsto_{\textsf{r}} \texttt{r14v} \ast \texttt{r15}\mapsto_{\textsf{r}} \texttt{r15v}}_{\rtv}$
-$\specline{\mathsf{ContextAt}(old,\_) \ast \mathsf{ContextAt}(new,[\mathsf{rbxv}',\ldots,\rtv']) }_{\rtv}$
-mov 0[rdi], rbx ... ;; also save rsp, rbp, r12, r13, r14, to offsets 8, 16, 24, 32, and 40 from rdi (*\label{line:start_save}*)
-mov 48[rdi], r15
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,\Xi,m) \ast [\rtv'](\mathcal{I}\texttt{ASpace}(\theta',\Xi',m') \ast \texttt{Pother})  \ast  \texttt{rsi}\mapsto_{\textsf{r}} new \ast \texttt{rdi}\mapsto_{\textsf{r}} old \ast \texttt{rbx}\mapsto_{\textsf{r}} \texttt{rbxv} }_{\rtv}$
-$\specline{ \texttt{rsp}\mapsto_{\textsf{r}} \texttt{rspv} \ast \texttt{rbp}\mapsto_{\textsf{r}} \texttt{rbpv} \ast \texttt{r12}\mapsto_{\textsf{r}} \texttt{r12v} \ast \texttt{r13}\mapsto_{\textsf{r}} \texttt{r13v} \ast \texttt{r14}\mapsto_{\textsf{r}} \texttt{r14v} \ast \texttt{r15}\mapsto_{\textsf{r}} \texttt{r15v}}_{\rtv}$
-$\specline{\mathsf{ContextAt}(old,[\mathsf{rbxv,\ldots,\rtv}]) \ast \mathsf{ContextAt}(new,[\mathsf{rbxv}',\ldots,\rtv']) }_{\rtv}$
-mov 56[%rdi], %cr3 $\specline{\ldots \ast \texttt{rdi+56} \mapsto_{\textsf{v}} \rtv \ast \ldots }_{\rtv}$   (*\label{line:end_save}*) 
-mov rbx, 0[rsi] (*\label{line:start_restore}*)
-mov rsp, 8[rsi] ;; Switch to new stack, which may not be mapped in the current address space!
-... ;; load rbp, r12, r13, r14, from offsets 16, 24, 32, and 40 from rsi
-mov r15, 48[rsi]
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,\Xi,m) \ast [\rtv'](\mathcal{I}\texttt{ASpace}(\theta',\Xi',m') \ast \texttt{Pother})  \ast  \texttt{rsi}\mapsto_{\textsf{r}} new \ast \texttt{rdi}\mapsto_{\textsf{r}} old \ast \texttt{rbx}\mapsto_{\textsf{r}} \texttt{rbxv}' }_{\rtv}$
-$\specline{\texttt{rsp}\mapsto_{\textsf{r}} \texttt{rspv}' \ast \texttt{rbp}\mapsto_{\textsf{r}} \texttt{rbpv}' \ast \texttt{r12}\mapsto_{\textsf{r}} \texttt{r12v}' \ast \texttt{r13}\mapsto_{\textsf{r}} \texttt{r13v}' \ast \texttt{r14}\mapsto_{\textsf{r}} \texttt{r14v}' \ast \texttt{r15}\mapsto_{\textsf{r}} \texttt{r15v}'}_{\rtv}$
-$\specline{\mathsf{ContextAt}(old,[\mathsf{rbxv,\ldots,\rtv}]) \ast \mathsf{ContextAt}(new,[\mathsf{rbxv}',\ldots,\rtv']) }_{\rtv}$
-mov cr3, 56[rsi] ;; <-- Switch to the new address space (*\label{line:end_restore}*)
-$\specline{ [\rtv](\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,\Xi,m) \ast \mathsf{ContextAt}(old,[\mathsf{rbxv,\ldots,\rtv}]) \ast \mathsf{ContextAt}(new,[\mathsf{rbxv}',\ldots,\rtv']))  }_{\rtv}$ (*\label{line:modality_switch}*)
-$\specline{ \mathcal{I}\texttt{ASpace}(\theta',\Xi',m') \ast \texttt{Pother} \ast  \texttt{rsi}\mapsto_{\textsf{r}} new \ast \texttt{rdi}\mapsto_{\textsf{r}} old \ast \ldots }_{\rtv'}$
-\end{lstlisting}
-\vspace{-1em}
-\caption{Basic task switch code that switches address spaces.}
-\label{fig:swtch}
-\end{figure}
+% \begin{figure}\footnotesize
+% \begin{lstlisting}[mathescape,escapeinside={(*}{*)}]
+% $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,\Xi,m) \ast [\rtv'](\mathcal{I}\texttt{ASpace}(\theta',\Xi',m') \ast \texttt{Pother}) \ast \texttt{rsi}\mapsto_{\textsf{r}} new \ast \texttt{rdi}\mapsto_{\textsf{r}} old \ast \texttt{rbx}\mapsto_{\textsf{r}} \texttt{rbxv} }_{\rtv}$
+% $\specline{\texttt{rsp}\mapsto_{\textsf{r}} \texttt{rspv} \ast \texttt{rbp}\mapsto_{\textsf{r}} \texttt{rbpv} \ast \texttt{r12}\mapsto_{\textsf{r}} \texttt{r12v} \ast \texttt{r13}\mapsto_{\textsf{r}} \texttt{r13v} \ast \texttt{r14}\mapsto_{\textsf{r}} \texttt{r14v} \ast \texttt{r15}\mapsto_{\textsf{r}} \texttt{r15v}}_{\rtv}$
+% $\specline{\mathsf{ContextAt}(old,\_) \ast \mathsf{ContextAt}(new,[\mathsf{rbxv}',\ldots,\rtv']) }_{\rtv}$
+% mov 0[rdi], rbx ... ;; also save rsp, rbp, r12, r13, r14, to offsets 8, 16, 24, 32, and 40 from rdi (*\label{line:start_save}*)
+% mov 48[rdi], r15
+% $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,\Xi,m) \ast [\rtv'](\mathcal{I}\texttt{ASpace}(\theta',\Xi',m') \ast \texttt{Pother})  \ast  \texttt{rsi}\mapsto_{\textsf{r}} new \ast \texttt{rdi}\mapsto_{\textsf{r}} old \ast \texttt{rbx}\mapsto_{\textsf{r}} \texttt{rbxv} }_{\rtv}$
+% $\specline{ \texttt{rsp}\mapsto_{\textsf{r}} \texttt{rspv} \ast \texttt{rbp}\mapsto_{\textsf{r}} \texttt{rbpv} \ast \texttt{r12}\mapsto_{\textsf{r}} \texttt{r12v} \ast \texttt{r13}\mapsto_{\textsf{r}} \texttt{r13v} \ast \texttt{r14}\mapsto_{\textsf{r}} \texttt{r14v} \ast \texttt{r15}\mapsto_{\textsf{r}} \texttt{r15v}}_{\rtv}$
+% $\specline{\mathsf{ContextAt}(old,[\mathsf{rbxv,\ldots,\rtv}]) \ast \mathsf{ContextAt}(new,[\mathsf{rbxv}',\ldots,\rtv']) }_{\rtv}$
+% mov 56[%rdi], %cr3 $\specline{\ldots \ast \texttt{rdi+56} \mapsto_{\textsf{v}} \rtv \ast \ldots }_{\rtv}$   (*\label{line:end_save}*) 
+% mov rbx, 0[rsi] (*\label{line:start_restore}*)
+% mov rsp, 8[rsi] ;; Switch to new stack, which may not be mapped in the current address space!
+% ... ;; load rbp, r12, r13, r14, from offsets 16, 24, 32, and 40 from rsi
+% mov r15, 48[rsi]
+% $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,\Xi,m) \ast [\rtv'](\mathcal{I}\texttt{ASpace}(\theta',\Xi',m') \ast \texttt{Pother})  \ast  \texttt{rsi}\mapsto_{\textsf{r}} new \ast \texttt{rdi}\mapsto_{\textsf{r}} old \ast \texttt{rbx}\mapsto_{\textsf{r}} \texttt{rbxv}' }_{\rtv}$
+% $\specline{\texttt{rsp}\mapsto_{\textsf{r}} \texttt{rspv}' \ast \texttt{rbp}\mapsto_{\textsf{r}} \texttt{rbpv}' \ast \texttt{r12}\mapsto_{\textsf{r}} \texttt{r12v}' \ast \texttt{r13}\mapsto_{\textsf{r}} \texttt{r13v}' \ast \texttt{r14}\mapsto_{\textsf{r}} \texttt{r14v}' \ast \texttt{r15}\mapsto_{\textsf{r}} \texttt{r15v}'}_{\rtv}$
+% $\specline{\mathsf{ContextAt}(old,[\mathsf{rbxv,\ldots,\rtv}]) \ast \mathsf{ContextAt}(new,[\mathsf{rbxv}',\ldots,\rtv']) }_{\rtv}$
+% mov cr3, 56[rsi] ;; <-- Switch to the new address space (*\label{line:end_restore}*)
+% $\specline{ [\rtv](\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,\Xi,m) \ast \mathsf{ContextAt}(old,[\mathsf{rbxv,\ldots,\rtv}]) \ast \mathsf{ContextAt}(new,[\mathsf{rbxv}',\ldots,\rtv']))  }_{\rtv}$ (*\label{line:modality_switch}*)
+% $\specline{ \mathcal{I}\texttt{ASpace}(\theta',\Xi',m') \ast \texttt{Pother} \ast  \texttt{rsi}\mapsto_{\textsf{r}} new \ast \texttt{rdi}\mapsto_{\textsf{r}} old \ast \ldots }_{\rtv'}$
+% \end{lstlisting}
+% \vspace{-1em}
+% \caption{Basic task switch code that switches address spaces.}
+% \label{fig:swtch}
+% \end{figure}
 
-\subsection{Change of Address Space}
-A critical piece of \emph{trusted} code in current verified OS kernels is the assembly code to change the current address space; current verified OS kernels currently 
-lack effective ways to specify and reason about this low-level operation, for reasons outlined in Section \ref{sec:relwork}.
+% \subsection{Change of Address Space}
+% A critical piece of \emph{trusted} code in current verified OS kernels is the assembly code to change the current address space; current verified OS kernels currently 
+% lack effective ways to specify and reason about this low-level operation, for reasons outlined in Section \ref{sec:relwork}.
 
-Figure \ref{fig:swtch} gives simplified code for a basic task switch, the heart of an OS scheduler implementation. This is code that saves the context (registers and stack)
-of the running thread (here in a structure pointed to by \lstinline|rdi|'s value shown in Lines \ref{line:start_save}--\ref{line:end_save} of Figure \ref{fig:swtch}) and restores the context of 
-an existing thread (from \lstinline|rsi| shown in abbreviated Lines \ref{line:start_restore}--\ref{line:end_restore}), including the corresponding change of address space for a target thread in another process.
-This code assumes the System V AMD64 ABI calling convention, where the normal registers not mentioned are caller-save, and therefore saved on the stack of the thread
-that calls this code, as well as on the new stack of the thread that is restored, thus only the callee-save registers and \texttt{cr3} must be 
-restored.\footnote{We are simplifying by ignoring non-integer registers (e.g., floating point, vector registers),
-and the caller-save registers should be initialized to 0 to avoid leaking information across processes, but this captures the key challenges.}
-With the addition of a return instruction, this code would satisfy the C function signature\footnote{This is the function in UNIX 6th Edition 
-with the infamous ``You are not expected to understand this'' comment~\cite{lions1996lions}.}
-\centerline{\lstinline[language=C]|void swtch(context_t* save, context_t* restore);|}\\
-A call to this code begins executing one thread (until just before Line \ref{line:end_save}) in one address space ($\rtv$), whose information will be saved in a structure at address $old$,
-and finishes execution executing a different thread in a different address space (Line \ref{line:end_restore} on) whose information is initially in $new$.
+% Figure \ref{fig:swtch} gives simplified code for a basic task switch, the heart of an OS scheduler implementation. This is code that saves the context (registers and stack)
+% of the running thread (here in a structure pointed to by \lstinline|rdi|'s value shown in Lines \ref{line:start_save}--\ref{line:end_save} of Figure \ref{fig:swtch}) and restores the context of 
+% an existing thread (from \lstinline|rsi| shown in abbreviated Lines \ref{line:start_restore}--\ref{line:end_restore}), including the corresponding change of address space for a target thread in another process.
+% This code assumes the System V AMD64 ABI calling convention, where the normal registers not mentioned are caller-save, and therefore saved on the stack of the thread
+% that calls this code, as well as on the new stack of the thread that is restored, thus only the callee-save registers and \texttt{cr3} must be 
+% restored.\footnote{We are simplifying by ignoring non-integer registers (e.g., floating point, vector registers),
+% and the caller-save registers should be initialized to 0 to avoid leaking information across processes, but this captures the key challenges.}
+% With the addition of a return instruction, this code would satisfy the C function signature\footnote{This is the function in UNIX 6th Edition 
+% with the infamous ``You are not expected to understand this'' comment~\cite{lions1996lions}.}
+% \centerline{\lstinline[language=C]|void swtch(context_t* save, context_t* restore);|}\\
+% A call to this code begins executing one thread (until just before Line \ref{line:end_save}) in one address space ($\rtv$), whose information will be saved in a structure at address $old$,
+% and finishes execution executing a different thread in a different address space (Line \ref{line:end_restore} on) whose information is initially in $new$.
 
-Because this code does not directly update the instruction pointer, it is worth explaining \emph{how} this switches threads: by switching address spaces and stacks. 
-This is meant to be called with a return address for the current thread stored on the current stack when called. 
-The precondition of the return address on the initial stack requires the callee-save register values at the time of the call: those stored in the first 
-half of the code.
-Likewise, part of the invariant of the stack of the second thread, the one being restored, is that the return address on \emph{that} stack requires the saved 
-callee-save registers stored in that context to be in registers as its precondition.
+% Because this code does not directly update the instruction pointer, it is worth explaining \emph{how} this switches threads: by switching address spaces and stacks. 
+% This is meant to be called with a return address for the current thread stored on the current stack when called. 
+% The precondition of the return address on the initial stack requires the callee-save register values at the time of the call: those stored in the first 
+% half of the code.
+% Likewise, part of the invariant of the stack of the second thread, the one being restored, is that the return address on \emph{that} stack requires the saved 
+% callee-save registers stored in that context to be in registers as its precondition.
 
-The wrinkle, and the importance of the modal treatment of assertions, is that the target thread's precondition is \emph{relative to its address space}, 
-not the address space of the calling thread, which is reflected by
-the other-space modality 
-$[\rtv'](\mathcal{I}\texttt{ASpace}(\theta,\Xi,m) \ast \texttt{Pother})$
-in the specfication. 
-The precondition of this code,
-in context, would include that the initial stack pointer (before \lstinline|rsp| is updated)
-has a return address expecting the then-current callee-save register values and 
-suitably updated (i.e., post-return) stack in the \emph{current} (initial) address space;
-this would be part of \textsf{P} in the precondition.
-The specification also requires that
-the stack pointer saved in the context to restore expects the same of the saved registers and stack 
-\emph{in the other address space}. 
-The other-space modality plays a critical role here; \textsf{Pother} would contain these assumptions in the other
-address space.
-\looseness=-1
+% The wrinkle, and the importance of the modal treatment of assertions, is that the target thread's precondition is \emph{relative to its address space}, 
+% not the address space of the calling thread, which is reflected by
+% the other-space modality 
+% $[\rtv'](\mathcal{I}\texttt{ASpace}(\theta,\Xi,m) \ast \texttt{Pother})$
+% in the specfication. 
+% The precondition of this code,
+% in context, would include that the initial stack pointer (before \lstinline|rsp| is updated)
+% has a return address expecting the then-current callee-save register values and 
+% suitably updated (i.e., post-return) stack in the \emph{current} (initial) address space;
+% this would be part of \textsf{P} in the precondition.
+% The specification also requires that
+% the stack pointer saved in the context to restore expects the same of the saved registers and stack 
+% \emph{in the other address space}. 
+% The other-space modality plays a critical role here; \textsf{Pother} would contain these assumptions in the other
+% address space.
+% \looseness=-1
 
-% Lines 10--16 save the current context into memory (in the current address space).
-% Line 22 saves the initial page table root.
-% Lines 33--38 begin restoring the target context, including the stack pointer (line 33),
-% which may not be mapped in the address space at that time: it is the stack for the context being
-% loaded into the CPU.
-% The actual address switch occurs on line 45, which is verified with our modal rule for updating \lstinline|cr3|,
-% and thus shifts resources in and out of other-space modalities as appropriate.
+% % Lines 10--16 save the current context into memory (in the current address space).
+% % Line 22 saves the initial page table root.
+% % Lines 33--38 begin restoring the target context, including the stack pointer (line 33),
+% % which may not be mapped in the address space at that time: it is the stack for the context being
+% % loaded into the CPU.
+% % The actual address switch occurs on line 45, which is verified with our modal rule for updating \lstinline|cr3|,
+% % and thus shifts resources in and out of other-space modalities as appropriate.
 
-The postcondition is analagous to the precondition, but interpreted \emph{in the new address space}: the then-current (updated) stack would have a return address expecting the new (restored) register values (again, in \textsf{Pother}),
-and the saved context's invariant captures the precondition for restoring its execution \emph{in the previous address space} (as part of \textsf{P}).
+% The postcondition is analagous to the precondition, but interpreted \emph{in the new address space}: the then-current (updated) stack would have a return address expecting the new (restored) register values (again, in \textsf{Pother}),
+% and the saved context's invariant captures the precondition for restoring its execution \emph{in the previous address space} (as part of \textsf{P}).
 
-Immediately after the page table switch, assertions about the saved and restored contexts are
-guarded by a modality for the retiring
-address space \rtv{} (Line \ref{line:modality_switch}), per
-\textsc{WriteToRegCtlFromRegModal} (Figure \ref{fig:wpdamd}),
-because
-there is no guarantee that the data structures of the previous address space are mapped in the new address space.
-The ability to transfer that points-to information out of that modality is specific to a given kernel's design. 
-Kernels that map kernel memory into all address spaces would need invariants
-that justified moving those assertions out of the other-space modality.
-% Following Spectre and Meltdown, this kernel design became less prevalent because speculative execution of accesses to kernel addresses could leak information even if the access did eventually cause a fault (the user/kernel mode permission check was done after fetching data from memory).
+% Immediately after the page table switch, assertions about the saved and restored contexts are
+% guarded by a modality for the retiring
+% address space \rtv{} (Line \ref{line:modality_switch}), per
+% \textsc{WriteToRegCtlFromRegModal} (Figure \ref{fig:wpdamd}),
+% because
+% there is no guarantee that the data structures of the previous address space are mapped in the new address space.
+% The ability to transfer that points-to information out of that modality is specific to a given kernel's design. 
+% Kernels that map kernel memory into all address spaces would need invariants
+% that justified moving those assertions out of the other-space modality.
+% % Following Spectre and Meltdown, this kernel design became less prevalent because speculative execution of accesses to kernel addresses could leak information even if the access did eventually cause a fault (the user/kernel mode permission check was done after fetching data from memory).
 Thus many modern kernels have reverted to the older kernel design where the kernel inhabits its own unique address space, and user processes have only enough extra material mapped in their address spaces to switch into the kernel (CPUs do not speculate past updates to \texttt{cr3}).
-\looseness=-1
+% \looseness=-1
 
-While prior work has verified context switches within a single address space~\cite{ni2007contexts}, and context switches
-without any code before or after~\cite{syeda2020formal} (i.e., not reasoning about the impact of address space change
-on what data was accessible), this is the first verification handling both.
-\looseness=-1
+% While prior work has verified context switches within a single address space~\cite{ni2007contexts}, and context switches
+% without any code before or after~\cite{syeda2020formal} (i.e., not reasoning about the impact of address space change
+% on what data was accessible), this is the first verification handling both.
+% \looseness=-1
 
-\begin{comment}
-\[  
-$\specline{\exists (\entryf ,\;\entrytr,\; \entrytw,\; \entryo,\;\textsf{pte\_addr },\paddr) \; \ldotp\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast  \texttt{r14}\mapsto_{\textsf{r}} \_ \ast \texttt{rdi}\mapsto_{r} \vaddr \ast \texttt{rax}\mapsto_{\textsf{r}} \textsf{ pte\_addr} \; \ast }_{\rtv}$
-$\specline{ \ulcorner  \texttt{addr\_L1 }(\vaddr, \entryo) = \paddr \urcorner \ast \ulcorner \texttt{entry\_present } \entryf \land \texttt{entry\_present } \entrytr \land  \texttt{entry\_present } \entrytw \urcorner \; \ast}_{\rtv}$
-$\specline{\nfpointsto{\mask\vaddr\maskfour\rtv}{\mask\vaddr\maskfouroff\rtv}\entryf\qone\naddr \; \ast \nfpointsto{\mask\vaddr\maskthree\entryf}{\mask\vaddr\maskthreeoff\entryf}\entrytr\qtwo\naddr \ast}_{\rtv}$
-$\specline{  \nfpointsto{\mask\vaddr\masktwo\entrytr}{\mask\vaddr\masktwooff\entrytr}\paddr\qthree\entryo \;\ast \texttt{pte\_addr} \mapsto_{\texttt{vpte}} \paddr \;(\texttt{wzero 64}) \ast \texttt{rax}\mapsto_{\textsf{r}} \texttt{pte\_addr}  }_{\rtv}$
-mov r14, rax ;; Save that before another call
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast  \texttt{r14}\mapsto_{\textsf{r}} \texttt{pte\_addr} \ast \texttt{rdi}\mapsto_{\textsf{r}} \vaddr \ast \texttt{rax}\mapsto_{\textsf{r}} \textsf{ pte\_addr} \; \ast }_{\rtv}$
-$\specline{ \nfpointsto{\mask\vaddr\maskfour\rtv}{\mask\vaddr\maskfouroff\rtv}\entryf\qone\naddr \ast \ulcorner \texttt{entry\_present } \entryf \land \texttt{entry\_present } \entrytr \land  \texttt{entry\_present } \entrytw \urcorner \ast}_{\rtv}$
-$\specline{  \nfpointsto{\mask\vaddr\maskthree\entryf}{\mask\vaddr\maskthreeoff\entryf}\entrytr\qtwo\naddr \ast \nfpointsto{\mask\vaddr\masktwo\entrytr}{\mask\vaddr\masktwooff\entrytr}\paddr\qthree\entryo \;\ast}_{\rtv}$
-$\specline{\texttt{pte\_addr} \mapsto_{\texttt{vpte}} \paddr \;(\texttt{wzero 64}) \ast \texttt{rax}\mapsto_{\textsf{r}} \texttt{pte\_addr}  }_{\rtv}$
-call alloc_phys_page_or_panic
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast  \texttt{r14}\mapsto_{\textsf{r}} \texttt{pte\_addr} \ast \texttt{rdi}\mapsto_{\textsf{r}} \vaddr \;\ast \nfpointsto{\mask\vaddr\maskfour\rtv}{\mask\vaddr\maskfouroff\rtv}\entryf\qone\naddr \ast}_{\rtv}$
-$\specline{  \nfpointsto{\mask\vaddr\maskthree\entryf}{\mask\vaddr\maskthreeoff\entryf}\entrytr\qtwo\naddr \ast \nfpointsto{\mask\vaddr\masktwo\entrytr}{\mask\vaddr\masktwooff\entrytr}\paddr\qthree\naddr \;\ast}_{\rtv}$
-$\specline{\texttt{pte\_addr} \mapsto_{\texttt{vpte}} \paddr\; (\texttt{wzero 64})  \ast \ulcorner \texttt{entry\_present } \entryf \land \texttt{entry\_present } \entrytr \land  \texttt{entry\_present } \entrytw \urcorner}_{\rtv}$
-$\specline{\exists \texttt{ fpaddr} \ldotp \ulcorner \texttt{aligned fpaddr} \urcorner \ast \texttt{rax}\mapsto_{\textsf{r}} \texttt{fpaddr+3} \ast \texttt{fpaddr} \mapsto_{\textsf{p}} (\texttt{wzero 64}) \ast \ulcorner \texttt{entry\_present (fpaddr+3)}\urcorner}_{\rtv}$
-;; Calculate new L1 entry
-mov [r14], rax ;; store the page table entry, mapping the page
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast  \texttt{r14}\mapsto_{\textsf{r}} \texttt{pte\_addr} \ast \texttt{rdi}\mapsto_{\textsf{r}} \vaddr \;\ast \nfpointsto{\mask\vaddr\maskfour\rtv}{\mask\vaddr\maskfouroff\rtv}\entryf\qone\naddr \ast}_{\rtv}$
-$\specline{  \nfpointsto{\mask\vaddr\maskthree\entryf}{\mask\vaddr\maskthreeoff\entryf}\entrytr\qtwo\naddr \ast \nfpointsto{\mask\vaddr\masktwo\entrytr}{\mask\vaddr\masktwooff\entrytr}\paddr\qthree\entryo \;\ast}_{\rtv}$
-$\specline{\texttt{pte\_addr} \mapsto_{\texttt{vpte}} \paddr \;(\texttt{fpaddr+3}) \; \ast \ulcorner \texttt{entry\_present } \entryf \land \texttt{entry\_present } \entrytr \land  \texttt{entry\_present } \entrytw \urcorner }_{\rtv}$
-$\specline{\ulcorner \texttt{aligned fpaddr} \urcorner \ast \texttt{rax}\mapsto_{\textsf{r}} \texttt{fpaddr+3} \ast \texttt{fpaddr} \mapsto_{\textsf{p}} (\texttt{wzero 64}) \ast \ulcorner \texttt{entry\_present fpaddr+3}\urcorner}_{\rtv}$
-$\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\; \sqsubseteq $
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast  \texttt{r14}\mapsto_{\textsf{r}} \texttt{pte\_addr} \ast \texttt{rdi}\mapsto_{\textsf{r}} \vaddr \ast }_{\rtv}$
-$\specline{\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\vaddr,\entryf,\entrytr,\entrytw,\fpaddr+3) \ast \ulcorner \theta \;!!\;\vaddr = \texttt{None}\urcorner \; \ast}_{\rtv}$
-$\specline{\ulcorner \texttt{aligned fpaddr} \urcorner \ast \texttt{rax}\mapsto_{\textsf{r}} \texttt{fpaddr+3} \ast \texttt{fpaddr} \mapsto_{\textsf{p}} (\texttt{wzero 64}) }_{\rtv}$
-$\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\; \sqsubseteq $
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace} (<[\vaddr:=\texttt{fpaddr}]> \theta,m) \ast}_{\rtv}$
-$\specline{\ulcorner \texttt{aligned fpaddr} \urcorner \ast \texttt{fpaddr} \mapsto_{\textsf{p}} \textsf{ wzero 64} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  \ast\sumwalkabs\vaddr\qfrac\fpaddr}_{\rtv}$
-$\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\; \sqsubseteq $
-  $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace} (<[\vaddr:=\texttt{fpaddr}]> \theta,m) \ast \vaddr \mapsto_{\textsf{vpte}}\; \{\qfrac\} \;\fpaddr \textsf{ wzero 64}}_{\rtv}$
-$\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\; \sqsubseteq $
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace} (<[\vaddr:=\texttt{fpaddr}]> \theta,m) \ast \vaddr \mapsto_{\textsf{v}}\; \{\qfrac\} \textsf{wzero 64}}_{\rtv}$
-\end{comment}
+% \begin{comment}
+% \[  
+% $\specline{\exists (\entryf ,\;\entrytr,\; \entrytw,\; \entryo,\;\textsf{pte\_addr },\paddr) \; \ldotp\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast  \texttt{r14}\mapsto_{\textsf{r}} \_ \ast \texttt{rdi}\mapsto_{r} \vaddr \ast \texttt{rax}\mapsto_{\textsf{r}} \textsf{ pte\_addr} \; \ast }_{\rtv}$
+% $\specline{ \ulcorner  \texttt{addr\_L1 }(\vaddr, \entryo) = \paddr \urcorner \ast \ulcorner \texttt{entry\_present } \entryf \land \texttt{entry\_present } \entrytr \land  \texttt{entry\_present } \entrytw \urcorner \; \ast}_{\rtv}$
+% $\specline{\nfpointsto{\mask\vaddr\maskfour\rtv}{\mask\vaddr\maskfouroff\rtv}\entryf\qone\naddr \; \ast \nfpointsto{\mask\vaddr\maskthree\entryf}{\mask\vaddr\maskthreeoff\entryf}\entrytr\qtwo\naddr \ast}_{\rtv}$
+% $\specline{  \nfpointsto{\mask\vaddr\masktwo\entrytr}{\mask\vaddr\masktwooff\entrytr}\paddr\qthree\entryo \;\ast \texttt{pte\_addr} \mapsto_{\texttt{vpte}} \paddr \;(\texttt{wzero 64}) \ast \texttt{rax}\mapsto_{\textsf{r}} \texttt{pte\_addr}  }_{\rtv}$
+% mov r14, rax ;; Save that before another call
+% $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast  \texttt{r14}\mapsto_{\textsf{r}} \texttt{pte\_addr} \ast \texttt{rdi}\mapsto_{\textsf{r}} \vaddr \ast \texttt{rax}\mapsto_{\textsf{r}} \textsf{ pte\_addr} \; \ast }_{\rtv}$
+% $\specline{ \nfpointsto{\mask\vaddr\maskfour\rtv}{\mask\vaddr\maskfouroff\rtv}\entryf\qone\naddr \ast \ulcorner \texttt{entry\_present } \entryf \land \texttt{entry\_present } \entrytr \land  \texttt{entry\_present } \entrytw \urcorner \ast}_{\rtv}$
+% $\specline{  \nfpointsto{\mask\vaddr\maskthree\entryf}{\mask\vaddr\maskthreeoff\entryf}\entrytr\qtwo\naddr \ast \nfpointsto{\mask\vaddr\masktwo\entrytr}{\mask\vaddr\masktwooff\entrytr}\paddr\qthree\entryo \;\ast}_{\rtv}$
+% $\specline{\texttt{pte\_addr} \mapsto_{\texttt{vpte}} \paddr \;(\texttt{wzero 64}) \ast \texttt{rax}\mapsto_{\textsf{r}} \texttt{pte\_addr}  }_{\rtv}$
+% call alloc_phys_page_or_panic
+% $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast  \texttt{r14}\mapsto_{\textsf{r}} \texttt{pte\_addr} \ast \texttt{rdi}\mapsto_{\textsf{r}} \vaddr \;\ast \nfpointsto{\mask\vaddr\maskfour\rtv}{\mask\vaddr\maskfouroff\rtv}\entryf\qone\naddr \ast}_{\rtv}$
+% $\specline{  \nfpointsto{\mask\vaddr\maskthree\entryf}{\mask\vaddr\maskthreeoff\entryf}\entrytr\qtwo\naddr \ast \nfpointsto{\mask\vaddr\masktwo\entrytr}{\mask\vaddr\masktwooff\entrytr}\paddr\qthree\naddr \;\ast}_{\rtv}$
+% $\specline{\texttt{pte\_addr} \mapsto_{\texttt{vpte}} \paddr\; (\texttt{wzero 64})  \ast \ulcorner \texttt{entry\_present } \entryf \land \texttt{entry\_present } \entrytr \land  \texttt{entry\_present } \entrytw \urcorner}_{\rtv}$
+% $\specline{\exists \texttt{ fpaddr} \ldotp \ulcorner \texttt{aligned fpaddr} \urcorner \ast \texttt{rax}\mapsto_{\textsf{r}} \texttt{fpaddr+3} \ast \texttt{fpaddr} \mapsto_{\textsf{p}} (\texttt{wzero 64}) \ast \ulcorner \texttt{entry\_present (fpaddr+3)}\urcorner}_{\rtv}$
+% ;; Calculate new L1 entry
+% mov [r14], rax ;; store the page table entry, mapping the page
+% $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast  \texttt{r14}\mapsto_{\textsf{r}} \texttt{pte\_addr} \ast \texttt{rdi}\mapsto_{\textsf{r}} \vaddr \;\ast \nfpointsto{\mask\vaddr\maskfour\rtv}{\mask\vaddr\maskfouroff\rtv}\entryf\qone\naddr \ast}_{\rtv}$
+% $\specline{  \nfpointsto{\mask\vaddr\maskthree\entryf}{\mask\vaddr\maskthreeoff\entryf}\entrytr\qtwo\naddr \ast \nfpointsto{\mask\vaddr\masktwo\entrytr}{\mask\vaddr\masktwooff\entrytr}\paddr\qthree\entryo \;\ast}_{\rtv}$
+% $\specline{\texttt{pte\_addr} \mapsto_{\texttt{vpte}} \paddr \;(\texttt{fpaddr+3}) \; \ast \ulcorner \texttt{entry\_present } \entryf \land \texttt{entry\_present } \entrytr \land  \texttt{entry\_present } \entrytw \urcorner }_{\rtv}$
+% $\specline{\ulcorner \texttt{aligned fpaddr} \urcorner \ast \texttt{rax}\mapsto_{\textsf{r}} \texttt{fpaddr+3} \ast \texttt{fpaddr} \mapsto_{\textsf{p}} (\texttt{wzero 64}) \ast \ulcorner \texttt{entry\_present fpaddr+3}\urcorner}_{\rtv}$
+% $\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\; \sqsubseteq $
+% $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast  \texttt{r14}\mapsto_{\textsf{r}} \texttt{pte\_addr} \ast \texttt{rdi}\mapsto_{\textsf{r}} \vaddr \ast }_{\rtv}$
+% $\specline{\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\vaddr,\entryf,\entrytr,\entrytw,\fpaddr+3) \ast \ulcorner \theta \;!!\;\vaddr = \texttt{None}\urcorner \; \ast}_{\rtv}$
+% $\specline{\ulcorner \texttt{aligned fpaddr} \urcorner \ast \texttt{rax}\mapsto_{\textsf{r}} \texttt{fpaddr+3} \ast \texttt{fpaddr} \mapsto_{\textsf{p}} (\texttt{wzero 64}) }_{\rtv}$
+% $\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\; \sqsubseteq $
+% $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace} (<[\vaddr:=\texttt{fpaddr}]> \theta,m) \ast}_{\rtv}$
+% $\specline{\ulcorner \texttt{aligned fpaddr} \urcorner \ast \texttt{fpaddr} \mapsto_{\textsf{p}} \textsf{ wzero 64} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  \ast\sumwalkabs\vaddr\qfrac\fpaddr}_{\rtv}$
+% $\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\; \sqsubseteq $
+%   $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace} (<[\vaddr:=\texttt{fpaddr}]> \theta,m) \ast \vaddr \mapsto_{\textsf{vpte}}\; \{\qfrac\} \;\fpaddr \textsf{ wzero 64}}_{\rtv}$
+% $\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\; \sqsubseteq $
+% $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace} (<[\vaddr:=\texttt{fpaddr}]> \theta,m) \ast \vaddr \mapsto_{\textsf{v}}\; \{\qfrac\} \textsf{wzero 64}}_{\rtv}$
+% \end{comment}

### 2025-07-28 programlogic.tex
remove stale TODOs, add appendix and TODO after live discussion
minor space tweaks, fix misspelling

@@
 \end{lstlisting}
-\todo[inline,color=cyan]{Ismail, we can drop the maddr from the l4M52 etc, and the l4e etc. from the l3off etc., right?}
 \vspace{-1em}
@@
 % what invariants should hold, then assert that those invariants do hold.}
-\add{We mus also ensure that different address spaces can have independent ghost maps ---
+\add{We must also ensure that different address spaces can have independent ghost maps ---
 which we resolve with an additional unique global ghost map (with ghost name $\delta{s}$) from address-space identifiers (page table roots
@@
 % role in our development.}
-  \[
+\\\centerline{$
   \begin{array}{l}
@@
     \end{array}
-  \]
+$}\\
 Our Hoare doubles $\{\Phi\}_\textsf{rtv}\;\textsf{e}$ state that the expression (i.e., sequence of instructions)
@@
 }
+\add{There are only two changes specific to our approach.}
+\looseness=-1
 
-\add{There are only two changes specific to our approach.}
 \replace{However, }{First,}

### 2025-07-28 syntaxsemantics.tex
minor space tweaks, fix misspelling

@@
 For example, $\kw{w}_{12}$ is a 12-bit word, which can be obtained for example truncating away 52 bits of a 64-bit word ($\kw{w}_{64}$).
-
 % \subsection{State}
@@
 Of particular note, \lstinline|cr3|, the page table register, is included in the machine state.
-% As one might have already anticipated from the syntax we introduce, we do not bind any value of an evaluated expression.
-All the indices and accessed values are treated as globally referenced.
-In align with this design choice, our expression is a stream of instructions, which is not evaluated to a value to be bound, but changes the machine state through the indices (e.g.
-$\kw{r}\in\kw{greg}$) -- to the global maps.
+\looseness=-1
 

### 2025-07-28 background.tex
space tweaks

@@
 Some prior \iris-based work~\cite{dang2019rustbelt,dang2022compass} has constructed derived modalities in the style we propose, indexed
-by thread IDs. However in that setting, the intepretation of those modalities was fully fixed ahead of time (to refer to essentially buffers in operationalized versions of C11
+by thread IDs. However \replace{in that setting, the}{their} intepretation of those modalities was fully fixed ahead of time (to refer to essentially buffers in operationalized versions of C11
 concurrency). In this setting, while our modalities will be indexed by page table roots, it is possible to modify the address translation for an address
@@
 \add{This is essential to updating page mappings for the current address space in use by a CPU.}
-%
-%
-%Work on modal logic extends back for decades with many widely-varied variations; we discuss here only the pieces most relevant to our work.
-%Broadly speaking modal logics incorporate \emph{modal operators}, which take as arguments a proposition expected to be true in another time~\cite{pnueli1977temporal}, place~\cite{gordon2019modal}\todo{other cites}, or circumstance~\cite{hintikka1962knowledge,halpern1985guide}, and result in a proposition true in the \emph{current} time, place, or circumstance at which the truth of the use of the modal operator is being evaluated.
-Classic examples include modal necessity $\square P$ describing that $P$ is \emph{necessarily} true, $\mathsf{G}~P$ meaning $P$ is true \emph{globally} (i.e., forever from this time onwards), or $K_i(P)$ describing that a particular participant $i$ \emph{knows} that $P$ is true.
-The latter is an example of of \emph{multimodal} logic, where there is an indexed family of modalities (modal operators) parameterized by some dimension of interest (there, participants).
-%A closely related variant of multimodal logic is \emph{dynamic logic}~\cite{pratt1976semantical}, a logic of weakest preconditions~\cite{dijkstra-75} which works with modalities of the form $[p](P)$, which states that \emph{in the current program state}, \emph{if} program $p$ is run then afterwards $P$ will hold (modulo non-termination).
-%This same idea is used to encode Hoare triples in the Iris program logic~\cite{krebbers2017essence}, using the same encoding as in Pratt's original presentation, where a Hoare triple $\{P\}C\{Q\}$ is encoded as $P\rightarrow[C](Q)$.
-Unlike classic work on dynamic logic~\cite{harel2000dynamic}, Iris applies these ideas in a \emph{substructural} setting (separation logic) where distributivity laws over substructural connectives must be considered.
-For both the dynamic modality and the later modality $\blacktriangleright$ used for guarded recursive predicates, these modalities satisfy axioms of the form $M(P)\ast Q\vdash M(P\ast Q)$, but not the reverse.
-%Iris is not the first combination of modal and substructural logic~\cite{dovsen1992modal,restall1993modalities,d1997grafting,kamide2002kripke,licata2017fibrational}, but is certainly the most heavily tested.
-%A hallmark of a unary logical operator $M$ being a modality is if $M$ satisfies an property akin to $(\upvarphi\rightarrow\psi)\rightarrow M(\upvarphi)\rightarrow\psi$, which roughly states that modus ponens holds under the modality.\footnote{Afficionados of modal logic will note that this property is not quite Axiom K (which requires the initial implication to also hold under $M$), but follows from K and a necessitation rule $\upvarphi\rightarrow M(\upvarphi)$.
-Non-necessitive modalities typically satisfy the weaker property we call out above.}
-%Another pillar of our work is hybrid logic, a branch of modal logic with \emph{names} (called \emph{nominals} for states in Kripke models~\cite{blackburn1995hybrid,goranko1996hierarchies,areces2001hybrid,gargov1993modal}, in contrast to the typical flavor of modality which refers to an \emph{unspecified} (in the assertion) set of alternative circumstances.
-%
-%Our work draws on ideas from hybrid modal logics, applied in the context of Iris's higher-order separation logic~\cite{jung2018iris}.
-We define our modalities and prove their rules directly within Iris, so there is no model-theoretic novelty in our work.
-Instead, we show that targeted use of modalities combining the ideas of \emph{named resources} with the power of substructural reasoning enable clear specification of general reasoning principles for data structures.
-Along the way we take advantage of the fact that modalities need not be injections between the \emph{same} logics, but in fact the propositions that modal operators contextualize can come from \emph{other} logics; in our case we exploit the fact that Iris itself is parameterized by a choice of step-indexed~\cite{ahmed-appel-virga-02} \textsf{BI}-algebra~\cite{ohearn1999bunched} to allow convenient specifications in embedded data-structure-relative logics.
+\looseness=-1

### 2025-07-28 examples_C.tex
integrate the assembly appendix by writing a brief new intro
space tweaks

@@
 \subsection{Traversing Live Page Tables}
-\label{sec:traversing}
+\label{sec:traversingC}
 We build up to the main task of mapping a new page after traversing the page tables in the software.
@@
 \emph{generated from this source code}.
-Listings of the assembly fragments with inline assertions appear in Appendix ???. %\ref{apdx:asm}.
-\todo[inline]{Need to add the assembly appendix.}
+Listings of the assembly fragments with inline assertions appear in Appendix \ref{sec:experiment_appendix}. %\ref{apdx:asm}.
 Whether in C or assembly,
@@
 $}\\
-This supports memory access rules much like Figure \ref{fig:wpdamd}'s rules (which are proven sound using soundness of virtual pte-points-to rules as lemmas!)
+This supports memory access rules much like Figure \ref{fig:wpdamd}'s rules (which are proven
+sound using the virtual pte-points-to rules as lemmas!)
 while exposing the physical location being modified.
@@
 identity mapping resources for \lstinline|entry| still pulled out (it was removed by the caller).
+\looseness=-1
 

### 2025-07-29 examples_C.tex
space tweaks, fix remaining todos to guage space

@@
 % \end{lstlisting}
-\todo[inline]{
-Ismail, why does the implication on line 13 talk about entryv? Is that really what the Rocq proof talks about, after overwriting?
-Are we missing some existential quantification of the value in the conditional branch and lines 19-21?
-}
-\todo[inline,color=cyan]{
-  reading carefully, it says that if entry\_present of pte\_initialize of something holds (which it always will)
-  then the 512 entries are there. So this is basically ``True implies'', but in a shape that's compatible
-  with the other branch/path's invariants
-}
 \vspace{-1em}
@@
 allocates a physical page for use as the next-level table.
-\lstinline|pte_initialize| is shown in Figure \ref{pteinitializespecC}: it simply calls
+\lstinline|pte_initialize| (Figure \ref{pteinitializespecC}) calls
 \lstinline|kalloc| to allocate a physical page (Figure \ref{pteinitializespecC} Line \ref{line:call_to_kallocC}),
@@
 by updating the current address space's entry.
-\todo[inline]{still feel like this was a little too high level}
 

### 2025-07-29 relwork.tex
space tweaks
space tweaks, fix remaining todos to guage space

@@
 \label{sec:relwork}
-\todo[inline]{Colin must incorporate stuff from response}
-
 We will discuss two streams of related work: OS verification, and 
@@
 While Kolanski and Klein verify the critical step of updating an already-located L1 entry to map a new page
-(ARM assembly corresponding roughly to Lines \ref{line:l1entry_storeC} and \ref{line:l1entry_set_present}
+(ARM assembly corresponding roughly to Lines \ref{line:l1entry_storeC} and \ref{line:l1entry_setpresent}
 in our Figure \ref{fig:mapping_codeC}), they ignore the essential code
-preceding that step, which as our \lstinline|walkpgdir| and \lstinline|pte_get_next_table| verifications
+preceding that step --- which as our \lstinline|walkpgdir| and \lstinline|pte_get_next_table| verifications
 demonstrate, side-steps a significant amount of complexity and critical reasoning tasks.
@@
 }
+\looseness=-1
 

### 2025-07-29 background.tex
still need to polish relwork, a little less than half a page over the revision page limit

@@
 \includegraphics[width=\textwidth]{addr_translation.png}
-\vspace{-1em}
+\vspace{-2em}
           \caption{x86-64 \replace{page table lookups}{hardware address translation}.}
@@
 
-The page tables are managed by the OS, typically by a \emph{virtual memory manager} (VMM).\footnote{Not to be confused with Virtual 
+The page tables are managed by the \replace{OS, typically by a}{OS kernel's} \emph{virtual memory manager} (VMM).\footnote{Not to be confused with Virtual 
 Machine Monitor. We focus on non-hypervisor scenarios, but hardware virtualization extensions for both 
@@
 own access of the tables}.
-
 \add{
@@
 
-We suppress many details of \iris assertions for brevity, but briefly note
+We suppress \replace{many details of \iris assertions}{some technical\iris details} for brevity, but briefly note
 a few recurring details used heavily in \iris but not necessarily in traditional
@@
 Because \iris is an embedded framework in \rocq, proofs in \iris-derived
-logics (including ours) often encapsulate raw \rocq assertions: $\ulcorner P \urcorner$ is an embedding
+logics \del{(including ours)} often encapsulate raw \rocq assertions: $\ulcorner P \urcorner$ is an embedding
 of the \rocq assertion $P$ into an \iris assertion (used for things like equality and other
@@
 \del{We introduce other details as they arise.}
+\looseness=-1
 

### 2025-07-29 examples_C.tex
still need to polish relwork, a little less than half a page over the revision page limit

@@
 $\specline{\mathsf{ContextAt}(save,\_) \ast \mathsf{ContextAt}(restore,[\mathsf{rbxv}',\ldots,\rtv']) }_{\rtv}$(*\label{line:end_swtch_pre}*)
-mov 0[rdi], rbx(*\label{line:start_save}*)
-... ;; also save rsp, rbp, r12, r13, r14, to offsets 8, 16, 24, 32, and 40 from rdi
-mov 48[rdi], r15
+mov 0[rdi], rbx(*\label{line:start_save}*) ... ;; also save rsp, rbp, r12, r13, r14, to offsets 8, 16, 24, 32, and 40 from rdi
+... mov 48[rdi], r15
 $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,\Xi,m) \ast [\rtv'](\mathcal{I}\texttt{ASpace}(\theta',\Xi',m') \ast \texttt{Pother})  \ast  \texttt{rsi}\mapsto_{\textsf{r}} restore \ast \texttt{rdi}\mapsto_{\textsf{r}} save \ast \texttt{rbx}\mapsto_{\textsf{r}} \texttt{rbxv} }_{\rtv}$
@@
     \underbrace{\left(\ulcorner\textsf{present\_L}(\vale,\mathsf{level})\urcorner \wand \forall_{\textsf{i}\in\textsf{0..511}} \ldotp \ghostmaptoken{\textsf{id}}{((\mathsf{entry\_page}\;\vale) + \textsf{i * 8})}{\textsf{level-1}}\right)}_{\text{\textcircled{4} Indexing into next level of tables}} \\
-  \textsf{ where } \\
-   \textsf{ASpace\_Lookup}_{\textsf{id}}(\ptablestore,\Xi,m) \stackrel{\triangle}{=} \lambda\textsf{ cr3val} \ldotp \; \exists \gammaPred,\gammaPred' \; \ldotp \ulcorner m \; !!\; \textsf{cr3val} = \textsf{Some } \gammaPred \urcorner \ast
+  \textsf{ where } \quad
+  \textsf{present\_L}(\vale,\mathsf{level})\stackrel{\triangle}{=} \mathsf{entry\_present}(\vale)\land \mathsf{level} > 1\\
+   \textsf{ASpace\_Lookup}_{\textsf{id}}(\ptablestore,\Xi,m) \stackrel{\triangle}{=} \lambda\textsf{ cr3val} \ldotp \; \exists \gammaPred,\gammaPred'\ldotp \ulcorner m \; !!\; \textsf{cr3val} = \textsf{Some } \gammaPred \urcorner \ast
    % \ownGhost\gammaPred{\authfull{\ptableabswalk\ptablestore}} \ast  \ownGhostpv\gammaPred{\authfull{\pvmapping\Xi}}
-   \ptableabswalk{\delta,\ptablestore} \ast \pvmapping{\delta',\Xi}\\
-  \textsf{present\_L}(\vale,\mathsf{level})\stackrel{\triangle}{=} \mathsf{entry\_present}(\vale)\land \mathsf{level} > 1
-
+   \ptableabswalk{\delta,\ptablestore} \ast \pvmapping{\delta',\Xi}
 \end{array}
 \]
-\vspace{-1em}
+\vspace{-1.5em}
 \caption{Address space invariant of Figure \ref{fig:peraspaceinvariant} extended with a ghost map bookkeeping identity mappings. }

### 2025-07-29 programlogic.tex
still need to polish relwork, a little less than half a page over the revision page limit

@@
 The existence of such a token implies that the actual map $\theta$ tracked by a corresponding $\mathsf{GhostMap}(\gamma,\theta)$
-  resource indeed maps $k$ to $v$. These properties are captured by key \iris rules in Figure \ref{fig:ghostmaps}.\footnote{\iris ghost maps lack established notation in the literature,
-  but the syntax we use captures the details of \texttt{iris.base\_logic.lib.ghost\_map}.}
+  resource indeed maps $k$ to $v$. These properties are captured by key \iris rules in Figure \ref{fig:ghostmaps}.\footnote{\iris ghost maps lack established notation\add{;}
+  \del{in the literature, but} the syntax we use captures the details of \texttt{iris.base\_logic.lib.ghost\_map}.}
   There are other rules, but these two are most important for explaining ghost maps.
@@
 \begin{array}{l}
-  \mathcal{I}\textsf{ASpace}(\ptablestore,m)\stackrel{\triangle}{=} \textsf{ASpace\_Lookup}(\ptablestore,m) \ast \\
- \qquad\qquad\qquad \bigast{(\vaddr, \textsf{paddr})\in \ptablestore}{\exists\;(\textsf{l4e l3e l2e, l1e, paddr})\ldotp \textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\vaddr\textsf{, l4e, l3e, l2e, l1e, paddr})} \\
+  \mathcal{I}\textsf{ASpace}(\ptablestore,m)\stackrel{\triangle}{=} \textsf{ASpace\_Lookup}(\ptablestore,m) \ast 
+  \bigast{(\vaddr, \textsf{pa})\in \ptablestore}{\exists\;(\textsf{l4e l3e l2e, l1e, pa})\ldotp \textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\vaddr\textsf{, l4e, l3e, l2e, l1e, pa})} \\
   \textsf{ where } 
@@
 We place the authorative ownership of the ghost \add{map} translation $\mathcal{A}\textsf{PTableWalk}$ in a per-address-space invariant
-$\mathcal{I}$\textsf{ASpace} (Figure \ref{fig:peraspaceinvariant}). 
-$\mathcal{I}$\textsf{ASpace} alone allows changes to the page tables
-that preserve overall virtual-to-physical translations.
-When combined with the
-fragment stored in the virtual points-to (Figure \ref{fig:virtualpointstosharing}),
-changes in the virtual-to-physical translations become possible without the need to track all virtual points-to assertions.
+$\mathcal{I}$\textsf{ASpace} (Figure \ref{fig:peraspaceinvariant}), 
+\replace{$\mathcal{I}$\textsf{ASpace} alone allows}{allowing} changes to the page tables
+that preserve overall virtual-to-physical translations \add{in isolation},
+\add{and also allowing changes to specific the virtual-to-physical translations}
+\replace{W}{w}hen combined with the
+\replace{fragment}{token} stored in the \add{relevant} virtual points-to (Figure \ref{fig:virtualpointstosharing}).
+\del{changes in the virtual-to-physical translations become possible without the need to track all virtual points-to assertions.}
+\looseness=-1
 

### 2025-07-29 relwork.tex
still need to polish relwork, a little less than half a page over the revision page limit

@@
 which is not true of page table updates.
-Other work on OS verification either never progressed far enough to address VMM verification
-(\textsc{Verisoft XT}~\cite{cohen2009vcc,cohen2010local,dahlweid2009vcc,cohen2013SOFSEM}), or uses memory-safe languages to enable safe cohabitation 
-of a single address space by all processes (\textsc{Singularity}~\cite{Fahndrich2006language,Hunt2007singularity,Hunt2007sealing,Barnett2011specsharp}, \textsc{Verve}~\cite{Yang2010Verve},
-and \textsc{Tock}~\cite{levy2017multiprogramming}). The latter approach ensures memory safety, but ignores other functional uses of virtual memory
+Other work on OS verification either never progressed \del{far enough} to \del{address} VMM verification
+(\textsc{Verisoft XT}~\cite{cohen2009vcc,cohen2010local,dahlweid2009vcc,cohen2013SOFSEM}),
+or uses memory-safe languages
+\replace{to enable safe cohabitation of a single address space by all processes}{for process isolation instead of address translation}
+(\textsc{Singularity}~\cite{Fahndrich2006language,Hunt2007singularity,Hunt2007sealing,Barnett2011specsharp}, \textsc{Verve}~\cite{Yang2010Verve},
+and \textsc{Tock}~\cite{levy2017multiprogramming})\replace{. The latter approach ensures}{, ensuring} memory safety, but ignor\replace{es}{ing} other functional uses of virtual memory
 hardware, like swapping~\cite{Denning1970VM} or exploiting copy-on-write techniques for dynamic migration of virtual machines~\cite{clark2005live}.

### 2025-07-29 examples_C.tex
okay one more space improvement for tonight

@@
 that \textsf{entryp} is part of the page table invariant on Line \ref{line:precondition_entry_out}).
-
 This comes into play with a key subtlety of \lstinline|pte_get_next_table|'s
@@
 but its postcondition does not yield new virtual points-to assertions!
-Instead it merely computes the base virtual address of the next table, and returns adequate tokens
-(discussed in Section \ref{subsec:identitymappingsC}, explicit on Line \ref{line:childrenC})
+It merely computes the base virtual address of the next table, and returns adequate tokens
+% (discussed in Section \ref{subsec:identitymappingsC}, explicit on
+(Line \ref{line:childrenC})
 for the \emph{caller} to construct a vpte-points-to for any entry of the next table level.

### 2025-07-29 examples_C.tex
Fix fig 11

@@
 \begin{array}{l}
-  \mathcal{I}\textsf{ASpace}_{\textsf{id}}(\ptablestore,\Xi,m)\stackrel{\triangle}{=} \textsf{ASpace\_Lookup}_{\textsf{id}}(\ptablestore,\Xi,m) \ast \mathsf{GhostMap}(\mathsf{id},\Xi)\ast\\
+  \mathcal{I}\textsf{ASpace}_{\textsf{id}}(\ptablestore,\Xi,m)\stackrel{\triangle}{=} \textsf{ASpace\_Lookup}_{\textsf{id}}(\ptablestore,\Xi,m) \ast\\  %\mathsf{GhostMap}(\mathsf{id},\Xi)\ast\\
   \left(\bigast{(\vaddr, \textsf{paddr})\in \ptablestore}{\exists\;(\textsf{l4e, l3e, l2e, l1e, paddr})\ldotp \textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\vaddr\textsf{, l4e, l3e, l2e, l1e, paddr})}\right)\ast \\

### 2025-07-29 examples.tex
simplify tautological implications to save space, but note proof details in appendix. delete obvious invariants that we do not necessarily need to repeat

@@
 moves into the page table invariant.
+
+\add{Note that the C presentation in Figure \ref{fig:calltopteinitializeC}
+omitted the precondition on the implication of Figure \ref{fig:calltopteinitialize}'s Line \ref{line:page_of_caps},
+which is logically equivalent to \textsf{True} since \textsf{entry\_present} checks if the present bit is set in an entry,
+and \textsf{pte\_initialize} sets that bit. The actual invariant has this form here, and in the postcondition
+of \lstinline|pte_initialize| (Figure \ref{pteinitializespec}), to match the conditional form from earlier in
+\lstinline|pte_get_next_table| (which is also provably true when the check of the present bit
+determines that the entry was already valid/present).
+Our proof discharges the conditional at the join point, rather than eagerly in each branch.
+}
 

### 2025-07-29 examples_C.tex
simplify tautological implications to save space, but note proof details in appendix. delete obvious invariants that we do not necessarily need to repeat

@@
     pte_initialize(entry); // Allocate a zeroed physical page and install in entry(*\label{line:call_to_pte_initializeC}\label{line:page_of_capsC}*)
-    $\specline{ \ldots \ast \ulcorner  \textsf{qfrac} = 1 \land \lnot(\textsf{entry\_present entryv})  \urcorner \ast \ulcorner  \textsf{qfrac} = 1 \land \lnot(\textsf{entry\_present (pfn\_set (entryv nextpaddr))}  \urcorner   }_{\rtv}$
-    $\specline{ \begin{array}{l}\textsf{entry\_present (pte\_initialized (pfn\_set(entryv nextpaddr)))} \wand  \\ \;\;\;\;\;\;\;\;\  \forall_{i\in \textsf{0 ...
-511} } \ldotp  \ghostmaptoken{\textsf{id}}{((\textsf{table\_root (pte\_initialized (pfn\_set(entryv nextpaddr)))}) + \textsf{i * 8})}{\textsf{v-1}} \end{array}  }$ (*\label{line:page_of_capsC}*)
     entry->present = 1; // Install newly-allocated level n table, mark present(*\label{line:install_new_entryC}*)
@@
   $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \lnot(\textsf{entry\_present entryv}) \ast \textsf{entryp+KERNBASE} \mapsto_{\mathsf{vpte,1}} \mathsf{entryp} \; \textsf{pfn\_set(entryv nextpaddr)}}$
-  $\specline{\ghostmaptoken{\delta{}s}{\rtv}{\delta} \ast \begin{array}{l} \ulcorner\textsf{entry\_present (pte\_initialize(pfn\_set(entry\_val nextpaddr)),level)}\urcorner \wand \\  \;\;\;\;\;\;\; \forall_{i\in\textsf{0..511}} \; \textsf{table\_root (pte\_initialized (pfn\_set (entry\_val nextpaddr))) + i * 8} \mapsto_{\textsf{id}} \textsf{level-1}    \end{array}  }_{\rtv}$(*\label{line:initialize_post}*)
+  $\specline{\ghostmaptoken{\delta{}s}{\rtv}{\delta} \ast \forall_{i\in\textsf{0..511}}\ldotp ()\textsf{table\_root (pte\_initialized (pfn\_set (entry\_val nextpaddr))) + i * 8}) \mapsto_{\textsf{id}} \textsf{level-1}   }_{\rtv}$(*\label{line:initialize_post}*)A
 }

### 2025-07-29 examples_C.tex
bus draft 25 pages

@@
 \subsubsection{From L$n+1$ Entries to L$n$ Tables}
-We will discuss access to the level 4 table later (Section \ref{wlkpgdirC}). However, for subsequent levels, the base address of level $n$ must be
+We discuss access to the level 4 table later (Section \ref{wlkpgdirC}). However, for subsequent levels, the base address of level $n$ must be
 fetched from the appropriate entry in the table of level $n+1$.
@@
 table entry.
+\looseness=-1
 
@@
 We can now discuss \lstinline|pte_get_next_table|'s proof of correctness.
-While at first glance this code does not sound terribly complex, beyond needing some care to distinguish physical and virtual addresses,
+While at first glance this code may look like its subtlety is mostly care to distinguish physical and virtual addresses,
 it has a highly nontrivial correctness argument, which depends critically on detailed invariants on how access to page table
 entries is shared between parts of the kernel. No prior work has engaged with this problem.
+\looseness=-1
 
@@
 Even given the slight adaptation of our assembly-level proof for the C-level presentation in Figure \ref{walkpgdirC},
-the proof outline in the figure omits may assertions that hold at various points in time.
-But by Line \ref{line:ex_l3_vpteC}, it should be clear that the proof is on track to accumulate a set of similar assertions for each level.
-Figure \ref{fig:rwalkC} expands the abbreviated postcondition to include the full set of facts that are accumulated in this way.
-$\mathsf{R}_\mathsf{walk}$ together with $\mathsf{R}_{\mathsf{l1e}}$ should look quite similar to \lstinline|L4_L1_PointsTo| (Figure \ref{fig:strongvirtualpointsto}),
-and this is quite intentional. \lstinline|walkpgdir|'s execution observes enough
-to determine that there is most of an address translation for virtual address \lstinline|va|,
+the proof outline in the figure omits some repeat intermediate assertions for readability.
+But by Line \ref{line:ex_l3_vpteC}, it should be clear that the proof accumulates a set of similar assertions for each level.
+Figure \ref{fig:rwalkC} expands the abbreviated postcondition to the full set of facts that are accumulated in this way.
+$\mathsf{R}_\mathsf{walk}$ together with $\mathsf{R}_{\mathsf{l1e}}$
+nearly entail \lstinline|L4_L1_PointsTo| (Figure \ref{fig:strongvirtualpointsto}) within the logic,
+forming the basis of the construction of a new virtual points-to for virtual address \textsf{va}.
+\lstinline|walkpgdir|'s execution observes most evidence of an address translation for \lstinline|va|,
 at least down to a possibly-invalid L1 entry (which \lstinline|walkpgdir|'s caller, \lstinline|vaspace_mappage|, will check). 
@@
 and the pure assertions in $\mathsf{R}_\mathsf{walk}$ ensure the address arithmetic works.
-$\mathsf{R}_{\mathsf{l1e}}$ also includes the self-conditional fractional ownership of the L1 entry (per Section \ref{sec:selfconditional})
-for the caller to initialize the entry if it is empty, making it possible for the caller to complete virtual points-to assertions
+$\mathsf{R}_{\mathsf{l1e}}$ includes the self-conditional fractional ownership of the L1 entry (Section \ref{sec:selfconditional})
+for the caller to initialize the entry if it is empty, so the caller can complete a virtual points-to assertions
 for a newly-mapped page, as we discuss next.
+\looseness=-1
 
@@
 Lines \ref{line:l1entry_storeC} and \ref{line:l1entry_setpresent} are verified using the pte-points-to
-memory store rule.\footnote{In the assembly, \lstinline|fpaddr| is already known to be word-aligned, so it is simply bitwise-and-ed with 1
-(setting the present flag) and a single store occurs.}
+memory store rule.\footnote{In the assembly proof, this is a bitwise-and of the word-aligned\lstinline|fpaddr| with 1
+(setting the present flag), yielding a single.
+\looseness=-1
+}
 \looseness=-1

### 2025-07-29 relwork.tex
bus draft 25 pages

@@
 
-\subsection{OS Kernel Verification}
+\paragraph{OS Kernel Verification}
 There has been relatively little prior work on formal verification of virtual memory.
@@
 \add{Thus} they \emph{trust}
-that the particular page table manipulations do not, for example, unmap kernel code\footnote{\add{A ``fun'' bug one author encountered in
-an industrial prototype kernel.}}
+that the particular page table manipulations do not, for example, unmap kernel code
+% \footnote{\add{A ``fun'' bug one author encountered in
+% an industrial prototype kernel.}}
  (which can crash the machine even if done ``temporarily'').
@@
 }
+\add{
+ Our work improves significantly on the technical capabilities of ther logic and
+ evaluates on kernel code that is more complete and more challenging than theirs.
+}
 \looseness=-1
@@
 \add{
- Our work improves significantly on the technical capabilities of ther logic and
- evaluates on kernel code that is more complete and more challenging than theirs.
  Our modal approach makes it possible to specify address space changes cleanly, which their logic cannot do at all.
- Our use of virtual pte-points-to assertions enables nearly the same proof rules\todo{if this stays, we should typeset one of the vpte-pointsto rules to justify this}
+ Our use of virtual pte-points-to assertions enables nearly the same proof rules
  as standard memory accesses, and constructing virtual points-to information within the logic
  (c.f.\ the logical entailment between $\textsf{R}_\textsf{walk}$ and $\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$)
- rather than reasoning semantically about when the model state supports new virtual points-to assertions.
+ whereas Kolanski and Klein must reason semantically about when the model state supports new virtual points-to assertions.
 }
@@
 \add{
-While Kolanski and Klein verify the critical step of updating an already-located L1 entry to map a new page
+Kolanski and Klein verify the critical step of updating an already-located L1 entry to map a new page
 (ARM assembly corresponding roughly to Lines \ref{line:l1entry_storeC} and \ref{line:l1entry_setpresent}
-in our Figure \ref{fig:mapping_codeC}), they ignore the essential code
+in our Figure \ref{fig:mapping_codeC}), but ignore the essential code
 preceding that step --- which as our \lstinline|walkpgdir| and \lstinline|pte_get_next_table| verifications
 demonstrate, side-steps a significant amount of complexity and critical reasoning tasks.
-We have verified the entirety of the software page table walk up to mapping a new page, aside from a trusted physical memory allocator.\footnote{
- \add{Which works quite similarly to verified \lstinline|malloc|~\cite{Chlipala2013Bedrock,wickerson2010explicit}}
+We have verified the entirety of the software page table walk up to mapping a new page, aside from a trusted physical memory allocator resembling
+\lstinline|malloc|~\cite{Chlipala2013Bedrock,wickerson2010explicit}.
 }
@@
 }
-\todo[inline]{author response talks about syeda and klein here briefly, and it fits, but we also want the longer discussion}
+% \todo[inline]{author response talks about syeda and klein here briefly, and it fits, but we also want the longer discussion}
 
-\paragraph{Translation Lookaside Buffers}
-As noted in Section \ref{sec:backgroundonmachinemodel}, we do not formally model or reason about TLBs.
+\del{Translation Lookaside Buffers}
+As noted in Section \ref{sec:backgroundonmachinemodel}, we do not formally model or reason about \add{translation lookaside buffers} \add{(}TLBs\add{)}.
 \replace{TLB flushes are required only when addresses are removed from a virtual address space, or when changing
@@
 
-\subsection{Program Logics with Modalities}
-Modalities have long been a staple of program logics, at least as far back as Dijkstra's weakest precondition calculus~\cite{dijkstra1975guarded}
+\paragraph{Program Logics with Modalities}
+Modalities have long been a staple of program logics, at least \replace{as far back as}{since} Dijkstra's weakest precondition calculus~\cite{dijkstra1975guarded}
 and Pratt's observation the Hoare triples
@@
 for impredicative and recursive features of logics and type systems~\cite{Appel2007,hobor2010theory,birkedal2011step,birkedal2013intensional,jung2018iris}.
+\looseness=-1
 

### 2025-07-29 relwork.tex
polish relwork, conclusion

@@
 \lstinline|malloc|~\cite{Chlipala2013Bedrock,wickerson2010explicit}.
+ As a consequence of tackling this larger verification challenge, our work is the first to formally specify large
+ segments of the self-referential portion of an OS kernel's virtual memory management invariants (per Sections \ref{sec:p2vC} and \ref{sec:selfconditional}),
+ and to reason about converting from physical addresses to virtual addresses efficiently.
+\looseness=-1
 }
- As a consequence of tackling this larger verification challenge, we are also the first to formally specify large
- segments of the self-referential portion of an OS kernel's virtual memory management invariants (per Sections \ref{sec:p2vC} and \ref{sec:selfconditional}),
- and to reason about converting back from physical addresses to virtual addresses efficiently.
-}
-\looseness=-1
 
@@
 entries can be verified in a VM-ignorant program logic, and that proofs in that logic can be embedded into the VM-aware logic 
-(essentially by interpreting ``normal'' points-to relations as virtual points-to facts).}{
+(essentially by interpreting ``normal'' points-to relations as virtual points-to facts).
+}{
  Kolanski and Klein prove that updates to any memory locations that are not part of the page tables
- do not affect the interpretation of other memory addresses.
+ do not affect the interpretation of other memory addresses, just like on real hardware.
  This implies that programs that do not modify memory mappings could be reasoned about without
@@
 }
-While we have not \add{formally} proven such a result \add[of our machine model, it holds (just as for real hardware)],
-an analagous result \replace{{should} hold}{holds} of our work: consider that the doubles for the \texttt{mov} instructions
-that access memory behave just as one would expect for a VM-ignorant logic~\cite{Chlipala2013Bedrock}.
-With our general approach to virtual points-to assertions being inspired by Kolanski and Klein, \emph{both}
- our approach and theirs could in principle be extended to account for pageable points-to assertions by adding \del{additional}
-disjunctions to an extended points-to definition; embedding ``regular'' separation logic into such a variant
-is the appropriate next step to extend reasoning to usermode programs running with a kernel that may demand-page the program's
+\replace{While we have not proven {such a result},
+an analagous result \replace{{should} hold}{holds} of our work:}{
+ An analagous result should holds of our model (though we have not proven it).
+}
+\replace{ consider that the doubles for the}{Informally this is visible in the rules for}
+\texttt{mov} instructions,
+\replace{that access memory behave just as one would expect for}{which are nearly identical to rules in a} VM-ignorant logic~\cite{Chlipala2013Bedrock,ni2007contexts}.
+\replace{With our general approach to virtual points-to assertions being inspired by Kolanski and Klein, \emph{both}
+ our approach and theirs could in principle}{
+  In principle both our approach and Kolanski and Klein's could
+ } be extended to account for pageable points-to assertions by adding \del{additional}
+disjunctions to an extended points-to definition\add{,}\del{; embedding ``regular'' separation logic into such a variant
+is} \add{which would be}
+ the appropriate next step to extend reasoning to usermode programs running with a kernel that may demand-page the program's
 memory.
+\looseness=-1
 
-\add{
- No prior work has dealt with combined address space changes and context switching. \textsc{XCAP}~\cite{ni2007contexts} deals
- only with usermode threading (in a single process). \textsc{CertiKOS} and \textsc{seL4} trust assembly primitives for
- context switches, in addition to not modeling address translation for executing code.
-}
 % \todo[inline]{author response talks about syeda and klein here briefly, and it fits, but we also want the longer discussion}
@@
 
-The only other attempt to verify a change of address space was by \citet{syeda2018program,syeda2020formal}, whose focus was on TLB maintenance (which we
-do not model). However, they verified only the 4 instructions to switch address spaces and update the TLB on an ARM processor, in isolation
+The only \replace{other attempt to verify a change of address space was by}{work to tackle TLB code verification was}
+\citet{syeda2018program,syeda2020formal},
+\replace{whose focus was on TLB maintenance (which we do not model)}{who are also the only prior work on verifying address-space-aware context switching}.
+However, they verified only the 4 instructions to switch address spaces and update the TLB on an ARM processor, in isolation
 (i.e., not the full context switch including changing stacks with address spaces).
@@
 and \add{an extension of} theirs, \add{a substantial project} which we leave to future work\replace{,}{.}\del{ as combining both the models and reasoning principles would be highly non-trivial.}
+\add{
+ No other prior work has considered address space changes during context switching. \textsc{XCAP}~\cite{ni2007contexts}
+ and Bedrock~\cite{Chlipala2011Bedrock,Chlipala2013Bedrock,Chlipala2015webapp} deal
+ only with usermode threading (in a single process). \textsc{CertiKOS} and \textsc{seL4} trust assembly primitives for
+ context switches, and do  not model address translation for executing code.
+ \looseness=-1
+}
 

### 2025-07-30 background.tex
actually commit the draft revision

@@
 \replace{
-ARMv8 (a.k.a.\ texttt{aarch64}), RISC-V, and PowerPC use similar hierarchical page tables for address translation.
-RISC-V's \texttt{sv48} paging configuration~\cite[\S 4.5]{riscv-privileged} and AArch64's 4-level paging configuration
+ARMv8 (a.k.a. aarch64), RISC-V, and PowerPC use similar hierarchical page tables for address translation.
+RISC-V's sv48 paging configuration~?? and AArch64's 4-level paging configuration
 both split 64-bit virtual addresses at the same points to index into the respective tables, so most of what follows
@@
 
-We suppress \replace{many details of \iris assertions}{some technical\iris details} for brevity, but briefly note
+We suppress \replace{many details of Iris assertions}{some technical \iris details} for brevity, but briefly note
 a few recurring details used heavily in \iris but not necessarily in traditional
@@
 But while propositional modal logics are no more expressive than first-order logic (with just two variables!),
-the primary goal of a modal logic is not expressive power but intuitive readable specifications.
+the primary goal of a modal logic is not raw expressive power but intuitive specifications and proof rules.
 Temporal logics are used because they simplify specifications (vs. having explicit time variables

### 2025-07-30 examples_C.tex
actually commit the draft revision

@@
     pteaddr->pfn = PTE_ADDR_TO_PFN(fpaddr); /*Store updated entry back to L1 entry (*\label{line:l1entry_storeC}*)*/
-    $\specline{ \textsf{l1\_entry+KERNBASE} \mapsto_{\textsf{vpte,1}} \textsf{l1\_entry} \; (\textsf{set\_pfn}(\textsf{l1e\_val}, \textsf{PTE\_ADDR\_TO\_PFN}(\textsf{fpaddr}))) \ast  \ulcorner \lnot(\textsf{entry\_present l1e\_val}) \urcorner \ast \textsf{P}}$
+    $\specline{ \textsf{l1\_entry+KERNBASE} \mapsto_{\textsf{vpte,1}} \textsf{l1\_entry} \; (\textsf{set\_pfn}(\textsf{l1e\_val}, \textsf{ADDR\_TO\_PFN}(\textsf{fpaddr}))) \ast  \ulcorner \lnot(\textsf{entry\_present l1e\_val}) \urcorner}$
     /*Set the present bit in entry*/
@@
   }
-  $\specline{ \begin{array}{l}\textsf{l1\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l1\_entry} \; \textsf{l1e\_val} \ast \ulcorner (\textsf{entry\_present l1e\_val}) \urcorner \ast \textsf{P} \\
- \lor \; \;  \textsf{l1\_entry+KERNBASE} \mapsto_{\textsf{vpte,1}} \textsf{l1\_entry} \; (\textsf{set\_pfn}(\textsf{l1e\_val}, \textsf{PTE\_ADDR\_TO\_PFN}(\textsf{fpaddr}))) \ast  \ulcorner \lnot(\textsf{entry\_present l1e\_val}) \urcorner \ast P \end{array} }$
+  $\specline{ \begin{array}{l}\textsf{l1\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l1\_entry} \; \textsf{l1e\_val} \ast \ulcorner (\textsf{entry\_present l1e\_val}) \urcorner \\
+ \lor \; \;  \textsf{l1\_entry+KERNBASE} \mapsto_{\textsf{vpte,1}} \textsf{l1\_entry} \; (\textsf{set\_pfn}(\textsf{l1e\_val}, \textsf{ADDR\_TO\_PFN}(\textsf{fpaddr}))) \ast  \ulcorner \lnot(\textsf{entry\_present l1e\_val}) \urcorner  \end{array} }$
 }

### 2025-07-30 programlogic.tex
actually commit the draft revision

@@
 }
-\del{In Figure \ref{fig:peraspaceinvariant},
+\del{In Figure ??,
 $\theta$ is the logical map from virtual addresses to the physical addresses they should translate to, 
 for the \emph{current} address space, corresponding to the  currently installed page tables (the ones
-indicated by \lstinline|cr3|). In Figure \ref{fig:virtualpointstosharing}, the existentially quantified $\delta$
+indicated by \texttt{cr3}). In Figure ??, the existentially quantified $\delta$
 is the ghost name for this map (again, for the current address space).

### 2025-07-30 relwork.tex
actually commit the draft revision

@@
 that the particular page table manipulations do not, for example, unmap kernel code
-% \footnote{\add{A ``fun'' bug one author encountered in
-% an industrial prototype kernel.}}
- (which can crash the machine even if done ``temporarily'').
+ (which can crash the machine even if done ``temporarily'').\footnote{\add{A ``fun'' bug one author encountered in an industrial prototype kernel.}}
 % \footnote{A bug

### 2025-07-30 examples.tex
the magic of sed (i.e., replacing all the mathcal I with regular math mode I)

@@
 ;;pte_t *pte_get_next_table(pte_t *entry) { ... // setting up the stack for a pte_t *next elided
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{entry}\},m)  \ast \texttt{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast  \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{next} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{\_} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{\_} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
+$\specline{\textsf{P} \ast  I\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{entry}\},m)  \ast \texttt{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast  \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{next} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{\_} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{\_} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
 $\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present entry\_val})\urcorner}_{\rtv}$ (*\label{line:get_next_vpte_precondition}*)
@@
 ;; pte_t *next;
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{entry}\},m)  \ast \texttt{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast  \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{next} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{\_} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{\_} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
+$\specline{\textsf{P} \ast  I\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{entry}\},m)  \ast \texttt{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast  \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{next} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{\_} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{\_} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
 $\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present entry\_val})\urcorner}_{\rtv}$ (*\label{line:get_next_vpte_precondition}*)
@@
 mov    -0x8[rbp],rdi
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{entry\},m)   \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{\_}  \ast  \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{next} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
+$\specline{\textsf{P} \ast  I\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{entry\},m)   \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{\_}  \ast  \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{next} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
 $\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present entry\_val})\urcorner}_{\rtv}$
@@
 mov     rdi, r8
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast  \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{next} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
+$\specline{\textsf{P} \ast  I\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast  \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{next} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
 $\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present entry\_val})\urcorner}_{\rtv}$
@@
 mov    [r8],rdi (*\label{line:read_entry_contents}*)
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast  \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{next} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val}  \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
+$\specline{\textsf{P} \ast  I\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast  \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{next} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val}  \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
 $\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present entry\_val})  \urcorner }_{\rtv}$
@@
 jne    161 <pte_get_next_table+0xa1> (* \label{line:check_entry_present_jump} *) ;; Jump if the present bit is not zero
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry} \ast \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{next} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val \& 0x1} }_{\rtv}$
+$\specline{\textsf{P} \ast  I\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry} \ast \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{next} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val \& 0x1} }_{\rtv}$
 $\specline{  \textsf{entry} \mapsto_{\textsf{id}} \textsf{v} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  \ast \texttt{rax}  \mapsto_{\textsf{r}} \textsf{entry\_val \& 0x1} \ast \ulcorner\textsf{entry\_val \& 0x1}=\textsf{0x0}\urcorner}_{\rtv}$
@@
 sub    0x10, rdi ;; Store the value of rbp minus 16 bytes (address of (*\textsf{next}*)) into rdi (*\label{line:pass_addrof_next}*)
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry} \ast \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{next} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{rbp - 16} }_{\rtv}$
+$\specline{\textsf{P} \ast  I\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry} \ast \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{next} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{rbp - 16} }_{\rtv}$
 $\specline{  \textsf{entry} \mapsto_{\textsf{id}} \textsf{v} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta} }_{\rtv}$
@@
 callq  70 <pte_initialize> ;;pte_initialize(entry);(* \label{line:call_to_pte_initialize} *)
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  }_{\rtv}$
+$\specline{\textsf{P} \ast  I\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  }_{\rtv}$
 $\specline{  \textsf{entry} \mapsto_{\textsf{id}} \textsf{v} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta} }_{\rtv}$
@@
 ... ;;now we know that entry is initialized, so we satisfy the condition to access children list
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  }_{\rtv}$
+$\specline{\textsf{P} \ast  I\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  }_{\rtv}$
 $\specline{  \textsf{entry} \mapsto_{\textsf{id}} \textsf{v} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta} }_{\rtv}$
@@
 \begin{array}{l}
-  \mathcal{I}\textsf{ASpace}_{\textsf{id}}(\ptablestore,\Xi,m)\stackrel{\triangle}{=} \textsf{ASpace\_Lookup}_{\textsf{id}}(\ptablestore,\Xi,m) \ast \mathsf{GhostMap}(\mathsf{id},\Xi)\ast\\
+   I\textsf{ASpace}_{\textsf{id}}(\ptablestore,\Xi,m)\stackrel{\triangle}{=} \textsf{ASpace\_Lookup}_{\textsf{id}}(\ptablestore,\Xi,m) \ast \mathsf{GhostMap}(\mathsf{id},\Xi)\ast\\
   \left(\bigast{(\vaddr, \textsf{paddr})\in \ptablestore}{\exists\;(\textsf{l4e, l3e, l2e, l1e, paddr})\ldotp \textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\vaddr\textsf{, l4e, l3e, l2e, l1e, paddr})}\right)\ast \\
@@
  \begin{array}{l}
-   \specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry} \textsf{\_} \ast \ulcorner \Xi !!
+   \specline{\textsf{P} \ast  I\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry} \textsf{\_} \ast \ulcorner \Xi !!
 \textsf{entry} = \textsf{Some V} \urcorner  }_{\rtv}  \\
 \sqsubseteq \\
-   \specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast  \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{next} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv} \\
+   \specline{\textsf{P} \ast  I\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast  \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{next} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv} \\
 \specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present entry\_val})\urcorner}_{\rtv} \\
 \mathsf{mov} \;   [ \mathsf{r8} ],\mathsf{rdi}\\
- \specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast  \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{next} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv} \\
+ \specline{\textsf{P} \ast  I\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast  \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{next} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv} \\
 \specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present entry\_val})\urcorner}_{\rtv} \\
@@
   \begin{lstlisting}[mathescape,escapeinside={(*}{*)}]
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)   \ast \texttt{rbp-16} \mapsto_{\textsf{r}} \textsf{\_} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry+KERNBASE}  \ast \texttt{rax} \mapsto_{\textsf{r}} \textsf{\_} \ast \lnot(\textsf{entry\_present entry\_val} }_{\rtv}$
+$\specline{\textsf{P} \ast  I\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)   \ast \texttt{rbp-16} \mapsto_{\textsf{r}} \textsf{\_} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry+KERNBASE}  \ast \texttt{rax} \mapsto_{\textsf{r}} \textsf{\_} \ast \lnot(\textsf{entry\_present entry\_val} }_{\rtv}$
 $\specline{\textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,1}} \textsf{entry entryv} \ast \ghostmaptoken{\textsf{id}}{(\mathsf{entry})}{\textsf{level}} }$
@@
 mov    -0x10[rbp],rax; mov    -0x8[rbp],rdi; mov    rax,[rdi]
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)   \ast \texttt{rbp-16} \mapsto_{\textsf{r}} \textsf{\_} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry+KERNBASE} \ast \texttt{rax} \mapsto_{\textsf{r}} \textsf{nextpaddr} \ast \lnot(\textsf{entry\_present entry\_val}) }$
+$\specline{\textsf{P} \ast  I\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)   \ast \texttt{rbp-16} \mapsto_{\textsf{r}} \textsf{\_} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry+KERNBASE} \ast \texttt{rax} \mapsto_{\textsf{r}} \textsf{nextpaddr} \ast \lnot(\textsf{entry\_present entry\_val}) }$
 $\specline{ \textsf{entry+KERNBASE} \mapsto_{\mathsf{vpte,1}} \mathsf{entry} \; \textsf{pfn\_set(entryv nextpaddr)}}$
@@
 ... ;; Continued from Figure (*\ref{fig:calltopteinitialize}*); assertions below specialized to non-allocating path for clarity
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \}),m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{rcx}  \mapsto_{\textsf{r}} \textsf{\_}  \ast \textsf{entry} \mapsto_{\textsf{id}} \textsf{\_} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
+$\specline{\textsf{P} \ast  I\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \}),m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{rcx}  \mapsto_{\textsf{r}} \textsf{\_}  \ast \textsf{entry} \mapsto_{\textsf{id}} \textsf{\_} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
 $\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{(pte\_initialized (entry\_val.pfn))} \urcorner }_{\rtv}$
@@
 ... ;; Stack setup
-$\specline{\textsf{P} \ast \ulcorner\textsf{rtv+KERNBASE}=\textsf{pml4}\urcorner \ast \forall_{i\in\textsf{0..511}}\ghostmaptoken{\textsf{id}}{(\textsf{rtv}+i*8)}{\textsf{4}} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}}_{\rtv}$
+$\specline{\textsf{P} \ast \ulcorner\textsf{rtv+KERNBASE}=\textsf{pml4}\urcorner \ast \forall_{i\in\textsf{0..511}}\ghostmaptoken{\textsf{id}}{(\textsf{rtv}+i*8)}{\textsf{4}} \ast  I\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}}_{\rtv}$
 ;;pte_t *pml4_entry = &pml4[PML4EX(va)]; // Virtual address of L4 entry(*\label{line:start_pml4_calc}*)
@@
 mov    rsi,-0x18[rbp] ;; Store virt. addr. to local var pml4_entry; logical pml4_entry is physical!
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{pml4\_entry}\},m) \ast \forall_{i\in\textsf{0..511}\setminus\{\textsf{PML4EX(va)}\}}\ghostmaptoken{\textsf{id}}{(\textsf{rtv}+i*8)}{\textsf{4}}\ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
+$\specline{\textsf{P} \ast  I\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{pml4\_entry}\},m) \ast \forall_{i\in\textsf{0..511}\setminus\{\textsf{PML4EX(va)}\}}\ghostmaptoken{\textsf{id}}{(\textsf{rtv}+i*8)}{\textsf{4}}\ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
 $\specline{\textsf{entry\_present(l4e\_val)} \wand \forall_{i\in\textsf{0..511}} \; \ghostmaptoken{\textsf{id}}{\textsf{table\_root(l4e\_val)+ i * 8}}{\textsf{3}} }_{\rtv} $
@@
 ;;save the physical next table address in rax
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{pml4\_entry}\},m)  \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
+$\specline{\textsf{P} \ast  I\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{pml4\_entry}\},m)  \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
 $\specline{\forall_{i\in\textsf{0..511}} \ghostmaptoken{\textsf{id}}{\textsf{table\_root(l4e\_val'.pfn) + i * 8}}{\textsf{3}}  \ast \ulcorner\textsf{pdp-KERNBASE}=\textsf{table\_root(l4e\_val'.pfn)}\urcorner} $
@@
 $\specline{\ulcorner\textsf{pdp}+\textsf{PDPEX(va)*8} = \textsf{pdp\_entry} \land \textsf{table\_root(l4e\_val'.pfn)}  = \textsf{pdp} \urcorner }_{\rtv}$
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{pml4\_entry},\textsf{pdp\_entry}\},m)  \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
+$\specline{\textsf{P} \ast  I\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{pml4\_entry},\textsf{pdp\_entry}\},m)  \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
 $\specline{\forall_{i\in\textsf{0..511}\setminus\{\textsf{PDPEX(va)}\}} \; \textsf{table\_root(pml4\_entry}_{\textsf{pfn}}) \textsf{ + i * 8} \mapsto_{\textsf{id}} \textsf{3} }_{\rtv}$
@@
 % ... ;; Stack setup
-% $\specline{\textsf{P} \ast \ulcorner\textsf{rtv+KERNBASE}=\textsf{pml4}\urcorner \ast \forall_{i\in\textsf{0..511}}\ghostmaptoken{\textsf{id}}{(\textsf{rtv}+i*8)}{\textsf{4}} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}}_{\rtv}$
+% $\specline{\textsf{P} \ast \ulcorner\textsf{rtv+KERNBASE}=\textsf{pml4}\urcorner \ast \forall_{i\in\textsf{0..511}}\ghostmaptoken{\textsf{id}}{(\textsf{rtv}+i*8)}{\textsf{4}} \ast  I\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}}_{\rtv}$
 % ;;set up the stack for root address and virtual address    
@@
 % mov    rsi,-0x18[rbp] ;; Store to local variable pml4_entry; logical pml4_entry is physical, program variable is virtual
-% $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{pml4\_entry}\},m) \ast \forall_{i\in\textsf{0..511}\setminus\{\textsf{PML4EX(va)}\}}\ghostmaptoken{\textsf{id}}{(\textsf{rtv}+i*8)}{\textsf{4}}\ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
+% $\specline{\textsf{P} \ast  I\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{pml4\_entry}\},m) \ast \forall_{i\in\textsf{0..511}\setminus\{\textsf{PML4EX(va)}\}}\ghostmaptoken{\textsf{id}}{(\textsf{rtv}+i*8)}{\textsf{4}}\ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
 % $\specline{\textsf{entry\_present(l4e\_val)} \wand \forall_{i\in\textsf{0..511}} \; \ghostmaptoken{\textsf{id}}{\textsf{table\_root(l4e\_val)+ i * 8}}{\textsf{3}} }_{\rtv} $
@@
 % ;;save the physical next table address in rax
-% $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{pml4\_entry}\},m)  \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
+% $\specline{\textsf{P} \ast  I\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{pml4\_entry}\},m)  \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
 % $\specline{\forall_{i\in\textsf{0..511}} \ghostmaptoken{\textsf{id}}{\textsf{table\_root(l4e\_val'.pfn) + i * 8}}{\textsf{3}}  \ast \ulcorner\textsf{pdp-KERNBASE}=\textsf{table\_root(l4e\_val'.pfn)}\urcorner} $
@@
 % $\specline{\ulcorner\textsf{pdp}+\textsf{PDPEX(va)*8} = \textsf{pdp\_entry} \land \textsf{table\_root(l4e\_val'.pfn)}  = \textsf{pdp} \urcorner }_{\rtv}$
-% $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{pml4\_entry},\textsf{pdp\_entry}\},m)  \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
+% $\specline{\textsf{P} \ast  I\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{pml4\_entry},\textsf{pdp\_entry}\},m)  \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
 % $\specline{\forall_{i\in\textsf{0..511}\setminus\{\textsf{PDPEX(va)}\}} \; \textsf{table\_root(pml4\_entry}_{\textsf{pfn}}) \textsf{ + i * 8} \mapsto_{\textsf{id}} \textsf{3} }_{\rtv}$
@@
 % ;;pte_t *pt = pte_get_next_table(pd_entry);
-% $\specline{ \left( \begin{array}{l} \textsf{pdp}+\textsf{PDPEX(va)} = \textsf{pdp\_entry} \land \textsf{table\_root(pml4\_entry}_{\textsf{pfn}})  = \textsf{pdp} \\ \textsf{pd}+\textsf{PDEX(va)} = \textsf{pd\_entry} \land \textsf{table\_root(pml3\_entry}_{\textsf{pfn}})  = \textsf{pd} \\ \textsf{pt}+\textsf{PDPEX(va)} = \textsf{pt\_entry} \land \textsf{table\_root(pml2\_entry}_{\textsf{pfn}})  = \textsf{pt} \\  \textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{pml4\_entry},\textsf{pdp\_entry},\textsf{pd\_entry},\textsf{pt\_entry}\},m)  \\ \ghostmaptoken{\delta{}s}{\rtv}{\delta} \\
+% $\specline{ \left( \begin{array}{l} \textsf{pdp}+\textsf{PDPEX(va)} = \textsf{pdp\_entry} \land \textsf{table\_root(pml4\_entry}_{\textsf{pfn}})  = \textsf{pdp} \\ \textsf{pd}+\textsf{PDEX(va)} = \textsf{pd\_entry} \land \textsf{table\_root(pml3\_entry}_{\textsf{pfn}})  = \textsf{pd} \\ \textsf{pt}+\textsf{PDPEX(va)} = \textsf{pt\_entry} \land \textsf{table\_root(pml2\_entry}_{\textsf{pfn}})  = \textsf{pt} \\  \textsf{P} \ast  I\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{pml4\_entry},\textsf{pdp\_entry},\textsf{pd\_entry},\textsf{pt\_entry}\},m)  \\ \ghostmaptoken{\delta{}s}{\rtv}{\delta} \\
 % \ghostmaptoken{\textsf{id}}{(\mathsf{pml4\_entry})}{\textsf{4}}\ast \ghostmaptoken{\textsf{id}}{(\mathsf{plm3\_entry})}{\textsf{3}} \ast \ghostmaptoken{\textsf{id}}{(\mathsf{plm2\_entry})}{\textsf{2}}  \\ \forall_{i\in\textsf{0..511}} \; \textsf{table\_root(pml4\_entry}_{\textsf{pfn}}) \textsf{ + i * 8} \mapsto_{\textsf{id}} \textsf{3}  \\ \forall_{i\in\textsf{0..511}} \; \textsf{table\_root(pdp\_entry}_{\textsf{pfn}}) \textsf{ + i * 8} \mapsto_{\textsf{id}} \textsf{2}  \\ \forall_{i\in\textsf{0..511}} \; \textsf{table\_root(pd\_entry}_{\textsf{pfn}}) \textsf{ + i * 8} \mapsto_{\textsf{id}} \textsf{1}  \\  \textsf{pml4\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l4e\_val} \\  \textsf{pdp\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l3e\_val} \\  \textsf{pd\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l2e\_val} \\   \textsf{pt\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l1e\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l1e\_val})\urcorner    \end{array} \right)  =\textsf{R}_{\textsf{walk}}}$
@@
   \vspace{-1em}
-$\begin{array}{l}\left( \begin{array}{l} \ulcorner \textsf{pdp}+\textsf{PDPEX(va)} = \textsf{pdp\_entry} \land \textsf{table\_root(pml4\_entry}_{\textsf{pfn}})  = \textsf{pdp} \\ \land \textsf{pd}+\textsf{PDEX(va)} = \textsf{pd\_entry} \land \textsf{table\_root(pml3\_entry}_{\textsf{pfn}})  = \textsf{pd} \\ \land \textsf{pt}+\textsf{PDPEX(va)} = \textsf{pt\_entry} \land \textsf{table\_root(pml2\_entry}_{\textsf{pfn}})  = \textsf{pt} \urcorner  \ast \\  \textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{pml4\_entry},\textsf{pdp\_entry},\textsf{pd\_entry},\textsf{pt\_entry}\},m)  \\ \ghostmaptoken{\delta{}s}{\rtv}{\delta} \ast
+$\begin{array}{l}\left( \begin{array}{l} \ulcorner \textsf{pdp}+\textsf{PDPEX(va)} = \textsf{pdp\_entry} \land \textsf{table\_root(pml4\_entry}_{\textsf{pfn}})  = \textsf{pdp} \\ \land \textsf{pd}+\textsf{PDEX(va)} = \textsf{pd\_entry} \land \textsf{table\_root(pml3\_entry}_{\textsf{pfn}})  = \textsf{pd} \\ \land \textsf{pt}+\textsf{PDPEX(va)} = \textsf{pt\_entry} \land \textsf{table\_root(pml2\_entry}_{\textsf{pfn}})  = \textsf{pt} \urcorner  \ast \\  \textsf{P} \ast  I\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{pml4\_entry},\textsf{pdp\_entry},\textsf{pd\_entry},\textsf{pt\_entry}\},m)  \\ \ghostmaptoken{\delta{}s}{\rtv}{\delta} \ast
       \ghostmaptoken{\textsf{id}}{(\mathsf{pml4\_entry})}{\textsf{4}}\ast \ghostmaptoken{\textsf{id}}{(\mathsf{plm3\_entry})}{\textsf{3}} \ast \ghostmaptoken{\textsf{id}}{(\mathsf{plm2\_entry})}{\textsf{2}}  \ast \\  \textsf{pml4\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{pml4\_entry} \; \textsf{l4e\_val}  \ast \\  \textsf{pdp\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{pdp\_entry}\;\textsf{l3e\_val}  \ast \\  \textsf{pd\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{pd\_entry} \; \textsf{l2e\_val}  \end{array} \right)  =\textsf{R}_{\textsf{walk} }
@@
 %...
-%$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)  \ast \texttt{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{\_} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{\_} \ast \sumpv\paddr\qfrac\true  \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
+%$\specline{\textsf{P} \ast  I\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)  \ast \texttt{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{\_} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{\_} \ast \sumpv\paddr\qfrac\true  \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
 %mov    rax,-0x18[rbp]
@@
 ... ;;setting up the stack      
- $\specline{\textsf{P} \ast \textsf{pml4} \mapsto_{\textsf{id}} \textsf{\_} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta} \ast \ulcorner\theta \; !!\; \vaddr = \texttt{None}\urcorner}_{\rtv}$
+ $\specline{\textsf{P} \ast \textsf{pml4} \mapsto_{\textsf{id}} \textsf{\_} \ast  I\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta} \ast \ulcorner\theta \; !!\; \vaddr = \texttt{None}\urcorner}_{\rtv}$
 ;; pte_t *pteaddr = walkpgdir(pml4, a);
@@
 % \begin{lstlisting}[mathescape,escapeinside={(*}{*)}]
-% $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,\Xi,m) \ast [\rtv'](\mathcal{I}\texttt{ASpace}(\theta',\Xi',m') \ast \texttt{Pother}) \ast \texttt{rsi}\mapsto_{\textsf{r}} new \ast \texttt{rdi}\mapsto_{\textsf{r}} old \ast \texttt{rbx}\mapsto_{\textsf{r}} \texttt{rbxv} }_{\rtv}$
+% $\specline{\textsf{P} \ast  I\texttt{ASpace}(\theta,\Xi,m) \ast [\rtv']( I\texttt{ASpace}(\theta',\Xi',m') \ast \texttt{Pother}) \ast \texttt{rsi}\mapsto_{\textsf{r}} new \ast \texttt{rdi}\mapsto_{\textsf{r}} old \ast \texttt{rbx}\mapsto_{\textsf{r}} \texttt{rbxv} }_{\rtv}$
 % $\specline{\texttt{rsp}\mapsto_{\textsf{r}} \texttt{rspv} \ast \texttt{rbp}\mapsto_{\textsf{r}} \texttt{rbpv} \ast \texttt{r12}\mapsto_{\textsf{r}} \texttt{r12v} \ast \texttt{r13}\mapsto_{\textsf{r}} \texttt{r13v} \ast \texttt{r14}\mapsto_{\textsf{r}} \texttt{r14v} \ast \texttt{r15}\mapsto_{\textsf{r}} \texttt{r15v}}_{\rtv}$
@@
 % mov 48[rdi], r15
-% $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,\Xi,m) \ast [\rtv'](\mathcal{I}\texttt{ASpace}(\theta',\Xi',m') \ast \texttt{Pother})  \ast  \texttt{rsi}\mapsto_{\textsf{r}} new \ast \texttt{rdi}\mapsto_{\textsf{r}} old \ast \texttt{rbx}\mapsto_{\textsf{r}} \texttt{rbxv} }_{\rtv}$
+% $\specline{\textsf{P} \ast  I\texttt{ASpace}(\theta,\Xi,m) \ast [\rtv']( I\texttt{ASpace}(\theta',\Xi',m') \ast \texttt{Pother})  \ast  \texttt{rsi}\mapsto_{\textsf{r}} new \ast \texttt{rdi}\mapsto_{\textsf{r}} old \ast \texttt{rbx}\mapsto_{\textsf{r}} \texttt{rbxv} }_{\rtv}$
 % $\specline{ \texttt{rsp}\mapsto_{\textsf{r}} \texttt{rspv} \ast \texttt{rbp}\mapsto_{\textsf{r}} \texttt{rbpv} \ast \texttt{r12}\mapsto_{\textsf{r}} \texttt{r12v} \ast \texttt{r13}\mapsto_{\textsf{r}} \texttt{r13v} \ast \texttt{r14}\mapsto_{\textsf{r}} \texttt{r14v} \ast \texttt{r15}\mapsto_{\textsf{r}} \texttt{r15v}}_{\rtv}$
@@
 % mov r15, 48[rsi]
-% $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,\Xi,m) \ast [\rtv'](\mathcal{I}\texttt{ASpace}(\theta',\Xi',m') \ast \texttt{Pother})  \ast  \texttt{rsi}\mapsto_{\textsf{r}} new \ast \texttt{rdi}\mapsto_{\textsf{r}} old \ast \texttt{rbx}\mapsto_{\textsf{r}} \texttt{rbxv}' }_{\rtv}$
+% $\specline{\textsf{P} \ast  I\texttt{ASpace}(\theta,\Xi,m) \ast [\rtv']( I\texttt{ASpace}(\theta',\Xi',m') \ast \texttt{Pother})  \ast  \texttt{rsi}\mapsto_{\textsf{r}} new \ast \texttt{rdi}\mapsto_{\textsf{r}} old \ast \texttt{rbx}\mapsto_{\textsf{r}} \texttt{rbxv}' }_{\rtv}$
 % $\specline{\texttt{rsp}\mapsto_{\textsf{r}} \texttt{rspv}' \ast \texttt{rbp}\mapsto_{\textsf{r}} \texttt{rbpv}' \ast \texttt{r12}\mapsto_{\textsf{r}} \texttt{r12v}' \ast \texttt{r13}\mapsto_{\textsf{r}} \texttt{r13v}' \ast \texttt{r14}\mapsto_{\textsf{r}} \texttt{r14v}' \ast \texttt{r15}\mapsto_{\textsf{r}} \texttt{r15v}'}_{\rtv}$
@@
 % mov cr3, 56[rsi] ;; <-- Switch to the new address space (*\label{line:end_restore}*)
-% $\specline{ [\rtv](\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,\Xi,m) \ast \mathsf{ContextAt}(old,[\mathsf{rbxv,\ldots,\rtv}]) \ast \mathsf{ContextAt}(new,[\mathsf{rbxv}',\ldots,\rtv']))  }_{\rtv}$ (*\label{line:modality_switch}*)
-% $\specline{ \mathcal{I}\texttt{ASpace}(\theta',\Xi',m') \ast \texttt{Pother} \ast  \texttt{rsi}\mapsto_{\textsf{r}} new \ast \texttt{rdi}\mapsto_{\textsf{r}} old \ast \ldots }_{\rtv'}$
+% $\specline{ [\rtv](\textsf{P} \ast  I\texttt{ASpace}(\theta,\Xi,m) \ast \mathsf{ContextAt}(old,[\mathsf{rbxv,\ldots,\rtv}]) \ast \mathsf{ContextAt}(new,[\mathsf{rbxv}',\ldots,\rtv']))  }_{\rtv}$ (*\label{line:modality_switch}*)
+% $\specline{  I\texttt{ASpace}(\theta',\Xi',m') \ast \texttt{Pother} \ast  \texttt{rsi}\mapsto_{\textsf{r}} new \ast \texttt{rdi}\mapsto_{\textsf{r}} old \ast \ldots }_{\rtv'}$
 % \end{lstlisting}
@@
 % the other-space modality 
-% $[\rtv'](\mathcal{I}\texttt{ASpace}(\theta,\Xi,m) \ast \texttt{Pother})$
+% $[\rtv']( I\texttt{ASpace}(\theta,\Xi,m) \ast \texttt{Pother})$
 % in the specfication. 
@@
 % \[  
-% $\specline{\exists (\entryf ,\;\entrytr,\; \entrytw,\; \entryo,\;\textsf{pte\_addr },\paddr) \; \ldotp\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast  \texttt{r14}\mapsto_{\textsf{r}} \_ \ast \texttt{rdi}\mapsto_{r} \vaddr \ast \texttt{rax}\mapsto_{\textsf{r}} \textsf{ pte\_addr} \; \ast }_{\rtv}$
+% $\specline{\exists (\entryf ,\;\entrytr,\; \entrytw,\; \entryo,\;\textsf{pte\_addr },\paddr) \; \ldotp\textsf{P} \ast  I\texttt{ASpace}(\theta,m) \ast  \texttt{r14}\mapsto_{\textsf{r}} \_ \ast \texttt{rdi}\mapsto_{r} \vaddr \ast \texttt{rax}\mapsto_{\textsf{r}} \textsf{ pte\_addr} \; \ast }_{\rtv}$
 % $\specline{ \ulcorner  \texttt{addr\_L1 }(\vaddr, \entryo) = \paddr \urcorner \ast \ulcorner \texttt{entry\_present } \entryf \land \texttt{entry\_present } \entrytr \land  \texttt{entry\_present } \entrytw \urcorner \; \ast}_{\rtv}$
@@
 % mov r14, rax ;; Save that before another call
-% $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast  \texttt{r14}\mapsto_{\textsf{r}} \texttt{pte\_addr} \ast \texttt{rdi}\mapsto_{\textsf{r}} \vaddr \ast \texttt{rax}\mapsto_{\textsf{r}} \textsf{ pte\_addr} \; \ast }_{\rtv}$
+% $\specline{\textsf{P} \ast  I\texttt{ASpace}(\theta,m) \ast  \texttt{r14}\mapsto_{\textsf{r}} \texttt{pte\_addr} \ast \texttt{rdi}\mapsto_{\textsf{r}} \vaddr \ast \texttt{rax}\mapsto_{\textsf{r}} \textsf{ pte\_addr} \; \ast }_{\rtv}$
 % $\specline{ \nfpointsto{\mask\vaddr\maskfour\rtv}{\mask\vaddr\maskfouroff\rtv}\entryf\qone\naddr \ast \ulcorner \texttt{entry\_present } \entryf \land \texttt{entry\_present } \entrytr \land  \texttt{entry\_present } \entrytw \urcorner \ast}_{\rtv}$
@@
 % call alloc_phys_page_or_panic
-% $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast  \texttt{r14}\mapsto_{\textsf{r}} \texttt{pte\_addr} \ast \texttt{rdi}\mapsto_{\textsf{r}} \vaddr \;\ast \nfpointsto{\mask\vaddr\maskfour\rtv}{\mask\vaddr\maskfouroff\rtv}\entryf\qone\naddr \ast}_{\rtv}$
+% $\specline{\textsf{P} \ast  I\texttt{ASpace}(\theta,m) \ast  \texttt{r14}\mapsto_{\textsf{r}} \texttt{pte\_addr} \ast \texttt{rdi}\mapsto_{\textsf{r}} \vaddr \;\ast \nfpointsto{\mask\vaddr\maskfour\rtv}{\mask\vaddr\maskfouroff\rtv}\entryf\qone\naddr \ast}_{\rtv}$
 % $\specline{  \nfpointsto{\mask\vaddr\maskthree\entryf}{\mask\vaddr\maskthreeoff\entryf}\entrytr\qtwo\naddr \ast \nfpointsto{\mask\vaddr\masktwo\entrytr}{\mask\vaddr\masktwooff\entrytr}\paddr\qthree\naddr \;\ast}_{\rtv}$
@@
 % mov [r14], rax ;; store the page table entry, mapping the page
-% $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast  \texttt{r14}\mapsto_{\textsf{r}} \texttt{pte\_addr} \ast \texttt{rdi}\mapsto_{\textsf{r}} \vaddr \;\ast \nfpointsto{\mask\vaddr\maskfour\rtv}{\mask\vaddr\maskfouroff\rtv}\entryf\qone\naddr \ast}_{\rtv}$
+% $\specline{\textsf{P} \ast  I\texttt{ASpace}(\theta,m) \ast  \texttt{r14}\mapsto_{\textsf{r}} \texttt{pte\_addr} \ast \texttt{rdi}\mapsto_{\textsf{r}} \vaddr \;\ast \nfpointsto{\mask\vaddr\maskfour\rtv}{\mask\vaddr\maskfouroff\rtv}\entryf\qone\naddr \ast}_{\rtv}$
 % $\specline{  \nfpointsto{\mask\vaddr\maskthree\entryf}{\mask\vaddr\maskthreeoff\entryf}\entrytr\qtwo\naddr \ast \nfpointsto{\mask\vaddr\masktwo\entrytr}{\mask\vaddr\masktwooff\entrytr}\paddr\qthree\entryo \;\ast}_{\rtv}$
@@
 % $\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\; \sqsubseteq $
-% $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast  \texttt{r14}\mapsto_{\textsf{r}} \texttt{pte\_addr} \ast \texttt{rdi}\mapsto_{\textsf{r}} \vaddr \ast }_{\rtv}$
+% $\specline{\textsf{P} \ast  I\texttt{ASpace}(\theta,m) \ast  \texttt{r14}\mapsto_{\textsf{r}} \texttt{pte\_addr} \ast \texttt{rdi}\mapsto_{\textsf{r}} \vaddr \ast }_{\rtv}$
 % $\specline{\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\vaddr,\entryf,\entrytr,\entrytw,\fpaddr+3) \ast \ulcorner \theta \;!!\;\vaddr = \texttt{None}\urcorner \; \ast}_{\rtv}$
@@
 % $\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\; \sqsubseteq $
-% $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace} (<[\vaddr:=\texttt{fpaddr}]> \theta,m) \ast}_{\rtv}$
+% $\specline{\textsf{P} \ast  I\texttt{ASpace} (<[\vaddr:=\texttt{fpaddr}]> \theta,m) \ast}_{\rtv}$
 % $\specline{\ulcorner \texttt{aligned fpaddr} \urcorner \ast \texttt{fpaddr} \mapsto_{\textsf{p}} \textsf{ wzero 64} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  \ast\sumwalkabs\vaddr\qfrac\fpaddr}_{\rtv}$
 % $\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\; \sqsubseteq $
-%   $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace} (<[\vaddr:=\texttt{fpaddr}]> \theta,m) \ast \vaddr \mapsto_{\textsf{vpte}}\; \{\qfrac\} \;\fpaddr \textsf{ wzero 64}}_{\rtv}$
+%   $\specline{\textsf{P} \ast  I\texttt{ASpace} (<[\vaddr:=\texttt{fpaddr}]> \theta,m) \ast \vaddr \mapsto_{\textsf{vpte}}\; \{\qfrac\} \;\fpaddr \textsf{ wzero 64}}_{\rtv}$
 % $\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\; \sqsubseteq $
-% $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace} (<[\vaddr:=\texttt{fpaddr}]> \theta,m) \ast \vaddr \mapsto_{\textsf{v}}\; \{\qfrac\} \textsf{wzero 64}}_{\rtv}$
+% $\specline{\textsf{P} \ast  I\texttt{ASpace} (<[\vaddr:=\texttt{fpaddr}]> \theta,m) \ast \vaddr \mapsto_{\textsf{v}}\; \{\qfrac\} \textsf{wzero 64}}_{\rtv}$
 % \end{comment}

### 2025-07-30 examples_C.tex
the magic of sed (i.e., replacing all the mathcal I with regular math mode I)

@@
 \begin{lstlisting}[mathescape,escapeinside={(*}{*)}]
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,\Xi,m) \ast [\rtv'](\mathcal{I}\texttt{ASpace}(\theta',\Xi',m') \ast \texttt{Pother}) \ast \texttt{rsi}\mapsto_{\textsf{r}} restore \ast \texttt{rdi}\mapsto_{\textsf{r}} save\ast \texttt{rbx}\mapsto_{\textsf{r}} \texttt{rbxv} }_{\rtv}$(*\label{line:start_swtch_pre}*)
+$\specline{\textsf{P} \ast  I\texttt{ASpace}(\theta,\Xi,m) \ast [\rtv']( I\texttt{ASpace}(\theta',\Xi',m') \ast \texttt{Pother}) \ast \texttt{rsi}\mapsto_{\textsf{r}} restore \ast \texttt{rdi}\mapsto_{\textsf{r}} save\ast \texttt{rbx}\mapsto_{\textsf{r}} \texttt{rbxv} }_{\rtv}$(*\label{line:start_swtch_pre}*)
 $\specline{\texttt{rsp}\mapsto_{\textsf{r}} \texttt{rspv} \ast \texttt{rbp}\mapsto_{\textsf{r}} \texttt{rbpv} \ast \texttt{r12}\mapsto_{\textsf{r}} \texttt{r12v} \ast \texttt{r13}\mapsto_{\textsf{r}} \texttt{r13v} \ast \texttt{r14}\mapsto_{\textsf{r}} \texttt{r14v} \ast \texttt{r15}\mapsto_{\textsf{r}} \texttt{r15v}}_{\rtv}$
@@
 ... mov 48[rdi], r15
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,\Xi,m) \ast [\rtv'](\mathcal{I}\texttt{ASpace}(\theta',\Xi',m') \ast \texttt{Pother})  \ast  \texttt{rsi}\mapsto_{\textsf{r}} restore \ast \texttt{rdi}\mapsto_{\textsf{r}} save \ast \texttt{rbx}\mapsto_{\textsf{r}} \texttt{rbxv} }_{\rtv}$
+$\specline{\textsf{P} \ast  I\texttt{ASpace}(\theta,\Xi,m) \ast [\rtv']( I\texttt{ASpace}(\theta',\Xi',m') \ast \texttt{Pother})  \ast  \texttt{rsi}\mapsto_{\textsf{r}} restore \ast \texttt{rdi}\mapsto_{\textsf{r}} save \ast \texttt{rbx}\mapsto_{\textsf{r}} \texttt{rbxv} }_{\rtv}$
 $\specline{ \texttt{rsp}\mapsto_{\textsf{r}} \texttt{rspv} \ast \texttt{rbp}\mapsto_{\textsf{r}} \texttt{rbpv} \ast \texttt{r12}\mapsto_{\textsf{r}} \texttt{r12v} \ast \texttt{r13}\mapsto_{\textsf{r}} \texttt{r13v} \ast \texttt{r14}\mapsto_{\textsf{r}} \texttt{r14v} \ast \texttt{r15}\mapsto_{\textsf{r}} \texttt{r15v}}_{\rtv}$
@@
 mov r15, 48[rsi]
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,\Xi,m) \ast [\rtv'](\mathcal{I}\texttt{ASpace}(\theta',\Xi',m') \ast \texttt{Pother})  \ast  \texttt{rsi}\mapsto_{\textsf{r}} restore \ast \texttt{rdi}\mapsto_{\textsf{r}} save \ast \texttt{rbx}\mapsto_{\textsf{r}} \texttt{rbxv}' }_{\rtv}$
+$\specline{\textsf{P} \ast  I\texttt{ASpace}(\theta,\Xi,m) \ast [\rtv']( I\texttt{ASpace}(\theta',\Xi',m') \ast \texttt{Pother})  \ast  \texttt{rsi}\mapsto_{\textsf{r}} restore \ast \texttt{rdi}\mapsto_{\textsf{r}} save \ast \texttt{rbx}\mapsto_{\textsf{r}} \texttt{rbxv}' }_{\rtv}$
 $\specline{\texttt{rsp}\mapsto_{\textsf{r}} \texttt{rspv}' \ast \texttt{rbp}\mapsto_{\textsf{r}} \texttt{rbpv}' \ast \texttt{r12}\mapsto_{\textsf{r}} \texttt{r12v}' \ast \texttt{r13}\mapsto_{\textsf{r}} \texttt{r13v}' \ast \texttt{r14}\mapsto_{\textsf{r}} \texttt{r14v}' \ast \texttt{r15}\mapsto_{\textsf{r}} \texttt{r15v}'}_{\rtv}$
@@
 mov cr3, 56[rsi] ;; <-- Switch to the new address space (*\label{line:end_restore}*)
-$\specline{ [\rtv](\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,\Xi,m) \ast \mathsf{ContextAt}(save,[\mathsf{rbxv,\ldots,\rtv}]) \ast \mathsf{ContextAt}(restore,[\mathsf{rbxv}',\ldots,\rtv']))  }_{\rtv}$ (*\label{line:modality_switchC}*)
-$\specline{ \mathcal{I}\texttt{ASpace}(\theta',\Xi',m') \ast \texttt{Pother} \ast  \texttt{rsi}\mapsto_{\textsf{r}} restore \ast \texttt{rdi}\mapsto_{\textsf{r}} save \ast \ldots }_{\rtv'}$
+$\specline{ [\rtv](\textsf{P} \ast  I\texttt{ASpace}(\theta,\Xi,m) \ast \mathsf{ContextAt}(save,[\mathsf{rbxv,\ldots,\rtv}]) \ast \mathsf{ContextAt}(restore,[\mathsf{rbxv}',\ldots,\rtv']))  }_{\rtv}$ (*\label{line:modality_switchC}*)
+$\specline{  I\texttt{ASpace}(\theta',\Xi',m') \ast \texttt{Pother} \ast  \texttt{rsi}\mapsto_{\textsf{r}} restore \ast \texttt{rdi}\mapsto_{\textsf{r}} save \ast \ldots }_{\rtv'}$
 \end{lstlisting}
@@
 % the other-space modality
-% $[\rtv'](\mathcal{I}\texttt{ASpace}(\theta,\Xi,m) \ast \texttt{Pother})$
+% $[\rtv']( I\texttt{ASpace}(\theta,\Xi,m) \ast \texttt{Pother})$
 % in the specfication. 
@@
   pte_t *next;
-  $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{entryp}\mapsto\textsf{v}\},m) \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta} \ast \mathsf{entry} \mapsto_r \mathsf{entryp+KERNBASE} \ast \ghostmaptoken{\textsf{id}}{(\mathsf{entryp})}{\textsf{level}}}_{\rtv}$(*\label{line:precondition_entry_out}*)
+  $\specline{\textsf{P} \ast  I\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{entryp}\mapsto\textsf{v}\},m) \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta} \ast \mathsf{entry} \mapsto_r \mathsf{entryp+KERNBASE} \ast \ghostmaptoken{\textsf{id}}{(\mathsf{entryp})}{\textsf{level}}}_{\rtv}$(*\label{line:precondition_entry_out}*)
   $\specline{ \textsf{entryp+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entryp entryv} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present entryv})\urcorner}_{\rtv}$ (*\label{line:get_next_vpte_preconditionC}*)
@@
   } (*\label{line:end_of_allocation_pathC}*)
-  $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entryp}\mapsto\textsf{v} \}),m) \ast \textsf{entryp} \mapsto_{\textsf{id}} \textsf{\_} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
+  $\specline{\textsf{P} \ast  I\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entryp}\mapsto\textsf{v} \}),m) \ast \textsf{entryp} \mapsto_{\textsf{id}} \textsf{\_} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
   $\specline{ \textsf{entryp+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entryp} \; \textsf{(pte\_initialized (entryv.pfn))} \urcorner }_{\rtv}$
@@
 \begin{array}{l}
-  \mathcal{I}\textsf{ASpace}_{\textsf{id}}(\ptablestore,\Xi,m)\stackrel{\triangle}{=} \textsf{ASpace\_Lookup}_{\textsf{id}}(\ptablestore,\Xi,m) \ast\\  %\mathsf{GhostMap}(\mathsf{id},\Xi)\ast\\
+   I\textsf{ASpace}_{\textsf{id}}(\ptablestore,\Xi,m)\stackrel{\triangle}{=} \textsf{ASpace\_Lookup}_{\textsf{id}}(\ptablestore,\Xi,m) \ast\\  %\mathsf{GhostMap}(\mathsf{id},\Xi)\ast\\
   \left(\bigast{(\vaddr, \textsf{paddr})\in \ptablestore}{\exists\;(\textsf{l4e, l3e, l2e, l1e, paddr})\ldotp \textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\vaddr\textsf{, l4e, l3e, l2e, l1e, paddr})}\right)\ast \\
@@
 /* @param entry: virtual address of a non-present entry to initialize with a new physical page */
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \lnot(\textsf{entry\_present entryv}) \ast \textsf{entryp+KERNBASE} \mapsto_{\textsf{vpte,1}} \textsf{entryp entryv} \ast \ghostmaptoken{\textsf{id}}{(\mathsf{entryp})}{\textsf{level}} }$
+$\specline{\textsf{P} \ast  I\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \lnot(\textsf{entry\_present entryv}) \ast \textsf{entryp+KERNBASE} \mapsto_{\textsf{vpte,1}} \textsf{entryp entryv} \ast \ghostmaptoken{\textsf{id}}{(\mathsf{entryp})}{\textsf{level}} }$
 void pte_initialize(pte_t *entry) {
@@
   entry->pfn = PTE_ADDR_TO_PFN((uintptr_t) local);(*\label{line:kalloc_install}*)
-  $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \lnot(\textsf{entry\_present entryv}) \ast \textsf{entryp+KERNBASE} \mapsto_{\mathsf{vpte,1}} \mathsf{entryp} \; \textsf{pfn\_set(entryv nextpaddr)}}$
+  $\specline{\textsf{P} \ast  I\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \lnot(\textsf{entry\_present entryv}) \ast \textsf{entryp+KERNBASE} \mapsto_{\mathsf{vpte,1}} \mathsf{entryp} \; \textsf{pfn\_set(entryv nextpaddr)}}$
   $\specline{\ghostmaptoken{\delta{}s}{\rtv}{\delta} \ast \forall_{i\in\textsf{0..511}}\ldotp ()\textsf{table\_root (pte\_initialized (pfn\_set (entry\_val nextpaddr))) + i * 8}) \mapsto_{\textsf{id}} \textsf{level-1}   }_{\rtv}$(*\label{line:initialize_post}*)A
@@
 pte_t *walkpgdir(pte_t *l4, const void *va) {
-  $\specline{\textsf{P} \ast \ulcorner\textsf{l4p}=\textsf{rtv}\urcorner\ast\ulcorner\textsf{rtv+KERNBASE}=\textsf{l4}\urcorner \ast \ghostmaptoken{\textsf{id}}{(\textsf{l4p}+8*\textsf{L4Offset}(\textsf{va}))}{\textsf{4}} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}}_{\rtv}$(*\label{line:walkpgdir_pre}*)
+  $\specline{\textsf{P} \ast \ulcorner\textsf{l4p}=\textsf{rtv}\urcorner\ast\ulcorner\textsf{rtv+KERNBASE}=\textsf{l4}\urcorner \ast \ghostmaptoken{\textsf{id}}{(\textsf{l4p}+8*\textsf{L4Offset}(\textsf{va}))}{\textsf{4}} \ast  I\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}}_{\rtv}$(*\label{line:walkpgdir_pre}*)
   pte_t *l4_entry = &l4[L4Offset(va)]; // Virtual address of L4 entry(*\label{line:start_l4_calcC}*)
-  $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{l4p+8*L4Offset(va)}\},m) \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
+  $\specline{\textsf{P} \ast  I\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{l4p+8*L4Offset(va)}\},m) \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
   $\specline{\textsf{entry\_present(l4e\_val)} \wand \forall_{i\in\textsf{0..511}} \; \ghostmaptoken{\textsf{id}}{(\textsf{table\_root(l4e\_val)+ i * 8})}{\textsf{3}} }_{\rtv} $
@@
   /*pte_get_next_table may have allocated a new page which updates entry value*/
-  $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{l4p+8*L4Offset(va)}\},m)  \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
+  $\specline{\textsf{P} \ast  I\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{l4p+8*L4Offset(va)}\},m)  \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
   $\specline{\forall_{i\in\textsf{0..511}} \ghostmaptoken{\textsf{id}}{(\textsf{table\_root(l4e\_val'.pfn) + i * 8})}{\textsf{3}}  \ast \ulcorner\textsf{l3-KERNBASE}=\textsf{table\_root(l4e\_val'.pfn)}\urcorner} $(*\label{line:l3tokens}*)
@@
   $\specline{\ulcorner\textsf{l3}+\textsf{L3Offset(va)*8} = \textsf{l3p} \land \textsf{table\_root(l4e\_val'.pfn)}  = \textsf{l3} \urcorner }_{\rtv}$
-  $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{l4p+8*L4Offset(va)},\textsf{l3p+8*L3Offset(va)}\},m)  \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
+  $\specline{\textsf{P} \ast  I\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{l4p+8*L4Offset(va)},\textsf{l3p+8*L3Offset(va)}\},m)  \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
   $\specline{\forall_{i\in\textsf{0..511}\setminus\{\textsf{L3Offset(va)}\}} \; \ghostmaptoken{\textsf{id}}{(\textsf{table\_root(l4e\_val.pfn}) \textsf{ + i * 8})} {\textsf{3}} }_{\rtv}$
@@
 % ... ;; Stack setup
-% $\specline{\textsf{P} \ast \ulcorner\textsf{rtv+KERNBASE}=\textsf{l4}\urcorner \ast \forall_{i\in\textsf{0..511}}\ghostmaptoken{\textsf{id}}{(\textsf{rtv}+i*8)}{\textsf{4}} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}}_{\rtv}$
+% $\specline{\textsf{P} \ast \ulcorner\textsf{rtv+KERNBASE}=\textsf{l4}\urcorner \ast \forall_{i\in\textsf{0..511}}\ghostmaptoken{\textsf{id}}{(\textsf{rtv}+i*8)}{\textsf{4}} \ast  I\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}}_{\rtv}$
 % ;;set up the stack for root address and virtual address    
@@
 % mov    rsi,-0x18[rbp] ;; Store to local variable l4_entry; logical l4_entry is physical, program variable is virtual
-% $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{l4\_entry}\},m) \ast \forall_{i\in\textsf{0..511}\setminus\{\textsf{L4EX(va)}\}}\ghostmaptoken{\textsf{id}}{(\textsf{rtv}+i*8)}{\textsf{4}}\ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
+% $\specline{\textsf{P} \ast  I\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{l4\_entry}\},m) \ast \forall_{i\in\textsf{0..511}\setminus\{\textsf{L4EX(va)}\}}\ghostmaptoken{\textsf{id}}{(\textsf{rtv}+i*8)}{\textsf{4}}\ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
 % $\specline{\textsf{entry\_present(l4e\_val)} \wand \forall_{i\in\textsf{0..511}} \; \ghostmaptoken{\textsf{id}}{\textsf{table\_root(l4e\_val)+ i * 8}}{\textsf{3}} }_{\rtv} $
@@
 % ;;save the physical next table address in rax
-% $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{l4\_entry}\},m)  \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
+% $\specline{\textsf{P} \ast  I\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{l4\_entry}\},m)  \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
 % $\specline{\forall_{i\in\textsf{0..511}} \ghostmaptoken{\textsf{id}}{\textsf{table\_root(l4e\_val'.pfn) + i * 8}}{\textsf{3}}  \ast \ulcorner\textsf{pdp-KERNBASE}=\textsf{table\_root(l4e\_val'.pfn)}\urcorner} $
@@
 % $\specline{\ulcorner\textsf{pdp}+\textsf{PDPEX(va)*8} = \textsf{pdp\_entry} \land \textsf{table\_root(l4e\_val'.pfn)}  = \textsf{pdp} \urcorner }_{\rtv}$
-% $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{l4\_entry},\textsf{pdp\_entry}\},m)  \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
+% $\specline{\textsf{P} \ast  I\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{l4\_entry},\textsf{pdp\_entry}\},m)  \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
 % $\specline{\forall_{i\in\textsf{0..511}\setminus\{\textsf{PDPEX(va)}\}} \; \textsf{table\_root(l4\_entry}_{\textsf{pfn}}) \textsf{ + i * 8} \mapsto_{\textsf{id}} \textsf{3} }_{\rtv}$
@@
 % ;;pte_t *pt = pte_get_next_table(pd_entry);
-% $\specline{ \left( \begin{array}{l} \textsf{pdp}+\textsf{PDPEX(va)} = \textsf{pdp\_entry} \land \textsf{table\_root(l4\_entry}_{\textsf{pfn}})  = \textsf{pdp} \\ \textsf{pd}+\textsf{PDEX(va)} = \textsf{pd\_entry} \land \textsf{table\_root(pml3\_entry}_{\textsf{pfn}})  = \textsf{pd} \\ \textsf{pt}+\textsf{PDPEX(va)} = \textsf{pt\_entry} \land \textsf{table\_root(pml2\_entry}_{\textsf{pfn}})  = \textsf{pt} \\  \textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{l4\_entry},\textsf{pdp\_entry},\textsf{pd\_entry},\textsf{pt\_entry}\},m)  \\ \ghostmaptoken{\delta{}s}{\rtv}{\delta} \\
+% $\specline{ \left( \begin{array}{l} \textsf{pdp}+\textsf{PDPEX(va)} = \textsf{pdp\_entry} \land \textsf{table\_root(l4\_entry}_{\textsf{pfn}})  = \textsf{pdp} \\ \textsf{pd}+\textsf{PDEX(va)} = \textsf{pd\_entry} \land \textsf{table\_root(pml3\_entry}_{\textsf{pfn}})  = \textsf{pd} \\ \textsf{pt}+\textsf{PDPEX(va)} = \textsf{pt\_entry} \land \textsf{table\_root(pml2\_entry}_{\textsf{pfn}})  = \textsf{pt} \\  \textsf{P} \ast  I\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{l4\_entry},\textsf{pdp\_entry},\textsf{pd\_entry},\textsf{pt\_entry}\},m)  \\ \ghostmaptoken{\delta{}s}{\rtv}{\delta} \\
 % \ghostmaptoken{\textsf{id}}{(\mathsf{l4\_entry})}{\textsf{4}}\ast \ghostmaptoken{\textsf{id}}{(\mathsf{plm3\_entry})}{\textsf{3}} \ast \ghostmaptoken{\textsf{id}}{(\mathsf{plm2\_entry})}{\textsf{2}}  \\ \forall_{i\in\textsf{0..511}} \; \textsf{table\_root(l4\_entry}_{\textsf{pfn}}) \textsf{ + i * 8} \mapsto_{\textsf{id}} \textsf{3}  \\ \forall_{i\in\textsf{0..511}} \; \textsf{table\_root(pdp\_entry}_{\textsf{pfn}}) \textsf{ + i * 8} \mapsto_{\textsf{id}} \textsf{2}  \\ \forall_{i\in\textsf{0..511}} \; \textsf{table\_root(pd\_entry}_{\textsf{pfn}}) \textsf{ + i * 8} \mapsto_{\textsf{id}} \textsf{1}  \\  \textsf{l4\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l4e\_val} \\  \textsf{pdp\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l3e\_val} \\  \textsf{pd\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l2e\_val} \\   \textsf{pt\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l1e\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l1e\_val})\urcorner    \end{array} \right)  =\textsf{R}_{\textsf{walk}}}$
@@
     \land \textsf{table\_root(l2e\_val}_{\textsf{pfn}})  = \textsf{l1} \urcorner  \ast \\
-    % \textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{l4p},\textsf{l3p},\textsf{l2p},\textsf{l1p}\},m)  \ast  
-    \textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{l4p+8*L4Offset(va)},\textsf{l3p+8*L3Offset(va)},\textsf{l2p+8*L2Offset(va)},\textsf{l1p+8*L1Offset(va)}\},m)  \ast  \\
+    % \textsf{P} \ast  I\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{l4p},\textsf{l3p},\textsf{l2p},\textsf{l1p}\},m)  \ast  
+    \textsf{P} \ast  I\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{l4p+8*L4Offset(va)},\textsf{l3p+8*L3Offset(va)},\textsf{l2p+8*L2Offset(va)},\textsf{l1p+8*L1Offset(va)}\},m)  \ast  \\
       \ghostmaptoken{\textsf{id}}{(\mathsf{l4p+8*L4Offset(va)})}{\textsf{4}}\ast \ghostmaptoken{\textsf{id}}{(\mathsf{l3p+8*L3Offset(va)})}{\textsf{3}} \ast \\ \ghostmaptoken{\textsf{id}}{(\mathsf{l2p+8*L2Offset(va)})}{\textsf{2}}  \ast \ghostmaptoken{\textsf{id}}{(\mathsf{l1p+8*L1Offset(va)})}{\textsf{1}} \ast \\
@@
 void vaspace_mappage(pte_t *l4, void *va,uintptr_t fpaddr ) {
-  $\specline{\textsf{P} \ast \ghostmaptoken{\mathsf{\textsf{id}}}{(\textsf{l4p+8*L4Offset(va)})}{4} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta} \ast \ulcorner\theta \; !!\; \vaddr = \texttt{None}\urcorner}_{\rtv}$ (*\label{line:mappage_pre}*)
+  $\specline{\textsf{P} \ast \ghostmaptoken{\mathsf{\textsf{id}}}{(\textsf{l4p+8*L4Offset(va)})}{4} \ast  I\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta} \ast \ulcorner\theta \; !!\; \vaddr = \texttt{None}\urcorner}_{\rtv}$ (*\label{line:mappage_pre}*)
   pte_t *pteaddr = walkpgdir(l4, va); (*\label{line:call_walkpgdirC}*)

### 2025-07-30 modal_abs_vmm_paper.tex
fixing arxiv mov instrs
enable todonotes
the magic of sed (i.e., replacing all the mathcal I with regular math mode I)

--- 2023-07-26 vmm_arxiv.tex
@@
 \usepackage{array}
-%
+\usepackage{todonotes}
 \usepackage{proof}
@@
 We place the authorative ownership $\mathcal{A}\textsf{PTableWalk}$ of the ghost-map in a per-address-space invariant
-($\mathcal{I}$\textsf{ASpace} in Figure \ref{fig:peraspaceinvariant}) whose ownership, when combined with the
+($ I$\textsf{ASpace} in Figure \ref{fig:peraspaceinvariant}) whose ownership, when combined with the
 fraction stored in the virtual points-to, permits changes in the virtual-to-physical translations.
@@
 \begin{array}{l}
-  \mathcal{I}\textsf{ASpace}(\ptablestore,m)\stackrel{\triangle}{=} \textsf{ASpace\_Lookup}(\ptablestore,m) \ast \\
+   I\textsf{ASpace}(\ptablestore,m)\stackrel{\triangle}{=} \textsf{ASpace\_Lookup}(\ptablestore,m) \ast \\
  \qquad\qquad\qquad \bigast{(\vaddr, \textsf{paddr})\in \ptablestore}{\exists\;(\textsf{l4e l3e l2e, l1e, paddr})\ldotp \textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\vaddr\textsf{, l4e, l3e, l2e, l1e, paddr})} \\
@@
 
-$\mathcal{I}\textsf{ASpace}(\Theta,m)$ is then the resources for the current address space:
+$ I\textsf{ASpace}(\Theta,m)$ is then the resources for the current address space:
 It ensures that the current address space is valid (it has an entry in $m$, whose corresponding
@@
 The invariant for the currently active address space (solid round square labelled as 
-$\mathcal{I}$\textsf{ASpace} in Figure \ref{fig:logicaladdrspace}) asserts authorative ownership (i.e. 
+$ I$\textsf{ASpace} in Figure \ref{fig:logicaladdrspace}) asserts authorative ownership (i.e. 
 update capability), $\ownGhost\gammaPred{\authfull{\ptableabswalk\ptablestore}}$, of the ghost map $\Theta$
@@
 shared invariant, and for each memory access,
-a fragment of this corresponding to $\mathcal{I}\textsf{ASpace}(\Theta,m)$
+a fragment of this corresponding to $ I\textsf{ASpace}(\Theta,m)$
 for the current address space should be extracted from this global resource, used to prove correctness of the
@@
 \inferrule[WriteToRegFromVirtMem]{
-  \{P \ast \mathcal{I}\textsf{ASpace}\ast \mathcal{I}\textsf{ASpace} \ast r_d \mapsto_{r}  \textsf{v} \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{v,rtv}} \textsf{v} \}_{\textsf{rtv}}\;\overline{is}
+  \{P \ast  I\textsf{ASpace}\ast  I\textsf{ASpace} \ast r_d \mapsto_{r}  \textsf{v} \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{v,rtv}} \textsf{v} \}_{\textsf{rtv}}\;\overline{is}
 }{
-  \{P \ast \mathcal{I}\textsf{ASpace}\ast r_d \mapsto_{r}  \textsf{rvd} \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{v}} \textsf{v} \}_{\textsf{rtv}}
+  \{P \ast  I\textsf{ASpace}\ast r_d \mapsto_{r}  \textsf{rvd} \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{v}} \textsf{v} \}_{\textsf{rtv}}
   \textsf{mov}~\textsf{r}_d,~[\textsf{r}_a];\;\overline{is}
@@
 \inferrule[WriteToVirtMemFromReg]{
-  \{P \ast \mathcal{I}\textsf{ASpace}\ast r_s \mapsto_{r}\{q\}  \textsf{rvs}  \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{v}} \textsf{v} \}_{\textsf{rtv}}\;\overline{ is}
+  \{P \ast  I\textsf{ASpace}\ast r_s \mapsto_{r}\{q\}  \textsf{rvs}  \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{v}} \textsf{v} \}_{\textsf{rtv}}\;\overline{ is}  
 }{
-  \{P \ast \mathcal{I}\textsf{ASpace}\ast r_s \mapsto_{r}\{q\}  \textsf{rvs}   \ast r_a \mapsto_{r}\{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{v}} \textsf{rvs} \}_{\textsf{rtv}}
+  \{P \ast  I\textsf{ASpace}\ast r_s \mapsto_{r}\{q\}  \textsf{rvs}   \ast r_a \mapsto_{r}\{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{v}} \textsf{rvs} \}_{\textsf{rtv}}
   \textsf{ mov}~[\textsf{r}_a],~\textsf{r}_s;\;\overline{is}
@@
 \inferrule[WriteToRegCtlFromRegModal]{
-  \{[\textsf{rtv}](P \ast \mathcal{I}\textsf{ASpace})\ast \mathcal{I}\textsf{ASpace} \ast R\ast r_s \mapsto_{r}\{q\}  \textsf{ rvs}  \}_{\textsf{rvs}} \overline{is}
+  \{[\textsf{rtv}](P \ast  I\textsf{ASpace})\ast  I\textsf{ASpace} \ast R\ast r_s \mapsto_{r}\{q\}  \textsf{ rvs}  \}_{\textsf{rvs}} \overline{is}
 }{
-  \{P \ast \mathcal{I}\textsf{ASpace} \ast [\textsf{rvs}](R \ast\mathcal{I}\textsf{ASpace})\ast r_s \mapsto_{r}\{q\}  \textsf{ rvs}   \}_{\textsf{rtv}}
+  \{P \ast  I\textsf{ASpace} \ast [\textsf{rvs}](R \ast I\textsf{ASpace})\ast r_s \mapsto_{r}\{q\}  \textsf{ rvs}   \}_{\textsf{rtv}}
   \textsf{mov}~\textsf{cr3},~r_s;\;\overline{is}
@@
 virtual points-to itself (in order to permit the physical page tables to be freely modified
-as long as they preserve virtual-to-physical translations), the rule requires $\mathcal{I}\textsf{ASpace}$
+as long as they preserve virtual-to-physical translations), the rule requires $ I\textsf{ASpace}$
 for the current address space to be carried through.
@@
   \begin{lstlisting}
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m)  \ast  \texttt{r14}\mapsto_{\textsf{r}} \_ \ast \texttt{rdi}\mapsto_{r} \vaddr \ast \texttt{rax}\mapsto_{\textsf{r}} \_ \ast \ulcorner \texttt{aligned va} \land  \theta \; !!\; \vaddr = \texttt{None}\urcorner}_{\rtv}$
+$\specline{\textsf{P} \ast  I\texttt{ASpace}(\theta,m)  \ast  \texttt{r14}\mapsto_{\textsf{r}} \_ \ast \texttt{rdi}\mapsto_{r} \vaddr \ast \texttt{rax}\mapsto_{\textsf{r}} \_ \ast \ulcorner \texttt{aligned va} \land  \theta \; !!\; \vaddr = \texttt{None}\urcorner}_{\rtv}$
 call ensure_L1_page
-$\specline{\exists (\entryf ,\;\entrytr,\; \entrytw,\; \entryo,\;\textsf{pte\_addr },\paddr) \; \ldotp\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast  \texttt{r14}\mapsto_{\textsf{r}} \_ \ast \texttt{rdi}\mapsto_{r} \vaddr \ast \texttt{rax}\mapsto_{\textsf{r}} \textsf{ pte\_addr} \; \ast }_{\rtv}$
+$\specline{\exists (\entryf ,\;\entrytr,\; \entrytw,\; \entryo,\;\textsf{pte\_addr },\paddr) \; \ldotp\textsf{P} \ast  I\texttt{ASpace}(\theta,m) \ast  \texttt{r14}\mapsto_{\textsf{r}} \_ \ast \texttt{rdi}\mapsto_{r} \vaddr \ast \texttt{rax}\mapsto_{\textsf{r}} \textsf{ pte\_addr} \; \ast }_{\rtv}$
 $\specline{ \ulcorner  \texttt{addr\_L1 }(\vaddr, \entryo) = \paddr \urcorner \ast \ulcorner \texttt{entry\_present } \entryf \land \texttt{entry\_present } \entrytr \land  \texttt{entry\_present } \entrytw \urcorner \; \ast}_{\rtv}$
@@
 ;; Returns the virtual address of the L1 entry in rax
-mov %
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast  \texttt{r14}\mapsto_{\textsf{r}} \texttt{pte\_addr} \ast \texttt{rdi}\mapsto_{\textsf{r}} \vaddr \ast \texttt{rax}\mapsto_{\textsf{r}} \textsf{ pte\_addr} \; \ast }_{\rtv}$
+mov %r14, rax ;; Save that before another call
+$\specline{\textsf{P} \ast  I\texttt{ASpace}(\theta,m) \ast  \texttt{r14}\mapsto_{\textsf{r}} \texttt{pte\_addr} \ast \texttt{rdi}\mapsto_{\textsf{r}} \vaddr \ast \texttt{rax}\mapsto_{\textsf{r}} \textsf{ pte\_addr} \; \ast }_{\rtv}$
 $\specline{ \nfpointsto{\mask\vaddr\maskfour\rtv}{\mask\vaddr\maskfouroff\rtv}\entryf\qone\naddr \ast \ulcorner \texttt{entry\_present } \entryf \land \texttt{entry\_present } \entrytr \land  \texttt{entry\_present } \entrytw \urcorner \ast}_{\rtv}$
@@
 call alloc_phys_page_or_panic
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast  \texttt{r14}\mapsto_{\textsf{r}} \texttt{pte\_addr} \ast \texttt{rdi}\mapsto_{\textsf{r}} \vaddr \;\ast}_{\rtv}$
+$\specline{\textsf{P} \ast  I\texttt{ASpace}(\theta,m) \ast  \texttt{r14}\mapsto_{\textsf{r}} \texttt{pte\_addr} \ast \texttt{rdi}\mapsto_{\textsf{r}} \vaddr \;\ast}_{\rtv}$
 $\specline{ \nfpointsto{\mask\vaddr\maskfour\rtv}{\mask\vaddr\maskfouroff\rtv}\entryf\qone\naddr \ast}_{\rtv}$ 
@@
 ;; update the page table entry, mapping the page
-mov (%
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast  \texttt{r14}\mapsto_{\textsf{r}} \texttt{pte\_addr} \ast \texttt{rdi}\mapsto_{\textsf{r}} \vaddr \;\ast}_{\rtv}$
+mov [r14], rax
+$\specline{\textsf{P} \ast  I\texttt{ASpace}(\theta,m) \ast  \texttt{r14}\mapsto_{\textsf{r}} \texttt{pte\_addr} \ast \texttt{rdi}\mapsto_{\textsf{r}} \vaddr \;\ast}_{\rtv}$
 $\specline{ \nfpointsto{\mask\vaddr\maskfour\rtv}{\mask\vaddr\maskfouroff\rtv}\entryf\qone\naddr \ast}_{\rtv}$ 
@@
 $\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\; \sqsubseteq $
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast  \texttt{r14}\mapsto_{\textsf{r}} \texttt{pte\_addr} \ast \texttt{rdi}\mapsto_{\textsf{r}} \vaddr \ast }_{\rtv}$
+$\specline{\textsf{P} \ast  I\texttt{ASpace}(\theta,m) \ast  \texttt{r14}\mapsto_{\textsf{r}} \texttt{pte\_addr} \ast \texttt{rdi}\mapsto_{\textsf{r}} \vaddr \ast }_{\rtv}$
 $\specline{\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\vaddr,\entryf,\entrytr,\entrytw,\fpaddr+3) \ast \ulcorner \theta \;!!\;\vaddr = \texttt{None}\urcorner \; \ast}_{\rtv}$
@@
 $\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\; \sqsubseteq $
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace} (<[\vaddr:=\texttt{fpaddr}]> \theta,m) \ast}_{\rtv}$
+$\specline{\textsf{P} \ast  I\texttt{ASpace} (<[\vaddr:=\texttt{fpaddr}]> \theta,m) \ast}_{\rtv}$
 $\specline{\ulcorner \texttt{aligned fpaddr} \urcorner \ast \texttt{fpaddr} \mapsto_{\textsf{p}} \textsf{ wzero 64} \ast \sumapaces\rtv\delta  \ast\sumwalkabs\vaddr\qfrac\fpaddr}_{\rtv}$
 $\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\; \sqsubseteq $
-  $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace} (<[\vaddr:=\texttt{fpaddr}]> \theta,m) \ast \vaddr \mapsto_{\textsf{vpte}}\; \{\qfrac\} \;\fpaddr \textsf{ wzero 64}}_{\rtv}$
+  $\specline{\textsf{P} \ast  I\texttt{ASpace} (<[\vaddr:=\texttt{fpaddr}]> \theta,m) \ast \vaddr \mapsto_{\textsf{vpte}}\; \{\qfrac\} \;\fpaddr \textsf{ wzero 64}}_{\rtv}$
 $\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\; \sqsubseteq $
-  $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace} (<[\vaddr:=\texttt{fpaddr}]> \theta,m) \ast \vaddr \mapsto_{\textsf{v}}\; \{\qfrac\} \textsf{wzero 64}}_{\rtv}$
+  $\specline{\textsf{P} \ast  I\texttt{ASpace} (<[\vaddr:=\texttt{fpaddr}]> \theta,m) \ast \vaddr \mapsto_{\textsf{v}}\; \{\qfrac\} \textsf{wzero 64}}_{\rtv}$
 \end{lstlisting}
@@
 ;; Save yielding context
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast [\rtv'+56](\mathcal{I}\texttt{ASpace}(\theta,m) \ast \texttt{Pother})}_{\rtv}$
+$\specline{\textsf{P} \ast  I\texttt{ASpace}(\theta,m) \ast [\rtv'+56]( I\texttt{ASpace}(\theta,m) \ast \texttt{Pother})}_{\rtv}$
 $\specline{ \texttt{rsi}\mapsto_{\textsf{r}} \rtv' \ast \texttt{rdi}\mapsto_{\textsf{r}} \_ \ast \texttt{rbx}\mapsto_{\textsf{r}} \texttt{rbxv} \ast  \texttt{rsp}\mapsto_{\textsf{r}} \texttt{rspv} \ast \texttt{rbp}\mapsto_{\textsf{r}} \texttt{rbpv} }_{\rtv}$
@@
 $\specline{\texttt{rdi+40} \mapsto_{\textsf{v}} \_ \ast \texttt{rdi+48} \mapsto_{\textsf{v}} \_\ast \texttt{rdi+56} \mapsto_{\textsf{v}} \_}_{\rtv}$
-mov 0[%
-mov 8[%
-mov 16[%
-mov 24[%
-mov 32[%
-mov 40[%
-mov 48[%
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast [\rtv'+56](\mathcal{I}\texttt{ASpace}(\theta,m) \ast \texttt{Pother})}_{\rtv}$
+mov 0[rdi], rbx
+mov 8[rdi], rsp
+mov 16[rdi], rbp
+mov 24[rdi], r12
+mov 32[rdi], r13
+mov 40[rdi], r14
+mov 48[rdi], r15$\specline{\textsf{P} \ast  I\texttt{ASpace}(\theta,m) \ast [\rtv'+56]( I\texttt{ASpace}(\theta,m) \ast \texttt{Pother})}_{\rtv}$
 $\specline{ \texttt{rdi}\mapsto_{\textsf{r}} \_ \ast \texttt{rbx}\mapsto_{\textsf{r}} \texttt{rbxv} \ast  \texttt{rsp}\mapsto_{\textsf{r}} \texttt{rspv} \ast \texttt{rbp}\mapsto_{\textsf{r}} \texttt{rbpv} }_{\rtv}$
@@
 mov 56[%
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast [\rtv'+56](\mathcal{I}\texttt{ASpace}(\theta,m) \ast \texttt{Pother})}_{\rtv}$
+$\specline{\textsf{P} \ast  I\texttt{ASpace}(\theta,m) \ast [\rtv'+56]( I\texttt{ASpace}(\theta,m) \ast \texttt{Pother})}_{\rtv}$
 $\specline{ \texttt{rdi}\mapsto_{\textsf{r}} \_ \ast \texttt{rbx}\mapsto_{\textsf{r}} \texttt{rbxv} \ast  \texttt{rsp}\mapsto_{\textsf{r}} \texttt{rspv} \ast \texttt{rbp}\mapsto_{\textsf{r}} \texttt{rbpv} }_{\rtv}$
@@
 ;; current address space!
-mov %
-mov %
-mov %
-mov %
-mov %
-mov %
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast [\rtv'+56](\mathcal{I}\texttt{ASpace}(\theta,m) \ast \texttt{Pother})}_{\rtv}$
+mov rsp, 8[rsi] 
+mov rbp, 16[rsi]
+mov r12, 24[rsi]
+mov r13, 32[rsi]
+mov r14, 40[rsi]
+mov r15, 48[rsi]
+$\specline{\textsf{P} \ast  I\texttt{ASpace}(\theta,m) \ast [\rtv'+56]( I\texttt{ASpace}(\theta,m) \ast \texttt{Pother})}_{\rtv}$
 $\specline{ \texttt{rdi}\mapsto_{\textsf{r}} \_ \ast \texttt{rbx}\mapsto_{\textsf{r}} \rtv'+0 \ast  \texttt{rsp}\mapsto_{\textsf{r}} \rtv'+8 \ast \texttt{rbp}\mapsto_{\textsf{r}} \rtv'+16 }_{\rtv}$
@@
 ;; Switch to the new address space
-mov %
-$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast [\rtv'+56](\mathcal{I}\texttt{ASpace}(\theta,m) \ast \texttt{Pother}) }_{\rtv}$
+mov cr3, 56[rsi]
+$\specline{\textsf{P} \ast  I\texttt{ASpace}(\theta,m) \ast [\rtv'+56]( I\texttt{ASpace}(\theta,m) \ast \texttt{Pother}) }_{\rtv}$
 $\specline{ \texttt{rdi}\mapsto_{\textsf{r}} \_ \ast \texttt{rbx}\mapsto_{r} \rtv'+0 \ast  \texttt{rsp}\mapsto_{\textsf{r}} \rtv'+8 \ast \texttt{rbp}\mapsto_{\textsf{r}} \rtv'+16 }_{\rtv}$
@@
 $\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\; \sqsubseteq $
-$\specline{ [\rtv](\texttt{Pcurrent}) \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast \texttt{Pother}}_{\rtv'+56}$
+$\specline{ [\rtv](\texttt{Pcurrent}) \ast  I\texttt{ASpace}(\theta,m) \ast \texttt{Pother}}_{\rtv'+56}$
 \end{lstlisting}
@@
 The wrinkle, and the importance of the modal treatment of assertions, is that the target thread's precondition is \emph{relative to its address space}, not the address space of the calling thread shown as
-\[[\rtv'+56](\mathcal{I}\texttt{ASpace}(\theta,m) \ast \texttt{Pother})\]
+\[[\rtv'+56]( I\texttt{ASpace}(\theta,m) \ast \texttt{Pother})\]
 in the specfication. 

### 2025-07-30 programlogic.tex
the magic of sed (i.e., replacing all the mathcal I with regular math mode I)

@@
 \begin{array}{l}
-  \mathcal{I}\textsf{ASpace}(\ptablestore,m)\stackrel{\triangle}{=} \textsf{ASpace\_Lookup}(\ptablestore,m) \ast 
+   I\textsf{ASpace}(\ptablestore,m)\stackrel{\triangle}{=} \textsf{ASpace\_Lookup}(\ptablestore,m) \ast 
   \bigast{(\vaddr, \textsf{pa})\in \ptablestore}{\exists\;(\textsf{l4e l3e l2e, l1e, pa})\ldotp \textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\vaddr\textsf{, l4e, l3e, l2e, l1e, pa})} \\
@@
 We place the authorative ownership of the ghost \add{map} translation $\mathcal{A}\textsf{PTableWalk}$ in a per-address-space invariant
-$\mathcal{I}$\textsf{ASpace} (Figure \ref{fig:peraspaceinvariant}), 
-\replace{$\mathcal{I}$\textsf{ASpace} alone allows}{allowing} changes to the page tables
+$ I$\textsf{ASpace} (Figure \ref{fig:peraspaceinvariant}), 
+\replace{$ I$\textsf{ASpace} alone allows}{allowing} changes to the page tables
 that preserve overall virtual-to-physical translations \add{in isolation},
@@
 }
-$\mathcal{I}\textsf{ASpace}(\theta,m)$ then \add{performs 3 roles:
+$ I\textsf{ASpace}(\theta,m)$ then \add{performs 3 roles:
 it associates the current address space's root with an appropriate \textsc{Iris} ghost name $\delta$;
@@
 % The invariant for the currently active address space (solid round square labelled as 
-% $\mathcal{I}$\textsf{ASpace} in Figure \ref{fig:logicaladdrspace}) asserts authorative ownership (i.e. 
+% $ I$\textsf{ASpace} in Figure \ref{fig:logicaladdrspace}) asserts authorative ownership (i.e. 
 % update capability), $\ownGhost\gammaPred{\authfull{\ptableabswalk\ptablestore}}$, of the ghost map $\Theta$
@@
 % shared invariant, and for each memory access,
-% a fragment of this corresponding to $\mathcal{I}\textsf{ASpace}(\Theta,m)$
+% a fragment of this corresponding to $ I\textsf{ASpace}(\Theta,m)$
 % for the current address space should be extracted from this global resource, used to prove correctness of the
@@
   \left\{ \begin{array}{l}
-    P \ast \mathcal{I}\textsf{ASpace}\ast  \textsf{rip} \mapsto_{\textsf{r}} \; {\textsf{iv} + \textsf{MovLen}(\textsf{r}_d,[\textsf{r}_a])} \ast \arcr
+    P \ast  I\textsf{ASpace}\ast  \textsf{rip} \mapsto_{\textsf{r}} \; {\textsf{iv} + \textsf{MovLen}(\textsf{r}_d,[\textsf{r}_a])} \ast \arcr
     r_d \mapsto_{r}  \textsf{ v} \ast r_a \mapsto_{r} \; \{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{v}} \;\{q\} \textsf{ v}
@@
 }{
-  \{P \ast \mathcal{I}\textsf{ASpace}\ast \textsf{rip} \mapsto_{\textsf{r}} \textsf{ iv} \ast r_d \mapsto_{r}  \textsf{ rvd} \ast r_a \mapsto_{r} \;\{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{v}} \textsf{ v} \}_{\textsf{rtv}}\;
+  \{P \ast  I\textsf{ASpace}\ast \textsf{rip} \mapsto_{\textsf{r}} \textsf{ iv} \ast r_d \mapsto_{r}  \textsf{ rvd} \ast r_a \mapsto_{r} \;\{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{v}} \textsf{ v} \}_{\textsf{rtv}}\;
   \textsf{mov}~\textsf{r}_d,~[\textsf{r}_a];\;\overline{is}
@@
 \inferrule[\footnotesize WriteToVirtMemFromReg]{
-  \{P \ast \mathcal{I}\textsf{ASpace}\ast \textsf{rip} \mapsto_{\textsf{r}} \textsf{ iv} + \textsf{MovLen}(\textsf{r}_d,[\textsf{r}_a])  \ast r_s \mapsto_{r}\;\{q_1\}  \textsf{ rvs}  \ast r_a \mapsto_{r} \; \{q_2\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{v}} \textsf{ v} \}_{\textsf{rtv}}\;\overline{ is}
+  \{P \ast  I\textsf{ASpace}\ast \textsf{rip} \mapsto_{\textsf{r}} \textsf{ iv} + \textsf{MovLen}(\textsf{r}_d,[\textsf{r}_a])  \ast r_s \mapsto_{r}\;\{q_1\}  \textsf{ rvs}  \ast r_a \mapsto_{r} \; \{q_2\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{v}} \textsf{ v} \}_{\textsf{rtv}}\;\overline{ is}
 }{
-  \{P \ast \mathcal{I}\textsf{ASpace}\ast \textsf{rip} \mapsto_{\textsf{r}} \textsf{ iv} \ast r_s \mapsto_{r} \;\{q_1\}  \textsf{ rvs}   \ast r_a \mapsto_{r} \; \{q_2\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{v}} \textsf{ rvs} \}_{\textsf{rtv}}
+  \{P \ast  I\textsf{ASpace}\ast \textsf{rip} \mapsto_{\textsf{r}} \textsf{ iv} \ast r_s \mapsto_{r} \;\{q_1\}  \textsf{ rvs}   \ast r_a \mapsto_{r} \; \{q_2\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{v}} \textsf{ rvs} \}_{\textsf{rtv}}
   \textsf{ mov}~[\textsf{r}_a],~\textsf{r}_s;\;\overline{is}
@@
 \inferrule[\footnotesize WriteToRegCtlFromRegModal]{
-  \{[\textsf{rtv}](P \ast \mathcal{I}\textsf{ASpace})\ast \textsf{rip} \mapsto_{\textsf{r}} \textsf{ iv + 4} \ast \mathcal{I}\textsf{ASpace} \ast R\ast r_s \mapsto_{r} \;\{q\}  \textsf{ rvs}  \}_{\textsf{rvs}} \;\overline{is}
+  \{[\textsf{rtv}](P \ast  I\textsf{ASpace})\ast \textsf{rip} \mapsto_{\textsf{r}} \textsf{ iv + 4} \ast  I\textsf{ASpace} \ast R\ast r_s \mapsto_{r} \;\{q\}  \textsf{ rvs}  \}_{\textsf{rvs}} \;\overline{is}
 }{
-  \{P \ast \mathcal{I}\textsf{ASpace} \ast \textsf{rip} \mapsto_{\textsf{r}} \textsf{ iv} \ast [\textsf{rvs}](R \ast\mathcal{I}\textsf{ASpace})\ast r_s \mapsto_{r}\; \{q\}  \textsf{ rvs}   \}_{\textsf{rtv}} \;
+  \{P \ast  I\textsf{ASpace} \ast \textsf{rip} \mapsto_{\textsf{r}} \textsf{ iv} \ast [\textsf{rvs}](R \ast I\textsf{ASpace})\ast r_s \mapsto_{r}\; \{q\}  \textsf{ rvs}   \}_{\textsf{rtv}} \;
   \textsf{mov}~\textsf{cr3},~r_s;\;\overline{is}
@@
 Figure \ref{fig:wpdamd} includes two  rules for accessing memory at an address stored in a register $r_a$. 
-\add{Setting aside $P$, $\mathcal{I}\textsf{ASpace}$, and the instruction pointer \lstinline|rip|,}
+\add{Setting aside $P$, $ I\textsf{ASpace}$, and the instruction pointer \lstinline|rip|,}
 \del{Ultimately, any memory access needs to ensure the relevant address translation would succeed,
@@
 virtual points-to itself (\replace{in order to permit the physical page tables to be freely modified
-as long as they preserve virtual-to-physical translations}{per the discussion of Section \ref{sec:sharingpages}}), the rule requires $\mathcal{I}\textsf{ASpace}$
+as long as they preserve virtual-to-physical translations}{per the discussion of Section \ref{sec:sharingpages}}), the rule requires $ I\textsf{ASpace}$
 for the current address space to be carried through.
@@
 the page table walk for the relevant address ($\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$)
-from the invariant ($\mathcal{I}\mathsf{ASpace}$), prove that the page table walk succeeds
+from the invariant ($ I\mathsf{ASpace}$), prove that the page table walk succeeds
 and that memory or registers are updated appropriately, before re-packing the invariant and virtual points-to resources.
@@
 \lstinline|mov| to $\textsf{rvs}$ afterwards. The global effects of this rule are reflected in moving
-\del{bare} assertions \add{for the current address space ($P$ and $\mathcal{I}\mathsf{ASpace}$)} under an other-space modality for \add{the initial
+\del{bare} assertions \add{for the current address space ($P$ and $ I\mathsf{ASpace}$)} under an other-space modality for \add{the initial
 (outgoing) address space} $\rtv$, and moving the new address space's assertions out of

### 2025-07-30 examples_C.tex
delete stray A that was probably from a vim/hx edit more mixup

@@
   $\specline{\textsf{P} \ast  I\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \lnot(\textsf{entry\_present entryv}) \ast \textsf{entryp+KERNBASE} \mapsto_{\mathsf{vpte,1}} \mathsf{entryp} \; \textsf{pfn\_set(entryv nextpaddr)}}$
-  $\specline{\ghostmaptoken{\delta{}s}{\rtv}{\delta} \ast \forall_{i\in\textsf{0..511}}\ldotp ()\textsf{table\_root (pte\_initialized (pfn\_set (entry\_val nextpaddr))) + i * 8}) \mapsto_{\textsf{id}} \textsf{level-1}   }_{\rtv}$(*\label{line:initialize_post}*)A
+  $\specline{\ghostmaptoken{\delta{}s}{\rtv}{\delta} \ast \forall_{i\in\textsf{0..511}}\ldotp ()\textsf{table\_root (pte\_initialized (pfn\_set (entry\_val nextpaddr))) + i * 8}) \mapsto_{\textsf{id}} \textsf{level-1}   }_{\rtv}$(*\label{line:initialize_post}*)
 }

### 2025-07-30 programlogic.tex
slightly refine the why-modalities argument more

@@
 \add{
-  Many approaches which could handle these problems in principle, such as tagging pointers with the relevant address space.
+  Many approaches could handle these problems in principle, such as tagging pointers with the relevant address space.
   But such approaches introduce other complexities.
@@
   over \emph{functions} from address space identity to assertions rather than just assertions.
+}
+
+\add{
   Using a modal approach resolves all of these challenges cleanly and uniformly.
-  Specifications that are not \emph{about} address space manipulation need not mention an address spaces at all:
-  one can use standard separation logic data structure specifications without restructuring,
-  and every assertion can be stated for either the current or specific other address spaces.
-  In short, modalities make it possible for specifications to mention address spaces when necessary,
-  without requiring mention when only one address space is at hand.
+  Specifications that are not \emph{about} address space manipulation need not mention address spaces in assertions.
+  One can use standard separation logic data structure specifications without restructuring or
+  adding explicit address space
+  tracking,
+  but every assertion can be still stated for either the current or specific other address spaces as needed.
+  In short, modalities make it possible for specifications to mention address spaces when
+  it is important to the code, and not when it is unimportant to the code.
+  \looseness=-1
 }

### 2025-08-16 background.tex
check in some last-minute revision-tweaks that were never committed

@@
 with assertions of the form $[\ell](P)$ indicating that $P$ is true in the specific alternate circumstance (Kripke world)
- named by the \emph{nominal} $\ell$. Note that a distinctive property of hybrid logics is that, rather than hiding
+ named by the \replace{\emph{nominal}}{term} $\ell$. Note that a distinctive property of hybrid logics is that, rather than hiding
 the points at which a modal assertion is evaluated inside the modality's definition, the choice of what world a modalized
@@
 
-This explicit naming of alternate worlds increases the power of \add{propositional} modal logics~\cite{blackburn1995hybrid}, and is actually
+This explicit naming of alternate worlds increases the power of \add{propositional} modal logics~\cite{blackburn1995hybrid}, and is \del{actually}
 necessary for completeness in classical separation logics~\cite{brotherston2014parametric}.

### 2025-08-16 intro.tex
check in some last-minute revision-tweaks that were never committed

@@
 In writing the authoritative reference on the internals of the Solaris kernel, McDougall and Mauro went so far as to claim that ``\emph{the virtual memory sub-system can be considered the core of a Solaris instance, and the implementation of Solaris virtual memory affects just about every other subsystem in the operating system}''~\cite{mcdougall2006solaris}.
-This makes rich support for verification the virtual memory management subsystem of an OS kernel critical to the correctness of every other piece of an OS or any software running atop it.
+This makes \del{rich} support for verification the virtual memory management subsystem of an OS kernel critical to the correctness of every other piece of an OS or any software running atop it.
 

### 2025-08-17 background.tex
add keywords and CCS; remove all uses of \del

@@
 Because \iris is an embedded framework in \rocq, proofs in \iris-derived
-logics \del{(including ours)} often encapsulate raw \rocq assertions: $\ulcorner P \urcorner$ is an embedding
+logics often encapsulate raw \rocq assertions: $\ulcorner P \urcorner$ is an embedding
 of the \rocq assertion $P$ into an \iris assertion (used for things like equality and other
@@
 \rocq terms of a particular type, rather than being drawn from a fixed vocabulary.
-\del{We introduce other details as they arise.}
 \looseness=-1
@@
 
-An important limitation \del{here} is that to date, every separation logic
+An important limitation is that to date, every separation logic
 has assumed that all pointer addresses are meant for use in a single address space, which avoided the
@@
 
-This explicit naming of alternate worlds increases the power of \add{propositional} modal logics~\cite{blackburn1995hybrid}, and is \del{actually}
+This explicit naming of alternate worlds increases the power of \add{propositional} modal logics~\cite{blackburn1995hybrid}, and is 
 necessary for completeness in classical separation logics~\cite{brotherston2014parametric}.

### 2025-08-17 intro.tex
add keywords and CCS; remove all uses of \del

--- 2025-07-27 intro.tex
@@
 In writing the authoritative reference on the internals of the Solaris kernel, McDougall and Mauro went so far as to claim that ``\emph{the virtual memory sub-system can be considered the core of a Solaris instance, and the implementation of Solaris virtual memory affects just about every other subsystem in the operating system}''~\cite{mcdougall2006solaris}.
-This makes rich support for verification the virtual memory management subsystem of an OS kernel critical to the correctness of every other piece of an OS or any software running atop it.
+This makes support for verification the virtual memory management subsystem of an OS kernel critical to the correctness of every other piece of an OS or any software running atop it.
 

### 2025-08-17 programlogic.tex
add keywords and CCS; remove all uses of \del

@@
   resource indeed maps $k$ to $v$. These properties are captured by key \iris rules in Figure \ref{fig:ghostmaps}.\footnote{\iris ghost maps lack established notation\add{;}
-  \del{in the literature, but} the syntax we use captures the details of \texttt{iris.base\_logic.lib.ghost\_map}.}
+   the syntax we use captures the details of \texttt{iris.base\_logic.lib.ghost\_map}.}
   There are other rules, but these two are most important for explaining ghost maps.
   \textsc{GhostMapUpdate} says that ownership of the actual ghost map with ghost name $\gamma$ and map contents $\theta$,
-  and a \del{full (non-fractional) }token witnessing that $\theta$ maps \textsf{pa} to \textsf{va} permits an update to the ghost map's state,
+  and a token witnessing that $\theta$ maps \textsf{pa} to \textsf{va} permits an update to the ghost map's state,
   changing the map and replacing the token to represent the new value.
@@
 \replace{fragment}{token} stored in the \add{relevant} virtual points-to (Figure \ref{fig:virtualpointstosharing}).
-\del{changes in the virtual-to-physical translations become possible without the need to track all virtual points-to assertions.}
 \looseness=-1
@@
 }
-\del{In Figure ??,
-$\theta$ is the logical map from virtual addresses to the physical addresses they should translate to, 
-for the \emph{current} address space, corresponding to the  currently installed page tables (the ones
-indicated by \texttt{cr3}). In Figure ??, the existentially quantified $\delta$
-is the ghost name for this map (again, for the current address space).
-Because our logic is the first to address the existence of multiple address spaces,
-it must ensure that $\theta$ actually corresponds to the intended ghost map for the current address space.
-This requires looking up the ghost name (i.e., $\delta$) for the current address space,
-which is handled by $m$ --- another logical map (tied to ghost state shortly)
-from various page table roots to ghost names for the per-address-space ghost maps discussed above.
-}
 \add{
@@
 stores the physical resources for the current address space mappings \add{(via the iterated $\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$)}.
-\del{ensuring that the current address space is valid (has an entry in $m$ whose corresponding
-authoritative map matches $\theta$), and asserting ownership of the appropriate fractional ownership
-of the physical page table walk for each virtual-to-physical translation in $\Theta$ (for the current address space).}
-
-\del{
-Virtual points-to assertions must also ensure that they are tied to a valid address space, and because they
-use ghost map tokens to ensure the virtual-to-physical translation exists in a specific ghost map, they must also
-name that map. In Figure \ref{fig:virtualpointstosharing}, this is done with a ghost map token for \emph{another} ghost map,
-one shared across the kernel, holding $m$. Our virtual points-to assertion existentially quantifies the address-space-specific
-ghost map $\delta$ mapping virtual addresses to physical addresses, and ensures it exists with a token for the distinguished
-global (unique) ghost map named by $\delta{s}$.}
-% the ghost map of virtual-to-physical translations for a particular address space (the ghost name $\delta$ in Figure \ref{fig:virtualpointstosharing}).
-% % for each address space must be locateable from a representation of the specific address space:
-% % This means we also require a way to locate the map for the current address space from 
-% We support lookup by the current root page table pointer stored in \lstinline|cr3|.
-\looseness=-1
-
-% Figure \ref{fig:logicaladdrspace} depicts the relationships between these pieces, for a single
-% address space.
-% The invariant for the currently active address space (solid round square labelled as 
-% $ I$\textsf{ASpace} in Figure \ref{fig:logicaladdrspace}) asserts authorative ownership (i.e. 
-% update capability), $\ownGhost\gammaPred{\authfull{\ptableabswalk\ptablestore}}$, of the ghost map $\Theta$
-% describing virtual-to-physical address translations
-% -- shown as a solid-round-head arrow to $\mathcal{A}\textsf{PTableWalk}$.
-% Each virtual points-to assertion (dashed round square labelled as \textsf{Virtual PointsTo}
-%  in Figure \ref{fig:peraspaceinvariant})
-% asserts fragmental ownership of the ghost map $\Theta$, $\sumwalkabs\vaddr\qfrac\paddr$,
-% shown via the dashed-diamond-head arrow to $\mathcal{A}\textsf{PTableWalk}$. Specifically, it asserts that
-% the virtual address the virtual points-to assertion concerns corresponds to some physical address.
-% There exists only one persistent authorative ownership 
-% ($\ownGhost\gammaPred{\authfull{\ptableabswalk\ptablestore}}$) which collects the fragmental 
-% ghost page-table-walk mapping ($\sumwalkabs\vaddr\qfrac\paddr$),
-%  shown as double-ended solid arrow 1 to N imposing the update protocol.
-
-% \begin{figure}
-%    \includegraphics[width=0.75\columnwidth]{logical_addr_space.png}
-%   \caption{Logical Resources of Address-Space Abstraction (1 Address Space)}
-%   \label{fig:logicaladdrspace}
-%   \end{figure}
-
-% \paragraph{From A Single Address Space to Many}
-% In principle the resources from each individual address space should be collected into a single
-% shared invariant, and for each memory access,
-% a fragment of this corresponding to $ I\textsf{ASpace}(\Theta,m)$
-% for the current address space should be extracted from this global resource, used to prove correctness of the
-% memory access, and put back.
-% In this paper we focus on the middle section, explicity identifying resources for each
-% address space. The reason for this is that in practice, this global resource is also
-% the place where kernel-specific assumptions, such as guaranteeing that certain virtual address range
-% was mapped in all address spaces, would be enforced. This paper focuses on the
-% reasoning principles behind the virtual memory access, and we leave the use of this within a larger
-% kernel to future work. We do, however, deal with the existence of multiple address spaces in examples (Section \ref{sec:experiment}).
 
@@
 \textsf{Fact}s; these include other-space assertions, physical memory points-tos, and register assertions.
-\textsf{Fact}s can \add{freely} move in and out of other address space modalities\del{ freely}.
+\textsf{Fact}s can \add{freely} move in and out of other address space modalities.
 \looseness=-1
@@
 \add{Setting aside $P$, $ I\textsf{ASpace}$, and the instruction pointer \lstinline|rip|,}
-\del{Ultimately, any memory access needs to ensure the relevant address translation would succeed,
-which can be ensured by what we informally call the page-table-walk points-to collection
-($\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$ in Figure ??). }
 \TirNameStyle{WriteToRegFromVirtMem} and \TirNameStyle{WriteToVirtMemFromReg}
-\del{each use a virtual-pointsto assertion ($\textsf{vaddr} \mapsto_{\textsf{v}} \textsf{v}$),
-and }
 are nearly-standard (assembly) separation logic rules for memory accesses~\add{\cite{Chlipala2013Bedrock,ni2007contexts}}.
@@
 \replace{model updates to}{increment} the instruction pointer \lstinline|rip| \add{by the length of the encoded instruction.}
-\del{using a metafunction \textsf{MovLen}.}
 \textsf{MovLen} returns how long the instruction encoding for the corresponding \lstinline|mov| is;
@@
 \lstinline|mov| to $\textsf{rvs}$ afterwards. The global effects of this rule are reflected in moving
-\del{bare} assertions \add{for the current address space ($P$ and $ I\mathsf{ASpace}$)} under an other-space modality for \add{the initial
+assertions \add{for the current address space ($P$ and $ I\mathsf{ASpace}$)} under an other-space modality for \add{the initial
 (outgoing) address space} $\rtv$, and moving the new address space's assertions out of

### 2025-08-17 relwork.tex
add keywords and CCS; remove all uses of \del

@@
 The original \textsc{Verisoft} project~\cite{alkassar2008verisoft,alkassar2010pervasive,alkassar2008formal,dalinger2005verification,hillebrand2005address,alkassar2008formal,starostin2010formal} 
-relied on custom hardware\del{,} which\del{, among other things,} always ran kernel code with virtual memory disabled, removing the circularity that is a key challenge of verifying
+relied on custom hardware which always ran kernel code with virtual memory disabled, removing the circularity that is a key challenge of verifying
 VMM code for real hardware: \replace{A}{a}t that point page tables become a basic partial map data structure to represent user program address translations,
@@
 \add{targets real hardware that uses virtual addressing in the kernel, but unsoundly
-uses hardware models that do not.} \del{also treats page tables this way, but
-unsoundly given that \replace{other projects}{their} target hardware \del{that} \emph{does} run the kernel with address translation:}
+uses hardware models that do not.} 
 \add{Thus} they \emph{trust}
@@
 which is not true of page table updates.
-Other work on OS verification either never progressed \del{far enough} to \del{address} VMM verification
+Other work on OS verification either never progressed to VMM verification
 (\textsc{Verisoft XT}~\cite{cohen2009vcc,cohen2010local,dahlweid2009vcc,cohen2013SOFSEM}),
@@
   In principle both our approach and Kolanski and Klein's could
- } be extended to account for pageable points-to assertions by adding \del{additional}
-disjunctions to an extended points-to definition\add{,}\del{; embedding ``regular'' separation logic into such a variant
-is} \add{which would be}
+ } be extended to account for pageable points-to assertions by adding 
+disjunctions to an extended points-to definition\add{,} \add{which would be}
  the appropriate next step to extend reasoning to usermode programs running with a kernel that may demand-page the program's
@@
 
-% \todo[inline]{author response talks about syeda and klein here briefly, and it fits, but we also want the longer discussion}
-
-\del{Translation Lookaside Buffers}
 As noted in Section \ref{sec:backgroundonmachinemodel}, we do not formally model or reason about \add{translation lookaside buffers} \add{(}TLBs\add{)}.
@@
 so is not flexible enough to extend directly to a full context switching primitive as in Figure \ref{fig:swtchC}\replace{,}{.}
-\del{in addition to inheriting other limitations from {the work they extend, discussed above}.}
 \add{Their logic adds an assertion tracking known-inconsistent addresses (i.e., recently-unmapped by a table update or change in page table root)
@@
 The right \replace{eventual}{general} solution would be to combine our work
-and \add{an extension of} theirs, \add{a substantial project} which we leave to future work\replace{,}{.}\del{ as combining both the models and reasoning principles would be highly non-trivial.}
+and \add{an extension of} theirs, \add{a substantial project} which we leave to future work\replace{,}{.}
 \add{

### 2025-08-17 background.tex
various wordsmithing to reduce overage; guessing at camera-ready options, have cut overage roughly in half; next target is sections 4.2/4.3 which has a large gap after figure 7

@@
 The entries \add{(e.g., L4 Entry in Figure \ref{fig:pagetables})} of each table are 64 bits wide,
-but each points to a physical address aligned to 4KB (4096 byte) boundaries, which leaves 12 bits to spare to 
-control a validity bit (called the \emph{present} bit), a read-write bit (which permits write access through the entry if and only if it is set),
+but each points to a physical address aligned to 4KB (4096 byte) boundaries, leaving 12 bits to spare to 
+control a validity bit (called the \emph{present} bit), a read-write bit (enabling write access),
 and a range of additional bits which can be used to control caching, write-through, and more.
 This paper will only consider the present bit (0).
+\looseness=-1
 
@@
 is how the OS ensures memory isolation between processes.
-
 If an instruction is executed that accesses a virtual address that either has no mapping, or does not have a mapping permitting 
@@
 and the faulting program should be terminated. But this can also be used for
-\replace{
-specialized functionality and optimizations, such as \emph{paging} (saving room in physical RAM and deferring
-unnecessary IO by only reading program code from disk when it is accessed, or even swapping memory that has not
-recently been accessed to disk, to read back in when a page fault indicates access).
-}{
 other purposes such as demand paging to save on IO and better-manage physical memory~\cite{doeppner2010operating}.
-}
+\looseness=-1
 
@@
 
-Virtual memory concerns propagate to the OS scheduler, which deals with multiple 
-address spaces, so must keep track of which virtual addresses are valid (and in what way) in which address spaces. 
+Virtual memory affects the OS scheduler, which deals with multiple 
+address spaces, so must track which virtual addresses are valid in which address spaces. 
 Some virtual addresses are valid in only a single address space (e.g., a code address for a particular usermode 
@@
 a certain range of virtual addresses (corresponding to the kernel's code and data) are valid in every address space.
+\looseness=-1
 
@@
 also includes a \emph{view shift} operator $\vs$ which models updates to ghost state: $A \vs B$
-means resources satisfying $A$ may be transformed into resources satisfying $B$, intuitively by updating only ghost state.\footnote{\iris
-experts may note that this is \emph{technically} mildly misleading given how \iris's update modalities and weakest precondition
-are defined, but
-this is adequate intuition for non-experts in \iris to follow their use.
-}
+means resources satisfying $A$ may be transformed into resources satisfying $B$, intuitively by updating only ghost state
+(a slight simplification of \iris's update modalities, but adequate intuition for non-\iris-experts).
+% .\footnote{\iris
+% experts may note that this is \emph{technically} mildly misleading given how \iris's update modalities and weakest precondition
+% are defined, but
+% this is adequate intuition for non-experts in \iris to follow their use.
+% }
 View shifts are essentially logical entailment plus ghost updates.

### 2025-08-17 syntaxsemantics.tex
add keywords and CCS; remove all uses of \del
various wordsmithing to reduce overage; guessing at camera-ready options, have cut overage roughly in half; next target is sections 4.2/4.3 which has a large gap after figure 7

@@
 sequence with an additional instruction (\iseq\instr\instrs).
-We model (and later, give program logic rules for) instructions for \del{basic} register \replace{moves}{loads and stores}, and reading and writing memory.
+We model (and later, give program logic rules for) instructions for register \replace{moves}{loads and stores}, and reading and writing memory.
 The latter require page table walks.
-% Figure \ref{fig:coq_addr_translation} gives a slightly simplified version of our address translation code in monadic style:
-% starting from the root page table address \textsf{rtv}, the virtual address \textsf{w} is either resolved to a physical address or an error is returned.
-% The full set of possible errors is not relevant to this paper; our logic is tailored for kernel code which should not fault,
-% so our proof rules guarantee no memory failures occur --- that the page table walk for any dereferenced address will succeed.
-
-% Page-table walk shown in Figure \ref{fig:pagetables} is realized with physical memory \texttt{load}s and \textsf{store}s for the entries in the 
-% page-tables which are simply modelled as lookup and updates to an instance of a memory map with 64 bit entries. In the address translation,
-% we see a physical memory $\mathsf{load}$ (i.e. physical memory map ($\sigma.\mathcal{M}$) for each level) for each level of page-table-walk 
-% concluded with alignment of the returned address for each \textit{successful} page-table-walk traversal which is what our reasoning principles 
-% only consider.
-
 The most important instructions that we model are memory-accessing variants of the x86-64 \lstinline|mov| instruction, which we format in Intel syntax
@@
 Thus, a store to memory is $\textsf{mov}~[r_m]~r_r$, and a load from memory is $\textsf{mov}~r_r~[r_m]$.
-% \[
-% \begin{array}{l}
-%   \hbox{(\TirNameStyle{WriteToMemFromReg})} \qquad
-%   \textsf{mov}~[r_m]~r_r \\
-%   \hbox{(\TirNameStyle{WriteToRegFromMem})} \qquad 
-%   \textsf{mov}~r_r~[r_m]
-%   \end{array}
-%   \]
 Operationally, each first translates the virtual address stored in the register $r_m$ to a physical address \add{(a page table walk)},
@@
 or updates $r_r$ with the contents of that memory (for load).
-% The semantics of each is realized by first translating the virtual-memory address of a memory location stored in $r_m$ to a physical memory location
-% per the description in Section \ref{sec:backgroundonmachinemodel},
-% then either updating those memory contents with the contents of register $r_r$ (\TirNameStyle{WriteToMemFromReg}) 
-% or loading the value at that physical memory location into the register $r_r$
-% (\TirNameStyle{WriteToRegFromMem}). 
 Our formalization includes
@@
 and \lstinline|pop|.
-The semantics model updates to the instruction pointer, though we elide that register (\lstinline|rip|)
-from specifications in the paper for clarity.
+% The semantics model updates to the instruction pointer, though we elide that register (\lstinline|rip|)
+% from specifications in the paper for clarity.
 \looseness=-1

### 2025-08-17 examples.tex
okay apparently I missed one use of \del

@@
 
-\del{
-A PTE points-to, as we already brifly mentioned, is defined just like the normal virtual points-to of Figure \ref{fig:virtualpointstosharing}, except the physical address (here, \textsf{pa}) is explicit in the assertion
-rather than existentially quantified:\\
-\centerline{$
-    \vaddr\mapsto_{\textsf{vpte,q}} \; \paddr \; \vpage : \mathsf{vProp}~\Sigma \stackrel{\triangle}{=} 
-    \exists \delta\ldotp
-	(\lambda \mathit{cr3val}\ldotp
-	\ghostmaptoken{\delta{}s}{\mathit{cr3val}}{\delta}) \ast 
-  \fracghostmaptoken{\delta}{\vaddr}{\paddr}{\qfrac} \ast \paddr \mapsto_{\mathsf{p}} \vpage
-$}\\
-This supports rules for accessing memory
-at that virtual address, but exposing the physical location being modified
-makes this convenient to use for page table modifications, since we must ensure
-the modified data is at the specific physical location that will be used as the L1 entry.
-}
 % After obtaining a virtual address \textsf{pte\_addr} in \textsf{rax} backed 

### 2025-08-17 examples_C.tex
a couple more space tweaks for a couple more lines. sad to lose the "you are not expected to understand this" footnote, but it can go in the talk. next up is an awkward gap before the 5.2 heading

@@
 Confusingly, a single dynamic execution of this function begins execution in one thread,
-and returns in another thread --- because the execution switches stacks, and thus returns on the second thread's stack.\footnote{This is the function in UNIX 6th Edition 
-with the infamous ``You are not expected to understand this'' comment~\cite{lions1996lions},
-though Doeppner~\cite{doeppner2010operating} and others, offer detailed explanations.}
+and returns in another thread --- because the execution switches stacks, and thus returns on the second thread's stack.%\footnote{This is the function in UNIX 6th Edition 
+% with the infamous ``You are not expected to understand this'' comment~\cite{lions1996lions},
+% though Doeppner~\cite{doeppner2010operating} and others, offer detailed explanations.}
 Hence this is used, for example, when the scheduler has chosen a new thread to execute for a voluntary (non-preempted)
@@
 Lines \ref{line:l1entry_storeC} and \ref{line:l1entry_setpresent} are verified using the pte-points-to
-memory store rule.\footnote{In the assembly proof, this is a bitwise-and of the word-aligned\lstinline|fpaddr| with 1
-(setting the present flag), yielding a single.
-\looseness=-1
-}
+memory store rule. In the assembly proof, this is a bitwise-and of the word-aligned \lstinline|fpaddr| with 1
+(setting the present flag), yielding a single store.
 \looseness=-1

### 2025-08-17 programlogic.tex
a couple more space tweaks for a couple more lines. sad to lose the "you are not expected to understand this" footnote, but it can go in the talk. next up is an awkward gap before the 5.2 heading

@@
 \footnotesize
-\[\begin{array}{c}
+% \[\begin{array}{c}
+\centerline{$
     \mbox{$[r](\mathsf{P}) \;: \mathsf{vProp} \; \Sigma \; \stackrel{\triangle}{=} \; \lambda \_,\; \mathsf{P}~\mathsf{r} \vspace{0.5em}$}
@@
     \inferrule*{ }{\textsf{Fact}\;(\textsf{r}\;\mapsto_{\textsf{r}}\{q\}\;\textsf{rv})}
-    \\[1.3em]
+  $}
+\vspace{0.5em}
+\centerline{$
     \inferrule{ }{\textsf{Fact}\;(\textsf{w} \mapsto_{\textsf{p}} \{q\} \textsf{ v} )}
@@
     \inferrule{ }{[r](\textsf{P} \ast \textsf{Q}) \dashv\vdash ([r](\textsf{P}) \ast [r](\textsf{Q})}\vspace{0.5em}
-    \\[1.3em]
+  $}
+\vspace{0.5em}
+\centerline{$
     \inferrule{ }{[r](\textsf{P}\land \textsf{Q}) \dashv\vdash [r](\textsf{P})\land [r](\textsf{Q})}
@@
     \inferrule{ }{\textsf{Fact P} \vdash  [r](\textsf{P}) \dashv\vdash (\textsf{P})} %(\textsf{vProp }\Sigma))\\
-\end{array}\]
+$}
+% \end{array}\]
 \vspace{-1em}
@@
 new current address space)}.
-
-Note that the \emph{global} aspect is important, as a na\"ive frame rule would be unsound for updates to \lstinline|cr3|,
-because one could frame out assertions in one address space, switch address spaces, and bring those assertions from the \emph{old}
+The \emph{global} aspect is important. A na\"ive frame rule would be unsound for \lstinline|cr3| updates:
+one could frame out assertions in one address space, switch address spaces, and bring those assertions from the \emph{old}
 address space back into the \emph{new} address space, where they may not hold. 

### 2025-08-18 examples_C.tex
down to limit, fixed a couple margin over-runs

@@
 We will ignore non-integer registers; others are handled similarly.
-
 In Figure \ref{fig:swtchC}, Lines \ref{line:start_save}--\ref{line:end_save} store the callee-save registers (per the System V AMD64 ABI calling conventions) of the calling
@@
 and vice versa. Thus the precondition has assertions for the new thread under an other-space modality for the new address space,
-and the postcondition has assertions for the old thread under an other-space modality for the old address space.\footnote{
-  The \textsf{ContextAt} assertions \emph{both} end up under the other-space modality for the old thread.
-  A real kernel would want to transfer both out, but justifying this is highly kernel-specific,
-  and particularly post-Spectre-and-Meltdown is quite varied.
-  \looseness=-1
-}
+and the postcondition has assertions for the old thread under an other-space modality for the old address space.%\footnote{
+  % The \textsf{ContextAt} assertions \emph{both} end up under the other-space modality for the old thread.
+  % A real kernel would want to transfer both out, but justifying this is highly kernel-specific,
+  % and particularly post-Spectre-and-Meltdown is quite varied.
+  %\looseness=-1
+%}
+Both \textsf{ContextAt} assertions end up under the old other-space modality, but in a real kernel would
+transfer out based on kernel-specific invariants.
 \looseness=-1
@@
 \begin{lstlisting}[language=C,mathescape,escapeinside={(*}{*)}]
-/* @param entry: virtual address of level n+1 entry which should point to the base of a level n table
+/* @param entry: virtual address of level n+1 entry which should point to a level-n table base
    @return next: a virtual pointer to the base of a valid level n table */
@@
     /*Now we know that entry is initialized, so we satisfy the condition to access children list */(*\label{line:now_we_know}*)
-    $\specline{ \textsf{entryp+KERNBASE} \mapsto_{\textsf{vpte,qfrac}}  \textsf{entryp} \; \textsf{pte\_initialized(pfn\_set(entryv nextpaddr))} \ast \ulcorner  \textsf{qfrac} = 1 \land \lnot(\textsf{entry\_present entryv})  \urcorner }_{\rtv}$
+    $\specline{ \textsf{entryp+KERNBASE} \mapsto_{\textsf{vpte,qfrac}}  \textsf{entryp} \; \textsf{pte\_initialized(pfn\_set(entryv nextpaddr))} \ast \ulcorner  \textsf{qfrac} = 1 \urcorner }_{\rtv}$
     $\specline{\forall_{i\in \textsf{0 ...
@@
 On Line \ref{line:read_entry_contentsC}, the code checks the present bit of the entry.
-If the bit is unset, there is no level-$n$ table, so one must be allocated via \lstinline|pte_initialize| (explained shortly\footnote{
-  Briefly, it allocates a fresh physical page, and initializes the memory pointed to by \lstinline|entry| with that physical address.
-}) and marked present.
+If the bit is unset, there is no level-$n$ table, so one must be allocated via \lstinline|pte_initialize| (explained shortly,
+but it essentially
+allocates a fresh physical page, and initializes the memory pointed to by \lstinline|entry| with that physical address) and marked present.
 By Line \ref{line:finalpieceS} the entry is known to be valid and contain the physical address of
@@
 \begin{figure}\footnotesize
-$\begin{array}{l}\left( \begin{array}{l}
+$\begin{array}{l}
+\textsf{R}_{\textsf{walk} }=
+\left( \begin{array}{l}
   \ulcorner
@@
       (\textsf{l2p+KERNBASE+8*L2Offset(va)}) \mapsto_{\textsf{vpte,qfrac2}} \textsf{l2p+8*L2Offset(va)} \; \textsf{l2e\_val}
-    \end{array} \right)  =\textsf{R}_{\textsf{walk} } 
-    \\ \left(\begin{array}{l}  (\textsf{l1p+KERNBASE*8*L1Offset(va)}) \mapsto_{\textsf{vpte,qfrac1}} \textsf{l1p+8*L1Offset(va)}\;\textsf{l1e\_val} \ast  \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l1e\_val})\urcorner \end{array}\right) =\textsf{R}_{\textsf{l1e}}\end{array} $
+    \end{array} \right)   
+    \\
+    \textsf{R}_{\textsf{l1e}}=(\textsf{l1p+KERNBASE*8*L1Offset(va)}) \mapsto_{\textsf{vpte,qf1}} \textsf{l1p+8*L1Offset(va)}\;\textsf{l1e\_val} \ast  \ulcorner  \textsf{qf1} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l1e\_val})\urcorner\end{array} $
 \vspace{-1em}
-\caption{\textsf{R}$_{\textsf{walk}}$: Resources obtained from invariant for walking page-table directory with \emph{non-existing entries} via calls to \textsf{pte\_get\_next\_table} in Figure \ref{fig:calltopteinitializeC}}
+\caption{\textsf{R}$_{\textsf{walk}}$: Resources obtained from invariant during a software page table walk in Figure \ref{fig:calltopteinitializeC}}
 \label{fig:rwalkC}
@@
     $\specline{ \textsf{l1\_entry+KERNBASE} \mapsto_{\textsf{vpte,1}} \textsf{l1\_entry} \; (\textsf{set\_pfn}(\textsf{l1e\_val}, \textsf{ADDR\_TO\_PFN}(\textsf{fpaddr}))) \ast  \ulcorner \lnot(\textsf{entry\_present l1e\_val}) \urcorner}$
-    /*Set the present bit in entry*/
-    pteaddr->present = 1;(*\label{line:l1entry_setpresent}*)
+    pteaddr->present = 1;(*\label{line:l1entry_setpresent}*) // Set the present bit in entry
   }
-  $\specline{ \begin{array}{l}\textsf{l1\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l1\_entry} \; \textsf{l1e\_val} \ast \ulcorner (\textsf{entry\_present l1e\_val}) \urcorner \\
- \lor \; \;  \textsf{l1\_entry+KERNBASE} \mapsto_{\textsf{vpte,1}} \textsf{l1\_entry} \; (\textsf{set\_pfn}(\textsf{l1e\_val}, \textsf{ADDR\_TO\_PFN}(\textsf{fpaddr}))) \ast  \ulcorner \lnot(\textsf{entry\_present l1e\_val}) \urcorner  \end{array} }$
+  $\specline{ \begin{array}{l}(\textsf{l1\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l1\_entry} \; \textsf{l1e\_val} \ast \ulcorner (\textsf{entry\_present l1e\_val}) \urcorner) \lor \\
+ \textsf{l1\_entry+KERNBASE} \mapsto_{\textsf{vpte,1}} \textsf{l1\_entry} \; (\textsf{set\_pfn}(\textsf{l1e\_val}, \textsf{ADDR\_TO\_PFN}(\textsf{fpaddr}))) \ast  \ulcorner \lnot(\textsf{entry\_present l1e\_val}) \urcorner  \end{array} }$
 }
@@
 For brevity this example is specialized to the case where $\vaddr$ is known to not be mapped:
-the precondition on Line \ref{line:mappage_pre} includes $\theta \; !!\; \vaddr = \texttt{None}$.\footnote{
-  The generalization to handling either case and returning an error is straightforward.
-}
+the precondition on Line \ref{line:mappage_pre} includes $\theta \; !!\; \vaddr = \texttt{None}$;
+generalization to returning an error if it is already mapped is straightforward.
+% .\footnote{
+%   The generalization to handling either case and returning an error is straightforward.
+% }
 The precondition on Line \ref{line:mappage_pre}, directly entails

### 2025-08-18 relwork.tex
down to limit, fixed a couple margin over-runs

@@
 that the particular page table manipulations do not, for example, unmap kernel code
- (which can crash the machine even if done ``temporarily'').\footnote{\add{A ``fun'' bug one author encountered in an industrial prototype kernel.}}
+ (which can crash the machine even if done ``temporarily'').%\footnote{\add{A ``fun'' bug one author encountered in an industrial prototype kernel.}}
 % \footnote{A bug

### 2025-08-21 examples_C.tex
add conditional for arxiv version
nearly fixed all margin complaints from camera-ready

@@
 $\specline{\mathsf{ContextAt}(save,\_) \ast \mathsf{ContextAt}(restore,[\mathsf{rbxv}',\ldots,\rtv']) }_{\rtv}$(*\label{line:end_swtch_pre}*)
-mov 0[rdi], rbx(*\label{line:start_save}*) ... ;; also save rsp, rbp, r12, r13, r14, to offsets 8, 16, 24, 32, and 40 from rdi
+mov 0[rdi], rbx(*\label{line:start_save}*) ... ;; also save rsp, rbp, r12, r13, r14, to offsets 8, 16, ... and 40 from rdi
 ... mov 48[rdi], r15
@@
 \emph{generated from this source code}.
-Listings of the assembly fragments with inline assertions appear in Appendix \ref{sec:experiment_appendix}. %\ref{apdx:asm}.
+Listings of the assembly fragments with inline assertions appear in
+\ifARXIV
+Appendix \ref{sec:experiment_appendix}.
+\else
+our technical report~\cite{kuru2025modal}.
+\fi
 Whether in C or assembly,
@@
     entry->present = 1; // Install newly-allocated level n table, mark present(*\label{line:install_new_entryC}*)
-    /*Now we know that entry is initialized, so we satisfy the condition to access children list */(*\label{line:now_we_know}*)
+    /*Now we know that entry is initialized, and the condition to access children list holds*/(*\label{line:now_we_know}*)
     $\specline{ \textsf{entryp+KERNBASE} \mapsto_{\textsf{vpte,qfrac}}  \textsf{entryp} \; \textsf{pte\_initialized(pfn\_set(entryv nextpaddr))} \ast \ulcorner  \textsf{qfrac} = 1 \urcorner }_{\rtv}$
@@
 511} } \ldotp  \ghostmaptoken{\textsf{id}}{((\textsf{table\_root (pte\_initialized (entryv.pfn)))}) + \textsf{i * 8})}{\textsf{v-1}}  }$ (*\label{line:childrenC}*)
-  uintptr_t next_phys_addr = PTE_PFN_TO_ADDR(entry->pfn); // Fetch physical address of next table (*\label{line:finalpieceS}*)(*\label{line:extract_pfn}*)
+  uintptr_t next_phys_addr = PTE_PFN_TO_ADDR(entry->pfn); // Fetch physical addr of next table (*\label{line:finalpieceS}*)(*\label{line:extract_pfn}*)
   uintptr_t next_virt_addr = (uintptr_t) P2V(next_phys_addr); // Convert to virtual address (*\label{line:p2vC}*)
@@
   \begin{lstlisting}[language=C,mathescape,escapeinside={(*}{*)}]
-/* @param entry: virtual address of a non-present entry to initialize with a new physical page */
-$\specline{\textsf{P} \ast  I\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \lnot(\textsf{entry\_present entryv}) \ast \textsf{entryp+KERNBASE} \mapsto_{\textsf{vpte,1}} \textsf{entryp entryv} \ast \ghostmaptoken{\textsf{id}}{(\mathsf{entryp})}{\textsf{level}} }$
+/* @param entry: virtual address of non-present entry to initialize with a new physical page*/
+$\specline{I\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \lnot(\textsf{entry\_present entryv}) \ast \textsf{entryp+KERNBASE} \mapsto_{\textsf{vpte,1}} \textsf{entryp entryv} \ast \ghostmaptoken{\textsf{id}}{(\mathsf{entryp})}{\textsf{level}} }$
 void pte_initialize(pte_t *entry) {
@@
   entry->pfn = PTE_ADDR_TO_PFN((uintptr_t) local);(*\label{line:kalloc_install}*)
-  $\specline{\textsf{P} \ast  I\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \lnot(\textsf{entry\_present entryv}) \ast \textsf{entryp+KERNBASE} \mapsto_{\mathsf{vpte,1}} \mathsf{entryp} \; \textsf{pfn\_set(entryv nextpaddr)}}$
+  $\specline{ I\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \lnot(\textsf{entry\_present entryv}) \ast \textsf{entryp+KERNBASE} \mapsto_{\mathsf{vpte,1}} \mathsf{entryp} \; \textsf{pfn\_set(entryv nextpaddr)}}$
   $\specline{\ghostmaptoken{\delta{}s}{\rtv}{\delta} \ast \forall_{i\in\textsf{0..511}}\ldotp ()\textsf{table\_root (pte\_initialized (pfn\_set (entry\_val nextpaddr))) + i * 8}) \mapsto_{\textsf{id}} \textsf{level-1}   }_{\rtv}$(*\label{line:initialize_post}*)

### 2025-08-21 relwork.tex
nearly fixed all margin complaints from camera-ready

--- 2025-08-17 relwork.tex
@@
 that the particular page table manipulations do not, for example, unmap kernel code
- (which can crash the machine even if done ``temporarily'').\footnote{\add{A ``fun'' bug one author encountered in an industrial prototype kernel.}}
-% \footnote{A bug
-% one author has personally encountered in a research kernel, though not one discussed here.}).
+ (which can crash the machine even if done ``temporarily'').
 This is true for \textsc{seL4}~\cite{Klein2009seL4,seL4TOCS,Sewell2013translation}, whose formal machine model omits address translation,
@@
 an analagous result \replace{{should} hold}{holds} of our work:}{
- An analagous result should holds of our model (though we have not proven it).
+ An analagous result should hold of our model (though we have not proven it).
 }

### 2025-08-21 examples_C.tex
narrow down some figures, hopefully for the last time

@@
   pte_t *l4_entry = &l4[L4Offset(va)]; // Virtual address of L4 entry(*\label{line:start_l4_calcC}*)
-  $\specline{\textsf{P} \ast  I\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{l4p+8*L4Offset(va)}\},m) \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
+  $\specline{\textsf{P} \ast  I\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{l4p+8*L4Offset(va)}\},m) \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  \ast \ulcorner  \textsf{qf4} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l4e\_val})\urcorner }_{\rtv}$
   $\specline{\textsf{entry\_present(l4e\_val)} \wand \forall_{i\in\textsf{0..511}} \; \ghostmaptoken{\textsf{id}}{(\textsf{table\_root(l4e\_val)+ i * 8})}{\textsf{3}} }_{\rtv} $
-  $\specline{ \textsf{l4p+KERNBASE+L4Offset(va)} \mapsto_{\textsf{vpte,qfrac4}} \textsf{(l4p+L4Offset(va)) l4e\_val} \ast \ulcorner  \textsf{qfrac4} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l4e\_val})\urcorner}_{\rtv}$
+  $\specline{ \textsf{l4p+KERNBASE+L4Offset(va)} \mapsto_{\textsf{vpte,qf4}} \textsf{(l4p+L4Offset(va)) l4e\_val}}_{\rtv}$    
   pte_t *l3 = pte_get_next_table(l4_entry); (*\label{line:first_getnext_callC}*) 
   /*pte_get_next_table may have allocated a new page which updates entry value*/
-  $\specline{\textsf{P} \ast  I\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{l4p+8*L4Offset(va)}\},m)  \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
+  $\specline{\textsf{P} \ast  I\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{l4p+8*L4Offset(va)}\},m)  \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  \ast \ulcorner  \textsf{qf4} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l4e\_val})\urcorner }_{\rtv}$
   $\specline{\forall_{i\in\textsf{0..511}} \ghostmaptoken{\textsf{id}}{(\textsf{table\_root(l4e\_val'.pfn) + i * 8})}{\textsf{3}}  \ast \ulcorner\textsf{l3-KERNBASE}=\textsf{table\_root(l4e\_val'.pfn)}\urcorner} $(*\label{line:l3tokens}*)
-  $\specline{ \textsf{l4p+KERNBASE+L4Offset(va)} \mapsto_{\textsf{vpte,qfrac4}} \textsf{(l4p+L4Offset(va)) l4e\_val} \ast \ulcorner  \textsf{qfrac4} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l4e\_val})\urcorner}_{\rtv}$(*\label{line:first_pte_pointstoC}*)
+  $\specline{ \textsf{l4p+KERNBASE+L4Offset(va)} \mapsto_{\textsf{vpte,qf4}} \textsf{(l4p+L4Offset(va)) l4e\_val}}_{\rtv}$(*\label{line:first_pte_pointstoC}*)
   pte_t *l3_entry = &l3[L3Offset(va)]; /* Virtual address of L3 entry*/(*\label{line:l3offset}*)
@@
   $\specline{\forall_{i\in\textsf{0..511}\setminus\{\textsf{L3Offset(va)}\}} \; \ghostmaptoken{\textsf{id}}{(\textsf{table\_root(l4e\_val.pfn}) \textsf{ + i * 8})} {\textsf{3}} }_{\rtv}$
-  $\specline{ \textsf{l4p+KERNBASE+L4Offset(va)} \mapsto_{\textsf{vpte,qfrac4}} \textsf{(l4p+L4Offset(va)) l4e\_val} \ast \ulcorner  \textsf{qfrac4} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l4e\_val})\urcorner}_{\rtv}$(*\label{line:ex_l4_vpteC}*)
-  $\specline{ \textsf{l3p+KERNBASE+L3Offset(va)} \mapsto_{\textsf{vpte,qfrac3}} \textsf{(l3p+L3Offset(va)) l3e\_val} \ast \ulcorner  \textsf{qfrac3} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l3e\_val})\urcorner}_{\rtv}$(*\label{line:ex_l3_vpteC}*)
+  $\specline{ \textsf{l4p+KERNBASE+L4Offset(va)} \mapsto_{\textsf{vpte,qf4}} \textsf{(l4p+L4Offset(va)) l4e\_val} \ast \ulcorner  \textsf{qf4} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l4e\_val})\urcorner}_{\rtv}$(*\label{line:ex_l4_vpteC}*)
+  $\specline{ \textsf{l3p+KERNBASE+L3Offset(va)} \mapsto_{\textsf{vpte,qf3}} \textsf{(l3p+L3Offset(va)) l3e\_val} \ast \ulcorner  \textsf{qf3} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l3e\_val})\urcorner}_{\rtv}$(*\label{line:ex_l3_vpteC}*)
   $\specline{\textsf{entry\_present(l3e\_val)} \wand \forall_{i\in\textsf{0..511}} \;\ghostmaptoken{\textsf{id}}{(\textsf{table\_root(l3e\_val.pfn) + i * 8})}{ \textsf{2}} } _{\rtv} $
@@
   pte_t *l1 = pte_get_next_table(l2_entry);(*\label{line:getnextl2}*)
-  return &l1[L1Offset(va)]; // Return virtual address of L1 entry with a similar index computation(*\label{line:returnl1entryaddr}*)
+  return &l1[L1Offset(va)]; // Return virtual addr of L1 entry with a similar index computation(*\label{line:returnl1entryaddr}*)
 } $\specline{\textsf{R}_{\textsf{walk}} \ast \textsf{R}_{\textsf{l1e}} }_{\rtv}$
@@
 \vspace{-1em}
-\caption{Walking page-table directory via calls to \textsf{pte\_get\_next\_table} in Figure \ref{fig:calltopteinitializeC}}
+\caption{Walking page-table via calls to \textsf{pte\_get\_next\_table} in Figure \ref{fig:calltopteinitializeC}}
 \label{walkpgdirC}
@@
       \ghostmaptoken{\textsf{id}}{(\mathsf{l4p+8*L4Offset(va)})}{\textsf{4}}\ast \ghostmaptoken{\textsf{id}}{(\mathsf{l3p+8*L3Offset(va)})}{\textsf{3}} \ast \\ \ghostmaptoken{\textsf{id}}{(\mathsf{l2p+8*L2Offset(va)})}{\textsf{2}}  \ast \ghostmaptoken{\textsf{id}}{(\mathsf{l1p+8*L1Offset(va)})}{\textsf{1}} \ast \\
-      \ghostmaptoken{\delta{}s}{\rtv}{\delta} \ast (\textsf{l4p+KERNBASE+8*L4Offset(va)}) \mapsto_{\textsf{vpte,qfrac4}} (\textsf{l4p+8*L4Offset(va)}) \; \textsf{l4e\_val}  \ast \\
-      (\textsf{l3p+KERNBASE+8*L3Offset(va)}) \mapsto_{\textsf{vpte,qfrac3}} \textsf{l3p+8*L3Offset(va)}\;\textsf{l3e\_val}  \ast\\
-      (\textsf{l2p+KERNBASE+8*L2Offset(va)}) \mapsto_{\textsf{vpte,qfrac2}} \textsf{l2p+8*L2Offset(va)} \; \textsf{l2e\_val}
+      \ghostmaptoken{\delta{}s}{\rtv}{\delta} \ast (\textsf{l4p+KERNBASE+8*L4Offset(va)}) \mapsto_{\textsf{vpte,qf4}} (\textsf{l4p+8*L4Offset(va)}) \; \textsf{l4e\_val}  \ast \\
+      (\textsf{l3p+KERNBASE+8*L3Offset(va)}) \mapsto_{\textsf{vpte,qf3}} \textsf{l3p+8*L3Offset(va)}\;\textsf{l3e\_val}  \ast\\
+      (\textsf{l2p+KERNBASE+8*L2Offset(va)}) \mapsto_{\textsf{vpte,qf2}} \textsf{l2p+8*L2Offset(va)} \; \textsf{l2e\_val}
     \end{array} \right)   
@@
   }
-  $\specline{ \begin{array}{l}(\textsf{l1\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l1\_entry} \; \textsf{l1e\_val} \ast \ulcorner (\textsf{entry\_present l1e\_val}) \urcorner) \lor \\
- \textsf{l1\_entry+KERNBASE} \mapsto_{\textsf{vpte,1}} \textsf{l1\_entry} \; (\textsf{set\_pfn}(\textsf{l1e\_val}, \textsf{ADDR\_TO\_PFN}(\textsf{fpaddr}))) \ast  \ulcorner \lnot(\textsf{entry\_present l1e\_val}) \urcorner  \end{array} }$
+$\specline{ \begin{array}{l}(\textsf{l1\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l1\_entry} \; \textsf{l1e\_val} \ast \ulcorner (\textsf{entry\_present l1e\_val}) \urcorner) \lor ( \ulcorner \lnot(\textsf{entry\_present l1e\_val}) \urcorner \\
+ \ast \textsf{l1\_entry+KERNBASE} \mapsto_{\textsf{vpte,1}} \textsf{l1\_entry} \; (\textsf{set\_pfn}(\textsf{l1e\_val}, \textsf{ADDR\_TO\_PFN}(\textsf{fpaddr}))) )  \end{array} }$
 }

### 2025-08-22 relwork.tex
fix relwork re: syeda and klein

@@
 and their memory access rules require accessed memory to lie outside that set.
-However their logic is proven sound against a generalized semantics where any page table change adds \emph{all}
-addresses to the inconsistent set, which is too imprecise for finer-grained unmapping.
+% However their logic is proven sound against a generalized semantics where
+% any page table change adds \emph{all}
+% addresses to the inconsistent set,
+% which is too imprecise for finer-grained unmapping.
+However in their logic,
+reasoning about updates to the inconsistent set (from a page table update)
+requires direct reasoning
+directly about the physical memory heap, which limits modularity.
 }

### 2025-08-22 background.tex
finish removing replace commands

@@
 \vspace{-2em}
-          \caption{x86-64 \replace{page table lookups}{hardware address translation}.}
+          \caption{x86-64 hardware address translation.}
         \label{fig:pagetables}
@@
 shows,
-address translation proceeds by repeatedly taking designated slices of the virtual address and indexing into \replace{a}{successive} table\add{s} \add{of 512 8-byte entries (making each table 4KB in size)}.
+address translation proceeds by repeatedly taking designated slices of the virtual address and indexing into {successive} table\add{s} \add{of 512 8-byte entries (making each table 4KB in size)}.
 The final lookup in the page tables gives the base physical address of a 4KB page of physical memory \add{accessible to the running
@@
 but for brevity and consistency, we simply number them, in keeping with the newer 5th level. Our formalization only deals with 4-level page tables, but is straightforwardly extensible to 5.} 
-The translation process or algorithm is \replace{sometimes}{often} referred to as a \emph{page-table walk}.
+The translation process or algorithm is {often} referred to as a \emph{page-table walk}.
 While Figure \ref{fig:pagetables} and most of our constants (how many levels, which virtual address bits index which
-table levels) are specific to \replace{the}{common} x86-64 \replace{architecture}{CPUs},
-\replace{
-ARMv8 (a.k.a. aarch64), RISC-V, and PowerPC use similar hierarchical page tables for address translation.
-RISC-V's sv48 paging configuration~?? and AArch64's 4-level paging configuration
-both split 64-bit virtual addresses at the same points to index into the respective tables, so most of what follows
-is equally applicable to those architectures (the only difference is that page table entries place control bits of write access, etc.,
-in different orders).
-}{
+table levels) are specific to {common} x86-64 {CPUs},
   it is straightforward to adapt our approach to ARMv8/\texttt{aarch64}, RISC-V, or PowerPC, which use nearly-identical
@@
   or to 5-level paging present in newer x86-64 or RISC-V designs (one step would be iterated an additional time).
-}
 
@@
 
-The page tables are managed by the \replace{OS, typically by a}{OS kernel's} \emph{virtual memory manager} (VMM).\footnote{Not to be confused with Virtual 
+The page tables are managed by the {OS kernel's} \emph{virtual memory manager} (VMM).\footnote{Not to be confused with Virtual 
 Machine Monitor. We focus on non-hypervisor scenarios, but hardware virtualization extensions for both 
@@
 
-We suppress \replace{many details of Iris assertions}{some technical \iris details} for brevity, but briefly note
+We suppress {some technical \iris details} for brevity, but briefly note
 a few recurring details used heavily in \iris but not necessarily in traditional
@@
 with assertions of the form $[\ell](P)$ indicating that $P$ is true in the specific alternate circumstance (Kripke world)
- named by the \replace{\emph{nominal}}{term} $\ell$. Note that a distinctive property of hybrid logics is that, rather than hiding
+ named by the {term} $\ell$. Note that a distinctive property of hybrid logics is that, rather than hiding
 the points at which a modal assertion is evaluated inside the modality's definition, the choice of what world a modalized
@@
 Some prior \iris-based work~\cite{dang2019rustbelt,dang2022compass} has constructed derived modalities in the style we propose, indexed
-by thread IDs. However \replace{in that setting, the}{their} intepretation of those modalities was fully fixed ahead of time (to refer to essentially buffers in operationalized versions of C11
+by thread IDs. However their intepretation of those modalities was fully fixed ahead of time (to refer to essentially buffers in operationalized versions of C11
 concurrency). In this setting, while our modalities will be indexed by page table roots, it is possible to modify the address translation for an address

### 2025-08-22 programlogic.tex
finish removing replace commands

@@
 
-We describe a program logic (a separation logic) along the lines suggested \replace{earlier}{above}, where every assertion is relative
+We describe a program logic (a separation logic) along the lines suggested {above}, where every assertion is relative
 to an address space in which it is interpreted, allowing us to define \emph{virtual points-to} assertions that make claims
@@
 This supports rules for updating not only typical data in memory that happens to be subject to address translation, but \add{also}
-manipulation of the page tables themselves via virtual addresses (as demanded by all modern hardware) \replace{also}{and}
+manipulation of the page tables themselves via virtual addresses (as demanded by all modern hardware) {and}
 via virtual points-to assertions.
@@
 carry over to our derived logic.
-% \add{Since \textsf{vProp}s are ``just'' \textsc{Iris} predicates on a choice of page table root,\todo{this section may be subsumed under first addition to this section, don't need both}
-% it may not be obvious what the gain is over simply using \textsc{Iris} directly.
-% The gain is not expressive power.
-% Every modal logic can be translated straightforwardly to a logic without modalities
-% (nearly always a small variation on the \emph{standard translation} from propositional modal logic
-% to first-order logic~\cite{blackburn2006modal}\footnote{\add{Itself similar to verification condition generation
-% for weakest precondition calculi and Hoare logics.}}) that can directly address the logic's models ---
-% but these translations, in addition to making specifications far more verbose,
-% generally do not preserve the appealing reasoning principles of modal logics.
-% This becomes \replace{most}{more} apparent when discussing our proof rules.
-% }
 However, we must still build up from existing \iris primitives to provide new primitives that depend on the address space --- primarily the notion
@@
 value of the register $\rv$.
-The fraction $\qfrac$ with value 1 asserts the unique ownership of the register mapping and grants update permission \replace{on}{to} it;
+The fraction $\qfrac$ with value 1 asserts the unique ownership of the register mapping and grants update permission {to} it;
 otherwise, any value $0 < \qfrac <1$ represents partial ownership, granting read-only permission on the mapping.\footnote{
@@
   step of Figure \ref{fig:pagetables}}.\footnote{\add{Note the offsets mentioned in Figure \ref{fig:pagetables} are 9-bit indexes into the 512 entries; the byte offset is that times 8.}}
-  \replace{
-    Then {at} that physical location {is} the appropriate entry in the L4 table \textsf{l4e}.
-  }{
     Thus Line \ref{line:l4pointsto} asserts that the physical address built from the table base and offset points to the L4 entry \textsf{l4e}.
-  }
   Subsequent levels of the page table walk \add{assertion (Lines \ref{line:l3pointsto}--\ref{line:l1pointsto})} work similarly.
@@
 }
-Our solution \replace{removes}{in the next subsection improves on theirs, removing} some restrictions in this definition by further abstracting the handling of address translation.
+Our solution in the next subsection improves on theirs, removing some restrictions in this definition by further abstracting the handling of address translation.
 \looseness=-1
@@
 yielding Figure \ref{fig:virtualpointstosharing}'s definition.
-This new definition guarantees that the ghost map \add{$\theta$} \replace{contains an mapping from}{maps} the virtual address ($\vaddr$) to a physical address ($\paddr$),
+This new definition guarantees that the ghost map \add{$\theta$} maps the virtual address ($\vaddr$) to a physical address ($\paddr$),
 and thus that the per-address-space invariant \add{described next} will contain the physical resources that guarantee \add{that} the hardware resolves the translation.
@@
 $ I$\textsf{ASpace} (Figure \ref{fig:peraspaceinvariant}), 
-\replace{$ I$\textsf{ASpace} alone allows}{allowing} changes to the page tables
+{allowing} changes to the page tables
 that preserve overall virtual-to-physical translations \add{in isolation},
 \add{and also allowing changes to specific the virtual-to-physical translations}
-\replace{W}{w}hen combined with the
-\replace{fragment}{token} stored in the \add{relevant} virtual points-to (Figure \ref{fig:virtualpointstosharing}).
+when combined with the
+{token} stored in the \add{relevant} virtual points-to (Figure \ref{fig:virtualpointstosharing}).
 \looseness=-1
@@
 soundness of an individual virtual memory access.
-\replace{But}{However,} such larger invariants would contain many kernel-specific properties that are orthogonal
+However, such larger invariants would contain many kernel-specific properties that are orthogonal
 to the fundamental reasoning principles that are the focus of this paper.
@@
 
-\replace{However, }{First,}
+First,
 because we split the physical resources for the page table walk from the
-virtual points-to itself (\replace{in order to permit the physical page tables to be freely modified
-as long as they preserve virtual-to-physical translations}{per the discussion of Section \ref{sec:sharingpages}}), the rule requires $ I\textsf{ASpace}$
+virtual points-to itself ({per the discussion of Section \ref{sec:sharingpages}}), the rule requires $ I\textsf{ASpace}$
 for the current address space to be carried through.
@@
 
-\replace{T}{Second, t}he memory access rules --- as with all rules in our logic ---
-\replace{model updates to}{increment} the instruction pointer \lstinline|rip| \add{by the length of the encoded instruction.}
+Second, the memory access rules --- as with all rules in our logic ---
+{increment} the instruction pointer \lstinline|rip| \add{by the length of the encoded instruction.}
 \textsf{MovLen} returns how long the instruction encoding for the corresponding \lstinline|mov| is;
@@
 both the precondition and the postcondition assert that $a\mapsto_\mathsf{v} x$ in the current address space, but
-\replace{we have no basis for concluding that address translation is preserved by the change of address space}{
-  the new address space may map $a$ to another value}. So, this derivation clearly leads to an unsound conclusion. 
+  the new address space may map $a$ to another value. So, this derivation clearly leads to an unsound conclusion. 
 The essential problem is that the frame rule is motivated by local reasoning about local updates, but

### 2025-08-22 pack_camera_ready.sh
add script to package cameraready zip

@@ -1,0 +1,2 @@
+#!/usr/bin/env bash
+zip cameraready.zip addr_translation.png *.cls *.sty paper.tex paper.bbl intro.tex background.tex syntaxsemantics.tex programlogic.tex examples_C.tex relwork.tex vmm.bib ACM-Reference-Format.bst
